### 2022-05-09 frontmatter.tex
add working hepthesis template

%% Title
\titlepage[of Homerton College]{%
  A dissertation submitted to the University of Cambridge\\ for the degree of Doctor of Philosophy}

%% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
  %\thispagestyle{empty}
  \LHCb is a \bphysics detector experiment which will take data at
  the \SI{14}{\Hz} \LHC accelerator at \CERN from 2007 onward\dots
\end{abstract}

%% Declaration
\begin{declaration}
  This dissertation is the result of my own work, except where explicit
  reference is made to the work of others, and has not been submitted
  for another qualification to this or any other university. This
  dissertation does not exceed the word limit for the respective Degree
  Committee.
  \vspace*{1cm}
  \begin{flushright}
    Andy Buckley
  \end{flushright}
\end{declaration}

%% Acknowledgements
\begin{acknowledgements}
  Of the many people who deserve thanks, some are particularly prominent,
  such as my supervisor\dots
\end{acknowledgements}

%% Preface
\begin{preface}
  This thesis describes my research on various aspects of the \LHCb
  particle physics program, centred around the \LHCb detector and \LHC
  accelerator at \CERN in Geneva.

  \noindent
  For this example, I'll just mention \ChapterRef{chap:SomeStuff}
  and \ChapterRef{chap:MoreStuff}.
\end{preface}

%% ToC
\tableofcontents

%% Strictly optional!
\frontquote{%
  Writing in English is the most ingenious torture\\
  ever devised for sins committed in previous lives.}%
  {James Joyce}
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}

### 2022-05-09 frontmatter.tex
try to reduce template bloat

@@
   %\thispagestyle{empty}
-  \LHCb is a \bphysics detector experiment which will take data at
-  the \SI{14}{\Hz} \LHC accelerator at \CERN from 2007 onward\dots
+  \atlas\ is a detector experiment at \cern. $14~\eV{T}$.
 \end{abstract}
@@
   \begin{flushright}
-    Andy Buckley
+    Your Name
   \end{flushright}
@@
 \begin{preface}
-  This thesis describes my research on various aspects of the \LHCb
-  particle physics program, centred around the \LHCb detector and \LHC
-  accelerator at \CERN in Geneva.
+  This thesis describes my research on various hoozywotsists.
 
   \noindent
-  For this example, I'll just mention \ChapterRef{chap:SomeStuff}
-  and \ChapterRef{chap:MoreStuff}.
+  For this example, I'll just mention Chapter~\ref{chapter:intro}.
+  and Chapter~\ref{chapter:intro}..
 \end{preface}
@@
 
-%% Strictly optional!
-\frontquote{%
-  Writing in English is the most ingenious torture\\
-  ever devised for sins committed in previous lives.}%
-  {James Joyce}
-%% I don't want a page number on the following blank page either.
 \thispagestyle{empty}

### 2022-05-10 main/frontmatter.tex
rename tex to main; simplify main; document Makefile better

--- 2022-05-09 frontmatter.tex
@@
-%% Title
 \titlepage[of Homerton College]{%
@@
 
-%% Abstract
-\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
-  %\thispagestyle{empty}
-  \atlas\ is a detector experiment at \cern. $14~\eV{T}$.
+\begin{abstract}
+\atlas\ is a detector experiment at \cern. $14~\eV[T]$.
 \end{abstract}
 
-%% Declaration
 \begin{declaration}
@@
 
-%% Acknowledgements
 \begin{acknowledgements}
@@
 such as my supervisor\dots
+This thesis was made in \LaTeXe{} using the ``hepthesis'' class~\cite{hepthesis}.
 \end{acknowledgements}
 
-%% Preface
 \begin{preface}
-  This thesis describes my research on various hoozywotsists.
+This thesis describes my research on various hoozywotsists.%
+\footnote{test.}
 
@@
 
-%% ToC
 \tableofcontents

### 2022-05-11 main/chapter_2ljets.tex
set out chapters, add front quote

\chapter{A search for new phenomena}
\label{chapter:2ljets}

\chapterquote{Laws were made to be broken.}%
{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830

Hello

\clearpage

Hello again

### 2022-05-11 main/chapter_2ljets.tex
begin adding sections and quotes

@@
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
-
-Hello
+\chapterquote{An experiment is never a failure solely because it fails to
+achieve predicted results. An experiment is a failure only when it also
+fails adequately to test the hypothesis in question, when the data it
+produces don’t prove anything one way or another}%
+{Robert M. Pirsig~\cite{pirsig1999zen}}
 
@@
 Hello again
+
+Hello

### 2022-05-11 main/chapter_experiment.tex
set out chapters, add front quote
begin adding sections and quotes

--- 2022-05-11 main/chapter_2ljets.tex
@@
-\chapter{A search for new phenomena}
-\label{chapter:2ljets}
+\chapter{The ATLAS experiment}
+\label{chapter:experiment}
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
+\chapterquote{%
+if observing outer space gives us a view of the past, observing inner space
+would surely give us a glimpse into the future - would be interesting if NASA
+made a telescope for that%
+}%
+{Ken M}
 
 Hello
-
-\clearpage
-
-Hello again

### 2022-05-11 main/chapter_searches.tex
set out chapters, add front quote
begin adding sections and quotes

--- 2022-05-11 main/chapter_2ljets.tex
@@
-\chapter{A search for new phenomena}
-\label{chapter:2ljets}
+\chapter{Searches in ATLAS}
+\label{chapter:intro}
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
+\chapterquote{%
+Within this theory, statistical methods of great practical usefulness have been
+developed, and its statements can and frequently do contribute in a vague way
+to the interpretation of data. \ldots%
+}%
+{John W. Pratt~\cite{pratt1961testing}}
 
-Hello
+\section{Data analysis}
+
+\subsection{Frequentist methods}
+
+\chapterquote{%
+\ldots But this book, by its very excellence, its thoroughness, lucidity and
+precision, intensifies my growing feeling that nevertheless the theory is
+arbitrary, be it however ``objective,'' and the problems it solves, however
+precisely it may solve them, are not even simplified theoretical counterparts
+of the real problems to which it is applied.%
+}%
+{John W. Pratt~\cite{pratt1961testing}}
+
+\subsection{Post-frequentist methods}
+
+\section{Searches}
 

### 2022-05-11 main/chapter_theory.tex
set out chapters, add front quote
begin adding sections and quotes

--- 2022-05-11 main/chapter_2ljets.tex
@@
-\chapter{A search for new phenomena}
-\label{chapter:2ljets}
+\chapter{The Standard Model and beyond}
+\label{chapter:theory}
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
+\chapterquote{%
+The idea that theorems follow from the postulates does not correspond to simple
+observation.
+If the Pythagorean theorem were found to not follow from the postulates, we
+would again search for a way to alter the postulates until it was true.
+Euclid's postulates came from the Pythagorean theorem, not the other way.}%
+{Richard Hamming~\cite{hamming1980unreasonable}}
+
+\chapterquote{%
+One geometry can not be more true than another; it can only be more convenient.
+Geometry is not true, it is advantageous.}%
+{Robert M. Pirsig~\cite{pirsig1999zen}}
+
+% Standard Model past present future
 
@@
 
-\clearpage
+\section{Particles}
 
-Hello again
+\section{The Standard Model}
+
+\section{Supersymmetric models}
+
+\subsection{Minimal supersymmetric models}
+
+\subsection{Guage Mediated Supersymmetry Breaking}

### 2022-05-11 main/frontmatter.tex
set out chapters, add front quote
begin adding sections and quotes

@@
 \titlepage[of Homerton College]{%
-  A dissertation submitted to the University of Cambridge\\ for the degree of Doctor of Philosophy}
+A dissertation submitted to the University of Cambridge\\
+for the degree of Doctor of Philosophy%
+}
 
@@
 \begin{flushright}
-Your Name
+Rupert Tombs
 \end{flushright}
@@
 such as my supervisor\dots
-This thesis was made in \LaTeXe{} using the ``hepthesis'' class~\cite{hepthesis}.
+\\
+This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
 \end{acknowledgements}
@@
 
+\pagequote{%
+Begin reading at Chapter N.
+Do \textbf{not} read the quotations that appear at the\\
+beginning of the chapter.%
+}%
+{Donald E. Knuth~\cite{knuth1975art}}
+
+% the following blank page should also have no page number
 \thispagestyle{empty}

### 2022-05-12 main/chapter_2ljets.tex
use epigraph; add hoare quote; se bibtex tidy

--- 2022-05-11 main/chapter_2ljets.tex
@@
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
-
-Hello
-
+\begin{epigraphs}
+\qitem{%
+An experiment is never a failure solely because it fails to
+achieve predicted results. An experiment is a failure only when it also
+fails adequately to test the hypothesis in question, when the data it
+produces don’t prove anything one way or another.%
+}%
+{Robert M. Pirsig~\cite{pirsig1999zen}}
+\qitem{%
+But there is one feature I notice that is generally missing in cargo cult
+science. \ldots\
+% That is the idea that we all hope you have learned in studying science in
+% school--we never say explicitly what this is, but just hope that you catch
+% on by all the examples of scientific investigation. It is interesting,
+% therefore, to bring it out now and speak of it explicitly.
+It's a kind of scientific integrity, a principle of scientific thought that
+corresponds to a kind of utter honesty --- a kind of leaning over backwards.%
+}%
+{Richard Feynman~\cite{feynman1974cargo}}
+\end{epigraphs}
 \clearpage

### 2022-05-12 main/chapter_experiment.tex
use epigraph; add hoare quote; se bibtex tidy

@@
 
-\chapterquote{%
+\epigraph{%
 if observing outer space gives us a view of the past, observing inner space
-would surely give us a glimpse into the future - would be interesting if NASA
+would surely give us a glimpse into the future --- would be interesting if NASA
 made a telescope for that%

### 2022-05-12 main/chapter_searches.tex
use epigraph; add hoare quote; se bibtex tidy

--- 2022-05-11 main/chapter_2ljets.tex
@@
-\chapter{A search for new phenomena}
-\label{chapter:2ljets}
+\chapter{Searches in ATLAS}
+\label{chapter:intro}
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
+\epigraph{%
+One way is to make it so simple that there are \emph{obviously} no deficiencies and the other way is to make it so complicated that there are
+no \emph{obvious} deficiencies.%
+}%
+{Tony Hoare~\cite{hoare2007emperor}}
 
-Hello
+\section{Data analysis}
+
+\section{Frequentist theory}
+
+\epigraph{%
+Within this theory, statistical methods of great practical usefulness have been
+developed, and its statements can and frequently do contribute in a vague way
+to the interpretation of data. \ldots%
+}%
+{John W. Pratt~\cite{pratt1961testing}}
+
+\section{Post-frequentist practice}
+
+\epigraph{%
+\ldots\ But this book, by its very excellence, its thoroughness, lucidity and
+precision, intensifies my growing feeling that nevertheless the theory is
+arbitrary, be it however ``objective,'' and the problems it solves, however
+precisely it may solve them, are not even simplified theoretical counterparts
+of the real problems to which it is applied.%
+}%
+{John W. Pratt~\cite{pratt1961testing}}
+
+\section{Searches}
 

### 2022-05-12 main/chapter_theory.tex
use epigraph; add hoare quote; se bibtex tidy

@@
 
-\chapterquote{%
-The idea that theorems follow from the postulates does not correspond to simple
-observation.
-If the Pythagorean theorem were found to not follow from the postulates, we
-would again search for a way to alter the postulates until it was true.
+\begin{epigraphs}
+\qitem{%
+% The idea that theorems follow from the postulates does not correspond to simple
+% observation.
+%If the Pythagorean theorem were found to not follow from the postulates, we
+%would again search for a way to alter the postulates until it was true.
 Euclid's postulates came from the Pythagorean theorem, not the other way.}%
 {Richard Hamming~\cite{hamming1980unreasonable}}
-
-\chapterquote{%
+\qitem{%
 One geometry can not be more true than another; it can only be more convenient.
@@
 {Robert M. Pirsig~\cite{pirsig1999zen}}
+\end{epigraphs}
 

### 2022-05-12 main/frontmatter.tex
use epigraph; add hoare quote; se bibtex tidy

@@
 \begin{preface}
+\epigraph{%
+Begin reading at Chapter N. Do \emph{not} read the quotations
+that appear at the beginning of the chapter.%
+}%
+{Donald E. Knuth~\cite{knuth1975art}}
+
 This thesis describes my research on various hoozywotsists.%
@@
 
-\pagequote{%
-Begin reading at Chapter N.
-Do \textbf{not} read the quotations that appear at the\\
-beginning of the chapter.%
-}%
-{Donald E. Knuth~\cite{knuth1975art}}
-
 % the following blank page should also have no page number

### 2022-05-13 main/chapter_2ljets.tex
add 2ljets summary plot; replace broken bib file

@@
 \end{epigraphs}
-\clearpage
 
-Hello again
-
-Hello
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
+\caption{%
+Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
+for which the lower panel shows $S_\mathrm{ATLAS}$ from
+Equation~\ref{eqn:significance_atlas}.
+Control, validation and signal regions are shown from left to right, with the
+regions within each category ordered approximately by their typical $\ptmiss$.
+The `Top' category contains $t\bar t$ and $tW$ processes, and
+`Other' contains fake/non-prompt lepton, higgs, triboson, , $t\bar tZ$ and other
+rare top processes.%
+}
+\label{fig:ewk_summary}
+\end{figure}

### 2022-05-13 main/chapter_2ljets.tex
add exclusion plots

@@
 
+Our central result is Figure~\ref{fig:2ljets_summary}, which shows the names,
+data, and fitted background expectations with error bars in all main regions.
+This figure illustrates the background-model part of a \clown{likelihood}
+function which is extended with signals to test alternative models.
+Exclusion contours testing the addition of C1N2 and GMSB signals to these
+backgrounds are shown in Figure~\ref{fig:2ljets_contours}.
+
 \begin{figure}[tp]
@@
 regions within each category ordered approximately by their typical $\ptmiss$.
+Likelihoods from validation regions are not included in the fit.
 The `Top' category contains $t\bar t$ and $tW$ processes, and
-`Other' contains fake/non-prompt lepton, higgs, triboson, , $t\bar tZ$ and other
+`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
 rare top processes.%
 }
-\label{fig:ewk_summary}
+\label{fig:2ljets_summary}
 \end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{\textwidth}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/2ljets_contours_c1n2.pdf}
+    \caption{C1N2 contours}
+\end{subfigure}
+\\[1em]
+\begin{subfigure}{\textwidth}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/2ljets_contours_gmsb.pdf}
+    \caption{GMSB contours}
+\end{subfigure}
+\caption{%
+Contours from the $\twoljets$ electroweak analysis for
+C1N2 (top) and GMSB (bottom) models.
+Space below the solid red line is labelled as excluded and its dotted
+neighbours show the result if all signal cross-sections are varied up and down
+by theoretical error bars.
+The yellow band shows the prior $\pm1$-sigma region of exclusion contours
+from asymptotic approximations to the distribution of the test statistic.
+Grey areas are observed limits from the $2\ell$ parts
+of~\cite{SUSY-2016-24} and~\cite{SUSY-2013-11}.
+Exclusion is defined by the $\mathrm{CLs}$ prescription at the $0.05$ level
+in asymptotic approximations.
+All contours are interpolated from a sparse grid.
+}
+\label{fig:2ljets_contours}
+\end{figure}

### 2022-05-13 main/chapter_2ljets.tex
add discovery region results table

@@
 backgrounds are shown in Figure~\ref{fig:2ljets_contours}.
+Interpretations of discovery region data are displayed in
+Table~\ref{tab:2ljets_discovery}.
+
+% summary plot
 
@@
 \end{figure}
+
+% exclusion plots
 
@@
 by theoretical error bars.
-The yellow band shows the prior $\pm1$-sigma region of exclusion contours
-from asymptotic approximations to the distribution of the test statistic.
-Grey areas are observed limits from the $2\ell$ parts
+The yellow band shows the $\pm1$-sigma region of exclusion contours
+from asymptotic approximations to the prior distribution of the test statistic.
+Grey areas are observed limits from the two-lepton parts
 of~\cite{SUSY-2016-24} and~\cite{SUSY-2013-11}.
-Exclusion is defined by the $\mathrm{CLs}$ prescription at the $0.05$ level
+Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
 in asymptotic approximations.
@@
 \end{figure}
+
+% upper limits
+\begin{table}[tp]
+\adjustbox{max width=0.98\textwidth}{%
+\centering
+\begin{tabular*}{\textwidth}{lccccccc}
+{Region} &
+\multicolumn{1}{c}{Fitted} &
+\multicolumn{1}{l}{Data} &
+$\langle A\epsilon{ \sigma}\rangle_{\mathrm{obs}}^{95}~\mathrm{fb}$ &
+\multicolumn{1}{c}{$S_{\mathrm{obs}}^{95}$}  &
+\multicolumn{1}{c}{$S_{\mathrm{exp}}^{95}$} &
+$\mathrm{CLb}$ &
+$p(s=0)~(Z)$  \\[1.5ex]
+DR-OffShell      & $22.1\pm2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50~(0.0)$ \\[.5ex]
+DR-Low           & $22\pm4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50~(0.0)$ \\[.5ex]
+DR-Int           & $35\pm4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23~(0.8)$ \\[.5ex]
+DR-High          & $3.9\pm0.5$  & 0  & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50~(0.0)$ \\[.5ex]
+DR-$\ell\ell bb$ & $0.51\pm0.20$  & 0  & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50~(0.0)$ \\[.5ex]
+\end{tabular*}
+}
+\caption{%
+Upper limit results in discovery region.
+The fitted yield is in the background-only model constrained by the data in each region.
+Limits are intended to reflect constrains on additive signal contributions.
+\\[0.5em]
+Left to right: the observed $95\%$ $\mathrm{CLs}$ upper limit on the visible cross-section
+$\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
+its corresponding signal expectation $S_\mathrm{obs}^{95}$,
+expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
+as would be obtained were the test statistic given by its central of $\pm 1$-sigma variations,
+$\mathrm{CLb}$ evaluated with the signal expectation set to its observed upper limit,
+and the discovery $p$-value capped at $0.5$ with its equivalent significance.
+\\[0.5em]
+Upper limits use the one-sided profile likelihood test statistic.
+The discovery $p$-value uses a profile likelihood test statistic in a one-sided test.
+All $p$-values are estimated by simulation of alternative data.
+}
+\label{tab:2ljets_discovery}
+\end{table}

### 2022-05-13 main/chapter_2ljets.tex
fix linespacing

@@
+\begin{singlespacing}
 \chapter{A search for new phenomena}
 \label{chapter:2ljets}
-
+%
 \begin{epigraphs}
@@
 \end{epigraphs}
+\end{singlespacing}
 
@@
 data, and fitted background expectations with error bars in all main regions.
-This figure illustrates the background-model part of a \clown{likelihood}
-function which is extended with signals to test alternative models.
+This figure illustrates the background part of a \clown{likelihood}
+which is extended with signals to test alternative models.
 Exclusion contours testing the addition of C1N2 and GMSB signals to these
@@
 Table~\ref{tab:2ljets_discovery}.
+
+\textls[1000]{foo bar}
 
@@
 \begin{table}[tp]
-\adjustbox{max width=0.98\textwidth}{%
+\adjustbox{max width=0.96\textwidth}{%
 \centering
@@
 All $p$-values are estimated by simulation of alternative data.
+{\color{red} TODO add labels and references to asymptotic formulae paper}
 }

### 2022-05-13 main/chapter_experiment.tex
fix linespacing

@@
+\begin{singlespacing}
 \chapter{The ATLAS experiment}
 \label{chapter:experiment}
-
-\epigraph{%
+%
+\begin{epigraphs}
+\qitem{%
 if observing outer space gives us a view of the past, observing inner space
@@
 {Ken M}
+\end{epigraphs}
+\end{singlespacing}
 

### 2022-05-13 main/chapter_searches.tex
add 2ljets summary plot; replace broken bib file
add exclusion plots
add discovery region results table
fix linespacing

@@
+\begin{singlespacing}
 \chapter{Searches in ATLAS}
 \label{chapter:intro}
-
-\epigraph{%
-One way is to make it so simple that there are \emph{obviously} no deficiencies and the other way is to make it so complicated that there are
-no \emph{obvious} deficiencies.%
+%
+\begin{epigraphs}
+\qitem{%
+One way is to make it so simple that there are \emph{obviously} no deficiencies
+and the other way is to make it so complicated that there are no \emph{obvious}
+deficiencies.%
 }%
 {Tony Hoare~\cite{hoare2007emperor}}
+\end{epigraphs}
+\end{singlespacing}
 
@@
 \section{Frequentist theory}
-
-\epigraph{%
+\begin{epigraphs}
+\qitem{%
 Within this theory, statistical methods of great practical usefulness have been
@@
 {John W. Pratt~\cite{pratt1961testing}}
+\end{epigraphs}
 
 \section{Post-frequentist practice}
-
-\epigraph{%
+\begin{epigraphs}
+\qitem{%
 \ldots\ But this book, by its very excellence, its thoroughness, lucidity and
@@
 {John W. Pratt~\cite{pratt1961testing}}
+\end{epigraphs}
+
+% significancen measures
+The ATLAS-recommended~\cite{atlas_significance} significance measure is
+\begin{equation}
+\label{eqn:significance_atlas}
+S_\mathrm{ATLAS}(n, \mu, \sigma)
+= \mathrm{sgn}(n - \mu) \sqrt{2\left(
+    n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
+    - \frac{\mu^2}{\sigma^2}\log\left(
+        1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
+    \right)
+\right)}
+~.
+\end{equation}
 

### 2022-05-16 main/chapter_2ljets.tex
add susy signal plots; various tweaks: todo command, ulem for wavy underline

@@
 
+% begin with physics content and results
+
 Our central result is Figure~\ref{fig:2ljets_summary}, which shows the names,
@@
 Exclusion contours testing the addition of C1N2 and GMSB signals to these
-backgrounds are shown in Figure~\ref{fig:2ljets_contours}.
+backgrounds are shown in Figures~\ref{fig:2ljets_contours_c1n2}
+and~\ref{fig:2ljets_contours_gmsb}.
 Interpretations of discovery region data are displayed in
 Table~\ref{tab:2ljets_discovery}.
-
-\textls[1000]{foo bar}
 
@@
 
+% TODO signal region plots here
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.48\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
+\quad
+\includegraphics[width=0.45\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
+\caption{%
+Supersymmetric signal processes in the $\twoljets$ electroweak analysis.
+\\[0.5em]
+Left: C1N2, where the initial $\chargino{1}\textrm{--}\neutralino{2}$ pair
+is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
+weak bosons in decays are bounded by the mass splitting
+$m(\chargino{1}, \neutralino{2}) - m(\neutralino{1})$.
+We explore the parameters
+$m(\chargino{1}, \neutralino{2})$ and $m(\neutralino{1})$.
+\\[0.5em]
+Right: GMSB, where the initial $\neutralino{1}\textrm{--}\neutralino{1}$ pair
+is produced by soft decays from pairs including $\chargino{1}$,
+$\neutralino{2}$ or $\neutralino{1}$.
+Although the $h$ and $Z$ bosons decay to many states, we target
+$Z\rightarrow \ell\ell$ with
+$h/Z\rightarrow bb/jj$.
+We explore the parameters
+$m(\neutralino{1})$ and $B(\neutralino{1} \rightarrow h \tilde{G})$ with fixed
+$m(\chargino{1}, \neutralino{2}) - m(\neutralino{1}) = 1~\eV[G]$.
+}
+\label{fig:2ljets_summary}
+\end{figure}
+
 % exclusion plots
@@
 \centering
-\begin{subfigure}{\textwidth}
-    \centering
-    \includegraphics[width=0.7\textwidth]{figures/2ljets_contours_c1n2.pdf}
-    \caption{C1N2 contours}
-\end{subfigure}
-\\[1em]
-\begin{subfigure}{\textwidth}
-    \centering
-    \includegraphics[width=0.7\textwidth]{figures/2ljets_contours_gmsb.pdf}
-    \caption{GMSB contours}
-\end{subfigure}
+\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
 \caption{%
-Contours from the $\twoljets$ electroweak analysis for
-C1N2 (top) and GMSB (bottom) models.
+Contours for the C1N2 model in the $\twoljets$ electroweak analysis.
 Space below the solid red line is labelled as excluded and its dotted
@@
 }
-\label{fig:2ljets_contours}
+\label{fig:2ljets_contours_c1n2}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
+\caption{%
+Contours for the GMSB model in the $\twoljets$ electroweak analysis.
+Space below the solid red line is labelled as excluded and its dotted
+neighbours show the result if all signal cross-sections are varied up and down
+by theoretical error bars.
+The yellow band shows the $\pm1$-sigma region of exclusion contours
+from asymptotic approximations to the prior distribution of the test statistic.
+Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
+in asymptotic approximations.
+All contours are interpolated from a sparse grid.
+}
+\label{fig:2ljets_contours_gmsb}
 \end{figure}
@@
 % upper limits
+% definitely want this after pictures
+\FloatBarrier
 \begin{table}[tp]
-\adjustbox{max width=0.96\textwidth}{%
 \centering
@@
 $\mathrm{CLb}$ &
-$p(s=0)~(Z)$  \\[1.5ex]
-DR-OffShell      & $22.1\pm2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50~(0.0)$ \\[.5ex]
-DR-Low           & $22\pm4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50~(0.0)$ \\[.5ex]
-DR-Int           & $35\pm4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23~(0.8)$ \\[.5ex]
-DR-High          & $3.9\pm0.5$  & 0  & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50~(0.0)$ \\[.5ex]
-DR-$\ell\ell bb$ & $0.51\pm0.20$  & 0  & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50~(0.0)$ \\[.5ex]
+$p(s=0)$  \\[1.5ex]
+DR-OffShell      & $22.1\pm2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$ \\[.5ex]
+DR-Low           & $22\pm4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$ \\[.5ex]
+DR-Int           & $35\pm4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$ \\[.5ex]
+DR-High          & $3.9\pm0.5$  & 0  & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50$ \\[.5ex]
+DR-$\ell\ell bb$ & $0.51\pm0.20$  & 0  & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
 \end{tabular*}
-}
 \caption{%
@@
 \\[0.5em]
-Left to right: the observed $95\%$ $\mathrm{CLs}$ upper limit on the visible cross-section
+Left to right:
+the region name,
+the post-fit background expectation,
+the number of data observed,
+the observed $95\%$ $\mathrm{CLs}$ upper limit on the visible cross-section
 $\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
 its corresponding signal expectation $S_\mathrm{obs}^{95}$,
-expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
-as would be obtained were the test statistic given by its central of $\pm 1$-sigma variations,
-$\mathrm{CLb}$ evaluated with the signal expectation set to its observed upper limit,
-and the discovery $p$-value capped at $0.5$ with its equivalent significance.
+the \clown{expected} $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
+as would be obtained were the test statistic given by its central or
+$\pm1$-sigma variations,
+$\mathrm{CLb}$ evaluated with the signal expectation set to the observed upper limit,
+and the discovery $p$-value (capped at $0.5$) with its equivalent significance.
 \\[0.5em]
@@
 All $p$-values are estimated by simulation of alternative data.
-{\color{red} TODO add labels and references to asymptotic formulae paper}
+\TODO{add labels and references to asymptotic formulae paper}
 }
@@
 \end{table}
+
+\paragraph{My contributions}
+\begin{itemize}
+\item Design, implementation and execution of the electroweak part of the analysis.
+\item Analysis Contact from January to October 2020.
+\item Production of main data inputs for all three analyses
+and the RJR $3\ell$ search \TODO{cite}.
+I produced all systematic variations for the background samples and all
+electroweak signal samples.
+Other group members assisted with the central background sample.
+\item Integration with ATLAS combinations and pMSSM scan efforts by
+performing their validation studies and serializing the analysis results.
+\item Preparation of electroweak results for publication in the paper, and the
+public HEPData record~\cite{maguire2017hepdata}.
+\end{itemize}
+
+\section{Context}
+% previous results on 2(/3)L from ATLAS
+% other constraints on these models from ATLAS/CMS
+% previous region design, motivation for this work
+
+\section{Design}
+
+% CRs go in with relevant SR category
+\subsection{High}
+
+\subsection{Intermediate}
+
+\subsection{Low}
+
+\subsection{Off-shell}
+
+\subsection{Discovery}
+
+\section{Modelling}
+% backgorund samples
+% matrix method fakes
+
+\subsection{Fake/non-prompt leptons}
+
+\section{Validation}
+
+\section{Systematics}
+
+\section{Results}

### 2022-05-16 main/chapter_theory.tex
fix linespacing
add susy signal plots; various tweaks: todo command, ulem for wavy underline

--- 2022-05-11 main/chapter_theory.tex
@@
+\begin{singlespacing}
 \chapter{The Standard Model and beyond}
 \label{chapter:theory}
-
-\chapterquote{%
+%
+\begin{epigraphs}
+\qitem{%
 The idea that theorems follow from the postulates does not correspond to simple
@@
 {Richard Hamming~\cite{hamming1980unreasonable}}
-
-\chapterquote{%
+\qitem{%
 One geometry can not be more true than another; it can only be more convenient.
@@
 {Robert M. Pirsig~\cite{pirsig1999zen}}
+\end{epigraphs}
+\end{singlespacing}
 

### 2022-05-16 main/chapter_2ljets.tex
add signal yield example plots; fix figures gitignore; align +- in dr table

@@
 
-Our central result is Figure~\ref{fig:2ljets_summary}, which shows the names,
-data, and fitted background expectations with error bars in all main regions.
-This figure illustrates the background part of a \clown{likelihood}
-which is extended with signals to test alternative models.
-Exclusion contours testing the addition of C1N2 and GMSB signals to these
-backgrounds are shown in Figures~\ref{fig:2ljets_contours_c1n2}
-and~\ref{fig:2ljets_contours_gmsb}.
-Interpretations of discovery region data are displayed in
+Our main result is the content of Figure~\ref{fig:2ljets_summary}, which shows
+the data in all main regions of this analysis, along with the regions' names,
+and post-fit background expectations.
+This figure illustrates the background-only part of a \clown{likelihood}
+model which is extended with signals to test alternative models.
+
+Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
+Figure~\ref{fig:2ljets_signal_diagrams}.
+Some example effects of these signals in signal regions are displayed in
+Figure~\ref{fig:2ljets_signal_examples}.
+
+At each parameter point, each signal model is added to the background model
+and tested against data in the $\mathrm{CLs}$ prescription.
+Contours showing which regions of signal model parameters are labelled as
+excluded are shown for C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
+GMSB in~\ref{fig:2ljets_contours_gmsb}.
+
+Data in discovery regions are presented and interpreted in
 Table~\ref{tab:2ljets_discovery}.
@@
 
-% TODO signal region plots here
+% signal diagrams
 
@@
 \includegraphics[width=0.48\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
-\quad
+\hfill
 \includegraphics[width=0.45\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
@@
 }
-\label{fig:2ljets_summary}
+\label{fig:2ljets_signal_diagrams}
+\end{figure}
+
+% signal region plots
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
+    \caption{SR-High-8}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
+    \caption{SR-$\ell\ell bb$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
+    \caption{SR-Int}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
+    \caption{SR-Low}
+\end{subfigure}
+\caption{%
+Example signal region histograms with benchmark signal sample yields overlaid
+as dotted lines.
+Physically, signal yields would add on top of the backgrounds.
+Data are shown; regions in the the top two plots observe zero data, and the
+Poisson error bars for $0$ events are hidden for aesthetic reasons.
+}
+\label{fig:2ljets_signal_examples}
 \end{figure}
@@
 \centering
-\begin{tabular*}{\textwidth}{lccccccc}
-{Region} &
-\multicolumn{1}{c}{Fitted} &
-\multicolumn{1}{l}{Data} &
+% custom separator for aligned +-
+% https://stackoverflow.com/a/2132998
+\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
+Region &
+\multicolumn{2}{c}{Fitted} &
+Data &
 $\langle A\epsilon{ \sigma}\rangle_{\mathrm{obs}}^{95}~\mathrm{fb}$ &
-\multicolumn{1}{c}{$S_{\mathrm{obs}}^{95}$}  &
-\multicolumn{1}{c}{$S_{\mathrm{exp}}^{95}$} &
+$S_{\mathrm{obs}}^{95}$  &
+$S_{\mathrm{exp}}^{95}$ &
 $\mathrm{CLb}$ &
 $p(s=0)$  \\[1.5ex]
-DR-OffShell      & $22.1\pm2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$ \\[.5ex]
-DR-Low           & $22\pm4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$ \\[.5ex]
-DR-Int           & $35\pm4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$ \\[.5ex]
-DR-High          & $3.9\pm0.5$  & 0  & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50$ \\[.5ex]
-DR-$\ell\ell bb$ & $0.51\pm0.20$  & 0  & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
+DR-OffShell      & $22.1$ & $2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$ \\[.5ex]
+DR-Low           & $22$ & $4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$ \\[.5ex]
+DR-Int           & $35$ & $4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$ \\[.5ex]
+DR-High          & $3.9$ & $0.5$  & 0  & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50$ \\[.5ex]
+DR-$\ell\ell bb$ & $0.51$ & $0.20$  & 0  & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
 \end{tabular*}

### 2022-05-16 main/chapter_2ljets.tex
add censored presel figures; begin event variables list

@@
 
+We use two signal models in the $\twoljets$ electroweak analysis.
+One is an MSSM chargino-neutralino model (C1N2) which acts to produce
+$\chargino{1}\textrm{--}\neutralino{2}$ pairs which decay via $W$ and $Z$
+channels respectively to jets and a lepton pair.
+The other is a GMSB model which decays via Higgs and $Z$ bosons in variable
+branching fractions.
 Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
@@
 
-At each parameter point, each signal model is added to the background model
+At each parameter point of a signal model, it is added to the background model
 and tested against data in the $\mathrm{CLs}$ prescription.
-Contours showing which regions of signal model parameters are labelled as
-excluded are shown for C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
+Contours showing which regions where those tests label the signal model
+parameters as excluded excluded are shown for
+C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
 GMSB in~\ref{fig:2ljets_contours_gmsb}.
@@
 Control, validation and signal regions are shown from left to right, with the
-regions within each category ordered approximately by their typical $\ptmiss$.
+regions within each category ordered approximately by their typical $\etmiss$.
 Likelihoods from validation regions are not included in the fit.
@@
 
+\clearpage
+\FloatBarrier
 \section{Context}
@@
 
+\section{Event variables}
+% lepton, jet definitions
+
+Uncomfortably many event variables are used to define the regions of this
+analysis.
+Definitions of these variables are collected here.
+All variables with a double object subscript (such as $\mll$, $\rjj$) refer to
+the first two of those particles when ordered in decreasing $\pt$.
+\begin{itemize}
+\item $\etmiss$: missing transverse momentum.
+\item $\etmisssig$: object-based missing transverse momentum significance.
+It scales $\etmiss$ by an uncertainty derived from the resolutions of
+reconstructed objects~\cite{atlas_met_significance}.
+\item $\mll$: invariant mass of the hardest two leptons.
+\item $\mjj$: invariant mass of the two hardest jets.
+\item $\mbb$: invariant mass of the two hardest b-tagged jets.
+\item $\mjetone$: mass of the hardest jet.
+% TODO
+% \item $\Rll$: estimated angular distance in $\eta,\phi$ between the lepton pair.
+% \item $\Rjj$: estimated angular distance in $\eta,\phi$ between the two hardest jets.
+% \item $\Delta\phi(p_{\ell\ell}, \Etmiss)$: estimated distance in azimuthal
+% angle between the vector sum of the lepton pair and the missing transverse momentum.
+% \item $\mttwo(\ell, \ell; 0)$: the `stransverse mass' between the two
+% leptons and the missing transverse momentum.
+% If two massive particles are pair produced and each decays to a
+% lepton and something massless and invisible,
+% then this estimates a greatest lower bound on the invariant
+% mass of the pair produced particles.
+% $\mttwo(\ell, \ell; 0)$ is useful for excluding standard model backgrounds;
+% di-leptonic ttbar ($tt\rightarrow WW\rightarrow\ell\nu\ell\nu$)
+% and $WW\rightarrow\ell\nu\ell\nu$ have a kinematic endpoint
+% at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.
+% \item $\nbtag$: number of b-tagged jets with $\pt > 20$ GeV,
+% where the tag is taken on the mv2c10 classifier at the $77\%$ working point.
+% \item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
+% \item $\pt(j_1)$: transverse momentum of the hardest jet.
+% \item $\Delta\phi(\mathrm{jet}_1,\etmiss)$: azimuthal angle between
+% the hardest jet and missing transverse momentum vector estimates.
+\end{itemize}
+Most of these are physical properties of particles, or collection of particles.
+Clearly, the stated values are not exact, but estimated values based on
+approximate reconstructions.
+While I think this distinction is important to remember, I have not spelled it
+out on every line.
+Brevity also has value.
+
+\FloatBarrier
 \section{Design}
+
+The most impactful decision in a search is the design of its regions.
+Like all decision in life, these designs must be made to balance conflicting
+and vaguely-specified desires, and are informed with incomplete information.
+
+This section describes the designs and rationales behind the control, validation
+and signal regions of the $\twoljets$ analysis.
+
+Splitting regions on $\etmisssig$ is the core idea behind the design.
+The basic motivation for this design are displayed in
+Figure~\ref{fig:2ljets_presel_met};
+$\etmisssig$ does better than $\etmiss$ in separating background and signal
+contributions.
+But there is more --- $\etmisssig$ also does a decent job of separating signal
+models from each other;
+signals with larger mass differences in their decays to invisibles tend to have
+larger $\etmisssig$.
+Binning regions in $\etmisssig$ therefore also improves sensitivity to
+individual parameter points and allows targeted selections on other, locally-relevant,
+event variables.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
+\hfill
+\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
+\caption{
+Illustration of how $\etmisssig$ beats $\etmiss$ alone in sensitivity to
+example signal models.
+At large $\etmisssig$, $t\bar t$ and other backgrounds are suppressed, leaving
+quite pure diboson backgrounds and greater significance measures.
+Signal models with smaller mass splittings tend to have the bulk of their
+events at smaller $\etmiss$.
+\\[0.5em]
+This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
+signal leptons with
+$\mll \in (71, 111)~\eV[G]$,
+$\mjj \in (60, 110)~\eV[G]$,
+$\nbtag \leq 1$,
+$\etmiss > 100~\eV[G]$, and
+$\etmisssig > 6$.
+\\[0.5em]
+``Significance'' displayed in the lower subplot uses the Binomial significance
+$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
+uncertainty and yields taken to the right of a given value.
+Large values indicate expected sensitivity to the signal model.
+}
+\label{fig:2ljets_presel_met}
+\end{figure}
 
@@
 \section{Modelling}
-% backgorund samples
-% matrix method fakes
+% TODO backgorund samples
+% TODO matrix method fakes
+% TODO GMSB reweighting
+% TODO off-shell plus/minus split
+% TODO generator filtering
 

### 2022-05-16 main/chapter_2ljets.tex
improve todo macro

@@
 % https://stackoverflow.com/a/2132998
-\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
+\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lccrrcc}
 Region &
@@
 its corresponding signal expectation $S_\mathrm{obs}^{95}$,
-the \clown{expected} $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
+the expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
 as would be obtained were the test statistic given by its central or
@@
 \TODO{add labels and references to asymptotic formulae paper}
+\TODO{reference pdg rounding}
 }

### 2022-05-17 main/chapter_2ljets.tex
finish variables cheatsheet

@@
 Control, validation and signal regions are shown from left to right, with the
-regions within each category ordered approximately by their typical $\etmiss$.
+regions within each category ordered approximately by their typical $\met$.
 Likelihoods from validation regions are not included in the fit.
@@
 \item Design, implementation and execution of the electroweak part of the analysis.
-\item Analysis Contact from January to October 2020.
+\item Service as `analysis contact' from January to October 2020.
 \item Production of main data inputs for all three analyses
 and the RJR $3\ell$ search \TODO{cite}.
-I produced all systematic variations for the background samples and all
-electroweak signal samples.
-Other group members assisted with the central background sample.
+Production of all systematic variations for the background samples;
+other group members assisted with the central background sample.
+Production of all electroweak signal samples and their systematic variations.
 \item Integration with ATLAS combinations and pMSSM scan efforts by
 performing their validation studies and serializing the analysis results.
-\item Preparation of electroweak results for publication in the paper, and the
+\item Preparation of electroweak results for publication in the paper \TODO{cite} and the
 public HEPData record~\cite{maguire2017hepdata}.
@@
 
-Uncomfortably many event variables are used to define the regions of this
-analysis.
-Definitions of these variables are collected here.
-All variables with a double object subscript (such as $\mll$, $\rjj$) refer to
-the first two of those particles when ordered in decreasing $\pt$.
-\begin{itemize}
-\item $\etmiss$: missing transverse momentum.
-\item $\etmisssig$: object-based missing transverse momentum significance.
-It scales $\etmiss$ by an uncertainty derived from the resolutions of
+\subsection{$\met$ significance}
+\label{sec:metsig}
+It scales $\met$ by an uncertainty derived from the resolutions of
 reconstructed objects~\cite{atlas_met_significance}.
-\item $\mll$: invariant mass of the hardest two leptons.
-\item $\mjj$: invariant mass of the two hardest jets.
-\item $\mbb$: invariant mass of the two hardest b-tagged jets.
-\item $\mjetone$: mass of the hardest jet.
-% TODO
-% \item $\Rll$: estimated angular distance in $\eta,\phi$ between the lepton pair.
-% \item $\Rjj$: estimated angular distance in $\eta,\phi$ between the two hardest jets.
-% \item $\Delta\phi(p_{\ell\ell}, \Etmiss)$: estimated distance in azimuthal
+
+\subsection{Stranverse mass $\mttwo$}
+\label{sec:mt2}
 % angle between the vector sum of the lepton pair and the missing transverse momentum.
-% \item $\mttwo(\ell, \ell; 0)$: the `stransverse mass' between the two
 % leptons and the missing transverse momentum.
@@
 % at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.
-% \item $\nbtag$: number of b-tagged jets with $\pt > 20$ GeV,
+
+\subsection{b-tagging}
+\label{sec:btagging}
+% with $\pt > 20$ GeV,
 % where the tag is taken on the mv2c10 classifier at the $77\%$ working point.
-% \item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
-% \item $\pt(j_1)$: transverse momentum of the hardest jet.
-% \item $\Delta\phi(\mathrm{jet}_1,\etmiss)$: azimuthal angle between
-% the hardest jet and missing transverse momentum vector estimates.
+
+\subsection{Cheat-sheet}
+Uncomfortably many event variables are used to define the regions of this
+analysis.
+To aid future revision of their meanings, definitions and references for these
+variables are collected here.
+\begin{itemize}
+\item $\met$: missing transverse momentum
+\item $\metsig$: object-based missing transverse momentum significance;\\
+described in Section~\ref{sec:metsig}
+\item $\ptjone$: transverse momentum of the hardest jet%
+\vspace{0.5em}
+\item $\mll$: invariant mass of the hardest two leptons
+\item $\mjj$: invariant mass of the two hardest jets
+\item $\mbb$: invariant mass of the two hardest b-tagged jets
+\item $\mjetone$: mass of the hardest jet
+\item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
+described in Section~\ref{sec:mt2}
+\vspace{0.5em}
+\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
+\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
+\vspace{0.5em}
+\item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
+\item $\nbtag$: number of b-tagged jets;
+described in Section~\ref{sec:btagging}
+\vspace{0.5em}
+\item $\dphillmet$: azimuthal angle between the hardest jet and $\ptmiss$
+\item $\Delta\phi(\mathrm{jet}_1,\met)$: azimuthal angle between
+the hardest jet and $\ptmiss$
 \end{itemize}
-Most of these are physical properties of particles, or collection of particles.
-Clearly, the stated values are not exact, but estimated values based on
-approximate reconstructions.
-While I think this distinction is important to remember, I have not spelled it
-out on every line.
-Brevity also has value.
+All of these are physical properties of particles or collections of particles.
+Please remember that none is an exact or true value;
+all are are estimates based on approximate reconstructions.
+While this distinction is easy to drop in lax language, is necessary to
+understand the shapes of noisy variables such as $\mjj$ and indeed the
+existence of $\metsig$.
 
@@
 
-Splitting regions on $\etmisssig$ is the core idea behind the design.
+Splitting regions on $\metsig$ is the core idea behind the design.
 The basic motivation for this design are displayed in
 Figure~\ref{fig:2ljets_presel_met};
-$\etmisssig$ does better than $\etmiss$ in separating background and signal
+$\metsig$ does better than $\met$ in separating background and signal
 contributions.
-But there is more --- $\etmisssig$ also does a decent job of separating signal
+But there is more --- $\metsig$ also does a decent job of separating signal
 models from each other;
 signals with larger mass differences in their decays to invisibles tend to have
-larger $\etmisssig$.
-Binning regions in $\etmisssig$ therefore also improves sensitivity to
+larger $\metsig$.
+Binning regions in $\metsig$ therefore also improves sensitivity to
 individual parameter points and allows targeted selections on other, locally-relevant,
@@
 \caption{
-Illustration of how $\etmisssig$ beats $\etmiss$ alone in sensitivity to
+Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
 example signal models.
-At large $\etmisssig$, $t\bar t$ and other backgrounds are suppressed, leaving
+At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
 quite pure diboson backgrounds and greater significance measures.
 Signal models with smaller mass splittings tend to have the bulk of their
-events at smaller $\etmiss$.
+events at smaller $\met$.
 \\[0.5em]
@@
 $\nbtag \leq 1$,
-$\etmiss > 100~\eV[G]$, and
-$\etmisssig > 6$.
+$\met > 100~\eV[G]$, and
+$\metsig > 6$.
 \\[0.5em]

### 2022-05-18 main/chapter_2ljets.tex
begin context, author credits

@@
 and post-fit background expectations.
-This figure illustrates the background-only part of a \clown{likelihood}
+This figure illustrates the background-only part of a \heplikelihood
 model which is extended with signals to test alternative models.
@@
 
-\paragraph{My contributions}
+\paragraph{Contributions}
+My contributions the the analysis are listed here.
 \begin{itemize}
@@
 \end{itemize}
+Except where specified, I produced all figures presented in this thesis.
+Many plotting scripts, however, are modified from versions inherited from
+many \atlas\ members past and present to whom I am grateful.
+
+Of the large collaboration involved in this project, a great deal of credit is
+due to the following colleagues, labelled with their primary focus within
+the $\twoljets$ search effort:
+Knut Oddvar Vadla (electroweak),
+Sarah Williams (electroweak),
+Jason Lea Oliver (RJR),
+Abhishek Sharma (RJR),
+Jonathan Long (strong),
+Arka Santra (strong),
+Matt Zhang (strong),
+Yumeng Cao (various),
+Eirik Gramstad (fake/non-prompt leptons),
+Benjamin Henry Hooberman (leadership).
+Cheers.
 
@@
 % previous region design, motivation for this work
+Previous searches studied this two-lepton + jets + $\ptmiss$ selection
+in different or smaller data-sets in \atlas.
+This project serves to update those results with the full LHC Run 2 data-set,
+which contains $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
+$13~\eV[T]$ collision energy.
+
+Increased quantity of data allows more options for this search.
+Together with lessons learned from the previous searches, this is not only
+an update but a chance for improvement.
+
+This paper focusses solely on $\twoljets$ data and analyses them in the three
+distinct RJR, electroweak, and strong parts.
+Each part has precedent from previous searches which were conducted and
+published separately with other work.
+A decision was made for full Run 2 to group the searches by targetted physics,
+rather than by method or targetted physical models.
+
+The remainder of this section describes the specific history and motivations
+for each of the RJR, electroweak and strong sub-projects in this search.
+
+\subsection{RJR}
+Recursive Jigsaw Reconstruction, abbreviated as RJR, is a method for
+interpreting particle data as arising from user-specified decay trees.
+The $\twoljets$ RJR analysis uses two C1N2 decay trees which are displayed in
+Figure~\ref{fig:2ljets_rjr_decay_trees}.
+
+\subsection{Electroweak}
+
+\subsection{Strong}
 
@@
 
-\subsection{b-tagging}
+\subsection{$b$-tagging}
 \label{sec:btagging}

### 2022-05-18 main/chapter_2ljets.tex
add rjr diagram

@@
 \subsection{RJR}
-Recursive Jigsaw Reconstruction, abbreviated as RJR, is a method for
+Recursive Jigsaw Reconstruction (RJR), is a method for
 interpreting particle data as arising from user-specified decay trees.
@@
 Figure~\ref{fig:2ljets_rjr_decay_trees}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.99\textwidth]{figures/2ljets_rjr_trees.pdf}
+\caption{%
+Decay trees for the $\twoljets$ RJR analysis.
+Both diagrams represent processes also targetted in the electroweak analysis.
+(left) A fully-resolved decay tree with its particles labelled.
+(right) Only the $Z\rightarrow\ell\ell$ (\underline{v}ector boson)
+is resolved, but the supersymmetric system recoils off
+hard QCD jets labelled ``ISR'', boosting the \underline{i}nvisible particles,
+to generate visibly large $\met$ in models with small mass-splittings.
+The same approach is used for SR-OffShell selections in the electroweak
+analysis.
+}
+\label{fig:2ljets_contours_gmsb}
+\end{figure}
 

### 2022-05-18 main/chapter_2ljets.tex
expand rjr references

@@
 \subsection{RJR}
-Recursive Jigsaw Reconstruction (RJR), is a method for
-interpreting particle data as arising from user-specified decay trees.
-The $\twoljets$ RJR analysis uses two C1N2 decay trees which are displayed in
-Figure~\ref{fig:2ljets_rjr_decay_trees}.
+The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
+particle data by reconciling them with user-specified decay trees.
+Decay trees used in the $\twoljets$ RJR analysis,
+displayed in Figure~\ref{fig:2ljets_rjr_decay_trees},
+target C1N2-like scenarios similar to those also considered in the $\twoljets$
+electroweak search.
+The RJR algorithm works by making decisions of how to assign four-momenta
+to different `systems' --- nodes in the decay graphs -- and evaluates event
+variables in their respective rest
+frames~\cite{jackson2017sparticles, jackson2017rjr}.
+
+Since the year 2017, RJR event variables have been demonstrated as effective
+for interpreting Supersymmetric particle
+physics~\cite{santoni2018probing}
+and used in numerous \atlas\
+searches~\cite{SUSY-2016-07, SUSY-2016-15, SUSY-2016-16} \TODO{more}.
 
@@
 \centering
-\includegraphics[width=0.99\textwidth]{figures/2ljets_rjr_trees.pdf}
+\includegraphics[width=0.9\textwidth]{figures/2ljets_rjr_trees.pdf}
 \caption{%
@@
 }
-\label{fig:2ljets_contours_gmsb}
+\label{fig:2ljets_rjr_decay_trees}
 \end{figure}

### 2022-05-18 main/chapter_2ljets.tex
expand atlas references and rjr section

--- 2022-05-18 main/chapter_2ljets.tex
@@
 % begin with physics content and results
+My main result in this thesis is displayed in Figure~\ref{fig:2ljets_summary}.
 
-Our main result is the content of Figure~\ref{fig:2ljets_summary}, which shows
-the data in all main regions of this analysis, along with the regions' names,
-and post-fit background expectations.
-This figure illustrates the background-only part of a \heplikelihood
-model which is extended with signals to test alternative models.
+It shows unblinded regions from the electroweak part of
+Reference~\cite{atlas2022searches}, which reports searches
+``for new phenomena in events with two leptons, jets, and missing
+transverse momentum''.
+For brevity, I will name that search $\twoljets$.
+
+This section begins with a splash summary of interpretations of the results in
+Figure~\ref{fig:2ljets_summary},
+before cutting through the gory details later on.
+
+In a large number orthogonal signal regions, where non-standard signals would
+encourage excess data above backgrounds,
+no significant excess is observed above \emph{post-fit} background expectations.
+Indeed there are some significant deficits --- data below the background noise.
+
+Background parts of a \heplikelihood model are illustrated behind data.
+To test non-standard (supersymmetric) models against observation, signal
+contributions are added on top of those backgrounds.
 
@@
 One is an MSSM chargino-neutralino model (C1N2) which acts to produce
-$\chargino{1}\textrm{--}\neutralino{2}$ pairs which decay via $W$ and $Z$
-channels respectively to jets and a lepton pair.
-The other is a GMSB model which decays via Higgs and $Z$ bosons in variable
-branching fractions.
-Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
+$\chargino{1}\textrm{--}\neutralino{2}$ pairs, which decay via
+$W\rightarrow q\bar q$ and
+$Z\rightarrow \ell^\pm \ell^\mp$ respectively..
+The other is a GMSB model which, after a variety of initial states produces
+$\neutralino{1}\textrm{--}\neutralino{1}$ pairs.
+These neutralinos decay,
+via Higgs and $Z$ bosons in variable branching fractions,
+to final states again containing $\ell^\pm \ell^\mp$ and jet-seeding partons.
+Diagrams illustrating these C1N2 and GMSB signal models are shown in
 Figure~\ref{fig:2ljets_signal_diagrams}.
-Some example effects of these signals in signal regions are displayed in
+
+Effects of some of these these signals in example signal regions of the
+$\twoljets$ electroweak analysis are displayed in
 Figure~\ref{fig:2ljets_signal_examples}.
@@
 and tested against data in the $\mathrm{CLs}$ prescription.
-Contours showing which regions where those tests label the signal model
-parameters as excluded excluded are shown for
+Contours showing which parameters of those signal models are labelled as
+excluded are displayed for
 C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
@@
 Grey areas are observed limits from the two-lepton parts
-of~\cite{SUSY-2016-24} and~\cite{SUSY-2013-11}.
+of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
 Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
@@
 
-\paragraph{Contributions}
+\FloatBarrier
+\subsection{Contributions}
 My contributions the the analysis are listed here.
@@
 \FloatBarrier
-\section{Context}
+\section{Origins}
 % previous results on 2(/3)L from ATLAS
@@
 % previous region design, motivation for this work
-Previous searches studied this two-lepton + jets + $\ptmiss$ selection
-in different or smaller data-sets in \atlas.
-This project serves to update those results with the full LHC Run 2 data-set,
-which contains $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
-$13~\eV[T]$ collision energy.
+Previous analyses of \atlas\ data have explored similar selections of
+two leptons, jets, and missing transverse momentum.
+Quite a few analyses, in fact.
+This project exists to supplement those results with information from the full
+LHC Run 2 data-set which was completed in 2018 and contains
+$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at a collision energy
+of $13~\eV[T]$.
+Leverage of the increased data quantity is one key aim.
+Analysis quality also might be improved if our designs can take lessons from
+the experiences of previous work.
 
-Increased quantity of data allows more options for this search.
-Together with lessons learned from the previous searches, this is not only
-an update but a chance for improvement.
+% electroweak
+% Run 1 data~\cite{atlas_2l_SUSY_2013_11},
+% partial Run 2 data~\cite{atlas_23l_SUSY_2016_24},
+% atlas_rjr_23l_SUSY_2017_03,
+% atlas_rjr_mimic_SUSY_2018_06
 
-This paper focusses solely on $\twoljets$ data and analyses them in the three
-distinct RJR, electroweak, and strong parts.
-Each part has precedent from previous searches which were conducted and
-published separately with other work.
-A decision was made for full Run 2 to group the searches by targetted physics,
-rather than by method or targetted physical models.
+The $\twoljets$ analysis is a trinity,
+of which all three parts expand upon \atlas\ results which previously appeared
+in separate publications.
+These three parts are named `RJR', `electroweak', and `strong', each of which
+is described in its historical context in this section.
 
-The remainder of this section describes the specific history and motivations
-for each of the RJR, electroweak and strong sub-projects in this search.
+Recursive Jigsaw Reconstruction (RJR) is an algorithm for constructing useful
+event variables.
+The $\twoljets$ RJR analysis, detailed in Section~\ref{sec:2ljets_origins_rjr},
+uses RJR variables to define its regions in a signal-independent search.
+
+My primary contributions are to the $\twoljets$ electroweak analysis,
+detailed in Section~\ref{sec:2ljets_origins_electroweak},
+which uses conventional (non-RJR) event variables to define a collection of
+orthogonal regions in a search for effects from the electroweak sector of
+supersymmetric models.
+
+Unsurprisingly, the strong part of the $\twoljets$ analysis,
+detailed in Section~\ref{sec:2ljets_origins_strong},
+searches for effects from the strong sector.
+It also uses conventional event variables and with an emphasis on the mass
+spectrum of lepton pairs.
 
 \subsection{RJR}
-Recursive Jigsaw Reconstruction (RJR), is a method for
-interpreting particle data as arising from user-specified decay trees.
-The $\twoljets$ RJR analysis uses two C1N2 decay trees which are displayed in
-Figure~\ref{fig:2ljets_rjr_decay_trees}.
+\label{sec:2ljets_origins_rjr}
 
@@
 analysis.
+\emph{This figure was made by the RJR team.}
 }
-\label{fig:2ljets_contours_gmsb}
+\label{fig:2ljets_rjr_decay_trees}
 \end{figure}
 
+The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
+particle data by reconciling them with user-specified decay trees, and works
+by making decisions of how to assign four-momenta to different `systems',
+which are nodes of the chosen the decay graph.
+Event variables are then evaluated in those systems' rest
+frames~\cite{jackson2017sparticles, jackson2017rjr}.
+
+In this way, RJR variables attain intelligence in how parent particles are
+reconstructed from their visible decay products and approximate independence
+from uninteresting boosts, particularly those from the emission of QCD jets
+outside of the supersymmetric decay tree.
+
+Over the last several years, RJR event variables have been demonstrated as
+effective tools for interpreting supersymmetric particle physics
+data~\cite{santoni2018probing},
+have been used in numerous \atlas\ analyses~\cite{%
+atlas_rjr_SUSY_2016_07,
+atlas_rjr_SUSY_2016_15,
+atlas_rjr_SUSY_2016_16,
+atlas_rjr_23l_SUSY_2017_03
+}, and have influenced the designs of other
+analyses~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+
+Decay trees used in the $\twoljets$ RJR analysis,
+which displayed in Figure~\ref{fig:2ljets_rjr_decay_tres},
+target C1N2-like scenarios similar to those also considered in the $\twoljets$
+electroweak search.
+
 \subsection{Electroweak}
+\label{sec:2ljets_origins_electroweak}
 
 \subsection{Strong}
+\label{sec:2ljets_origins_strong}
 

### 2022-05-18 main/chapter_2ljets.tex
tweaks; add 2ljets references

@@
 \caption{%
-Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
-for which the lower panel shows $S_\mathrm{ATLAS}$ from
+Data of the $\twoljets$ electroweak analysis with \emph{post-fit}
+backgrounds~\cite{atlas2022searches}.
+The lower panel shows $S_\mathrm{ATLAS}$ from
 Equation~\ref{eqn:significance_atlas}.
@@
 \caption{%
-Supersymmetric signal processes in the $\twoljets$ electroweak analysis.
+Supersymmetric signal processes in the $\twoljets$ electroweak
+analysis~\cite{atlas2022searches}.
 \\[0.5em]
@@
 Example signal region histograms with benchmark signal sample yields overlaid
-as dotted lines.
+as dotted lines~\cite{atlas2022searches}.
 Physically, signal yields would add on top of the backgrounds.
@@
 \caption{%
-Contours for the C1N2 model in the $\twoljets$ electroweak analysis.
+Contours for the C1N2 model in the $\twoljets$ electroweak
+analysis~\cite{atlas2022searches}.
 Space below the solid red line is labelled as excluded and its dotted
@@
 \caption{%
-Contours for the GMSB model in the $\twoljets$ electroweak analysis.
+Contours for the GMSB model in the $\twoljets$ electroweak
+analysis~\cite{atlas2022searches}.
 Space below the solid red line is labelled as excluded and its dotted
@@
 \caption{%
-Upper limit results in discovery region.
+Upper limit results in $\twoljets$ electroweak discovery
+regions~\cite{atlas2022searches}.
 The fitted yield is in the background-only model constrained by the data in each region.
@@
 
-Unsurprisingly, the strong part of the $\twoljets$ analysis,
-detailed in Section~\ref{sec:2ljets_origins_strong},
+Unsurprisingly, the strong part of the $\twoljets$ analysis is
+detailed in Section~\ref{sec:2ljets_origins_strong} and
 searches for effects from the strong sector.
-It also uses conventional event variables and with an emphasis on the mass
-spectrum of lepton pairs.
+It also uses conventional event variables, and focuses on features in the
+distribution of the invariant masses of lepton pairs.
 
@@
 \caption{%
-Decay trees for the $\twoljets$ RJR analysis.
+Decay trees for the $\twoljets$ RJR analysis~\cite{atlas2022searches}.
 Both diagrams represent processes also targetted in the electroweak analysis.
 (left) A fully-resolved decay tree with its particles labelled.
-(right) Only the $Z\rightarrow\ell\ell$ (\underline{v}ector boson)
+(right) Only the $Z\rightarrow\ell\ell$ (\underline{V}ector boson)
 is resolved, but the supersymmetric system recoils off
-hard QCD jets labelled ``ISR'', boosting the \underline{i}nvisible particles,
+hard QCD jets labelled ``ISR'', boosting the \underline{I}nvisible particles,
 to generate visibly large $\met$ in models with small mass-splittings.

### 2022-05-18 main/chapter_2ljets.tex
rjr former excesses and conventional comparison

@@
 Physically, signal yields would add on top of the backgrounds.
-Data are shown; regions in the the top two plots observe zero data, and the
+Data are shown; regions in the the top two plots observe zero data, and
 Poisson error bars for $0$ events are hidden for aesthetic reasons.
@@
 analysis~\cite{atlas2022searches}.
+\\[0.5em]
 Space below the solid red line is labelled as excluded and its dotted
@@
 by theoretical error bars.
+\\[0.5em]
 The yellow band shows the $\pm1$-sigma region of exclusion contours
@@
 LHC Run 2 data-set which was completed in 2018 and contains
-$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at a collision energy
-of $13~\eV[T]$.
+$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13~\eV[T]$.
 Leverage of the increased data quantity is one key aim.
@@
 
-In this way, RJR variables attain intelligence in how parent particles are
-reconstructed from their visible decay products and approximate independence
-from uninteresting boosts, particularly those from the emission of QCD jets
-outside of the supersymmetric decay tree.
+In this way, RJR variables get some intelligence in how parent particles are
+reconstructed from their visible decay products, as well as approximate
+independence from uninteresting boosts, particularly those from the emission of
+QCD jets outside of the supersymmetric decay tree.
 
-Over the last several years, RJR event variables have been demonstrated as
-effective tools for interpreting supersymmetric particle physics
+Over the last several years, RJR event variables have been demonstrated to be
+effective for interpreting supersymmetric particle physics
 data~\cite{santoni2018probing},
-have been used in numerous \atlas\ analyses~\cite{%
+and have been used in numerous \atlas\ analyses~\cite{%
 atlas_rjr_SUSY_2016_07,
@@
 atlas_rjr_SUSY_2016_16,
-atlas_rjr_23l_SUSY_2017_03
-}, and have influenced the designs of other
-analyses~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+atlas_rjr_23l_SUSY_2017_03,
+atlas_rjr_3l_SUSY_2019_09
+} \TODO{more}.
+They have also influenced the designs of other analyses which do not use
+RJR variables themselves~\cite{atlas_rjr_mimic_SUSY_2018_06}.
 
-Decay trees used in the $\twoljets$ RJR analysis,
-which displayed in Figure~\ref{fig:2ljets_rjr_decay_tres},
-target C1N2-like scenarios similar to those also considered in the $\twoljets$
-electroweak search.
+Two decay trees are used in the $\twoljets$ RJR analysis, with signal region
+each.
+These trees target C1N2-like scenarios similar to those also considered in the
+$\twoljets$ electroweak search,
+and illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees}.
+
+\paragraph{Former excesses}
+Excess data were observed in several signal regions of the partial Run 2
+search ``for chargino--neutralino production using recursive
+jigsaw reconstruction in final states with two or three charged
+leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
+which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
+$\sqrt{s} = 13~\eV[T]$.
+That search reported notable excesses in four signal regions.
+Of these, two required two charged leptons
+($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
+and other two required three
+($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).
+
+Although the largest local excess was reported as ``$3.0$ standard deviations'',
+which is not particularly large, the pattern of four related regions with
+excess data can understandably draw attention.
+
+Both $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$
+were have been revisited with the full Run 2 data-set and updated background
+modelling.
+The update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
+These regions have also been approximated without direct use of RJR event
+variables;
+these approximations also find data ``in agreement'' with the background
+model~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+
+The $\twoljets$ RJR serves to analysis reproduce
+$\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$,
+with the full Run 2 data-set and updated background modelling,
+and asks whether the excesses persist.
+They do not.
+
+\paragraph{RJR versus conventional}
+Disappearing excesses have encouraged doubts about RJR variables;
+the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06}
+is public evidence of this.
+
+A more precise problem is that that phase space selected by RJR regions tends
+to be poorly (inaccurately or imprecisely) modelled by simulations.
+This is plausibly explained how RJR variables are not simply related to
+conventional, lab-frame variables.
+Simulations have been tuned to model conventional variables well, and
+`systematic' variations also are evaluated and derived in conventional variables.
+Not all functions of the data are well understood.
+
+A lack of reliable simulations motivates reliance on `data-driven' background
+estimates, which, from my perspective, are hard to interpret or trust.
+For these reasons, the $\twoljets$ electroweak analysis uses only conventional
+event variables.
 

### 2022-05-18 main/chapter_2ljets.tex
clear up chapter start

@@
 
+We begin with a quick summary of results from my work on this search project;
+the coming section briefly presents the plots and tables which I personally
+find first when looking at an \atlas\ search paper.
+The the meat of this chapter (or if you prefer, its gory details) follows in
+Section~\ref{sec:2ljets_context} onwards.
+
+\subsection{Splash summary}
+
 % begin with physics content and results
-My main result in this thesis is displayed in Figure~\ref{fig:2ljets_summary}.
+The sum of my work presented in this thesis is displayed in
+Figure~\ref{fig:2ljets_summary}.
+It illustrated the unblinded results in regions of the electroweak part of the
+recent \atlas\ search ``for new phenomena in events with two leptons, jets, and
+missing transverse momentum''~\cite{atlas2022searches}.
+For brevity, I call this search $\twoljets$.
 
-It shows unblinded regions from the electroweak part of
-Reference~\cite{atlas2022searches}, which reports searches
-``for new phenomena in events with two leptons, jets, and missing
-transverse momentum''.
-For brevity, I will name that search $\twoljets$.
-
-This section begins with a splash summary of interpretations of the results in
-Figure~\ref{fig:2ljets_summary},
-before cutting through the gory details later on.
-
-In a large number orthogonal signal regions, where non-standard signals would
-encourage excess data above backgrounds,
-no significant excess is observed above \emph{post-fit} background expectations.
+In a large number orthogonal signal regions, no significant excess is observed
+above \emph{post-fit} backgrounds.
 Indeed there are some significant deficits --- data below the background noise.
@@
 \subsection{Contributions}
-My contributions the the analysis are listed here.
+
+This subsection claims attributions of work on the $\twoljets$ analysis.
+
 \begin{itemize}
@@
 \FloatBarrier
-\section{Origins}
+\section{Context}
+\label{sec:2ljets_context}
 % previous results on 2(/3)L from ATLAS
@@
 The $\twoljets$ analysis is a trinity,
-of which all three parts expand upon \atlas\ results which previously appeared
-in separate publications.
+of which all three parts update previously published \atlas\ results.
 These three parts are named `RJR', `electroweak', and `strong', each of which
@@
 which uses conventional (non-RJR) event variables to define a collection of
-orthogonal regions in a search for effects from the electroweak sector of
-supersymmetric models.
+orthogonal regions, and performs a search for effects from the electroweak
+sectors of supersymmetric models.
 
@@
 
-\paragraph{RJR versus conventional}
+\paragraph{Versus conventional}
 Disappearing excesses have encouraged doubts about RJR variables;

### 2022-05-20 main/chapter_2ljets.tex
add quote sources

@@
 }%
-{Robert M. Pirsig~\cite{pirsig1999zen}}
+{Robert~M.~Pirsig,
+\textit{Zen and the Art of Motorcycle Maintenance},
+1974~\cite{pirsig1999zen}}
 \qitem{%
@@
 }%
-{Richard Feynman~\cite{feynman1974cargo}}
+{Richard~Feynman,
+\textit{Cargo Cult Science},
+1974~\cite{feynman1974cargo}}
 \end{epigraphs}
@@
 
-We begin with a quick summary of results from my work on this search project;
-the coming section briefly presents the plots and tables which I personally
-find first when looking at an \atlas\ search paper.
+My core contribution of this thesis is to the recent \atlas\
+search ``for new phenomena in events with two leptons, jets, and
+missing transverse momentum''~\cite{atlas2022searches}.
+For brevity, I will call this search $\twoljets$.
+
+There are certain plots and tables which, personally, I first seek when
+meeting an \atlas\ search paper.
+This chapter begins by presenting those results in brief in
+Section~\ref{sec:2ljets_splash}.
 The the meat of this chapter (or if you prefer, its gory details) follows in
@@
 \subsection{Splash summary}
+\label{sec:2ljets_splash}
+% begin with physics content and results
 
-% begin with physics content and results
-The sum of my work presented in this thesis is displayed in
-Figure~\ref{fig:2ljets_summary}.
-It illustrated the unblinded results in regions of the electroweak part of the
-recent \atlas\ search ``for new phenomena in events with two leptons, jets, and
-missing transverse momentum''~\cite{atlas2022searches}.
-For brevity, I call this search $\twoljets$.
+Unblinded results in all main regions of the $\twoljets$ electroweak analysis
+are presented in of the Figure~\ref{fig:2ljets_summary}.
+Since this figure appears to represent the data and background estimates,
+on which all derived results depend, it could be seen as the keystone of this
+thesis.
+There are, however, caveats to be developed later in this chapter.
 
@@
 Both $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$
-were have been revisited with the full Run 2 data-set and updated background
-modelling.
-The update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
+have been revisited with the full Run 2 data-set and updated background
+modelling;
+the update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
 These regions have also been approximated without direct use of RJR event
 variables;
-these approximations also find data ``in agreement'' with the background
+those approximations also find data ``in agreement'' with the background
 model~\cite{atlas_rjr_mimic_SUSY_2018_06}.
@@
 
+\clearpage
 \subsection{Cheat-sheet}
@@
 \vspace{0.5em}
-\item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
+\item $\njet$: number of signal jets
 \item $\nbtag$: number of b-tagged jets;
@@
 \vspace{0.5em}
-\item $\dphillmet$: azimuthal angle between the hardest jet and $\ptmiss$
-\item $\Delta\phi(\mathrm{jet}_1,\met)$: azimuthal angle between
-the hardest jet and $\ptmiss$
+\item $\dphillmet$: azimuthal angle between the lepton pair and $\ptmiss$
+\item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
 \end{itemize}
@@
 contributions.
-But there is more --- $\metsig$ also does a decent job of separating signal
-models from each other;
+But there is more --- $\metsig$ also decently separates signal models from one
+another;
 signals with larger mass differences in their decays to invisibles tend to have

### 2022-05-20 main/chapter_experiment.tex
add susy signal plots; various tweaks: todo command, ulem for wavy underline
add quote sources

@@
 }%
-{Ken M}
+{Ken~M,
+\textit{internet comment},
+2012~\cite{kenm2012inner}}
 \end{epigraphs}
@@
 Hello
+
+\section{Inner tracker}
+
+\section{Calorimeters}
+
+\section{Muon system}
+
+\section{Simulation}
+% how do we predict data?
+
+\section{Data}
+% discuss collisions, pileup, met, metsig

### 2022-05-20 main/chapter_searches.tex
add susy signal plots; various tweaks: todo command, ulem for wavy underline
add quote sources

@@
 }%
-{Tony Hoare~\cite{hoare2007emperor}}
+{Tony~Hoare,
+\textit{The Emperor’s Old Clothes},
+1981~\cite{hoare2007emperor}}
 \end{epigraphs}
 \end{singlespacing}
+
+\section{Searches}
+% histograms, CR VR SR, systematics
 
@@
 }%
-{John W. Pratt~\cite{pratt1961testing}}
+{John~W.~Pratt,
+\textit{Review: Testing Statistical Hypotheses},
+1961~\cite{pratt1961testing}}
 \end{epigraphs}
@@
 
-\section{Searches}
-
 \clearpage

### 2022-05-20 main/chapter_theory.tex
add quote sources

@@
 Euclid's postulates came from the Pythagorean theorem, not the other way.}%
-{Richard Hamming~\cite{hamming1980unreasonable}}
+{Richard~Hamming,
+\textit{The Unreasonable Effectiveness of Mathematics},
+1980~\cite{hamming1980unreasonable}}
 \qitem{%
@@
 Geometry is not true, it is advantageous.}%
-{Robert M. Pirsig~\cite{pirsig1999zen}}
+{Robert~M.~Pirsig,
+\textit{Zen and the Art of Motorcycle Maintenance},
+1974~\cite{pirsig1999zen}}
 \end{epigraphs}

### 2022-05-20 main/chapter_2ljets.tex
add other rjr searches

@@
 data~\cite{santoni2018probing},
-and have been used in numerous \atlas\ analyses~\cite{%
+and have been used in numerous \atlas\ searches~\cite{%
 atlas_rjr_SUSY_2016_07,
@@
 atlas_rjr_23l_SUSY_2017_03,
+atlas_rjr_SUSY_2018_12,
+atlas_rjr_EXOT_2019_19,
 atlas_rjr_3l_SUSY_2019_09
-} \TODO{more}.
+}.
 They have also influenced the designs of other analyses which do not use

### 2022-05-20 main/chapter_2ljets.tex
tidy up doubts section

@@
 
-\paragraph{Versus conventional}
-Disappearing excesses have encouraged doubts about RJR variables;
+\paragraph{Doubts}
+Disappearing excesses have encouraged some doubts about RJR variables in
+certain social circles;
 the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06}
@@
 
-A more precise problem is that that phase space selected by RJR regions tends
-to be poorly (inaccurately or imprecisely) modelled by simulations.
-This is plausibly explained how RJR variables are not simply related to
-conventional, lab-frame variables.
+A more precise frame problem with RJR variables is that they select phase space
+which tends to be poorly modelled by simulations.
+This is plausibly explained by the fact that  RJR variables are non-trivially
+related to conventional, lab-frame variables.
 Simulations have been tuned to model conventional variables well, and
-`systematic' variations also are evaluated and derived in conventional variables.
+`systematic' variations are evaluated and derived in conventional variables.
 Not all functions of the data are well understood.
+These effects cause Monte Carlo simulations to give predictions in RJR regions
+which are inaccurate and/or carry large uncertainties.
 
@@
 estimates, which, from my perspective, are hard to interpret or trust.
-For these reasons, the $\twoljets$ electroweak analysis uses only conventional
-event variables.
+For these reasons, the $\twoljets$ electroweak analysis does not use any
+RJR event variables.
 

### 2022-05-20 main/chapter_2ljets.tex
para to subsub

@@
 
-\paragraph{Former excesses}
+\subsubsection{Former excesses}
 Excess data were observed in several signal regions of the partial Run 2
@@
 
-\paragraph{Doubts}
+\subsubsection{Doubts}
 Disappearing excesses have encouraged some doubts about RJR variables in

### 2022-05-21 main/chapter_2ljets.tex
add metsig and mt2 descriptions

@@
 One is an MSSM chargino-neutralino model (C1N2) which acts to produce
-$\chargino{1}\textrm{--}\neutralino{2}$ pairs, which decay via
+$\chargino_1\textrm{--}\neutralino_2$ pairs, which decay via
 $W\rightarrow q\bar q$ and
@@
 The other is a GMSB model which, after a variety of initial states produces
-$\neutralino{1}\textrm{--}\neutralino{1}$ pairs.
+$\neutralino_1\textrm{--}\neutralino_1$ pairs.
 These neutralinos decay,
@@
 \\[0.5em]
-Left: C1N2, where the initial $\chargino{1}\textrm{--}\neutralino{2}$ pair
+Left: C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
 is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
 weak bosons in decays are bounded by the mass splitting
-$m(\chargino{1}, \neutralino{2}) - m(\neutralino{1})$.
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
 We explore the parameters
-$m(\chargino{1}, \neutralino{2})$ and $m(\neutralino{1})$.
+$m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
 \\[0.5em]
-Right: GMSB, where the initial $\neutralino{1}\textrm{--}\neutralino{1}$ pair
-is produced by soft decays from pairs including $\chargino{1}$,
-$\neutralino{2}$ or $\neutralino{1}$.
+Right: GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
+is produced by soft decays from pairs including $\chargino_1$,
+$\neutralino_2$ or $\neutralino_1$.
 Although the $h$ and $Z$ bosons decay to many states, we target
@@
 We explore the parameters
-$m(\neutralino{1})$ and $B(\neutralino{1} \rightarrow h \tilde{G})$ with fixed
-$m(\chargino{1}, \neutralino{2}) - m(\neutralino{1}) = 1~\eV[G]$.
+$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with fixed
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1~\eV[G]$.
 }
@@
 \label{sec:metsig}
-It scales $\met$ by an uncertainty derived from the resolutions of
-reconstructed objects~\cite{atlas_met_significance}.
+
+Most common Standard Model processes at the LHC do not produce hard
+invisible particles.
+Our supersymmetric signals, however, do.
+This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
+important features for separating signal and background events in all parts
+of the $\twoljets$ search.
+
+The experimental resolution of $\ptmiss$, however, is imprecise.
+Importantly also, the uncertainty on any inferred $\ptmiss$ varies
+significantly from one event to another.
+
+A muon, for example, will tend to have its momentum measured more precisely
+than a jet of similar energy.
+Unless, as discussed in Chapter~\ref{chapter:experiment}, the muon is hard
+enough that its curvature through \atlas' magnetic fields is poorly tracked.
+Unless, in the right circumstances, that jet originates from a $b$ quark,
+such that the weak decays of its hadrons tend to lose momentum in
+neutrinos or soft muons. (The $\metsig$ variable used in this analysis is not
+aware of $b$-tagging information, but perhaps it could benefit from some.)
+
+By incorporating an uncertainty in $\ptmiss$ tailored to each event, the
+variable $\metsig$ helps to separate those events with real and well-measured
+$\met$ from those where the measures $\ptmiss$ is more likely to result from
+mismeasurements.
+
+In the context of $\twoljets$ selections, $\metsig$ suppresses $t\bar t$
+contributions more than $\met$ alone, plausibly since $t\bar t$ events are
+accompanied with more hadronic activity than other sources of $\ptmiss$.
+What remains of the standard model in the high-$\metsig$ tail for $\twoljets$
+selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
+and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
+is not reconstructed.
+\TODO{plot of diboson decomposition here?}
+These effects are demonstrated from simulations in
+Figure~\ref{fig:2ljets_presel_met}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
+\hfill
+\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
+\caption{
+Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
+example signal models.
+At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
+quite pure diboson backgrounds and greater significance measures.
+Signal models with smaller mass splittings tend to have the bulk of their
+events at smaller $\met$.
+\\[0.5em]
+This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
+signal leptons with
+$\mll \in (71, 111)~\eV[G]$,
+$\mjj \in (60, 110)~\eV[G]$,
+$\nbtag \leq 1$,
+$\met > 100~\eV[G]$, and
+$\metsig > 6$.
+\\[0.5em]
+``Significance'' displayed in the lower subplot uses the Binomial significance
+$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
+uncertainty and yields taken to the right of a given value.
+Large values indicate expected sensitivity to the signal model.
+}
+\label{fig:2ljets_presel_met}
+\end{figure}
+
+Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix, $S$,
+in the transverse plane, and used to define
+\begin{equation}
+\metsig
+=
+\ptmiss S^{-1} \ptmiss
+,
+\end{equation}
+which can be rearranged into a ratio of $(\met)^2$ to a function of elements
+of the covariance matrix~\cite{atlas_met_significance}.
+This is the square of a Mahalanobis (standard-deviations) distance between
+the origin and $\ptmiss$~\cite{mahalanobis1936generalised}.
+
+Just as $\ptmiss$ is constructed from a (negative) sum of contributions from
+objects in an event,
+$S$ is a sum of covariances assigned to those objects;
+both means and variances add linearly.
 
@@
 \label{sec:mt2}
+% TODO
 % angle between the vector sum of the lepton pair and the missing transverse momentum.
@@
 
+Just as the slepton is a supersymmetric partner to a lepton, the stransverse
+mass  $\mttwo$ is a partner to the `transverse mass' $m_T$
+(to be defined shortly),
+that is useful in searches for supersymmetric particles.
+
+The stransverse mass of an event is a greatest lower bound on the mass of a
+pair-produced resonance, that applies when each partner decays semi-invisibly.
+For example, an on-shell $\tilde\ell^+\tilde\ell^-$ pair decaying as
+$\tilde\ell^+ \rightarrow \ell^+ \neutralino_1
+,~
+\tilde\ell^- \rightarrow \ell^- \neutralino_1$ has
+$\mttwo(\ell^+, \ell^-, m(\neutralino_1)) \leq m(\tilde\ell)$, plus some error
+margin for experimental resolution.
+Such cases are common since, R-parity conservation implies pair production,
+and the idea that dark matter could comprise the lightest supersymmetric
+particle motivates invisible decays.
+
+Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
+standard model processes; dileptonic $t\bar t$ and $W^\pm W^\mp$ decays both
+satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
+The $\twoljets$ electroweak search depends on this effect;
+most selections require $\mttwo(\ell^+, \ell^-, 0) > 80~\eV[G]$, or similar,
+to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
+
+To be precise, the stransverse mass explores how to split the observed
+$\ptmiss$ between the two visible decay products in an event, and can be
+defined as
+\begin{equation}
+\mttwo(a_\mu, b_\mu, m)
+=
+\min_{\vec p\,+\,\vec q\,=\,\ptmiss}
+\max
+\begin{Bmatrix}
+m_T(a_\mu + \{\!\sqrt{m^2 - \vec p_T^{\:2}},\,\vec p\,\}_\mu)\\
+m_T(b_\mu + \{\!\sqrt{m^2 - \vec q_T^{\:2}},\,\vec q\,\}_\mu)\\
+\end{Bmatrix}
+,
+\end{equation}
+in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$~\cite{lester1999measuring}.
+As an extension, one can assign different masses $m_p$ and $m_q$ to the
+the two invisible products~\cite{lester2015bisection}.
+
+Numerical evaluation of $\mttwo$ can be performed efficiently with interval
+bisection algorithms, which seek the least resonance mass for which the
+transverse masses allow any compatible missing momentum
+assignment~\cite{cheng2008minimal, lester2015bisection}.
+Public code for calculating $\mttwo$ is available as a Python
+library~\cite{gillam2021mt2}, to which I contributed a faster and more robust
+implementation of the bisection algorithm by
+Lester and Nachman~\cite{lester2015bisection} during this project.
+
 \subsection{$b$-tagging}
@@
 
-\clearpage
 \subsection{Cheat-sheet}
@@
 
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
-\hfill
-\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
-\caption{
-Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
-example signal models.
-At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
-quite pure diboson backgrounds and greater significance measures.
-Signal models with smaller mass splittings tend to have the bulk of their
-events at smaller $\met$.
-\\[0.5em]
-This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
-signal leptons with
-$\mll \in (71, 111)~\eV[G]$,
-$\mjj \in (60, 110)~\eV[G]$,
-$\nbtag \leq 1$,
-$\met > 100~\eV[G]$, and
-$\metsig > 6$.
-\\[0.5em]
-``Significance'' displayed in the lower subplot uses the Binomial significance
-$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
-uncertainty and yields taken to the right of a given value.
-Large values indicate expected sensitivity to the signal model.
-}
-\label{fig:2ljets_presel_met}
-\end{figure}
-
 % CRs go in with relevant SR category

### 2022-05-21 main/chapter_2ljets.tex
slash after heplikelihood
add todo citations, fiddle with claimed contributions

@@
 
-Background parts of a \heplikelihood model are illustrated behind data.
+Background parts of a \heplikelihood\ model are illustrated behind data.
 To test non-standard (supersymmetric) models against observation, signal
@@
 
-This subsection claims attributions of work on the $\twoljets$ analysis.
+In this subsection, I claims attributions of work on the $\twoljets$ analysis
+and related \atlas\ efforts.
 
@@
 \item Service as `analysis contact' from January to October 2020.
-\item Production of main data inputs for all three analyses
-and the RJR $3\ell$ search \TODO{cite}.
+\item Production of main data inputs for all three $\twoljets$ analyses
+and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
 Production of all systematic variations for the background samples;
@@
 Production of all electroweak signal samples and their systematic variations.
-\item Integration with ATLAS combinations and pMSSM scan efforts by
-performing their validation studies and serializing the analysis results.
-\item Preparation of electroweak results for publication in the paper \TODO{cite} and the
-public HEPData record~\cite{maguire2017hepdata}.
+\item \TODO{?Evaluation of Monte Carlo $Z$+jets uncertainties for RJR signal regions
+which indicated they exceeded $100\%$}
+\item \TODO{Various collaborations with group members on scripts to conduct
+tasks which were common between the searches.}
+\item Integration of the electroweak analysis with ATLAS combinations and
+pMSSM scan efforts, by performing their validation studies and serializing the
+analysis results.
+\item Preparation of electroweak results for publication in the
+paper~\cite{atlas2022searches} and the public HEPData
+record~\cite{maguire2017hepdata}.
 \end{itemize}

### 2022-05-21 main/chapter_2ljets.tex
add draft rjr summary figure

@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_rjr_23l_2l_summary.png}
+    \caption{$2\ell$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_rjr_23l_3l_summary.png}
+    \caption{$3\ell$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.99\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_rjr_summary_log.pdf}
+    \caption{$\twoljets$}
+\end{subfigure}
+\caption{%
+\TODO{summary plots with the previous 2/3l excesses and $\twoljets$.}
+\TODO{reproduced from / made by the RJR team}
+}
+\label{fig:2ljets_signal_examples}
+\end{figure}
+
 \subsubsection{Doubts}

### 2022-05-23 main/chapter_2ljets.tex
write rjr summary caption

@@
     \includegraphics[width=\textwidth]{figures/2ljets_rjr_23l_2l_summary.png}
-    \caption{$2\ell$}
+    \caption{$2\ell$ RJR~\cite{atlas_rjr_23l_SUSY_2017_03}}
 \end{subfigure}
@@
     \includegraphics[width=\textwidth]{figures/2ljets_rjr_23l_3l_summary.png}
-    \caption{$3\ell$}
+    \caption{$3\ell$ RJR~\cite{atlas_rjr_23l_SUSY_2017_03}}
 \end{subfigure}
 \\[0.5em]
-\begin{subfigure}{0.99\textwidth}
+\begin{subfigure}{0.8\textwidth}
     \centering
     \includegraphics[width=\textwidth]{figures/2ljets_rjr_summary_log.pdf}
-    \caption{$\twoljets$}
+    \caption{$\twoljets$ RJR~\cite{atlas2022searches}}
 \end{subfigure}
 \caption{%
-\TODO{summary plots with the previous 2/3l excesses and $\twoljets$.}
-\TODO{reproduced from / made by the RJR team}
+Summary plots for RJR analyses:
+(a) and (b)~two- and three- lepton selections respectively reproduced
+from~\cite{atlas_rjr_23l_SUSY_2017_03},
+(c)~all regions from the $\twoljets$ RJR search~\cite{atlas2022searches}.
+Note the excesses in
+$\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$ of (a),
+and in
+$\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$ of (b).
+\textit{Sub-figure (c) was made by the RJR team.}
 }

### 2022-05-24 main/chapter_2ljets.tex
fix a sentence
move b tagging to where objects will be
first runthough trigger sec

@@
 
-A more precise frame problem with RJR variables is that they select phase space
-which tends to be poorly modelled by simulations.
+A more precisely framed problem with RJR variables is that they tend to select
+phase space which is poorly modelled by simulations.
 This is plausibly explained by the fact that  RJR variables are non-trivially
@@
 % lepton, jet definitions
+
+\subsection{Trigger}
+There would be no event variables if not for the trigger.
+When living our daily lives, we are bombarded with sensory data and must pay
+attention only only those gems few of any use, to avoid exhausting our
+processing capacity.
+Running \atlas\ is no different.
+
+Each trigger in \atlas' arsenal activates on a selection comprising some
+combination of sufficiently hard objects.
+The set of targeted objects includes all those used in this analysis and more:
+electron, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
+decaying tau leptons~\cite{atlas_PERF_2007_01}.
+
+The list of available triggers in \atlas\ is designed to balance data rates
+against physical interest.
+It includes a variety of general purpose selections, which are used for
+$\twoljets$, along with more specialized targets such as vector boson fusion
+or long lived particles.
+The list varies through time, so we select triggers specifically in
+each data-taking period.
+
+The loosest triggers are `prescaled' --- their data rate is reduced by randomly
+accepting only a small fraction of events.
+For $\twoljets$, we use only the `lowest' (loosest, highest rate) un-prescaled
+triggers selecting two charged leptons~\cite{atlas_twiki_lowest_unprescaled}.
+Those charged leptons are selected in the
+high-level trigger (described in Section~\ref{sec:atlas_trigger})
+primarily requiring large $\pt$ and certain quality criteria which vary by
+trigger.
+
+In total, we use $14$ different two-lepton trigger selections, each of which
+selects for an electron pair, a muon pair, or an emu (electron-muon pair).
+To describe those triggers let $\pt^{\ell_i}$ be the transverse momentum of the
+$i\mathrm{th}$ hardest lepton.
+
+In short, all triggers require $\pt^{\ell_i}$ values to exceed
+$8\textrm{--}26~\eV[G]$.
+To detail all trigger requirements and their variations with data period
+would be too tedious and not enlightening, but their $\pt^{\ell_i}$ requirements
+are summarized here.
+\begin{itemize}
+\item Electron pair: requirements vary between
+\begin{itemize}
+\item $\pt^{e_2} > 12~\eV[G]$ and
+\item $\pt^{e_2} > 17~\eV[G]$.
+\end{itemize}
+\item Muon pair: symmetric requirements vary between
+\begin{itemize}
+\item $\pt^{\mu_2} > 10~\eV[G]$ and
+\item $\pt^{\mu_2} > 14~\eV[G]$.
+\end{itemize}
+Asymmetric requirements vary between
+\begin{itemize}
+\item $\pt^{\mu_1} > 18~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$, and
+\item $\pt^{\mu_1} > 22~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$.
+\end{itemize}
+\item Emu pair: various asymmetric requirements include
+\begin{itemize}
+\item $\pt^{e_1} > 17~\eV[G]~\&~\pt^{\mu_1} > 14~\eV[G]$,
+\item $\pt^{e_1} > 7~\eV[G]~\&~\pt^{\mu_1} > 24~\eV[G]$, and
+\item $\pt^{e_1} > 26~\eV[G]~\&~\pt^{\mu_1} > 8~\eV[G]$.
+\end{itemize}
+\end{itemize}
+
+\subsubsection{Alternative triggers}
+\textit{Knut Vadla} \TODO
+
+\subsection{Leptons}
+Charged leptons are essential to the $\twoljets$ analysis.
+
+\subsection{Jets}
+
+\subsubsection{$b$-tagging}
+\label{sec:btagging}
+
+\subsection{$\met$}
 
@@
 \label{sec:mt2}
-% TODO
-% angle between the vector sum of the lepton pair and the missing transverse momentum.
-% leptons and the missing transverse momentum.
-% If two massive particles are pair produced and each decays to a
-% lepton and something massless and invisible,
-% then this estimates a greatest lower bound on the invariant
-% mass of the pair produced particles.
-% $\mttwo(\ell, \ell; 0)$ is useful for excluding standard model backgrounds;
-% di-leptonic ttbar ($tt\rightarrow WW\rightarrow\ell\nu\ell\nu$)
-% and $WW\rightarrow\ell\nu\ell\nu$ have a kinematic endpoint
-% at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.
 
@@
 Lester and Nachman~\cite{lester2015bisection} during this project.
-
-\subsection{$b$-tagging}
-\label{sec:btagging}
-% with $\pt > 20$ GeV,
-% where the tag is taken on the mv2c10 classifier at the $77\%$ working point.
 

### 2022-05-24 main/chapter_2ljets.tex
rename to features, add welcome, improve itemize

@@
 
-\section{Event variables}
+\section{Event features}
 % lepton, jet definitions
+
+From its name, the $\twoljets$ analysis analyses features of events with two
+charged leptons (electrons or muons --- not taus), and at least one jet.
+Though not nominal, the a $\met$ requirement is also implicit from the context
+of $R$-parity conserving supersymmetric models.
+This section serves to describe how events are selected for the $\twoljets$
+analyses, and which event variables are extracted for use in its selections.
+
+Triggers and object definitions are shared between the three $\twoljets$
+searches.
+Higher level event variables and selections will only be described for the
+$\twoljets$ electroweak search.
 
@@
 
-In short, all triggers require $\pt^{\ell_i}$ values to exceed
-$8\textrm{--}26~\eV[G]$.
+In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
+the range $8\textrm{--}26~\eV[G]$.
 To detail all trigger requirements and their variations with data period
@@
 \end{itemize}
-\item Muon pair: symmetric requirements vary between
+\item Muon pair: various requirements include
 \begin{itemize}
-\item $\pt^{\mu_2} > 10~\eV[G]$ and
-\item $\pt^{\mu_2} > 14~\eV[G]$.
-\end{itemize}
-Asymmetric requirements vary between
-\begin{itemize}
+\item $\pt^{\mu_2} > 10~\eV[G]$,
+\item $\pt^{\mu_2} > 14~\eV[G]$,
 \item $\pt^{\mu_1} > 18~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$, and
@@
 \end{itemize}
+Trigger requirements have increased in hardness with time.
 

### 2022-05-24 main/chapter_2ljets.tex
fill in alt triggers

@@
 accepting only a small fraction of events.
-For $\twoljets$, we use only the `lowest' (loosest, highest rate) un-prescaled
+Prescaling adds complication which we avoid;
+for $\twoljets$, we use only the `lowest' (loosest, highest rate) un-prescaled
 triggers selecting two charged leptons~\cite{atlas_twiki_lowest_unprescaled}.
@@
 \subsubsection{Alternative triggers}
-\textit{Knut Vadla} \TODO
+Why should we only use two-lepton triggers?
+Alternatives could be one-lepton triggers, for cases where the other is not
+reconstructed at trigger level, or perhaps also triggers for jets and $\met$.
+The answer is that the benefits of additional triggers were assessed to not
+outweigh the cost of using using them.
+
+The benefit of an additional trigger could be increased signal yields, if that
+trigger do not also harmfully increase inseparable backgrounds.
+A study, \textit{which was performed by Knut Vadla}, compared the efficiencies
+of including one-lepton triggers.
+In a signal-region-like selection requiring two leptons, two jets and
+$\met > 200~\eV[G]$, this study showed that two-lepton triggers already accept
+in excess of $90\%$ of events from benchmark C1N2 models.
+Therefore any additional trigger can therefore have at most a $\approx10\%$;
+improvement, and adding one-lepton triggers was found to increase efficiency by
+only a few percentage points, depending on the model.
+We considered that to be too little.
+
+Costs of triggers are attributable to human effort; adding triggers requires
+software work to include them and their associated Monte Carlo weights, along
+with additional studies to validation of background estimates and evaluation of
+systematic uncertainties.
+Similarly, our implementation of the Matrix Method fake/non-prompt lepton
+background estimate (see Section~\ref{sec:2ljets_mm_fakes}) is not equipped to
+handle one-lepton triggers.
+
+Although jet and $\met$ triggers were not studied in detail, they require
+hard objects, such as $\ptjone \gtrsim 400~\eV[G]$ or
+$\met \gtrsim 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
+Leptons are striking signals which \atlas\ can trigger on with much softer
+requirements; lepton triggers allow us to use softer jets and $\met$,
+which are relied upon throughout the $\twoljets$ analyses.
 
@@
 \subsection{Fake/non-prompt leptons}
+\label{sec:2ljets_mm_fakes}
 

### 2022-05-24 main/chapter_2ljets.tex
fix gtrsim; other tweaks

@@
 in excess of $90\%$ of events from benchmark C1N2 models.
-Therefore any additional trigger can therefore have at most a $\approx10\%$;
-improvement, and adding one-lepton triggers was found to increase efficiency by
+
+Therefore, any additional trigger has at most a $\approx10\%$;
+improvement. Adding one-lepton triggers was found to increase efficiency by
 only a few percentage points, depending on the model.
@@
 
-Costs of triggers are attributable to human effort; adding triggers requires
+Costs of triggers are attributable to human effort;
 software work to include them and their associated Monte Carlo weights, along
-with additional studies to validation of background estimates and evaluation of
-systematic uncertainties.
+with studies to validate background estimates and evaluate systematic
+uncertainties.
 Similarly, our implementation of the Matrix Method fake/non-prompt lepton
@@
 Although jet and $\met$ triggers were not studied in detail, they require
-hard objects, such as $\ptjone \gtrsim 400~\eV[G]$ or
-$\met \gtrsim 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
+hard objects: examples are approximately $\ptjone > 400~\eV[G]$,
+$p_\mathrm{T}^{\,j_3} > 200~\eV[G]$, or
+$\met > 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
 Leptons are striking signals which \atlas\ can trigger on with much softer

### 2022-05-24 main/chapter_2ljets.tex
clarify why no jet met triggers

@@
 \end{itemize}
-Trigger requirements have increased in hardness with time.
+Trigger requirements increase in hardness with time.
 
@@
 $\met > 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
-Leptons are striking signals which \atlas\ can trigger on with much softer
-requirements; lepton triggers allow us to use softer jets and $\met$,
-which are relied upon throughout the $\twoljets$ analyses.
+Leptons, however, are striking signals which are produced as smaller background
+rates, as they suppressed by electroweak couplings.
+Therefore, \atlas\ can trigger on leptons with much softer requirements;
+lepton triggers allow us to use softer jets and $\met$, which are relied upon
+throughout the $\twoljets$ analyses.
 

### 2022-05-24 main/chapter_2ljets.tex
add descriptions of muon and electron id

@@
 Section~\ref{sec:2ljets_splash}.
-The the meat of this chapter (or if you prefer, its gory details) follows in
+The meat of this chapter (or if you prefer, its gory detail) follows in
 Section~\ref{sec:2ljets_context} onwards.
@@
 \end{itemize}
-Trigger requirements increase in hardness with time.
+Quite generally throughout the data used in $\twoljets$, trigger requirements
+have increased in hardness with time.
 
@@
 
-The benefit of an additional trigger could be increased signal yields, if that
-trigger do not also harmfully increase inseparable backgrounds.
-A study, \textit{which was performed by Knut Vadla}, compared the efficiencies
-of including one-lepton triggers.
+The benefit of an additional trigger could be in increased signal yields if
+that trigger do not also harmfully increase inseparable backgrounds.
+A study, \textit{which was performed by Knut Vadla}, investigated the effects
+of of including one-lepton triggers.
 In a signal-region-like selection requiring two leptons, two jets and
@@
 in excess of $90\%$ of events from benchmark C1N2 models.
-
 Therefore, any additional trigger has at most a $\approx10\%$;
@@
 only a few percentage points, depending on the model.
-We considered that to be too little.
+We considered that to be too little to be worthwhile.
 
@@
 \subsection{Leptons}
-Charged leptons are essential to the $\twoljets$ analysis.
+In the context of $\twoljets$, `leptons' means `electron-or-muons';
+this is because taus, which decay before reconstruction, and are not used
+directly in $\twoljets$, and neutral leptons (the neutrinos) go missing.
+
+Reconstructed lepton objects feature momenta, charges, and flavours.
+Each is included only if it satisfies certain kinematic and
+is identified with certain `quality' requirements.
+
+Muons are identified by fitting tracks to hits in the inner tracker and muon
+spectrometer, and their quality requirements use goodness-of-fit metrics of
+those fits along with low level details of detector interactions, to primarily
+to reduce backgrounds from non-prompt muons from hadronic
+decays~\cite{atlas_muon_quality_MUON_2018_03}.
+
+Electron identification combines data from the inner tracker and
+calorimeters. Electron quality criteria use numerous features of detector
+interactions, track fitting metrics, matching of those tracks to calorimeters,
+calorimeter shower shapes, and leakage into the hadronic calorimeter to reduce
+backgrounds from photons, hadrons, and non-prompt electrons from hadronic
+decays~\cite{atlas_egamma_quality_EGAM_2018_01}.
+
+\TODO{signal objects have isolation criteria}
+
+\TODO{lepton pairs SF-OS for Z}
 

### 2022-05-25 main/chapter_2ljets.tex
edit leptons, add isolation paragraph

@@
 Each is included only if it satisfies certain kinematic and
-is identified with certain `quality' requirements.
+is identified with `quality' criteria.
+For both electron and muons, these criteria are standardized with labels
+`loose', `medium', and `tight'.
+These are nested, in the sense that tight implies medium,
+and medium implies loose.
+Other, more specialized, criteria exist, but they are not relevant $\twoljets$.
 
 Muons are identified by fitting tracks to hits in the inner tracker and muon
-spectrometer, and their quality requirements use goodness-of-fit metrics of
-those fits along with low level details of detector interactions, to primarily
-to reduce backgrounds from non-prompt muons from hadronic
-decays~\cite{atlas_muon_quality_MUON_2018_03}.
+spectrometer.
+Their quality criteria use goodness-of-fit metrics from those fits, along with
+low-level details of detector interactions.
+Tighter requirements primarily act to to reduce from non-prompt backgrounds,
+which are muons from hadronic decays~\cite{atlas_muon_quality_MUON_2018_03}.
 
 Electron identification combines data from the inner tracker and
-calorimeters. Electron quality criteria use numerous features of detector
-interactions, track fitting metrics, matching of those tracks to calorimeters,
-calorimeter shower shapes, and leakage into the hadronic calorimeter to reduce
-backgrounds from photons, hadrons, and non-prompt electrons from hadronic
+calorimeters.
+Electron quality criteria use numerous features of detector
+interactions, track fitting metrics, the matching of tracks to calorimeter
+clusters, calorimeter shower shapes, and leakage into the hadronic calorimeter.
+Tighter requirements reduce backgrounds, which include misidentified objects
+--- photons and hadrons --- as well as non-prompt electrons from hadronic
 decays~\cite{atlas_egamma_quality_EGAM_2018_01}.
+
+Isolation of leptons from nearby hadronic activity is another feature which
+can be used to reduce misidentification rates.
+Standard isolation definitions exist for both electrons and muons in a similar
+structure to the quality criteria.
 

### 2022-05-26 main/chapter_2ljets.tex
conclude draft of leptons section

@@
 \subsection{Trigger}
+\label{sec:2ljets_trigger}
 There would be no event variables if not for the trigger.
@@
 
-\TODO{signal objects have isolation criteria}
+Leptons in $\twoljets$ are further sliced into two classes:
+`baseline' and `signal'.
+Baseline leptons and are used in background estimation methods and to ensure
+orthogonality with other \atlas\ analyses, which study events with differing
+lepton counts.
+Baseline requirements are softer and lower quality than signal requirements.
 
-\TODO{lepton pairs SF-OS for Z}
+Any event used in regions of the $\twoljets$ electroweak analysis is required
+to have exactly two baseline leptons and exactly two signal leptons, meaning
+that the event has two reconstructed leptons which pass both sets of criteria.
+This vetoes events with soft or low quality third leptons, ensuring that those
+three-lepton events can be used independently in other studies.
+
+Baseline leptons are required only to have  $\pt > 3~\eV[G]$ for muons and
+$\pt > 4.5~\eV[G]$ for electrons.
+Signal leptons are all required to have $\pt > 10~\eV[G]$, along with tighter
+quality and isolation requirements, details of which can be found in the
+paper~\cite{atlas2022searches}.
+
+All leptons used in regions of the $\twoljets$ electroweak analysis are signal
+leptons with $\pt > 25~\eV[G]$.
+This harder requirement reduces backgrounds, moves our selections away from the
+lowest trigger thresholds, and matches the precedent set by previous
+work~\cite{atlas_23l_SUSY_2016_24}.
 

### 2022-05-26 main/chapter_2ljets.tex
begin jets and btagging

@@
 
+\TODO{reference for trigger details, eta range?}
+
 \subsubsection{Alternative triggers}
@@
 
-Although jet and $\met$ triggers were not studied in detail, they require
-hard objects: examples are approximately $\ptjone > 400~\eV[G]$,
-$p_\mathrm{T}^{\,j_3} > 200~\eV[G]$, or
+Although jet and $\met$ triggers were not studied in detail, they can be
+quickly ruled out since they hard objects: examples of these triggers require
+approximately $\ptjone > 400~\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200~\eV[G]$, or
 $\met > 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
@@
 
+Like leptons, jets are identified with `baseline' and `signal' requirements
+with increasingly tight requirements on quality metrics and kinematics.
+
+\TODO{anti-kt 0.4}
+
+Baseline jets require $\pt > 20~\eV[G]$, and signal jets require
+$\pt > 30~\eV[G]$.
+The $\twoljets$ electroweak analysis uses only signal jets unless those jets
+are $b$-tagged.
+
 \subsubsection{$b$-tagging}
 \label{sec:btagging}
+Tagging adds additional labels to jets, and is particularly useful for
+encoding flavour information derived from data within the jet cluster.
+In particular, $b$-tagging attempts to identify jets originating from $b$
+quarks, which have distinct features due to the long lifetimes and decay
+channels of $B$ hadrons.
+
+\TODO{context: charm quarks, large radius (jumbo) jets top higgs}
 

### 2022-05-26 main/chapter_2ljets.tex
fill in more on jets

@@
 Signal leptons are all required to have $\pt > 10~\eV[G]$, along with tighter
-quality and isolation requirements, details of which can be found in the
+quality and isolation requirements, details of which are given in the
 paper~\cite{atlas2022searches}.
+Electrons are accepted within $|\eta| < 2.47$.
+Baseline muons have $|\eta| < 2.7$, and signal muons have $|\eta| < 2.6$.
 
@@
 
-Like leptons, jets are identified with `baseline' and `signal' requirements
-with increasingly tight requirements on quality metrics and kinematics.
+Like leptons, hadronic jets are identified with `baseline' and `signal'
+requirements with increasingly tight requirements on quality metrics and
+kinematics.
 
-\TODO{anti-kt 0.4}
+Jets in $\twoljets$ use the anti-$k_t$ algorithm with the standard small radius
+parameter of $R=0.4$~\cite{jet_anti_kt}.
+Jets are formed from ``topological'' clusters of energy deposits in the
+calorimeters~\cite{atlas_jet_topo_PERF_2014_07}.
 
-Baseline jets require $\pt > 20~\eV[G]$, and signal jets require
+This topological clustering algorithm is not the latest \atlas\ jet algorithm;
+it has been supplanted in recent years by a `particle flow' method, which
+uses tracking information to improve the reconstruction of charged
+hadrons~\cite{atlas_jet_pflow_PERF_2015_09}.
+We did not switch to these improved jets because the change would have required
+further delays.
+
+Baseline jets are defined require $\pt > 20~\eV[G]$, and signal jets require
 $\pt > 30~\eV[G]$.
+Again, details of quality criteria and angular selections
 The $\twoljets$ electroweak analysis uses only signal jets unless those jets

### 2022-05-26 main/chapter_2ljets.tex
begin serious b-tagging section

@@
 
-My core contribution of this thesis is to the recent \atlas\
+My core contribution in this thesis is to the recent \atlas\
 search ``for new phenomena in events with two leptons, jets, and
@@
 
-Jets in $\twoljets$ use the anti-$k_t$ algorithm with the standard small radius
+Jets in $\twoljets$ use the anti-$k_t$ algorithm with the standard small-radius
 parameter of $R=0.4$~\cite{jet_anti_kt}.
-Jets are formed from ``topological'' clusters of energy deposits in the
+Jets are formed from `topological' clusters of energy deposits in the
 calorimeters~\cite{atlas_jet_topo_PERF_2014_07}.
 
-This topological clustering algorithm is not the latest \atlas\ jet algorithm;
-it has been supplanted in recent years by a `particle flow' method, which
-uses tracking information to improve the reconstruction of charged
+This topological clustering algorithm is no longer the \atlas-recommended jet
+algorithm; it has been supplanted in recent years by a `particle flow' method,
+which uses tracking information to improve the reconstruction of charged
 hadrons~\cite{atlas_jet_pflow_PERF_2015_09}.
-We did not switch to these improved jets because the change would have required
-further delays.
+We did not attempt to switch to these improved jets because the change in
+policy occurred during the years this analysis was in development, and the
+change would have required more work and even more delays.
 
-Baseline jets are defined require $\pt > 20~\eV[G]$, and signal jets require
-$\pt > 30~\eV[G]$.
-Again, details of quality criteria and angular selections
-The $\twoljets$ electroweak analysis uses only signal jets unless those jets
-are $b$-tagged.
+Baseline jets are defined to require $\pt > 20~\eV[G]$ with $|\eta| < 2.8$;
+signal jets additionally require $\pt > 30~\eV[G]$.
+To reduce the impact of pile-up jets, a variety of quality criteria are used,
+depending on the jet $\pt$ and $|\eta|$.
+Details of these quality criteria and angular selections are given in the
+paper~\cite{atlas2022searches}.
 
@@
 \label{sec:btagging}
-Tagging adds additional labels to jets, and is particularly useful for
-encoding flavour information derived from data within the jet cluster.
+
+Tagging adds labels to objects.
+For jets in $\twoljets$, $b$-tagging is used to identify jets which are
+more likely to have originated from $b$ quarks;
+these tags serve to improve our separation of $t\bar t$ backgrounds and
+of signals producing $Z/h \rightarrow b\bar b$ decays.
+
+Signal $b$-tagged jets require $\pt > 20~\eV[G]$, $|\eta| < 2.5$ and a tag from
+the `MV2c10' algorithm at its $77\%$ efficiency working point.
+\TODO{cite, summarize}
+
+encode information derived from data within the jet cluster.
 In particular, $b$-tagging attempts to identify jets originating from $b$

### 2022-05-26 main/chapter_2ljets.tex
\,\eV over ~\eV; minor figure caption improvement

@@
 $m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with fixed
-$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1~\eV[G]$.
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1\,\eV[G]$.
 }
@@
 LHC Run 2 data-set which was completed in 2018 and contains
-$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13~\eV[T]$.
+$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
 Leverage of the increased data quantity is one key aim.
@@
 which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
-$\sqrt{s} = 13~\eV[T]$.
+$\sqrt{s} = 13\,\eV[T]$.
 That search reported notable excesses in four signal regions.
@@
 In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
-the range $8\textrm{--}26~\eV[G]$.
+the range $8\textrm{--}26\,\eV[G]$.
 To detail all trigger requirements and their variations with data period
@@
 \begin{itemize}
-\item $\pt^{e_2} > 12~\eV[G]$ and
-\item $\pt^{e_2} > 17~\eV[G]$.
+\item $\pt^{e_2} > 12\,\eV[G]$ and
+\item $\pt^{e_2} > 17\,\eV[G]$.
 \end{itemize}
@@
 \begin{itemize}
-\item $\pt^{\mu_2} > 10~\eV[G]$,
-\item $\pt^{\mu_2} > 14~\eV[G]$,
-\item $\pt^{\mu_1} > 18~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$, and
-\item $\pt^{\mu_1} > 22~\eV[G]~\&~\pt^{\mu_2} > 8~\eV[G]$.
+\item $\pt^{\mu_2} > 10\,\eV[G]$,
+\item $\pt^{\mu_2} > 14\,\eV[G]$,
+\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
+\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
 \end{itemize}
@@
 \begin{itemize}
-\item $\pt^{e_1} > 17~\eV[G]~\&~\pt^{\mu_1} > 14~\eV[G]$,
-\item $\pt^{e_1} > 7~\eV[G]~\&~\pt^{\mu_1} > 24~\eV[G]$, and
-\item $\pt^{e_1} > 26~\eV[G]~\&~\pt^{\mu_1} > 8~\eV[G]$.
+\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
+\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
+\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
 \end{itemize}
@@
 In a signal-region-like selection requiring two leptons, two jets and
-$\met > 200~\eV[G]$, this study showed that two-lepton triggers already accept
+$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
 in excess of $90\%$ of events from benchmark C1N2 models.
@@
 quickly ruled out since they hard objects: examples of these triggers require
-approximately $\ptjone > 400~\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200~\eV[G]$, or
-$\met > 100~\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
+approximately $\ptjone > 400\,\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200\,\eV[G]$, or
+$\met > 100\,\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
 Leptons, however, are striking signals which are produced as smaller background
@@
 
-Baseline leptons are required only to have  $\pt > 3~\eV[G]$ for muons and
-$\pt > 4.5~\eV[G]$ for electrons.
-Signal leptons are all required to have $\pt > 10~\eV[G]$, along with tighter
+Baseline leptons are required only to have  $\pt > 3\,\eV[G]$ for muons and
+$\pt > 4.5\,\eV[G]$ for electrons.
+Signal leptons are all required to have $\pt > 10\,\eV[G]$, along with tighter
 quality and isolation requirements, details of which are given in the
@@
 All leptons used in regions of the $\twoljets$ electroweak analysis are signal
-leptons with $\pt > 25~\eV[G]$.
+leptons with $\pt > 25\,\eV[G]$.
 This harder requirement reduces backgrounds, moves our selections away from the
@@
 
-Baseline jets are defined to require $\pt > 20~\eV[G]$ with $|\eta| < 2.8$;
-signal jets additionally require $\pt > 30~\eV[G]$.
+Baseline jets are defined to require $\pt > 20\,\eV[G]$ with $|\eta| < 2.8$;
+signal jets additionally require $\pt > 30\,\eV[G]$.
 To reduce the impact of pile-up jets, a variety of quality criteria are used,
@@
 
-Signal $b$-tagged jets require $\pt > 20~\eV[G]$, $|\eta| < 2.5$ and a tag from
+Signal $b$-tagged jets require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a tag from
 the `MV2c10' algorithm at its $77\%$ efficiency working point.
@@
 \\[0.5em]
-This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
-signal leptons with
-$\mll \in (71, 111)~\eV[G]$,
-$\mjj \in (60, 110)~\eV[G]$,
+This selection requires two same-flavor opposite-sign signal leptons with
+$\pt^{\ell_2} > 25\,\eV[G]$,
+$\mll \in (71, 111)\,\eV[G]$,
+$\mjj \in (60, 110)\,\eV[G]$,
 $\nbtag \leq 1$,
-$\met > 100~\eV[G]$, and
+$\met > 100\,\eV[G]$, and
 $\metsig > 6$.
@@
 The $\twoljets$ electroweak search depends on this effect;
-most selections require $\mttwo(\ell^+, \ell^-, 0) > 80~\eV[G]$, or similar,
+most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
 to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.

### 2022-05-27 main/chapter_2ljets.tex
clarify mt2

@@
 Tagging adds labels to objects.
+Unlike how electrons and muons are very cleanly separated however, jet tagging
+is a more noisy affair.
+
 For jets in $\twoljets$, $b$-tagging is used to identify jets which are
 more likely to have originated from $b$ quarks;
-these tags serve to improve our separation of $t\bar t$ backgrounds and
-of signals producing $Z/h \rightarrow b\bar b$ decays.
+these tags serve to improve our separation of $t\bar t$ backgrounds
+(since their weak decays almost always produce $b$ quarks), and
+of signals producing $Z/h \rightarrow b\bar b$ decays in the GMSB model.
 
@@
 \TODO{context: charm quarks, large radius (jumbo) jets top higgs}
+
+\subsection{Overlap removal}
 
@@
 =
-\min_{\vec p\,+\,\vec q\,=\,\ptmiss}
+\min_{\vec p + \vec q=\ptmiss}
 \max
@@
 bisection algorithms, which seek the least resonance mass for which the
-transverse masses allow any compatible missing momentum
-assignment~\cite{cheng2008minimal, lester2015bisection}.
-Public code for calculating $\mttwo$ is available as a Python
+transverse masses allow any compatible assignment such that
+$\vec p + \vec q = \ptmiss$.
+Testing this compatibility reduces to testing for the intersection of two
+ellipses, one for each transverse
+mass~\cite{cheng2008minimal, lester2015bisection}.
+
+Public code for calculating $\mttwo$ is available in a Python
 library~\cite{gillam2021mt2}, to which I contributed a faster and more robust

### 2022-05-27 main/chapter_2ljets.tex
S to \Omega, include metsig ratio form

@@
 the `MV2c10' algorithm at its $77\%$ efficiency working point.
+This tagging algorithm uses
+\cite{atlas_jet_mv2c10_2017, atlas_jet_tagging_2019}
 \TODO{cite, summarize}
+%
+boosted decision tree~\cite{friedman2002stochastic}
 
@@
 
-Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix, $S$,
-in the transverse plane, and used to define
+Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix
+$\Omega$ in the transverse plane, and used to define
 \begin{equation}
@@
 =
-\ptmiss S^{-1} \ptmiss
+\ptmiss \Omega^{-1} \ptmiss
+=
+\frac{(\met)^2}{\sigma_L^2(1 - \rho_{LT}^2)}
 ,
 \end{equation}
-which can be rearranged into a ratio of $(\met)^2$ to a function of elements
-of the covariance matrix~\cite{atlas_met_significance}.
+in which $\sigma_L^2$ is the variance in the direction of $\ptmiss$ and
+$\rho_{LT}$ is the correlation coefficient between the parallel and
+perpendicular components~\cite{atlas_met_significance}.
 This is the square of a Mahalanobis (standard-deviations) distance between
@@
 objects in an event,
-$S$ is a sum of covariances assigned to those objects;
+$\Omega$ is a sum of covariances assigned to those objects;
 both means and variances add linearly.

### 2022-05-27 main/chapter_2ljets.tex
cite mv2c10

@@
 Signal $b$-tagged jets require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a tag from
-the `MV2c10' algorithm at its $77\%$ efficiency working point.
+the `MV2c10' boosted decision tree
+algorithm~\cite{
+friedman2002stochastic,
+atlas_jet_mv2c10_2017,
+atlas_jet_tagging_2019
+} at its $77\%$ efficiency working point, meaning that a true $b$-jet sampled
+from the distribution of its $t\bar t$ testing set has a $77\%$ probability of
+being tagged.
+The `c10' suffix means that the its training data include a $7\%$ fraction of
+$c$-jets~\cite{atlas_twiki_mv2}.
+
 This tagging algorithm uses
-\cite{atlas_jet_mv2c10_2017, atlas_jet_tagging_2019}
-\TODO{cite, summarize}
-%
-boosted decision tree~\cite{friedman2002stochastic}
+\TODO{summarize}
 

### 2022-05-27 main/chapter_2ljets.tex
flesh out btagging

--- 2022-05-27 main/chapter_2ljets.tex
@@
 Likelihoods from validation regions are not included in the fit.
-The `Top' category contains $t\bar t$ and $tW$ processes, and
-`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
+The `Top' category contains $\ttbar$ and $tW$ processes, and
+`Other' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and other
 rare top processes.%
@@
 is a more noisy affair.
+Taggers are available for various jet properties, such charm flavor ($c$-jet),
+bottom flavor ($b$-jet).
+Of these, $\twoljets$ uses only $b$-tagging.
 
-For jets in $\twoljets$, $b$-tagging is used to identify jets which are
-more likely to have originated from $b$ quarks;
-these tags serve to improve our separation of $t\bar t$ backgrounds
-(since their weak decays almost always produce $b$ quarks), and
-of signals producing $Z/h \rightarrow b\bar b$ decays in the GMSB model.
+Larger radius jets can also use taggers for heavy decays, such as those from
+$W$, Higgs or top resonance, but large radius were considered too much
+additional work to considered in $\twoljets$.
 
-Signal $b$-tagged jets require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a tag from
-the `MV2c10' algorithm at its $77\%$ efficiency working point.
-This tagging algorithm uses
-\cite{atlas_jet_mv2c10_2017, atlas_jet_tagging_2019}
-\TODO{cite, summarize}
-%
-boosted decision tree~\cite{friedman2002stochastic}
+Jets originating from $b$ quarks are distinct due to the long lifetimes and
+long decay trees of $B$-hadrons, both of which leave signals in \atlas'
+inner trackers.
 
-encode information derived from data within the jet cluster.
-In particular, $b$-tagging attempts to identify jets originating from $b$
-quarks, which have distinct features due to the long lifetimes and decay
-channels of $B$ hadrons.
+For jets in $\twoljets$, $b$-tagging is useful to separate events due to
+$\ttbar$ backgrounds (since their weak decays almost always produce $b$
+quarks), and to select for signals producing $Z/h \rightarrow b\bar b$ decays
+in the GMSB model.
 
-\TODO{context: charm quarks, large radius (jumbo) jets top higgs}
+Signal $b$-jets in $\twoljets$ require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a
+tag from the `MV2c10' boosted decision tree
+algorithm~\cite{
+friedman2002stochastic,
+atlas_jet_mv2c10_2017,
+atlas_jet_tagging_2019
+} at its $77\%$ efficiency working point.
+This efficiency means that a true $b$-jet sampled from its testing set has a
+$77\%$ probability of being tagged.
+It is trained on a mixture of $\ttbar$ and $\zjets$ simulated data,
+and the `c10' suffix means that those data include a $7\%$ fraction of
+$c$-jets~\cite{atlas_twiki_mv2}.
+
+This tagging algorithm combines various lower-level features, including the
+basic kinematics $\pt$ and $|\eta|$, and more intricate features of
+observable secondary vertices including reconstructed hadronic decay
+trees~\cite{atlas_jet_mv2c10_2017}.
+
+Alternative working points of $70\%$ and $85\%$ efficiency are used in similar
+\atlas\ searches.
+These were trialled and found to make negligible differences to $\twoljets$.
+
+An alternative $b$-tagging algorithm named `DL1' is now recommended by \atlas,
+but its performance is ``similar'' to MV2, and the two tag
+``highly correlated'' samples~\cite{atlas_jet_tagging_2019}, so we did not
+attempt to switch.
 
@@
 
-In the context of $\twoljets$ selections, $\metsig$ suppresses $t\bar t$
-contributions more than $\met$ alone, plausibly since $t\bar t$ events are
+In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
+contributions more than $\met$ alone, plausibly since $\ttbar$ events are
 accompanied with more hadronic activity than other sources of $\ptmiss$.
@@
 example signal models.
-At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
+At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
 quite pure diboson backgrounds and greater significance measures.
@@
 Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
-standard model processes; dileptonic $t\bar t$ and $W^\pm W^\mp$ decays both
+standard model processes; dileptonic $\ttbar$ and $W^\pm W^\mp$ decays both
 satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
@@
 most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
-to reduce $t\bar t$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
+to reduce $\ttbar$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
 

### 2022-05-27 main/chapter_2ljets.tex
add trigger listing point

@@
 \end{itemize}
+Please do not read these all in detail.
+The point of listing them is that trigger activations are not trivial,
+and motivate lepton selections which exceed these trigger thresholds
+to avoid trigger activation effects.
+
 Quite generally throughout the data used in $\twoljets$, trigger requirements

### 2022-05-27 main/chapter_2ljets.tex
label and ref rjr summaries; minor cleanup

@@
 
+Summary plots for $\twoljets$-RJR and the former 2 and 3 lepton analysis,
+including these signal regions with former excesses,
+are displayed in Figure~\ref{fig:2ljets_rjr_summaries}.
+
 \begin{figure}[tp]
@@
 }
-\label{fig:2ljets_signal_examples}
+\label{fig:2ljets_rjr_summaries}
 \end{figure}
@@
 are summarized here.
+Please do not read them all in detail.
 \begin{itemize}
@@
 \end{itemize}
-Please do not read these all in detail.
-The point of listing them is that trigger activations are not trivial,
-and motivate lepton selections which exceed these trigger thresholds
-to avoid trigger activation effects.
+The point of listing these is that demonstrate trigger activations are
+non-trivial.
+This motivates lepton selections which exceed these trigger thresholds,
+so as to avoid the need to model their jagged activations.
 
@@
 \subsection{Jets}
-
 Like leptons, hadronic jets are identified with `baseline' and `signal'
@@
 \label{sec:btagging}
-
 Tagging adds labels to objects.
@@
 \label{sec:mt2}
-
 Just as the slepton is a supersymmetric partner to a lepton, the stransverse
@@
 \section{Design}
-
 The most impactful decision in a search is the design of its regions.
@@
 \section{Results}
+
+\section{Unblinding}

### 2022-05-27 main/chapter_2ljets.tex
add overlap removal

@@
 
-\section{Event features}
-% lepton, jet definitions
-
+\section{Events and their features}
 From its name, the $\twoljets$ analysis analyses features of events with two
@@
 \end{itemize}
-The point of listing these is that demonstrate trigger activations are
+The point of listing these is to demonstrate trigger activations are
 non-trivial.
@@
 \subsection{Overlap removal}
-
-\subsection{$\met$}
-
-\subsection{$\met$ significance}
+Leptons and jets are independently reconstructed.
+This means that, despite the various quality and isolation requirement, there
+can remain examples where detector responses can be attributed to more than
+one object.
+For example, a track could be attributed to both a muon and an electron,
+or a calorimeter energy cluster could be attributed to both an electron and a
+photon. (Although we do not directly use photons in $\twoljets$, we do not veto
+events containing them --- photons contribute to $\ptmiss$.)
+
+Now we know about both leptons and hadronic jets, we can also improve our
+rejection of non-prompt leptons from hadron decays, since leptons near to jets
+become suspect.
+
+To remove duplicates and further reduce non-prompt lepton contributions, a
+standard overlap removal procedure is applied.
+This procedure, which is detailed in the paper~\cite{atlas2022searches},
+comprises a sequence of rules which modify the collection of event objects.
+Some rules decide which to keep from an overlapping pair,
+and others remove leptons overlapping with jets, based on their $\Delta R$
+and $\pt$.
+
+\subsection{$\ptmiss$}
+
+\subsubsection{$\met$ significance}
 \label{sec:metsig}

### 2022-05-27 main/chapter_2ljets.tex
fill in ptmiss; add sec prefixes

@@
 \subsubsection{$b$-tagging}
-\label{sec:btagging}
+\label{sec:2ljets_btagging}
 Tagging adds labels to objects.
@@
 \subsection{Overlap removal}
+\label{sec:2ljets_overlap_removal}
 Leptons and jets are independently reconstructed.
@@
 
-To remove duplicates and further reduce non-prompt lepton contributions, a
-standard overlap removal procedure is applied.
+To remove duplicates and further reduce non-prompt lepton contributions, we
+apply a standard overlap removal procedure.
 This procedure, which is detailed in the paper~\cite{atlas2022searches},
@@
 \subsection{$\ptmiss$}
-
-\subsubsection{$\met$ significance}
-\label{sec:metsig}
-
+Every event has $\ptmiss$.
+Unlike leptons, jets, and other objects which we must select for, the $\ptmiss$
+of an event is a natural result of the conservation of momentum, which \atlas'
+barrel-shaped construction allows us to measure with reasonable precision.
+
+The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
+Please note that $\met$ is not an energy.
+Transverse energy does not exist.
+One may of course define their own directed `transverse energy', but it will
+not be conserved (in the standard model) unless it is equivalent to a conserved
+momentum.
+Indeed, momentum-balanced events do have energy imbalances;
+for example, in a single-top event $pp \rightarrow tq$, the $t$ and $q$ are
+produced back-to-back with equal momenta $p$, but the two particles have very
+different energies since they have very different masses:
+$E_t^2 = m_t^2 + p^2$, $E_q^2 = m_q^2 + p^2$, and $m_t \gg m_q$.
+% hobbyhorse
+
+The $\ptmiss$ vector is defined such that the vector sum of $\ptmiss$ with
+all reconstructed objects (after overlap removal) has zero transverse momentum.
+In addition to the leptons and jets used directly in $\twoljets$, this includes
+photons, tau leptons, and a `soft term' comprising tracks not associated to
+any other objects~\cite{atlas_met}.
+
+Since signal models in $\twoljets$ do generate hard, invisible particles,
+by selecting for large $\met$ we select events in which either there were real
+invisible particles or something was badly mismeasured.
+The standard selection for $\twoljets$-electroweak is $\met > 100\,\eV[G]$.
+This is a nice round number, and corresponds roughly to where the $\met$-less
+background $\zjets$ tails off.
+This $100\,\eV[G]$ also corresponds roughly to several standard deviations of
+a typical $\met$ noise scale of $\approx 10--30\,\eV[G]$, which I observe
+by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections;
+$\metsig$ is defined in the next section.
+
+\subsection{$\met$ significance}
+\label{sec:2ljets_metsig}
 Most common Standard Model processes at the LHC do not produce hard
@@
 \begin{equation}
-\metsig
+\metsig^2
 =
@@
 \subsection{Stranverse mass $\mttwo$}
-\label{sec:mt2}
+\label{sec:2ljets_mt2}
 Just as the slepton is a supersymmetric partner to a lepton, the stransverse
@@
 \item $\metsig$: object-based missing transverse momentum significance;\\
-described in Section~\ref{sec:metsig}
+described in Section~\ref{sec:2ljets_metsig}
 \item $\ptjone$: transverse momentum of the hardest jet%
@@
 \item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
-described in Section~\ref{sec:mt2}
+described in Section~\ref{sec:2ljets_mt2}
 \vspace{0.5em}
@@
 \item $\nbtag$: number of b-tagged jets;
-described in Section~\ref{sec:btagging}
+described in Section~\ref{sec:2ljets_btagging}
 \vspace{0.5em}

### 2022-05-27 main/chapter_2ljets.tex
hyphenate sub searches

@@
 
-Unblinded results in all main regions of the $\twoljets$ electroweak analysis
+Unblinded results in all main regions of the $\twoljets$-electroweak analysis
 are presented in of the Figure~\ref{fig:2ljets_summary}.
@@
 
-We use two signal models in the $\twoljets$ electroweak analysis.
+We use two signal models in the $\twoljets$-electroweak analysis.
 One is an MSSM chargino-neutralino model (C1N2) which acts to produce
@@
 Effects of some of these these signals in example signal regions of the
-$\twoljets$ electroweak analysis are displayed in
+$\twoljets$-electroweak analysis are displayed in
 Figure~\ref{fig:2ljets_signal_examples}.
@@
 \caption{%
-Data of the $\twoljets$ electroweak analysis with \emph{post-fit}
+Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
 backgrounds~\cite{atlas2022searches}.
@@
 \caption{%
-Supersymmetric signal processes in the $\twoljets$ electroweak
+Supersymmetric signal processes in the $\twoljets$-electroweak
 analysis~\cite{atlas2022searches}.
@@
 \caption{%
-Contours for the C1N2 model in the $\twoljets$ electroweak
+Contours for the C1N2 model in the $\twoljets$-electroweak
 analysis~\cite{atlas2022searches}.
@@
 \caption{%
-Contours for the GMSB model in the $\twoljets$ electroweak
+Contours for the GMSB model in the $\twoljets$-electroweak
 analysis~\cite{atlas2022searches}.
@@
 \caption{%
-Upper limit results in $\twoljets$ electroweak discovery
+Upper limit results in $\twoljets$-electroweak discovery
 regions~\cite{atlas2022searches}.
@@
 event variables.
-The $\twoljets$ RJR analysis, detailed in Section~\ref{sec:2ljets_origins_rjr},
+The $\twoljets$-RJR analysis, detailed in Section~\ref{sec:2ljets_origins_rjr},
 uses RJR variables to define its regions in a signal-independent search.
 
-My primary contributions are to the $\twoljets$ electroweak analysis,
+My primary contributions are to the $\twoljets$-electroweak analysis,
 detailed in Section~\ref{sec:2ljets_origins_electroweak},
@@
 \caption{%
-Decay trees for the $\twoljets$ RJR analysis~\cite{atlas2022searches}.
+Decay trees for the $\twoljets$-RJR analysis~\cite{atlas2022searches}.
 Both diagrams represent processes also targetted in the electroweak analysis.
@@
 
-Two decay trees are used in the $\twoljets$ RJR analysis, with signal region
+Two decay trees are used in the $\twoljets$-RJR analysis, with signal region
 each.
 These trees target C1N2-like scenarios similar to those also considered in the
-$\twoljets$ electroweak search,
+$\twoljets$-electroweak search,
 and illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees}.
@@
 
-The $\twoljets$ RJR serves to analysis reproduce
+The $\twoljets$-RJR serves to analysis reproduce
 $\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$,
@@
     \includegraphics[width=\textwidth]{figures/2ljets_rjr_summary_log.pdf}
-    \caption{$\twoljets$ RJR~\cite{atlas2022searches}}
+    \caption{$\twoljets$-RJR~\cite{atlas2022searches}}
 \end{subfigure}
@@
 from~\cite{atlas_rjr_23l_SUSY_2017_03},
-(c)~all regions from the $\twoljets$ RJR search~\cite{atlas2022searches}.
+(c)~all regions from the $\twoljets$-RJR search~\cite{atlas2022searches}.
 Note the excesses in
@@
 estimates, which, from my perspective, are hard to interpret or trust.
-For these reasons, the $\twoljets$ electroweak analysis does not use any
+For these reasons, the $\twoljets$-electroweak analysis does not use any
 RJR event variables.
@@
 Higher level event variables and selections will only be described for the
-$\twoljets$ electroweak search.
+$\twoljets$-electroweak search.
 
@@
 
-Any event used in regions of the $\twoljets$ electroweak analysis is required
+Any event used in regions of the $\twoljets$-electroweak analysis is required
 to have exactly two baseline leptons and exactly two signal leptons, meaning
@@
 
-All leptons used in regions of the $\twoljets$ electroweak analysis are signal
+All leptons used in regions of the $\twoljets$-electroweak analysis are signal
 leptons with $\pt > 25\,\eV[G]$.
@@
 This $100\,\eV[G]$ also corresponds roughly to several standard deviations of
-a typical $\met$ noise scale of $\approx 10--30\,\eV[G]$, which I observe
-by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections;
-$\metsig$ is defined in the next section.
+a typical $\met$ noise scale of $\approx 10\textrm{--}30\,\eV[G]$, which I observe
+by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections,
+and where $\metsig$ is defined in the next section.
 
@@
 \label{sec:2ljets_mt2}
-Just as the slepton is a supersymmetric partner to a lepton, the stransverse
+Just as the slepton is a supersymmetric partner to a lepton, the `stransverse'
 mass  $\mttwo$ is a partner to the `transverse mass' $m_T$
@@
 satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
-The $\twoljets$ electroweak search depends on this effect;
+The $\twoljets$-electroweak search depends on this effect;
 most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,

### 2022-05-27 main/chapter_2ljets.tex
newpage cheatsheet; clarify mt2; clarify rjr

@@
 `systematic' variations are evaluated and derived in conventional variables.
-Not all functions of the data are well understood.
+Modelling these features accurately does not imply that all other functions of
+the data are well understood.
 These effects cause Monte Carlo simulations to give predictions in RJR regions
@@
 A lack of reliable simulations motivates reliance on `data-driven' background
-estimates, which, from my perspective, are hard to interpret or trust.
-For these reasons, the $\twoljets$-electroweak analysis does not use any
-RJR event variables.
+estimates in RJR selections, which, from my perspective, are hard to interpret
+or trust.
+Indeed I argue that all `data-driven' methods rely not only on data but on
+hypothetical extrapolations which may not be well motivated, particularly when
+studying complicated features such as RJR variables.
+For these reasons, I choose to not use any RJR event variables in the the
+$\twoljets$-electroweak analysis.
 
@@
 \end{equation}
-in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$~\cite{lester1999measuring}.
+in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$ and $\{E,\,\vec p\,\}_\mu$ denotes
+the four-vector with energy $E$ and momentum $\vec p$~\cite{lester1999measuring}.
 As an extension, one can assign different masses $m_p$ and $m_q$ to the

### 2022-05-27 main/chapter_2ljets.tex
change tq to Zg, remove -and their features

@@
 
-\section{Events and their features}
+\section{Events}
 From its name, the $\twoljets$ analysis analyses features of events with two
@@
 Indeed, momentum-balanced events do have energy imbalances;
-for example, in a single-top event $pp \rightarrow tq$, the $t$ and $q$ are
-produced back-to-back with equal momenta $p$, but the two particles have very
-different energies since they have very different masses:
-$E_t^2 = m_t^2 + p^2$, $E_q^2 = m_q^2 + p^2$, and $m_t \gg m_q$.
+for example, in a $Z\mathrm{+jet}$ event $pp \rightarrow Zg$, the $Z$ and $g$
+are produced back-to-back with equal momenta $p$, but the two particles have
+very different energies since they have very different masses:
+$E_Z^2 = m_Z^2 + p^2$, $E_g^2 = 0 + p^2$.
 % hobbyhorse

### 2022-05-27 main/chapter_2ljets.tex
add citations for egamma and muon triggers

@@
 \label{sec:2ljets_origins_electroweak}
+% susy motivation
+% 2/3l, older 2l context
+% update with run 2
+% multi-bin fit
+% orthogonal regions
 
@@
 \label{sec:2ljets_origins_strong}
+% previous papers
+% three signal models
+% mll, kinematic end point
 
@@
 This motivates lepton selections which exceed these trigger thresholds,
-so as to avoid the need to model their jagged activations.
+so as to avoid the need to model their jagged activations and variable
+efficiencies near to those thresholds~\cite{
+atlas_trigger_egamma_run2,
+atlas_trigger_muon_run2
+}.
 
@@
 have increased in hardness with time.
-
-\TODO{reference for trigger details, eta range?}
 

### 2022-05-27 main/chapter_2ljets.tex
Run 2 to Run~2

@@
 This project exists to supplement those results with information from the full
-LHC Run 2 data-set which was completed in 2018 and contains
+LHC Run~2 data-set which was completed in 2018 and contains
 $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
@@
 % Run 1 data~\cite{atlas_2l_SUSY_2013_11},
-% partial Run 2 data~\cite{atlas_23l_SUSY_2016_24},
+% partial Run~2 data~\cite{atlas_23l_SUSY_2016_24},
 % atlas_rjr_23l_SUSY_2017_03,
@@
 \subsubsection{Former excesses}
-Excess data were observed in several signal regions of the partial Run 2
+Excess data were observed in several signal regions of the partial Run~2
 search ``for chargino--neutralino production using recursive
@@
 Both $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$
-have been revisited with the full Run 2 data-set and updated background
+have been revisited with the full Run~2 data-set and updated background
 modelling;
@@
 $\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$,
-with the full Run 2 data-set and updated background modelling,
+with the full Run~2 data-set and updated background modelling,
 and asks whether the excesses persist.

### 2022-05-27 main/chapter_2ljets.tex
mv2c10 add rnnip and smt notes

@@
 
-\subsubsection{$b$-tagging}
+\subsection{$b$-jets}
 \label{sec:2ljets_btagging}
@@
 and the `c10' suffix means that those data include a $7\%$ fraction of
-$c$-jets~\cite{atlas_twiki_mv2}.
+$c$-jets, and that the tagger does not use newer inputs from the
+RNNIP (Recurrent Neural Network Impact Parameter) tagger, nor the
+SMT (Soft Muon Tagger)~\cite{atlas_twiki_mv2, atlas_jet_mv2c10_2017}.
 
@@
 
-\subsection{$\ptmiss$}
+\subsection{$\met$}
 Every event has $\ptmiss$.

### 2022-05-29 main/chapter_2ljets.tex
swap order to rjr strong ewk for context

@@
 Supersymmetric signal processes in the $\twoljets$-electroweak
-analysis~\cite{atlas2022searches}.
+analysis.
+\textit{Figures reproduced from the paper}~\cite{atlas2022searches}.
 \\[0.5em]
@@
 
+The $\twoljets$-strong analysis is
+detailed in Section~\ref{sec:2ljets_origins_strong} and
+searches for effects from the strong sector.
+It also uses conventional event variables, and focuses on features in the
+distribution of the invariant masses of lepton pairs.
+
 My primary contributions are to the $\twoljets$-electroweak analysis,
@@
 sectors of supersymmetric models.
-
-Unsurprisingly, the strong part of the $\twoljets$ analysis is
-detailed in Section~\ref{sec:2ljets_origins_strong} and
-searches for effects from the strong sector.
-It also uses conventional event variables, and focuses on features in the
-distribution of the invariant masses of lepton pairs.
 
@@
 
+\subsection{Strong}
+\label{sec:2ljets_origins_strong}
+% previous papers
+% three signal models
+% mll, kinematic end point
+
 \subsection{Electroweak}
@@
 % orthogonal regions
-
-\subsection{Strong}
-\label{sec:2ljets_origins_strong}
-% previous papers
-% three signal models
-% mll, kinematic end point
 

### 2022-05-30 main/chapter_2ljets.tex
omega to sigma

@@
 Uncertainty in $\ptmiss$ is modelled for $\metsig$ by a covariance matrix
-$\Omega$ in the transverse plane, and used to define
+$\Sigma$ in the transverse plane, and used to define
 \begin{equation}
@@
 =
-\ptmiss \Omega^{-1} \ptmiss
+\ptmiss \Sigma^{-1} \ptmiss
 =
@@
 objects in an event,
-$\Omega$ is a sum of covariances assigned to those objects;
+$\Sigma$ is a sum of covariances assigned to those objects;
 both means and variances add linearly.

### 2022-05-30 main/chapter_2ljets.tex
begin strong context; add signal diagram figures

@@
 analysis.
-\textit{Figures reproduced from the paper}~\cite{atlas2022searches}.
+\textit{Figures reproduced from the paper%
+}~\cite{atlas2022searches, atlas_susy_feynman}.
 \\[0.5em]
@@
 \label{sec:2ljets_origins_strong}
+% intro
+Strong supersymmetry, producing super-partners of quarks and gluons, would be
+produced at much larger cross-sections than electroweak supersymmetry at the
+LHC.
+Squarks and gluinos are therefore excluded with masses up to several
+$\eV[T]$ using hadronic data alone, depending on which simplified model is
+considered~\cite{atlas_susy_strong_0l}.
+
+The $\twoljets$-strong search complements the search for strongly interacting
+supersymmetric effects by including cases where their decays produce pairs
+of charged leptons.
+Like the RJR and electroweak parts, jets and invisible lightest supersymmetric
+particles are also generated.
+
+% mll, kinematic end point
+Unlike the RJR and electroweak parts, not all of these leptons are produced
+by $Z$ boson decays.
+As illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
+featuring sleptons can also produce lepton pairs.
+Lepton pairs produced in these slepton interactions are of the same flavor
+and opposite charge, like those from $Z$ boson decays, but they have different
+mass dilepton mass characteristics.
+
+Rather than peaking at a resonance mass, these pairs have an $\mll$
+distribution with a signature end-point which depends on the masses of other
+supersymmetric particles in the
+$\neutralino_2 \ldots \slepton \ldots \neutralino_1$
+decay chain~\cite{paige1996determining}:
+\begin{equation}
+\max \mll
+= m(\neutralino_2)
+\sqrt{1 - \frac{m(\slepton\,)^2}{m(\neutralino_2)^2}}
+\sqrt{1 - \frac{m(\neutralino_1)^2}{m(\slepton\,)^2}}
+.
+\end{equation}
+
 % previous papers
 % three signal models
-% mll, kinematic end point
+\begin{figure}[t]
+\centering
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqllllN1N1_N2.pdf}
+    \caption{gluino-slepton}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqZZN1N1.pdf}
+    \caption{gluino-$Z^{(*)}$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_sqsq_qqZZN1N1.pdf}
+    \caption{squark-$Z^{(*)}$}
+\end{subfigure}
+\caption{%
+Supersymmetric signal processes in the $\twoljets$-strong search.
+\textit{Figures reproduced from the paper%
+}~\cite{atlas2022searches, atlas_susy_feynman}.
+\\[0.5em]
+\TODO{expand{
+}
+\label{fig:2ljets_strong_signal_diagrams}
+\end{figure}
+
+% three exclusion plots
 

### 2022-05-30 main/chapter_2ljets.tex
use subfigure for ewk; add some todos; some cleanup

@@
 \centering
-\includegraphics[width=0.48\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
+\begin{subfigure}{0.48\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
+    \caption{C1N2}
+\end{subfigure}
 \hfill
-\includegraphics[width=0.45\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
+    \caption{GMSB}
+\end{subfigure}
 \caption{%
@@
 
-The $\twoljets$-strong search complements the search for strongly interacting
-supersymmetric effects by including cases where their decays produce pairs
+The $\twoljets$-strong search target strongly interacting supersymmetric effects in cases where their decays produce pairs
 of charged leptons.
@@
 \hfill
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.47\textwidth}
     \centering
@@
 \\[0.5em]
-\TODO{expand{
+\TODO{expand}
 }
@@
 % three exclusion plots
+
+\TODO{data driven backgrounds: flavour symmetry, photon method z+jets, 40GeV cut}
 

### 2022-05-30 main/chapter_2ljets.tex
add lit context for strong

@@
 \label{sec:2ljets_origins_strong}
-% intro
 Strong supersymmetry, producing super-partners of quarks and gluons, would be
@@
 
-The $\twoljets$-strong search target strongly interacting supersymmetric effects in cases where their decays produce pairs
-of charged leptons.
-Like the RJR and electroweak parts, jets and invisible lightest supersymmetric
-particles are also generated.
-
-% mll, kinematic end point
+The $\twoljets$-strong search targets strongly interacting supersymmetric
+effects in cases where their decays produce pairs of charged leptons.
+Like the RJR and electroweak $\twoljets$ searches, the $\twoljets$-strong
+signals produce jets and invisible lightest supersymmetric particles.
 Unlike the RJR and electroweak parts, not all of these leptons are produced
@@
 
-% previous papers
-% three signal models
+Various previous incarnations of this search have been conducted with similar
+selections and target models on in older data, including partial Run~2~\cite{
+atlas_susy_strong_2l_partial_run2_1,
+atlas_susy_strong_2l_partial_run2,
+cms_susy_2016_strong_2l_run2_1,
+cms_susy_2016_strong_2l_run2
+} and Run~1~\cite{
+atlas_susy_strong_2l_run1,
+cms_susy_2016_strong_2l_run1
+} in \atlas\ and \cms.
+One can cherry-pick excesses in these Run~1 searches,
+of ``$3.0$ standard deviations'' in \atlas~\cite{atlas_susy_strong_2l_run1}
+and ``$2.6$ standard deviations'' in \cms~\cite{cms_susy_2016_strong_2l_run1}.
+These excesses were not reproduced in their early Run~2 updates.
+
+% signal models
 \begin{figure}[t]
@@
 
-% three exclusion plots
+\TODO{three exclusion plots}
 

### 2022-05-30 main/chapter_2ljets.tex
add strong exclusion plots initial

@@
 % signal models
-\begin{figure}[t]
+\begin{figure}[tp]
 \centering
@@
 
-\TODO{three exclusion plots}
+% exclusion plots
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_gluino_slepton.pdf}
+    \caption{gluino-slepton}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_gluino_z.pdf}
+    \caption{gluino-$Z^{(*)}$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_squark_z.pdf}
+    \caption{squark-$Z^{(*)}$}
+\end{subfigure}
+\caption{%
+Contours for signal models of the $\twoljets$-strong search.
+\textit{Figures reproduced from the paper%
+}~\cite{atlas2022searches}.
+\\[0.5em]
+\TODO{expand}
+}
+\label{fig:2ljets_strong_contours}
+\end{figure}
+
+Figure~\ref{fig:2ljets_strong_contours} \TODO{sentence}
 

### 2022-05-30 main/chapter_2ljets.tex
figure caption improvements

@@
 {Robert~M.~Pirsig,
-\textit{Zen and the Art of Motorcycle Maintenance},
+\emph{Zen and the Art of Motorcycle Maintenance},
 1974~\cite{pirsig1999zen}}
@@
 {Richard~Feynman,
-\textit{Cargo Cult Science},
+\emph{Cargo Cult Science},
 1974~\cite{feynman1974cargo}}
@@
 analysis.
-\textit{Figures reproduced from the paper%
+\emph{Figures are reproduced from the paper%
 }~\cite{atlas2022searches, atlas_susy_feynman}.
 \\[0.5em]
-Left: C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
+(a): C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
 is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
@@
 \\[0.5em]
-Right: GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
+(b): GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
 is produced by soft decays from pairs including $\chargino_1$,
@@
 \caption{%
-Decay trees for the $\twoljets$-RJR analysis~\cite{atlas2022searches}.
+Decay trees for the $\twoljets$-RJR analysis.
+\emph{This figure was made by the RJR team.}~\cite{atlas2022searches}.
 Both diagrams represent processes also targetted in the electroweak analysis.
@@
 analysis.
-\emph{This figure was made by the RJR team.}
 }
@@
 Summary plots for RJR analyses:
-(a) and (b)~two- and three- lepton selections respectively reproduced
-from~\cite{atlas_rjr_23l_SUSY_2017_03},
+(a) and (b)~two- and three- lepton selections respectively,
+\emph{reproduced from}~\cite{atlas_rjr_23l_SUSY_2017_03},
 (c)~all regions from the $\twoljets$-RJR search~\cite{atlas2022searches}.
@@
 $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$ of (b).
-\textit{Sub-figure (c) was made by the RJR team.}
+\emph{Sub-figure (c) was made by the RJR team.}
 }
@@
 Supersymmetric signal processes in the $\twoljets$-strong search.
-\textit{Figures reproduced from the paper%
+\emph{Figures reproduced from the paper%
 }~\cite{atlas2022searches, atlas_susy_feynman}.
 \\[0.5em]
-\TODO{expand}
+All models produce same-flavour, opposite-sign lepton pairs along with jets
+and $\met$.
+The gluino-slepton model (a) has a kinematic endpoint in $\mll$
+given in Equation~\ref{eqn:2ljets_strong_max_mll}.
+The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
+resonances, which may be forced off-shell for small mass splittings
+$m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
 }
@@
 Contours for signal models of the $\twoljets$-strong search.
-\textit{Figures reproduced from the paper%
+\emph{Figures were produced by the strong team%
 }~\cite{atlas2022searches}.
@@
 that trigger do not also harmfully increase inseparable backgrounds.
-A study, \textit{which was performed by Knut Vadla}, investigated the effects
+A study, \emph{which was performed by Knut Vadla}, investigated the effects
 of of including one-lepton triggers.

### 2022-05-30 main/chapter_2ljets.tex
add strong contours caption

@@
 Excess data were observed in several signal regions of the partial Run~2
-search ``for chargino--neutralino production using recursive
+search ``for chargino-neutralino production using recursive
 jigsaw reconstruction in final states with two or three charged
@@
 \begin{equation}
+\label{eqn:2ljets_strong_max_mll}
 \max \mll
@@
 \\[0.5em]
-\TODO{expand}
+Each model point uses fit results from the signal region with the
+``best expected sensitivity'';
+jagged edges arise from the switching between contours set by these different
+signal regions.
+The grey shaded contours showing previous exclusion are taken
+from~\cite{atlas_susy_strong_2l_partial_run2}.
 }

### 2022-05-30 main/chapter_2ljets.tex
fill in strong context todos

@@
 
+% signal models
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqllllN1N1_N2.pdf}
+    \caption{gluino-slepton}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.47\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqZZN1N1.pdf}
+    \caption{gluino-$Z^{(*)}$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_strong_sqsq_qqZZN1N1.pdf}
+    \caption{squark-$Z^{(*)}$}
+\end{subfigure}
+\caption{%
+Supersymmetric signal processes in the $\twoljets$-strong search.
+\emph{Figures reproduced from the paper%
+}~\cite{atlas2022searches, atlas_susy_feynman}.
+\\[0.5em]
+All models produce same-flavour, opposite-sign lepton pairs along with jets
+and $\met$.
+The gluino-slepton model (a) has a kinematic endpoint in $\mll$
+given in Equation~\ref{eqn:2ljets_strong_max_mll}.
+The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
+resonances, which may be forced off-shell for small mass splittings
+$m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
+}
+\label{fig:2ljets_strong_signal_diagrams}
+\end{figure}
+
 Rather than peaking at a resonance mass, these pairs have an $\mll$
@@
 
-% signal models
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqllllN1N1_N2.pdf}
-    \caption{gluino-slepton}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.47\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/2ljets_strong_gogo_qqqqZZN1N1.pdf}
-    \caption{gluino-$Z^{(*)}$}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.45\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{figures/2ljets_strong_sqsq_qqZZN1N1.pdf}
-    \caption{squark-$Z^{(*)}$}
-\end{subfigure}
-\caption{%
-Supersymmetric signal processes in the $\twoljets$-strong search.
-\emph{Figures reproduced from the paper%
-}~\cite{atlas2022searches, atlas_susy_feynman}.
-\\[0.5em]
-All models produce same-flavour, opposite-sign lepton pairs along with jets
-and $\met$.
-The gluino-slepton model (a) has a kinematic endpoint in $\mll$
-given in Equation~\ref{eqn:2ljets_strong_max_mll}.
-The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
-resonances, which may be forced off-shell for small mass splittings
-$m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
-}
-\label{fig:2ljets_strong_signal_diagrams}
-\end{figure}
-
 % exclusion plots
@@
 
-Figure~\ref{fig:2ljets_strong_contours} \TODO{sentence}
-
-\TODO{data driven backgrounds: flavour symmetry, photon method z+jets, 40GeV cut}
+To target its signal models, $\twoljets$-strong search defines seven
+overlapping signal regions.
+Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
+and the others are binned in $\mll$ to give broad sensitivity to the various
+kinematic edges across the grid of signal model parameters.
+
+The largest Standard Model background to this search is dileptonic $\ttbar$,
+in part because no $b$-jet veto is used.
+All three signal models produce same-flavor lepton pairs, but $\ttbar$
+produces same-flavor ($ee$/$\mu\mu$) and different-flavor ($e\mu$) pairs at
+equal rates.
+This difference in flavor symmetry is exploited to estimate backgrounds in
+$\twoljets$-strong,
+and its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
+The estimate works by taking different-flavor data, weighting them by factors
+to calibrate for differences in the leptons' reconstruction, and histogramming
+them in same-flavour regions.
+Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
+$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.
+
+Post-fit results match the data to within $1$-sigma in all $\twoljets$-strong
+regions.
+
+Exclusion contours from the $\twoljets$-strong search are reproduced in
+Figure~\ref{fig:2ljets_strong_contours}.
+Besides their increased reach, the jagged edges to parts of these contour are
+striking.
+These arise from the combination of results from overlapping signal regions.
+Because they overlap, they are not independent and cannot be trivially
+combined.
+Instead, each model in the signal grid is assigned to the signal region
+with the ``best expected sensitivity'' by some measure.
+An exclusion contour is constructed for each signal region, then those contours
+are joined such that models assigned to each region use its corresponding
+contour.
+
+Personally, I find this hard to interpret.
+If I had a pet theory which landed near to one of these sharp
+region-transition edges, what should I make of these data?
+Could some combination of the regions not have combined their information to
+produce more precise contours?
+Since model predictions change continuously, should our inferences not also?
+To aid interpretations and to allow combinations of regions'
+information, I chose to design the $\twoljets$-electroweak with only orthogonal
+regions, and without any manual switching of regions between models.
 

### 2022-05-30 main/chapter_2ljets.tex
begin 2ljets context

@@
 } in \atlas\ and \cms.
-One can cherry-pick excesses in these Run~1 searches,
+From these Run~1 searches, one can cherry-pick excesses
 of ``$3.0$ standard deviations'' in \atlas~\cite{atlas_susy_strong_2l_run1}
 and ``$2.6$ standard deviations'' in \cms~\cite{cms_susy_2016_strong_2l_run1}.
-These excesses were not reproduced in their early Run~2 updates.
+These excesses of data above estimated backgrounds were not reproduced in their
+early Run~2 updates.
 
@@
 % susy motivation
+
+As motivated in the previous sections, I personal tastes for how searches
+should be constructed.
+One taste is for simulation-driven background estimates, since they are
+interpretable as representations of the Standard Model and they can exceed the
+precision available in data alone.
+Another is for orthogonal regions, such that their likelihoods are independent
+and the information in their data can be cleanly combined.
+
+The core idea behind the design of the $\twoljets$-electroweak analysis is to
+update the regions of the partial Run~2 search~\cite{atlas_23l_SUSY_2016_24},
+(its $2\ell + jets$ channel).
+A particular goal was to improve sensitivity by replacing their overlapping
+`SR2-Int' and `SR2-High' regions with orthogonal, combined selections.
+\emph{I am grateful to Sarah Williams for suggesting this approach.}
+
+% run1
+\TODO{run 1?}
+\cite{atlas_2l_SUSY_2013_11}
+
+\TODO{CMS papers, compare exclusion}
+
+\TODO{sr2 int high figure}
+
 % 2/3l, older 2l context

### 2022-05-30 main/chapter_2ljets.tex
add 23l figure, crop pngs

@@
 The core idea behind the design of the $\twoljets$-electroweak analysis is to
-update the regions of the partial Run~2 search~\cite{atlas_23l_SUSY_2016_24},
-(its $2\ell + jets$ channel).
+update the regions of the partial Run~2 search in its $\twoljets$
+channel~\cite{atlas_23l_SUSY_2016_24}.
 A particular goal was to improve sensitivity by replacing their overlapping
@@
 \emph{I am grateful to Sarah Williams for suggesting this approach.}
+As in the $\twoljets$-strong search, each signal region is chosen from an
+expected sensitivity for each signal point.
+These regions and their resulting exclusion contours are displayed in
+Figure~\ref{fig:2ljets_23l_stuff}.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.62\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_23l_sr2inthigh.png}
+    \caption{SR2-Int and SR2-High}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.52\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_23l_contours_fig_08d.png}
+    \caption{Combined exclusion}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.46\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{figures/2ljets_23l_chosen_regions.png}
+    \caption{Chosen regions}
+\end{subfigure}
+\caption{%
+Figures reproduced from the partial Run~2 search with $2/3\ell$
+selections~\cite{atlas_23l_SUSY_2016_24, hepdata.81996}.
+\\[0.5em]
+(a) Histogrammed distribution of $\met$ in and below SR2-Int and SR2-High with
+example signal models.
+SR2-Int requires $\met > 150\,\eV[G]$ and SR2-High requires $\met > 250\,\eV[G]$.
+The rightmost bin incluedes overflow.
+\\[0.5em]
+(b) Contours for the combined $\twoljets$ and $3\ell$ channels.
+Each signal point chooses a signal region which is displayed in (c).
+\\[0.5em]
+(c) Signal regions selected for each signal point in the grid.
+Two-lepton selections occupy the largest mass splittings.
+The $3\ell$ regions are orthogonal and combined.
+The separation between SR2-low and $3\ell$ is not trivial.
+}
+\label{fig:2ljets_23l_stuff}
+\end{figure}
 

### 2022-05-30 main/chapter_2ljets.tex
tweaks, Int High > int high

@@
 A particular goal was to improve sensitivity by replacing their overlapping
-`SR2-Int' and `SR2-High' regions with orthogonal, combined selections.
+`SR2-int' and `SR2-high' regions with orthogonal, combined selections.
 \emph{I am grateful to Sarah Williams for suggesting this approach.}
-As in the $\twoljets$-strong search, each signal region is chosen from an
-expected sensitivity for each signal point.
-These regions and their resulting exclusion contours are displayed in
-Figure~\ref{fig:2ljets_23l_stuff}.
+
+As in the $\twoljets$-strong search, the partial Run~2 search chose one
+$\twoljets$ signal region from an expected sensitivity for each signal point.
+Two regions and the exclusion contours from that search are displayed in
+Figure~\ref{fig:2ljets_23l_stuff}/
 
@@
     \includegraphics[width=\textwidth]{figures/2ljets_23l_sr2inthigh.png}
-    \caption{SR2-Int and SR2-High}
+    \caption{SR2-int and SR2-high}
 \end{subfigure}
@@
 \\[0.5em]
-(a) Histogrammed distribution of $\met$ in and below SR2-Int and SR2-High with
+(a) Histogrammed distribution of $\met$ in and below SR2-int and SR2-high with
 example signal models.
-SR2-Int requires $\met > 150\,\eV[G]$ and SR2-High requires $\met > 250\,\eV[G]$.
-The rightmost bin incluedes overflow.
+SR2-int requires $\met > 150\,\eV[G]$ and SR2-high requires $\met > 250\,\eV[G]$.
+The rightmost bin includes overflow.
 \\[0.5em]
@@
 Each signal point chooses a signal region which is displayed in (c).
+The grey region shows exclusion from a previous search in Run~1
+data~\cite{atlas_2l_SUSY_2013_11}.
 \\[0.5em]
@@
 
-% run1
-\TODO{run 1?}
-\cite{atlas_2l_SUSY_2013_11}
-
 \TODO{CMS papers, compare exclusion}
-
-\TODO{sr2 int high figure}
-
-% 2/3l, older 2l context
-% update with run 2
-% multi-bin fit
-% orthogonal regions
 

### 2022-05-30 main/chapter_2ljets.tex
add gmsb paragraph

@@
 update the regions of the partial Run~2 search in its $\twoljets$
-channel~\cite{atlas_23l_SUSY_2016_24}.
+channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for signal models
+producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
 A particular goal was to improve sensitivity by replacing their overlapping
@@
 Two regions and the exclusion contours from that search are displayed in
-Figure~\ref{fig:2ljets_23l_stuff}/
+Figure~\ref{fig:2ljets_23l_stuff}.
 
@@
 
-\TODO{CMS papers, compare exclusion}
+Additionally, $\twoljets$ widens its scope to target a simplified GMSB model,
+which has been illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
+These two models behave similarly since both feature decays via massive bosons
+($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay trees
+to pairs of invisible supersymmetric particles.
+Although widening sensitivity to both models does sacrifice sensitivity to
+either one, the increased degree of model independence is itself of value for
+future reinterpretations of the data.
+
+Searches by \cms\ have found comparable exclusions contours for the
+C1N2 and GMSB models.
+\TODO{CMS papers}
 

### 2022-05-30 main/chapter_2ljets.tex
add cms run 2 references
grammar check; fix some warnings

@@
 Section~\ref{sec:2ljets_splash}.
-The meat of this chapter (or if you prefer, its gory detail) follows in
+The meat of this chapter (or, if you prefer, its gory detail) follows in
 Section~\ref{sec:2ljets_context} onwards.
@@
 
-In a large number orthogonal signal regions, no significant excess is observed
+In numerous orthogonal signal regions, no significant excess is observed
 above \emph{post-fit} backgrounds.
-Indeed there are some significant deficits --- data below the background noise.
+Indeed, there are some significant deficits --- data below the background noise.
 
@@
 $W\rightarrow q\bar q$ and
-$Z\rightarrow \ell^\pm \ell^\mp$ respectively..
-The other is a GMSB model which, after a variety of initial states produces
+$Z\rightarrow \ell^\pm \ell^\mp$ respectively.
+The other is a GMSB model which, from a variety of initial states, produces
 $\neutralino_1\textrm{--}\neutralino_1$ pairs.
@@
 
-Effects of some of these these signals in example signal regions of the
+Effects of some of these signals in example signal regions of the
 $\twoljets$-electroweak analysis are displayed in
@@
 Physically, signal yields would add on top of the backgrounds.
-Data are shown; regions in the the top two plots observe zero data, and
+Data are shown; regions in the top two plots observe zero data, and
 Poisson error bars for $0$ events are hidden for aesthetic reasons.
@@
 \\[0.5em]
-Space below the solid red line is labelled as excluded and its dotted
+Space below the solid red line is labelled as excluded, and its dotted
 neighbours show the result if all signal cross-sections are varied up and down
@@
 
-In this subsection, I claims attributions of work on the $\twoljets$ analysis
+In this subsection, I claim attributions of work on the $\twoljets$ analysis
 and related \atlas\ efforts.
@@
 \begin{itemize}
-\item Design, implementation and execution of the electroweak part of the analysis.
+\item Design, implementation, and execution of the electroweak part of the analysis.
 \item Service as `analysis contact' from January to October 2020.
@@
 Many plotting scripts, however, are modified from versions inherited from
-many \atlas\ members past and present to whom I am grateful.
+many \atlas\ members past and present, to whom I am grateful.
 
@@
 This project exists to supplement those results with information from the full
-LHC Run~2 data-set which was completed in 2018 and contains
+LHC Run~2 data-set, which was completed in 2018 and contains
 $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
@@
 \emph{This figure was made by the RJR team.}~\cite{atlas2022searches}.
-Both diagrams represent processes also targetted in the electroweak analysis.
+Both diagrams represent processes which are targetted in the electroweak
+analysis.
 (left) A fully-resolved decay tree with its particles labelled.
@@
 independence from uninteresting boosts, particularly those from the emission of
-QCD jets outside of the supersymmetric decay tree.
+QCD jets outside the supersymmetric decay tree.
 
@@
 ($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
-and other two required three
+and the other two required three
 ($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).
@@
 A more precisely framed problem with RJR variables is that they tend to select
-phase space which is poorly modelled by simulations.
+phase space, which is poorly modelled by simulations.
 This is plausibly explained by the fact that  RJR variables are non-trivially
@@
 or trust.
-Indeed I argue that all `data-driven' methods rely not only on data but on
+Indeed, I argue that all `data-driven' methods rely not only on data but on
 hypothetical extrapolations which may not be well motivated, particularly when
 studying complicated features such as RJR variables.
-For these reasons, I choose to not use any RJR event variables in the the
+For these reasons, I choose to not use any RJR event variables in the
 $\twoljets$-electroweak analysis.
@@
 featuring sleptons can also produce lepton pairs.
-Lepton pairs produced in these slepton interactions are of the same flavor
+Lepton pairs produced in these slepton interactions are of the same flavour
 and opposite charge, like those from $Z$ boson decays, but they have different
@@
 The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
-resonances, which may be forced off-shell for small mass splittings
+resonances, which may be forced off-shell for small mass-splittings
 $m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
@@
 in part because no $b$-jet veto is used.
-All three signal models produce same-flavor lepton pairs, but $\ttbar$
-produces same-flavor ($ee$/$\mu\mu$) and different-flavor ($e\mu$) pairs at
+All three signal models produce same-flavour lepton pairs, but $\ttbar$
+produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
 equal rates.
-This difference in flavor symmetry is exploited to estimate backgrounds in
+This difference in flavour symmetry is exploited to estimate backgrounds in
 $\twoljets$-strong,
 and its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
-The estimate works by taking different-flavor data, weighting them by factors
+The estimate works by taking different-flavour data, weighting them by factors
 to calibrate for differences in the leptons' reconstruction, and histogramming
@@
 
-As motivated in the previous sections, I personal tastes for how searches
+As motivated in the previous sections, I have personal tastes for how searches
 should be constructed.
 One taste is for simulation-driven background estimates, since they are
-interpretable as representations of the Standard Model and they can exceed the
+interpretable as representations of the Standard Model, and they can exceed the
 precision available in data alone.
@@
 \centering
-\begin{subfigure}{0.62\textwidth}
+\begin{subfigure}{0.6\textwidth}
     \centering
@@
 \end{subfigure}
-\hfill
+% \hfill
 \begin{subfigure}{0.46\textwidth}
@@
 SR2-int requires $\met > 150\,\eV[G]$ and SR2-high requires $\met > 250\,\eV[G]$.
-The rightmost bin includes overflow.
 \\[0.5em]
 (b) Contours for the combined $\twoljets$ and $3\ell$ channels.
-Each signal point chooses a signal region which is displayed in (c).
+Each signal point chooses a signal region, which is displayed in (c).
 The grey region shows exclusion from a previous search in Run~1
@@
 (c) Signal regions selected for each signal point in the grid.
-Two-lepton selections occupy the largest mass splittings.
+Two-lepton selections occupy the largest mass-splittings.
 The $3\ell$ regions are orthogonal and combined.
-The separation between SR2-low and $3\ell$ is not trivial.
 }
@@
 
-Searches by \cms\ have found comparable exclusions contours for the
-C1N2 and GMSB models.
-\TODO{CMS papers}
-
 \section{Events}
@@
 charged leptons (electrons or muons --- not taus), and at least one jet.
-Though not nominal, the a $\met$ requirement is also implicit from the context
+Though not nominal, a $\met$ requirement is also implicit from the context
 of $R$-parity conserving supersymmetric models.
@@
 
+Searches have been performed in \cms\ for similar
+C1N2~\cite{cms_susy_2022_c1n2} and GMSB models~\cite{
+cms_susy_2018_partial_run2_combination,
+cms_susy_2022_gmsb
+}, with comparable results.
+
 \subsection{Trigger}
@@
 When living our daily lives, we are bombarded with sensory data and must pay
-attention only only those gems few of any use, to avoid exhausting our
+attention only those gems few of any use, to avoid exhausting our
 processing capacity.
@@
 $\twoljets$, along with more specialized targets such as vector boson fusion
-or long lived particles.
+or long-lived particles.
 The list varies through time, so we select triggers specifically in
@@
 selects for an electron pair, a muon pair, or an emu (electron-muon pair).
-To describe those triggers let $\pt^{\ell_i}$ be the transverse momentum of the
+To describe those triggers, let $\pt^{\ell_i}$ be the transverse momentum of the
 $i\mathrm{th}$ hardest lepton.
@@
 This motivates lepton selections which exceed these trigger thresholds,
-so as to avoid the need to model their jagged activations and variable
+to avoid the need to model their jagged activations and variable
 efficiencies near to those thresholds~\cite{
@@
 
-Quite generally throughout the data used in $\twoljets$, trigger requirements
+Quite generally, throughout the data used in $\twoljets$, trigger requirements
 have increased in hardness with time.
@@
 The answer is that the benefits of additional triggers were assessed to not
-outweigh the cost of using using them.
+outweigh the cost of using them.
 
 The benefit of an additional trigger could be in increased signal yields if
-that trigger do not also harmfully increase inseparable backgrounds.
+that trigger does not also harmfully increase inseparable backgrounds.
 A study, \emph{which was performed by Knut Vadla}, investigated the effects
-of of including one-lepton triggers.
+of including one-lepton triggers.
 In a signal-region-like selection requiring two leptons, two jets and
@@
 low-level details of detector interactions.
-Tighter requirements primarily act to to reduce from non-prompt backgrounds,
+Tighter requirements primarily act to reduce from non-prompt backgrounds,
 which are muons from hadronic decays~\cite{atlas_muon_quality_MUON_2018_03}.
@@
 Like leptons, hadronic jets are identified with `baseline' and `signal'
-requirements with increasingly tight requirements on quality metrics and
+requirements, with increasingly tight requirements on quality metrics and
 kinematics.
@@
 
-\subsection{$b$-jets}
+\subsection{\textit{b}-jets}
 \label{sec:2ljets_btagging}
 Tagging adds labels to objects.
-Unlike how electrons and muons are very cleanly separated however, jet tagging
+Unlike how electrons and muons are very cleanly separated, however, jet tagging
 is a more noisy affair.
-Taggers are available for various jet properties, such charm flavor ($c$-jet),
-bottom flavor ($b$-jet).
+Taggers are available for various jet properties, such charm flavour ($c$-jet),
+bottom flavour ($b$-jet).
 Of these, $\twoljets$ uses only $b$-tagging.
@@
 To remove duplicates and further reduce non-prompt lepton contributions, we
-apply a standard overlap removal procedure.
+apply a standard overlap-removal procedure.
 This procedure, which is detailed in the paper~\cite{atlas2022searches},
@@
 
-\subsection{$\met$}
+\subsection{\texorpdfstring{$\met$}{ETmiss}}
 Every event has $\ptmiss$.
@@
 
-\subsection{$\met$ significance}
+\subsection{\texorpdfstring{$\met$ significance}{Etmiss significance}}
 \label{sec:2ljets_metsig}
-Most common Standard Model processes at the LHC do not produce hard
+Most common Standard Model processes at the LHC do not produce hard,
 invisible particles.
@@
 variable $\metsig$ helps to separate those events with real and well-measured
-$\met$ from those where the measures $\ptmiss$ is more likely to result from
+$\met$ from those where the measured $\ptmiss$ is more likely to result from
 mismeasurements.
@@
 contributions more than $\met$ alone, plausibly since $\ttbar$ events are
-accompanied with more hadronic activity than other sources of $\ptmiss$.
-What remains of the standard model in the high-$\metsig$ tail for $\twoljets$
+accompanied by more hadronic activity than other sources of $\ptmiss$.
+What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
 selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
@@
 quite pure diboson backgrounds and greater significance measures.
-Signal models with smaller mass splittings tend to have the bulk of their
+Signal models with smaller mass-splittings tend to have the bulk of their
 events at smaller $\met$.
 \\[0.5em]
-This selection requires two same-flavor opposite-sign signal leptons with
+This selection requires two same-flavour opposite-sign signal leptons with
 $\pt^{\ell_2} > 25\,\eV[G]$,
@@
 
-\subsection{Stranverse mass $\mttwo$}
+\subsection{Stranverse mass \texorpdfstring{$\mttwo$}{mT2}}
 \label{sec:2ljets_mt2}
@@
 margin for experimental resolution.
-Such cases are common since, R-parity conservation implies pair production,
+Such cases are common, since R-parity conservation implies pair production,
 and the idea that dark matter could comprise the lightest supersymmetric
@@
 the four-vector with energy $E$ and momentum $\vec p$~\cite{lester1999measuring}.
-As an extension, one can assign different masses $m_p$ and $m_q$ to the
+As an extension, one can assign different masses $m_p$ and $m_q$ to
 the two invisible products~\cite{lester2015bisection}.
@@
 bisection algorithms, which seek the least resonance mass for which the
-transverse masses allow any compatible assignment such that
+transverse masses allow any compatible assignment which also satisfies
 $\vec p + \vec q = \ptmiss$.
@@
 analysis.
-To aid future revision of their meanings, definitions and references for these
+To aid future revision of their meanings, definitions, and references for these
 variables are collected here.
@@
 Please remember that none is an exact or true value;
-all are are estimates based on approximate reconstructions.
-While this distinction is easy to drop in lax language, is necessary to
+all are estimates based on approximate reconstructions.
+While this distinction is easy to drop in lax language, it is necessary to
 understand the shapes of noisy variables such as $\mjj$ and indeed the
@@
 Splitting regions on $\metsig$ is the core idea behind the design.
-The basic motivation for this design are displayed in
+The basic motivation for this design is displayed in
 Figure~\ref{fig:2ljets_presel_met};

### 2022-05-30 main/chapter_2ljets.tex
\approx to approximately

@@
 This $100\,\eV[G]$ also corresponds roughly to several standard deviations of
-a typical $\met$ noise scale of $\approx 10\textrm{--}30\,\eV[G]$, which I observe
+a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$, which I observe
 by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections,

### 2022-05-31 main/chapter_2ljets.tex
expan ewk with problems

@@
 LHC Run~2 data-set, which was completed in 2018 and contains
-$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13\,\eV[T]$.
+$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
+$\sqrt{s} = 13\,\eV[T]$.
 Leverage of the increased data quantity is one key aim.
@@
 \label{sec:2ljets_origins_electroweak}
-% susy motivation
-
-As motivated in the previous sections, I have personal tastes for how searches
-should be constructed.
-One taste is for simulation-driven background estimates, since they are
-interpretable as representations of the Standard Model, and they can exceed the
-precision available in data alone.
-Another is for orthogonal regions, such that their likelihoods are independent
-and the information in their data can be cleanly combined.
-
-The core idea behind the design of the $\twoljets$-electroweak analysis is to
+As motivated in the previous sections, I have personal preferences for
+simulation-driven backgrounds and orthogonal multi-bin regions.
+This is because simulation-driven estimates are interpretable as
+Standard Model predictions,
+and orthogonal regions have independent likelihoods which can be cleanly
+combined to leverage the information in their data.
+
+The main idea behind the design of the $\twoljets$-electroweak analysis is to
 update the regions of the partial Run~2 search in its $\twoljets$
-channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for signal models
+channel~\cite{atlas_23l_SUSY_2016_24} which also searched for signal models
 producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
 A particular goal was to improve sensitivity by replacing their overlapping
-`SR2-int' and `SR2-high' regions with orthogonal, combined selections.
+`SR2-int' and `SR2-high' regions with orthogonal selections.
 \emph{I am grateful to Sarah Williams for suggesting this approach.}
@@
 As in the $\twoljets$-strong search, the partial Run~2 search chose one
-$\twoljets$ signal region from an expected sensitivity for each signal point.
-Two regions and the exclusion contours from that search are displayed in
-Figure~\ref{fig:2ljets_23l_stuff}.
+signal region from an expected sensitivity for each signal point,
+as displayed in
+Figure~\ref{fig:2ljets_23l_stuff}
+with the two regions and exclusion contours.
 
@@
 
-Additionally, $\twoljets$ widens its scope to target a simplified GMSB model,
-which has been illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
-These two models behave similarly since both feature decays via massive bosons
-($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay trees
-to pairs of invisible supersymmetric particles.
+Additionally, $\twoljets$ widens its scope to target a GMSB simplified model
+illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
+The C1N2 and GMSB models behave similarly since both feature decays via massive
+bosons ($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay
+trees to pairs of invisible supersymmetric particles.
 Although widening sensitivity to both models does sacrifice sensitivity to
@@
 future reinterpretations of the data.
+We are happy to sacrifice some precision for generality.
+
+Multi-bin fits are not special.
+They are standard \atlas\ practice, and there could have been no complication
+in replacing the overlapping SR2-int and SR2-high with orthogonally binned
+regions.
+However, my design decisions for $\twoljets$-electroweak
+caused problems for the clarity of communication and accuracy of background
+modelling.
+These issues are developed further through the remainder of this chapter.
+
+A multi-binned histogram is usually and clearly illustrated as slicing up one
+variable.
+My design of $\twoljets$-electroweak violates this.
+Instead, it is designed as a tree whose branches recursively split various
+variables.
+This leads to specifications which are challenging to present and understand,
+and which require region names with numerous suffixes to specify where they
+live in that tree.
+
+Worse, I disregarded the idea that every region should have a large
+background expectation.
+Normally, this is justified by the limitations of approximate statistical
+methods, which work
+``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
+But the Poisson likelihood works for infinitesimally small
+expectations~\cite{skilling2009cosmology}, and in small-bin limit one has
+continuous fits (sometimes called ``extended''~\cite{barlow1990extended}),
+which are standard and easily analysed.
+
+The real reason to avoid small backgrounds is, in this context, to allow honest
+assignments of background uncertainties.
+Although main simulated samples may have good precision,
+many `systematic' variations and `data-driven' estimates are also required.
+Many of these alternative estimates have far worse precision, and can fail
+totally as the number of simulated events accepted plunges towards zero;
+they can require $\approx10$ data events to work properly.
+
+Unfortunately, $\twoljets$-electroweak selections were finalized before I
+appreciated this problem that background variations were dubious.
+Various workaround are invoked to assign plausible variations despite small
+yields, which we shall in the remainder of this chapter.
+
+As seen in the splash summary of Section~\ref{sec:2ljets_splash}, this search
+finds agreement with backgrounds except for some local deficits of data below
+backgrounds.
+Exclusion contours exceed their expected range because of these deficits,
+but they can only be trusted if the background model is trusted.
+
+Searches have been performed in \cms\ for similar
+C1N2~\cite{cms_susy_2022_c1n2} and GMSB models~\cite{
+cms_susy_2018_partial_run2_combination,
+cms_susy_2022_gmsb
+}, with comparable results.
 
@@
 $\twoljets$-electroweak search.
-
-Searches have been performed in \cms\ for similar
-C1N2~\cite{cms_susy_2022_c1n2} and GMSB models~\cite{
-cms_susy_2018_partial_run2_combination,
-cms_susy_2022_gmsb
-}, with comparable results.
 
@@
 
-\subsection{\texorpdfstring{$\met$ significance}{Etmiss significance}}
+\subsection{\texorpdfstring{$\met$}{ETmiss} significance}
 \label{sec:2ljets_metsig}

### 2022-05-31 main/chapter_2ljets.tex
expand design intro

@@
 \FloatBarrier
-\section{Context}
+\section{Summaries}
 \label{sec:2ljets_context}
@@
 But the Poisson likelihood works for infinitesimally small
-expectations~\cite{skilling2009cosmology}, and in small-bin limit one has
-continuous fits (sometimes called ``extended''~\cite{barlow1990extended}),
+expectations~\cite{skilling2009cosmology}, and in the small-bin limit one has
+continuous distributions
+(sometimes called ``extended''~\cite{barlow1990extended}),
 which are standard and easily analysed.
@@
 \section{Events}
-From its name, the $\twoljets$ analysis analyses features of events with two
-charged leptons (electrons or muons --- not taus), and at least one jet.
+This section describes the events selected for the $\twoljets$ searches, the
+reconstructed particle objects derived from those events and various
+important event variables.
+
+From its name, it is clear that $\twoljets$ analyses features of events with
+two charged leptons (electrons or muons --- not taus), and at least one jet.
 Though not nominal, a $\met$ requirement is also implicit from the context
@@
 }.
-
-Quite generally, throughout the data used in $\twoljets$, trigger requirements
+Throughout the data used in $\twoljets$, trigger requirements
 have increased in hardness with time.
@@
 
-Splitting regions on $\metsig$ is the core idea behind the design.
-The basic motivation for this design is displayed in
+The clearest idea behind this design is to split regions on $\metsig$.
+Motivation for this design is displayed in
 Figure~\ref{fig:2ljets_presel_met};
 $\metsig$ does better than $\met$ in separating background and signal
-contributions.
+contributions for models with various mass splittings above $m(Z)$.
 But there is more --- $\metsig$ also decently separates signal models from one
@@
 Binning regions in $\metsig$ therefore also improves sensitivity to
-individual parameter points and allows targeted selections on other, locally-relevant,
-event variables.
-
-% CRs go in with relevant SR category
+individual parameter points and allows targeted selections on other,
+locally-relevant, event variables.
+
+Regions in three blocks of $\metsig$ are named High, Intermediate (or Int),
+and Low,
+which target models with decreasing mass splittings and lepton pairs on the
+$Z$ resonance mass.
+Separate regions target $\mll < m(Z)$ for models with small mass splittings,
+which force decays to proceed via low-mass off-shell weak bosons.
+
+Control and validation regions are defined to constrain the background
+modelling in all regions.
+
+All main regions are defined in Table~\TODO{tab}.
+High regions are discussed in Section~\ref{sec:2ljets_high}.
+Intermediate regions are discussed in Section~\ref{sec:2ljets_int}.
+Low regions are discussed in Section~\ref{sec:2ljets_low}.
+Off-shell regions are discussed in Section~\ref{sec:2ljets_high}.
+Finally, these are combined into model-independent `Discovery' regions as
+discussed in Section~\ref{2ljets_disco}.
+
 \subsection{High}
+\label{sec:2ljets_high}
 
 \subsection{Intermediate}
+\label{sec:2ljets_int}
 
 \subsection{Low}
+\label{sec:2ljets_low}
 
 \subsection{Off-shell}
+\label{sec:2ljets_offshell}
 
 \subsection{Discovery}
+\label{sec:2ljets_disco}
 

### 2022-05-31 main/chapter_2ljets.tex
add region defn tables for Hight and Int

@@
 (left) A fully-resolved decay tree with its particles labelled.
-(right) Only the $Z\rightarrow\ell\ell$ (\underline{V}ector boson)
+(right) Only the $Z\rightarrow\ell\ell$ \underline{V}ector boson)
 is resolved, but the supersymmetric system recoils off
@@
 
-All main regions are defined in Table~\TODO{tab}.
+\TODO{define presel, preselection SFOS 25 1j met>100 metsig>6}
+
 High regions are discussed in Section~\ref{sec:2ljets_high}.
@@
 Off-shell regions are discussed in Section~\ref{sec:2ljets_high}.
+Control and validation regions are included in these sections near to their
+targetted regions.
 Finally, these are combined into model-independent `Discovery' regions as
-discussed in Section~\ref{2ljets_disco}.
+discussed in Section~\ref{sec:2ljets_disco}.
 
@@
 
+Table~\ref{tab:2ljets_high}.
+
+\begin{table}[tp]
+\footnotesize
+\centering
+\begin{tabular}{lcccccccc}
+& $\njet$
+& $\nbtag$
+& $\metsig$
+& $m_X/\eV[G]$
+& $\Delta R_X$
+\\[1em]
+SR-High
+& $\geq2$
+& $\leq 1$
+& $18\textrm{--}21\textrm{--}\infty$
+& $\mjj \in 60\textrm{--}110$
+&
+$\rjj \in 0\textrm{--}0.8\textrm{--}1.6$
+\\[.5em]
+\quad VR-High
+& $\geq2$
+& $\leq 1$
+& $> 18$
+& $\heartsuit \mjj \in 20\textrm{--}60 \mid 60\textrm{--}110$
+& $\rjj < 1.6$
+\\[.5em]
+\quad VR-High-R
+& $\geq 2$
+& $\leq 1$
+& $> 18$
+& $\heartsuit \mjj > 20$
+& $\heartsuit \rjj > 1.6$
+\\[1em]
+SR-High-1J
+&
+$\hphantom{\geq}~1$
+& $\leq 1$
+& $> 12$
+& $\mjetone \in 60\textrm{--}110$
+&
+\\[.5em]
+\quad VR-High-1J
+& $\hphantom{\geq}~1$
+& $\leq 1$
+& $> 12$
+& $\heartsuit \mjetone \in 20\textrm{--}60 \mid 60\textrm{--}110$
+&
+\\[1em]
+SR-$\ell\ell bb$
+& $\geq 2$
+& $\geq 2$
+& $> 18$
+& $\mbb \in 60\textrm{--}150$
+&
+\\[.5em]
+\quad VR-$\ell\ell bb$
+& $\geq 2$
+& $\geq 2$
+& $\heartsuit 12\textrm{--}18$
+& $\mbb \in 60\textrm{--}150$
+&
+\end{tabular}
+\small
+\\[1em]
+Common: presel, $\mll \in 71\textrm{--}111\,\eV[G]$, $\mttwo > 80\,\eV[G]$
+\caption{%
+% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
+%All regions except VR-SS require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\eV[G]\ and $\met>100~\eV[G]$.
+%Instead, VR-SS requires same-charge leptons.
+% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
+% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+Differences between regions are marked $\heartsuit$.
+\TODO{}
+}
+\label{tab:2ljets_high}
+\end{table}
+
 \subsection{Intermediate}
 \label{sec:2ljets_int}
+
+Table~\ref{tab:2ljets_int}.
+
+\begin{table}[tp]
+\footnotesize
+\centering
+\begin{tabular}{lcccccccc}
+& $\njet$
+& $\nbtag$
+& $\metsig$
+& $m_X/\eV[G]$
+& $\ptjone/\eV[G]$
+\\[1em]
+SR-Int
+& $\geq 2$
+& $\hphantom{\geq}~0$
+& $12\textrm{--}15\textrm{--}18$
+& $\mjj \in 60\textrm{--}110$
+& $> 60$
+\\[.5em]
+\quad VR-Int
+& $\geq 2$
+& $\hphantom{\geq}~0$
+& $12\textrm{--}18$
+& $\mjj \in 60\textrm{--}110$
+& $\heartsuit < 60$
+\\[1em]
+CR-VZ
+& $\geq 2$
+& $\hphantom{\geq}~0$
+& $12\textrm{--}18$
+& $\heartsuit \mjj \in 20\textrm{--}60 \mid 60\textrm{--}110$
+& $\heartsuit $
+\\[.5em]
+CR-tt
+& $\geq 2$
+& $\heartsuit \geq 1$
+& $\heartsuit 9\textrm{--}12$
+& $\heartsuit \mjj > 20$
+& $> 60$
+\end{tabular}
+\small
+\\[0.5em]
+Common: presel, $\mll \in 81\textrm{--}101\,\eV[G]$, $\mttwo > 80\,\eV[G]$
+\caption{%
+% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
+%All regions except VR-SS require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\eV[G]\ and $\met>100~\eV[G]$.
+%Instead, VR-SS requires same-charge leptons.
+% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
+% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+Differences between regions are marked $\heartsuit$.
+\TODO{}
+}
+\label{tab:2ljets_int}
+\end{table}
 

### 2022-05-31 main/chapter_2ljets.tex
add top q observation example

@@
 which are standard and easily analysed.
+
+Events appearing in a region with minimal background is evidence in favour
+of new contributions, and that only strengthens as the background approaches
+zero.
+For example, the top quark was observed by D0 in 1995 using seven regions all with
+background expectations of $1$ or less~\cite{abachi1995observation},
+for a total of $3.79\pm0.55$ expected and $17$ observed.
+If they could do that with 20th century technology, we can do it now.
 

### 2022-05-31 main/chapter_2ljets.tex
dzero command

--- 2022-05-31 main/chapter_2ljets.tex
@@
 which are standard and easily analysed.
+
+Events appearing in a region with minimal background is evidence in favour
+of new contributions, and that only strengthens as the background approaches
+zero.
+For example, the top quark was observed by \dzero\ using seven regions all with
+background expectations of $1$ or less, for a total of $3.79\pm0.55$ expected
+and $17$ observed~\cite{abachi1995observation}.
+If they could do that with technology from the year $1995$, we can do it now.
 

### 2022-05-31 main/chapter_2ljets.tex
add low region def table
add low region defn table
add preselection table
add list of disco regions
fix quotes
reduce gev noise

@@
 variables are collected here.
+All units are $\eV[G]$ where appropriate.
 \begin{itemize}
@@
 \item $\mjj$: invariant mass of the two hardest jets
-\item $\mbb$: invariant mass of the two hardest b-tagged jets
+\item $\mbb$: invariant mass of the two hardest $b$-tagged jets
 \item $\mjetone$: mass of the hardest jet
-\item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
+\item $\mttwoll$: `stransverse' mass of the lepton pair with massless invisibles;\\
 described in Section~\ref{sec:2ljets_mt2}
@@
 
+Table~\ref{tab:2ljets_presel}
 \TODO{define presel, preselection SFOS 25 1j met>100 metsig>6}
@@
 
+\begin{table}[tp]
+\centering
+\begin{tabular}{c}
+Preselection:
+\\[1em]
+$n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$
+\\[.5em]
+$\njet \geq 1$
+\\[.5em]
+same flavour, opposite charge $\ell^+\ell^-$
+\\[.5em]
+$p_\mathrm{T}^{\,\ell_2} > 25\,\eV[G]$
+\\[.5em]
+$\met > 100\,\eV[G]$
+\\[.5em]
+$\metsig > 6$
+\end{tabular}
+\caption{%
+``Preselection'': requirements ensured prior to all region definitions in the
+$\twoljets$-electroweak search.
+}
+\label{tab:2ljets_presel}
+\end{table}
+
 \subsection{High}
@@
 \centering
-\begin{tabular}{lcccccccc}
+\begin{tabular}{lccccc}
 & $\njet$
@@
 & $\metsig$
-& $m_X/\eV[G]$
-& $\Delta R_X$
+& $m_X$
+& $\rjj$
 \\[1em]
@@
 & $18\textrm{--}21\textrm{--}\infty$
-& $\mjj \in 60\textrm{--}110$
-&
-$\rjj \in 0\textrm{--}0.8\textrm{--}1.6$
+& $\mjj\!:~  60\textrm{--}110$
+& $0\textrm{--}0.8\textrm{--}1.6$
 \\[.5em]
@@
 & $> 18$
-& $\heartsuit \mjj \in 20\textrm{--}60 \mid 60\textrm{--}110$
-& $\rjj < 1.6$
+& $\uwave{\mjj\!:~  20\textrm{--}60 \mid 60\textrm{--}110}$
+& $< 1.6$
 \\[.5em]
@@
 & $> 18$
-& $\heartsuit \mjj > 20$
-& $\heartsuit \rjj > 1.6$
+& $\uwave{\mjj\!:~  > 20}$
+& $\uwave{> 1.6}$
 \\[1em]
 SR-High-1J
-&
-$\hphantom{\geq}~1$
+& $\hphantom{\geq}~1$
 & $\leq 1$
 & $> 12$
-& $\mjetone \in 60\textrm{--}110$
+& $\mjetone\!:~  60\textrm{--}110$
 &
@@
 & $> 12$
-& $\heartsuit \mjetone \in 20\textrm{--}60 \mid 60\textrm{--}110$
+& $\uwave{\mjetone\!:~  20\textrm{--}60 \mid 60\textrm{--}110}$
 &
@@
 & $> 18$
-& $\mbb \in 60\textrm{--}150$
+& $\mbb\!:~  60\textrm{--}150$
 &
@@
 & $\geq 2$
-& $\heartsuit 12\textrm{--}18$
-& $\mbb \in 60\textrm{--}150$
+& $\uwave{12\textrm{--}18}$
+& $\mbb\!:~  60\textrm{--}150$
 &
@@
 \\[1em]
-Common: presel, $\mll \in 71\textrm{--}111\,\eV[G]$, $\mttwo > 80\,\eV[G]$
+Common: Preselection,
+$\mll \in 71\textrm{--}111\,\eV[G]$,
+$\mttwo > 80\,\eV[G]$.
 \caption{%
@@
 % The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
-Differences between regions are marked $\heartsuit$.
-\TODO{}
+Differences between regions are \uwave{underlined}.
+\TODO{caption}
 }
@@
 & $\metsig$
-& $m_X/\eV[G]$
-& $\ptjone/\eV[G]$
+& $\mjj$
+& $\ptjone$
 \\[1em]
@@
 & $12\textrm{--}15\textrm{--}18$
-& $\mjj \in 60\textrm{--}110$
+& $60\textrm{--}110$
 & $> 60$
@@
 & $12\textrm{--}18$
-& $\mjj \in 60\textrm{--}110$
-& $\heartsuit < 60$
+& $60\textrm{--}110$
+& $\uwave{< 60}$
 \\[1em]
@@
 & $12\textrm{--}18$
-& $\heartsuit \mjj \in 20\textrm{--}60 \mid 60\textrm{--}110$
-& $\heartsuit $
+& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{\hphantom{< 60}}$
 \\[.5em]
@@
 & $\geq 2$
-& $\heartsuit \geq 1$
-& $\heartsuit 9\textrm{--}12$
-& $\heartsuit \mjj > 20$
+& $\uwave{\geq 1}$
+& $\uwave{9\textrm{--}12}$
+& $\uwave{20}$
 & $> 60$
@@
 \small
-\\[0.5em]
-Common: presel, $\mll \in 81\textrm{--}101\,\eV[G]$, $\mttwo > 80\,\eV[G]$
+\\[1em]
+Common: Preselection,
+$\njet \geq 2$,
+$\mll \in 81\textrm{--}101\,\eV[G]$,
+$\mttwo > 80\,\eV[G]$.
 \caption{%
@@
 
+Table~\ref{tab:2ljets_low}
+
+\begin{table}[tp]
+\footnotesize
+\centering
+\begin{tabular}{lcccccccc}
+& $\metsig$
+& $\mjj$
+& $\mttwo$
+& $\rll$
+& $\dphillmet$
+\\[1em]
+SR-Low
+& $6\textrm{--}9\textrm{--}12$
+& $60\textrm{--}110$
+& $> 80$
+& $< 1$
+&
+\\[.5em]
+\quad VR-Low
+& $6\textrm{--}12$
+& $60\textrm{--}110$
+& $>80$
+& $1\textrm{--}1.4$
+&
+\\[1em]
+SR-Low-2
+& $6\textrm{--}9$
+& $60\textrm{--}110$
+& $< 80$
+& $< 1.6$
+& $< 0.6$
+\\[.5em]
+\quad VR-Low-2
+& $6\textrm{--}9$
+& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $< 80$
+& $< 1.6$
+& $< 0.6$
+\\[1em]
+CR-Z
+& $6\textrm{--}9$
+& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $> 80$
+& $\uwave{\hphantom{< 1.6}}$
+& $\uwave{\hphantom{< 0.6}}$
+\end{tabular}
+\small
+\\[1em]
+Common: Preselection,
+$\njet = 2$,
+$\nbtag = 0$,
+$\mll \in 81\textrm{--}101\,\eV[G]$.
+\caption{%
+% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
+%All regions except VR-SS-EWK require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\GeV\ and $\met>100~\GeV$.
+%Instead, VR-SS-EWK requires same-charge leptons.
+% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
+% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+Differences between regions are \uwave{underlined}.
+\TODO{caption}
+}
+\label{tab:2ljets_low}
+\end{table}
+
 \subsection{Off-shell}
@@
 
+Table~\ref{tab:2ljets_offshell}
+
+\begin{table}[tp]
+\footnotesize
+\centering
+\begin{tabular}{lccccc}
+& $\metsig$
+& $\mll$
+& $\mttwo$
+& $\ptjone$
+& $\dphijmet$
+\\[1em]
+SR-OffShell
+& $>9$
+& $12\textrm{--}40\textrm{--}71$
+& $>100$
+& $>100$
+& $>2$
+\\[.5em]
+\quad VR-OffShell
+& $>9$
+&  12--71
+& $\uwave{80\textrm{--}100}$
+& $> 100$
+& $> 2$
+\\[1em]
+CR-DY
+& $\uwave{6\textrm{--}9}$
+& $12\textrm{--}71$
+& $> 100$
+& $\uwave{\hphantom{> 100}}$
+& $\uwave{\hphantom{> 2}}$
+\end{tabular}
+\small
+\\[1em]
+Common: Preselection,
+$\njet \geq 2$,
+$\nbtag = 0$.
+\caption{%
+Differences between regions are \uwave{underlined}.
+\TODO{caption}
+}
+\label{tab:2ljets_offshell}
+\end{table}
+
 \subsection{Discovery}
 \label{sec:2ljets_disco}
+
+\begin{itemize}
+\item DR-High: SR-High-8
+\item DR-$\ell\ell bb$: SR-$\ell\ell bb$
+\item DR-Int: SR-Int
+\item DR-Low: SR-Low
+\item DR-OffShell: SR-OffShell
+\end{itemize}
 

### 2022-05-31 main/chapter_2ljets.tex
further reduce gev noise
full font size, condense tables
add region def captions

@@
 variables are collected here.
-All units are $\eV[G]$ where appropriate.
+All dimensionful quantities are in $\eV[G]$.
 \begin{itemize}
@@
 targetted regions.
-Finally, these are combined into model-independent `Discovery' regions as
-discussed in Section~\ref{sec:2ljets_disco}.
+Finally, how these are combined into `Discovery' regions for
+model-independent interpretations is discussed in
+Section~\ref{sec:2ljets_disco}.
 
@@
 \\[.5em]
+same flavour, opposite charge leptons
+\\[.5em]
 $\njet \geq 1$
 \\[.5em]
-same flavour, opposite charge $\ell^+\ell^-$
+$p_\mathrm{T}^{\,\ell_2} > 25$
 \\[.5em]
-$p_\mathrm{T}^{\,\ell_2} > 25\,\eV[G]$
-\\[.5em]
-$\met > 100\,\eV[G]$
+$\met > 100$
 \\[.5em]
@@
 \caption{%
-``Preselection'': requirements ensured prior to all region definitions in the
+`Preselection' requirements ensured prior to all region definitions in the
 $\twoljets$-electroweak search.
@@
 \begin{table}[tp]
-\footnotesize
 \centering
@@
 \\[.5em]
-\quad VR-High
+\: VR-High
 & $\geq2$
@@
 \\[.5em]
-\quad VR-High-R
+\: VR-High-R
 & $\geq 2$
@@
 \\[.5em]
-\quad VR-High-1J
+\: VR-High-1J
 & $\hphantom{\geq}~1$
@@
 \\[.5em]
-\quad VR-$\ell\ell bb$
+\: VR-$\ell\ell bb$
 & $\geq 2$
@@
 \end{tabular}
-\small
 \\[1em]
 Common: Preselection,
-$\mll \in 71\textrm{--}111\,\eV[G]$,
-$\mttwo > 80\,\eV[G]$.
+$\mll \in 71\textrm{--}111$, and
+$\mttwoll > 80$.
 \caption{%
-% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
-%All regions except VR-SS require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\eV[G]\ and $\met>100~\eV[G]$.
-%Instead, VR-SS requires same-charge leptons.
-% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
-% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+High region definitions in the $\twoljets$-electroweak analysis.
+En-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$.
+Concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
+with boundaries at $a$, $b$, and $c$.
+The mid bar `$\mid$' indicates logical or.
 Differences between regions are \uwave{underlined}.
-\TODO{caption}
 }
@@
 \\[.5em]
-\quad VR-Int
+\: VR-Int
 & $\geq 2$
@@
 \end{tabular}
-\small
 \\[1em]
@@
 $\njet \geq 2$,
-$\mll \in 81\textrm{--}101\,\eV[G]$,
-$\mttwo > 80\,\eV[G]$.
+$\mll \in 81\textrm{--}101$, and
+$\mttwoll > 80$.
 \caption{%
-% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
-%All regions except VR-SS require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\eV[G]\ and $\met>100~\eV[G]$.
-%Instead, VR-SS requires same-charge leptons.
-% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
-% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+Intermediate region definitions in the $\twoljets$-electroweak analysis.
+En-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$.
+Concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
+with boundaries at $a$, $b$, and $c$.
+The mid bar `$\mid$' indicates logical or.
 Differences between regions are \uwave{underlined}.
-\TODO{caption}
 }
@@
 & $\mjj$
-& $\mttwo$
+& $\mttwoll$
 & $\rll$
@@
 \\[.5em]
-\quad VR-Low
+\: VR-Low
 & $6\textrm{--}12$
@@
 & $>80$
-& $1\textrm{--}1.4$
+& $\uwave{1\textrm{--}1.4}$
 &
@@
 \\[.5em]
-\quad VR-Low-2
+\: VR-Low-2
 & $6\textrm{--}9$
@@
 $\njet = 2$,
-$\nbtag = 0$,
-$\mll \in 81\textrm{--}101\,\eV[G]$.
+$\nbtag = 0$, and
+$\mll \in 81\textrm{--}101$.
 \caption{%
-% Overview of the High and Intermediate signal, control, and validations regions of the electroweak search.
-%All regions except VR-SS-EWK require exactly two opposite-charge, same-flavour signal leptons with $\pt>25$~\GeV\ and $\met>100~\GeV$.
-%Instead, VR-SS-EWK requires same-charge leptons.
-% The main requirements that distinguish the control and validation regions from the signal regions are highlighted in boldface.
-% The bracketed requirements denote the binning used in model-dependent fits, as described in the text.
+Low region definitions in the $\twoljets$-electroweak analysis.
+En-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$.
+Concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
+with boundaries at $a$, $b$, and $c$.
+The mid bar `$\mid$' indicates logical or.
 Differences between regions are \uwave{underlined}.
-\TODO{caption}
 }
@@
 & $\mll$
-& $\mttwo$
+& $\mttwoll$
 & $\ptjone$
@@
 \\[.5em]
-\quad VR-OffShell
+\: VR-OffShell
 & $>9$
@@
 \end{tabular}
-\small
 \\[1em]
 Common: Preselection,
-$\njet \geq 2$,
+$\njet \geq 2$, and
 $\nbtag = 0$.
 \caption{%
+Off-shell region definitions in the $\twoljets$-electroweak analysis.
+En-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$.
+Concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
+with boundaries at $a$, $b$, and $c$.
+The mid bar `$\mid$' indicates logical or.
 Differences between regions are \uwave{underlined}.
-\TODO{caption}
 }

### 2022-05-31 main/chapter_2ljets.tex
add files for figures: high design and disco significance

@@
 Table~\ref{tab:2ljets_high}.
+
+\TODO{}
+\begin{verbatim}
+2ljets_region_design_hist1d_mll_SRHigh.png
+2ljets_region_design_hist1d_nbtag_SRHigh.png
+\end{verbatim}
 
@@
 
+\TODO{}
+\begin{verbatim}
+2ljets_disco_High_C1N2_WZ_500p0_200p0_2L2J.png
+2ljets_disco_High_C1N2_WZ_600_0_2L2J.png
+2ljets_disco_Int_C1N2_WZ_250p0_100p0_2L2J.png
+2ljets_disco_Int_C1N2_WZ_400p0_200p0_2L2J.png
+2ljets_disco_llbb_C1N2N1_GGMHinoZh50_700.png
+\end{verbatim}
+
 \section{Modelling}

### 2022-06-01 main/chapter_2ljets.tex
add high figure; minor tidyup

@@
 
-\subsection{Stranverse mass \texorpdfstring{$\mttwo$}{mT2}}
+\subsection{\texorpdfstring{$\mttwo$}{mT2}}
 \label{sec:2ljets_mt2}
@@
 Table~\ref{tab:2ljets_high}.
-
-\TODO{}
-\begin{verbatim}
-2ljets_region_design_hist1d_mll_SRHigh.png
-2ljets_region_design_hist1d_nbtag_SRHigh.png
-\end{verbatim}
 
@@
 
+Figure~\ref{fig:2ljets_high_mll_b}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_mll_SRHigh.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_nbtag_SRHigh.png}
+\caption{}
+\end{subfigure}
+\caption{%
+SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$.
+The pure diboson background has very similar shapes to the signals.
+Tighter cuts would therefore reduce background and signal yields comparably,
+reducing sensitivity.
+The significance dropouts in (a) are errors in plotting.
+`Significance Z' is the binomial significance with $30\%$ relative background
+uncertainty~\cite{cousins2008evaluation}.
+Error bars are statistical.
+}
+\label{fig:2ljets_high_mll_b}
+\end{figure}
+
 \subsection{Intermediate}

### 2022-06-01 main/chapter_2ljets.tex
add disco plots

@@
 \label{sec:2ljets_high}
-
 Table~\ref{tab:2ljets_high}.
@@
 \label{sec:2ljets_int}
-
 Table~\ref{tab:2ljets_int}.
@@
 
-\TODO{}
-\begin{verbatim}
-2ljets_disco_High_C1N2_WZ_500p0_200p0_2L2J.png
-2ljets_disco_High_C1N2_WZ_600_0_2L2J.png
-2ljets_disco_Int_C1N2_WZ_250p0_100p0_2L2J.png
-2ljets_disco_Int_C1N2_WZ_400p0_200p0_2L2J.png
-2ljets_disco_llbb_C1N2N1_GGMHinoZh50_700.png
-\end{verbatim}
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_High_C1N2_WZ_500p0_200p0_2L2J.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_High_C1N2_WZ_600_0_2L2J.png}
+\caption{}
+\end{subfigure}
+\caption{%
+Evaluation of candidate High discovery regions.
+The region SR-High-8, labelled `High8\_1+\_2' here, is chosen for its
+simplicity and good sensitivity to both benchmark signal models.
+Upper plots show background and signal yields.
+Lower plots show various measures of projected
+sensitivity~\cite{cousins2008evaluation}.
+Yields are derived in an early iteration of the analysis and do not exactly
+match those in the final result.
+}
+\label{fig:2ljets_disco_trials_high}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_llbb_C1N2N1_GGMHinoZh50_600.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_llbb_C1N2N1_GGMHinoZh50_700.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
+\caption{%
+Evaluation of candidate High discovery regions.
+The region SR-$\ell\ell bb$, labelled `SRllbb' here, is chosen for
+sensitivity, simplicity, and to avoid overlapping with DR-High.
+Upper plots show background and signal yields.
+Lower plots show various measures of projected
+sensitivity~\cite{cousins2008evaluation}.
+Yields are derived in an early iteration of the analysis and do not exactly
+match those in the final result.
+}
+\label{fig:2ljets_disco_trials_llbb}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_Int_C1N2_WZ_250p0_100p0_2L2J.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_Int_C1N2_WZ_400p0_200p0_2L2J.png}
+\caption{}
+\end{subfigure}
+\caption{%
+Evaluation of candidate High discovery regions.
+The region SR-Int, labelled `Int\_1+\_2' here, is chosen for its simplicity
+and broad sensitivity.
+Upper plots show background and signal yields.
+Lower plots show various measures of projected
+sensitivity~\cite{cousins2008evaluation}.
+Yields are derived in an early iteration of the analysis and do not exactly
+match those in the final result.
+}
+\label{fig:2ljets_disco_trials_int}
+\end{figure}
 

### 2022-06-01 main/chapter_2ljets.tex
add prefit plot

@@
 
+Figure~\ref{fig:2ljets_disco_trials_high}
+Figure~\ref{fig:2ljets_disco_trials_llbb}
+Figure~\ref{fig:2ljets_disco_trials_int}
+Figure~\ref{fig:2ljets_disco_prefit}
+
 \begin{figure}[tp]
 \centering
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.45\textwidth}
 \centering
@@
 \hfill
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.45\textwidth}
 \centering
@@
 \centering
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.45\textwidth}
 \centering
@@
 \end{subfigure}
-\\[.5em]
 \caption{%
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRHigh.png}
+\caption{DR-High}
+\end{subfigure}
+\quad\quad
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRllbb.png}
+\caption{DR-$\ell\ell bb$}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRInt.png}
+\caption{DR-Int}
+\end{subfigure}
+\quad\quad
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRLow.png}
+\caption{DR-Low}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DROffShell.png}
+\caption{DR-OffShell}
+\end{subfigure}
+\caption{%
+Pre-fit yields in discovery regions.
+Errors are statistical.
+}
+\label{fig:2ljets_disco_prefit}
+\end{figure}
+
 \section{Modelling}

### 2022-06-01 main/chapter_2ljets.tex
add region def plots; add high figure

@@
 \label{sec:2ljets_high}
-Table~\ref{tab:2ljets_high}.
+Table~\ref{tab:2ljets_high}
+Figure~\ref{fig:2ljets_high_mll_b}
+Figure~\ref{fig:2ljets_high_region}
 
@@
 
-Figure~\ref{fig:2ljets_high_mll_b}
-
 \begin{figure}[tp]
@@
 \label{fig:2ljets_high_mll_b}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh8.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh16.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHighR.png}
+\caption{}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_high_region}
 \end{figure}

### 2022-06-01 main/chapter_2ljets.tex
add more region def figures

@@
 }, with comparable results.
+
+\TODO{other atlas C1N2 and GMSB searches (combinations, 3L offshell exclusion)}
 
@@
 Figure~\ref{fig:2ljets_high_region}
+Figure~\ref{fig:2ljets_high_1j_region}
+Figure~\ref{fig:2ljets_high_llbb_region}
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh4.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh4.png}
+\caption{\TODO{mj1}}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh4.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh4.png}
+\caption{\TODO{mj1}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_high_1j_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRllbb.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRllbb.png}
+\caption{\TODO{mbb}}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRllbb.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRllbb.png}
+\caption{\TODO{mbb}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_high_llbb_region}
+\end{figure}
+
 \subsection{Intermediate}
 \label{sec:2ljets_int}
-Table~\ref{tab:2ljets_int}.
+Table~\ref{tab:2ljets_int}
+Figure~\ref{fig:2ljets_int_region}
+Figure~\ref{fig:2ljets_int_cr_region}
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRInt.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRInt.png}
+\caption{\TODO{mjj}}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRInt.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRInt.png}
+\caption{\TODO{mjj}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_int_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZZ.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
+\caption{}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_int_cr_region}
+\end{figure}
+
 \subsection{Low}
 \label{sec:2ljets_low}
-
 Table~\ref{tab:2ljets_low}
+Figure~\ref{fig:2ljets_low_region}
+Figure~\ref{fig:2ljets_low_cr_region}
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRLow.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRLow2.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow2.png}
+\caption{}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_low_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZ.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZ.png}
+\caption{\TODO{mjj}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+\TODO{version with sr filled in, conspicuous dip at W mass}
+}
+\label{fig:2ljets_low_cr_region}
+\end{figure}
+
 \subsection{Off-shell}
 \label{sec:2ljets_offshell}
-
 Table~\ref{tab:2ljets_offshell}
+Figure~\ref{fig:2ljets_offshell_region}
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SROffShell.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SROffShell.png}
+\caption{\TODO{mll}}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VROffShell.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VROffShell.png}
+\caption{\TODO{mll}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_offshell_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
+\caption{\TODO{mll}}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_offshell_cr_region}
+\end{figure}
+
 \subsection{Discovery}

### 2022-06-01 main/chapter_2ljets.tex
add plots in other (non-metsig) variables

@@
 
+\subsubsection{High}
+
 \begin{figure}[tp]
@@
 
+\subsubsection{Mono-jet}
+
 \begin{figure}[tp]
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh4.png}
-\caption{\TODO{mj1}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjetone_SRHigh4.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh4.png}
-\caption{\TODO{mj1}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjetone_VRHigh4.png}
+\caption{}
 \end{subfigure}
@@
 
+\subsubsection{\texorpdfstring{$\ell\ell bb$}{\textit{llbb}}}
+
 \begin{figure}[tp]
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRllbb.png}
-\caption{\TODO{mbb}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mbb_SRllbb.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRllbb.png}
-\caption{\TODO{mbb}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mbb_VRllbb.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRInt.png}
-\caption{\TODO{mjj}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRInt.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRInt.png}
-\caption{\TODO{mjj}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRInt.png}
+\caption{}
 \end{subfigure}
@@
 \centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZZ.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRtt.png}
 \caption{}
@@
 Figure~\ref{fig:2ljets_low_region}
+Figure~\ref{fig:2ljets_low2_region}
 Figure~\ref{fig:2ljets_low_cr_region}
@@
 \centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRLow.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRLow.png}
+\caption{}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_low_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\centering
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRLow2.png}
@@
 \end{subfigure}
-\\[.5em]
-\begin{subfigure}{0.45\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow.png}
-\caption{}
-\end{subfigure}
-\hfill
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRLow2.png}
+\caption{}
+\end{subfigure}
+\\[.5em]
 \begin{subfigure}{0.45\textwidth}
@@
 \end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRLow2.png}
+\caption{}
+\end{subfigure}
 \caption{%
@@
 }
-\label{fig:2ljets_low_region}
+\label{fig:2ljets_low2_region}
 \end{figure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZ.png}
-\caption{\TODO{mjj}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZ.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SROffShell.png}
-\caption{\TODO{mll}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mll_SROffShell.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VROffShell.png}
-\caption{\TODO{mll}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mll_VROffShell.png}
+\caption{}
 \end{subfigure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
-\caption{\TODO{mll}}
+\includegraphics[width=\textwidth]{figures/2ljets_def_mll_CRDY.png}
+\caption{}
 \end{subfigure}

### 2022-06-01 main/chapter_2ljets.tex
add srlow crz union figures

@@
 Figure~\ref{fig:2ljets_low_cr_region}
+Figure~\ref{fig:2ljets_low_sr_or_cr_region}
+Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
 
@@
 \TODO{caption}
-\TODO{version with sr filled in, conspicuous dip at W mass}
 }
 \label{fig:2ljets_low_cr_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.9\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ.pdf}
+\caption{}
+\end{subfigure}
+\caption{%
+The union of CR-Z and SR-Low, which exactly fills its gap at
+$\mjj \in 60\textrm{--}110\,\eV[G]$.
+Data have a substantial dip in exactly that mass window.
+}
+\label{fig:2ljets_low_sr_or_cr_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.9\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ_mg5.pdf}
+\caption{}
+\end{subfigure}
+\caption{%
+The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
+MadGraph.
+This sample has reduced statistical precision, decreased yields in the signal
+region, and increased yields (which match data better) in the high-$\mjj$
+overflow bin.
+}
+\label{fig:2ljets_low_sr_or_cr_region_alt}
 \end{figure}

### 2022-06-02 main/chapter_2ljets.tex
add short captions

@@
 \includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
-\caption{%
+\caption[
+Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
+backgrounds
+]{%
 Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
@@
 \end{subfigure}
-\caption{%
-Supersymmetric signal processes in the $\twoljets$-electroweak
-analysis.
+\caption[
+Supersymmetric signal processes in the $\twoljets$-electroweak analysis
+]{%
+Supersymmetric signal processes in the $\twoljets$-electroweak analysis.
 \emph{Figures are reproduced from the paper%
@@
 \end{subfigure}
-\caption{%
+\caption[
+Example signal region histograms with benchmark signal sample yields overlaid
+]{%
 Example signal region histograms with benchmark signal sample yields overlaid
@@
 \includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
-\caption{%
+\caption[
+Contours for the C1N2 model in the $\twoljets$-electroweak analysis
+]{%
 Contours for the C1N2 model in the $\twoljets$-electroweak
@@
 \includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
-\caption{%
+\caption[
+Contours for the GMSB model in the $\twoljets$-electroweak analysis
+]{%
 Contours for the GMSB model in the $\twoljets$-electroweak
@@
 \end{tabular*}
-\caption{%
+\caption[
+Upper limit results in $\twoljets$-electroweak discovery regions
+]{%
 Upper limit results in $\twoljets$-electroweak discovery
@@
 \includegraphics[width=0.9\textwidth]{figures/2ljets_rjr_trees.pdf}
-\caption{%
+\caption[
+Decay trees for the $\twoljets$-RJR analysis
+]{%
 Decay trees for the $\twoljets$-RJR analysis.
@@
 \end{subfigure}
-\caption{%
+\caption[
+Summary plots for RJR analyses
+]{%
 Summary plots for RJR analyses:
@@
 \end{subfigure}
-\caption{%
+\caption[
+Supersymmetric signal processes in the $\twoljets$-strong search
+]{%
 Supersymmetric signal processes in the $\twoljets$-strong search.
@@
 \end{subfigure}
-\caption{%
+\caption[
+Contours for signal models of the $\twoljets$-strong search
+]{%
 Contours for signal models of the $\twoljets$-strong search.
@@
 \end{subfigure}
-\caption{%
+\caption[
+Figures reproduced from the partial Run~2 search with $2/3\ell$ selections
+]{%
 Figures reproduced from the partial Run~2 search with $2/3\ell$
@@
 \includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
-\caption{
-Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
-example signal models.
+\caption[
+Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
+models
+]{
+Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
+models.
 At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
@@
 $\mttwoll > 80$.
-\caption{%
+\caption[
+High region definitions in the $\twoljets$-electroweak analysis
+]{%
 High region definitions in the $\twoljets$-electroweak analysis.
@@
 \end{subfigure}
-\caption{%
+\caption[
+SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$
+]{%
 SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$.
@@
 $\mttwoll > 80$.
-\caption{%
+\caption[
+Intermediate region definitions in the $\twoljets$-electroweak analysis
+]{%
 Intermediate region definitions in the $\twoljets$-electroweak analysis.
@@
 $\mll \in 81\textrm{--}101$.
-\caption{%
+\caption[
+Low region definitions in the $\twoljets$-electroweak analysis
+]{%
 Low region definitions in the $\twoljets$-electroweak analysis.
@@
 \end{subfigure}
-\caption{%
+\caption[
+The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
+MadGraph
+]{%
 The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
@@
 $\nbtag = 0$.
-\caption{%
+\caption[
+Off-shell region definitions in the $\twoljets$-electroweak analysis
+]{%
 Off-shell region definitions in the $\twoljets$-electroweak analysis.
@@
 \end{subfigure}
-\caption{%
+\caption[
+Evaluation of candidate High discovery regions
+]{%
 Evaluation of candidate High discovery regions.
@@
 \end{subfigure}
-\caption{%
-Evaluation of candidate High discovery regions.
+\caption[
+Evaluation of candidate $\ell\ell bb$ discovery regions
+]{%
+Evaluation of candidate $\ell\ell bb$ discovery regions.
 The region SR-$\ell\ell bb$, labelled `SRllbb' here, is chosen for
@@
 \end{subfigure}
-\caption{%
-Evaluation of candidate High discovery regions.
+\caption[
+Evaluation of candidate Intermediate discovery regions
+]{%
+Evaluation of candidate Intermediate discovery regions.
 The region SR-Int, labelled `Int\_1+\_2' here, is chosen for its simplicity
@@
 \end{subfigure}
-\caption{%
+\caption[
+Pre-fit yields in discovery regions
+]{%
 Pre-fit yields in discovery regions.

### 2022-06-02 main/chapter_2ljets.tex
make an llbb command

@@
 \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
-\caption{SR-$\ell\ell bb$}
+\caption{SR-$\llbb$}
 \end{subfigure}
@@
 DR-High & $3.9$ & $0.5$ & 0 & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50$ \\[.5ex]
-DR-$\ell\ell bb$ & $0.51$ & $0.20$ & 0 & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
+DR-$\llbb$ & $0.51$ & $0.20$ & 0 & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
 \end{tabular*}
@@
 \\[1em]
-SR-$\ell\ell bb$
+SR-$\llbb$
 & $\geq 2$
@@
 \\[.5em]
-\: VR-$\ell\ell bb$
+\: VR-$\llbb$
 & $\geq 2$
@@
 
-\subsubsection{\texorpdfstring{$\ell\ell bb$}{\textit{llbb}}}
+\subsubsection{\texorpdfstring{$\llbb$}{\textit{llbb}}}
 
@@
 \item DR-High: SR-High-8
-\item DR-$\ell\ell bb$: SR-$\ell\ell bb$
+\item DR-$\llbb$: SR-$\llbb$
 \item DR-Int: SR-Int
@@
 \caption[
-Evaluation of candidate $\ell\ell bb$ discovery regions
+Evaluation of candidate $\llbb$ discovery regions
 ]{%
-Evaluation of candidate $\ell\ell bb$ discovery regions.
-The region SR-$\ell\ell bb$, labelled `SRllbb' here, is chosen for
+Evaluation of candidate $\llbb$ discovery regions.
+The region SR-$\llbb$, labelled `SRllbb' here, is chosen for
 sensitivity, simplicity, and to avoid overlapping with DR-High.
@@
 \includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRllbb.png}
-\caption{DR-$\ell\ell bb$}
+\caption{DR-$\llbb$}
 \end{subfigure}

### 2022-06-02 main/chapter_2ljets.tex
add aux low plots; add pre post figure

@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
+Figure~\ref{fig:2ljets_low_pre_post}
 
@@
 \end{subfigure}
-\caption{%
+\caption[
+The union of CR-Z and SR-Low
+]{%
 The union of CR-Z and SR-Low, which exactly fills its gap at
@@
 \label{fig:2ljets_low_sr_or_cr_region_alt}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+]{%
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+(a) fitted to control regions, and (b) fitted to control and signal regions.
+The arrows point in to the signal region, which shows a large peak in the
+diboson background near $m(W)$.
+Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
+the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
+Pre-fit yields are preseented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
+and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
+}
+\label{fig:2ljets_low_pre_post}
 \end{figure}

### 2022-06-02 main/chapter_2ljets.tex
_X to _*
add crz extensions figure

@@
 & $\metsig$
-& $m_X$
+& $m_*$
 & $\rjj$
@@
 Figure~\ref{fig:2ljets_low_pre_post}
+Figure~\ref{fig:2ljets_low_crz_extensions}
 
@@
 the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
-Pre-fit yields are preseented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
+Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
 and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
@@
 \label{fig:2ljets_low_pre_post}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Et.png}
+\caption{}
+\end{subfigure}
+\caption[
+Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$
+]{%
+Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
+The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
+upper bound of $\met$.
+Systematic uncertainties are extrapolated from CR-Z to the extended tail.
+The tail overlaps with other regions, but shows good agreement with data.
+Sub-figure (b) is reproduced from auxiliary material to the
+paper~\cite{atlas2022searches}.
+Each last bin includes overflow.
+}
+\label{fig:2ljets_low_crz_extensions}
 \end{figure}

### 2022-06-02 main/chapter_2ljets.tex
add vrlow pre post fit

@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
-Figure~\ref{fig:2ljets_low_pre_post}
+Figure~\ref{fig:2ljets_low_crz_pre_post}
+Figure~\ref{fig:2ljets_low_vrlow_pre_post}
 Figure~\ref{fig:2ljets_low_crz_extensions}
@@
 }
-\label{fig:2ljets_low_pre_post}
+\label{fig:2ljets_low_crz_pre_post}
 \end{figure}
@@
 \centering
+\includegraphics[width=\textwidth]{figures/2ljets_prefit_VR_Low_EWK_met_Sign_no_sr.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_postfit_VR_Low_EWK_met_Sign.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+Comparison of VR-Low before and after fitting to the signal regions
+]{%
+Comparison of VR-Low (a) before and (b) after fitting to the signal regions.
+All systematic variations are included after fitting to the control regions.
+The signal region deficit pulls the $\zjets$ below data in the validation
+region, whose likelihood is not included in the \heplikelihood.
+Figure (b) is reproduced from paper~\cite{atlas2022searches}.
+Each last bin includes overflow.
+}
+\label{fig:2ljets_low_vrlow_pre_post}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
 \includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
@@
 \caption[
-Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$
+Extension from CR-Z into higher $\metsig$ and $\met$
 ]{%

### 2022-06-02 main/chapter_2ljets.tex
move some low plots to relevant sections

@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region}
-Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
-Figure~\ref{fig:2ljets_low_crz_pre_post}
-Figure~\ref{fig:2ljets_low_vrlow_pre_post}
-Figure~\ref{fig:2ljets_low_crz_extensions}
 
@@
 
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.9\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ_mg5.pdf}
-\caption{}
-\end{subfigure}
-\caption[
-The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
-MadGraph
-]{%
-The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
-MadGraph.
-This sample has reduced statistical precision, decreased yields in the signal
-region, and increased yields (which match data better) in the high-$\mjj$
-overflow bin.
-}
-\label{fig:2ljets_low_sr_or_cr_region_alt}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
-\caption{}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
-\caption{}
-\end{subfigure}
-\caption[
-Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
-]{%
-Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
-(a) fitted to control regions, and (b) fitted to control and signal regions.
-The arrows point in to the signal region, which shows a large peak in the
-diboson background near $m(W)$.
-Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
-the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
-Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
-and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
-}
-\label{fig:2ljets_low_crz_pre_post}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_prefit_VR_Low_EWK_met_Sign_no_sr.png}
-\caption{}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_postfit_VR_Low_EWK_met_Sign.pdf}
-\caption{}
-\end{subfigure}
-\caption[
-Comparison of VR-Low before and after fitting to the signal regions
-]{%
-Comparison of VR-Low (a) before and (b) after fitting to the signal regions.
-All systematic variations are included after fitting to the control regions.
-The signal region deficit pulls the $\zjets$ below data in the validation
-region, whose likelihood is not included in the \heplikelihood.
-Figure (b) is reproduced from paper~\cite{atlas2022searches}.
-Each last bin includes overflow.
-}
-\label{fig:2ljets_low_vrlow_pre_post}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
-\caption{}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.49\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Et.png}
-\caption{}
-\end{subfigure}
-\caption[
-Extension from CR-Z into higher $\metsig$ and $\met$
-]{%
-Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
-The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
-upper bound of $\met$.
-Systematic uncertainties are extrapolated from CR-Z to the extended tail.
-The tail overlaps with other regions, but shows good agreement with data.
-Sub-figure (b) is reproduced from auxiliary material to the
-paper~\cite{atlas2022searches}.
-Each last bin includes overflow.
-}
-\label{fig:2ljets_low_crz_extensions}
-\end{figure}
-
 \subsection{Off-shell}
@@
 
-\section{Systematics}
+Figure~\ref{fig:2ljets_low_crz_extensions}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Et.png}
+\caption{}
+\end{subfigure}
+\caption[
+Extension from CR-Z into higher $\metsig$ and $\met$
+]{%
+Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
+The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
+upper bound of $\met$.
+Systematic uncertainties are extrapolated from CR-Z to the extended tail.
+The tail overlaps with other regions, but shows good agreement with data.
+Sub-figure (b) is reproduced from auxiliary material to the
+paper~\cite{atlas2022searches}.
+Each last bin includes overflow.
+}
+\label{fig:2ljets_low_crz_extensions}
+\end{figure}
+
+\section{Variations}
+% "systematic"
+
+Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
+Figure~\ref{fig:2ljets_low_crz_pre_post}
+Figure~\ref{fig:2ljets_low_vrlow_pre_post}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.9\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ_mg5.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
+MadGraph
+]{%
+The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
+MadGraph.
+This sample has reduced statistical precision, decreased yields in the signal
+region, and increased yields (which match data better) in the high-$\mjj$
+overflow bin.
+}
+\label{fig:2ljets_low_sr_or_cr_region_alt}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+]{%
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+(a) fitted to control regions, and (b) fitted to control and signal regions.
+The arrows point in to the signal region, which shows a large peak in the
+diboson background near $m(W)$.
+Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
+the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
+Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
+and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
+}
+\label{fig:2ljets_low_crz_pre_post}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_prefit_VR_Low_EWK_met_Sign_no_sr.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_postfit_VR_Low_EWK_met_Sign.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+Comparison of VR-Low before and after fitting to the signal regions
+]{%
+Comparison of VR-Low (a) before and (b) after fitting to the signal regions.
+All systematic variations are included after fitting to the control regions.
+The signal region deficit pulls the $\zjets$ below data in the validation
+region, whose likelihood is not included in the \heplikelihood.
+Figure (b) is reproduced from paper~\cite{atlas2022searches}.
+Each last bin includes overflow.
+}
+\label{fig:2ljets_low_vrlow_pre_post}
+\end{figure}
 

### 2022-06-02 main/chapter_2ljets.tex
add vrss figure

@@
 
+Figure~\ref{fig:2ljets_vrss}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWK_met_Sign.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWKMC_met_Sign.png}
+\caption{}
+\end{subfigure}
+\caption[
+Same-sign validation with Matrix Method and Monte Carlo fake estimates
+]{%
+Same-sign validation with (a) Matrix Method and (b) Monte Carlo fake estimates.
+Errors are statistical summed in quadrature with variations in Matrix Method
+sample weights.
+Those Matrix Method variations are coupled between bins, meaning that the
+directed differences between bins are less significant than an independent
+combination.
+% thesis_project/2ljets/vrss_plots/PAM_FNP_50pc.pdf
+The figure in (a) was included in an early draft of the paper as evidence to
+the quality of the Matrix Method fake/non-prompt estimate.
+On the decision of our editorial board, it was removed after validation studies
+showed that adding an additional $50\%$ uncertainty to fake/non-prompt yields
+made an insignificant difference to the results.
+This variation was coupled between all bins.
+Differences were assessed by the resulting change in exclusion contours,
+which mostly moved by less than their plotted line width.
+}
+\label{fig:2ljets_vrss}
+\end{figure}
+
 \section{Validation}

### 2022-06-02 main/chapter_2ljets.tex
standardize 0.5em; improve vrss caption

--- 2022-06-02 main/chapter_2ljets.tex
@@
 $n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$
-\\[.5em]
+\\[0.5em]
 same flavour, opposite charge leptons
-\\[.5em]
+\\[0.5em]
 $\njet \geq 1$
-\\[.5em]
+\\[0.5em]
 $p_\mathrm{T}^{\,\ell_2} > 25$
-\\[.5em]
+\\[0.5em]
 $\met > 100$
-\\[.5em]
+\\[0.5em]
 $\metsig > 6$
@@
 & $0\textrm{--}0.8\textrm{--}1.6$
-\\[.5em]
+\\[0.5em]
 \: VR-High
@@
 & $< 1.6$
-\\[.5em]
+\\[0.5em]
 \: VR-High-R
@@
 &
-\\[.5em]
+\\[0.5em]
 \: VR-High-1J
@@
 &
-\\[.5em]
+\\[0.5em]
 \: VR-$\llbb$
@@
 \end{subfigure}
-\\[.5em]
+\\[0.5em]
 \begin{subfigure}{0.45\textwidth}
@@
 & $> 60$
-\\[.5em]
+\\[0.5em]
 \: VR-Int
@@
 & $\uwave{\hphantom{< 60}}$
-\\[.5em]
+\\[0.5em]
 CR-tt
@@
 &
-\\[.5em]
+\\[0.5em]
 \: VR-Low
@@
 & $< 0.6$
-\\[.5em]
+\\[0.5em]
 \: VR-Low-2
@@
 & $>2$
-\\[.5em]
+\\[0.5em]
 \: VR-OffShell
@@
 \end{subfigure}
-\\[.5em]
+\\[0.5em]
 \begin{subfigure}{0.4\textwidth}
@@
 \label{sec:2ljets_mm_fakes}
+
+Figure~\ref{fig:2ljets_vrss}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWK_met_Sign.png}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWKMC_met_Sign.png}
+\caption{}
+\end{subfigure}
+\caption[
+Same-sign validation with Matrix Method and Monte Carlo fake estimates
+]{%
+Same-sign validation with (a) Matrix Method and (b) Monte Carlo fake estimates.
+Backgrounds are pre-fit.
+Errors are statistical summed in quadrature with variations in Matrix Method
+sample weights.
+Those weight variations are coupled between bins, meaning that the
+directed differences between bins are less significant than an independent
+combination.
+The Monte Carlo estimate is clearly more accurate.
+\\[0.5em]
+% thesis_project/2ljets/vrss_plots/PAM_FNP_50pc.pdf
+Part (a) was included in an early draft of the paper.
+It was removed after validation studies showed that adding an additional
+$50\%$ uncertainty to fake/non-prompt yields, coupled between all bins, did not
+significantly change our exclusion contours;
+most contours shrank by less than their plotted line width.
+The largest change was a $2\%$ reduction in the
+$B(\neutralino_1 \rightarrow h \gravitino)$
+limit at $m(\neutralino_1) \approx 500\,\eV[G]$.
+}
+\label{fig:2ljets_vrss}
+\end{figure}
 

### 2022-06-02 main/chapter_2ljets.tex
add vrlow zjets alt figure

@@
 Figure~\ref{fig:2ljets_low_crz_pre_post}
+Figure~\ref{fig:2ljets_vrlow_alt}
 Figure~\ref{fig:2ljets_low_vrlow_pre_post}
@@
 \centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow.pdf}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow_mg5.pdf}
+\caption{}
+\end{subfigure}
+\caption[
+Pre-fit yields in VR-Low with standard and alternative $\zjets$ simulations
+]{%
+Pre-fit yields in VR-Low with (a) standard and (b) alternative $\zjets$
+simulations.
+Errors are statistical.
+}
+\label{fig:2ljets_vrlow_alt}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
 \includegraphics[width=\textwidth]{figures/2ljets_prefit_VR_Low_EWK_met_Sign_no_sr.png}

### 2022-06-02 main/chapter_2ljets.tex
add VR-Low-ISR figure; standardize figure widths

@@
 \hfill
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \centering
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \hfill
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.49\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \centering
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.48\textwidth}
 \centering
@@
 \centering
-\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
-\hfill
-\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
+\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_logy.png}
+\hfill
+\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_sig_logy.png}
 \caption[
@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region}
+Figure~\ref{fig:2ljets_vrlow_old}
 
@@
 \label{fig:2ljets_low_sr_or_cr_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_srlowisr.png}
+\caption{SR-Low-ISR (blind)}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_vrlowisr.png}
+\caption{VR-Low-ISR}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_crlowisr.png}
+\caption{CR-Low-ISR}
+\end{subfigure}
+\caption{%
+\TODO{caption}
+}
+\label{fig:2ljets_vrlow_old}
 \end{figure}

### 2022-06-02 main/chapter_2ljets.tex
add split samples figure

@@
 Figure~\ref{fig:2ljets_low_crz_extensions}
+Figure~\ref{fig:2ljets_splits}
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZZ.png}
+\caption{CR-VZ, diboson}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZ.png}
+\caption{CR-Z, $\zjets$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRtt.png}
+\caption{CR-tt, rare top}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_SRllbb.png}
+\caption{SR-$\llbb$, rare top}
+\end{subfigure}
+\caption[
+Subcomponents of diboson, $\zjets$ and rare top processes.
+]{%
+Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) rare top processes.
+Target samples broken down by simulated sample identity codes.
+The `no filter' selection comprises $\zjets$ samples with unconstrained jet
+flavour.
+Errors are statistical.
+}
+\label{fig:2ljets_splits}
+\end{figure}
+
 \section{Variations}

### 2022-06-02 main/chapter_2ljets.tex
add summary slide figure

@@
 
+Figure~\ref{fig:2ljets_region_summary}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.99\textwidth]{figures/2Ljets_pam_ewkslide.png}
+\caption[
+Region design summary slide
+]{%
+Region design summary slide adapted from an internal presentation.
+Colours represent primary background contributions in our plotting scheme.
+}
+\label{fig:2ljets_region_summary}
+\end{figure}
+
 \begin{table}[tp]
@@
 Target samples broken down by simulated sample identity codes.
+The large three-lepton component in CR-VZ motivated its renaming from CR-ZZ.
 The `no filter' selection comprises $\zjets$ samples with unconstrained jet

### 2022-06-02 main/chapter_2ljets.tex
fill in some todo captions

@@
 Table~\ref{tab:2ljets_presel}
-\TODO{define presel, preselection SFOS 25 1j met>100 metsig>6}
+\TODO{define presel, preselection SFOS 25 1J met>100 metsig>6}
 
@@
 Figure~\ref{fig:2ljets_high_region}
-Figure~\ref{fig:2ljets_high_1j_region}
+Figure~\ref{fig:2ljets_high_1J_region}
 Figure~\ref{fig:2ljets_high_llbb_region}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_mll_SRHigh.png}
-\caption{}
+\caption{SR-High, $\mll$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_nbtag_SRHigh.png}
-\caption{}
+\caption{SR-High, $\nbtag$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh8.png}
-\caption{}
+\caption{SR-High-8, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh16.png}
-\caption{}
+\caption{SR-High-16, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh.png}
-\caption{}
+\caption{VR-High-8, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHighR.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{VR-High-16, $\metsig$}
+\end{subfigure}
+\caption[
+Pre-fit distributions of $\metsig$ in High signal and validation regions
+]{%
+Pre-fit distributions of $\metsig$ in High signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRHigh4.png}
-\caption{}
+\caption{SR-High-1J, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjetone_SRHigh4.png}
-\caption{}
+\caption{SR-High-1J, $\mjetone$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRHigh4.png}
-\caption{}
+\caption{VR-High-1J, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjetone_VRHigh4.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
-}
-\label{fig:2ljets_high_1j_region}
+\caption{SR-High-1J, $\mjetone$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in mono-jet signal and validation regions
+]{%
+Pre-fit distributions in mono-jet signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
+}
+\label{fig:2ljets_high_1J_region}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRllbb.png}
-\caption{}
+\caption{SR-$\llbb$, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mbb_SRllbb.png}
-\caption{}
+\caption{SR-$\llbb$, $\mbb$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRllbb.png}
-\caption{}
+\caption{VR-$\llbb$, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mbb_VRllbb.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{VR-$\llbb$, $\mbb$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in $\llbb$ signal and validation regions
+]{%
+Pre-fit distributions in $\llbb$ signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRInt.png}
-\caption{}
+\caption{SR-Int, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRInt.png}
-\caption{}
+\caption{SR-Int, $\mjj$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRInt.png}
-\caption{}
+\caption{VR-Int, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRInt.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{VR-Int, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in Intermediate signal and validation regions
+]{%
+Pre-fit distributions in Intermediate signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZZ.png}
-\caption{}
+\caption{CR-VZ, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZZ.png}
-\caption{}
+\caption{CR-VZ, $\mjj$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
-\caption{}
+\caption{CR-tt, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRtt.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{CR-tt, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in the control regions CR-tt and CR-VZ
+]{%
+Pre-fit distributions in the control regions CR-tt and CR-VZ.
+Signal regions are unblinded.
+Errors are statistical.
 }

### 2022-06-02 main/chapter_2ljets.tex
add split background plots; add more missing captions

@@
 Pre-fit distributions in the control regions CR-tt and CR-VZ.
-Signal regions are unblinded.
 Errors are statistical.
@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region}
-Figure~\ref{fig:2ljets_vrlow_old}
+
+\subsubsection{Low-ISR}
+\label{sec:2ljets_low_isr}
+Figure~\ref{fig:2ljets_low_isr_old}
 
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRLow.png}
-\caption{}
+\caption{SR-Low, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRLow.png}
-\caption{}
+\caption{SR-Low, $\mjj$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow.png}
-\caption{}
+\caption{VR-Low, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRLow.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{VR-Low, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in Low signal and validation regions
+]{%
+Pre-fit distributions in Low signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRLow2.png}
-\caption{}
+\caption{SR-Low-2, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_SRLow2.png}
-\caption{}
+\caption{SR-Low-2, $\mjj$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VRLow2.png}
-\caption{}
+\caption{VR-Low-2, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRLow2.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{VR-Low-2, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in Low-2 signal and validation regions
+]{%
+Pre-fit distributions in Low-2 signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZ.png}
-\caption{}
+\caption{CR-Z, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZ.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+\caption{CR-Z, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in the control region CR-Z
+]{%
+Pre-fit distributions in the control region CR-Z.
+Errors are statistical.
 }
@@
 Data have a substantial dip in exactly that mass window.
+Errors are statistical.
 }
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_srlowisr.png}
-\caption{SR-Low-ISR (blind)}
+\caption{SR-Low-ISR, $\ptisr$ (blind)}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_vrlowisr.png}
-\caption{VR-Low-ISR}
+\caption{VR-Low-ISR, $\mwisr$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_crlowisr.png}
-\caption{CR-Low-ISR}
-\end{subfigure}
-\caption{%
-\TODO{caption}
-}
-\label{fig:2ljets_vrlow_old}
+\caption{CR-Low-ISR, $\mwisr$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in draft Low-ISR regions
+]{%
+Pre-fit distributions in draft Low-ISR regions which were removed and never
+unblinded.
+These regions required a third, hard jet directed away from $\ptmiss$.
+As seen from (b) and (c), the background modelling is poor and in conflict
+with normalization factors in other control regions.
+Variables are defined in the text of Section~\ref{sec:2ljets_low_isr}.
+Errors are statistical.
+}
+\label{fig:2ljets_low_isr_old}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SROffShell.png}
-\caption{}
+\caption{SR-OffShell, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mll_SROffShell.png}
-\caption{}
+\caption{SR-OffShell, $\mll$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_VROffShell.png}
-\caption{}
+\caption{VR-OffShell, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mll_VROffShell.png}
-\caption{}
+\caption{VR-OffShell, $\mll$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in OffShell signal and validation regions
+]{%
+Pre-fit distributions in OffShell signal and validation regions.
+Signal regions are unblinded.
+Errors are statistical.
+}
+\label{fig:2ljets_offshell_region}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
+\caption{CR-DY, $\metsig$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mll_CRDY.png}
+\caption{CR-DY, $\mll$}
 \end{subfigure}
 \caption{%
-\TODO{caption}
-}
-\label{fig:2ljets_offshell_region}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
-\caption{}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_mll_CRDY.png}
-\caption{}
-\end{subfigure}
-\caption{%
-\TODO{caption}
+Pre-fit distributions in the control region CR-DY.
+Errors are statistical.
 }

### 2022-06-06 main/chapter_2ljets.tex
add illustrative presel plots; minor tidyup

@@
 \begin{tabular}{c}
-Preselection:
+Pre-selection:
 \\[1em]
@@
 \caption{%
-`Preselection' requirements ensured prior to all region definitions in the
+`Pre-selection' requirements ensured prior to all region definitions in the
 $\twoljets$-electroweak search.
@@
 \\[1em]
-Common: Preselection,
+Common: Pre-selection,
 $\mll \in 71\textrm{--}111$, and
@@
 \\[1em]
-Common: Preselection,
+Common: Pre-selection,
 $\njet \geq 2$,
@@
 \\[1em]
-Common: Preselection,
+Common: Pre-selection,
 $\njet = 2$,
@@
 \\[1em]
-Common: Preselection,
+Common: Pre-selection,
 $\njet \geq 2$, and
@@
 \centering
-\begin{subfigure}{0.9\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ_mg5.pdf}
-\caption{}
-\end{subfigure}
+\includegraphics[width=0.9\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ_mg5.pdf}
 \caption[
@@
 \includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
-\caption{}
+\caption{Fitted to control regions only}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
-\caption{}
+\caption{Fitted to control and signal regions}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow.pdf}
-\caption{}
+\caption{Standard}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow_mg5.pdf}
-\caption{}
+\caption{Alternative}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_prefit_VR_Low_EWK_met_Sign_no_sr.png}
-\caption{}
+\caption{Fitted to control regions only}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_postfit_VR_Low_EWK_met_Sign.pdf}
-\caption{}
+\caption{Fitted to control and signal regions}
 \end{subfigure}
@@
 All systematic variations are included after fitting to the control regions.
-The signal region deficit pulls the $\zjets$ below data in the validation
+The signal region deficit pulls $\zjets$ yields below data in the validation
 region, whose likelihood is not included in the \heplikelihood.
-Figure (b) is reproduced from paper~\cite{atlas2022searches}.
-Each last bin includes overflow.
 }
@@
 
+\section{Unblinding}
+
+Figure~\ref{fig:2ljets_prepre}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_prepre_met_Sign_VRpresel_2j_MCfakeslogY.pdf}
+\caption{$\metsig$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_prepre_mt2leplsp_0_VRpresel_2j_MCfakeslogY.pdf}
+\caption{$\mttwoll$}
+\end{subfigure}
+\caption[
+Histograms in a very loose pre-selection like those inspected to validate
+background modelling prior to designing the $\twoljets$ regions.
+]{%
+Histograms in a very loose pre-selection like those inspected to validate
+background modelling prior to designing the $\twoljets$ regions.
+This selection requires `Pre-selection' from Table~\ref{tab:2ljets_presel} and
+$\njet \geq 2$.
+Errors are statistical.
+Each last bin contains overflow.
+}
+\label{fig:2ljets_prepre}
+\end{figure}
+
 \section{Results}
-
-\section{Unblinding}

### 2022-06-06 main/chapter_2ljets.tex
add summary tables from unblinding and no_sr

@@
 Figure~\ref{fig:2ljets_prepre}
+Figure~\ref{fig:2ljets_summary_unblinding}
 
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_unblinding_log.png}
+\caption[
+First unblinded summary table from 28th September 2020
+]{%
+First unblinded summary table from 28th September 2020.
+Backgrounds are fitted to data in control and signal regions.
+Errors include all variations that were evaluated at the time.
+Data and backgrounds were later reproduced with updated software and tighter
+lepton definitions, leading to a general reduction in yields.
+}
+\label{fig:2ljets_summary_unblinding}
+\end{figure}
+
 \section{Results}
+
+Figure~\ref{fig:2ljets_summary_no_sr}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.pdf}
+\caption[
+Alternative summary table with backgrounds fitted to control regions only
+]{%
+Alternative summary table with backgrounds fitted to control regions only.
+A version fitted also to signal regions is shown in
+Figure~\ref{fig:2ljets_summary}.
+Fitting to signal regions is policy of the \atlas\ SUSY group, but deviates
+from the intentions of the significance measure.
+Because the post-fit yields in SR-Low achieve a small negative significance,
+all mentions of its deficit were removed from the paper.
+}
+\label{fig:2ljets_summary_no_sr}
+\end{figure}

### 2022-06-06 main/chapter_2ljets.tex
remove atlas internal tag from no_sr summary plot
add fit corrmatrix

@@
 Figure~\ref{fig:2ljets_summary_no_sr}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.pdf}
+Figure~\ref{fig:2ljets_fit_corrmatrix}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.png}
 \caption[
@@
 \end{figure}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_corrmatrix.pdf}
+\caption[
+Correlation matrix from the background-only fit to control and signal regions
+]{%
+Correlation matrix from the background-only fit to control and signal regions,
+as estimated by HistFitter.
+Parameters with smaller levels of correlation are excluded.
+}
+\label{fig:2ljets_fit_corrmatrix}
+\end{figure}

### 2022-06-06 main/chapter_2ljets.tex
add parameter pull plot

@@
 Figure~\ref{fig:2ljets_fit_corrmatrix}
+Figure~\ref{fig:2ljets_fit_parameters}
 
@@
 \end{figure}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[height=0.9\textheight]{figures/2ljets_fit_parameters.pdf}
+\caption[
+Parameter values and errors from the background only-fit to control and signal
+regions.
+]{%
+Parameter values and errors from the background only-fit to control and signal
+regions.
+Those named $\mu\_*$ are unconstrained normalization factors,
+those named $\gamma\_*$ describe the combined statistical errors of simulated
+samples,
+and those named $\alpha\_*$ parametrize other variations.
+The red arrow highlights the $\zjets$ generator comparison parameter
+$\alpha\mathrm{\_Zjets\_alt}$ whose post-fit value is
+% 1.8598e+00 +/-  6.87e-01
+$1.86 \pm 0.69$;
+since it exceeds $1.5$, it was otherwise excluded from this plot.
+}
+\label{fig:2ljets_fit_parameters}
+\end{figure}

### 2022-06-06 main/chapter_2ljets.tex
add multi contours figure

@@
 Figure~\ref{fig:2ljets_fit_parameters}
+Figure~\ref{fig:2ljets_fit_muti_contourss}
 
@@
 \end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_c1n2.pdf}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_gmsb.pdf}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Projected exclusion from sets of orthogonal signal regions
+]{%
+Projected exclusion from sets of orthogonal signal regions, demonstrating that
+their combination has greater sensitivity reach than any individually.
+The label `High' indicates the bins of SR-High, and `HighNo4' excludes
+SR-High-1J to show the locations where mono-jet events have an impact.
+Similarly, `$\llbb$', `Int', `Low', and `OffShell' are shown to target their
+appropriate parameter spaces.
+Although figures were produced before unblinding, with older samples and region
+designs, the general message has not changed.
+Values of CLs from mean data in the combined fit are are overlaid on the grid.
+}
+\label{fig:2ljets_fit_muti_contourss}
+\end{figure}

### 2022-06-06 main/chapter_2ljets.tex
add figure and some text for signal theory sys

@@
 % "systematic"
-
 Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
@@
 \label{fig:2ljets_low_vrlow_pre_post}
+\end{figure}
+
+% extrapolation
+
+The total theory variation is a combination of seven components representing
+\TODO{what are they?}.
+A relative uncertainty due to each, $r$, is evaluated at each benchmark
+point $\vec y_i$, and extrapolated elsewhere as an exponentially weighted
+combination
+\begin{equation}
+\label{eqn:2ljets_signal_extrap}
+r(\vec x) = \frac{
+\sum_i r_i e^{-d(\vec x, \vec y_i)}
+}{
+\sum_i e^{-d(\vec x, \vec y_i)}
+}
+\end{equation}
+in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
+physical understanding into the extrapolation.
+
+Parent mass $m(\chargino_1, \neutralino_2) = m_a$
+and mass splitting $m_a - m_b$ are the most
+characteristic features in C1N2.
+From some experimentation, I found that a Manhattan distance in those features
+extrapolated nicely.
+Since the distance is dimensionful, it needs to be scaled to a reference value,
+which I chose as $100\,\eV[G]$; therefore
+\begin{equation}
+d_\mathrm{C1N2}(\vec x, \vec y) =
+\frac{1}{100\,\eV[G]}
+\big(
+|m_a^{\vec x} - m_b^{\vec y}|
++
+|(m_a^{\vec x} - m_b^{\vec x}) - (m_a^{\vec y} - m_b^{\vec y})|
+\big)
+.
+\end{equation}
+To put the Higgs branching fraction $B$ on a similar scale to the parent mass
+for GMSB models, its distance measure has a separate scale factor;
+\begin{equation}
+d_\mathrm{GMSB}(\vec x, \vec y) =
+\frac{1}{100\,\eV[G]}
+|m_a^{\vec x} - m_b^{\vec y}|
++
+5
+|B^{\vec x} - B^{\vec y}|
+.
+\end{equation}
+Some example extrapolations according to these rules are displayed in
+Figure~\ref{fig:2ljets_signal_sys_extrap}.
+Variations tend to be small, around $\approx 20\%$, where and only where
+yields are large enough for good statistical precision.
+
+Relative uncertainty $r$ is evaluated for each component, and the total
+uncertainty is their sum in quadrature, capped at $100\%$.
+This is done independently in each selection region without considering the
+directions of coherent variations; all variations are combined as described
+and varied together a single parameter in the fit.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SRHigh4.pdf}
+\caption{C1N2, SR-High-1J}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SROffShell_1.pdf}
+\caption{C1N2, SR-OffShell\_a}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRHigh8_2.pdf}
+\caption{GMSB, SR-High-1J}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRllbb.pdf}
+\caption{GMSB, SR-OffShell\_a}
+\end{subfigure}
+\caption[
+Example extrapolated signal theoretical uncertainties
+]{%
+Example extrapolated signal theoretical uncertainties.
+Variations are evaluated for signals at the marked points and extrapolated to
+the each parameter plane according to Equation~\ref{eqn:2ljets_signal_extrap}.
+To account for its drastically different kinematics, the C1N2 off-shell space
+is treated independently.
+}
+\label{fig:2ljets_signal_sys_extrap}
 \end{figure}

### 2022-06-07 main/chapter_2ljets.tex
add fnp estiamte tables

@@
 Figure~\ref{fig:2ljets_vrss}
+Table~\ref{tab:2ljets_fnp_negative_mm}
+Table~\ref{tab:2ljets_fnp_estimates}
 
@@
 \end{figure}
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{lccc}
+& MC Prompt (stat) & MC FNP (stat) & MM FNP (stat, sys)
+\\[0.5em]
+CR-Z & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
+\\
+VR-Int & $18.7 \pm 0.9$ & $0.2 \pm 0.2$ & $-0.3 \pm 0.1 \pm 0.3$
+\\
+VR-High & $19.7 \pm 0.7$ & $0.04 \pm 0.04$ & $-0.2 \pm 0.2 \pm 0.5$
+\\
+VR-High-1J & $51.9 \pm 1.9$ & $0.003 \pm 0.002$ & $-1.9 \pm 0.2 \pm 0.4$
+\\
+SR-Low-b & $15.4 \pm 1.0$ & $0.001 \pm 0.000$ & $-0.08 \pm 0.05 \pm 0.06$
+\\
+SR-Int-a & $22.3 \pm 0.8$ & $0 \pm 0$ & $-0.4 \pm 0.2 \pm 0.6$
+\\
+SR-Int-b & $11.4 \pm 0.5$ & $0.002 \pm 0.002$ & $-0.3 \pm 0.1 \pm 0.3$
+\\
+SR-High-8-a & $2.4 \pm 0.2$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
+\\
+SR-High-8-b & $2.4 \pm 0.3$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
+\\
+SR-High-1J & $1.0 \pm 0.1$ & $0.004 \pm 0.003$ & $0 \pm 0 \pm 0$
+\\
+SR-$\llbb$ & $0.7 \pm 0.1$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
+\\
+\end{tabular}
+\caption[%
+Background yields from prompt and fake/non-prompt sources in bins with
+non-positive Matrix Method estimates
+]{%
+Background yields from prompt and fake/non-prompt sources in bins with
+non-positive Matrix Method estimates.
+The `MC FNP' yields are selected using truth information in sumulated samples.
+The `MM FNP' yields are from the Matrix Method, with statistical errors from
+the sample weights and systematic variations from the larger absolute change
+from variations of those weights.
+}
+\label{tab:2ljets_fnp_negative_mm}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{lcccccc}
+& Prompt (stat)
+& mMC
+& mMM
+& mMean
+& \quad\underline{mMax}\hphantom{foo}
+\\[0.5em]
+CR-Z & $143.6 \pm 4.6$
+& $0.1^{+0.2}_{-0.1}$
+& $0.01^{+3.2}_{-0.0}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.1^{+3.2}_{-0.1}$
+\\[0.2em]
+VR-Int & $18.7 \pm 0.9$
+& $0.2^{+0.4}_{-0.2}$
+& $0.01^{+0.46}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.2^{+0.5}_{-0.2}$
+\\[0.2em]
+VR-High & $19.7 \pm 0.7$
+& $0.04^{+0.08}_{-0.04}$
+& $0.01^{+0.73}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.04^{+0.73}_{-0.04}$
+\\[0.2em]
+VR-High-1J & $51.9 \pm 1.9$
+& $0.003^{+0.005}_{-0.003}$
+& $0.01^{+0.59}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.59}_{-0.02}$
+\\[0.2em]
+SR-Low-b & $15.4 \pm 1.0$
+& $0.001^{+0.001}_{-0.001}$
+& $0.01^{+0.11}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.17}_{-0.02}$
+\\[0.2em]
+SR-Int-a & $22.3 \pm 0.8$
+& $0^{+0}_{-0}$
+& $0.01^{+0.72}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.72}_{-0.02}$
+\\[0.2em]
+SR-Int-b & $11.4 \pm 0.5$
+& $0.002^{+0.003}_{-0.002}$
+& $0.01^{+0.43}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.43}_{-0.02}$
+\\[0.2em]
+SR-High-8-a & $2.4 \pm 0.2$
+& $0^{+0}_{-0}$
+& $0.01^{+0}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.14}_{-0.02}$
+\\[0.2em]
+SR-High-8-b & $2.4 \pm 0.3$
+& $0^{+0}_{-0}$
+& $0.01^{+0}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.14}_{-0.02}$
+\\[0.2em]
+SR-High-1J & $1.0 \pm 0.1$
+& $0.004^{+0.007}_{-0.004}$
+& $0.01^{+0}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.14}_{-0.02}$
+\\[0.2em]
+SR-$\llbb$ & $0.7 \pm 0.1$
+& $0^{+0}_{-0}$
+& $0.01^{+0}_{-0.01}$
+& $0.02^{+0.14}_{-0.02}$
+& $0.02^{+0.14}_{-0.02}$
+\\
+\end{tabular}
+\caption[%
+Replacement fake estimates with up and down variations
+]{%
+Replacement fake estimates with up and down variations.
+Those variations cover statistical and systematic sources.
+The method `mMax' is used for the $\twoljets$-electroweak analysis;
+as described in the text, it uses the largest values of central, statistical
+and positive systematic variations from other methods, and a down variation
+of zero.
+}
+\label{tab:2ljets_fnp_estimates}
+\end{table}
 
@@
 \includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SROffShell_1.pdf}
-\caption{C1N2, SR-OffShell\_a}
+\caption{C1N2, SR-OffShell-a}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRllbb.pdf}
-\caption{GMSB, SR-OffShell\_a}
+\caption{GMSB, SR-OffShell-a}
 \end{subfigure}
@@
 The red arrow highlights the $\zjets$ generator comparison parameter
-$\alpha\mathrm{\_Zjets\_alt}$ whose post-fit value is
+$\alpha\mathrm{\_Zjets-alt}$ whose post-fit value is
 % 1.8598e+00 +/-  6.87e-01

### 2022-06-07 main/chapter_2ljets.tex
add zjets alt table

@@
 % TODO generator filtering
+
+Table~\ref{tab:2ljets_zjets_alt}
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{lcc}
+& Central
+& Relative variations
+\\[0.5em]
+CR-DY
+& $59.64$
+& $1-0.1+0.1$
+\\[0.2em]
+\underline{CR-Z}
+& $101.95$
+& $1+0.2-0.2$
+\\[0.2em]
+CR-tt
+& $38.07$
+& $1+0.5-0.5$
+\\[0.2em]
+CR-VZ
+& $5.03$
+& $1+1.3-1.0$
+\\[0.5em]
+VR-Low
+& $24.19$
+& $1-0.3+0.3$
+\\[0.2em]
+VR-Low-2
+& $17.05$
+& $1+0.2-0.2$
+\\[0.5em]
+\underline{SR-Low-a}
+& $14.30$
+& $1-0.2+0.2$
+\\[0.2em]
+\underline{SR-Low-b}
+& $4.17$
+& $1-0.3+0.3$
+\\[0.2em]
+SR-Low-2
+& $1.51$
+& $1+1.5-1.0$
+\\[0.2em]
+\end{tabular}
+\caption[%
+Alternative $\zjets$ variations in selected regions
+]{%
+Alternative $\zjets$ variations in selected regions.
+The `Central' column shows the central $\zjets$ yield from the main sample.
+The other column shows the relative scale factors applied to this central yield
+for up and down variations.
+These are evaluated from the central and alternative samples as described in
+the text.
+Note how up and down variations can move in different directions.
+}
+\label{tab:2ljets_zjets_alt}
+\end{table}
 

### 2022-06-07 main/chapter_2ljets.tex
add first draft high region def text

@@
 % begin with physics content and results
-
 Unblinded results in all main regions of the $\twoljets$-electroweak analysis
@@
 
+\subsection{Jet pairs}
+\TODO{why leading two jets mjj / rjj}
+
 \subsection{Cheat-sheet}
@@
 \section{Design}
-The most impactful decision in a search is the design of its regions.
-Like all decision in life, these designs must be made to balance conflicting
-and vaguely-specified desires, and are informed with incomplete information.
-
-This section describes the designs and rationales behind the control, validation
-and signal regions of the $\twoljets$ analysis.
+Perhaps the most impactful decision in a search is the design of its regions.
+Like most decision, these designs must be made to balance conflicting and
+vaguely-specified desires, and are informed with incomplete information.
+
+This section describes the designs and rationales behind the control,
+validation and signal regions of the $\twoljets$ analysis.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.99\textwidth]{figures/2Ljets_pam_ewkslide.png}
+\caption[
+Region design summary slide
+]{%
+Region design summary slide adapted from an internal presentation.
+Colours represent primary background contributions in our plotting scheme.
+}
+\label{fig:2ljets_region_summary}
+\end{figure}
 
@@
 contributions for models with various mass splittings above $m(Z)$.
-But there is more --- $\metsig$ also decently separates signal models from one
-another;
+Furthermore, $\metsig$ also works to separate signal models from each other;
 signals with larger mass differences in their decays to invisibles tend to have
@@
 
-Control and validation regions are defined to constrain the background
-modelling in all regions.
-
-Table~\ref{tab:2ljets_presel}
-\TODO{define presel, preselection SFOS 25 1J met>100 metsig>6}
+Following the conventional style descibed in
+Section~\ref{sec:sec:searches_searches},
+control regions are defined to normalize specific background samples,
+and validation regions are defined to test their extrapolation towards
+signal regions before unblinding.
+
+All regions are defined within a loose `Pre-selection' requirement which
+requires the basic $Z\rightarrow \ell^+\ell^-$ plus jets pluts $\met$
+conditions that we target in the $\twoljets$-electroweak.
+These Pre-selection requirements are defined in Table~\ref{tab:2ljets_presel}.
 
@@
 
-Figure~\ref{fig:2ljets_region_summary}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.99\textwidth]{figures/2Ljets_pam_ewkslide.png}
-\caption[
-Region design summary slide
-]{%
-Region design summary slide adapted from an internal presentation.
-Colours represent primary background contributions in our plotting scheme.
-}
-\label{fig:2ljets_region_summary}
-\end{figure}
+A high-level summary of the region design is shown
+Figure~\ref{fig:2ljets_region_summary}.
+The design comprises four control regions and thirteen orthogonal signal region
+bins, with validation regions targeting each group of signal regions.
 
@@
 \begin{tabular}{c}
-Pre-selection:
+Pre-selction
 \\[1em]
@@
 \\[0.5em]
+$\metsig > 6$
+\\[0.5em]
 $\met > 100$
-\\[0.5em]
-$\metsig > 6$
 \end{tabular}
-\caption{%
-`Pre-selection' requirements ensured prior to all region definitions in the
+\caption[%
+Pre-selection requirements ensured prior to all region definitions in the
+$\twoljets$-electroweak search
+]{%
+Pre-selection requirements ensured prior to all region definitions in the
 $\twoljets$-electroweak search.
+Although $\metsig > 6$ implies that most events already have large $\met$
+$\met > 100\,\eV[G]$ requirement is included to ensure that events with small
+$\met$ and smaller uncertainties on it are excluded.
 }
@@
 \label{sec:2ljets_high}
-Table~\ref{tab:2ljets_high}
-Figure~\ref{fig:2ljets_high_mll_b}
-Figure~\ref{fig:2ljets_high_region}
-Figure~\ref{fig:2ljets_high_1J_region}
-Figure~\ref{fig:2ljets_high_llbb_region}
+The highest $\metsig$ selections, of around $\metsig > 18$, contain almost pure
+diboson backgrounds in relatively small quantities, along with the bulk masses
+of signal samples at the heaviest resonance masses.
+This is therefore the most promising territory for signal regions which are
+highly sensitive to the most extreme signals.
+Their backgrounds, however, must be constrained by control regions in less
+extreme parameter spaces with more data.
+
+Diboson backgrounds and C1N2 signals have more in common than just $\met$.
+Notably, both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
+and do not enhance the rate of heavy-flavour jets, as displayed in
+Figure~\ref{fig:2ljets_high_mll_b}.
+Since signals and backgrounds have such similar shapes in $\mll$ and $\nbtag$,
+sensitivity is only decreased by rejecting portions of these distributions.
+Although most $\twoljets$-electroweak regions use tighter selection on these
+variables to reduce backgrounds, this observation allows us to win
+additional sensitivity in High regions by loose requirements.
+
+The $\mll \in 71\textrm{--}111$ window loosely selects the $m(Z)$ peak to
+scoop up nearby rare of mismeasured events.
+Similarly, allowing up to one $b$-tag improves yields without boosting
+backgrounds since the $\metsig$ requirement already removes most $\ttbar$
+processes.
+
+The standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement captures
+both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
+respectively.
+
+The high regions, defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
+SR-High-1J, and SR-$\llbb$.
+To maximize projected sensitivity, SR-High is split into four bins with
+splits at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and second at
+$\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
+For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
+The regions SR-High-8 and SR-High-16 are displayed in
+Figure~\ref{fig:2ljets_high_region}, with their -a and -b bins visible across
+the $\metsig$ distribution.
+
+One major difference between signals and diboson backgrounds is that signal
+jet pairs mostly arise from boosted $W\rightarrow jj$ decays which more often
+produce nearby jets.
+Selecting small $\rjj$ in SR-High therefore picks out these boosted cases.
+
+Extremely boosted $W$ bosons can be clustered into single jets.
+Although large-radius jets are typically used to capture these, we only use
+small, $R=0.4$, jets in the $\twoljets$ analysis.
+And those small jets can work for the extreme boosts that our signals produce.
+To catch these extreme events is the goal of SR-High-1J, which is displayed
+in Figure~\ref{fig:2ljets_high_1J_region}.
+
+All other regions require two jets or more, so this single-jet selection has
+the liberty to reduce its $\metsig$ requirement to $\metsig > 12$
+without overlapping any other regions.
+
+Finally, SR-$\llbb$ is designed to target the GMSB model, which produces decays
+via $ZZ$ and $Zh$ bosons pairs.
+Since the $h\rightarrow bb$ branching fraction is large, SR-$\llbb$ finds
+sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
+The low edge of this $\mbb \in 60\textrm{--}150\,\eV[G]$ window is chosen not
+only to accommodate \atlas' noisy reconstruction of $b$-jets, but also to
+capture $Z\rightarrow bb$ decays, which become significant signal models with
+small $B(\neutralino_1 \rightarrow h \gravitino)$.
+
+As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in SR-$\llbb$
+contain a significant proportion of rare top quark processes;
+studies presented in Section~\ref{sec:2ljets_validation} show that these are
+mainly $t\bar t Z$.
 
@@
 \section{Validation}
-
+\label{sec:2ljets_validation}
 Figure~\ref{fig:2ljets_low_crz_extensions}

### 2022-06-07 main/chapter_2ljets.tex
remove subsub; rearrange figure locations; minor fixes

@@
 
-The $\mll \in 71\textrm{--}111$ window loosely selects the $m(Z)$ peak to
-scoop up nearby rare of mismeasured events.
-Similarly, allowing up to one $b$-tag improves yields without boosting
-backgrounds since the $\metsig$ requirement already removes most $\ttbar$
-processes.
-
-The standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement captures
-both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
-respectively.
-
 The high regions, defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
 SR-High-1J, and SR-$\llbb$.
-To maximize projected sensitivity, SR-High is split into four bins with
-splits at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and second at
-$\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
-For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
-The regions SR-High-8 and SR-High-16 are displayed in
-Figure~\ref{fig:2ljets_high_region}, with their -a and -b bins visible across
-the $\metsig$ distribution.
-
-One major difference between signals and diboson backgrounds is that signal
-jet pairs mostly arise from boosted $W\rightarrow jj$ decays which more often
-produce nearby jets.
-Selecting small $\rjj$ in SR-High therefore picks out these boosted cases.
-
-Extremely boosted $W$ bosons can be clustered into single jets.
-Although large-radius jets are typically used to capture these, we only use
-small, $R=0.4$, jets in the $\twoljets$ analysis.
-And those small jets can work for the extreme boosts that our signals produce.
-To catch these extreme events is the goal of SR-High-1J, which is displayed
-in Figure~\ref{fig:2ljets_high_1J_region}.
-
-All other regions require two jets or more, so this single-jet selection has
-the liberty to reduce its $\metsig$ requirement to $\metsig > 12$
-without overlapping any other regions.
-
-Finally, SR-$\llbb$ is designed to target the GMSB model, which produces decays
-via $ZZ$ and $Zh$ bosons pairs.
-Since the $h\rightarrow bb$ branching fraction is large, SR-$\llbb$ finds
-sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
-The low edge of this $\mbb \in 60\textrm{--}150\,\eV[G]$ window is chosen not
-only to accommodate \atlas' noisy reconstruction of $b$-jets, but also to
-capture $Z\rightarrow bb$ decays, which become significant signal models with
-small $B(\neutralino_1 \rightarrow h \gravitino)$.
-
-As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in SR-$\llbb$
-contain a significant proportion of rare top quark processes;
-studies presented in Section~\ref{sec:2ljets_validation} show that these are
-mainly $t\bar t Z$.
 
@@
 
-\subsubsection{High}
-
 \begin{figure}[tp]
@@
 
+The $\mll \in 71\textrm{--}111$ window loosely selects the $m(Z)$ peak to
+scoop up nearby rare of mismeasured events.
+Similarly, allowing up to one $b$-tag improves yields without boosting
+backgrounds since the $\metsig$ requirement already removes most $\ttbar$
+processes.
+
+The standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement captures
+both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
+respectively.
+
+To maximize projected sensitivity, SR-High is split into four bins with
+splits at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and second at
+$\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
+For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
+The regions SR-High-8 and SR-High-16 are displayed in
+Figure~\ref{fig:2ljets_high_region}, with their -a and -b bins visible across
+the $\metsig$ distribution.
+
 \begin{figure}[tp]
@@
 
-\subsubsection{Mono-jet}
+One major difference between signals and diboson backgrounds is that signal
+jet pairs mostly arise from boosted $W\rightarrow jj$ decays which more often
+produce nearby jets.
+Selecting small $\rjj$ in SR-High therefore picks out these boosted cases.
+
+Extremely boosted $W$ bosons can be clustered into single jets.
+Although large-radius jets are typically used to capture these, we only use
+small, $R=0.4$, jets in the $\twoljets$ analysis.
+And those small jets can work for the extreme boosts that our signals produce.
+To catch these extreme events is the goal of SR-High-1J, which is displayed
+in Figure~\ref{fig:2ljets_high_1J_region}.
 
@@
 
-\subsubsection{\texorpdfstring{$\llbb$}{\textit{llbb}}}
+All other regions require two jets or more, so this single-jet selection has
+the liberty to reduce its $\metsig$ requirement to $\metsig > 12$
+without overlapping any other regions.
+
+Finally, SR-$\llbb$ is designed to target the GMSB model, which produces decays
+via $ZZ$ and $Zh$ bosons pairs.
+Since the $h\rightarrow bb$ branching fraction is large, SR-$\llbb$ finds
+sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
+The low edge of this $\mbb \in 60\textrm{--}150\,\eV[G]$ window is chosen not
+only to accommodate \atlas' noisy reconstruction of $b$-jets, but also to
+capture $Z\rightarrow bb$ decays, which become significant signal models with
+small $B(\neutralino_1 \rightarrow h \gravitino)$.
+
+As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in SR-$\llbb$
+contain a significant proportion of rare top quark processes;
+studies presented in Section~\ref{sec:2ljets_validation} show that these are
+mainly $t\bar t Z$.
 
@@
 \caption[
-Histograms in a very loose pre-selection like those inspected to validate
-background modelling prior to designing the $\twoljets$ regions.
+Histograms in a very loose pre-selection
 ]{%

### 2022-06-07 main/chapter_2ljets.tex
fix overleaf warnings

@@
 \begin{tabular*}{\textwidth}{lr@{$~\pm~$}lccrrcc}
-Region &
-\multicolumn{2}{c}{Fitted} &
-Data &
-$\langle A\epsilon{ \sigma}\rangle_{\mathrm{obs}}^{95}~\mathrm{fb}$ &
-$S_{\mathrm{obs}}^{95}$ &
-$S_{\mathrm{exp}}^{95}$ &
-$\mathrm{CLb}$ &
-$p(s=0)$ \\[1.5ex]
-DR-OffShell & $22.1$ & $2.7$ & 21 & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$ \\[.5ex]
-DR-Low & $22$ & $4$ & 18 & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$ \\[.5ex]
-DR-Int & $35$ & $4$ & 38 & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$ \\[.5ex]
-DR-High & $3.9$ & $0.5$ & 0 & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0.00$ & $0.50$ \\[.5ex]
-DR-$\llbb$ & $0.51$ & $0.20$ & 0 & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$ \\[.5ex]
+& \multicolumn{2}{c}{Fitted}
+& Data
+& $A\epsilon\sigma_{\mathrm{obs}}^{95}~\mathrm{fb}$
+& $S_{\mathrm{obs}}^{95}$
+& $S_{\mathrm{exp}}^{95}$
+& $\mathrm{CLb}$
+& $p_{s\!=\!0}$
+\\[1.5ex]
+DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
+\\[.5ex]
+DR-Low & $22$ & $4$ & $18$ & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$
+\\[.5ex]
+DR-Int & $35$ & $4$ & $38$ & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$
+\\[.5ex]
+DR-High & $3.9$ & $0.5$ & $0$ & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0$ & $0.50$
+\\[.5ex]
+DR-$\llbb$ & $0.51$ & $0.20$ & $0$ & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$
+\\[.5ex]
 \end{tabular*}
@@
 All $p$-values are estimated by simulation of alternative data.
+\\
 \TODO{add labels and references to asymptotic formulae paper}
+\\
 \TODO{reference pdg rounding}
@@
 Following the conventional style descibed in
-Section~\ref{sec:sec:searches_searches},
+Section~\ref{sec:searches_searches},
 control regions are defined to normalize specific background samples,
@@
 \centering
-\begin{subfigure}{0.4\textwidth}
+\begin{subfigure}{0.38\textwidth}
 \centering
@@
 \quad\quad
-\begin{subfigure}{0.4\textwidth}
+\begin{subfigure}{0.38\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.4\textwidth}
+\begin{subfigure}{0.38\textwidth}
 \centering
@@
 \centering
-\includegraphics[height=0.9\textheight]{figures/2ljets_fit_parameters.pdf}
+\includegraphics[height=0.81\textheight]{figures/2ljets_fit_parameters.pdf}
 \caption[
 Parameter values and errors from the background only-fit to control and signal
-regions.
+regions
 ]{%
@@
 The red arrow highlights the $\zjets$ generator comparison parameter
-$\alpha\mathrm{\_Zjets-alt}$ whose post-fit value is
+$\alpha\mathrm{\_Zjets\_alt}$ whose post-fit value is
 % 1.8598e+00 +/-  6.87e-01

### 2022-06-08 main/chapter_2ljets.tex
center columns
high validation

@@
 % https://stackoverflow.com/a/2132998
-\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lccrrcc}
+\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
 & \multicolumn{2}{c}{Fitted}
@@
 
+\begin{table}[tp]
+\centering
+\begin{tabular}{c}
+Pre-selction
+\\[1em]
+$n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$
+\\[0.5em]
+same flavour, opposite charge leptons
+\\[0.5em]
+$\njet \geq 1$
+\\[0.5em]
+$p_\mathrm{T}^{\,\ell_2} > 25$
+\\[0.5em]
+$\metsig > 6$
+\\[0.5em]
+$\met > 100$
+\end{tabular}
+\caption[%
+Pre-selection requirements ensured prior to all region definitions in the
+$\twoljets$-electroweak search
+]{%
+Pre-selection requirements ensured prior to all region definitions in the
+$\twoljets$-electroweak search.
+Although $\metsig > 6$ implies that most events already have large $\met$
+$\met > 100\,\eV[G]$ requirement is included to ensure that events with small
+$\met$ and smaller uncertainties on it are excluded.
+}
+\label{tab:2ljets_presel}
+\end{table}
+
 High regions are discussed in Section~\ref{sec:2ljets_high}.
@@
 
-\begin{table}[tp]
-\centering
-\begin{tabular}{c}
-Pre-selction
-\\[1em]
-$n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$
-\\[0.5em]
-same flavour, opposite charge leptons
-\\[0.5em]
-$\njet \geq 1$
-\\[0.5em]
-$p_\mathrm{T}^{\,\ell_2} > 25$
-\\[0.5em]
-$\metsig > 6$
-\\[0.5em]
-$\met > 100$
-\end{tabular}
-\caption[%
-Pre-selection requirements ensured prior to all region definitions in the
-$\twoljets$-electroweak search
-]{%
-Pre-selection requirements ensured prior to all region definitions in the
-$\twoljets$-electroweak search.
-Although $\metsig > 6$ implies that most events already have large $\met$
-$\met > 100\,\eV[G]$ requirement is included to ensure that events with small
-$\met$ and smaller uncertainties on it are excluded.
-}
-\label{tab:2ljets_presel}
-\end{table}
-
+% presel table should come before High
+\FloatBarrier
 \subsection{High}
@@
 Differences between regions are \uwave{underlined}.
+\\[0.5em]
+The $\rll$ bins of SR-High are labelled SR-High-8 and SR-High-16, and their
+$\metsig$ binns are suffixed -a and -b for $\metsig \in 18\textrm{--}21$
+and $\metsig > 21$ respectively.
 }
@@
 \label{fig:2ljets_high_llbb_region}
+\end{figure}
+
+\subsubsection{Validation}
+
+As always, validation regions are defined near to each signal region, under
+the constraint of being somehow `nearby' and having enough data to test
+background predictions.
+
+For SR-High, inverting the selections on $\mjj$ or $\rjj$ satisfies
+these criteria, so one validation region is designed on each.
+An apparent excess appears at small $\mjj$ in VR-High, which is shown in
+Figure~\ref{fig:2ljets_high_mjj_vrhigh}.
+Various validation studies which explored similar selections and different
+simulated samples found no clear explanation for this effect.
+It therefore motivates the second validation region VR-High-R, and is one
+reason for including a lower bound of $\mjj > 20\,\eV[G]$ in all regions.
+As seen in Figures~\ref{fig:2ljets_high_region}
+and~\ref{fig:2ljets_high_mjj_vrhigh}, both VR-High and VR-High-R otherwise show
+acceptable pre-fit modelling even without systematic uncertainties.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRHigh.png}
+\caption{VR-High, $\mjj$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_VRHighR.png}
+\caption{VR-High-R, $\mjj$}
+\end{subfigure}
+\caption[
+Pre-fit $\mjj$ distributions in validation regions VR-High and VR-High-R
+]{%
+Pre-fit $\mjj$ distributions in validation regions VR-High and VR-High-R.
+Signal regions are unblinded.
+Errors are statistical.
+The apparent excess at low $\mjj$ in VR-High is not well understood.
+}
+\label{fig:2ljets_high_mjj_vrhigh}
+\end{figure}
+
+The distribution of mono-jet masses $\mjetone$ is sharply speaked towards zero,
+so cutting this out while inveritng the $\mjetone$ selection of SR-High-1J
+more clealrly inspects the modelling of massive jets, which appears to be good.
+Note that the region $\mjetone > 110$ is also checked in this validation region
+and seen to contain no data.
+This is good for the standard model.
+Since background estimates are near zero up there, non-zero data would have
+ruled out our modelling in favour of any model making small positive
+predictions.
+
+The small yields beside SR-$\llbb$ in other event variables left $\metsig$ as
+the only appropriate caldidate to change for VR-$\llbb$.
+Unfortunately, this region mostly checks $\ttbar$ backgrounds, and not the
+rare $\ttbar Z$ backgrounds which we would most like to validate.
+Early work observed good $\ttbar Z$ modelling in three- and four- lepton
+selections, but these were excluded from the $\twoljets$ analysis for
+orthogonality from other \atlas\ searches with which our results are to be
+combined.
+
+Total backgrounds predict data well in VR-$\llbb$.
+What was not appreciated soon enough, however, was an excess in the $\metsig$
+tail of this region, which is reproduced from the paper in
+Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper},
+The highest bins have very small background predictions and positive data.
+Had this been noticed before unblinding, more validation might have
+been advisable.
+``In the end'', one \atlas\ author comments,
+``it seems fine since there's not events in the SR!''~\cite{comment2022vrllbb}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.48\textwidth]{figures/2ljets_vr_llbb_met_sig.pdf}
+\caption[
+Post-fit $\metsig$ distribution in VR-$\llbb$
+]{%
+Post-fit $\metsig$ distribution in VR-$\llbb$.
+Backgrounds are to control and signal regions.
+This region is also displayed pre-fit with coarser bins in
+Figure~\ref{fig:2ljets_high_llbb_region}.
+}
+\label{fig:2ljets_high_metsig_vrllbb_paper}
 \end{figure}

### 2022-06-08 main/chapter_2ljets.tex
add first intermediate design text; fix inverted mjj statement; add floatbarriers

@@
 % summary plot
-
 \begin{figure}[tp]
@@
 % signal diagrams
-
 \begin{figure}[t]
@@
 % signal region plots
-
 \begin{figure}[tp]
@@
 % exclusion plots
-
 \begin{figure}[tp]
@@
 % atlas_rjr_mimic_SUSY_2018_06
-
 The $\twoljets$ analysis is a trinity,
@@
 
+\FloatBarrier
 \subsection{RJR}
@@
 
+\FloatBarrier
 \subsection{Strong}
@@
 
+\FloatBarrier
 \subsection{Electroweak}
@@
 update the regions of the partial Run~2 search in its $\twoljets$
-channel~\cite{atlas_23l_SUSY_2016_24} which also searched for signal models
+channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for signal models
 producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
@@
 \subsection{Jet pairs}
-\TODO{why leading two jets mjj / rjj}
+Di-jet assignments, for variables like $\mjj$ and $\rjj$, are always chosen as
+the two hardest jets.
+Why?
+
+This choice matches the convention of SR2-int and SR2-high from the previous
+search~\cite{atlas_23l_SUSY_2016_24}, but my initial reaction was that since we
+are looking for jet pairs near to $m(W)$ or $m(Z)$, we should surely choose
+candidate jet pairs by their mass.
+
+Two alternative di-jet assignments were therefore considered in early
+development of region designs:
+the two jets with invariant mass closest to $m(W)$, and the two jets with the
+minimum $\rjj^\mathrm{alternative}$ between them.
+The result was bland.
+All assignments performed similarly, and all left similar fractions of signal
+yields outside of the mass window.
+The simplicity and precedent for the two-hardest jet assignment therefore swung
+the decision.
+
+However, alternatives did select different subsets of signal yields;
+the alternative jet assignments had some potential to define secondary signal
+regions to catch those signal events which fall outside of the primary
+$\mjj$ window.
+One can, for example, select $\mjj \notin 60\textrm{--}110\,\eV[G]$ and
+$\mjj^\mathrm{alternative} \in 60\textrm{--}110\,\eV[G]$ for some additional
+sensitivity.
+However, this comes at the cost of complexity, particularly in the boundaries
+with control and validation regions, and is therefore not used here.
 
@@
 & $> 18$
-& $\uwave{\mjj\!:~  20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{\mjj\!:~  20\textrm{--}60 \mid 110\textrm{--}\infty}$
 & $< 1.6$
@@
 & $> 12$
-& $\uwave{\mjetone\!:~  20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{\mjetone\!:~  20\textrm{--}60 \mid 110\textrm{--}\infty}$
 &
@@
 
-\subsubsection{Validation}
-
-As always, validation regions are defined near to each signal region, under
+Validation regions, as always, are defined near to each signal region, under
 the constraint of being somehow `nearby' and having enough data to test
 background predictions.
-
 For SR-High, inverting the selections on $\mjj$ or $\rjj$ satisfies
@@
 Figure~\ref{fig:2ljets_high_mjj_vrhigh}.
-Various validation studies which explored similar selections and different
-simulated samples found no clear explanation for this effect.
+
+Various validation studies explored similar selections and different
+simulated samples, and found no clear explanation for this effect.
 It therefore motivates the second validation region VR-High-R, and is one
@@
 As seen in Figures~\ref{fig:2ljets_high_region}
-and~\ref{fig:2ljets_high_mjj_vrhigh}, both VR-High and VR-High-R otherwise show
+and~\ref{fig:2ljets_high_mjj_vrhigh}, VR-High and VR-High-R otherwise show
 acceptable pre-fit modelling even without systematic uncertainties.
@@
 
-The distribution of mono-jet masses $\mjetone$ is sharply speaked towards zero,
-so cutting this out while inveritng the $\mjetone$ selection of SR-High-1J
-more clealrly inspects the modelling of massive jets, which appears to be good.
+The distribution of mono-jet masses $\mjetone$ is sharply peaked towards zero,
+so cutting this out while inverting the $\mjetone$ selection of SR-High-1J
+more clearly inspects the modelling of massive jets, which appears to be good.
 Note that the region $\mjetone > 110$ is also checked in this validation region
@@
 The small yields beside SR-$\llbb$ in other event variables left $\metsig$ as
-the only appropriate caldidate to change for VR-$\llbb$.
+the only appropriate candidate to change for VR-$\llbb$.
 Unfortunately, this region mostly checks $\ttbar$ backgrounds, and not the
@@
 been advisable.
-``In the end'', one \atlas\ author comments,
+``In the end'', comments one \atlas\ author,
 ``it seems fine since there's not events in the SR!''~\cite{comment2022vrllbb}.
@@
 
+\FloatBarrier
 \subsection{Intermediate}
 \label{sec:2ljets_int}
-Table~\ref{tab:2ljets_int}
-Figure~\ref{fig:2ljets_int_region}
-Figure~\ref{fig:2ljets_int_cr_region}
+
+Intermediate signal regions are defined to sit in the $\metsig$ range of
+$12$ to $18$.
+As seen from Figure~\ref{fig:2ljets_presel_met}, this space suffers from
+$\ttbar$ background contributions, but it also receives substantial signal
+yields from some of the lighter signal models.
+The larger rates of background processes in this space also make it ripe for
+control regions.
+
+Intermediate signal regions are defined in Table~\ref{tab:2ljets_int},
+along with their associated validation regions and the control regions in this
+space.
+To suppress the increased backgrounds, the intermediate $\mll$ window is
+tightened to reduce non-$Z$ backgrounds including $\ttbar$ and diboson
+processes.
+Requiring zero $b$-tags further reduces $\ttbar$ backgrounds.
 
@@
 & $12\textrm{--}18$
-& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{20\textrm{--}60 \mid 110\textrm{--}\infty}$
 & $\uwave{\hphantom{< 60}}$
@@
 
+The less extreme signal models targeted here have less boosted
+$W\rightarrow jj$ distributions, and so it was found that a $\ptjone$ selection
+was effective at separating signals;
+a $\ptjone > 60\,\eV[G]$ requirement split SR-Int from its validation region,
+VR-Intm and is seen from Figure~\ref{fig:2ljets_int_region} to retain most
+signal contributions.
+Following our standard pattern set with other signal regions, SR-Int is binned
+in $\metsig$ in intervals of $3$ units to improve signal sensitivity.
+
 \begin{figure}[tp]
@@
 
+The control region CR-VZ is defined on the $\mjj$ side-band to
+SR-Int and VR-Int.
+As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
+intermediate space give CR-VZ a very pure diboson background, which is very
+valuable or constraining the modelling of High regions.
+Also included in Table~\ref{tab:2ljets_int} and
+Figure~\ref{fig:2ljets_int_cr_region} is CR-tt, which uses various other
+selections of $b$-tagged data to constrain $\ttbar$ backgrounds.
+Both control regions show good pre-fit agreement with statistical uncertainties
+only.
+
 \begin{figure}[tp]
@@
 
+Some figures label CR-VZ as `CR-ZZ'; the region was renamed late in this
+analysis project when it was show to contain substantial fractions of $WZ$
+as well as $ZZ$, as illustrated in Figure~\ref{fig:2ljets_splits} of
+Section~\ref{sec:2ljets_validation}.
+
+\FloatBarrier
 \subsection{Low}
@@
 & $6\textrm{--}9$
-& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{20\textrm{--}60 \mid 110\textrm{--}\infty}$
 & $< 80$
@@
 & $6\textrm{--}9$
-& $\uwave{20\textrm{--}60 \mid 60\textrm{--}110}$
+& $\uwave{20\textrm{--}60 \mid 110\textrm{--}\infty}$
 & $> 80$
@@
 
+\FloatBarrier
 \subsection{Off-shell}
@@
 
+\FloatBarrier
 \subsection{Discovery}

### 2022-06-08 main/chapter_2ljets.tex
spelling grammar check

@@
 All assignments performed similarly, and all left similar fractions of signal
-yields outside of the mass window.
+yields outside the mass window.
 The simplicity and precedent for the two-hardest jet assignment therefore swung
@@
 the alternative jet assignments had some potential to define secondary signal
-regions to catch those signal events which fall outside of the primary
+regions to catch those signal events which fall outside the primary
 $\mjj$ window.
@@
 
-A high-level summary of the region design is shown
+A high-level summary of the region design is shown in
 Figure~\ref{fig:2ljets_region_summary}.
@@
 with boundaries at $a$, $b$, and $c$.
-The mid bar `$\mid$' indicates logical or.
+The mid-bar `$\mid$' indicates logical or.
 Differences between regions are \uwave{underlined}.
@@
 The $\mll \in 71\textrm{--}111$ window loosely selects the $m(Z)$ peak to
-scoop up nearby rare of mismeasured events.
+scoop up nearby rare or mismeasured events.
 Similarly, allowing up to one $b$-tag improves yields without boosting
-backgrounds since the $\metsig$ requirement already removes most $\ttbar$
+backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
 processes.
@@
 To maximize projected sensitivity, SR-High is split into four bins with
-splits at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and second at
+one split at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and a second at
 $\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
@@
 One major difference between signals and diboson backgrounds is that signal
-jet pairs mostly arise from boosted $W\rightarrow jj$ decays which more often
+jet pairs mostly arise from boosted $W\rightarrow jj$ decays, which more often
 produce nearby jets.
@@
 $\ttbar$ background contributions, but it also receives substantial signal
-yields from some of the lighter signal models.
+yields from some lighter signal models.
 The larger rates of background processes in this space also make it ripe for
@@
 To suppress the increased backgrounds, the intermediate $\mll$ window is
-tightened to reduce non-$Z$ backgrounds including $\ttbar$ and diboson
+tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and diboson
 processes.
@@
 a $\ptjone > 60\,\eV[G]$ requirement split SR-Int from its validation region,
-VR-Intm and is seen from Figure~\ref{fig:2ljets_int_region} to retain most
+VR-Int and is seen from Figure~\ref{fig:2ljets_int_region} to retain most
 signal contributions.
@@
 selections of $b$-tagged data to constrain $\ttbar$ backgrounds.
-Both control regions show good pre-fit agreement with statistical uncertainties
+Both control regions show good agreement pre-fit and with statistical uncertainties
 only.
-
 \begin{figure}[tp]
@@
 \label{sec:2ljets_low}
+
+With the tighter selections inherited from Intermediate space, $\ttbar$ yields
+remain small in Low regions, which have $\metsig \in 6\textrm{--}12$.
+Down here, however, $\zjets$ backgrounds become large, as their fake $\met$
+from mismeasured jets is no longer eliminated.
+
+\TODO{read int again, remind myself why some of these decisions are made}
+
 Table~\ref{tab:2ljets_low}

### 2022-06-10 main/chapter_2ljets.tex
add mt2 comment; add srlow n-1 plots

@@
 \begin{tabular}{c}
-Pre-selction
+Pre-selection:
 \\[1em]
@@
 
+A explained in Section~\ref{2ljets_mt2}, requiring $\mttwo$ above $m(W)$
+effectively removes $\ttbar$ and other backgrounds.
+The relatively low requirement $\mttwoll > 80\,\eV[G]$ is chosen here and
+elsewhere in the $\twoljets$-electroweak design to reduce these backgrounds
+while minimize lost signal yields.
+
 To maximize projected sensitivity, SR-High is split into four bins with
@@
 
-With the tighter selections inherited from Intermediate space, $\ttbar$ yields
-remain small in Low regions, which have $\metsig \in 6\textrm{--}12$.
-Down here, however, $\zjets$ backgrounds become large, as their fake $\met$
-from mismeasured jets is no longer eliminated.
-
-\TODO{read int again, remind myself why some of these decisions are made}
+In Low regions, $\metsig \in 6\textrm{--}12$, $\zjets$ becomes a large
+background contribution, since its fake $\met$, which arises from
+from mismeasurements or rare interactions, is no longer eliminated.
+Selections inherited from Intermediate regions to suppress
+$\ttbar$ backgrounds also need supplementing.
 
 Table~\ref{tab:2ljets_low}
+Figure~\ref{fig:2ljets_low_minus}
 Figure~\ref{fig:2ljets_low_region}
@@
 Figure~\ref{fig:2ljets_low_sr_or_cr_region}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_Rll_SRLow_noRll.pdf}
+\caption{SR-Low$/\{\rll\}$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_mt2leplsp_0_SRLow_noRll_nomt2.pdf}
+\caption{SR-Low$/\{\rll,\mttwoll\}$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi.pdf}
+\caption{SR-Low-2$/\{\dphillmet\}$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi_mg5.pdf}
+\caption{SR-Low-2$/\{\dphillmet\}$, alternative}
+\end{subfigure}
+\caption[
+Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections
+]{%
+Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections.
+Each removes requirements listed in Table~\ref{tab:2ljets_low}:
+(a) SR-Low removes $\rll$,
+(b) SR-Low removes $\rll$ and $\mttwoll$,
+and
+(c, d) SR-Low-2 removes $\dphillmet$.
+\\[0.5em]
+Signals trend towards small $\rll$, plausibly due to their more boosted
+$Z\rightarrow \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
+The mass of signals seen at small $\mttwoll$ in (b) motivates the design of
+SR-Low-2.
+Signals in SR-Low-2 spike at small $\dphillmet$ in (c); better statistical
+precision is provided by an alternative $\zjets$ sample in (d).
+}
+\label{fig:2ljets_low_minus}
+\end{figure}
 

### 2022-06-13 main/chapter_2ljets.tex
begin fleshing out low text

@@
 
-In Low regions, $\metsig \in 6\textrm{--}12$, $\zjets$ becomes a large
-background contribution, since its fake $\met$, which arises from
+In Low regions, which have $\metsig \in 6\textrm{--}12$, the $\zjets$
+background becomes a large since its fake $\met$, which arises from
 from mismeasurements or rare interactions, is no longer eliminated.
-Selections inherited from Intermediate regions to suppress
-$\ttbar$ backgrounds also need supplementing.
-
-Table~\ref{tab:2ljets_low}
-Figure~\ref{fig:2ljets_low_minus}
+Similarly, backgrounds from $\ttbar$ increase; the selections inherited from
+Intermediate regions to suppress $\ttbar$ need supplementing.
+
+Angular selections on lepton variables $\rll$ and $\dphillmet$ are used to
+reject $\ttbar$ and provide signal sensitivity in Low regions, which are
+defined in Table~\ref{tab:2ljets_low}.
+Motivations for these angular selections are displayed in
+Figure~\ref{fig:2ljets_low_minus}, and discussed here.
+
+As elsewhere, binning of SR-Low in $\metsig \in 6\textrm{--}9\textrm{--}12$
+increases sensitivity.
+%TODO
+
+Uniquely in Low regions and signal models of small mass splittings, it is
+seen (from Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}) that some
+substantial signal masses accumulate at small $\mttwoll$.
+To capture these signals, SR-Low-2 is defined as the only
+$\twoljets$-electroweak region with $\mttwoll < 80\,\eV[G]$.
+
+That threshold of $80\,\eV[G]$ was chosen since after other selections for
+SR-Low-2, almost all signal and background samples are in the spike near $0$;
+a lower requirement would make little difference, and $80\,\eV[G]$ matches
+other, usually inverted, selections.
+
+The exclusion of $\metsig \in 9\textrm{--}12$ in SR-Low-2 is not particularly
+well motivated.
+This space contains about $2.3$ background expectation (mostly diboson) and
+$\approx 1$ from some benchmark signals, so its inclusion as a single combined
+bin not have increased sensitivity.
+Its inclusion as an orthogonal bin with small yields would have added excessive
+complication, however.
+
 Figure~\ref{fig:2ljets_low_region}
@@
 \caption{SR-Low$/\{\rll,\mttwoll\}$}
+\label{fig:2ljets_low_minus_norll_nomt2_mt2}
 \end{subfigure}

### 2022-06-13 main/chapter_2ljets.tex
finish main low section; rearrange figures

@@
 
-As elsewhere, binning of SR-Low in $\metsig \in 6\textrm{--}9\textrm{--}12$
-increases sensitivity.
-%TODO
-
-Uniquely in Low regions and signal models of small mass splittings, it is
-seen (from Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}) that some
-substantial signal masses accumulate at small $\mttwoll$.
-To capture these signals, SR-Low-2 is defined as the only
-$\twoljets$-electroweak region with $\mttwoll < 80\,\eV[G]$.
-
-That threshold of $80\,\eV[G]$ was chosen since after other selections for
-SR-Low-2, almost all signal and background samples are in the spike near $0$;
-a lower requirement would make little difference, and $80\,\eV[G]$ matches
-other, usually inverted, selections.
-
-The exclusion of $\metsig \in 9\textrm{--}12$ in SR-Low-2 is not particularly
-well motivated.
-This space contains about $2.3$ background expectation (mostly diboson) and
-$\approx 1$ from some benchmark signals, so its inclusion as a single combined
-bin not have increased sensitivity.
-Its inclusion as an orthogonal bin with small yields would have added excessive
-complication, however.
-
-Figure~\ref{fig:2ljets_low_region}
-Figure~\ref{fig:2ljets_low2_region}
-Figure~\ref{fig:2ljets_low_cr_region}
-Figure~\ref{fig:2ljets_low_sr_or_cr_region}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_Rll_SRLow_noRll.pdf}
-\caption{SR-Low$/\{\rll\}$}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_mt2leplsp_0_SRLow_noRll_nomt2.pdf}
-\caption{SR-Low$/\{\rll,\mttwoll\}$}
-\label{fig:2ljets_low_minus_norll_nomt2_mt2}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi.pdf}
-\caption{SR-Low-2$/\{\dphillmet\}$}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi_mg5.pdf}
-\caption{SR-Low-2$/\{\dphillmet\}$, alternative}
-\end{subfigure}
-\caption[
-Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections
-]{%
-Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections.
-Each removes requirements listed in Table~\ref{tab:2ljets_low}:
-(a) SR-Low removes $\rll$,
-(b) SR-Low removes $\rll$ and $\mttwoll$,
-and
-(c, d) SR-Low-2 removes $\dphillmet$.
-\\[0.5em]
-Signals trend towards small $\rll$, plausibly due to their more boosted
-$Z\rightarrow \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
-The mass of signals seen at small $\mttwoll$ in (b) motivates the design of
-SR-Low-2.
-Signals in SR-Low-2 spike at small $\dphillmet$ in (c); better statistical
-precision is provided by an alternative $\zjets$ sample in (d).
-}
-\label{fig:2ljets_low_minus}
-\end{figure}
-
-\subsubsection{Low-ISR}
-\label{sec:2ljets_low_isr}
-Figure~\ref{fig:2ljets_low_isr_old}
-
 \begin{table}[tp]
@@
 
+Boosted $Z\rightarrow \ell^+\ell^-$ decays will tend to have small $\rll$.
+Similarly to $\rjj$ in the High regions, this can usefully select signals,
+but here also works to reject other $\ell^+\ell^-$ production mechanisms which
+produce leptons is less correlated directions, particularly dileptonic
+$\ttbar$.
+As elsewhere, binning of SR-Low in $\metsig \in 6\textrm{--}9\textrm{--}12$
+to form SR-Low-a and SR-Low-b respectively, increases sensitivity.
+
+Unblinded, pre-fit histograms of SR-Low are shown in
+Figure~\ref{fig:2ljets_low_region}.
+Note that these plots do not include systematic uncertainties, so the meaning
+of the observed data deficits cannot yet be fully interpreted.
+This figure also includes VR-Low, which observes the high $\rll$ edge and
+predicts the data well.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_Rll_SRLow_noRll.pdf}
+\caption{SR-Low$/\{\rll\}$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_mt2leplsp_0_SRLow_noRll_nomt2.pdf}
+\caption{SR-Low$/\{\rll,\mttwoll\}$}
+\label{fig:2ljets_low_minus_norll_nomt2_mt2}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi.pdf}
+\caption{SR-Low-2$/\{\dphillmet\}$}
+\label{fig:2ljets_low_minus_dphi}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi_mg5.pdf}
+\caption{SR-Low-2$/\{\dphillmet\}$, alternative}
+\label{fig:2ljets_low_minus_dphi_alt}
+\end{subfigure}
+\caption[
+Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections
+]{%
+Relaxed versions of SR-Low and SR-Low-2 to motivate their angular selections.
+Each removes requirements listed in Table~\ref{tab:2ljets_low}:
+(a) SR-Low removes $\rll$,
+(b) SR-Low removes $\rll$ and $\mttwoll$,
+and
+(c, d) SR-Low-2 removes $\dphillmet$.
+\\[0.5em]
+Signals trend towards small $\rll$, plausibly due to their more boosted
+$Z\rightarrow \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
+The mass of signals seen at small $\mttwoll$ in (b) motivates the design of
+SR-Low-2.
+Signals in SR-Low-2 spike at small $\dphillmet$ in (c); better statistical
+precision is provided by an alternative $\zjets$ sample in (d).
+}
+\label{fig:2ljets_low_minus}
+\end{figure}
+
 \begin{figure}[tp]
@@
 
+Uniquely to Low regions, and signal models of small mass splittings, it is
+seen (in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}) that substantial
+signal yields exist at small $\mttwoll$.
+To capture these signals, SR-Low-2 is defined as the only
+$\twoljets$-electroweak region with $\mttwoll < 80\,\eV[G]$.
+That threshold of $80\,\eV[G]$ is not very important.
+After the other selections for SR-Low-2, almost all signal and background
+samples are in the spike near $0$;
+a lower requirement would make little difference, and $80\,\eV[G]$ matches
+other, usually inverted, selections.
+
+As shown in Figure~\ref{fig:2ljets_low_minus_dphi}, signal models peak at small
+$\dphillmet$, and backgrounds do not; this motivates the $\dphillmet < 0.6$
+requirement, and can be understood since a boosted resonance decaying to
+$Z+\mathrm{invisible}$ will tend to send the $Z\rightarrow \ell^+\ell^-$ and
+$\ptmiss$ in the same direction.
+
+The distribution $\zjets$ here may be surprising, since its $\ptmiss$ is
+usually expected to align with a mismeasured jet.
+This is explained by the context of a prior $\mttwoll < 80\,\eV[G]$
+requirement:
+if $\ptmiss$ is facing away from both leptons, then they have large transverse
+masses.
+Following its definition as a $\min-\max$ of transverse masses,
+given in Section~\ref{sec:2ljets_mt2},
+the $\mttwoll$ requirement therefore rejects those events, leaving only the
+rarer cases of $\zjets$ with small $\dphillmet$.
+
+Excluding $\metsig \in 9\textrm{--}12$ in SR-Low-2 is also not crucial.
+This space contains about $2.3$ background expectation (mostly diboson) and
+$\approx 1$ from some benchmark signals, so its inclusion as a single combined
+bin not have increased sensitivity.
+Its inclusion as an orthogonal bin with small yields would have added excessive
+complication, however.
+
+Histograms of SR-Low-2 are shown in Figure~\ref{fig:2ljets_low2_region}.
+To get sufficiently large backgrounds, the validation region VR-Low-2 uses
+the $\mjj$ side-band, and sees good agreement.
+
 \begin{figure}[tp]
@@
 
+Sadly, the $\zjets$ background in SR-Low-2 has terrible statistical precision
+in the central simulated sample due to large weights.
+This was not as bad when the region was first designed, but became exaggerated
+after reprocessing with the combination lepton definition.
+The region is plotted with an alternative sample in
+Figure~\ref{fig:2ljets_low_minus_dphi_alt}, which has better precision.
+
+% CR
+The $\zjets$ control region CR-Z is displayed in
+Figure~\ref{fig:2ljets_low_cr_region}.
+Pre-fit agreement with data is quite good, although the overflow bin has a
+notable excess.
+Standardizing cuts to remove high-$\mjj$ tails (such as requiring
+$\mjj < 500\,\eV[G]$) could have made control and validation regions like CR-Z
+kinematically closer the the signal regions they are supposed to target,
+and at least in this case, I believe such a cut would have improved the
+analysis.
+
 \begin{figure}[tp]
@@
 
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.9\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ.pdf}
-\caption{}
-\end{subfigure}
+The gap at in CR-Z at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
+Their union therefore has a continuous $\mjj$ distribution, which is displayed
+in Figure~\ref{fig:2ljets_low_sr_or_cr_region}.
+Looking at the signal region alone can give an impression that $\zjets$ is
+overestimated.
+This plot contradicts that; most of the $\zjets$ distribution is fine;
+the strange effect happens where diboson becomes large near $m(W)$.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ.pdf}
 \caption[
@@
 \end{figure}
+
+\subsubsection{Low-ISR}
+\label{sec:2ljets_low_isr}
+Figure~\ref{fig:2ljets_low_isr_old}
+% TODO
 

### 2022-06-13 main/chapter_2ljets.tex
clarify low, mention 2jet selection

@@
 
-Angular selections on lepton variables $\rll$ and $\dphillmet$ are used to
-reject $\ttbar$ and provide signal sensitivity in Low regions, which are
-defined in Table~\ref{tab:2ljets_low}.
-Motivations for these angular selections are displayed in
-Figure~\ref{fig:2ljets_low_minus}, and discussed here.
+Angular selections on lepton variables $\rll$ and $\dphillmet$ characterize
+the Low regions, which are defined in Table~\ref{tab:2ljets_low}.
+These angular variables usefully reject backgrounds including $\ttbar$, and
+provide signal sensitivity to benchmark signal models, as illustrated in
+Figure~\ref{fig:2ljets_low_minus};
+signals with with small mass splittings of around $100\textrm{--}150\,\eV[G]$
+are targetted.
+
+To remove some observed mis-modelling of backgrounds, Low regions require
+exactly two jets.
+Abandoned region designs which used more jets are discussed below in
+Section~\ref{sec:2ljets_low_isr}.
 
@@
 Figure~\ref{fig:2ljets_low_isr_old}
-% TODO
+
+The Low regions presented in Section~\ref{sec:2ljets_low_isr} resulted from a
+hasty redesign shortly before unblinding.
 

### 2022-06-13 main/chapter_2ljets.tex
low-ISR writing and extra cr figure

@@
 \label{sec:2ljets_low_isr}
-Figure~\ref{fig:2ljets_low_isr_old}
-
-The Low regions presented in Section~\ref{sec:2ljets_low_isr} resulted from a
+
+The Low regions presented in Section~\ref{sec:2ljets_low} resulted from a
 hasty redesign shortly before unblinding.
+
+Before that redesign, an multi-jet signal region SR-Low-ISR was included,
+with associated validation and control regions,
+which required $\njet \geq 3$ and offered some meagre extra sensitivity.
+Here, `ISR' stands for Initial-State Radiation, following the naming
+of previous \atlas searches including partial Run-2 2/3-lepton
+search which we are updating~\cite{atlas_23l_SUSY_2016_24}.
+
+Although initial-state radiation typically refers to soft emissions in
+parton showers~\cite{corcella2000initial, bewick2022initial},
+our ISR explicitly includes hard emissions from the perturbative matrix element
+simulation;
+ISR is selected to give a hard transverse boost to the supersymmetric
+particles, making their resulting $\met$, and other kinematics, more distinctive.
+
+As an event variable, we follow \cite{atlas_23l_SUSY_2016_24} to define the
+ISR momentum as the vector sum of all jets not assigned to the
+$W\rightarrow jj$ decay.
+This ISR construction does not select the two hardest for the $W$ candidate,
+unlike our other regions, but instead takes the two closest in $\phi$ to the
+vector $\vec \pt^{\ell\ell} + \ptmiss$;
+the two jets selected in this way are used to define the mass $\mwisr$.
+
+The region SR-Low-ISR required $\ptisr > 300\,\eV[G]$ for a substantial
+boost and $\dphiisrmet > 2.6$ to catch invisible particles boosted away from the
+ISR, along with standard selections $\mwisr \in 60\textrm{--}110\,\eV[G]$,
+$\metsig \in 6\textrm{--}12$, and preselection, as well as $\mjj > 110\,\eV[G]$
+for orthogonality from other signal regions.
+
+The main ISR regions are shown in Figure~\ref{fig:2ljets_low_isr_old}; no data
+are shown in the signal region since it was never unblinded.
+From the validation and control regions, however, there is a clear deficit
+of data below backgrounds by about $30\%$.
+Here, VR-Low-ISR differs from SR-Low-ISR by requiring
+$\ptisr \in 200\textit{--}300\,\eV[G]$, and CR-Low-ISR differs by requiring
+$\ptisr > 200\,\eV[G]$ and
+$\mwisr \in 20\textrm{--}60 \mid 110\textrm{--}\infty$.
 
@@
 \caption{SR-Low-ISR, $\ptisr$ (blind)}
+\label{fig:2ljets_low_isr_old_sr}
 \end{subfigure}
@@
 with normalization factors in other control regions.
-Variables are defined in the text of Section~\ref{sec:2ljets_low_isr}.
+Variables and regions are defined in the text of
+Section~\ref{sec:2ljets_low_isr}.
 Errors are statistical.
@@
 \end{figure}
+
+Additional control regions were designed in an attempt to improve background
+normalization local to this region.
+These are CR-Low-ISR-Z and CR-Low-ISR-DF, which select $\zjets$ and $\ttbar$
+backgrounds and find smaller deficits of $20\%$ and $10\%$ respectively,
+as shown in Figure~\ref{fig:2ljets_low_isr_old_more}.
+To select $\zjets$, CR-Low-ISR-Z differs from CR-Low-ISR by requiring
+$\dphiisrmet \in 1\textrm{--}2.6$ and $\metsig \in 6\textrm{--}9$.
+To select $\ttbar$, CR-Low-ISR-DF differs from CR-Low-ISR by requiring
+different-flavour leptons and the simpler $\mwisr > 20\,\eV[G]$, since it
+has no signal regions to avoid in different-flavour data.
+No nearby selection was found to form a pure diboson control region near here.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_crlowisr_z.png}
+\caption{CR-Low-ISR-Z, $\mwisr$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_old_crlowisr_df.png}
+\caption{CR-Low-ISR-DF, $\mwisr$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in extra control regions for the draft Low-ISR
+]{%
+Pre-fit distributions in extra control regions for the draft Low-ISR.
+These regions required a third, hard jet directed away from $\ptmiss$.
+Variables and regions are defined in the text of
+Section~\ref{sec:2ljets_low_isr}.
+Errors are statistical.
+}
+\label{fig:2ljets_low_isr_old_more}
+\end{figure}
+
+A complicated likelihood model was constructed which normalized diboson,
+$\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,
+CR-Low-ISR-Z, CR-Low-ISR-DF.
+After this fit, the validation region VR-Low-ISR had a high background
+of $56.7 \pm 4.4$ for $47$ data, which is only a $1$-sigma deficit.
+
+This complication is a great deal of cost for what is only a weakly sensitive
+signal region, reaching only a projected significance $Z\approx 1$ for the
+(pre-fit) signals shown in Figure~\ref{fig:2ljets_low_isr_old_sr}.
+On the very sound advice of \atlas authors commenting on the internal
+documentation, this ISR construction was entirely removed.
+
+Modelling multi-jet QCD is hard.
+This deficit was therefore blamed on multi-jet QCD modelling, and resolved by
+retreating to two-jet events.
+I find it interesting, however, that the two-jet SR-Low did eventually observe
+a comparable deficit when unblinded.
 

### 2022-06-13 main/chapter_2ljets.tex
grammar check

@@
 
-A explained in Section~\ref{2ljets_mt2}, requiring $\mttwo$ above $m(W)$
+As explained in Section~\ref{sec:2ljets_mt2}, requiring $\mttwo$ above $m(W)$
 effectively removes $\ttbar$ and other backgrounds.
@@
 Figure~\ref{fig:2ljets_low_minus};
-signals with with small mass splittings of around $100\textrm{--}150\,\eV[G]$
+signals with small mass-splittings of around $100\textrm{--}150\,\eV[G]$
 are targetted.
@@
 
-Uniquely to Low regions, and signal models of small mass splittings, it is
+Uniquely to Low regions, and signal models with small mass-splitting, it is
 seen (in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}) that substantial
@@
 This was not as bad when the region was first designed, but became exaggerated
-after reprocessing with the combination lepton definition.
+after reprocessing with a new lepton definition and updated software.
 The region is plotted with an alternative sample in
@@
 $\mjj < 500\,\eV[G]$) could have made control and validation regions like CR-Z
-kinematically closer the the signal regions they are supposed to target,
+kinematically closer to the signal regions they are supposed to target,
 and at least in this case, I believe such a cut would have improved the
@@
 The main ISR regions are shown in Figure~\ref{fig:2ljets_low_isr_old}; no data
-are shown in the signal region since it was never unblinded.
+are shown in the signal region, since it was never unblinded.
 From the validation and control regions, however, there is a clear deficit
@@
 ]{%
-Pre-fit distributions in draft Low-ISR regions which were removed and never
+Pre-fit distributions in draft Low-ISR regions, which were removed and never
 unblinded.

### 2022-06-13 main/chapter_2ljets.tex
shrink crz+srlow figure to .8
clarify isr captions

@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ.pdf}
+\includegraphics[width=0.8\textwidth]{figures/2ljets_low_mjj_SRLow_or_CRZ.pdf}
 \caption[
@@
 \caption[
-Pre-fit distributions in draft Low-ISR regions
+Pre-fit distributions in draft Low-ISR regions, which were removed
 ]{%
@@
 unblinded.
-These regions required a third, hard jet directed away from $\ptmiss$.
+These regions required hard jets directed away from the leptons and $\ptmiss$.
 As seen from (b) and (c), the background modelling is poor and in conflict
@@
 \caption[
-Pre-fit distributions in extra control regions for the draft Low-ISR
-]{%
-Pre-fit distributions in extra control regions for the draft Low-ISR.
-These regions required a third, hard jet directed away from $\ptmiss$.
+Pre-fit distributions in extra draft Low-ISR control regions, which were
+removed
+]{%
+Pre-fit distributions in extra draft Low-ISR control regions, which were
+removed.
+These regions required hard jets directed away from the leptons and $\ptmiss$.
 Variables and regions are defined in the text of

### 2022-06-13 main/chapter_2ljets.tex
\atlas\

@@
 Here, `ISR' stands for Initial-State Radiation, following the naming
-of previous \atlas searches including partial Run-2 2/3-lepton
+of previous \atlas\ searches including partial Run-2 2/3-lepton
 search which we are updating~\cite{atlas_23l_SUSY_2016_24}.
@@
 (pre-fit) signals shown in Figure~\ref{fig:2ljets_low_isr_old_sr}.
-On the very sound advice of \atlas authors commenting on the internal
+On the very sound advice of \atlas\ authors commenting on the internal
 documentation, this ISR construction was entirely removed.

### 2022-06-13 main/chapter_2ljets.tex
remove duplicate near
clarify interesting sentence

@@
 has no signal regions to avoid in different-flavour data.
-No nearby selection was found to form a pure diboson control region near here.
+No nearby selection was found to form a pure diboson control region.
 
@@
 retreating to two-jet events.
-I find it interesting, however, that the two-jet SR-Low did eventually observe
-a comparable deficit when unblinded.
+It is interesting , however, that the two-jet SR-Low did eventually observe a
+comparable deficit.
 

### 2022-06-14 main/chapter_2ljets.tex
clarify interesting sentence more
add c1n2 200,100 note; remove atlas tag from some figures

@@
 \label{sec:2ljets_low}
-
 In Low regions, which have $\metsig \in 6\textrm{--}12$, the $\zjets$
@@
 Intermediate regions to suppress $\ttbar$ need supplementing.
+
+These regions target signal models with small mass splittings which still
+exceed the masses of electroweak bosons.
+In particular, we aim to test C1N2($200$, $100$) as a light, on-shell
+benchmark which has not previously been excluded in this channel.
 
@@
 retreating to two-jet events.
-It is interesting , however, that the two-jet SR-Low did eventually observe a
+It is interesting , however, that the two-jet SR-Low eventually observed a
 comparable deficit.

### 2022-06-14 main/chapter_2ljets.tex
write offshell region section

@@
 \label{sec:2ljets_offshell}
-Table~\ref{tab:2ljets_offshell}
-Figure~\ref{fig:2ljets_offshell_region}
+Previous iterations of our $\twoljets$ search did not target `off-shell'
+scenarios~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}, in which
+models have mass splittings below $m(W)$ or $m(Z)$, such that decays must
+proceed below their pole masses.
+Models with smaller off-shell mass splittings are described as more
+`compressed'.
+Separate searches have, however, targetted compressed scenarios in
+specialized and carefully designed analyses~\cite{
+atlas_susy_compressed_2l_2016_partial_run2,
+atlas_susy_compressed_2l_2018_run2
+} in selections with two leptons.
+
+Since compressed signals generate softer kinematics and \atlas\ is sensitive
+to soft leptons, three-lepton searches have stronger sensitivity to
+compressed scenarios~\cite{atlas_rjr_3l_SUSY_2019_09}.
+These searches consider the same C1N2 $WZ$ signal models as us, but target the
+case where the $W$ decays leptonically $W\rightarrow\ell\nu$.
+For comparison, impressive off-shell exclusion contours are reproduced
+from the \atlas\ Run-2 $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09} in
+Figure~\ref{fig:ljets_offshell_3l_exclusion}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/2ljets_compressed_3l_ins1866951_fig_16a.png}
+\caption[
+Contours showing off-shell C1N2 exclusion from the \atlas\ Run-2 $3\ell$ search
+]{%
+Contours showing off-shell C1N2 exclusion in the three lepton channel,
+reproduced from the \atlas\ Run-2 $3\ell$
+search~\cite{atlas_rjr_3l_SUSY_2019_09, hepdata.95751}.
+Off-shell models are excluded in that large lobe squeezed between the
+$m(\neutralino_2) = m(\neutralino_1)$ and
+$m(\neutralino_2) = m(\neutralino_1) - m_Z$ diagonal lines.
+}
+\label{fig:ljets_offshell_3l_exclusion}
+\end{figure}
+
+Despite the expert input from these off-shell and compressed searches, it was
+decided that $\twoljets$-electroweak should chime in to the off-shell
+conversation.
+Although specialized searches benefit from softer leptons and $\met$
+triggers~\cite{atlas_susy_compressed_2l_2018_run2},
+we still have some sensitivity since standard model backgrounds drop off
+below $\mll \approx m(Z)$.
+
+This is the purpose of SR-OffShell and its associated validation and control
+regions that are defined in Table~\ref{tab:2ljets_offshell}.
+Histograms of SR-OffShell and VR-OffShell are displayed in
+Figure~\ref{fig:2ljets_offshell_region}, and the off-shell control region
+CR-DY is displayed in Figure~\ref{fig:2ljets_offshell_cr_region}.
 
@@
 
+Our off-shell regime is characterized by $\mll < m(Z)$.
+The $W$ must also be off shell, the softer hadronic activity from light $W^*$
+events is not reliably reconstructed as signal jets.
+
+Early OffShell designs used the `ISR' construction, which is described
+Section~\ref{sec:2ljets_low} for Low regions,
+and which seeks hard jets which boost the supersymmetric particles and their
+decay products.
+
+After some iterations, however, the same intent was reproduced with a simpler
+and more effective selection.
+As seen in Table~\ref{tab:2ljets_offshell}, this design is to simply ask for
+one hard jet facing away from $\ptmiss$; this is specified as
+$\ptjone > 100\,\eV[G]$ and $\dphijmet > 2$.
+In comparison to ISR variables, this design improves yields by not needing jets
+for a $W\rightarrow jj$ candidate.
+
+An upper limit of $\mjj < 71\,\eV[G]$ is chosen for orthogonality from all
+on-shell regions.
+Our simulations of $\zjets$ are split into on- and off-shell parts at
+$40\,\eV[G]$, so that forms a natural breakpoint for binning which might help
+to separate concerns about systematic variations.
+As always, this binning aids sensitivity since models' contributions to the
+the bins vary with their mass splittings.
+
+Hadronic resonances which decay to lepton pairs,
+such as the bottomonium $\Upsilon$,
+become important below $10\,\eV[G]$, and we did not have simulations available
+for the $\mll < 10\,\eV[G]$ range.
+The requirement $\mll > 12\,\eV[G]$ is therefore imposed to robustly remove
+this un-modelled regime with a margin for error.
+
+Unlike on-shell regions, the standard $\mttwoll > 80\,\eV[G]$ requirement did
+not adequately reduce off-shell $\ttbar$ backgrounds.
+Pushing it up to $\mttwoll > 100\,\eV[G]$ was found to work well for the signal
+region, and the $\mttwoll \in 80\textrm{--}100\,\eV[G]$ nicely forms a the
+nearby validation region.
+Part of the light $\mll$ background arises from leptonically decaying
+$Z\rightarrow \tau\tau$ events; these are also removed by requiring large
+$\mttwoll$.
+
+Away from the $m(Z)$ peak, the full dilepton spectrum is not dominated by the
+$Z$ resonance, but depends on the full Drell-Yan, $\zyjets$,
+process~\cite{drell1970massive}.
+That is, on- and off-shell $\mll$ distributions are physically different.
+Since also $\zyjets$ simulations are culturally not trusted, we decided
+to introduce separate normalization factors for $\zyjets$ in on- and off-shell
+regions.
+
 \begin{figure}[tp]
@@
 \end{figure}
+
+The control region CR-DY (named after Drell and Yan) is defined to normalize off-shell $\zyjets$ near to
+SR-OffShell.
+Its reduced $\metsig$ requirement and relaxed angular selections were required
+to increase yields.
+As seen in Figure~\ref{fig:2ljets_offshell_cr_region}, pre-fit backgrounds
+agree well with data.
+Contributions of $\zyjets$ in SR-OffShell are small anyway, as shown in
+Figure~\ref{fig:2ljets_offshell_region}.
 

### 2022-06-14 main/chapter_2ljets.tex
grammar check

@@
 Despite the expert input from these off-shell and compressed searches, it was
-decided that $\twoljets$-electroweak should chime in to the off-shell
+decided that $\twoljets$-electroweak should join the off-shell
 conversation.
@@
 Our off-shell regime is characterized by $\mll < m(Z)$.
-The $W$ must also be off shell, the softer hadronic activity from light $W^*$
-events is not reliably reconstructed as signal jets.
+The $W$ boson must also be off shell, the softer hadronic activity from
+light $W^*$ events is not reliably reconstructed as signal jets.
 
@@
 Our simulations of $\zjets$ are split into on- and off-shell parts at
-$40\,\eV[G]$, so that forms a natural breakpoint for binning which might help
+$40\,\eV[G]$, so that forms a natural breakpoint for binning, which might help
 to separate concerns about systematic variations.
-As always, this binning aids sensitivity since models' contributions to the
-the bins vary with their mass splittings.
+As always, this binning aids sensitivity, since models' contributions to the
+bins vary with their mass splittings.
 
@@
 Pushing it up to $\mttwoll > 100\,\eV[G]$ was found to work well for the signal
-region, and the $\mttwoll \in 80\textrm{--}100\,\eV[G]$ nicely forms a the
+region, and the $\mttwoll \in 80\textrm{--}100\,\eV[G]$ nicely forms the
 nearby validation region.

### 2022-06-15 main/chapter_2ljets.tex
improve crdy histogram; add numerical simplicity rant

@@
 
+My understanding of search design is as conflict between simplicity and
+precision.
+Yes, an extra bin with higher $\metsig$ might catch some extreme signals,
+and yes, an RJR construction might pick out candidate decay trees without
+extreme $\met$.
+But if these delay the result without comparable benefits, they will act to
+slow down our scientific progress.
+
+Furthermore, simplicity encourages interpretability and trust.
+A selection at $100\,\eV[G]$ may be slightly outperformed in precision by one
+at $100.314\,\eV[G]$, but the former is preferable --- it is easily
+communicated and perhaps understood or reproduced by readers.
+The latter could be overly tuned, perhaps by a learning algorithm, to specific
+simulations.
+Perhaps it cut out a highly weighted sample which, although annoying, was an
+honest representation of our modelling.
+Denying ourselves the ability to over-fit simulations keeps us honest.
+
+Numerical simplicity is valuable.
+I stated this as a design goal in internal documentation, but that statement was
+removed after one \atlas\ author commented
+``maybe this paragraph is a little
+\emph{too honest}? (The `numerical simplicity' bit)''
+(emphasis added)~\cite{comment2020toohonest}.
+
 This section describes the designs and rationales behind the control,
-validation and signal regions of the $\twoljets$ analysis.
+validation, and signal regions of the $\twoljets$ analysis.
 
@@
 
-Following the conventional style descibed in
+Following the conventional style described in
 Section~\ref{sec:searches_searches},
@@
 All regions are defined within a loose `Pre-selection' requirement which
-requires the basic $Z\rightarrow \ell^+\ell^-$ plus jets pluts $\met$
+requires the basic $Z\rightarrow \ell^+\ell^-$ plus jets plus $\met$
 conditions that we target in the $\twoljets$-electroweak.
@@
 The $\rll$ bins of SR-High are labelled SR-High-8 and SR-High-16, and their
-$\metsig$ binns are suffixed -a and -b for $\metsig \in 18\textrm{--}21$
+$\metsig$ bins are suffixed -a and -b for $\metsig \in 18\textrm{--}21$
 and $\metsig > 21$ respectively.
@@
 The significance dropouts in (a) are errors in plotting.
-`Significance Z' is the binomial significance with $30\%$ relative background
-uncertainty~\cite{cousins2008evaluation}.
+`Significance Z' is the binomial significance~\cite{cousins2008evaluation}
+with a $30\%$ relative background uncertainty.
 Error bars are statistical.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.png}
+\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.pdf}
 \caption{CR-DY, $\metsig$}
@@
 
-The control region CR-DY (named after Drell and Yan) is defined to normalize off-shell $\zyjets$ near to
-SR-OffShell.
+The control region CR-DY (named after Drell and Yan) is defined to normalize
+off-shell $\zyjets$ near to SR-OffShell.
 Its reduced $\metsig$ requirement and relaxed angular selections were required
@@
 designs, the general message has not changed.
-Values of CLs from mean data in the combined fit are are overlaid on the grid.
+Values of CLs from mean data in the combined fit are overlaid on the grid.
 }

### 2022-06-15 main/chapter_2ljets.tex
remove May; improve intro paragraph

@@
 
-This section describes the designs and rationales behind the control,
-validation, and signal regions of the $\twoljets$ analysis.
+The remainder of this section describes the designs of the control,
+validation, and signal regions of the $\twoljets$ analysis, along with some
+reasoning behind those designs.
 

### 2022-06-15 main/chapter_2ljets.tex
improve numerical simp paragraph
yet another internal doc sentence rearrangement

@@
 honest representation of our modelling.
-Denying ourselves the ability to over-fit simulations keeps us honest.
+Restricting ourselves to round numbers not only clarifies our communications,
+but keeps us honest by denying ourselves the ability to over-fit simulations.
 
 Numerical simplicity is valuable.
-I stated this as a design goal in internal documentation, but that statement was
-removed after one \atlas\ author commented
+In internal documentation, I stated that it was a motivator behind the design,
+but that statement was removed after one \atlas\ author commented:
 ``maybe this paragraph is a little

### 2022-06-22 main/chapter_2ljets.tex
write discovery section

@@
 
+Although discovery regions are model-independent in principle, our designs
+remain inspired by our simulated signal models.
+Various approaches are taken to discovery regions in \atlas\ searches.
+Although it is normal to have one per signal
+regions~\cite{SUSY-2018-02},
+some searches study
+dozens of discovery regions comprising signal region bins~\cite{SUSY-2018-36},
+or other modifications to signal
+regions~\cite{atlas_susy_compressed_2l_2018_run2}.
+
+Personally, I do not see the benefit from tens of discovery regions.
+Yes, given a pet model, one can choose the region with its best expected
+sensitivity and interpret the model with the data in that region.
+But one cannot easily combine data from multiple regions if their background
+modelling is meaningfully related.
+And our background modelling will always be related if our discovery regions
+share control region data, or other coupled systematic variations.
+
+That is, finely binning discovery regions can harm sensitivity since their
+interpretations are decoupled.
+Since we do couple the background models in exclusion studies,
+opposite is true (to some extent) there.
+
+We therefore chose to design a small number of discovery regions, each of which
+comprises a union of one or more signal regions.
+
+This decision turned out well, not because of the above statistical
+pontification, but since the practical cost of running the standard statistical
+software to analyse each discovery region (and to perform their many validation
+studies), turned out to be so large~\cite{verkerke2003roofit}.
+
+Given our existing division into High, Low, Int, and OffShell
+categories, the assignment one discovery region to each was simple.
+The $\llbb$ category within High also stands out for targetting physically
+different events, so a DR-$\llbb$ was also assigned.
+
+Our five discovery regions are defined as unions on signal regions as follows:
 \begin{itemize}
-\item DR-High: SR-High-8
-\item DR-$\llbb$: SR-$\llbb$
-\item DR-Int: SR-Int
-\item DR-Low: SR-Low
-\item DR-OffShell: SR-OffShell
+\item DR-High is SR-High-8 (the union of SR-High-8-a and -b),
+\item DR-$\llbb$ is is SR-$\llbb$,
+\item DR-Int is SR-Int (the union of SR-Int-a and -b),
+\item DR-Low is SR-Low (the union of SR-Low-a and -b)
+\item DR-OffShell is SR-OffShell (the union of SR-OffShell-a and -b).
 \end{itemize}
-
-Figure~\ref{fig:2ljets_disco_trials_high}
-Figure~\ref{fig:2ljets_disco_trials_llbb}
-Figure~\ref{fig:2ljets_disco_trials_int}
-Figure~\ref{fig:2ljets_disco_prefit}
+Although it was not a design requirement, these give discovery regions are
+all orthogonal.
+
+Most signal regions are included as parts of these discovery regions, with the
+exceptions of SR-High-1J, SR-High-16, and SR-Low-2.
+These exceptions were excluded since they have either questionable background
+modelling or poor sensitivity to candidate signal models.
+
+To arrive at this design, I proposed many combinations of signal regions and
+compared metrics for their projected sensitivity to simulated signal models.
+The final choices are not the most sensitive to all models --- rather, they
+were chosen for broad sensitivity to varieties of signals.
+These studies are displayed in Figures~\ref{fig:2ljets_disco_trials_high},
+~\ref{fig:2ljets_disco_trials_llbb}~and~\ref{fig:2ljets_disco_trials_int}
+for High, $\llbb$ and Int categories respectively.
+Similar figures were produced for the other categories, but they are used older
+region definitions, are identical in style, and are not illuminating
 
@@
 sensitivity~\cite{cousins2008evaluation}.
-Yields are derived in an early iteration of the analysis and do not exactly
+Yields were derived in an early iteration of the analysis and do not exactly
 match those in the final result.
@@
 \end{figure}
+
+As with other designs, simplicity has value.
+Although sharper sensitivity to special cases could have been achieved
+with more complicated selections, the reward of this discovery region design
+is that it can be stated clearly as the bullet pointed list given above.
+
+The five discovery regions are displayed in
+Figure~\ref{fig:2ljets_disco_prefit}
+with pre-fit backgrounds and statistical uncertainties only.
+The unblinded data in this figure repeat the SR-High and SR-Low deficits
+mentioned previously.
+to make inferences from these data, it is necessary to assign honestly
+uncertain background models; the next section tackles this crucial problem of
+background modelling.
 
@@
 
+\FloatBarrier
 \section{Modelling}

### 2022-06-27 main/chapter_2ljets.tex
begin modelling section; add hist* and pyhf references

@@
 \section{Modelling}
-% TODO backgorund samples
-% TODO matrix method fakes
-% TODO GMSB reweighting
-% TODO off-shell plus/minus split
-% TODO generator filtering
-
-Table~\ref{tab:2ljets_zjets_alt}
-
-\begin{table}[tp]
-\centering
-\begin{tabular}{lcc}
-& Central
-& Relative variations
-\\[0.5em]
-CR-DY
-& $59.64$
-& $1-0.1+0.1$
-\\[0.2em]
-\underline{CR-Z}
-& $101.95$
-& $1+0.2-0.2$
-\\[0.2em]
-CR-tt
-& $38.07$
-& $1+0.5-0.5$
-\\[0.2em]
-CR-VZ
-& $5.03$
-& $1+1.3-1.0$
-\\[0.5em]
-VR-Low
-& $24.19$
-& $1-0.3+0.3$
-\\[0.2em]
-VR-Low-2
-& $17.05$
-& $1+0.2-0.2$
-\\[0.5em]
-\underline{SR-Low-a}
-& $14.30$
-& $1-0.2+0.2$
-\\[0.2em]
-\underline{SR-Low-b}
-& $4.17$
-& $1-0.3+0.3$
-\\[0.2em]
-SR-Low-2
-& $1.51$
-& $1+1.5-1.0$
-\\[0.2em]
-\end{tabular}
-\caption[%
-Alternative $\zjets$ variations in selected regions
-]{%
-Alternative $\zjets$ variations in selected regions.
-The `Central' column shows the central $\zjets$ yield from the main sample.
-The other column shows the relative scale factors applied to this central yield
-for up and down variations.
-These are evaluated from the central and alternative samples as described in
-the text.
-Note how up and down variations can move in different directions.
-}
-\label{tab:2ljets_zjets_alt}
-\end{table}
+
+So far, we have seen many histograms of background samples overlaid with
+observed data.
+Those backgrounds may or may not have appeared to predicted the data well, but
+no conclusions can be drawn until uncertainties are assigned to the background
+and signal modelling.
+
+To interpret the results seriously, we need to understand our uncertainties ---
+we do not know enough about the Standard Model to make definitive predictions,
+but must distribute our unit allotment of probability over various plausible
+background predictions.
+
+With such a complicated experiment as \atlas, in a universe with such varied
+particles as ours, there are many different factors of uncertainty to consider.
+As described in Section~\ref{sec:searches_practice}, the effect of each
+factor is studied to assign regularized variations to the Poisson expectations
+of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
+For the $\twoljets$ search, we used HistFitter to build our every HistFactory
+models~\cite{Besjes_2015,baak2015histfitter}.
+
+Thankfully, many standard variations are evaluated automatically with \atlas\
+software by simulating alternative samples with varied parameters, then
+comparing the resulting histograms.
+Such standard variations relate to performance and calibration effects, which
+have been carefully studied by many researchers in \atlas, to whom I am deeply
+grateful.
+They cover variations on our objects --- jets, leptons and $\ptmiss$, as well
+as flavour tagging and pileup effects.
+
+Statistical uncertainties on simulated samples are also semi-automatic.
+In each bin, its statistical uncertainty is controlled by a single parameter,
+which intends to cover the uncertainties on all backgrounds samples
+contributing in that bin.
+An alternative prescription exists which gives one parameter per sample per
+bin, but this greatly increases the complexity of the model and was not
+considered necessary.
+
+The $\twoljets$-electroweak search implements a total of $136$ modifiers its
+background-only model, and more for each signal model.
+Of those I evaluated more manually, some of the more interesting or important
+will be discussed in the remainder of this section ---
+Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
+alternative simulated samples in Section~\ref{sec:2ljets_sample_comparisons},
+signal modelling uncertainties in Section~\ref{sec:2ljets_signal_variations}
+and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.
 
@@
 & Prompt (stat)
-& mMC
-& mMM
-& mMean
-& \quad\underline{mMax}\hphantom{foo}
+& MC
+& MM
+& Mean
+& \quad\underline{Max}\hphantom{foo}
 \\[0.5em]
@@
 
-\section{Validation}
-\label{sec:2ljets_validation}
-Figure~\ref{fig:2ljets_low_crz_extensions}
-Figure~\ref{fig:2ljets_splits}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
-\caption{}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Et.png}
-\caption{}
-\end{subfigure}
-\caption[
-Extension from CR-Z into higher $\metsig$ and $\met$
-]{%
-Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
-The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
-upper bound of $\met$.
-Systematic uncertainties are extrapolated from CR-Z to the extended tail.
-The tail overlaps with other regions, but shows good agreement with data.
-Sub-figure (b) is reproduced from auxiliary material to the
-paper~\cite{atlas2022searches}.
-Each last bin includes overflow.
-}
-\label{fig:2ljets_low_crz_extensions}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZZ.png}
-\caption{CR-VZ, diboson}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZ.png}
-\caption{CR-Z, $\zjets$}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRtt.png}
-\caption{CR-tt, rare top}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_SRllbb.png}
-\caption{SR-$\llbb$, rare top}
-\end{subfigure}
-\caption[
-Subcomponents of diboson, $\zjets$ and rare top processes.
-]{%
-Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) rare top processes.
-Target samples broken down by simulated sample identity codes.
-The large three-lepton component in CR-VZ motivated its renaming from CR-ZZ.
-The `no filter' selection comprises $\zjets$ samples with unconstrained jet
-flavour.
-Errors are statistical.
-}
-\label{fig:2ljets_splits}
-\end{figure}
-
-\section{Variations}
-% "systematic"
+\subsection{Sample comparisons}
+\label{sec:2ljets_sample_comparisons}
 Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
@@
 Figure~\ref{fig:2ljets_low_vrlow_pre_post}
+
+Table~\ref{tab:2ljets_zjets_alt}
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{lcc}
+& Central
+& Relative variations
+\\[0.5em]
+CR-DY
+& $59.64$
+& $1-0.1+0.1$
+\\[0.2em]
+\underline{CR-Z}
+& $101.95$
+& $1+0.2-0.2$
+\\[0.2em]
+CR-tt
+& $38.07$
+& $1+0.5-0.5$
+\\[0.2em]
+CR-VZ
+& $5.03$
+& $1+1.3-1.0$
+\\[0.5em]
+VR-Low
+& $24.19$
+& $1-0.3+0.3$
+\\[0.2em]
+VR-Low-2
+& $17.05$
+& $1+0.2-0.2$
+\\[0.5em]
+\underline{SR-Low-a}
+& $14.30$
+& $1-0.2+0.2$
+\\[0.2em]
+\underline{SR-Low-b}
+& $4.17$
+& $1-0.3+0.3$
+\\[0.2em]
+SR-Low-2
+& $1.51$
+& $1+1.5-1.0$
+\\[0.2em]
+\end{tabular}
+\caption[%
+Alternative $\zjets$ variations in selected regions
+]{%
+Alternative $\zjets$ variations in selected regions.
+The `Central' column shows the central $\zjets$ yield from the main sample.
+The other column shows the relative scale factors applied to this central yield
+for up and down variations.
+These are evaluated from the central and alternative samples as described in
+the text.
+Note how up and down variations can move in different directions.
+}
+\label{tab:2ljets_zjets_alt}
+\end{table}
 
@@
 
+\subsection{Signals}
+\label{sec:2ljets_signal_variations}
+
+% TODO GMSB reweighting
+% TODO off-shell plus/minus split
+% TODO xsec variations
+
+\subsection{Theory}
+\label{sec:2ljets_theory_variations}
+
 % extrapolation
-
 The total theory variation is a combination of seven components representing
@@
 
+\section{Validation}
+\label{sec:2ljets_validation}
+Figure~\ref{fig:2ljets_low_crz_extensions}
+Figure~\ref{fig:2ljets_splits}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Sign.pdf}
+\caption{}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_met_EWK_met_Et.png}
+\caption{}
+\end{subfigure}
+\caption[
+Extension from CR-Z into higher $\metsig$ and $\met$
+]{%
+Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
+The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
+upper bound of $\met$.
+Systematic uncertainties are extrapolated from CR-Z to the extended tail.
+The tail overlaps with other regions, but shows good agreement with data.
+Sub-figure (b) is reproduced from auxiliary material to the
+paper~\cite{atlas2022searches}.
+Each last bin includes overflow.
+}
+\label{fig:2ljets_low_crz_extensions}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZZ.png}
+\caption{CR-VZ, diboson}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZ.png}
+\caption{CR-Z, $\zjets$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRtt.png}
+\caption{CR-tt, rare top}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_SRllbb.png}
+\caption{SR-$\llbb$, rare top}
+\end{subfigure}
+\caption[
+Subcomponents of diboson, $\zjets$ and rare top processes.
+]{%
+Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) rare top processes.
+Target samples broken down by simulated sample identity codes.
+The large three-lepton component in CR-VZ motivated its renaming from CR-ZZ.
+The `no filter' selection comprises $\zjets$ samples with unconstrained jet
+flavour.
+Errors are statistical.
+}
+\label{fig:2ljets_splits}
+\end{figure}
+
 \section{Unblinding}

### 2022-06-27 main/chapter_searches.tex
squeeze significance to fit width
splay out big significance measure a little
add first draft high region def text
begin modelling section; add hist* and pyhf references

@@
 \section{Searches}
+\label{sec:searches_searches}
 % histograms, CR VR SR, systematics
@@
 \section{Frequentist theory}
+\label{sec:searches_frequentism}
 \begin{epigraphs}
@@
 \section{Post-frequentist practice}
+\label{sec:searches_practice}
 \begin{epigraphs}
@@
 
+% histfactory
+
+To model the consequences of these many factors of uncertainty, each factor is
+assigned a real-valued parameter which when varied changed the expected yields
+in the hisogram bins of the search.
+In this way, we form a statistical model which is a function
+\begin{equation}
+H_{\!f}(\vec \theta) \rightarrow (\vec \mu, \log f)
+\end{equation}
+from the parameters $\theta$ to Poisson expectations $\vec\mu$ the histogram
+bins and a regularization term $\log f$ which is maximized in fitting.
+
+HistFactory~\cite{cranmer2012histfactory}
+pyhf~\cite{heinrich2021pyhf}
+HistFitter~\cite{Besjes_2015,baak2015histfitter}
+
 % significancen measures
-The ATLAS-recommended~\cite{atlas_significance} significance measure is
-\begin{equation}
+The ATLAS-recommended significance measure~\cite{atlas_significance} is
+\begin{align}
 \label{eqn:significance_atlas}
-S_\mathrm{ATLAS}(n, \mu, \sigma)
-= \mathrm{sgn}(n - \mu) \sqrt{2\left(
+S_\mathrm{\atlas}(n; \mu, \sigma) =~&
+\mathrm{sgn}(n - \mu) \times \sqrt{2} \\[0.2em] \nonumber
+&
+\times \sqrt{
 n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
@@
 \right)
-\right)}
-~.
-\end{equation}
+}
+.
+\end{align}
 

### 2022-06-27 main/chapter_2ljets.tex
smash out some matrix method theory and opinions

@@
 
+% TODO jetM
+\TODO{JET MASS}
+
 \subsection{Fake/non-prompt leptons}
 \label{sec:2ljets_mm_fakes}
+
+\TODO{composition (INT)}
+% components
+% lots of non-prompt, a little fake hadrons
+%
+
+Historical precedent leads us to use the `Matrix Method' to estimate
+backgrounds including fake and non-prompt leptons, so named because it works by
+estimating a linear relationship between dilepton event rates.
+\TODO{previous 2ljets MM}
+
+In data, one measures tightly ($t$) and loosely ($l$) identified leptons, leading
+to four categories of events with two ordered leptons as labelled $tt$, $tl$,
+$lt$, and $ll$ for tight-tight, tight-loose, and so on.
+The Matrix Method assumes a linear relationship between these selections
+and those of real ($r$) and fake/non-prompt ($f$) leptons:
+\begin{equation}
+\begin{pmatrix}
+\mu^\textrm{data}_{tt} \\
+\mu^\textrm{data}_{tl} \\
+\mu^\textrm{data}_{tl} \\
+\mu^\textrm{data}_{ll} \\
+\end{pmatrix}
+=
+M
+\begin{pmatrix}
+\mu^\textrm{truth}_{rr} \\
+\mu^\textrm{truth}_{fr} \\
+\mu^\textrm{truth}_{rf} \\
+\mu^\textrm{truth}_{ff} \\
+\end{pmatrix}
+,
+\end{equation}
+where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the matrix of
+transition probabilities after which this method is
+named~\cite{ATLAS-CONF-2014-058}.
+
+Each element of $M$ is a probability for a truth-level (real/fake) event to be
+also in a corresponding data-level (tight/loose) category.
+For example, its top right element $M_{{tt}, {rr}}$ is the
+probability that an event with two real leptons has them both be identified
+tightly.
+The leptons are further assumed to be reconstructed independently, such that
+$M_{{tt}, {rr}} =
+p(t\!\mid\!r,1)
+\times p(t\!\mid\!r,2)$
+in which the numerical index points to the first and second leptons
+respectively.
+
+Transition probabilities $p(*\!\mid\!*,i)$ are estimated by studies in
+separate selections and assigned as functions of the lepton
+flavour, $\pt$ and $|\eta|$.
+Given those estimated transition probabilities and some observed data
+in a selection of two loose leptons,
+$n^\textrm{data}_{tt}$,
+$n^\textrm{data}_{tl}$,
+$n^\textrm{data}_{tl}$,
+and $n^\textrm{data}_{ll}$,
+the Matrix Method inverts $M$ to estimate truth level rates
+$\bar \mu^\textrm{truth}_{**}$.
+(Recall that tight implies loose, so this selection of loose leptons includes
+some tight leptons.)
+Finally, these truth-level estimates are converted back to data-level
+estimates in the tight-tight selection of our search regions back through the
+independent transition probabilities as
+\begin{align}
+% rr
+\bar \mu^\textrm{data}_{tt\mid rr} &=
+p(t\!\mid\!r,1)
+\times p(t\!\mid\!r,2)
+\times \bar \mu^\textrm{truth}_{rr}~~, \\
+% fr
+\bar \mu^\textrm{data}_{tt\mid fr} &=
+p(t\!\mid\!f,1)
+\times p(t\!\mid\!r,2)
+\times \bar \mu^\textrm{truth}_{fr}~~, \\
+% rf
+\bar \mu^\textrm{data}_{tt\mid rf} &=
+p(t\!\mid\!r,1)
+\times p(t\!\mid\!f,2)
+\times \bar \mu^\textrm{truth}_{rf}~~,~~\textrm{and} \\
+% ff
+\bar \mu^\textrm{data}_{tt\mid ff} &=
+p(t\!\mid\!f,1)
+\times p(t\!\mid\!f,2)
+\times \bar \mu^\textrm{truth}_{ff}~~.
+\end{align}
+
+To assign fake/non-prompt estimates in our regions, this method is
+implemented by assigning weights to observed data events, which are
+histogrammed in the tighter selections of our search
+regions~\cite{twoljets2018int}.
+Systematic variations are assigned by comparing histograms when those weights
+are varied according to various uncertainties in the Matrix Method procedure.
+
+Yes, that is a histogramming of real data with weights, and those data include
+tight-tight events which comprise the signal data of our search regions,
+in which we are searching for anomalies.
+Since it includes real data, this method looks at all unblinded data, including
+signal regions.
+And because it knows the data we are supposed to be predicting, it is no
+prediction at all!
+
+That last statement was an exaggeration.
+Although it does cheat by using the data we should be blind to, it uses those
+data in a diluted way, such that they which might end up not making much
+difference.
+
+In fact, real data tend to be assigned negative Matrix Method weights.
+For example, CR-VZ receives $157$ weighted data in its Matrix Method sample,
+of which $135$ pass the signal lepton requirements (tight-tight)%
+\footnote{%
+I am not sure why this differs from the $194$ actual data observed in CR-VZ.
+The two do overlap in event identity numbers, so there is presumably some
+tighter selection in the Matrix Method which should be accounted for in its
+weights.
+},
+and the remainder have one loose lepton (zero loose-loose lepton events
+are observed here).
+The signal events have an average weight of $-0.0227$, and the others
+have an average weight of $+0.238$ (one can check these recover the central
+fake/non-prompt background of $2.2$).
+Such negative signal weights suggest that a local \emph{excess} of data
+would act to \emph{decrease} the Matrix Method estimate in the region.
+Yes this backwards dependency is backwards, but this gives some relief that it
+should not hide real local excesses.
+
+Although the $\twoljets$-electroweak regions were designed without using the
+Matrix Method fake estimate, and before it was finalized, we are fortunate that
+most search regions do not have substantial Matrix Method yields.
+I consider those that do to be dubious.
+
+% TODO cite past searches
+% TODO thank Eirik
 

### 2022-06-27 main/chapter_2ljets.tex
expand fnp intro, thank eirik, improve text

@@
 
-\TODO{composition (INT)}
-% components
-% lots of non-prompt, a little fake hadrons
-%
-
-Historical precedent leads us to use the `Matrix Method' to estimate
-backgrounds including fake and non-prompt leptons, so named because it works by
-estimating a linear relationship between dilepton event rates.
-\TODO{previous 2ljets MM}
+Given our requirements on lepton quality, isolation, and hardness, it is
+expected that the rates of fake leptons, or non-prompt leptons (which do not
+originating from the hard scatter) should be small.
+And indeed more details estimates indicate that they are small.
+However, just like other backgrounds, it is important that we assign
+honest and honestly uncertain estimates for these fake/non-prompt backgrounds.
+
+Although my style of spelling out fake/non-prompt every time
+(as opposed to just `fake') may be awkward, it is supported by fact that this
+the non-prompt part dominates.
+Indeed, figures in our internal documentation~\cite{twoljets2018int} suggest
+that, in simulations, non-prompt hadron decays produce in excess of $90\%$ of
+this background in a loose selections near to our regions, and across the
+entire $\metsig$ spectrum.
+Non-prompt muons are mostly from decays of heavy-flavour ($b$ or $c$) hadrons,
+and electrons arise in roughly equal parts from light flavour hadrons.
+Tau decays form the next largest part, with prompt fakes due to misidentified
+photons or leptons in the sub-$\%$ weeds.
+
+However, simulations are not totally accurate.
+In particular, do not use simulations of pure hadronic processes, so do not
+really know the rates at which these strong processes can fake leptons.
+Although prompt strong effects may only rarely fake leptons, they have very
+large cross-sections are not modelled well.
+This motivates the use of data-driven fake estimates to replace the need for
+such specialized simulations of poorly understood physics.
+
+Precedent from previous searches in this channel~\cite{
+atlas_2l_SUSY_2013_11,
+atlas_23l_SUSY_2016_24
+}
+leads us to use the `Matrix Method'~\cite{ATLAS-CONF-2014-058} to estimate
+backgrounds of fake and non-prompt leptons, so named because it works by
+estimating a linear relationship between dilepton event rates, which is
+written as a matrix multiplication.%
+\footnote{%
+To be explicit, the Matrix Method does not use Matrix Elements from quantum
+scattering, and is not related to the similarly-named
+Matrix Element Method~\cite{gainer2013matrix}.
+}
+
+\emph{
+I am deeply grateful to Eirik~Gramstad for developing the Matrix Method for our
+$\twoljets$ searches, documenting it, providing its samples, and showing us how
+to use them.
+Although I have reservations about the method itself (expounded below), these
+are not criticisms of his excellent work spent in using it.
+}
 
@@
 \begin{equation}
+\label{eqn:2ljets_matrix_method}
 \begin{pmatrix}
@@
 flavour, $\pt$ and $|\eta|$.
-Given those estimated transition probabilities and some observed data
-in a selection of two loose leptons,
+Given those estimated transition probabilities, the Matrix Method begins by
+taking some observed data
 $n^\textrm{data}_{tt}$,
@@
 $n^\textrm{data}_{tl}$,
-and $n^\textrm{data}_{ll}$,
-the Matrix Method inverts $M$ to estimate truth level rates
-$\bar \mu^\textrm{truth}_{**}$.
+and $n^\textrm{data}_{ll}$
+in a selection of two loose leptons.
 (Recall that tight implies loose, so this selection of loose leptons includes
 some tight leptons.)
+The Matrix method uses these data as estimate $\mu^\textrm{data}_{**}$,
+then inverts $M$ to estimate the truth level rates
+$\bar \mu^\textrm{truth}_{**}$ through Equation~\ref{eqn:2ljets_matrix_method}.
+
 Finally, these truth-level estimates are converted back to data-level
@@
 
-To assign fake/non-prompt estimates in our regions, this method is
-implemented by assigning weights to observed data events, which are
-histogrammed in the tighter selections of our search
+To assign fake/non-prompt estimates in our regions, the Matrix Method is
+implemented by assigning weights to observed data events, and histogramming
+them in the tighter selections of our search
 regions~\cite{twoljets2018int}.
@@
 
-Yes, that is a histogramming of real data with weights, and those data include
-tight-tight events which comprise the signal data of our search regions,
-in which we are searching for anomalies.
-Since it includes real data, this method looks at all unblinded data, including
-signal regions.
+Yes, actual data are histogrammed in the Matrix Method, and those data include
+tight-tight events which comprise the data of our search regions, and those
+are the same data in which we are searching for anomalies.
+Since it includes actual data, this method unblinds all regions,
+including signal regions, early.
 And because it knows the data we are supposed to be predicting, it is no
@@
 
-That last statement was an exaggeration.
+That last statement was a slight exaggeration.
 Although it does cheat by using the data we should be blind to, it uses those
-data in a diluted way, such that they which might end up not making much
-difference.
+data in a diluted way, such that they can end up not making much difference.
 
 In fact, real data tend to be assigned negative Matrix Method weights.
-For example, CR-VZ receives $157$ weighted data in its Matrix Method sample,
-of which $135$ pass the signal lepton requirements (tight-tight)%
+For example, CR-VZ receives $157$ data in its Matrix Method sample,%
 \footnote{%
-I am not sure why this differs from the $194$ actual data observed in CR-VZ.
+This count of $157$ differs from the $194$ actual data observed in CR-VZ.
+I do not know why.
 The two do overlap in event identity numbers, so there is presumably some
@@
 weights.
-},
+}
+of which $135$ pass the signal lepton requirements (tight-tight),
 and the remainder have one loose lepton (zero loose-loose lepton events
-are observed here).
-The signal events have an average weight of $-0.0227$, and the others
+are observed in CR-VZ).
+The tight-tight events have an average weight of $-0.0227$, and the others
 have an average weight of $+0.238$ (one can check these recover the central
@@
 would act to \emph{decrease} the Matrix Method estimate in the region.
-Yes this backwards dependency is backwards, but this gives some relief that it
-should not hide real local excesses.
+Yes this backwards dependency is backward, but at least it gives some relief
+hide real local excesses would not be hidden by increased fake estimates.
 
@@
 
-% TODO cite past searches
-% TODO thank Eirik
-
+% TODO
 Figure~\ref{fig:2ljets_vrss}

### 2022-06-27 main/chapter_2ljets.tex
some spelling grammar check

@@
 Since we do couple the background models in exclusion studies,
-opposite is true (to some extent) there.
+the opposite is true (to some extent) there.
 
@@
 observed data.
-Those backgrounds may or may not have appeared to predicted the data well, but
-no conclusions can be drawn until uncertainties are assigned to the background
-and signal modelling.
+Those backgrounds may or may not have appeared to have predicted the data well,
+but no conclusions can be drawn until uncertainties are assigned to the
+background and signal modelling.
 
@@
 With such a complicated experiment as \atlas, in a universe with such varied
-particles as ours, there are many different factors of uncertainty to consider.
+particles as ours, there are many and varied factors of uncertainty to consider.
 As described in Section~\ref{sec:searches_practice}, the effect of each
@@
 
-The $\twoljets$-electroweak search implements a total of $136$ modifiers its
+The $\twoljets$-electroweak search implements a total of $136$ modifiers in its
 background-only model, and more for each signal model.
@@
 originating from the hard scatter) should be small.
-And indeed more details estimates indicate that they are small.
+Indeed, more detailed estimates indicate that they are small.
 However, just like other backgrounds, it is important that we assign
@@
 Although my style of spelling out fake/non-prompt every time
-(as opposed to just `fake') may be awkward, it is supported by fact that this
+(as opposed to just `fake') may be awkward, it is supported by the fact that this
 the non-prompt part dominates.
@@
 that, in simulations, non-prompt hadron decays produce in excess of $90\%$ of
-this background in a loose selections near to our regions, and across the
+this background in a loose selection near to our regions, and across the
 entire $\metsig$ spectrum.
@@
 where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the matrix of
-transition probabilities after which this method is
+transition probabilities, after which this method is
 named~\cite{ATLAS-CONF-2014-058}.
@@
 The two do overlap in event identity numbers, so there is presumably some
-tighter selection in the Matrix Method which should be accounted for in its
+tighter selection in the Matrix Method, which should be accounted for in its
 weights.
@@
 non-positive Matrix Method estimates.
-The `MC FNP' yields are selected using truth information in sumulated samples.
+The `MC FNP' yields are selected using truth information in simulated samples.
 The `MM FNP' yields are from the Matrix Method, with statistical errors from

### 2022-07-20 main/chapter_2ljets.tex
smash out some more text on fake estimates

@@
 
+\subsubsection{The Matrix Method}
 Precedent from previous searches in this channel~\cite{
@@
 \end{equation}
-where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the matrix of
-transition probabilities, after which this method is
-named~\cite{ATLAS-CONF-2014-058}.
+where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the
+(left-stochastic) matrix of transition probabilities, after which this method
+is named~\cite{ATLAS-CONF-2014-058}.
 
@@
 \end{align}
-
-To assign fake/non-prompt estimates in our regions, the Matrix Method is
-implemented by assigning weights to observed data events, and histogramming
-them in the tighter selections of our search
-regions~\cite{twoljets2018int}.
+To make fake/non-prompt estimates in our regions, the Matrix Method is
+implemented by assigning weights to observed data events that are then
+histogrammed them in the tighter selections of our search
+regions much like other samples~\cite{twoljets2018int}.
 Systematic variations are assigned by comparing histograms when those weights
@@
 Yes this backwards dependency is backward, but at least it gives some relief
-hide real local excesses would not be hidden by increased fake estimates.
+that real local excesses would not be hidden by increased fake estimates.
+Unless, that is, the processes driving that excess also produced loose leptons
+at rates similar to (or larger than) the standard model backgrounds.
+Loose data do have positive weights and they would drive up the fake/non-prompt
+background.
 
@@
 
-% TODO
-Figure~\ref{fig:2ljets_vrss}
-Table~\ref{tab:2ljets_fnp_negative_mm}
-Table~\ref{tab:2ljets_fnp_estimates}
+\subsubsection{Fake/non-prompt estimates in $\twoljets$-electroweak regions}
+Many histograms were inspected to validate the Matrix Method estimates in
+$\twoljets$-electroweak regions.
+To give a broad summary across all of these regions, we define an additional
+validation region, named VR-SS, to inspect similar data with same-charged
+(same sign, SS) leptons.
+This selection enhances fake/non-prompt contributions relative backgrounds
+since it suppresses contributions $Z\rightarrow \ell^\pm\ell^\mp$ processes,
+whose leptons should be oppositely charged if accurately reconstructed,
+but does not suppress fake/non-prompt backgrounds because their lepton charges
+are expected to be less dependent:
+fake/non-prompt backgrounds produce $\ell^\mp\ell^\mp$ and $\ell^\pm\ell^\pm$
+vents at more similar rates.
+
+For these reasons, VR-SS is defined as a modification to the pre-selection
+requirements of Table~\ref{tab:2ljets_presel}, in which the requirement
+of opposite charge leptons is changed to a requirement of same charge leptons
+(same sign, SS). Additionally, VR-SS requires $\mll \in (12, 111)~\eV[G]$,
+to cover both on- and off-shell selections.
+Histogrammed $\metsig$ distributions in VR-SS are displayed in
+Figure~\ref{fig:2ljets_vrss},
+with fake/non-prompt estimates taken from the Matrix Method in
+Figure~\ref{fig:2ljets_vrss_mm},
+and from background simulations in
+Figure~\ref{fig:2ljets_vrss_mc}.
+When using the Matrix Method, we use parton-level `truth' information to
+identify and remove remove fake/non-prompt contributions from simulations.
+This is why the simulated backgrounds (not `FNP leptons') increase by around
+$30\%$ when switching off the Matrix Method; the difference between the
+two orange histograms is the simulated fake/non-prompt estimate.
+
+From this figure, it is clear that the Matrix Method has overestimated
+the background rate in this region, and that the raw simulations predict
+the data more accurately.
+The difference is less substantial than it may appear, however,
+because the Matrix Method error bars are highly correlated --- they combine
+(in quadrature)
+statistical uncertainties with variations due to uncertainties in the
+Matrix Method weights.
+Matrix Method weight variations are parametrized with a single parameter
+which couples the yields in all bins, so moving it down by $1$-sigma in a fit
+would easily pull the backgrounds back into agreement with data.
+
+Drafts of the $\twoljets$ paper included
+Figure~\ref{fig:2ljets_vrss_mm} and a similar plot for the
+$\twoljets$-strong search, and repeated this point that the deviation was not
+very substantial.
+These plots were removed at the request of internal reviewers, who argued that
+they did not need to be shown since the fake/non-prompt background is not
+very important.
+To demonstrate that it was not important, we were first instructed to repeat
+the exclusion fits of the $\twoljets$-electroweak analysis with an additional
+uncertainty, to further cover the deviation observed in this figure, which
+scaled the Matrix Method yields by a $\pm 50\%$ scale factor.
+The utility of this is somewhat questionable since that variation is coupled
+between all bins, so it could not vary to explain an excess in one region
+without also increasing the fitted Matrix Method backgrounds in all other
+regions, and it would struggle to make any significant difference.
+Indeed, the result was an insignificant change to all exclusion contours;
+the contours were almost indistinguishable from those shown in
+Figures~\ref{fig:2ljets_contours_c1n2} and~\ref{fig:2ljets_contours_gmsb}.
+
+Overlaying these modified contours highlighted some minor differences, the
+largest being $2\%$ reduction in the limit on
+$B(\neutralino_1 \rightarrow h \gravitino)$ at
+$m(\neutralino_1) \approx 500\,\eV[G]$.
+Different interpolation options induce larger variations, as do different
+random seeds for the fits, so this really is an insignificant difference.
+
+Although we nominally use the Matrix Method to estimate fakes, its results
+are unusable in numerous regions, since they are either negative or zero.
+Details of these eleven problematic regions are displayed in
+Table~\ref{tab:2ljets_fnp_negative_mm}.
+For comparison, this table also includes the simulated background without
+fake/non-prompt leptons under the heading `MC Prompt' and the, and only
+the simulated fake/non-prompt contributions under the heading `MC FNP'.
+Negative or void estimates are an easy consequence of the use of signed
+weights of data used by the Matrix Method, but we require alternative
+estimates to assign a sensible background model.
+
+To replace these failed Matrix Method estimates, our strategy is to take three
+alternative estimates that assign central and high variations to the
+fake/non-prompt background estimate, and consider a maximum across them for
+each region.
+More precisely, we take the maximum central value and add it to the maximum
+difference (between central and high) for the combined high variation, and
+sets the low ($-1$-sigma) variation to zero.
+These three alternative estimates are reported with their combination in
+Table~\ref{tab:2ljets_fnp_estimates}.
+
+These three estimates are labelled `MC', `MM', `Mean', and we combine them into
+the `Max' estimate.
+Each is described here.
+\begin{itemize}
+\item The `MC' estimate takes the simulated estimate that was previously reported
+in Table~\ref{tab:2ljets_fnp_negative_mm}.
+\item The `MM' estimate modifies the Matrix Method result,
+also from Table~\ref{tab:2ljets_fnp_negative_mm}, by shifting its central
+value to a small positive value chosen baselessly as $0.01$ and by adding
+its statistical and systematic uncertainties linearly
+(not in quadrature to conservatively cover a correlated variation)
+to assign a high variation.
+\end{itemize}
+After I had proposed a combination of these two estimates, reviewers
+suggested additionally using the statistical properties of samples in looser
+selections.
+This became the third `Mean' estimate.
+\begin{itemize}
+\item The `Mean' follows an idea that the smallest assignable value should
+be comparable to typical Matrix Method weights, and those typical values
+are extracted from a region requiring pre-selection,
+from Table~\ref{tab:2ljets_presel},
+$\mll \in 71\textrm{--}101$,
+$\mjj \in 60\textrm{--}100$,
+$\njet \geq 2$, and
+$\nbtag \leq 1$.
+Its central value takes the mean weight of Matrix Method samples in this
+selection, which is $\mu^\textrm{M} = 0.0202$.
+Statistical errors are set to the root of the mean squared weight,
+which is $\sigma^\textrm{M}_\textrm{stat} = 0.1132$,
+and systematic errors are set as the larger absolute deviation on the mean
+from the up and down variations in the Matrix Method weights, which comes out
+$\sigma^\textrm{M}_\textrm{sys} = 0.0226$.
+Finally, the `Mean' high variation is their linear sum;
+$\mu^\textrm{M} + \sigma^\textrm{M}_\textrm{stat}
+\sigma^\textrm{M}_\textrm{sys} = 0.1560$.
+\end{itemize}
+All low variations are set to zero.
+
+As seen from Table~\ref{tab:2ljets_fnp_estimates}, even the high variations of
+the Max method are negligibly small next to the central simulated yields
+in most regions.
+The exception regions are all in the High category, and are those regions with
+small enough expected yields such that the Mean method dominates.
+
+An awkward fact in the history of this fake/non-prompt estimate is that it
+was reworked after unblinding to include the Mean estimate and more regions;
+with a reprocessing of our background simulations, changes in the Matrix Method
+estimates left more regions with non-positive estimate than previously.
+To assign them without being biased by the unblinded signal region data,
+known to contain no significant excesses, therefore took some barefaced
+honesty.
+Thankfully, we do not need to consider what might have happened had an excess
+been observed before the assignment of these fake estimates.%
+\footnote{%
+Of course $p$-values \emph{do} depend on what might have happened with
+alternative data
+(those with test statistics greater than or equal to that observed),
+it would be wasteful and unproductive to consider them seriously.
+I am grateful that \atlas' statistical conventions allow us to ignore these
+problems.
+}
 
@@
 \caption{}
+\label{fig:2ljets_vrss_mm}
 \end{subfigure}
@@
 \caption{}
+\label{fig:2ljets_vrss_mc}
 \end{subfigure}
@@
 ]{%
-Background yields from prompt and fake/non-prompt sources in bins with
-non-positive Matrix Method estimates.
+Background yields from prompt (two real, prompt leptons) and fake/non-prompt
+sources in bins with non-positive Matrix Method estimates.
 The `MC FNP' yields are selected using truth information in simulated samples.
@@
 ]{%
-Replacement fake estimates with up and down variations.
+Replacement fake estimates with up and down variations for the regions with
+unusable Matrix Method yields that were reported in
+Table~\ref{tab:2ljets_fnp_negative_mm}.
 Those variations cover statistical and systematic sources.
-The method `mMax' is used for the $\twoljets$-electroweak analysis;
+The `Max' method is used for the $\twoljets$-electroweak analysis;
 as described in the text, it uses the largest values of central, statistical
@@
 of zero.
+`Prompt' labels the simulated background with two true real, prompt leptons,
+and these are included for comparison.
 }

### 2022-07-21 main/chapter_2ljets.tex
begin jetM validation chunk

@@
 
+\emph{%
+Thanks are due to Knut~Vadla for important contributions in the initial
+exploration of signal region designs,
+to Ben Hooberman for suggesting the investigation of massive jets,
+and to both, and the whole $\twoljets$, collaboration in their finalization.
+}
+
 % presel table should come before High
@@
 in Figure~\ref{fig:2ljets_high_1J_region}.
+The use of $\mjetone$ with small-radius jets for this region is unusual,
+and required modelling of additional uncertainties that we discuss in
+Section~\ref{sec:2ljets_jet_mass}.
 
@@
 
-The distribution of mono-jet masses $\mjetone$ is sharply peaked towards zero,
+The distribution of mono-jet masses ($\mjetone$) is sharply peaked towards zero,
 so cutting this out while inverting the $\mjetone$ selection of SR-High-1J
@@
 
-% TODO jetM
-\TODO{JET MASS}
-
 \subsection{Fake/non-prompt leptons}
@@
 However, just like other backgrounds, it is important that we assign
-honest and honestly uncertain estimates for these fake/non-prompt backgrounds.
-
-Although my style of spelling out fake/non-prompt every time
+honest and honestly uncertain estimates for these fake/non-prompt (FNP)
+backgrounds.
+
+Although my style of spelling out FNP or FNP every time
 (as opposed to just `fake') may be awkward, it is supported by the fact that this
@@
 The Matrix Method assumes a linear relationship between these selections
-and those of real ($r$) and fake/non-prompt ($f$) leptons:
+and those of real ($r$) and FNP ($f$) leptons:
 \begin{equation}
@@
 \end{align}
-To make fake/non-prompt estimates in our regions, the Matrix Method is
+To make FNP estimates in our regions, the Matrix Method is
 implemented by assigning weights to observed data events that are then
@@
 have an average weight of $+0.238$ (one can check these recover the central
-fake/non-prompt background of $2.2$).
+FNP background of $2.2$).
 Such negative signal weights suggest that a local \emph{excess} of data
@@
 at rates similar to (or larger than) the standard model backgrounds.
-Loose data do have positive weights and they would drive up the fake/non-prompt
+Loose data do have positive weights and they would drive up the FNP
 background.
@@
 \subsubsection{Fake/non-prompt estimates in $\twoljets$-electroweak regions}
-Many histograms were inspected to validate the Matrix Method estimates in
-$\twoljets$-electroweak regions.
-To give a broad summary across all of these regions, we define an additional
-validation region, named VR-SS, to inspect similar data with same-charged
-(same sign, SS) leptons.
-This selection enhances fake/non-prompt contributions relative backgrounds
-since it suppresses contributions $Z\rightarrow \ell^\pm\ell^\mp$ processes,
-whose leptons should be oppositely charged if accurately reconstructed,
-but does not suppress fake/non-prompt backgrounds because their lepton charges
-are expected to be less dependent:
-fake/non-prompt backgrounds produce $\ell^\mp\ell^\mp$ and $\ell^\pm\ell^\pm$
-vents at more similar rates.
-
-For these reasons, VR-SS is defined as a modification to the pre-selection
-requirements of Table~\ref{tab:2ljets_presel}, in which the requirement
-of opposite charge leptons is changed to a requirement of same charge leptons
-(same sign, SS). Additionally, VR-SS requires $\mll \in (12, 111)~\eV[G]$,
+To validate the Matrix Method estimates in $\twoljets$-electroweak regions
+when preparing the search, we plotted and inspected many histograms in
+selections designed to enhance the FNP backgrounds.
+All of these selections leverage the expected independence of the two leptons
+in FNP events --- important backgrounds producing leptons through
+$Z\rightarrow \ell^\pm\ell^\mp$ must have leptons of the same flavour and
+opposite charge, and those are quite accurately reconstructed.
+If one lepton is FNP, however, its charge and flavour should be
+independent of the other lepton.
+Selections requiring different-flavour or same-charge leptons,
+in otherwise $Z$-like scenarios, should therefore be
+FNP-enhanced.
+
+Alternative versions of all core control, validation regions were inspected
+with their lepton requirements change to require either different-flavour
+or same-charge leptons.
+The different-flavour versions are contain increased $\ttbar$ backgrounds,
+due to their flavour symmetry,
+and both different-flavour and same-sign versions have drastically reduced
+expected backgrounds.
+None of these flavour- or charge- flipped regions showed any significant
+deviations from the simulated backgrounds, but most also contained very small
+predicted FNP yields.
+
+To validate FNP yields in a region with sufficiently many data for precision,
+a additional validations were also inspected with different-flavour or
+same-charge requirements, and all showed reasonably agreement with data
+within uncertainties.
+In particular, a region named VR-SS is defined to inspect data with
+same-charged (same sign, SS) leptons in a loose preselection that covers all
+of our main regions;%
+\footnote{%
+Like many of the regions plotted in designing this search, VR-SS is not a
+`core' validation region in the sense that it is not reported in the paper.
+Perhaps a fourth categorization would be useful for such regions to supplement
+CR, VR, and SR, but it should be clear that any VR not included in the main
+region definitions is not core and not reported in the paper.
+}
+VR-SS requires the pre-selections of Table~\ref{tab:2ljets_presel},
+but replaces its requirement of opposite-charge leptons with a requirement of
+same-charge leptons, and adds a selections of $\mll \in (12, 111)~\eV[G]$
 to cover both on- and off-shell selections.
+
 Histogrammed $\metsig$ distributions in VR-SS are displayed in
 Figure~\ref{fig:2ljets_vrss},
-with fake/non-prompt estimates taken from the Matrix Method in
+with FNP estimates taken from the Matrix Method in
 Figure~\ref{fig:2ljets_vrss_mm},
@@
 When using the Matrix Method, we use parton-level `truth' information to
-identify and remove remove fake/non-prompt contributions from simulations.
+identify and remove remove FNP contributions from simulations.
 This is why the simulated backgrounds (not `FNP leptons') increase by around
 $30\%$ when switching off the Matrix Method; the difference between the
-two orange histograms is the simulated fake/non-prompt estimate.
+two orange histograms is the simulated FNP estimate.
 
@@
 
-Drafts of the $\twoljets$ paper included
-Figure~\ref{fig:2ljets_vrss_mm} and a similar plot for the
-$\twoljets$-strong search, and repeated this point that the deviation was not
+Early drafts of the $\twoljets$ paper included
+Figure~\ref{fig:2ljets_vrss_mm}, and a similar plot for the
+$\twoljets$-strong search, and stated this point that the deviation was not
 very substantial.
 These plots were removed at the request of internal reviewers, who argued that
-they did not need to be shown since the fake/non-prompt background is not
+they did not need to be shown since the FNP background is not
 very important.
-To demonstrate that it was not important, we were first instructed to repeat
-the exclusion fits of the $\twoljets$-electroweak analysis with an additional
-uncertainty, to further cover the deviation observed in this figure, which
-scaled the Matrix Method yields by a $\pm 50\%$ scale factor.
-The utility of this is somewhat questionable since that variation is coupled
-between all bins, so it could not vary to explain an excess in one region
+To demonstrate that it was not important, we first repeated the exclusion fits
+with an additional uncertainty, to further cover the deviation observed in this
+figure, which scaled the Matrix Method yields by a $\pm 50\%$ scale factor.
+The utility of this variation is somewhat questionable since it is coupled
+between all bins, so it cannot vary to fit an excess in one region
 without also increasing the fitted Matrix Method backgrounds in all other
-regions, and it would struggle to make any significant difference.
+regions.
+That is, it is a global change, and if constrained by data then the data
+constraint would dominate, leaving little difference.
 Indeed, the result was an insignificant change to all exclusion contours;
@@
 For comparison, this table also includes the simulated background without
-fake/non-prompt leptons under the heading `MC Prompt' and the, and only
-the simulated fake/non-prompt contributions under the heading `MC FNP'.
+FNP leptons under the heading `MC Prompt' and the, and only
+the simulated FNP contributions under the heading `MC FNP'.
 Negative or void estimates are an easy consequence of the use of signed
@@
 alternative estimates that assign central and high variations to the
-fake/non-prompt background estimate, and consider a maximum across them for
+FNP background estimate, and consider a maximum across them for
 each region.
@@
 
-An awkward fact in the history of this fake/non-prompt estimate is that it
+An awkward fact in the history of this combined FNP estimate is that it
 was reworked after unblinding to include the Mean estimate and more regions;
 with a reprocessing of our background simulations, changes in the Matrix Method
-estimates left more regions with non-positive estimate than previously.
+estimates left more of our regions with non-positive estimate than prior to
+unblinding the data.
 To assign them without being biased by the unblinded signal region data,
-known to contain no significant excesses, therefore took some barefaced
-honesty.
+which were known to contain no significant excesses, therefore took some
+barefaced honesty.
 Thankfully, we do not need to consider what might have happened had an excess
@@
 }
+Despite any awkwardness, the its does appear that FNP backgrounds are indeed
+small and that they do not greatly impact the result, especially when compared
+to other systematic variations.
 
@@
 It was removed after validation studies showed that adding an additional
-$50\%$ uncertainty to fake/non-prompt yields, coupled between all bins, did not
+$50\%$ uncertainty to FNP yields, coupled between all bins, did not
 significantly change our exclusion contours;
@@
 \caption[%
-Background yields from prompt and fake/non-prompt sources in bins with
+Background yields from prompt and FNP sources in bins with
 non-positive Matrix Method estimates
 ]{%
-Background yields from prompt (two real, prompt leptons) and fake/non-prompt
+Background yields from prompt (two real, prompt leptons) and FNP
 sources in bins with non-positive Matrix Method estimates.
@@
 
+\subsection{Jet mass}
+\label{sec:2ljets_jet_mass}
+An extraordinary feature of this analysis is the use of $\mjetone$ for
+SR-High-1J and VR-High-1J which, as described in Section~\ref{sec:2ljets_high},
+require a boosted, small-radius jet with a large mass
+($\mjetone \in 60\textrm{--}110$).
+This signal region is important for extending our reach to the most massive
+models with large mass splittings, but probes an extreme case whose modelling
+in simulations is not well validated.
+
+A maximally crude display of the jet mass variable is given in
+Figure~\ref{fig:2ljets_jetm_c1n2_600_0} for a benchmark C1N2 model in
+the fast simulation, showing that most jets have small masses but that
+a substantial fraction appear in a bump around $m(W)$.
+To check that a similar bump exists in data and background simulations,
+I conducted a study into single-lepton events to search for $\ttbar$ events
+in which one top decays leptonically, and the other to a boosted hadronic $W$
+boson that is reconstructed into a single small-radius jet.
+Since our main samples require two leptons, this study required independent
+inputs derived from different sources, so to reduce the required computing time
+I opted to include only $\ttbar$ and single-top backgrounds.
+To verify the origins of high-mass jets, partonic truth information is used
+to label those events containing a $W$ boson decaying to two quarks that are
+within $\Delta R < 0.4$ of the hardest reconstructed jet.
+The resulting bump is visible in Figure~\ref{fig:2ljets_jetm_ttbar_mw} in a
+selection described in its figure caption, and indeed contains the targetted
+boosted $W$ boson. This positive result was an early motivation for our
+continued pursuit of the 1J regions.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/2ljets_jetm_c1n2_600_0_tbrowser.png}
+\caption[
+Jet masses in raw samples of a benchmark C1N2 signal model
+]{%
+Jet masses in raw samples of a benchmark C1N2 signal model with
+$m(\chargino_1,\neutralino_2) = 600\,\eV[G]$ and
+$m(\neutralino_1) = 0$.
+All jets in all events in this file are histogrammed without weights,
+and the signal model is simulated with fast simulation.
+}
+\label{fig:2ljets_jetm_c1n2_600_0}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/2Ljets_hist1d_jetM0_ttbar_merged_tweak.pdf}
+\caption[
+Jet mass distribution in a single-lepton selection looking for
+$t\bar t \rightarrow \ell^\pm\nu W(J) \bar bb$
+]{%
+Jet mass distribution in a single-lepton selection looking for
+$t\bar t \rightarrow \ell^\pm\nu W bb$, where the $W$ boson decays hadronically
+into a single small-radius jet.
+A bump is observed around $m(W)$, showing that these jet algorithms do
+reconstruct the hadronic $W$ boson decays bosons in their full ATLAS
+simulation.
+This selection requires one signal lepton (electron or muon) with
+$\pt > 25\,\eV[G]$, $\njet = 3$, $\nbtag = 2$, $\ptjone > 250\,\eV[G]$,
+and that the hardest jet is not $b$-tagged.
+The `$\ttbar$ Merged' sample selects events in which both quarks from the
+$W\rightarrow q\bar q$ decay are within $\Delta R < 0.4$ of the hardest
+reconstructed jet.
+The discrepancy in low masses is attributable to missing $W\mathrm{+jets}$
+processes that were not processed for this study.
+Backgrounds are normalized to data in the high-mass region near to $m(W)$.
+}
+\label{fig:2ljets_jetm_ttbar_mw}
+\end{figure}
+
+% problem of jet mass for small radius
+Our inclusion of SR-High-1J and VR-High-1J regions induces two related problems.
+First is their unusual use of the jet mass variable $\mjetone$ with
+small-radius ($R=0.4$), topologically clustered jets~\cite{atlas_jet_topo_PERF_2014_07}.
+Usually, hadronic decays of massive objects
+(such as $W$, $Z$, or $t$ resonances) are searched for with large-radius jets
+with $R=0.8$, since these require less extreme boosts to capture their targets.
+The focus from ATLAS' performance operations is therefore reasonably to
+understand these common use-cases first.
+Second is the use of fast simulation~\cite{SOFT-2010-01} for our signal models,
+which uses approximate descriptions of the detector response and does not
+guarantee to model the substructure that determines a jet's mass.
+
+Official jet mass scale uncertainties for small-radius jets in the full ATLAS
+simulation were released late into the development of this search
+(25 April 2021), and these uncertainties explicitly do not support the
+fast simulation~\cite{atlas_twiki_smallrjms}.
+Once released, these uncertainties were evaluated for the diboson and $\ttbar$
+contributions, which are the primary backgrounds in both 1J regions.
+
+% TODO
+% alternative for fast simulation
+
+% also apply this method to minor backgrounds
+
+% AFII vs FullSim comparison, 15% variations
+
+Jet mass resolution uncertainties were evaluated by smearing simulations
+to increase their resolution error (decrease their precision) by $20\%$
+and comparing the smeared results to standard results according to a standard
+prescription~\cite{atlas_twiki_jmr}.
+
 \subsection{Sample comparisons}

### 2022-07-21 main/chapter_2ljets.tex
correct and clarift some jetm modelling

@@
 A maximally crude display of the jet mass variable is given in
-Figure~\ref{fig:2ljets_jetm_c1n2_600_0} for a benchmark C1N2 model in
-the fast simulation, showing that most jets have small masses but that
-a substantial fraction appear in a bump around $m(W)$.
-To check that a similar bump exists in data and background simulations,
-I conducted a study into single-lepton events to search for $\ttbar$ events
-in which one top decays leptonically, and the other to a boosted hadronic $W$
-boson that is reconstructed into a single small-radius jet.
+Figure~\ref{fig:2ljets_jetm_c1n2_600_0} for a benchmark C1N2 model, showing
+that most  jets have small masses but that a substantial fraction appear in a
+bump around $m(W)$.
+However, all of our signal models are simulated with the fast ATLAS simulation
+ATLFAST-II~\cite{SOFT-2010-01} that uses approximate descriptions of the
+detector response and does not guarantee to accurately model the substructures
+that determines the jet's reconstructed mass.
+Jet mass modelling should be well described since masses are part of the
+reconstructed energy, and so also for standard mass variables such as $\mjj$.
+However, momenta are typically more important for these variables, so
+errors could be hidden.
+Inclusion of these 1J regions in the $\twoljets$-electroweak search therefore
+required validation of these ATLFAST-II mass distributions.
+
+To check that a similar bump exists in data and fully simulated (FullSim)
+background simulations, I conducted a study into a standard model process
+production boosted hadronic $W$ bosons in $\ttbar$ events.
+Although non-$\ttbar$ $WW$ processes do occur, the semileptonic $\ttbar$
+process additionally produces two $b$ quarks in its dominant decay, and their
+$b$-tagged jets that can be selected to reduce backgrounds.
+
 Since our main samples require two leptons, this study required independent
-inputs derived from different sources, so to reduce the required computing time
-I opted to include only $\ttbar$ and single-top backgrounds.
+single-lepton inputs derived from different sources, so to reduce the required
+computing time I included only $\ttbar$ and single-top backgrounds generated
+with the full \atlas simulation.
 To verify the origins of high-mass jets, partonic truth information is used
-to label those events containing a $W$ boson decaying to two quarks that are
-within $\Delta R < 0.4$ of the hardest reconstructed jet.
-The resulting bump is visible in Figure~\ref{fig:2ljets_jetm_ttbar_mw} in a
-selection described in its figure caption, and indeed contains the targetted
-boosted $W$ boson. This positive result was an early motivation for our
-continued pursuit of the 1J regions.
+to label those events containing a $W$ boson decaying to two quarks within
+$\Delta R < 0.4$ of the hardest reconstructed jet.
+The resulting bump does exist, as shown in
+Figure~\ref{fig:2ljets_jetm_ttbar_mw}.
+This positive result was an early motivation for our continued pursuit of the
+1J regions.
 
@@
 % problem of jet mass for small radius
-Our inclusion of SR-High-1J and VR-High-1J regions induces two related problems.
-First is their unusual use of the jet mass variable $\mjetone$ with
-small-radius ($R=0.4$), topologically clustered jets~\cite{atlas_jet_topo_PERF_2014_07}.
-Usually, hadronic decays of massive objects
-(such as $W$, $Z$, or $t$ resonances) are searched for with large-radius jets
-with $R=0.8$, since these require less extreme boosts to capture their targets.
-The focus from ATLAS' performance operations is therefore reasonably to
+Not only does fast simulation pose a problem to the 1J regions, but our use
+of $\mjetone$ in small-radius ($R=0.4$), with the older
+`topologically clustered' jet technology~\cite{atlas_jet_topo_PERF_2014_07}
+is also unusual.
+% TODO citations for large-R jet searches
+Hadronic decays of massive objects
+(such as $W$, $Z$, or $t$ resonances) in \atlas are usually searched for with
+large-radius jets with $R=0.8$, since these require less extreme boosts to
+capture their targets.
+The focus from ATLAS' performance operations is therefore reasonable to
 understand these common use-cases first.
-Second is the use of fast simulation~\cite{SOFT-2010-01} for our signal models,
-which uses approximate descriptions of the detector response and does not
-guarantee to model the substructure that determines a jet's mass.
 

### 2022-07-21 main/chapter_2ljets.tex
add afii fullsim signal comparisons to jetm modelling; remove atlas labels; some other tweaks

@@
 tasks which were common between the searches.}
-\item Integration of the electroweak analysis with ATLAS combinations and
+\item Integration of the electroweak analysis with \atlas\ combinations and
 pMSSM scan efforts, by performing their validation studies and serializing the
@@
 the $\twoljets$ search effort:
-Knut Oddvar Vadla (electroweak),
-Sarah Williams (electroweak),
-Jason Lea Oliver (RJR),
-Abhishek Sharma (RJR),
-Jonathan Long (strong),
-Arka Santra (strong),
-Matt Zhang (strong),
-Yumeng Cao (various),
-Eirik Gramstad (fake/non-prompt leptons),
-Benjamin Henry Hooberman (leadership).
+Knut~Oddvar Vadla (electroweak, core),
+Sarah~Williams (electroweak),
+Jason~Oliver (RJR),
+Abhishek~Sharma (RJR),
+Jonathan~Long (strong),
+Arka~Santra (strong, RECAST),
+Matt~Zhang (strong),
+Yumeng~Cao (electroweak, strong),
+Eirik~Gramstad (fake/non-prompt leptons),
+Benjamin~Hooberman (leadership).
 Cheers.
@@
 From these Run~1 searches, one can cherry-pick excesses
-of ``$3.0$ standard deviations'' in \atlas~\cite{atlas_susy_strong_2l_run1}
+of ``$3.0$ standard deviations'' in \atlas\ \cite{atlas_susy_strong_2l_run1}
 and ``$2.6$ standard deviations'' in \cms~\cite{cms_susy_2016_strong_2l_run1}.
@@
 bump around $m(W)$.
-However, all of our signal models are simulated with the fast ATLAS simulation
-ATLFAST-II~\cite{SOFT-2010-01} that uses approximate descriptions of the
+However, all of our signal models are simulated with the fast \atlas\ simulation
+\afii\ \cite{SOFT-2010-01} that uses approximate descriptions of the
 detector response and does not guarantee to accurately model the substructures
@@
 Inclusion of these 1J regions in the $\twoljets$-electroweak search therefore
-required validation of these ATLFAST-II mass distributions.
+required validation of these \afii\ mass distributions.
 
@@
 computing time I included only $\ttbar$ and single-top backgrounds generated
-with the full \atlas simulation.
+with the full \atlas\ simulation.
 To verify the origins of high-mass jets, partonic truth information is used
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/2Ljets_hist1d_jetM0_ttbar_merged_tweak.pdf}
+\includegraphics[width=0.9\textwidth]{figures/2Ljets_hist1d_jetM0_ttbar_merged_tweak.png}
 \caption[
@@
 A bump is observed around $m(W)$, showing that these jet algorithms do
-reconstruct the hadronic $W$ boson decays bosons in their full ATLAS
+reconstruct the hadronic $W$ boson decays bosons in their full \atlas\
 simulation.
+\\
 This selection requires one signal lepton (electron or muon) with
@@
 is also unusual.
-% TODO citations for large-R jet searches
 Hadronic decays of massive objects
-(such as $W$, $Z$, or $t$ resonances) in \atlas are usually searched for with
+(such as $W$, $Z$, or $t$ resonances) in \atlas\ are usually searched for with
 large-radius jets with $R=0.8$, since these require less extreme boosts to
 capture their targets.
-The focus from ATLAS' performance operations is therefore reasonable to
+The focus from \atlas' performance operations is therefore reasonable to
 understand these common use-cases first.
 
-Official jet mass scale uncertainties for small-radius jets in the full ATLAS
-simulation were released late into the development of this search
+Official jet mass scale (JMS) uncertainties for small-radius jets in the full
+\atlas\ simulation were released late into the development of this search
 (25 April 2021), and these uncertainties explicitly do not support the
 fast simulation~\cite{atlas_twiki_smallrjms}.
-Once released, these uncertainties were evaluated for the diboson and $\ttbar$
-contributions, which are the primary backgrounds in both 1J regions.
-
-% TODO
+These variations comprise six different parameters representing a reported
+difference between data and simulations, differences between simulations from
+different generators, uncertainties in tracking, and statistical uncertainties.
+Once released, JMS variations were evaluated and used for the diboson and
+$\ttbar$ contributions, which are the primary backgrounds in both 1J regions.
+
 % alternative for fast simulation
-
-% also apply this method to minor backgrounds
+To reproduce all signal models with full simulation would have cost substantial
+computing and human resources, so a non-standard alternative assignment was
+required to model JMS variations in \afii\ samples.
+For the kinematic regions to which our regions are sensitive, specifically
+$\ptjone > 300\,\eV[G]$ and $\mjetone/\ptjone < 0.3$, the official
+JMS relative uncertainties are always less than $10\%$.
+As a conservative approximation, I therefore assigned alternative JMS
+variations by scaling the observed $\mjetone$ values up and down by $10\%$
+and recording the resulting histogram yields as variations.
+
+In both official and alternative assignments, the JMS variations result in
+large relative yield variations ranging from around $20\%$ to $40\%$ in samples
+with good statistical precision, and both methods assign similar variations
+when applied to the FullSim samples.
+To avoid the expense of reprocessing minor backgrounds with the official JMS
+variations, this alternative prescription is used also for minor backgrounds.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_C1N2_WZ_600_0_2L2J_presel_jetM[0]_26.png}
+\caption{C1N2 $(600, 0)$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_C1N2_WZ_500p0_200p0_2L2J_presel_jetM[0]_38.png}
+\caption{C1N2 $(500, 200)$}
+\end{subfigure}
+\\
+\centering
+\begin{subfigure}{0.48\textwidth}
+\includegraphics[width=\textwidth]{figures/2ljets_C1N2_WZ_200p0_100p0_2L2J_presel_jetM[0]_50.png}
+\caption{C1N2 $(200, 100)$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_C1N2N1_GGMHinoZh50_600_BF50_presel_jetM[0]_62.png}
+\caption{GMSB $(500, 50\%)$}
+\end{subfigure}
+\caption[
+Signal samples in \afii\ and FullSim simulations in a loose preselection
+]{%
+Signal samples in \afii\ and FullSim simulations in a loose preselection
+requiring the pre-selections of Table~\ref{tab:2ljets_presel} and
+$\mll \in 71\textrm{--}111\,\eV[G]$.
+No $m(W)$ bump for C1N2 $(200, 100)$ is a positive sign that the fast \afii\
+simulations do not anomalously produce massive jets.
+Other samples appear to show a slight bias of \afii\ (AF2) towards lower jet
+masses, indicated by the FullSim excesses to the right of each peak.
+This could be explained by a reduced level of background energy deposits within
+jet clusters, but would also be covered in a large part by a little smearing of
+the \afii\ result to describe other noise.
+}
+\label{fig:2ljets_high_mll_b}
+\end{figure}
 
 % AFII vs FullSim comparison, 15% variations
+To further check the correspondence between \afii\ and FullSim signal modelling,
+\emph{Arka~Santra prepared for us FullSim versions of four representative
+signal models}.
+These alternative detector simulations are compared in
+Figure~\ref{fig:2ljets_jetm_afii_fullsim} in a loose pre-selection,
+showing quite good modelling of the $\mjetone$ distributions.
+The ratio subplots do indicate local differences, particularly of \afii\ yields
+below FullSim yields on the right-hand slopes of peaks, but average out
+across the $\mjetone \in 60\textrm{--}110\,\eV[G]$ window.
+On consideration of these results and similar plots in tighter selections ,
+we assign a $15\%$ relative variation to all \afii\ samples in 1J regions
+to cover the divergence of \afii\ from FullSim in $\mjetone$.
 
@@
 prescription~\cite{atlas_twiki_jmr}.
+These apply equally to \afii\ and FullSim samples.
+
+\emph{This jet mass modelling was discussed with, and approved by, the \atlas\
+Jet/Etmiss conveners. I owe special thanks to Chris~Delitzsch for assistance
+with the official jet mass scale variations.}
 

### 2022-07-22 main/chapter_2ljets.tex
language check

@@
 \includegraphics[width=\textwidth]{figures/2ljets_low_absdPhiPllMet_SRLow2_noDphi_mg5.pdf}
-\caption{SR-Low-2$/\{\dphillmet\}$, alternative}
+\caption{SR-Low-2$/\{\dphillmet\}$ (alt)}
 \label{fig:2ljets_low_minus_dphi_alt}
@@
 \label{sec:2ljets_mm_fakes}
-
 Given our requirements on lepton quality, isolation, and hardness, it is
@@
 The tight-tight events have an average weight of $-0.0227$, and the others
-have an average weight of $+0.238$ (one can check these recover the central
-FNP background of $2.2$).
+have an average weight of $+0.238$ (one can check that these recover the
+central FNP background of $2.2$).
 Such negative signal weights suggest that a local \emph{excess} of data
 would act to \emph{decrease} the Matrix Method estimate in the region.
-Yes this backwards dependency is backward, but at least it gives some relief
+Yes, this backwards dependency is backward, but at least it gives some relief
 that real local excesses would not be hidden by increased fake estimates.
@@
 at rates similar to (or larger than) the standard model backgrounds.
-Loose data do have positive weights and they would drive up the FNP
+Loose data do have positive weights, and they would drive up the FNP
 background.
@@
 or same-charge leptons.
-The different-flavour versions are contain increased $\ttbar$ backgrounds,
+The different-flavour versions contain enhanced $\ttbar$ fractions,
 due to their flavour symmetry,
@@
 To validate FNP yields in a region with sufficiently many data for precision,
-a additional validations were also inspected with different-flavour or
+additional validations were also inspected with different-flavour or
 same-charge requirements, and all showed reasonably agreement with data
@@
 but replaces its requirement of opposite-charge leptons with a requirement of
-same-charge leptons, and adds a selections of $\mll \in (12, 111)~\eV[G]$
+same-charge leptons, and adds a requirement of $\mll \in (12, 111)~\eV[G]$
 to cover both on- and off-shell selections.
@@
 When using the Matrix Method, we use parton-level `truth' information to
-identify and remove remove FNP contributions from simulations.
+identify and remove FNP contributions from simulations.
 This is why the simulated backgrounds (not `FNP leptons') increase by around
@@
 Overlaying these modified contours highlighted some minor differences, the
-largest being $2\%$ reduction in the limit on
+largest being a $2\%$ reduction in the limit on
 $B(\neutralino_1 \rightarrow h \gravitino)$ at
@@
 For comparison, this table also includes the simulated background without
-FNP leptons under the heading `MC Prompt' and the, and only
+FNP leptons under the heading `MC Prompt' and the and only
 the simulated FNP contributions under the heading `MC FNP'.
@@
 }
-Despite any awkwardness, the its does appear that FNP backgrounds are indeed
+Despite any awkwardness, it does appear that FNP backgrounds are indeed
 small and that they do not greatly impact the result, especially when compared
@@
 To avoid the expense of reprocessing minor backgrounds with the official JMS
-variations, this alternative prescription is used also for minor backgrounds.
+variations, this alternative prescription is also used for minor backgrounds.
 
@@
 }
-\label{fig:2ljets_high_mll_b}
+\label{fig:2ljets_jetm_afii_fullsim}
 \end{figure}
@@
 % extrapolation
-The total theory variation is a combination of seven components representing
+The total theory variation is a combination of seven components, representing
 \TODO{what are they?}.
@@
 Figure~\ref{fig:2ljets_summary}.
-Fitting to signal regions is policy of the \atlas\ SUSY group, but deviates
+Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
 from the intentions of the significance measure.

### 2022-07-22 main/chapter_2ljets.tex
add hepdata paragraph

@@
 \section{Modelling}
-
 So far, we have seen many histograms of background samples overlaid with
@@
 and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.
+
+% Thanks to Jon Long, Alessia Murrone
+This section only covers some of the the most peculiar or important aspects
+of our signal and background modelling, and does not specify most exact
+details.
+Full details are publically available through the $\twoljets$
+HEPData~\cite{hepdata.116034, Maguire_2017} archive our json-serialized pyhf
+workspaces~\cite{cranmer2012histfactory, heinrich2021pyhf}.
+\emph{Thanks to Jon~Long for collabouration in preparing this HEPData archive,
+and to Alessia~Murrone for invaluable feedback on, and validation of, its
+drafts.}
 

### 2022-07-22 main/chapter_2ljets.tex
begin sample comparisons

@@
 
-% Thanks to Jon Long, Alessia Murrone
 This section only covers some of the the most peculiar or important aspects
@@
 \label{sec:2ljets_sample_comparisons}
+The Standard Model is a distillation of past data, and all of our simulations
+of it have been tuned to those data, in both their treatment of fundamental
+physics and of experimental hardware.%
+\footnote{%
+Amusingly, tuning to data specific to each experiment means that each
+experiment can simulate different fundamental physics~\cite{
+herwig2000parameter,
+Altarelli:473529
+}, not only different detector responses.
+All Standard Models are not the same.
+}
+Each tune yields a point estimate that does not incorporate remaining
+uncertainties, so those uncertainties must be quantified by other means.
+Some uncertainties are included as weights, particularly those for
+uncertainties in the
+PDFs (parton distribution functions) and phyiscal scales
+(renormalization and factorization scales, $\mu_R$ and $\mu_F$).
+Our inclusion of these weighty variations through standard prescriptions is
+described in Section~\ref{sec:2ljets_theory_variations}.
+
+To include other uncertainties that are not described by weights,
+the standard approach is to compare the results of different simulation
+algorithms, and we implement this for our major backgrounds $\zjets$, $\ttbar$,
+and, to some extent diboson.
+Each generator (simulation algorithm), or component thereof, uses different
+approximate methods in its simulations, so the variations between simulations
+indicate something about a reasonable uncertainty scale.
+
+Clearly, it could happen that two algorithms agree in their tune despite
+neither being confident --- as an illustration, imagine a wide error bar and
+two post-fit results right in its centre.
+This effect would yield an underestimated uncertainty, so does not invalidate
+the comparison approach.
+On the contrary, it highlights that it is a bare minimum and that larger
+uncertainties may well be reasonable given our imprecise constraints on the
+Standard Model.
+
 Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
@@
 
-Table~\ref{tab:2ljets_zjets_alt}
+For implementations in the HistFactory model, these variations reduce to
+the relative scale factors presented in Table~\ref{tab:2ljets_zjets_alt}.
+Note that up and down variations of the sample comparison parameter do not
+mean up and down variations of bin yields, but to the coupled variations
+in directions given by the signs of the relative variations given in the
+various regions.
 
@@
 ]{%
-Alternative $\zjets$ variations in selected regions.
+Alternative $\zjets$ variations in selected regions that have significant
+$\zjets$ contributions.
+In other regions the $\zjets$ sample and its variations are negligible.
 The `Central' column shows the central $\zjets$ yield from the main sample.
@@
 the text.
-Note how up and down variations can move in different directions.
+Note how up and down variations can move in different directions, and these
+different movements are coupled in the background model.
 }

### 2022-07-22 main/chapter_2ljets.tex
add ROOT TBrowser note

@@
 Finally, the `Mean' high variation is their linear sum;
-$\mu^\textrm{M} + \sigma^\textrm{M}_\textrm{stat}
+$\mu^\textrm{M} + \sigma^\textrm{M}_\textrm{stat} +
 \sigma^\textrm{M}_\textrm{sys} = 0.1560$.
@@
 $m(\neutralino_1) = 0$.
-All jets in all events in this file are histogrammed without weights,
+All jets in all events in this file are histogrammed,
 and the signal model is simulated with fast simulation.
+\\
+Yes, this really is a screenshot of a ROOT TBrowser~\cite{ROOT}, with no
+weights on its samples, nor plotting frills such as error bars or labels.
+A similar view was our first, and surprising, indication that mono-jet masses
+might be useful, and although this chronology is not at all necessary to this
+thesis, I find it an amusing reminder of the power of quick-and-dirty checks.
 }

### 2022-07-22 main/chapter_2ljets.tex
expand zjets alt section

@@
 uncertainties in the
-PDFs (parton distribution functions) and phyiscal scales
+PDFs (parton distribution functions) and physical scales
 (renormalization and factorization scales, $\mu_R$ and $\mu_F$).
@@
 Clearly, it could happen that two algorithms agree in their tune despite
-neither being confident --- as an illustration, imagine a wide error bar and
-two post-fit results right in its centre.
+neither being confident.
+To illustrate this point, imagine a wide error bar and
+two post-fit results right in its centre;
+both have achieved the best fit, but both also could allow large variations.
 This effect would yield an underestimated uncertainty, so does not invalidate
-the comparison approach.
-On the contrary, it highlights that it is a bare minimum and that larger
-uncertainties may well be reasonable given our imprecise constraints on the
-Standard Model.
-
-Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
-Figure~\ref{fig:2ljets_low_crz_pre_post}
-Figure~\ref{fig:2ljets_vrlow_alt}
-Figure~\ref{fig:2ljets_low_vrlow_pre_post}
-
-For implementations in the HistFactory model, these variations reduce to
-the relative scale factors presented in Table~\ref{tab:2ljets_zjets_alt}.
+the comparison approach but highlights that it is a bare minimum and that
+larger uncertainties may well be reasonable given our imprecise constraints on
+the Standard Model.
+
+As mentioned in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison
+is hugely important to our results in SR-Low.
+Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
+in the union of CR-Z and SR-Low, which form a continuous region in $\mjj$.
+Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
+Bothmann:2019yzt,
+ATL-PHYS-PUB-2016-003
+}
+at next-to-leading order for up to $4$ partonic jets, and leading order for
+$5$ partonic jets.
+In Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} we compare this to our
+alternative sample that simulates its hard scatter
+with \madgraph\ (MG5\_aMC@NLO~2.2.3~\cite{Alwall:2014hca}) at leading order with
+up to four partonic jets, plugged into \pythia~8~\cite{Sjostrand:2014zea} for
+showering and hadronization.
+These plots show statistical error bars only, and the larger statistical error
+bars on the alternative sample are due to it comprising around one sixth as
+many simulated events as the central sample.
+
+For their implementation in the background model in HistFactory, these
+variations reduce to the relative scale factors that are presented in
+Table~\ref{tab:2ljets_zjets_alt}.
 Note that up and down variations of the sample comparison parameter do not
@@
 various regions.
-
+These variations are evaluated according to a standard recipe prescribed by
+the ATLAS SUSY group~\cite{atlas_twiki_susy_theory_comparison} that says to
+\begin{itemize}
+\item evaluate the difference between the two samples $\delta_{12}$.
+\item evaluate the statistical uncertainty on the alternative sample $\sigma_2$.
+\item combine these components in quadrature, so
+$\sigma_{12} = \sqrt{\delta_{12}^2 + \sigma_2^2}$, and
+\item if the deviation is statistically significant, then preserve the direction
+of the uncertainty, so $\sigma = \mathrm{sgn}\sigma_{12}$.
+\end{itemize}
+An `if A then B' statement does not define what to do if A is false, so
+this recipe does not define what sign to give $\sigma$ if the deviation is not
+statistically significant.
+Since all variations have direction
+--- they are either positive or negative or zero ---
+we are unable to give no direction to any deviation, and we choose to
+we use the same, signed, definition of $\sigma$ for all regions.
+
+As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
+$\sigma_2$ is a substantial contribution to these variations;
+the statistical uncertainty in the alternative sample does significantly
+expand this variation.
+Indeed, it is the larger contribution in SR-Low-a and SR-Low-b, which are the
+regions most sensitive to $\zjets$ variations.
+Our unprescribed choice to match their directions therefore matters.
+It may have been reasonable to treat them differently, perhaps by adding
+independent parameters in the separate regions.
+I believe such a change would have improved the plausibility of our background
+modelling, and I regret not finding the capacity to push it at the time.
+
+This SUSY recipe contradicts the advice of the ATLAS Physics Modelling Group
+(PMG), who say to quote only the absolute value of the two
+estimates~\cite{atlas_twiki_pmg_theory_comparison} in for two-sample
+comparisons.
+Since our search is in the SUSY group we follow their instructions.
+
+Figure~\ref{fig:2ljets_low_crz_pre_post}
+Figure~\ref{fig:2ljets_vrlow_alt}
+Figure~\ref{fig:2ljets_low_vrlow_pre_post}
+
+% in branch hackmg5
+% python get_ewk_Zjets_alt.py
+%                 CRDY,     2.30,     5.97
+%                  CRZ,    16.75,     6.68
+%                 CRZZ,     6.35,     1.61
+%                 CRtt,    18.06,     4.29
+%           SRHigh16_1,     0.02,     0.00
+%           SRHigh16_2,     0.01,     0.00
+%              SRHigh4,     0.02,     0.00
+%            SRHigh8_1,     0.02,     0.00
+%            SRHigh8_2,     0.02,     0.00
+%              SRInt_1,     0.46,     0.76
+%              SRInt_2,     0.09,     0.20
+%               SRLow2,     1.63,     1.69
+%              SRLow_1,     2.20,     2.36
+%              SRLow_2,     0.45,     1.02
+%         SROffShell_1,     0.12,     0.30
+%         SROffShell_2,     0.95,     0.95
+%               SRllbb,     0.03,     0.00
+%               VRHigh,     0.13,     0.14
+%              VRHigh4,     0.49,     0.39
+%              VRHighR,     0.26,     0.34
+%                VRInt,     0.32,     0.23
+%                VRLow,     7.34,     3.24
+%               VRLow2,     1.71,     3.21
+%           VROffShell,     0.09,     1.06
+%               VRllbb,     0.31,     0.36
 \begin{table}[tp]
 \centering
-\begin{tabular}{lcc}
+\begin{tabular}{lcccc}
 & Central
 & Relative variations
+& $|\delta_{12}|$
+& $\sigma_2$
 \\[0.5em]
@@
 & $1-0.1+0.1$
+& $2.3$
+& $6.0$
 \\[0.2em]
@@
 & $1+0.2-0.2$
+& $16.8$
+& $6.7$
 \\[0.2em]
@@
 & $1+0.5-0.5$
+& $18.1$
+& $4.29$
 \\[0.2em]
@@
 & $1+1.3-1.0$
+& $6.4$
+& $1.6$
 \\[0.5em]
@@
 & $1-0.3+0.3$
+& $7.3$
+& $3.2$
 \\[0.2em]
@@
 & $1+0.2-0.2$
+& $1.7$
+7 $3.2$
 \\[0.5em]
@@
 & $1-0.2+0.2$
+& $2.2$
+& $2.4$
 \\[0.2em]
@@
 & $1-0.3+0.3$
+& $0.5$
+& $0.8$
 \\[0.2em]
@@
 & $1+1.5-1.0$
+& $1.6$
+& $1.7$
 \\[0.2em]
@@
 These are evaluated from the central and alternative samples as described in
-the text.
-Note how up and down variations can move in different directions, and these
-different movements are coupled in the background model.
+the text of Section~\ref{sec:2ljets_sample_comparisons};
+$\delta_{12}$ is the difference of the alternative from the central sample
+and $\sigma_2$ is the statistical uncertainty on the alternative sample.
+Note that the up and down variations can move in different directions, and
+these different movements are coupled in the background model.
 }

### 2022-07-22 main/chapter_2ljets.tex
expand zjets alt; expand intro to excuse stupid structure; add some section labels

@@
 
-My core contribution in this thesis is to the recent \atlas\
-search ``for new phenomena in events with two leptons, jets, and
-missing transverse momentum''~\cite{atlas2022searches}.
-For brevity, I will call this search $\twoljets$.
+My core contributions in this thesis are to the recent \atlas\
+searches ``for new phenomena in events with two leptons, jets, and
+missing transverse momentum''~\cite{atlas2022searches},
+particularly to its electroweak part.
+For brevity, I will call this publication $\twoljets$ and its electroweak
+subset $\twoljets$-electroweak.
+
+This chapter follows a questionable order.
+It may be clearer, and is certainly more ordinary,
+to first present preparations (blind to data),
+followed later by the results (unblinded).
+Instead, this chapter first reports the unblinded results, then follows with
+in-depth discussions of individual elements, partially in light of those
+results.
+I choose this order for two reasons.
+First, it avoids iterating over those elements twice, once before and again
+after unblinding.
+The analysis has many little and loosely related pieces that I attempt to
+to cover locally to reduce jumps between far-flung corners of the
+chapter.
+Second, and sadly, many design decisions can only be explained in light of the
+unblinded results, since important aspects were reworked in the
+$19$~months between its unblinding and eventual publication.
 
 There are certain plots and tables which, personally, I first seek when
-meeting an \atlas\ search paper.
-This chapter begins by presenting those results in brief in
-Section~\ref{sec:2ljets_splash}.
-The meat of this chapter (or, if you prefer, its gory detail) follows in
-Section~\ref{sec:2ljets_context} onwards.
+meeting an \atlas\ search paper, particularly the summary plot with all signal
+regions and the model-dependent exclusion contours.
+This chapter begins by presenting those results in brief in Section
+~\ref{sec:2ljets_splash}.
+The meat of the chapter follows.
+\begin{itemize}
+\item Context from previous searches and inspirations for the design goals
+of the $\twoljets$-electroweak search are described in
+Section~\ref{sec:2ljets_context}.
+\item Data, their event variables and the reconstructed objects from which
+they derive are described in Section~\ref{sec:2ljets_events}.
+\item Region design for the $\twoljets$-electroweak search is described and
+explained in Section~\ref{sec:2ljets_design}.
+\item The modelling of our uncertain predictions of background and signal
+contributions in those regions is covered in
+Section~\ref{sec:2ljets_modelling}.
+\item Some validation studies are presented in Section~\ref{sec:2ljets_validation}.
+\item The unblinding procedure of this search is discussed in
+Section~\ref{sec:2ljets_unblinding}, and
+\item we recap the search results in Section~\ref{sec:2ljets_results}.
+\end{itemize}
 
@@
 \section{Events}
+\label{sec:2ljets_events}
+\TODO{make sure we say 136ifb from 2015 to 2018 run 2 data. Good run lists etc}
+
 This section describes the events selected for the $\twoljets$ searches, the
@@
 \section{Design}
+\label{sec:2ljets_design}
 Perhaps the most impactful decision in a search is the design of its regions.
@@
 \section{Modelling}
+\label{sec:2ljets_modelling}
 So far, we have seen many histograms of background samples overlaid with
@@
 
+\subsubsection{Alternative \texorpdfstring{$\zjets$}{Z+jets}}
+\label{sec:2ljets_zjets_alt}
 As mentioned in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison
@@
 showering and hadronization.
-These plots show statistical error bars only, and the larger statistical error
-bars on the alternative sample are due to it comprising around one sixth as
-many simulated events as the central sample.
+This \madgraph\ alternative sample simulates only $\mll > 40\,\eV[G]$, so in
+regions allowing $\mll$ outside of this we evaluate relative variations and
+scale them up to the central yield.
+The plots in Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} show statistical
+error bars only; the alternative sample has larger statistical errors since
+it contains around one sixth as many simulated events as the central sample,
+and that difference is not countered by their different weight distributions
+in this region.
 
@@
 \item if the deviation is statistically significant, then preserve the direction
-of the uncertainty, so $\sigma = \mathrm{sgn}\sigma_{12}$.
+of the uncertainty, so $\sigma = \mathrm{sgn}(\delta_{12})\sigma_{12}$.
 \end{itemize}
@@
 we are unable to give no direction to any deviation, and we choose to
-we use the same, signed, definition of $\sigma$ for all regions.
+use the same, signed, definition of $\sigma$ for all regions.
 
 As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
-$\sigma_2$ is a substantial contribution to these variations;
+$\sigma_2$ is a substantial contribution to these variations, and so
 the statistical uncertainty in the alternative sample does significantly
 expand this variation.
-Indeed, it is the larger contribution in SR-Low-a and SR-Low-b, which are the
-regions most sensitive to $\zjets$ variations.
+Indeed, $\sigma_2$ is the larger contribution in SR-Low-a and SR-Low-b, which
+are the regions most sensitive to $\zjets$ variations.
 Our unprescribed choice to match their directions therefore matters.
@@
 I believe such a change would have improved the plausibility of our background
-modelling, and I regret not finding the capacity to push it at the time.
+modelling, and I regret not finding the capacity to push for it at the time.
 
@@
 comparisons.
-Since our search is in the SUSY group we follow their instructions.
-
-Figure~\ref{fig:2ljets_low_crz_pre_post}
-Figure~\ref{fig:2ljets_vrlow_alt}
-Figure~\ref{fig:2ljets_low_vrlow_pre_post}
-
+Since our search is in the SUSY group, we follow their instructions.
+
+Including this $\zjets$ sample comparison greatly improves our model's post-fit
+agreement with the SR-Low deficits.
+It therefore seems proper to disclose that this variation was added after
+unblinding, and after we all knew about those deficits.
+Prior to unblinding, we in the $\twoljets$ team had not planned on including it
+in addition to other theoretical uncertainties, and I do not know whether or
+not senior members in the collaboration would have required its inclusion
+without the deficit.
+The \atlas\ SUSY recommendation is to consider the comparison if
+``the scale ($+$ PDF) uncertainties are too small in your
+analysis''~\cite{atlas_twiki_susy_v_jets},
+but I know of no official advice on how to judge what is ``too small''.
+
+\begin{table}[tp]
 % in branch hackmg5
@@
 %               VRllbb,     0.31,     0.36
-\begin{table}[tp]
 \centering
@@
 & $1.7$
-7 $3.2$
+& $3.2$
 \\[0.5em]
@@
 
+Auxiliary data to the $\twoljets$ paper~\cite{atlas2022searches} include
+Figure~\ref{fig:2ljets_low_crz_pre_post_post}~\cite{hepdata.116034.v1/t101},
+which appears to show good agreement in SR-Low and CR-Z, a relatively
+small error bar in SR-Low of around $\pm10$ events per $100\,\eV[G]$,
+and a curious dip in the $\zjets$ background in the signal region.
+The figure caption also admits that
+``All EWK search control and signal regions are included in the fit.''
+
+Fitting only to control region (not signal regions) results in
+Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
+the signal region, with diboson yields peaking near to $m(W)$ above a smooth
+varying $\zjets$ background.
+Fitting to the signal region slashes the $\zjets$ background in SR-Low by
+around $80\%$. How?
+This sample comparison is the largest variation in VR-Low, and we have seen
+from Table~\ref{tab:2ljets_zjets_alt} that SR-Low variations are around $30\%$.
+A few factors contribute:
+\begin{itemize}
+\item The $\zjets$ sample is normalized to CR-Z with an unconstrained
+normalization factor. This means that variations to the $\zjets$ sample in CR-Z
+are absorbed by scaling all of its modifications in other regions.
+\item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
+in CR-Z match to negative variations in SR-Low bins; the two yields are
+anticorrelated.
+With normalization to CR-Z, this anticorrelation increases the effective
+variation in SR-Low to $\pm40\%$.
+\item Post-fit, the parameter controlling this $\zjets$ sample comparison
+modifier is pulled to around $2\sigma$. This is seen in
+Figure~\ref{fig:2ljets_fit_parameters}, and will be discussed further in
+Section~\ref{sec:2ljets_results}.
+\end{itemize}
+A two-sigma pull on a $40\%$ scaling therefore explains this $80\%$ change.
+Validation regions do not constrain our fits, and are also affected by
+this effect; in VR-Low, small differences between the central and alternative
+$\zjets$ samples are visible in Figure~\ref{fig:2ljets_vrlow_alt}, but
+after fitting to the signal regions,
+Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
+is almost wiped out, leaving a deficit below data.
+
+To avoid this extreme effect, I searched for alternative configurations
+to the background model, such as constraining the parameter to $1$-sigma.
+More senior members of the \atlas\ SUSY group decided that we should keep it as
+described, reasoning along the lines that
+a) this is a real and significant deficit in SR-Low that must appear as either
+a pull on systematic variations or as a deviation from data, and
+b) we prefer the pull on systematic variations.
+With this decision, post-fit results showed no deficit;
+this $\zjets$ sample comparison is the main reason why
+Figure~\ref{fig:2ljets_summary} shows no significant in SR-Low-a or SR-Low-b.
+Internal reviewers were then confused as to why we mentioned a deficit in
+SR-Low, so, much to my ire, all mentions of an SR-Low deficit were erased from
+the paper.
+
+To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
+excessive diboson background should also take some blame for this poor
+prediction in SR-Low.
+We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
+not similarly squashed.
+
 \begin{figure}[tp]
@@
 \caption{Fitted to control regions only}
+\label{fig:2ljets_low_crz_pre_post_pre}
 \end{subfigure}
@@
 \caption{Fitted to control and signal regions}
+\label{fig:2ljets_low_crz_pre_post_post}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow.pdf}
-\caption{Standard}
+\caption{Central}
 \end{subfigure}
@@
 \caption[
-Pre-fit yields in VR-Low with standard and alternative $\zjets$ simulations
-]{%
-Pre-fit yields in VR-Low with (a) standard and (b) alternative $\zjets$
+Pre-fit yields in VR-Low with central and alternative $\zjets$ simulations
+]{%
+Pre-fit yields in VR-Low with (a) central and (b) alternative $\zjets$
 simulations.
 Errors are statistical.
+Results in these regions with all uncertainties, and fitted to data in
+control or signal regions are presented, in
+Figure~\ref{fig:2ljets_low_vrlow_pre_post}.
 }
@@
 region, whose likelihood is not included in the \heplikelihood.
+Pre-fit results with statistical uncertainties only are presented in
+Figure~\ref{fig:2ljets_vrlow_alt}.
+Sorry for the different order of backgrounds; this figure uses
+conventions from the paper, while Figure~\ref{fig:2ljets_vrlow_alt}
+uses the conventions we used in development.
 }
@@
 \end{figure}
+
+\subsubsection{Alternative diboson}
+\label{sec:2ljets_diboson_alt}
+\TODO{}
+
+\subsubsection{Alternative \texorpdfstring{$\ttbar$}{ttbar}}
+\label{sec:2ljets_ttbar_alt}
+\TODO{}
 
@@
 \section{Unblinding}
-
+\label{sec:2ljets_unblinding}
 Figure~\ref{fig:2ljets_prepre}
@@
 \section{Results}
-
+\label{sec:2ljets_results}
 Figure~\ref{fig:2ljets_summary_no_sr}

### 2022-07-22 main/chapter_2ljets.tex
clarify erasure of SRLow deficit

@@
 b) we prefer the pull on systematic variations.
-With this decision, post-fit results showed no deficit;
-this $\zjets$ sample comparison is the main reason why
-Figure~\ref{fig:2ljets_summary} shows no significant in SR-Low-a or SR-Low-b.
-Internal reviewers were then confused as to why we mentioned a deficit in
-SR-Low, so, much to my ire, all mentions of an SR-Low deficit were erased from
+
+With this decision, post-fit results, as presented in
+Figure~\ref{fig:2ljets_summary},
+indeed show no significant deviation from data in SR-Low-a or SR-Low-b,
+and this $\zjets$ sample comparison is the main reason why.
+Internal reviewers were, however, confused as to why we mentioned a deficit in
+SR-Low that did not appear in the results plots or tables.
+Much to my ire, all mentions of an SR-Low deficit were therefore removed from
 the paper.
+Despite bring real and significant, this deficit was erased from public view,
+and to my knowledge it is only reported through the HistFactory models in our
+HEPData~\cite{hepdata.116034}.
 

### 2022-07-22 main/chapter_2ljets.tex
more rework of zjets alt comparison

@@
 
-To avoid this extreme effect, I searched for alternative configurations
-to the background model, such as constraining the parameter to $1$-sigma.
+It is not clear that such a large extrapolation from the difference between
+two simulations is physically reasonable, especially since a large fraction
+of its effect is due to the statistical error on a small alternative sample.
+To avoid this questionable effect, I searched for alternative configurations
+of the background model, such as constraining the parameter to $1$-sigma.
 More senior members of the \atlas\ SUSY group decided that we should keep it as
 described, reasoning along the lines that
-a) this is a real and significant deficit in SR-Low that must appear as either
+a) this is a real and significant deficit in SR-Low, so must appear as either
 a pull on systematic variations or as a deviation from data, and
@@
 
-With this decision, post-fit results, as presented in
+With this decision, post-fit results, such as the summary plot of
 Figure~\ref{fig:2ljets_summary},
@@
 Despite bring real and significant, this deficit was erased from public view,
-and to my knowledge it is only reported through the HistFactory models in our
-HEPData~\cite{hepdata.116034}.
+and to my knowledge it is only reported through the HistFactory
+models in our HEPData~\cite{hepdata.116034}.
 

### 2022-07-22 main/chapter_2ljets.tex
clarify extended fits Poisson stuff

@@
 expectations~\cite{skilling2009cosmology}, and in the small-bin limit one has
-continuous distributions
-(sometimes called ``extended''~\cite{barlow1990extended}),
-which are standard and easily analysed.
+continuous distributions with Poisson normalizations
+(sometimes called ``extended'' in high-energy
+physics~\cite{barlow1990extended})
+that are standard and easily analysed.
 

### 2022-07-25 main/chapter_2ljets.tex
explain why zero low

@@
 \end{itemize}
-All low variations are set to zero.
+All low variations are set to zero for these replacement FNP estimates, to
+encode that the yield is compatible with zero, but more importantly to give
+a slow extrapolation on the low side (from the small central value to zero)
+and reduce the risk from non-physical negative yields in the HistFactory model.
 

### 2022-07-25 main/chapter_2ljets.tex
magician fake

@@
 \end{align}
+That is, the Matrix Method starts from data, flips them over to truth
+estimates, then flops those back to data-level estimates. Flip, flop.
+
 To make FNP estimates in our regions, the Matrix Method is
@@
 tight-tight events which comprise the data of our search regions, and those
-are the same data in which we are searching for anomalies.
+are the same data in which we are searching for anomalous signals.
 Since it includes actual data, this method unblinds all regions,
@@
 Although it does cheat by using the data we should be blind to, it uses those
-data in a diluted way, such that they can end up not making much difference.
+data in a diluted way, such that, as far as I know, they might end up not
+making much difference.
+It is not a magician's card trick, but an honest (if flawed) attempt to make
+useful estimates.
 
@@
 This count of $157$ differs from the $194$ actual data observed in CR-VZ.
-I do not know why.
+I do not know why, and do not have access to the source data or code.
 The two do overlap in event identity numbers, so there is presumably some
-tighter selection in the Matrix Method, which should be accounted for in its
+tighter selection in the Matrix Method that should be accounted for in its
 weights.

### 2022-07-25 main/chapter_2ljets.tex
lots more diboson variation stuff

@@
 \times p(t\!\mid\!r,2)
-\times \bar \mu^\textrm{truth}_{rr}~~, \\
+\times \bar \mu^\textrm{truth}_{rr}, \\
 % fr
@@
 \times p(t\!\mid\!r,2)
-\times \bar \mu^\textrm{truth}_{fr}~~, \\
+\times \bar \mu^\textrm{truth}_{fr}, \\
 % rf
@@
 \times p(t\!\mid\!f,2)
-\times \bar \mu^\textrm{truth}_{rf}~~,~~\textrm{and} \\
+\times \bar \mu^\textrm{truth}_{rf},~\textrm{and} \\
 % ff
@@
 \times p(t\!\mid\!f,2)
-\times \bar \mu^\textrm{truth}_{ff}~~.
+\times \bar \mu^\textrm{truth}_{ff}.
 \end{align}
@@
 implemented by assigning weights to observed data events that are then
-histogrammed them in the tighter selections of our search
-regions much like other samples~\cite{twoljets2018int}.
+histogrammed them in the tighter selections of our search regions,
+much like other samples~\cite{twoljets2018int}.
 Systematic variations are assigned by comparing histograms when those weights
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWK_met_Sign.png}
+\caption{}
+\label{fig:2ljets_vrss_mm}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWKMC_met_Sign.png}
+\caption{}
+\label{fig:2ljets_vrss_mc}
+\end{subfigure}
+\caption[
+Same-sign validation with Matrix Method and Monte Carlo fake estimates
+]{%
+Same-sign validation with (a) Matrix Method and (b) Monte Carlo fake estimates.
+Backgrounds are pre-fit.
+Errors are statistical summed in quadrature with variations in Matrix Method
+sample weights.
+Those weight variations are coupled between bins, meaning that the
+directed differences between bins are less significant than an independent
+combination.
+The Monte Carlo estimate is clearly more accurate.
+\\[0.5em]
+% thesis_project/2ljets/vrss_plots/PAM_FNP_50pc.pdf
+Part (a) was included in an early draft of the paper.
+It was removed after validation studies showed that adding an additional
+$50\%$ uncertainty to FNP yields, coupled between all bins, did not
+significantly change our exclusion contours;
+most contours shrank by less than their plotted line width.
+The largest change was a $2\%$ reduction in the
+$B(\neutralino_1 \rightarrow h \gravitino)$
+limit at $m(\neutralino_1) \approx 500\,\eV[G]$.
+}
+\label{fig:2ljets_vrss}
+\end{figure}
+
 Overlaying these modified contours highlighted some minor differences, the
@@
 Table~\ref{tab:2ljets_fnp_estimates}.
-
-These three estimates are labelled `MC', `MM', `Mean', and we combine them into
-the `Max' estimate.
-Each is described here.
-\begin{itemize}
-\item The `MC' estimate takes the simulated estimate that was previously reported
-in Table~\ref{tab:2ljets_fnp_negative_mm}.
-\item The `MM' estimate modifies the Matrix Method result,
-also from Table~\ref{tab:2ljets_fnp_negative_mm}, by shifting its central
-value to a small positive value chosen baselessly as $0.01$ and by adding
-its statistical and systematic uncertainties linearly
-(not in quadrature to conservatively cover a correlated variation)
-to assign a high variation.
-\end{itemize}
-After I had proposed a combination of these two estimates, reviewers
-suggested additionally using the statistical properties of samples in looser
-selections.
-This became the third `Mean' estimate.
-\begin{itemize}
-\item The `Mean' follows an idea that the smallest assignable value should
-be comparable to typical Matrix Method weights, and those typical values
-are extracted from a region requiring pre-selection,
-from Table~\ref{tab:2ljets_presel},
-$\mll \in 71\textrm{--}101$,
-$\mjj \in 60\textrm{--}100$,
-$\njet \geq 2$, and
-$\nbtag \leq 1$.
-Its central value takes the mean weight of Matrix Method samples in this
-selection, which is $\mu^\textrm{M} = 0.0202$.
-Statistical errors are set to the root of the mean squared weight,
-which is $\sigma^\textrm{M}_\textrm{stat} = 0.1132$,
-and systematic errors are set as the larger absolute deviation on the mean
-from the up and down variations in the Matrix Method weights, which comes out
-$\sigma^\textrm{M}_\textrm{sys} = 0.0226$.
-Finally, the `Mean' high variation is their linear sum;
-$\mu^\textrm{M} + \sigma^\textrm{M}_\textrm{stat} +
-\sigma^\textrm{M}_\textrm{sys} = 0.1560$.
-\end{itemize}
-All low variations are set to zero for these replacement FNP estimates, to
-encode that the yield is compatible with zero, but more importantly to give
-a slow extrapolation on the low side (from the small central value to zero)
-and reduce the risk from non-physical negative yields in the HistFactory model.
-
-As seen from Table~\ref{tab:2ljets_fnp_estimates}, even the high variations of
-the Max method are negligibly small next to the central simulated yields
-in most regions.
-The exception regions are all in the High category, and are those regions with
-small enough expected yields such that the Mean method dominates.
-
-An awkward fact in the history of this combined FNP estimate is that it
-was reworked after unblinding to include the Mean estimate and more regions;
-with a reprocessing of our background simulations, changes in the Matrix Method
-estimates left more of our regions with non-positive estimate than prior to
-unblinding the data.
-To assign them without being biased by the unblinded signal region data,
-which were known to contain no significant excesses, therefore took some
-barefaced honesty.
-Thankfully, we do not need to consider what might have happened had an excess
-been observed before the assignment of these fake estimates.%
-\footnote{%
-Of course $p$-values \emph{do} depend on what might have happened with
-alternative data
-(those with test statistics greater than or equal to that observed),
-it would be wasteful and unproductive to consider them seriously.
-I am grateful that \atlas' statistical conventions allow us to ignore these
-problems.
-}
-Despite any awkwardness, it does appear that FNP backgrounds are indeed
-small and that they do not greatly impact the result, especially when compared
-to other systematic variations.
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWK_met_Sign.png}
-\caption{}
-\label{fig:2ljets_vrss_mm}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_vrss_VR_SS_EWKMC_met_Sign.png}
-\caption{}
-\label{fig:2ljets_vrss_mc}
-\end{subfigure}
-\caption[
-Same-sign validation with Matrix Method and Monte Carlo fake estimates
-]{%
-Same-sign validation with (a) Matrix Method and (b) Monte Carlo fake estimates.
-Backgrounds are pre-fit.
-Errors are statistical summed in quadrature with variations in Matrix Method
-sample weights.
-Those weight variations are coupled between bins, meaning that the
-directed differences between bins are less significant than an independent
-combination.
-The Monte Carlo estimate is clearly more accurate.
-\\[0.5em]
-% thesis_project/2ljets/vrss_plots/PAM_FNP_50pc.pdf
-Part (a) was included in an early draft of the paper.
-It was removed after validation studies showed that adding an additional
-$50\%$ uncertainty to FNP yields, coupled between all bins, did not
-significantly change our exclusion contours;
-most contours shrank by less than their plotted line width.
-The largest change was a $2\%$ reduction in the
-$B(\neutralino_1 \rightarrow h \gravitino)$
-limit at $m(\neutralino_1) \approx 500\,\eV[G]$.
-}
-\label{fig:2ljets_vrss}
-\end{figure}
 
@@
 
+These three estimates are labelled `MC', `MM', `Mean', and we combine them into
+the `Max' estimate.
+Each is described here.
+\begin{itemize}
+\item The `MC' estimate takes the simulated estimate that was previously reported
+in Table~\ref{tab:2ljets_fnp_negative_mm}.
+\item The `MM' estimate modifies the Matrix Method result,
+also from Table~\ref{tab:2ljets_fnp_negative_mm}, by shifting its central
+value to a small positive value chosen baselessly as $0.01$ and by adding
+its statistical and systematic uncertainties linearly
+(not in quadrature to conservatively cover a correlated variation)
+to assign a high variation.
+\end{itemize}
+After I had proposed a combination of these two estimates, reviewers
+suggested additionally using the statistical properties of samples in looser
+selections.
+This became the third `Mean' estimate.
+\begin{itemize}
+\item The `Mean' follows an idea that the smallest assignable value should
+be comparable to typical Matrix Method weights, and those typical values
+are extracted from a region requiring pre-selection,
+from Table~\ref{tab:2ljets_presel},
+$\mll \in 71\textrm{--}101$,
+$\mjj \in 60\textrm{--}100$,
+$\njet \geq 2$, and
+$\nbtag \leq 1$.
+Its central value takes the mean weight of Matrix Method samples in this
+selection, which is $\mu^\textrm{M} = 0.0202$.
+Statistical errors are set to the root of the mean squared weight,
+which is $\sigma^\textrm{M}_\textrm{stat} = 0.1132$,
+and systematic errors are set as the larger absolute deviation on the mean
+from the up and down variations in the Matrix Method weights, which comes out
+$\sigma^\textrm{M}_\textrm{sys} = 0.0226$.
+Finally, the `Mean' high variation is their linear sum;
+$\mu^\textrm{M} + \sigma^\textrm{M}_\textrm{stat} +
+\sigma^\textrm{M}_\textrm{sys} = 0.1560$.
+\end{itemize}
+All low variations are set to zero for these replacement FNP estimates, to
+encode that the yield is compatible with zero, but more importantly to give
+a slow extrapolation on the low side (from the small central value to zero)
+and reduce the risk from non-physical negative yields in the HistFactory model.
+
+As seen from Table~\ref{tab:2ljets_fnp_estimates}, even the high variations of
+the Max method are negligibly small next to the central simulated yields
+in most regions.
+The exception regions are all in the High category, and are those regions with
+small enough expected yields such that the Mean method dominates.
+
+An awkward fact in the history of this combined FNP estimate is that it
+was reworked after unblinding to include the Mean estimate and more regions;
+with a reprocessing of our background simulations, changes in the Matrix Method
+estimates left more of our regions with non-positive estimate than prior to
+unblinding the data.
+To assign them without being biased by the unblinded signal region data,
+which were known to contain no significant excesses, therefore took some
+barefaced honesty.
+Thankfully, we do not need to consider what might have happened had an excess
+been observed before the assignment of these fake estimates.%
+\footnote{%
+Although $p$-values \emph{do} depend on what might have happened with
+alternative data
+(those with test statistics greater than or equal to that observed),
+it would be wasteful and unproductive to consider them seriously.
+I am grateful that \atlas' statistical conventions allow us to ignore these
+problems.
+}
+Despite any awkwardness, it does appear that FNP backgrounds are indeed
+small and that they do not greatly impact the result, especially when compared
+to other systematic variations.
+
 \subsection{Jet mass}
@@
 Bothmann:2019yzt,
+Gleisberg:2008ta,
 ATL-PHYS-PUB-2016-003
@@
 in this region.
+
+Since \sherpa implements both the hard scatter and shower models, we are
+considering variations due to the simultaneous change of both of these
+components together.
 
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
+\caption{Fitted to control regions only}
+\label{fig:2ljets_low_crz_pre_post_pre}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
+\caption{Fitted to control and signal regions}
+\label{fig:2ljets_low_crz_pre_post_post}
+\end{subfigure}
+\caption[
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+]{%
+Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+(a) fitted to control regions, and (b) fitted to control and signal regions.
+The arrows point in to the signal region, which shows a large peak in the
+diboson background near $m(W)$.
+Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
+the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
+Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
+and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
+}
+\label{fig:2ljets_low_crz_pre_post}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow.pdf}
+\caption{Central}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow_mg5.pdf}
+\caption{Alternative}
+\end{subfigure}
+\caption[
+Pre-fit yields in VR-Low with central and alternative $\zjets$ simulations
+]{%
+Pre-fit yields in VR-Low with (a) central and (b) alternative $\zjets$
+simulations.
+Errors are statistical.
+Results in these regions with all uncertainties, and fitted to data in
+control or signal regions are presented, in
+Figure~\ref{fig:2ljets_low_vrlow_pre_post}.
+}
+\label{fig:2ljets_vrlow_alt}
+\end{figure}
+
 It is not clear that such a large extrapolation from the difference between
@@
 Despite bring real and significant, this deficit was erased from public view,
-and to my knowledge it is only reported through the HistFactory
-models in our HEPData~\cite{hepdata.116034}.
+and to my knowledge is only reported through the inner details of
+the HistFactory models shared in our HEPData~\cite{hepdata.116034}.
 
@@
 excessive diboson background should also take some blame for this poor
-prediction in SR-Low.
+prediction in SR-Low;
+its diboson yield peaks in the signal region on $\mjj \approx m(W)$, but
+we do not expect $W\rightarrow jj$ in this process since one boson is needed
+for the leptons and the other for the $\met$ that we select.
 We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
-not similarly squashed.
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
-\caption{Fitted to control regions only}
-\label{fig:2ljets_low_crz_pre_post_pre}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
-\caption{Fitted to control and signal regions}
-\label{fig:2ljets_low_crz_pre_post_post}
-\end{subfigure}
-\caption[
-Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
-]{%
-Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
-(a) fitted to control regions, and (b) fitted to control and signal regions.
-The arrows point in to the signal region, which shows a large peak in the
-diboson background near $m(W)$.
-Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
-the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
-Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
-and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
-}
-\label{fig:2ljets_low_crz_pre_post}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow.pdf}
-\caption{Central}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_vrlow_met_Sign_VRLow_mg5.pdf}
-\caption{Alternative}
-\end{subfigure}
-\caption[
-Pre-fit yields in VR-Low with central and alternative $\zjets$ simulations
-]{%
-Pre-fit yields in VR-Low with (a) central and (b) alternative $\zjets$
-simulations.
-Errors are statistical.
-Results in these regions with all uncertainties, and fitted to data in
-control or signal regions are presented, in
-Figure~\ref{fig:2ljets_low_vrlow_pre_post}.
-}
-\label{fig:2ljets_vrlow_alt}
-\end{figure}
+not similarly squashed by sample comparison variations.
 
@@
 \label{sec:2ljets_diboson_alt}
-\TODO{}
+No alternative samples were available for comparison with our central diboson
+samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
+electroweak vertices.
+To my knowledge, this is simply a limitation from what \atlas\ has decided to
+spend its resources on, and quite reasonably since full simulation is
+computationally expensive.
+We did, however, have access to the cheaper alternative of set of small,
+truth-level samples, that were simulated with the same \sherpa\ setup but with
+varied parameters to assess uncertainties in the modelling of parton showers.
+
+These small truth samples include variations of three parameters:
+\begin{itemize}
+% CKKW: merging scale
+\item $Q_\textrm{cut}^\textrm{CKKW}$, the energy scale for merging
+matrix elements with parton showers under the CKKW
+scheme~\cite{Catani_2001_CKKW,Hoeche:2009rj}.
+It is varied from its default $20\,\eV[G]$ up to $30\,\eV[G]$ and down
+to $15\,\eV[G]$.
+% CSS_KIN: CSS is the combined matrix element + shower
+\item \texttt{CSS\_KIN}, a binary parameter that tells \sherpa\ how to treat
+kinematic recoils in its parton showers~\cite{sherpa_222_manual}.
+It has two modes, one default~\cite{Hoeche:2009rj}, and one
+alternative~\cite{Hoeche:2009xc}
+that is recommended for studies of systematic variations.
+These differ in how they resolve the problem that an isolated massless parton
+cannot conserve four-momentum when splitting into two non-collinear massless
+children (for example gluon emission $q\rightarrow gq$);
+each scheme achieves conservation by recoiling off different parts of the
+parton shower.
+% resummation: parton showing including all orders in QCD
+% https://link.springer.com/article/10.1007/JHEP01(2018)118
+% "radiative corrections determined in QCD perturbation theory must be resummed
+% to all orders."
+% https://sherpa-team.gitlab.io/sherpa/tmp-update-extamp-examples/manual/parameters/matrix-elements.html
+% "In parton shower matched/merged calculations a third perturbative scale is
+% present, the resummation or parton shower starting scale. It can be set by
+% the user in the third argument like"
+% METS is a scale setter
+% QSF: scale factor for the resummation scale; relates to parton showers; quadratic
+% vary up and down by a factor of 4, but QSF is quadratic in the scale so this is
+% an implementation details of the usual 2x 0.5x form
+\item $S_Q$, a scale factor on the resummation scale $\mu_{Q^2}$ that is set
+from the hard scatter kinematics in each event and affects emissions in the
+parton shower.
+Its central value is $1$, and it is varied up and down by factors of $4$.
+Since \sherpa\ defines its scales in a quadratic form, this matches the typical
+convention of considering scale variations over factors of
+$2$~\cite{sherpa_222_manual,Bothmann:2019yzt}.
+\end{itemize}
+Truth-level background yields in our regions were calculated with the
+SimpleAnalysis~\cite{Lorenz:2752450,simpleanalysis_cern}
+software package for
+our central diboson samples and these varied alternatives.
+SimpleAnalysis histograms truth-level simulations and optionally smears them
+with calibrated noise to approximate reconstruction-level results.
+\emph{I am grateful to Jon~Long for preparing our initial
+SimpleAnalysis implementations and sharing software to evaluate these
+variations.}
+
+Unfortunately, a total failure of statistical precision means that no
+meaningful result can be derived from these variations.
+First our regions receive very few raw events in each variation,
+for example $1$ sample in SR-$\llbb$, around a dozen in
+each SR-High bin, and a maximum of $597$ in CR-VZ.
+Second, those samples carry weights in almost-equal mixtures of $+1$ and $-1$,
+so they further lose precision in cancellation between the positive and
+negative halves.
+
+Variations for these samples were assigned by a scheme that follows a
+convention set by other parts of the $\twoljets$ electroweak analysis.
+With given central, up, and down variations of the yield, it works as
+follows.
+\begin{enumerate}
+\item If both up and down variations lie to the same side of central, then
+the variation is half the difference from up and down.
+\item Otherwise (opposite sides of central), the variation is half of the
+maximum absolute difference between central and up or down.
+\item Any variation, from the above two bullets, that is smaller than the
+largest statistical uncertainty on any of the samples is increased to the size
+of that statistical uncertainty.
+\item Directions of variation preserved by maintaining their signs, and
+each variation is coupled between bins with its own single parameter.
+\end{enumerate}
+Point 1 acts to mitigate large variations where the variations lie far from
+the central sample,
+and all remainders are dominated by the statistical uncertainty due to point 3.
+
+In the signal regions, most of these statistical variations are around
+$100\%$ relative to the central yield, with a minimum across all variations
+of $26\%$ and a maximum of $960\%$.
+For $Q_\textrm{cut}^\textrm{CKKW}$ and $S_Q$ their signs appear randomly
+distributed in positive and negative directions in approximately equal
+proportions, but for \texttt{CSS\_KIN} most are correlated in the positive
+direction, with the only SR-OffShell-a moving in the negative direction, by
+$46\%$.
+Variations in control regions are typically smaller, ranging from $12\%$ to
+$50\%$, but remain limited by statistical precision.
+
+To include these large variations moving in wacky directions could allow the
+background model to easily fit deviations on the basis of our ignorance
+from statistical imprecision.
+To mitigate this, we instead take the variations from looser regions
+VR-sys-main and VR-sys-OffShell, that target on- and off-shell regions
+respectively and hope to improve the statistical precision.
+Both require the pre-selection from Table~\ref{tab:2ljets_presel};
+VR-sys-main also requires
+$\mjj \in 60\textrm{--}110\,\eV[G]$ and $\nbtag \leq 1$,
+and VR-sys-OffShell requires
+$\nbtag = 0$, $\ptjone > 100\,\eV[G]$ and $\dphijmet > 2$ to target the
+OffShell regions.
+No extra lepton selections are made for these regions since these variations
+relate primarily to jets through parton showers, and not to leptons, and
+relaxing cuts can only improve statistical precision.
+
+In these looser regions VR-sys-main and VR-sys-OffShell, each variation
+reduces to around $10\%$ and remains dominated by the statistical uncertainty.
+Although local effects could occur, this a reasonable indication that these
+parameters should not much affect our signal regions.
+We therefore apply the variations measured in VR-sys-main to all on-shell
+regions,
+and the variations measured in VR-sys-OffShell to off-shell regions,
+with one parameter coupling all regions for each of
+$Q_\textrm{cut}^\textrm{CKKW}$,
+\texttt{CSS\_KIN}, and
+$S_Q$.
+
+This assignment varies CR-VZ and on-shell other regions identically,
+and since the diboson background is normalized to CR-VZ this means that
+these variations have no effect in on-shell regions, and cannot move to
+fit the SR-Low or SR-High-8 deficits.
+Had we considered complete alternative diboson samples, we would certainly
+had had some meaningful variations to apply in those regions, but I have no
+way of knowing how large those variations would be.
 
@@
 \label{sec:2ljets_theory_variations}
-
+\TODO{put this above alternative samples}
 % extrapolation

### 2022-07-25 main/chapter_2ljets.tex
expand conclusion on diboson; add some todos

@@
 fit the SR-Low or SR-High-8 deficits.
+
 Had we considered complete alternative diboson samples, we would certainly
-had had some meaningful variations to apply in those regions, but I have no
-way of knowing how large those variations would be.
-
+have had some meaningful variations to apply in those regions, but I do not
+know how large those variations should be.
+This lack of a real alternative sample is particularly disappointing since
+diboson is our most important background.
+In hindsight, it may have been sensible to push for more resources to be
+allocated to resolve this problem.
+
+\TODO{Other theory variations; these are less important / interesting.
+bullet point quickly.}
 \subsubsection{Alternative \texorpdfstring{$\ttbar$}{ttbar}}
@@
 
+\TODO{Signal theory variations}
 \subsection{Signals}
@@
 \label{sec:2ljets_theory_variations}
-\TODO{put this above alternative samples}
+\TODO{put this above alternative samples?}
+\TODO{this appears to be signal theory}
 % extrapolation

### 2022-07-25 main/chapter_2ljets.tex
enumerate subsubsections; fix a math title

@@
 
-\subsubsection{Fake/non-prompt estimates in $\twoljets$-electroweak regions}
+\subsubsection{%
+Estimates in \texorpdfstring{$\twoljets$}{2Ljets}-electroweak regions%
+}
 To validate the Matrix Method estimates in $\twoljets$-electroweak regions

### 2022-07-27 main/chapter_2ljets.tex
flawed->imperfect; senior members to committee; relocate a figure

@@
 making much difference.
-It is not a magician's card trick, but an honest (if flawed) attempt to make
+It is not a magician's card trick, but an honest (if imperfect) attempt to make
 useful estimates.
@@
 in addition to other theoretical uncertainties, and I do not know whether or
-not senior members in the collaboration would have required its inclusion
+not committee decisions in the collaboration would have required its inclusion
 without the deficit.
@@
 
-It is not clear that such a large extrapolation from the difference between
-two simulations is physically reasonable, especially since a large fraction
-of its effect is due to the statistical error on a small alternative sample.
-To avoid this questionable effect, I searched for alternative configurations
-of the background model, such as constraining the parameter to $1$-sigma.
-More senior members of the \atlas\ SUSY group decided that we should keep it as
-described, reasoning along the lines that
-a) this is a real and significant deficit in SR-Low, so must appear as either
-a pull on systematic variations or as a deviation from data, and
-b) we prefer the pull on systematic variations.
-
-With this decision, post-fit results, such as the summary plot of
-Figure~\ref{fig:2ljets_summary},
-indeed show no significant deviation from data in SR-Low-a or SR-Low-b,
-and this $\zjets$ sample comparison is the main reason why.
-Internal reviewers were, however, confused as to why we mentioned a deficit in
-SR-Low that did not appear in the results plots or tables.
-Much to my ire, all mentions of an SR-Low deficit were therefore removed from
-the paper.
-Despite bring real and significant, this deficit was erased from public view,
-and to my knowledge is only reported through the inner details of
-the HistFactory models shared in our HEPData~\cite{hepdata.116034}.
-
-To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
-excessive diboson background should also take some blame for this poor
-prediction in SR-Low;
-its diboson yield peaks in the signal region on $\mjj \approx m(W)$, but
-we do not expect $W\rightarrow jj$ in this process since one boson is needed
-for the leptons and the other for the $\met$ that we select.
-We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
-not similarly squashed by sample comparison variations.
-
 \begin{figure}[tp]
@@
 \end{figure}
+
+It is not clear that such a large extrapolation from the difference between
+two simulations is physically reasonable, especially since a large fraction
+of its effect is due to the statistical error on a small alternative sample.
+To avoid this questionable effect, I searched for alternative configurations
+of the background model, such as constraining the parameter to $1$-sigma.
+Committee decisions within the \atlas\ SUSY group dictated that we should keep
+it as described, reasoning along the lines that
+a) this is a real and significant deficit in SR-Low, so must appear as either
+a pull on systematic variations or as a deviation from data, and
+b) we prefer the pull on systematic variations.
+
+With this decision, post-fit results (such as the summary plot of
+Figure~\ref{fig:2ljets_summary})
+indeed show no significant deviation from data in SR-Low-a or SR-Low-b,
+and this $\zjets$ sample comparison is the main reason why.
+Internal reviewers were, however, confused as to why we mentioned a deficit in
+SR-Low that did not appear in the results plots or tables.
+Much to my ire, the paper editors therefore removed all mentions of an SR-Low
+deficit from the paper.
+Despite bring real and significant, this deficit was erased from public view,
+and to my knowledge is only reported through inner details of
+the HistFactory models shared in our HEPData archive~\cite{hepdata.116034}.
+
+To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
+excessive diboson background should also take some blame for this poor
+prediction in SR-Low;
+its diboson yields peak in the signal region on $\mjj \approx m(W)$, and
+we do not expect large $W\rightarrow jj$ contributions in this diboson process,
+since one boson is needed for the leptons and the other is needed for the
+$\met$ that we select.
+We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
+not similarly squashed by sample comparison variations.
 

### 2022-07-27 main/chapter_2ljets.tex
change to other variatiosn subsection; add ttbar and topother

@@
 will be discussed in the remainder of this section ---
+\TODO{update this}
 Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
-alternative simulated samples in Section~\ref{sec:2ljets_sample_comparisons},
-signal modelling uncertainties in Section~\ref{sec:2ljets_signal_variations}
-and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.
+some alternative samples in Section~\ref{sec:2ljets_sample_comparisons},
+various less impactful effects in
+Section~\ref{sec:2ljets_other_background_variations},
+and signal modelling uncertainties in
+Section~\ref{sec:2ljets_signal_variations}.
 
@@
 workspaces~\cite{cranmer2012histfactory, heinrich2021pyhf}.
-\emph{Thanks to Jon~Long for collabouration in preparing this HEPData archive,
+\emph{Thanks to Jon~Long for collaboration in preparing this HEPData archive,
 and to Alessia~Murrone for invaluable feedback on, and validation of, its
@@
 alternative sample that simulates its hard scatter
-with \madgraph\ (MG5\_aMC@NLO~2.2.3~\cite{Alwall:2014hca}) at leading order with
+with \madgraph\ (\amcatnlo~2.2.3~\cite{Alwall:2014hca}) at leading order with
 up to four partonic jets, plugged into \pythia~8~\cite{Sjostrand:2014zea} for
@@
 
-\TODO{Other theory variations; these are less important / interesting.
-bullet point quickly.}
-\subsubsection{Alternative \texorpdfstring{$\ttbar$}{ttbar}}
-\label{sec:2ljets_ttbar_alt}
-\TODO{}
-
+\subsection{Other background variations}
+\label{sec:2ljets_other_background_variations}
+Many variations were evaluated, but few of them significantly impact the
+results or have interesting details, so they are not covered in detail in
+this thesis.
+This section briefly lists some of the remaining variations.
+
+\paragraph{Alternative $\ttbar$}
+Several $\ttbar$ variations model alternative simulations of the
+hard scatter (central \powheg\ versus \amcatnlo),
+parton shower (central \pythia8 versus \herwig7).
+Initial state radiation variations are modelled from an alternative
+sample using $h_\textrm{damp} = 517.5\,\eV[G] \approx 3m(t)$,
+which damps real emissions and defaults to $m(t)$~\cite{CMS-PAS-TOP-16-021},
+with weights to vary an ``eigentune'' parameter named
+\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}.
+Modifications to final state radiation are evaluated by weighting to a doubled
+and halved renormalization scale for the final state only, while keeping
+its central value in simulation of initial state radiation.
+
+All of these follow standard prescriptions given by the \atlas\ physics
+modelling group~\cite{atlas_twiki_top}, and use the method described in
+Section~\ref{sec:2ljets_zjets_alt} that combines absolute deviations with
+statistical uncertainties into directed variations.
+In CR-tt, the parton shower change induces a $\pm10\%$ relative variations,
+and other effects are below $1\%$; much larger variations are measured in
+regions with less statistical precision, notably of $20\textrm{--}50\%$ changes
+in SR-OffShell and SR-Low-2.
+
+\paragraph{Alternative rare top} Variations in rare top processes, which matter for SR-$\llbb$, use
+alternative simulations of the hard scatter and parton shower.
+Only alternative $t\bar{t}Z$ simulations were available, but they form around
+$70\%$ of the rare top background, so their relative variations are evaluated
+and scaled to the total rare top yield.
+
+The central $t\bar{t}Z$ sample uses \amcatnlo+\pythia8, and it is compared
+against \pythia~2.2.1 and \amcatnlo+\herwig7 simulations, each of which
+finds a $\pm21\%$ variation in SR-$\llbb$.
+
+\paragraph{Scale variations}
+\TODO{Scale variations}
+
+\paragraph{Parton distribution variations}
+\TODO{PDF}
+
+\subsection{Signal theory variations}
+\label{sec:2ljets_signal_variations}
 \TODO{Signal theory variations}
-\subsection{Signals}
-\label{sec:2ljets_signal_variations}
-
 % TODO GMSB reweighting
@@
 
-\subsection{Theory}
 \label{sec:2ljets_theory_variations}

### 2022-07-27 main/chapter_2ljets.tex
add xs, sclae, pdf variations; add miscellanea

@@
 `SR2-int' and `SR2-high' regions with orthogonal selections.
-\emph{I am grateful to Sarah Williams for suggesting this approach.}
+\emph{Special thanks to Sarah~Williams for suggesting this approach.}
 
@@
 These variations are evaluated according to a standard recipe prescribed by
-the ATLAS SUSY group~\cite{atlas_twiki_susy_theory_comparison} that says to
+the ATLAS SUSY group~\cite{atlas_twiki_susytheoretical} that says to
 \begin{itemize}
@@
 
-This SUSY recipe contradicts the advice of the ATLAS Physics Modelling Group
+This SUSY recipe contradicts the advice of the \atlas\ Physics Modelling Group
 (PMG), who say to quote only the absolute value of the two
-estimates~\cite{atlas_twiki_pmg_theory_comparison} in for two-sample
+estimates~\cite{atlas_twiki_pmg_theory} in for two-sample
 comparisons.
@@
 ``the scale ($+$ PDF) uncertainties are too small in your
-analysis''~\cite{atlas_twiki_susy_v_jets},
+analysis''~\cite{atlas_twiki_susytheoretical},
 but I know of no official advice on how to judge what is ``too small''.
@@
 with calibrated noise to approximate reconstruction-level results.
-\emph{I am grateful to Jon~Long for preparing our initial
+\emph{Thanks to Jon~Long for preparing our initial
 SimpleAnalysis implementations and sharing software to evaluate these
@@
 
-\paragraph{Alternative rare top} Variations in rare top processes, which matter for SR-$\llbb$, use
+\paragraph{Alternative rare top}
+Variations in rare top processes, which matter in SR-$\llbb$, use
 alternative simulations of the hard scatter and parton shower.
 Only alternative $t\bar{t}Z$ simulations were available, but they form around
-$70\%$ of the rare top background, so their relative variations are evaluated
-and scaled to the total rare top yield.
-
-The central $t\bar{t}Z$ sample uses \amcatnlo+\pythia8, and it is compared
-against \pythia~2.2.1 and \amcatnlo+\herwig7 simulations, each of which
+$70\%$ of the rare top background, so their relative variations were evaluated
+and are scaled to the total rare top yield.
+
+The central $t\bar{t}Z$ sample uses \amcatnlo$+$\pythia8, and it is compared
+against \pythia~2.2.1 (for a different hard scatter, but also parton shower)
+and \amcatnlo$+$\herwig7 (for different parton shower only), each of which
 finds a $\pm21\%$ variation in SR-$\llbb$.
+In CR-tt,
+where rate top processes contribute around $5\%$ of the total background,
+the rare top variation are
+$\pm8\%$ (hard scatter) and
+$\pm13\%$ (parton shower).
+Most other regions have larger fractional errors due to our limited
+statistical precision on negligibly small yields.
+
+\paragraph{Cross-sections}
+As described in the paper~\cite{atlas2022searches},
+background samples that are not normalized to data are assigned cross-section
+uncertainties as global scale-factors.
+These are $\pm10\%$ for Higgs processes,
+$\pm32\%$ for triboson,
+$\pm6\%$ for single-top,
+and $\pm13\%$ for rare top.
+\emph{Thanks to Jon Long for researching and assigning these uncertainties.}
 
 \paragraph{Scale variations}
-\TODO{Scale variations}
+Scale variations are a standard way to assess the uncertainties in finite-order
+predictions, and by convention involve varying scale parameters up and down
+by factors of $2$.
+We follow the recommendations of the
+\atlas\ SUSY group~\cite{atlas_twiki_susytheoretical}, so implement
+these variations for the
+renormalization scale $\mu_R$ and factorization scale $\mu_F$ such that they
+are independent for each sample and coupled between all regions.
+This prescription introduces three parameters for the isolated variation of
+$\mu_R$, the isolated variation of $\mu_F$, and their coherent variations
+up and down together.
+This recipe again contradicts the PMG, who say
+that a single scale variation should be assigned by a single `envelope',
+given by the maximum difference over the various coherent and incoherent scale
+changes in both $\mu_R$ and $\mu_F$~\cite{atlas_twiki_pmg_theory}.
+
+These scale variations are implemented by weighting the central simulations
+by weights that have been provided by the generator.
+In each region, the central, up, and down variations are assigned by the
+resulting yields under each weighting, with no symmetrization or
+consideration of statistical uncertainties.
+
+Scale variations mostly move all regions in the same direction, with increased
+scales tending to reduce yields.
+Variations in $\mu_R$ are typically in the $\approx 10\textrm{--}30\%$ range,
+making them larger than the $\approx 1\textrm{--}10\%$ variations due to
+$\mu_F$.
+Given that it takes a maximum, the envelope prescription gives a single larger
+variation, but that may not mean more flexibility in the model since the SUSY
+prescription has three degrees of freedom.
+
+Since these variations are highly correlated between regions, samples that are
+normalized to control regions (diboson, $\ttbar$, $\zjets$, and $\zyjets$)
+absorb some of their effect into their unconstrained normalization factors.
+For other samples, the effect of scale variations is already included in
+the uncertainty on their total cross-section, so their scale variations are
+normalized to the sum of varied weights in very loose pre-selections,
+similarly reducing their effects.
 
 \paragraph{Parton distribution variations}
-\TODO{PDF}
+Like scale variations, uncertainties in parton distribution functions (PDFs)
+are assigned by weights stored by the generator.
+These weights, however, do not correspond to simple parameters variations,
+but different replica PDFs in a Monte Carlo ensemble
+that is intended to model the uncertainties left after fitting to data
+in the NNPDF methodology~\cite{NNPDF:2017mvq,BALL2011112,BALL2012608}.
+
+The number of PDF replicas varies between samples, but is between
+$100$ and $138$ for all except $\zyjets$, that has just one central and two
+alternatives.
+Remarkably, both SUSY and PMG prescriptions agree on how to assign variations
+from these replica weights, since both say to follow the PDF4LHC
+recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, p.
+Since PDF4LHC gives two alternative prescriptions, one using mean and variance
+and the other using quantiles, we evaluate and consider both.
+Both are negligible, making $\approx 1\%$ differences before normalization
+to control regions, but are included for completeness.
+Since there is nothing particularly special about the central PDF weight,
+this prescription allows the central scale-factor to differ from unity,
+and this is also a $\approx 1\%$ effect.
+
+\paragraph{Miscellanea}
+\begin{itemize}
+\item A $\pm30\%$ variation on $\zjets$ is assigned in CR-tt only to cover our
+uncertainty from extrapolation over $b$-tagged
+jets~\cite{atlas_twiki_susytheoretical}.
+\item Variations for the the \sherpa\ parton shower parameters
+$Q_\textrm{cut}^\textrm{CKKW}$ and
+$S_Q$ are assigned for $\zjets$.
+Unlike the alternative samples used for diboson as described in
+Section~\ref{sec:2ljets_diboson_alt}, these are assigned as functions of the
+jet kinematics in each event by a parametrized method that avoids expensive
+alternative simulations~\cite{Anders:2125718, atlas_twiki_susytheoretical}.
+\emph{Thanks to Jon~Long and Yumeng~Cao for collaboration in preparing these
+variations}.
+\end{itemize}
 

### 2022-07-27 main/chapter_2ljets.tex
autoformat and quick proofread

@@
 Despite bring real and significant, this deficit was erased from public view,
-and to my knowledge is only reported through inner details of
+and to my knowledge, is only reported through inner details of
 the HistFactory models shared in our HEPData archive~\cite{hepdata.116034}.
@@
 electroweak vertices.
-To my knowledge, this is simply a limitation from what \atlas\ has decided to
-spend its resources on, and quite reasonably since full simulation is
+This limitation is simply a result of how \atlas\ has decided to allocate
+its limited resources on, and quite reasonably so, since full simulation is
 computationally expensive.
-We did, however, have access to the cheaper alternative of set of small,
+We did, however, have access to the cheaper alternative set of small,
 truth-level samples, that were simulated with the same \sherpa\ setup but with
@@
 meaningful result can be derived from these variations.
-First our regions receive very few raw events in each variation,
+First, our regions receive very few raw events in each variation,
 for example $1$ sample in SR-$\llbb$, around a dozen in
@@
 the variation is half the difference from up and down.
-\item Otherwise (opposite sides of central), the variation is half of the
+\item Otherwise, opposite sides of central, the variation is half of the
 maximum absolute difference between central and up or down.
@@
 
-In these looser regions VR-sys-main and VR-sys-OffShell, each variation
+In the looser regions VR-sys-main and VR-sys-OffShell, each variation
 reduces to around $10\%$ and remains dominated by the statistical uncertainty.
-Although local effects could occur, this a reasonable indication that these
+Although local effects could occur, this is a reasonable indication that these
 parameters should not much affect our signal regions.
@@
 this thesis.
-This section briefly lists some of the remaining variations.
+This section briefly lists some remaining variations.
 
@@
 jets~\cite{atlas_twiki_susytheoretical}.
-\item Variations for the the \sherpa\ parton shower parameters
+\item Variations for the \sherpa\ parton shower parameters
 $Q_\textrm{cut}^\textrm{CKKW}$ and

### 2022-07-28 main/chapter_2ljets.tex
add event counts for vr-sys

@@
 
-In the looser regions VR-sys-main and VR-sys-OffShell, each variation
+In the looser regions VR-sys-main and VR-sys-OffShell,
+which receive $1784$ and $7015$ truth events respectively, each variation
 reduces to around $10\%$ and remains dominated by the statistical uncertainty.

### 2022-07-28 main/chapter_2ljets.tex
weaken too honest complaint

@@
 Numerical simplicity is valuable.
-In internal documentation, I stated that it was a motivator behind the design,
-but that statement was removed after one \atlas\ author commented:
-``maybe this paragraph is a little
-\emph{too honest}? (The `numerical simplicity' bit)''
-(emphasis added)~\cite{comment2020toohonest}.
+I wrote in a draft of our \atlas-internal documentation that numerical
+simplicity was a design goal, but am guilty of removing that statement after a
+senior reviewer suggested that it was ``too honest''.
+% ~\cite{comment2020toohonest}. % Could cite, probably should not.
 

### 2022-07-28 main/chapter_2ljets.tex
begin signal variations sections; srllbb command; various tweaks

@@
 \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
-\caption{SR-$\llbb$}
+\caption{\srllbb}
 \end{subfigure}
@@
 The high regions, defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
-SR-High-1J, and SR-$\llbb$.
+SR-High-1J, and \srllbb.
 
@@
 \\[1em]
-SR-$\llbb$
+\srllbb
 & $\geq 2$
@@
 
-Finally, SR-$\llbb$ is designed to target the GMSB model, which produces decays
+Finally, \srllbb\ is designed to target the GMSB model, which produces decays
 via $ZZ$ and $Zh$ bosons pairs.
-Since the $h\rightarrow bb$ branching fraction is large, SR-$\llbb$ finds
+Since the $h\rightarrow bb$ branching fraction is large, \srllbb\ finds
 sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
@@
 
-As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in SR-$\llbb$
+As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in \srllbb
 contain a significant proportion of rare top quark processes;
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_SRllbb.png}
-\caption{SR-$\llbb$, $\metsig$}
+\caption{\srllbb, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mbb_SRllbb.png}
-\caption{SR-$\llbb$, $\mbb$}
+\caption{\srllbb, $\mbb$}
 \end{subfigure}
@@
 
-The small yields beside SR-$\llbb$ in other event variables left $\metsig$ as
+The small yields beside \srllbb\ in other event variables left $\metsig$ as
 the only appropriate candidate to change for VR-$\llbb$.
@@
 \item DR-High is SR-High-8 (the union of SR-High-8-a and -b),
-\item DR-$\llbb$ is is SR-$\llbb$,
+\item DR-$\llbb$ is is \srllbb,
 \item DR-Int is SR-Int (the union of SR-Int-a and -b),
@@
 Evaluation of candidate $\llbb$ discovery regions.
-The region SR-$\llbb$, labelled `SRllbb' here, is chosen for
+The region \srllbb, labelled `SRllbb' here, is chosen for
 sensitivity, simplicity, and to avoid overlapping with DR-High.
@@
 of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
-For the $\twoljets$ search, we used HistFitter to build our every HistFactory
+For the $\twoljets$ search, we used \histfitter\ to build our every HistFactory
 models~\cite{Besjes_2015,baak2015histfitter}.
@@
 \\
-SR-$\llbb$ & $0.7 \pm 0.1$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
+\srllbb & $0.7 \pm 0.1$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
 \\
@@
 \\[0.2em]
-SR-$\llbb$ & $0.7 \pm 0.1$
+\srllbb & $0.7 \pm 0.1$
 & $0^{+0}_{-0}$
@@
 First, our regions receive very few raw events in each variation,
-for example $1$ sample in SR-$\llbb$, around a dozen in
+for example $1$ sample in \srllbb, around a dozen in
 each SR-High bin, and a maximum of $597$ in CR-VZ.
@@
 \paragraph{Alternative rare top}
-Variations in rare top processes, which matter in SR-$\llbb$, use
+Variations in rare top processes, which matter in \srllbb, use
 alternative simulations of the hard scatter and parton shower.
@@
 and \amcatnlo$+$\herwig7 (for different parton shower only), each of which
-finds a $\pm21\%$ variation in SR-$\llbb$.
+finds a $\pm21\%$ variation in \srllbb.
 In CR-tt,
@@
 \begin{itemize}
-\item A $\pm30\%$ variation on $\zjets$ is assigned in CR-tt only to cover our
+\item A $\pm30\%$ variation is applied to $\zjets$ in CR-tt to cover our
 uncertainty from extrapolation over $b$-tagged
@@
 
-\subsection{Signal theory variations}
+\subsection{Signal variations}
 \label{sec:2ljets_signal_variations}
-\TODO{Signal theory variations}
+Signal uncertainties are a strange case.
+Like Standard Model backgrounds, our signal models are simulated by
+approximations to their ideal QFT, and those approximations do have
+uncertainties that we can describe.
+On backgrounds, large uncertainties are conservative, since
+flexibility to fit outlying data reduces the chance of mistaken claims of
+discoveries.
+On signals, however, that effect is reversed.
+In maximum \heplikelihood\ fits, a more flexible
+signal model will always fit any observed data better, and in the limit of
+infinite flexibility, an unconstrained signal will always fit
+any data better than a constrained background, making it look better to certain
+interpretations.%
+\footnote{%
+This superiority of flexible models is not fundamental, but is a property of
+\heplikelihood\ fitting methods.
+If we want to punish uncertain predictions, then we should consider
+likelihoods (not \emph{maximum} \heplikelihood s), as described in
+Section~\ref{searches:searces_data_analysis}.
+}
+
+In an apparent notional effort to address this problem, we treat the signal
+cross-section specially.
+Rather than fitting signals cross-section within a constraint, we analyse each
+signal model three times with its cross-section fixed to central, up, and down
+variations, and report an exclusion contour for each.
+These alternative contours have been shown as the three observed limits
+in Figures~\ref{fig:2ljets_contours_c1n2} and~\ref{fig:2ljets_contours_gmsb}.
+This approach is entirely standard, and is ingrained in \histfitter\ with the
+special parameter name
+\texttt{$(\alpha\_)$SigXSec}~\cite{histfitter2014,histfittergithub}.
+
+% TODO xsec variations
+
 % TODO GMSB reweighting
 % TODO off-shell plus/minus split
-% TODO xsec variations
 
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_SRllbb.png}
-\caption{SR-$\llbb$, rare top}
+\caption{\srllbb, rare top}
 \end{subfigure}
@@
 Correlation matrix from the background-only fit to control and signal regions,
-as estimated by HistFitter.
+as estimated by \histfitter.
 Parameters with smaller levels of correlation are excluded.

### 2022-07-28 main/chapter_2ljets.tex
add xsec variations references and description

@@
 parton shower (central \pythia8 versus \herwig7).
-Initial state radiation variations are modelled from an alternative
-sample using $h_\textrm{damp} = 517.5\,\eV[G] \approx 3m(t)$,
-which damps real emissions and defaults to $m(t)$~\cite{CMS-PAS-TOP-16-021},
-with weights to vary an ``eigentune'' parameter named
-\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}.
+Initial state radiation variations are evaluated from an alternative
+sample with $h_\textrm{damp} = 517.5\,\eV[G] \approx 3m(t)$,
+which damps real emissions with weights to vary an ``eigentune'' parameter
+named \texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}..
+The central sample uses a default value $h_\textrm{damp} = 172.5\,\eV[G]$ to
+approximate $m(t)$~\cite{CMS-PAS-TOP-16-021}.
+
 Modifications to final state radiation are evaluated by weighting to a doubled
@@
 Like scale variations, uncertainties in parton distribution functions (PDFs)
-are assigned by weights stored by the generator.
+are assigned by weights stored Fuks:2012qxby the generator.
 These weights, however, do not correspond to simple parameters variations,
@@
 uncertainties that we can describe.
-On backgrounds, large uncertainties are conservative, since
+On backgrounds, however, large uncertainties are conservative, since
 flexibility to fit outlying data reduces the chance of mistaken claims of
 discoveries.
-On signals, however, that effect is reversed.
+That effect reverses for signals.
 In maximum \heplikelihood\ fits, a more flexible
@@
 likelihoods (not \emph{maximum} \heplikelihood s), as described in
-Section~\ref{searches:searces_data_analysis}.
+Section~\ref{sec:searches_searches}.
 }
@@
 In an apparent notional effort to address this problem, we treat the signal
-cross-section specially.
-Rather than fitting signals cross-section within a constraint, we analyse each
-signal model three times with its cross-section fixed to central, up, and down
-variations, and report an exclusion contour for each.
-These alternative contours have been shown as the three observed limits
+cross-section specially by not fitting it within a constraint, but fixing it
+to central, up, and down values for three repeated analyses.
+We then derive and report separate exclusion contours for each cross-section
+variation;
+such alternative contours are shown as the three observed limits
 in Figures~\ref{fig:2ljets_contours_c1n2} and~\ref{fig:2ljets_contours_gmsb}.
-This approach is entirely standard, and is ingrained in \histfitter\ with the
-special parameter name
-\texttt{$(\alpha\_)$SigXSec}~\cite{histfitter2014,histfittergithub}.
-
-% TODO xsec variations
-
-% TODO GMSB reweighting
-% TODO off-shell plus/minus split
-
+This approach is entirely standard, and is indeed ingrained into \histfitter\
+through the special parameter name
+\texttt{$(\alpha\_)$SigXSec}~\cite{histfitter2014, histfittergithub}.
+
+Standard experimental uncertainties, which relate to tagging and
+reconstruction of final state objects, are evaluated and included for signals
+exactly as they were for backgrounds.
+The remainder of this section describes signal-specific variations.
+
+\paragraph{Cross-section variations}
+\label{sec:2ljets_xsec_variations}
+We scale all signal models to cross-sections calculated with the
+\textsc{Resummino} package at NLO accuracy in QCD with NLL
+(next-to-leading logarithmic) accuracy in the gaugino pair production of our
+signal initial states~\cite{Fuks:2012qx, Fuks:2013vua}.
+These cross-sections are independent of their subsequent decays, so depend only
+on the initial state masses ---
+$m(\chargino_1, \neutralino_2)$ for C1N2, and
+$m(\neutralino_1)$ for GMSB.
+
+Uncertainties in these cross-sections are assigned from an envelope
+(using symmetrized maximum deviations)
+over scale variations~\cite{Fuks:2012qx} and parton distribution variations
+with the
+\textsc{CTEQ6.6}~\cite{Nadolsky:2008cw} and
+\textsc{MSTW2008nlo90cl}~\cite{Martin:2009iq} PDF sets.
+These uncertainties increase with parent mass from around
+$\pm4\%$ at $100\,\eV[G]$ to around
+$\pm11\%$ at $800\,\eV[G]$ for both C1N2 and GMSB models~\cite{
+atlas_twiki_xsec_c1n2,
+atlas_twiki_xsec_hino
+}.
+
+\paragraph{Theory variations}
 \label{sec:2ljets_theory_variations}

### 2022-07-28 main/chapter_2ljets.tex
expand footnote waffle
revise uncertainty footnote waffle

@@
 \heplikelihood\ fitting methods.
-If we want to punish uncertain predictions, then we should consider
-likelihoods (not \emph{maximum} \heplikelihood s), as described in
+If we want to punish models that make uncertain predictions, then we should
+consider likelihoods (not \emph{maximum} \heplikelihood s), as described in
 Section~\ref{sec:searches_searches}.
+The important distinction is between uncertain predictions from the model,
+and our uncertainty in what the model prediction is.
+Our ignorance of the a model's predictions should not lead us to excluding it.
 }

### 2022-08-02 main/chapter_2ljets.tex
expand fnp with likelihood methods and fake factor method

@@
 }
-leads us to use the `Matrix Method'~\cite{ATLAS-CONF-2014-058} to estimate
+leads us to use the `Matrix Method'~\cite{
+ATLAS-CONF-2014-058,
+D0:1999qdf
+}%
+to estimate
 backgrounds of fake and non-prompt leptons, so named because it works by
@@
 This count of $157$ differs from the $194$ actual data observed in CR-VZ.
-I do not know why, and do not have access to the source data or code.
+I do not know why, and do not have access to the source data or code to
+learn why.
 The two do overlap in event identity numbers, so there is presumably some
@@
 at rates similar to (or larger than) the standard model backgrounds.
-Loose data do have positive weights, and they would drive up the FNP
-background.
+Loose data have positive weights, so they would drive up the FNP estimate.
 
@@
 I consider those that do to be dubious.
+
+The Matrix Method's approximation that
+$\mu^\textrm{data}_{**} = n^\textrm{data}_{**}$ can be effective but is not
+exact.
+More accurately, the two are are related through a Poisson likelihood function
+that alternative FNP estimation methods can use~\cite{
+Erdmann:2021pqi,
+Gillam:2014xua
+}.
+Our criticism that the Matrix Method prematurely unblinds data is not new, and
+the fake-factor is a notable alternative that avoids this problem by estimating
+prompt contributions from simulations, not data~\cite{
+Alison2015,
+ATLAS:2017ztq
+}.
 

### 2022-08-02 main/chapter_2ljets.tex
wrap up signal theory variations

@@
 
-\paragraph{Cross-section variations}
+\subsubsection{Cross-section variations}
 \label{sec:2ljets_xsec_variations}
@@
 (using symmetrized maximum deviations)
-over scale variations~\cite{Fuks:2012qx} and parton distribution variations
+over scale variations and parton distribution variations~\cite{Fuks:2012qx}
 with the
@@
 
-\paragraph{Theory variations}
+\subsubsection{Theory variations}
 \label{sec:2ljets_theory_variations}
-\TODO{put this above alternative samples?}
-\TODO{this appears to be signal theory}
-% extrapolation
-The total theory variation is a combination of seven components, representing
-\TODO{what are they?}.
-A relative uncertainty due to each, $r$, is evaluated at each benchmark
-point $\vec y_i$, and extrapolated elsewhere as an exponentially weighted
-combination
+Another signal variation is assigned to cover sources of uncertainty from
+scale variations ($\mu_R$, $\mu_F$, and the merging scale
+$Q_\textrm{cut}^\textrm{CKKW}$)) and modelling of radiations and parton
+showers (`eigentune' parameters
+\texttt{Var1},
+\texttt{Var2},
+\texttt{Var3a},
+\texttt{Var3b},
+\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}).
+Following the \atlas\ SUSY group
+prescription~\cite{atlas_twiki_susytheoretical},
+we consider only a coherent scaling of the scales $\mu_R$ and $\mu_F$ together,
+so this theory variation includes a total of seven components.
+
+Scale variations have appeared here twice:
+first for cross-section variations and
+second for these theory variations.
+They therefore contribute both in each fit model and in the varied
+cross-sections for varied exclusion contours.
+This is questionable, but we shall see here that, thankfully, this scale
+double-counting makes very little difference.
+
+\paragraph{Extrapolation}
+To evaluate each variation requires a new simulation with different parameters,
+so to reduce computational costs we use truth-level only
+(through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
+variations at only a handful of benchmark signal models.
+From these benchmark signal models, we evaluate a relative uncertainty due
+due to each component, and extrapolate those variations from these benchmark
+examples to others in the parameter plane.
+
+Extrapolation is performed as follows.
+Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
+for a given source of uncertainty $j$ its relative variation has been evaluated
+as $r_{ij}$.
+To extrapolate these variations across each parameter plane to another signal
+model at $\vec x$, we use an exponentially weighted combination
 \begin{equation}
 \label{eqn:2ljets_signal_extrap}
-r(\vec x) = \frac{
-\sum_i r_i e^{-d(\vec x, \vec y_i)}
+r_j(\vec x) = \frac{
+\sum_i r_{ij} e^{-d(\vec x, \vec y_i)}
 }{
@@
 in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
-physical understanding into the extrapolation.
+physical understanding into the extrapolation,
+and the sum is over all benchmark points.
 
@@
 Figure~\ref{fig:2ljets_signal_sys_extrap}.
-Variations tend to be small, around $\approx 20\%$, where and only where
-yields are large enough for good statistical precision.
+This figure illustrates the common theme that our theory variations are small
+if and only if signal yields are large enough for good statical precision.
 
@@
 \end{figure}
+
+\paragraph{Evaluation}
+Statistical precision has limited the utility of most variations described in
+this chapter.
+These are no different.
+To reduce statistical noise here, I generated truth samples $10$~times as large
+as our central, reconstructed samples.
+These sample sizes vary roughly with cross-section from $300\,000$ to
+$2.8$~million events per sample, and due to the $\sqrt{n}$ scaling of
+Monte Carlo errors, meant approximately $\sqrt{10}\approx 3.2$ times reduced
+noise.
+However, this remains insufficient since each resulting variation is a
+sum in quadrature of variations in $7$ pairs of alternative samples
+whose combination will inflate the uncertainty by roughly a factor of
+$\sqrt{7}$.
+Where signal models' yields has significant in a signal region,
+each variation tends to be around $\approx 20\%$ relative error,
+and are compatible with the central samples within statistical uncertainties.
+
+Yet another unique sample-comparison prescription now assigns variations as
+half of the absolute difference between up and down yields, irrespective of the
+central yield.
+This differs from the methods previously described in
+Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_diboson_alt},
+and is used for consistency with other $\twoljets$ work.
+
+We evaluate and extrapolate each source of uncertainty individually,
+and combine all sources by summation in quadrature in each region to form a
+single total theory variation~\cite{atlas_twiki_susytheoretical}.
+This sum loses all sense of direction, so we choose to give up and down
+variations the same $+$ and $-$ signs in all bins.
+Since the theory variation parameter is coupled across all bins, it therefore
+implements a form of non-uniform normalization factor.
+
+Why the double counting of scale variations does not matter is now apparent:
+this variation is dominated by statistical noise, and the inclusion of one
+extra source for the scale variations simply scales that noise up by
+a factor of $\approx \sqrt{7/6}$.
+
+\emph{%
+Special thanks to Sara~Alderweireldt for sharing \textsc{SignalSystTools},
+whose excellent scripts were adapted to simulate these signal
+samples, and thanks to Jon~Long for contributing to this study.
+}
 

### 2022-08-02 main/chapter_2ljets.tex
write validation; add intro sentence for unblinding

@@
 \subsection{Leptons}
+\label{sec:2ljets_events_leptons}
 In the context of $\twoljets$, `leptons' means `electron-or-muons';
@@
 through the special parameter name
-\texttt{$(\alpha\_)$SigXSec}~\cite{histfitter2014, histfittergithub}.
+\texttt{SigXSec}~\cite{histfitter2014, histfittergithub}.
 
@@
 \label{sec:2ljets_validation}
-Figure~\ref{fig:2ljets_low_crz_extensions}
-Figure~\ref{fig:2ljets_splits}
+Many additional validation studies supported the $\twoljets$ search prior to
+its publication.
+Many of these, comprising hundreds of histograms in various regions or looser
+selections in various event variables, are available to \atlas\ members in
+our internal documentation~\cite{twoljets2018int}.
+This section highlights some interesting results from these validation studies.
+
+\paragraph{CR-Z-met}
+Concerns about our modelling of $\zjets$ and diboson backgrounds have been
+raised due to our deficits in SR-Low and SR-High, which compound with the
+anticipated difficulty in modelling the $\met$ tails of $\zjets$.
+To further check these backgrounds, Figure~\ref{fig:2ljets_low_crz_extensions}
+displays histograms in a `CR-Z-met', which is modifies CR-Z
+(from Table~\ref{tab:2ljets_low}) by removing its upper bound on $\metsig$.
+The arrow in Figure~\ref{fig:2ljets_low_crz_extensions_metsig} indicates this
+upper bound at $\metsig = 9$, and shows excellent agreement with data across
+the entire distribution, both where $\zjets$ dominates and where diboson
+dominates.
+The corresponding $\met$ distribution that is shown in
+Figure~\ref{fig:2ljets_low_crz_extensions_met}
+similarly shows excellent agreement.
+
+Like our other plots intended for publication, these figures use the full fit
+model with all variations for statistical and systematic uncertainties,
+and are fitted to data in control and signal regions, but the signal regions
+have little effect since the pulled variations are normalized to this control
+region.
+To avoid the work of evaluating variations in CR-Z-met (and all of its bins),
+its variations are linearly extrapolated from CR-Z.
+
+Good modelling in CR-Z-met is a positive sign for our simulations prior to
+their modification in the fit.
+Noting that CR-Z excludes $m(W)$, however, this does not conflict with the idea
+of poor modelling in the signal regions.
 
@@
 \caption{}
+\label{fig:2ljets_low_crz_extensions_metsig}
 \end{subfigure}
@@
 \caption{}
+\label{fig:2ljets_low_crz_extensions_met}
 \end{subfigure}
@@
 Systematic uncertainties are extrapolated from CR-Z to the extended tail.
-The tail overlaps with other regions, but shows good agreement with data.
+The tail overlaps with other regions and shows good agreement with data.
 Sub-figure (b) is reproduced from auxiliary material to the
 paper~\cite{atlas2022searches}.
-Each last bin includes overflow.
+Each rightmost bin includes overflow.
 }
@@
 
+\paragraph{Sub-processes}
+\TODO{sort out top other label}
+Our high-level sample labels such as $\ttbar$, $\zjets$, diboson and
+rare-top (`Top other') each include many sub-samples that are physically
+distinct and non-interfering.
+Highlights from a study of the contributions from these sub-samples are
+presented in Figure~\ref{fig:2ljets_splits}.
+
+Although it might be expected that our $Z$-like two-lepton selection would
+favour the $ZZ$-like diboson sub-sample, with the other $Z$ decaying to
+neutrinos,
+Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the three-lepton
+$WZ$-like process has almost equal yields.
+All of those $WZ$ events therefore have a missing third lepton, that could be
+softer than the few $\eV[G]$ needed for our baseline selection
+(see Section~\ref{sec:2ljets_events_leptons}),
+or outside of our angular acceptance.
+(Overlap removal cannot explain this since we count baseline leptons before
+it.)
+Although we would hope that such failed identifications would be rare,
+the effect is enhanced by the fact that $WZ$ has a larger cross-section than
+$ZZ$ by around a factor of three~\cite{Campbell:2011bn}, that is enhanced
+by the sum over $W^+Z$ and $W^-Z$ processes.
+
+Our \sherpa\ diboson samples include the interference of electroweak processes
+for each fixed final state, so for example the separation of $ZZ$ and $WW$ is
+not exact since they interfere.
+These histograms are therefore split on charged lepton number in their
+final states.
+
+Some of our $\zjets$ samples are filtered by hadron flavour
+(requiring at least one selected hadron with $|\eta| < 4$), and
+Figure~\ref{fig:2ljets_splits_zjets} shows CR-Z with its $\zjets$ yields
+split by these filters.
+Heavy flavour samples are enhanced, with the filtered yields comprising
+$22\%$ bottom and $30\%$ charm flavours despite the $\nbtag = 0$ veto in CR-Z;
+this is sensible, since we also require $\met$, and extra non-prompt $\met$ is
+generated  in the weak decays of of heavy-flavoured hadrons.
+
+The remaining parts of Figure~\ref{fig:2ljets_splits_diboson} show that
+of the rare-top sample, $\ttbar Z$ is the largest contribution,
+at $73\%$ in \srllbb, but that contributions from other rare processes
+such as $tW\kern-0.4exZ$, $t\ell\ell g$ and $\ttbar W$ are significant.
+Here, the plain `Top other' label refers to all other rare-top contributions.
+
 \begin{figure}[tp]
@@
 \caption{CR-VZ, diboson}
+\label{fig:2ljets_splits_diboson}
 \end{subfigure}
@@
 \caption{CR-Z, $\zjets$}
+\label{fig:2ljets_splits_zjets}
 \end{subfigure}
@@
 \caption{CR-tt, rare top}
+\label{fig:2ljets_splits_topother1}
 \end{subfigure}
@@
 \caption{\srllbb, rare top}
+\label{fig:2ljets_splits_topother2}
 \end{subfigure}
@@
 \label{sec:2ljets_unblinding}
+This section is not a discussion of our unblinded results.
+It is a discussion of our unblinding procedures,
+their extent, their implementation, and their effect on the results.
+
 Figure~\ref{fig:2ljets_prepre}

### 2022-08-02 main/chapter_2ljets.tex
clarify specialized topother
kern all WZ

@@
 The C1N2 and GMSB models behave similarly since both feature decays via massive
-bosons ($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay
+bosons ($W\kern-0.4exZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay
 trees to pairs of invisible supersymmetric particles.
@@
 Some figures label CR-VZ as `CR-ZZ'; the region was renamed late in this
-analysis project when it was show to contain substantial fractions of $WZ$
+analysis project when it was show to contain substantial fractions of $W\kern-0.4exZ$
 as well as $ZZ$, as illustrated in Figure~\ref{fig:2ljets_splits} of
@@
 compressed scenarios~\cite{atlas_rjr_3l_SUSY_2019_09}.
-These searches consider the same C1N2 $WZ$ signal models as us, but target the
+These searches consider the same C1N2 $W\kern-0.4exZ$ signal models as us, but target the
 case where the $W$ decays leptonically $W\rightarrow\ell\nu$.
@@
 Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the three-lepton
-$WZ$-like process has almost equal yields.
-All of those $WZ$ events therefore have a missing third lepton, that could be
+$W\kern-0.4exZ$-like process has almost equal yields.
+All of those $W\kern-0.4exZ$ events therefore have a missing third lepton, that could be
 softer than the few $\eV[G]$ needed for our baseline selection
@@
 Although we would hope that such failed identifications would be rare,
-the effect is enhanced by the fact that $WZ$ has a larger cross-section than
+the effect is enhanced by the fact that $W\kern-0.4exZ$ has a larger cross-section than
 $ZZ$ by around a factor of three~\cite{Campbell:2011bn}, that is enhanced
-by the sum over $W^+Z$ and $W^-Z$ processes.
+by the sum over $W^+\kern-0.4exZ$ and $W^-\kern-0.4exZ$ processes.
 
@@
 such as $tW\kern-0.4exZ$, $t\ell\ell g$ and $\ttbar W$ are significant.
-Here, the plain `Top other' label refers to all other rare-top contributions.
+Here, the plain `Top other' label refers to all other rare-top contributions
+not included in the bracketed specializations.
 

### 2022-08-02 main/chapter_2ljets.tex
switch to commands for control region names

@@
 \\[1em]
-CR-VZ
+\crvz
 & $\geq 2$
@@
 \\[0.5em]
-CR-tt
+\crtt
 & $\geq 2$
@@
 
-The control region CR-VZ is defined on the $\mjj$ side-band to
+The control region \crvz\ occupies the $\mjj$ side-band to
 SR-Int and VR-Int.
 As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
-intermediate space give CR-VZ a very pure diboson background, which is very
+intermediate space give \crvz\ a very pure diboson background, which is very
 valuable or constraining the modelling of High regions.
 Also included in Table~\ref{tab:2ljets_int} and
-Figure~\ref{fig:2ljets_int_cr_region} is CR-tt, which uses various other
+Figure~\ref{fig:2ljets_int_cr_region} is \crtt, which uses various other
 selections of $b$-tagged data to constrain $\ttbar$ backgrounds.
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZZ.png}
-\caption{CR-VZ, $\metsig$}
+\caption{\crvz, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZZ.png}
-\caption{CR-VZ, $\mjj$}
+\caption{\crvz, $\mjj$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
-\caption{CR-tt, $\metsig$}
+\caption{\crtt, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRtt.png}
-\caption{CR-tt, $\mjj$}
+\caption{\crtt, $\mjj$}
 \end{subfigure}
@@
 ]{%
-Pre-fit distributions in the control regions CR-tt and CR-VZ.
+Pre-fit distributions in the control regions \crtt\ and \crvz.
 Errors are statistical.
@@
 
-Some figures label CR-VZ as `CR-ZZ'; the region was renamed late in this
+Some figures label \crvz\ as `\crz Z'; the region was renamed late in this
 analysis project when it was show to contain substantial fractions of $W\kern-0.4exZ$
@@
 \\[1em]
-CR-Z
+\crz
 & $6\textrm{--}9$
@@
 % CR
-The $\zjets$ control region CR-Z is displayed in
+The $\zjets$ control region \crz\ is displayed in
 Figure~\ref{fig:2ljets_low_cr_region}.
@@
 Standardizing cuts to remove high-$\mjj$ tails (such as requiring
-$\mjj < 500\,\eV[G]$) could have made control and validation regions like CR-Z
+$\mjj < 500\,\eV[G]$) could have made control and validation regions like \crz
 kinematically closer to the signal regions they are supposed to target,
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZ.png}
-\caption{CR-Z, $\metsig$}
+\caption{\crz, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZ.png}
-\caption{CR-Z, $\mjj$}
+\caption{\crz, $\mjj$}
 \end{subfigure}
@@
 ]{%
-Pre-fit distributions in the control region CR-Z.
+Pre-fit distributions in the control region \crz.
 Errors are statistical.
@@
 
-The gap at in CR-Z at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
+The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
 Their union therefore has a continuous $\mjj$ distribution, which is displayed
@@
 ]{%
-The union of CR-Z and SR-Low, which exactly fills its gap at
+The union of \crz\ and SR-Low, which exactly fills its gap at
 $\mjj \in 60\textrm{--}110\,\eV[G]$.
@@
 Figure~\ref{fig:2ljets_offshell_region}, and the off-shell control region
-CR-DY is displayed in Figure~\ref{fig:2ljets_offshell_cr_region}.
+\crdy\ is displayed in Figure~\ref{fig:2ljets_offshell_cr_region}.
 
@@
 \\[1em]
-CR-DY
+\crdy
 & $\uwave{6\textrm{--}9}$
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRDY.pdf}
-\caption{CR-DY, $\metsig$}
+\caption{\crdy, $\metsig$}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_def_mll_CRDY.png}
-\caption{CR-DY, $\mll$}
-\end{subfigure}
-\caption{%
-Pre-fit distributions in the control region CR-DY.
+\caption{\crdy, $\mll$}
+\end{subfigure}
+\caption[
+Pre-fit distributions in the control region CR-DY
+]{%
+Pre-fit distributions in the control region \crdy.
 Errors are statistical.
@@
 
-The control region CR-DY (named after Drell and Yan) is defined to normalize
+The control region \crdy\ (named after Drell and Yan) is defined to normalize
 off-shell $\zyjets$ near to SR-OffShell.
@@
 In fact, real data tend to be assigned negative Matrix Method weights.
-For example, CR-VZ receives $157$ data in its Matrix Method sample,%
+For example, \crvz\ receives $157$ data in its Matrix Method sample,%
 \footnote{%
-This count of $157$ differs from the $194$ actual data observed in CR-VZ.
+This count of $157$ differs from the $194$ actual data observed in \crvz.
 I do not know why, and do not have access to the source data or code to
@@
 and the remainder have one loose lepton (zero loose-loose lepton events
-are observed in CR-VZ).
+are observed in \crvz).
 The tight-tight events have an average weight of $-0.0227$, and the others
@@
 \\[0.5em]
-CR-Z & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
+\crz\ & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
 \\
@@
 \\[0.5em]
-CR-Z & $143.6 \pm 4.6$
+\crz\ & $143.6 \pm 4.6$
 & $0.1^{+0.2}_{-0.1}$
@@
 Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
-in the union of CR-Z and SR-Low, which form a continuous region in $\mjj$.
+in the union of \crz\ and SR-Low, which form a continuous region in $\mjj$.
 Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
@@
 \\[0.5em]
-CR-DY
+\crdy
 & $59.64$
@@
 \\[0.2em]
-\underline{CR-Z}
+\underline{\crz}
 & $101.95$
@@
 \\[0.2em]
-CR-tt
+\crtt
 & $38.07$
@@
 \\[0.2em]
-CR-VZ
+\crvz
 & $5.03$
@@
 ]{%
-The union of CR-Z and SR-Low with an alternative $\zjets$ simulation using
+The union of \crz\ and SR-Low with an alternative $\zjets$ simulation using
 MadGraph.
@@
 Figure~\ref{fig:2ljets_low_crz_pre_post_post}~\cite{hepdata.116034.v1/t101},
-which appears to show good agreement in SR-Low and CR-Z, a relatively
+which appears to show good agreement in SR-Low and \crz, a relatively
 small error bar in SR-Low of around $\pm10$ events per $100\,\eV[G]$,
@@
 \begin{itemize}
-\item The $\zjets$ sample is normalized to CR-Z with an unconstrained
-normalization factor. This means that variations to the $\zjets$ sample in CR-Z
+\item The $\zjets$ sample is normalized to \crz\ with an unconstrained
+normalization factor. This means that variations to the $\zjets$ sample in \crz
 are absorbed by scaling all of its modifications in other regions.
 \item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
-in CR-Z match to negative variations in SR-Low bins; the two yields are
+in \crz\ match to negative variations in SR-Low bins; the two yields are
 anticorrelated.
-With normalization to CR-Z, this anticorrelation increases the effective
+With normalization to \crz, this anticorrelation increases the effective
 variation in SR-Low to $\pm40\%$.
@@
 ]{%
-Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
+Distribution of $\mjj$ in \crz\ and SR-Low with all systematic variations
 (a) fitted to control regions, and (b) fitted to control and signal regions.
@@
 for example $1$ sample in \srllbb, around a dozen in
-each SR-High bin, and a maximum of $597$ in CR-VZ.
+each SR-High bin, and a maximum of $597$ in \crvz.
 Second, those samples carry weights in almost-equal mixtures of $+1$ and $-1$,
@@
 
-This assignment varies CR-VZ and on-shell other regions identically,
-and since the diboson background is normalized to CR-VZ this means that
+This assignment varies \crvz\ and on-shell other regions identically,
+and since the diboson background is normalized to \crvz\ this means that
 these variations have no effect in on-shell regions, and cannot move to
@@
 statistical uncertainties into directed variations.
-In CR-tt, the parton shower change induces a $\pm10\%$ relative variations,
+In \crtt, the parton shower change induces a $\pm10\%$ relative variations,
 and other effects are below $1\%$; much larger variations are measured in
@@
 finds a $\pm21\%$ variation in \srllbb.
-In CR-tt,
+In \crtt,
 where rate top processes contribute around $5\%$ of the total background,
@@
 \begin{itemize}
-\item A $\pm30\%$ variation is applied to $\zjets$ in CR-tt to cover our
+\item A $\pm30\%$ variation is applied to $\zjets$ in \crtt\ to cover our
 uncertainty from extrapolation over $b$-tagged
@@
 
-\paragraph{CR-Z-met}
+\paragraph{\crz-met}
 Concerns about our modelling of $\zjets$ and diboson backgrounds have been
@@
 To further check these backgrounds, Figure~\ref{fig:2ljets_low_crz_extensions}
-displays histograms in a `CR-Z-met', which is modifies CR-Z
+displays histograms in a `\crz-met', which is modifies \crz
 (from Table~\ref{tab:2ljets_low}) by removing its upper bound on $\metsig$.
@@
 region.
-To avoid the work of evaluating variations in CR-Z-met (and all of its bins),
-its variations are linearly extrapolated from CR-Z.
-
-Good modelling in CR-Z-met is a positive sign for our simulations prior to
+To avoid the work of evaluating variations in \crz-met (and all of its bins),
+its variations are linearly extrapolated from \crz.
+
+Good modelling in \crz-met is a positive sign for our simulations prior to
 their modification in the fit.
-Noting that CR-Z excludes $m(W)$, however, this does not conflict with the idea
+Noting that \crz\ excludes $m(W)$, however, this does not conflict with the idea
 of poor modelling in the signal regions.
@@
 ]{%
-Extension from CR-Z into higher (a) $\metsig$ and (b) $\met$.
-The arrow indicates the CR-Z  upper bound of $\metsig$; it has no hard
+Extension from \crz\ into higher (a) $\metsig$ and (b) $\met$.
+The arrow indicates the \crz\  upper bound of $\metsig$; it has no hard
 upper bound of $\met$.
-Systematic uncertainties are extrapolated from CR-Z to the extended tail.
+Systematic uncertainties are extrapolated from \crz\ to the extended tail.
 The tail overlaps with other regions and shows good agreement with data.
@@
 (requiring at least one selected hadron with $|\eta| < 4$), and
-Figure~\ref{fig:2ljets_splits_zjets} shows CR-Z with its $\zjets$ yields
+Figure~\ref{fig:2ljets_splits_zjets} shows \crz\ with its $\zjets$ yields
 split by these filters.
 Heavy flavour samples are enhanced, with the filtered yields comprising
-$22\%$ bottom and $30\%$ charm flavours despite the $\nbtag = 0$ veto in CR-Z;
+$22\%$ bottom and $30\%$ charm flavours despite the $\nbtag = 0$ veto in \crz;
 this is sensible, since we also require $\met$, and extra non-prompt $\met$ is
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZZ.png}
-\caption{CR-VZ, diboson}
+\caption{\crvz, diboson}
 \label{fig:2ljets_splits_diboson}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZ.png}
-\caption{CR-Z, $\zjets$}
+\caption{\crz, $\zjets$}
 \label{fig:2ljets_splits_zjets}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRtt.png}
-\caption{CR-tt, rare top}
+\caption{\crtt, rare top}
 \label{fig:2ljets_splits_topother1}
@@
 Target samples broken down by simulated sample identity codes.
-The large three-lepton component in CR-VZ motivated its renaming from CR-ZZ.
+The large three-lepton component in \crvz\ motivated its renaming from \crz Z.
 The `no filter' selection comprises $\zjets$ samples with unconstrained jet

### 2022-08-02 main/chapter_2ljets.tex
rare top to rare-top

@@
 `Other' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and other
-rare top processes.%
+\topother\ processes.%
 }
@@
 As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in \srllbb
-contain a significant proportion of rare top quark processes;
+contain a significant proportion of \topother\ quark processes;
 studies presented in Section~\ref{sec:2ljets_validation} show that these are
@@
 
-\paragraph{Alternative rare top}
-Variations in rare top processes, which matter in \srllbb, use
+\paragraph{Alternative \topother}
+Variations in \topother\ processes, which matter in \srllbb, use
 alternative simulations of the hard scatter and parton shower.
 Only alternative $t\bar{t}Z$ simulations were available, but they form around
-$70\%$ of the rare top background, so their relative variations were evaluated
-and are scaled to the total rare top yield.
+$70\%$ of the \topother\ background, so their relative variations were evaluated
+and are scaled to the total \topother\ yield.
 
@@
 where rate top processes contribute around $5\%$ of the total background,
-the rare top variation are
+the \topother\ variation are
 $\pm8\%$ (hard scatter) and
@@
 $\pm6\%$ for single-top,
-and $\pm13\%$ for rare top.
+and $\pm13\%$ for \topother.
 \emph{Thanks to Jon Long for researching and assigning these uncertainties.}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRtt.png}
-\caption{\crtt, rare top}
+\caption{\crtt, \topother}
 \label{fig:2ljets_splits_topother1}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_SRllbb.png}
-\caption{\srllbb, rare top}
+\caption{\srllbb, \topother}
 \label{fig:2ljets_splits_topother2}
@@
 \caption[
-Subcomponents of diboson, $\zjets$ and rare top processes.
-]{%
-Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) rare top processes.
+Subcomponents of diboson, $\zjets$ and rare-top processes.
+]{%
+Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) \topother\ processes.
 Target samples broken down by simulated sample identity codes.

### 2022-08-02 main/chapter_2ljets.tex
remove top other todo

@@
 \paragraph{Sub-processes}
-\TODO{sort out top other label}
 Our high-level sample labels such as $\ttbar$, $\zjets$, diboson and

### 2022-08-02 main/chapter_2ljets.tex
auto check

@@
 and our uncertainty in what the model prediction is.
-Our ignorance of the a model's predictions should not lead us to excluding it.
+Our ignorance of a model's predictions should not lead us to excluding it.
 }
@@
 Variations are evaluated for signals at the marked points and extrapolated to
-the each parameter plane according to Equation~\ref{eqn:2ljets_signal_extrap}.
+each parameter plane according to Equation~\ref{eqn:2ljets_signal_extrap}.
 To account for its drastically different kinematics, the C1N2 off-shell space
@@
 (see Section~\ref{sec:2ljets_events_leptons}),
-or outside of our angular acceptance.
-(Overlap removal cannot explain this since we count baseline leptons before
+or outside our angular acceptance.
+(Overlap removal cannot explain this, since we count baseline leptons before
 it.)
@@
 this is sensible, since we also require $\met$, and extra non-prompt $\met$ is
-generated  in the weak decays of of heavy-flavoured hadrons.
+generated  in the weak decays of heavy-flavoured hadrons.
 

### 2022-08-02 main/chapter_2ljets.tex
diboson to \diboson; clarify overlap removal

@@
 What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
-selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
+selections is almost pure \diboson\ -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
 and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
 is not reconstructed.
-\TODO{plot of diboson decomposition here?}
+\TODO{plot of \diboson\ decomposition here?}
 These effects are demonstrated from simulations in
@@
 At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
-quite pure diboson backgrounds and greater significance measures.
+quite pure \diboson\ backgrounds and greater significance measures.
 Signal models with smaller mass-splittings tend to have the bulk of their
@@
 The highest $\metsig$ selections, of around $\metsig > 18$, contain almost pure
-diboson backgrounds in relatively small quantities, along with the bulk masses
+\diboson\ backgrounds in relatively small quantities, along with the bulk masses
 of signal samples at the heaviest resonance masses.
@@
 SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$.
-The pure diboson background has very similar shapes to the signals.
+The pure \diboson\ background has very similar shapes to the signals.
 Tighter cuts would therefore reduce background and signal yields comparably,
@@
 
-One major difference between signals and diboson backgrounds is that signal
+One major difference between signals and \diboson\ backgrounds is that signal
 jet pairs mostly arise from boosted $W\rightarrow jj$ decays, which more often
@@
 To suppress the increased backgrounds, the intermediate $\mll$ window is
-tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and diboson
+tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and \diboson
 processes.
@@
 As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
-intermediate space give \crvz\ a very pure diboson background, which is very
+intermediate space give \crvz\ a very pure \diboson\ background, which is very
 valuable or constraining the modelling of High regions.
@@
 Excluding $\metsig \in 9\textrm{--}12$ in SR-Low-2 is also not crucial.
-This space contains about $2.3$ background expectation (mostly diboson) and
+This space contains about $2.3$ background expectation (mostly \diboson) and
 $\approx 1$ from some benchmark signals, so its inclusion as a single combined
@@
 This plot contradicts that; most of the $\zjets$ distribution is fine;
-the strange effect happens where diboson becomes large near $m(W)$.
+the strange effect happens where \diboson\ becomes large near $m(W)$.
 
@@
 has no signal regions to avoid in different-flavour data.
-No nearby selection was found to form a pure diboson control region.
+No nearby selection was found to form a pure \diboson\ control region.
 
@@
 
-A complicated likelihood model was constructed which normalized diboson,
+A complicated likelihood model was constructed which normalized \diboson,
 $\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,
@@
 different generators, uncertainties in tracking, and statistical uncertainties.
-Once released, JMS variations were evaluated and used for the diboson and
+Once released, JMS variations were evaluated and used for the \diboson\ and
 $\ttbar$ contributions, which are the primary backgrounds in both 1J regions.
@@
 algorithms, and we implement this for our major backgrounds $\zjets$, $\ttbar$,
-and, to some extent diboson.
+and, to some extent \diboson.
 Each generator (simulation algorithm), or component thereof, uses different
@@
 Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
-the signal region, with diboson yields peaking near to $m(W)$ above a smooth
+the signal region, with \diboson\ yields peaking near to $m(W)$ above a smooth
 varying $\zjets$ background.
@@
 The arrows point in to the signal region, which shows a large peak in the
-diboson background near $m(W)$.
+\diboson\ background near $m(W)$.
 Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
@@
 To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
-excessive diboson background should also take some blame for this poor
+excessive \diboson\ background should also take some blame for this poor
 prediction in SR-Low;
-its diboson yields peak in the signal region on $\mjj \approx m(W)$, and
-we do not expect large $W\rightarrow jj$ contributions in this diboson process,
+its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
+we do not expect large $W\rightarrow jj$ contributions in this \diboson\ process,
 since one boson is needed for the leptons and the other is needed for the
 $\met$ that we select.
-We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
+We shall see in Section~\ref{sec:2ljets_\diboson_alt} why \diboson\ yields are
 not similarly squashed by sample comparison variations.
 
-\subsubsection{Alternative diboson}
-\label{sec:2ljets_diboson_alt}
-No alternative samples were available for comparison with our central diboson
+\subsubsection{Alternative \diboson}
+\label{sec:2ljets_\diboson_alt}
+No alternative samples were available for comparison with our central \diboson
 samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
@@
 software package for
-our central diboson samples and these varied alternatives.
+our central \diboson\ samples and these varied alternatives.
 SimpleAnalysis histograms truth-level simulations and optionally smears them
@@
 This assignment varies \crvz\ and on-shell other regions identically,
-and since the diboson background is normalized to \crvz\ this means that
+and since the \diboson\ background is normalized to \crvz\ this means that
 these variations have no effect in on-shell regions, and cannot move to
@@
 
-Had we considered complete alternative diboson samples, we would certainly
+Had we considered complete alternative \diboson\ samples, we would certainly
 have had some meaningful variations to apply in those regions, but I do not
@@
 This lack of a real alternative sample is particularly disappointing since
-diboson is our most important background.
+\diboson\ is our most important background.
 In hindsight, it may have been sensible to push for more resources to be
@@
 Since these variations are highly correlated between regions, samples that are
-normalized to control regions (diboson, $\ttbar$, $\zjets$, and $\zyjets$)
+normalized to control regions (\diboson, $\ttbar$, $\zjets$, and $\zyjets$)
 absorb some of their effect into their unconstrained normalization factors.
@@
 $S_Q$ are assigned for $\zjets$.
-Unlike the alternative samples used for diboson as described in
-Section~\ref{sec:2ljets_diboson_alt}, these are assigned as functions of the
+Unlike the alternative samples used for \diboson\ as described in
+Section~\ref{sec:2ljets_\diboson_alt}, these are assigned as functions of the
 jet kinematics in each event by a parametrized method that avoids expensive
@@
 This differs from the methods previously described in
-Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_diboson_alt},
+Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_\diboson_alt},
 and is used for consistency with other $\twoljets$ work.
@@
 \paragraph{\crz-met}
-Concerns about our modelling of $\zjets$ and diboson backgrounds have been
+Concerns about our modelling of $\zjets$ and \diboson\ backgrounds have been
 raised due to our deficits in SR-Low and SR-High, which compound with the
@@
 upper bound at $\metsig = 9$, and shows excellent agreement with data across
-the entire distribution, both where $\zjets$ dominates and where diboson
+the entire distribution, both where $\zjets$ dominates and where \diboson
 dominates.
@@
 \paragraph{Sub-processes}
-Our high-level sample labels such as $\ttbar$, $\zjets$, diboson and
+Our high-level sample labels such as $\ttbar$, $\zjets$, \diboson\ and
 rare-top (`Top other') each include many sub-samples that are physically
@@
 Although it might be expected that our $Z$-like two-lepton selection would
-favour the $ZZ$-like diboson sub-sample, with the other $Z$ decaying to
+favour the $ZZ$-like \diboson\ sub-sample, with the other $Z$ decaying to
 neutrinos,
-Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the three-lepton
+Figure~\ref{fig:2ljets_splits_\diboson} shows that in fact the three-lepton
 $W\kern-0.4exZ$-like process has almost equal yields.
@@
 (Overlap removal cannot explain this, since we count baseline leptons before
-it.)
+applying it.)
 Although we would hope that such failed identifications would be rare,
@@
 
-Our \sherpa\ diboson samples include the interference of electroweak processes
+Our \sherpa\ \diboson\ samples include the interference of electroweak processes
 for each fixed final state, so for example the separation of $ZZ$ and $WW$ is
@@
 
-The remaining parts of Figure~\ref{fig:2ljets_splits_diboson} show that
+The remaining parts of Figure~\ref{fig:2ljets_splits_\diboson} show that
 of the rare-top sample, $\ttbar Z$ is the largest contribution,
@@
 \includegraphics[width=\textwidth]{figures/2ljets_splits_met_Sign_CRZZ.png}
-\caption{\crvz, diboson}
-\label{fig:2ljets_splits_diboson}
+\caption{\crvz, \diboson}
+\label{fig:2ljets_splits_\diboson}
 \end{subfigure}
@@
 \caption[
-Subcomponents of diboson, $\zjets$ and rare-top processes.
-]{%
-Subcomponents of (a) diboson, (b) $\zjets$ and (c, d) \topother\ processes.
+Subcomponents of \diboson, $\zjets$ and rare-top processes.
+]{%
+Subcomponents of (a) \diboson, (b) $\zjets$ and (c, d) \topother\ processes.
 Target samples broken down by simulated sample identity codes.

### 2022-08-02 main/chapter_2ljets.tex
meant to means

@@
 $2.8$~million events per sample, and due to the $\sqrt{n}$ scaling of
-Monte Carlo errors, meant approximately $\sqrt{10}\approx 3.2$ times reduced
-noise.
+Monte Carlo errors, means approximately $\sqrt{10}$ times reduced
+standard deviation.
 However, this remains insufficient since each resulting variation is a

### 2022-08-03 main/chapter_2ljets.tex
add unblinding pros

@@
 
+Blindness to data is useful, but not necessary.
+All that is necessary is that our predictions are independent of the data
+they are supposed to predict.
+Probabilistically, that independence makes our Poisson likelihood assignments
+justifiable, and scientifically it allows our modelling to be useful if it
+makes accurate predictions.
+If one knows what the data are, then it takes utter honesty ---
+Feynman's ``leaning over backwards''~\cite{feynman1974cargo} ---
+to assign predictions as if the data were unknown.
+Blindness is a useful and wilful ignorance that automatically implements this
+honesty without this backbreaking work.
+
+% pro
+Naturally, \atlas\ has an unblinding procedure that requires committee
+review and approval based on internal documentation.
+In the \atlas\ SUSY world, searches are nominally performed blind to signal
+regions.
+Control and validation regions are consumed in the design and modelling of
+signal regions, so need to be inspected.
+Control and validation region data are inspected, but this is not a problem
+since they should also have negligibly small yields (contaminations) from
+interesting signal models.
+With negligible signal contaminations, control and validation region data can
+be consumed to constrain the background modelling in addition to all other that
+on which the Standard Model stands, and without being influenced by the signal
+models to be tested in signal regions.
+
+The \atlas\ SUSY approval procedure~\cite{atlas_twiki_susyapprove}
+requires various criteria including, and not limited to, that the search has:
+\begin{itemize}
+\item defined all control, validation and signal regions,
+\item evaluated signal contaminations in control and validation,
+\item estimated all regions' simulated backgrounds,
+\item outlined procedures to estimate non-simulated (data-driven) backgrounds,
+\item evaluated variations due to experimental uncertainties in all regions
+according to recommended procedures,
+\item stated clearly how all other (theoretical) uncertainties will be treated.
+\end{itemize}
+So many aspects must be finalized blind, and for those that are not
+finalized blind must have procedures defined such that they are not
+influenced by the signal region data.
+Positively, disallowing dependence of procedures on signal region data
+eliminates one major source of unblinding error,
+and requiring knowledge of signal contaminations supports the use of control
+and validation region data.
+
+% con
+\TODO{flaws in that procedure:
+explicit allowance of data-driven methods using SR data,
+questionable enforcement of the stated procedures
+}
+
 Figure~\ref{fig:2ljets_prepre}

### 2022-08-03 main/chapter_2ljets.tex
discuss unblinding dis and impact

@@
 \label{sec:2ljets_low_isr}
-
 The Low regions presented in Section~\ref{sec:2ljets_low} resulted from a
@@
 \subsubsection{The Matrix Method}
+\label{sec:2ljets_matrix_method}
 Precedent from previous searches in this channel~\cite{
@@
 }
+\label{sec:2ljets_matrix_method_ewk}
 To validate the Matrix Method estimates in $\twoljets$-electroweak regions
@@
 but I know of no official advice on how to judge what is ``too small''.
+Unblinding procedures are discussed further in
+Section~\ref{sec:2ljets_unblinding}.
 
@@
 $\met$ that we select.
-We shall see in Section~\ref{sec:2ljets_\diboson_alt} why \diboson\ yields are
+We shall see in Section~\ref{sec:2ljets_diboson_alt} why \diboson\ yields are
 not similarly squashed by sample comparison variations.
@@
 \subsubsection{Alternative \diboson}
-\label{sec:2ljets_\diboson_alt}
+\label{sec:2ljets_diboson_alt}
 No alternative samples were available for comparison with our central \diboson
@@
 Unlike the alternative samples used for \diboson\ as described in
-Section~\ref{sec:2ljets_\diboson_alt}, these are assigned as functions of the
+Section~\ref{sec:2ljets_diboson_alt}, these are assigned as functions of the
 jet kinematics in each event by a parametrized method that avoids expensive
@@
 This differs from the methods previously described in
-Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_\diboson_alt},
+Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_diboson_alt},
 and is used for consistency with other $\twoljets$ work.
@@
 \item defined all control, validation and signal regions,
-\item evaluated signal contaminations in control and validation,
+\item evaluated signal contaminations in control and validation regions,
 \item estimated all regions' simulated backgrounds,
@@
 influenced by the signal region data.
-Positively, disallowing dependence of procedures on signal region data
+Encouragingly, disallowing dependence of procedures on signal region data
 eliminates one major source of unblinding error,
@@
 % con
-\TODO{flaws in that procedure:
-explicit allowance of data-driven methods using SR data,
-questionable enforcement of the stated procedures
-}
-
-Figure~\ref{fig:2ljets_prepre}
-Figure~\ref{fig:2ljets_summary_unblinding}
-
+Unfortunately, these requirements limit only when data-driven estimation
+procedures, are defined, and do not limit the content of those procedures.
+Indeed, the rules explicitly allow methods like the Matrix Method
+(described in Section~\ref{sec:2ljets_matrix_method})
+that use signal region data to affect the background `estimates' in signal
+regions, in violation of blinding principles.
+No sensible research would employ an estimation procedure that clearly set
+a prediction directly from the data it is supposed to predict, but our use of
+the Matrix Method is proof that sufficiently obfuscated violations are allowed.
+
+A second unfortunate flaw is that these rules were not strictly enforced.
+Although we performed the approval procedure as written, some of its promises
+were later violated, at the request of senior committees, without consequence.
+The data-driven FNP estimation procedure should have been frozen, but was in
+fact extended with the `Mean' method after unblinding,
+as seen in Section~\ref{sec:2ljets_matrix_method_ewk}).
+Theoretical uncertainties were always to be treated according to the
+\atlas\ SUSY recommendations~\cite{atlas_twiki_susytheoretical}, but those
+recommendations have ambiguities that were resolved after unblinding,
+notably for the alternative $\zjets$ sample comparison as discussed in
+Section~\ref{sec:2ljets_zjets_alt}, among other less impactful sample
+comparisons.
+
+% looser selections
 \begin{figure}[tp]
@@
 
+When first introduced to this search project, I was shown plots resembling
+(but not identical to) those in Figure~\ref{fig:2ljets_prepre}, which present
+data and pre-fit backgrounds in loose selections.
+Such plots were discussed in group meetings and used to catch bugs in the
+background simulations.
+But they unblind the data, notably in high-$\met$ where we expect signals.
+When commenting on this, I was told that it does not count as unblinding
+because we are in a loose selection.
+Any signal region must be within this loose selection, and those
+highest bins contain only a few events that agree with the backgrounds well.
+From such plots, we knew before designing the regions that either there was no
+excess at high $\met$ or that a signal region excess must coincide with a
+deficit in its surrounding phase space (or visa versa).
+Since we fill surrounding phase space with validation regions, one could in
+fact have predicted an SR-High deficit from the subtle excess in VR-High
+that is shown in Figure~\ref{fig:2ljets_high_mjj_vrhigh}.%
+\footnote{%
+Lacking this intelligence, at the time of unblinding I was actually concerned
+that we were missing a background (perhaps from pileup jets) that would
+explain low-$\mjj$ excess in VR-High, and might lead to a false excess
+in SR-High.
+My naive past self therefore found the observed $0$ events in SR-High-8
+quite hilarious.
+}
+If blinding is to be taken seriously, then restrictions must be placed on such
+pre-selection plots.
+
+% unblinded results
 \begin{figure}[tp]
@@
 \end{figure}
+
+Our initial unblinded results are displayed in the summary plot of
+Figure~\ref{fig:2ljets_summary_unblinding}.
+This figure includes all backgrounds and variations that had been evaluated
+at the time, and is included for comparison against the final result
+that has been displayed in Figure~\ref{fig:2ljets_summary}.
+We knew when unblinding that our simulations and data needed to be reprocessed
+with the lepton definitions given Section~\ref{sec:2ljets_events_leptons},
+which had previously required harder baseline leptons,
+and had expected only minor changes.
+As seen from these two figures, the reprocessing reduced our yields of data
+slightly; most noticeably, the one event in SR-$\llbb$ disappeared.
+Other than aesthetic differences, post-unblinding activities have had little
+impact.
+The reported significance of deficits in SR-Low-a and SR-Low-b has reduced,
+and the shrinking $\zjets$ background has caused a $1$-sigma excess in VR-Low.
+Excesses have also grown in VR-High-* regions.
+Little has changed in the SR-High-8 bins with $0$ data since, as explained in
+Section~\ref{sec:2ljets_diboson_alt}, the added \diboson\ sample comparison
+is ineffective.
+
+Despite my concerns expressed in this section, our post-unblinding changes are
+relatively minor.
+They do not change the broad message that there are no excesses here, and at
+most amount to a misleadingly small background estimate in SR-Low bins if the
+post-fit yields that we present are misinterpreted as predictions.
+Pre-fit yields, which can plausibly be interpreted as predictions, and other
+additional results are presented in the next section.
 

### 2022-08-03 main/chapter_2ljets.tex
some unblinding readthrough tweaks

@@
 directions of coherent variations; all variations are combined as described
-and varied together a single parameter in the fit.
+and varied together in all regions under the control of a single parameter in
+the fit.
 
@@
 % con
-Unfortunately, these requirements limit only when data-driven estimation
-procedures, are defined, and do not limit the content of those procedures.
-Indeed, the rules explicitly allow methods like the Matrix Method
+Unfortunately, these requirements limit only \emph{when} data-driven estimation
+procedures are defined, and do not limit the \emph{what} those procedures do.
+Indeed, the rules explicitly allow for methods like the Matrix Method
 (described in Section~\ref{sec:2ljets_matrix_method})
 that use signal region data to affect the background `estimates' in signal
-regions, in violation of blinding principles.
-No sensible research would employ an estimation procedure that clearly set
+regions in violation of blinding principles.
+No effective research would employ an estimation procedure that clearly set
 a prediction directly from the data it is supposed to predict, but our use of
-the Matrix Method is proof that sufficiently obfuscated violations are allowed.
+the Matrix Method is proof by example that sufficiently obfuscated violations
+are allowed here.
 
@@
 }
-If blinding is to be taken seriously, then restrictions must be placed on such
-pre-selection plots.
+If blinding is to be taken seriously, then restrictions must be placed on the
+use of such pre-selection plots.
 
@@
 
-Despite my concerns expressed in this section, our post-unblinding changes are
-relatively minor.
+Despite my concerns expressed in this section, these post-unblinding changes
+are relatively minor.
 They do not change the broad message that there are no excesses here, and at
@@
 post-fit yields that we present are misinterpreted as predictions.
-Pre-fit yields, which can plausibly be interpreted as predictions, and other
-additional results are presented in the next section.
+Pre-fit yields, which can plausibly be interpreted as predictions, are
+presented with other extra results in the next section.
 

### 2022-08-03 main/chapter_2ljets.tex
begin fleshing out results

@@
 \label{sec:2ljets_results}
-Figure~\ref{fig:2ljets_summary_no_sr}
-Figure~\ref{fig:2ljets_fit_corrmatrix}
-Figure~\ref{fig:2ljets_fit_parameters}
-Figure~\ref{fig:2ljets_fit_muti_contourss}
+
+Main results from the $\twoljets$-electroweak search have already been
+presented in Section~\ref{sec:2ljets_splash} and sprinkled throughout this
+chapter.
+This section reports some ancillary results.
+
+\TODO{Discuss and tie these together.}
+
+A pre-fit (fitted to control regions but not signal regions) summary plot
+is displayed in Figure~\ref{fig:2ljets_summary_no_sr}.
+
+The correlation matrix of parameters, estimated from the inverse Hessian at
+the best fit point and reduced to its most correlated parameters, is displayed
+in Figure~\ref{fig:2ljets_fit_corrmatrix}.
+
+All fitted parameters and their marginal error bars are displayed in
+Figure~\ref{fig:2ljets_fit_parameters}.
+
+Preliminary projected exclusion contours that separate the use only subsets
+of our signal regions to assess their effects are displayed in
+Figure~\ref{fig:2ljets_fit_muti_contourss}.
 
@@
 Parameters with smaller levels of correlation are excluded.
+Apologies for the small font size.
 }

### 2022-08-03 main/chapter_2ljets.tex
minor proofread changes

@@
 ]{%
-Histograms in a very loose pre-selection like those inspected to validate
+Histograms in a very loose pre-selection, like those inspected to validate
 background modelling prior to designing the $\twoljets$ regions.
@@
 excess at high $\met$ or that a signal region excess must coincide with a
-deficit in its surrounding phase space (or visa versa).
+deficit in its surrounding phase space (or vice versa).
 Since we fill surrounding phase space with validation regions, one could in
@@
 We knew when unblinding that our simulations and data needed to be reprocessed
-with the lepton definitions given Section~\ref{sec:2ljets_events_leptons},
+with the lepton definitions given in Section~\ref{sec:2ljets_events_leptons},
 which had previously required harder baseline leptons,

### 2022-08-03 main/chapter_2ljets.tex
add cls clsb clb commands

@@
 At each parameter point of a signal model, it is added to the background model
-and tested against data in the $\mathrm{CLs}$ prescription.
+and tested against data in the $\cls$ prescription.
 Contours showing which parameters of those signal models are labelled as
@@
 of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
-Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
+Exclusion is defined by the $95\%$ $\cls$ prescription
 in asymptotic approximations.
@@
 from asymptotic approximations to the prior distribution of the test statistic.
-Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
+Exclusion is defined by the $95\%$ $\cls$ prescription
 in asymptotic approximations.
@@
 the number of data observed,
-the observed $95\%$ $\mathrm{CLs}$ upper limit on the visible cross-section
+the observed $95\%$ $\cls$ upper limit on the visible cross-section
 $\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
@@
 designs, the general message has not changed.
-Values of CLs from mean data in the combined fit are overlaid on the grid.
+Values of $\cls$ from mean data in the combined fit are overlaid on the grid.
 }

### 2022-08-03 main/chapter_2ljets.tex
use \clb

@@
 & $S_{\mathrm{exp}}^{95}$
-& $\mathrm{CLb}$
+& $\clb$
 & $p_{s\!=\!0}$
@@
 $\pm1$-sigma variations,
-$\mathrm{CLb}$ evaluated with the signal expectation set to the observed upper limit,
+$\clb$ evaluated with the signal expectation set to the observed upper limit,
 and the discovery $p$-value (capped at $0.5$) with its equivalent significance.

### 2022-08-04 main/chapter_2ljets.tex
flesh out summary cov and parameters

@@
 \label{sec:2ljets_results}
-
 Main results from the $\twoljets$-electroweak search have already been
@@
 
-\TODO{Discuss and tie these together.}
-
-A pre-fit (fitted to control regions but not signal regions) summary plot
-is displayed in Figure~\ref{fig:2ljets_summary_no_sr}.
-
-The correlation matrix of parameters, estimated from the inverse Hessian at
-the best fit point and reduced to its most correlated parameters, is displayed
-in Figure~\ref{fig:2ljets_fit_corrmatrix}.
-
-All fitted parameters and their marginal error bars are displayed in
-Figure~\ref{fig:2ljets_fit_parameters}.
-
+\paragraph{Summary}
+As I have argued incessantly, pre-fit predictions are useful because they can
+actually be predictions to be tested by the data.
+Our post-fit summary plot has previously been displayed in
+Figure~\ref{fig:2ljets_summary}.
+Its alternative version that does not fit to signal region data is shown here
+in Figure~\ref{fig:2ljets_summary_no_sr}.
+Since the significance measure is meaningful for this pre-signal-region plot,
+we see that despite the inflated uncertainties in SR-Low, its bins still have
+substantial deficits.
+Validation region data are notably better predicted by this prediction;
+as the post-fit result tunes itself to signal region data, coupled variations
+lead to worse predictions in the validation regions (that are not included
+directly in either fit).
+
+\paragraph{Correlations}
+All parameters are initially independent.
+If the \heplikelihood\ is related to a probability distribution of parameters,
+however, then they gain interdependencies when coupled through the likelihoods
+of control and signal region data.
+The the correlation matrix of parameters, estimated from an approximate Hessian
+matrix at the best fit \heplikelihood, is displayed in
+Figure~\ref{fig:2ljets_fit_corrmatrix}.
+To reduce this matrix to a more readable size, only the most correlated
+parameters are shown.
+
+Mixed contributions of $\ttbar$, $\zjets$ and \diboson\ in the control regions
+leads to anticorrelations between their normalization factors
+``\texttt{mu\_*}'' that are seen in the bottom right corner of this correlation
+plot.
+Since statistical errors ``\texttt{gamma\_stat\_*}'' in control regions are not
+absorbed into the sample-specific normalization factors, they are also
+anticorrelated with normalization factors.
+Similarly, the extra uncertainty for $\zjets$ in CR-tt causes an
+anticorrelation between its parameter ``\texttt{alpha\_Zjets\_bextrap}''
+and the $\ttbar$ normalization.
+The failure of statistical precision in SR-Low-2 causes large systematic
+variations that appear as the anticorrelations between
+\texttt{gamma\_stat\_SRLow2\_cuts\_bin\_0} and numerous experimental
+uncertainties.
+Correlations between the $\zyjets$ normalization and $\ttbar$ uncertainties
+are due to large variations in the alternative $\ttbar$ samples in \crdy.
+Electron identification variations ``\texttt{alpha\_el\_chargeid\_*}'' have
+some impact in SR-Low,  so are correlated with the $\zjets$ alternative sample
+comparison ``\texttt{alpha\_Zjets\_alt}''.
+Note that the up and down variations of parameters need not correspond to up
+and down variations of yields in any region, so the signs of these correlations
+are no necessarily meaningful.
+
+\paragraph{Parameters}
+Correlations between parameters do not indicate their central values.
+To inspect individual parameters in the same fit to control and signal regions,
+Figure~\ref{fig:2ljets_fit_parameters} displays all fitted parameters and their
+marginal error bars.
+
+Parameters named ``\texttt{$\alpha$\_*}'' are centred on zero with unit initial width,
+and others are initially relative to $1$ with variable error bars.
+Internally, plots like this are considered good if all parameters sit on
+the central vertical line, with the motivations that normalization factors
+should be close to $1$ (the predicted cross-sections are accurate) and that
+our data should not compete with the precision of the studies that defined
+our systematic variations.
+That is, if we constrain a parameter by squeezing its error bar, then our data
+could be better used to calibrate that parameter throughout the Collaboration.
+
+I added the red arrow to indicate the $\zjets$ alternative sample comparison,
+named ``\texttt{$\alpha$\_Zjets\_alt}'' here, that has disappeared off the
+right-hand edge due to its best-fit value of $1.86 \pm 0.69$.
+(see Section~\ref{sec:2ljets_zjets_alt}).
+Its anticorrelation with the electron charge identification variations is
+explained by their movement to negative values;
+positive \texttt{$\alpha$\_Zjets\_alt} and negative
+\texttt{$\alpha$\_el\_chargeid\_*} both reduce yields in SR-Low.
+
+Several of the $\ttbar$ alternative-sample comparisons are also pulled away
+from zero.
+Some of this effect is explained by the drastic variation of
+\texttt{$\alpha$\_Zjets\_alt}, that nearly doubles $\zjets$ yields in CR-tt
+(see Table~\ref{tab:2ljets_zjets_alt}).
+Much of the remainder is due to the alternating deficit and excess in
+SR-OffShell-a and SR-OffShell-b respectively;
+in particular, the
+final state radiation (\texttt{$\alpha$\_ttbar\_fsr}) and
+hard scatter (\texttt{$\alpha$\_ttbar\_hs}) variations both move $\ttbar$
+yields in opposite directions in SR-OffShell-a and SR-OffShell-b, so the
+movements of these parameters help to fit both deviations simultaneously.
+
+\paragraph{Contours}
 Preliminary projected exclusion contours that separate the use only subsets

### 2022-08-04 main/chapter_2ljets.tex
add cls obs contours; add contuor results text

@@
 \subsection{Contributions}
-
 In this subsection, I claim attributions of work on the $\twoljets$ analysis
@@
 Full details are publically available through the $\twoljets$
-HEPData~\cite{hepdata.116034, Maguire_2017} archive our json-serialized pyhf
+\hepdata~\cite{hepdata.116034, Maguire_2017} archive our json-serialized pyhf
 workspaces~\cite{cranmer2012histfactory, heinrich2021pyhf}.
-\emph{Thanks to Jon~Long for collaboration in preparing this HEPData archive,
+\emph{Thanks to Jon~Long for collaboration in preparing this \hepdata\ archive,
 and to Alessia~Murrone for invaluable feedback on, and validation of, its
@@
 and to my knowledge, is only reported through inner details of
-the HistFactory models shared in our HEPData archive~\cite{hepdata.116034}.
+the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}.
 
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.png}
+\caption[
+Alternative summary table with backgrounds fitted to control regions only
+]{%
+Alternative summary table with backgrounds fitted to control regions only.
+A version fitted also to signal regions is shown in
+Figure~\ref{fig:2ljets_summary}.
+Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
+from the intentions of the significance measure.
+Because the post-fit yields in SR-Low achieve a small negative significance,
+all mentions of its deficit were removed from the paper.
+}
+\label{fig:2ljets_summary_no_sr}
+\end{figure}
+
 \paragraph{Correlations}
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_corrmatrix.pdf}
+\caption[
+Correlation matrix from the background-only fit to control and signal regions
+]{%
+Correlation matrix from the background-only fit to control and signal regions,
+as estimated by \histfitter.
+Parameters with smaller levels of correlation are excluded.
+Apologies for the small font size.
+}
+\label{fig:2ljets_fit_corrmatrix}
+\end{figure}
+
 \paragraph{Parameters}
@@
 
-\paragraph{Contours}
-Preliminary projected exclusion contours that separate the use only subsets
-of our signal regions to assess their effects are displayed in
-Figure~\ref{fig:2ljets_fit_muti_contourss}.
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.png}
-\caption[
-Alternative summary table with backgrounds fitted to control regions only
-]{%
-Alternative summary table with backgrounds fitted to control regions only.
-A version fitted also to signal regions is shown in
-Figure~\ref{fig:2ljets_summary}.
-Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
-from the intentions of the significance measure.
-Because the post-fit yields in SR-Low achieve a small negative significance,
-all mentions of its deficit were removed from the paper.
-}
-\label{fig:2ljets_summary_no_sr}
-\end{figure}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_corrmatrix.pdf}
-\caption[
-Correlation matrix from the background-only fit to control and signal regions
-]{%
-Correlation matrix from the background-only fit to control and signal regions,
-as estimated by \histfitter.
-Parameters with smaller levels of correlation are excluded.
-Apologies for the small font size.
-}
-\label{fig:2ljets_fit_corrmatrix}
-\end{figure}
-
 \begin{figure}[tp]
@@
 $1.86 \pm 0.69$;
-since it exceeds $1.5$, it was otherwise excluded from this plot.
+since it exceeds $1.5$, it otherwise disappeared off the edge of this plot.
 }
@@
 \end{figure}
+
+\paragraph{Contours}
+Exclusion contours, which are displayed in
+Figures~\ref{fig:2ljets_contours_c1n2} and~\ref{fig:2ljets_contours_gmsb},
+derive from fits with models that extend the background-only model with
+additive signal samples.
+To reduce the computational cost of building the hundreds of models%
+\footnote{%
+This is worded precisely. The dominant cost was not in fitting the models, nor
+in histogramming the data (thanks to histogram caching), but in forming the
+HistFactory model from those histograms.
+I do not know why that model-building step took days of computer time per
+model, but it did reduce to only hours after upgrading the \histfitter\ version
+from \textsc{v0.64.0} to \textsc{v0.66.0}.
+}%
+across these grids, we prune the number of small irrelevant parameters
+with \histfitter's pruning~\cite{histfittergithub} mechanism in ``method 2''.
+This pruning method removes variations if the difference between their up and
+down values do not exceed a threshold.
+With pruning enabled, its default threshold is set to $1\%$ (relative to the
+central yield), but to reduce its impact we use a threshold of $0.1\%$.
+Validation studies comparing each of these pruning thresholds, against the
+un-pruned version, and against a custom pruning implemented by myself, found no
+significant differences.%
+\footnote{%
+Pruning is, however, the reason why the signal patches (differences from
+background-only to signal) in our serialized \hepdata\ models are so large;
+they implement many small changes due to different decisions made by the
+pruning implementation.
+}
+
+We further reduce computational costs by sing the asymptotic approximations
+(as opposed to the expensive `toy throwing' approximation) for all statistical
+tests on these exclusion fits.
+
+To some extent, our published exclusion contours are an artistic choice.
+What we know are the values of statistics in an irregular grid of
+signal points,
+and all contours are painted by heuristic interpolation algorithms.
+We follow standard \atlas\ SUSY conventions and perform that interpolation
+with the \textsc{Scipy} implementation of
+Radial Basis Functions~\cite{2020SciPy-NMeth,fasshauer2007meshfree},
+interpolating significance-transformed $\cls$ values.
+That is we interpolate in $\cls$ values transformed through the inverse
+cumulative distribution of the standard normal distribution.
+Contours are drawn where this interpolated function meets
+$\cls = 0.05$.
+\emph{%
+Thanks to Larry~Lee for sharing contour preparation scripts that were adapted
+to produce our results.%
+}
+
+This interpolation algorithm has some tunable parameters, for which all
+contours of the $\twoljets$-electroweak search use
+\texttt{epsilon=10},
+\texttt{smoothing=1},
+and the ``multiquadric'' radial basis function.
+These settings were tuned to keep the thumbs-up wobble in the C1N2 off-shell
+region of Figure~\ref{fig:2ljets_contours_c1n2},
+which is physical and due to our increased sensitivity for mass splittings
+below $m(W)$,
+while sufficiently smoothing wobbles elsewhere.
+Some unattractive wobbles remain at high GMSB masses in
+Figure~\ref{fig:2ljets_contours_gmsb}, but I decided it was preferable to keep
+the simplicity of having the same interpolation settings for both contours.
+
+Initially, the contours showed more glaring wobbles.
+These were removed by repeating the fits of anomalous signal points until they
+behaved.
+In total, this hack was used for two points near the highest
+$m(\chargino_1,\neutralino_2)$ edge in C1N2 and four across the whole
+GMSB spectrum.
+Different random seeds causing failed fits explain this unstable behaviour,
+and I failed to control those random seeds when running the fits through
+\histfitter.
+
+Auxiliary results published in our \hepdata archive~\cite{hepdata.116034}
+include the observed upper limits on the signal cross-section at each point.
+We additionally display the $\cls$ values, from which observed contour derives,
+in Figure~\ref{fig:2ljets_fit_contours_cls}.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_c1n2_Output_Contours_cls_obs.png}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_gmsb_Output_Contours_cls_obs.png}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Exclusion contours with observed $\cls$ values
+]{%
+Exclusion contours with observed $\cls$ values.
+The observed exclusion contour is interpolated from these values.
+Parameter points are at the lower left corners of the text representing each
+$\cls$ value.
+}
+\label{fig:2ljets_fit_contours_cls}
+\end{figure}
+
+\TODO{consider moving this to region design?}
+Preliminary projected exclusion contours that separate the use only subsets
+of our signal regions to assess their effects are displayed in
+Figure~\ref{fig:2ljets_fit_muti_contourss}.
 
@@
 designs, the general message has not changed.
-Values of $\cls$ from mean data in the combined fit are overlaid on the grid.
+Values of $\cls$ from expected (not observed --- the label is wrong) data in
+the combined fit are overlaid on the grid.
 }
@@
 \end{figure}
+
+% TODO move figures
+
+\paragraph{Discovery}
+\TODO{pruning $1\%$}

### 2022-08-04 main/chapter_2ljets.tex
minor cleanup, spelling grammar check

@@
 of control and signal region data.
-The the correlation matrix of parameters, estimated from an approximate Hessian
+The correlation matrix of parameters, estimated from an approximate Hessian
 matrix at the best fit \heplikelihood, is displayed in
@@
 \footnote{%
-Pruning is, however, the reason why the signal patches (differences from
+Pruning is, however, the reason the signal patches (differences from
 background-only to signal) in our serialized \hepdata\ models are so large;
@@
 
-We further reduce computational costs by sing the asymptotic approximations
+We further reduce computational costs by using the asymptotic approximations
 (as opposed to the expensive `toy throwing' approximation) for all statistical
@@
 interpolating significance-transformed $\cls$ values.
-That is we interpolate in $\cls$ values transformed through the inverse
+That is, we interpolate in $\cls$ values transformed through the inverse
 cumulative distribution of the standard normal distribution.
@@
 \emph{%
-Thanks to Larry~Lee for sharing contour preparation scripts that were adapted
-to produce our results.%
+Thanks to Larry~Lee for sharing scripts that were adapted to prepare the
+contours in our results.%
 }
@@
 include the observed upper limits on the signal cross-section at each point.
-We additionally display the $\cls$ values, from which observed contour derives,
-in Figure~\ref{fig:2ljets_fit_contours_cls}.
+We additionally display the $\cls$ values, from which the observed contour
+derives, in Figure~\ref{fig:2ljets_fit_contours_cls}.
 
@@
 
-% TODO move figures
-
 \paragraph{Discovery}

### 2022-08-04 main/chapter_2ljets.tex
improve pruning footnote

@@
 \footnote{%
-Pruning is, however, the reason the signal patches (differences from
-background-only to signal) in our serialized \hepdata\ models are so large;
-they implement many small changes due to different decisions made by the
-pruning implementation.
+Pruning is, however, the reason our signal patches (that describe differences
+from background-only to signal models) in our serialized \hepdata\ models are
+so large; they implement many small changes due to different decisions made by
+the pruning algorithm.
 }

### 2022-08-04 main/chapter_2ljets.tex
minor text tweaks; change to pdf conouts so grey numbers can be zoomed on

@@
 Some unattractive wobbles remain at high GMSB masses in
-Figure~\ref{fig:2ljets_contours_gmsb}, but I decided it was preferable to keep
-the simplicity of having the same interpolation settings for both contours.
-
-Initially, the contours showed more glaring wobbles.
-These were removed by repeating the fits of anomalous signal points until they
-behaved.
+Figure~\ref{fig:2ljets_contours_gmsb}, but I decided that it was preferable to
+keep the simplicity of having the same interpolation settings for both
+contours.
+
+Our initial contours showed more glaring and unphysical wobbles.
+Those erroneous wobbles were removed by repeating the fits of anomalous signal
+points until they behaved.
 In total, this hack was used for two points near the highest
-$m(\chargino_1,\neutralino_2)$ edge in C1N2 and four across the whole
+$m(\chargino_1,\neutralino_2)$ edge in C1N2, and four across the whole
 GMSB spectrum.
@@
 
-Auxiliary results published in our \hepdata archive~\cite{hepdata.116034}
+Auxiliary results published in our \hepdata\ archive~\cite{hepdata.116034}
 include the observed upper limits on the signal cross-section at each point.
@@
 derives, in Figure~\ref{fig:2ljets_fit_contours_cls}.
+By zooming in, one can see that not all points with $\cls < 0.05$ are
+labelled as excluded;
+this is direct a result of our smoothing choices.
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_c1n2_Output_Contours_cls_obs.png}
+\includegraphics[width=\textwidth]{figures/2ljets_c1n2_Output_Contours_cls_obs.pdf}
 \caption{C1N2}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_gmsb_Output_Contours_cls_obs.png}
+\includegraphics[width=\textwidth]{figures/2ljets_gmsb_Output_Contours_cls_obs.pdf}
 \caption{GMSB}

### 2022-08-04 main/chapter_2ljets.tex
which/that, improve matrix method non-prediction paragraph

@@
 Yes, actual data are histogrammed in the Matrix Method, and those data include
-tight-tight events which comprise the data of our search regions, and those
-are the same data in which we are searching for anomalous signals.
+tight-tight events that comprise the data of our search regions, the very same
+data in which we are searching for anomalous signals.
 Since it includes actual data, this method unblinds all regions,
 including signal regions, early.
-And because it knows the data we are supposed to be predicting, it is no
-prediction at all!
+And because it knows and uses the data we are supposed to be predicting, it is
+no prediction at all!
 

### 2022-08-04 main/chapter_2ljets.tex
replace some 0s with zero

@@
 Data are shown; regions in the top two plots observe zero data, and
-Poisson error bars for $0$ events are hidden for aesthetic reasons.
+Poisson error bars for zero events are hidden for aesthetic reasons.
 }
@@
 After the other selections for SR-Low-2, almost all signal and background
-samples are in the spike near $0$;
+samples are in the spike near zero;
 a lower requirement would make little difference, and $80\,\eV[G]$ matches
@@
 in SR-High.
-My naive past self therefore found the observed $0$ events in SR-High-8
+My naive past self therefore found the observed zero events in SR-High-8
 quite hilarious.
@@
 Excesses have also grown in VR-High-* regions.
-Little has changed in the SR-High-8 bins with $0$ data since, as explained in
+Little has changed in the SR-High-8 bins with zero data since, as explained in
 Section~\ref{sec:2ljets_diboson_alt}, the added \diboson\ sample comparison

### 2022-08-05 main/chapter_2ljets.tex
move orthogonal region impact to design section

@@
 % That is the idea that we all hope you have learned in studying science in
-% school--we never say explicitly what this is, but just hope that you catch
+% school --- we never say explicitly what this is, but just hope that you catch
 % on by all the examples of scientific investigation. It is interesting,
@@
 I choose this order for two reasons.
-First, it avoids iterating over those elements twice, once before and again
+Firstly, it avoids iterating over those elements twice, once before and again
 after unblinding.
@@
 chapter.
-Second, and sadly, many design decisions can only be explained in light of the
+Secondly, and sadly, many design decisions can only be explained in light of the
 unblinded results, since important aspects were reworked in the
-$19$~months between its unblinding and eventual publication.
+nineteen months between its unblinding and eventual publication.
 
@@
 Section~\ref{sec:2ljets_modelling}.
-\item Some validation studies are presented in Section~\ref{sec:2ljets_validation}.
+\item Some validation studies are presented in
+Section~\ref{sec:2ljets_validation}.
 \item The unblinding procedure of this search is discussed in
 Section~\ref{sec:2ljets_unblinding}, and
-\item we recap the search results in Section~\ref{sec:2ljets_results}.
+\item and some additional results are presented in
+Section~\ref{sec:2ljets_results}.
+\TODO{Conclusions}
 \end{itemize}
@@
 Uniquely to Low regions, and signal models with small mass-splitting, it is
-seen (in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}) that substantial
+shown in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2} that substantial
 signal yields exist at small $\mttwoll$.
@@
 Figure~\ref{fig:2ljets_offshell_region}.
+
+\subsection{Impact}
+\label{sec:2ljets_imapct}
+We have now seen the definitions of OffShell, Low, Intermediate and High
+regions, how their designs intend to target certain regions of parameter space,
+and how example signal models can create visible effects in those regions.
+To further demonstrate that these regions work as intended,
+Figure~\ref{fig:2ljets_fit_multi_contours} presents projected exclusion
+contours generated from fits with subsets of the signal regions.
+For each contour, fits are performed with only the control regions and one
+subset of the signal regions as labelled and described in the figure caption.
+
+Combining all regions provides greater exclusion than any individual contour,
+and avoids the sharp corners that could have been assigned by stitching those
+individual parts together.
+That confirms that our design of combined, orthogonal regions works as
+intended.
+In most locations, however, most exclusion is attributable to only one subset
+of the regions.
+That shows that the combination is a relatively minor refinement that may or
+may not be worth the cost of implementing it.
+
+At the highest $m(\chargino_1,\neutralino_2)$ edge of the C1N2 contours,
+the inclusion of SR-High-1J (in addition to all other SR-High bins) is shown
+to extend the exclusion by around $60\,\eV[G]$.
+Across the rest of the C1N2 grid, each set of regions is impacting its targeted
+parameter space.
+Sensitivity to the C1N2 ($200$, $100$) point was a goal of the SR-Low design,
+and although it alone does not quite achieve expected exclusion there, the
+combination with other regions does.
+
+For GMSB, only SR-$\llbb$ has sensitivity to large
+$B(\neutralino_1 \rightarrow h \tilde{G})$ since it is able to capture the
+resulting $h\rightarrow bb$ decays.
+At its highest masses, the main SR-High regions retain sensitivity to its
+$ZZ$ decays, and SR-High-1J again visibly extends the projected exclusion.
+At its lowest masses, SR-Int and SR-Low both contribute.
+These light GMSB signal models have substantial contaminations in VR-Low, which
+are seen in Figure~\ref{fig:2ljets_low_region} to exceed $50\%$ of the central
+background.
+That SR-Int also provides sensitivity here is some comfort that this signal
+contamination does not entirely invalidate the result.
+We do not consider off-shell GMSB models, so SR-OffShell regions do not exclude
+any points alone here.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_c1n2.pdf}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_gmsb.pdf}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Projected exclusion from sets of orthogonal signal regions
+]{%
+Projected exclusion from sets of orthogonal signal regions, demonstrating that
+their combination has greater sensitivity reach than any individually.
+The label `High' indicates the bins of SR-High, and `HighNo4' excludes
+SR-High-1J to show the locations where mono-jet events have an impact.
+Similarly, `$\llbb$', `Int', `Low', and `OffShell' are shown to target their
+appropriate parameter spaces.
+Although figures were produced before unblinding, with older samples and region
+designs, their later updates do not change the general message.
+Values of $\cls$ from expected (not observed --- the label is wrong) data in
+the combined fit are overlaid on the grid.
+}
+\label{fig:2ljets_fit_multi_contours}
+\end{figure}
 
@@
 
-\TODO{consider moving this to region design?}
-Preliminary projected exclusion contours that separate the use only subsets
-of our signal regions to assess their effects are displayed in
-Figure~\ref{fig:2ljets_fit_muti_contourss}.
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_c1n2.pdf}
-\caption{C1N2}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_gmsb.pdf}
-\caption{GMSB}
-\end{subfigure}
-\caption[
-Projected exclusion from sets of orthogonal signal regions
-]{%
-Projected exclusion from sets of orthogonal signal regions, demonstrating that
-their combination has greater sensitivity reach than any individually.
-The label `High' indicates the bins of SR-High, and `HighNo4' excludes
-SR-High-1J to show the locations where mono-jet events have an impact.
-Similarly, `$\llbb$', `Int', `Low', and `OffShell' are shown to target their
-appropriate parameter spaces.
-Although figures were produced before unblinding, with older samples and region
-designs, the general message has not changed.
-Values of $\cls$ from expected (not observed --- the label is wrong) data in
-the combined fit are overlaid on the grid.
-}
-\label{fig:2ljets_fit_muti_contourss}
-\end{figure}
-
 \paragraph{Discovery}

### 2022-08-08 main/chapter_2ljets.tex
add discovery extra results figures

@@
 as our central, reconstructed samples.
-These sample sizes vary roughly with cross-section from $300\,000$ to
-$2.8$~million events per sample, and due to the $\sqrt{n}$ scaling of
+These sample sizes vary roughly with cross-section from $300$k to
+$2.8$m events per sample, and due to the $\sqrt{n}$ scaling of
 Monte Carlo errors, means approximately $\sqrt{10}$ times reduced
@@
 \TODO{pruning $1\%$}
+
+Figure~\ref{fig:2ljets_test_stat_distributions}
+
+Figure~\ref{fig:2ljets_test_stat_distributions}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_upper_limit_scan_drhigh.pdf}
+\caption{DR-High}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_upper_limit_scan_drlow.pdf}
+\caption{DR-Low}
+\end{subfigure}
+\caption[
+Upper limit scans for DR-High and DR-Low
+]{%
+Upper limit scans for (a) DR-High and (a) DR-Low.
+Solid black lines interpolate the $p$-values $\clsb$ and $\clb$ estimated at
+each marked point with the observed data.
+The upper limit is reported where the interpolation of their ratio $\cls$
+crosses $0.05$.
+Coloured backgrounds illustrate what would have been observed with data at
+$\pm1$-sigma quantiles of the background-only distribution.
+Each point uses $500$k ``toy'' samples for the signal-plus-background
+distribution, and $1$m samples for the backgroun-only distribution.
+Individual distributions are inspected in
+Figure~\ref{fig:2ljets_test_stat_distributions}.
+}
+\label{fig:2ljets_upper_limit_scans}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.50\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_drhigh3.pdf}
+\caption{DR-High, $\mu_\textrm{Discovery}=3$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.46\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_drlow10.pdf}
+\caption{DR-Low, $\mu_\textrm{Discovery}=10$}
+\end{subfigure}
+\caption[
+Test statistic distributions for points in the DR-High and DR-Low scans
+]{%
+Test statistic distributions for points in the (a) DR-High and (b) DR-Low
+scans, with $\mu_\textrm{Discovery}=3$ and $\mu_\textrm{Discovery}=10$
+respectively.
+Axis limits are set from the minimum and maximum test statistics recorded.
+Note the different binning: histograms contain normalized densisites with
+different widths between the two distributions.
+The black vertical line indicates the observed test statistic,
+and the $p$-values $\clb$ and $\clsb$ are tail areas to its right.
+If the two distributions were predictive, their relative densities at that
+observation would indicate their likelihood ratio.
+Results are:
+% clb =  0.001322
+% clsb =  5.7e-05
+% cls =  0.0431164901664
+(a, DR-High) $\clb = 1.3\times 10^{-3}$ and $\clsb = 5.7\times 10^{-5}$,
+% clb =  0.087502
+% clsb =  0.007667
+% cls =  0.0876208543805
+(b, DR-Low) $\clb = 8.8\times 10^{-2}$ and $\clsb = 7.7\times 10^{-3}$.
+}
+\label{fig:2ljets_test_stat_distributions}
+\end{figure}

### 2022-08-08 main/chapter_2ljets.tex
begin writing on odd discovery region results

@@
 \paragraph{Discovery}
-\TODO{pruning $1\%$}
-
-Figure~\ref{fig:2ljets_test_stat_distributions}
-
-Figure~\ref{fig:2ljets_test_stat_distributions}
+Our discovery region results,
+which have been presented in Table~\ref{tab:2ljets_discovery},
+feature, and veil, some interesting quirks.
 
@@
 \end{subfigure}
+\\[.5ex]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_upper_limit_scan_droffshell.pdf}
+\caption{DR-OffShell}
+\end{subfigure}
 \caption[
@@
 ]{%
-Upper limit scans for (a) DR-High and (a) DR-Low.
+Upper limit scans for (a) DR-High, (b) DR-Low, and (c) DR-OffShell.
 Solid black lines interpolate the $p$-values $\clsb$ and $\clb$ estimated at
@@
 crosses $0.05$.
+\\
 Coloured backgrounds illustrate what would have been observed with data at
 $\pm1$-sigma quantiles of the background-only distribution.
-Each point uses $500$k ``toy'' samples for the signal-plus-background
-distribution, and $1$m samples for the backgroun-only distribution.
+Each point uses $500$k ``toy'' samples for the background-only
+distribution, and $1$m samples for the signal-plus-background distribution.
 Individual distributions are inspected in
 Figure~\ref{fig:2ljets_test_stat_distributions}.
+\\
+\emph{These figures were produced with \histfitter\ tools through
+my ``Bailed statistics'' software~\cite{bailedstatisticsgithub}.}
 }
@@
 
+Perhaps what should be the most meaningful result is the observed upper limit
+$S_{\mathrm{obs}}^{95}$.
+The analysis of each region sets its upper limit by scanning different values
+of the signal expectation $\mu_\textrm{Discovery}$ and interpolating to
+estimate the point where $\cls(\mu_\textrm{Discovery}^*) = 0.05$.
+That interpolated point determines
+$\mu_\textrm{Discovery}^* = S_{\mathrm{obs}}^{95}$, and the reported value
+of $\clb$ is also reported at that signal expectation
+$\mu_\textrm{Discovery}^*$, since the test statistic is a function of
+$\mu_\textrm{Discovery}$.
+Three examples of these scans are visualized in
+Figure~\ref{fig:2ljets_upper_limit_scans}.
+
+Some of our discovery regions have small yields for which the asymptotic
+approximations~\cite{Cowan:2010js} to test statistic distributions are not
+believed to work.
+For consistency, we therefore use the alternative toy simulation
+approximation for all discovery regions.
+Sampling of the test statistic distributions
+(with \histfitter-\roofit\ software)
+is extremely computationally expensive, so cut those costs by pruning our
+discovery regions models \histfitter's ``method 2'' at a $1\%$ tolerance.
+As discussed previously, this makes negligible differences to the results.
+The exact pruned models are shared in our \hepdata\
+archive~\cite{hepdata.116034}.
+Not only do these simulations take computational time, they also suffer
+torrential memory leaks that halt progress beyond small sample sizes.
+To work around this, we evaluate and combine many small samples, with different
+random seeds, with software that I name
+``Bailed statistics''~\cite{bailedstatisticsgithub}%
+\footnote{%
+To share software that assists in spending computational resources to
+calculate statistics that are so deeply flawed causes me some shame.
+In a vain attempt to absolve myself of this sin, and to positively encourage
+honest reporting of the data, I include with it the following note:
+\emph{%
+``Hypotheses compare through the relative likelihoods they assign to data.
+A p-value is a cumulative distribution function evaluated at data; CLs is a
+ratio of p-values.
+Please present results of this software accurately and clearly.
+Examples of false or unclear presentations of a p-value or CLs include:
+--- as a probability that an hypothesis is true or false,
+--- as a probability of compatibility with an hypothesis,
+--- as a probability that data occurred at random, by chance, or by statistical
+fluctuation,
+--- as a likelihood or importance of data,
+--- as necessary for an optimal or rational decision rule.
+Please also respect that the association of a p-value or CLs with the words
+`test', `limit', `confidence', `significance', `exclusion', `evidence',
+`observation' or `discovery' is nominal,
+and may not reflect the words' meanings in English.''
+}%
+}
+as it bails the leaks overboard.
+Despite this workaround, the simulations took several weeks of parallel
+processing on around $1500$ processor cores, accumulating to around a million
+core hours.
+
 \begin{figure}[tp]
@@
 \caption{DR-High, $\mu_\textrm{Discovery}=3$}
+\label{fig:2ljets_test_stat_distributions_drhigh}
 \end{subfigure}
@@
 \end{subfigure}
+\\[.5ex]
+\begin{subfigure}{0.46\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_droffshell14.pdf}
+\caption{DR-OffShell, $\mu_\textrm{Discovery}=14$}
+\end{subfigure}
 \caption[
@@
 ]{%
-Test statistic distributions for points in the (a) DR-High and (b) DR-Low
-scans, with $\mu_\textrm{Discovery}=3$ and $\mu_\textrm{Discovery}=10$
-respectively.
+Test statistic distributions for points in the (a) DR-High, (b) DR-Low,
+and (c) DR-OffShell scans at points close to their reported upper limits:
+$\mu_\textrm{Discovery}=3$, $\mu_\textrm{Discovery}=10$,
+and $\mu_\textrm{Discovery}=14$ respectively.
+\\
 Axis limits are set from the minimum and maximum test statistics recorded.
-Note the different binning: histograms contain normalized densisites with
+Note the different binning: histograms contain normalized densities with
 different widths between the two distributions.
@@
 observation would indicate their likelihood ratio.
+\\
 Results are:
@@
 % cls =  0.0431164901664
-(a, DR-High) $\clb = 1.3\times 10^{-3}$ and $\clsb = 5.7\times 10^{-5}$,
+(a) $\clb = 1.3\times 10^{-3}$ and $\clsb = 5.7\times 10^{-5}$,
 % clb =  0.087502
@@
 % cls =  0.0876208543805
-(b, DR-Low) $\clb = 8.8\times 10^{-2}$ and $\clsb = 7.7\times 10^{-3}$.
+(b) $\clb = 8.8\times 10^{-2}$ and $\clsb = 7.7\times 10^{-3}$,
+% clb =  0.67508
+% clsb =  0.03661
+% cls =  0.0542306097055
+and (c) $\clb = 0.68$ and $\clsb = 0.037$.
+\\
+\emph{These figures were adapted from plots produced with \histfitter\ tools
+through my ``Bailed statistics'' software~\cite{bailedstatisticsgithub}.}
 }
@@
 \end{figure}
+
+Many test statistic distributions are involved here --- two (background-only
+and signal-plus-background) at each scanned point, and one additional
+distribution per region (with yet another test statistic) for the discovery
+$p$-value $p_{s\!=\!0}$.
+From these, a few examples are presented in
+Figure~\ref{fig:2ljets_test_stat_distributions}
+to investigate curious features of the results.
+
+Our first curiosity is that the reported $S_{\mathrm{obs}}^{95} = 3.0$
+for DR-High is rounded up from a result of $2.93$ that was evaluated with these
+simulation methods.
+This makes sense --- the limit for $n=0$ should arguably always be $3$ since
+if a signal adds $\mu$ events, then the Poisson likelihood for zero data always
+scales by $e^{-\mu} \approx 0.05$, and so any likelihood ratio between
+background-only and signal-plus-background models is $0.05$ for $\mu = 3$.
+By taking a ratio of $p$-values, $\cls$ approximates that likelihood ratio:
+we see from Figure~\ref{fig:2ljets_test_stat_distributions_drhigh} that the
+observed data does land in a spike at the end of the distribution that
+corresponds to to pseudo-data with $n=0$; the peaks to its left correspond
+to $n=1$, $n=2$ and so on, with widths due to the other parameters varied in
+toy sampling.
+
+The $n=0$ case has no visible width since its likelihood is linear
+(adding signal just adds a constant to the $-\log$ likelihood objectives being
+minimized), but its distribution is not quite a delta function since the test
+statistic depends on the noisy termination of numerical optimization
+algorithms.
+Resulting $\clb$ and $\clsb$ values depend on this effect ---
+with $3.9$ expected background events and zero data we expect
+$\clb = \mathrm{Poisson}(0\mid 3.9) = e^{-3.9} \approx 2.0\times 10^{-2}$.
+And that is $15$ times larger than the evaluated $\clb = 1.3\times 10^{-3}$
+because, as shown in Figure~\ref{fig:2ljets_test_stat_distributions_drhigh_n0},
+the observed test statistic happened to fall quite far to the right
+of that $n=0$ noise distribution.
+This would appear to be an anomalously large significance if $\clb$ were
+misinterpreted as a measure of compatibility with the background model.
+Thankfully it is remedied by the $\cls$-prescribed ratio with $\clsb$,
+since both $p$-values suffer from similar effects.
+Statistical precision is harmed, however --- although $1$m samples should
+give us around $e^{-7.9} \times 10^6 \approx 1000$ signal samples with
+$n=0$, only $57$ toy samples land to the right of the observed test statistic.
+And with $\sqrt{n}$ precision, that easily explains the difference between
+limits at $3.0$ and $2.93$ despite the huge sample size.
+
+\TODO{}
+% real world costs to pretending that our uncertainties are auxiliary data
+% not only idle Bayesian-versus-Frequentist philosophy
+% other examples - cite 4.6 limit.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/2ljets_drhigh_plot_n0.pdf}
+\caption[
+Test statistic distribution for DR-High close to 3
+]{%
+Test statistic distribution for DR-High close to $3$.
+This zooms in to the narrow spike of toy samples that should have $n=0$
+observed data in the discovery region.
+Because the likelihood function for zero data,
+$\mathrm{Poisson}(0\mid x) = e^{-x}$, is $\log$-linear,
+this would analytically be a delta spike, but it acquires width from the finite
+precision of numerical optimization algorithms.
+The observed test statistic falling far to the right of that numerical noise
+reduces our statistical precision in estimating $\clb$ and $\clsb$.
+}
+\label{fig:2ljets_test_stat_distributions_drhigh_n0}
+\end{figure}

### 2022-08-08 main/chapter_2ljets.tex
splash out more words on perverse discovery results. Begin discussion with bfod quote

@@
 Full details are publically available through the $\twoljets$
-\hepdata~\cite{hepdata.116034, Maguire_2017} archive our json-serialized pyhf
+\hepdata~\cite{hepdata.116034, Maguire_2017} archive our json-serialized \pyhf\
 workspaces~\cite{cranmer2012histfactory, heinrich2021pyhf}.
@@
 \\
+For visual effect, the test statistic for $\mu_\textrm{discovery}=16$
+in DR-Low is edited from an anomalous value near $-1$ to $4.6$.
+This aesthetic change does not affect the results.
+\\
 \emph{These figures were produced with \histfitter\ tools through
@@
 The analysis of each region sets its upper limit by scanning different values
-of the signal expectation $\mu_\textrm{Discovery}$ and interpolating to
-estimate the point where $\cls(\mu_\textrm{Discovery}^*) = 0.05$.
+of the signal expectation $\mu_\textrm{discovery}$ and interpolating to
+estimate the point where $\cls(\mu_\textrm{discovery}^*) = 0.05$.
 That interpolated point determines
-$\mu_\textrm{Discovery}^* = S_{\mathrm{obs}}^{95}$, and the reported value
+$\mu_\textrm{discovery}^* = S_{\mathrm{obs}}^{95}$, and the reported value
 of $\clb$ is also reported at that signal expectation
-$\mu_\textrm{Discovery}^*$, since the test statistic is a function of
-$\mu_\textrm{Discovery}$.
+$\mu_\textrm{discovery}^*$, since the test statistic is a function of
+$\mu_\textrm{discovery}$.
 Three examples of these scans are visualized in
@@
 torrential memory leaks that halt progress beyond small sample sizes.
-To work around this, we evaluate and combine many small samples, with different
-random seeds, with software that I name
+To work around this, we evaluate and combine many small samples (with different
+random seeds), using software that I share under the name
 ``Bailed statistics''~\cite{bailedstatisticsgithub}%
@@
 }
-as it bails the leaks overboard.
+as it bails those memory leaks overboard.
 Despite this workaround, the simulations took several weeks of parallel
@@
 \includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_drhigh3.pdf}
-\caption{DR-High, $\mu_\textrm{Discovery}=3$}
+\caption{DR-High, $\mu_\textrm{discovery}=3$}
 \label{fig:2ljets_test_stat_distributions_drhigh}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_drlow10.pdf}
-\caption{DR-Low, $\mu_\textrm{Discovery}=10$}
+\caption{DR-Low, $\mu_\textrm{discovery}=10$}
+\label{fig:2ljets_test_stat_distributions_drlow}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/2ljets_test_stat_distribution_droffshell14.pdf}
-\caption{DR-OffShell, $\mu_\textrm{Discovery}=14$}
+\caption{DR-OffShell, $\mu_\textrm{discovery}=14$}
 \end{subfigure}
@@
 and (c) DR-OffShell scans at points close to their reported upper limits:
-$\mu_\textrm{Discovery}=3$, $\mu_\textrm{Discovery}=10$,
-and $\mu_\textrm{Discovery}=14$ respectively.
+$\mu_\textrm{discovery}=3$, $\mu_\textrm{discovery}=10$,
+and $\mu_\textrm{discovery}=14$ respectively.
 \\
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/2ljets_drhigh_plot_n0.pdf}
+\caption[
+Test statistic distribution for DR-High close to 3
+]{%
+Test statistic distribution for DR-High close to $3$.
+This zooms in to the narrow spike of toy samples that should have $n=0$
+observed data in the discovery region.
+Because the likelihood function for zero data,
+$\mathrm{Poisson}(0\mid x) = e^{-x}$, is $\log$-linear,
+this would analytically be a delta spike.
+It acquires this width from the finite precision of numerical optimization
+algorithms.
+The observed test statistic falling far to the right of that numerical noise
+reduces our statistical precision in estimating $\clb$ and $\clsb$.
+}
+\label{fig:2ljets_test_stat_distributions_drhigh_n0}
+\end{figure}
+
 The $n=0$ case has no visible width since its likelihood is linear
@@
 of that $n=0$ noise distribution.
-This would appear to be an anomalously large significance if $\clb$ were
-misinterpreted as a measure of compatibility with the background model.
+This small $\clb$ value could appear to be an anomalously large significance
+if that $p$-value were  misinterpreted as a measure of compatibility with the
+background model.
 Thankfully it is remedied by the $\cls$-prescribed ratio with $\clsb$,
@@
 $n=0$, only $57$ toy samples land to the right of the observed test statistic.
-And with $\sqrt{n}$ precision, that easily explains the difference between
-limits at $3.0$ and $2.93$ despite the huge sample size.
+And with $\approx1/\sqrt{57}$ relative statistical precision, that easily
+explains the difference between limits at $3.0$ and $2.93$ despite the huge,
+and likewise expensive, toy sample size.
+
+Although I have found no \atlas\ search reporting a discovery region limit
+below $3$, there are published examples with zero data but larger observed
+limits~\cite{SUSY-2018-36, atlas_susy_strong_0l}.
+The most extreme is $\mathrm{SR}3\ell_{[90,110]}$
+of~\cite{SUSY-2018-36}, which reports a background expectation of
+$1.08 \pm 0.18$ and expected limits of
+$S_{\mathrm{exp}}^{95} = 3.8^{+1.3}_{-0.5}$, but zero observed data with a
+whopping $S_{\mathrm{obs}}^{95} = 4.8$.
+Using their \pyhf\ models published through \hepdata\ I have failed to
+reproduce this effect outside of the \histfitter-\roofit ecosystem, but would
+not be surprised if similar complications with the test statistic distribution
+played a part.
+
+Not only is that limit more than $3$, it is more than the expected limit
+despite the fact that increasing data should increase the observed limit.
+The same effect has occurred in our results for DR-OffShell --- its observation
+of $21$ data is below the post-fit background is $22.1$, but its observed limit
+exceeds the central expected limit.
+Another public example of this effect is
+$\mathrm{SR}0\textrm{-}\mathrm{ZZ^{tight}_{bveto}}$ of~\cite{SUSY-2018-02}.
+\emph{Thanks to Tina~Potter discussions of this problem and for sharing
+its details of its modelling.}
+
+Internal discussions of this error led me to perform numerous validation
+studies that reproduced this finding with \pyhf\ and with various alternative
+constructions of the models.
+Various explanations were suggested in these discussions, including problems
+with asymmetric systematic variations or asymmetries in the Poisson
+distribution.
+Fundamentally, a sufficient explanation of this problem is our refusal to
+compare models with likelihood ratios.
+By instead using \emph{maximum} likelihood ratios, and modifications thereof,
+we guarantee that our tests can be suboptimal because they are not protected by
+the Neyman–Pearson lemma~\cite{neyman1933ix}.
+In this world of model comparison through $p$-values, sub-optimality means
+that the test statistic puts some data in the wrong
+order~\cite{cousins2008evaluation},
+and so by knowingly using sub-optimal test statistics we also know that the
+results need not behave as if the data had a sensible ordering.
+Limits moving in the wrong directions can simply be explained by flawed
+test statistics ranking data in a bad order.
+
+A second sufficient explanation for these paradoxical limit results is also
+that the toy sampling method does not assign legitimate $p$-values because its
+test statistic distributions are assigned by first fitting to the
+data~\cite{CMS-NOTE-2011-005}.
+Just as the Matrix Method is not predictive because it is conditioned on the
+data it attempts to estimate,
+these test statistic distributions are not predictive because they are
+conditioned on the data whose distribution they are supposed to estimate.
+Results in DR-Low are greatly impacted by this effect since its
+pre-fit background
+(fitted to control regions but not the discovery region)
+is $43 \pm 11$, but its post-fit background (from whose central value
+the background-only test statistic distribution is sampled) is
+$22 \pm 4$ as reported in Table~\ref{tab:2ljets_discovery},
+between which the data rate has almost halved.
+
+Test statistic distributions for DR-Low,
+which are shown in Figure~\ref{fig:2ljets_test_stat_distributions_drlow},
+show some other quirks: negative values and a bump at $-1$.
+The maximum likelihood ratio test statistics that we
+employ~\cite{Cowan:2010js} compare a background model against a strictly more
+flexible signal model, and should therefore, to my understanding, be always
+non-negative.
+The observed negative values are therefore quite unsettling, particularly since
+the observed test statistic landed in the bump at $-1$ for
+$\mu_\textrm{discovery} = 16$ in the upper limit scan.
+Plausible explanations could include rogue values of sampled alternative data
+(real data and auxiliary data that constrain model parameters), but given this
+result for the observed test statistic, I am inclined to blame failures in
+the numerical optimization.
+
+It is undoubtedly arguable that the pathological test statistic distributions
+observed in DR-Low indicate a bug that should be fixed.
+However, the statistical software packages that generated this result
+--- \roofit\ through \histfitter ---
+define in silicon the standard statistical methods for \atlas\ SUSY searches.
+To diverge from them would make our result incomparable.
+What they do is the definition of correct.
+All \pyhf\ can do is imitate.
+There is no thing as a bug.
+
+\section{Discussion}
+\begin{epigraphs}
+\qitem{%
+When you're building a video game world, you're building with ideas, and that
+can be like working with quick-set cement. You mold your ideas into a certain
+shape that can be played with, and in the process of playing with them they
+begin to harden and set until they're immutable like rock, and at that point
+you can't change the world --- not without breaking it into pieces and starting
+fresh with new ideas.%
+}%
+{Bennett~Foddy,
+\emph{Getting Over It with Bennett~Foddy},
+2017~\cite{foddy2017getting}}
+\end{epigraphs}
 
 \TODO{}
-% real world costs to pretending that our uncertainties are auxiliary data
-% not only idle Bayesian-versus-Frequentist philosophy
-% other examples - cite 4.6 limit.
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/2ljets_drhigh_plot_n0.pdf}
-\caption[
-Test statistic distribution for DR-High close to 3
-]{%
-Test statistic distribution for DR-High close to $3$.
-This zooms in to the narrow spike of toy samples that should have $n=0$
-observed data in the discovery region.
-Because the likelihood function for zero data,
-$\mathrm{Poisson}(0\mid x) = e^{-x}$, is $\log$-linear,
-this would analytically be a delta spike, but it acquires width from the finite
-precision of numerical optimization algorithms.
-The observed test statistic falling far to the right of that numerical noise
-reduces our statistical precision in estimating $\clb$ and $\clsb$.
-}
-\label{fig:2ljets_test_stat_distributions_drhigh_n0}
-\end{figure}

### 2022-08-08 main/chapter_2ljets.tex
sg checks

@@
 observed data does land in a spike at the end of the distribution that
-corresponds to to pseudo-data with $n=0$; the peaks to its left correspond
+corresponds to pseudo-data with $n=0$; the peaks to its left correspond
 to $n=1$, $n=2$ and so on, with widths due to the other parameters varied in
@@
 background model.
-Thankfully it is remedied by the $\cls$-prescribed ratio with $\clsb$,
+Thankfully, it is remedied by the $\cls$-prescribed ratio with $\clsb$,
 since both $p$-values suffer from similar effects.
@@
 Using their \pyhf\ models published through \hepdata\ I have failed to
-reproduce this effect outside of the \histfitter-\roofit ecosystem, but would
+reproduce this effect outside the \histfitter-\roofit ecosystem, but would
 not be surprised if similar complications with the test statistic distribution
@@
 
-Not only is that limit more than $3$, it is more than the expected limit
+Not only is that limit more than $3$, it is more than the expected limit,
 despite the fact that increasing data should increase the observed limit.
@@
 order~\cite{cousins2008evaluation},
-and so by knowingly using sub-optimal test statistics we also know that the
+and so by knowingly using suboptimal test statistics we also know that the
 results need not behave as if the data had a sensible ordering.
@@
 All \pyhf\ can do is imitate.
-There is no thing as a bug.
+There is no such thing as a bug.
 

### 2022-08-09 main/chapter_2ljets.tex
expand spacing on triple plots; rephrease is the defn of
improve figure layout; add note on histogram ranges

@@
 Section~\ref{sec:2ljets_results}.
-\TODO{Conclusions}
+\item A concluding discussion is given in
+Section~\ref{sec:2ljets_discussion}.
 \end{itemize}
@@
 \end{subfigure}
-\\[.5ex]
+\\[1.5ex]
 \begin{subfigure}{0.48\textwidth}
@@
 \end{subfigure}
-\\[.5ex]
+\\[1.5ex]
 \begin{subfigure}{0.46\textwidth}
@@
 
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/2ljets_drhigh_plot_n0.pdf}
-\caption[
-Test statistic distribution for DR-High close to 3
-]{%
-Test statistic distribution for DR-High close to $3$.
-This zooms in to the narrow spike of toy samples that should have $n=0$
-observed data in the discovery region.
-Because the likelihood function for zero data,
-$\mathrm{Poisson}(0\mid x) = e^{-x}$, is $\log$-linear,
-this would analytically be a delta spike.
-It acquires this width from the finite precision of numerical optimization
-algorithms.
-The observed test statistic falling far to the right of that numerical noise
-reduces our statistical precision in estimating $\clb$ and $\clsb$.
-}
-\label{fig:2ljets_test_stat_distributions_drhigh_n0}
-\end{figure}
-
 The $n=0$ case has no visible width since its likelihood is linear
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/2ljets_drhigh_plot_n0.pdf}
+\caption[
+Test statistic distribution for DR-High close to 3
+]{%
+Test statistic distribution for DR-High close to $3$.
+This zooms in to the narrow spike of toy samples that should have $n=0$
+observed data in the discovery region.
+Because the likelihood function for zero data,
+$\mathrm{Poisson}(0\mid x) = e^{-x}$, is $\log$-linear,
+this would analytically be a delta spike.
+It acquires this width from the finite precision of numerical optimization
+algorithms.
+The observed test statistic falling far to the right of that numerical noise
+reduces our statistical precision in estimating $\clb$ and $\clsb$.
+}
+\label{fig:2ljets_test_stat_distributions_drhigh_n0}
+\end{figure}
+
 Although I have found no \atlas\ search reporting a discovery region limit
@@
 the numerical optimization.
-
-It is undoubtedly arguable that the pathological test statistic distributions
+Histogram ranges in Figure~\ref{fig:2ljets_test_stat_distributions} are
+set from the smallest and largest toy test statistics, and their large extent
+indicates rare outliers that may also result from failed fits.
+
+It is certainly arguable that the pathological test statistic distributions
 observed in DR-Low indicate a bug that should be fixed.
@@
 To diverge from them would make our result incomparable.
-What they do is the definition of correct.
+What they do defines what is correct.
 All \pyhf\ can do is imitate.
@@
 \section{Discussion}
+\label{sec:2ljets_discussion}
 \begin{epigraphs}

### 2022-08-09 main/chapter_2ljets.tex
resolve todos for: asymptotic formulae ref, pdg rounding, claimed contributions, forward refs

@@
 \\[0.5em]
-Upper limits use the one-sided profile likelihood test statistic.
-The discovery $p$-value uses a profile likelihood test statistic in a one-sided test.
+Upper limits use the one-sided profile likelihood test
+statistic~\cite{Cowan:2010js}.
+The discovery $p$-value uses a profile likelihood test statistic in a one-sided
+test.
 All $p$-values are estimated by simulation of alternative data.
-\\
-\TODO{add labels and references to asymptotic formulae paper}
-\\
-\TODO{reference pdg rounding}
+Values with errors are rounded to one or two significant figures according
+the the PDG rounding scheme~\cite{pdg2020review}.
 }
@@
 Production of all electroweak signal samples and their systematic variations.
-\item \TODO{?Evaluation of Monte Carlo $Z$+jets uncertainties for RJR signal regions
-which indicated they exceeded $100\%$}
-\item \TODO{Various collaborations with group members on scripts to conduct
-tasks which were common between the searches.}
+\item Evaluation of $Z$+jets uncertainties for RJR signal regions through
+from simulations. These exceeded $100\%$ relative variation and supported the
+use of data-driven methods.
+\item Various collaborations with all other group members on common core tasks
+that are shared between the searches of the $\twoljets$ project.
 \item Integration of the electroweak analysis with \atlas\ combinations and
@@
 accompanied by more hadronic activity than other sources of $\ptmiss$.
+This effect in the $\metsig$ distribution is demonstrated in simulations in
+Figure~\ref{fig:2ljets_presel_met}.
 What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
@@
 and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
-is not reconstructed.
-\TODO{plot of \diboson\ decomposition here?}
-These effects are demonstrated from simulations in
-Figure~\ref{fig:2ljets_presel_met}.
+is not reconstructed; these subcomponents of the \diboson\ process are discussed
+later in Section~\ref{sec:2ljets_validation} and displayed in
+Figure~\ref{fig:2ljets_splits_\diboson}.
 
@@
 background-only model, and more for each signal model.
-Of those I evaluated more manually, some of the more interesting or important
-will be discussed in the remainder of this section ---
-\TODO{update this}
+Of those not evaluated automatically, some of the more interesting or important
+will be discussed in the remainder of this section.
 Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},

### 2022-08-09 main/chapter_2ljets.tex
comment on unique exclusion and other combinations results

@@
 
-Searches have been performed in \cms\ for similar
+The \cms\ Collaboration has published similar searches for
 C1N2~\cite{cms_susy_2022_c1n2} and GMSB models~\cite{
@@
 cms_susy_2022_gmsb
-}, with comparable results.
-
-\TODO{other atlas C1N2 and GMSB searches (combinations, 3L offshell exclusion)}
+}, with comparable exclusion limits in similar selections of data.
+Other published \atlas\ results in the full Run~2 dataset put constraints on
+the same models that we study in $\twoljets$.
+Since we designed these searches use orthogonal datasets, they may be
+combined for more sensitive future results.
+
+For the C1N2 model, these public Run~2 \atlas\ searches explore the
+all-hadronic~\cite{SUSY-2018-41},
+$2/3\ell$ compressed~\cite{atlas_susy_compressed_2l_2018_run2},
+$3\ell$ on-shell~\cite{SUSY-2018-06}, and
+$3\ell$/same-sign~\cite{SUSY-2018-09} channels.
+In comparison to these other searches, our exclusion contour of
+Figure~\ref{fig:2ljets_contours_c1n2} excludes a unique area of parameter
+space within $m(\chargino_1,\neutralino_2) \in 550\textrm{--}700\,\eV[G]$
+and $m(\neutralino_1) \in 200\textrm{--}300\,\eV[G]$.
+The all-hadronic search is most sensitive at the highest parent masses,
+and three-lepton data of in the on-shell and compressed searches are most
+sensitive at intermediate and small (off-shell/compressed) mass splittings.
+For the GMSB model, our result complements limits from the
+$4\ell$~\cite{SUSY-2018-02} and
+all-hadronic~\cite{SUSY-2018-41} searches.
+In addition to the exclusion contours that we present in
+Figure~\ref{fig:2ljets_contours_gmsb}, the all-hadronic search again dominates
+the heaviest parameter space, with $m(\neutralino_1) > 600\,\eV[G]$,
+and the $4\ell$ search has more sensitivity at
+$m(\neutralino_1) < 250\,\eV[G]$.
+This leaves the area that our $\twoljets$-electroweak search uniquely excludes
+to have with large Higgs branching fractions (up to $90\%$) at intermediate
+masses with $m(\neutralino_1) \in 250\textrm{--}600\,\eV[G]$.
 

### 2022-08-09 main/chapter_2ljets.tex
add dataset info; begin typing discussion

@@
 & Data
-& $A\epsilon\sigma_{\mathrm{obs}}^{95}~\mathrm{fb}$
+& $A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$
 & $S_{\mathrm{obs}}^{95}$
@@
 \label{sec:2ljets_events}
-\TODO{make sure we say 136ifb from 2015 to 2018 run 2 data. Good run lists etc}
-
 This section describes the events selected for the $\twoljets$ searches, the
@@
 
+Our dataset comprises $136.0\,\mathrm{fb}^{-1}$ ($\pm 1.7\%$) of integrated
+luminosity observed during normal operations of the \atlas\ detector in the
+years $2015$ through to $2018$ of the LHC Run~2~\cite{DAPR-2018-01}.
 Triggers and object definitions are shared between the three $\twoljets$
@@
 However, the statistical software packages that generated this result
---- \roofit\ through \histfitter ---
+--- \roofit\ through \histfitter\ ---
 define in silicon the standard statistical methods for \atlas\ SUSY searches.
@@
 
-\TODO{}
+From \atlas' large and valuable Run~2 dataset, our $\twoljets$ searches
+have extracted a sharper image of the high energy processes that simultaneously
+produce leptons, jets and missing transverse momentum in proton-proton
+collisions at the Large Hadron Collider.
+The core message is clear: we see no indication of any additive process
+atop the Standard Model.
+
+These searches do not exist in isolation. They extend and strengthen the
+results of numerous previous \atlas\ searches that have used similar or smaller
+datasets and found similar conclusions.
+By providing refined and updated interpretations, we have advanced \atlas' SUSY
+research programme in this $2\ell$+jets+$\met$ channel.
+My focus has been on the $\twoljets$-electroweak search within this project.
+Its result is a central pillar of the visible effects of both C1N2 and GMSB
+simplified supersymmetric models, and uniquely contributes exclusion
+(under the $95\%$ $\cls$ prescription) of certain regions of the models'
+parameter spaces.
+
+% model (in)dependent
+Although designed to study potential supersymmetric effects, the results
+themselves are open to wider interpretation.
+Serialized \heplikelihood\ models and region definitions that we publish
+through the paper~\cite{atlas2022searches} and its \hepdata\
+archive~\cite{hepdata.116034} are resources that remain useful for
+reinterpretation in future research testing different physical hypotheses
+--- both of the orthogonal multi-bin selections designed for model-dependent
+sensitivity and through our model-independent discovery regions.
+
+% limitations
+% overly ambitious design
+% lack: resources to study, methods to analyse
+% cost: delays, opportunity cost
+% higher quality result by simpler more robust design published sooner
+% accelerate science by doing it now, not in two years.

### 2022-08-09 main/chapter_2ljets.tex
add last of discussion. autocheck

@@
 $\pt^{\ell_2} > 25\,\eV[G]$,
-$\mll \in (71, 111)\,\eV[G]$,
-$\mjj \in (60, 110)\,\eV[G]$,
+$\mll \in 71\textrm{--}111\,\eV[G]$,
+$\mjj \in 60\textrm{--}110\,\eV[G]$,
 $\nbtag \leq 1$,
@@
 Tau decays form the next largest part, with prompt fakes due to misidentified
-photons or leptons in the sub-$\%$ weeds.
+photons or leptons in the sub-percent weeds.
 
@@
 but replaces its requirement of opposite-charge leptons with a requirement of
-same-charge leptons, and adds a requirement of $\mll \in (12, 111)~\eV[G]$
+same-charge leptons, and adds a requirement of $\mll \in 12\textrm{--}111\,\eV[G]$
 to cover both on- and off-shell selections.
@@
 prescription~\cite{atlas_twiki_jmr}.
-These apply equally to \afii\ and FullSim samples.
+These apply equally to both \afii\ and FullSim samples.
 
 \emph{This jet mass modelling was discussed with, and approved by, the \atlas\
-Jet/Etmiss conveners. I owe special thanks to Chris~Delitzsch for assistance
-with the official jet mass scale variations.}
+Jet/Etmiss conveners. Special thanks to Chris~Delitzsch for assistance
+preparing and sharing the official jet mass scale variations.}
 
@@
 \caption[
-Test statistic distribution for DR-High close to 3
+Test statistic distribution for DR-High close to $3$
 ]{%
@@
 whopping $S_{\mathrm{obs}}^{95} = 4.8$.
-Using their \pyhf\ models published through \hepdata\ I have failed to
-reproduce this effect outside the \histfitter-\roofit ecosystem, but would
-not be surprised if similar complications with the test statistic distribution
-played a part.
+Using their \pyhf\ models published through \hepdata, I have failed to
+reproduce this effect outside the \histfitter-\roofit system, finding instead
+a limit close to $3$. It would not surprise me if similar complications with
+the test statistic distribution played a part.
 
@@
 
-These searches do not exist in isolation. They extend and strengthen the
+These searches do not exist in isolation; they extend and strengthen the
 results of numerous previous \atlas\ searches that have used similar or smaller
@@
 % limitations
-% overly ambitious design
-% lack: resources to study, methods to analyse
-% cost: delays, opportunity cost
-% higher quality result by simpler more robust design published sooner
-% accelerate science by doing it now, not in two years.
+Such successes do not, however, deny that there is room for improvement.
+Although I brazenly philosophized in
+Section~\ref{sec:2ljets_origins_electroweak} that the limitations of dubious
+statistical practices should not prevent us from using large numbers of small
+signal regions, their overly ambitious design in the $\twoljets$-electroweak
+did come with costs:
+% mm fakes
+in replacing unusable fake/non-prompt assignments
+(Section~\ref{sec:2ljets_mm_fakes}),
+% precise alternative simulations
+in failures of the statistical precision of alternative simulated samples
+(Section~\ref{sec:2ljets_sample_comparisons})
+and struggles to reconcile their standard prescriptions for coupled
+(correlated) variations with physical intuition between our many bins,
+% computational
+and computational costs particularly in the `toy' simulations that assign our
+statistical results (Section~\ref{sec:2ljets_results}).
+The root of this problem that in diverging from standard designs we
+had to use tools and rules outside the domains that their designers
+intended.
+No design can be expected to work perfectly outside its expected context,
+and this divergence incurred us costs in either engineering workarounds or
+leaving dubious features in the final result.
+
+This thesis has supplemented many details of such features that are not
+explicitly declared in the main publication~\cite{atlas2022searches},
+and might otherwise have remained hidden in \atlas' internal documentation.
+
+Perhaps we \emph{should not} be restrained by crude statistical
+methods, but the fact is that we \emph{are} limited to immutable standard
+prescriptions.
+And by ignoring that truth, my designs cost this search project complications
+that caused real delays when problematic data
+--- substantial deficits ---
+were observed in our complicated and rigidly-modelled signal regions.
+Although these regions are sensitive, as demonstrated by the extent of their
+projected and observed exclusion contours, delays are themselves a meaningful
+cost.
+Simpler designs, with fewer and larger regions, could be cheaper to model with
+our limited simulation capacity and more robust to unexpected data, without
+sacrificing much sensitivity.
+Simplicity in design may therefore have afforded a higher quality result by
+reducing these delays and publishing its result years sooner,
+and so implemented \cern's goal of
+``accelerating science''~\cite{Radford:1551933}.
+
+% this extra line activates enlarged spacing for the final paragraph lol

### 2022-08-09 main/chapter_2ljets.tex
hugely; more singlespacing for epigraphs

@@
 explains the difference between limits at $3.0$ and $2.93$ despite the huge,
-and likewise expensive, toy sample size.
+and hugely expensive, toy sample size.
 
@@
 
+\begin{singlespacing}
 \section{Discussion}

### 2022-08-09 main/frontmatter.tex
fix linespacing
add quote sources
hugely; more singlespacing for epigraphs

@@
 \begin{preface}
-\epigraph{%
+\begin{singlespacing}
+\begin{epigraphs}
+\qitem{%
 Begin reading at Chapter N. Do \emph{not} read the quotations
@@
 }%
-{Donald E. Knuth~\cite{knuth1975art}}
+{Donald~E.~Knuth,
+\textit{The Art of Computer Programming: Volume~1},
+1975~\cite{knuth1975art}}
+\end{epigraphs}
+\end{singlespacing}
 
@@
 \noindent
-For this example, I'll just mention Chapter~\ref{chapter:intro}.
-and Chapter~\ref{chapter:intro}..
+For this example, I'll just mention Chapter~\ref{chapter:searches}.
+and Chapter~\ref{chapter:searches}..
 \end{preface}

### 2022-08-09 main/chapter_introduction.tex
add introduction chapter 0

\setcounter{chapter}{-1}
\begin{singlespacing}
\chapter{Introduction}
\label{chapter:introduction}
%
\begin{epigraphs}
\qitem{%
\TODO{}}%
{\TODO{},
\textit{\TODO{}},
\TODO{}}
\end{epigraphs}
\end{singlespacing}

Helloasdasdasd

### 2022-08-09 main/chapter_2ljets.tex
add example pointing to drlow test stat distribution
add missing sherpa backslash

@@
 
-Since \sherpa implements both the hard scatter and shower models, we are
+Since \sherpa\ implements both the hard scatter and shower models, we are
 considering variations due to the simultaneous change of both of these
@@
 results need not behave as if the data had a sensible ordering.
+For example, Figure~\ref{fig:2ljets_test_stat_distributions_drhigh} shows
+overlapping bumps from toys with two and three events; their overlap
+contains examples where our toy sampling scheme reverses the natural
+ordering that larger $n$ is more signal-like.
 Limits moving in the wrong directions can simply be explained by flawed

### 2022-08-09 main/chapter_2ljets.tex
add a missing roofit backslack

@@
 Using their \pyhf\ models published through \hepdata, I have failed to
-reproduce this effect outside the \histfitter-\roofit system, finding instead
+reproduce this effect outside the \histfitter-\roofit\ system, finding instead
 a limit close to $3$. It would not surprise me if similar complications with

### 2022-08-10 main/chapter_2ljets.tex
start 2ljets readthrough 1; crop rjr summary figures

@@
 
-My core contributions in this thesis are to the recent \atlas\
-searches ``for new phenomena in events with two leptons, jets, and
-missing transverse momentum''~\cite{atlas2022searches},
+My main messages that this thesis reports are my contributions to the recently
+published \atlas\ searches ``for new phenomena in events with two leptons,
+jets, and missing transverse momentum''~\cite{atlas2022searches},
 particularly to its electroweak part.
 For brevity, I will call this publication $\twoljets$ and its electroweak
-subset $\twoljets$-electroweak.
-
-This chapter follows a questionable order.
-It may be clearer, and is certainly more ordinary,
-to first present preparations (blind to data),
-followed later by the results (unblinded).
-Instead, this chapter first reports the unblinded results, then follows with
-in-depth discussions of individual elements, partially in light of those
+part $\twoljets$-electroweak.
+
+This chapter follows an unusual order.
+A more ordinary ordering would first present preparations made in the search
+project -- blind to data ---
+followed later by the results --- unblinded.
+Instead, this chapter first reports the unblinded results and follows those
+with in-depth discussions of individual elements, partially in light of those
 results.
-I choose this order for two reasons.
-Firstly, it avoids iterating over those elements twice, once before and again
-after unblinding.
-The analysis has many little and loosely related pieces that I attempt to
-to cover locally to reduce jumps between far-flung corners of the
-chapter.
-Secondly, and sadly, many design decisions can only be explained in light of the
-unblinded results, since important aspects were reworked in the
+This unconventional ordering makes sense for two reasons.
+Firstly, it avoids iterating over those elements twice, once before unblinding
+and again after;
+search project has many little and loosely-related pieces that I attempt to
+to cover locally to reduce the reader's need to jump between far-flung corners
+of the chapter.
+Secondly, and sadly, many design decisions can only be explained in light of
+the unblinded results, since major aspects of the search were reworked in the
 nineteen months between its unblinding and eventual publication.
 
-There are certain plots and tables which, personally, I first seek when
-meeting an \atlas\ search paper, particularly the summary plot with all signal
-regions and the model-dependent exclusion contours.
-This chapter begins by presenting those results in brief in Section
-~\ref{sec:2ljets_splash}.
+When first meeting an \atlas\ search paper,
+there are certain plots and tables that I, personally, prefer to find before
+delving into its written details.
+Specifically, these include the summary plot with all signal regions,
+and the model-dependent exclusion contours with their comparisons to previous
+results.
+This chapter begins by presenting those major plots and tables in brief in
+Section~\ref{sec:2ljets_splash}.
 The meat of the chapter follows.
@@
 they derive are described in Section~\ref{sec:2ljets_events}.
-\item Region design for the $\twoljets$-electroweak search is described and
+\item Region designs for the $\twoljets$-electroweak search are described and
 explained in Section~\ref{sec:2ljets_design}.
@@
 \item The unblinding procedure of this search is discussed in
-Section~\ref{sec:2ljets_unblinding}, and
-\item and some additional results are presented in
+Section~\ref{sec:2ljets_unblinding},
+\item Some additional details of our results are presented in
 Section~\ref{sec:2ljets_results}.
@@
 % begin with physics content and results
-Unblinded results in all main regions of the $\twoljets$-electroweak analysis
-are presented in of the Figure~\ref{fig:2ljets_summary}.
+The summary plot in Figure~\ref{fig:2ljets_summary} displays unblinded results
+in all main regions of the $\twoljets$-electroweak analysis.
 Since this figure appears to represent the data and background estimates,
@@
 thesis.
-There are, however, caveats to be developed later in this chapter.
-
+There are, however, caveats to be developed later in this chapter ---
+the backgrounds plotted are not exactly estimates, but are adapted and fitted
+to the observed data.
 In numerous orthogonal signal regions, no significant excess is observed
-above \emph{post-fit} backgrounds.
-Indeed, there are some significant deficits --- data below the background noise.
-
-Background parts of a \heplikelihood\ model are illustrated behind data.
-To test non-standard (supersymmetric) models against observation, signal
-contributions are added on top of those backgrounds.
-
+above the displayed post-fit backgrounds.
+Indeed, there are some significant deficits of data are below the assigned
+backgrounds.
+
+% summary plot
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
+\caption[
+Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
+backgrounds
+]{%
+Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
+backgrounds~\cite{atlas2022searches}.
+The lower panel shows $S_\mathrm{ATLAS}$ from
+Equation~\ref{eqn:significance_atlas}.
+Control, validation and signal regions are shown from left to right, with the
+regions within each category ordered approximately by their typical $\met$.
+Validation region likelihoods do not directly constrain the fit.
+The ``Top'' category contains $\ttbar$ and $tW$ processes, and
+``Other'' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and
+\topother\ processes.%
+}
+\label{fig:2ljets_summary}
+\end{figure}
+
+To test non-standard (supersymmetric) models against observation
+we add signal contributions on top of the modelled backgrounds.
 We use two signal models in the $\twoljets$-electroweak analysis.
-One is an MSSM chargino-neutralino model (C1N2) which acts to produce
-$\chargino_1\textrm{--}\neutralino_2$ pairs, which decay via
-$W\rightarrow q\bar q$ and
-$Z\rightarrow \ell^\pm \ell^\mp$ respectively.
+One is an MSSM chargino-neutralino model (C1N2) that acts to produce
+$\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
+that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
+respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
 The other is a GMSB model which, from a variety of initial states, produces
-$\neutralino_1\textrm{--}\neutralino_1$ pairs.
-These neutralinos decay,
-via Higgs and $Z$ bosons in variable branching fractions,
-to final states again containing $\ell^\pm \ell^\mp$ and jet-seeding partons.
+$\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
+bosons, in variable branching fractions,
+to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
+and invisible gravitinos.
 Diagrams illustrating these C1N2 and GMSB signal models are shown in
 Figure~\ref{fig:2ljets_signal_diagrams}.
-
-Effects of some of these signals in example signal regions of the
-$\twoljets$-electroweak analysis are displayed in
+Example effects of these signal models are displayed in signal regions of the
+$\twoljets$-electroweak search are displayed in
 Figure~\ref{fig:2ljets_signal_examples}.
-
+All of these regions require two leptons, jets, and
+missing transverse momentum, and the signal expectations would be added on top
+of the backgrounds if Nature realized these supersymmetric models.
+
+% signal diagrams
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Supersymmetric signal processes in the $\twoljets$-electroweak analysis
+]{%
+Supersymmetric signal processes in the $\twoljets$-electroweak analysis.
+\emph{Figures are reproduced from the paper%
+}~\cite{atlas2022searches, atlas_susy_feynman}.
+\\[0.5em]
+(a) C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
+is produced through an $s$-channel $W^{\pm}$ resonance and the masses of
+weak bosons in decays are bounded by the mass splitting
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
+We explore the parameters
+$m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
+\\[0.5em]
+(b) GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
+is produced by soft decays from pairs including $\chargino_1$,
+$\neutralino_2$ or $\neutralino_1$.
+Although the Higgs and $Z$ bosons can decay to many states, we target events
+with $Z\rightarrow \ell\ell$ and $h/Z\rightarrow bb/jj$.
+We explore the parameters
+$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with a tiny
+fixed mass splitting of
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1\,\eV[G]$.%
+}
+\label{fig:2ljets_signal_diagrams}
+\end{figure}
+
+% signal region plots
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
+\caption{SR-High-8}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
+\caption{\srllbb}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
+\caption{SR-Int}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
+\caption{SR-Low}
+\end{subfigure}
+\caption[
+Example signal region histograms with benchmark signal sample yields overlaid
+]{%
+Example signal region histograms with benchmark signal sample yields overlaid
+as dotted lines~\cite{atlas2022searches}.
+Physically, signal yields would add on top of the backgrounds.
+Unblinded data are shown --- regions in the top two plots observe zero data
+and the error bars for zero events are hidden for aesthetic reasons.%
+}
+\label{fig:2ljets_signal_examples}
+\end{figure}
+
+Both signal models have two free parameters that our search varies.
 At each parameter point of a signal model, it is added to the background model
@@
 GMSB in~\ref{fig:2ljets_contours_gmsb}.
-
-Data in discovery regions are presented and interpreted in
-Table~\ref{tab:2ljets_discovery}.
-
-% summary plot
-\begin{figure}[tp]
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
-\caption[
-Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
-backgrounds
-]{%
-Data of the $\twoljets$-electroweak analysis with \emph{post-fit}
-backgrounds~\cite{atlas2022searches}.
-The lower panel shows $S_\mathrm{ATLAS}$ from
-Equation~\ref{eqn:significance_atlas}.
-Control, validation and signal regions are shown from left to right, with the
-regions within each category ordered approximately by their typical $\met$.
-Likelihoods from validation regions are not included in the fit.
-The `Top' category contains $\ttbar$ and $tW$ processes, and
-`Other' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and other
-\topother\ processes.%
-}
-\label{fig:2ljets_summary}
-\end{figure}
-
-% signal diagrams
-\begin{figure}[t]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_c1n2_llqqn1n1_wz.pdf}
-\caption{C1N2}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_n1n1_hhggzz.pdf}
-\caption{GMSB}
-\end{subfigure}
-\caption[
-Supersymmetric signal processes in the $\twoljets$-electroweak analysis
-]{%
-Supersymmetric signal processes in the $\twoljets$-electroweak analysis.
-\emph{Figures are reproduced from the paper%
-}~\cite{atlas2022searches, atlas_susy_feynman}.
-\\[0.5em]
-(a): C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
-is produced through a $s$-channel $W^{\pm}$ resonance and the masses of
-weak bosons in decays are bounded by the mass splitting
-$m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
-We explore the parameters
-$m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
-\\[0.5em]
-(b): GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
-is produced by soft decays from pairs including $\chargino_1$,
-$\neutralino_2$ or $\neutralino_1$.
-Although the $h$ and $Z$ bosons decay to many states, we target
-$Z\rightarrow \ell\ell$ with
-$h/Z\rightarrow bb/jj$.
-We explore the parameters
-$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with fixed
-$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = 1\,\eV[G]$.
-}
-\label{fig:2ljets_signal_diagrams}
-\end{figure}
-
-% signal region plots
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
-\caption{SR-High-8}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
-\caption{\srllbb}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
-\caption{SR-Int}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
-\caption{SR-Low}
-\end{subfigure}
-\caption[
-Example signal region histograms with benchmark signal sample yields overlaid
-]{%
-Example signal region histograms with benchmark signal sample yields overlaid
-as dotted lines~\cite{atlas2022searches}.
-Physically, signal yields would add on top of the backgrounds.
-Data are shown; regions in the top two plots observe zero data, and
-Poisson error bars for zero events are hidden for aesthetic reasons.
-}
-\label{fig:2ljets_signal_examples}
-\end{figure}
+In both of these figures, space to the lower left of the contours is
+labelled as excluded, and space on the other side of the observed limit contour
+is not excluded.
+The observed contours are somewhat larger than expected due to observed
+deficits in data.
 
@@
 \caption[
-Contours for the C1N2 model in the $\twoljets$-electroweak analysis
+Contours for the C1N2 model in the $\twoljets$-electroweak search
 ]{%
 Contours for the C1N2 model in the $\twoljets$-electroweak
-analysis~\cite{atlas2022searches}.
+search~\cite{atlas2022searches}.
 \\[0.5em]
 Space below the solid red line is labelled as excluded, and its dotted
-neighbours show the result if all signal cross-sections are varied up and down
-by theoretical error bars.
+neighbours show alternative results that are obtained if all signal
+cross-sections are varied up and down by theoretical error bars.
 \\[0.5em]
 The yellow band shows the $\pm1$-sigma region of exclusion contours
-from asymptotic approximations to the prior distribution of the test statistic.
-Grey areas are observed limits from the two-lepton parts
-of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
+from asymptotic approximations to the distribution of the test statistics.
+Grey areas are observed limits from the two-lepton results from previous
+\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
 Exclusion is defined by the $95\%$ $\cls$ prescription
-in asymptotic approximations.
+with asymptotic approximations~\cite{Cowan:2010js}.
 All contours are interpolated from a sparse grid.
@@
 \caption[
-Contours for the GMSB model in the $\twoljets$-electroweak analysis
+Contours for the GMSB model in the $\twoljets$-electroweak search
 ]{%
 Contours for the GMSB model in the $\twoljets$-electroweak
-analysis~\cite{atlas2022searches}.
+search~\cite{atlas2022searches}.
 \\[0.5em]
 Space below the solid red line is labelled as excluded, and its dotted
-neighbours show the result if all signal cross-sections are varied up and down
-by theoretical error bars.
+neighbours show alternative results that are obtained if all signal
+cross-sections are varied up and down by theoretical error bars.
 \\[0.5em]
 The yellow band shows the $\pm1$-sigma region of exclusion contours
-from asymptotic approximations to the prior distribution of the test statistic.
+from asymptotic approximations to the distribution of the test statistics.
+Grey areas are observed limits from the two-lepton results from previous
+\atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.
 Exclusion is defined by the $95\%$ $\cls$ prescription
-in asymptotic approximations.
+with asymptotic approximations~\cite{Cowan:2010js}.
 All contours are interpolated from a sparse grid.
@@
 \end{figure}
+
+Signal regions of the $\twoljets$-electroweak analysis are combined to form
+five discovery regions.
+Results of their model-independent interpretations are presented in
+Table~\ref{tab:2ljets_discovery}.
+The same data deficits drive stronger bounds on possible signals that manifest
+as smaller signal expectations limits $S_{\mathrm{obs}}^{95}$.
 
@@
 regions~\cite{atlas2022searches}.
-The fitted yield is in the background-only model constrained by the data in each region.
+Each fitted yield is in the background-only model constrained by the data in
+all control regions and the discovery region.
 Limits are intended to reflect constrains on additive signal contributions.
@@
 test.
-All $p$-values are estimated by simulation of alternative data.
+All $p$-values are estimated by ``toy'' simulations of alternative data.
 Values with errors are rounded to one or two significant figures according
-the the PDG rounding scheme~\cite{pdg2020review}.
+the the PDG rounding scheme~\cite{pdg2020review}.%
 }
@@
 \subsection{Contributions}
-In this subsection, I claim attributions of work on the $\twoljets$ analysis
-and related \atlas\ efforts.
-
+Many people contributed directly and indirectly to the $\twoljets$ searches.
+I claim personal responsibility for the following parts of the $\twoljets$
+search and related \atlas\ efforts.
 \begin{itemize}
-\item Design, implementation, and execution of the electroweak part of the analysis.
-\item Service as `analysis contact' from January to October 2020.
-\item Production of main data inputs for all three $\twoljets$ analyses
+\item Design, implementation, and execution of the electroweak part of the
+analysis.
+\item Service as Analysis Contact (a management role) for the $\twoljets$
+project from January to October 2020.
+\item Production of main data inputs for all three $\twoljets$ searches
 and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
-Production of all systematic variations for the background samples;
+Production of all experimental systematic variations for the background
+samples;
 other group members assisted with the central background sample.
@@
 paper~\cite{atlas2022searches} and the public HEPData
-record~\cite{maguire2017hepdata}.
+archive~\cite{maguire2017hepdata}.
 \end{itemize}
@@
 Many plotting scripts, however, are modified from versions inherited from
-many \atlas\ members past and present, to whom I am grateful.
+many past and present \atlas\ members to whom I am grateful.
 
@@
 two leptons, jets, and missing transverse momentum.
-Quite a few analyses, in fact.
-This project exists to supplement those results with information from the full
-LHC Run~2 data-set, which was completed in 2018 and contains
+Quite a few analyses, in fact ---
+this search project exists to supplement those results with information from
+the full LHC Run~2 data-set that was collected between in the years $2015$
+through to $2018$ and contains
 $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
 $\sqrt{s} = 13\,\eV[T]$.
-Leverage of the increased data quantity is one key aim.
-Analysis quality also might be improved if our designs can take lessons from
+Leveraging the increased data quantity is one key contribution of $\twoljets$.
+Analysis quality also might be improved if our designs take lessons from
 the experiences of previous work.
 
-% electroweak
-% Run 1 data~\cite{atlas_2l_SUSY_2013_11},
-% partial Run~2 data~\cite{atlas_23l_SUSY_2016_24},
-% atlas_rjr_23l_SUSY_2017_03,
-% atlas_rjr_mimic_SUSY_2018_06
-The $\twoljets$ analysis is a trinity,
-of which all three parts update previously published \atlas\ results.
-These three parts are named `RJR', `electroweak', and `strong', each of which
-is described in its historical context in this section.
+The $\twoljets$ analysis is a trinity in which all three parts update
+previously published \atlas\ results.
+These three parts are named `RJR', `electroweak', and `strong',
+and this section describes each in its historical context.
 
 Recursive Jigsaw Reconstruction (RJR) is an algorithm for constructing useful
-event variables.
+event variables that target given decay trees.
 The $\twoljets$-RJR analysis, detailed in Section~\ref{sec:2ljets_origins_rjr},
@@
 detailed in Section~\ref{sec:2ljets_origins_strong} and
-searches for effects from the strong sector.
-It also uses conventional event variables, and focuses on features in the
-distribution of the invariant masses of lepton pairs.
-
-My primary contributions are to the $\twoljets$-electroweak analysis,
-detailed in Section~\ref{sec:2ljets_origins_electroweak},
-which uses conventional (non-RJR) event variables to define a collection of
-orthogonal regions, and performs a search for effects from the electroweak
-sectors of supersymmetric models.
+searches for effects from supersymmetric models that are active in the strong
+sector.
+It uses conventional (non-RJR) event variables and focuses on features in the
+distribution of the invariant masses of lepton pairs that can have distinct
+features that depend on the masses of decaying supersymmetric resonances.
+
+My primary contributions are to the $\twoljets$-electroweak search,
+detailed in Section~\ref{sec:2ljets_origins_electroweak}, that seeks effects
+from supersymmetric models that are active in the electroweak
+sector by using conventional event variables to define a collection of
+orthogonal selection regions.
 
@@
 (left) A fully-resolved decay tree with its particles labelled.
-(right) Only the $Z\rightarrow\ell\ell$ \underline{V}ector boson)
+(right) Only the $Z\rightarrow\ell\ell~$ \underline{V}ector boson
 is resolved, but the supersymmetric system recoils off
-hard QCD jets labelled ``ISR'', boosting the \underline{I}nvisible particles,
-to generate visibly large $\met$ in models with small mass-splittings.
-The same approach is used for SR-OffShell selections in the electroweak
-analysis.
+hard QCD jets labelled `ISR' that boost the \underline{I}nvisible system,
+to generate visibly large $\met$ in models that would otherwise not,
+due to their small mass-splittings.
+A similar concept is approached by the SR-OffShell selections in the
+$\twoljets$-electroweak search.
 }
@@
 The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
-particle data by reconciling them with user-specified decay trees, and works
+particle data by reconciling them with user-specified decay trees, and operates
 by making decisions of how to assign four-momenta to different `systems',
-which are nodes of the chosen the decay graph.
-Event variables are then evaluated in those systems' rest
+which are represented as nodes on the chosen decay graph.
+Various RJR event variables are then defined in those systems' rest
 frames~\cite{jackson2017sparticles, jackson2017rjr}.
 
-In this way, RJR variables get some intelligence in how parent particles are
+In this way, RJR variables have some intelligence in how parent particles are
 reconstructed from their visible decay products, as well as approximate
 independence from uninteresting boosts, particularly those from the emission of
-QCD jets outside the supersymmetric decay tree.
+QCD jets outside the main supersymmetric decay tree.
 
@@
 }.
-They have also influenced the designs of other analyses which do not use
-RJR variables themselves~\cite{atlas_rjr_mimic_SUSY_2018_06}.
-
-Two decay trees are used in the $\twoljets$-RJR analysis, with signal region
-each.
-These trees target C1N2-like scenarios similar to those also considered in the
-$\twoljets$-electroweak search,
-and illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees}.
+They have also influenced the designs of other analyses that do not themselves
+use RJR variables~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+
+Two decay trees are used in the $\twoljets$-RJR analysis, and from each
+one signal region is defined and explored.
+These trees are illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees} and
+target C1N2-like scenarios similar to those also targetted by the
+$\twoljets$-electroweak search.
 
 \subsubsection{Former excesses}
-Excess data were observed in several signal regions of the partial Run~2
+Excess data were observed in several signal regions of the past partial Run~2
 search ``for chargino-neutralino production using recursive
@@
 $\sqrt{s} = 13\,\eV[T]$.
-That search reported notable excesses in four signal regions.
-Of these, two required two charged leptons
+That search reported notable excesses in four signal regions, of which
+two required two charged leptons
 ($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
@@
 ($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).
-
+These excesses are visible in their summary plots that we reproduce in
+Figures~\ref{fig:2ljets_rjr_summaries_2017_2l}
+and~\ref{fig:2ljets_rjr_summaries_2017_3l}.
 Although the largest local excess was reported as ``$3.0$ standard deviations'',
@@
 
-Both $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$
+Both of these three-lepton regions with past excesses,
+$\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$,
 have been revisited with the full Run~2 data-set and updated background
 modelling;
-the update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
+their update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
 These regions have also been approximated without direct use of RJR event
 variables;
-those approximations also find data ``in agreement'' with the background
-model~\cite{atlas_rjr_mimic_SUSY_2018_06}.
-
-The $\twoljets$-RJR serves to analysis reproduce
+those approximations also find data ``in agreement'' with background
+modelling~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+
+The $\twoljets$-RJR search serves to reproduce
 $\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$,
@@
 They do not.
-
-Summary plots for $\twoljets$-RJR and the former 2 and 3 lepton analysis,
-including these signal regions with former excesses,
-are displayed in Figure~\ref{fig:2ljets_rjr_summaries}.
+As seen in Figure~\ref{fig:2ljets_rjr_summaries_2ljets},
+neither reproduced region shows any significant deviation of data from its
+fitted background.
 
@@
 \caption{$2\ell$ RJR~\cite{atlas_rjr_23l_SUSY_2017_03}}
+\label{fig:2ljets_rjr_summaries_2017_2l}
 \end{subfigure}
@@
 \caption{$3\ell$ RJR~\cite{atlas_rjr_23l_SUSY_2017_03}}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.8\textwidth}
+\label{fig:2ljets_rjr_summaries_2017_3l}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.9\textwidth}
 \centering
@@
 \caption{$\twoljets$-RJR~\cite{atlas2022searches}}
-\end{subfigure}
-\caption[
-Summary plots for RJR analyses
-]{%
-Summary plots for RJR analyses:
+\label{fig:2ljets_rjr_summaries_2ljets}
+\end{subfigure}
+\caption[
+Summary plots for RJR searches
+]{%
+Summary plots for RJR searches:
 (a) and (b)~two- and three- lepton selections respectively,
@@
 Disappearing excesses have encouraged some doubts about RJR variables in
-certain social circles;
-the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06}
+certain social circles.
+The existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06},
+which approximates the $3$ lepton result of~\cite{atlas_rjr_23l_SUSY_2017_03}
+with non-RJR ``laboratory-frame'' event variables,
 is public evidence of this.
@@
 A more precisely framed problem with RJR variables is that they tend to select
-phase space, which is poorly modelled by simulations.
+regions of phase space that are poorly modelled by simulations.
 This is plausibly explained by the fact that RJR variables are non-trivially
 related to conventional, lab-frame variables.
-Simulations have been tuned to model conventional variables well, and
-`systematic' variations are evaluated and derived in conventional variables.
-Modelling these features accurately does not imply that all other functions of
-the data are well understood.
-These effects cause Monte Carlo simulations to give predictions in RJR regions
-which are inaccurate and/or carry large uncertainties.
+All of our simulations have been tuned to model conventional variables well,
+and systematic variations are also evaluated and derived in those conventional
+variables.
+Unfortunately, modelling the distributions of standard event variables
+accurately does not necessarily imply that all other functions of the data are
+well modelled.
+These effects might contribute to the fact that our Monte Carlo simulations
+give predictions in RJR regions that are inaccurate and/or carry large
+uncertainties.
 
@@
 or trust.
-Indeed, I argue that all `data-driven' methods rely not only on data but on
-hypothetical extrapolations which may not be well motivated, particularly when
-studying complicated features such as RJR variables.
-For these reasons, I choose to not use any RJR event variables in the
-$\twoljets$-electroweak analysis.
+Indeed,  all `data-driven' methods rely not only on data but on
+hypothetical extrapolations that may not be well motivated, particularly when
+studying complicated non-linear event variables such as RJR variables.
+Although RJR variables can appear to offer appealing separations of
+supersymmetric signals from Standard Model backgrounds,
+these practical issues motivate my decision to use any RJR event variables in
+the $\twoljets$-electroweak analysis.
 
@@
 Supersymmetric signal processes in the $\twoljets$-strong search.
-\emph{Figures reproduced from the paper%
+\emph{These figures are reproduced from the paper%
 }~\cite{atlas2022searches, atlas_susy_feynman}.
@@
 $\neutralino_2 \ldots \slepton \ldots \neutralino_1$
-decay chain~\cite{paige1996determining}:
+decay chain~\cite{paige1996determining}
 \begin{equation}
 \label{eqn:2ljets_strong_max_mll}
-\max \mll
+\max(\mll)
 = m(\neutralino_2)
@@
 and the signal model is simulated with fast simulation.
-\\
+\\[0.5em]
 Yes, this really is a screenshot of a ROOT TBrowser~\cite{ROOT}, with no
@@
 simulation.
-\\
+\\[0.5em]
 This selection requires one signal lepton (electron or muon) with
@@
 crosses $0.05$.
-\\
+\\[0.5em]
 Coloured backgrounds illustrate what would have been observed with data at
@@
 Figure~\ref{fig:2ljets_test_stat_distributions}.
-\\
+\\[0.5em]
 For visual effect, the test statistic for $\mu_\textrm{discovery}=16$
@@
 This aesthetic change does not affect the results.
-\\
+\\[0.5em]
 \emph{These figures were produced with \histfitter\ tools through
@@
 and $\mu_\textrm{discovery}=14$ respectively.
-\\
+\\[0.5em]
 Axis limits are set from the minimum and maximum test statistics recorded.
@@
 observation would indicate their likelihood ratio.
-\\
+\\[0.5em]
 Results are:
@@
 and (c) $\clb = 0.68$ and $\clsb = 0.037$.
-\\
+\\[0.5em]
 \emph{These figures were adapted from plots produced with \histfitter\ tools

### 2022-08-10 main/chapter_2ljets.tex
first proofread of summaires section; add linking paragraph

@@
 LHC.
-Squarks and gluinos are therefore excluded with masses up to several
-$\eV[T]$ using hadronic data alone, depending on which simplified model is
-considered~\cite{atlas_susy_strong_0l}.
+Squarks and gluinos in simplified models are therefore excluded with masses up
+to several $\eV[T]$ using hadronic data alone~\cite{atlas_susy_strong_0l}.
 
 The $\twoljets$-strong search targets strongly interacting supersymmetric
-effects in cases where their decays produce pairs of charged leptons.
-Like the RJR and electroweak $\twoljets$ searches, the $\twoljets$-strong
-signals produce jets and invisible lightest supersymmetric particles.
-Unlike the RJR and electroweak parts, not all of these leptons are produced
-by $Z$ boson decays.
-As illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
+effects in cases where their decays include pairs of charged leptons.
+Like the RJR and electroweak $\twoljets$ searches, signal models targeted by
+the $\twoljets$-strong search produce jets and invisible LSPs.
+Unlike the RJR and electroweak parts, however, not all of these leptons are
+produced by $Z$ boson decays;
+as illustrated in Figure~\ref{fig:2ljets_strong_signal_diagrams}, decay trees
 featuring sleptons can also produce lepton pairs.
@@
 
-% signal models
 \begin{figure}[tp]
@@
 \caption[
-Supersymmetric signal processes in the $\twoljets$-strong search
-]{%
-Supersymmetric signal processes in the $\twoljets$-strong search.
+Diagrams for supersymmetric signal models targeted by the $\twoljets$-strong
+search
+]{%
+Diagrams for supersymmetric signal models targeted by the $\twoljets$-strong
+search.
 \emph{These figures are reproduced from the paper%
@@
 \\[0.5em]
-All models produce same-flavour, opposite-sign lepton pairs along with jets
+All three produce same-flavour, opposite-sign lepton pairs along with jets
 and $\met$.
@@
 The gluino-$Z^{(*)}$ and squark-$Z^{(*)}$ models decay via $Z$ boson
-resonances, which may be forced off-shell for small mass-splittings
+resonances that may be forced off-shell for small mass-splittings
 $m(\neutralino_2) - m(\neutralino_1) < m(Z)$.
@@
 
-Rather than peaking at a resonance mass, these pairs have an $\mll$
-distribution with a signature end-point which depends on the masses of other
-supersymmetric particles in the
+Rather than peaking at a resonance mass, the lepton pairs in the
+`gluino-slepton' signal model have an $\mll$ distribution that features a
+signature end-point that depends on the masses of other supersymmetric
+particles in the
 $\neutralino_2 \ldots \slepton \ldots \neutralino_1$
-decay chain~\cite{paige1996determining}
+decay chain~\cite{paige1996determining}:
 \begin{equation}
@@
 
-Various previous incarnations of this search have been conducted with similar
+Previous incarnations of this search have been conducted with similar
 selections and target models on in older data, including partial Run~2~\cite{
@@
 
-% exclusion plots
+To target its signal models, the $\twoljets$-strong search defines seven
+overlapping signal regions.
+Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
+and the others are binned in $\mll$ for sensitivity to the various
+kinematic edges across the grid of signal model parameters.
+
+The largest Standard Model background to this search is dileptonic $\ttbar$,
+partly because no $b$-tagging veto is applied.
+All three signal models produce same-flavour lepton pairs, but $\ttbar$
+produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
+equal rates ---
+the $\twoljets$ search exploits this difference in flavour symmetry to derive
+a data-driven background estimate for all of its flavour-symmetric backgrounds,
+and have its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
+The background estimate works by taking different-flavour data, weighting them
+by factors to calibrate for differences in the leptons' reconstruction, and
+histogramming them in same-flavour regions.
+Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
+$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.
+
+All post-fit results in $\twoljets$-strong regions match the data to within
+$1$-sigma.
+
+Exclusion contours from the $\twoljets$-strong search are reproduced in
+Figure~\ref{fig:2ljets_strong_contours} and show a substantially increased
+area of exclusion, improving by around $400\,\eV[G]$ in the largest limits on
+$m(\tilde g)$ over the partial Run~2
+results~\cite{atlas_susy_strong_2l_partial_run2}.
+However, the jagged edges along these contour are striking.
+This jaggedness arises from the method employed to combine exclusion contours
+derived from individual signal regions.
+Because they overlap, they are not independent and cannot be trivially
+combined.
+Instead, each model in the signal grid is assigned to the signal region
+with the ``best expected sensitivity'' by some
+measure~\cite{atlas2022searches}.
+This combination method constructs an exclusion contour for each signal region,
+then joins those contours such that models assigned to each region use its
+corresponding contour.
+
 \begin{figure}[tp]
@@
 Contours for signal models of the $\twoljets$-strong search.
-\emph{Figures were produced by the strong team%
+\emph{These figures were produced by the strong team%
 }~\cite{atlas2022searches}.
@@
 from~\cite{atlas_susy_strong_2l_partial_run2}.
+No previous contour is available for the exact squark-$Z^{(*)}$ model studied
+here.
 }
@@
 
-To target its signal models, $\twoljets$-strong search defines seven
-overlapping signal regions.
-Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
-and the others are binned in $\mll$ to give broad sensitivity to the various
-kinematic edges across the grid of signal model parameters.
-
-The largest Standard Model background to this search is dileptonic $\ttbar$,
-in part because no $b$-jet veto is used.
-All three signal models produce same-flavour lepton pairs, but $\ttbar$
-produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
-equal rates.
-This difference in flavour symmetry is exploited to estimate backgrounds in
-$\twoljets$-strong,
-and its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
-The estimate works by taking different-flavour data, weighting them by factors
-to calibrate for differences in the leptons' reconstruction, and histogramming
-them in same-flavour regions.
-Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
-$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.
-
-Post-fit results match the data to within $1$-sigma in all $\twoljets$-strong
-regions.
-
-Exclusion contours from the $\twoljets$-strong search are reproduced in
-Figure~\ref{fig:2ljets_strong_contours}.
-Besides their increased reach, the jagged edges to parts of these contour are
-striking.
-These arise from the combination of results from overlapping signal regions.
-Because they overlap, they are not independent and cannot be trivially
-combined.
-Instead, each model in the signal grid is assigned to the signal region
-with the ``best expected sensitivity'' by some measure.
-An exclusion contour is constructed for each signal region, then those contours
-are joined such that models assigned to each region use its corresponding
-contour.
-
-Personally, I find this hard to interpret.
-If I had a pet theory which landed near to one of these sharp
+Personally, I find these jagged contours hard to interpret.
+If I had a pet theory that landed on one of these sharp
 region-transition edges, what should I make of these data?
-Could some combination of the regions not have combined their information to
-produce more precise contours?
-Since model predictions change continuously, should our inferences not also?
-To aid interpretations and to allow combinations of regions'
-information, I chose to design the $\twoljets$-electroweak with only orthogonal
-regions, and without any manual switching of regions between models.
+I wouldn't know.
+It seems certain that some combination of the regions could have combined their
+information to produce a continuous and more informative contour.
+Where model predictions change continuously in these signal spaces, it seems
+that our inferences should also.
+To aid interpretations and to allow justifiable combinations of regions'
+information, I therefore chose to design the $\twoljets$-electroweak with only
+orthogonal regions, and without any switching of regions between models.
 
@@
 simulation-driven backgrounds and orthogonal multi-bin regions.
-This is because simulation-driven estimates are interpretable as
-Standard Model predictions,
-and orthogonal regions have independent likelihoods which can be cleanly
-combined to leverage the information in their data.
-
-The main idea behind the design of the $\twoljets$-electroweak analysis is to
+Simulation-driven estimates are interpretable as Standard Model predictions,
+relying less on loosely justified extrapolations,
+and orthogonal regions have independent likelihood functions that can be
+cleanly combined, by multiplication, to leverage the joint information in their
+data.
+
+The $\twoljets$-electroweak search aims to
 update the regions of the partial Run~2 search in its $\twoljets$
-channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for signal models
-producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
-A particular goal was to improve sensitivity by replacing their overlapping
+channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for the same C1N2
+signal models that produce chargino-neutralino pairs decaying via $W$ and $Z$
+bosons.
+A particular design goal was to improve sensitivity by replacing their overlapping
 `SR2-int' and `SR2-high' regions with orthogonal selections.
@@
 
-As in the $\twoljets$-strong search, the partial Run~2 search chose one
-signal region from an expected sensitivity for each signal point,
-as displayed in
-Figure~\ref{fig:2ljets_23l_stuff}
-with the two regions and exclusion contours.
+Similarly to the $\twoljets$-strong search, the partial Run~2
+search~\cite{atlas_23l_SUSY_2016_24} chose one signal region for each signal
+point from a measure expected sensitivity measure and drew contours around
+results from the chosen fit results.
+The two signal regions SR2-int and SR2-high differ only by a selection of
+$\met > 250\,\eV[G]$ and are displayed in Figure~\ref{fig:2ljets_23l_stuff},
+along with the combined exclusion contour (additionally including an `SR2-low'
+region and a three-lepton result) and the map of which region is chosen at
+which signal point.
 
@@
 illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
-The C1N2 and GMSB models behave similarly since both feature decays via massive
-bosons ($W\kern-0.4exZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay
-trees to pairs of invisible supersymmetric particles.
-Although widening sensitivity to both models does sacrifice sensitivity to
-either one, the increased degree of model independence is itself of value for
-future reinterpretations of the data.
+The C1N2 and GMSB models behave similarly since both feature gaugino decays via
+massive bosons ($W\kern-0.4exZ$, $ZZ$ or $Zh$) to leptons and jets, in
+similarly structures decay trees to invisible supersymmetric pairs.
+Although widening sensitivity to both models must sacrifice some sensitivity to
+each individually, the increased degree of model independence is itself of
+value for future reinterpretations of our data.
 We are happy to sacrifice some precision for generality.
+Since the $W$ and $Z$ boson masses are similar, this sacrifice turns out to be
+quite small anyway.
 
 Multi-bin fits are not special.
-They are standard \atlas\ practice, and there could have been no complication
-in replacing the overlapping SR2-int and SR2-high with orthogonally binned
-regions.
-However, my design decisions for $\twoljets$-electroweak
-caused problems for the clarity of communication and accuracy of background
-modelling.
-These issues are developed further through the remainder of this chapter.
+They are standard \atlas\ practice, and there should have been no complication
+had I designed copied SR2-int and SR2-high for $\twoljets$-electroweak but with
+with minor tweaks to change them into orthogonally binned regions.
+However, my design decisions for $\twoljets$-electroweak became much more
+ambitious, increasing to a total of $13$ signal regions with complicated
+definitions.
+Although this multi-bin design squeezed out some extra sensitivity,
+it also caused problems --- for the clarity of communication and the accuracy
+of our background modelling;
+these issues shall be developed throughout the remainder of this chapter.
 
@@
 My design of $\twoljets$-electroweak violates this.
-Instead, it is designed as a tree whose branches recursively split various
-variables.
-This leads to specifications which are challenging to present and understand,
+Instead, it is designed more like a tree whose branches recursively split
+various variables.
+This leads to specifications which are challenging to present or understand,
 and which require region names with numerous suffixes to specify where they
@@
 
-Worse, I disregarded the idea that every region should have a large
-background expectation.
-Normally, this is justified by the limitations of approximate statistical
-methods, which work
+Worse, I disregarded the popular idea that every region should have a large
+background expectation, preferably of ten events or more.
+This requirement is typically justified by the limitations of approximate
+statistical methods that work
 ``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
@@
 physics~\cite{barlow1990extended})
-that are standard and easily analysed.
+that are standard elsewhere in high energy physics, and are easily analysed
+by the standard likelihood fitting methods.
 
@@
 If they could do that with technology from the year $1995$, we can do it now.
+This is a fact, but only if we can really trust our predictions of zero
+backgrounds.
 
@@
 Although main simulated samples may have good precision,
-many `systematic' variations and `data-driven' estimates are also required.
-Many of these alternative estimates have far worse precision, and can fail
-totally as the number of simulated events accepted plunges towards zero;
-they can require $\approx10$ data events to work properly.
-
-Unfortunately, $\twoljets$-electroweak selections were finalized before I
-appreciated this problem that background variations were dubious.
-Various workaround are invoked to assign plausible variations despite small
-yields, which we shall in the remainder of this chapter.
-
-As seen in the splash summary of Section~\ref{sec:2ljets_splash}, this search
-finds agreement with backgrounds except for some local deficits of data below
-backgrounds.
+many variations for systematic uncertainties and data-driven estimates are also
+required.
+Many methods to evaluate these variations have far worse precision than the
+central simulations, and can fail totally as their number of Monte Carlo events
+accepted plunges towards zero.
+In practice, our methods to assign background uncertainties can easily require
+tens of expected events to work properly.
+
+Unfortunately, the $\twoljets$-electroweak regions designs were finalized
+before unblinding and before I appreciated this problem with background
+uncertainties in the smaller regions.
+Various dubious workaround are invoked to assign plausible variations despite
+small yields, which we shall in the remainder of this chapter.
+As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
+$\twoljets$-electroweak search observes agreement with post-fit backgrounds
+except for some local deficits of data below backgrounds.
 Exclusion contours exceed their expected range because of these deficits,
@@
 sensitive at intermediate and small (off-shell/compressed) mass splittings.
+
 For the GMSB model, our result complements limits from the
@@
 masses with $m(\neutralino_1) \in 250\textrm{--}600\,\eV[G]$.
+Exclusion of the largest $B(\neutralino_1 \rightarrow h \tilde{G})$ values
+requires sensitivity to double-Higgs decay signatures --- \cms\ have published
+exclusions in the region combining $hh\rightarrow \gamma\gamma+\met$ and
+$hh\rightarrow b\bar bb\bar b$ with other channels to exclude
+$B(\neutralino_1 \rightarrow h \tilde{G}) = 1$ up to
+$m(\neutralino_1) = 750\,\eV[G]$~\cite{
+cms_susy_2018_partial_run2_combination,
+CMS-SUS-16-045,
+CMS-SUS-16-044
+}.
+
+For reasons described in this section, the $\twoljets$-electroweak search is
+designed with large number of orthogonal regions, and with those regions its
+results offer some extensions over previous and orthogonal constraints.
+Before describing those regions' definitions and modelling we apply to their
+expected yields, the following sections now return to the basic details of the
+data and event variables they use.
 
@@
 \label{sec:2ljets_events}
-This section describes the events selected for the $\twoljets$ searches, the
-reconstructed particle objects derived from those events and various
+This section describes the events selected for the three $\twoljets$ searches,
+the reconstructed particle objects derived from those events and various
 important event variables.

### 2022-08-10 main/chapter_2ljets.tex
proofread events section

@@
 From its name, it is clear that $\twoljets$ analyses features of events with
-two charged leptons (electrons or muons --- not taus), and at least one jet.
-Though not nominal, a $\met$ requirement is also implicit from the context
+two charged leptons (those are electrons or muons --- not taus),
+and at least one jet.
+Though not in the name, a $\met$ requirement is also implicit from the context
 of $R$-parity conserving supersymmetric models.
@@
 searches.
-Higher level event variables and selections will only be described for the
+Higher level event variables and selections will only be described here for the
 $\twoljets$-electroweak search.
@@
 \label{sec:2ljets_trigger}
-There would be no event variables if not for the trigger.
-When living our daily lives, we are bombarded with sensory data and must pay
-attention only those gems few of any use, to avoid exhausting our
+There would be no event variables if not for the trigger reducing the data to
+a manageable size.
+When living our lives as humans, we are bombarded with sensory data and must
+pay attention only those gems few of any use, to avoid exhausting our
 processing capacity.
@@
 or long-lived particles.
-The list varies through time, so we select triggers specifically in
-each data-taking period.
-
-The loosest triggers are `prescaled' --- their data rate is reduced by randomly
-accepting only a small fraction of events.
-Prescaling adds complication which we avoid;
-for $\twoljets$, we use only the `lowest' (loosest, highest rate) un-prescaled
-triggers selecting two charged leptons~\cite{atlas_twiki_lowest_unprescaled}.
-Those charged leptons are selected in the
-high-level trigger (described in Section~\ref{sec:atlas_trigger})
-primarily requiring large $\pt$ and certain quality criteria which vary by
-trigger.
+That trigger list has varied through time, so we select triggers individually
+in each data-taking period.
+
+The loosest triggers are `prescaled', meaning that their data rate is reduced
+by randomly accepting only a small fraction of events.
+Interpreting prescaled data adds some complications that we avoid in
+$\twoljets$, by using only un-prescaled triggers.
+We use use only the `lowest' (loosest, highest rate) triggers selecting two
+charged leptons~\cite{atlas_twiki_lowest_unprescaled}.
+Those charged leptons are selected in the high-level trigger,
+described in Section~\ref{sec:atlas_trigger}, which
+primarily requires large $\pt$ and certain quality criteria.
 
@@
 selects for an electron pair, a muon pair, or an emu (electron-muon pair).
-To describe those triggers, let $\pt^{\ell_i}$ be the transverse momentum of the
-$i\mathrm{th}$ hardest lepton.
-
-In short, all triggers require $\pt^{\ell_i}$ values to exceed thresholds in
+To describe those triggers, we use $\pt^{\ell_i}$ for the transverse momentum
+of the $i\mathrm{th}$ hardest lepton.
+All triggers require $\pt^{\ell_i}$ values to exceed thresholds in
 the range $8\textrm{--}26\,\eV[G]$.
@@
 \end{itemize}
+Throughout the data used in $\twoljets$, trigger requirements
+have increased in hardness with time.
 The point of listing these is to demonstrate trigger activations are
@@
 }.
-Throughout the data used in $\twoljets$, trigger requirements
-have increased in hardness with time.
+To avoid these thresholds, all leptons selected for the $\twoljets$-electroweak
+regions are required to have $\pt > 25\,\eV[G]$.
 
@@
 that trigger does not also harmfully increase inseparable backgrounds.
-A study, \emph{which was performed by Knut Vadla}, investigated the effects
+A study, \emph{performed by Knut Vadla}, investigated the effects
 of including one-lepton triggers.
@@
 in excess of $90\%$ of events from benchmark C1N2 models.
-Therefore, any additional trigger has at most a $\approx10\%$;
+Therefore, any additional trigger has at most a $10\%$;
 improvement. Adding one-lepton triggers was found to increase efficiency by
 only a few percentage points, depending on the model.
-We considered that to be too little to be worthwhile.
-
-Costs of triggers are attributable to human effort;
-software work to include them and their associated Monte Carlo weights, along
-with studies to validate background estimates and evaluate systematic
+The $\twoljets$ team considered that to be too little to be worthwhile.
+
+Costs of triggers are primarily attributable to human effort:
+software work to include them and their associated Monte Carlo weights, and
+extra studies to validate background estimates and evaluate systematic
 uncertainties.
 Similarly, our implementation of the Matrix Method fake/non-prompt lepton
-background estimate (see Section~\ref{sec:2ljets_mm_fakes}) is not equipped to
-handle one-lepton triggers.
+background estimate (described in Section~\ref{sec:2ljets_mm_fakes}) is not
+equipped to handle one-lepton triggers, so further work would have been
+required to replace it.
 
 Although jet and $\met$ triggers were not studied in detail, they can be
-quickly ruled out since they hard objects: examples of these triggers require
+quickly ruled out since they only trigger on hard objects:
+examples of these triggers require
 approximately $\ptjone > 400\,\eV[G]$, $p_\mathrm{T}^{\,j_3} > 200\,\eV[G]$, or
@@
 rates, as they suppressed by electroweak couplings.
-Therefore, \atlas\ can trigger on leptons with much softer requirements;
-lepton triggers allow us to use softer jets and $\met$, which are relied upon
+\atlas\ can therefore trigger on leptons with much softer requirements;
+lepton triggers allow us to use softer jets and $\met$, that are relied upon
 throughout the $\twoljets$ analyses.
+Although $\twoljets$-electroweak requires $\met > 100\,\eV[G]$ in all of its
+regions, the same is not true of $\twoljets$-RJR not various validation
+studies.
 
@@
 In the context of $\twoljets$, `leptons' means `electron-or-muons';
-this is because taus, which decay before reconstruction, and are not used
-directly in $\twoljets$, and neutral leptons (the neutrinos) go missing.
+this is because taus decay before reconstruction so are used only indirectly
+through their decay products, and neutral leptons
+--- neutrinos --- go missing and contribute to the $\met$.
 
 Reconstructed lepton objects feature momenta, charges, and flavours.
-Each is included only if it satisfies certain kinematic and
-is identified with `quality' criteria.
+Each lepton is accepted for use in $\twoljets$ only if it satisfies certain
+kinematic and is identified with `quality' criteria.
 For both electron and muons, these criteria are standardized with labels
-`loose', `medium', and `tight'.
-These are nested, in the sense that tight implies medium,
+`loose', `medium', and `tight';
+these labels are nested in the sense that tight implies medium
 and medium implies loose.
-Other, more specialized, criteria exist, but they are not relevant $\twoljets$.
+Other, more specialized, criteria exist in \atlas, but they are not relevant
+$\twoljets$.
 
@@
 Baseline leptons and are used in background estimation methods and to ensure
-orthogonality with other \atlas\ analyses, which study events with differing
-lepton counts.
+orthogonality with other \atlas\ analyses that study events with differing
+lepton counts, such that our results might be combined in future to deliver an
+improved total sensitivity.
 Baseline requirements are softer and lower quality than signal requirements.
@@
 
-Baseline leptons are required only to have $\pt > 3\,\eV[G]$ for muons and
+Baseline leptons must have $\pt > 3\,\eV[G]$ for muons and
 $\pt > 4.5\,\eV[G]$ for electrons.
-Signal leptons are all required to have $\pt > 10\,\eV[G]$, along with tighter
+Signal leptons must have $\pt > 10\,\eV[G]$, along with tighter
 quality and isolation requirements, details of which are given in the
@@
 This topological clustering algorithm is no longer the \atlas-recommended jet
-algorithm; it has been supplanted in recent years by a `particle flow' method,
-which uses tracking information to improve the reconstruction of charged
+algorithm; it has been supplanted in recent years by a `particle flow' method
+that uses tracking information to improve the reconstruction of charged
 hadrons~\cite{atlas_jet_pflow_PERF_2015_09}.
 We did not attempt to switch to these improved jets because the change in
-policy occurred during the years this analysis was in development, and the
-change would have required more work and even more delays.
+recommendation occurred during the years this analysis was in development, and
+the change would have required more work and further delays for an unknown
+amount of benefit.
 
@@
 Tagging adds labels to objects.
-Unlike how electrons and muons are very cleanly separated, however, jet tagging
-is a more noisy affair.
+Unlike how leptons are tagged as electrons and muons with very clean
+separation, jet tagging is a rather more noisy affair given \atlas' limited
+ability to reconstruct the individual hadrons within jets.
 Taggers are available for various jet properties, such charm flavour ($c$-jet),
 bottom flavour ($b$-jet).
-Of these, $\twoljets$ uses only $b$-tagging.
+Of these, $\twoljets$ uses only $b$-tagging, which is particularly useful to
+reject $\ttbar$ events and select for $h\rightarrow b\bar b$ decays.
 
 Larger radius jets can also use taggers for heavy decays, such as those from
-$W$, Higgs or top resonance, but large radius were considered too much
-additional work to considered in $\twoljets$.
+$W$, Higgs or top resonance~\cite{JETM-2018-06}, but large radius were
+considered too much additional work to considered in $\twoljets$,
+particularly since they would require a large set of independently evaluated
+systematic variations in addition to those on our standard jets.
 
@@
 inner trackers.
-
 For jets in $\twoljets$, $b$-tagging is useful to separate events due to
@@
 } at its $77\%$ efficiency working point.
-This efficiency means that a true $b$-jet sampled from its testing set has a
-$77\%$ probability of being tagged.
-It is trained on a mixture of $\ttbar$ and $\zjets$ simulated data,
-and the `c10' suffix means that those data include a $7\%$ fraction of
+This algorithm is trained on a mixture of simulated $\ttbar$ and $\zjets$ data,
+and the reported efficiency means that a true $b$-jet sampled from the
+algorithm's testing dataset has an estimated $77\%$ probability of being
+tagged.
+The `c10' suffix means that those data include a $7\%$ fraction of
 $c$-jets, and that the tagger does not use newer inputs from the
@@
 SMT (Soft Muon Tagger)~\cite{atlas_twiki_mv2, atlas_jet_mv2c10_2017}.
-
 This tagging algorithm combines various lower-level features, including the
@@
 Alternative working points of $70\%$ and $85\%$ efficiency are used in similar
-\atlas\ searches.
-These were trialled and found to make negligible differences to $\twoljets$.
+\atlas\ searches;
+these alternatives were trialled and found to make negligible differences in
+$\twoljets$.
 
@@
 photon. (Although we do not directly use photons in $\twoljets$, we do not veto
-events containing them --- photons contribute to $\ptmiss$.)
-
-Now we know about both leptons and hadronic jets, we can also improve our
-rejection of non-prompt leptons from hadron decays, since leptons near to jets
-become suspect.
-
+events containing them --- photons are accounted for in the momentum balance
+of $\ptmiss$.)
+
+Once we know about all of the reconstructed leptons and hadronic jets, we can
+also improve our rejection of non-prompt leptons from hadron decays, since
+leptons near to jets become suspect.
 To remove duplicates and further reduce non-prompt lepton contributions, we
 apply a standard overlap-removal procedure.
-This procedure, which is detailed in the paper~\cite{atlas2022searches},
-comprises a sequence of rules which modify the collection of event objects.
-Some rules decide which to keep from an overlapping pair,
-and others remove leptons overlapping with jets, based on their $\Delta R$
+This procedure, which is detailed in the $\twoljets$
+paper~\cite{atlas2022searches},
+comprises a sequence of rules that modify the collection of event objects.
+Some rules decide which object to keep from an overlapping pair,
+and others remove leptons overlapping with jets based on their $\Delta R$
 and $\pt$.
@@
 The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
-Please note that $\met$ is not an energy.
-Transverse energy does not exist.
-One may of course define their own directed `transverse energy', but it will
+Please note that $\met$ is not an energy; transverse energy does not exist.
+One may, of course, define a directed `transverse energy', but it will
 not be conserved (in the standard model) unless it is equivalent to a conserved
@@
 Indeed, momentum-balanced events do have energy imbalances;
-for example, in a $Z\mathrm{+jet}$ event $pp \rightarrow Zg$, the $Z$ and $g$
-are produced back-to-back with equal momenta $p$, but the two particles have
-very different energies since they have very different masses:
-$E_Z^2 = m_Z^2 + p^2$, $E_g^2 = 0 + p^2$.
+for example, in a $Z\mathrm{+jet}$ event $pp \rightarrow Zg$, with the $Z$ and
+$g$ emitted back-to-back in the transverse plane with equal momenta
+$p = |\pm \vec p|$, the two particles have different energies since they
+have different masses: $E_Z^2 = m_Z^2 + p^2$, $E_g^2 = 0 + p^2$.
 % hobbyhorse
@@
 The $\ptmiss$ vector is defined such that the vector sum of $\ptmiss$ with
-all reconstructed objects (after overlap removal) has zero transverse momentum.
+all reconstructed objects, after overlap removal, has zero transverse momentum.
 In addition to the leptons and jets used directly in $\twoljets$, this includes
@@
 
-Since signal models in $\twoljets$ do generate hard, invisible particles,
-by selecting for large $\met$ we select events in which either there were real
-invisible particles or something was badly mismeasured.
-The standard selection for $\twoljets$-electroweak is $\met > 100\,\eV[G]$.
+Signal models in $\twoljets$ generate hard, invisible particles.
+By selecting for large $\met$, however, we select events in which either there
+were real invisible particles or something was badly mismeasured.
+And the rate of bad mismeasurements from Standard Model processes
+with large cross-sections can be substantial.
+A standard selection requirement in $\twoljets$-electroweak is
+$\met > 100\,\eV[G]$.
 This is a nice round number, and corresponds roughly to where the $\met$-less
-background $\zjets$ tails off.
-This $100\,\eV[G]$ also corresponds roughly to several standard deviations of
-a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$, which I observe
-by inspecting $\met / \metsig$ in $\twoljets$-electroweak selections,
-and where $\metsig$ is defined in the next section.
+background $\zjets$ tails off,
+and also corresponds roughly to several standard deviations of
+a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$;
+I observe that noise scale by inspecting $\met / \metsig$ in
+$\twoljets$-electroweak selections, where $\metsig$ is defined in the next
+section.
 
@@
 
-The experimental resolution of $\ptmiss$, however, is imprecise.
-Importantly also, the uncertainty on any inferred $\ptmiss$ varies
-significantly from one event to another.
+The experimental resolution of $\ptmiss$, however, is imprecise, and
+importantly that resolution varies significantly from one event to another.
 
@@
 than a jet of similar energy.
-Unless, as discussed in Chapter~\ref{chapter:experiment}, the muon is hard
+Unless, as discussed in Chapter~\ref{chapter:experiment}, that muon is hard
 enough that its curvature through \atlas' magnetic fields is poorly tracked.
@@
 such that the weak decays of its hadrons tend to lose momentum in
-neutrinos or soft muons. (The $\metsig$ variable used in this analysis is not
-aware of $b$-tagging information, but perhaps it could benefit from some.)
+neutrinos or soft muons.
 
@@
 mismeasurements.
+(The $\metsig$ variable used in this analysis is not directly
+aware of $b$-tagging information, but perhaps it could benefit from some.)
 
 In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
-contributions more than $\met$ alone, plausibly since $\ttbar$ events are
+contributions more than $\met$ alone, plausibly because $\ttbar$ events are
 accompanied by more hadronic activity than other sources of $\ptmiss$.
@@
 What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
-selections is almost pure \diboson\ -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
+selections is almost pure \diboson\ --- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
 and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
@@
 uncertainty and yields taken to the right of a given value.
-Large values indicate expected sensitivity to the signal model.
+Larger values indicate greater expected sensitivity to the signal model.
 }
@@
 Just as $\ptmiss$ is constructed from a (negative) sum of contributions from
-objects in an event,
-$\Sigma$ is a sum of covariances assigned to those objects;
-both means and variances add linearly.
+objects in an event, this $\Sigma$ is a sum of covariances assigned to those
+objects; both means and variances add linearly.
 
@@
 mass $\mttwo$ is a partner to the `transverse mass' $m_T$
-(to be defined shortly),
-that is useful in searches for supersymmetric particles.
+(to be defined shortly)
+which is useful in searches for supersymmetric particles.
 
 The stransverse mass of an event is a greatest lower bound on the mass of a
-pair-produced resonance, that applies when each partner decays semi-invisibly.
+pair-produced resonance that applies when each partner decays semi-invisibly.
 For example, an on-shell $\tilde\ell^+\tilde\ell^-$ pair decaying as
@@
 
-Lacking direct observation of sleptons, $\mttwo$ remains useful in separating
-standard model processes; dileptonic $\ttbar$ and $W^\pm W^\mp$ decays both
-satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ (since $0 \approx m(\nu)$).
-The $\twoljets$-electroweak search depends on this effect;
-most selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$, or similar,
+Lacking direct observation of the sleptons, or any other supersymmetric
+resonances, at the LHCs, $\mttwo$ remains useful for separating standard model
+processes since, with ideal reconstruction,
+dileptonic $\ttbar$ and $W^\pm W^\mp$ decays both
+satisfy $\mttwo(\ell^+, \ell^-, 0) \leq m(W)$ since $m(\nu) \approx 0$.
+The $\twoljets$-electroweak search depends on this effect:
+most of its selections require $\mttwo(\ell^+, \ell^-, 0) > 80\,\eV[G]$,
+or similar,
 to reduce $\ttbar$, $W^\pm W^\mp$ and $\tau^\pm\tau^\mp$ backgrounds.
 
-To be precise, the stransverse mass explores how to split the observed
-$\ptmiss$ between the two visible decay products in an event, and can be
-defined as
+More precisely, the stransverse mass can be defined as
 \begin{equation}
@@
 \end{equation}
-in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$ and $\{E,\,\vec p\,\}_\mu$ denotes
-the four-vector with energy $E$ and momentum $\vec p$~\cite{lester1999measuring}.
+in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$ and $\{E,\,\vec p\,\}_\mu$
+denotes the four-vector with energy $E$ and momentum
+$\vec p$~\cite{lester1999measuring}.
+That is, can be seen as exploring the attribution of the observed $\ptmiss$
+between the two visible decay products in an event, finding a best mass over
+all possible assignments.
 As an extension, one can assign different masses $m_p$ and $m_q$ to
@@
 Numerical evaluation of $\mttwo$ can be performed efficiently with interval
-bisection algorithms, which seek the least resonance mass for which the
-transverse masses allow any compatible assignment which also satisfies
+bisection algorithms that seek the least resonance mass for which the
+transverse masses allow any compatible assignment that also satisfies
 $\vec p + \vec q = \ptmiss$.
@@
 library~\cite{gillam2021mt2}, to which I contributed a faster and more robust
-implementation of the bisection algorithm by
+implementation of the bisection algorithm of
 Lester and Nachman~\cite{lester2015bisection} during this project.
@@
 \subsection{Jet pairs}
-Di-jet assignments, for variables like $\mjj$ and $\rjj$, are always chosen as
-the two hardest jets.
-Why?
-
-This choice matches the convention of SR2-int and SR2-high from the previous
-search~\cite{atlas_23l_SUSY_2016_24}, but my initial reaction was that since we
-are looking for jet pairs near to $m(W)$ or $m(Z)$, we should surely choose
-candidate jet pairs by their mass.
+In $\twoljets$, di-jet assignments for variables like $\mjj$ and $\rjj$ are
+always chosen as the two hardest jets. Why?
+
+One reason is that this choice matches the convention of SR2-int and SR2-high
+from the previous search~\cite{atlas_23l_SUSY_2016_24}.
+However, my initial reaction was that since we are looking for jet pairs near
+to $m(W)$ or $m(Z)$, we should surely choose candidate jet pairs by their mass.
 
 Two alternative di-jet assignments were therefore considered in early
-development of region designs:
+development of my $\twoljets$-electroweak region designs:
 the two jets with invariant mass closest to $m(W)$, and the two jets with the
@@
 The result was bland.
-All assignments performed similarly, and all left similar fractions of signal
-yields outside the mass window.
-The simplicity and precedent for the two-hardest jet assignment therefore swung
-the decision.
-
-However, alternatives did select different subsets of signal yields;
+All assignments performed similarly, and all left similar amounts of signal
+outside of their selected mass windows.
+The simplicity of, and precedent for, the two-hardest jet assignment therefore
+swung the decision in its favour.
+
+However, alternatives did select different subsets of signal yields:
 the alternative jet assignments had some potential to define secondary signal
@@
 analysis.
-To aid future revision of their meanings, definitions, and references for these
-variables are collected here.
-All dimensionful quantities are in $\eV[G]$.
+For future reference, their meanings, definitions, and references for these
+variables are all collected here.
+All dimensionful quantities are always reported in $\eV[G]$ units in this
+thesis.
 \begin{itemize}
 \item $\met$: missing transverse momentum
-\item $\metsig$: object-based missing transverse momentum significance;\\
+\item $\metsig$: object-based missing transverse momentum significance,\\
 described in Section~\ref{sec:2ljets_metsig}
@@
 \item $\mjetone$: mass of the hardest jet
-\item $\mttwoll$: `stransverse' mass of the lepton pair with massless invisibles;\\
+\item $\mttwoll$: `stransverse' mass of the lepton pair with massless
+invisibles,\\
 described in Section~\ref{sec:2ljets_mt2}
@@
 \item $\njet$: number of signal jets
-\item $\nbtag$: number of b-tagged jets;
+\item $\nbtag$: number of b-tagged jets,
 described in Section~\ref{sec:2ljets_btagging}
@@
 Please remember that none is an exact or true value;
-all are estimates based on approximate reconstructions.
+all of these event variables are estimates derived from approximate
+reconstructions.
 While this distinction is easy to drop in lax language, it is necessary to
@@
 \label{sec:2ljets_design}
-Perhaps the most impactful decision in a search is the design of its regions.
-Like most decision, these designs must be made to balance conflicting and
+Perhaps the most impactful decision in an \atlas\ SUSY search is the design of
+its regions.
+Like most decisions, these designs must be made to balance conflicting and
 vaguely-specified desires, and are informed with incomplete information.
@@
 Some figures label \crvz\ as `\crz Z'; the region was renamed late in this
-analysis project when it was show to contain substantial fractions of $W\kern-0.4exZ$
-as well as $ZZ$, as illustrated in Figure~\ref{fig:2ljets_splits} of
+analysis project when it was show to contain substantial fractions of
+$W\kern-0.4exZ$ as well as $ZZ$, as illustrated in
+Figure~\ref{fig:2ljets_splits} of
 Section~\ref{sec:2ljets_validation}.

### 2022-08-10 main/chapter_2ljets.tex
more proofread

@@
 The latter could be overly tuned, perhaps by a learning algorithm, to specific
-simulations.
-Perhaps it cut out a highly weighted sample which, although annoying, was an
+simulations; perhaps it cut out a highly weighted sample that, although
+annoying, was an
 honest representation of our modelling.
@@
 Numerical simplicity is valuable.
-I wrote in a draft of our \atlas-internal documentation that numerical
+In a draft of our \atlas-internal documentation I wrote that numerical
 simplicity was a design goal, but am guilty of removing that statement after a
-senior reviewer suggested that it was ``too honest''.
+reviewer suggested that it was ``too honest''.
 % ~\cite{comment2020toohonest}. % Could cite, probably should not.
@@
 
-The clearest idea behind this design is to split regions on $\metsig$.
-Motivation for this design is displayed in
-Figure~\ref{fig:2ljets_presel_met};
+The first idea behind this design is to split regions on $\metsig$, as
+motivated by the histograms displayed in Figure~\ref{fig:2ljets_presel_met};
 $\metsig$ does better than $\met$ in separating background and signal
-contributions for models with various mass splittings above $m(Z)$.
-Furthermore, $\metsig$ also works to separate signal models from each other;
+contributions for models with various mass splittings.
+Furthermore, $\metsig$ also works to separate signal models from each other:
 signals with larger mass differences in their decays to invisibles tend to have
 larger $\metsig$.
-Binning regions in $\metsig$ therefore also improves sensitivity to
-individual parameter points and allows targeted selections on other,
-locally-relevant, event variables.
-
-Regions in three blocks of $\metsig$ are named High, Intermediate (or Int),
-and Low,
-which target models with decreasing mass splittings and lepton pairs on the
-$Z$ resonance mass.
-Separate regions target $\mll < m(Z)$ for models with small mass splittings,
-which force decays to proceed via low-mass off-shell weak bosons.
+Initial binning in $\metsig$ therefore offers improves sensitivity to
+various parameter points.
+In each $\metsig$ bin, the $\twoljets$-electroweak design further targets
+signal sensitivity with other, locally-relevant event variables.
+
+Three of our signal $\metsig$ bins are named
+`High',
+`Intermediate' (or `Int'), and
+`Low',
+to label both the size of $\metsig$ they require and the size of
+mass-splittings they target.
+These High, Int, and Low selections all target lepton pairs on the $Z$
+resonance mass.
+Separate `OffShell' regions target $\mll < m(Z)$ events that arise
+in models with small mass splittings that force decays to proceed via
+low-mass off-shell weak bosons.
 
@@
 
-All regions are defined within a loose `Pre-selection' requirement which
-requires the basic $Z\rightarrow \ell^+\ell^-$ plus jets plus $\met$
-conditions that we target in the $\twoljets$-electroweak.
+All $\twoljets$-electroweak regions are defined within a loose `Pre-selection'
+requirement of the basic $Z\rightarrow \ell^+\ell^-$ plus jets plus $\met$
+conditions that we target.
 These Pre-selection requirements are defined in Table~\ref{tab:2ljets_presel}.
@@
 $\twoljets$-electroweak search.
-Although $\metsig > 6$ implies that most events already have large $\met$
-$\met > 100\,\eV[G]$ requirement is included to ensure that events with small
-$\met$ and smaller uncertainties on it are excluded.
+Although $\metsig > 6$ implies that most events already have large $\met$, the
+$\met > 100\,\eV[G]$ requirement is included to exclude any rare events with
+small $\met$ and precise reconstructions.
 }
@@
 
+In the remainder of this section,
 High regions are discussed in Section~\ref{sec:2ljets_high}.
@@
 to Ben Hooberman for suggesting the investigation of massive jets,
-and to both, and the whole $\twoljets$, collaboration in their finalization.
+and the the whole $\twoljets$ team for collaboration in their finalization.
 }
@@
 \label{sec:2ljets_high}
-The highest $\metsig$ selections, of around $\metsig > 18$, contain almost pure
-\diboson\ backgrounds in relatively small quantities, along with the bulk masses
+The highest $\metsig$ selections, starting around $\metsig > 18$, contain
+pure \diboson\ backgrounds in relatively small quantities, along with the bulk
 of signal samples at the heaviest resonance masses.
-This is therefore the most promising territory for signal regions which are
+This is therefore the most promising territory for signal regions that are
 highly sensitive to the most extreme signals.
-Their backgrounds, however, must be constrained by control regions in less
-extreme parameter spaces with more data.
-
-Diboson backgrounds and C1N2 signals have more in common than just $\met$.
-Notably, both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
+Their backgrounds, however, must be constrained by control regions in
+less-extreme parameter spaces with sufficiently many data to precisely
+normalize their scale factors.
+
+Diboson backgrounds and C1N2 signals have more in common than just $\met$;
+both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
 and do not enhance the rate of heavy-flavour jets, as displayed in
@@
 variables to reduce backgrounds, this observation allows us to win
-additional sensitivity in High regions by loose requirements.
+additional sensitivity in High regions by loose requirements
+of $\mll \in 71\textrm{--}111$ and $\nbtag \leq 1$.
+This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
+mismeasured events away from the central peak.
+Similarly, allowing up to one $b$-tag improves yields without boosting
+backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
+processes.
 
@@
 
-The $\mll \in 71\textrm{--}111$ window loosely selects the $m(Z)$ peak to
-scoop up nearby rare or mismeasured events.
-Similarly, allowing up to one $b$-tag improves yields without boosting
-backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
-processes.
-
-The standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement captures
-both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
-respectively.
+Here, the standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement
+captures both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
+respectively and is applied similarly to all other on-shell
+$\twoljets$-electroweak signal regions.
 
@@
 Although large-radius jets are typically used to capture these, we only use
-small, $R=0.4$, jets in the $\twoljets$ analysis.
-And those small jets can work for the extreme boosts that our signals produce.
+small, $R=0.4$, jets in the $\twoljets$ analysis, and those small jets can
+be good enough to work for the extreme boosts that our signals produce.
 To catch these extreme events is the goal of SR-High-1J, which is displayed
@@
 The use of $\mjetone$ with small-radius jets for this region is unusual,
-and required modelling of additional uncertainties that we discuss in
+and required modelling of additional uncertainties that we discuss later in
 Section~\ref{sec:2ljets_jet_mass}.
@@
 
-All other regions require two jets or more, so this single-jet selection has
-the liberty to reduce its $\metsig$ requirement to $\metsig > 12$
+All other regions require two jets or more, so this single-jet region design
+has the freedom to reduce its $\metsig$ requirement to $\metsig > 12$
 without overlapping any other regions.
@@
 sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
-The low edge of this $\mbb \in 60\textrm{--}150\,\eV[G]$ window is chosen not
-only to accommodate \atlas' noisy reconstruction of $b$-jets, but also to
-capture $Z\rightarrow bb$ decays, which become significant signal models with
-small $B(\neutralino_1 \rightarrow h \gravitino)$.
-
-As shown in Figure~\ref{fig:2ljets_high_llbb_region}, backgrounds in \srllbb
-contain a significant proportion of \topother\ quark processes;
-studies presented in Section~\ref{sec:2ljets_validation} show that these are
-mainly $t\bar t Z$.
+The low edge of this $\mbb \in 60\textrm{--}150\,\eV[G]$ window is far from the
+Higgs boson mass at around $125\,\eV[G]$, and is chosen not only to
+accommodate \atlas' noisy reconstruction of $b$-jets, but also to
+capture $Z\rightarrow bb$ decays, which become significant for signal models
+with small branching fractions $B(\neutralino_1 \rightarrow h \gravitino)$.
+
+As shown in Figure~\ref{fig:2ljets_high_llbb_region}, the Standard Model
+background in \srllbb\ contains a significant proportion of \topother\
+processes; studies presented in Section~\ref{sec:2ljets_validation} show that
+these are mainly $t\bar t Z$.
 
@@
 Figure~\ref{fig:2ljets_high_mjj_vrhigh}.
-
 Various validation studies explored similar selections and different
 simulated samples, and found no clear explanation for this effect.
-It therefore motivates the second validation region VR-High-R, and is one
-reason for including a lower bound of $\mjj > 20\,\eV[G]$ in all regions.
+The resulting doubt in our modelling of light jet pairs also therefore motivates
+the second validation region VR-High-R, and the inclusion of a lower bound of
+$\mjj > 20\,\eV[G]$ in all regions.
 As seen in Figures~\ref{fig:2ljets_high_region}
 and~\ref{fig:2ljets_high_mjj_vrhigh}, VR-High and VR-High-R otherwise show
-acceptable pre-fit modelling even without systematic uncertainties.
+acceptable pre-fit modelling even without the inclusion of systematic
+uncertainties in their error bars.
 
@@
 Note that the region $\mjetone > 110$ is also checked in this validation region
-and seen to contain no data.
-This is good for the standard model.
+and seen to contain no data --- a good sign for the standard model.
 Since background estimates are near zero up there, non-zero data would have
-ruled out our modelling in favour of any model making small positive
-predictions.
+ruled out our modelling in favour of any plausible model that made small
+positive predictions.
 
@@
 orthogonality from other \atlas\ searches with which our results are to be
-combined.
+combined in future.
 
@@
 What was not appreciated soon enough, however, was an excess in the $\metsig$
-tail of this region, which is reproduced from the paper in
-Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper},
+tail of this region, which is shown in the $\twoljets$ paper and reproduced
+Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper}.
 The highest bins have very small background predictions and positive data.
-Had this been noticed before unblinding, more validation might have
-been advisable.
+Had this been noticed before unblinding, more validation studies or altered
+modelling might have been necessary.
 ``In the end'', comments one \atlas\ author,
@@
 ]{%
-Post-fit $\metsig$ distribution in VR-$\llbb$.
-Backgrounds are to control and signal regions.
+Post-fit $\metsig$ distribution in VR-$\llbb$ with all statistical and
+systematic uncertainties~\cite{atlas2022searches}.
+Backgrounds are fitted to control and signal regions.
 This region is also displayed pre-fit with coarser bins in
@@
 
-Intermediate signal regions are defined to sit in the $\metsig$ range of
+Our Intermediate signal regions sit in the $\metsig$ range of
 $12$ to $18$.
@@
 To suppress the increased backgrounds, the intermediate $\mll$ window is
-tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and \diboson
+tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and \diboson\
 processes.
-Requiring zero $b$-tags further reduces $\ttbar$ backgrounds.
+Requiring zero $b$-tags further reduces $\ttbar$ backgrounds here.
 
@@
 The less extreme signal models targeted here have less boosted
-$W\rightarrow jj$ distributions, and so it was found that a $\ptjone$ selection
-was effective at separating signals;
-a $\ptjone > 60\,\eV[G]$ requirement split SR-Int from its validation region,
-VR-Int and is seen from Figure~\ref{fig:2ljets_int_region} to retain most
-signal contributions.
+$W\rightarrow jj$ distributions, but a $\ptjone$ selection is found to be
+effective as an additional discriminator of signals from backgrounds to make
+up for this loss;
+a $\ptjone > 60\,\eV[G]$ requirement splits SR-Int from its validation region
+VR-Int, and is seen from Figure~\ref{fig:2ljets_int_region} to keep most
+signal contributions in the signal region.
 Following our standard pattern set with other signal regions, SR-Int is binned
@@
 In Low regions, which have $\metsig \in 6\textrm{--}12$, the $\zjets$
-background becomes a large since its fake $\met$, which arises from
+background becomes large since its fake $\met$, which arises from
 from mismeasurements or rare interactions, is no longer eliminated.
-Similarly, backgrounds from $\ttbar$ increase; the selections inherited from
-Intermediate regions to suppress $\ttbar$ need supplementing.
-
-These regions target signal models with small mass splittings which still
-exceed the masses of electroweak bosons.
-In particular, we aim to test C1N2($200$, $100$) as a light, on-shell
-benchmark which has not previously been excluded in this channel.
-
-Angular selections on lepton variables $\rll$ and $\dphillmet$ characterize
-the Low regions, which are defined in Table~\ref{tab:2ljets_low}.
+Backgrounds from $\ttbar$ similarly increase and the selections inherited from
+Intermediate regions are no longer sufficient.
+
+These Low regions target signal models with small mass splittings,
+of around $100\textrm{--}150\,\eV[G]$ that still exceed the masses of
+electroweak bosons.
+In particular, we aim to test C1N2($200$, $100$), which is a light, on-shell
+benchmark that has not previously been excluded in this
+channel~\cite{atlas_23l_SUSY_2016_24}.
+
+Angular selections on lepton variables $\rll$ and $\dphillmet$ are powerful
+in these Low regions, and characterize their definitions that
+are given in Table~\ref{tab:2ljets_low}.
 These angular variables usefully reject backgrounds including $\ttbar$, and
 provide signal sensitivity to benchmark signal models, as illustrated in
-Figure~\ref{fig:2ljets_low_minus};
-signals with small mass-splittings of around $100\textrm{--}150\,\eV[G]$
-are targetted.
-
-To remove some observed mis-modelling of backgrounds, Low regions require
-exactly two jets.
-Abandoned region designs which used more jets are discussed below in
-Section~\ref{sec:2ljets_low_isr}.
+Figure~\ref{fig:2ljets_low_minus}.
 
@@
 $\ttbar$.
-As elsewhere, binning of SR-Low in $\metsig \in 6\textrm{--}9\textrm{--}12$
-to form SR-Low-a and SR-Low-b respectively, increases sensitivity.
+As elsewhere, sensitivity is increased by binning SR-Low with
+$\metsig \in 6\textrm{--}9\textrm{--}12$ to form SR-Low-a and SR-Low-b
+respectively.
+
+To remove some observed mis-modelling of backgrounds, Low regions require
+exactly two jets.
+Abandoned region designs which used more jets are discussed below in
+Section~\ref{sec:2ljets_low_isr}.
 
@@
 
-Uniquely to Low regions, and signal models with small mass-splitting, it is
-shown in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2} that substantial
-signal yields exist at small $\mttwoll$.
+Uniquely to Low regions and signal models with small mass-splitting, it is
+found, and shown in Figure~\ref{fig:2ljets_low_minus_norll_nomt2_mt2}, that
+substantial signal yields exist above backgrounds at small $\mttwoll$.
 To capture these signals, SR-Low-2 is defined as the only
 $\twoljets$-electroweak region with $\mttwoll < 80\,\eV[G]$.
-That threshold of $80\,\eV[G]$ is not very important.
+That precise threshold of $80\,\eV[G]$ is not very important, and used only to make it
+a simple inversion of the usual $\mttwoll > 80\,\eV[G]$ requirement.
 After the other selections for SR-Low-2, almost all signal and background
 samples are in the spike near zero;
-a lower requirement would make little difference, and $80\,\eV[G]$ matches
-other, usually inverted, selections.
+a lower requirement would make little difference and work just as well,
 
@@
 $\dphillmet$, and backgrounds do not; this motivates the $\dphillmet < 0.6$
-requirement, and can be understood since a boosted resonance decaying to
-$Z+\mathrm{invisible}$ will tend to send the $Z\rightarrow \ell^+\ell^-$ and
-$\ptmiss$ in the same direction.
+requirement, and can be plausibly understood from scenarios where a the boosted
+supersymmetric resonance decays to $Z+\mathrm{invisible}$ and sends both
+the $Z\rightarrow \ell^+\ell^-$ and invisible particle, which becomes
+$\ptmiss$, in the same direction.
 
@@
 usually expected to align with a mismeasured jet.
-This is explained by the context of a prior $\mttwoll < 80\,\eV[G]$
+This is explained by the context of the $\mttwoll < 80\,\eV[G]$
 requirement:
@@
 masses.
-Following its definition as a $\min-\max$ of transverse masses,
-given in Section~\ref{sec:2ljets_mt2},
+Following its definition as a $\min-\max$ of transverse masses from
+Section~\ref{sec:2ljets_mt2},
 the $\mttwoll$ requirement therefore rejects those events, leaving only the
@@
 This space contains about $2.3$ background expectation (mostly \diboson) and
-$\approx 1$ from some benchmark signals, so its inclusion as a single combined
-bin not have increased sensitivity.
-Its inclusion as an orthogonal bin with small yields would have added excessive
-complication, however.
+around $1$ expected event from some benchmark signals, so its inclusion as a
+single combined bin might not have increased sensitivity.
+Its inclusion as an orthogonal bin with small yields would also have added
+unnecessary complication with little benefit.
 
@@
 Standardizing cuts to remove high-$\mjj$ tails (such as requiring
-$\mjj < 500\,\eV[G]$) could have made control and validation regions like \crz
+$\mjj < 500\,\eV[G]$) could have made control and validation regions like \crz\
 kinematically closer to the signal regions they are supposed to target,
-and at least in this case, I believe such a cut would have improved the
-analysis.
+and such a cut may well have improved the analysis in this case.
 
@@
 
-The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
-Their union therefore has a continuous $\mjj$ distribution, which is displayed
+The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low;
+their union therefore has a continuous $\mjj$ distribution that is displayed
 in Figure~\ref{fig:2ljets_low_sr_or_cr_region}.
@@
 overestimated.
-This plot contradicts that; most of the $\zjets$ distribution is fine;
-the strange effect happens where \diboson\ becomes large near $m(W)$.
+This plot contradicts that: most of the $\zjets$ distribution is fine, and
+the strange data deficit happens where \diboson\ becomes large near $m(W)$.
 
@@
 hasty redesign shortly before unblinding.
-
 Before that redesign, an multi-jet signal region SR-Low-ISR was included,
-with associated validation and control regions,
-which required $\njet \geq 3$ and offered some meagre extra sensitivity.
+with associated validation and control regions
+that required $\njet \geq 3$ and offered some meagre extra sensitivity.
 Here, `ISR' stands for Initial-State Radiation, following the naming
-of previous \atlas\ searches including partial Run-2 2/3-lepton
-search which we are updating~\cite{atlas_23l_SUSY_2016_24}.
-
-Although initial-state radiation typically refers to soft emissions in
+of previous \atlas\ searches including partial Run-2 $2\ell$/$3\ell$
+search which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
+$\twoljets$-electroweak search.
+
+Although initial-state radiation often refers to soft emissions in
 parton showers~\cite{corcella2000initial, bewick2022initial},
 our ISR explicitly includes hard emissions from the perturbative matrix element
-simulation;
-ISR is selected to give a hard transverse boost to the supersymmetric
-particles, making their resulting $\met$, and other kinematics, more distinctive.
-
-As an event variable, we follow \cite{atlas_23l_SUSY_2016_24} to define the
-ISR momentum as the vector sum of all jets not assigned to the
-$W\rightarrow jj$ decay.
-This ISR construction does not select the two hardest for the $W$ candidate,
-unlike our other regions, but instead takes the two closest in $\phi$ to the
-vector $\vec \pt^{\ell\ell} + \ptmiss$;
-the two jets selected in this way are used to define the mass $\mwisr$.
-
-The region SR-Low-ISR required $\ptisr > 300\,\eV[G]$ for a substantial
-boost and $\dphiisrmet > 2.6$ to catch invisible particles boosted away from the
-ISR, along with standard selections $\mwisr \in 60\textrm{--}110\,\eV[G]$,
+simulation.
+Our ISR is selected to give a hard transverse boost to the supersymmetric
+particles, making their resulting $\met$ more distinctive.
+
+To select ISR with event variables, we follow \cite{atlas_23l_SUSY_2016_24}
+to define the ISR momentum as the vector sum of all jets not assigned to the
+$W\rightarrow jj$ decay;
+unlike our other regions, this construction does not select the two hardest for
+the $W$ candidate, , but instead takes the two closest in $\phi$ to the
+vector $\vec \pt^{\ell\ell} + \ptmiss$.
+The two jets selected in this way are used to define the mass $\mwisr$.
+
+The abandoned region SR-Low-ISR required $\ptisr > 300\,\eV[G]$ for a
+substantial boost, and $\dphiisrmet > 2.6$ to catch invisible particles boosted
+away from the ISR,
+along with standard selections $\mwisr \in 60\textrm{--}110\,\eV[G]$,
 $\metsig \in 6\textrm{--}12$, and preselection, as well as $\mjj > 110\,\eV[G]$
@@
 
-The main ISR regions are shown in Figure~\ref{fig:2ljets_low_isr_old}; no data
-are shown in the signal region, since it was never unblinded.
+The main ISR regions are shown in Figure~\ref{fig:2ljets_low_isr_old}.
+No data are shown in this signal region, since it was never unblinded.
 From the validation and control regions, however, there is a clear deficit
@@
 On the very sound advice of \atlas\ authors commenting on the internal
-documentation, this ISR construction was entirely removed.
+documentation, this ISR construction was entirely removed from the search
+along with all of its associated regions.
 
@@
 This deficit was therefore blamed on multi-jet QCD modelling, and resolved by
-retreating to two-jet events.
-It is interesting , however, that the two-jet SR-Low eventually observed a
-comparable deficit.
+retreating to the SR-Low requirement of exactly jets.
+It is interesting, however, that the two-jet SR-Low eventually observed a
+comparable deficit upon unblinding.
 
@@
 scenarios~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}, in which
-models have mass splittings below $m(W)$ or $m(Z)$, such that decays must
-proceed below their pole masses.
-Models with smaller off-shell mass splittings are described as more
-`compressed'.
-Separate searches have, however, targetted compressed scenarios in
-specialized and carefully designed analyses~\cite{
+models have mass splittings below $m(W)$ or $m(Z)$ such that their decays must
+proceed below the weak bosons' pole masses.
+Such models with off-shell mass splittings are described as
+`compressed', and separate, specialized and carefully designed, searches
+have targetted for compressed scenarios in our two-lepton channel~\cite{
 atlas_susy_compressed_2l_2016_partial_run2,
 atlas_susy_compressed_2l_2018_run2
-} in selections with two leptons.
-
-Since compressed signals generate softer kinematics and \atlas\ is sensitive
-to soft leptons, three-lepton searches have stronger sensitivity to
-compressed scenarios~\cite{atlas_rjr_3l_SUSY_2019_09}.
+}.
+
+Since compressed signals generate soft kinematics and \atlas\ is more sensitive
+to soft leptons than soft jets, three-lepton searches have stronger sensitivity
+to compressed scenarios~\cite{atlas_rjr_3l_SUSY_2019_09}.
 These searches consider the same C1N2 $W\kern-0.4exZ$ signal models as us, but target the
 case where the $W$ decays leptonically $W\rightarrow\ell\nu$.
-For comparison, impressive off-shell exclusion contours are reproduced
+For comparison with our off-shell exclusion,
+shown in Figure~\ref{fig:2ljets_contours_c1n2},
+the impressive off-shell exclusion contours are reproduced
 from the \atlas\ Run-2 $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09} in
@@
 
-This is the purpose of SR-OffShell and its associated validation and control
-regions that are defined in Table~\ref{tab:2ljets_offshell}.
+To target compressed models is the purpose of SR-OffShell and its associated
+validation and control regions, all of which are defined in
+Table~\ref{tab:2ljets_offshell}.
 Histograms of SR-OffShell and VR-OffShell are displayed in
@@
 
-Early OffShell designs used the `ISR' construction, which is described
+Early OffShell designs used the `ISR' construction that was described
 Section~\ref{sec:2ljets_low} for Low regions,
@@
 decay products.
-
-After some iterations, however, the same intent was reproduced with a simpler
-and more effective selection.
-As seen in Table~\ref{tab:2ljets_offshell}, this design is to simply ask for
-one hard jet facing away from $\ptmiss$; this is specified as
+After some iterations, however, the same intent was reproduced with the simpler
+and more effective selection that we present here.
+
+As seen in Table~\ref{tab:2ljets_offshell}, our OffShell region design
+simply asks for one hard jet facing away from $\ptmiss$, specified as
 $\ptjone > 100\,\eV[G]$ and $\dphijmet > 2$.
-In comparison to ISR variables, this design improves yields by not needing jets
-for a $W\rightarrow jj$ candidate.
+In comparison to the ISR construction, this design improves yields by not
+needing reconstructed jets for a $W\rightarrow jj$ candidate.
 
@@
 
-Hadronic resonances which decay to lepton pairs,
-such as the bottomonium $\Upsilon$,
-become important below $10\,\eV[G]$, and we did not have simulations available
-for the $\mll < 10\,\eV[G]$ range.
+Hadronic resonances that decay to lepton pairs,
+such as the various charmonium ($J/\psi$) bottomonium ($\Upsilon$) states,
+become important below around $10\,\eV[G]$, and we did not have simulations
+available for the $\zyjets$ process below $\mll < 10\,\eV[G]$.
 The requirement $\mll > 12\,\eV[G]$ is therefore imposed to robustly remove
-this un-modelled regime with a margin for error.
-
-Unlike on-shell regions, the standard $\mttwoll > 80\,\eV[G]$ requirement did
-not adequately reduce off-shell $\ttbar$ backgrounds.
+this un-modelled regime with a $2\,\eV[G]$ margin to avoid edge effects.
+
+Unlike on-shell regions, the standard $\mttwoll > 80\,\eV[G]$ requirement does
+not adequately reduce OffShell $\ttbar$ backgrounds.
 Pushing it up to $\mttwoll > 100\,\eV[G]$ was found to work well for the signal
@@
 $Z$ resonance, but depends on the full Drell-Yan, $\zyjets$,
-process~\cite{drell1970massive}.
-That is, on- and off-shell $\mll$ distributions are physically different.
-Since also $\zyjets$ simulations are culturally not trusted, we decided
-to introduce separate normalization factors for $\zyjets$ in on- and off-shell
-regions.
+process~\cite{drell1970massive};
+on- and off-shell $\mll$ distributions are physically different.
+Since also $\zyjets$ simulations are culturally not trusted in our regime of
+large $\met$, we decided to introduce separate normalization factors for
+$\zyjets$ in on- and off-shell regions.
 
@@
 That shows that the combination is a relatively minor refinement that may or
-may not be worth the cost of implementing it.
+may not be worth the cost of implementing it in different contexts.
 
@@
 remain inspired by our simulated signal models.
-Various approaches are taken to discovery regions in \atlas\ searches.
-Although it is normal to have one per signal
-regions~\cite{SUSY-2018-02},
-some searches study
-dozens of discovery regions comprising signal region bins~\cite{SUSY-2018-36},
-or other modifications to signal
+Various approaches have been taken to discovery regions in \atlas\ searches
+--- although it is normal to have one per signal regions~\cite{SUSY-2018-02},
+some searches study dozens of discovery regions comprising signal region
+bins~\cite{SUSY-2018-36}, or other modifications to signal
 regions~\cite{atlas_susy_compressed_2l_2018_run2}.
 
-Personally, I do not see the benefit from tens of discovery regions.
+Personally, I see little benefit from having tens of discovery regions.
 Yes, given a pet model, one can choose the region with its best expected
@@
 
-We therefore chose to design a small number of discovery regions, each of which
-comprises a union of one or more signal regions.
-
-This decision turned out well, not because of the above statistical
-pontification, but since the practical cost of running the standard statistical
+We therefore design just five discovery regions for the
+$\twoljets$-electroweak, each of which comprises a union of one or more
+signal regions.
+This decision turned out well, not because of the statistical pontifications
+above, but because the practical cost of performing the standard statistical
 software to analyse each discovery region (and to perform their many validation
 studies), turned out to be so large~\cite{verkerke2003roofit}.
+Having many more discovery regions would have further exaggerated this
+computational and human cost.
 
@@
 \begin{itemize}
-\item DR-High is SR-High-8 (the union of SR-High-8-a and -b),
+\item DR-High is SR-High-8 (the union of SR-High-8-a and SR-High-8-b),
 \item DR-$\llbb$ is is \srllbb,
-\item DR-Int is SR-Int (the union of SR-Int-a and -b),
-\item DR-Low is SR-Low (the union of SR-Low-a and -b)
-\item DR-OffShell is SR-OffShell (the union of SR-OffShell-a and -b).
+\item DR-Int is SR-Int (the union of SR-Int-a and SR-Int-b),
+\item DR-Low is SR-Low (the union of SR-Low-a and SR-Low-b)
+\item DR-OffShell is SR-OffShell (the union of SR-OffShell-a and SR-OffShell-b).
 \end{itemize}
-Although it was not a design requirement, these give discovery regions are
-all orthogonal.
+Although orthogonality was not a design requirement for these discovery
+regions, they do happen to be all orthogonal.
 
@@
 were chosen for broad sensitivity to varieties of signals.
-These studies are displayed in Figures~\ref{fig:2ljets_disco_trials_high},
+Some plots from this study are displayed in
+Figures~\ref{fig:2ljets_disco_trials_high},
 ~\ref{fig:2ljets_disco_trials_llbb}~and~\ref{fig:2ljets_disco_trials_int}
@@
 Similar figures were produced for the other categories, but they are used older
-region definitions, are identical in style, and are not illuminating
+region definitions, are identical in style, and are not particularly
+illuminating.
 
@@
 Evaluation of candidate Intermediate discovery regions.
-The region SR-Int, labelled `Int\_1+\_2' here, is chosen for its simplicity
+The region SR-Int, labelled `Int\_1$+$\_2' here, is chosen for its simplicity
 and broad sensitivity.
@@
 Although sharper sensitivity to special cases could have been achieved
-with more complicated selections, the reward of this discovery region design
+with more complicated selections, a nice reward of this discovery region design
 is that it can be stated clearly as the bullet pointed list given above.
 
-The five discovery regions are displayed in
+Histograms of our five discovery regions are displayed in
 Figure~\ref{fig:2ljets_disco_prefit}
@@
 mentioned previously.
-to make inferences from these data, it is necessary to assign honestly
+To make inferences from these data, it is necessary to assign honestly
 uncertain background models; the next section tackles this crucial problem of
@@
 Those backgrounds may or may not have appeared to have predicted the data well,
-but no conclusions can be drawn until uncertainties are assigned to the
-background and signal modelling.
+but no conclusions can be drawn until uncertainties are assigned in our
+modelling of background and signal contributions.
 
@@
 of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
-For the $\twoljets$ search, we used \histfitter\ to build our every HistFactory
+For the $\twoljets$ search, we use \histfitter\ to build and analyse those
 models~\cite{Besjes_2015,baak2015histfitter}.
 
-Thankfully, many standard variations are evaluated automatically with \atlas\
-software by simulating alternative samples with varied parameters, then
-comparing the resulting histograms.
-Such standard variations relate to performance and calibration effects, which
-have been carefully studied by many researchers in \atlas, to whom I am deeply
+Thankfully, many standard variations, corresponding to experimental
+uncertainties, are evaluated automatically with \atlas\ software by simulating
+alternative samples with varied parameters, then comparing the resulting
+histograms.
+Such standard variations relate to performance and calibration effects that
+have been carefully studied by many \atlas\ researchers, to whom I am deeply
 grateful.
@@
 Statistical uncertainties on simulated samples are also semi-automatic.
-In each bin, its statistical uncertainty is controlled by a single parameter,
-which intends to cover the uncertainties on all backgrounds samples
-contributing in that bin.
-An alternative prescription exists which gives one parameter per sample per
-bin, but this greatly increases the complexity of the model and was not
-considered necessary.
+We choose the ``staterror'' scheme, for which each bin has its statistical
+uncertainty controlled by a single constrained parameter that intends to cover
+the statistical uncertainties on all of its contributing
+samples~\cite{cranmer2012histfactory}.
+An alternative ``shapefactor'' prescription exists which gives one parameter
+per sample per bin, but this greatly increases the complexity of the model and
+was considered unnecessary here.
 
@@
 Of those not evaluated automatically, some of the more interesting or important
-will be discussed in the remainder of this section.
-Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
+will be discussed in the remainder of this section:
+Fake/non-prompt lepton backgrounds are in Section~\ref{sec:2ljets_mm_fakes},
 some alternative samples in Section~\ref{sec:2ljets_sample_comparisons},
@@
 Section~\ref{sec:2ljets_other_background_variations},
-and signal modelling uncertainties in
+and signal uncertainties in
 Section~\ref{sec:2ljets_signal_variations}.
@@
 expected that the rates of fake leptons, or non-prompt leptons (which do not
-originating from the hard scatter) should be small.
+originate from the hard scatter) should be small.
 Indeed, more detailed estimates indicate that they are small.
@@
 
-Although my style of spelling out FNP or FNP every time
-(as opposed to just `fake') may be awkward, it is supported by the fact that this
+Although my style of spelling out fake/non-prompt or FNP every time
+(as opposed to just `fake') may be awkward, it is supported by the fact that
 the non-prompt part dominates.
-Indeed, figures in our internal documentation~\cite{twoljets2018int} suggest
-that, in simulations, non-prompt hadron decays produce in excess of $90\%$ of
-this background in a loose selection near to our regions, and across the
-entire $\metsig$ spectrum.
-Non-prompt muons are mostly from decays of heavy-flavour ($b$ or $c$) hadrons,
-and electrons arise in roughly equal parts from light flavour hadrons.
+Indeed, figures \emph{kindly provided by Eirik~Gramstad} in our internal
+documentation~\cite{twoljets2018int} suggest that, in simulations, non-prompt
+hadron decays produce in excess of $90\%$ of this background in a loose
+selection similar to our regions, and across the entire $\metsig$ spectrum.
+Non-prompt muons mostly arise from decays of heavy-flavour ($b$ or $c$)
+hadrons, and electrons arise in roughly equal parts from light flavour hadrons.
 Tau decays form the next largest part, with prompt fakes due to misidentified
@@
 In particular, do not use simulations of pure hadronic processes, so do not
-really know the rates at which these strong processes can fake leptons.
+really know the predicted rates at which these strong processes would fake
+leptons.
 Although prompt strong effects may only rarely fake leptons, they have very
@@
 backgrounds of fake and non-prompt leptons, so named because it works by
-estimating a linear relationship between dilepton event rates, which is
+estimating a linear relationship between dilepton event rates that is
 written as a matrix multiplication.%
 \footnote{%
-To be explicit, the Matrix Method does not use Matrix Elements from quantum
-scattering, and is not related to the similarly-named
-Matrix Element Method~\cite{gainer2013matrix}.
+The Matrix Method is not related to the Matrix Elements from quantum scattering
+theory, nor to the similarly-named
+Matrix Element Method~\cite{gainer2013matrix} for signal discrimination.%
 }
@@
 \emph{
-I am deeply grateful to Eirik~Gramstad for developing the Matrix Method for our
+Thanks to Eirik~Gramstad for developing the Matrix Method for our
 $\twoljets$ searches, documenting it, providing its samples, and showing us how
 to use them.
-Although I have reservations about the method itself (expounded below), these
+Although I have criticisms of the method itself (expounded below), these
 are not criticisms of his excellent work spent in using it.
@@
 respectively.
-
-Transition probabilities $p(*\!\mid\!*,i)$ are estimated by studies in
+These transition probabilities $p(*\!\mid\!*,i)$ are estimated by studies in
 separate selections and assigned as functions of the lepton
 flavour, $\pt$ and $|\eta|$.
-Given those estimated transition probabilities, the Matrix Method begins by
+Given estimated transition probabilities, the Matrix Method begins by
 taking some observed data
@@
 That is, the Matrix Method starts from data, flips them over to truth
-estimates, then flops those back to data-level estimates. Flip, flop.
+estimates, then flops those back to data-level estimates. Flip\ldots flop.
 
@@
 Systematic variations are assigned by comparing histograms when those weights
-are varied according to various uncertainties in the Matrix Method procedure.
-
-Yes, actual data are histogrammed in the Matrix Method, and those data include
-tight-tight events that comprise the data of our search regions, the very same
-data in which we are searching for anomalous signals.
+are varied according to various uncertainties in the Matrix Method procedure,
+and statistical errors are assigned in the standard way from the samples
+weights.
+
+Yes, actual data are histogrammed by the Matrix Method to assign its fake
+estimates, and those data include tight-tight events that enter our search
+regions --- the very same data in which we are searching for anomalous signals.
 Since it includes actual data, this method unblinds all regions,
@@
 
-In fact, real data tend to be assigned negative Matrix Method weights.
+Real, tight-tight data are assigned negative Matrix Method weights.
 For example, \crvz\ receives $157$ data in its Matrix Method sample,%
@@
 Unless, that is, the processes driving that excess also produced loose leptons
-at rates similar to (or larger than) the standard model backgrounds.
+at rates similar to (or larger than) the Standard Model backgrounds.
 Loose data have positive weights, so they would drive up the FNP estimate.
@@
 most search regions do not have substantial Matrix Method yields.
-I consider those that do to be dubious.
+I consider results in those that do to be dubious.
 
 The Matrix Method's approximation that
-$\mu^\textrm{data}_{**} = n^\textrm{data}_{**}$ can be effective but is not
+$\mu^\textrm{data}_{**} = n^\textrm{data}_{**}$ can be effective, but is not
 exact.
@@
 }.
-Our criticism that the Matrix Method prematurely unblinds data is not new, and
+This criticism that the Matrix Method prematurely unblinds data is not new, and
 the fake-factor is a notable alternative that avoids this problem by estimating
@@
 Selections requiring different-flavour or same-charge leptons,
-in otherwise $Z$-like scenarios, should therefore be
-FNP-enhanced.
+in otherwise $Z$-like scenarios, should therefore have enhanced FNP
+backgrounds.
 
@@
 or same-charge leptons.
-The different-flavour versions contain enhanced $\ttbar$ fractions,
+The different-flavour versions also contain enhanced $\ttbar$ fractions
 due to their flavour symmetry,
@@
 This is why the simulated backgrounds (not `FNP leptons') increase by around
-$30\%$ when switching off the Matrix Method; the difference between the
+$30\%$ when switching off the Matrix Method: the difference between the
 two orange histograms is the simulated FNP estimate.
 
-From this figure, it is clear that the Matrix Method has overestimated
-the background rate in this region, and that the raw simulations predict
-the data more accurately.
-The difference is less substantial than it may appear, however,
+From Figure~\ref{fig:2ljets_vrss} it is clear that the Matrix Method has
+overestimated the background rate in VR-SS, and that the raw simulations
+predict the data more accurately.
+The difference is less significant than it may appear, however,
 because the Matrix Method error bars are highly correlated --- they combine
-(in quadrature)
 statistical uncertainties with variations due to uncertainties in the
-Matrix Method weights.
-Matrix Method weight variations are parametrized with a single parameter
-which couples the yields in all bins, so moving it down by $1$-sigma in a fit
-would easily pull the backgrounds back into agreement with data.
+Matrix Method weights that are coupled with a single parameter in all bins,
+so moving it down by $1$-sigma in a fit would easily pull the backgrounds back
+into agreement with the data.
 
@@
 These plots were removed at the request of internal reviewers, who argued that
-they did not need to be shown since the FNP background is not
-very important.
-To demonstrate that it was not important, we first repeated the exclusion fits
+the figures were unnecessary since our FNP background is unimportant.
+To demonstrate that it was unimportant, we first repeated the exclusion fits
 with an additional uncertainty, to further cover the deviation observed in this
@@
 $m(\neutralino_1) \approx 500\,\eV[G]$.
-Different interpolation options induce larger variations, as do different
-random seeds for the fits, so this really is an insignificant difference.
+Different contour-interpolation options induce larger variations, as do
+different random seeds for the fits, so this really is an insignificant
+difference.
 
 Although we nominally use the Matrix Method to estimate fakes, its results
-are unusable in numerous regions, since they are either negative or zero.
+are unusable in numerous regions because they are either negative or zero.
 Details of these eleven problematic regions are displayed in
@@
 sets the low ($-1$-sigma) variation to zero.
-These three alternative estimates are reported with their combination in
-Table~\ref{tab:2ljets_fnp_estimates}.
 
@@
 
-These three estimates are labelled `MC', `MM', `Mean', and we combine them into
-the `Max' estimate.
+These three alternative estimates are labelled `MC', `MM', `Mean', and
+reported with their combination the final `Max' estimate in
+Table~\ref{tab:2ljets_fnp_estimates}.
 Each is described here.
@@
 \end{itemize}
-After I had proposed a combination of these two estimates, reviewers
-suggested additionally using the statistical properties of samples in looser
-selections.
-This became the third `Mean' estimate.
+After I had proposed a combination of these two estimates, and long after
+unblinding, internal reviewers suggested additionally using the statistical
+properties of samples in looser selections.
+This became the `Mean' estimate.
 \begin{itemize}
@@
 in most regions.
-The exception regions are all in the High category, and are those regions with
-small enough expected yields such that the Mean method dominates.
+The exceptions to this rule are all regions the High category,
+which have small enough expected yields for the high variation of the Mean
+method to appear relatively large.
 
@@
 barefaced honesty.
-Thankfully, we do not need to consider what might have happened had an excess
-been observed before the assignment of these fake estimates.%
+We are lucky that we do not need to consider what might have happened had an
+excess been observed before the assignment of these fake estimates.%
 \footnote{%
@@
 require a boosted, small-radius jet with a large mass
-($\mjetone \in 60\textrm{--}110$).
+$\mjetone \in 60\textrm{--}110$.
 This signal region is important for extending our reach to the most massive
@@
 Figure~\ref{fig:2ljets_jetm_c1n2_600_0} for a benchmark C1N2 model, showing
-that most  jets have small masses but that a substantial fraction appear in a
+that most  jets have small masses, but that a substantial fraction appear in a
 bump around $m(W)$.
 However, all of our signal models are simulated with the fast \atlas\ simulation
-\afii\ \cite{SOFT-2010-01} that uses approximate descriptions of the
+\afii\ \cite{SOFT-2010-01}, which uses approximate descriptions of the
 detector response and does not guarantee to accurately model the substructures
-that determines the jet's reconstructed mass.
+that determine jets reconstructed masses.
 Jet mass modelling should be well described since masses are part of the
@@
 However, momenta are typically more important for these variables, so
-errors could be hidden.
+problems could remain hidden from usual use-cases.
 Inclusion of these 1J regions in the $\twoljets$-electroweak search therefore
 required validation of these \afii\ mass distributions.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/2ljets_jetm_c1n2_600_0_tbrowser.png}
+\caption[
+Jet masses in raw samples of a benchmark C1N2 signal model
+]{%
+Jet masses in raw samples of a benchmark C1N2 signal model with
+$m(\chargino_1,\neutralino_2) = 600\,\eV[G]$ and
+$m(\neutralino_1) = 0$.
+All jets in all events in this file are histogrammed,
+and the signal model is simulated with fast simulation.
+\\[0.5em]
+Yes, this really is a screenshot of a ROOT TBrowser~\cite{ROOT}, with no
+weights on its samples, nor plotting frills such as error bars or labels.
+A similar view was our first, and surprising, indication that mono-jet masses
+might be useful, and although this chronology is not at all necessary to this
+thesis, I find it an amusing reminder of the power of quick-and-dirty checks.
+}
+\label{fig:2ljets_jetm_c1n2_600_0}
+\end{figure}
 
@@
 computing time I included only $\ttbar$ and single-top backgrounds generated
-with the full \atlas\ simulation.
-To verify the origins of high-mass jets, partonic truth information is used
+with the full \atlas\ simulation in this study.
+To verify the origins of high-mass jets, partonic truth information was used
 to label those events containing a $W$ boson decaying to two quarks within
 $\Delta R < 0.4$ of the hardest reconstructed jet.
-The resulting bump does exist, as shown in
+The predicted bump does exist and is shown in
 Figure~\ref{fig:2ljets_jetm_ttbar_mw}.
@@
 1J regions.
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.9\textwidth]{figures/2ljets_jetm_c1n2_600_0_tbrowser.png}
-\caption[
-Jet masses in raw samples of a benchmark C1N2 signal model
-]{%
-Jet masses in raw samples of a benchmark C1N2 signal model with
-$m(\chargino_1,\neutralino_2) = 600\,\eV[G]$ and
-$m(\neutralino_1) = 0$.
-All jets in all events in this file are histogrammed,
-and the signal model is simulated with fast simulation.
-\\[0.5em]
-Yes, this really is a screenshot of a ROOT TBrowser~\cite{ROOT}, with no
-weights on its samples, nor plotting frills such as error bars or labels.
-A similar view was our first, and surprising, indication that mono-jet masses
-might be useful, and although this chronology is not at all necessary to this
-thesis, I find it an amusing reminder of the power of quick-and-dirty checks.
-}
-\label{fig:2ljets_jetm_c1n2_600_0}
-\end{figure}
 
@@
 processes that were not processed for this study.
-Backgrounds are normalized to data in the high-mass region near to $m(W)$.
+A scale factor is applied to all backgrounds to normalize them to the data in
+the high-mass region near to $m(W)$.
 }
@@
 Not only does fast simulation pose a problem to the 1J regions, but our use
-of $\mjetone$ in small-radius ($R=0.4$), with the older
+of $\mjetone$ in small-radius $R=0.4$ jets with the older
 `topologically clustered' jet technology~\cite{atlas_jet_topo_PERF_2014_07}
@@
 large-radius jets with $R=0.8$, since these require less extreme boosts to
-capture their targets.
-The focus from \atlas' performance operations is therefore reasonable to
-understand these common use-cases first.
+capture their targets, and he focus from \atlas' performance operations is
+therefore reasonable to understand these common use-cases first.
 
@@
 To illustrate this point, imagine a wide error bar and
-two post-fit results right in its centre;
+two post-fit results right in its centre ---
 both have achieved the best fit, but both also could allow large variations.
@@
 larger uncertainties may well be reasonable given our imprecise constraints on
-the Standard Model.
+predictions from the Standard Model.
 

### 2022-08-10 main/chapter_2ljets.tex
modelling proofread

@@
 \label{sec:2ljets_zjets_alt}
-As mentioned in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison
-is hugely important to our results in SR-Low.
+As seen in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison is
+important to our results in SR-Low.
 Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
@@
 
-Since \sherpa\ implements both the hard scatter and shower models, we are
-considering variations due to the simultaneous change of both of these
-components together.
+Since \sherpa\ implements both the hard scatter and parton shower models, by
+comparing these two samples we are considering variations due to the
+simultaneous change of both of hard scatter and parton shower components
+together.
 
@@
 Note that up and down variations of the sample comparison parameter do not
-mean up and down variations of bin yields, but to the coupled variations
-in directions given by the signs of the relative variations given in the
-various regions.
+necessarily mean up and down variations of bin yields, but to the coupled
+variations in directions given by the signs of the relative variations given in
+the various regions.
 These variations are evaluated according to a standard recipe prescribed by
@@
 As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
-$\sigma_2$ is a substantial contribution to these variations, and so
+$\sigma_2$ is a substantial contribution to these variations, so
 the statistical uncertainty in the alternative sample does significantly
@@
 It may have been reasonable to treat them differently, perhaps by adding
-independent parameters in the separate regions.
-I believe such a change would have improved the plausibility of our background
+independent parameters to describe the statistical errors in the separate
+regions.
+Such a change may well have improved the plausibility of our background
 modelling, and I regret not finding the capacity to push for it at the time.
@@
 but I know of no official advice on how to judge what is ``too small''.
-Unblinding procedures are discussed further in
+Our unblinding procedure is discussed further in
 Section~\ref{sec:2ljets_unblinding}.
@@
 ``All EWK search control and signal regions are included in the fit.''
-
 Fitting only to control region (not signal regions) results in
@@
 Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
-is almost wiped out, leaving a deficit below data.
+is almost wiped out, leaving an apparent post-fit excess of data in VR-Low.
 
@@
 of its effect is due to the statistical error on a small alternative sample.
-To avoid this questionable effect, I searched for alternative configurations
+To avoid this questionable effect, I explored alternative configurations
 of the background model, such as constraining the parameter to $1$-sigma.
-Committee decisions within the \atlas\ SUSY group dictated that we should keep
+Committee decisions within the \atlas\ SUSY group stated that we should keep
 it as described, reasoning along the lines that
@@
 and this $\zjets$ sample comparison is the main reason why.
-Internal reviewers were, however, confused as to why we mentioned a deficit in
+Internal reviewers were, there, confused as to why we mentioned a deficit in
 SR-Low that did not appear in the results plots or tables.
@@
 and to my knowledge, is only reported through inner details of
-the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}.
+the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}
+(and now, I hope, this thesis).
 
@@
 its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
-we do not expect large $W\rightarrow jj$ contributions in this \diboson\ process,
-since one boson is needed for the leptons and the other is needed for the
-$\met$ that we select.
+we do not expect large $W\rightarrow jj$ contributions in this \diboson\
+process since one boson is needed for the leptons and the other is needed for
+the $\met$ that we select.
 We shall see in Section~\ref{sec:2ljets_diboson_alt} why \diboson\ yields are
@@
 \label{sec:2ljets_diboson_alt}
-No alternative samples were available for comparison with our central \diboson
-samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
+No complete alternative samples were available for comparison with our central
+\diboson\ samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
 electroweak vertices.
 This limitation is simply a result of how \atlas\ has decided to allocate
-its limited resources on, and quite reasonably so, since full simulation is
-computationally expensive.
+its limited resources on, and quite reasonably so since full simulation is
+computationally expensive and \diboson\ processes are relatively rare.
 We did, however, have access to the cheaper alternative set of small,
@@
 proportions, but for \texttt{CSS\_KIN} most are correlated in the positive
-direction, with the only SR-OffShell-a moving in the negative direction, by
+direction, with the only SR-OffShell-a moving in the negative direction by
 $46\%$.
@@
 respectively and hope to improve the statistical precision.
-Both require the pre-selection from Table~\ref{tab:2ljets_presel};
+Both of these loose VR-sys-* regions require the pre-selection from
+Table~\ref{tab:2ljets_presel};
 VR-sys-main also requires
@@
 which damps real emissions with weights to vary an ``eigentune'' parameter
-named \texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}..
+named \texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}.
 The central sample uses a default value $h_\textrm{damp} = 172.5\,\eV[G]$ to
@@
 scales tending to reduce yields.
-Variations in $\mu_R$ are typically in the $\approx 10\textrm{--}30\%$ range,
-making them larger than the $\approx 1\textrm{--}10\%$ variations due to
+Variations in $\mu_R$ are typically around the $10\textrm{--}30\%$ range,
+making them larger than the approximately $1\textrm{--}10\%$ variations due to
 $\mu_F$.
@@
 variation, but that may not mean more flexibility in the model since the SUSY
-prescription has three degrees of freedom.
+prescription has three independent degrees of freedom.
 
@@
 Like scale variations, uncertainties in parton distribution functions (PDFs)
-are assigned by weights stored Fuks:2012qxby the generator.
-These weights, however, do not correspond to simple parameters variations,
+are assigned by weights stored by the generator.
+These weights, however, do not correspond to simple parameter variations,
 but different replica PDFs in a Monte Carlo ensemble
-that is intended to model the uncertainties left after fitting to data
-in the NNPDF methodology~\cite{NNPDF:2017mvq,BALL2011112,BALL2012608}.
+that is intended to model the uncertainties left after fitting with the
+\textsc{NNPDF} methodology~\cite{NNPDF:2017mvq,BALL2011112,BALL2012608}.
 
@@
 Remarkably, both SUSY and PMG prescriptions agree on how to assign variations
-from these replica weights, since both say to follow the PDF4LHC
-recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, p.
+from these replica weights --- both say to follow the PDF4LHC
+recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, pp.46,
+Section~``6.2. Formulae for the calculation of $PDF$ and $PDF+\alpha_s$
+uncertainties).
 Since PDF4LHC gives two alternative prescriptions, one using mean and variance
 and the other using quantiles, we evaluate and consider both.
-Both are negligible, making $\approx 1\%$ differences before normalization
-to control regions, but are included for completeness.
+Both assignments of PDF variations are negligible in all regions, making
+approximately $1\%$ differences before normalization to control regions, but are
+included for completeness.
 Since there is nothing particularly special about the central PDF weight,
-this prescription allows the central scale-factor to differ from unity,
-and this is also a $\approx 1\%$ effect.
+this prescription allows the central scale-factor to differ from $1$,
+and this is also a $1\%$-level effect.
 
@@
 This superiority of flexible models is not fundamental, but is a property of
-\heplikelihood\ fitting methods.
-If we want to punish models that make uncertain predictions, then we should
+best-fit methods.
+If we want to punish models that make uncertain predictions, then we could
 consider likelihoods (not \emph{maximum} \heplikelihood s), as described in
@@
 prescription~\cite{atlas_twiki_susytheoretical},
-we consider only a coherent scaling of the scales $\mu_R$ and $\mu_F$ together,
-so this theory variation includes a total of seven components.
-
-Scale variations have appeared here twice:
+these scale variations consider only a coherent movement $\mu_R$ and $\mu_F$
+together, so this theory variation includes a total of seven components.
+
+Scale variations now appeared twice for signals:
 first for cross-section variations and
@@
 cross-sections for varied exclusion contours.
-This is questionable, but we shall see here that, thankfully, this scale
-double-counting makes very little difference.
+This is dubious, but thankfully we shall see shortly that this double-counting
+of scale variations makes very little difference.
 
@@
 To evaluate each variation requires a new simulation with different parameters,
-so to reduce computational costs we use truth-level only
+so to reduce costs we use truth-level only
 (through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
@@
 
-Extrapolation is performed as follows.
+My extrapolation scheme is fairly arbitrary, and is performed as follows.
 Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
@@
 
-Relative uncertainty $r$ is evaluated for each component, and the total
-uncertainty is their sum in quadrature, capped at $100\%$.
+We interpolate an uncertainty for each of the seven components and combine them
+all into a single total variation by summation quadrature, capped at a $100\%$
+relative error.
 This is done independently in each selection region without considering the
@@
 To account for its drastically different kinematics, the C1N2 off-shell space
-is treated independently.
+is treated separately.
 }
@@
 this chapter.
-These are no different.
+Signal variations are no different.
 To reduce statistical noise here, I generated truth samples $10$~times as large
 as our central, reconstructed samples.
-These sample sizes vary roughly with cross-section from $300$k to
-$2.8$m events per sample, and due to the $\sqrt{n}$ scaling of
-Monte Carlo errors, means approximately $\sqrt{10}$ times reduced
-standard deviation.
-However, this remains insufficient since each resulting variation is a
+These sample sizes vary roughly with the signal cross-section from $300$k to
+$2.8$m events per sample.
+Due to the $\sqrt{n}$ scaling of Monte Carlo errors this means approximately
+$\sqrt{10}$ times reduced statistical noise.
+This remains insufficient, however, since each resulting variation is a
 sum in quadrature of variations in $7$ pairs of alternative samples
@@
 Where signal models' yields has significant in a signal region,
-each variation tends to be around $\approx 20\%$ relative error,
-and are compatible with the central samples within statistical uncertainties.
+each variation tends to be around $20\%$ relative error,
+and all are compatible with the central samples within statistical
+uncertainties.
 
@@
 
-We evaluate and extrapolate each source of uncertainty individually,
-and combine all sources by summation in quadrature in each region to form a
-single total theory variation~\cite{atlas_twiki_susytheoretical}.
-This sum loses all sense of direction, so we choose to give up and down
-variations the same $+$ and $-$ signs in all bins.
-Since the theory variation parameter is coupled across all bins, it therefore
-implements a form of non-uniform normalization factor.
+Combining all of these variations in quadrature loses all sense of direction,
+so we choose to give up and down variations the same $+$ and $-$ signs in all
+bins.
+Since the theory variation parameter is coupled across all bins, its effect is
+therefore to implement a slightly non-uniform normalization factor.
 
 Why the double counting of scale variations does not matter is now apparent:
-this variation is dominated by statistical noise, and the inclusion of one
-extra source for the scale variations simply scales that noise up by
+these signal variations are dominated by statistical noise, and the inclusion
+of one extra source for the scale variations simply scales that noise up by
 a factor of $\approx \sqrt{7/6}$.

### 2022-08-11 main/chapter_2ljets.tex
proofread validation

@@
 later in Section~\ref{sec:2ljets_validation} and displayed in
-Figure~\ref{fig:2ljets_splits_\diboson}.
+Figure~\ref{fig:2ljets_splits_diboson}.
 
@@
 \label{sec:2ljets_disco}
-
 Although discovery regions are model-independent in principle, our designs
@@
 \label{sec:2ljets_validation}
-Many additional validation studies supported the $\twoljets$ search prior to
+Many validation studies supported the $\twoljets$-electroweak search prior to
 its publication.
-Many of these, comprising hundreds of histograms in various regions or looser
-selections in various event variables, are available to \atlas\ members in
-our internal documentation~\cite{twoljets2018int}.
-This section highlights some interesting results from these validation studies.
+Results of these, mostly comprising hundreds of histograms in various regions
+or looser selections in various event variables, are available only to \atlas\
+members in our internal documentation~\cite{twoljets2018int}.
+This section presents some highlights these validation studies.
 
 \paragraph{\crz-met}
-Concerns about our modelling of $\zjets$ and \diboson\ backgrounds have been
+Concerns about our modelling of $\zjets$ and \diboson\ backgrounds were
 raised due to our deficits in SR-Low and SR-High, which compound with the
@@
 To further check these backgrounds, Figure~\ref{fig:2ljets_low_crz_extensions}
-displays histograms in a `\crz-met', which is modifies \crz
+displays histograms in a selection named `\crz-met', which modifies \crz\
 (from Table~\ref{tab:2ljets_low}) by removing its upper bound on $\metsig$.
 The arrow in Figure~\ref{fig:2ljets_low_crz_extensions_metsig} indicates this
-upper bound at $\metsig = 9$, and shows excellent agreement with data across
-the entire distribution, both where $\zjets$ dominates and where \diboson
-dominates.
-The corresponding $\met$ distribution that is shown in
+removed upper bound at $\metsig = 9$, and shows excellent agreement with data
+across the entire distribution and in all relative mixtures of $\zjets$ and
+\diboson\ contributions.
+The corresponding $\met$ distribution \crz-met that is shown in
 Figure~\ref{fig:2ljets_low_crz_extensions_met}
-similarly shows excellent agreement.
-
-Like our other plots intended for publication, these figures use the full fit
-model with all variations for statistical and systematic uncertainties,
-and are fitted to data in control and signal regions, but the signal regions
-have little effect since the pulled variations are normalized to this control
-region.
-To avoid the work of evaluating variations in \crz-met (and all of its bins),
-its variations are linearly extrapolated from \crz.
-
-Good modelling in \crz-met is a positive sign for our simulations prior to
-their modification in the fit.
-Noting that \crz\ excludes $m(W)$, however, this does not conflict with the idea
-of poor modelling in the signal regions.
+also shows excellent agreement.
 
@@
 
+Like our other plots intended for publication, these figures use the full
+background model with all variations for statistical and systematic
+uncertainties,
+and are fitted to data in control and signal regions.
+The signal regions have little effect on $\zjets$ here, however, since its
+variations are normalized to \crz.
+To avoid the work of evaluating variations in \crz-met, its variations are
+linearly extrapolated from \crz.
+
+Good modelling in \crz-met, as observed here, is a positive sign for our
+simulations prior to their modification in the fit.
+Since \crz\ excludes an $\mjj \approx m(W)$ window, however, this does not
+rule out of poor modelling in certain signal regions.
+
 \paragraph{Sub-processes}
@@
 
-Although it might be expected that our $Z$-like two-lepton selection would
-favour the $ZZ$-like \diboson\ sub-sample, with the other $Z$ decaying to
-neutrinos,
-Figure~\ref{fig:2ljets_splits_\diboson} shows that in fact the three-lepton
-$W\kern-0.4exZ$-like process has almost equal yields.
-All of those $W\kern-0.4exZ$ events therefore have a missing third lepton, that could be
-softer than the few $\eV[G]$ needed for our baseline selection
-(see Section~\ref{sec:2ljets_events_leptons}),
-or outside our angular acceptance.
-(Overlap removal cannot explain this, since we count baseline leptons before
-applying it.)
-Although we would hope that such failed identifications would be rare,
-the effect is enhanced by the fact that $W\kern-0.4exZ$ has a larger cross-section than
-$ZZ$ by around a factor of three~\cite{Campbell:2011bn}, that is enhanced
-by the sum over $W^+\kern-0.4exZ$ and $W^-\kern-0.4exZ$ processes.
-
-Our \sherpa\ \diboson\ samples include the interference of electroweak processes
-for each fixed final state, so for example the separation of $ZZ$ and $WW$ is
-not exact since they interfere.
-These histograms are therefore split on charged lepton number in their
-final states.
-
-Some of our $\zjets$ samples are filtered by hadron flavour
-(requiring at least one selected hadron with $|\eta| < 4$), and
-Figure~\ref{fig:2ljets_splits_zjets} shows \crz\ with its $\zjets$ yields
-split by these filters.
-Heavy flavour samples are enhanced, with the filtered yields comprising
-$22\%$ bottom and $30\%$ charm flavours despite the $\nbtag = 0$ veto in \crz;
-this is sensible, since we also require $\met$, and extra non-prompt $\met$ is
-generated  in the weak decays of heavy-flavoured hadrons.
-
-The remaining parts of Figure~\ref{fig:2ljets_splits_\diboson} show that
-of the rare-top sample, $\ttbar Z$ is the largest contribution,
-at $73\%$ in \srllbb, but that contributions from other rare processes
-such as $tW\kern-0.4exZ$, $t\ell\ell g$ and $\ttbar W$ are significant.
-Here, the plain `Top other' label refers to all other rare-top contributions
-not included in the bracketed specializations.
-
 \begin{figure}[tp]
@@
 \caption{\crvz, \diboson}
-\label{fig:2ljets_splits_\diboson}
+\label{fig:2ljets_splits_diboson}
 \end{subfigure}
@@
 
+Although one might expect our selected \diboson\ sample to comprise mostly
+$ZZ$, with one $Z\rightarrow \ell\ell$ and the other invisibly to
+$Z\rightarrow \nu\nu$ to generate our required $Z$-like lepton pair with
+$\met$.
+Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the three-lepton
+$W\kern-0.4exZ$-like process has similar equal yields.
+Those $W\kern-0.4exZ$ events must therefore have a missing third lepton from
+$W\rightarrow \ell\nu$ decays.
+To avoid selection, that third lepton could be softer than the few $\eV[G]$
+needed for our baseline selection
+(see Section~\ref{sec:2ljets_events_leptons}),
+or outside our angular acceptance.
+(Overlap removal cannot explain this since we count baseline leptons before
+applying overlap removal.)
+Although we would hope that such failed lepton identifications would be rare,
+the effect is enhanced by the fact that $W\kern-0.4exZ$ has a larger
+cross-section than $ZZ$ by around a factor of three~\cite{Campbell:2011bn},
+due in part to the sum over both $W^+\kern-0.4exZ$ and $W^-\kern-0.4exZ$
+processes.
+
+Our \sherpa\ \diboson\ samples include the interference of electroweak
+processes for each fixed final state.
+One consequence of this is that there is no exact separation between $ZZ$ and
+$WW$ to $\ell\ell\nu\nu$ final states. They interfere.
+These histograms are therefore split on charged lepton number in their
+final states and we cannot exactly attribute the intermediate boson states.
+
+Some of our $\zjets$ samples are filtered by hadron flavour
+(requiring at least one selected hadron with $|\eta| < 4$), and
+Figure~\ref{fig:2ljets_splits_zjets} shows \crz\ with its $\zjets$ yields
+split by these filters.
+Heavy flavour samples are enhanced, with the filtered yields comprising
+$22\%$ bottom and $30\%$ charm flavours despite the $b$-jet veto ($\nbtag = 0$)
+in \crz;
+this is sensible since we also require $\met$ and the weak decays of
+heavy-flavoured hadrons generate extra, non-prompt $\met$.
+
+The remaining parts of Figure~\ref{fig:2ljets_splits} show that $\ttbar Z$ is
+the largest contribution to our \topother\ sample,
+at $73\%$ in \srllbb, and that contributions from other rare processes such as
+$tW\kern-0.4exZ$, $t\ell\ell g$ and $\ttbar W$ are also significant.
+Here, the plain `Top other' label refers to all other \topother\ contributions
+not included in the bracketed specializations.
+
 \section{Unblinding}
@@
 It is a discussion of our unblinding procedures,
-their extent, their implementation, and their effect on the results.
+their extent, their implementation, and their impact on our published results.
 

### 2022-08-11 main/chapter_2ljets.tex
begin unblinding proofread

@@
 
-Blindness to data is useful, but not necessary.
-All that is necessary is that our predictions are independent of the data
+Blindness to data is useful, but not necessary ---
+all that is necessary is that our predictions are independent of the data
 they are supposed to predict.
@@
 justifiable, and scientifically it allows our modelling to be useful if it
-makes accurate predictions.
+makes good predictions.
 If one knows what the data are, then it takes utter honesty ---
 Feynman's ``leaning over backwards''~\cite{feynman1974cargo} ---
-to assign predictions as if the data were unknown.
-Blindness is a useful and wilful ignorance that automatically implements this
-honesty without this backbreaking work.
+to assign predictions as if those data were not known.
+Blindness is therefore a useful, wilful ignorance that automatically implements
+this honesty without the (backbreaking) mental work of acting ``as if'' blind.
 

### 2022-08-11 main/chapter_2ljets.tex
proofread results and conclusion; add disco region refs

@@
 factor is studied to assign regularized variations to the Poisson expectations
-of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
+of the histogram bins with a \histfactory\ model~\cite{cranmer2012histfactory}.
 For the $\twoljets$ search, we use \histfitter\ to build and analyse those
@@
 a slow extrapolation on the low side (from the small central value to zero)
-and reduce the risk from non-physical negative yields in the HistFactory model.
+and reduce the risk from non-physical negative yields in the \histfactory\
+model.
 
@@
 
-For their implementation in the background model in HistFactory, these
+For their implementation in the background model in \histfactory, these
 variations reduce to the relative scale factors that are presented in
@@
 and to my knowledge, is only reported through inner details of
-the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}
+the \histfactory\ models shared in our \hepdata\ archive~\cite{hepdata.116034}
 (and now, I hope, this thesis).
@@
 Naturally, \atlas\ has an unblinding procedure that requires committee
-review and approval based on internal documentation.
+review and approval in accordance with internal documentation.
 In the \atlas\ SUSY world, searches are nominally performed blind to signal
 regions.
-Control and validation regions are consumed in the design and modelling of
-signal regions, so need to be inspected.
-Control and validation region data are inspected, but this is not a problem
-since they should also have negligibly small yields (contaminations) from
-interesting signal models.
+For these searches, their control and validation regions must consumed in the
+design and modelling of signal regions, so they must be inspected.
+This is not a problem since control and validation regions should also
+have negligibly small yields (contaminations) from interesting signal models,
 With negligible signal contaminations, control and validation region data can
-be consumed to constrain the background modelling in addition to all other that
-on which the Standard Model stands, and without being influenced by the signal
-models to be tested in signal regions.
-
-The \atlas\ SUSY approval procedure~\cite{atlas_twiki_susyapprove}
-requires various criteria including, and not limited to, that the search has:
+be consumed to constrain the background modelling, in composition with to all
+other data on which the Standard Model stands, and without being influenced by
+possible data from the target signal models that are to be tested in signal
+regions.
+
+Before allowing unblinding, the \atlas\ SUSY approval
+procedure~\cite{atlas_twiki_susyapprove} requires various criteria including,
+and not limited to, that a search has:
 \begin{itemize}
@@
 \end{itemize}
-So many aspects must be finalized blind, and for those that are not
-finalized blind must have procedures defined such that they are not
-influenced by the signal region data.
-Encouragingly, disallowing dependence of procedures on signal region data
-eliminates one major source of unblinding error,
+Many aspects of a search must therefore be finalized blind, and those aspects
+that are not finalized blind must have procedures defined such that they are
+not influenced by the signal region data.
+Encouragingly, this disallowance of dependence of procedures on signal region
+data eliminates one major source of unblinding error,
 and requiring knowledge of signal contaminations supports the use of control
@@
 
-% con
 Unfortunately, these requirements limit only \emph{when} data-driven estimation
-procedures are defined, and do not limit the \emph{what} those procedures do.
-Indeed, the rules explicitly allow for methods like the Matrix Method
+procedures are defined, and do not limit the \emph{what} those procedures
+\emph{do}.
+The rules explicitly allow for methods like the Matrix Method
 (described in Section~\ref{sec:2ljets_matrix_method})
-that use signal region data to affect the background `estimates' in signal
-regions in violation of blinding principles.
-No effective research would employ an estimation procedure that clearly set
-a prediction directly from the data it is supposed to predict, but our use of
+that use signal region data to determine their `estimates' in signal
+regions, in violation of basic blinding principles.
+Only the procedure is blind to data, not its action.
+No scientific research would employ an estimation procedure that clearly set
+its prediction directly from the data it is supposed to predict, but our use of
 the Matrix Method is proof by example that sufficiently obfuscated violations
-are allowed here.
-
-A second unfortunate flaw is that these rules were not strictly enforced.
+are permitted here.
+
+A second unfortunate flaw is that no mechanism enforces these rules after
+unblinding.
 Although we performed the approval procedure as written, some of its promises
-were later violated, at the request of senior committees, without consequence.
-The data-driven FNP estimation procedure should have been frozen, but was in
-fact extended with the `Mean' method after unblinding,
-as seen in Section~\ref{sec:2ljets_matrix_method_ewk}).
+were later violated on the instruction of committee decisions and without
+negative consequences.
+For example, by these rules the data-driven FNP estimation procedure should
+have been fixed, but it in fact extended with the `Mean' method after
+unblinding, as described in Section~\ref{sec:2ljets_matrix_method_ewk}.
 Theoretical uncertainties were always to be treated according to the
 \atlas\ SUSY recommendations~\cite{atlas_twiki_susytheoretical}, but those
-recommendations have ambiguities that were resolved after unblinding,
-notably for the alternative $\zjets$ sample comparison as discussed in
-Section~\ref{sec:2ljets_zjets_alt}, among other less impactful sample
-comparisons.
+recommendations have ambiguities that were resolved after unblinding.
+For example, the decision to include the various alternative sample comparisons
+was made after unblinding, and debates on our treatment of the alternative
+$\zjets$ sample comparison made direct reference to the anomalous unblinded
+signal region data, as described in Section~\ref{sec:2ljets_zjets_alt}.
+On balance, however, we are fortunate that the many post-unblinding validation
+studies found no actionable flaws. We cannot know what would have changed if a
+validation study had found reasons to further adapt our modelling to data after
+unblinding.
 
@@
 because we are in a loose selection.
-Any signal region must be within this loose selection, and those
+However, any signal region must be within this loose selection, and those
 highest bins contain only a few events that agree with the backgrounds well.
-From such plots, we knew before designing the regions that either there was no
-excess at high $\met$ or that a signal region excess must coincide with a
-deficit in its surrounding phase space (or vice versa).
-Since we fill surrounding phase space with validation regions, one could in
-fact have predicted an SR-High deficit from the subtle excess in VR-High
+From such plots, we therefore knew before designing the regions that either
+there was no excess at high $\met$ or that a signal region excess must
+coincide with a deficit in its surrounding phase space (or vice versa).
+Since we fill surrounding phase space with validation regions, an excess in a
+validation region therefore predicts a deficit in its signal region.
+One could have predicted an SR-High deficit from the excess in VR-High
 that is shown in Figure~\ref{fig:2ljets_high_mjj_vrhigh}.%
@@
 }
-If blinding is to be taken seriously, then restrictions must be placed on the
-use of such pre-selection plots.
+For blinding to be done properly, restrictions must be placed on the usage of
+such pre-selection plots.
 
@@
 Errors include all variations that were evaluated at the time.
-Data and backgrounds were later reproduced with updated software and tighter
+Data and backgrounds were later rederived with updated software and tighter
 lepton definitions, leading to a general reduction in yields.
@@
 We knew when unblinding that our simulations and data needed to be reprocessed
-with the lepton definitions given in Section~\ref{sec:2ljets_events_leptons},
-which had previously required harder baseline leptons,
-and had expected only minor changes.
+with the lepton definitions given in Section~\ref{sec:2ljets_events_leptons}.
+Our previous baseline lepton definition had required a larger $\pt$,
+and we had expected this change to make minimal differences.
 As seen from these two figures, the reprocessing reduced our yields of data
-slightly; most noticeably, the one event in SR-$\llbb$ disappeared.
-Other than aesthetic differences, post-unblinding activities have had little
-impact.
+slightly; perhaps most noticeably, the one event in SR-$\llbb$ disappeared.
+Other than aesthetic differences, and post-unblinding activities have on
+balance had relatively little impact to the post-fit results.
 The reported significance of deficits in SR-Low-a and SR-Low-b has reduced,
 and the shrinking $\zjets$ background has caused a $1$-sigma excess in VR-Low.
-Excesses have also grown in VR-High-* regions.
+Excesses have also grown in VR-High-* regions as the model has further
+adapted towards the signal region data.
 Little has changed in the SR-High-8 bins with zero data since, as explained in
-Section~\ref{sec:2ljets_diboson_alt}, the added \diboson\ sample comparison
-is ineffective.
+Section~\ref{sec:2ljets_diboson_alt}, the extra \diboson\ sample comparison
+is impotent.
 
@@
 post-fit yields that we present are misinterpreted as predictions.
-Pre-fit yields, which can plausibly be interpreted as predictions, are
-presented with other extra results in the next section.
+Pre-fit yields, which can plausibly be interpreted as more like predictions,
+are presented with other extra results in the next section.
 
@@
 \paragraph{Summary}
-As I have argued incessantly, pre-fit predictions are useful because they can
-actually be predictions to be tested by the data.
+As I have argued quite incessantly throughout this thesis, pre-fit predictions
+are useful because they can be predictions to be tested by the data.
 Our post-fit summary plot has previously been displayed in
-Figure~\ref{fig:2ljets_summary}.
-Its alternative version that does not fit to signal region data is shown here
-in Figure~\ref{fig:2ljets_summary_no_sr}.
-Since the significance measure is meaningful for this pre-signal-region plot,
-we see that despite the inflated uncertainties in SR-Low, its bins still have
-substantial deficits.
-Validation region data are notably better predicted by this prediction;
+Figure~\ref{fig:2ljets_summary}, and an alternative version that does not fit
+to signal region data is displayed in Figure~\ref{fig:2ljets_summary_no_sr}.
+The significance measure is more meaningful for this pre-signal-region plot
+since it expects pre-fit uncertainties, and we see here that despite the
+inflated $\zjets$ uncertainties in SR-Low its bins still have substantial
+deficits.
+Validation region data are notably better predicted pre-fit ---
 as the post-fit result tunes itself to signal region data, coupled variations
-lead to worse predictions in the validation regions (that are not included
-directly in either fit).
+lead to worse predictions in the validation regions since their likelihoods are
+not included directly in any fits.
 
@@
 \paragraph{Correlations}
-All parameters are initially independent.
+All parameters of our \histfactory\ models are initially independent.
 If the \heplikelihood\ is related to a probability distribution of parameters,
@@
 
-Mixed contributions of $\ttbar$, $\zjets$ and \diboson\ in the control regions
-leads to anticorrelations between their normalization factors
+In this correlation matrix, mixed contributions of $\ttbar$, $\zjets$ and
+\diboson\ in the control regions leads to anticorrelations between their
+normalization factors
 ``\texttt{mu\_*}'' that are seen in the bottom right corner of this correlation
@@
 The failure of statistical precision in SR-Low-2 causes large systematic
-variations that appear as the anticorrelations between
+variations that appear as the anticorrelations between its statistical error
 \texttt{gamma\_stat\_SRLow2\_cuts\_bin\_0} and numerous experimental
@@
 comparison ``\texttt{alpha\_Zjets\_alt}''.
+
 Note that the up and down variations of parameters need not correspond to up
-and down variations of yields in any region, so the signs of these correlations
-are no necessarily meaningful.
+and down variations of yields in any region, so the signs of these
+(anti)correlations are no necessarily meaningful.
 
@@
 and others are initially relative to $1$ with variable error bars.
-Internally, plots like this are considered good if all parameters sit on
-the central vertical line, with the motivations that normalization factors
-should be close to $1$ (the predicted cross-sections are accurate) and that
-our data should not compete with the precision of the studies that defined
-our systematic variations.
+Within the \atlas\ SUSY group, plots like this are considered good if all
+parameters sit on the central vertical line, with the motivations that
+normalization factors should be close to $1$ (the predicted cross-sections are
+accurate) and that our data should not compete with the precision of the
+studies that defined our systematic variations.
 That is, if we constrain a parameter by squeezing its error bar, then our data
@@
 
-I added the red arrow to indicate the $\zjets$ alternative sample comparison,
-named ``\texttt{$\alpha$\_Zjets\_alt}'' here, that has disappeared off the
-right-hand edge due to its best-fit value of $1.86 \pm 0.69$.
+Here, I added the big red arrow to indicate the $\zjets$ alternative sample
+comparison, named ``\texttt{$\alpha$\_Zjets\_alt}'', that has disappeared off
+the right-hand edge due to its best-fit value of $1.86 \pm 0.69$.
 (see Section~\ref{sec:2ljets_zjets_alt}).
 Its anticorrelation with the electron charge identification variations is
-explained by their movement to negative values;
+explained by their movement to negative values:
 positive \texttt{$\alpha$\_Zjets\_alt} and negative
@@
 \paragraph{Contours}
-Exclusion contours, which are displayed in
+Exclusion contours, displayed in
 Figures~\ref{fig:2ljets_contours_c1n2} and~\ref{fig:2ljets_contours_gmsb},
@@
 This is worded precisely. The dominant cost was not in fitting the models, nor
-in histogramming the data (thanks to histogram caching), but in forming the
-HistFactory model from those histograms.
+in histogramming the data (thanks to caching), but in forming the
+\histfactory\ model from those histograms.
 I do not know why that model-building step took days of computer time per
@@
 from \textsc{v0.64.0} to \textsc{v0.66.0}.
-}%
-across these grids, we prune the number of small irrelevant parameters
-with \histfitter's pruning~\cite{histfittergithub} mechanism in ``method 2''.
+}
+across the signal parameter planes, we prune the number of small irrelevant
+parameters with \histfitter's pruning~\cite{histfittergithub} mechanism in
+``method 2''.
 This pruning method removes variations if the difference between their up and
@@
 With pruning enabled, its default threshold is set to $1\%$ (relative to the
-central yield), but to reduce its impact we use a threshold of $0.1\%$.
-Validation studies comparing each of these pruning thresholds, against the
-un-pruned version, and against a custom pruning implemented by myself, found no
-significant differences.%
+central yield), but to minimize concerns about its impact we use a threshold
+of $0.1\%$.
+Validation studies that compared each of these pruning thresholds, the
+un-pruned version, and a custom pruning implemented by myself, found no
+significant differences to post-fit results.%
 \footnote{%
-Pruning is, however, the reason our signal patches (that describe differences
+Pruning is, however, the reason our signal patches (which encode differences
 from background-only to signal models) in our serialized \hepdata\ models are
-so large; they implement many small changes due to different decisions made by
-the pruning algorithm.
+so large --- they implement many small changes due to different decisions made
+by the pruning algorithm.
 }
@@
 with the \textsc{Scipy} implementation of
-Radial Basis Functions~\cite{2020SciPy-NMeth,fasshauer2007meshfree},
-interpolating significance-transformed $\cls$ values.
+Radial Basis Function~\cite{2020SciPy-NMeth,fasshauer2007meshfree}
+interpolation, applied to significance-transformed $\cls$ values.
 That is, we interpolate in $\cls$ values transformed through the inverse
 cumulative distribution of the standard normal distribution.
-Contours are drawn where this interpolated function meets
+Contours are drawn where this interpolated function meets the equivalent of
 $\cls = 0.05$.
@@
 
-This interpolation algorithm has some tunable parameters, for which all
-contours of the $\twoljets$-electroweak search use
+This interpolation algorithm has some tunable parameters, for which after some
+experimentation I decided that all contours of the $\twoljets$-electroweak
+search use
 \texttt{epsilon=10},
@@
 and the ``multiquadric'' radial basis function.
-These settings were tuned to keep the thumbs-up wobble in the C1N2 off-shell
+These settings were tuned to keep the `thumbs-up' wobble in the C1N2 off-shell
 region of Figure~\ref{fig:2ljets_contours_c1n2},
-which is physical and due to our increased sensitivity for mass splittings
-below $m(W)$,
+which is physical and due to our increased sensitivity for some mass splittings
+below $m(Z)$,
 while sufficiently smoothing wobbles elsewhere.
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_c1n2_Output_Contours_cls_obs.pdf}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_gmsb_Output_Contours_cls_obs.pdf}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Exclusion contours with observed $\cls$ values
+]{%
+Exclusion contours with observed $\cls$ values.
+The observed exclusion contour is interpolated from these values.
+Parameter points are at the lower left corners of the text representing each
+$\cls$ value.
+}
+\label{fig:2ljets_fit_contours_cls}
+\end{figure}
+
 Our initial contours showed more glaring and unphysical wobbles.
-Those erroneous wobbles were removed by repeating the fits of anomalous signal
-points until they behaved.
+Those erroneous wobbles were removed by repeating the exclusion fits for signal
+points with anomalous results until their results appeared to behave.
 In total, this hack was used for two points near the highest
@@
 GMSB spectrum.
-Different random seeds causing failed fits explain this unstable behaviour,
-and I failed to control those random seeds when running the fits through
-\histfitter.
+Different random seeds in the optimization algorithms explain this unstable
+behaviour; they clearly affect the fit results, and my attempts control all
+random seeds failed when executing these fits through \histfitter.
 
@@
 labelled as excluded;
-this is direct a result of our smoothing choices.
-
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_c1n2_Output_Contours_cls_obs.pdf}
-\caption{C1N2}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_gmsb_Output_Contours_cls_obs.pdf}
-\caption{GMSB}
-\end{subfigure}
-\caption[
-Exclusion contours with observed $\cls$ values
-]{%
-Exclusion contours with observed $\cls$ values.
-The observed exclusion contour is interpolated from these values.
-Parameter points are at the lower left corners of the text representing each
-$\cls$ value.
-}
-\label{fig:2ljets_fit_contours_cls}
-\end{figure}
+this is direct a result of our artistic interpolation choices.
 
@@
 
-Perhaps what should be the most meaningful result is the observed upper limit
-$S_{\mathrm{obs}}^{95}$.
-The analysis of each region sets its upper limit by scanning different values
+What should berhaps be the most meaningful result of each discovery region is
+is observed upper limit $S_{\mathrm{obs}}^{95}$.
+The analysis of each region sets this upper limit by scanning different values
 of the signal expectation $\mu_\textrm{discovery}$ and interpolating to
@@
 believed to work.
-For consistency, we therefore use the alternative toy simulation
-approximation for all discovery regions.
+For consistency between our five discovery regions, we therefore use the
+alternative toy simulation approximation for all of them.
 Sampling of the test statistic distributions
@@
 As discussed previously, this makes negligible differences to the results.
-The exact pruned models are shared in our \hepdata\
-archive~\cite{hepdata.116034}.
-Not only do these simulations take computational time, they also suffer
+The exact \histfactory\ models used in our discovery fits are shared in our
+\hepdata\ archive~\cite{hepdata.116034}.%
+Although many other searches' discovery regions can be approximately reproduced
+from their published \pyhf\ models of signal regions for exclusion fits,
+$\twoljets$ is the first \atlas\ search to directly publish its exact discovery
+region models on \hepdata~\cite{
+SUSY-2019-08,
+hepdata.90607,
+SUSY-2018-10,
+hepdata.97041,
+SUSY-2018-04,
+hepdata.92006,
+SUSY-2018-32,
+hepdata.89413,
+atlas2022searches,
+hepdata.116034,
+atlas_rjr_3l_SUSY_2019_09,
+hepdata.95751,
+SUSY-2018-36,
+hepdata.99806,
+atlas_rjr_mimic_SUSY_2018_06,
+hepdata.91127,
+SUSY-2018-09,
+hepdata.91214,
+SUSY-2018-02,
+hepdata.103062,
+atlas_susy_compressed_2l_2018_run2,
+hepdata.91374,
+SUSY-2018-33,
+hepdata.91760,
+SUSY-2018-31,
+hepdata.89408,
+atlas_susy_strong_0l,
+hepdata.95664
+}.
+
+Not only do these simulations take computational time, they suffer
 torrential memory leaks that halt progress beyond small sample sizes.
@@
 To share software that assists in spending computational resources to
-calculate statistics that are so deeply flawed causes me some shame.
+calculate statistics of this kind causes me some pain.
 In a vain attempt to absolve myself of this sin, and to positively encourage
-honest reporting of the data, I include with it the following note:
+honest reporting results, I include with it the following note:
 \emph{%
@@
 Despite this workaround, the simulations took several weeks of parallel
-processing on around $1500$ processor cores, accumulating to around a million
-core hours.
+processing on around $1500$ processor cores, accumulating to more than a
+million core hours.
 
@@
 
-Many test statistic distributions are involved here --- two (background-only
-and signal-plus-background) at each scanned point, and one additional
-distribution per region (with yet another test statistic) for the discovery
+Many test statistic distributions are involved in these results ---
+two (background-only and signal-plus-background) at each scanned point,
+and one additional distribution per region for the discovery
 $p$-value $p_{s\!=\!0}$.
-From these, a few examples are presented in
+Some examples of these distributions are displayed in
 Figure~\ref{fig:2ljets_test_stat_distributions}
-to investigate curious features of the results.
+to help investigate curious features of the results.
 
@@
 for DR-High is rounded up from a result of $2.93$ that was evaluated with these
-simulation methods.
+toy simulation methods.
 This makes sense --- the limit for $n=0$ should arguably always be $3$ since
@@
 background-only and signal-plus-background models is $0.05$ for $\mu = 3$.
-By taking a ratio of $p$-values, $\cls$ approximates that likelihood ratio:
+And by taking a ratio of $p$-values for zero data, $\cls$ approximates that
+likelihood ratio:
 we see from Figure~\ref{fig:2ljets_test_stat_distributions_drhigh} that the
 observed data does land in a spike at the end of the distribution that
-corresponds to pseudo-data with $n=0$; the peaks to its left correspond
+corresponds only to pseudo-data with $n=0$; the peaks to its left correspond
 to $n=1$, $n=2$ and so on, with widths due to the other parameters varied in
@@
 minimized), but its distribution is not quite a delta function since the test
-statistic depends on the noisy termination of numerical optimization
-algorithms.
-Resulting $\clb$ and $\clsb$ values depend on this effect ---
+statistic depends not exactly on the theoretical best fit, but on the noisy
+termination of numerical optimization algorithms.
+The resulting $\clb$ and $\clsb$ values depend on this effect ---
 with $3.9$ expected background events and zero data we expect
@@
 of that $n=0$ noise distribution.
-This small $\clb$ value could appear to be an anomalously large significance
-if that $p$-value were  misinterpreted as a measure of compatibility with the
+This tiny $\clb$ $p$-value could appear to be an anomalously large
+significance if one misinterpreted it as a measure of compatibility with the
 background model.
-Thankfully, it is remedied by the $\cls$-prescribed ratio with $\clsb$,
-since both $p$-values suffer from similar effects.
+Thankfully, this problem is somewhat remedied by the $\cls$-prescribed ratio
+against $\clsb$ since both $p$-values suffer from similar effects.
 Statistical precision is harmed, however --- although $1$m samples should
@@
 The most extreme is $\mathrm{SR}3\ell_{[90,110]}$
-of~\cite{SUSY-2018-36}, which reports a background expectation of
-$1.08 \pm 0.18$ and expected limits of
+of the trilepton resonance search~\cite{SUSY-2018-36}, which reports a
+background expectation of $1.08 \pm 0.18$ and expected limits of
 $S_{\mathrm{exp}}^{95} = 3.8^{+1.3}_{-0.5}$, but zero observed data with a
 whopping $S_{\mathrm{obs}}^{95} = 4.8$.
-Using their \pyhf\ models published through \hepdata, I have failed to
-reproduce this effect outside the \histfitter-\roofit\ system, finding instead
-a limit close to $3$. It would not surprise me if similar complications with
-the test statistic distribution played a part.
+Using their \pyhf\ models published through \hepdata~\cite{hepdata.99806},
+I have failed to reproduce this effect outside the \histfitter-\roofit\ system,
+finding instead a limit close to $3$.
+It would not surprise me if similar complications with the test statistic
+distribution played a part.
 
@@
 Another public example of this effect is
-$\mathrm{SR}0\textrm{-}\mathrm{ZZ^{tight}_{bveto}}$ of~\cite{SUSY-2018-02}.
+$\mathrm{SR}0\textrm{-}\mathrm{ZZ^{tight}_{bveto}}$ of the four-lepton
+search~\cite{SUSY-2018-02}.
 \emph{Thanks to Tina~Potter discussions of this problem and for sharing
@@
 Limits moving in the wrong directions can simply be explained by flawed
-test statistics ranking data in a bad order.
+test statistics ranking data in the wrong order.
 
 A second sufficient explanation for these paradoxical limit results is also
-that the toy sampling method does not assign legitimate $p$-values because its
+that the toy sampling method does not assign legitimate $p$-values, since its
 test statistic distributions are assigned by first fitting to the
@@
 employ~\cite{Cowan:2010js} compare a background model against a strictly more
-flexible signal model, and should therefore, to my understanding, be always
-non-negative.
+flexible signal model, so should (to my understanding), be always non-negative.
 The observed negative values are therefore quite unsettling, particularly since
@@
 the numerical optimization.
-Histogram ranges in Figure~\ref{fig:2ljets_test_stat_distributions} are
+
+Histogram ranges for all plots in
+Figure~\ref{fig:2ljets_test_stat_distributions} are
 set from the smallest and largest toy test statistics, and their large extent
-indicates rare outliers that may also result from failed fits.
-
-It is certainly arguable that the pathological test statistic distributions
+can also be explained by failed fits causing rare outliers.
+
+It is arguable that the pathological test statistic distributions
 observed in DR-Low indicate a bug that should be fixed.
@@
 --- \roofit\ through \histfitter\ ---
-define in silicon the standard statistical methods for \atlas\ SUSY searches.
-To diverge from them would make our result incomparable.
+define in stone (silicon) the standard statistical methods for \atlas\ SUSY
+searches.
+To diverge from them would make our result incomparable with all preceding
+work.
 What they do defines what is correct.
@@
 
-From \atlas' large and valuable Run~2 dataset, our $\twoljets$ searches
+From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
 have extracted a sharper image of the high energy processes that simultaneously
@@
 
-These searches do not exist in isolation; they extend and strengthen the
+These searches do not exist in isolation --- they extend and strengthen the
 results of numerous previous \atlas\ searches that have used similar or smaller
@@
 By providing refined and updated interpretations, we have advanced \atlas' SUSY
-research programme in this $2\ell$+jets+$\met$ channel.
-My focus has been on the $\twoljets$-electroweak search within this project.
-Its result is a central pillar of the visible effects of both C1N2 and GMSB
-simplified supersymmetric models, and uniquely contributes exclusion
-(under the $95\%$ $\cls$ prescription) of certain regions of the models'
-parameter spaces.
+research programme in this two-lepton plus jets plus $\met$ channel.
+My focus has been on the $\twoljets$-electroweak search within this project,
+and its result tests a central pillar of the visible effects of both C1N2 and
+GMSB simplified supersymmetric models, and uniquely contributes exclusion
+contours in certain regions of the models' parameter spaces.
 
@@
 statistical practices should not prevent us from using large numbers of small
-signal regions, their overly ambitious design in the $\twoljets$-electroweak
-did come with costs:
+signal regions, their overly ambitious design for the $\twoljets$-electroweak
+regions did incur costs:
 % mm fakes
@@
 (Section~\ref{sec:2ljets_sample_comparisons})
-and struggles to reconcile their standard prescriptions for coupled
+in struggles to reconcile their standard prescriptions for coupled
 (correlated) variations with physical intuition between our many bins,
 % computational
-and computational costs particularly in the `toy' simulations that assign our
-statistical results (Section~\ref{sec:2ljets_results}).
-The root of this problem that in diverging from standard designs we
+and in computational costs particularly in the `toy' simulations that assign
+our statistical results (Section~\ref{sec:2ljets_results}).
+The root of this problem is that by diverging from standard designs we
 had to use tools and rules outside the domains that their designers
@@
 were observed in our complicated and rigidly-modelled signal regions.
-Although these regions are sensitive, as demonstrated by the extent of their
-projected and observed exclusion contours, delays are themselves a meaningful
-cost.
+Although these regions are sensitive to new signals, as demonstrated by the
+extent of their projected and observed exclusion contours, delays are
+themselves a meaningful cost.
 Simpler designs, with fewer and larger regions, could be cheaper to model with
@@
 sacrificing much sensitivity.
-Simplicity in design may therefore have afforded a higher quality result by
-reducing these delays and publishing its result years sooner,
-and so implemented \cern's goal of
+Greater simplicity in design may therefore have afforded a higher quality
+result by reducing these delays and publishing its result years sooner,
+and so might have reached towards \cern's goal of
 ``accelerating science''~\cite{Radford:1551933}.

### 2022-08-11 main/chapter_2ljets.tex
autocheck

@@
 A more ordinary ordering would first present preparations made in the search
-project -- blind to data ---
+project --- blind to data ---
 followed later by the results --- unblinded.
@@
 The ``Top'' category contains $\ttbar$ and $tW$ processes, and
-``Other'' contains fake/non-prompt lepton, higgs, triboson, $\ttbar Z$, and
+``Other'' contains fake/non-prompt lepton, Higgs, triboson, $\ttbar Z$, and
 \topother\ processes.%
@@
 Values with errors are rounded to one or two significant figures according
-the the PDG rounding scheme~\cite{pdg2020review}.%
+the PDG rounding scheme~\cite{pdg2020review}.%
 }
@@
 Many plotting scripts, however, are modified from versions inherited from
-many past and present \atlas\ members to whom I am grateful.
+many past and present \atlas\ members, to whom I am grateful.
 
@@
 \label{sec:2ljets_origins_strong}
-Strong supersymmetry, producing super-partners of quarks and gluons, would be
-produced at much larger cross-sections than electroweak supersymmetry at the
+Strongly supersymmetric particles, the super-partners of quarks and gluons, would
+be produced at much larger cross-sections than their electroweak cousins at the
 LHC.
@@
 Additionally, $\twoljets$ widens its scope to target a GMSB simplified model
-illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
+that is illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
 The C1N2 and GMSB models behave similarly since both feature gaugino decays via
@@
 This section describes the events selected for the three $\twoljets$ searches,
-the reconstructed particle objects derived from those events and various
+the reconstructed particle objects derived from those events, and various
 important event variables.
@@
 When living our lives as humans, we are bombarded with sensory data and must
-pay attention only those gems few of any use, to avoid exhausting our
+pay attention only to those few gems of any use, to avoid exhausting our
 processing capacity.
@@
 $\twoljets$, by using only un-prescaled triggers.
-We use use only the `lowest' (loosest, highest rate) triggers selecting two
+We use only the `lowest' (loosest, highest rate) triggers that select two
 charged leptons~\cite{atlas_twiki_lowest_unprescaled}.
@@
 have increased in hardness with time.
-The point of listing these is to demonstrate trigger activations are
+The point to listing these requirements is to show trigger activations are
 non-trivial.
@@
 
-Once we know about all of the reconstructed leptons and hadronic jets, we can
+Once we know about all the reconstructed leptons and hadronic jets, we can
 also improve our rejection of non-prompt leptons from hadron decays, since
@@
 The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
-Please note that $\met$ is not an energy; transverse energy does not exist.
+Note that despite its $E$, $\met$ is not an energy; transverse energy does not exist.
 One may, of course, define a directed `transverse energy', but it will
@@
 
-Signal models in $\twoljets$ generate hard, invisible particles.
+All signal models in $\twoljets$ generate hard, invisible particles.
 By selecting for large $\met$, however, we select events in which either there
@@
 All assignments performed similarly, and all left similar amounts of signal
-outside of their selected mass windows.
+outside their selected mass windows.
 The simplicity of, and precedent for, the two-hardest jet assignment therefore
@@
 to Ben Hooberman for suggesting the investigation of massive jets,
-and the the whole $\twoljets$ team for collaboration in their finalization.
+and to the whole $\twoljets$ team for collaboration in their finalization.
 }
@@
 
-The control region \crvz\ occupies the $\mjj$ side-band to
+The control region \crvz\ occupies the $\mjj$ sideband to
 SR-Int and VR-Int.
@@
 from mismeasurements or rare interactions, is no longer eliminated.
-Backgrounds from $\ttbar$ similarly increase and the selections inherited from
+Backgrounds from $\ttbar$ similarly increase, and the selections inherited from
 Intermediate regions are no longer sufficient.
@@
 $\dphillmet$, and backgrounds do not; this motivates the $\dphillmet < 0.6$
-requirement, and can be plausibly understood from scenarios where a the boosted
+requirement, and can be plausibly understood from scenarios where a boosted
 supersymmetric resonance decays to $Z+\mathrm{invisible}$ and sends both
@@
 The distribution $\zjets$ here may be surprising, since its $\ptmiss$ is
-usually expected to align with a mismeasured jet.
+expected to usually align with a mismeasured jet.
 This is explained by the context of the $\mttwoll < 80\,\eV[G]$
@@
 To get sufficiently large backgrounds, the validation region VR-Low-2 uses
-the $\mjj$ side-band, and sees good agreement.
+the $\mjj$ sideband, and sees good agreement.
 
@@
 hasty redesign shortly before unblinding.
-Before that redesign, an multi-jet signal region SR-Low-ISR was included,
+Before that redesign, a multi-jet signal region SR-Low-ISR was included,
 with associated validation and control regions
@@
 unlike our other regions, this construction does not select the two hardest for
-the $W$ candidate, , but instead takes the two closest in $\phi$ to the
+the $W$ candidate, but instead takes the two closest in $\phi$ to the
 vector $\vec \pt^{\ell\ell} + \ptmiss$.
@@
 
-Early OffShell designs used the `ISR' construction that was described
+Early OffShell designs used the `ISR' construction that was described in
 Section~\ref{sec:2ljets_low} for Low regions,
@@
 For each contour, fits are performed with only the control regions and one
-subset of the signal regions as labelled and described in the figure caption.
+subset of the signal regions, as labelled and described in the figure caption.
 
@@
 Given our existing division into High, Low, Int, and OffShell
-categories, the assignment one discovery region to each was simple.
+categories, the assignment of one discovery region to each was simple.
 The $\llbb$ category within High also stands out for targetting physically
@@
 
-Our five discovery regions are defined as unions on signal regions as follows:
+Our five discovery regions are defined as unions of signal regions as follows:
 \begin{itemize}
 \item DR-High is SR-High-8 (the union of SR-High-8-a and SR-High-8-b),
-\item DR-$\llbb$ is is \srllbb,
+\item DR-$\llbb$ is \srllbb,
 \item DR-Int is SR-Int (the union of SR-Int-a and SR-Int-b),
@@
 
-This section only covers some of the the most peculiar or important aspects
+This section only covers some of the most peculiar or important aspects
 of our signal and background modelling, and does not specify most exact
@@
 exact.
-More accurately, the two are are related through a Poisson likelihood function
+More accurately, the two are related through a Poisson likelihood function
 that alternative FNP estimation methods can use~\cite{
@@
 in most regions.
-The exceptions to this rule are all regions the High category,
+The exceptions to this rule are all regions in the High category,
 which have small enough expected yields for the high variation of the Mean
@@
 detector response and does not guarantee to accurately model the substructures
-that determine jets reconstructed masses.
+that determine the jets' reconstructed masses.
 Jet mass modelling should be well described since masses are part of the
@@
 large-radius jets with $R=0.8$, since these require less extreme boosts to
-capture their targets, and he focus from \atlas' performance operations is
+capture their targets, and the focus from \atlas' performance operations is
 therefore reasonable to understand these common use-cases first.
@@
 across the $\mjetone \in 60\textrm{--}110\,\eV[G]$ window.
-On consideration of these results and similar plots in tighter selections ,
+On consideration of these results and similar plots in tighter selections,
 we assign a $15\%$ relative variation to all \afii\ samples in 1J regions
@@
 This effect would yield an underestimated uncertainty, so does not invalidate
-the comparison approach but highlights that it is a bare minimum and that
+the comparison approach, but highlights that it is a bare minimum and that
 larger uncertainties may well be reasonable given our imprecise constraints on
@@
 This \madgraph\ alternative sample simulates only $\mll > 40\,\eV[G]$, so in
-regions allowing $\mll$ outside of this we evaluate relative variations and
+regions allowing $\mll$ outside this we evaluate relative variations and
 scale them up to the central yield.
@@
 Table~\ref{tab:2ljets_zjets_alt}.
-Note that up and down variations of the sample comparison parameter do not
-necessarily mean up and down variations of bin yields, but to the coupled
-variations in directions given by the signs of the relative variations given in
-the various regions.
+Note that up and down variations of a parameter do not
+necessarily mean up and down variations of bin yields.
+Coupled variations move in directions given by the signs of their relative
+variations, which are listed for the various regions.
 These variations are evaluated according to a standard recipe prescribed by
-the ATLAS SUSY group~\cite{atlas_twiki_susytheoretical} that says to
+the \atlas\ SUSY group~\cite{atlas_twiki_susytheoretical} that says to
 \begin{itemize}
@@
 $\zjets$ contributions.
-In other regions the $\zjets$ sample and its variations are negligible.
+In other regions, the $\zjets$ sample and its variations are negligible.
 The `Central' column shows the central $\zjets$ yield from the main sample.
@@
 we do not expect large $W\rightarrow jj$ contributions in this \diboson\
-process since one boson is needed for the leptons and the other is needed for
+process, since one boson is needed for the leptons and the other is needed for
 the $\met$ that we select.
@@
 This limitation is simply a result of how \atlas\ has decided to allocate
-its limited resources on, and quite reasonably so since full simulation is
+its limited resources on, and quite reasonably so, since full simulation is
 computationally expensive and \diboson\ processes are relatively rare.
@@
 cannot conserve four-momentum when splitting into two non-collinear massless
-children (for example gluon emission $q\rightarrow gq$);
+children (for example, gluon emission $q\rightarrow gq$);
 each scheme achieves conservation by recoiling off different parts of the
@@
 Truth-level background yields in our regions were calculated with the
-SimpleAnalysis~\cite{Lorenz:2752450,simpleanalysis_cern}
+\simpleanalysis~\cite{Lorenz:2752450,simpleanalysis_cern}
 software package for
 our central \diboson\ samples and these varied alternatives.
-SimpleAnalysis histograms truth-level simulations and optionally smears them
+\simpleanalysis\ histograms truth-level simulations and optionally smears them
 with calibrated noise to approximate reconstruction-level results.
 \emph{Thanks to Jon~Long for preparing our initial
-SimpleAnalysis implementations and sharing software to evaluate these
+\simpleanalysis\ implementations and sharing software to evaluate these
 variations.}
@@
 
-This assignment varies \crvz\ and on-shell other regions identically,
+This assignment varies yields in \crvz\ and other on-shell regions identically,
 and since the \diboson\ background is normalized to \crvz\ this means that
@@
 so to reduce costs we use truth-level only
-(through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
+(through \simpleanalysis~\cite{simpleanalysis_cern}) and evaluate the
 variations at only a handful of benchmark signal models.
@@
 members in our internal documentation~\cite{twoljets2018int}.
-This section presents some highlights these validation studies.
+This section presents some highlights of these validation studies.
 
@@
 or outside our angular acceptance.
-(Overlap removal cannot explain this since we count baseline leptons before
+(Overlap removal cannot explain this, since we count baseline leptons before
 applying overlap removal.)
@@
 These histograms are therefore split on charged lepton number in their
-final states and we cannot exactly attribute the intermediate boson states.
+final states, and we cannot exactly attribute the intermediate boson states.
 
@@
 regions.
-For these searches, their control and validation regions must consumed in the
+For these searches, their control and validation regions are used in the
 design and modelling of signal regions, so they must be inspected.
@@
 With negligible signal contaminations, control and validation region data can
-be consumed to constrain the background modelling, in composition with to all
+be consumed to constrain the background modelling, in composition with all
 other data on which the Standard Model stands, and without being influenced by
@@
 Other than aesthetic differences, and post-unblinding activities have on
-balance had relatively little impact to the post-fit results.
+balance had relatively little impact on the post-fit results.
 The reported significance of deficits in SR-Low-a and SR-Low-b has reduced,
@@
 All parameters of our \histfactory\ models are initially independent.
-If the \heplikelihood\ is related to a probability distribution of parameters,
+If the \heplikelihood\ relates to a probability distribution of parameters,
 however, then they gain interdependencies when coupled through the likelihoods
@@
 un-pruned version, and a custom pruning implemented by myself, found no
-significant differences to post-fit results.%
+significant differences in post-fit results.%
 \footnote{%
@@
 Different random seeds in the optimization algorithms explain this unstable
-behaviour; they clearly affect the fit results, and my attempts control all
+behaviour; they clearly affect the fit results, and my attempts to control all
 random seeds failed when executing these fits through \histfitter.
@@
 
-What should berhaps be the most meaningful result of each discovery region is
+What should perhaps be the most meaningful result of each discovery region is
 is observed upper limit $S_{\mathrm{obs}}^{95}$.
@@
 In a vain attempt to absolve myself of this sin, and to positively encourage
-honest reporting results, I include with it the following note:
+honest reporting of results, I include with it the following note:
 \emph{%
@@
 Thankfully, this problem is somewhat remedied by the $\cls$-prescribed ratio
-against $\clsb$ since both $p$-values suffer from similar effects.
+against $\clsb$, since both $p$-values suffer from similar effects.
 Statistical precision is harmed, however --- although $1$m samples should
@@
 search~\cite{SUSY-2018-02}.
-\emph{Thanks to Tina~Potter discussions of this problem and for sharing
+\emph{Thanks to Tina~Potter for discussions of this problem and for sharing
 its details of its modelling.}
@@
 our statistical results (Section~\ref{sec:2ljets_results}).
-The root of this problem is that by diverging from standard designs we
+The root of this problem is that by diverging from standard designs, we
 had to use tools and rules outside the domains that their designers

### 2022-08-15 main/chapter_2ljets.tex
tweaks for Chris' excellent feedback; fix y label on drhigh n0 plot

@@
 and again after;
-search project has many little and loosely-related pieces that I attempt to
+this search project has many little and loosely-related pieces that I attempt
 to cover locally to reduce the reader's need to jump between far-flung corners
@@
 we add signal contributions on top of the modelled backgrounds.
-We use two signal models in the $\twoljets$-electroweak analysis.
+
+The $\twoljets$-electroweak search uses two signal models,
+which we name C1N1 and GMSB and illustrate with diagrams in
+Figure~\ref{fig:2ljets_signal_diagrams}.
 One is an MSSM chargino-neutralino model (C1N2) that acts to produce
@@
 and invisible gravitinos.
-Diagrams illustrating these C1N2 and GMSB signal models are shown in
-Figure~\ref{fig:2ljets_signal_diagrams}.
 Example effects of these signal models are displayed in signal regions of the
@@
 Both signal models have two free parameters that our search varies.
-At each parameter point of a signal model, it is added to the background model
-and tested against data in the $\cls$ prescription.
-Contours showing which parameters of those signal models are labelled as
-excluded are displayed for
-C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
-GMSB in~\ref{fig:2ljets_contours_gmsb}.
+At each parameter point of a signal model, the search proceed by evaluating
+event rates (yields) from that signal model with those parameters,
+and considering what would happen of those event yields were added to a
+Standard Model background.
+Each signal-plus-background model is then tested against observed data in the
+$\cls$ prescription.
+From these tests, contours showing which parameters of those signal models are
+labelled as excluded are derived, and displayed in
+Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and
+in~\ref{fig:2ljets_contours_gmsb} for GMSB.
 In both of these figures, space to the lower left of the contours is
@@
 The observed contours are somewhat larger than expected due to observed
-deficits in data.
+deficits in data that we have shown in Figure~\ref{fig:2ljets_summary}.
 
@@
 neighbours show alternative results that are obtained if all signal
-cross-sections are varied up and down by theoretical error bars.
+cross-sections are varied up and down by factors intended to describe the size
+theoretical uncertainties.
 \\[0.5em]
@@
 The same data deficits drive stronger bounds on possible signals that manifest
-as smaller signal expectations limits $S_{\mathrm{obs}}^{95}$.
+as smaller limits on the signal expectations, which are labelled
+$S_{\mathrm{obs}}^{95}$ in the figure.
 
 % upper limits
-% definitely want this after pictures
-\FloatBarrier
 \begin{table}[tp]
@@
 the observed $95\%$ $\cls$ upper limit on the visible cross-section
-$\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
+$A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$,
 its corresponding signal expectation $S_\mathrm{obs}^{95}$,
@@
 $\clb$ evaluated with the signal expectation set to the observed upper limit,
-and the discovery $p$-value (capped at $0.5$) with its equivalent significance.
+and the discovery $p$-value capped at $0.5$.
+The cross-section and signal yield limits are related by a factor of our
+integrated luminosity $139.0\,\mathrm{fb}^{-1}$.
 \\[0.5em]
@@
 All $p$-values are estimated by ``toy'' simulations of alternative data.
-Values with errors are rounded to one or two significant figures according
-the PDG rounding scheme~\cite{pdg2020review}.%
+Values with errors are rounded according the PDG rounding
+scheme~\cite{pdg2020review}.%
 }
@@
 Leveraging the increased data quantity is one key contribution of $\twoljets$.
-Analysis quality also might be improved if our designs take lessons from
-the experiences of previous work.
+How well those data are used depends on how well the search designed, so
+$\twoljets$ further benefits from lessons learned from the experiences of
+past analyses.
 
@@
 Disappearing excesses have encouraged some doubts about RJR variables in
-certain social circles.
-The existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06},
-which approximates the $3$ lepton result of~\cite{atlas_rjr_23l_SUSY_2017_03}
+certain social circles;
+the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06},
+which approximates the $3$-lepton result of~\cite{atlas_rjr_23l_SUSY_2017_03}
 with non-RJR ``laboratory-frame'' event variables,
@@
 Although RJR variables can appear to offer appealing separations of
-supersymmetric signals from Standard Model backgrounds,
-these practical issues motivate my decision to use any RJR event variables in
-the $\twoljets$-electroweak analysis.
+supersymmetric signals from Standard Model backgrounds, these practical issues
+motivate my decision to not use any RJR event variables in the
+$\twoljets$-electroweak analysis.
 
@@
 
-The $\twoljets$-strong search targets strongly interacting supersymmetric
-effects in cases where their decays include pairs of charged leptons.
+The $\twoljets$-strong search is designed to target strongly interacting
+supersymmetric effects in cases where their decays include pairs of charged
+leptons.
 Like the RJR and electroweak $\twoljets$ searches, signal models targeted by
@@
 and the others are binned in $\mll$ for sensitivity to the various
-kinematic edges across the grid of signal model parameters.
+kinematic endpoints that arise across the grid of various signal model
+parameters.
 
@@
 produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
-equal rates ---
-the $\twoljets$ search exploits this difference in flavour symmetry to derive
+equal rates.
+The $\twoljets$ search exploits this difference in flavour symmetry to derive
 a data-driven background estimate for all of its flavour-symmetric backgrounds,
@@
 Personally, I find these jagged contours hard to interpret.
-If I had a pet theory that landed on one of these sharp
+If I were interested in a theory that landed on one of these sharp
 region-transition edges, what should I make of these data?
@@
 expectations~\cite{skilling2009cosmology}, and in the small-bin limit one has
-continuous distributions with Poisson normalizations
+continuous of `unbinned' distributions with Poisson normalizations
 (sometimes called ``extended'' in high-energy
 physics~\cite{barlow1990extended})
-that are standard elsewhere in high energy physics, and are easily analysed
-by the standard likelihood fitting methods.
+that are standard elsewhere in high-energy physics,
+and are easily analysed by the standard likelihood fitting
+methods~\cite{cowan1998statistical}.
+Examples uses of unbinned modelling include analyses of hadronic
+resonances in
+\lhcb~\cite{Tullythesis,lhcb2018563}, searches for exotic effects in
+\atlas~\cite{EXOT-2018-08},
+and notably the observation of the
+Higgs boson~\cite{HIGG-2012-27}.
 
@@
 For example, the top quark was observed by \dzero\ using seven regions all with
-background expectations of $1$ or less, for a total of $3.79\pm0.55$ expected
-and $17$ observed~\cite{abachi1995observation}.
+background expectations of one event or less, for a total of $3.79\pm0.55$
+expected and seventeen observed~\cite{abachi1995observation}.
 If they could do that with technology from the year $1995$, we can do it now.
@@
 uncertainties in the smaller regions.
-Various dubious workaround are invoked to assign plausible variations despite
-small yields, which we shall in the remainder of this chapter.
+Various workaround are invoked to assign plausible variations despite
+small yields and failing statistical precision, which shall be seen in the
+remainder of this chapter.
 As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
@@
 all-hadronic~\cite{SUSY-2018-41},
-$2/3\ell$ compressed~\cite{atlas_susy_compressed_2l_2018_run2},
+$2\ell/3\ell$ compressed~\cite{atlas_susy_compressed_2l_2018_run2},
 $3\ell$ on-shell~\cite{SUSY-2018-06}, and
@@
 
-From its name, it is clear that $\twoljets$ analyses features of events with
-two charged leptons (those are electrons or muons --- not taus),
-and at least one jet.
-Though not in the name, a $\met$ requirement is also implicit from the context
+The `$\twoljets$' name is intended to indicate interest in events with two
+charged leptons (those are electrons or muons --- not taus), and at least one
+jet.
+Though not in the name, an $\met$ requirement is also implicit from the context
 of $R$-parity conserving supersymmetric models.
-This section serves to describe how events are selected for the $\twoljets$
-analyses, and which event variables are extracted for use in its selections.
+This Section~\ref{sec:2ljets_events} serves to describe how events are selected
+for the $\twoljets$ analyses, and which event variables are extracted for use
+in its selections.
 
@@
 \label{sec:2ljets_trigger}
-There would be no event variables if not for the trigger reducing the data to
-a manageable size.
+There would be no event variables without events, and \atlas\ only records
+when triggered.
 When living our lives as humans, we are bombarded with sensory data and must
@@
 processing capacity.
-Running \atlas\ is no different.
+Running \atlas\ is no different --- its triggering is necessary to reduce the
+bombardment of collision data to a manageable size.
 
@@
 are summarized here.
-Please do not read them all in detail.
+Please do not read them all in detail ---
+the point of this listing is to show trigger activations are varied and
+non-trivial.
 \begin{itemize}
-\item Electron pair: requirements vary between
-\begin{itemize}
-\item $\pt^{e_2} > 12\,\eV[G]$ and
-\item $\pt^{e_2} > 17\,\eV[G]$.
+\item Electron pair: requirements vary between $\pt^{e_2} > 12\,\eV[G]$ and
+$\pt^{e_2} > 17\,\eV[G]$.
+\item Muon pair: various requirements include
+$\pt^{\mu_2} > 10\,\eV[G]$,
+$\pt^{\mu_2} > 14\,\eV[G]$,
+$\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
+$\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
+\item Emu pair: various asymmetric requirements include
+$\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
+$\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
+$\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
 \end{itemize}
-\item Muon pair: various requirements include
-\begin{itemize}
-\item $\pt^{\mu_2} > 10\,\eV[G]$,
-\item $\pt^{\mu_2} > 14\,\eV[G]$,
-\item $\pt^{\mu_1} > 18\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$, and
-\item $\pt^{\mu_1} > 22\,\eV[G]~\&~\pt^{\mu_2} > 8\,\eV[G]$.
-\end{itemize}
-\item Emu pair: various asymmetric requirements include
-\begin{itemize}
-\item $\pt^{e_1} > 17\,\eV[G]~\&~\pt^{\mu_1} > 14\,\eV[G]$,
-\item $\pt^{e_1} > 7\,\eV[G]~\&~\pt^{\mu_1} > 24\,\eV[G]$, and
-\item $\pt^{e_1} > 26\,\eV[G]~\&~\pt^{\mu_1} > 8\,\eV[G]$.
-\end{itemize}
-\end{itemize}
-Throughout the data used in $\twoljets$, trigger requirements
-have increased in hardness with time.
-The point to listing these requirements is to show trigger activations are
-non-trivial.
-This motivates lepton selections which exceed these trigger thresholds,
-to avoid the need to model their jagged activations and variable
-efficiencies near to those thresholds~\cite{
+Trigger requirements have increased in hardness with time throughout the data
+used in $\twoljets$.
+
+This complicated collection of trigger requirements motivates lepton selections
+that exceed these trigger thresholds,
+such that they avoid the need to accurately model their jagged activation
+thresholds and the variable efficiencies near to those thresholds~\cite{
 atlas_trigger_egamma_run2,
@@
 }.
-To avoid these thresholds, all leptons selected for the $\twoljets$-electroweak
-regions are required to have $\pt > 25\,\eV[G]$.
+To avoid the most of these thresholds, all leptons selected for the
+$\twoljets$-electroweak regions are required to have $\pt > 25\,\eV[G]$.
 
@@
 only a few percentage points, depending on the model.
-The $\twoljets$ team considered that to be too little to be worthwhile.
+We in the $\twoljets$ team considered that to be too little to be worthwhile.
 
@@
 \label{sec:2ljets_events_leptons}
-In the context of $\twoljets$, `leptons' means `electron-or-muons';
-this is because taus decay before reconstruction so are used only indirectly
-through their decay products, and neutral leptons
---- neutrinos --- go missing and contribute to the $\met$.
+As mentioned at the beginning of this Section~\ref{sec:2ljets_events},
+`leptons' means `electron-or-muons' in the context of $\twoljets$,.
+This is because taus leptons decay before reconstruction so are used only
+indirectly through their decay products, and neutral leptons
+--- neutrinos --- go missing and contribute only to our $\ptmiss$.
 
@@
 recommendation occurred during the years this analysis was in development, and
-the change would have required more work and further delays for an unknown
-amount of benefit.
+the change would have required more work and further delays for an unknown,
+but probably small, benefit.
 
@@
 Tagging adds labels to objects.
-Unlike how leptons are tagged as electrons and muons with very clean
-separation, jet tagging is a rather more noisy affair given \atlas' limited
+Unlike lepton flavor tagging, which can very cleanly separate electrons from
+muons, jet tagging is a rather more noisy affair due to \atlas' limited
 ability to reconstruct the individual hadrons within jets.
@@
 $W$, Higgs or top resonance~\cite{JETM-2018-06}, but large radius were
-considered too much additional work to considered in $\twoljets$,
+considered too much additional work to be used in $\twoljets$,
 particularly since they would require a large set of independently evaluated
@@
 Signal $b$-jets in $\twoljets$ require $\pt > 20\,\eV[G]$, $|\eta| < 2.5$ and a
-tag from the `MV2c10' boosted decision tree
-algorithm~\cite{
+tag from the `\textsc{MV2c10}' algorithm~\cite{
 friedman2002stochastic,
@@
 } at its $77\%$ efficiency working point.
-This algorithm is trained on a mixture of simulated $\ttbar$ and $\zjets$ data,
-and the reported efficiency means that a true $b$-jet sampled from the
+Notice that this $\pt > 20\,\eV[G]$ requirement is softer than our standard jet
+selection of $\pt > 30\,\eV[G]$. Selections that veto $b$-tags therefore act on
+a softer collection of jets.
+
+This \textsc{MV2c10} algorithm uses a boosted decision tree that is trained on
+a mixture of simulated $\ttbar$ and $\zjets$ data.
+The reported efficiency means that a true $b$-jet sampled from the
 algorithm's testing dataset has an estimated $77\%$ probability of being
 tagged.
-The `c10' suffix means that those data include a $7\%$ fraction of
-$c$-jets, and that the tagger does not use newer inputs from the
-RNNIP (Recurrent Neural Network Impact Parameter) tagger, nor the
-SMT (Soft Muon Tagger)~\cite{atlas_twiki_mv2, atlas_jet_mv2c10_2017}.
-This tagging algorithm combines various lower-level features, including the
-basic kinematics $\pt$ and $|\eta|$, and more intricate features of
-observable secondary vertices including reconstructed hadronic decay
-trees~\cite{atlas_jet_mv2c10_2017}.
+It uses low-level event variables including basic kinematics and more intricate
+features of reconstructed hadronic decay
+trees~\cite{atlas_jet_mv2c10_2017, atlas_twiki_mv2}.
 
@@
 $\met > 100\,\eV[G]$.
-This is a nice round number, and corresponds roughly to where the $\met$-less
-background $\zjets$ tails off,
+This is a nice round number (in these conventional units), and corresponds
+roughly to where the $\met$-less background $\zjets$ tails off,
 and also corresponds roughly to several standard deviations of
@@
 than a jet of similar energy.
-Unless, as discussed in Chapter~\ref{chapter:experiment}, that muon is hard
-enough that its curvature through \atlas' magnetic fields is poorly tracked.
-Unless, in the right circumstances, that jet originates from a $b$ quark,
-such that the weak decays of its hadrons tend to lose momentum in
-neutrinos or soft muons.
+But there are exceptions, and exceptions to the exceptions.
+One exception, as discussed in Chapter~\ref{chapter:experiment}, is if the muon
+is hard enough that its curvature through \atlas' magnetic fields is
+imprecisely reconstructed.
+A further exception to that rule can occur, however, if the jet originates from a
+$b$ quark, such that the weak decays of its hadrons tend to lose momentum in
+neutrinos or soft muons that are not reconstructed.
 
@@
 =
-\ptmiss \Sigma^{-1} \ptmiss
+\ptmiss\,{}^T \Sigma^{-1} \ptmiss
 =
@@
 perpendicular components~\cite{atlas_met_significance}.
-This is the square of a Mahalanobis (standard-deviations) distance between
-the origin and $\ptmiss$~\cite{mahalanobis1936generalised}.
+This is the square of a `Mahalanobis' (standard-deviations)
+distance~\cite{mahalanobis1936generalised} between the origin and $\ptmiss$.
 
 Just as $\ptmiss$ is constructed from a (negative) sum of contributions from
-objects in an event, this $\Sigma$ is a sum of covariances assigned to those
-objects; both means and variances add linearly.
+objects in an event, this $\Sigma$ is a sum of covariance matrices assigned to
+those objects; both means and variances add linearly.
 
@@
 in which $m_T^2(p_\mu) = p_0^2 - p_1^2 - p_2^2$ and $\{E,\,\vec p\,\}_\mu$
-denotes the four-vector with energy $E$ and momentum
+denotes a four-vector with energy $E$ and momentum
 $\vec p$~\cite{lester1999measuring}.
@@
 Lester and Nachman~\cite{lester2015bisection} during this project.
+\emph{Thanks to Tom~Gillam and Christopher~G.~Lester for collaboration on
+refining implementation.}
 
@@
 \end{itemize}
-All of these are physical properties of particles or collections of particles.
-Please remember that none is an exact or true value;
-all of these event variables are estimates derived from approximate
-reconstructions.
-While this distinction is easy to drop in lax language, it is necessary to
-understand the shapes of noisy variables such as $\mjj$ and indeed the
-existence of $\metsig$.
+All of these are physical properties of particles or collections of particles
+that are reconstructed from data.
+Note, however, that these are not identical to the same quantities that could
+be defined from hypothetically true states of those particles ---
+the reconstruction for a given `truth' state can vary with noise.
+While I admit to dropping this distinction in lax language, it is needed to
+explain the broad shapes of noisy variables such as $\mjj$ and indeed the
+existence of $\metsig$, which relies on estimates of the reconstruction noise
+and is ill-defined at a particle-truth level.
 
@@
 
-My understanding of search design is as conflict between simplicity and
-precision.
+In my understanding of search design, there is a deep conflict between
+simplicity and precision.
 Yes, an extra bin with higher $\metsig$ might catch some extreme signals,
@@
 extreme $\met$.
-But if these delay the result without comparable benefits, they will act to
-slow down our scientific progress.
+But if these delay our result without comparable benefits, they will act to
+slow down scientific progress in the area.
 
@@
 honest representation of our modelling.
-Restricting ourselves to round numbers not only clarifies our communications,
+Restricting ourselves to round numbers (in the units with which we communicate)
+not only clarifies our communications,
 but keeps us honest by denying ourselves the ability to over-fit simulations.
@@
 reviewer suggested that it was ``too honest''.
-% ~\cite{comment2020toohonest}. % Could cite, probably should not.
 
@@
 \\[1em]
-$n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$
-\\[0.5em]
-same flavour, opposite charge leptons
-\\[0.5em]
-$\njet \geq 1$
-\\[0.5em]
-$p_\mathrm{T}^{\,\ell_2} > 25$
-\\[0.5em]
-$\metsig > 6$
-\\[0.5em]
-$\met > 100$
+$n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$,
+\\[0.5em]
+same flavour, opposite charge leptons,
+\\[0.5em]
+$\njet \geq 1$,
+\\[0.5em]
+$p_\mathrm{T}^{\,\ell_2} > 25$,
+\\[0.5em]
+$\metsig > 6$, and
+\\[0.5em]
+$\met > 100$.
 \end{tabular}
@@
 
-% presel table should come before High
-\FloatBarrier
 \subsection{High}
@@
 with boundaries at $a$, $b$, and $c$.
-The mid-bar `$\mid$' indicates logical or.
+The mid-bar `$\mid$' indicates logical `or'.
 Differences between regions are \uwave{underlined}.
@@
 modelling might have been necessary.
-``In the end'', comments one \atlas\ author,
-``it seems fine since there's not events in the SR!''~\cite{comment2022vrllbb}.
+``In the end'', commented one \atlas\ author~\cite{comment2022vrllbb},
+``it seems fine since there's not events in the SR!''
 
@@
 
-The control region \crvz\ occupies the $\mjj$ sideband to
-SR-Int and VR-Int.
+The control region \crvz\ occupies the $\mjj$ sideband of SR-Int and VR-Int.
 As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
@@
 
-The control region \crdy\ (named after Drell and Yan) is defined to normalize
-off-shell $\zyjets$ near to SR-OffShell.
+The control region \crdy\ (named after Sidney D. Drell and Tung-Mow Yan) is
+defined to normalize off-shell $\zyjets$ near to SR-OffShell.
 Its reduced $\metsig$ requirement and relaxed angular selections were required
@@
 agree well with data.
-Contributions of $\zyjets$ in SR-OffShell are small anyway, as shown in
+Contributions of $\zyjets$ in SR-OffShell are small, as shown in
 Figure~\ref{fig:2ljets_offshell_region}.
@@
 modelling is meaningfully related.
-And our background modelling will always be related if our discovery regions
-share control region data, or other coupled systematic variations.
-
-That is, finely binning discovery regions can harm sensitivity since their
+Our background modelling will in fact always be related, since our discovery
+regions share control region data or other coupled systematic variations.
+
+Finely binning discovery regions can harm sensitivity since their
 interpretations are decoupled.
@@
 above, but because the practical cost of performing the standard statistical
-software to analyse each discovery region (and to perform their many validation
-studies), turned out to be so large~\cite{verkerke2003roofit}.
+methods to analyse each discovery region (and to perform their many validation
+studies), turned out to be prohibitively large~\cite{verkerke2003roofit}.
 Having many more discovery regions would have further exaggerated this
@@
 Evaluation of candidate High discovery regions.
-The region SR-High-8, labelled `High8\_1+\_2' here, is chosen for its
+The region SR-High-8, labelled `\texttt{High8\_1+\_2}' here, is chosen for its
 simplicity and good sensitivity to both benchmark signal models.
@@
 samples~\cite{cranmer2012histfactory}.
-An alternative ``shapefactor'' prescription exists which gives one parameter
-per sample per bin, but this greatly increases the complexity of the model and
+An alternative ``\texttt{shapefactor}'' prescription
+exists~\cite{cranmer2012histfactory} that gives one parameter
+per sample per bin, but greatly increases the complexity of the model and
 was considered unnecessary here.
@@
 large cross-sections are not modelled well.
-This motivates the use of data-driven fake estimates to replace the need for
+This motivates the use of data-driven FNP estimates to replace the need for
 such specialized simulations of poorly understood physics.
@@
 Although I have criticisms of the method itself (expounded below), these
-are not criticisms of his excellent work spent in using it.
+are not criticisms of his excellent work with it.
 }
@@
 That is, the Matrix Method starts from data, flips them over to truth
-estimates, then flops those back to data-level estimates. Flip\ldots flop.
+estimates, then flops those back to data-level estimates.
+% Flip\ldots flop.
 
@@
 Pre-fit results with statistical uncertainties only are presented in
-Figure~\ref{fig:2ljets_vrlow_alt}.
-Sorry for the different order of backgrounds; this figure uses
-conventions from the paper, while Figure~\ref{fig:2ljets_vrlow_alt}
-uses the conventions we used in development.
+Figure~\ref{fig:2ljets_vrlow_alt}
+The ordering of background samples differs between this figure,
+which uses the convention we chose for the paper, and
+Figure~\ref{fig:2ljets_vrlow_alt} which uses the convention we used in
+development.
 }
@@
 
-To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
-excessive \diboson\ background should also take some blame for this poor
-prediction in SR-Low;
+A possible interpretation of Figure~\ref{fig:2ljets_low_crz_pre_post} could be
+that the excessive \diboson\ background should also take some blame for this
+poor prediction in SR-Low;
 its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
@@
 parton shower.
-Its central value is $1$, and it is varied up and down by factors of $4$.
+Its central value is one, and it is varied up and down by factors of four.
 Since \sherpa\ defines its scales in a quadratic form, this matches the typical
@@
 
-To include these large variations moving in wacky directions could allow the
+To include these large variations moving in strange directions could allow the
 background model to easily fit deviations on the basis of our ignorance
@@
 from these replica weights --- both say to follow the PDF4LHC
-recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, pp.46,
-Section~``6.2. Formulae for the calculation of $PDF$ and $PDF+\alpha_s$
-uncertainties).
+recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, pp. 46,
+Section~``6.2. Formulae for the calculation of PDF and PDF+$\alpha_s$
+uncertainties'').
 Since PDF4LHC gives two alternative prescriptions, one using mean and variance
@@
 extrapolated nicely.
-Since the distance is dimensionful, it needs to be scaled to a reference value,
-which I chose as $100\,\eV[G]$; therefore
+Being a linear sum, the exponentiated Manhattan distance beneficially
+separating into a product of terms, so contours of constant $r_j(\vec x)$ are
+segmented into straight lines that relate to the chosen coordinates.
+Alternatives, such as the Euclidean distance, produced curved contours that did
+not align as well with my intuitions.
+Since the mass distance is dimensionful, it needs to be scaled to a reference
+value, which I chose as $100\,\eV[G]$; therefore
 \begin{equation}
@@
 One consequence of this is that there is no exact separation between $ZZ$ and
-$WW$ to $\ell\ell\nu\nu$ final states. They interfere.
+$WW$ to $\ell\ell\nu\nu$ final states --- they interfere.
 These histograms are therefore split on charged lepton number in their
@@
 Only the procedure is blind to data, not its action.
-No scientific research would employ an estimation procedure that clearly set
-its prediction directly from the data it is supposed to predict, but our use of
-the Matrix Method is proof by example that sufficiently obfuscated violations
-are permitted here.
+No research would employ an estimation procedure that clearly set its
+prediction directly from the data it is supposed to predict, but our use of the
+Matrix Method is proof by example that sufficiently obfuscated violations are
+permitted here.
 
@@
 data and pre-fit backgrounds in loose selections.
-Such plots were discussed in group meetings and used to catch bugs in the
-background simulations.
+Such plots were discussed in group meetings and usefully helped to catch bugs
+in the background simulations.
 But they unblind the data, notably in high-$\met$ where we expect signals.
@@
 and all contours are painted by heuristic interpolation algorithms.
-We follow standard \atlas\ SUSY conventions and perform that interpolation
+We follow current \atlas\ SUSY conventions and perform that interpolation
 with the \textsc{Scipy} implementation of
@@
 \emph{%
-Thanks to Larry~Lee for sharing scripts that were adapted to prepare the
-contours in our results.%
+Thanks to Larry~Lee for sharing scripts through \histfitter\ that were adapted
+to prepare the contours in our results.%
 }
@@
 I have failed to reproduce this effect outside the \histfitter-\roofit\ system,
-finding instead a limit close to $3$.
+finding instead a limit compatible with $3$.
 It would not surprise me if similar complications with the test statistic
-distribution played a part.
+distribution played a part in this result.
 
@@
 despite the fact that increasing data should increase the observed limit.
-The same effect has occurred in our results for DR-OffShell --- its observation
+The same effect has occurred in our results for DR-OffShell: its observation
 of $21$ data is below the post-fit background is $22.1$, but its observed limit
-exceeds the central expected limit.
-Another public example of this effect is
-$\mathrm{SR}0\textrm{-}\mathrm{ZZ^{tight}_{bveto}}$ of the four-lepton
+of $14.3$ events exceeds the central expected limit of $12.3$;
+results are reproduced from Table~\ref{tab:2ljets_discovery}.
+Another recent example with this effect is
+$\mathrm{SR}0\textrm{-}\mathrm{ZZ^{tight}_{bveto}}$ of the $4\ell$
 search~\cite{SUSY-2018-02}.
 \emph{Thanks to Tina~Potter for discussions of this problem and for sharing
-its details of its modelling.}
+its details of the $4\ell$ region models.}
 
@@
 ordering that larger $n$ is more signal-like.
-Limits moving in the wrong directions can simply be explained by flawed
+Limits moving in the wrong directions can simply be explained by imperfect
 test statistics ranking data in the wrong order.

### 2022-08-15 main/chapter_2ljets.tex
add a missing this
we have only one LHC

@@
 Lacking direct observation of the sleptons, or any other supersymmetric
-resonances, at the LHCs, $\mttwo$ remains useful for separating standard model
+resonances, at the LHC, $\mttwo$ remains useful for separating standard model
 processes since, with ideal reconstruction,
@@
 \emph{Thanks to Tom~Gillam and Christopher~G.~Lester for collaboration on
-refining implementation.}
+refining this implementation.}
 

### 2022-08-15 main/chapter_2ljets.tex
respond to Dan's awesome comments

@@
 
-My main messages that this thesis reports are my contributions to the recently
+The main messages that this thesis reports are my contributions to the recently
 published \atlas\ searches ``for new phenomena in events with two leptons,
@@
 $\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
-bosons, in variable branching fractions,
+bosons, with variable branching fractions,
 to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
@@
 labelled as excluded are derived, and displayed in
-Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and
-in~\ref{fig:2ljets_contours_gmsb} for GMSB.
+Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and in
+Figure~\ref{fig:2ljets_contours_gmsb} for GMSB.
 In both of these figures, space to the lower left of the contours is
@@
 
-\FloatBarrier
 \subsection{Contributions}
@@
 \item Production of main data inputs for all three $\twoljets$ searches
-and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
+and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09}.
 Production of all experimental systematic variations for the background
-samples;
-other group members assisted with the central background sample.
+samples --- other group members assisted with the central background sample.
 Production of all electroweak signal samples and their systematic variations.
-\item Evaluation of $Z$+jets uncertainties for RJR signal regions through
-from simulations. These exceeded $100\%$ relative variation and supported the
-use of data-driven methods.
+\item Evaluation of $Z$+jets uncertainties for RJR signal regions from
+simulations. Their variations exceeded $100\%$ relative variation and supported
+the use of data-driven methods.
 \item Various collaborations with all other group members on common core tasks
@@
 It uses conventional (non-RJR) event variables and focuses on features in the
-distribution of the invariant masses of lepton pairs that can have distinct
-features that depend on the masses of decaying supersymmetric resonances.
+distribution of the invariant masses of lepton pairs.
+That dilepton mass distribution can have distinct features that would stand
+out from Standard Model backgrounds and depend on the masses of decaying
+supersymmetric resonances.
 
@@
 
-\FloatBarrier
 \subsection{RJR}
 \label{sec:2ljets_origins_rjr}
+The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
+particle data by reconciling them with user-specified decay trees, and operates
+by making decisions of how to assign four-momenta to different `systems',
+which are represented as nodes on the chosen decay graph.
+Various RJR event variables are then defined in those systems' rest
+frames~\cite{jackson2017sparticles, jackson2017rjr}.
+
+In this way, RJR variables have some intelligence in how parent particles are
+reconstructed from their visible decay products, as well as approximate
+independence from uninteresting boosts, particularly those from the emission of
+QCD jets outside the main supersymmetric decay tree.
+
+Over the last several years, RJR event variables have been demonstrated to be
+effective for interpreting supersymmetric particle physics
+data~\cite{santoni2018probing},
+and have been used in numerous \atlas\ searches~\cite{
+atlas_rjr_SUSY_2016_07,
+atlas_rjr_SUSY_2016_15,
+atlas_rjr_SUSY_2016_16,
+atlas_rjr_23l_SUSY_2017_03,
+atlas_rjr_SUSY_2018_12,
+atlas_rjr_EXOT_2019_19,
+atlas_rjr_3l_SUSY_2019_09
+}.
+They have also influenced the designs of other analyses that do not themselves
+use RJR variables~\cite{atlas_rjr_mimic_SUSY_2018_06}.
+
+Two decay trees are used in the $\twoljets$-RJR analysis, and from each
+one signal region is defined and explored.
+These trees are illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees} and
+target C1N2-like scenarios similar to those also targetted by the
+$\twoljets$-electroweak search.
 
@@
 \end{figure}
-
-The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
-particle data by reconciling them with user-specified decay trees, and operates
-by making decisions of how to assign four-momenta to different `systems',
-which are represented as nodes on the chosen decay graph.
-Various RJR event variables are then defined in those systems' rest
-frames~\cite{jackson2017sparticles, jackson2017rjr}.
-
-In this way, RJR variables have some intelligence in how parent particles are
-reconstructed from their visible decay products, as well as approximate
-independence from uninteresting boosts, particularly those from the emission of
-QCD jets outside the main supersymmetric decay tree.
-
-Over the last several years, RJR event variables have been demonstrated to be
-effective for interpreting supersymmetric particle physics
-data~\cite{santoni2018probing},
-and have been used in numerous \atlas\ searches~\cite{%
-atlas_rjr_SUSY_2016_07,
-atlas_rjr_SUSY_2016_15,
-atlas_rjr_SUSY_2016_16,
-atlas_rjr_23l_SUSY_2017_03,
-atlas_rjr_SUSY_2018_12,
-atlas_rjr_EXOT_2019_19,
-atlas_rjr_3l_SUSY_2019_09
-}.
-They have also influenced the designs of other analyses that do not themselves
-use RJR variables~\cite{atlas_rjr_mimic_SUSY_2018_06}.
-
-Two decay trees are used in the $\twoljets$-RJR analysis, and from each
-one signal region is defined and explored.
-These trees are illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees} and
-target C1N2-like scenarios similar to those also targetted by the
-$\twoljets$-electroweak search.
 
@@
 \label{sec:2ljets_origins_strong}
-Strongly supersymmetric particles, the super-partners of quarks and gluons, would
-be produced at much larger cross-sections than their electroweak cousins at the
-LHC.
+Strongly interacting supersymmetric particles, the super-partners of quarks and
+gluons, would be produced at much larger cross-sections than their electroweak
+cousins at the LHC.
 Squarks and gluinos in simplified models are therefore excluded with masses up
@@
 Previous incarnations of this search have been conducted with similar
-selections and target models on in older data, including partial Run~2~\cite{
+selections and signal models in older data, including partial Run~2~\cite{
 atlas_susy_strong_2l_partial_run2_1,
@@
 overlapping signal regions.
-Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
+Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to
+$m(Z)\approx 91.2\,\eV[G]$~\cite{pdg2020review},
 and the others are binned in $\mll$ for sensitivity to the various
@@
 The largest Standard Model background to this search is dileptonic $\ttbar$,
-partly because no $b$-tagging veto is applied.
+partly because it employes no veto for $b$-tagged jets.
 All three signal models produce same-flavour lepton pairs, but $\ttbar$
@@
 
-\FloatBarrier
 \subsection{Electroweak}
@@
 We are happy to sacrifice some precision for generality.
-Since the $W$ and $Z$ boson masses are similar, this sacrifice turns out to be
-quite small anyway.
+Since the $W$ and $Z$ boson masses are similar, this sacrifice turned out to
+have quite a minor impact.
 
@@
 They are standard \atlas\ practice, and there should have been no complication
-had I designed copied SR2-int and SR2-high for $\twoljets$-electroweak but with
-with minor tweaks to change them into orthogonally binned regions.
+had I imitated SR2-int and SR2-high for $\twoljets$-electroweak but with minor
+tweaks to change them into orthogonally binned regions.
 However, my design decisions for $\twoljets$-electroweak became much more
@@
 Unfortunately, the $\twoljets$-electroweak regions designs were finalized
-before unblinding and before I appreciated this problem with background
+before unblinding, and before I appreciated this problem with background
 uncertainties in the smaller regions.
-Various workaround are invoked to assign plausible variations despite
-small yields and failing statistical precision, which shall be seen in the
-remainder of this chapter.
+Various workarounds are therefore invoked to assign plausible variations
+despite small yields and failing statistical precision, which shall be seen in
+the remainder of this chapter.
 As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
@@
 This Section~\ref{sec:2ljets_events} serves to describe how events are selected
-for the $\twoljets$ analyses, and which event variables are extracted for use
+for the $\twoljets$ searches, and which event variables are extracted for use
 in its selections.
 
-Our dataset comprises $136.0\,\mathrm{fb}^{-1}$ ($\pm 1.7\%$) of integrated
+Our dataset comprises $139.0\,\mathrm{fb}^{-1}$ ($\pm 1.7\%$) of integrated
 luminosity observed during normal operations of the \atlas\ detector in the
@@
 The set of targeted objects includes all those used in this analysis and more:
-electron, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
+electrons, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
 decaying tau leptons~\cite{atlas_PERF_2007_01}.
@@
 such that they avoid the need to accurately model their jagged activation
-thresholds and the variable efficiencies near to those thresholds~\cite{
+thresholds and varying efficiencies near to those thresholds~\cite{
 atlas_trigger_egamma_run2,
@@
 in excess of $90\%$ of events from benchmark C1N2 models.
-Therefore, any additional trigger has at most a $10\%$;
-improvement. Adding one-lepton triggers was found to increase efficiency by
-only a few percentage points, depending on the model.
-We in the $\twoljets$ team considered that to be too little to be worthwhile.
+Therefore, any additional trigger has at most a $10\%$ improvement.
+Adding one-lepton triggers was found to increase acceptance rates by only a
+few percentage points.
+Especially since this must come with a comparable increase in background rates,
+the benefit is too little to be considered worthwhile.
 
@@
 Each lepton is accepted for use in $\twoljets$ only if it satisfies certain
-kinematic and is identified with `quality' criteria.
+kinematic selections and is identified with `quality' criteria.
 For both electron and muons, these criteria are standardized with labels
@@
 orthogonality with other \atlas\ analyses that study events with differing
-lepton counts, such that our results might be combined in future to deliver an
+lepton counts, such that our results can be combined in future to deliver an
 improved total sensitivity.
@@
 
-Jets in $\twoljets$ use the anti-$k_t$ algorithm with the standard small-radius
-parameter of $R=0.4$~\cite{jet_anti_kt}.
+Jets in $\twoljets$ are constructed with the anti-$k_t$ algorithm with a
+standard small radius parameter of $R=0.4$~\cite{jet_anti_kt}.
 Jets are formed from `topological' clusters of energy deposits in the
@@
 Once we know about all the reconstructed leptons and hadronic jets, we can
-also improve our rejection of non-prompt leptons from hadron decays, since
-leptons near to jets become suspect.
+also improve our rejection of non-prompt leptons from hadron decays by
+leveraging the tendancy for those non-prompt leptons to be near to jets.
 To remove duplicates and further reduce non-prompt lepton contributions, we
@@
 All signal models in $\twoljets$ generate hard, invisible particles.
-By selecting for large $\met$, however, we select events in which either there
-were real invisible particles or something was badly mismeasured.
-And the rate of bad mismeasurements from Standard Model processes
-with large cross-sections can be substantial.
+By selecting for large $\met$, however, we select events with either
+real invisible particles or bad mismeasurements.
+And the rate of badly measured events from Standard Model processes with large
+cross-sections can be substantial.
 A standard selection requirement in $\twoljets$-electroweak is
@@
 and also corresponds roughly to several standard deviations of
-a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$;
-I observe that noise scale by inspecting $\met / \metsig$ in
+a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$.
+(I observe that noise scale by inspecting $\met / \metsig$ in
 $\twoljets$-electroweak selections, where $\metsig$ is defined in the next
-section.
+section.)
 
@@
 
-However, alternatives did select different subsets of signal yields:
+However, alternative jet pair assignments did select different subsets of
+signal yields:
 the alternative jet assignments had some potential to define secondary signal
@@
 not only clarifies our communications,
-but keeps us honest by denying ourselves the ability to over-fit simulations.
+but keeps us honest by restricting our ability to over-fit to simulations.
 
@@
 \label{sec:2ljets_high}
-The highest $\metsig$ selections, starting around $\metsig > 18$, contain
+Our highest $\metsig$ selections, starting around $\metsig > 18$, contain
 pure \diboson\ backgrounds in relatively small quantities, along with the bulk
@@
 additional sensitivity in High regions by loose requirements
-of $\mll \in 71\textrm{--}111$ and $\nbtag \leq 1$.
+of $\mll \in 71\textrm{--}111\,\eV[G]$ and $\nbtag \leq 1$.
 This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
@@
 elsewhere in the $\twoljets$-electroweak design to reduce these backgrounds
-while minimize lost signal yields.
-
-To maximize projected sensitivity, SR-High is split into four bins with
-one split at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and a second at
-$\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
+while minimizing lost signal yields.
+
+To maximize our projected sensitivity, SR-High is split into four bins with
+one division at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and a second
+division at $\metsig = 21$ to split each of those into
+SR-High-*-a and SR-High-*-b.
 For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
@@
 
-Total backgrounds predict data well in VR-$\llbb$.
+Overall, the total background yield predict the data well in VR-$\llbb$.
 What was not appreciated soon enough, however, was an excess in the $\metsig$
@@
 Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper}.
-The highest bins have very small background predictions and positive data.
+The highest bins have small background predictions and positive data.
 Had this been noticed before unblinding, more validation studies or altered
@@
 
-\FloatBarrier
 \subsection{Intermediate}
@@
 As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
-intermediate space give \crvz\ a very pure \diboson\ background, which is very
-valuable or constraining the modelling of High regions.
+intermediate space give \crvz\ a very pure \diboson\ background, which is
+valuable to constrain our modelling by extrapolation to High regions.
 Also included in Table~\ref{tab:2ljets_int} and
@@
 
-\FloatBarrier
 \subsection{Low}
@@
 
-These Low regions target signal models with small mass splittings,
-of around $100\textrm{--}150\,\eV[G]$ that still exceed the masses of
+These Low regions target signal models with small mass splittings of
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1)
+\approx 100\textrm{--}150\,\eV[G]$ that still exceed the masses of
 electroweak bosons.
-In particular, we aim to test C1N2($200$, $100$), which is a light, on-shell
-benchmark that has not previously been excluded in this
-channel~\cite{atlas_23l_SUSY_2016_24}.
+In particular, we aim to test C1N2~($200$, $100$)
+(meaning $m(\chargino_1, \neutralino_2) = 200\,\eV[G]$ and
+$m(\neutralino_1) = 100\,\eV[G]$), which is a light on-shell benchmark that
+has not previously been excluded in this channel~\cite{atlas_23l_SUSY_2016_24}.
 
@@
 but here also works to reject other $\ell^+\ell^-$ production mechanisms which
-produce leptons is less correlated directions, particularly dileptonic
+produce leptons in less correlated directions, particularly dileptonic
 $\ttbar$.
 As elsewhere, sensitivity is increased by binning SR-Low with
-$\metsig \in 6\textrm{--}9\textrm{--}12$ to form SR-Low-a and SR-Low-b
-respectively.
+to form SR-Low-a with $\metsig \in 6\textrm{--}9$ and
+SR-Low-b $\metsig \in 9\textrm{--}12$.
 
@@
 notable excess.
-Standardizing cuts to remove high-$\mjj$ tails (such as requiring
+Standardizing selections to remove high-$\mjj$ tails (such as requiring
 $\mjj < 500\,\eV[G]$) could have made control and validation regions like \crz\
 kinematically closer to the signal regions they are supposed to target,
-and such a cut may well have improved the analysis in this case.
+and such a selection may well have improved the analysis in this case.
 
@@
 CR-Low-ISR-Z, CR-Low-ISR-DF.
-After this fit, the validation region VR-Low-ISR had a high background
+After this fit, the validation region VR-Low-ISR had a high background yield
 of $56.7 \pm 4.4$ for $47$ data, which is only a $1$-sigma deficit.
@@
 (pre-fit) signals shown in Figure~\ref{fig:2ljets_low_isr_old_sr}.
-On the very sound advice of \atlas\ authors commenting on the internal
+On the very sound advice of \atlas\ authors commenting on our internal
 documentation, this ISR construction was entirely removed from the search
@@
 This deficit was therefore blamed on multi-jet QCD modelling, and resolved by
-retreating to the SR-Low requirement of exactly jets.
+retreating to the SR-Low requirement of exactly two jets.
 It is interesting, however, that the two-jet SR-Low eventually observed a
@@
 
-\FloatBarrier
 \subsection{Off-shell}
@@
 These searches consider the same C1N2 $W\kern-0.4exZ$ signal models as us, but target the
-case where the $W$ decays leptonically $W\rightarrow\ell\nu$.
+case where the $W$ boson decays leptonically $W\rightarrow\ell\nu$.
 For comparison with our off-shell exclusion,
@@
 subset of the signal regions, as labelled and described in the figure caption.
+\emph{Thanks to Tina~Potter for suggesting this impact study.}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_c1n2.pdf}
+\caption{C1N2}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_gmsb.pdf}
+\caption{GMSB}
+\end{subfigure}
+\caption[
+Projected exclusion from sets of orthogonal signal regions
+]{%
+Projected exclusion from sets of orthogonal signal regions, demonstrating that
+their combination has greater sensitivity reach than any individually.
+The label `High' indicates the bins of SR-High, and `HighNo4' excludes
+SR-High-1J to show the locations where mono-jet events have an impact.
+Similarly, `$\llbb$', `Int', `Low', and `OffShell' are shown to target their
+appropriate parameter spaces.
+Although figures were produced before unblinding, with older samples and region
+designs, their later updates do not change the general message.
+Values of $\cls$ from expected data in the combined fit are overlaid.
+}
+\label{fig:2ljets_fit_multi_contours}
+\end{figure}
 
@@
 parameter space.
-Sensitivity to the C1N2 ($200$, $100$) point was a goal of the SR-Low design,
+Sensitivity to the C1N2~($200$, $100$) point was a goal of the SR-Low design,
 and although it alone does not quite achieve expected exclusion there, the
@@
 
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_c1n2.pdf}
-\caption{C1N2}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_multi_contours_gmsb.pdf}
-\caption{GMSB}
-\end{subfigure}
-\caption[
-Projected exclusion from sets of orthogonal signal regions
-]{%
-Projected exclusion from sets of orthogonal signal regions, demonstrating that
-their combination has greater sensitivity reach than any individually.
-The label `High' indicates the bins of SR-High, and `HighNo4' excludes
-SR-High-1J to show the locations where mono-jet events have an impact.
-Similarly, `$\llbb$', `Int', `Low', and `OffShell' are shown to target their
-appropriate parameter spaces.
-Although figures were produced before unblinding, with older samples and region
-designs, their later updates do not change the general message.
-Values of $\cls$ from expected (not observed --- the label is wrong) data in
-the combined fit are overlaid on the grid.
-}
-\label{fig:2ljets_fit_multi_contours}
-\end{figure}
-
-\FloatBarrier
 \subsection{Discovery}
@@
 However, simulations are not totally accurate.
-In particular, do not use simulations of pure hadronic processes, so do not
+In particular, we do not use simulations of pure hadronic processes, so do not
 really know the predicted rates at which these strong processes would fake
@@
 Although prompt strong effects may only rarely fake leptons, they have very
-large cross-sections are not modelled well.
+large cross-sections and are not exoected to be modelled well.
 This motivates the use of data-driven FNP estimates to replace the need for
@@
 D0:1999qdf
-}%
-to estimate
-backgrounds of fake and non-prompt leptons, so named because it works by
-estimating a linear relationship between dilepton event rates that is
+}
+to estimate backgrounds of fake and non-prompt leptons, so named because it
+works by estimating a linear relationship between dilepton event rates that is
 written as a matrix multiplication.%
@@
 implemented by assigning weights to observed data events that are then
-histogrammed them in the tighter selections of our search regions,
+histogrammed in the tighter selections of our search regions,
 much like other samples~\cite{twoljets2018int}.
@@
 have an average weight of $+0.238$ (one can check that these recover the
-central FNP background of $2.2$).
+central FNP background yield of $2.2$).
 Such negative signal weights suggest that a local \emph{excess} of data
@@
 
-Alternative versions of all core control, validation regions were inspected
+Alternative versions of all main control, validation regions were inspected
 with their lepton requirements change to require either different-flavour
@@
 To check that a similar bump exists in data and fully simulated (FullSim)
-background simulations, I conducted a study into a standard model process
+background simulations, I conducted a study into a Standard Model process
 production boosted hadronic $W$ bosons in $\ttbar$ events.
@@
 \begin{itemize}
-\item evaluate the difference between the two samples $\delta_{12}$.
-\item evaluate the statistical uncertainty on the alternative sample $\sigma_2$.
+\item evaluate the difference between the two samples $\delta_{12}$,
+\item evaluate the statistical uncertainty on the alternative sample
+$\sigma_2$,
 \item combine these components in quadrature, so
@@
 and this $\zjets$ sample comparison is the main reason why.
-Internal reviewers were, there, confused as to why we mentioned a deficit in
-SR-Low that did not appear in the results plots or tables.
+Internal reviewers were, therefore, confused as to why we mentioned a deficit
+in SR-Low that did not appear in the results plots or tables.
 Much to my ire, the paper editors therefore removed all mentions of an SR-Low
 deficit from the paper.
-Despite bring real and significant, this deficit was erased from public view,
-and to my knowledge, is only reported through inner details of
+Despite bring real and significant, this deficit was erased from public view
+and, to my knowledge, is only reported through inner details of
 the \histfactory\ models shared in our \hepdata\ archive~\cite{hepdata.116034}
@@
 
-Unfortunately, a total failure of statistical precision means that no
-meaningful result can be derived from these variations.
+Unfortunately, a total failure of statistical precision for these \diboson\
+variations means that no meaningful result can be derived.
 First, our regions receive very few raw events in each variation,
@@
 The number of PDF replicas varies between samples, but is between
-$100$ and $138$ for all except $\zyjets$, that has just one central and two
-alternatives.
+$100$ and $138$ for all samples except $\zyjets$, which has just one central
+and two alternatives.
 Remarkably, both SUSY and PMG prescriptions agree on how to assign variations
@@
 
-Scale variations now appeared twice for signals:
-first for cross-section variations and
-second for these theory variations.
+Scale variations have now appeared twice for signals:
+first for cross-section variations and second for these theory variations.
 They therefore contribute both in each fit model and in the varied
@@
 But they unblind the data, notably in high-$\met$ where we expect signals.
-When commenting on this, I was told that it does not count as unblinding
-because we are in a loose selection.
+When commenting on this, the resonse was that it does not count as unblinding
+because we are inspecting a loose selection.
 However, any signal region must be within this loose selection, and those
@@
 The exact \histfactory\ models used in our discovery fits are shared in our
-\hepdata\ archive~\cite{hepdata.116034}.%
+\hepdata\ archive~\cite{hepdata.116034}.
 Although many other searches' discovery regions can be approximately reproduced
@@
 The $n=0$ case has no visible width since its likelihood is linear
-(adding signal just adds a constant to the $-\log$ likelihood objectives being
-minimized), but its distribution is not quite a delta function since the test
+(adding signal just adds a constant to the negative $\log$ likelihood
+objectives being minimized), but its distribution is not quite a delta function since the test
 statistic depends not exactly on the theoretical best fit, but on the noisy
@@
 against $\clsb$, since both $p$-values suffer from similar effects.
-Statistical precision is harmed, however --- although $1$m samples should
+Statistical precision is harmed, however --- although a million samples should
 give us around $e^{-7.9} \times 10^6 \approx 1000$ signal samples with

### 2022-08-16 main/chapter_2ljets.tex
autocheck

@@
 
-To test non-standard (supersymmetric) models against observation
-we add signal contributions on top of the modelled backgrounds.
+To test non-standard (supersymmetric) models against observation,
+we add signal yields on top of the modelled backgrounds.
 
@@
 event rates (yields) from that signal model with those parameters,
-and considering what would happen of those event yields were added to a
+and considering what would happen if those event yields were added to a
 Standard Model background.
@@
 All $p$-values are estimated by ``toy'' simulations of alternative data.
-Values with errors are rounded according the PDG rounding
+Values with errors are rounded according to the PDG rounding
 scheme~\cite{pdg2020review}.%
@@
 \\[0.5em]
-All three produce same-flavour, opposite-sign lepton pairs along with jets
+All three produce same-flavour, opposite-sign lepton pairs, along with jets
 and $\met$.
@@
 They are standard \atlas\ practice, and there should have been no complication
-had I imitated SR2-int and SR2-high for $\twoljets$-electroweak but with minor
+had I imitated SR2-int and SR2-high for $\twoljets$-electroweak, but with minor
 tweaks to change them into orthogonally binned regions.
@@
 the same models that we study in $\twoljets$.
-Since we designed these searches use orthogonal datasets, they may be
+Since we designed these searches to use orthogonal datasets, they may be
 combined for more sensitive future results.
@@
 important event variables.
+Definitions of the $\twoljets$-electroweak event variables are collected
+Section~\ref{sec:2ljets_cheatsheet}, which is intended as a useful bookmark
+for future reminders, and may also clarify occasions where we make forward
+references to variables before giving their precise definitions.
 
@@
 that trigger does not also harmfully increase inseparable backgrounds.
-A study, \emph{performed by Knut Vadla}, investigated the effects
+A study, \emph{performed by Knut~Vadla}, investigated the effects
 of including one-lepton triggers.
@@
 Tagging adds labels to objects.
-Unlike lepton flavor tagging, which can very cleanly separate electrons from
+Unlike lepton flavour tagging, which can very cleanly separate electrons from
 muons, jet tagging is a rather more noisy affair due to \atlas' limited
@@
 tagged.
-It uses low-level event variables including basic kinematics and more intricate
+It uses low-level event variables, including basic kinematics and more intricate
 features of reconstructed hadronic decay
@@
 also improve our rejection of non-prompt leptons from hadron decays by
-leveraging the tendancy for those non-prompt leptons to be near to jets.
+leveraging the tendency for those non-prompt leptons to be near to jets.
 To remove duplicates and further reduce non-prompt lepton contributions, we
@@
 \subsection{Cheat-sheet}
+\label{sec:2ljets_cheatsheet}
 Uncomfortably many event variables are used to define the regions of this
@@
 Numerical simplicity is valuable.
-In a draft of our \atlas-internal documentation I wrote that numerical
+In a draft of our \atlas-internal documentation, I wrote that numerical
 simplicity was a design goal, but am guilty of removing that statement after a
@@
 exploration of signal region designs,
-to Ben Hooberman for suggesting the investigation of massive jets,
+to Ben~Hooberman for suggesting the investigation of massive jets,
 and to the whole $\twoljets$ team for collaboration in their finalization.
@@
 \label{sec:2ljets_int}
-
 Our Intermediate signal regions sit in the $\metsig$ range of
@@
 $2.8$m events per sample.
-Due to the $\sqrt{n}$ scaling of Monte Carlo errors this means approximately
+Due to the $\sqrt{n}$ scaling of Monte Carlo errors, this means approximately
 $\sqrt{10}$ times reduced statistical noise.
@@
 Our high-level sample labels such as $\ttbar$, $\zjets$, \diboson\ and
-rare-top (`Top other') each include many sub-samples that are physically
+rare-top (`Top~other') each include many sub-samples that are physically
 distinct and non-interfering.
@@
 $tW\kern-0.4exZ$, $t\ell\ell g$ and $\ttbar W$ are also significant.
-Here, the plain `Top other' label refers to all other \topother\ contributions
+Here, the plain `Top~other' label refers to all other \topother\ contributions
 not included in the bracketed specializations.
@@
 anticorrelated with normalization factors.
-Similarly, the extra uncertainty for $\zjets$ in CR-tt causes an
+Similarly, the extra uncertainty for $\zjets$ in \crtt\ causes an
 anticorrelation between its parameter ``\texttt{alpha\_Zjets\_bextrap}''
@@
 --- as a probability of compatibility with an hypothesis,
---- as a probability that data occurred at random, by chance, or by statistical
-fluctuation,
+--- as a probability that data occurred at random, by chance, or by statistical fluctuation,
 --- as a likelihood or importance of data,
@@
 distribution.
-Fundamentally, a sufficient explanation of this problem is our refusal to
-compare models with likelihood ratios.
+Fundamentally, a sufficient explanation of this retrograde limit movement
+is our refusal to compare models with likelihood ratios.
 By instead using \emph{maximum} likelihood ratios, and modifications thereof,

### 2022-08-16 main/chapter_2ljets.tex
begin quick summary of other work in intro; not sure it should go here

@@
 theory, nor to the similarly-named
-Matrix Element Method~\cite{gainer2013matrix} for signal discrimination.%
+Matrix Element Method~\cite{Gainer:2013iya} for signal discrimination.%
 }
@@
 
-% this extra line activates enlarged spacing for the final paragraph lol
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-08-16 main/chapter_introduction.tex
begin quick summary of other work in intro; not sure it should go here

@@
 
-Helloasdasdasd
+Data Intensive Science in High-Energy Physics
+
+\section{Other work}
+
+\subsection{Testing symmetries in data}
+Data-driven tests of symmetries:
+Machine Learning algorithms to ~\cite{lester2021stressed}
+~\cite{Lester:2019bso}
+~\cite{tombs2021which}
+~\cite{lester2022hunting}
+
+\subsection{Higgs to tau-tau Nested Sampling}
+Proof-of-concept implementation of the Nested Sampling algorithm~\cite{
+skilling2004nested,
+skilling2006nested,
+skilling2010foundations
+}
+to classify events mediated by Higgs or Z bosons in their leptonic decays
+through tau lepton pairs in \atlas.
+In this scheme, Nested Sampling approximates likelihoods at given data for
+approximate models of the Higgs and Z models by `marginalizing',
+or integrating over, all freedoms in their production in LHC proton-proton
+collisions.
+Models of their processes were constructed from theory for the
+helicity-dependent leptonic decays of tau leptons~\cite{Hays2014tau}
+(with their two neutrinos integrated out analytically)
+and Z bosons~\cite{Thomson2011EWK},
+and parametrized from numerical parton distribution
+functions~\cite{Buckley:2014ana} for the models' differing distributions of
+longitudinal boosts.
+
+Unlike the similar Matrix Element Method~\cite{
+Fiedler:2010sg,
+Gainer:2013iya,
+PhysRevD.83.074010
+},
+which estimates likelihoods by fixing certain observed quantities and
+integrating remaining freedoms with adaptive Monte Carlo algorithms~\cite{
+Fiedler:2010sg,
+Gainer:2013iya,
+Lepage:1977sw,
+Ohl:1998jn
+},
+with Nested Sampling we begin in ignorance of the data in each event and
+gradually explore its constraints with random walks.
+Euclidean coordinates projected to the surface of the unit ball,
+in each of the respective rest frames of the heavy boson and tau leptons, allow
+those walks to avoid the gimble lock of spherical polar coordinates, and are
+used to describe Lorentz transformations to transform the observable leptons
+to the lab frame.
+
+This method worked in testing on smeared parton-level simulations from
+\madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
+\TODO{Should these posters be included as figures?}
+However, real events have substantial additional noise from QCD
+radiation~\cite{PhysRevD.83.074010} and pileup, and this method of approximate,
+parametrized modelling does not easily scale to include such extra, complicated
+effects.
+For this special case it also cost a few second of computing time per event,
+which is not competitive with modern Machine Learning algorithms that, once
+trained, can be rapidly evaluated over many events and should approximate
+the same result.
+Although this study shows that we \emph{can} use theoretical modelling
+with theoretically motivated algorithms for practical classification, it does
+not show that we \emph{should}.
+Empirically motivated function approximators such as
+Boosted Decision Trees~\cite{xgboost} or
+Neural Networks~\cite{MurphyKevinP.2012Mlap} are more scalable to complicated
+uses and more computationally efficient in evaluation.
+
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-08-16 main/chapter_introduction.tex
add some words on symmetries

@@
 \section{Other work}
-
-\subsection{Testing symmetries in data}
-Data-driven tests of symmetries:
-Machine Learning algorithms to ~\cite{lester2021stressed}
-~\cite{Lester:2019bso}
-~\cite{tombs2021which}
-~\cite{lester2022hunting}
+\TODO{This does not belong in an introduction and will be moved.}
 
@@
 \madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
-\TODO{Should these posters be included as figures?}
+\TODO{Should these posters be included as figures? Maybe appendix}
 However, real events have substantial additional noise from QCD
@@
 
+\subsection{Testing symmetries in data}
+Data-driven tests of physical symmetries are a speciality of the Cambridge
+\atlas\ group.
+Each test considers a theoretically motivated symmetry that states that
+distributions of data should be identical when transformed under its operation.
+This means that it can be possible to disprove a symmetry by searching only
+for differences between data and transformed versions of those data, without
+necessarily requiring other theoretical inputs.
+
+% emus
+Charge symmetry of electron-muon (`emu') pairs is a key example among these
+tests.
+Theoretically, it proposes that, in the absence of biases, data
+from proton-proton collisions at the LHC would include equal distributions
+of $e^+\mu^-$ and $e^-\mu^+$ events~\cite{Lester:2016qdv}.
+Real data can have emu biases --- one clear (but minor) example is that our
+detectors are made of matter containing negatively charged electrons, not
+positrons, so high energy electrons positrons have subtly different rates of
+energy loss in that detector matter.
+However, \cite{Lester:2016qdv} argues that all biases at the LHC should
+conspire to favour $e^-\mu^+$ events.
+Therefore defining an asymmetry
+\begin{equation}
+A^\mathrm{emu} = \frac{
+\sigma(e^+\mu^- + X)
+}{
+\sigma(e^-\mu^+ + X)
+}
+,
+\end{equation}
+one can predict $A^\mathrm{emu} \leq 1$ from the Standard Model at the LHC.
+Negative values could be explained by experimental biases, but
+$A^\mathrm{emu} > 1$ measured in data might indicate new physics, since it and
+can be caused by models of new physics that might invoke R-parity violating
+supersymmetry~\cite{Lester:2016qdv} or leptoquarks~\cite{EXOT-2018-29}.
+After its extensive development~\cite{Brunt:2674708,Pacey:2747774},
+this test has been performed in \atlas\ data, and found ``[n]o significant
+model-independent evidence''~\cite{EXOT-2018-29} for $A^\mathrm{emu} > 1$.
+
+% parity violation
+Parity symmetry is a second example. Nature violates parity is as first shown
+by the Wu experiment~\cite{PhysRev.105.1413} in the weak section, but unknown
+effects could possibly cause unexpected new forms of parity asymmetry that
+could be seen at the LHC. In particular, \TODO{}
+
+Data-driven tests of symmetries:
+\TODO{write, add some self-supervision references from wir}
+Machine Learning algorithms to ~\cite{lester2021stressed}
+~\cite{Lester:2019bso}
+~\cite{tombs2021which}
+~\cite{lester2022hunting}
+
 % this comment activates enlarged spacing for the final paragraph lol

### 2022-08-17 main/chapter_introduction.tex
expand on which is real side project and symmetries

--- 2022-08-09 main/chapter_introduction.tex
@@
 
-Helloasdasdasd
+Data Intensive Science in High-Energy Physics
+
+\section{Other work}
+\TODO{This does not belong in an introduction and will be moved.}
+
+\subsection{Testing symmetries in data}
+Data-driven tests of physical symmetries are a speciality of the Cambridge
+\atlas\ group.
+Each test considers a theoretically motivated symmetry that states that
+distributions of data should be identical when transformed under its operation.
+This means that it can be possible to disprove a symmetry by searching only
+for differences between data and transformed versions of those data, without
+necessarily requiring other theoretical inputs.
+
+% emus
+Charge symmetry of electron-muon, `emu', pairs is a key example among these
+tests.
+Theoretically, it proposes that, in the absence of biases, data
+from proton-proton collisions at the LHC would include equal distributions
+of $e^+\mu^-$ and $e^-\mu^+$ events~\cite{Lester:2016qdv}.
+Real data can have emu biases --- one clear (but minor) example is that our
+detectors are made of matter containing negatively charged electrons, not
+positrons, so high energy electrons positrons have subtly different rates of
+energy loss in that detector matter.
+However, \cite{Lester:2016qdv} argues that any biases at the LHC should
+conspire only to favour the rate of $e^-\mu^+$ over the rate of $e^+\mu^-$
+events.
+Therefore defining an asymmetry
+\begin{equation}
+A^\mathrm{emu} = \frac{
+\sigma(e^+\mu^- + X)
+}{
+\sigma(e^-\mu^+ + X)
+}
+,
+\end{equation}
+one can predict $A^\mathrm{emu} \leq 1$ from the Standard Model at the LHC.
+Negative values could be explained by experimental biases, but
+$A^\mathrm{emu} > 1$ measured in data might indicate new physics, since it and
+can be caused by models of new physics that might invoke R-parity violating
+supersymmetry~\cite{Lester:2016qdv} or leptoquarks~\cite{EXOT-2018-29}.
+After its extensive development~\cite{Brunt:2674708,Pacey:2747774},
+this test has been performed in \atlas\ data, and found no ``significant
+model-independent evidence''~\cite{EXOT-2018-29} for $A^\mathrm{emu} > 1$.
+
+% parity violation
+Parity symmetry, which changes the signs of all spatial coordinates
+$\vec x\rightarrow -\vec x$, is a second key example.
+Nature does violate parity symmetry, as first shown by the Wu
+experiment~\cite{PhysRev.105.1413}, in the weak sector, but that is not
+directly observable at the LHC due to its non-polarized beams and
+polarization-insensitive detectors~\cite{Lester:2019bso}.
+Unknown physical effects could, however, generate forms of parity asymmetry
+that could be seen at the LHC.
+Although the charge symmetry of the emu test has known sources of bias,
+no such biases are so clearly expected for parity.
+
+General parity symmetry can therefore be tested at the LHC by considering a
+parity-odd event variable and comparing its distribution to its parity-flipped
+version, which simply swaps its sign.
+% alpha from lester schott
+One such test has been performed with LHC data~\cite{Lester:2019bso}.
+This test used \cms\ open data and the parity-odd event variable
+\begin{equation}
+\label{eqn:cms_alpha}
+\alpha
+= \arcsin\!
+\left(
+\frac{
+\vec p^{\,j_1} \times \vec p^{\,j_2}
+}{
+|\vec p^{\,j_1} \times \vec p^{\,j_2}|
+}
+\cdot
+\frac{
+\vec p^{\,j_3}
+}{
+|\vec p^{\,j_3}|
+}
+\right)
+,
+\end{equation}
+in which $p^{\,j_i}$ is the momentum of the $i$th hardest jet in a multi-jet
+event,
+and did not find evidence for parity violation.
+But that does not prove that no parity-violating effects are occuring in these
+data, only that it did not appear in the distribution of $\alpha$.
+
+Can a single parity-odd event variable be sensitive to all norms of parity
+violation?
+The answer, it turns out, is `no' if that variable is also required to be
+continuous~\cite{lesterChiralMeasurements2021}.
+In this framework, therefore, a joint analysis of multiple variables is
+required for any sufficiently complicated data.
+Furthermore, it is not generally easy to construct such universally-sensitive
+sets of event variables~\cite{
+Gripaios:2020hya,
+Lester:2020jrg
+}.
+
+Rather than beginning from universally sensitive event variables, an
+alternative strategy is to build candidate variables to the data
+themselves~\cite{lester2021stressed}.
+If done wrong, such an idea could simply reinforce variations due to simple
+statistical noise.
+It can, however, be formulated as a rigorous test if one employs standard
+Machine Learning practice and splits the data into independent training and
+testing sets.
+Variables learned (by function approximation algorithms such as
+Boosted Decision Trees~\cite{xgboost} or
+Neural Networks~\cite{MurphyKevinP.2012Mlap})
+from the testing set can correspond to hypotheses about possible asymmetries
+in the data distribution, and if those hypotheses make accurate predictions in
+independent testing data, then they are evidence against the symmetry.
+
+My contribution to this field is to adapt this idea of learning parity
+variables,
+\emph{which was initiated by Christopher~G.~Lester}~\cite{lester2021stressed},
+into a probabilistic description, in which the algorithms predict
+`which is real?' between real data an their parity-transformed
+copies~\cite{tombs2021which}.
+This approach derives an objective function
+(which is a special case of the binary
+cross-entropy~\cite{MurphyKevinP.2012Mlap}),
+from which learning algorithms
+can approximate the odds ratio between the data distribution and its
+parity-transformed version.
+Importantly, it also allows one to model inefficiencies or holes in the
+distribution, which exist in practice from kinematic selections and imperfect
+hardware, and incorporate those into the training and testing without biasing
+the result.
+We have also demonstrated this method on simulated multi-jet data in an
+\atlas-like detector~\cite{lester2022hunting}.
+By simulating parity-violating events from a Lorentz-violating extension
+to the Standard Model, we show that this method build variables that are much
+more sensitive than the $\alpha$ of Equation~\ref{eqn:cms_alpha} while
+retaining its other attractive symmetries by encoding them into our function
+architectures~\cite{lester2021stressed, tombs2021which}.
+
+This work demonstrates that the `which is real?' method is ready to test for
+parity violation in real LHC data.
+Much like the emu work, however, its realization would require great diligence
+to handle important considerations of systematic uncertainties and statistical
+interpretations before it could be published through \atlas.
+
+The `which is real?' method is general, and can test not only parity, but
+symmetry under any transformation of the data, irrespective of whether it is
+discrete or continuous.
+In the context of modern machine learning, it is an example of
+`self-supervised' learning, in which the training objective derives from the
+data and not from additional labels~\cite{
+multitaskself2017,
+devlin2019bert
+}, and quite similar to the `CLIP' objective, which also attempts to identify
+a real examples from collections of alternatives~\cite{pmlr-v139-radford21a}.
+What is special in our work is our application for scientific self-supervised
+testing.
+
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-08-17 main/draft.tex
add junk tex file
expand on which is real side project and symmetries

--- 2022-05-11 main/chapter_2ljets.tex
@@
-\chapter{A search for new phenomena}
-\label{chapter:2ljets}
+As $\sigma \rightarrow 0$, this tends towards the the significance-transformed,
+sign-flipped Poisson maximum likelihood ratio
+\begin{equation}
+S_\mathrm{Poisson}(n, \mu, \sigma)
+= \mathrm{sgn}(n - \mu) \sqrt{2\left(
+    n\log\left(\frac{n}{\mu}\right) - n + \mu
+\right)}
+~.
+\end{equation}
 
-\chapterquote{Laws were made to be broken.}%
-{Christopher North, 1785--1854}%: Blackwood's Magazine May 1830
+\subsection{Higgs to tau-tau Nested Sampling}
+Proof-of-concept implementation of the Nested Sampling algorithm~\cite{
+skilling2004nested,
+skilling2006nested,
+skilling2010foundations
+}
+to classify events mediated by Higgs or Z bosons in their leptonic decays
+through tau lepton pairs in \atlas.
+In this scheme, Nested Sampling approximates likelihoods at given data for
+approximate models of the Higgs and Z models by `marginalizing',
+or integrating over, all freedoms in their production in LHC proton-proton
+collisions.
+Models of their processes were constructed from theory for the
+helicity-dependent leptonic decays of tau leptons~\cite{Hays2014tau}
+(with their two neutrinos integrated out analytically)
+and Z bosons~\cite{Thomson2011EWK},
+and parametrized from numerical parton distribution
+functions~\cite{Buckley:2014ana} for the models' differing distributions of
+longitudinal boosts.
 
-Hello
+Unlike the similar Matrix Element Method~\cite{
+Fiedler:2010sg,
+Gainer:2013iya,
+PhysRevD.83.074010
+},
+which estimates likelihoods by fixing certain observed quantities and
+integrating remaining freedoms with adaptive Monte Carlo algorithms~\cite{
+Fiedler:2010sg,
+Gainer:2013iya,
+Lepage:1977sw,
+Ohl:1998jn
+},
+with Nested Sampling we begin in ignorance of the data in each event and
+gradually explore its constraints with random walks.
+Euclidean coordinates projected to the surface of the unit ball,
+in each of the respective rest frames of the heavy boson and tau leptons, allow
+those walks to avoid the gimble lock of spherical polar coordinates, and are
+used to describe Lorentz transformations to transform the observable leptons
+to the lab frame.
 
-\clearpage
-
-Hello again
+This method worked in testing on smeared parton-level simulations from
+\madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
+\TODO{Should these posters be included as figures? Maybe appendix}
+However, real events have substantial additional noise from QCD
+radiation~\cite{PhysRevD.83.074010} and pileup, and this method of approximate,
+parametrized modelling does not easily scale to include such extra, complicated
+effects.
+For this special case it also cost a few second of computing time per event,
+which is not competitive with modern Machine Learning algorithms that, once
+trained, can be rapidly evaluated over many events and should approximate
+the same result.
+Although this study shows that we \emph{can} use theoretical modelling
+with theoretically motivated algorithms for practical classification, it does
+not show that we \emph{should}.
+Empirically motivated function approximators such as
+Boosted Decision Trees~\cite{xgboost} or
+Neural Networks~\cite{MurphyKevinP.2012Mlap} are more scalable to complicated
+uses and more computationally efficient in evaluation.

### 2022-08-18 main/chapter_2ljets.tex
expand on which is real side project and symmetries
align and line
add searches cartoons; add prob command

@@
 
-\clearpage
-\FloatBarrier
 \section{Summaries}
@@
 \\[0.5em]
-$\metsig > 6$, and
+\hphantom{~and}$\metsig > 6$, and
 \\[0.5em]
@@
 $M_{{tt}, {rr}} =
-p(t\!\mid\!r,1)
-\times p(t\!\mid\!r,2)$
+\prob{t}{r,1}
+\times \prob{t}{r,2}$
 in which the numerical index points to the first and second leptons
 respectively.
-These transition probabilities $p(*\!\mid\!*,i)$ are estimated by studies in
+These transition probabilities $\prob{*}{*,i}$ are estimated by studies in
 separate selections and assigned as functions of the lepton
@@
 \bar \mu^\textrm{data}_{tt\mid rr} &=
-p(t\!\mid\!r,1)
-\times p(t\!\mid\!r,2)
+\prob{t}{r,1}
+\times \prob{t}{r,2}
 \times \bar \mu^\textrm{truth}_{rr}, \\
@@
 \bar \mu^\textrm{data}_{tt\mid fr} &=
-p(t\!\mid\!f,1)
-\times p(t\!\mid\!r,2)
+\prob{t}{f,1}
+\times \prob{t}{r,2}
 \times \bar \mu^\textrm{truth}_{fr}, \\
@@
 \bar \mu^\textrm{data}_{tt\mid rf} &=
-p(t\!\mid\!r,1)
-\times p(t\!\mid\!f,2)
+\prob{t}{r,1}
+\times \prob{t}{f,2}
 \times \bar \mu^\textrm{truth}_{rf},~\textrm{and} \\
@@
 \bar \mu^\textrm{data}_{tt\mid ff} &=
-p(t\!\mid\!f,1)
-\times p(t\!\mid\!f,2)
+\prob{t}{f,1}
+\times \prob{t}{f,2}
 \times \bar \mu^\textrm{truth}_{ff}.
@@
 estimates, then flops those back to data-level estimates.
-% Flip\ldots flop.
+Flip\ldots\ flop.
 

### 2022-08-18 main/chapter_searches.tex
begin signal variations sections; srllbb command; various tweaks
hugely; more singlespacing for epigraphs
add some words on symmetries
add searches cartoons; add prob command

@@
 \chapter{Searches in ATLAS}
-\label{chapter:intro}
+\label{chapter:searches}
 %
@@
 
-\section{Searches}
-\label{sec:searches_searches}
-% histograms, CR VR SR, systematics
+\section{Data analysis}
+\label{sec:searces_data_analysis}
 
-\section{Data analysis}
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.25\textwidth]{figures/searches_beans.pdf}
+\caption[
+How many beans are in the jar?
+]{%
+How many beans are in the jar? This jar of jelly beans is a vehicle for our
+discussion of inference and decision theory.
+}
+\label{fig:significant_beans}
+\end{figure}
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/searches_baton_roue_bayes.jpg}
+\caption[
+Bayesian analysis produces two outputs: the evidence and posterior
+]{%
+Bayesian analysis produces two outputs: the
+evidence $\prob{d}{h}$ and
+posterior $\prob{x}{d,h}$ for hypothesis $h$, data $d$, and parameters $x$.
+Posterior-only, semi-Bayesian methods are legitimately criticized because they
+do not challenge the hypotheses that assign their priors $\prob{x}{h}$.
+Since the evidence is a likelihood function over hypotheses, considering the
+evidence can challenge those priors.
+\\[0.5em]
+Cartoon adapted from ``Baton roue'' by
+Corentin~Penloup~\cite{penloup2011baton}.
+}
+\label{fig:searches_baton_roue_bayes}
+\end{figure}
+
+\begin{singlespacing}
 \section{Frequentist theory}
@@
 \end{epigraphs}
+\end{singlespacing}
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.55\textwidth]{figures/searches_significant_shrink.png}
+\\
+\begin{footnotesize}
+`So, uh, we did the green study again and got no link. It was probably a{-}{-}'
+\\
+`RESEARCH CONFLICTED ON GREEN JELLY BEAN ACNE LINK; MORE~STUDY~RECOMMENDED!'
+\end{footnotesize}
+\caption[
+Significant by Randall~Munroe
+]{%
+Significant by Randall~Munroe~\cite{xkcd2011significant}.
+}
+\label{fig:searches_significant}
+\end{figure}
+
+\begin{singlespacing}
 \section{Post-frequentist practice}
@@
 \end{epigraphs}
+\end{singlespacing}
 
+% significance measures
+The ATLAS-recommended significance measure~\cite{atlas_significance} is
+\begin{align}
+\label{eqn:significance_atlas}
+S_\mathrm{\atlas}(n; \mu, \sigma) =~&
+\sqrt{2} \times
+\mathrm{sign}(n - \mu) \times
+\\[0.2em] \nonumber
+&
+\sqrt{
+n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
+- \frac{\mu^2}{\sigma^2}\log\left(
+1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
+\right)
+}
+,
+\end{align}
+in an appropriate limit where $0\log(0x) = 0$ for the $n=0$ case.
+
+\section{Searches}
+\label{sec:searches_searches}
+% histograms, CR VR SR, systematics
 % histfactory
@@
 
-HistFactory~\cite{cranmer2012histfactory}
-pyhf~\cite{heinrich2021pyhf}
-HistFitter~\cite{Besjes_2015,baak2015histfitter}
-
-% significancen measures
-The ATLAS-recommended significance measure~\cite{atlas_significance} is
-\begin{align}
-\label{eqn:significance_atlas}
-S_\mathrm{\atlas}(n; \mu, \sigma) =~&
-\mathrm{sgn}(n - \mu) \times \sqrt{2} \\[0.2em] \nonumber
-&
-\times \sqrt{
-n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
-- \frac{\mu^2}{\sigma^2}\log\left(
-1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
-\right)
-}
-.
-\end{align}
+\histfactory~\cite{cranmer2012histfactory}
+\pyhf~\cite{heinrich2021pyhf}
+\histfitter~\cite{Besjes_2015,baak2015histfitter}
 

### 2022-08-18 main/chapter_searches.tex
write intro to searches chapter

@@
 \end{singlespacing}
+
+Our task in scientific data analysis is to distil an informative
+message from data and to report that message to the wider community.
+From any given data there are many possible messages that could be extracted,
+and those messages will have different values to different recipients of our
+messages --- our data analysis is not an automatic process, but a result
+of many judgements that are informed by theory, convention, and intuition.
+
+To clarify our later description of the $\twoljets$ search in
+Chapter~\ref{chapter:2ljets}, this chapter introduces the basic theory and
+practice of data analysis and its manifestations in the \atlas\ SUSY community.
+
+Theoretical data analysis in the tightly coupled fields of probability theory
+and decision theory is introduced in Section~\ref{sec:searces_data_analysis}.
+The attempted alternative formulation of frequentist theory is discussed in
+Section~\ref{sec:searches_frequentist}.
+In practice, our field of research applies modified frequentist methods to
+avoid up some of their more drastic flaws; these practices are
+described in Section~\ref{sec:searches_practice}.
+Searches for Supersymmetric signals in \atlas\ follow a standard strategy for
+their data analysis; its procedures, nomenclature, and parametric modelling
+are introduced in Section~\ref{sec:searches_searches}.
 
@@
 \section{Frequentist theory}
-\label{sec:searches_frequentism}
+\label{sec:searches_frequentist}
 \begin{epigraphs}
@@
 \caption[
-Significant by Randall~Munroe
+``Significant'' by Randall~Munroe
 ]{%
-Significant by Randall~Munroe~\cite{xkcd2011significant}.
+``Significant'' by Randall~Munroe~\cite{xkcd2011significant}.
 }

### 2022-08-18 main/chapter_searches.tex
scribble on probability garbage

@@
 and decision theory is introduced in Section~\ref{sec:searces_data_analysis}.
-The attempted alternative formulation of frequentist theory is discussed in
-Section~\ref{sec:searches_frequentist}.
-In practice, our field of research applies modified frequentist methods to
-avoid up some of their more drastic flaws; these practices are
-described in Section~\ref{sec:searches_practice}.
+An attempted alternative formulation, frequentist theory, is discussed in
+Section~\ref{sec:searches_frequentist} for its relevance to our methods.
+Standard practice in this field of research does not strictly follow
+frequentist methods, but patches up some of their pathologies with modified
+approaches that we describe in Section~\ref{sec:searches_practice}.
 Searches for Supersymmetric signals in \atlas\ follow a standard strategy for
@@
 \label{sec:searces_data_analysis}
+How many jelly beans are in the jar of Figure~\ref{fig:significant_beans}?
+Your answer depends not only on what you observe from the figure, but also on
+what else you know about cartoon jars of beans, and on what reward is on offer.
+Each answer is a decision that depends both on what the possible rewards are,
+and on which numbers of beans you subjectively infer to be likely.
+We shall return to this example after developing some probability and decision
+theory.
 
@@
 discussion of inference and decision theory.
+Please write your guess on this page.
+The first correct answer wins a gold star.
 }
@@
 
+We begin with what probability is.
+At its most abstract, a probability a proportion of stuff within a larger
+whole, and probabilities are useful because they can be manipulated with an
+algebra that corresponds to how one naturally manipulates such proportions.
+
+\paragraph{Probability, abstract}
+A probability $\prob{b}{a}$ is a number that quantifies the proportion
+of $b$ within $a$, and can be expressed as a ratio of measures
+as~\cite{axioms1010038}
+\begin{equation}
+\label{eqn:searches_prob_ratio}
+\prob{b}{a} = \frac{m(b, a)}{m(a)},
+\end{equation}
+in which $m(x)$ is a measure (a non-negative quantity assigned to its
+argument, such as mass of tea in a cup), and the comma ``,'' means `and'
+(logical and or set intersection or any other equivalent operation).
+To continue that example, $\prob{b}{a}$ could be a proportion of tea within
+a spoon within a cup, where $m(b \land a)$ measures the volume of tea in a
+teaspoon that is partially submerged in the cup.%
+\footnote{%
+All probabilities are conditional in this definition; $\probc{b}$ alone does
+not exist unless the conditioning information $a$ is implicit from context.%
+}
+
+Logic is an important application of probability for which $a$ and $b$ are
+interpreted as logical propositions, and
+$\prob{b}{a}$ is the degree to which $a$ implies $b$, with the limiting cases
+that $\prob{b}{a} = 1$ when $a$ implies $b$ and $\prob{b}{a} = 0$ when $a$
+disproves $b$.
+Indeed Cox's theorem derives the laws of probability by requiring compatibility
+with Boolean logic~\cite{
+cox1946probability,
+cox1961algebra,
+garrett1998nand,
+jaynes2003probability
+keynes1920treatise,
+}.
+Probability works for logic and it works \emph{for tea, too}.
+
+Probability algebra has two rules that can derive from the measure ratio of
+Equation~\ref{eqn:searches_prob_ratio}:
+a \textbf{product rule}
+\begin{equation}
+\prob{c, b}{a} = \prob{c}{b, a}\times \prob{b}{a},
+\end{equation}
+and a \textbf{sum rule}
+\begin{equation}
+\prob{c\textrm{ or }b}{a} = \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a}.
+\end{equation}
+To avoid of this subtraction we prefer split any problem into orthogonal or
+disjoint states; for such cases, the normalization condition is that
+\begin{equation}
+\sum_i \prob{b_i}{a} = 1.
+\end{equation}
+These rules are basic and well known, but we state them here to support the
+following statements of how probabilities can be applied for practical
+results.%
+\footnote{%
+The formulation with sums and products is not strictly unique, but all
+alternatives are equivalent up to an invertible
+transformation~\cite{axioms1010038}.
+This choice is usually best because it uses ordinary addition and
+multiplication.
+In numerical analysis, however, it is often practical to use $\log$
+probabilities (to avoid overflows outside the range representable by floating
+point numbers(, thus changing multiplication to addition, and addition to
+the `$\log$-add-$\exp$' operation.
+}
+
+\paragraph{Probability, applied}
+% science snake
+
+Probabilistic data analysis proceeds as follows:
+\begin{enumerate}
+\item Predict new data $d$ with probability $\prob{d}{\ldots}$.
+\item Update your predictions with the observed data to $\prob{x}{d, \ldots}$.
+\item Go to 1.
+\end{enumerate}
+
+% evidence
+
+% bayes
+Bayes' theorem is an infamous and necessary result of probability.
+
+% posterior
+
+\paragraph{Probability assignments}
+infamously arbitrary and subjective. This is reality. They are.
+Consider this simple mathematical example with two propositions
+$a=\textrm{``}x\textrm{ is even''}$ and
+$b=\textrm{``}x\textrm{ equals }2\textrm{''}$.
+Clearly $\prob{a}{b} = 1$ since $2$ is indeed even.
+But what is $\prob{b}{a}$? We don't know. Perhaps one could argue symmetry
+across all even numbers, but there are infinitely many of them so that
+probability distribution could not be normalized.
+
+\TODO{\ldots}
+
+\paragraph{Decision theory}
+
+% summary
+
+Figure~\ref{fig:searches_baton_roue_bayes}
+
 \begin{figure}[tp]
 \centering
-\includegraphics[width=0.6\textwidth]{figures/searches_baton_roue_bayes.jpg}
+\includegraphics[width=0.65\textwidth]{figures/searches_baton_roue_bayes.jpg}
 \caption[
@@
 \end{singlespacing}
+
+Figure~\ref{fig:searches_significant}
 

### 2022-08-18 main/chapter_searches.tex
fix a cite

@@
 garrett1998nand,
-jaynes2003probability
-keynes1920treatise,
+jaynes2003probability,
+keynes1920treatise
 }.

### 2022-08-19 main/chapter_searches.tex
add more prob words; add a sig bkg figure to explain prior likelihood difference

@@
 
-We begin with what probability is.
-At its most abstract, a probability a proportion of stuff within a larger
-whole, and probabilities are useful because they can be manipulated with an
-algebra that corresponds to how one naturally manipulates such proportions.
+Statistical data analysis suffers from dogma --- prescriptions and recipes that
+students are taught to memorize and follow without knowing their reasons or
+limitations.
+From a standard undergraduate physics education, my experience has been
+error propagation formulae,
+least squares fitting routines,
+chi-squared per degree of freedom,
+and myriad statistical tests and confidence intervals.
+Many of these are effective approximations, but all also fail if misused
+outside of their domain.%
+\footnote{%
+Multiplicative error propagation is an example.
+For $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
+the perturbative approximation is
+\(
+\widetilde{\sigma}_{ab}
+= \mu_a \mu_b \sqrt{\sigma_a^2/\mu_a^2 + \sigma_b^2/\mu_b^2}
+% = \sigma_a^2 \mu_b^2 + \sigma_b^2 \mu_a^2
+\),
+but calculation with their moments derives
+\(
+\sigma_{ab}
+= \sqrt{(\mu_a^2 + \sigma_a^2)(\mu_b^2 + \sigma_b^2) - \mu_a^2\mu_b^2}
+% = \sigma_a^2 \mu_b^2 + \sigma_b^2 \mu_a^2 + \sigma_a^2 \sigma_b^2
+\)
+exactly if independent.
+Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, the
+propagation formula works only for small relative errors.
+}
+All effective methods of data analysis can be derived upwards from the roots of
+probability and decision theory, and as elsewhere in physics, by knowing the
+derivations we can also know when each approximation should work and when it
+needs replacing.
+
+We begin with what probability is: at its most abstract, a probability a
+proportion of stuff within a larger whole.
+Probabilities useful because proportions have applications, and because they
+can be manipulated and interrelated with a linear and intuitive algebra.
 
@@
 A probability $\prob{b}{a}$ is a number that quantifies the proportion
-of $b$ within $a$, and can be expressed as a ratio of measures
-as~\cite{axioms1010038}
+of $b$ within $a$, and can be expressed as a ratio~\cite{axioms1010038}
 \begin{equation}
@@
 in which $m(x)$ is a measure (a non-negative quantity assigned to its
-argument, such as mass of tea in a cup), and the comma ``,'' means `and'
+argument, such as mass of tea in a cup), and the comma `,' means `and'
 (logical and or set intersection or any other equivalent operation).
 To continue that example, $\prob{b}{a}$ could be a proportion of tea within
-a spoon within a cup, where $m(b \land a)$ measures the volume of tea in a
+a spoon within a cup, where $m(b, a)$ measures the volume of tea in a
 teaspoon that is partially submerged in the cup.%
@@
 
-Probability algebra has two rules that can derive from the measure ratio of
-Equation~\ref{eqn:searches_prob_ratio}:
+Two algebraic rules derive from Equation~\ref{eqn:searches_prob_ratio}:
 a \textbf{product rule}
+\begin{align}
+\prob{c, b}{a} &= \prob{c}{b, a}\times \prob{b}{a},
+\label{eqn:searches_product_rule}
+\intertext{%
+that describes nested proportions
+(perhaps a spoon within a teacup within a sink),
+and a \textbf{sum rule}%
+}
+\prob{c\vee b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
+\label{eqn:searches_sum_rule}
+\end{align}
+that describes combination in a union $\vee$ that coincides with logical `or'.
+The subtraction avoids double-counting of any overlap, and is often avoided by
+splitting any problem into orthogonal, or disjoint states.
+For disjoint cases with labels $b_i$, normalization therefore requires that
 \begin{equation}
-\prob{c, b}{a} = \prob{c}{b, a}\times \prob{b}{a},
-\end{equation}
-and a \textbf{sum rule}
-\begin{equation}
-\prob{c\textrm{ or }b}{a} = \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a}.
-\end{equation}
-To avoid of this subtraction we prefer split any problem into orthogonal or
-disjoint states; for such cases, the normalization condition is that
-\begin{equation}
+\label{eqn:searches_normalization}
 \sum_i \prob{b_i}{a} = 1.
@@
 In numerical analysis, however, it is often practical to use $\log$
-probabilities (to avoid overflows outside the range representable by floating
-point numbers(, thus changing multiplication to addition, and addition to
+probabilities to avoid overflows outside the range representable by floating
+point numbers, thus changing multiplication to addition and addition to
 the `$\log$-add-$\exp$' operation.
@@
 % science snake
+Science involves making predictions, testing those predictions against data,
+and reporting what is learned from the results.
+Various probabilities are involved in this prediction and learning that have
+conventional names to communicate what we intend to do with them.
+Most important for prediction is the \textbf{likelihood}
+\begin{align}
+L(a) &= \prob{b}{a}
+\intertext{%
+that explores the probabilities assigned to a fixed result $b$ from different
+contexts $a$, or how well different hypotheses predicted a result.
+A likelihood is clearly a probability, but it is not a probability distribution
+--- only the left hand argument $b$ forms a normalized distribution.
+Alternatively, fixing the context gives a \textbf{prior}%
+}
+\pi(b) &= \prob{b}{a}
+\end{align}
+for fixed $a$, which is a normalized distribution over disjoint states $b$.
 
-Probabilistic data analysis proceeds as follows:
+Likelihood and prior are evidently related --- they explore two indices into
+the same object, two sides of the same coin, and are assigned by the same
+principles.
+But they are profoundly different objects that work for different purposes,
+and clarity in their distinction will help in our later understanding of
+practical methods.
+For observed data, we use likelihoods to compare which hypotheses predicted
+them better than others, and before observing data we use priors to describe
+what data might result from different models to design our experiments.
+
+Take for example an overflow histogram bin that requires $\met > 100\,\eV[G]$,
+and is part of a figure with stacked histograms from background
+and (supersymmetric) signal samples.
+If those histograms are normalized to unit area, they are prior distributions
+for events from the background and signal hypotheses.
+The likelihood function for this overflow bin is
+$\prob{\met > 100\,\eV[G]}{a}$, and depends on the hypothetical origin of that
+event: $a = \textrm{`background'}$ or $a = \textrm{`signal'}$.
+\TODO{transformation behaviours are important}
+
+Figure~\ref{fig:searches_sig_bkg_prior_likelihood}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=\textwidth]{figures/searches_sig_bkg_prior_likelihood.pdf}
+\caption[
+Background and signal prior densities
+]{%
+Background and signal prior densities in $\eV[G]$ and $\eV[G]^{1/2}$ units.
+When changing coordinates, the probability is a fluid of constant total volume.
+In each bin a likelihood function is indexed by the background--signal labels
+and is independent of the prior units since the area in each bin is invariant.
+}
+\label{fig:searches_sig_bkg_prior_likelihood}
+\end{figure}
+
+% evidence: lower level events with different $\met$ values, bin likelihood
+
+% likelihood ratio
+\begin{equation}
+\hphantom{\quad \quad \mid\mid \mathrm{context}}
+\frac{\prob{S}{d}}{\prob{B}{d}}
+=
+\!\!
+\underbrace{\frac{\prob{d}{S}}{\prob{d}{B}}}_\textrm{likelihood ratio}
+\!\!\times~
+\frac{\probc{S}}{\probc{B}}
+\quad \quad \mid\mid \mathrm{context}
+\end{equation}
+
+Density is a misleading term. Prior mass cannot be compressed, it can only be
+moved and redistributed. It is more like depth of tea than density of air.
+
+Probabilistic learning proceeds as follows:
 \begin{enumerate}
-\item Predict new data $d$ with probability $\prob{d}{\ldots}$.
-\item Update your predictions with the observed data to $\prob{x}{d, \ldots}$.
+\item Predict new data $d$ with probability $\prob{d}{\ldots, h}$.
+\item Update your predictions with the observed data to $\prob{x}{d, \ldots, h}$.
 \item Go to 1.
 \end{enumerate}
-
-% evidence
 

### 2022-08-19 main/chapter_2ljets.tex
add stops after paragraph headers

@@
 
-\paragraph{Alternative $\ttbar$}
+\paragraph{Alternative $\ttbar$.}
 Several $\ttbar$ variations model alternative simulations of the
@@
 
-\paragraph{Alternative \topother}
+\paragraph{Alternative \topother.}
 Variations in \topother\ processes, which matter in \srllbb, use
@@
 
-\paragraph{Cross-sections}
+\paragraph{Cross-sections.}
 As described in the paper~\cite{atlas2022searches},
@@
 
-\paragraph{Scale variations}
+\paragraph{Scale variations.}
 Scale variations are a standard way to assess the uncertainties in finite-order
@@
 
-\paragraph{Parton distribution variations}
+\paragraph{Parton distribution variations.}
 Like scale variations, uncertainties in parton distribution functions (PDFs)
@@
 
-\paragraph{Miscellanea}
+\paragraph{Miscellanea.}
 \begin{itemize}
@@
 
-\paragraph{Extrapolation}
+\paragraph{Extrapolation.}
 To evaluate each variation requires a new simulation with different parameters,
@@
 
-\paragraph{Evaluation}
+\paragraph{Evaluation.}
 Statistical precision has limited the utility of most variations described in
@@
 
-\paragraph{\crz-met}
+\paragraph{\crz-met.}
 Concerns about our modelling of $\zjets$ and \diboson\ backgrounds were
@@
 
-\paragraph{Sub-processes}
+\paragraph{Sub-processes.}
 Our high-level sample labels such as $\ttbar$, $\zjets$, \diboson\ and
@@
 
-\paragraph{Summary}
+\paragraph{Summary.}
 As I have argued quite incessantly throughout this thesis, pre-fit predictions
@@
 
-\paragraph{Correlations}
+\paragraph{Correlations.}
 All parameters of our \histfactory\ models are initially independent.
@@
 
-\paragraph{Parameters}
+\paragraph{Parameters.}
 Correlations between parameters do not indicate their central values.
@@
 
-\paragraph{Contours}
+\paragraph{Contours.}
 Exclusion contours, displayed in
@@
 
-\paragraph{Discovery}
+\paragraph{Discovery.}
 Our discovery region results,

### 2022-08-19 main/chapter_searches.tex
add stops after paragraph headers

@@
 
-\paragraph{Probability, abstract}
+\paragraph{Probability, abstract.}
 A probability $\prob{b}{a}$ is a number that quantifies the proportion
@@
 
-\paragraph{Probability, applied}
+\paragraph{Probability, applied.}
 % science snake
@@
 
-\paragraph{Probability assignments}
+\paragraph{Probability assignments.}
 infamously arbitrary and subjective. This is reality. They are.
@@
 
-\paragraph{Decision theory}
+\paragraph{Decision theory.}
 

### 2022-08-20 main/chapter_searches.tex
add some more bayes

@@
 conventional names to communicate what we intend to do with them.
-Most important for prediction is the \textbf{likelihood}
+Most important for prediction is a \textbf{likelihood}
 \begin{align}
-L(a) &= \prob{b}{a}
+L(a) &= \prob{b}{a} \\
+\label{eqn:searches_likelihood}
 \intertext{%
@@
 \pi(b) &= \prob{b}{a}
+\label{eqn:searches_prior}
 \end{align}
@@
 
-Take for example an overflow histogram bin that requires $\met > 100\,\eV[G]$,
-and is part of a figure with stacked histograms from background
-and (supersymmetric) signal samples.
-If those histograms are normalized to unit area, they are prior distributions
-for events from the background and signal hypotheses.
-The likelihood function for this overflow bin is
-$\prob{\met > 100\,\eV[G]}{a}$, and depends on the hypothetical origin of that
-event: $a = \textrm{`background'}$ or $a = \textrm{`signal'}$.
-\TODO{transformation behaviours are important}
-
-Figure~\ref{fig:searches_sig_bkg_prior_likelihood}
+To illustrate this, consider a histogram that bins event yields in
+$\met$, with distributions from background and (supersymmetric) signal samples
+overlaid.
+Example such histograms are displayed in
+Figure~\ref{fig:searches_sig_bkg_prior_likelihood}.
+Since these histograms are normalized to unit area, they show prior
+probability distributions along the bin axis; there is one prior
+for events from each of the background and signal hypotheses.
+Priors distribute unit probability explore along the bin axis.
+In each bin, however, a likelihood function compares the assignments from
+the two hypotheses: signal and background.
+Likelihoods explore vertically down the sample axis.
 
@@
 Background and signal prior densities in $\eV[G]$ and $\eV[G]^{1/2}$ units.
-When changing coordinates, the probability is a fluid of constant total volume.
+When changing coordinates, probability mass moves as a fluid of constant total
+volume.
 In each bin a likelihood function is indexed by the background--signal labels
@@
 
-% evidence: lower level events with different $\met$ values, bin likelihood
+Identical distributions are drawn in two different coordinates
+on the left and right of Figure~\ref{fig:searches_sig_bkg_prior_likelihood};
+they are identical in content, but differ in appearance because they use the
+standard approach of representing each prior distribution as a density.%
+\footnote{%
+Density is a somewhat misleading term here.
+Probability mass does not compress, it moves and redistributes --- it is more
+like depth of tea than density of air.%
+}
+To preserve the area of each bin, which shows its probability mass,
+bin heights therefore change to adapt to their widths that change due to the
+differing values of $\met$ and $\sqrt{\met}$.
+
+We typically approximate such histograms by Monte Carlo simulation methods,
+since simulations do not generate bin indices but event variables.
+We connect these concepts with probabilities here.
+First, some notation; we have hypotheses $h$ (signal or background),
+a parameter $x$ that we histogram ($x = \met$),
+and we observe a datum $d$ that datum indexes the histogram bins --- supposing
+an event is observed in the overflow bin, it says
+$d = `x > 100\,\eV[G]\textrm{'}$.
+This observed bin has a likelihood function
+\begin{equation}
+L(x) = \prob{d}{x, h} = \prob{d}{x} =
+\left\{
+\begin{matrix}
+1 & \textrm{if }~x > 100\,\eV[G], \\
+0 & \textrm{otherwise} \\
+\end{matrix}
+\right.
+\end{equation}
+that corresponds to Equation~\ref{eqn:searches_likelihood} by
+$b \leftarrow d$ and $a \leftarrow `x, h\textrm{'}$,
+and simplifies $\prob{d}{x, h} = \prob{d}{x}$ since binning is independent of
+hypothesis.
+Simulating events from background and signal hypotheses samples from their
+respective priors
+\begin{equation}
+\pi(x) = \prob{x}{h}
+\end{equation}
+that correspond to Equation~\ref{eqn:searches_prior} by
+$b \leftarrow x$ and $a \leftarrow h$.
+Although we previously used the same labels $a$ and $b$, those are only
+labels --- the words `prior' and `likelihood' state our intent to index into
+the left and right arguments respectively.
+
+When we fill a histogram by summing weights of simulated samples, we are
+calculating another likelihood over hypotheses $h$ alone, not $`x, h\textrm{'}$
+jointly.
+To disambiguate the two, we call this abstracted, $x$-independent likelihood
+the \textbf{evidence} $Z(h)$ for hypothesis $h$.
+We calculate this evidence
+\begin{align}
+Z
+= \prob{d}{h} &= \sum_i L(x_i) \,\pi(x_i) \\
+&= \int\! L(x) \,\mathrm{d}\pi(x),
+\label{eqn:searches_evidence_integral_measure}
+\end{align}
+where Equation~\ref{eqn:searches_evidence_integral_measure} is notation for
+integrating the likelihood function over the prior measure~\cite{
+billingsley2008probability,
+skilling2010foundations
+}, and relates most closely to our practice of simulating from the model and
+counting events that hit the non-zero likelihood in the bin.
+We shall shortly derive this identity.
+The name `evidence' relates to its application for model comparison.
+Supposing we have this one event, did it come from the background or signal
+sample?
+The answer depends on the signal and background cross-sections in our collider,
+but larger evidence for signal does influence any conclusion to move in its
+favour.
+
+% posterior
+Perhaps we now want to see where in the overflow bin this event landed, and
+compare the signal and background distributions there.
+This is an inspection of the \textbf{posterior}
+\begin{equation}
+p(x) = \prob{x}{d, h} = \frac{L(x)\,\pi(x)}{Z},
+\end{equation}
+which is a prior probability distribution specialized by the data;
+it simply uses the likelihood to filter simulated samples
+(perhaps in `skimming an ntuple')
+and re-normalizes them to a probability distribution within the constraint.
+
+We finally have the four parts of `Bayesian' data analysis:
+the two inputs \textbf{likelihood} and \textbf{prior},
+and two outputs \textbf{posterior} and \textbf{evidence}.
+These relate through Bayes' rule, which is a consequence of the
+produce rule of Equation~\ref{eqn:searches_product_rule} and the
+commutativity of `and' --- that `$a$ and $b$' is equivalent to
+`$b$ and $a$';
+the joint probability of data and parameters has two identical
+forms, $\prob{d, x}{h}$ and $\prob{x, d}{h}$, that each factor into two
+pieces:
+\begin{align}
+\prob{d}{x, h}\times \prob{x}{h} &= \prob{x}{d, h}\times \prob{d}{h},
+\intertext{all of which which we have named:}
+\underbrace{L(x)}_\textbf{likelihood}
+\times
+\overbrace{\pi(x)}^\textbf{prior}
+~\,
+&=
+\overbrace{p(x)}^\textbf{posterior}
+\times
+\underbrace{Z}_\textbf{evidence}
+.
+% \textbf{likelihood} \times \textbf{prior}
+% &= \textbf{posterior} \times \textbf{evidence} .\nonumber
+\end{align}
 
@@
 
-Density is a misleading term. Prior mass cannot be compressed, it can only be
-moved and redistributed. It is more like depth of tea than density of air.
-
 Probabilistic learning proceeds as follows:
@@
 % bayes
-Bayes' theorem is an infamous and necessary result of probability.
+\TODO{move to connection to frequentist section}
+Bayes' theorem is an infamous and necessary result, and it is not seriously
+disputed. What has been disputed is whether probability can be applied to
+logic, or any application other than the random happenings of events.
 

### 2022-08-20 main/chapter_searches.tex
autocheck; shrink xkcd to fit page

@@
 error propagation formulae,
-least squares fitting routines,
+least-squares fitting routines,
 chi-squared per degree of freedom,
@@
 Many of these are effective approximations, but all also fail if misused
-outside of their domain.%
+outside their domain.%
 \footnote{%
@@
 disproves $b$.
-Indeed Cox's theorem derives the laws of probability by requiring compatibility
+Indeed, Cox's theorem derives the laws of probability by requiring compatibility
 with Boolean logic~\cite{
@@
 }.
-Probability works for logic and it works \emph{for tea, too}.
+Probability works for logic, and it works \emph{for tea, too}.
 
@@
 \centering
-\includegraphics[width=0.55\textwidth]{figures/searches_significant_shrink.png}
+\includegraphics[width=0.46\textwidth]{figures/searches_significant_shrink.png}
 \\
@@
 
+\TODO{minimize energy not max logf}
 To model the consequences of these many factors of uncertainty, each factor is
-assigned a real-valued parameter which when varied changed the expected yields
-in the hisogram bins of the search.
+assigned a real-valued parameter, which when varied changed the expected yields
+in the histogram bins of the search.
 In this way, we form a statistical model which is a function
@@
 \end{equation}
-from the parameters $\theta$ to Poisson expectations $\vec\mu$ the histogram
-bins and a regularization term $\log f$ which is maximized in fitting.
+from the parameters $\theta$ to Poisson expectations $\vec\mu$ of the histogram
+bins and a regularization term $\log f$ that is maximized in fitting.
 

### 2022-08-22 main/chapter_searches.tex
some data analysis words

@@
 can be manipulated and interrelated with a linear and intuitive algebra.
+Probabilisitc data analysis follows the ordinary structure of science ---
+write down models that make predictions, test those predictions against data,
+then update your models to refine their predictions.
+For this process, likelihoods are particularly important because they describe
+our predictions. But not everything is a likelihood.
+The remainder of this section uses a standard example from High Energy Physics
+to introdule this probabilistic structure and other parts of its technical
+language.
 
@@
 When changing coordinates, probability mass moves as a fluid of constant total
-volume.
-In each bin a likelihood function is indexed by the background--signal labels
-and is independent of the prior units since the area in each bin is invariant.
+volume, changing its shape in this density view to keep the area of the
+rectangle in each bin constant.
+A likelihood function indexes each bin in the background--signal labels,
+and is independent of the coordinates units since the area in each bin is
+invariant.
 }
@@
 differing values of $\met$ and $\sqrt{\met}$.
+This coordinate-dependence of the prior density means that its maximum is
+not always meaningful.
+Since the probability mass in each bin is constant, the likelihood is
+coordinate-independent.
 
@@
 \prob{d}{x, h}\times \prob{x}{h} &= \prob{x}{d, h}\times \prob{d}{h},
-\intertext{all of which which we have named:}
+\intertext{all of which we have named:}
 \underbrace{L(x)}_\textbf{likelihood}
@@
 \end{align}
+This is (the infamous) Bayes' rule, derived as a consequence of the
+symmetry of logical `and' and the product rule of probability
+algebra.
 

### 2022-08-22 main/chapter_searches.tex
add bayes Laplace references and some more words

@@
 hypothesis.
+\TODO{binary filter --- others same but not binary}
+
 Simulating events from background and signal hypotheses samples from their
@@
 algebra.
+\TODO{here state how we should use this}.
+
+The infamy of Bayes' rule stems from a fixation on the posterior distributions
+from unjustified prior assignments, which has been present since its earliest
+classical developments by Bayes and
+Laplace~\cite{bayes1763lii, laplace1774stigler}.
+Both Bayes and Laplace assume prior densities that are uniform in their chosen
+coordinates, a practice that is fairly criticized as arbitrary
+\TODO{zero evidence because for uniform on the real line --- a normalized prior
+with any inkling of the scale of the actual parameter will be infinitely
+preferred in its evidence.}
 
@@
 \TODO{\ldots}
+
+Succinctly stated by Jaynes that
+``any function with a single rounded maximum, raised to a higher and higher
+power, goes into a Gaussian function''~\cite{jaynes2003probability}.
 

### 2022-08-23 main/chapter_searches.tex
add discussion of evidence

@@
 
-We typically approximate such histograms by Monte Carlo simulation methods,
-since simulations do not generate bin indices but event variables.
-We connect these concepts with probabilities here.
-First, some notation; we have hypotheses $h$ (signal or background),
+One typically approximates such histograms by Monte Carlo simulation methods,
+since simulations do not generate bin indices but event variables that derive
+from the sampled events.
+To connect these concept with probabilities, we begin with some notation;
+we have hypotheses $h$ (signal or background),
 a parameter $x$ that we histogram ($x = \met$),
@@
 $b \leftarrow d$ and $a \leftarrow `x, h\textrm{'}$,
-and simplifies $\prob{d}{x, h} = \prob{d}{x}$ since binning is independent of
-hypothesis.
-\TODO{binary filter --- others same but not binary}
+and simplifies $\prob{d}{x, h} = \prob{d}{x}$ since such binning is independent
+of hypothesis.%
+\footnote{%
+Independence of the likelihood function from prior hypotheses is so common
+that it should usually be assumed unless otherwise stated.
+}
+This example of a binary selection highlights how the likelihood acts as a
+filter that rejects events that are contradicted by the data.
+Other likelihoods that are not binary act similarly, but as inefficient filters
+that reject events with some probability.
 
@@
 the \textbf{evidence} $Z(h)$ for hypothesis $h$.
-We calculate this evidence
+The evidence is calculable as
 \begin{align}
@@
 The name `evidence' relates to its application for model comparison.
-Supposing we have this one event, did it come from the background or signal
-sample?
+Supposing we have this one event with $\met > 100\,\eV[G]$, did it come from
+the background or signal sample?
 The answer depends on the signal and background cross-sections in our collider,
@@
 favour.
+
+This expression integral, or average over the prior, means that evidence $Z$
+for model arises from a larger fraction of samples from it hitting large values
+of the likelihood function.
+In contrast, one might consider a maximum likelihood approach that constrained
+an optimization to regions of nonzero prior in each model.
+Such optimizations can give a useful upper bound on $Z$, but this example
+highlights a flaw since both models support the $\met > 100\,\eV[G]$ bin
+and both support the same maximum likelihood of
+$L(\varepsilon + 100\,\eV[G]) = 1$.
 
@@
 \times
+~\,
 \overbrace{\pi(x)}^\textbf{prior}
@@
 algebra.
+All influence from data is conducted through the likelihood function
 \TODO{here state how we should use this}.

### 2022-08-23 main/chapter_searches.tex
some text on use of bayes rule

@@
 algebra.
-All influence from data is conducted through the likelihood function
-\TODO{here state how we should use this}.
 
-The infamy of Bayes' rule stems from a fixation on the posterior distributions
-from unjustified prior assignments, which has been present since its earliest
-classical developments by Bayes and
-Laplace~\cite{bayes1763lii, laplace1774stigler}.
-Both Bayes and Laplace assume prior densities that are uniform in their chosen
-coordinates, a practice that is fairly criticized as arbitrary
-\TODO{zero evidence because for uniform on the real line --- a normalized prior
-with any inkling of the scale of the actual parameter will be infinitely
-preferred in its evidence.}
+How Bayes' rule \emph{should} be used in scientific data analysis is to extract
+useful information from the data.
+If the likelihood function itself is interpretable, then it should be reported
+directly.
+When it is not, perhaps due to a complicated parametric modelling, then one
+should abstract those parameters away by reporting evidence values for
+predictive models, which must of course be clearly stated and interpretable if
+they are to be useful.
+Data act only through the likelihood function, and reporting a posterior
+can act to obfuscate those data.
 
+Posterior probabilities are, however, useful when digesting data to make
+inferences about the world.
+Reporting likelihoods and evidence values allows readers to assign priors as
+they wish.
+For model comparison between background $h=B$ and signal $h=S$, this is clearly
+expressed in the ratio form
 % likelihood ratio
@@
 \!\!
-\underbrace{\frac{\prob{d}{S}}{\prob{d}{B}}}_\textrm{likelihood ratio}
+\frac{\prob{d}{S}}{\prob{d}{B}}
 \!\!\times~
@@
 \quad \quad \mid\mid \mathrm{context}
+,
 \end{equation}
+where we use the double bar, $\mid\mid$, to refer to any contextual information
+a reader might wish to include.
+Whenever the evidence ratio is independent of that context, it factors out from
+the prior (odds) ratio.
+In this sense, likelihood (or evidence) ratios are scale factors by which prior
+odds are updated.
 
-Probabilistic learning proceeds as follows:
+The infamy of Bayes' rule stems from a fixation on the posterior distributions
+from unjustified prior assignments.
+Classical works by Bayes and Laplace~\cite{bayes1763lii, laplace1774stigler}
+began this focus;
+both assume prior densities that are uniform in their chosen coordinates, a
+practice that can yield practical results in posterior inferences.
+But for our purposes of reporting scientific data, such practices are fairly
+criticized as arbitrary when they influce our results.
+\TODO{references for definition of Bayesian analysis as computation of
+posterior without even giving a symbool to the evidence}
+Figure~\ref{fig:searches_baton_roue_bayes}
+
+\TODO{zero evidence because for uniform on the real line --- a normalized prior
+with any inkling of the scale of the actual parameter will be infinitely
+preferred in its evidence.}
+oceeds as follows:
 \begin{enumerate}
@@
 \paragraph{Decision theory.}
-
-% summary
-
-Figure~\ref{fig:searches_baton_roue_bayes}
 

### 2022-08-23 main/chapter_2ljets.tex
more probability posteriors etc, ref cartoon

@@
 continuous of `unbinned' distributions with Poisson normalizations
-(sometimes called ``extended'' in high-energy
-physics~\cite{barlow1990extended})
-that are standard elsewhere in high-energy physics,
+(sometimes called ``extended'' in High-Energy
+Physics~\cite{barlow1990extended})
+that are standard elsewhere in High-Energy Physics,
 and are easily analysed by the standard likelihood fitting

### 2022-08-23 main/chapter_searches.tex
more probability posteriors etc, ref cartoon

@@
 Z
-= \prob{d}{h} &= \sum_i L(x_i) \,\pi(x_i) \\
+= \prob{d}{h} &= \sum_i L(x_i) \,\pi(x_i)
+\label{eqn:searches_evidence_integral_sum}
+\\
 &= \int\! L(x) \,\mathrm{d}\pi(x),
@@
 We shall shortly derive this identity.
-The name `evidence' relates to its application for model comparison.
+The name `evidence' relates to its application for model
+comparison~\cite{mackay2003information}.
 Supposing we have this one event with $\met > 100\,\eV[G]$, did it come from
@@
 the prior (odds) ratio.
-In this sense, likelihood (or evidence) ratios are scale factors by which prior
-odds are updated.
+In this sense, likelihood (or evidence) ratios are scale factors that update
+any odds ratio by the same multiplication.
 
@@
 from unjustified prior assignments.
-Classical works by Bayes and Laplace~\cite{bayes1763lii, laplace1774stigler}
-began this focus;
-both assume prior densities that are uniform in their chosen coordinates, a
-practice that can yield practical results in posterior inferences.
-But for our purposes of reporting scientific data, such practices are fairly
-criticized as arbitrary when they influce our results.
-\TODO{references for definition of Bayesian analysis as computation of
-posterior without even giving a symbool to the evidence}
-Figure~\ref{fig:searches_baton_roue_bayes}
-
-\TODO{zero evidence because for uniform on the real line --- a normalized prior
-with any inkling of the scale of the actual parameter will be infinitely
-preferred in its evidence.}
-oceeds as follows:
-\begin{enumerate}
-\item Predict new data $d$ with probability $\prob{d}{\ldots, h}$.
-\item Update your predictions with the observed data to $\prob{x}{d, \ldots, h}$.
-\item Go to 1.
-\end{enumerate}
-
-% bayes
-\TODO{move to connection to frequentist section}
-Bayes' theorem is an infamous and necessary result, and it is not seriously
-disputed. What has been disputed is whether probability can be applied to
-logic, or any application other than the random happenings of events.
-
-% posterior
-
-\paragraph{Probability assignments.}
-infamously arbitrary and subjective. This is reality. They are.
-Consider this simple mathematical example with two propositions
-$a=\textrm{``}x\textrm{ is even''}$ and
-$b=\textrm{``}x\textrm{ equals }2\textrm{''}$.
-Clearly $\prob{a}{b} = 1$ since $2$ is indeed even.
-But what is $\prob{b}{a}$? We don't know. Perhaps one could argue symmetry
-across all even numbers, but there are infinitely many of them so that
-probability distribution could not be normalized.
-
-\TODO{\ldots}
-
-Succinctly stated by Jaynes that
-``any function with a single rounded maximum, raised to a higher and higher
-power, goes into a Gaussian function''~\cite{jaynes2003probability}.
-
-\paragraph{Decision theory.}
+Classical works by
+Bayes~\cite{bayes1763lii} and
+Laplace~\cite{laplace1774stigler} began this focus ---
+both assume prior densities that are uniform in chosen coordinates, a practice
+that can yield practical results in posterior inferences.
+But to report scientific data, such practices are fairly criticized as
+arbitrary when they influce the reported results.
+Prior assignments are arbitrary.
+You are welcome to propose any prediction in any circumstances, and the
+scientific response is to test whether or not its predictions are compatible
+with our observations of nature.
+Our emphasis on the likelihood functions and evidence values that quantify
+those predictions avoids the pitfall of seeking uninformative, neutral, or
+unbiased priors by embracing priors as active and explicit predictions.
+This is the modern, practical approach to Bayesian data analysis~\cite{
+mackay2003information,
+skilling2004nested,
+skilling2006nested,
+sivia2006data,
+skilling2010foundations
+}.
 
@@
 \end{figure}
+
+Some sources continue to define Bayesian analysis as the computation of
+posterior probabilities given data, often without naming the evidence, and
+expressing it only as the sum from
+Equation~\ref{eqn:searches_evidence_integral_sum}
+--- as just a normalizing constant~\cite{
+gelman1995bayesian,
+gelman2008objections,
+DAgostini:1994fjx,
+DAgostini:2010hil,
+cowan1998statistical,
+pdg2020review
+}.
+Bayes' rule has two outputs, and both outputs have uses.
+Cutting out the evidence hamstrings Bayesian analysis, leaving only the
+semi-Bayesian posterior-only approach and its clear flaws.
+A bicycle is stable with two wheels and falls over if one either
+is removed;
+Figure~\ref{fig:searches_baton_roue_bayes} illustrates this analogy.
+Although it is possible to ride a unicycle, it is a circus trick and not often
+a practical tool.
+
+The essence of this form of Probabilistic data analysis is most succinctly
+summarized by Skilling~\cite{skilling2008rant}:
+``we SHOULD use Bayes by assigning subjective priors and inspecting evidence
+values.''
+This agrees exactly with the standard practice of histogramming simulations,
+as illustrated in Figure~\ref{fig:searches_sig_bkg_prior_likelihood}.
+Signal and background models define priors, and their (normalized) bin yields
+are the evidence values that we use to compare the models.
+Finally corresponding posteriors are the normalized shapes those samples take
+within the observed histogram bin, which we do not use to compare the models,
+but which could in principle be used for secondary interpretations.
+
+\paragraph{Probability assignments.}
+Probability algebra manipulates probabilities, but does not state how they
+should be initially assigned.
+Probability assignments can, in general, be freely assigned and refined from
+calibration data. But there are unambiguous cases that are commonly used in
+High-Energy, notably the Poisson and Gaussian distributions.
+To avoid turning this thesis into a textbook, we only sketch the derivations
+of these, and refer to real textbooks for their detailed
+derivations~\cite{jaynes2003probability}.
+
+Both Poisson and
+Binary accept/reject experiments are described by the
+The Large Hadron Collider creates $40$~million proton collisions per second,
+only around $1000$ are accepted by \atlas' trigger system per
+second~\cite{atlas2020trigger}.
+This acceptance rate of $250$ parts per million is sufficient to invoke the
+Poisson distribution for event counts.
+Each collision event has some probability $p$ of passing the trigger
+\begin{equation}
+\label{eqn:searches_poisson}
+\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp{(-\lambda)}
+\end{equation}
+
+Succinctly stated by Jaynes that
+``any function with a single rounded maximum, raised to a higher and higher
+power, goes into a Gaussian function''~\cite{jaynes2003probability}.
+\begin{equation}
+\label{eqn:searches_gaussian}
+\mathrm{Gaussian}(x\mid \mu, \sigma) =
+\frac{1}{\sqrt{2\pi \sigma^2}}
+\exp{\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)}
+\end{equation}
+This is not the famous Central Limit Theorem, which is related to the
+distribution of sums, but can be used to derive it: the distribution of a sum
+is a convolution integral, which becomes a product in Fourier-transformed
+space, and that product can converge to a Gaussian function by this rule.
+Jaynes gives a beautiful alternative proof of the
+Central Limit Theorem~\cite{jaynes2003probability}
+(Section ``7.16 The central limit theorem'').
+
+\paragraph{Decision theory.}
+
+% connection to frequentist section
+Bayes' theorem is an infamous and necessary result, and it is not seriously
+disputed. What has been disputed is whether probability can be applied to
+logic, or any application other than the random happenings of events.
 

### 2022-08-23 main/draft.tex
more probability posteriors etc, ref cartoon

@@
 uses and more computationally efficient in evaluation.
+
+infamously arbitrary and subjective. This is reality. They are.
+Consider this simple mathematical example with two propositions
+$a=\textrm{``}x\textrm{ is even''}$ and
+$b=\textrm{``}x\textrm{ equals }2\textrm{''}$.
+Clearly $\prob{a}{b} = 1$ since $2$ is indeed even.
+But what is $\prob{b}{a}$? We don't know. Perhaps one could argue symmetry
+across all even numbers, but there are infinitely many of them so that
+probability distribution could not be normalized.
+
+\TODO{\ldots}

### 2022-08-23 main/chapter_searches.tex
mostly adding citations; some content on binomial

@@
 should be initially assigned.
-Probability assignments can, in general, be freely assigned and refined from
-calibration data. But there are unambiguous cases that are commonly used in
-High-Energy, notably the Poisson and Gaussian distributions.
+In general, such probability assignments are free and unconstrained, and might
+informed by calibration data or loosely motivated opinions.
+But there are examples of unambiguously derived that are commonly used in
+High-Energy Physics, notably the Poisson and Gaussian distributions.
 To avoid turning this thesis into a textbook, we only sketch the derivations
 of these, and refer to real textbooks for their detailed
-derivations~\cite{jaynes2003probability}.
+derivations~\cite{jaynes2003probability}, both of which use the exponential
+identity that
+\begin{equation}
+\label{eqn:searches_exponential}
+\exp(x) = \displaystyle \lim_{N \to \infty} (1 - \frac{x}{N})^N
+.
+\end{equation}
 
-Both Poisson and
-Binary accept/reject experiments are described by the
-The Large Hadron Collider creates $40$~million proton collisions per second,
-only around $1000$ are accepted by \atlas' trigger system per
-second~\cite{atlas2020trigger}.
-This acceptance rate of $250$ parts per million is sufficient to invoke the
-Poisson distribution for event counts.
-Each collision event has some probability $p$ of passing the trigger
+The Large Hadron Collider collides $40$~million proton bunches per second,
+but \atlas' trigger system accepts only only around $1000$ of
+those~\cite{atlas2020trigger}.
+This is a Binomial scenario with a large fixed number $N$ of trials
+(bunch collisions)
+and a small probability of acceptance (triggering),
+due to a notable collision event.
+The Binomial distribution for $n$ triggered events is well known as
+\begin{equation}
+\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
+,
+\end{equation}
+which is a product of the probabilities to the sequence of $N$ accept/reject
+data, summed over the permutations that could have led to the total $n$.
+With our extremely small trigger rate, we are interested in the limit of small
+$p$ and large $N$ which result in a mean acceptance number of $\lambda = Np$.
+In the large $N$ limit, this tends towards the Poisson
+distribution~\cite{jaynes2003probability}
 \begin{equation}
 \label{eqn:searches_poisson}
-\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp{(-\lambda)}
+\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp(-\lambda)
+,
 \end{equation}
+which is used ubiquitously in the analysis of our histograms.
 
+The Gaussian distribution has many justifications:
+as a practical approximation it permits analysis in linear algebra,
+as the unique distribution is uniquely spherically symmetric and factors into
+independent dimensions~\cite{
+jaynes2003probability,
+herschel1850normal,
+maxwell1860normal,
+muller1959note,
+marsaglia1972normal
+},
+and as the maximum entropy~\cite{PhysRev.106.620} distribution for a
+known mean and variance.
+\TODO{}
 Succinctly stated by Jaynes that
@@
 \frac{1}{\sqrt{2\pi \sigma^2}}
-\exp{\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)}
+\exp\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)
 \end{equation}

### 2022-08-24 main/chapter_searches.tex
add gaussian stuff

@@
 \label{eqn:searches_exponential}
-\exp(x) = \displaystyle \lim_{N \to \infty} (1 - \frac{x}{N})^N
+\exp(x) = \displaystyle \lim_{N \to \infty}
+\left(1 - \frac{x}{N}\right)^N
 .
@@
 those~\cite{atlas2020trigger}.
-This is a Binomial scenario with a large fixed number $N$ of trials
+This means we have scenario with a large fixed number $N$ of trials
 (bunch collisions)
-and a small probability of acceptance (triggering),
+and a small probability of binary acceptance (triggering),
 due to a notable collision event.
-The Binomial distribution for $n$ triggered events is well known as
+The summed total number $n$ of accepted events is therefore described by the
+Binomial distribution for $n$;
 \begin{equation}
@@
 \end{equation}
-which is a product of the probabilities to the sequence of $N$ accept/reject
-data, summed over the permutations that could have led to the total $n$.
+which is a product of the probabilities (produce rule) in the sequence of $N$
+accept/reject decisions, summed (sum rule) over all of that sequence
+permutations that could have led to the total $n$.
 With our extremely small trigger rate, we are interested in the limit of small
 $p$ and large $N$ which result in a mean acceptance number of $\lambda = Np$.
-In the large $N$ limit, this tends towards the Poisson
-distribution~\cite{jaynes2003probability}
+Using the exponential identity of Equation~\ref{eqn:searches_exponential}, one
+can show that this limit tends towards the Poisson
+distribution~\cite{jaynes2003probability};
 \begin{equation}
@@
 The Gaussian distribution has many justifications:
-as a practical approximation it permits analysis in linear algebra,
-as the unique distribution is uniquely spherically symmetric and factors into
-independent dimensions~\cite{
+as the unique distribution that is spherically symmetric and factors linearly
+into independent dimensions~\cite{
 jaynes2003probability,
@@
 },
+as a practical approximation that permits analysis in linear algebra,
 and as the maximum entropy~\cite{PhysRev.106.620} distribution for a
 known mean and variance.
-\TODO{}
-Succinctly stated by Jaynes that
-``any function with a single rounded maximum, raised to a higher and higher
-power, goes into a Gaussian function''~\cite{jaynes2003probability}.
+This maximum entropy statement means that for a distribution with known mean
+and variance in a given coordinates, its best Gaussian approximation matches
+those moments.\footnote{%
+By `best', this means minimal Kullback-Leibler divergence
+\(
+H(p\leftarrow q) = \int \log(p(x)/q(x)) \,\mathrm{d}p(x)
+\)
+from the Gaussian approximation $q$ to the target $p$ with known moments.
+Such an approximation may not be \emph{good}, but it is the best in its class.%
+}
+As a limiting approximation, the Gaussian form arises from products of many
+functions, such as combinations of many independent measurements that multiply
+their likelihood functions.
+At a maximum with $x=\mu$, the first derivative of a function vanishes and its
+second derivative is negative, so this product looks something like
+\begin{equation}
+f(x)^N = \displaystyle \lim_{N \to \infty}
+(1 - \kappa (x - \mu)^2)^N,
+\end{equation}
+which, again by Equation~\ref{eqn:searches_exponential}, goes into the Gaussian
+form
 \begin{equation}
 \label{eqn:searches_gaussian}
-\mathrm{Gaussian}(x\mid \mu, \sigma) =
+\mathrm{Gaussian}(x\mid \mu, \sigma)
+\,\mathrm{d}x =
 \frac{1}{\sqrt{2\pi \sigma^2}}
-\exp\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)
+\exp\!\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)
+\mathrm{d}x
 \end{equation}
+where the prefactor is for normalization to a probability distribution over
+$x$, and we include the $\mathrm{d}x$ terms as reminders that it is a density
+function in that coordinate.
+
+Here, $\kappa$ and $\sigma$ both relate to the second derivative of the target
+function; this relationship extends to multiple dimensions in the Hessian matrix
+and its inverse, and allows one to estimate error bars from second derivatives
+--- a practical tool that the methods discussed in this thesis employ.
+
 This is not the famous Central Limit Theorem, which is related to the
-distribution of sums, but can be used to derive it: the distribution of a sum
-is a convolution integral, which becomes a product in Fourier-transformed
-space, and that product can converge to a Gaussian function by this rule.
-Jaynes gives a beautiful alternative proof of the
-Central Limit Theorem~\cite{jaynes2003probability}
-(Section ``7.16 The central limit theorem'').
+distribution of sums.
+But can be used to derive it: the distribution of a sum is a convolution, which
+becomes a product in Fourier-transformed space, and that product can converge
+to a Gaussian function by this same derivation if the mean and variance exist.
+A beautiful alternative proof of the Central Limit Theorem is given by
+Jaynes~\cite{jaynes2003probability}
+(Section ``7.16 The central limit theorem''), and would be spoiled by a
+reproduction here.
 
@@
 
+\TODO{}
 % connection to frequentist section

### 2022-08-24 main/chapter_searches.tex
add ml practice paragraph

@@
 
-The essence of this form of Probabilistic data analysis is most succinctly
-summarized by Skilling~\cite{skilling2008rant}:
+In the words of Skilling~\cite{skilling2008rant},
 ``we SHOULD use Bayes by assigning subjective priors and inspecting evidence
 values.''
-This agrees exactly with the standard practice of histogramming simulations,
+This framing of probabilistic data analysis agrees exactly with out standard
+practice of histogramming simulations,
 as illustrated in Figure~\ref{fig:searches_sig_bkg_prior_likelihood}.
@@
 The Large Hadron Collider collides $40$~million proton bunches per second,
-but \atlas' trigger system accepts only only around $1000$ of
-those~\cite{atlas2020trigger}.
-This means we have scenario with a large fixed number $N$ of trials
+but \atlas' trigger system accepts only around $1000$ of
+those per second~\cite{atlas2020trigger}.
+This means that we have scenario with a large fixed number $N$ of trials
 (bunch collisions)
@@
 \(
-H(p\leftarrow q) = \int \log(p(x)/q(x)) \,\mathrm{d}p(x)
+H(p\leftarrow q) = \sum_i p_i \log(p_i/q_i)
 \)
 from the Gaussian approximation $q$ to the target $p$ with known moments.
-Such an approximation may not be \emph{good}, but it is the best in its class.%
+Such an approximation may not be \emph{good}, but it is best in class.%
 }
@@
 their likelihood functions.
-At a maximum with $x=\mu$, the first derivative of a function vanishes and its
-second derivative is negative, so this product looks something like
+At a maximum $\check{x}$, the first derivative of a function vanishes and its
+second derivative is negative, after Taylor expanding about its maximum this
+product looks like
 \begin{equation}
 f(x)^N = \displaystyle \lim_{N \to \infty}
-(1 - \kappa (x - \mu)^2)^N,
+\left(
+1 - \kappa (x - \check{x})^2
+\right)^N,
 \end{equation}
@@
 Here, $\kappa$ and $\sigma$ both relate to the second derivative of the target
-function; this relationship extends to multiple dimensions in the Hessian matrix
-and its inverse, and allows one to estimate error bars from second derivatives
---- a practical tool that the methods discussed in this thesis employ.
+function; this relationship extends to multiple dimensions in the Hessian
+matrix and its inverse, and allows one to estimate error bars from second
+derivatives --- a practical tool that the methods discussed in this thesis
+employ.
 
@@
 Jaynes~\cite{jaynes2003probability}
-(Section ``7.16 The central limit theorem''), and would be spoiled by a
+(Section ``7.16 The central limit theorem''), but would be spoiled by a
 reproduction here.
+
+In the current year, it would be remiss to not mention machine learning
+algorithms.
+Among various other applications, learning algorithms demonstrate practical
+successes in probability assignments, such as assigning
+probabilities to labels for classification~\cite{MurphyKevinP.2012Mlap},
+or to assign probabilities to words for text
+generation~\cite{radford2019language}.
+Machine learning tunes flexible functions by consuming data and testing their
+predictions in the manner we have described; their training and testing works
+roughly as follows:
+\begin{enumerate}
+\item Predict new data with $\prob{d_i}{f(d_{i-1}, \ldots), h}$.
+Keep this evidence as a testing score for comparison against other models.
+(Practically, keep its equivalent $\log$ mean over independent samples.)
+\item Use those data to update your state
+$f(d_{i-1}, \ldots) \rightarrow f(d_i, d_{i-1}, \ldots)$ to improve your
+predictions.
+Unlike ideal posterior updating, this does not exactly employ all information
+in the data, only a function of it --- one usually cannot afford to consume all
+aspects of data, so although theoretically suboptimal this approximation is
+vastly superior in practice.
+\item Go to 1.
+\end{enumerate}
+The sequence of data \TODO{}
 

### 2022-08-24 main/chapter_searches.tex
add some words on decision theory

@@
 Theoretical data analysis in the tightly coupled fields of probability theory
-and decision theory is introduced in Section~\ref{sec:searces_data_analysis}.
+and decision theory is introduced in Section~\ref{sec:searches_data_analysis}.
 An attempted alternative formulation, frequentist theory, is discussed in
@@
 \section{Data analysis}
-\label{sec:searces_data_analysis}
-How many jelly beans are in the jar of Figure~\ref{fig:significant_beans}?
+\label{sec:searches_data_analysis}
+How many jelly beans are in the jar of Figure~\ref{fig:searches_beans}?
 Your answer depends not only on what you observe from the figure, but also on
@@
 }
-\label{fig:significant_beans}
+\label{fig:searches_beans}
 \end{figure}
@@
 For $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
-the perturbative approximation is
+the standard perturbative approximation is
 \(
@@
 \),
-but calculation with their moments derives
+but a short calculation with their moments derives the exact result
 \(
@@
 \)
-exactly if independent.
+if $\probc{a}$ and $\probc{b}$ are independent.
 Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, the
@@
 can be manipulated and interrelated with a linear and intuitive algebra.
-Probabilisitc data analysis follows the ordinary structure of science ---
+Probabilistic data analysis follows the ordinary structure of science ---
 write down models that make predictions, test those predictions against data,
@@
 The remainder of this section uses a standard example from High Energy Physics
-to introdule this probabilistic structure and other parts of its technical
+to introduce this probabilistic structure and other parts of its technical
 language.
@@
 All probabilities are conditional in this definition; $\probc{b}$ alone does
-not exist unless the conditioning information $a$ is implicit from context.%
+not exist unless some conditioning information is implicit from context.%
 }
@@
 $\prob{b}{a}$ is the degree to which $a$ implies $b$, with the limiting cases
-that $\prob{b}{a} = 1$ when $a$ implies $b$ and $\prob{b}{a} = 0$ when $a$
-disproves $b$.
+that $\prob{b}{a} = 1$ when $a$ implies $b$, and $\prob{b}{a} = 0$ when $a$
+contradicts $b$.
 Indeed, Cox's theorem derives the laws of probability by requiring compatibility
@@
 =
-\!\!
 \frac{\prob{d}{S}}{\prob{d}{B}}
-\!\!\times~
+\times
 \frac{\probc{S}}{\probc{B}}
@@
 successes in probability assignments, such as assigning
-probabilities to labels for classification~\cite{MurphyKevinP.2012Mlap},
+probabilities to labels for classification,
 or to assign probabilities to words for text
-generation~\cite{radford2019language}.
+generation, or any of myriad other
+predictions~\cite{MurphyKevinP.2012Mlap, radford2019language}.
 Machine learning tunes flexible functions by consuming data and testing their
@@
 \end{enumerate}
-The sequence of data \TODO{}
+These sequential data $d_i$ might be labelled the `training', `validation',
+and `testing' sets, but other splittings are possible.
+This perspective highlights the importance of independent training and testing
+sets --- since $\prob{d}{d, \ldots} = 1$, if any memory of $d_i$ is kept in
+the $f(\ldots, d_i, \ldots)$ state, then it is a memory and not a prediction.
+Independent testing data permit the evidence interpretation of predictive
+probabilities.
 
 \paragraph{Decision theory.}
+Now that we have thoroughly covered probability theory, we are ready to return
+to the jelly bean game that is illustrated in Figure~\ref{fig:searches_beans}
+and asks you to guess how many beans are in the jar.
+Given all you know about illustrations of jars and beans, you could make a
+vague guess without looking.
+That guess should only be refined after inspecting the image and updating your
+model to account for it.
+Although I do not propose that human brains \emph{do} use probability
+algebra~\cite{jaynes1988brain}, this inference process could, in principle,
+be modelled numerically by distributing probability over the positive integers
+and updating those probabilities in light of a likelihood function for the
+image data.
+Any guess, however, does not depend only on that probability distribution,
+but on the risks and rewards associated with guessing.
 
-\TODO{}
+As written in the caption to Figure~\ref{fig:searches_beans}, there is a
+notional reward to the first exactly answer.
+You are therefore incentivised to answer a) quickly, before other good guesses,
+and b) exactly, with no reward for being close to the true answer.
+If you are the first to answer, then it makes sense to write what you think
+is the most probable number.
+But others following you, seeing your answer, should prefer to guess
+differently if they want to win (and they think any other guess is plausibly
+correct).
+
+To try hard, you could even dissect the vector graphics of the figure and count
+its bean objects, but that may not be optimal use of your resources --- the
+dubious promise of a gold star may be worth little to you.
+
+\TODO{expected loss: risk}
+minimize the expected risk~\cite{
+wald1947bayes,
+wald1950bayes,
+jaynes2003probability
+}
+\begin{equation}
+\label{eqn:searches_bayes_decision_rule}
+R(t) = \int\! c(t, x) \,\mathrm{d}p(x)
+,
+\end{equation}
+for some cost function $c(t, x)$ for decisions $t$ and states of the world
+$x$ that have prior probabilities $p(x)$.
+Although alternative ideas (such as minimax, minimizing a maximum
+loss~\cite{savage1951review}) may empirically work well, this theory finds that
+restricting oneself to Bayes decision rules cannot deny access to optimal
+decisions.
+
 % connection to frequentist section
 Bayes' theorem is an infamous and necessary result, and it is not seriously
-disputed. What has been disputed is whether probability can be applied to
-logic, or any application other than the random happenings of events.
+disputed.
+What has been disputed is whether probability can be applied to logic, or any
+application other than the random happenings of events.
+\TODO{also ignore the results decision theory and attempt to supplant them with
+alternative prescriptions}
 

### 2022-08-24 main/chapter_searches.tex
add more on decision theory, start sketching how to link to next section

@@
 
-\TODO{expected loss: risk}
-minimize the expected risk~\cite{
+In decision theory, we want a mathematical procedure that identifies the
+optimal decision to make in a given context.
+Although this theory has a long history, we point specifically to the result
+by Wald~\cite{
 wald1947bayes,
@@
 }
+that the optimal solution to any decision problem can be framed as the
+minimization of an expected cost function, or risk:
 \begin{equation}
@@
 \end{equation}
-for some cost function $c(t, x)$ for decisions $t$ and states of the world
-$x$ that have prior probabilities $p(x)$.
+where we mean `expected' in the mathematical mean sense of a `mean' average.
+This cost function $c(t, x)$ evaluates the badness of outcomes for a given
+decision $t$ and states of the world $x$ to which context must assign prior
+probabilities $p(x)$.
+For a given cost function and prior, one should choose the decision $t$ that
+minimizes $R(t)$.
+
 Although alternative ideas (such as minimax, minimizing a maximum
-loss~\cite{savage1951review}) may empirically work well, this theory finds that
-restricting oneself to Bayes decision rules cannot deny access to optimal
-decisions.
+loss~\cite{savage1951review}) can empirically work well, this the mathematical
+statement of this theorem states that restricting oneself to Bayes decision
+rules cannot deny access to optimal decisions.
+Alternatives can at best be equivalent.
+
+This theory does not, however, define how to construct that cost function.
+Like the prior, it must come from context.
+Indeed the two are intertwined --- decision making machines and organisms
+are capable of learning to make good decisions without constructing either
+explicitly.
+Towards our goal of usefully reporting data, such decision theory has been
+applied to argue for ways to quote properties of posterior distributions.
+For example, a minimizing the expected absolute error $|x - \check{x}$ leads to
+reporting the posterior median, and minimizing the expected square error
+$(x - \check{x})^2$ leads to reporting the posterior mean.
+Had the rules of our bean counting game rewarded you based on such distances,
+these results could have been good guesses. But those weren't the rules.
+
+\TODO{why?
+not explicitly constructed cost functions for what to report...
+report for contrast with attempted alternative formulations that for social...
+reasons retain popularity despite contradicting decision theory...
+link to frequentist ideas
+Wald worked firmly in that camp - quote
+}
 

### 2022-08-24 main/chapter_searches.tex
some autochecking and tweaks

@@
 labels --- the words `prior' and `likelihood' state our intent to index into
-the left and right arguments respectively.
+the left and right arguments, respectively.
 
@@
 
-This expression integral, or average over the prior, means that evidence $Z$
-for model arises from a larger fraction of samples from it hitting large values
-of the likelihood function.
-In contrast, one might consider a maximum likelihood approach that constrained
-an optimization to regions of nonzero prior in each model.
+This expression as an integral --- an average over the prior, means that
+the evidence $Z$ depends on what fraction of samples hits large values of
+the likelihood function, and not only on how large that likelihood gets.
+In contrast, one could consider a maximum likelihood approach that would
+constrain a numerical optimization to regions of nonzero prior in each model.
 Such optimizations can give a useful upper bound on $Z$, but this example
-highlights a flaw since both models support the $\met > 100\,\eV[G]$ bin
-and both support the same maximum likelihood of
-$L(\varepsilon + 100\,\eV[G]) = 1$.
+highlights a flaw when both models support the $\met > 100\,\eV[G]$ bin.
+Then, both support the same maximum likelihood of
+$L(\varepsilon + 100\,\eV[G]) = 1$, despite different evidence values.
 
@@
 But to report scientific data, such practices are fairly criticized as
-arbitrary when they influce the reported results.
+arbitrary when they influence the reported results.
 Prior assignments are arbitrary.
@@
 
-In the words of Skilling~\cite{skilling2008rant},
+Phrased succinctly by Skilling~\cite{skilling2008rant},
 ``we SHOULD use Bayes by assigning subjective priors and inspecting evidence
 values.''
-This framing of probabilistic data analysis agrees exactly with out standard
+This framing of probabilistic data analysis agrees exactly with our standard
 practice of histogramming simulations,
@@
 are the evidence values that we use to compare the models.
-Finally corresponding posteriors are the normalized shapes those samples take
-within the observed histogram bin, which we do not use to compare the models,
+Finally, corresponding posteriors are the normalized shapes those samples take
+within the observed histogram bin, with which we do not compare the models,
 but which could in principle be used for secondary interpretations.
@@
 In general, such probability assignments are free and unconstrained, and might
-informed by calibration data or loosely motivated opinions.
+be informed by calibration data or loosely motivated opinions.
 But there are examples of unambiguously derived that are commonly used in
@@
 those per second~\cite{atlas2020trigger}.
-This means that we have scenario with a large fixed number $N$ of trials
+This means that we have a scenario with a large fixed number $N$ of trials
 (bunch collisions)
@@
 which is a product of the probabilities (produce rule) in the sequence of $N$
-accept/reject decisions, summed (sum rule) over all of that sequence
-permutations that could have led to the total $n$.
-With our extremely small trigger rate, we are interested in the limit of small
-$p$ and large $N$ which result in a mean acceptance number of $\lambda = Np$.
+accept/reject decisions, summed (sum rule) over all permutations of that sequence
+that could have led to the total $n$.
+With our tiny trigger rate, we are interested in the limit of small
+$p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
 Using the exponential identity of Equation~\ref{eqn:searches_exponential}, one
@@
 This maximum entropy statement means that for a distribution with known mean
-and variance in a given coordinates, its best Gaussian approximation matches
+and variance in given coordinates, its best Gaussian approximation matches
 those moments.\footnote{%
@@
 their likelihood functions.
-At a maximum $\check{x}$, the first derivative of a function vanishes and its
-second derivative is negative, after Taylor expanding about its maximum this
-product looks like
+At a maximum $\check{x}$, the first derivative of a function vanishes and the
+second derivative is negative.
+Taylor expanding about this maximum to the leading order, the product looks like
 \begin{equation}
@@
 notional reward to the first exactly answer.
-You are therefore incentivised to answer a) quickly, before other good guesses,
+You are therefore incentivized to answer a) quickly, before other good guesses,
 and b) exactly, with no reward for being close to the true answer.
@@
 Although alternative ideas (such as minimax, minimizing a maximum
-loss~\cite{savage1951review}) can empirically work well, this the mathematical
+loss~\cite{savage1951review}) can empirically work well, the mathematical
 statement of this theorem states that restricting oneself to Bayes decision
@@
 Like the prior, it must come from context.
-Indeed the two are intertwined --- decision making machines and organisms
+Indeed, the two are intertwined --- decision-making machines and organisms
 are capable of learning to make good decisions without constructing either

### 2022-08-24 main/chapter_searches.tex
finish up decision theory and link to frequentist sec; some tweaks

@@
 those moments.\footnote{%
-By `best', this means minimal Kullback-Leibler divergence
+By `best', this statement means a minimal Kullback-Leibler divergence
 \(
@@
 predictions.
-Unlike ideal posterior updating, this does not exactly employ all information
-in the data, only a function of it --- one usually cannot afford to consume all
-aspects of data, so although theoretically suboptimal this approximation is
-vastly superior in practice.
+Unlike ideal posterior updating, this encodes only a function of the data, not
+its whole --- one usually cannot afford to consume all aspects of data, so
+although theoretically suboptimal this approximation is vastly superior in
+practice. Learning only part of the data is not wrong, just less precise
+than might be possible with a divinely perfect learning algorithm.
 \item Go to 1.
@@
 
-\TODO{why?
-not explicitly constructed cost functions for what to report...
-report for contrast with attempted alternative formulations that for social...
-reasons retain popularity despite contradicting decision theory...
-link to frequentist ideas
-Wald worked firmly in that camp - quote
-}
+When arguing that we should report likelihoods and evidence values for our
+data, we have not explicitly constructed cost functions, nor have we made
+quantitative predictions of future events.
+But we can qualitatively consider them in this decision theory framework ---
+we imagine that future scientists will want to reuse our data, and would be
+pleased (attain a small cost) if they were able to reinterpret the results for
+their own models.
+We can therefore reason that likelihoods would have more utility than
+alternatives such as (point) posterior estimates.
 
 % connection to frequentist section
-Bayes' theorem is an infamous and necessary result, and it is not seriously
-disputed.
-What has been disputed is whether probability can be applied to logic, or any
-application other than the random happenings of events.
-\TODO{also ignore the results decision theory and attempt to supplant them with
-alternative prescriptions}
+An amusing feature of Wald's theory is that it was formulated from a
+perspective on probability in which,
+``In most applications, however, not even the existence of an a priori
+distribution can be postulated''~\cite{wald1950bayes}.
+That might seem problematic, since the risk is the expectation value of a cost
+function over a prior (or \emph{a priori}) distribution.
+Wald was nonetheless interested in these decision rules for their
+``completeness'' that he had proven, and named them ``Bayes'' decision rules
+despite approaching the theory from a stalwartly non-Bayesian perspective.
+
+Although Bayes' rule is not seriously disputed, the system in which Wald was
+working defined probability as a frequency in random sampling, which cannot
+reasonably apply to subjective, non-random distributions of probability mass
+that many prior distributions are.
+This frequency-restricted, `frequentist', philosophy dominated 20th century
+statistical research, and influences practices that \atlas\ continues to
+employ.
+We therefore review some features of frequentist statistics in the following
+section.
 

### 2022-08-25 main/chapter_searches.tex
add some notes and references for frequentist section
add note on intuition, correct early mention of Bayes decision rules

@@
 minimizes $R(t)$.
+More important than this Wald's mathematical proof is that this coincides with
+intuition --- when considering decisions, it feels normal to imagine what might
+occur in each case and consider how good or bad the plausble outcomes are.
 
 Although alternative ideas (such as minimax, minimizing a maximum
-loss~\cite{savage1951review}) can empirically work well, the mathematical
-statement of this theorem states that restricting oneself to Bayes decision
-rules cannot deny access to optimal decisions.
-Alternatives can at best be equivalent.
+loss~\cite{savage1951review}) can work well in certain contexts, this theorem
+states that restricting oneself to minimizing an expected cost function cannot
+deny one access to optimal decisions.
+Alternative formulations can at best be equivalent.
 
@@
 
+\TODO{p-value}
 Figure~\ref{fig:searches_significant}
+\TODO{confidence interval}
+\TODO{maximum likelihood}
+\TODO{unbiased}
+
+\TODO{likelihood principle ~\cite{
+birnbaum1962foundations,
+savage1962foundations
+}
+}
 

### 2022-08-25 main/chapter_searches.tex
begin adding notes and reference urls for frequentist section

@@
 
+\TODO{
+Fisher:
+criteria for estimators,
+(consistency, efficiency, sufficiency)
+https://royalsocietypublishing.org/doi/pdf/10.1098/rsta.1922.0009
+testing, p-value,
+likelihood, also ratios of p-values (here named probable errors)
+https://digital.library.adelaide.edu.au/dspace/handle/2440/15169
+tail area = significance
+http://psychclassics.yorku.ca/Fisher/Methods/
+}
+
+\TODO{
+Fisher:
+``We can know nothing of the probability of hypotheses or of hypothetical
+quantities.''
+}
+
+\TODO{
+restrict our notion of a probability with a special case: only frequency
+of random events
+}
+
+\TODO{
+tossing a coin is considered random
+but if instead you are presented with a coin hidden under a cloth,
+no frequency interpretation --- it either is or is not heads
+you are therefore denied the ability to make uncertain predictions.
+more generally, you are not allowed to express an uncertainty in something
+unless some random process is involved
+}
+
+\TODO{
+Can keep likelihoods, but cannot compute evidence values unless the prior
+has a frequency interpretation.
+}
+
+\TODO{
+Its void is filled with various proposed alternatives that do have frequency
+interpretations and that sometimes yield useful results.
+}
+
+\TODO{
+maximum likelihood, p-values, confidence intervals, and biasedness of
+estimators.
+}
+
+\TODO{maximum likelihood}
 \TODO{p-value}
+\TODO{p-value:
+The chance of a small $p$-value is small.
+This is true, but not interesting because the value of any $p$-value is small.
+By construction $p$-values are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+$p$-value to have meaning.
+}
 Figure~\ref{fig:searches_significant}
 \TODO{confidence interval}
-\TODO{maximum likelihood}
 \TODO{unbiased}

### 2022-08-25 main/chapter_searches.tex
dump a bunch of notes from reading

@@
 likelihood, also ratios of p-values (here named probable errors)
+(introduced the term likelihood? cite?)
 https://digital.library.adelaide.edu.au/dspace/handle/2440/15169
@@
 quantities.''
+Fisher:
+``absurdly academic''
+Fisher:
+``There is something horrifying in the ideological movement represented by the doctrine
+that reasoning, properly speaking, cannot be applied to empirical data to lead to inferences
+valid in the real world.''
+}
+
+\TODO{
+Nayman-Pearson:
+Seek good test statistics --
+likelihood ratio intuitive for two-hypothesis comparison
+prove it optimal in a reasonable sense
+idea of long-run mostly not wrong
+recede to maximum likelihood ratios
+https://www.jstor.org/stable/2331945
+``for which hypothesis A is more probable and those for which it is less
+probable (sic!)''
+unbiased (unbiassed) test differs from unbiasedness of estimator
+``Then the two corresponding densities at the sample point
+are (Al) 6-42 x 10-20, (A2) 5 18*x 1O-4 and the ratio of the likelihood of Hypothesis
+Al to- that of A2 is 1P24 x 10-16. This ratio confirms the common-sense judgment
+that A2 is far more plausible than A1''
+^ Another posterior argument --- this likelihood ratio in in fact a change
+in plausibilities. For example, if A2 is an inconsistent, self-contradictory
+theory, it is totally implausible, and not more plausible despite predicting
+these data better.
+``The statistician can
+balance the numerical verdict of likelihood against the vaguer expectations de-
+rived from 'a priori considerations, and it must be left to his judgment to
+at what point the evidence in favour of alternative hypotheses becomes so con-
+vincing that Hypothesis A must be rejected.''
+Introduce maximum likelihood ratio as interesting without claiming it to be
+necessarily best.
+https://royalsocietypublishing.org/doi/10.1098/rsta.1933.0009
+Neyman-Pearson lemma
+https://www.cambridge.org/core/journals/canadian-mathematical-bulletin/article/a-selection-of-early-statistical-papers-of-j-neyman-university-of-california-press-1967-ix-429-pages-1495/75CE96368E6F34BD25D06BAF6DC8E69F
+}
+
+\TODO{
+discrete failure to match --- randomization
+K. D. Tocher (1950)
+``extraneous randomization''
+}
+
+\TODO{
+Given all this understanding, what I do not understand is their choice to
+convert this to a $p$-value.
+Seems to be this idea of long-run correctness guarantees.
+But their derivation does rely on long-run frequencies.
 }

### 2022-08-27 main/chapter_searches.tex
add some references for stats section; tweak some data analysis; improve chapter name

--- 2022-08-25 main/chapter_searches.tex
@@
 \begin{singlespacing}
-\chapter{Searches in ATLAS}
+\chapter{Data analysis and \atlas\ searches}
 \label{chapter:searches}
-%
 \begin{epigraphs}
@@
 When it is not, perhaps due to a complicated parametric modelling, then one
-should abstract those parameters away by reporting evidence values for
-predictive models, which must of course be clearly stated and interpretable if
-they are to be useful.
+should abstract those parameters away to summarize the likelihood.
+Properties such as local maxima or regions in which it exceeds meaningful
+thresholds can be useful.
+Evidence values for predictive models can also be useful if those models are
+clearly stated and relevant to readers' interests.
 Data act only through the likelihood function, and reporting a posterior
@@
 inferences about the world.
-Reporting likelihoods and evidence values allows readers to assign priors as
-they wish.
-For model comparison between background $h=B$ and signal $h=S$, this is clearly
-expressed in the ratio form
-% likelihood ratio
+Reporting likelihood functions and evidence values allows readers to assign
+priors as they wish.
+For model comparison, for example between background $h=B$ and signal $h=S$
+models, the analysis is conveniently linearized by expressing it in $\log$
+ratio form:
 \begin{equation}
 \hphantom{\quad \quad \mid\mid \mathrm{context}}
-\frac{\prob{S}{d}}{\prob{B}{d}}
+\phi(d) =
+\log\frac{\prob{S}{d}}{\prob{B}{d}}
 =
-\frac{\prob{d}{S}}{\prob{d}{B}}
-\times
-\frac{\probc{S}}{\probc{B}}
+\log\frac{\prob{d}{S}}{\prob{d}{B}}
++
+\log\frac{\probc{S}}{\probc{B}}
 \quad \quad \mid\mid \mathrm{context}
@@
 a reader might wish to include.
-Whenever the evidence ratio is independent of that context, it factors out from
-the prior (odds) ratio.
+When the evidence ratio is independent of that context, as it usually is,
+it factors out from the prior (odds) ratio.
 In this sense, likelihood (or evidence) ratios are scale factors that update
-any odds ratio by the same multiplication.
+any odds prior ratio by the same multiplication, and that on this $\log$ scale
+that means just adding a constant.
+
+From such $\log$ ratios one individual probabilities are recovered through the
+logistic function:
+\begin{equation}
+\label{eqn:searches_logistic}
+\prob{S}{d} = \frac{1}{1 + e^{-\phi(d)}}
+.
+\end{equation}
+Physically, this is analogous to the Boltzmann distribution where the energy
+of each model is its negative $\log$ probability, with unit inverse
+temperature~\cite{
+pmlr-v2-ranzato07a,
+skilling2017david
+}.
+With multiple models in play, it is also called the `softmax'
+function~\cite{MurphyKevinP.2012Mlap}.
 
@@
 practice of histogramming simulations,
-as illustrated in Figure~\ref{fig:searches_sig_bkg_prior_likelihood}.
-Signal and background models define priors, and their (normalized) bin yields
+as illustrated in Figure~\ref{fig:searches_sig_bkg_prior_likelihood}:
+signal and background models define priors, and their (normalized) bin yields
 are the evidence values that we use to compare the models.
-Finally, corresponding posteriors are the normalized shapes those samples take
-within the observed histogram bin, with which we do not compare the models,
-but which could in principle be used for secondary interpretations.
+Finally, their corresponding posteriors are the normalized shapes those samples
+take within the observed histogram bin, with which we do not compare the
+models, but which could in principle be used for secondary interpretations.
 
@@
 
-In decision theory, we want a mathematical procedure that identifies the
-optimal decision to make in a given context.
-Although this theory has a long history, we point specifically to the result
-by Wald~\cite{
+When considering what decision to make, it feels natural to consider what
+might result from each possible action and to judge that action on how good or
+bad its results would be.
+Mathematical decision theory agrees with this intuition---
+a famous theorem by Wald~\cite{
 wald1947bayes,
@@
 }
-that the optimal solution to any decision problem can be framed as the
-minimization of an expected cost function, or risk:
+shows that the optimal solution to any decision problem can be always be
+formulated as the minimization of an expected cost function, or risk:
 \begin{equation}
@@
 \end{equation}
-where we mean `expected' in the mathematical mean sense of a `mean' average.
+where we mean `expected' in the mathematical mean sense of the `mean' average.
 This cost function $c(t, x)$ evaluates the badness of outcomes for a given
@@
 minimizes $R(t)$.
-More important than this Wald's mathematical proof is that this coincides with
-intuition --- when considering decisions, it feels normal to imagine what might
-occur in each case and consider how good or bad the plausble outcomes are.
 
@@
 This theory does not, however, define how to construct that cost function.
-Like the prior, it must come from context.
-Indeed, the two are intertwined --- decision-making machines and organisms
-are capable of learning to make good decisions without constructing either
-explicitly.
+Like the prior, it must come from context, and indeed the two are intertwined
+in the risk;
+decision-making machines and organisms are capable of learning to make good
+decisions without constructing either cost or prior explicitly.
 Towards our goal of usefully reporting data, such decision theory has been
 applied to argue for ways to quote properties of posterior distributions.
-For example, a minimizing the expected absolute error $|x - \check{x}$ leads to
-reporting the posterior median, and minimizing the expected square error
-$(x - \check{x})^2$ leads to reporting the posterior mean.
+As examples, minimizing the expected absolute error $|x - \check{x}|$ leads
+to reporting the median, and minimizing the expected square error
+$(x - \check{x})^2$ leads to reporting the mean.
 Had the rules of our bean counting game rewarded you based on such distances,
@@
 their own models.
-We can therefore reason that likelihoods would have more utility than
-alternatives such as (point) posterior estimates.
+We can therefore reason that likelihoods would usually have more utility than
+alternatives such as point posterior averages.
 
@@
 ``completeness'' that he had proven, and named them ``Bayes'' decision rules
-despite approaching the theory from a stalwartly non-Bayesian perspective.
+despite approaching the theory from a stalwart non-Bayesian perspective.
 
@@
 \TODO{
-Fisher:
-criteria for estimators,
-(consistency, efficiency, sufficiency)
-https://royalsocietypublishing.org/doi/pdf/10.1098/rsta.1922.0009
-testing, p-value,
-likelihood, also ratios of p-values (here named probable errors)
-https://digital.library.adelaide.edu.au/dspace/handle/2440/15169
-tail area = significance
-http://psychclassics.yorku.ca/Fisher/Methods/
+Lehmann~\cite{lehmann2005testing};
+p. 26 poisson inadmissable example;
+L-unbiased with given loss function;
 }
@@
 Fisher:
-``We can know nothing of the probability of hypotheses or of hypothetical
-quantities.''
+criteria for estimators,
+(consistency, efficiency, sufficiency)
+~\cite{fisher1922estimators}
+testing, p-value,
+likelihood, also ratios of p-values (here named probable errors)
+(introduced the term likelihood? cite?)
+~\cite{fisher1921probable}
+tail area = significance
+~\cite{fisher1925smrw}
+}
+
+\TODO{
+Fisher:
+``We can know nothing of the probability of hypotheses or hypothetical
+quantities.
+On the other hand we may ascertain the likelihood of hypotheses
+and hypothetical quantities by calculation from observations''
+~\cite{fisher1921probable}
+Fisher:
+``absurdly academic''
+~\cite{fisher1956statistical}
+% \begin{epigraphs}
+% \qitem{%
+% We can know nothing of the probability of hypotheses or hypothetical
+% quantities.
+% On the other hand we may ascertain the likelihood of hypotheses and
+% hypothetical quantities by calculation from observations%
+% }%
+% {Ronald A. Fisher, 1921~\cite{fisher1921probable}}
+% \end{epigraphs}
+There is something horrifying in the ideological movement represented by the
+doctrine that reasoning, properly speaking, cannot be applied to empirical data
+to lead to inferences valid in the real world.
+~\cite{fisher1956statistical}
+}
+
+\begin{singlespacing}
+\begin{displayquote}
+We can know nothing of the probability of hypotheses or hypothetical
+quantities.
+On the other hand we may ascertain the likelihood of hypotheses
+and hypothetical quantities by calculation from observations
+\end{displayquote}
+\end{singlespacing}
+~\cite{fisher1921probable}
+
+\begin{singlespacing}
+\begin{displayquote}
+There is something horrifying in the ideological movement represented by the
+doctrine that reasoning, properly speaking, cannot be applied to empirical data
+to lead to inferences valid in the real world.
+\end{displayquote}
+\end{singlespacing}
+~\cite{fisher1956statistical}
+
+\TODO{
+Nayman-Pearson:
+Seek good test statistics --
+likelihood ratio intuitive for two-hypothesis comparison
+prove it optimal in a reasonable sense
+idea of long-run mostly not wrong
+recede to maximum likelihood ratios
+~\cite{neymanpearson1928max}
+``for which hypothesis A is more probable and those for which it is less
+probable (sic!)''
+unbiased (unbiassed) test differs from unbiasedness of estimator
+``Then the two corresponding densities at the sample point
+are (Al) 6-42 x 10-20, (A2) 5 18*x 1O-4 and the ratio of the likelihood of Hypothesis
+Al to- that of A2 is 1P24 x 10-16. This ratio confirms the common-sense judgment
+that A2 is far more plausible than A1''
+\^ Another posterior argument --- this likelihood ratio in in fact a change
+in plausibilities. For example, if A2 is an inconsistent, self-contradictory
+theory, it is totally implausible, and not more plausible despite predicting
+these data better.
+``The statistician can
+balance the numerical verdict of likelihood against the vaguer expectations de-
+rived from `a priori' considerations, and it must be left to his judgment to
+at what point the evidence in favour of alternative hypotheses becomes so con-
+vincing that Hypothesis A must be rejected.''
+Introduce maximum likelihood ratio as interesting without claiming it to be
+necessarily best.
+Neyman-Pearson lemma --- reproduce Fig 5?
+~\cite{neymanpearson19233lemma}
+}
+
+\TODO{
+discrete failure to match --- randomization.
+table of random numbers --- reputable table!
+K. D. Tocher (1950)
+``extraneous randomization''
+~\cite{tocher1950discontinuous}
+}
+
+% TODO Kolmogorov axioms, unbiasedness
+\TODO{
+Given all this understanding, what I do not understand is their choice to
+convert this to a $p$-value.
+Seems to be this idea of long-run correctness guarantees.
+But their derivation does rely on long-run frequencies.
 }

### 2022-08-27 main/draft.tex
add gaussian stuff
add some references for stats section; tweak some data analysis; improve chapter name

@@
 \TODO{\ldots}
+
+This is summarized most succinctly by Jaynes:
+``any function with a single rounded maximum, raised to a higher and higher
+power, goes into a Gaussian function''~\cite{jaynes2003probability}.
+
+More important than this Wald's mathematical proof is that this coincides with
+intuition --- when considering decisions, it feels normal to imagine what might
+occur in each case and consider how good or bad the plausble outcomes are.
+
+SMRW scan url http://psychclassics.yorku.ca/Fisher/Methods/

### 2022-08-27 main/chapter_searches.tex
paragraph to subsection

@@
 
-\paragraph{Probability, abstract.}
+\subsection{Probability, abstract}
 A probability $\prob{b}{a}$ is a number that quantifies the proportion
@@
 
-\paragraph{Probability, applied.}
+\subsection{Probability, applied}
 % science snake
@@
 
-\paragraph{Probability assignments.}
+\subsection{Probability assignments}
 Probability algebra manipulates probabilities, but does not state how they
@@
 
-\paragraph{Decision theory.}
+\subsection{Decision theory}
 Now that we have thoroughly covered probability theory, we are ready to return

### 2022-09-01 main/chapter_introduction.tex
add least squares note; add labels note; fix pa pb independent

@@
 data and not from additional labels~\cite{
+Noroozi2016jigsaw,
 multitaskself2017,

### 2022-09-01 main/chapter_searches.tex
add least squares note; add labels note; fix pa pb independent

@@
 \)
-if $\probc{a}$ and $\probc{b}$ are independent.
-Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, the
+if $a$ and $b$ are independent.
+Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, this
 propagation formula works only for small relative errors.
@@
 $x$, and we include the $\mathrm{d}x$ terms as reminders that it is a density
-function in that coordinate.
+function in $x$, not a probability exactly.
+On a $\log$ scale, the Gaussian form is simply quadratic in $\mu$ (or $x$);
+that quadratic form continues in its multivariate generalization, and
+justifies least-squares methods.
 
@@
 (Practically, keep its equivalent $\log$ mean over independent samples.)
+These $d_i$ may be only reductions of the data, such as class labels
+or index permutations~\cite{Noroozi2016jigsaw, multitaskself2017}.
+Other inputs that are not predicted, such as associated images that the machine
+uses to predict the labels, are included in the context $h$.
 \item Use those data to update your state
-$f(d_{i-1}, \ldots) \rightarrow f(d_i, d_{i-1}, \ldots)$ to improve your
-predictions.
+$f(d_{i-1}, \ldots) \rightarrow f(d_i, d_{i-1}, \ldots)$.
 Unlike ideal posterior updating, this encodes only a function of the data, not
@@
 although theoretically suboptimal this approximation is vastly superior in
-practice. Learning only part of the data is not wrong, just less precise
-than might be possible with a divinely perfect learning algorithm.
+practice.
+Learning only part of the data is not wrong, it is just less precise than a
+divinely perfect learning algorithm, which we do not currently possess.
 \item Go to 1.

### 2022-09-01 main/chapter_searches.tex
add likelihood prior transformation figure

@@
 \subsection{Probability, applied}
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_abacus_likelihood.pdf}
+\caption{%
+Likelihood: $\mathrm{Poisson}(0\mid x)$ in
+(left) $x$ and
+(right) $u = \log x$%
+}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_abacus_prior.pdf}
+\caption{%
+Prior density: exponential in
+(left) $x$ and
+(right) $u = \log x$%
+}
+\end{subfigure}
+\caption[
+Likelihoods and prior densities' different transformations
+]{%
+Likelihoods and prior densities' different transformations with coordinates.
+The $\mathrm{Poisson}(0\mid x)$ likelihood and exponential $e^{-x}\mathrm{d}x$
+prior have the same form, but the prior density changes to preserve its unit
+area when expressed relative to $u = \log x$.
+\\[.5em]
+Relative to $x$, both are identical (left).
+On changing to $\log$ coordinates (right), the likelihood slides horizontally,
+but the prior density changes value to preserve area.
+Both maximum likelihood points are equivalent (maximum likelihoods are robust),
+but the maximum prior density moves (maximum densities are slippery).
+\\[.5em]
+Particle lifetimes follow exponential distributions, and their maximum
+densities are at zero time.
+It is misleading, however, to describe their most probable lifetimes as zero
+---
+the `most probable' $\log$ lifetime is not $\log 0$, but equals the mean
+lifetime.
+\\[.5em]
+For zero Poisson data, the maximum likelihood rate is zero, but zero might not
+be the most likely rate (if, for example, an experiment has known backgrounds).
+}
+\label{fig:searches_abacus}
+\end{figure}
 % science snake
@@
 
+Since prior probabilities are normalized distributions, it is usually
+convenient to represent them as density functions.
+Densities are stated relative to coordinates (or occasionally measures),
+so changing coordinates changes their values as their unit probability mass
+is sloshed around.
+Likelihoods are not densities in the conditioning variable that they explore,
+so changing coordinates slides them horizontally like beads on an abacus,
+but does not change their values for equivalent states.
+Figure~\ref{fig:searches_abacus} illustrates this difference.
+
 Likelihood and prior are evidently related --- they explore two indices into
@@
 practical methods.
-For observed data, we use likelihoods to compare which hypotheses predicted
-them better than others, and before observing data we use priors to describe
-what data might result from different models to design our experiments.
+We use likelihoods to compare which hypotheses predicted observed data better
+than others.
+Prior to observing data we use priors to describe what might result from
+different models; priors on data can be useful in designing our experiments,
+and priors on other states describe other predictions.
 

### 2022-09-01 main/chapter_searches.tex
add figures for frequentist section --- likelihoods and tail areas

@@
 In contrast, one could consider a maximum likelihood approach that would
-constrain a numerical optimization to regions of nonzero prior in each model.
-Such optimizations can give a useful upper bound on $Z$, but this example
-highlights a flaw when both models support the $\met > 100\,\eV[G]$ bin.
+constrain a numerical optimization to regions of non-zero prior in each model.
+Such optimizations can give useful upper bounds on the evidence, but this
+example highlights a flaw when both models support the $\met > 100\,\eV[G]$
+bin.
 Then, both support the same maximum likelihood of
@@
 
-\TODO{
-Lehmann~\cite{lehmann2005testing};
-p. 26 poisson inadmissable example;
-L-unbiased with given loss function;
-}
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
+\caption[
+Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
+]{%
+Ratio thresholds of a Binomial likelihood function, reproduced from
+Fisher~\cite{fisher1956statistical}.
+Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum, as a practical way
+to communicate its profile.
+}
+\label{fig:searches_fisher_likelihood}
+\end{figure}
 
@@
 }
+Figure~\ref{fig:searches_fisher_likelihood}
 
@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/searches_np_curves_crop.png}
+\caption[
+Tail areas of prior data distributions from two alternative hypotheses,
+reproduced from Neyman and Pearson
+]{%
+Tail areas of prior data distributions from two alternative hypotheses,
+reproduced from Neyman and Pearson~\cite{neymanpearson19233lemma}.
+}
+\label{fig:searches_np_tails}
+\end{figure}
+
+Figure~\ref{fig:searches_np_tails}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/searches_atlas_higgs_4l_2207_00320.pdf}
+\caption[
+Signal and background distributions reproduced from the recent \atlas\
+measurement of Higgs boson decays to $4\ell$
+]{%
+Signal and background distributions reproduced from the recent \atlas\
+measurement of Higgs boson decays to $4\ell$~\cite{ATLAS:2022net}.
+}
+\label{fig:atlas_higgs_4l}
+\end{figure}
+
+Figure~\ref{fig:atlas_higgs_4l}
+
 \TODO{
@@
 estimators.
+}
+
+\TODO{
+Lehmann~\cite{lehmann2005testing};
+p. 26 poisson inadmissable example;
+L-unbiased with given loss function;
 }

### 2022-09-01 main/chapter_searches.tex
add fisher quote on communicating likelihoods

@@
 \begin{displayquote}
+For all purposes, and more particularly for the \textit{communication} of the
+relevant evidence supplied by a body of data, the values of the Mathematical
+Likelihood are better fitted to analyse, summarize, and communicate
+statistical evidence of types too weak to supply true probability statements
+\end{displayquote}
+\end{singlespacing}
+~\cite{fisher1956statistical}
+
+\begin{singlespacing}
+\begin{displayquote}
 There is something horrifying in the ideological movement represented by the

### 2022-09-02 main/chapter_searches.tex
begin frequentist section intro, add some references

@@
 The summed total number $n$ of accepted events is therefore described by the
-Binomial distribution for $n$;
-\begin{equation}
-\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
+binomial distribution for $n$;
+\begin{equation}
+\label{eqn:searches_binomial}
+\mathrm{binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
 ,
@@
 
+Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
+proportion of stuff within a larger whole, which when formalized as a ratio
+of measures and has broad applications.
+Frequentist theory confines itself to the special case where those measures
+are numbers of discrete randomly occuring events, such that a frequentist
+probability is a ratio of frequencies:
+\begin{equation}
+\label{eqn:searches_prob_frequency_ratio}
+\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
+\end{equation}
+where $n$ is the number observed to satisfy conditions $a$ and $b$, and
+$N$ have satisfied $b$.
+From Equation~\ref{eqn:searches_binomial}, this converges to the general
+probability in its infinite limit.
+
+Venn~\cite{venn1866logic},
+Boole~\cite{boole1854investigation},
+Kolmogorov~\cite{
+kolomogoroff1933de,
+kolomogoroff1950translated
+},
+
 \begin{figure}[tp]
@@
 Fisher~\cite{fisher1956statistical}.
-Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum, as a practical way
-to communicate its profile.
+Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum.
+Their extents could be reported in a table as a practical communication of
+the likelihood function's profile.
 }
@@
 
+\TODO{
+on ``likelihood'' models ---
+we are instructed to pretend that these are likelihoods on auxiliary data
+TODO cite.
+One is free to engage in such doublethink, just as one is free to proclaim that
+$2 + 2 = 5$.
+The fact remains that many constraints do not correspond to predictions of
+auxiliary data: counterexamples include
+scale variations,
+theoretical cross-section uncertainties,
+non-closure uncertainties (which cover differences between estimates).
+simulated sample camparisons,
+statistical errors on weighted simulations,
+extrapolation uncertainties,
+and
+all uncertainties in supersymmetric signal yields
+(for which we have yet no data).
+}
+
 \section{Searches}

### 2022-09-02 main/chapter_searches.tex
add some words on frequencies

@@
 Frequentist theory confines itself to the special case where those measures
-are numbers of discrete randomly occuring events, such that a frequentist
+are numbers of discrete randomly occurring events, such that a frequentist
 probability is a ratio of frequencies:
@@
 \end{equation}
-where $n$ is the number observed to satisfy conditions $a$ and $b$, and
+where $n$ is the number observed to satisfy conditions $a$ and $b$ when
 $N$ have satisfied $b$.
-From Equation~\ref{eqn:searches_binomial}, this converges to the general
-probability in its infinite limit.
+Interpreted with the Binomial distribution of
+Equation~\ref{eqn:searches_binomial}, this definition converges to our more
+general probability whenever it applies.
+
+We do think of many processes as random when considering scientific data, so
+this definition works widely for likelihood functions.
+Non-data hypotheses, however, are rarely considered as actually random events,
+so the frequency definition cannot not apply.
+Prior predictions often apply probabilities to express a subjective uncertainty
+predictions that are manifestly non-random.
+Under this frequency definition, those predictive probabilities are nonsense
+and must be rejected.
+The frequency definition does not change probability algebra, it only restricts
+it to a permitted scope.
+
+Coin tosses, for example, can be imagined as random.
+Earth's human population could toss a few billion coins to precisely
+establish a probability of landing heads from the ratio of
+$n$ (number of heads) to $N$ (number of tosses).
+But suppose you are given a red envelope containing a coin ---
+which way up is the coin in the envelope?
+You could guess or bet on the answer, and that decision might involve a tacitly
+assumed probability.
+A frequency, however, would only be permitted if the envelope had been packaged
+randomly.
+Was the coin first tossed to randomize its face, or placed deliberately?
+It is unclear.
+
+Tom Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
+stoppard1967rosencrantz
+}
+contains a clearer example ---
+Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads.
+They play a game with a third player, landing eight more heads.
+Is the next toss heads or tails?
+The answer is not random --- it is burned in print onto paper, so no frequency
+can be assigned.
+Were subjective probabilities permitted, you could predict the result, and that
+prediction would be subjectively informed by it being the one-hundred-and-first
+toss, our cultural entanglement with in base-10 numbers, and your reading of
+these words.
+Frequentist theory forbids probabilities of such non-random unknowns, and is
+left seeking alternatives.
 
@@
 TODO cite.
-One is free to engage in such doublethink, just as one is free to proclaim that
-$2 + 2 = 5$.
-The fact remains that many constraints do not correspond to predictions of
-auxiliary data: counterexamples include
+One is free to engage in such doublethink~\cite{orwell1949nnineteen},
+% just as one is free to proclaim that $2 + 2 = 5$
+but the fact remains that many constraints do not correspond to predictions of
+auxiliary data so are not likelihoods: examples include
 scale variations,
@@
 non-closure uncertainties (which cover differences between estimates).
-simulated sample camparisons,
+simulated sample comparisons,
 statistical errors on weighted simulations,
@@
 and
-all uncertainties in supersymmetric signal yields
+all uncertainties on supersymmetric signal yields
 (for which we have yet no data).
+energy based modelling ---
+I know of no tangible benefit from pretending that something is a likelihood
+when it is not.
 }

### 2022-09-02 main/chapter_searches.tex
add some history of frequentism text

@@
 $N$ have satisfied $b$.
-Interpreted with the Binomial distribution of
+Interpreted with the Binomial model of
 Equation~\ref{eqn:searches_binomial}, this definition converges to our more
@@
 contains a clearer example ---
-Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads.
+Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads
+(by ``un-, sub- or supernatural forces'').
 They play a game with a third player, landing eight more heads.
@@
 prediction would be subjectively informed by it being the one-hundred-and-first
-toss, our cultural entanglement with in base-10 numbers, and your reading of
-these words.
-Frequentist theory forbids probabilities of such non-random unknowns, and is
-left seeking alternatives.
-
-Venn~\cite{venn1866logic},
-Boole~\cite{boole1854investigation},
-Kolmogorov~\cite{
+toss, our cultural entanglement with round numbers in base-10, and your reading
+of these words.
+Frequentist theory, however, forbids such subjective probabilities of
+non-random unknowns, so is left seeking alternatives.
+
+As established in Section~\ref{sec:searches_data_analysis}, the original
+probability theory of Laplace and Bayes, did not require frequency
+interpretations; these were developed later in the 19th century by
+mathematicians including Venn~\cite{venn1866logic} and
+Boole~\cite{boole1854investigation}, in reactions that may well have been
+propelled by reasonable objections to flawed posterior-only methods.
+Modern frequentist theory has a rigorous set-theoretic formulation in the
+Kolmogorov axioms~\cite{
 kolomogoroff1933de,
-kolomogoroff1950translated
-},
+kolomogoroff1950translated,
+axioms1010038
+}.
+Applied frequentist methods, unlike frequentist theory, are constrained with
+considerably less precision --- the denial of certain (non-frequency)
+probabilities does not construct their replacements, and various surrogates
+have evolved.
+Modern frequentist methods are largely attributable to the authors
+R. A. Fisher (for likelihood, $p$-values, and estimation criteria), and
+J. Neyman and E. S. Pearson
+(for statistical hypothesis testing and confidence intervals).
+Together these do not form a harmonious or consistent framework ---
+the two groups have emphatic disagreements in print ---
+but do cover many practices that continue to survive in modern science
+and explain important features of the methods apply to LHC data, which will be
+introduced in Section~\ref{sec:searches_practice}.
+
+\TODO{Fisher's great contribution --- likelihood}
+\TODO{
+Fisher:
+criteria for estimators,
+(consistency, efficiency, sufficiency)
+~\cite{fisher1922estimators}
+testing, p-value,
+likelihood, also ratios of p-values (here named probable errors)
+(introduced the term likelihood? cite?)
+~\cite{fisher1921probable}
+tail area = significance
+~\cite{fisher1925smrw}
+}
+Figure~\ref{fig:searches_fisher_likelihood}
 
@@
 \end{figure}
-
-\TODO{
-Fisher:
-criteria for estimators,
-(consistency, efficiency, sufficiency)
-~\cite{fisher1922estimators}
-testing, p-value,
-likelihood, also ratios of p-values (here named probable errors)
-(introduced the term likelihood? cite?)
-~\cite{fisher1921probable}
-tail area = significance
-~\cite{fisher1925smrw}
-}
-Figure~\ref{fig:searches_fisher_likelihood}
 

### 2022-09-02 main/chapter_searches.tex
some proofreading

@@
 $n$ (number of heads) to $N$ (number of tosses).
-But suppose you are given a red envelope containing a coin ---
+Now suppose you are given a red envelope containing a coin;
 which way up is the coin in the envelope?
@@
 }
-contains a clearer example ---
-Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads
+opens with a clearer example ---
+the titular characters toss a coin ninety-two times, all landing heads
 (by ``un-, sub- or supernatural forces'').
-They play a game with a third player, landing eight more heads.
-Is the next toss heads or tails?
+They then play a game with a third player, landing eight more heads.
+Will the next toss be heads or tails?
 The answer is not random --- it is burned in print onto paper, so no frequency
@@
 As established in Section~\ref{sec:searches_data_analysis}, the original
-probability theory of Laplace and Bayes, did not require frequency
+probability theory of Laplace and Bayes did not require frequency
 interpretations; these were developed later in the 19th century by
@@
 Boole~\cite{boole1854investigation}, in reactions that may well have been
-propelled by reasonable objections to flawed posterior-only methods.
+propelled by reactions to flawed posterior-only methods.
 Modern frequentist theory has a rigorous set-theoretic formulation in the
@@
 axioms1010038
-}.
-Applied frequentist methods, unlike frequentist theory, are constrained with
-considerably less precision --- the denial of certain (non-frequency)
-probabilities does not construct their replacements, and various surrogates
-have evolved.
+}, but applied frequentist methods, are constrained with considerably less
+precision --- the denial of certain (non-frequency) probabilities does not
+construct their replacements, and various surrogates have evolved.
+
 Modern frequentist methods are largely attributable to the authors
@@
 (for statistical hypothesis testing and confidence intervals).
-Together these do not form a harmonious or consistent framework ---
-the two groups have emphatic disagreements in print ---
-but do cover many practices that continue to survive in modern science
-and explain important features of the methods apply to LHC data, which will be
-introduced in Section~\ref{sec:searches_practice}.
+Together these do not form a harmonious framework;
+the two groups have emphatic disagreements in print, but together they
+do cover many practices that survive in modern science and explain important
+features of the methods we apply to LHC data, which will be introduced in
+Section~\ref{sec:searches_practice}.
 

### 2022-09-03 main/chapter_searches.tex
some frequentist text

@@
 Was the coin first tossed to randomize its face, or placed deliberately?
-It is unclear.
+You do not know.
 
@@
 opens with a clearer example ---
-the titular characters toss a coin ninety-two times, all landing heads
-(by ``un-, sub- or supernatural forces'').
+the titular characters toss a coin ninety-two times, all landing heads.
 They then play a game with a third player, landing eight more heads.
 Will the next toss be heads or tails?
-The answer is not random --- it is burned in print onto paper, so no frequency
+The answer is not random --- it is burned onto paper in print, so no frequency
 can be assigned.
@@
 Frequentist theory, however, forbids such subjective probabilities of
-non-random unknowns, so is left seeking alternatives.
+non-random unknowns, so is left to search for alternative modes of inference.
 
@@
 
-Modern frequentist methods are largely attributable to the authors
-R. A. Fisher (for likelihood, $p$-values, and estimation criteria), and
-J. Neyman and E. S. Pearson
-(for statistical hypothesis testing and confidence intervals).
-Together these do not form a harmonious framework;
-the two groups have emphatic disagreements in print, but together they
-do cover many practices that survive in modern science and explain important
-features of the methods we apply to LHC data, which will be introduced in
-Section~\ref{sec:searches_practice}.
+Modern frequentist methods are largely attributable to work by
+Ronald A. Fisher (for likelihood, $p$-values, and estimation criteria~\cite{
+fisher1921probable,
+fisher1922estimators,
+fisher1925smrw,
+fisher1956statistical
+}),
+and Jerzy Neyman and Egon S. Pearson
+(for statistical hypothesis testing and confidence intervals~\cite{
+neymanpearson1933lemma,
+neymanpearson1928max
+})
+in the 1920s and 30s.
+Together, the Fisher and Neyman-Pearson methods do not form a harmonious
+framework, but comprise a loose collection of ideas and practices that survive
+in modern science, and explain important features of the methods we apply to
+LHC data, which Section~\ref{sec:searches_practice} develops further.
+
+Some literature names this Fisher/Neyman-Pearson assemblage
+``classical'' statistics~\cite{lehmann2011fisher, Feldman:1997qc}, but that is
+simply inaccurate for a 20th century divergence from 18th century
+ideas~\cite{bayes1763lii, laplace1774stigler}.
+Jaynes instead names them ``orthodox'~\cite{jaynes2003probability},
+which may be accurate in some contexts, but alternative
+(heretical?) ideas have taken hold under modern LHC authorities ---
+the orthodoxy has moved on.
+We therefore stick to the name `frequentist', as it is a firm common factor.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
+\caption[
+Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
+]{%
+Ratio thresholds of a Binomial likelihood function, reproduced from
+Fisher~\cite{fisher1956statistical}.
+Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum.
+Their extents could be reported in a table as a practical communication of
+the likelihood function's profile.
+}
+\label{fig:searches_fisher_likelihood}
+\end{figure}
+
+% likelihood
+Fisher coined the term `likelihood'~\cite{fisher1921probable}, in the usage
+we have employed, for the probability of observed data given varied
+contexts\footnote{%
+Fisher used the term `likelihood' for the function exploring
+contexts and `probability' for the distribution over
+outcomes~\cite{fisher1925smrw}.
+We choose the alternative word `prior' for his `probability' since the
+likelihood is also a probability, just not a probability \emph{distribution}.
+},
+and deserves credit for popularising its use in interpreting and communicating
+data.
+% figure here TODO
+% likelihood quotes
+Fisher forcefully argued against the use of Bayes' rule and prior
+probabilities, stating for example that
+\begin{singlespacing}
+\begin{displayquote}
+We can know nothing of the probability of hypotheses or hypothetical
+quantities.
+On the other hand we may ascertain the likelihood of hypotheses
+and hypothetical quantities by calculation from observations~\cite{
+fisher1921probable
+}
+\end{displayquote}
+\end{singlespacing}
+
+% p-values
+
+% estimators --- not unbiased!
+
+% savage likelihood principle
 
@@
 Figure~\ref{fig:searches_fisher_likelihood}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
-\caption[
-Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
-]{%
-Ratio thresholds of a Binomial likelihood function, reproduced from
-Fisher~\cite{fisher1956statistical}.
-Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum.
-Their extents could be reported in a table as a practical communication of
-the likelihood function's profile.
-}
-\label{fig:searches_fisher_likelihood}
-\end{figure}
 
@@
 Neyman-Pearson lemma --- reproduce Fig 5?
-~\cite{neymanpearson19233lemma}
+~\cite{neymanpearson1933lemma}
 }
@@
 Tail areas of prior data distributions from two alternative hypotheses,
-reproduced from Neyman and Pearson~\cite{neymanpearson19233lemma}.
+reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
 }

### 2022-09-03 main/chapter_searches.tex
rearrange some quotes, add some likelihood fisher content

@@
 in the 1920s and 30s.
-Together, the Fisher and Neyman-Pearson methods do not form a harmonious
-framework, but comprise a loose collection of ideas and practices that survive
-in modern science, and explain important features of the methods we apply to
-LHC data, which Section~\ref{sec:searches_practice} develops further.
+These Fisher and Neyman-Pearson schools do not form a unified framework, but
+do comprise a collection of ideas and practices that survive in modern science,
+and which explain some methods, to be introduced in
+Section~\ref{sec:searches_practice}, that we apply to modern LHC data.
 
@@
 ``classical'' statistics~\cite{lehmann2011fisher, Feldman:1997qc}, but that is
-simply inaccurate for a 20th century divergence from 18th century
+simply inaccurate for a 20th century innovation created to replace 18th century
 ideas~\cite{bayes1763lii, laplace1774stigler}.
-Jaynes instead names them ``orthodox'~\cite{jaynes2003probability},
+Jaynes instead names them ``orthodox''~\cite{jaynes2003probability},
 which may be accurate in some contexts, but alternative
-(heretical?) ideas have taken hold under modern LHC authorities ---
+(heretical?) ideas have taken hold in research at the LHC ---
 the orthodoxy has moved on.
-We therefore stick to the name `frequentist', as it is a firm common factor.
+We therefore stick with `frequentist'.
 
@@
 Fisher~\cite{fisher1956statistical}.
-Levels are drawn at $1/2$, $1/5$, and $1/15$ of its maximum.
-Their extents could be reported in a table as a practical communication of
-the likelihood function's profile.
+Horizontal lines mark regions where the likelihood exceeds $1/2$, $1/5$, or
+$1/15$ of its maximum respectively.
+The extents of these lines could be tabulated to usefully communicate
+aspects of its profile.
 }
@@
 
+\paragraph{Fisher.}
 % likelihood
-Fisher coined the term `likelihood'~\cite{fisher1921probable}, in the usage
-we have employed, for the probability of observed data given varied
-contexts\footnote{%
+Fisher coined the term `likelihood'~\cite{fisher1921probable} in the usage that
+we employ for the probability of observed data given varied
+contexts%
+\footnote{%
 Fisher used the term `likelihood' for the function exploring
@@
 and deserves credit for popularising its use in interpreting and communicating
-data.
-% figure here TODO
-% likelihood quotes
-Fisher forcefully argued against the use of Bayes' rule and prior
-probabilities, stating for example that
-\begin{singlespacing}
+data without the confounding influence of prior probabilities.
+In his own words,%
 \begin{displayquote}
+\small
+``%
 We can know nothing of the probability of hypotheses or hypothetical
@@
 On the other hand we may ascertain the likelihood of hypotheses
-and hypothetical quantities by calculation from observations~\cite{
-fisher1921probable
-}
+and hypothetical quantities by calculation from observations%
+''~\cite{fisher1921probable},
 \end{displayquote}
-\end{singlespacing}
+and furthermore we agree that in comparison to confidence intervals,
+\begin{displayquote}
+``%
+For all purposes, and more particularly for the \textit{communication} of the
+relevant evidence supplied by a body of data, the values of the Mathematical
+Likelihood are better fitted to analyse, summarize, and communicate
+statistical evidence of types too weak to supply true probability statements%
+''~\cite{fisher1956statistical}.
+\end{displayquote}
+For example, Figure~\ref{fig:searches_fisher_likelihood} displays a figure by
+Fisher~\cite{fisher1956statistical} that illustrates how data can be usefully
+reported by their likelihood's maximum and the regions in which it exceeds
+certain ratios relative to that maximum.%
+\footnote{%
+We use such $L(\check{x})\pm\mathrm{ratio}$ visualizations for
+$\pm1\sigma$ and $\pm2\sigma$ error bars on Poisson data in
+Figure~7 of~\cite{lester2022hunting}.
+When making error bar recommendations, \atlas' statistics committee appears
+to have not considered this option~\cite{cowan2011errorbars};
+I cannot explain why not, since our statistics experts must surely be familiar
+with Fisher's work.
+}
+
+% likelihood quotes
+In avoiding the pitfalls of posteriors, Fisher's use of likelihoods remains
+innovative and practical.
+However, the hard rejection of priors on hypotheses also prevents calculation
+of evidence values, without which likelihoods can become confounded
+in too many parameters (perhaps making them nuisances).
+To circumvent this, frequentist methods employ two key concepts:
+test statistics, which can reduce the data to remove unwanted parameters,
+and $p$-values, which divert the discourse
+from what the alternative hypotheses \emph{are}
+to what alternative data \emph{might have been}.
+
+% test statistics
+A test statistic is a function of data.
+All real-valued functions are permitted, but some are more useful than others.
+\TODO{TODO t-statistic}
 
 % p-values
+(probable error?), known in modern jargon as $p$-values
 
@@
 }
-Figure~\ref{fig:searches_fisher_likelihood}
 
 \TODO{
-Fisher:
-``We can know nothing of the probability of hypotheses or hypothetical
-quantities.
-On the other hand we may ascertain the likelihood of hypotheses
-and hypothetical quantities by calculation from observations''
-~\cite{fisher1921probable}
 Fisher:
@@
 ~\cite{fisher1956statistical}
-% \begin{epigraphs}
-% \qitem{%
-% We can know nothing of the probability of hypotheses or hypothetical
-% quantities.
-% On the other hand we may ascertain the likelihood of hypotheses and
-% hypothetical quantities by calculation from observations%
-% }%
-% {Ronald A. Fisher, 1921~\cite{fisher1921probable}}
-% \end{epigraphs}
 There is something horrifying in the ideological movement represented by the
@@
 }
-
-\begin{singlespacing}
-\begin{displayquote}
-We can know nothing of the probability of hypotheses or hypothetical
-quantities.
-On the other hand we may ascertain the likelihood of hypotheses
-and hypothetical quantities by calculation from observations
-\end{displayquote}
-\end{singlespacing}
-~\cite{fisher1921probable}
-
-\begin{singlespacing}
-\begin{displayquote}
-For all purposes, and more particularly for the \textit{communication} of the
-relevant evidence supplied by a body of data, the values of the Mathematical
-Likelihood are better fitted to analyse, summarize, and communicate
-statistical evidence of types too weak to supply true probability statements
-\end{displayquote}
-\end{singlespacing}
-~\cite{fisher1956statistical}
 
@@
 
+\paragraph{Neyman-Pearson.}
 \TODO{

### 2022-09-03 main/chapter_searches.tex
write a little on test statistics

@@
 ``classical'' statistics~\cite{lehmann2011fisher, Feldman:1997qc}, but that is
-simply inaccurate for a 20th century innovation created to replace 18th century
+simply inaccurate for a 20th century creations that innovated on 18th century
 ideas~\cite{bayes1763lii, laplace1774stigler}.
@@
 \begin{displayquote}
+\small
 ``%
@@
 innovative and practical.
-However, the hard rejection of priors on hypotheses also prevents calculation
-of evidence values, without which likelihoods can become confounded
-in too many parameters (perhaps making them nuisances).
+However, the hard rejection of priors on hypotheses also prevents evidence
+calculations, without which likelihoods alone become confounded in many
+parameters, perhaps making them nuisances.
 To circumvent this, frequentist methods employ two key concepts:
@@
 % test statistics
-A test statistic is a function of data.
-All real-valued functions are permitted, but some are more useful than others.
-\TODO{TODO t-statistic}
+A statistic is a function of data, and a test statistic is just a statistic
+that is used in statistical tests.
+For \atlas\ collision data, event variables are statistics; when working with
+data, one always uses statistics and is free to design them for their benefits
+in context.
+Just as we define event variables (like $\pt$) to remove nuisance
+dependencies (on $\phi$ rotations or longitudinal boosts),
+useful test statistics can also eliminate unwanted parameters.
+The quintessential example is $t$-statistic
+\begin{equation}
+t = (\bar{\mu} - \mu)\sqrt{\frac{n}{\bar{\sigma}^2}}
+,
+\end{equation}
+where $\mu$ is the known data mean, and $\bar{\mu}$ and $\bar{\sigma}^2$ are
+standard estimates of the mean and variance of the data sample respectively.
+By dividing out a standard deviation, $t$ is independent of any common
+scale factors on the data~\cite{student1908, fisher1925t, lehmann2011fisher}.
 
@@
 
-\TODO{
-Fisher:
-``absurdly academic''
-~\cite{fisher1956statistical}
+``%
 There is something horrifying in the ideological movement represented by the
 doctrine that reasoning, properly speaking, cannot be applied to empirical data
-to lead to inferences valid in the real world.
-~\cite{fisher1956statistical}
-}
-
-\begin{singlespacing}
-\begin{displayquote}
-There is something horrifying in the ideological movement represented by the
-doctrine that reasoning, properly speaking, cannot be applied to empirical data
-to lead to inferences valid in the real world.
-\end{displayquote}
-\end{singlespacing}
-~\cite{fisher1956statistical}
+to lead to inferences valid in the real world.%
+''~\cite{fisher1956statistical}
 

### 2022-09-03 main/chapter_searches.tex
some rearrangements

@@
 
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.46\textwidth]{figures/searches_significant_shrink.png}
+\\
+\begin{footnotesize}
+`So, uh, we did the green study again and got no link. It was probably a{-}{-}'
+\\
+`RESEARCH CONFLICTED ON GREEN JELLY BEAN ACNE LINK; MORE~STUDY~RECOMMENDED!'
+\end{footnotesize}
+\caption[
+``Significant'' by Randall~Munroe
+]{%
+``Significant'' by Randall~Munroe~\cite{xkcd2011significant}.
+}
+\label{fig:searches_significant}
+\end{figure}
+
 % p-values
+\TODO[
 (probable error?), known in modern jargon as $p$-values
+tail area = significance
+~\cite{fisher1925smrw}
+}
+
+\TODO{maximum likelihood}
+\TODO{p-value}
+\TODO{p-value:
+The chance of a small $p$-value is small.
+This is true, but not interesting because the value of any $p$-value is small.
+By construction $p$-values are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+$p$-value to have meaning.
+}
+Figure~\ref{fig:searches_significant}
 
 % estimators --- not unbiased!
+\TODO{
+(unbiased estimators)
+Lehmann~\cite{lehmann2005testing};
+p. 26 poisson inadmissable example;
+L-unbiased with given loss function;
+}
 
 % savage likelihood principle
-
-\TODO{Fisher's great contribution --- likelihood}
 \TODO{
@@
 ~\cite{fisher1922estimators}
-testing, p-value,
-likelihood, also ratios of p-values (here named probable errors)
-(introduced the term likelihood? cite?)
-~\cite{fisher1921probable}
-tail area = significance
-~\cite{fisher1925smrw}
+}
+
+\TODO{likelihood principle ~\cite{
+birnbaum1962foundations,
+savage1962foundations
+}
 }
@@
 
-% TODO Kolmogorov axioms, unbiasedness
 \TODO{
@@
 \TODO{
-tossing a coin is considered random
-but if instead you are presented with a coin hidden under a cloth,
-no frequency interpretation --- it either is or is not heads
-you are therefore denied the ability to make uncertain predictions.
-more generally, you are not allowed to express an uncertainty in something
-unless some random process is involved
-}
-
-\TODO{
-Can keep likelihoods, but cannot compute evidence values unless the prior
-has a frequency interpretation.
-}
-
-\TODO{
 Its void is filled with various proposed alternatives that do have frequency
@@
 \TODO{
-maximum likelihood, p-values, confidence intervals, and biasedness of
-estimators.
-}
-
-\TODO{
-Lehmann~\cite{lehmann2005testing};
-p. 26 poisson inadmissable example;
-L-unbiased with given loss function;
-}
-
-\TODO{maximum likelihood}
-\TODO{p-value}
-\TODO{p-value:
-The chance of a small $p$-value is small.
-This is true, but not interesting because the value of any $p$-value is small.
-By construction $p$-values are uniformly distributed on the unit interval
-under a given hypothesis.
-The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
-$0.6 < p < 0.65$.
-The frequency of $p < \varepsilon$ is identical to the frequency
-of $0.6 < p < 0.6 + \varepsilon$.
-An appropriate choice of test statistic is therefore necessary for the
-$p$-value to have meaning.
-}
-Figure~\ref{fig:searches_significant}
+maximum likelihood, p-values, confidence intervals
+}
+
 \TODO{confidence interval}
 \TODO{unbiased}
-
-\TODO{likelihood principle ~\cite{
-birnbaum1962foundations,
-savage1962foundations
-}
-}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.46\textwidth]{figures/searches_significant_shrink.png}
-\\
-\begin{footnotesize}
-`So, uh, we did the green study again and got no link. It was probably a{-}{-}'
-\\
-`RESEARCH CONFLICTED ON GREEN JELLY BEAN ACNE LINK; MORE~STUDY~RECOMMENDED!'
-\end{footnotesize}
-\caption[
-``Significant'' by Randall~Munroe
-]{%
-``Significant'' by Randall~Munroe~\cite{xkcd2011significant}.
-}
-\label{fig:searches_significant}
-\end{figure}
 

### 2022-09-04 main/chapter_searches.tex
add draft p-value eqn
more boring text on p-values

@@
 
+\TODO{estimators briefly here}
+\TODO{maximum likelihood}
+
+% estimators --- not unbiased!
+\TODO{
+(unbiased estimators)
+Lehmann~\cite{lehmann2005testing};
+p. 26 poisson inadmissable example;
+L-unbiased with given loss function;
+}
+
+% savage likelihood principle
+\TODO{
+Fisher:
+criteria for estimators,
+(consistency, efficiency, sufficiency)
+~\cite{fisher1922estimators}
+}
+
+\TODO{likelihood principle ~\cite{
+birnbaum1962foundations,
+savage1962foundations
+}
+}
+
 % likelihood quotes
@@
 \begin{equation}
-t = (\bar{\mu} - \mu)\sqrt{\frac{n}{\bar{\sigma}^2}}
+t(d) = \frac{\bar{\mu}(d) - \mu}{\bar{\sigma}(d)}\sqrt{n}
 ,
 \end{equation}
-where $\mu$ is the known data mean, and $\bar{\mu}$ and $\bar{\sigma}^2$ are
-standard estimates of the mean and variance of the data sample respectively.
-By dividing out a standard deviation, $t$ is independent of any common
-scale factors on the data~\cite{student1908, fisher1925t, lehmann2011fisher}.
-
+where $\mu$ is the known data mean, and $\bar{\mu}(d)$ and $\bar{\sigma}(d)$
+are standard estimates of the mean and standard deviation of the data sample
+respectively~\cite{student1908, fisher1925t}.
+By dividing out a (scale equivariant) standard deviation, $t$ is independent
+of common scale factors on the data, and that scale invariance has helped to
+drive the study and common use of this $t$ statistic~\cite{lehmann2011fisher}.
+
+One can usefully reduce observed data to just the test statistic
+$\tilde{t} = t(d)$ and compare likelihoods to compare how well different
+hypotheses predict the observed $\tilde{t}$.
+For reasons I cannot explain, Fisher advocates a different use ---
+rather than considering likelihoods and likelihood ratios, he recommends
+that researchers consider the ``probable error''~\cite{fisher1921probable},
+which is a cumulative distribution function in the negated test statistic.
+In modern jargon, a `probable error' at the observed test statistic
+$\tilde{t}$ and an hypothesis $h$ is a \textbf{$p$-value}
+\begin{align}
+P(h)
+&= \prob{t \geq \tilde{t}}{h} \\
+&= L(h) + \prob{t > \tilde{t}}{h}
+\label{eqn:searches_pvalue_likelihood}
+;
+\end{align}
+the $p$-value $P(h)$ is the prior probability of data
+\emph{greater or equal to} the observed data when ordered by the value of the
+test statistic.
+Equation~\ref{eqn:searches_pvalue_likelihood} emphasizes that it is just the
+likelihood plus some fluff from a tail of larger test statistics.
+
+\TODO{quote EX.II from ~\cite{fisher1921probable}
+}
+
+\TODO{p-value:
+The chance of a small $p$-value is small.
+This is true, but not interesting because the value of any $p$-value is small.
+By construction $p$-values are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+$p$-value to have meaning.
+}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/searches_atlas_higgs_4l_2207_00320.pdf}
+\caption[
+Signal and background distributions reproduced from the recent \atlas\
+measurement of Higgs boson decays to $4\ell$
+]{%
+Signal and background distributions reproduced from the recent \atlas\
+measurement of Higgs boson decays to $4\ell$~\cite{ATLAS:2022net}, with
+observed data.
+This example illustrates $p$-values pathological dependency on tail selections,
+and the benefits of likelihood ratios.
+}
+\label{fig:atlas_higgs_4l}
+\end{figure}
+
+Fisher claims that the $p$-value indicates the ``significance'' of an
+observation~\cite{fisher1925smrw}, with small or large values being
+suspiciously significant, and that
+``If $P$ is between $.1$ and $.9$ there is certainly no reason to suspect the
+hypothesis tested.''~\cite{fisher1925smrw}
+This may have been true for his intended applications, but it is not general.
+
+Consider for example the distributions of signal (Higgs boson) and background
+(mostly $ZZ^*$) samples in Figure~\ref{fig:atlas_higgs_4l}.
+If an event is observed with $m_{4\ell} = 125\,\eV[G]$, is it signal or
+background?
+It is obviously a Higgs boson candidate, but $m_{4\ell}$ is around the middle
+of both signal and background distributions, so with an $m_{4\ell}$ test
+statistic $P(\textrm{Higgs boson}) \approx P(ZZ^*) \approx 0.5$, so neither
+is ``suspect''.
+Local ratios between the two distributions clearly favour signal for this
+event, however; the $p$-values' failure is in their dependence on irrelevant
+tail areas, which are a feature of the histogram and not the data.
+In fact, this $ZZ^*$ density peaks for $ZZ$ above a
+$2 m(Z) \approx 180\,\eV[G]$
+edge, and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$;
+expanding the histogram to show either of these tails would change nothing
+local to our $m_{4\ell} = 125\,\eV[G]$ event, but would change the $p$-values.
+
+Better test statistics can soften this problem, and a key observation by
+Neyman and Pearson is an optimality of likelihood ratio test statistics.
+
+\TODO{
+``%
+There is something horrifying in the ideological movement represented by the
+doctrine that reasoning, properly speaking, cannot be applied to empirical data
+to lead to inferences valid in the real world.%
+''~\cite{fisher1956statistical}
+}
+
+\paragraph{Neyman-Pearson.}
 \begin{figure}[tp]
@@
 \end{figure}
-
-% p-values
-\TODO[
-(probable error?), known in modern jargon as $p$-values
-tail area = significance
-~\cite{fisher1925smrw}
-}
-
-\TODO{maximum likelihood}
-\TODO{p-value}
-\TODO{p-value:
-The chance of a small $p$-value is small.
-This is true, but not interesting because the value of any $p$-value is small.
-By construction $p$-values are uniformly distributed on the unit interval
-under a given hypothesis.
-The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
-$0.6 < p < 0.65$.
-The frequency of $p < \varepsilon$ is identical to the frequency
-of $0.6 < p < 0.6 + \varepsilon$.
-An appropriate choice of test statistic is therefore necessary for the
-$p$-value to have meaning.
-}
 Figure~\ref{fig:searches_significant}
 
-% estimators --- not unbiased!
-\TODO{
-(unbiased estimators)
-Lehmann~\cite{lehmann2005testing};
-p. 26 poisson inadmissable example;
-L-unbiased with given loss function;
-}
-
-% savage likelihood principle
-\TODO{
-Fisher:
-criteria for estimators,
-(consistency, efficiency, sufficiency)
-~\cite{fisher1922estimators}
-}
-
-\TODO{likelihood principle ~\cite{
-birnbaum1962foundations,
-savage1962foundations
-}
-}
-
-``%
-There is something horrifying in the ideological movement represented by the
-doctrine that reasoning, properly speaking, cannot be applied to empirical data
-to lead to inferences valid in the real world.%
-''~\cite{fisher1956statistical}
-
-\paragraph{Neyman-Pearson.}
 \TODO{
@@
 
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/searches_atlas_higgs_4l_2207_00320.pdf}
-\caption[
-Signal and background distributions reproduced from the recent \atlas\
-measurement of Higgs boson decays to $4\ell$
-]{%
-Signal and background distributions reproduced from the recent \atlas\
-measurement of Higgs boson decays to $4\ell$~\cite{ATLAS:2022net}.
-}
-\label{fig:atlas_higgs_4l}
-\end{figure}
-
-Figure~\ref{fig:atlas_higgs_4l}
-
 \TODO{

### 2022-09-06 main/chapter_searches.tex
pvalue ratio versus likelihood ratio

@@
 ratio form:
+\TODO{different choice for phi to avoid angle?}
 \begin{equation}
@@
 
-From such $\log$ ratios one individual probabilities are recovered through the
-logistic function:
+As data are reported more precisely, likelihoods will often tend towards zero
+--- in the continuum limit, the probability of any data value vanishes.
+Small likelihoods are themselves not significant unless other likelihoods
+are larger --- unless other hypotheses predicted \emph{better}.
+And these ($\log$) ratios can remain well defined in continuum limits even as
+both likelihoods tend towards zero~\cite{billingsley2008probability}.
+
+From such $\log$ ratios, a little rearrangement recovers the individual
+probabilities through the logistic function:
 \begin{equation}
@@
 ``classical'' statistics~\cite{lehmann2011fisher, Feldman:1997qc}, but that is
-simply inaccurate for a 20th century creations that innovated on 18th century
+inaccurate for 20th century creations that innovated on 18th century
 ideas~\cite{bayes1763lii, laplace1774stigler}.
 Jaynes instead names them ``orthodox''~\cite{jaynes2003probability},
-which may be accurate in some contexts, but alternative
+which may have been accurate in his environment, but different
 (heretical?) ideas have taken hold in research at the LHC ---
@@
 useful test statistics can also eliminate unwanted parameters.
-The quintessential example is $t$-statistic
+The quintessential example is the $t$-statistic
 \begin{equation}
@@
 \end{equation}
-where $\mu$ is the known data mean, and $\bar{\mu}(d)$ and $\bar{\sigma}(d)$
-are standard estimates of the mean and standard deviation of the data sample
-respectively~\cite{student1908, fisher1925t}.
-By dividing out a (scale equivariant) standard deviation, $t$ is independent
-of common scale factors on the data, and that scale invariance has helped to
-drive the study and common use of this $t$ statistic~\cite{lehmann2011fisher}.
+which shifts the sample mean $\bar{\mu}(d)$ by a known true mean $\mu$, and
+scales that difference down by the same standard deviation
+$\bar{\sigma}(d)$~\cite{student1908, fisher1925t}.
+By dividing out a scale equivariant estimate of the standard deviation, $t$ is
+independent of common scale factors on the data, and that scale invariance
+has helped to drive the study and common use of this
+$t$-statistic~\cite{lehmann2011fisher}.
 
 One can usefully reduce observed data to just the test statistic
-$\tilde{t} = t(d)$ and compare likelihoods to compare how well different
-hypotheses predict the observed $\tilde{t}$.
+$\tilde{t} = t(d)$, and compare likelihoods on that observed $\tilde{t}$
+to compare how well different hypotheses predict it.
 For reasons I cannot explain, Fisher advocates a different use ---
@@
 P(h)
-&= \prob{t \geq \tilde{t}}{h} \\
+&= \prob{t \geq \tilde{t}}{h}
+\label{eqn:searches_pvalue}
+\\
 &= L(h) + \prob{t > \tilde{t}}{h}
@@
 test statistic.
-Equation~\ref{eqn:searches_pvalue_likelihood} emphasizes that it is just the
-likelihood plus some fluff from a tail of larger test statistics.
-
-\TODO{quote EX.II from ~\cite{fisher1921probable}
-}
-
-\TODO{p-value:
-The chance of a small $p$-value is small.
-This is true, but not interesting because the value of any $p$-value is small.
-By construction $p$-values are uniformly distributed on the unit interval
-under a given hypothesis.
-The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
-$0.6 < p < 0.65$.
-The frequency of $p < \varepsilon$ is identical to the frequency
-of $0.6 < p < 0.6 + \varepsilon$.
-An appropriate choice of test statistic is therefore necessary for the
-$p$-value to have meaning.
-}
+Equation~\ref{eqn:searches_pvalue_likelihood} points out that the $p$-value is
+just is the likelihood plus some fluff from a tail of larger test statistics.
+
+Fisher illustrates the use of $p$-values in the second example
+of~\cite{fisher1921probable}, which models some data with a parameter $r$ and
+compares two hypotheses
+$r=0.3$ and
+$r=0.95$, finding $p$-values
+$P(r=0.3)=0.142$ and
+$P(r=0.95)=0.00014$.
+``Thus'', Fisher concludes, ``the value $0.95$\ldots
+is roughly $1000$ times less likely than $0.30$''
+(``when measured on the $r$ scale'').
+By writing `likely', Fisher is carefully not referencing the probability
+of parameters, but their likelihoods in our definition.
+But this factor of $1000$ is a $p$-value ratio, not a likelihood ratio,
+so how can this make sense?
+Plausibly, it is a data reduction; comparing
+Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
+shows that the $p$-value is a likelihood (only) if the result is reduced from
+$A: t =    \tilde{t}$ to
+$B: t \geq \tilde{t}$.
+And since this reduction is technically true --- $A$ implies $B$ --- the
+likelihood ratio on $B$ is logically valid and not \emph{necessarily}
+misleading.
+
+But $B: t \geq \tilde{t}$ is a logical statement (true for all test statistics
+greater than or equal to the observed value) and not a randomly occurring
+event, so its assignment frequentist probability is deeply dubious.
+Much more importantly, the reduction to a one-sided inequality has no tangible
+benefit; it reduces the information content of the reported data, but does not
+reduce the costs of reporting or interpreting those data.
+Both $A$ and $B$ are printed in the same number of characters, and $p$-values
+can be painfully expensive to evaluate.
+Furthermore, it can actively harm the result by diluting the likelihood ratio,
+since if
+\begin{equation}
+\phi = \log \frac{L(h_1)}{L(h_2)}
+\to
+\log \frac{
+L(h_1) + \prob{t > \tilde{t}}{h_1}
+}
+{
+L(h_2) + \prob{t > \tilde{t}}{h_2}
+}
+\end{equation}
+then large tail probabilities will only squash the $\log$ ratio toward zero.
 
@@
 
+\TODO{p-value:
+The chance of a small $p$-value is small.
+This is true, but not interesting because the value of any $p$-value is small.
+By construction $p$-values are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+$p$-value to have meaning.
+}
+
 \begin{figure}[tp]

### 2022-09-06 main/chapter_searches.tex
write on p-value ratios

@@
 likelihood ratio on $B$ is logically valid and not \emph{necessarily}
-misleading.
+misleading if clearly stated.
 
@@
 can be painfully expensive to evaluate.
-Furthermore, it can actively harm the result by diluting the likelihood ratio,
-since if
-\begin{equation}
-\phi = \log \frac{L(h_1)}{L(h_2)}
+Furthermore, it actively harms any result with small a test statistic by
+diluting the likelihood ratio, since if
+\begin{equation}
+\log \frac{L(h_1)}{L(h_2)}
 \to
@@
 \end{equation}
-then large tail probabilities will only squash the $\log$ ratio toward zero.
+then large tail probabilities will only squash the $\log$ likelihood ratio
+toward zero.
 

### 2022-09-06 main/chapter_searches.tex
minor tweaks on likelihoods and p-values

@@
 \end{equation}
-then large tail probabilities will only squash the $\log$ likelihood ratio
-toward zero.
+then large tail probabilities squash the $\log$ likelihood ratio toward zero.
 
@@
 hypothesis tested.''~\cite{fisher1925smrw}
-This may have been true for his intended applications, but it is not general.
+This may have been true for his intended applications, but it is not general;
+as for usual likelihood functions, small values are only meaningful in
+comparison against larger values.
 

### 2022-09-06 main/chapter_searches.tex
clean up bib; add some refs for p-values; add a note on fisher pvalue ratio

@@
 \caption[
-Ratio thresholds of a Binomial likelihood function, reproduced from Fisher
+Intervals exceeding ratios from the maximum of a Binomial likelihood function,
+reproduced from Fisher
 ]{%
-Ratio thresholds of a Binomial likelihood function, reproduced from
-Fisher~\cite{fisher1956statistical}.
-Horizontal lines mark regions where the likelihood exceeds $1/2$, $1/5$, or
-$1/15$ of its maximum respectively.
-The extents of these lines could be tabulated to usefully communicate
-aspects of its profile.
+Intervals exceeding ratios from the maximum of a Binomial likelihood function,
+reproduced from Fisher~\cite{fisher1956statistical}.
+Horizontal lines mark regions where the likelihood function exceeds
+$1/2$, $1/5$, and $1/15$ of its maximum respectively.
+The extents of these lines could be tabulated to usefully communicate aspects
+of its profile.
 }
@@
 % likelihood
-Fisher coined the term `likelihood'~\cite{fisher1921probable} in the usage that
+Fisher coined the term `likelihood'~\cite{
+fisher1912fitting,
+fisher1915frequency,
+fisher1921probable
+} in the usage that
 we employ for the probability of observed data given varied
@@
 so how can this make sense?
-Plausibly, it is a data reduction; comparing
+Plausibly, it results from a tacit data reduction; comparing
 Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
@@
 
-But $B: t \geq \tilde{t}$ is a logical statement (true for all test statistics
-greater than or equal to the observed value) and not a randomly occurring
-event, so its assignment frequentist probability is deeply dubious.
+However, $B: t \geq \tilde{t}$ is a logical statement (true for all test
+statistics greater than or equal to the observed value) and not a randomly
+occurring event; it therefore does not have a frequentist probability and so
+is probably not what Fisher intended.
 Much more importantly, the reduction to a one-sided inequality has no tangible
-benefit; it reduces the information content of the reported data, but does not
-reduce the costs of reporting or interpreting those data.
+benefit --- it reduces the information content of the reported data without
+reducing the associated costs of reporting or interpreting those data.
 Both $A$ and $B$ are printed in the same number of characters, and $p$-values
@@
 then large tail probabilities squash the $\log$ likelihood ratio toward zero.
+
+Fisher's $p$-value ratio of $1000$ is, therefore, a $\log$ likelihood
+ratio of $+7$ in favour of $r=0.3$ over $r=0.95$, which is a moderately strong
+indication of roughly $3.7\sigma$.
+But a likelihood ratio on the test statistic itself, and not the inclusive
+tail statement, would encode strictly more information and could have provide a
+more precise answer.
 
@@
 
+% p-values
+no!~\cite{
+schervish1996p,
+Cohen1994TheEI,
+goodman2008dirty,
+greenland2016no,
+Amrhein2017TheEI,
+Bernstein2016princess,
+Trafimow2015ban
+}
+
 \begin{singlespacing}

### 2022-09-09 main/chapter_searches.tex
fix many doi links; add cls references

@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/seacrhes_fisher_likelihood_binomial.png}
+\includegraphics[width=0.95\textwidth]{figures/searches_fisher_likelihood_binomial.png}
 \caption[
@@
 
-\paragraph{Fisher.}
+\subsection{Fisher}
+\label{sec:searches_fisher}
 % likelihood
@@
 \end{displayquote}
-and furthermore we agree that in comparison to confidence intervals,
+and furthermore we agree that in comparison to other prescriptions
+(specifically confidence intervals, to be introduced in
+Section~\ref{sec:searches_np})
 \begin{displayquote}
@@
 
-\paragraph{Neyman-Pearson.}
+\subsection{Neyman-Pearson}
+\label{sec:searches_np}
 \begin{figure}[tp]
@@
 Cohen1994TheEI,
+Amrhein2017TheEI,
 goodman2008dirty,
 greenland2016no,
-Amrhein2017TheEI,
 Bernstein2016princess,
-Trafimow2015ban
+Trafimow2015ban,
+Read2002cls
 }
@@
 \end{singlespacing}
+
+$\cls$~\cite{Read:451614, Read2002cls, ralcls}
 

### 2022-09-09 main/chapter_searches.tex
write a little on biasedness; add some refs; relocate closing Fisher quote

@@
 \begin{equation}
+\label{eqn:searches_log_odds}
 \hphantom{\quad \quad \mid\mid \mathrm{context}}
-\phi(d) =
+\Delta(d) =
 \log\frac{\prob{S}{d}}{\prob{B}{d}}
@@
 \label{eqn:searches_logistic}
-\prob{S}{d} = \frac{1}{1 + e^{-\phi(d)}}
+\prob{S}{d} = \frac{1}{1 + e^{-\Delta(d)}}
 .
@@
 Ronald A. Fisher (for likelihood, $p$-values, and estimation criteria~\cite{
+fisher1912fitting,
+fisher1915frequency,
 fisher1921probable,
@@
 fisher1915frequency,
-fisher1921probable
+fisher1921probable,
+fisher1922estimators
 } in the usage that
@@
 Fisher~\cite{fisher1956statistical} that illustrates how data can be usefully
-reported by their likelihood's maximum and the regions in which it exceeds
+reported by the likelihood's maximum and the regions in which it exceeds
 certain ratios relative to that maximum.%
@@
 }
-
-\TODO{estimators briefly here}
-\TODO{maximum likelihood}
-
-% estimators --- not unbiased!
-\TODO{
-(unbiased estimators)
-Lehmann~\cite{lehmann2005testing};
-p. 26 poisson inadmissable example;
-L-unbiased with given loss function;
-}
-
-% savage likelihood principle
-\TODO{
-Fisher:
-criteria for estimators,
-(consistency, efficiency, sufficiency)
-~\cite{fisher1922estimators}
-}
-
-\TODO{likelihood principle ~\cite{
+That the likelihood contains all relevant information in the data is true by
+definition in our formulation of Equation~\ref{eqn:searches_log_odds},
+but non-Bayesian arguments find the same conclusion~\cite{
 birnbaum1962foundations,
 savage1962foundations
-}
-}
+}.
+
+Maximum likelihoods are often useful for interpreting and communicating data.
+The location of a maximum likelihood can be a useful estimate for the
+parameters being considered, and Fisher showed that maximum likelihood
+estimates satisfy some desirable criteria~\cite{fisher1922estimators}.
+These criteria do not include \emph{unbiasedness}, which states that the mean
+value of an estimate (over the distribution for alternative data) should equal
+its known true value~\cite{pdg2020review}.
+Defining bias as a difference from a mean fails to be coordinate invariant,
+so an unbiased estimate of the variance $\bar{\sigma}^2$ is a biased
+estimate of the standard deviation $\bar{\sigma}$ --- the square of the mean
+is not the mean of the squares~\cite{Barlow:2019svl}.
+Fisher rejected the bias criterion for this
+reason~\cite{jaynes2003probability}, and other definitions absorb its
+arbitrariness into an explicit loss function~\cite{lehmann2005testing}.
+Requiring unbiasedness can actively harm our results by preventing the use of
+useful statistics,
+\footnote{%
+Problem~1.14 of~\cite{lehmann2005testing} has a delightful example of
+estimating an exponentiated a Poisson mean $e^\mu$ from $n$ data
+where the unique unbiased estimate is $(-1)^{n + 1}$.
+Section~17.3 of~\cite{jaynes2003probability} shows similar nonsense for
+unbiased estimates of $\mu^2$ and $\exp(-\mu)$.%
+}
+or by obfuscating those statistics with bias corrections.
+Remaining uses of unbiasedness in High Energy Physics~\cite{
+pdg2020review,
+Tullythesis,
+lhcb2018563,
+LHCb:2021trn,
+LHCb:2015yax
+}
+are largely explained by the linguistic trickery in its naming --- our
+results should not be biased by forces outside the data and relevant theory,
+and we would love to claim that to be true.
+But we should proudly proclaim that our results are `biased' under this flawed
+statistical definition if that makes them better by any reasonable measure.
 
 % likelihood quotes
-In avoiding the pitfalls of posteriors, Fisher's use of likelihoods remains
-innovative and practical.
+In avoiding the pitfalls of reported posteriors, Fisher's use of likelihoods
+remains innovative and practical.
 However, the hard rejection of priors on hypotheses also prevents evidence
@@
 Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
-shows that the $p$-value is a likelihood (only) if the result is reduced from
+shows that the $p$-value is a likelihood only if the result is reduced from
 $A: t =    \tilde{t}$ to
@@
 likelihood ratio on $B$ is logically valid and not \emph{necessarily}
-misleading if clearly stated.
+misleading if accurately reported.
 
@@
 can be painfully expensive to evaluate.
-Furthermore, it actively harms any result with small a test statistic by
+And it actively harms any result with small a test statistic by
 diluting the likelihood ratio, since if
@@
 Fisher's $p$-value ratio of $1000$ is, therefore, a $\log$ likelihood
-ratio of $+7$ in favour of $r=0.3$ over $r=0.95$, which is a moderately strong
+ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately strong
 indication of roughly $3.7\sigma$.
 But a likelihood ratio on the test statistic itself, and not the inclusive
-tail statement, would encode strictly more information and could have provide a
+tail statement, would encode strictly more information and could provide a
 more precise answer.
@@
 This may have been true for his intended applications, but it is not general;
-as for usual likelihood functions, small values are only meaningful in
-comparison against larger values.
+as for all other likelihoods, small values are only meaningful in comparison
+against larger values from alternative hypotheses.
 
@@
 (mostly $ZZ^*$) samples in Figure~\ref{fig:atlas_higgs_4l}.
-If an event is observed with $m_{4\ell} = 125\,\eV[G]$, is it signal or
+If an event is observed at $m_{4\ell} = 125\,\eV[G]$, is it signal or
 background?
@@
 tail areas, which are a feature of the histogram and not the data.
-In fact, this $ZZ^*$ density peaks for $ZZ$ above a
-$2 m(Z) \approx 180\,\eV[G]$
-edge, and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$;
+This $ZZ^*$ density peaks for $ZZ$ above a
+$2 m(Z) \approx 180\,\eV[G]$ kinematic edge,
+and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$;
 expanding the histogram to show either of these tails would change nothing
-local to our $m_{4\ell} = 125\,\eV[G]$ event, but would change the $p$-values.
+local to our $m_{4\ell} = 125\,\eV[G]$ event, but would change our $p$-values.
 
@@
 Neyman and Pearson is an optimality of likelihood ratio test statistics.
-
-\TODO{
-``%
-There is something horrifying in the ideological movement represented by the
-doctrine that reasoning, properly speaking, cannot be applied to empirical data
-to lead to inferences valid in the real world.%
-''~\cite{fisher1956statistical}
-}
 
@@
 \caption[
-Tail areas of prior data distributions from two alternative hypotheses,
-reproduced from Neyman and Pearson
+Tail areas of probability distributions of data from two alternative
+hypotheses, reproduced from Neyman and Pearson
 ]{%
-Tail areas of prior data distributions from two alternative hypotheses,
-reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
+Tail areas of probability distributions of data from two alternative
+hypotheses, reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
 }
@@
 \TODO{unbiased}
+
+\begin{displayquote}
+\small
+``%
+There is something horrifying in the ideological movement represented by the
+doctrine that reasoning, properly speaking, cannot be applied to empirical data
+to lead to inferences valid in the real world.%
+''~\cite{fisher1956statistical}
+\end{displayquote}
 

### 2022-09-10 main/chapter_2ljets.tex
fix - to + for energy
shorten foddy quote
tidy up some quotes; add confidence interval citations

@@
 \begin{Bmatrix}
-m_T(a_\mu + \{\!\sqrt{m^2 - \vec p_T^{\:2}},\,\vec p\,\}_\mu)\\
-m_T(b_\mu + \{\!\sqrt{m^2 - \vec q_T^{\:2}},\,\vec q\,\}_\mu)\\
+m_T(a_\mu + \{\!\sqrt{m^2 + \vec p_T^{\:2}},\,\vec p\,\}_\mu)\\
+m_T(b_\mu + \{\!\sqrt{m^2 + \vec q_T^{\:2}},\,\vec q\,\}_\mu)\\
 \end{Bmatrix}
@@
 we guarantee that our tests can be suboptimal because they are not protected by
-the Neyman–Pearson lemma~\cite{neyman1933ix}.
+the Neyman–Pearson lemma~\cite{neymanpearson1933lemma}.
 In this world of model comparison through $p$-values, sub-optimality means
@@
 \qitem{%
-When you're building a video game world, you're building with ideas, and that
-can be like working with quick-set cement. You mold your ideas into a certain
-shape that can be played with, and in the process of playing with them they
-begin to harden and set until they're immutable like rock, and at that point
-you can't change the world --- not without breaking it into pieces and starting
-fresh with new ideas.%
+% When you're building a video game world, you're building with ideas, and that
+% can be like working with quick-set cement.
+You mold your ideas into a certain shape that can be played with, and in the
+process of playing with them they begin to harden and set until they're
+immutable like rock, and at that point you can't change the world ---
+not without breaking it into pieces and starting  fresh with new ideas.%
 }%

### 2022-09-10 main/chapter_searches.tex
tidy up some quotes; add confidence interval citations

@@
 Some literature names this Fisher/Neyman-Pearson assemblage
-``classical'' statistics~\cite{lehmann2011fisher, Feldman:1997qc}, but that is
-inaccurate for 20th century creations that innovated on 18th century
-ideas~\cite{bayes1763lii, laplace1774stigler}.
+``classical'' statistics~\cite{
+Neyman1937Outline,
+lehmann2011fisher,
+Feldman:1997qc
+}, but that is simply inaccurate for 20th century creations that innovated on
+18th century ideas~\cite{bayes1763lii, laplace1774stigler}.
 Jaynes instead names them ``orthodox''~\cite{jaynes2003probability},
@@
 Requiring unbiasedness can actively harm our results by preventing the use of
-useful statistics,
+useful estimates,%
 \footnote{%
@@
 }
-or by obfuscating those statistics with bias corrections.
+or by obfuscating those estimates with bias corrections.
 Remaining uses of unbiasedness in High Energy Physics~\cite{
@@
 }
-are largely explained by the linguistic trickery in its naming --- our
+are largely explained by the linguistic trickery of its name --- our
 results should not be biased by forces outside the data and relevant theory,
-and we would love to claim that to be true.
+and we like to report that to be true.
 But we should proudly proclaim that our results are `biased' under this flawed
@@
 $P(r=0.95)=0.00014$.
-``Thus'', Fisher concludes, ``the value $0.95$\ldots
-is roughly $1000$ times less likely than $0.30$''
+``Thus'', Fisher concludes, ``the value $0.95$\ldots is roughly $1000$ times
+less likely than $0.30$''
 (``when measured on the $r$ scale'').
@@
 This may have been true for his intended applications, but it is not general;
-as for all other likelihoods, small values are only meaningful in comparison
-against larger values from alternative hypotheses.
+as for other likelihoods, small values are only meaningful in comparison
+against larger values from alternative hypotheses on the same data.
 
@@
 \label{sec:searches_np}
+Suspecting an hypothesis when its $p$-value lands below (or above) a threshold
+means that there is a prior probability $\alpha$ that any hypothesis will be
+suspect after an experiment.
+For example, in Fisher's suggestion of $P(h) \in 0.1--0.9$ being not suspect,
+then $\alpha = 0.2$ (if the distribution for is continuous or
+otherwise meets the edges at $0.1$ and $0.9$ in its cumulative distribution).
+And that chance is independent of the hypothesis, its alternatives, and even
+the nature of the experiment.
+Blindly followed, these rules mean that the data and test statistic could be
+anything at all and still incur suspicion for unrelated hypotheses with the
+same frequency.
+Your wall clock reads not less than $48$ minutes past the hour?
+The Standard Model is suspect.
+This peculiarity is illustrated at the wildly popular $\alpha = 0.05$ level by
+Figure~\ref{fig:searches_significant}.
+
 \begin{figure}[tp]
@@
 \caption[
-``Significant'' by Randall~Munroe
+``Significant'' from xkcd by Randall~Munroe
 ]{%
-``Significant'' by Randall~Munroe~\cite{xkcd2011significant}.
+``Significant'' from xkcd by Randall~Munroe~\cite{xkcd2011significant}.
 }
@@
 \end{figure}
-Figure~\ref{fig:searches_significant}
-
-\TODO{
-Nayman-Pearson:
-Seek good test statistics --
-likelihood ratio intuitive for two-hypothesis comparison
-prove it optimal in a reasonable sense
-idea of long-run mostly not wrong
-recede to maximum likelihood ratios
-~\cite{neymanpearson1928max}
+
+Neyman and Pearson formalized this $p$-value based suspicion into decision
+rules and thought about what it is that makes one test statistic better than
+another~\cite{neymanpearson1928max, neymanpearson1933lemma}.
+Inspired by the same Bayesian arguments that we used to suggest likelihood
+ratios for model comparison~\cite{neymanpearson1928max}, they considered test
+statistics based on likelihoods\TODO{}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/searches_np_curves_crop.png}
+\caption[
+Tail areas of probability distributions of data from two alternative
+hypotheses, reproduced from Neyman and Pearson
+]{%
+Tail areas of probability distributions of data from two alternative
+hypotheses, reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
+}
+\label{fig:searches_np_tails}
+\end{figure}
+
+Figure~\ref{fig:searches_np_tails}
+
+\TODO{extend}
 ``for which hypothesis A is more probable and those for which it is less
 probable (sic!)''
-unbiased (unbiassed) test differs from unbiasedness of estimator
-``Then the two corresponding densities at the sample point
-are (Al) 6-42 x 10-20, (A2) 5 18*x 1O-4 and the ratio of the likelihood of Hypothesis
-Al to- that of A2 is 1P24 x 10-16. This ratio confirms the common-sense judgment
-that A2 is far more plausible than A1''
-\^ Another posterior argument --- this likelihood ratio in in fact a change
-in plausibilities. For example, if A2 is an inconsistent, self-contradictory
-theory, it is totally implausible, and not more plausible despite predicting
-these data better.
-``The statistician can
-balance the numerical verdict of likelihood against the vaguer expectations de-
-rived from `a priori' considerations, and it must be left to his judgment to
-at what point the evidence in favour of alternative hypotheses becomes so con-
-vincing that Hypothesis A must be rejected.''
+to which Lehmann appends ``(sic!)''~\cite{lehmann2011fisher}.
+
+\TODO{consider changing our A, B to A1 A2}
+\begin{displayquote}
+``%
+Then the two corresponding densities at the sample point are
+(A1) $6.42 \times 10^{-20}$,
+(A2) $5.18 \times 10^{-4}$ and the ratio of the likelihood of Hypothesis
+A1 to that of A2 is $1.24 \times 10^{-16}$.
+This ratio confirms the common-sense judgment that A2 is far more plausible
+than A1''
+''~\cite{neymanpearson1933lemma}
+\end{displayquote}
+
+\begin{displayquote}
+``%
+The statistician can balance the numerical verdict of likelihood against the
+vaguer expectations derived from `a priori' considerations, and it must be left
+to his judgment to at what point the evidence in favour of alternative
+hypotheses becomes so convincing that Hypothesis A must be rejected.%
+''~\cite{neymanpearson1933lemma}
+\end{displayquote}
+
+\TODO{
+\^ Another posterior argument
 Introduce maximum likelihood ratio as interesting without claiming it to be
@@
 
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/searches_np_curves_crop.png}
-\caption[
-Tail areas of probability distributions of data from two alternative
-hypotheses, reproduced from Neyman and Pearson
-]{%
-Tail areas of probability distributions of data from two alternative
-hypotheses, reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
-}
-\label{fig:searches_np_tails}
-\end{figure}
-
-Figure~\ref{fig:searches_np_tails}
+\TODO{Confidence intervals~\cite{neyman1935Intervals, Neyman1937Outline}}
 
@@
 }
-
-\TODO{
-Given all this understanding, what I do not understand is their choice to
-convert this to a $p$-value.
-Seems to be this idea of long-run correctness guarantees.
-But their derivation does rely on long-run frequencies.
-}
-
-\TODO{
-restrict our notion of a probability with a special case: only frequency
-of random events
-}
-
-\TODO{
-Its void is filled with various proposed alternatives that do have frequency
-interpretations and that sometimes yield useful results.
-}
-
-\TODO{
-maximum likelihood, p-values, confidence intervals
-}
-
-\TODO{confidence interval}
-\TODO{unbiased}
 
@@
 }
+\TODO{examples of use and misuse in hep}
 
@@
 
-$\cls$~\cite{Read:451614, Read2002cls, ralcls}
+$\cls$~\cite{Read:451614, Read2002cls, Murray2010heretic}
 

### 2022-09-10 main/chapter_2ljets.tex
change rightarrow to to

@@
 $\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
-that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
+that decay via $W\to q\bar q$ and $Z\to \ell^\pm \ell^\mp$
 respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
@@
 Although the Higgs and $Z$ bosons can decay to many states, we target events
-with $Z\rightarrow \ell\ell$ and $h/Z\rightarrow bb/jj$.
+with $Z\to \ell\ell$ and $h/Z\to bb/jj$.
 We explore the parameters
-$m(\neutralino_1)$ and $B(\neutralino_1 \rightarrow h \tilde{G})$ with a tiny
+$m(\neutralino_1)$ and $B(\neutralino_1 \to h \tilde{G})$ with a tiny
 fixed mass splitting of
@@
 (left) A fully-resolved decay tree with its particles labelled.
-(right) Only the $Z\rightarrow\ell\ell~$ \underline{V}ector boson
+(right) Only the $Z\to\ell\ell~$ \underline{V}ector boson
 is resolved, but the supersymmetric system recoils off
@@
 masses with $m(\neutralino_1) \in 250\textrm{--}600\,\eV[G]$.
-Exclusion of the largest $B(\neutralino_1 \rightarrow h \tilde{G})$ values
+Exclusion of the largest $B(\neutralino_1 \to h \tilde{G})$ values
 requires sensitivity to double-Higgs decay signatures --- \cms\ have published
-exclusions in the region combining $hh\rightarrow \gamma\gamma+\met$ and
-$hh\rightarrow b\bar bb\bar b$ with other channels to exclude
-$B(\neutralino_1 \rightarrow h \tilde{G}) = 1$ up to
+exclusions in the region combining $hh\to \gamma\gamma+\met$ and
+$hh\to b\bar bb\bar b$ with other channels to exclude
+$B(\neutralino_1 \to h \tilde{G}) = 1$ up to
 $m(\neutralino_1) = 750\,\eV[G]$~\cite{
@@
 Of these, $\twoljets$ uses only $b$-tagging, which is particularly useful to
-reject $\ttbar$ events and select for $h\rightarrow b\bar b$ decays.
+reject $\ttbar$ events and select for $h\to b\bar b$ decays.
 
@@
 $\ttbar$ backgrounds (since their weak decays almost always produce $b$
-quarks), and to select for signals producing $Z/h \rightarrow b\bar b$ decays
+quarks), and to select for signals producing $Z/h \to b\bar b$ decays
 in the GMSB model.
@@
 Indeed, momentum-balanced events do have energy imbalances;
-for example, in a $Z\mathrm{+jet}$ event $pp \rightarrow Zg$, with the $Z$ and
+for example, in a $Z\mathrm{+jet}$ event $pp \to Zg$, with the $Z$ and
 $g$ emitted back-to-back in the transverse plane with equal momenta
@@
 What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
-selections is almost pure \diboson\ --- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
-and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
+selections is almost pure \diboson\ --- $ZZ$ with a hard $Z\to \nu\nu$,
+and $W^\pm Z$ where the charged lepton in $W^\pm\to\ell^\pm\nu$
 is not reconstructed; these subcomponents of the \diboson\ process are discussed
@@
 For example, an on-shell $\tilde\ell^+\tilde\ell^-$ pair decaying as
-$\tilde\ell^+ \rightarrow \ell^+ \neutralino_1
+$\tilde\ell^+ \to \ell^+ \neutralino_1
 ,~
-\tilde\ell^- \rightarrow \ell^- \neutralino_1$ has
+\tilde\ell^- \to \ell^- \neutralino_1$ has
 $\mttwo(\ell^+, \ell^-, m(\neutralino_1)) \leq m(\tilde\ell)$, plus some error
@@
 All $\twoljets$-electroweak regions are defined within a loose `Pre-selection'
-requirement of the basic $Z\rightarrow \ell^+\ell^-$ plus jets plus $\met$
+requirement of the basic $Z\to \ell^+\ell^-$ plus jets plus $\met$
 conditions that we target.
@@
 Diboson backgrounds and C1N2 signals have more in common than just $\met$;
-both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
+both also produce resonant $Z\to \ell^+\ell^-$ pairs
 and do not enhance the rate of heavy-flavour jets, as displayed in
@@
 Here, the standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement
-captures both $W\rightarrow jj$ and $Z\rightarrow jj$ for C1N2 and GMSB models
+captures both $W\to jj$ and $Z\to jj$ for C1N2 and GMSB models
 respectively and is applied similarly to all other on-shell
@@
 One major difference between signals and \diboson\ backgrounds is that signal
-jet pairs mostly arise from boosted $W\rightarrow jj$ decays, which more often
+jet pairs mostly arise from boosted $W\to jj$ decays, which more often
 produce nearby jets.
@@
 via $ZZ$ and $Zh$ bosons pairs.
-Since the $h\rightarrow bb$ branching fraction is large, \srllbb\ finds
+Since the $h\to bb$ branching fraction is large, \srllbb\ finds
 sensitivity by requiring a pair of $b$-tagged jets in a heavy mass window.
@@
 accommodate \atlas' noisy reconstruction of $b$-jets, but also to
-capture $Z\rightarrow bb$ decays, which become significant for signal models
-with small branching fractions $B(\neutralino_1 \rightarrow h \gravitino)$.
+capture $Z\to bb$ decays, which become significant for signal models
+with small branching fractions $B(\neutralino_1 \to h \gravitino)$.
 
@@
 The less extreme signal models targeted here have less boosted
-$W\rightarrow jj$ distributions, but a $\ptjone$ selection is found to be
+$W\to jj$ distributions, but a $\ptjone$ selection is found to be
 effective as an additional discriminator of signals from backgrounds to make
@@
 
-Boosted $Z\rightarrow \ell^+\ell^-$ decays will tend to have small $\rll$.
+Boosted $Z\to \ell^+\ell^-$ decays will tend to have small $\rll$.
 Similarly to $\rjj$ in the High regions, this can usefully select signals,
@@
 Signals trend towards small $\rll$, plausibly due to their more boosted
-$Z\rightarrow \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
+$Z\to \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
 The mass of signals seen at small $\mttwoll$ in (b) motivates the design of
@@
 supersymmetric resonance decays to $Z+\mathrm{invisible}$ and sends both
-the $Z\rightarrow \ell^+\ell^-$ and invisible particle, which becomes
+the $Z\to \ell^+\ell^-$ and invisible particle, which becomes
 $\ptmiss$, in the same direction.
@@
 to define the ISR momentum as the vector sum of all jets not assigned to the
-$W\rightarrow jj$ decay;
+$W\to jj$ decay;
 unlike our other regions, this construction does not select the two hardest for
@@
 These searches consider the same C1N2 $W\kern-0.4exZ$ signal models as us, but target the
-case where the $W$ boson decays leptonically $W\rightarrow\ell\nu$.
+case where the $W$ boson decays leptonically $W\to\ell\nu$.
 For comparison with our off-shell exclusion,
@@
 In comparison to the ISR construction, this design improves yields by not
-needing reconstructed jets for a $W\rightarrow jj$ candidate.
+needing reconstructed jets for a $W\to jj$ candidate.
 
@@
 Part of the light $\mll$ background arises from leptonically decaying
-$Z\rightarrow \tau\tau$ events; these are also removed by requiring large
+$Z\to \tau\tau$ events; these are also removed by requiring large
 $\mttwoll$.
@@
 For GMSB, only SR-$\llbb$ has sensitivity to large
-$B(\neutralino_1 \rightarrow h \tilde{G})$ since it is able to capture the
-resulting $h\rightarrow bb$ decays.
+$B(\neutralino_1 \to h \tilde{G})$ since it is able to capture the
+resulting $h\to bb$ decays.
 At its highest masses, the main SR-High regions retain sensitivity to its
@@
 in FNP events --- important backgrounds producing leptons through
-$Z\rightarrow \ell^\pm\ell^\mp$ must have leptons of the same flavour and
+$Z\to \ell^\pm\ell^\mp$ must have leptons of the same flavour and
 opposite charge, and those are quite accurately reconstructed.
@@
 The largest change was a $2\%$ reduction in the
-$B(\neutralino_1 \rightarrow h \gravitino)$
+$B(\neutralino_1 \to h \gravitino)$
 limit at $m(\neutralino_1) \approx 500\,\eV[G]$.
@@
 largest being a $2\%$ reduction in the limit on
-$B(\neutralino_1 \rightarrow h \gravitino)$ at
+$B(\neutralino_1 \to h \gravitino)$ at
 $m(\neutralino_1) \approx 500\,\eV[G]$.
@@
 Jet mass distribution in a single-lepton selection looking for
-$t\bar t \rightarrow \ell^\pm\nu W(J) \bar bb$
+$t\bar t \to \ell^\pm\nu W(J) \bar bb$
 ]{%
 Jet mass distribution in a single-lepton selection looking for
-$t\bar t \rightarrow \ell^\pm\nu W bb$, where the $W$ boson decays hadronically
+$t\bar t \to \ell^\pm\nu W bb$, where the $W$ boson decays hadronically
 into a single small-radius jet.
@@
 The `$\ttbar$ Merged' sample selects events in which both quarks from the
-$W\rightarrow q\bar q$ decay are within $\Delta R < 0.4$ of the hardest
+$W\to q\bar q$ decay are within $\Delta R < 0.4$ of the hardest
 reconstructed jet.
@@
 its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
-we do not expect large $W\rightarrow jj$ contributions in this \diboson\
+we do not expect large $W\to jj$ contributions in this \diboson\
 process, since one boson is needed for the leptons and the other is needed for
@@
 cannot conserve four-momentum when splitting into two non-collinear massless
-children (for example, gluon emission $q\rightarrow gq$);
+children (for example, gluon emission $q\to gq$);
 each scheme achieves conservation by recoiling off different parts of the
@@
 Although one might expect our selected \diboson\ sample to comprise mostly
-$ZZ$, with one $Z\rightarrow \ell\ell$ and the other invisibly to
-$Z\rightarrow \nu\nu$ to generate our required $Z$-like lepton pair with
+$ZZ$, with one $Z\to \ell\ell$ and the other invisibly to
+$Z\to \nu\nu$ to generate our required $Z$-like lepton pair with
 $\met$.
@@
 Those $W\kern-0.4exZ$ events must therefore have a missing third lepton from
-$W\rightarrow \ell\nu$ decays.
+$W\to \ell\nu$ decays.
 To avoid selection, that third lepton could be softer than the few $\eV[G]$

### 2022-09-10 main/chapter_searches.tex
change rightarrow to to
add np lemma stuff

@@
 priors as they wish.
-For model comparison, for example between background $h=B$ and signal $h=S$
-models, the analysis is conveniently linearized by expressing it in $\log$
+For model comparison, for example between background $h_1=B$ and signal $h_2=S$
+models, the analysis is conveniently linearised by expressing it in $\log$
 ratio form:
-\TODO{different choice for phi to avoid angle?}
 \begin{equation}
@@
 --- as just a normalizing constant~\cite{
+Neyman1937Outline,
 gelman1995bayesian,
@@
 \item Use those data to update your state
-$f(d_{i-1}, \ldots) \rightarrow f(d_i, d_{i-1}, \ldots)$.
+$f(d_{i-1}, \ldots) \to f(d_i, d_{i-1}, \ldots)$.
 Unlike ideal posterior updating, this encodes only a function of the data, not
@@
 A frequency, however, would only be permitted if the envelope had been packaged
-randomly.
+at `random'.
 Was the coin first tossed to randomize its face, or placed deliberately?
 You do not know.
+In Neyman's words,
+``I have put the word ``~random~'' in inverted commas because it is very
+difficult to define what is meant by it in practice''~\cite{
+Neyman1937Outline
+}.
 
@@
 kolomogoroff1950translated,
+Neyman1937Outline,
 axioms1010038
@@
 These criteria do not include \emph{unbiasedness}, which states that the mean
-value of an estimate (over the distribution for alternative data) should equal
-its known true value~\cite{pdg2020review}.
+value of an estimate --- over the distribution for alternative data --- should
+equal its known true value~\cite{
+sheynin1989aa,
+Neyman1937Outline,
+pdg2020review
+}.
 Defining bias as a difference from a mean fails to be coordinate invariant,
@@
 \begin{equation}
+\label{eqn:searches_t_statistic}
 t(d) = \frac{\bar{\mu}(d) - \mu}{\bar{\sigma}(d)}\sqrt{n}
@@
 }
-\label{fig:atlas_higgs_4l}
+\label{fig:searches_atlas_higgs_4l}
 \end{figure}
@@
 Consider for example the distributions of signal (Higgs boson) and background
-(mostly $ZZ^*$) samples in Figure~\ref{fig:atlas_higgs_4l}.
+(mostly $ZZ^*$) samples in Figure~\ref{fig:searches_atlas_higgs_4l}.
 If an event is observed at $m_{4\ell} = 125\,\eV[G]$, is it signal or
@@
 the nature of the experiment.
-Blindly followed, these rules mean that the data and test statistic could be
+
+Obediently followed, these rules mean that the data and test statistic could be
 anything at all and still incur suspicion for unrelated hypotheses with the
-same frequency.
+same prior probability.
 Your wall clock reads not less than $48$ minutes past the hour?
 The Standard Model is suspect.
-This peculiarity is illustrated at the wildly popular $\alpha = 0.05$ level by
-Figure~\ref{fig:searches_significant}.
+Figure~\ref{fig:searches_significant} illustrates this peculiarity and the
+associated difficulty in communicating the results.
 
@@
 another~\cite{neymanpearson1928max, neymanpearson1933lemma}.
-Inspired by the same Bayesian arguments that we used to suggest likelihood
-ratios for model comparison~\cite{neymanpearson1928max}, they considered test
-statistics based on likelihoods\TODO{}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/searches_np_curves_crop.png}
-\caption[
-Tail areas of probability distributions of data from two alternative
-hypotheses, reproduced from Neyman and Pearson
-]{%
-Tail areas of probability distributions of data from two alternative
-hypotheses, reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
-}
-\label{fig:searches_np_tails}
-\end{figure}
-
-Figure~\ref{fig:searches_np_tails}
-
-\TODO{extend}
-``for which hypothesis A is more probable and those for which it is less
-probable (sic!)''
-to which Lehmann appends ``(sic!)''~\cite{lehmann2011fisher}.
-
-\TODO{consider changing our A, B to A1 A2}
-\begin{displayquote}
-``%
-Then the two corresponding densities at the sample point are
-(A1) $6.42 \times 10^{-20}$,
-(A2) $5.18 \times 10^{-4}$ and the ratio of the likelihood of Hypothesis
-A1 to that of A2 is $1.24 \times 10^{-16}$.
-This ratio confirms the common-sense judgment that A2 is far more plausible
-than A1''
-''~\cite{neymanpearson1933lemma}
-\end{displayquote}
-
+Rather than a $p$-value arousing suspicion of an hypothesis, here it is used
+to decide on actions of `rejecting' or `accepting' an hypothesis.
+From the Bayesian arguments that we presenting in
+Section~\ref{sec:searches_data_analysis},
+Neyman and Pearson argued for the interest in likelihood ratios, and further
+considered their use as test statistics~\cite{neymanpearson1928max}.
+On one hand they argue for the practical usefulness of maximum likelihood
+ratios
+\begin{equation}
+\Lambda =
+\frac{L(h_1)}{\mathrm{max}_i\,L(h_i)}
+\end{equation}
+for a specified hypothesis $h_1$ against the largest likelihood from a set
+of alternatives.
+As descriptions of the likelihoods, maximum likelihood ratios can usefully
+communicate relevant information from the data.
+As test statistics, they also win generality by avoid context-specific
+motivations, such as scale invariance for the $t$-statistic of
+Equation~\ref{eqn:searches_t_statistic} --- any situation with predictive
+models can construct likelihoods and maximum likelihood ratios.
+``Without claiming that this method is necessarily the ``best'' to adopt,''
+Neyman and Pearson suggest maximum likelihood ratios as
+``one clearly defined method of discriminating between samples for which
+Hypothesis A is more probable and those for which it is less probable''~\cite{
+neymanpearson1933lemma}.
+To this quotation, Lehmann adds ``(sic!)''~\cite{lehmann2011fisher}, since it
+appears to conflate the likelihood for an hypothesis with its probability.
+
+But Neyman and Pearson's association of likelihood ratios with
+``more probable'' hypotheses may not be an error.
+They authors understood the Bayesian algebra of how large likelihood ratios
+make hypotheses \emph{more probable}, independent of whatever (non-zero) prior
+probabilities they are assigned.
+They state later in the same work that
 \begin{displayquote}
@@
 \end{displayquote}
+And elsewhere in an example with two hypotheses they state that that
+\begin{displayquote}
+``%
+the two corresponding densities at the sample point are
+($\mathrm{A}_1$) $6.42 \times 10^{-20}$,
+($\mathrm{A}_2$) $5.18 \times 10^{-4}$ and the ratio of the likelihood of
+Hypothesis $\mathrm{A}_1$ to that of $\mathrm{A}_2$ is $1.24 \times 10^{-16}$.
+This ratio confirms the common-sense judgment that $\mathrm{A}_2$ is far more
+plausible than $\mathrm{A}_1$%
+''~\cite{neymanpearson1928max},
+\end{displayquote}
+again relating the likelihood ratio to how ``plausible'' an hypothesis is.
+
+The famous Neyman-Pearson lemma does not prove this usefulness of likelihood
+ratios in probability algebra.
+That was already known.
+What it contributes is a utility for likelihood ratio test statistics
+in one-sided selections --- it states that likelihood ratios
+(not maximum likelihood ratios)
+are optimal in a specific sense within their framework of decisions
+based on small $p$-values.
+Precisely, it states that at a fixed probability $\alpha$ of rejecting a true
+hypothesis, a likelihood ratio test statistic maximizes the probability
+$\alpha'$ of rejecting the alternative, false hypothesis that lives in the
+denominator of that ratio.
+They illustrate this construction with a plot reproduced in
+Figure~\ref{fig:searches_np_tails}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/searches_np_curves_crop.png}
+\caption[
+Tail areas of prior probability distributions for data from two alternative
+hypotheses, reproduced from Neyman and Pearson
+]{%
+Tail areas of prior probability distributions for data from two alternative
+hypotheses, reproduced from Neyman and Pearson~\cite{neymanpearson1933lemma}.
+}
+\label{fig:searches_np_tails}
+\end{figure}
+
+Although the likelihood ratio local to the observed data is all that matters
+for the probabilistic analysis, this lemma is relevant to the construction
+of useful event variables for analysing our data.
+Figures~\ref{fig:searches_sig_bkg_prior_likelihood}
+and~\ref{fig:searches_atlas_higgs_4l} \TODO{}
+
+\TODO{Confidence intervals~\cite{
+clopper1934confidence,
+neyman1935Intervals,
+Neyman1937Outline
+}}
 
 \TODO{
-\^ Another posterior argument
-Introduce maximum likelihood ratio as interesting without claiming it to be
-necessarily best.
-Neyman-Pearson lemma --- reproduce Fig 5?
-~\cite{neymanpearson1933lemma}
-}
+discrete failure to match --- randomization.
+table of random numbers --- reputable table!
+K. D. Tocher (1950)
+``extraneous randomization''
+~\cite{tocher1950discontinuous}
+}
+
+\begin{displayquote}
+\small
+``%
+There is something horrifying in the ideological movement represented by the
+doctrine that reasoning, properly speaking, cannot be applied to empirical data
+to lead to inferences valid in the real world.%
+''~\cite{fisher1956statistical}
+\end{displayquote}
+
+% p-values
+no!~\cite{
+schervish1996p,
+Cohen1994TheEI,
+Amrhein2017TheEI,
+goodman2008dirty,
+greenland2016no,
+Bernstein2016princess,
+Trafimow2015ban,
+Read2002cls
+}
+\TODO{examples of use and misuse in hep}
 
@@
 }
-
-\TODO{Confidence intervals~\cite{neyman1935Intervals, Neyman1937Outline}}
-
-\TODO{
-discrete failure to match --- randomization.
-table of random numbers --- reputable table!
-K. D. Tocher (1950)
-``extraneous randomization''
-~\cite{tocher1950discontinuous}
-}
-
-\begin{displayquote}
-\small
-``%
-There is something horrifying in the ideological movement represented by the
-doctrine that reasoning, properly speaking, cannot be applied to empirical data
-to lead to inferences valid in the real world.%
-''~\cite{fisher1956statistical}
-\end{displayquote}
-
-% p-values
-no!~\cite{
-schervish1996p,
-Cohen1994TheEI,
-Amrhein2017TheEI,
-goodman2008dirty,
-greenland2016no,
-Bernstein2016princess,
-Trafimow2015ban,
-Read2002cls
-}
-\TODO{examples of use and misuse in hep}
 
@@
 \begin{equation}
-H_{\!f}(\vec \theta) \rightarrow (\vec \mu, \log f)
+H_{\!f}(\vec \theta) \to (\vec \mu, \log f)
 \end{equation}

### 2022-09-10 main/chapter_searches.tex
write up confidence intervals

@@
 Supposing we have this one event with $\met > 100\,\eV[G]$, did it come from
-the background or signal sample?
+the background or signal sample?%
+\footnote{%
+Questioning the origin of individual events is unusual in our context in
+\atlas, but is done in other research into rare effects such as neutrino
+interactions and black hole mergers~\cite{
+opera2010event,
+icecube2020event,
+Abbott2021event
+}.%
+}
 The answer depends on the signal and background cross-sections in our collider,
@@
 Frequentist theory confines itself to the special case where those measures
-are numbers of discrete randomly occurring events, such that a frequentist
+are numbers of randomly occurring events, such that a frequentist
 probability is a ratio of frequencies:
@@
 18th century ideas~\cite{bayes1763lii, laplace1774stigler}.
-Jaynes instead names them ``orthodox''~\cite{jaynes2003probability},
+Jaynes instead names them ``orthodox''~\cite{
+Jaynes1976intervals,
+jaynes2003probability
+},
 which may have been accurate in his environment, but different
@@
 to compare how well different hypotheses predict it.
-For reasons I cannot explain, Fisher advocates a different use ---
+For unclear reasons, Fisher advocates a different use ---
 rather than considering likelihoods and likelihood ratios, he recommends
@@
 For example, in Fisher's suggestion of $P(h) \in 0.1--0.9$ being not suspect,
-then $\alpha = 0.2$ (if the distribution for is continuous or
+then $\alpha = 0.2$.%
+\footnote{%
+Exactly matching $\alpha$ is impossible for discrete distributions when
+the probability masses do not add up to $\alpha$.
+Neyman accommodates this by allowing inequalities with $\alpha$~\cite{
+neyman1935Intervals
+},
+and other approaches introduce ``extraneous randomization'' to
+break up discrete distributions~\cite{tocher1950discontinuous}.
+This problem of inexact `coverage' continues to cause controversy~\cite{
+Read2002cls,
+Feldman:1997qc,
+cousins2008evaluation,
+Cranmer2006Statistical
+}.%
+}
 otherwise meets the edges at $0.1$ and $0.9$ in its cumulative distribution).
@@
 Neyman and Pearson argued for the interest in likelihood ratios, and further
-considered their use as test statistics~\cite{neymanpearson1928max}.
-On one hand they argue for the practical usefulness of maximum likelihood
-ratios
+considered their use as test statistics~\cite{neymanpearson1928max};
+they also considered \emph{maximum} likelihood ratios
 \begin{equation}
@@
 of alternatives.
-As descriptions of the likelihoods, maximum likelihood ratios can usefully
-communicate relevant information from the data.
+As descriptions of the likelihoods in play, maximum likelihood ratios can
+usefully communicate relevant information from the data.
 As test statistics, they also win generality by avoid context-specific
@@
 based on small $p$-values.
-Precisely, it states that at a fixed probability $\alpha$ of rejecting a true
-hypothesis, a likelihood ratio test statistic maximizes the probability
+Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis, a
+likelihood ratio test statistic maximizes the probability
 $\alpha'$ of rejecting the alternative, false hypothesis that lives in the
-denominator of that ratio.
+denominator of the ratio.
 They illustrate this construction with a plot reproduced in
@@
 
-Although the likelihood ratio local to the observed data is all that matters
-for the probabilistic analysis, this lemma is relevant to the construction
-of useful event variables for analysing our data.
+Only the likelihood ratio local to the observed data matters to the results of
+a probabilistic analysis, but this lemma is relevant to the construction
+of useful event variables when designing an analysis procedure.
+This is seen from the signal and background distributions of
 Figures~\ref{fig:searches_sig_bkg_prior_likelihood}
-and~\ref{fig:searches_atlas_higgs_4l} \TODO{}
-
-\TODO{Confidence intervals~\cite{
+and~\ref{fig:searches_atlas_higgs_4l};
+the event variable in Figure~\ref{fig:searches_sig_bkg_prior_likelihood} is
+like a likelihood ratio in that any right-handed selection catches a good
+signal purity relative to its background yield, but the signal bump in the
+middle of the background distribution in
+Figure~\ref{fig:searches_atlas_higgs_4l} would be better selected with a
+window around the Higgs boson mass.
+Re-ordering $m_{4\ell}$ by a likelihood ratio could turn this window into an
+equivalent and possibly simpler one-sided selection.
+
+Neyman and Pearson were also involved in the development of
+`confidence intervals', which are ranges of parameters that are accepted by the
+$p$-value decision rules~\cite{
 clopper1934confidence,
@@
 Neyman1937Outline
-}}
-
-\TODO{
-discrete failure to match --- randomization.
-table of random numbers --- reputable table!
-K. D. Tocher (1950)
-``extraneous randomization''
-~\cite{tocher1950discontinuous}
-}
-
+}.
+A parameter corresponds to an hypothesis which assigns a prior distribution over
+data, and those data for which an obeisant researcher accepts the hypothesis
+have probabilities that sum to $1 - \alpha$.
+This rule selects a set of favoured data for each hypothesis, and the
+confidence interval (when it can be constructed) contains hypotheses for which
+the observed data are in that favoured set.
+
+Confidence intervals are confusing.
+It is easy to be tricked into thinking that the confidence interval for given
+data `probably' contains the unknown parameter, but that is backwards.
+For a known true parameter $x$, the confidence intervals in random repeated
+experiments will contain that $x$ with probability $1 - \alpha$ ---
+the interval makes a subtle statement about unknown alternative data for
+known true parameters.
+For given data, the confidence interval can in fact exclude all possible
+parameters~\cite{
+pratt1961testing,
+Jaynes1976intervals
+}.
+The correct interpretation of a confidence interval is explained in an example
+by Pratt:
+\begin{displayquote}
+\small
+``%
+A method yielding true statements with probability $.95$, when applied to your
+experiment, yields the statement that your treatment effect is between $17$ and
+$29$, but no conclusion is possible about how probable it is that your
+treatment effect is between $17$ an $29$.%
+''~\cite{pratt1961testing}.
+\end{displayquote}
+Neyman and Pearson never claimed otherwise; they did precise mathematics, and
+confusion about their intervals arises externally.
+
+Although the confidence interval is a function of the data, so can be used
+for vague inferences, the recommendation by Fisher to communicate the
+likelihood function remains simpler and clearer.
+Fisher deserves the last word here also:
 \begin{displayquote}
@@
 % p-values
-no!~\cite{
+\TODO{no!}~\cite{
 schervish1996p,

### 2022-09-10 main/chapter_searches.tex
comma before fisher likelihood quote

@@
 (specifically confidence intervals, to be introduced in
-Section~\ref{sec:searches_np})
+Section~\ref{sec:searches_np}),
 \begin{displayquote}

### 2022-09-10 main/chapter_searches.tex
add another p-value criticism

@@
 Trafimow2015ban,
+wagenmakers2007practical,
 Read2002cls

### 2022-09-11 main/chapter_searches.tex
move some drafts to next section; some text tweaks

@@
 \footnote{%
-Exactly matching $\alpha$ is impossible for discrete distributions when
-the probability masses do not add up to $\alpha$.
+Matching $\alpha$ can be impossible in discrete distributions if
+the tail probability masses do not add exactly to $\alpha$.
 Neyman accommodates this by allowing inequalities with $\alpha$~\cite{
 neyman1935Intervals
-},
-and other approaches introduce ``extraneous randomization'' to
+}.
+Other approaches introduce ``extraneous randomization'' to
 break up discrete distributions~\cite{tocher1950discontinuous}.
@@
 of alternatives.
+
 As descriptions of the likelihoods in play, maximum likelihood ratios can
 usefully communicate relevant information from the data.
-As test statistics, they also win generality by avoid context-specific
+And ass test statistics, they also win generality by avoid context-specific
 motivations, such as scale invariance for the $t$-statistic of
@@
 neymanpearson1933lemma}.
-To this quotation, Lehmann adds ``(sic!)''~\cite{lehmann2011fisher}, since it
+To this quote, Lehmann appends ``(sic!)''~\cite{lehmann2011fisher}, since it
 appears to conflate the likelihood for an hypothesis with its probability.
-
 But Neyman and Pearson's association of likelihood ratios with
 ``more probable'' hypotheses may not be an error.
-They authors understood the Bayesian algebra of how large likelihood ratios
-make hypotheses \emph{more probable}, independent of whatever (non-zero) prior
-probabilities they are assigned.
-They state later in the same work that
+The authors understood the algebra of how large likelihood ratios make
+hypotheses \emph{more probable} by changing their posterior probabilities,
+independent of whatever (non-zero) prior probabilities they are assigned.
+Indeed they later state that
 \begin{displayquote}
@@
 \end{displayquote}
-And elsewhere in an example with two hypotheses they state that that
-\begin{displayquote}
+Elsewhere, they also consider an example with two hypotheses $\mathrm{A}_1$
+and $\mathrm{A}_2$, and with some data they find a $\log$ likelihood ratio of
+% the two corresponding densities at the sample point are
+% ($\mathrm{A}_1$) $6.42 \times 10^{-20}$,
+% ($\mathrm{A}_2$) $5.18 \times 10^{-4}$ and the ratio of the likelihood of
+% Hypothesis $\mathrm{A}_1$ to that of $\mathrm{A}_2$ is $1.24 \times 10^{-16}$.
+$\Delta = -37$, concluding that
 ``%
-the two corresponding densities at the sample point are
-($\mathrm{A}_1$) $6.42 \times 10^{-20}$,
-($\mathrm{A}_2$) $5.18 \times 10^{-4}$ and the ratio of the likelihood of
-Hypothesis $\mathrm{A}_1$ to that of $\mathrm{A}_2$ is $1.24 \times 10^{-16}$.
 This ratio confirms the common-sense judgment that $\mathrm{A}_2$ is far more
@@
 ''~\cite{neymanpearson1928max},
-\end{displayquote}
-again relating the likelihood ratio to how ``plausible'' an hypothesis is.
-
-The famous Neyman-Pearson lemma does not prove this usefulness of likelihood
-ratios in probability algebra.
-That was already known.
-What it contributes is a utility for likelihood ratio test statistics
-in one-sided selections --- it states that likelihood ratios
-(not maximum likelihood ratios)
-are optimal in a specific sense within their framework of decisions
-based on small $p$-values.
-Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis, a
-likelihood ratio test statistic maximizes the probability
-$\alpha'$ of rejecting the alternative, false hypothesis that lives in the
-denominator of the ratio.
-They illustrate this construction with a plot reproduced in
-Figure~\ref{fig:searches_np_tails}.
+which again relates the likelihood ratio to how plausible an hypothesis is.
 
@@
 
-Only the likelihood ratio local to the observed data matters to the results of
-a probabilistic analysis, but this lemma is relevant to the construction
-of useful event variables when designing an analysis procedure.
-This is seen from the signal and background distributions of
+The famous Neyman-Pearson lemma does not prove the usefulness of likelihood
+ratios in probability algebra.
+That was already known.
+What it contributes is a utility for likelihood ratio test statistics
+in one-sided inequalities --- it states that likelihood ratios
+(not maximum likelihood ratios)
+are optimal in a specific sense within their framework of decisions
+based on $p$-values.
+Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis, a
+likelihood ratio test statistic maximizes the probability
+$\alpha'$ of rejecting the alternative, false hypothesis that lives in the
+denominator of the ratio; they illustrate this construction with the diagram
+reproduced in Figure~\ref{fig:searches_np_tails}.
+
+Only the likelihood ratio local to the observed data matters directly in the
+probabilistic analysis, but this lemma is useful in the construction
+of event variables of the kind we use in \atlas\ searches.
+This can be seen from the signal and background distributions of
 Figures~\ref{fig:searches_sig_bkg_prior_likelihood}
@@
 the event variable in Figure~\ref{fig:searches_sig_bkg_prior_likelihood} is
-like a likelihood ratio in that any right-handed selection catches a good
+like a likelihood ratio in that any right-handed inequality selects a good
 signal purity relative to its background yield, but the signal bump in the
@@
 window around the Higgs boson mass.
-Re-ordering $m_{4\ell}$ by a likelihood ratio could turn this window into an
-equivalent and possibly simpler one-sided selection.
+Re-ordering $m_{4\ell}$ by a likelihood ratio could turn this selection window
+into an equivalent and simpler one-sided selection ---
+the Neyman-Pearson lemma states that there is no better ordering for optimizing
+the signal-to-background ratio available from a one-sided selection.
 
@@
 For a known true parameter $x$, the confidence intervals in random repeated
-experiments will contain that $x$ with probability $1 - \alpha$ ---
+experiments will contain that $x$ with probability $1 - \alpha$;
 the interval makes a subtle statement about unknown alternative data for
 known true parameters.
-For given data, the confidence interval can in fact exclude all possible
-parameters~\cite{
+Given confidence intervals can even exclude all plausible parameters~\cite{
 pratt1961testing,
@@
 Neyman and Pearson never claimed otherwise; they did precise mathematics, and
-confusion about their intervals arises externally.
-
-Although the confidence interval is a function of the data, so can be used
-for vague inferences, the recommendation by Fisher to communicate the
-likelihood function remains simpler and clearer.
+confusion about their intervals arises externally~\cite{jaynes2003probability}.
+
+Confidence intervals are functions of data, so can inform vague inferences if
+used as data for likelihood functions, as we did to interpret Fisher's ratio
+$p$-value ratio as a likelihood ratio.
+And again this cannot add to the data, it can only detract.
+Fisher's recommendation to communicate the data and likelihood function
+directly remains simpler and clearer.
 Fisher deserves the last word here also:
@@
 \end{displayquote}
-
-% p-values
-\TODO{no!}~\cite{
-schervish1996p,
-Cohen1994TheEI,
-Amrhein2017TheEI,
-goodman2008dirty,
-greenland2016no,
-Bernstein2016princess,
-Trafimow2015ban,
-wagenmakers2007practical,
-Read2002cls
-}
-\TODO{examples of use and misuse in hep}
-
-\TODO{p-value:
-The chance of a small $p$-value is small.
-This is true, but not interesting because the value of any $p$-value is small.
-By construction $p$-values are uniformly distributed on the unit interval
-under a given hypothesis.
-The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
-$0.6 < p < 0.65$.
-The frequency of $p < \varepsilon$ is identical to the frequency
-of $0.6 < p < 0.6 + \varepsilon$.
-An appropriate choice of test statistic is therefore necessary for the
-$p$-value to have meaning.
-}
 
@@
 \end{singlespacing}
+
+% p-values
+\TODO{no!}~\cite{
+schervish1996p,
+Cohen1994TheEI,
+Amrhein2017TheEI,
+goodman2008dirty,
+greenland2016no,
+Bernstein2016princess,
+Trafimow2015ban,
+wagenmakers2007practical,
+Read2002cls
+}
+\TODO{examples of use and misuse in hep}
+
+\TODO{p-value:
+The chance of a small $p$-value is small.
+This is true, but not interesting because the value of any $p$-value is small.
+By construction $p$-values are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+$p$-value to have meaning.
+}
 

### 2022-09-12 main/chapter_searches.tex
add confidence interval likelihood; clean up some text from yesterday

@@
 That was already known.
-What it contributes is a utility for likelihood ratio test statistics
-in one-sided inequalities --- it states that likelihood ratios
-(not maximum likelihood ratios)
+What it contributes is a relationship with one-sided inequalities ---
+it states that likelihood ratios (not maximum likelihood ratios)
 are optimal in a specific sense within their framework of decisions
 based on $p$-values.
-Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis, a
-likelihood ratio test statistic maximizes the probability
-$\alpha'$ of rejecting the alternative, false hypothesis that lives in the
-denominator of the ratio; they illustrate this construction with the diagram
-reproduced in Figure~\ref{fig:searches_np_tails}.
+Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis,
+the Neyman-Pearson lemma finds that a likelihood ratio test statistic maximizes
+the probability $\alpha'$ of rejecting the alternative, false hypothesis that
+lives in the denominator of the ratio; they illustrate this construction with
+the diagram reproduced in Figure~\ref{fig:searches_np_tails}.
 
@@
 Re-ordering $m_{4\ell}$ by a likelihood ratio could turn this selection window
-into an equivalent and simpler one-sided selection ---
-the Neyman-Pearson lemma states that there is no better ordering for optimizing
-the signal-to-background ratio available from a one-sided selection.
+into an equivalent and simpler one-sided selection, and the Neyman-Pearson
+lemma states that there is no ordering that further optimizes the selected
+signal-to-background ratio.
 
@@
 It is easy to be tricked into thinking that the confidence interval for given
-data `probably' contains the unknown parameter, but that is backwards.
+data `probably' contains the unknown parameter, but that gets it backwards.
 For a known true parameter $x$, the confidence intervals in random repeated
@@
 }.
-The correct interpretation of a confidence interval is explained in an example
-by Pratt:
+Pratt eloquently explains the correct interpretation of a confidence interval
+by example:
 \begin{displayquote}
@@
 
-Confidence intervals are functions of data, so can inform vague inferences if
-used as data for likelihood functions, as we did to interpret Fisher's ratio
-$p$-value ratio as a likelihood ratio.
-And again this cannot add to the data, it can only detract.
-Fisher's recommendation to communicate the data and likelihood function
-directly remains simpler and clearer.
-Fisher deserves the last word here also:
+Confidence intervals are functions of the data, so can inform vague inferences
+if used as data for likelihood functions, just as we did to interpret Fisher's
+$p$-value ratio as a likelihood ratio ---
+an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
+likelihood function
+\begin{equation}
+L(x) =
+\left\{
+\begin{matrix}
+1 - \alpha & \textrm{if}~x \in x_1\textrm{--}x_2, \\
+\alpha & \textrm{otherwise.} \\
+\end{matrix}
+\right.
+\end{equation}
+And again, this reduction cannot add to the data.
+It can only detract.
+The recommendation by Fisher to directly communicate the likelihood function
+remains simpler and clearer.
+In response to the idea that no conclusion can be drawn about the parameter
+iself, Fisher also deserves the last word:
 \begin{displayquote}

### 2022-09-12 main/chapter_searches.tex
add some references on particle stats history (cls, cl confusion)

@@
 In modern jargon, a `probable error' at the observed test statistic
-$\tilde{t}$ and an hypothesis $h$ is a \textbf{$p$-value}
+$\tilde{t}$ and an hypothesis $h$ is a \textbf{\pvalue}
 \begin{align}
@@
 \end{align}
-the $p$-value $P(h)$ is the prior probability of data
+the \pvalue\ $P(h)$ is the prior probability of data
 \emph{greater or equal to} the observed data when ordered by the value of the
 test statistic.
-Equation~\ref{eqn:searches_pvalue_likelihood} points out that the $p$-value is
+Equation~\ref{eqn:searches_pvalue_likelihood} points out that the \pvalue\ is
 just is the likelihood plus some fluff from a tail of larger test statistics.
 
-Fisher illustrates the use of $p$-values in the second example
+Fisher illustrates the use of \pvalues\ in the second example
 of~\cite{fisher1921probable}, which models some data with a parameter $r$ and
@@
 of parameters, but their likelihoods in our definition.
-But this factor of $1000$ is a $p$-value ratio, not a likelihood ratio,
+But this factor of $1000$ is a \pvalue\ ratio, not a likelihood ratio,
 so how can this make sense?
@@
 Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
-shows that the $p$-value is a likelihood only if the result is reduced from
+shows that the \pvalue\ is a likelihood only if the result is reduced from
 $A: t =    \tilde{t}$ to
@@
 
-Fisher's $p$-value ratio of $1000$ is, therefore, a $\log$ likelihood
+Fisher's \pvalue\ ratio of $1000$ is, therefore, a $\log$ likelihood
 ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately strong
@@
 observed data.
-This example illustrates $p$-values pathological dependency on tail selections,
+This example illustrates \pvalues\ pathological dependency on tail selections,
 and the benefits of likelihood ratios.
@@
 
-Fisher claims that the $p$-value indicates the ``significance'' of an
+Fisher claims that the \pvalue\ indicates the ``significance'' of an
 observation~\cite{fisher1925smrw}, with small or large values being
@@
 \label{sec:searches_np}
-Suspecting an hypothesis when its $p$-value lands below (or above) a threshold
+Suspecting an hypothesis when its \pvalue\ lands below (or above) a threshold
 means that there is a prior probability $\alpha$ that any hypothesis will be
@@
 
-Neyman and Pearson formalized this $p$-value based suspicion into decision
+Neyman and Pearson formalized this \pvalue\ based suspicion into decision
 rules and thought about what it is that makes one test statistic better than
 another~\cite{neymanpearson1928max, neymanpearson1933lemma}.
-Rather than a $p$-value arousing suspicion of an hypothesis, here it is used
+Rather than a \pvalue\ arousing suspicion of an hypothesis, here it is used
 to decide on actions of `rejecting' or `accepting' an hypothesis.
@@
 \end{displayquote}
-Elsewhere, they also consider an example with two hypotheses $\mathrm{A}_1$
-and $\mathrm{A}_2$, and with some data they find a $\log$ likelihood ratio of
+Elsewhere also, they consider an example with two hypotheses $\mathrm{A}_1$
+and $\mathrm{A}_2$, and find a $\log$ likelihood ratio of
 % the two corresponding densities at the sample point are
@@
 % Hypothesis $\mathrm{A}_1$ to that of $\mathrm{A}_2$ is $1.24 \times 10^{-16}$.
-$\Delta = -37$, concluding that
+$-37$, concluding that
 ``%
@@
 `confidence intervals', which are ranges of parameters that are accepted by the
-$p$-value decision rules~\cite{
+\pvalue\ decision rules~\cite{
 clopper1934confidence,
@@
 data, and those data for which an obeisant researcher accepts the hypothesis
-have probabilities that sum to $1 - \alpha$.
-This rule selects a set of favoured data for each hypothesis, and the
+have probabilities that sum to $\gamma = 1 - \alpha$.
+This $\gamma$ is known as the `confidence level'.
+Such a rule selects a set of favoured data for each hypothesis, and the
 confidence interval (when it can be constructed) contains hypotheses for which
@@
 if used as data for likelihood functions, just as we did to interpret Fisher's
-$p$-value ratio as a likelihood ratio ---
+\pvalue\ ratio as a likelihood ratio ---
 an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
@@
 
-\TODO{p-value:
-The chance of a small $p$-value is small.
-This is true, but not interesting because the value of any $p$-value is small.
-By construction $p$-values are uniformly distributed on the unit interval
+\TODO{p-value footnote:
+The chance of a small \pvalue\ is small.
+This is true, but not interesting because the value of any \pvalue\ is small.
+By construction \pvalues\ are uniformly distributed on the unit interval
 under a given hypothesis.
@@
 An appropriate choice of test statistic is therefore necessary for the
-$p$-value to have meaning.
-}
-
-$\cls$~\cite{Read:451614, Read2002cls, Murray2010heretic}
+\pvalue\ to have meaning.
+}
 
@@
 
-\section{Searches}
+$\cls$~\cite{
+zech1988cls,
+Bock:353201,
+Read:451614,
+Read2002cls,
+Murray2010heretic
+}
+\begin{equation}
+\cls
+=
+\frac{P(s + b)}{P(b)}
+=
+\frac{
+\prob{t \geq \tilde{t}}{s + b}
+}{
+\prob{t \geq \tilde{t}}{b}
+}
+.
+\end{equation}
+These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
+Some literature also defines $\clb = 1 - p_b$~\cite{
+Cowan:2010js
+},
+where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
+statistic.
+
+textbooks cited by pdg~\cite{
+eadie1971statistical,
+frodesen1979probability
+}
+
+\subsection{Searches}
 \label{sec:searches_searches}
@@
 \histfitter~\cite{Besjes_2015,baak2015histfitter}
-
-\clearpage
-
-Hello again

### 2022-09-13 main/chapter_2ljets.tex
update to pdg 2022; add older pdg references for CL, CLs

@@
 Values with errors are rounded according to the PDG rounding
-scheme~\cite{pdg2020review}.%
+scheme~\cite{pdg2022ynf}.%
 }
@@
 Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to
-$m(Z)\approx 91.2\,\eV[G]$~\cite{pdg2020review},
+$m(Z)\approx 91.2\,\eV[G]$~\cite{pdg2022ynf},
 and the others are binned in $\mll$ for sensitivity to the various

### 2022-09-13 main/chapter_searches.tex
update to pdg 2022; add older pdg references for CL, CLs

@@
 cowan1998statistical,
-pdg2020review
+pdg2022ynf
 }.
@@
 Neyman1937Outline,
-pdg2020review
+pdg2022ynf
 }.
@@
 where the unique unbiased estimate is $(-1)^{n + 1}$.
-Section~17.3 of~\cite{jaynes2003probability} shows similar nonsense for
+Section~17.3 of~\cite{jaynes2003probability} finds similar nonsense for
 unbiased estimates of $\mu^2$ and $\exp(-\mu)$.%
@@
 Remaining uses of unbiasedness in High Energy Physics~\cite{
-pdg2020review,
+pdg2022ynf,
 Tullythesis,
@@
 Neyman and Pearson were also involved in the development of
-`confidence intervals', which are ranges of parameters that are accepted by the
-\pvalue\ decision rules~\cite{
+`confidence intervals', which are essentially ranges of parameters that are
+accepted by the \pvalue\ decision rules~\cite{
 clopper1934confidence,
@@
 Neyman1937Outline
-}.
-A parameter corresponds to an hypothesis which assigns a prior distribution over
-data, and those data for which an obeisant researcher accepts the hypothesis
-have probabilities that sum to $\gamma = 1 - \alpha$.
+}.%
+\footnote{%
+Confidence are usually introduced by endpoints containing constrained
+probability mass, but the arbitrary choice of test statistic can always reduce
+that to this \pvalue\ form.%
+}
+A parameter corresponds to an hypothesis that assigns a prior distribution to
+plausible data, and those data for which an obeisant researcher accepts the
+hypothesis have probabilities that sum to $\gamma = 1 - \alpha$.
 This $\gamma$ is known as the `confidence level'.
@@
 Some literature also defines $\clb = 1 - p_b$~\cite{
-Cowan:2010js
+Cowan:2010js,
+pdg2022ynf
 },
@@
 statistic.
+
+\TODO{CL \pvalue~\cite{pdg1984}}
+\TODO{Helene ~\cite{pdg1988}}
 

### 2022-09-13 main/chapter_searches.tex
tidy up some citations; add CLs for Higgs searches at LEP

@@
 estimate of the standard deviation $\bar{\sigma}$ --- the square of the mean
-is not the mean of the squares~\cite{Barlow:2019svl}.
+is not the mean of the squares~\cite{barlow2019svl}.
 Fisher rejected the bias criterion for this
@@
 zech1988cls,
-Bock:353201,
-Read:451614,
+bock1998lower,
+read1997optimal,
 Read2002cls,
@@
 
-\TODO{CL \pvalue~\cite{pdg1984}}
-\TODO{Helene ~\cite{pdg1988}}
+\TODO{CL \pvalue~\cite{pdg1984, murray1997use}}
+\TODO{Helene~\cite{pdg1988}}
+\TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
 

### 2022-09-13 main/chapter_searches.tex
add bad p-value takes; use commands for atlas and lhcb

@@
 
+probability of compatibility~\cite{
+HIGG-2018-04,
+HIGG-2018-27,
+IceCube2013search,
+lhcb2021test,
+cern2021test
+}
+
+mix $p$-value with posterior tail~\cite{lhcb2017test}
+
 \subsection{Searches}
@@
 \histfitter~\cite{Besjes_2015,baak2015histfitter}
+
+lhcb2021test article: compatible with the SM prediction with a P value of 0.10\%
+cern:  probability of around 0.1\% that the data is compatible with the Standard Model predictions
+https://home.cern/news/news/physics/intriguing-new-result-lhcb-experiment-cern

### 2022-09-13 main/chapter_searches.tex
add pvalue misinterpretations paragraphs

@@
 
-% p-values
-\TODO{no!}~\cite{
+Data analysis in modern High Energy Physics does not exactly follow frequentist
+ideology, but is a hands-on, practical, and evolving mixture that has adapted
+to avoid some the more glaring flaws.
+We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
+can clearly express our scientific interest in how well competing models
+predict data, and how those likelihood can be manipulated in probability
+algebra.
+We have also seen in Section~\ref{sec:searches_frequentist} how 20th century
+writers championed likelihoods not on the data themselves but on inequalities,
+forming \pvalues\ and confidence intervals in a reductions that throw away
+information and obfuscate the results.
+This section reviews how these objects are applied generally in conventional
+analysis at the LHC and specifically towards the \atlas\ search that this
+thesis presents.
+
+Misinterpretations of \pvalues\ are widespread and well reported~\cite{
 schervish1996p,
@@
 Bernstein2016princess,
-Trafimow2015ban,
 wagenmakers2007practical,
-Read2002cls
-}
-\TODO{examples of use and misuse in hep}
-
-\TODO{p-value footnote:
-The chance of a small \pvalue\ is small.
-This is true, but not interesting because the value of any \pvalue\ is small.
-By construction \pvalues\ are uniformly distributed on the unit interval
-under a given hypothesis.
-The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
-$0.6 < p < 0.65$.
-The frequency of $p < \varepsilon$ is identical to the frequency
-of $0.6 < p < 0.6 + \varepsilon$.
-An appropriate choice of test statistic is therefore necessary for the
-\pvalue\ to have meaning.
+Trafimow2015ban
+}.
+For example, the idea of ``probability of compatibility'' with a model~\cite{
+HIGG-2018-04
+}
+appears to be spreading as a good meme~\cite{dawkins1989selfish} in
+High Energy Physics~\cite{
+HIGG-2018-27,
+HIGG-2018-51,
+EXOT-2018-08,
+HION-2018-19,
+HIGG-2017-09,
+langford2021combination,
+mastrandrea2019searches,
+white2019search,
+IceCube2013search,
+IceCube2014searches,
+gerasimov2021new
+},
+plausibly explained by vague but accurate statements like
+``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
+lhcb2021test
+}
+mutating to precise and inaccurate statements like
+``a probability of around $0.1\%$ that the data is compatible with the
+Standard Model predictions''~\cite{cern2021test}.
+
+There also exists a notion that a small \pvalue\ is significant because
+it is a rare fluctuation.
+All \pvalues\ are rare ---
+by construction they are uniformly distributed, so the
+probability of $P(h) < \varepsilon$ is equal to the probability of
+$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
+(if the data are sampled under $h$).
+All that matters to the interpretation is that an alternative hypothesis makes
+the observed result \emph{more probable}.
+
+% CLs
+$\cls$~\cite{
+zech1988cls,
+bock1998lower,
+read1997optimal,
+Read2002cls,
+Murray2010heretic
+}
+\begin{equation}
+\cls
+=
+\frac{P(s + b)}{P(b)}
+=
+\frac{
+\prob{t \geq \tilde{t}}{s + b}
+}{
+\prob{t \geq \tilde{t}}{b}
+}
+.
+\end{equation}
+These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
+Some literature also defines $\clb = 1 - p_b$~\cite{
+Cowan:2010js,
+pdg2022ynf
+},
+where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
+statistic.
+
+\TODO{CL \pvalue~\cite{pdg1984, murray1997use}}
+\TODO{Helene~\cite{pdg1988}}
+\TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
+
+textbooks cited by pdg~\cite{
+eadie1971statistical,
+frodesen1979probability
 }
@@
 
-$\cls$~\cite{
-zech1988cls,
-bock1998lower,
-read1997optimal,
-Read2002cls,
-Murray2010heretic
-}
-\begin{equation}
-\cls
-=
-\frac{P(s + b)}{P(b)}
-=
-\frac{
-\prob{t \geq \tilde{t}}{s + b}
-}{
-\prob{t \geq \tilde{t}}{b}
-}
-.
-\end{equation}
-These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
-Some literature also defines $\clb = 1 - p_b$~\cite{
-Cowan:2010js,
-pdg2022ynf
-},
-where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
-statistic.
-
-\TODO{CL \pvalue~\cite{pdg1984, murray1997use}}
-\TODO{Helene~\cite{pdg1988}}
-\TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
-
-textbooks cited by pdg~\cite{
-eadie1971statistical,
-frodesen1979probability
-}
-
 probability of compatibility~\cite{

### 2022-09-13 main/draft.tex
add pvalue misinterpretations paragraphs

@@
 SMRW scan url http://psychclassics.yorku.ca/Fisher/Methods/
+
+% Read2002cls
+% }
+\TODO{examples of use and misuse in hep}
+
+\TODO{p-value footnote:
+The chance of a small \pvalue\ is small.
+This is true, but not interesting because the value of any \pvalue\ is small.
+By construction \pvalues\ are uniformly distributed on the unit interval
+under a given hypothesis.
+The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
+$0.6 < p < 0.65$.
+The frequency of $p < \varepsilon$ is identical to the frequency
+of $0.6 < p < 0.6 + \varepsilon$.
+An appropriate choice of test statistic is therefore necessary for the
+\pvalue\ to have meaning.
+}

### 2022-09-13 main/chapter_searches.tex
motivate cls and arrange some citations

@@
 
-Misinterpretations of \pvalues\ are widespread and well reported~\cite{
+Misrepresentations of \pvalues\ (and other frequentist statistics) are
+widespread and widely reported~\cite{
 schervish1996p,
@@
 },
-plausibly explained by vague but accurate statements like
+as plausibly explained by mutations from vaguely accurate statements like
 ``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
@@
 }
-mutating to precise and inaccurate statements like
+to precisely inaccurate statements like
 ``a probability of around $0.1\%$ that the data is compatible with the
@@
 There also exists a notion that a small \pvalue\ is significant because
-it is a rare fluctuation.
+it is a rare fluctuation~\cite{murray1997use}.
 All \pvalues\ are rare ---
-by construction they are uniformly distributed, so the
+by construction, they are uniformly distributed, so the
 probability of $P(h) < \varepsilon$ is equal to the probability of
-$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
-(if the data are sampled under $h$).
-All that matters to the interpretation is that an alternative hypothesis makes
+$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set.
+If, that is, the data are sampled under $h$ ---
+what matters to the interpretation is whether an alternative hypothesis makes
 the observed result \emph{more probable}.
 
-% CLs
-$\cls$~\cite{
+For example, maximum likelihood ratios (and variations on their theme) are
+useful and commonly applied, but they can give \pvalues\ that appear
+pathological --- consider a Poisson background ($b$) added to an
+unconstrained signal yield ($s$).
+A sensible interpretation would find that `excess' data well above the
+background expectation favour the background-plus-signal hypothesis ($s + b$)
+---
+but if the signal is allowed to be negative, then a `deficit' of data below the
+background can have a maximum likelihood ratio that favours $s + b$.
+This scrambles excesses and deficits together in the same tail, so both get
+small \pvalues, and so a small \pvalue\ implies neither a preference for the
+signal hypothesis nor incompatibility with the background.
+
+As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
+on obfuscated data, so they can be interpreted in ratios.
+When setting upper limits on disfavoured signal models at the LHC,
+it is standard practice to use such \pvalue\ ratios name $\cls$~\cite{
+read2000modified,
+Read2002cls,
+Murray2010heretic,
+pdg2022ynf
+},
+which defined as
+\begin{equation}
+\cls
+=
+\frac{P(s + b)}{P(b)\hphantom{s +\;}}
+=
+\frac{
+\prob{t \geq \tilde{t}}{s + b}
+}{
+\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
+}
+.
+\end{equation}
+These individual \pvalues\ are sometimes labelled
+$\clsb = P(s + b)$ and $\clb = P(b)$.%
+\footnote{%
+Despite the $\mathrm{CL}$ labels, these are not confidence levels ---
+a confidence level $\gamma$ is a boundary at which \pvalues\ are selected, such
+as to set limits by solving for $x$ such that
+$\gamma = 1 - P(x)$~\cite{pdg1996}.
+In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
+at least 1984~\cite{pdg1984}, although earlier textbooks
+(cited by~\cite{pdg1984}) do make the distinction~\cite{
+eadie1971statistical,
+frodesen1979probability
+}.
+}
+Some literature also defines $\clb = 1 - p_b$~\cite{
+Cowan:2010js,
+pdg2022ynf
+},
+but $p_b$ may not quite be a \pvalue\ if it excludes equality with the observed
+test statistic.
+
+% history: bayes uniform prior in special cases
+% just an approximation
+% use for Higgs at LEP to avoid spurious rejection of bkg
+LEP Higgs boson searches~\cite{
+read1997optimal,
+bock1998lower,
+etde1998prospects,
+junk1999confidence,
+lep2000searches
+}
+
+credible intervals in simple Poisson or Gaussian cases~\cite{
+Helene1983upper,
+pdg1988,
+pdg2022ynf
+}
+
+origins~\cite{
+pdg1988,
 zech1988cls,
@@
 }
-\begin{equation}
-\cls
-=
-\frac{P(s + b)}{P(b)}
-=
-\frac{
-\prob{t \geq \tilde{t}}{s + b}
-}{
-\prob{t \geq \tilde{t}}{b}
-}
-.
-\end{equation}
-These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
-Some literature also defines $\clb = 1 - p_b$~\cite{
-Cowan:2010js,
-pdg2022ynf
-},
-where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
-statistic.
 
@@
 \TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
-
-textbooks cited by pdg~\cite{
-eadie1971statistical,
-frodesen1979probability
-}
 

### 2022-09-14 main/chapter_searches.tex
first addition of pvalue cls nonsense figures

@@
 High Energy Physics~\cite{
+HIGG-2017-09,
 HIGG-2018-27,
@@
 HION-2018-19,
-HIGG-2017-09,
-langford2021combination,
 mastrandrea2019searches,
 white2019search,
+langford2021combination,
 IceCube2013search,
@@
 probability of $P(h) < \varepsilon$ is equal to the probability of
-$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set.
-If, that is, the data are sampled under $h$ ---
-what matters to the interpretation is whether an alternative hypothesis makes
-the observed result \emph{more probable}.
-
+$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
+(if the data are indeed sampled under $h$).
+What matters to our interpretation is whether or not an alternative hypothesis
+makes the observed result \emph{more probable}.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_n.pdf}
+\caption{%
+(left) \pvalues\ for Poisson data,
+and (right) $\log$ likelihood ratios and $\log$ $\cls$ values between
+Poisson background and signal-plus-background models.%
+}
+\end{subfigure}
+\caption[
+Analysis of a minimal Poisson modelling example in the natural order.
+test statistics
+]{%
+Analysis of a minimal Poisson modelling example with $t(n) = \pm n$
+test statistics.
+}
+\label{fig:searches_sb_n}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.5\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
+\caption{%
+Data reordered by the maximum likelihood ratio test statistic with a background
+expectation $b=4.3$.
+}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_t.pdf}
+\caption{%
+(left) \pvalues\ for Poisson data,
+and (right) $\log$ likelihood ratios and $\log$ $\cls$ values between
+Poisson background and signal-plus-background models.%
+}
+\end{subfigure}
+\caption[
+Analysis of a minimal Poisson modelling example with maximum likelihood ratio
+test statistics
+]{%
+Analysis of a minimal Poisson modelling example with maximum likelihood ratio
+test statistics $t(n) = \pm \Lambda(n)$,
+where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$.
+}
+\label{fig:searches_sb_t}
+\end{figure}
+
+Figure~\ref{fig:searches_sb_n}
+Figure~\ref{fig:searches_sb_t}
+
+\TODO{n test stat and history first?}
 For example, maximum likelihood ratios (and variations on their theme) are
@@
 As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
-on obfuscated data, so they can be interpreted in ratios.
-When setting upper limits on disfavoured signal models at the LHC,
-it is standard practice to use such \pvalue\ ratios name $\cls$~\cite{
+on obfuscated data, so they are interpretable in likelihood ratios, and such
+\pvalue\ ratios are in fact standard practice at the LHC when setting upper
+limits on disfavoured signal models.
+The standard ratio is named $\cls$~\cite{
 read2000modified,
@@
 pdg2022ynf
-},
-which defined as
+}
+and defined as
 \begin{equation}
@@
 }
-.
-\end{equation}
-These individual \pvalues\ are sometimes labelled
-$\clsb = P(s + b)$ and $\clb = P(b)$.%
+,
+\end{equation}
+in which the individual \pvalues\ are labelled
+$\clb = P(b)$ (for background only) and
+$\clsb = P(s + b)$ (for signal plus background).%
 \footnote{%
-Despite the $\mathrm{CL}$ labels, these are not confidence levels ---
-a confidence level $\gamma$ is a boundary at which \pvalues\ are selected, such
-as to set limits by solving for $x$ such that
+Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
+not confidence levels ---
+a confidence level $\gamma$ is a decision boundary used, for example,
+to assign limits by solving for $x$ such that
 $\gamma = 1 - P(x)$~\cite{pdg1996}.
 In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
-at least 1984~\cite{pdg1984}, although earlier textbooks
-(cited by~\cite{pdg1984}) do make the distinction~\cite{
+at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
+by~\cite{pdg1984}) do make the distinction~\cite{
 eadie1971statistical,
 frodesen1979probability
-}.
+}.%
 }
@@
 \TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
+
+\TODO{Alas, slightly different test statistics }
 

### 2022-09-14 main/chapter_2ljets.tex
add atlas glossery ref, some fiddles

@@
 particles in the
-$\neutralino_2 \ldots \slepton \ldots \neutralino_1$
+$\neutralino_2 \ldots\ \slepton \ldots\ \neutralino_1$
 decay chain~\cite{paige1996determining}:

### 2022-09-14 main/chapter_searches.tex
add atlas glossery ref, some fiddles

@@
 $P(r=0.95)=0.00014$.
-``Thus'', Fisher concludes, ``the value $0.95$\ldots is roughly $1000$ times
+``Thus'', Fisher concludes, ``the value $0.95$\ldots\ is roughly $1000$ times
 less likely than $0.30$''
@@
 Misrepresentations of \pvalues\ (and other frequentist statistics) are
-widespread and widely reported~\cite{
+widespread and widely reported in areas of science that still use them~\cite{
 schervish1996p,
@@
 }.
-For example, the idea of ``probability of compatibility'' with a model~\cite{
-HIGG-2018-04
-}
-appears to be spreading as a good meme~\cite{dawkins1989selfish} in
-High Energy Physics~\cite{
+For example, the false presentation as a ``probability of compatibility'' with
+a model~\cite{HIGG-2018-04} appears to be spreading as a good
+meme~\cite{dawkins1989selfish} in High Energy Physics ~\cite{
 HIGG-2017-09,
@@
 gerasimov2021new
-},
-as plausibly explained by mutations from vaguely accurate statements like
+},%
+\footnote{%
+A draft of~\cite{HIGG-2018-51} stated that a \pvalue\ determined
+``the probability that the background-only hypothesis is compatible with the
+observed data''.
+That text was removed in response to my internal comments, but
+``probability of compatibility'' remains in its figure captions.%
+}%
+\footnote{%
+Reference~\cite{HIGG-2018-57} accurately describes a \pvalue\ as a tail
+probability, but also (repeatedly) states that a
+``probability of compatibility\ldots\ corresponds to a \pvalue''.%
+}%
+and is plausibly explained by mutations from vaguely accurate statements like
 ``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
@@
 
-There also exists a notion that a small \pvalue\ is significant because
-it is a rare fluctuation~\cite{murray1997use}.
-All \pvalues\ are rare ---
-by construction, they are uniformly distributed, so the
-probability of $P(h) < \varepsilon$ is equal to the probability of
-$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
+There also exists a notion that a small \pvalue\ means that the data
+represent a rare fluctuation~\cite{murray1997use, atlas2022pvalue}.%
+\footnote{%
+The \atlas\ Glossary of Terms~\cite{atlas2022pvalue} falsely states that
+``The \textbf{p-value} is the probability (ranging from $0$ to $1$) that the
+results observed could have occurred by chance, if the tested theory has no
+impact on the study.''%
+}
+All \pvalues\ are rare.
+By construction, they are uniformly distributed between $0$ and $1$, so the
+probability that $P(h) < \varepsilon$ is equal to the probability that
+$0.4 \leq P(h) < 0.4 + \varepsilon$ or that it lands in any
+$\varepsilon$-sized set
 (if the data are indeed sampled under $h$).
@@
 makes the observed result \emph{more probable}.
+
+\TODO{positive message:
+not meant only to disparage our lies, but to indicate that these explanations
+arise to fill a hole of confusion that we can fill by reporting interpretable
+statistics such as $\log$ likelihood ratios.
+}
 
@@
 (left) \pvalues\ for Poisson data,
-and (right) $\log$ likelihood ratios and $\log$ $\cls$ values between
-Poisson background and signal-plus-background models.%
+and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
+background and signal-plus-background models.%
 }
@@
 Analysis of a minimal Poisson modelling example with $t(n) = \pm n$
-test statistics.
+test statistics; $b = 4.3$ and $s = 3.0$.
+The $n$ test statistic is Neyman-Pearson optimal because it is monotonically
+(in fact linearly) related to the likelihood ratio.
+Each $\cls$ is a ratio between signal-plus-background and background-only
+\pvalues\ as defined in Equation~\ref{eqn:searches_cls}.
+By reducing the observed data to an inequality, the $p$-value selectively
+censors some of its information.
 }
@@
 \caption{%
-Data reordered by the maximum likelihood ratio test statistic with a background
-expectation $b=4.3$.
+Ordering the data by a maximum likelihood ratio test statistic.
 }
@@
 test statistics $t(n) = \pm \Lambda(n)$,
-where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$.
+where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$; $b = 4.3$ and $s = 3.0$.
+This test statistic jumbles up the data into an that makes $n=11$ appear to be
+weaker support for the signal than $n=9$.
+Each $\cls$ is a ratio between signal-plus-background and background-only
+\pvalues\ as defined in Equation~\ref{eqn:searches_cls}.
+By reducing the observed data to an inequality, the $p$-value selectively
+censors some of its information.
 }
@@
 \begin{equation}
+\label{eqn:searches_cls}
 \cls
@@
 https://home.cern/news/news/physics/intriguing-new-result-lhcb-experiment-cern
+
+\TODO{Statistical and systematic uncertainties}

### 2022-09-14 main/chapter_searches.tex
add some awkward words and drafts on cls; update figures with sky blue to be clear in greyscale

@@
 \),
-but a short calculation with their moments derives the exact result
+but a calculation with their moments derives the exact result
 \(
@@
 if $a$ and $b$ are independent.
-Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, this
+Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, the
 propagation formula works only for small relative errors.
@@
 
-Logic is an important application of probability for which $a$ and $b$ are
-interpreted as logical propositions, and
+Logic is an important application of probability, for which $a$ and $b$ are
+interpreted as logical propositions and
 $\prob{b}{a}$ is the degree to which $a$ implies $b$, with the limiting cases
@@
 Probability works for logic, and it works \emph{for tea, too}.
+
+Probability \emph{distributions} $\prob{b}{a}$ spread unit probability mass
+across states $b$, and need not be random distributions of fluctuating
+$b$s~\cite{jaynes2003probability,frankfurt2005on}, although random processes
+are useful applications.
 
@@
 \end{align}
-for fixed $a$, which is a normalized distribution over disjoint states $b$.
+for fixed $a$, which is a normalized probability distribution over $b$.
 
@@
 Neyman and Pearson were also involved in the development of
-`confidence intervals', which are essentially ranges of parameters that are
+`confidence intervals', which are ranges of parameters that are
 accepted by the \pvalue\ decision rules~\cite{
@@
 Confidence are usually introduced by endpoints containing constrained
-probability mass, but the arbitrary choice of test statistic can always reduce
-that to this \pvalue\ form.%
+probability mass, but this \pvalue\ form can always satisfy that through the
+free choice of test statistic and extends naturally beyond one dimension.%
 }
@@
 
-\TODO{positive message:
-not meant only to disparage our lies, but to indicate that these explanations
-arise to fill a hole of confusion that we can fill by reporting interpretable
-statistics such as $\log$ likelihood ratios.
-}
+Such misrepresentations of our results reinforces the need for clearer
+replacements, which $\log$ likelihood ratios can provide.
+And standard practice at the LHC contains the beginnings of such
+replacements under the name $\cls$~\cite{
+read2000modified,
+Read2002cls,
+junk1999confidence,
+read1997optimal,
+bock1998lower,
+etde1998prospects,
+lep2000searches,
+lep2003search,
+Murray2010heretic,
+CMS-NOTE-2011-005,
+pdg2022ynf
+},
+which is used for setting upper limits on disfavoured signal yields $s$
+above background yields $b$;
+\begin{equation}
+\label{eqn:searches_cls}
+\cls
+=
+\frac{P(s + b)}{P(b)\hphantom{s +\;}}
+=
+\frac{
+\prob{t \geq \tilde{t}}{s + b}
+}{
+\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
+}
+,
+\end{equation}
+in which the individual \pvalues\ are named
+$\clb = P(b)$ and
+$\clsb = P(s + b)$.%
+\footnote{%
+Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
+not confidence levels ---
+a confidence level $\gamma$ is a decision boundary used, for example,
+to assign limits by solving for $x$ such that
+$\gamma = 1 - P(x)$~\cite{pdg1996}.
+In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
+at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
+by~\cite{pdg1984}) do make the distinction~\cite{
+eadie1971statistical,
+frodesen1979probability
+}.%
+}%
+\footnote{%
+Some literature defines $\clb = 1 - p_b$~\cite{
+Cowan:2010js,
+pdg2022ynf
+};
+$p_b$ might not be a \pvalue\ if it excludes equality with the observed
+test statistic.
+}%
+As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
+on obfuscated data, so (with some caution) $\cls$ can be interpreted as a
+likelihood ratio.%
+\footnote{%
+Likelihood ratios are useful if both their numerator and denominator use the
+same data, so $\clsb$ and $\clb$ should use the same test statistic.
+Identical test statistics are used for all results in presented in this thesis,
+but differing test statistics have been allowed elsewhere~\cite{bock1998lower}.
+}%
+At a given confidence level $\gamma$ (typically $95\%$), limits are set to
+values of $s$ for which $\cls = 1 - \gamma$.
 
@@
 \includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_n.pdf}
-\caption{%
-(left) \pvalues\ for Poisson data,
-and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
-background and signal-plus-background models.%
-}
 \end{subfigure}
 \caption[
-Analysis of a minimal Poisson modelling example in the natural order.
-test statistics
+A Poisson modelling example in the natural order.
 ]{%
-Analysis of a minimal Poisson modelling example with $t(n) = \pm n$
-test statistics; $b = 4.3$ and $s = 3.0$.
+A Poisson modelling example with $t(n) = \pm n$
+test statistics, $b = 4.3$ and $s = 3.0$.
+(left) \pvalues,
+and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
+background and signal-plus-background models.
 The $n$ test statistic is Neyman-Pearson optimal because it is monotonically
-(in fact linearly) related to the likelihood ratio.
-Each $\cls$ is a ratio between signal-plus-background and background-only
-\pvalues\ as defined in Equation~\ref{eqn:searches_cls}.
+(linearly) related to the likelihood ratio.
+Each $\cls$ is a \pvalue ratio.
 By reducing the observed data to an inequality, the $p$-value selectively
@@
 \end{figure}
+
+Early forms of $\cls$ derived it from an identity with posterior tail
+probabilities in Poisson or Gaussian special cases~\cite{
+Helene1983upper,
+pdg1988,
+read2000modified,
+pdg2022ynf
+}
+with a contrived frequentist interpretation~\cite{zech1988cls}.
+For use in general cases, it is more recently described as a
+``modified frequentist'',
+``approximate confidence in the signal-only hypothesis''~\cite{
+read2000modified,
+Read2002cls
+}.
+High-profile usage in Higgs boson searches at the
+Large Electron–Positron Collider~\cite{
+read1997optimal,
+bock1998lower,
+etde1998prospects,
+junk1999confidence,
+lep2000searches,
+lep2003search
+}
+appears to have propelled $\cls$ to prominence.
+These early applications used data deficits to exclude their
+background-plus signal models with the test statistic $t(n) = -n$, but
+disliked how deficits led to small \pvalues\ for the background-only model.
+Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left);
+in contrast, $\cls$ (shown on the right of this figure) can only favour the
+background-only model.
+
+\TODO{also show actual likelihood ratio}
+\TODO{this implements a form of propaganda that speaks only truth while
+censoring all information that supports alternative ideas}
+\TODO{The one-sided nature of this evidence is paradoxical, but has
+resolutions.}
+\TODO{If an analysis were able to `is $n \geq 5$?', it could be answered in the
+affirmative or negative, and would avoid this bias.}
+\TODO{Otherwise, one can fall back to not trusting the reporter ---
+the statements `\atlas\ reports $n \geq 0$' and `$n \geq 0$' are not identical
+since the first might suggest that in fact $n = 0$, so wise likelihoods can be
+could recover the original data.
+One knows what the message says, but not whether it states the whole truth.
+}
 
@@
 \caption{%
-Ordering the data by a maximum likelihood ratio test statistic.
+Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
 }
@@
 weaker support for the signal than $n=9$.
-Each $\cls$ is a ratio between signal-plus-background and background-only
-\pvalues\ as defined in Equation~\ref{eqn:searches_cls}.
+Each $\cls$ is \pvalue\ ratio.
 By reducing the observed data to an inequality, the $p$-value selectively
@@
 
-Figure~\ref{fig:searches_sb_n}
 Figure~\ref{fig:searches_sb_t}
 
-\TODO{n test stat and history first?}
+\TODO{refresh}
 For example, maximum likelihood ratios (and variations on their theme) are
@@
 signal hypothesis nor incompatibility with the background.
-
-As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
-on obfuscated data, so they are interpretable in likelihood ratios, and such
-\pvalue\ ratios are in fact standard practice at the LHC when setting upper
-limits on disfavoured signal models.
-The standard ratio is named $\cls$~\cite{
-read2000modified,
-Read2002cls,
-Murray2010heretic,
-pdg2022ynf
-}
-and defined as
-\begin{equation}
-\label{eqn:searches_cls}
-\cls
-=
-\frac{P(s + b)}{P(b)\hphantom{s +\;}}
-=
-\frac{
-\prob{t \geq \tilde{t}}{s + b}
-}{
-\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
-}
-,
-\end{equation}
-in which the individual \pvalues\ are labelled
-$\clb = P(b)$ (for background only) and
-$\clsb = P(s + b)$ (for signal plus background).%
-\footnote{%
-Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
-not confidence levels ---
-a confidence level $\gamma$ is a decision boundary used, for example,
-to assign limits by solving for $x$ such that
-$\gamma = 1 - P(x)$~\cite{pdg1996}.
-In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
-at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
-by~\cite{pdg1984}) do make the distinction~\cite{
-eadie1971statistical,
-frodesen1979probability
-}.%
-}
-Some literature also defines $\clb = 1 - p_b$~\cite{
-Cowan:2010js,
-pdg2022ynf
-},
-but $p_b$ may not quite be a \pvalue\ if it excludes equality with the observed
-test statistic.
-
-% history: bayes uniform prior in special cases
-% just an approximation
-% use for Higgs at LEP to avoid spurious rejection of bkg
-LEP Higgs boson searches~\cite{
-read1997optimal,
-bock1998lower,
-etde1998prospects,
-junk1999confidence,
-lep2000searches
-}
-
-credible intervals in simple Poisson or Gaussian cases~\cite{
-Helene1983upper,
-pdg1988,
-pdg2022ynf
-}
-
-origins~\cite{
-pdg1988,
-zech1988cls,
-bock1998lower,
-read1997optimal,
-Read2002cls,
-Murray2010heretic
-}
-
-\TODO{CL \pvalue~\cite{pdg1984, murray1997use}}
-\TODO{Helene~\cite{pdg1988}}
-\TODO{LEP~\cite{etde1998prospects, lep2000searches, lep2003search}}
-
-\TODO{Alas, slightly different test statistics }
 

### 2022-09-15 main/chapter_searches.tex
add another painful todo
type some text on bias from pvalue ratios

@@
 Fisher's \pvalue\ ratio of $1000$ is, therefore, a $\log$ likelihood
-ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately strong
-indication of roughly $3.7\sigma$.
+ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately
+strong  indication of roughly $3.7\sigma$.
 But a likelihood ratio on the test statistic itself, and not the inclusive
@@
 Misrepresentations of \pvalues\ (and other frequentist statistics) are
-widespread and widely reported in areas of science that still use them~\cite{
+widespread and widely reported~\cite{
 schervish1996p,
@@
 Bernstein2016princess,
-wagenmakers2007practical,
-Trafimow2015ban
-}.
+wagenmakers2007practical
+}
+in areas of they are not banned~\cite{Trafimow2015ban}.
 For example, the false presentation as a ``probability of compatibility'' with
@@
 pdg2022ynf
-};
-$p_b$ might not be a \pvalue\ if it excludes equality with the observed
-test statistic.
+}, but such a $p_b$ might not be a \pvalue\ if it excludes equality with the
+observed test statistic.
 }%
@@
 (linearly) related to the likelihood ratio.
-Each $\cls$ is a \pvalue ratio.
+Each $\cls$ is a \pvalue\ ratio.
 By reducing the observed data to an inequality, the $p$-value selectively
@@
 
-Early forms of $\cls$ derived it from an identity with posterior tail
+Early variants of $\cls$ derived it from an identity with posterior tail
 probabilities in Poisson or Gaussian special cases~\cite{
@@
 }
-appears to have propelled $\cls$ to prominence.
-These early applications used data deficits to exclude their
+appears to have propelled $\cls$ to prominence;
+these early applications used data deficits to exclude their
 background-plus signal models with the test statistic $t(n) = -n$, but
-disliked how deficits led to small \pvalues\ for the background-only model.
-Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left);
-in contrast, $\cls$ (shown on the right of this figure) can only favour the
-background-only model.
-
-\TODO{also show actual likelihood ratio}
-\TODO{this implements a form of propaganda that speaks only truth while
-censoring all information that supports alternative ideas}
-\TODO{The one-sided nature of this evidence is paradoxical, but has
-resolutions.}
-\TODO{If an analysis were able to `is $n \geq 5$?', it could be answered in the
-affirmative or negative, and would avoid this bias.}
-\TODO{Otherwise, one can fall back to not trusting the reporter ---
-the statements `\atlas\ reports $n \geq 0$' and `$n \geq 0$' are not identical
-since the first might suggest that in fact $n = 0$, so wise likelihoods can be
-could recover the original data.
-One knows what the message says, but not whether it states the whole truth.
-}
+disliked how deficits led to small \pvalues\ for the background-only model,
+with which they might have confusingly set negative limits on the signal.
+Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left)
+for small data counts;
+in contrast, $\cls$ can only favour the background-only model with its negative
+values shown on the right of this figure.
+
+Although the statement in Section~\ref{sec:searches_fisher} that \pvalue\
+ratios are not ``not \emph{necessarily} misleading'',
+Figure~\ref{fig:searches_sb_n} shows a bias from the fact that
+no data can give a $\cls$ that favours the signal model;
+$\log\cls$ is non-positive with the $-n$ test statistic, and indeed non-negative
+for the opposite ordering $t(n) = n$.
+The $\log$ likelihood ratio on the actual data does not feature this bias:
+it favours background-only for deficits and background-plus-signal for
+excesses and is indeed linear in $n$:
+\begin{equation}
+\log \frac{L(s + b)}{L(b)}
+= -s + n\log(1 + \frac{s}{b})
+.
+\end{equation}
+Since likelihood ratio is monotonic in $n$, both $n$ and $-n$ are optimal
+test statistics by the criteria of the Neyman-Pearson lemma.
+
+By using \pvalue\ ratios with Neyman-Pearson optimal test statistic, we have
+constructed an analysis that can only ever favour one hypothesis:
+we have built a propaganda machine that only ever speaks truth but censors any
+information that supports alternative ideas.
+Reducing data to inequalities $t \geq \tilde{t}$ implements censorship,
+and the likelihood-ratio ordering guarantees the one-sided inferences.
+
+How can one escape from this trap?
+Likelihood ratios on the actual data escape immediately by bypassing the
+censor.
+Even inequalities need not be biased --- if against a predetermined threshold,
+say $5$ events, then models can predict whether or not $n \leq 5$, and its
+truth or falsehood in data can favour either side.%
+\footnote{%
+This reduction to a fixed inequality appears to be the condition for which a
+\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
+}
+Finally, a reader can fall back to not trusting the reporter ---
+the statements `\atlas\ reports $n \leq 5$' and logical `$n \leq 5$' are not
+identical, since one can infer from the former that in fact $n = 5$.
+They know what the message says, but not whether it states the whole truth.
 

### 2022-09-15 main/chapter_searches.tex
cleanup, prepare for next writing

@@
 }
-in areas of they are not banned~\cite{Trafimow2015ban}.
+in fields where they are not banned~\cite{Trafimow2015ban}.
 For example, the false presentation as a ``probability of compatibility'' with
@@
 disliked how deficits led to small \pvalues\ for the background-only model,
-with which they might have confusingly set negative limits on the signal.
+with which they might have confusingly set negative limits on the Higgs boson
+signal cross-section.
 Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left)
 for small data counts;
-in contrast, $\cls$ can only favour the background-only model with its negative
-values shown on the right of this figure.
+in contrast, $\cls$ can only favour the background-only model, due to its
+negative values that are shown on the right of the same figure.
 
@@
 \end{equation}
-Since likelihood ratio is monotonic in $n$, both $n$ and $-n$ are optimal
+Since the likelihood ratio is linear in $n$, both $n$ and $-n$ are optimal
 test statistics by the criteria of the Neyman-Pearson lemma.
@@
 information that supports alternative ideas.
-Reducing data to inequalities $t \geq \tilde{t}$ implements censorship,
-and the likelihood-ratio ordering guarantees the one-sided inferences.
-
-How can one escape from this trap?
+Reducing data to inequalities $t \geq \tilde{t}$ implements the censorship,
+and the likelihood-ratio ordering guarantees the one-sided inferences
+that appear in Figure~\ref{fig:searches_sb_n} (right) as one-sided
+$\log\cls$ values.
+
+How can one escape this propaganda trap?
 Likelihood ratios on the actual data escape immediately by bypassing the
 censor.
-Even inequalities need not be biased --- if against a predetermined threshold,
-say $5$ events, then models can predict whether or not $n \leq 5$, and its
-truth or falsehood in data can favour either side.%
+Even inequalities need not be biased --- if taken against a predetermined
+threshold, say $5$ events, then models can predict whether or not $n \leq 5$,
+and its truth or falsehood in data can favour either side.%
 \footnote{%
@@
 identical, since one can infer from the former that in fact $n = 5$.
-They know what the message says, but not whether it states the whole truth.
+They know what the message says, but not that it states the whole truth.
 
@@
 
-probability of compatibility~\cite{
-HIGG-2018-04,
-HIGG-2018-27,
-IceCube2013search,
-lhcb2021test,
-cern2021test
-}
-
-mix $p$-value with posterior tail~\cite{lhcb2017test}
-
-\subsection{Searches}
-\label{sec:searches_searches}
-% histograms, CR VR SR, systematics
-% histfactory
-
+\TODO{histfactory}
 \TODO{minimize energy not max logf}
+
 To model the consequences of these many factors of uncertainty, each factor is
@@
 
-lhcb2021test article: compatible with the SM prediction with a P value of 0.10\%
-cern:  probability of around 0.1\% that the data is compatible with the Standard Model predictions
-https://home.cern/news/news/physics/intriguing-new-result-lhcb-experiment-cern
-
-\TODO{Statistical and systematic uncertainties}
+\TODO{uncertainties: systematics, statistical}
+
+\subsection{\atlas\ SUSY searches}
+\label{sec:searches_searches}
+
+\TODO{CR VR SR}

### 2022-09-15 main/chapter_searches.tex
minor stats tweaks

@@
 For a known true parameter $x$, the confidence intervals in random repeated
-experiments will contain that $x$ with probability $1 - \alpha$;
+experiments will contain that $x$ with probability $\gamma$;
 the interval makes a subtle statement about unknown alternative data for
@@
 \begin{matrix}
-1 - \alpha & \textrm{if}~x \in x_1\textrm{--}x_2, \\
-\alpha & \textrm{otherwise.} \\
+\gamma & \textrm{if}~x \in x_1\textrm{--}x_2, \\
+1 - \gamma & \textrm{otherwise.} \\
 \end{matrix}
@@
 pdg2022ynf
-}, but such a $p_b$ might not be a \pvalue\ if it excludes equality with the
-observed test statistic.
+}, but such a $p_b$ might not be a \pvalue\ when it excludes equality with the
+observed test statistic~\cite{CMS-NOTE-2011-005}.
 }%

### 2022-09-15 main/chapter_2ljets.tex
write about other maximum likelihood test statistics; add some subsections

@@
 test statistic distributions are assigned by first fitting to the
-data~\cite{CMS-NOTE-2011-005}.
+data~\cite{cern2011procedure}.
 Just as the Matrix Method is not predictive because it is conditioned on the

### 2022-09-15 main/chapter_searches.tex
write about other maximum likelihood test statistics; add some subsections

@@
 \begin{equation}
+\label{eqn:searches_max_like_ratio}
 \Lambda =
@@
 
+\subsection{\texorpdfstring{$\cls$}{CLs}}
 Such misrepresentations of our results reinforces the need for clearer
@@
 Murray2010heretic,
-CMS-NOTE-2011-005,
+cern2011procedure,
 pdg2022ynf
@@
 }, but such a $p_b$ might not be a \pvalue\ when it excludes equality with the
-observed test statistic~\cite{CMS-NOTE-2011-005}.
+observed test statistic~\cite{cern2011procedure}.
 }%
@@
 \log \frac{L(s + b)}{L(b)}
-= -s + n\log(1 + \frac{s}{b})
+= -s + n\log\!\left(1 + \frac{s}{b}\right)
 .
@@
 
-How can one escape this propaganda trap?
-Likelihood ratios on the actual data escape immediately by bypassing the
-censor.
-Even inequalities need not be biased --- if taken against a predetermined
-threshold, say $5$ events, then models can predict whether or not $n \leq 5$,
-and its truth or falsehood in data can favour either side.%
-\footnote{%
-This reduction to a fixed inequality appears to be the condition for which a
-\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
-}
-Finally, a reader can fall back to not trusting the reporter ---
-the statements `\atlas\ reports $n \leq 5$' and logical `$n \leq 5$' are not
-identical, since one can infer from the former that in fact $n = 5$.
-They know what the message says, but not that it states the whole truth.
-
 \begin{figure}[tp]
 \centering
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{\textwidth}
 \centering
-\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
+\includegraphics[width=0.5\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
 \caption{%
@@
 
-Figure~\ref{fig:searches_sb_t}
-
-\TODO{refresh}
-For example, maximum likelihood ratios (and variations on their theme) are
-useful and commonly applied, but they can give \pvalues\ that appear
-pathological --- consider a Poisson background ($b$) added to an
-unconstrained signal yield ($s$).
-A sensible interpretation would find that `excess' data well above the
-background expectation favour the background-plus-signal hypothesis ($s + b$)
----
-but if the signal is allowed to be negative, then a `deficit' of data below the
-background can have a maximum likelihood ratio that favours $s + b$.
-This scrambles excesses and deficits together in the same tail, so both get
-small \pvalues, and so a small \pvalue\ implies neither a preference for the
-signal hypothesis nor incompatibility with the background.
-
-% significance measures
-The ATLAS-recommended significance measure~\cite{atlas_significance} is
-\begin{align}
-\label{eqn:significance_atlas}
-S_\mathrm{\atlas}(n; \mu, \sigma) =~&
-\sqrt{2} \times
-\mathrm{sign}(n - \mu) \times
-\\[0.2em] \nonumber
-&
-\sqrt{
-n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
-- \frac{\mu^2}{\sigma^2}\log\left(
-1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
-\right)
-}
-,
-\end{align}
-in an appropriate limit where $0\log(0x) = 0$ for the $n=0$ case.
+How can one escape this propaganda?
+Likelihood ratios on the actual data escape immediately by bypassing the
+censor.
+Even inequalities need not be biased --- if taken against a predetermined
+threshold, say $5$ events, then models can predict whether or not $n \leq 5$,
+and its truth or falsehood in data can favour either side.%
+\footnote{%
+This context of a fixed inequality is one condition where the
+\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
+}
+Finally, a reader can fall back to not trusting the reporter ---
+the statements `\atlas\ reports $n \leq 5$' and logical `$n \leq 5$' are not
+identical, since one can infer from the former that in fact $n = 5$.
+They know what the message says, but not that it states the whole truth.
+
+With more than one Poisson bin to model, the $n$ test statistic is no longer
+sufficient.
+In such general cases, maximum likelihood ratio test statistics,
+such as $\Lambda$ of Equation~\ref{eqn:searches_max_like_ratio}, often useful.
+In the single Poisson case, $\Lambda$ shuffles the ordering of
+data and leads to some odd features in the results, which are displayed in
+Figure~\ref{fig:searches_sb_t}.
+In particular, the choice $t(n) = -\Lambda(n)$ pushes data that are far from
+the background expectation into its tails, so can disfavour the background
+model for small $n$.
+Probability masses from small data also interrupt its behaviour for excesses,
+leading to jagged changes, notably that $\log\cls$ reduces as the data
+increase from $9$ to $11$ --- counter to intuition and logic, increasing excess
+data can reduce the reported favour for the signal model.
+
+Instead of $\Lambda$ itself, various modified maximum likelihood ratio test
+statistics enjoy standard usage at the LHC~\cite{cern2011procedure};
+these are variably defined as piecewise functions of the fitted signal
+parameter $s$, and can include likelihoods from the signal-plus-background
+model with constrained maximizations which aim to reduce the impact of
+unphysical results such as negative Poisson expectations~\cite{
+Feldman:1997qc,
+Cowan:2010js
+},
+and can help to better approximate a likelihood ratio ordering.
+
+\subsection{Like-lihood modelling}
 
@@
 
+\subsection{Significance}
+
+\TODO{significance}
+
+% significance measures
+The ATLAS-recommended significance measure~\cite{atlas_significance} is
+\begin{align}
+\label{eqn:significance_atlas}
+S_\mathrm{\atlas}(n; \mu, \sigma) =~&
+\sqrt{2} \times
+\mathrm{sign}(n - \mu) \times
+\\[0.2em] \nonumber
+&
+\sqrt{
+n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
+- \frac{\mu^2}{\sigma^2}\log\left(
+1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
+\right)
+}
+,
+\end{align}
+taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.
+
 \subsection{\atlas\ SUSY searches}

### 2022-09-15 main/chapter_searches.tex
add note on estimating p-values

@@
 
+A great benefit to these modified maximum likelihood ratios is that their
+sampling distributions have known approximations in asymptotic (Gaussian)
+limits, which assign \pvalues\ through well-behaved analytic
+expressions~\cite{Cowan:2010js}.
+Without such formulae, the remaining practical approach to estimating
+\pvalues\ is to generate Monte Carlo simulations of alternative data.
+Such simulations can be computationally expensive, and even hard to define for
+``likelihood'' models that blur the line between actual data and
+auxiliary model parameters.
+
 \subsection{Like-lihood modelling}
-
 \TODO{

### 2022-09-15 main/chapter_searches.tex
minor tweaks

@@
 \caption[
-A Poisson modelling example in the natural order.
+A Poisson model in the natural order
 ]{%
-A Poisson modelling example with $t(n) = \pm n$
+A Poisson model in the natural order of $t(n) = \pm n$
 test statistics, $b = 4.3$ and $s = 3.0$.
-(left) \pvalues,
+(left) $\log$ \pvalues,
 and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
@@
 it favours background-only for deficits and background-plus-signal for
-excesses and is indeed linear in $n$:
+excesses.
+Since
 \begin{equation}
@@
 = -s + n\log\!\left(1 + \frac{s}{b}\right)
-.
-\end{equation}
-Since the likelihood ratio is linear in $n$, both $n$ and $-n$ are optimal
-test statistics by the criteria of the Neyman-Pearson lemma.
+\end{equation}
+is linear in $n$, both $n$ and $-n$ are optimal in the Neyman-Pearson sense.
 
@@
 \caption{%
-(left) \pvalues\ for Poisson data,
+(left) $\log$ \pvalues\ for Poisson data,
 and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
@@
 \caption[
-Analysis of a minimal Poisson modelling example with maximum likelihood ratio
+Analysis of a minimal Poisson model with maximum likelihood ratio
 test statistics
 ]{%
-Analysis of a minimal Poisson modelling example with maximum likelihood ratio
+Analysis of a minimal Poisson model with maximum likelihood ratio
 test statistics $t(n) = \pm \Lambda(n)$,

### 2022-09-15 main/chapter_searches.tex
remove pvalue spacer
(s+b)/s, minor stats tweaks, begin experiment notes

@@
 =
-\frac{P(s + b)}{P(b)\hphantom{s +\;}}
+\frac{P(s + b)}{P(b)}
 =
@@
 same data, so $\clsb$ and $\clb$ should use the same test statistic.
-Identical test statistics are used for all results in presented in this thesis,
-but differing test statistics have been allowed elsewhere~\cite{bock1998lower}.
+Identical test statistics are used for all results of this work, but differing
+test statistics have been allowed elsewhere~\cite{bock1998lower}.
 }%
@@
 excesses.
-Since
+Since for Poisson models the $\log$ likelihood ratio is linear in $n$,
 \begin{equation}
 \log \frac{L(s + b)}{L(b)}
-= -s + n\log\!\left(1 + \frac{s}{b}\right)
-\end{equation}
-is linear in $n$, both $n$ and $-n$ are optimal in the Neyman-Pearson sense.
+= n\log\frac{s + b}{b} - s
+,
+\end{equation}
+both $n$ and $-n$ are optimal in the Neyman-Pearson sense.
 

### 2022-09-15 main/chapter_experiment.tex
yahoo!
add metsig and mt2 descriptions
first runthough trigger sec
fix mahalonobis name
begin detector chapter draft

@@
 {Ken~M,
-\textit{internet comment},
+\textit{Comment: Yahoo! News},
 2012~\cite{kenm2012inner}}
@@
 
-Hello
+\TODO{merge into theory chapter?}
+
+from the paper~\cite{atlas2022searches}
+
+\TODO{main detector}
+\begin{displayquote}
+``%
+The ATLAS detector~\cite{PERF-2007-01} is a multipurpose particle detector with
+a forward--backward symmetric cylindrical geometry and a near $4\pi$ coverage
+in solid angle.%
+\footnote{%
+``ATLAS uses a right-handed coordinate system with its origin at the nominal
+interaction point (IP) in the centre of the detector and the $z$-axis along the
+beam pipe.
+The $x$-axis points from the IP to the centre of the LHC ring, and the $y$-axis
+points upwards.
+Cylindrical coordinates $(r,\phi)$ are used in the transverse plane, $\phi$
+being the azimuthal angle around the $z$-axis.
+The pseudorapidity is defined in terms of the polar angle $\theta$ as
+$\eta = -\ln \tan(\theta/2)$.
+Angular distance is measured in units of
+$\Delta R \equiv \sqrt{(\Delta\eta)^{2} + (\Delta\phi)^{2}}$.%
+''~\cite{atlas2022searches}
+}
+It consists of an inner tracking detector surrounded by a thin superconducting
+solenoid providing a $2\,\mathrm{T}$ axial magnetic field, electromagnetic and
+hadronic calorimeters, and a muon spectrometer.%
+''~\cite{atlas2022searches}
+\end{displayquote}
 
 \section{Inner tracker}
+\label{sec:atlas_tracker}
+\TODO{inner tracker}
+\begin{displayquote}
+``%
+The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
+It consists of silicon pixel, silicon microstrip, and transition radiation
+tracking detectors.
+An additional layer of silicon pixels, the insertable
+B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
+''~\cite{atlas2022searches}
+\end{displayquote}
 
 \section{Calorimeters}
+\label{sec:atlas_calo}
+\TODO{calo}
+\begin{displayquote}
+``%
+Lead/liquid-argon (LAr) sampling calorimeters provide electromagnetic (EM)
+energy measurements with high granularity.
+A steel/scintillator-tile hadron calorimeter covers the central pseudorapidity
+range ($|\eta| < 1.7$).
+The endcap and forward regions are instrumented with LAr calorimeters for both
+the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
+''~\cite{atlas2022searches}
+\end{displayquote}
 
 \section{Muon system}
+\label{sec:atlas_muon}
+\TODO{muon system}
+\begin{displayquote}
+``%
+The muon spectrometer surrounds the calorimeters and is based on three large
+superconducting air-core toroidal magnets with eight coils each.
+The field integral of the toroids ranges between $2.0$ and
+$6.0\,\mathrm{T\kern-0.15ex m}$ across most of the detector.
+The muon spectrometer includes a system of precision chambers for tracking and
+fast detectors for triggering.%
+''~\cite{atlas2022searches}
+\end{displayquote}
 
-\section{Simulation}
+\section{Trigger}
+\label{sec:atlas_trigger}
+% L1 L2 HLT
+\TODO{trigger}
+\begin{displayquote}
+``%
+A two-level trigger system is used to select events.
+The first-level trigger is implemented in hardware and uses a subset of the
+detector information to accept events at a rate below $100\,\mathrm{kHz}$.
+This is followed by a software-based trigger that reduces the accepted event
+rate to $1\,\mathrm{kHz}$ on average depending on the data-taking conditions.%
+''~\cite{atlas2022searches}
+\end{displayquote}
+
+\section{Software}
+\label{sec:atlas_software}
 % how do we predict data?
+\TODO{software}
+\begin{displayquote}
+``%
+An extensive software suite~\cite{ATL-SOFT-PUB-2021-001} is used in the
+reconstruction and analysis of real and simulated data, in detector operations,
+and in the trigger and data acquisition systems of the experiment.%
+''~\cite{atlas2022searches}
+\end{displayquote}
 
 \section{Data}
+\label{sec:atlas_data}
 % discuss collisions, pileup, met, metsig

### 2022-09-15 main/chapter_searches.tex
autocheck tweaks

@@
 \caption{%
-Prior density: exponential in
+Prior: exponential density in
 (left) $x$ and
@@
 \caption[
-Likelihoods and prior densities' different transformations
+Likelihoods and prior densities' transformations with coordinates.
 ]{%
-Likelihoods and prior densities' different transformations with coordinates.
+Likelihoods and prior densities' transformations with coordinates.
 The $\mathrm{Poisson}(0\mid x)$ likelihood and exponential $e^{-x}\mathrm{d}x$
-prior have the same form, but the prior density changes to preserve its unit
-area when expressed relative to $u = \log x$.
+prior have the same form (left), but the prior density changes to preserve its unit
+area relative to $u = \log x$.
+On changing to $\log$ coordinates (right), the likelihood slides horizontally,
+but the prior density scales to preserve area.
+Both maximum likelihood points are equal, but the maximum prior density moves.
 \\[.5em]
-Relative to $x$, both are identical (left).
-On changing to $\log$ coordinates (right), the likelihood slides horizontally,
-but the prior density changes value to preserve area.
-Both maximum likelihood points are equivalent (maximum likelihoods are robust),
-but the maximum prior density moves (maximum densities are slippery).
-\\[.5em]
-Particle lifetimes follow exponential distributions, and their maximum
-densities are at zero time.
-It is misleading, however, to describe their most probable lifetimes as zero
----
+Particle lifetimes follow exponential distributions and their maximum
+densities are at zero time, but it is incorrect to describe their most probable
+lifetimes as zero ---
 the `most probable' $\log$ lifetime is not $\log 0$, but equals the mean
@@
 For zero Poisson data, the maximum likelihood rate is zero, but zero might not
-be the most likely rate (if, for example, an experiment has known backgrounds).
+be the most probable rate (if, for example, an experiment has known backgrounds).
 }
@@
 A likelihood function indexes each bin in the background--signal labels,
-and is independent of the coordinates units since the area in each bin is
+and is independent of the coordinates units, since the area in each bin is
 invariant.
@@
 For model comparison, for example between background $h_1=B$ and signal $h_2=S$
-models, the analysis is conveniently linearised by expressing it in $\log$
+models, the analysis is conveniently linearized by expressing it in $\log$
 ratio form:
@@
 are larger --- unless other hypotheses predicted \emph{better}.
-And these ($\log$) ratios can remain well defined in continuum limits even as
+And these ($\log$) ratios can remain well-defined in continuum limits even as
 both likelihoods tend towards zero~\cite{billingsley2008probability}.
@@
 You are welcome to propose any prediction in any circumstances, and the
-scientific response is to test whether or not its predictions are compatible
-with our observations of nature.
+scientific response is to test how well it predictts our observations of nature.
 Our emphasis on the likelihood functions and evidence values that quantify
@@
 so the frequency definition cannot not apply.
-Prior predictions often apply probabilities to express a subjective uncertainty
-predictions that are manifestly non-random.
+Prior predictions often apply probabilities to express subjective uncertainties
+that are manifestly non-random.
 Under this frequency definition, those predictive probabilities are nonsense
@@
 
-Tom Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
+Tom~Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
 stoppard1967rosencrantz
@@
 axioms1010038
-}, but applied frequentist methods, are constrained with considerably less
+}, but \emph{applied} frequentist methods are constrained with considerably less
 precision --- the denial of certain (non-frequency) probabilities does not
@@
 Modern frequentist methods are largely attributable to work by
-Ronald A. Fisher (for likelihood, $p$-values, and estimation criteria~\cite{
+Ronald~Fisher (for likelihood, $p$-values, and estimation criteria~\cite{
 fisher1912fitting,
@@
 }),
-and Jerzy Neyman and Egon S. Pearson
+and Jerzy~Neyman and Egon~Pearson
 (for statistical hypothesis testing and confidence intervals~\cite{
@@
 })
-in the 1920s and 30s.
+in the 1920s and '30s.
 These Fisher and Neyman-Pearson schools do not form a unified framework, but
@@
 Horizontal lines mark regions where the likelihood function exceeds
-$1/2$, $1/5$, and $1/15$ of its maximum respectively.
+$1/2$, $1/5$, and $1/15$ of its maximum, respectively.
 The extents of these lines could be tabulated to usefully communicate aspects
@@
 },
-and deserves credit for popularising its use in interpreting and communicating
+and deserves credit for popularizing its use in interpreting and communicating
 data without the confounding influence of prior probabilities.
@@
 }.
-Defining bias as a difference from a mean fails to be coordinate invariant,
+Defining bias as a difference from a mean fails to be coordinate-invariant,
 so an unbiased estimate of the variance $\bar{\sigma}^2$ is a biased
@@
 useful test statistics can also eliminate unwanted parameters.
-The quintessential example is the $t$-statistic
+The quintessential example is the $t$-statistic,
 \begin{equation}
@@
 $\bar{\sigma}(d)$~\cite{student1908, fisher1925t}.
-By dividing out a scale equivariant estimate of the standard deviation, $t$ is
-independent of common scale factors on the data, and that scale invariance
+By dividing out a (scale-equivariant) estimate of the standard deviation, $t$
+is independent of common scale factors on the data, and that scale invariance
 has helped to drive the study and common use of this
@@
 to decide on actions of `rejecting' or `accepting' an hypothesis.
-From the Bayesian arguments that we presenting in
+From the Bayesian arguments that we are presenting in
 Section~\ref{sec:searches_data_analysis},
@@
 independent of whatever (non-zero) prior probabilities they are assigned.
-Indeed they later state that
+Indeed, they later state that
 \begin{displayquote}
@@
 \footnote{%
-Confidence are usually introduced by endpoints containing constrained
+Confidence intervals are usually introduced by endpoints containing constrained
 probability mass, but this \pvalue\ form can always satisfy that through the
@@
 known true parameters.
-Given confidence intervals can even exclude all plausible parameters~\cite{
+Confidence intervals can even exclude all plausible parameters~\cite{
 pratt1961testing,
@@
 In response to the idea that no conclusion can be drawn about the parameter
-iself, Fisher also deserves the last word:
+itself, Fisher also deserves the last word:
 \begin{displayquote}
@@
 ideology, but is a hands-on, practical, and evolving mixture that has adapted
-to avoid some the more glaring flaws.
+to avoid some of the more glaring flaws.
 We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
 can clearly express our scientific interest in how well competing models
-predict data, and how those likelihood can be manipulated in probability
+predict data, and how those likelihoods can be manipulated in probability
 algebra.
@@
 writers championed likelihoods not on the data themselves but on inequalities,
-forming \pvalues\ and confidence intervals in a reductions that throw away
+forming \pvalues\ and confidence intervals in reductions that throw away
 information and obfuscate the results.
@@
 (if the data are indeed sampled under $h$).
-What matters to our interpretation is whether or not an alternative hypothesis
+What matters to our interpretation is whether an alternative hypothesis
 makes the observed result \emph{more probable}.
@@
 Even inequalities need not be biased --- if taken against a predetermined
-threshold, say $5$ events, then models can predict whether or not $n \leq 5$,
+threshold, say $5$ events, then models can predict whether $n \leq 5$,
 and its truth or falsehood in data can favour either side.%

### 2022-09-16 main/chapter_searches.tex
write some on likelihood models --- yields and energy

@@
 \subsection{Probability, applied}
+\label{sec:searches_probability_applied}
 \begin{figure}[tp]
@@
 these early applications used data deficits to exclude their
-background-plus signal models with the test statistic $t(n) = -n$, but
-disliked how deficits led to small \pvalues\ for the background-only model,
+background-plus signal models, sometimes with the test statistic $t(n) = -n$,
+and disliked how deficits led to small \pvalues\ for the background-only model,
 with which they might have confusingly set negative limits on the Higgs boson
@@
 
-Although the statement in Section~\ref{sec:searches_fisher} that \pvalue\
+Despite the statement in Section~\ref{sec:searches_fisher} that \pvalue\
 ratios are not ``not \emph{necessarily} misleading'',
@@
 \end{equation}
-both $n$ and $-n$ are optimal in the Neyman-Pearson sense.
+both $n$ and $-n$ can be optimal in the Neyman-Pearson sense.
 
@@
 identical, since one can infer from the former that in fact $n = 5$.
-They know what the message says, but not that it states the whole truth.
+One knows what the message says, but not that it states the whole truth.
 
@@
 },
-and can help to better approximate a likelihood ratio ordering.
-
-A great benefit to these modified maximum likelihood ratios is that their
-sampling distributions have known approximations in asymptotic (Gaussian)
-limits, which assign \pvalues\ through well-behaved analytic
+and can improve (in a context-dependent sense) the resulting ordering.
+
+These modified maximum likelihood ratios benefit from known approximations to
+their sampling distributions in asymptotic (Gaussian) limits,
+which can assign their \pvalues\ through well-behaved analytic
 expressions~\cite{Cowan:2010js}.
@@
 \pvalues\ is to generate Monte Carlo simulations of alternative data.
-Such simulations can be computationally expensive, and even hard to define for
-``likelihood'' models that blur the line between actual data and
-auxiliary model parameters.
+Such simulations can be computationally expensive, ill-defined for
+``likelihood'' models that blur the distinction between actual data and
+auxiliary constraints on the models' parameters.
 
 \subsection{Like-lihood modelling}
-\TODO{
-on ``likelihood'' models ---
-we are instructed to pretend that these are likelihoods on auxiliary data
-TODO cite.
-One is free to engage in such doublethink~\cite{orwell1949nnineteen},
-% just as one is free to proclaim that $2 + 2 = 5$
-but the fact remains that many constraints do not correspond to predictions of
-auxiliary data so are not likelihoods: examples include
+To represent the Standard Model and alternatives, LHC data analysis uses
+parametric models to describe the models' Poisson expectations of
+collision events and how those expectations change in response to their
+parameters.
+Parameters often correspond to interpretable, physical quantities and when
+those parameters are known only with some uncertainty, the models help to
+propagate uncertainty into variations of their outputs.
+
+Uncertainties on imperfectly-known yet non-random parameters, however,
+conflict with the frequentist denial of prior probabilities.
+To avoid this, all modelling is officially framed as not predicting parameters,
+but instead constraining them with likelihood functions that use
+either real control data or fictional auxiliary data.
+
+This transformation of priors to auxiliary likelihoods is generally possible
+--- in a prior density $f(x)\,\mathrm{d}x$ the function $f(x)$ is a likelihood
+(strictly, a Radon-Nikodym derivative~\cite{billingsley2008probability})
+that constrains the (possibly un-normalized or `improper') prior measure that
+is uniform in $x$~\cite{Cowan:2010js}.
+But as discussed in Section~\ref{sec:searches_probability_applied}, the shape
+of $f(x)$ is entirely determined by the coordinate, and is therefore a free
+choice made by the analyst.
+By assigning a form to the auxiliary likelihood, therefore, one is really
+choosing a \emph{regularizer} that should be indended to induce preferable
+behaviours in the parametric model and its `maximum likelihood' statistics.
+
+These regularized models with constraints from fictitious auxiliary data
+are usually often ``likelihood'' models~\cite{
+cranmer2012histfactory,
+baak2015histfitter,
+Besjes_2015,
+heinrich2021pyhf
+}.
+To distinguish them from real likelihood functions on actual data, this
+thesis uses the alternative spelling \textbf{\heplikelihood}.
+
+Many of our uncertainties do not correspond to predictions of auxiliary data:
+examples include
 scale variations,
 theoretical cross-section uncertainties,
-non-closure uncertainties (which cover differences between estimates).
+non-closure uncertainties (which cover differences between estimates),
 simulated sample comparisons,
@@
 (for which we have yet no data).
-energy based modelling ---
-I know of no tangible benefit from pretending that something is a likelihood
-when it is not.
-}
-
-\TODO{histfactory}
-\TODO{minimize energy not max logf}
-
-To model the consequences of these many factors of uncertainty, each factor is
-assigned a real-valued parameter, which when varied changed the expected yields
-in the histogram bins of the search.
-In this way, we form a statistical model which is a function
-\begin{equation}
-H_{\!f}(\vec \theta) \to (\vec \mu, \log f)
-\end{equation}
-from the parameters $\theta$ to Poisson expectations $\vec\mu$ of the histogram
-bins and a regularization term $\log f$ that is maximized in fitting.
+We describe these uncertainties by heuristically
+combining theoretical modelling, physical intuition, and convention,
+and not only by measuring random data.
+Believing that they are also likelihoods on auxiliary data is
+doublethink~\cite{orwell1949nnineteen}.
+
+Despite their troublesome nomenclature, \heplikelihood\ models are extremely
+useful and effective for describing the predictions of our hypotheses and
+extracting interpretable results.
+A general \heplikelihood\ model is a function $H$ that takes some parameters $x$
+and outputs two things:
+expected yields $y$ to predict Poisson data,
+and an objective value $E$ that should be minimized in a `fit' to choose the
+best prediction.
+That is,
+\begin{equation}
+H(x) = [\,y(x), \,E(x)\,]
+.
+\end{equation}
+The minimization objective $E$ is the summed negative $\log$ likelihood
+(or regularizer) for all actual and auxiliary data;
+we choose the letter $E$ for energy in the analogy to statistical physics where
+at low temperature $T = 1/\beta$ a system collapses to its minimuum energy
+state due to the Boltzmann factor
+\begin{equation}
+p(x) \propto e^{-\beta E(x)}
+.
+\end{equation}
+At unit temperature, this can define a propability distribution for the
+parametesr relative to an appropraite (prior)
+measure~\cite{cranmer2012histfactory, skilling2017david},
+but in practice such distributions do not always coincide with intuition or
+with the best fit value.
+
+\TODO{fit, expand for error}
+
+\TODO{software tools}
 
@@
 
-\TODO{uncertainties: systematics, statistical}
-
-\subsection{Significance}
+\TODO{uncertainties: systematics, statistical --- maybe susy step?}
+
+\TODO{toy procedure}
+
+\subsection{\atlas\ SUSY searches}
+\label{sec:searches_searches}
+
+\TODO{CR VR SR}
+
+\subsection{Gaussian significance}
 
@@
 taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.
-
-\subsection{\atlas\ SUSY searches}
-\label{sec:searches_searches}
-
-\TODO{CR VR SR}

### 2022-09-16 main/chapter_searches.tex
add some works on error bars and histfactory models

@@
 For $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
-the standard perturbative approximation is
+the standard perturbative approximation for their product is
 \(
@@
 Figure~7 of~\cite{lester2022hunting}.
-When making error bar recommendations, \atlas' statistics committee appears
-to have not considered this option~\cite{cowan2011errorbars};
-I cannot explain why not, since our statistics experts must surely be familiar
-with Fisher's work.
+When recommending error bar recipes, \atlas' statistics committee appears
+to have not considered this option~\cite{cowan2011errorbars}, although
+our statistics experts are surely familiar with Fisher's work.
 }
@@
 (or regularizer) for all actual and auxiliary data;
-we choose the letter $E$ for energy in the analogy to statistical physics where
-at low temperature $T = 1/\beta$ a system collapses to its minimuum energy
-state due to the Boltzmann factor
+we choose the letter $E$ for \textbf{energy}, to highlight the analogy with
+statistical physics and the Boltzmann factor
 \begin{equation}
@@
 \end{equation}
-At unit temperature, this can define a propability distribution for the
-parametesr relative to an appropraite (prior)
+with which a system collapses to its minimum energy at low temperature
+$T = 1/\beta$.
+At unit temperature, the Boltzmann factor can define a probability distribution
+for the parameters relative to an appropriate (prior)
 measure~\cite{cranmer2012histfactory, skilling2017david},
@@
 
-\TODO{fit, expand for error}
-
-\TODO{software tools}
-
-\histfactory~\cite{cranmer2012histfactory}
-\pyhf~\cite{heinrich2021pyhf}
+Error bars on the yields can be assigned by exploring around that best fit.
+In the Gaussian approximation, the inverse (Hessian) matrix of second
+derivatives gives a covariance matrix in the parameters from which linear
+projections can estimate the variance of any output.
+The second derivatives of exactly Gaussian forms are independent of the
+parameter point, but since we do not have exactly Gaussian models we evaluate
+these derivatives about the best fit point, either analytically or with finite
+difference approximations~\cite{cranmer2021building}.
+Other practical approaches explore parameter ranges in which the minimum energy
+exceeds thresholds, similar to Fisher's likelihood intervals.
+
+All results presented in this thesis use
+\histfactory~\cite{cranmer2012histfactory} \heplikelihood\ models that are
+constructed and analysed with
 \histfitter~\cite{Besjes_2015,baak2015histfitter}
-
-\TODO{uncertainties: systematics, statistical --- maybe susy step?}
-
-\TODO{toy procedure}
+and \pyhf~\cite{heinrich2021pyhf}.
+\histfactory\ models are specialized for working with discrete numbers of
+histogram bins.
+They constrain parameters with either Gaussian or Poisson likelihood
+functions, and assign expected yields through a variety of
+non-linear functions of the parameters.
+Each parameter's effect is generally described by
+\emph{central}, \emph{high}, and \emph{low}
+(corresponding to the best and $\pm1\sigma$)
+values of modifications to bin yields.%
+\footnote{%
+Because extrapolations from the variations are non-linear, the exact forms of
+the constraints do not directly matter.
+Numerical optimizers prefer Gaussian forms for their $\log$ quadratic shape, which is
+easily minimized in the common case that the parameter has no observable
+effect.%
+}
+
+Along with each \heplikelihood\ constraint, \histfactory\ also assigns a
+sampling distribution for alternative values of its (auxiliary) data.
+\TODO{
+toy procedure;
+not true $p$-value because it pre-fits;
+alternative approximation to asymptotics; not necessarily better;
+choose your poison}
 

### 2022-09-16 main/chapter_searches.tex
add words on toy sampling distributions

@@
 Other practical approaches explore parameter ranges in which the minimum energy
-exceeds thresholds, similar to Fisher's likelihood intervals.
+exceeds thresholds.
 
@@
 sampling distribution for alternative values of its (auxiliary) data.
-\TODO{
-toy procedure;
-not true $p$-value because it pre-fits;
-alternative approximation to asymptotics; not necessarily better;
-choose your poison}
+To approximate \pvalues\ without the asymptotic approximations,
+these distributions are used in a standard procedure for simulating
+alternative ``pseudo-data (toys)''~\cite{cern2011procedure}.
+This procedure works as follows:
+\begin{enumerate}
+\item fit to the observed data to find best-fit yields $\tilde{y}$, then
+\item sample alternative Poisson (actual) data with rates $\tilde{y}$ and
+alternative auxiliary (fictitious) data from their assigned distributions,
+then
+\item evaluate the test statistic for each simulated
+example~\cite{cern2011procedure}.
+\end{enumerate}
+Because this first item fits to the observed data, the test statistic
+distribution pathologically depends on the observed data $d$ ---
+the approximated \pvalue\ is not
+$\prob{t \geq t(d)}{h}$, but a depends on some function of the data
+$\prob{t \geq t(d)}{f(d, h)}$.
+And since the \pvalue\ is a censored likelihood, this is an error of early
+unblinding, or training on the test set in machine-learning speak.
+It can make little difference when the yields are insensitive to the observed
+data, but we shall see an example (of DR-Low) in
+Section~\ref{sec:2ljets_results} where they are sensitive, and where
+this makes a deficit in data appear much less significant that it should.
+
+Nonetheless, this procedure often simulates reasonable-looking test statistic
+distributions without relying on the asymptotic approximations.
+Toy simulation is an alternative approximation, and in given circumstances
+may be better or worse.
+Pick your poison.
 

### 2022-09-16 main/chapter_searches.tex
add note and ref for claimed good coverage

@@
 Nonetheless, this procedure often simulates reasonable-looking test statistic
-distributions without relying on the asymptotic approximations.
+distributions without relying on the asymptotic approximations, and purportedly
+with ``good coverage properties''~\cite{
+cern2011procedure,
+Cranmer2006Statistical
+}.
 Toy simulation is an alternative approximation, and in given circumstances

### 2022-09-16 main/chapter_searches.tex
autocheck tweaks

@@
 (strictly, a Radon-Nikodym derivative~\cite{billingsley2008probability})
-that constrains the (possibly un-normalized or `improper') prior measure that
+that constrains the (possibly non-normalized or `improper') prior measure that
 is uniform in $x$~\cite{Cowan:2010js}.
@@
 By assigning a form to the auxiliary likelihood, therefore, one is really
-choosing a \emph{regularizer} that should be indended to induce preferable
+choosing a \emph{regularizer} that should be intended to induce preferable
 behaviours in the parametric model and its `maximum likelihood' statistics.
@@
 the approximated \pvalue\ is not
-$\prob{t \geq t(d)}{h}$, but a depends on some function of the data
+$\prob{t \geq t(d)}{h}$, but depends on the data like
 $\prob{t \geq t(d)}{f(d, h)}$.
@@
 Section~\ref{sec:2ljets_results} where they are sensitive, and where
-this makes a deficit in data appear much less significant that it should.
+this makes a deficit in data appear much less significant than it should.
 

### 2022-09-16 main/chapter_searches.tex
comment out poison
additions on toy sampling; add some little todos

@@
 
+\TODO{cut this}
 Statistical data analysis suffers from dogma --- prescriptions and recipes that
@@
 
+\TODO{reduce this}
 In the current year, it would be remiss to not mention machine learning
@@
 histogram bins.
-They constrain parameters with either Gaussian or Poisson likelihood
+They constrain parameters with either Gaussian or Poisson%
+\footnote{%
+Poisson constraints typically use a continuous interpolation with
+$n!\rightarrow \Gamma(n + 1)$.%
+}
+likelihood
 functions, and assign expected yields through a variety of
@@
 this makes a deficit in data appear much less significant than it should.
+Sampling alternative data from exactly $\tilde{y}$, despite us actually being
+uncertain of the expected yields
+(demonstrated by us drawing error bars on them), is also suspicious.
 
 Nonetheless, this procedure often simulates reasonable-looking test statistic
-distributions without relying on the asymptotic approximations, and purportedly
-with ``good coverage properties''~\cite{
+distributions without relying on the asymptotic approximations, and reportedly
+has ``good coverage properties''~\cite{
 cern2011procedure,
@@
 may be better or worse.
-Pick your poison.
+% Pick your poison.
 

### 2022-09-16 main/chapter_searches.tex
add more compatibility cites; tidy up journals

@@
 HIGG-2018-51,
+HIGG-2019-14,
+EXOT-2016-36,
 EXOT-2018-08,
+EXOT-2019-15,
 HION-2018-19,
+chan2019search,
 mastrandrea2019searches,
@@
 IceCube2014searches,
-gerasimov2021new
+gerasimov2021new,
+roth2007fit
 },%

### 2022-09-17 main/chapter_searches.tex
tweak toy simulation commentary

@@
 \begin{enumerate}
-\item fit to the observed data to find best-fit yields $\tilde{y}$, then
-\item sample alternative Poisson (actual) data with rates $\tilde{y}$ and
+\item minimize $E(x)$ to find the best-fit parameters $\check{x}$, then
+\item sample alternative Poisson (actual) data with rates $y(\check(x))$ and
 alternative auxiliary (fictitious) data from their assigned distributions,
 then
-\item evaluate the test statistic for each simulated
-example~\cite{cern2011procedure}.
+\item evaluate the test statistic for each simulated `toy'
+sample~\cite{cern2011procedure}.
 \end{enumerate}
-Because this first item fits to the observed data, the test statistic
-distribution pathologically depends on the observed data $d$ ---
-the approximated \pvalue\ is not
+The \pvalue\ is then estimated as the fraction of toys with test statistics
+greater than or equal to the observed test statistic.
+Because this first item fits to the observed data, the sampled test statistic
+distribution depends on the observed data $d$ ---
+this problematic because it means that the approximated \pvalue\ is not
 $\prob{t \geq t(d)}{h}$, but depends on the data like
@@
 And since the \pvalue\ is a censored likelihood, this is an error of early
-unblinding, or training on the test set in machine-learning speak.
+unblinding, or, in machine-learning speak, training on the test set.
 It can make little difference when the yields are insensitive to the observed
@@
 Section~\ref{sec:2ljets_results} where they are sensitive, and where
-this makes a deficit in data appear much less significant than it should.
-Sampling alternative data from exactly $\tilde{y}$, despite us actually being
-uncertain of the expected yields
-(demonstrated by us drawing error bars on them), is also suspicious.
+this makes a deficit in data appear much less significant than it  otherwise
+would.
+Sampling alternative data from fixed $y(\check{x})$ is also suspicious because
+we are in fact uncertain of the expected yields
+(demonstrated by us drawing error bars on them).
 
 Nonetheless, this procedure often simulates reasonable-looking test statistic
-distributions without relying on the asymptotic approximations, and reportedly
+distributions without relying on asymptotic approximations, and reportedly
 has ``good coverage properties''~\cite{

### 2022-09-17 main/chapter_searches.tex
write up CR VR SR

@@
 \footnote{%
-Poisson constraints typically use a continuous interpolation with
+The Poisson likelihoods typically use a continuous interpolation with
 $n!\rightarrow \Gamma(n + 1)$.%
@@
 Nonetheless, this procedure often simulates reasonable-looking test statistic
-distributions without relying on asymptotic approximations, and reportedly
+distributions, does not rely on asymptotic approximations, and reportedly
 has ``good coverage properties''~\cite{
@@
 }.
-Toy simulation is an alternative approximation, and in given circumstances
-may be better or worse.
+Toy simulation is an approximation, and in given circumstances may be better or
+worse than the asymptotic formulae.
 % Pick your poison.
@@
 \label{sec:searches_searches}
-
-\TODO{CR VR SR}
+Searches for new effects from supersymmetric models in \atlas\ face the problem
+that their signals can appear in obscure, extreme, or otherwise less-studied
+corners of the data.
+In \atlas\ SUSY searches, our modelling of the  Standard Model is therefore
+treated with substantial suspicion, and we tune that modelling to better
+describe the backgrounds local to our data of interest before finalizing a
+search.
+Tuning and analysis are performed in a conventional and sensible procedure,
+but their pieces have somewhat esoteric names that we define here.
+
+\paragraph{CR, VR, SR.}
+\atlas\ SUSY searches see data only through discrete histogram bins,
+or `regions', which are defined by logical selections in event variables.
+To communicate their intended uses, we further categorize these bins into
+`control', `validation', and `signal' regions.
+
+\textbf{Control regions} (\textbf{CR}s), select large numbers data to
+constrain global normalization factors that scale the major
+Standard Model backgrounds;
+these normalization factors control the total cross-sections of
+background samples while preserving their relatives shapes between regions.
+Disregarding theoretical Standard Model cross-sections is justified, since
+the control region constraints are, by design, more precise than any previous
+estimate local to our regions.
+
+To represent the Standard Model, control regions must be insensitive to
+non-standard signals.
+Before applying their normalization factors to regions that are sensitive
+to signal hypotheses, data in \textbf{validation regions} (\textbf{VR}s) first
+test the quality of extrapolation.
+Each model is adjusted until validation region data are well predicted ---
+either by changing region definitions or restructuring the \heplikelihood\
+model, perhaps with extrapolation uncertainties or extra normalization
+factors.
+Likelihoods for validation region data are not included in the \heplikelihood\
+models, but their data are important and indirectly used to update the
+model in this manner.
+
+Consuming information from control and validation regions to refine the
+predictions of our background-only model is scientific learning, and can be
+reconciled with machine learning practice as fitting a model with training
+(CR) and validation (VR) data.
+Testing is performed through in prediction of data in the
+\textbf{signal regions} (\textbf{SR}s),
+in which signal yields should significantly add on top of the Standard Model
+backgrounds.
+Some signal regions are also labelled \textbf{DR} to indicate their use in
+specific `discovery' fits.
+
+\TODO{Figure to illustrate CR VR SR extrapolation}
+
+\paragraph{Fits.}
+\TODO{Background-only}
+
+\TODO{Exclusion (model-dependent signal}
+
+\TODO{Discovery (model-independent signal)}
 

### 2022-09-17 main/chapter_searches.tex
add energy-based modelling citations

@@
 pmlr-v2-ranzato07a,
-skilling2017david
+skilling2017david,
+lecun-06,
+pmlr-v2-ranzato07a
 }.

### 2022-09-17 main/chapter_searches.tex
change to paragraph for cr vr sr

@@
 
-\paragraph{CR, VR, SR.}
-\atlas\ SUSY searches see data only through discrete histogram bins,
+Our searches see data only through discrete histogram bins,
 or `regions', which are defined by logical selections in event variables.
@@
 
-\textbf{Control regions} (\textbf{CR}s), select large numbers data to
+\paragraph{Control regions (CRs).}
+Control regions select large numbers data to
 constrain global normalization factors that scale the major
@@
 
+\paragraph{Validation regions (VRs).}
 To represent the Standard Model, control regions must be insensitive to
@@
 
+\paragraph{Signal regions (SRs).}
 Consuming information from control and validation regions to refine the

### 2022-09-17 main/chapter_2ljets.tex
add note and ref for claimed good coverage
add cr vr sr figure; improve histfitter citation; add some notes

@@
 This requirement is typically justified by the limitations of approximate
-statistical methods that work
-``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
+statistical methods which work
+``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2015}.
 But the Poisson likelihood works for infinitesimally small
@@
 through the special parameter name
-\texttt{SigXSec}~\cite{histfitter2014, histfittergithub}.
+\texttt{SigXSec}~\cite{histfitter2015, histfittergithub}.
 

### 2022-09-17 main/chapter_searches.tex
add cr vr sr figure; improve histfitter citation; add some notes

@@
 To communicate their intended uses, we further categorize these bins into
-`control', `validation', and `signal' regions.
+`control', `validation', and `signal' regions;
+Figure~\ref{fig:searches_histfitter_regions} illustrates the uses of these
+three region types.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/searches_histfitter_regions.pdf}
+\caption[
+Control, validation, and signal regions
+]{%
+Control, validation, and signal regions and the extrapolation between them.
+Both observables are event variables in which models should change
+smoothly to justify the extrapolations that are indicated by arrows.
+Dashed lines illustrate the each region may be subclassified into
+disjoint (orthogonal) smaller bins.
+Adapted from \histfitter~\cite{histfitter2015}.
+}
+\label{fig:searches_histfitter_regions}
+\end{figure}
 
@@
 Before applying their normalization factors to regions that are sensitive
-to signal hypotheses, data in \textbf{validation regions} (\textbf{VR}s) first
+to signal hypotheses, data in validation regions first
 test the quality of extrapolation.
@@
 (CR) and validation (VR) data.
-Testing is performed through in prediction of data in the
-\textbf{signal regions} (\textbf{SR}s),
+Testing is done by predicting of data in the signal regions,
 in which signal yields should significantly add on top of the Standard Model
 backgrounds.
-Some signal regions are also labelled \textbf{DR} to indicate their use in
-specific `discovery' fits.
-
-\TODO{Figure to illustrate CR VR SR extrapolation}
-
-\paragraph{Fits.}
+Some signal regions are labelled as \textbf{discovery regions} (\textbf{DRs})
+to indicate their use in special `discovery' fits.
+
+\paragraph{Background-only fits.}
 \TODO{Background-only}
-
+% summary plots
+% fit to just CR, show VR and SR
+% conventional to also show SR
+% this is more like a posterior
+% so not easily interpreted
+
+\paragraph{Exclusion fits.}
 \TODO{Exclusion (model-dependent signal}
-
+% CLs
+
+\paragraph{Discovery fits.}
 \TODO{Discovery (model-independent signal)}
+% CLs and p-value
+% maximum likelihood ratio test statistics
+% despite likeliy superiority of $n$
 
@@
 taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.
+
+% tail area
+
+% maximum likelihood ratio
+
+% similarity and difference

### 2022-09-17 main/chapter_searches.tex
more bib tidying
type some on bkg-only and excl fits

@@
 observed data.
-This example illustrates \pvalues\ pathological dependency on tail selections,
-and the benefits of likelihood ratios.
+Because $m_{4\ell}$ would be a poor test statistic, this example illustrates
+\pvalues' pathological dependency on tail selections and the usefulness of
+likelihood ratios.
 }
@@
 }{
-\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
+\prob{t \geq \tilde{t}}{b}\hfill
 }
@@
 
+Various fits are useful to extract different results from search regions.
+We briefly describe the three classes of fit here, and shall see examples
+of their results in Chapter~\ref{chapter:2ljets}.
+
 \paragraph{Background-only fits.}
-\TODO{Background-only}
-% summary plots
-% fit to just CR, show VR and SR
-% conventional to also show SR
-% this is more like a posterior
-% so not easily interpreted
+With no signal contribution present, the background-only model is fitted
+to the control regions to make predictions in the signal regions.
+These background-only fits usefully validate that the \heplikelihood\ model
+works as intended: that its normalization factors are not too far from the
+Standard Model, that its tuned extrapolation to validation regions works,
+and that other parameters are behaving according to intuition.
+
+After fitting to observed signal region data, the background-only model
+it can be useful to check how its parameters have been `pulled' away from their
+central values.
+Post-fit expectations in signal regions are also available with error bars;
+because these depend on the data they are posterior results and not directly
+useful to judge the background-only model.
+Instead, they are an updated prediction for the expected yield in an
+hypothetical future repeat of the experiment.
+
+Searches for new signals require some concept of how their signals differ
+from the backgrounds;
+\atlas\ SUSY searches consider two such concepts with two flavours:
+exclusion fits, and discovery fits.
 
 \paragraph{Exclusion fits.}
+Model-dependent `exclusion' fits add simulated yields from precisely
+defined models to the \heplikelihood\ model.
+These fits can use multiple signal regions simultaneously, and couple the
+effect of parametrized modifications to the signals and backgrounds between
+the bins of those regions.
+% cls
+% parameter plane
+% exclusion contours
 \TODO{Exclusion (model-dependent signal}
@@
 \TODO{Discovery (model-independent signal)}
+
+model-independent `discovery' fits, which add a variable signal yield to one
+signal region in isolation.
+
 % CLs and p-value

### 2022-09-17 main/chapter_searches.tex
remove one energy clause

@@
 From such $\log$ ratios, a little rearrangement recovers the individual
-probabilities through the logistic function:
-\begin{equation}
-\label{eqn:searches_logistic}
-\prob{S}{d} = \frac{1}{1 + e^{-\Delta(d)}}
-.
-\end{equation}
-Physically, this is analogous to the Boltzmann distribution where the energy
-of each model is its negative $\log$ probability, with unit inverse
-temperature~\cite{
-pmlr-v2-ranzato07a,
-skilling2017david,
-lecun-06,
-pmlr-v2-ranzato07a
-}.
-With multiple models in play, it is also called the `softmax'
-function~\cite{MurphyKevinP.2012Mlap}.
+probabilities through the `logistic' function, also known as
+`softmax' if multiple hypotheses are in play~\cite{MurphyKevinP.2012Mlap}.
 
@@
 But there are examples of unambiguously derived that are commonly used in
-High-Energy Physics, notably the Poisson and Gaussian distributions.
+High Energy Physics, notably the Poisson and Gaussian distributions.
 To avoid turning this thesis into a textbook, we only sketch the derivations
@@
 with which a system collapses to its minimum energy at low temperature
-$T = 1/\beta$.
-At unit temperature, the Boltzmann factor can define a probability distribution
+$T = 1/\beta$~\cite{
+skilling2017david,
+lecun-06,
+pmlr-v2-ranzato07a
+}.
+At unit temperature, this can induce a probability distribution
 for the parameters relative to an appropriate (prior)
 measure~\cite{cranmer2012histfactory, skilling2017david},
-but in practice such distributions do not always coincide with intuition or
-with the best fit value.
+but in practice such distributions often coincide with neither intuition nor
+the best fit value.
 

### 2022-09-17 main/draft.tex
motivate cls and arrange some citations
begin detector chapter draft
remove one energy clause

@@
 }
+
+, and a small observed \pvalue\ does implies neither
+a preference for the signal hypothesis nor incompatibility with the background.
+
+% removed comment on jets vs muons --- think I discuss this elsewhere
+
+%% Jet / muon start of discussion sand
+% Jets are `messy', right?
+% Jets do tend to worse resolution --- they suffer from calorimeter noise
+% (which varies with statistical properties of hadronic showers
+% along with hardware effects), as well as variable
+% Jets include neutral hadrons which cannot be tracked, jet clustering may
+% incorrectly assign objects, and some hadrons can decay to neutrinos and soft
+% muons which would go undetected, and calorimetry is inherently variable.
+% This tendency, however, fails for very hard muons.
+
+\begin{equation}
+\label{eqn:searches_logistic}
+\prob{S}{d} = \frac{1}{1 + e^{-\Delta(d)}}
+.
+\end{equation}
+Physically, this is analogous to the Boltzmann distribution where the energy
+of each model is its negative $\log$ probability, with unit inverse
+temperature~\cite{
+skilling2017david,
+lecun-06,
+pmlr-v2-ranzato07a
+}.
+With multiple models in play, it is also called the `softmax'
+function.

### 2022-09-17 main/chapter_searches.tex
add words about discovery regions; very beginning of significance

@@
 the bins of those regions.
-% cls
-% parameter plane
-% exclusion contours
-\TODO{Exclusion (model-dependent signal}
-% CLs
+
+From exclusion fits, we extract statistical results with the $\cls$ method to
+mark individual models as excluded and set limits on their production
+cross-sections.
+For presentation on paper, articles consider two-dimensional planes of model
+parameters and present contiguous `exclusion contours' within which all models
+are marked as excluded at a given level.
 
 \paragraph{Discovery fits.}
-\TODO{Discovery (model-independent signal)}
-
-model-independent `discovery' fits, which add a variable signal yield to one
-signal region in isolation.
-
-% CLs and p-value
-% maximum likelihood ratio test statistics
-% despite likeliy superiority of $n$
+Model-dependence sacrifices some generality.
+Model-independent `discovery' fits are therefore performed to extract results
+that could be interpretable for any hypothesized model.
+These fits include the control regions and only one signal region, a
+discovery region, in which a flat signal yield is considered.
+The $\cls$ method is again used to set limits on that signal yield,
+and a `discovery' \pvalue\ is also computed with the intention to quantify
+the significance of a data excess were one observed.
+
+Despite the likely superiority of $t(n) = n$ test statistics for these
+single-bin signal-plus-background experiments, standard maximum likelihood
+ratio test statistics are used, and may explain quirks in the discovery
+region results that we shall see in the results of
+Chapter~\ref{chapter:2ljets}.
 
 \subsection{Gaussian significance}
-
-\TODO{significance}
+Standard-deviations distance is a more intuitive measure of compatibility than
+a \pvalue.
+
+\TODO{tail area}
+
+\TODO{log maximum likelihood ratio}
+
+\TODO{significance measures}
 
@@
 
-% tail area
-
-% maximum likelihood ratio
-
-% similarity and difference
+% similarity and difference?? TODO

### 2022-09-17 main/chapter_searches.tex
add some formulae for significance

@@
 \intertext{all of which we have named:}
-\underbrace{L(x)}_\textbf{likelihood}
+\underbrace{L(x)}_\textrm{likelihood}
 \times
 ~\,
-\overbrace{\pi(x)}^\textbf{prior}
+\overbrace{\pi(x)}^\textrm{prior}
 ~\,
 &=
-\overbrace{p(x)}^\textbf{posterior}
+\overbrace{p(x)}^\textrm{posterior}
 \times
-\underbrace{Z}_\textbf{evidence}
+\underbrace{Z}_\textrm{evidence}
 .
@@
 ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately
-strong  indication of roughly $3.7\sigma$.
-But a likelihood ratio on the test statistic itself, and not the inclusive
-tail statement, would encode strictly more information and could provide a
+strong indication.
+A likelihood ratio on the test statistic itself, and not the inclusive
+tail statement, would however encode more information and could provide a
 more precise answer.
@@
 
+defining $\chi = (x - \mu) / \sigma$, the Gaussian likelihood is
+
+\begin{equation}
+L(\chi) =
+\frac{1}{\sqrt{2\pi}}
+\exp\!\left(-\frac{1}{2}\chi^2\right)
+\end{equation}
+
+\TODO{log maximum likelihood ratio}
+$\log$ maximum likelihood ratio $\kappa$
+\begin{equation}
+\kappa = -\frac{1}{2}\chi^2
+\quad
+\Leftrightarrow
+\quad
+\chi = \pm\sqrt{-2\kappa}
+\end{equation}
+
 \TODO{tail area}
 
-\TODO{log maximum likelihood ratio}
-
+\begin{align}
+P(\chi)
+&=
+\frac{1}{\sqrt{2\pi}}
+\int_\chi^\infty
+\!
+\exp\!\left(-\frac{1}{2}t^2\right)
+\,\mathrm{d}t
+\\[0.5em]
+&=
+L(\chi)
+\times
+\left(
+\frac{1}{\chi}
++
+\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
+\right)
+\end{align}
 \TODO{significance measures}
+~\cite{
+jaynes2003probability,
+abramowitz1965handbook
+}
 
 % significance measures
-The ATLAS-recommended significance measure~\cite{atlas_significance} is
+ATLAS' recommended significance measure~\cite{atlas_significance} is
 \begin{align}

### 2022-09-17 main/chapter_searches.tex
remove significance section; this is already far too long; some tweaks

@@
 
-How can one escape this propaganda?
+How can one escape this trap?
 Likelihood ratios on the actual data escape immediately by bypassing the
-censor.
-Even inequalities need not be biased --- if taken against a predetermined
-threshold, say $5$ events, then models can predict whether $n \leq 5$,
-and its truth or falsehood in data can favour either side.%
+censor, but even inequalities need not be biased --- if taken against a
+predetermined threshold, say $5$ events, then models can predict whether
+$n \leq 5$, and the truth or falsehood in data can favour either side.%
 \footnote{%
@@
 }
-Finally, a reader can fall back to not trusting the reporter ---
-the statements `\atlas\ reports $n \leq 5$' and logical `$n \leq 5$' are not
-identical, since one can infer from the former that in fact $n = 5$.
-One knows what the message says, but not that it states the whole truth.
+Ultimately, the reader can fall back to not trusting the reporter ---
+the statements `\atlas\ reports $n \leq 5$' and `logically, $n \leq 5$' are not
+identical, since one might infer from the former that in fact $n = 5$.
+The message itself is data --- what it says may not be the whole truth.
 
@@
 Chapter~\ref{chapter:2ljets}.
-
-\subsection{Gaussian significance}
-Standard-deviations distance is a more intuitive measure of compatibility than
-a \pvalue.
-
-defining $\chi = (x - \mu) / \sigma$, the Gaussian likelihood is
-
-\begin{equation}
-L(\chi) =
-\frac{1}{\sqrt{2\pi}}
-\exp\!\left(-\frac{1}{2}\chi^2\right)
-\end{equation}
-
-\TODO{log maximum likelihood ratio}
-$\log$ maximum likelihood ratio $\kappa$
-\begin{equation}
-\kappa = -\frac{1}{2}\chi^2
-\quad
-\Leftrightarrow
-\quad
-\chi = \pm\sqrt{-2\kappa}
-\end{equation}
-
-\TODO{tail area}
-
-\begin{align}
-P(\chi)
-&=
-\frac{1}{\sqrt{2\pi}}
-\int_\chi^\infty
-\!
-\exp\!\left(-\frac{1}{2}t^2\right)
-\,\mathrm{d}t
-\\[0.5em]
-&=
-L(\chi)
-\times
-\left(
-\frac{1}{\chi}
-+
-\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
-\right)
-\end{align}
-\TODO{significance measures}
-~\cite{
-jaynes2003probability,
-abramowitz1965handbook
-}
-
-% significance measures
-ATLAS' recommended significance measure~\cite{atlas_significance} is
-\begin{align}
-\label{eqn:significance_atlas}
-S_\mathrm{\atlas}(n; \mu, \sigma) =~&
-\sqrt{2} \times
-\mathrm{sign}(n - \mu) \times
-\\[0.2em] \nonumber
-&
-\sqrt{
-n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
-- \frac{\mu^2}{\sigma^2}\log\left(
-1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
-\right)
-}
-,
-\end{align}
-taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.
-
-% similarity and difference?? TODO

### 2022-09-17 main/draft.tex
remove significance section; this is already far too long; some tweaks

@@
 function.
+
+% whooo removing a section
+
+\subsection{Gaussian significance}
+Gaussian standard-deviations distance is an intuitive measure of compatibility
+with a prediction.
+than
+a \pvalue.
+
+defining $\chi = (x - \mu) / \sigma$, the Gaussian likelihood is
+
+\begin{equation}
+L(\chi) =
+\frac{1}{\sqrt{2\pi}}
+\exp\!\left(-\frac{1}{2}\chi^2\right)
+\end{equation}
+
+\TODO{log maximum likelihood ratio}
+$\log$ maximum likelihood ratio $\kappa$
+\begin{equation}
+\kappa = -\frac{1}{2}\chi^2
+\quad
+\Leftrightarrow
+\quad
+\chi = \pm\sqrt{-2\kappa}
+\end{equation}
+
+\TODO{tail area}
+
+\begin{align}
+P(\chi)
+&=
+\frac{1}{\sqrt{2\pi}}
+\int_\chi^\infty
+\!
+\exp\!\left(-\frac{1}{2}t^2\right)
+\,\mathrm{d}t
+\\[0.5em]
+&=
+L(\chi)
+\times
+\left(
+\frac{1}{\chi}
++
+\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
+\right)
+\end{align}
+\TODO{significance measures}
+~\cite{
+jaynes2003probability,
+abramowitz1965handbook
+}
+
+% significance measures
+ATLAS' recommended significance measure~\cite{atlas_significance} is
+\begin{align}
+\label{eqn:significance_atlas}
+S_\mathrm{\atlas}(n; \mu, \sigma) =~&
+\sqrt{2} \times
+\mathrm{sign}(n - \mu) \times
+\\[0.2em] \nonumber
+&
+\sqrt{
+n\log\left(\frac{n(\mu + \sigma^2)}{\mu^2 + n\sigma^2}\right)
+- \frac{\mu^2}{\sigma^2}\log\left(
+1 + \frac{\sigma^2(n - \mu)}{\mu(\mu + \sigma^2)}
+\right)
+}
+,
+\end{align}
+taken in a limit where $0\log(0x) = 0$ for the $n=0$ case.
+
+% similarity and difference?? TODO

### 2022-09-17 main/chapter_searches.tex
add ending comment
tweaks with autochecker from overleaf

@@
 $\bar{\sigma}(d)$~\cite{student1908, fisher1925t}.
-By dividing out a (scale-equivariant) estimate of the standard deviation, $t$
+By dividing out a (scale equivariant) estimate of the standard deviation, $t$
 is independent of common scale factors on the data, and that scale invariance
@@
 meme~\cite{dawkins1989selfish} in High Energy Physics ~\cite{
+roth2007fit,
 HIGG-2017-09,
@@
 IceCube2014searches,
-gerasimov2021new,
-roth2007fit
+gerasimov2021new
 },%
@@
 where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$; $b = 4.3$ and $s = 3.0$.
-This test statistic jumbles up the data into an that makes $n=11$ appear to be
-weaker support for the signal than $n=9$.
+This test statistic jumbles the data into an order that makes $n=11$ appear to
+be weaker support for the signal than $n=9$.
 Each $\cls$ is \pvalue\ ratio.
@@
 }
-Ultimately, the reader can fall back to not trusting the reporter ---
+Ultimately, the reader just distrust the reporter ---
 the statements `\atlas\ reports $n \leq 5$' and `logically, $n \leq 5$' are not
@@
 Toy simulation is an approximation, and in given circumstances may be better or
-worse than the asymptotic formulae.
+worse than the asymptotic approximations.
 % Pick your poison.
@@
 smoothly to justify the extrapolations that are indicated by arrows.
-Dashed lines illustrate the each region may be subclassified into
+Dashed lines illustrate that each region may be subdivided into
 disjoint (orthogonal) smaller bins.
@@
 \paragraph{Control regions (CRs).}
-Control regions select large numbers data to
+Control regions select large numbers of data to
 constrain global normalization factors that scale the major
@@
 
-After fitting to observed signal region data, the background-only model
+After fitting the background-only model to observed signal region data,
 it can be useful to check how its parameters have been `pulled' away from their
@@
 Post-fit expectations in signal regions are also available with error bars;
-because these depend on the data they are posterior results and not directly
+because these depend on the data, they are posterior results and not directly
 useful to judge the background-only model.
@@
 Chapter~\ref{chapter:2ljets}.
+
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-09-18 main/chapter_searches.tex
kern pvalue; first wave of review edits

@@
 
-Our task in scientific data analysis is to distil an informative
-message from data and to report that message to the wider community.
-From any given data there are many possible messages that could be extracted,
-and those messages will have different values to different recipients of our
-messages --- our data analysis is not an automatic process, but a result
-of many judgements that are informed by theory, convention, and intuition.
-
-To clarify our later description of the $\twoljets$ search in
+Our task in scientific data analysis is to distil useful messages from the
+data, and to report that message to the wider community.
+Many different messages could be extracted from given data,
+and those messages will have various values to their recipients.
+Data analysis is not an automatic process, but a design with many judgements
+that are informed by theory, intuition, and convention to give value to their
+results.
+
+To clarify our later description of the \atlas\ SUSY $\twoljets$ search in
 Chapter~\ref{chapter:2ljets}, this chapter introduces the basic theory and
-practice of data analysis and its manifestations in the \atlas\ SUSY community.
+practice of data analysis and its conventional form in the \atlas\ SUSY
+community, parts of which can only be explained in their historical context.
 
@@
 Searches for Supersymmetric signals in \atlas\ follow a standard strategy for
-their data analysis; its procedures, nomenclature, and parametric modelling
+their data analysis, whose procedures, nomenclature, and parametric modelling
 are introduced in Section~\ref{sec:searches_searches}.
@@
 Your answer depends not only on what you observe from the figure, but also on
-what else you know about cartoon jars of beans, and on what reward is on offer.
-Each answer is a decision that depends both on what the possible rewards are,
-and on which numbers of beans you subjectively infer to be likely.
-We shall return to this example after developing some probability and decision
-theory.
+what else you know about cartoon jars of beans, and on what reward is offered.
+Each answer is a decision, which depends both on how you value the rewards,
+and on which numbers of beans you subjectively infer to be probable.
+We shall return to this example after developing its analysis in probability
+and decision theory.
 
@@
 
-\TODO{cut this}
 Statistical data analysis suffers from dogma --- prescriptions and recipes that
-students are taught to memorize and follow without knowing their reasons or
-limitations.
-From a standard undergraduate physics education, my experience has been
+one can memorize and follow while knowing neither their reasons nor their
+limitations:
+among these are
 error propagation formulae,
@@
 chi-squared per degree of freedom,
-and myriad statistical tests and confidence intervals.
-Many of these are effective approximations, but all also fail if misused
-outside their domain.%
+and myriad statistical tests with \pvalues\ and confidence intervals.
+Each of these can be an effective, but every one fail if misused outside
+its domain.%
 \footnote{%
-Multiplicative error propagation is an example.
-For $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
-the standard perturbative approximation for their product is
+Multiplicative error propagation is an example:
+for $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
+the standard rule says
 \(
@@
 \)
-if $a$ and $b$ are independent.
-Since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$, the
-propagation formula works only for small relative errors.
-}
-All effective methods of data analysis can be derived upwards from the roots of
-probability and decision theory, and as elsewhere in physics, by knowing the
-derivations we can also know when each approximation should work and when it
-needs replacing.
+if $a$ and $b$ are independent;
+s ince $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$,
+the propagation formula works only for small relative errors.
+}
+All effective methods of data analysis can be justified in probability and
+decision theory.
+As throughout physics, knowing a method's derivation allows one to judge when
+its approximations should hold and when it leaves room for improvement.
+
+Probabilistic data analysis follows the ordinary structure of science:
+write down predictive models, test those predictions against data,
+then update your models (or invent new ones) to refine their predictions
+to accommodate what has been learned.
+Likelihood functions are particularly important to this process because they
+describe our predictions.
+But as we shall see, not everything is a likelihood.
 
@@
 proportion of stuff within a larger whole.
-Probabilities useful because proportions have applications, and because they
-can be manipulated and interrelated with a linear and intuitive algebra.
-Probabilistic data analysis follows the ordinary structure of science ---
-write down models that make predictions, test those predictions against data,
-then update your models to refine their predictions.
-For this process, likelihoods are particularly important because they describe
-our predictions. But not everything is a likelihood.
+Probabilities are useful because proportions have applications
+(particularly to logical reasoning with uncertainty), and because
+they can be manipulated and interrelated with an intuitive linear algebra.
 The remainder of this section uses a standard example from High Energy Physics
+--- a histogram ---
 to introduce this probabilistic structure and other parts of its technical
@@
 \subsection{Probability, abstract}
-A probability $\prob{b}{a}$ is a number that quantifies the proportion
-of $b$ within $a$, and can be expressed as a ratio~\cite{axioms1010038}
+A probability $\prob{b}{a}$ is a number, which quantifies the proportion
+of $b$ within $a$ and can be expressed as a ratio~\cite{axioms1010038}
 \begin{equation}
@@
 (logical and or set intersection or any other equivalent operation).
-To continue that example, $\prob{b}{a}$ could be a proportion of tea within
-a spoon within a cup, where $m(b, a)$ measures the volume of tea in a
-teaspoon that is partially submerged in the cup.%
-\footnote{%
-All probabilities are conditional in this definition; $\probc{b}$ alone does
-not exist unless some conditioning information is implicit from context.%
-}
-
-Logic is an important application of probability, for which $a$ and $b$ are
-interpreted as logical propositions and
-$\prob{b}{a}$ is the degree to which $a$ implies $b$, with the limiting cases
-that $\prob{b}{a} = 1$ when $a$ implies $b$, and $\prob{b}{a} = 0$ when $a$
-contradicts $b$.
-Indeed, Cox's theorem derives the laws of probability by requiring compatibility
+For example, $\prob{b}{a}$ could quantify the proportion of tea within
+a spoon within a cup --- then $m(a)$ measures the content of the cup, and
+$m(b, a)$ measures the amount of tea in the teaspoon that is partially
+submerged in the cup.
+
+Note that, in this definition, all probabilities are conditional.
+Alone, no $\probc{b}$ exists unless some conditioning information is implicit
+from context.
+
+Logic is an important application of probability.
+To apply to logical reasoning, $a$ and $b$ are logical propositions and
+$\prob{b}{a}$ quantifies the degree to which $a$ implies $b$, with the
+limiting cases that $\prob{b}{a} = 1$ when $a$ implies $b$,
+and $\prob{b}{a} = 0$ when $a$ contradicts $b$.
+Indeed, Cox's theorem shows that probability algebra is uniquely consistent
 with Boolean logic~\cite{
@@
 Probability \emph{distributions} $\prob{b}{a}$ spread unit probability mass
-across states $b$, and need not be random distributions of fluctuating
-$b$s~\cite{jaynes2003probability,frankfurt2005on}, although random processes
-are useful applications.
+across states $b$.
+They and need not be random distributions of fluctuating $b$s~\cite{
+jaynes2003probability,
+frankfurt2005on
+},
+although random processes are another useful applications.
 
@@
 \intertext{%
-that describes nested proportions
-(perhaps a spoon within a teacup within a sink),
+describing nested proportions
+(a spoon in a teacup in a sink),
 and a \textbf{sum rule}%
 }
-\prob{c\vee b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
+\prob{c\;\textrm{or}\;b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
 \label{eqn:searches_sum_rule}
 \end{align}
-that describes combination in a union $\vee$ that coincides with logical `or'.
-The subtraction avoids double-counting of any overlap, and is often avoided by
-splitting any problem into orthogonal, or disjoint states.
-For disjoint cases with labels $b_i$, normalization therefore requires that
+describing combination in a union;
+the subtraction avoids double-counting of any overlap and is usually avoided by
+splitting any problem into disjoint, orthogonal states.
+For disjoint cases with labels $b_i$, normalization requires that
 \begin{equation}
@@
 \end{equation}
-These rules are basic and well known, but we state them here to support the
-following statements of how probabilities can be applied for practical
-results.%
+These basic rules are well known, but we state them here to support the
+following statements on how probabilities can applied for practical results.%
 \footnote{%
@@
 transformation~\cite{axioms1010038}.
-This choice is usually best because it uses ordinary addition and
-multiplication.
-In numerical analysis, however, it is often practical to use $\log$
-probabilities to avoid overflows outside the range representable by floating
-point numbers, thus changing multiplication to addition and addition to
-the `$\log$-add-$\exp$' operation.
+This choice is often best because it uses ordinary addition and
+multiplication, but in numerical analysis $\log$ probabilities can avoid
+overflows outside the ranges representable by floating-point numbers,
+thus changing multiplication to addition and addition to the
+`$\log$-add-$\exp$' operation.%
 }
@@
 \label{sec:searches_probability_applied}
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/searches_abacus_likelihood.pdf}
-\caption{%
-Likelihood: $\mathrm{Poisson}(0\mid x)$ in
-(left) $x$ and
-(right) $u = \log x$%
-}
-\end{subfigure}
-\\[.5em]
-\begin{subfigure}{\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/searches_abacus_prior.pdf}
-\caption{%
-Prior: exponential density in
-(left) $x$ and
-(right) $u = \log x$%
-}
-\end{subfigure}
-\caption[
-Likelihoods and prior densities' transformations with coordinates.
-]{%
-Likelihoods and prior densities' transformations with coordinates.
-The $\mathrm{Poisson}(0\mid x)$ likelihood and exponential $e^{-x}\mathrm{d}x$
-prior have the same form (left), but the prior density changes to preserve its unit
-area relative to $u = \log x$.
-On changing to $\log$ coordinates (right), the likelihood slides horizontally,
-but the prior density scales to preserve area.
-Both maximum likelihood points are equal, but the maximum prior density moves.
-\\[.5em]
-Particle lifetimes follow exponential distributions and their maximum
-densities are at zero time, but it is incorrect to describe their most probable
-lifetimes as zero ---
-the `most probable' $\log$ lifetime is not $\log 0$, but equals the mean
-lifetime.
-\\[.5em]
-For zero Poisson data, the maximum likelihood rate is zero, but zero might not
-be the most probable rate (if, for example, an experiment has known backgrounds).
-}
-\label{fig:searches_abacus}
-\end{figure}
-% science snake
 Science involves making predictions, testing those predictions against data,
@@
 for fixed $a$, which is a normalized probability distribution over $b$.
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_abacus_likelihood.pdf}
+\caption{%
+Likelihood: $\mathrm{Poisson}(0\mid x)$ in
+(left) $x$ and
+(right) $u = \log x$%
+}
+\end{subfigure}
+\\[.5em]
+\begin{subfigure}{\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/searches_abacus_prior.pdf}
+\caption{%
+Prior: exponential density in
+(left) $x$ and
+(right) $u = \log x$%
+}
+\end{subfigure}
+\caption[
+Likelihoods and prior densities' transformations with coordinates.
+]{%
+Likelihoods and prior densities' transformations with coordinates.
+The $\mathrm{Poisson}(0\mid x)$ likelihood and exponential $e^{-x}\mathrm{d}x$
+prior have the same form (left), but the prior density changes to preserve its unit
+area relative to $u = \log x$.
+On changing to $\log$ coordinates (right), the likelihood slides horizontally,
+but the prior density scales to preserve area.
+Both maximum likelihood points are equivalent, but the maximum prior density
+moves.
+\\[.5em]
+Particle lifetimes follow exponential distributions and their maximum
+densities are at zero time, but it is incorrect to describe their most probable
+lifetimes as zero ---
+the `most probable' $\log$ lifetime is not $\log 0$, but turns out to coincide
+with the mean lifetime.
+\\[.5em]
+For zero Poisson data, the maximum likelihood rate is zero, but zero might not
+be the most probable rate (if, for example, an experiment has known backgrounds).
+}
+\label{fig:searches_abacus}
+\end{figure}
 
@@
 Nonetheless, this procedure often simulates reasonable-looking test statistic
-distributions, does not rely on asymptotic approximations, and reportedly
+distributions, does not rely on asymptotic approximations, and purportedly
 has ``good coverage properties''~\cite{

### 2022-09-18 main/chapter_searches.tex
add CL glossary entry; minor tweaks

@@
 and myriad statistical tests with \pvalues\ and confidence intervals.
-Each of these can be an effective, but every one fail if misused outside
-its domain.%
+Each of these can be a useful approximation, but every one can also fail if
+misused outside its domain.%
 \footnote{%
@@
 if $a$ and $b$ are independent;
-s ince $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$,
+since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$,
 the propagation formula works only for small relative errors.
@@
 
-Confidence intervals are confusing.
+Confidence intervals are confusing.%
+\footnote{%
+Exemplifying such confusion,
+the \atlas\ Glossary of Terms~\cite{atlas2022glossary} states that the
+``Confidence Level (CL) is a statistical measure of the percentage of test
+results that can be expected to be within a specified range.''
+In fact it is the percentage of ranges
+(confidence intervals, which are test results)
+that should cover a true value.%
+}
 It is easy to be tricked into thinking that the confidence interval for given
@@
 There also exists a notion that a small \pvalue\ means that the data
-represent a rare fluctuation~\cite{murray1997use, atlas2022pvalue}.%
+represent a rare fluctuation~\cite{murray1997use, atlas2022glossary}.%
 \footnote{%
-The \atlas\ Glossary of Terms~\cite{atlas2022pvalue} falsely states that
+The \atlas\ Glossary of Terms~\cite{atlas2022glossary} falsely states that
 ``The \textbf{p-value} is the probability (ranging from $0$ to $1$) that the

### 2022-09-18 main/chapter_searches.tex
more confidence interval tweaks

@@
 }
-It is easy to be tricked into thinking that the confidence interval for given
-data `probably' contains the unknown parameter, but that gets it backwards.
-For a known true parameter $x$, the confidence intervals in random repeated
-experiments will contain that $x$ with probability $\gamma$;
+It is easy to stumble into thinking that a confidence interval
+`probably' contains the unknown parameter, but that gets it backwards.
+For a known true parameter $x$, the confidence intervals of random repeated
+experiments should contain $x$ with probability $\gamma$;
 the interval makes a subtle statement about unknown alternative data for
 known true parameters.
-Confidence intervals can even exclude all plausible parameters~\cite{
+For given data, the confidence interval can even exclude all possible
+parameters~\cite{
 pratt1961testing,
@@
 }.
-Pratt eloquently explains the correct interpretation of a confidence interval
-by example:
+Pratt succinctly explains the confidence interval by example:
 \begin{displayquote}
@@
 \end{displayquote}
-Neyman and Pearson never claimed otherwise; they did precise mathematics, and
-confusion about their intervals arises externally~\cite{jaynes2003probability}.
-
-Confidence intervals are functions of the data, so can inform vague inferences
-if used as data for likelihood functions, just as we did to interpret Fisher's
-\pvalue\ ratio as a likelihood ratio ---
+Neyman and Pearson never claimed otherwise; they did precise and rigorous
+abstract mathematics, and confusion about their intervals arises
+externally~\cite{jaynes2003probability}.
+
+Confidence intervals are functions of the data, so they can inform vague
+inferences if used as data for likelihood functions, just as we did to
+interpret Fisher's \pvalue\ ratio as a likelihood ratio ---
 an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
@@
 remains simpler and clearer.
-In response to the idea that no conclusion can be drawn about the parameter
-itself, Fisher also deserves the last word:
+In response to the idea that no conclusion can be drawn about the parameter,
+Fisher also has strong words here:
 \begin{displayquote}
@@
 \end{displayquote}
+Amen.
 

### 2022-09-18 main/chapter_searches.tex
more proofread edits

@@
 This choice is often best because it uses ordinary addition and
-multiplication, but in numerical analysis $\log$ probabilities can avoid
-overflows outside the ranges representable by floating-point numbers,
+multiplication, but $\log$ probabilities can avoid overflows outside the ranges
+representable by floating-point numbers in numerical analysis,
 thus changing multiplication to addition and addition to the
@@
 and reporting what is learned from the results.
-Various probabilities are involved in this prediction and learning that have
-conventional names to communicate what we intend to do with them.
+Various probabilities are involved in this prediction and learning, and those
+probabilities have conventional names that communicate our intended use for
+them.
 Most important for prediction is a \textbf{likelihood}
 \begin{align}
-L(a) &= \prob{b}{a} \\
 \label{eqn:searches_likelihood}
+L(a) &= \prob{b}{a}
 \intertext{%
-that explores the probabilities assigned to a fixed result $b$ from different
-contexts $a$, or how well different hypotheses predicted a result.
+which explores the probabilities assigned to a fixed result $b$
+(usually data) from different contexts $a$ (hypotheses).
 A likelihood is clearly a probability, but it is not a probability distribution
@@
 \end{align}
-for fixed $a$, which is a normalized probability distribution over $b$.
+normalized probability distribution over results $b$ for a fixed context $a$.
 
@@
 Likelihoods are not densities in the conditioning variable that they explore,
-so changing coordinates slides them horizontally like beads on an abacus,
-but does not change their values for equivalent states.
+so changing coordinates instead slides them horizontally like beads on an
+abacus, and does not change their values for equivalent states.
 Figure~\ref{fig:searches_abacus} illustrates this difference.
 
-Likelihood and prior are evidently related --- they explore two indices into
-the same object, two sides of the same coin, and are assigned by the same
-principles.
-But they are profoundly different objects that work for different purposes,
-and clarity in their distinction will help in our later understanding of
+Likelihood and prior are evidently related; they explore two indices into
+the same object --- two sides of the same coin --- and they are both assigned
+by the same principles.
+But they are profoundly different objects, which are applied for different
+purposes ---
+clarity in their distinction will help in our later understanding of
 practical methods.
-We use likelihoods to compare which hypotheses predicted observed data better
-than others.
-Prior to observing data we use priors to describe what might result from
-different models; priors on data can be useful in designing our experiments,
-and priors on other states describe other predictions.
+Likelihoods are useful to compare which hypotheses predicted observed data
+better than others.
+Priors on hypotheses express uncertain predictions, and
+priors on data are useful in experimental design
 
@@
 Since these histograms are normalized to unit area, they show prior
-probability distributions along the bin axis; there is one prior
-for events from each of the background and signal hypotheses.
-Priors distribute unit probability explore along the bin axis.
+probability distributions along the bin axis.
+There is one prior for events from each of the background and signal
+hypotheses ---
+priors distribute unit probability explore along the bin axis.
 In each bin, however, a likelihood function compares the assignments from
-the two hypotheses: signal and background.
-Likelihoods explore vertically down the sample axis.
+the two hypotheses: signal and background ---
+likelihoods explore vertically down the sample axis.
 
@@
 
-Identical distributions are drawn in two different coordinates
+Both distributions are equivalently drawn in two different coordinates
 on the left and right of Figure~\ref{fig:searches_sig_bkg_prior_likelihood};
-they are identical in content, but differ in appearance because they use the
-standard approach of representing each prior distribution as a density.%
+they are identical in content, but differ in appearance because they represent
+each prior distribution as a density in the standard way.%
 \footnote{%
-Density is a somewhat misleading term here.
-Probability mass does not compress, it moves and redistributes --- it is more
-like depth of tea than density of air.%
-}
+Density is a somewhat misleading term here ---
+probability mass is incompressible; it moves and redistributes more like
+depth of water than density of air.%
+}
+
 To preserve the area of each bin, which shows its probability mass,
@@
 differing values of $\met$ and $\sqrt{\met}$.
-This coordinate-dependence of the prior density means that its maximum is
-not always meaningful.
-Since the probability mass in each bin is constant, the likelihood is
+This coordinate-dependence of the prior density means that its point of
+maximum density is not always meaningful.
+But since the probability mass in each bin is constant, the likelihood is
 coordinate-independent.
 
-One typically approximates such histograms by Monte Carlo simulation methods,
-since simulations do not generate bin indices but event variables that derive
-from the sampled events.
-To connect these concept with probabilities, we begin with some notation;
+Standard methods approximate such histograms from Monte Carlo simulations,
+which do not directly indicate bin indices but state the values for event
+variables.
+To connect event variables to bin indices, we begin with some notation:
 we have hypotheses $h$ (signal or background),
-a parameter $x$ that we histogram ($x = \met$),
-and we observe a datum $d$ that datum indexes the histogram bins --- supposing
-an event is observed in the overflow bin, it says
+an event variable $x$ that we histogram ($x = \met$),
+and we observe a datum $d$ that indexes the histogram bins.
+Supposing, for example, that an event is observed in the overflow bin, it says
 $d = `x > 100\,\eV[G]\textrm{'}$.
@@
 \right.
+,
 \end{equation}
-that corresponds to Equation~\ref{eqn:searches_likelihood} by
+which corresponds to Equation~\ref{eqn:searches_likelihood} by
 $b \leftarrow d$ and $a \leftarrow `x, h\textrm{'}$,
-and simplifies $\prob{d}{x, h} = \prob{d}{x}$ since such binning is independent
-of hypothesis.%
+and simplifies by $\prob{d}{x, h} = \prob{d}{x}$ because at given $x$ the
+binning is independent of any origin from signal or background.%
 \footnote{%
@@
 This example of a binary selection highlights how the likelihood acts as a
-filter that rejects events that are contradicted by the data.
+filter that only rejects events that are contradicted by the data.
 Other likelihoods that are not binary act similarly, but as inefficient filters
@@
 Simulating events from background and signal hypotheses samples from their
-respective priors
-\begin{equation}
-\pi(x) = \prob{x}{h}
-\end{equation}
-that correspond to Equation~\ref{eqn:searches_prior} by
+respective priors $\pi(x) = \prob{x}{h}$, which
+correspond to Equation~\ref{eqn:searches_prior} by
 $b \leftarrow x$ and $a \leftarrow h$.
@@
 When we fill a histogram by summing weights of simulated samples, we are
-calculating another likelihood over hypotheses $h$ alone, not $`x, h\textrm{'}$
-jointly.
-To disambiguate the two, we call this abstracted, $x$-independent likelihood
-the \textbf{evidence} $Z(h)$ for hypothesis $h$.
-The evidence is calculable as
+calculating another likelihood over hypotheses $h$ alone in the bin,
+independent of, by averaging the likelihood over the prior.
+This abstracted, $x$-independent likelihood is labelled the \textbf{evidence}:
 \begin{align}
@@
 skilling2010foundations
-}, and relates most closely to our practice of simulating from the model and
-counting events that hit the non-zero likelihood in the bin.
-We shall shortly derive this identity.
+},
+can be imagined as simulating from the model $\mathrm{d}\pi$ and
+counting events that hit pass the likelihood filter.
 The name `evidence' relates to its application for model
 comparison~\cite{mackay2003information}.
-Supposing we have this one event with $\met > 100\,\eV[G]$, did it come from
-the background or signal sample?%
+
+Suppose that we have this one event with $\met > 100\,\eV[G]$ ---
+did it come from the background or signal?%
 \footnote{%
@@
 
-This expression as an integral --- an average over the prior, means that
-the evidence $Z$ depends on what fraction of samples hits large values of
-the likelihood function, and not only on how large that likelihood gets.
-In contrast, one could consider a maximum likelihood approach that would
-constrain a numerical optimization to regions of non-zero prior in each model.
-Such optimizations can give useful upper bounds on the evidence, but this
-example highlights a flaw when both models support the $\met > 100\,\eV[G]$
-bin.
-Then, both support the same maximum likelihood of
-$L(\varepsilon + 100\,\eV[G]) = 1$, despite different evidence values.
+Expressing the evidence $Z$ as an average over the prior highlights how it
+depends both on how large that likelihood gets,
+and on what fraction of samples hit large likelihood values.
+Maximum likelihood approaches,
+which compare models by the largest likelihoods their priors support,
+can be practical and give useful upper bounds on the evidence.
+This histogram example, however, highlights a limitation since both models
+support the same maximum likelihood of
+$L(\varepsilon + 100\,\eV[G]) = 1$.
+These evidence values here differ only by their prior masses in that bin.
 
 % posterior
-Perhaps we now want to see where in the overflow bin this event landed, and
+Perhaps we now wanted to see where in the overflow bin this event landed, and
 compare the signal and background distributions there.
-This is an inspection of the \textbf{posterior}
+That would be inspection of the \textbf{posterior}
 \begin{equation}
-p(x) = \prob{x}{d, h} = \frac{L(x)\,\pi(x)}{Z},
+p(x) = \prob{x}{d, h} = \frac{L(x)}{Z}\pi(x),
 \end{equation}
@@
 it simply uses the likelihood to filter simulated samples
-(perhaps in `skimming an ntuple')
-and re-normalizes them to a probability distribution within the constraint.
-
-We finally have the four parts of `Bayesian' data analysis:
+(like `skimming an ntuple'),
+then re-normalizes them to a probability distribution within the constraint.
+
+We now have the four parts of `Bayesian' data analysis:
 the two inputs \textbf{likelihood} and \textbf{prior},
@@
 commutativity of `and' --- that `$a$ and $b$' is equivalent to
-`$b$ and $a$';
-the joint probability of data and parameters has two identical
-forms, $\prob{d, x}{h}$ and $\prob{x, d}{h}$, that each factor into two
+`$b$ and $a$'.
+The joint probability of data and parameters has two identical
+forms, $\prob{d, x}{h}$ and $\prob{x, d}{h}$, which each factor into two
 pieces:
@@
 \intertext{all of which we have named:}
+\hphantom{\quad \quad \mid\mid \mathrm{context}}
 \underbrace{L(x)}_\textrm{likelihood}
@@
 \underbrace{Z}_\textrm{evidence}
+\quad \quad \mid\mid h
 .
@@
 
-How Bayes' rule \emph{should} be used in scientific data analysis is to extract
-useful information from the data.
-If the likelihood function itself is interpretable, then it should be reported
-directly.
-When it is not, perhaps due to a complicated parametric modelling, then one
-should abstract those parameters away to summarize the likelihood.
-Properties such as local maxima or regions in which it exceeds meaningful
-thresholds can be useful.
-Evidence values for predictive models can also be useful if those models are
-clearly stated and relevant to readers' interests.
-Data act only through the likelihood function, and reporting a posterior
-can act to obfuscate those data.
-
-Posterior probabilities are, however, useful when digesting data to make
-inferences about the world.
-Reporting likelihood functions and evidence values allows readers to assign
-priors as they wish.
-For model comparison, for example between background $h_1=B$ and signal $h_2=S$
-models, the analysis is conveniently linearized by expressing it in $\log$
-ratio form:
+Data only enter the inputs to Bayes' rule through the likelihood function, so
+the likelihood must contain all relevant information from the data.
+To summarize data for scientific communications, we therefore find a
+hierarchy of objects with decreasing complexity and increasing abstraction:
+\begin{itemize}
+\item the data themselves, with instructions for their interpretation,
+\item the likelihood function, or functions for varied interpretations, and
+\item evidence values for well-defined, predictive models.
+\end{itemize}
+Each is a compression of the data, which is easier to interpret in more refined
+contexts.
+Note that summaries of these objects also be useful, such as maxima
+of the likelihood function,
+or regions in which it exceeds meaningful thresholds.
+
+Posterior probabilities are model-dependent objects that can usefully
+digesting data to make inferences, but are not pure statements of the data ---
+reporting likelihood functions and evidence values allows readers to assign
+their own priors, to compute any posteriors they wish.
+
+For model comparison between two hypotheses $h_1$ and $h_2$
+(perhaps signal versus background),
+Bayes' rule is conveniently linearized in $\log$ ratio form:
 \begin{equation}
@@
 \hphantom{\quad \quad \mid\mid \mathrm{context}}
-\Delta(d) =
-\log\frac{\prob{S}{d}}{\prob{B}{d}}
+\log\!
+\overbrace{
+\frac{\prob{h_1}{d}}{\prob{h_2}{d}}
+}^\textrm{posterior odds}
 =
-\log\frac{\prob{d}{S}}{\prob{d}{B}}
+\log\!\!
+\underbrace{
+\frac{\prob{d}{h_1}}{\prob{d}{h_2}}
+}_\textrm{likelihood ratio}
+\!\!
 +
-\log\frac{\probc{S}}{\probc{B}}
+\log\!
+\overbrace{
+\frac{\probc{h_1}}{\probc{h_2}}
+}^\textrm{prior odds}
 \quad \quad \mid\mid \mathrm{context}
@@
 \end{equation}
-where we use the double bar, $\mid\mid$, to refer to any contextual information
-a reader might wish to include.
+where we use the double bar, $\mid\mid$, to separate to any common contextual
+information.
 When the evidence ratio is independent of that context, as it usually is,
-it factors out from the prior (odds) ratio.
-In this sense, likelihood (or evidence) ratios are scale factors that update
-any odds prior ratio by the same multiplication, and that on this $\log$ scale
-that means just adding a constant.
-
-As data are reported more precisely, likelihoods will often tend towards zero
---- in the continuum limit, the probability of any data value vanishes.
-Small likelihoods are themselves not significant unless other likelihoods
-are larger --- unless other hypotheses predicted \emph{better}.
-And these ($\log$) ratios can remain well-defined in continuum limits even as
-both likelihoods tend towards zero~\cite{billingsley2008probability}.
-
-From such $\log$ ratios, a little rearrangement recovers the individual
-probabilities through the `logistic' function, also known as
-`softmax' if multiple hypotheses are in play~\cite{MurphyKevinP.2012Mlap}.
-
-The infamy of Bayes' rule stems from a fixation on the posterior distributions
-from unjustified prior assignments.
-Classical works by
-Bayes~\cite{bayes1763lii} and
-Laplace~\cite{laplace1774stigler} began this focus ---
-both assume prior densities that are uniform in chosen coordinates, a practice
-that can yield practical results in posterior inferences.
-But to report scientific data, such practices are fairly criticized as
-arbitrary when they influence the reported results.
-Prior assignments are arbitrary.
-You are welcome to propose any prediction in any circumstances, and the
-scientific response is to test how well it predictts our observations of nature.
-Our emphasis on the likelihood functions and evidence values that quantify
-those predictions avoids the pitfall of seeking uninformative, neutral, or
-unbiased priors by embracing priors as active and explicit predictions.
-This is the modern, practical approach to Bayesian data analysis~\cite{
-mackay2003information,
-skilling2004nested,
-skilling2006nested,
-sivia2006data,
-skilling2010foundations
-}.
+it separates from the prior (odds) ratio,
+and the $\log$ likelihood (or evidence) ratios are constant additive terms,
+which update any $\log$ prior odds ratio by the same addition.
 
@@
 
+As data are reported more precisely, likelihoods will often tend towards zero,
+and in the continuum limit, the probability of any data value vanishes.
+Small likelihoods are themselves not significant unless other likelihoods
+are larger --- unless other hypotheses predicted \emph{better}.
+These ($\log$) ratios can, however, remain well-defined and useful in continuum
+limits even as both likelihoods tend towards
+zero~\cite{billingsley2008probability}.
+
+From a posterior $\log$ odds, a little rearrangement recovers the individual
+probabilities through the `logistic' function, also known as
+`softmax' when multiple hypotheses are in play~\cite{MurphyKevinP.2012Mlap}.
+
+The infamy of Bayes' rule stems from a fixation on the posterior distributions
+from unjustified prior assignments.
+Classical works by
+Bayes~\cite{bayes1763lii} and
+Laplace~\cite{laplace1774stigler} began this focus ---
+both assume prior densities that are uniform in chosen coordinates, a practice
+which can yield practical posterior inferences,
+but is not always justified or useful.
+But to report scientific data, such practices are fairly criticized as
+arbitrary when they influence the reported results.
+Prior assignments are arbitrary.
+You are welcome to propose any prediction in any circumstances, and the
+scientific response is to test how well it predicts our observations of nature.
+Our emphasis on the likelihood functions and evidence values that quantify
+those predictions avoids the pitfall of seeking uninformative, neutral, or
+unbiased priors by embracing priors as active and explicit predictions.
+This likelihood perspective the modern, practical approach to Bayesian data
+analysis~\cite{
+mackay2003information,
+skilling2004nested,
+skilling2006nested,
+sivia2006data,
+skilling2010foundations
+}.
+
 Some sources continue to define Bayesian analysis as the computation of
@@
 semi-Bayesian posterior-only approach and its clear flaws.
-A bicycle is stable with two wheels and falls over if one either
+A bicycle is stable with two wheels, and falls over if one either
 is removed;
 Figure~\ref{fig:searches_baton_roue_bayes} illustrates this analogy.
-Although it is possible to ride a unicycle, it is a circus trick and not often
-a practical tool.
+A unicycle is more of a circus trick than a practical transport.
 
@@
 values.''
-This framing of probabilistic data analysis agrees exactly with our standard
-practice of histogramming simulations,
-as illustrated in Figure~\ref{fig:searches_sig_bkg_prior_likelihood}:
+This framing of probabilistic data analysis agrees exactly with the standard
+practice of histogramming simulations in High Energy Physics:
 signal and background models define priors, and their (normalized) bin yields
@@
 take within the observed histogram bin, with which we do not compare the
-models, but which could in principle be used for secondary interpretations.
+models, but which could in principle be used for secondary investigations.
 

### 2022-09-18 main/chapter_searches.tex
tweak bike caption

@@
 Bayesian analysis produces two outputs: the
-evidence $\prob{d}{h}$ and
-posterior $\prob{x}{d,h}$ for hypothesis $h$, data $d$, and parameters $x$.
+evidence $Z = \prob{d}{h}$ and
+posterior $p(x) \prob{x}{d,h}$
+for hypothesis $h$, data $d$, and parameters $x$.
 Posterior-only, semi-Bayesian methods are legitimately criticized because they
-do not challenge the hypotheses that assign their priors $\prob{x}{h}$.
-Since the evidence is a likelihood function over hypotheses, considering the
-evidence can challenge those priors.
-\\[0.5em]
+do not challenge the hypotheses that assign their priors $\pi(x) \prob{x}{h}$.
+Since the evidence is a likelihood function over hypotheses, it can challenge
+them by comparing how well they predicted data.
 Cartoon adapted from ``Baton roue'' by

### 2022-09-18 main/chapter_searches.tex
add missing = signs

@@
 evidence $Z = \prob{d}{h}$ and
-posterior $p(x) \prob{x}{d,h}$
+posterior $p(x) = \prob{x}{d,h}$
 for hypothesis $h$, data $d$, and parameters $x$.
 Posterior-only, semi-Bayesian methods are legitimately criticized because they
-do not challenge the hypotheses that assign their priors $\pi(x) \prob{x}{h}$.
+do not challenge the hypotheses that assign their priors
+$\pi(x) = \prob{x}{h}$.
 Since the evidence is a likelihood function over hypotheses, it can challenge

### 2022-09-19 main/chapter_searches.tex
tweaks starting Monday woo

@@
 Statistical data analysis suffers from dogma --- prescriptions and recipes that
-one can memorize and follow while knowing neither their reasons nor their
+one memorizes and follows while knowing neither their reasons nor their
 limitations:
@@
 Multiplicative error propagation is an example:
-for $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
+for independent $a = \mu_a \pm \sigma_a$ and $b = \mu_b \pm \sigma_b$,
 the standard rule says
@@
 % = \sigma_a^2 \mu_b^2 + \sigma_b^2 \mu_a^2 + \sigma_a^2 \sigma_b^2
-\)
-if $a$ and $b$ are independent;
+\);
 since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$,
-the propagation formula works only for small relative errors.
+the propagation formula assumes at least one small relative error.
 }
@@
 $\pi(x) = \prob{x}{h}$.
-Since the evidence is a likelihood function over hypotheses, it can challenge
-them by comparing how well they predicted data.
+Since the evidence is a likelihood function over hypotheses, it does challenge
+hypotheses by comparing how well they predicted data.
 Cartoon adapted from ``Baton roue'' by

### 2022-09-19 main/chapter_searches.tex
more review edits

@@
 should be initially assigned.
-In general, such probability assignments are free and unconstrained, and might
-be informed by calibration data or loosely motivated opinions.
+Probability assignments are generally constrained only by normalization, but
+might be informed by calibration data or loosely motivated opinions,
+or by established theories such as quantum mechanics.
 But there are examples of unambiguously derived that are commonly used in
 High Energy Physics, notably the Poisson and Gaussian distributions.
-To avoid turning this thesis into a textbook, we only sketch the derivations
-of these, and refer to real textbooks for their detailed
-derivations~\cite{jaynes2003probability}, both of which use the exponential
-identity that
+We sketch their derivations here, and refer to textbooks for
+details~\cite{jaynes2003probability}.
+Both which use the exponential identity that
 \begin{equation}
 \label{eqn:searches_exponential}
-\exp(x) = \displaystyle \lim_{N \to \infty}
+\exp(x) =
+\lim_{N \to \infty}\!
 \left(1 - \frac{x}{N}\right)^N
@@
 
-The Large Hadron Collider collides $40$~million proton bunches per second,
-but \atlas' trigger system accepts only around $1000$ of
-those per second~\cite{atlas2020trigger}.
-This means that we have a scenario with a large fixed number $N$ of trials
+The Large Hadron Collider collides proton bunches at
+$40\,\mathrm{GHz}$,
+but \atlas' trigger system accepts only around
+$1\,\mathrm{kHz}$~\cite{atlas2020trigger}:
+our scenario has a fixed large number $N$ of trials
 (bunch collisions)
-and a small probability of binary acceptance (triggering),
-due to a notable collision event.
+and a small probability of acceptance (triggering).
 The summed total number $n$ of accepted events is therefore described by the
-binomial distribution for $n$;
+binomial distribution for $n$:
 \begin{equation}
 \label{eqn:searches_binomial}
-\mathrm{binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
+\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
 ,
 \end{equation}
-which is a product of the probabilities (produce rule) in the sequence of $N$
-accept/reject decisions, summed (sum rule) over all permutations of that sequence
-that could have led to the total $n$.
-With our tiny trigger rate, we are interested in the limit of small
+which is a product of the probabilities (product rule) in the sequence of $N$
+accept/reject decisions, summed over all possible permutations (sum rule) of
+that sequence for $n$ accepted.
+With the tiny trigger rate, we are interested in the limit of small
 $p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
-Using the exponential identity of Equation~\ref{eqn:searches_exponential}, one
-can show that this limit tends towards the Poisson
-distribution~\cite{jaynes2003probability};
+Using the exponential identity of Equation~\ref{eqn:searches_exponential},
+this limit tends towards the Poisson distribution~\cite{jaynes2003probability}:
 \begin{equation}
@@
 \end{equation}
-which is used ubiquitously in the analysis of our histograms.
+which is used ubiquitously to interpret histograms.
 
@@
 known mean and variance.
-This maximum entropy statement means that for a distribution with known mean
-and variance in given coordinates, its best Gaussian approximation matches
-those moments.\footnote{%
-By `best', this statement means a minimal Kullback-Leibler divergence
+For any distribution with known mean and variance (in given coordinates),
+its best%
+\footnote{%
+By `best', this statement means minimal Kullback-Leibler divergence
 \(
 H(p\leftarrow q) = \sum_i p_i \log(p_i/q_i)
-\)
-from the Gaussian approximation $q$ to the target $p$ with known moments.
+\)~\cite{
+skilling2010foundations
+}
+from approximation $q$ to target $p$ with known moments.
 Such an approximation may not be \emph{good}, but it is best in class.%
 }
+Gaussian approximation matches those moments.
+
 As a limiting approximation, the Gaussian form arises from products of many
-functions, such as combinations of many independent measurements that multiply
-their likelihood functions.
-At a maximum $\check{x}$, the first derivative of a function vanishes and the
-second derivative is negative.
-Taylor expanding about this maximum to the leading order, the product looks like
+likelihood functions from the combination of independent measurements.
+At its maximum $\check{x}$, the first derivative of a smooth function $f(x)$
+vanishes and its second derivative is negative;
+to leading order, the Taylor expansion of a product of $f$s looks like
 \begin{equation}
-f(x)^N = \displaystyle \lim_{N \to \infty}
+f(x)^N =
+\lim_{N \to \infty}\!
 \left(
@@
 \mathrm{d}x
+,
 \end{equation}
-where the prefactor is for normalization to a probability distribution over
-$x$, and we include the $\mathrm{d}x$ terms as reminders that it is a density
-function in $x$, not a probability exactly.
-On a $\log$ scale, the Gaussian form is simply quadratic in $\mu$ (or $x$);
-that quadratic form continues in its multivariate generalization, and
-justifies least-squares methods.
+which is normalized by its prefactor, and in which we include the
+$\mathrm{d}x$ terms as reminders that it is a density function in $x$.
+On a $\log$ scale, the Gaussian form is quadratic in $\mu$, so
+Gaussian likelihoods justify many least-squares methods.
 
@@
 function; this relationship extends to multiple dimensions in the Hessian
-matrix and its inverse, and allows one to estimate error bars from second
-derivatives --- a practical tool that the methods discussed in this thesis
-employ.
+matrix and its inverse, and permits the useful approximation of estimate error
+bars from second derivatives.
 
 This is not the famous Central Limit Theorem, which is related to the
-distribution of sums.
+distribution of sums of independent quantities with finite mean and variance.
 But can be used to derive it: the distribution of a sum is a convolution, which
-becomes a product in Fourier-transformed space, and that product can converge
+becomes a product in Fourier-transformed space, and that product converges
 to a Gaussian function by this same derivation if the mean and variance exist.
-A beautiful alternative proof of the Central Limit Theorem is given by
+Another beautiful alternative proof of the Central Limit Theorem is given by
 Jaynes~\cite{jaynes2003probability}
@@
 
-\TODO{reduce this}
 In the current year, it would be remiss to not mention machine learning
 algorithms.
-Among various other applications, learning algorithms demonstrate practical
-successes in probability assignments, such as assigning
-probabilities to labels for classification,
-or to assign probabilities to words for text
-generation, or any of myriad other
-predictions~\cite{MurphyKevinP.2012Mlap, radford2019language}.
-Machine learning tunes flexible functions by consuming data and testing their
-predictions in the manner we have described; their training and testing works
-roughly as follows:
+Among many applications, learning algorithms have practical successes in
+probability assignments, such as to predict labels in classification,
+or to predict next words in text generation~\cite{
+MurphyKevinP.2012Mlap,
+radford2019language
+}.
+Machine learning tunes flexible functions by consuming data and testing
+predictions in this probabilistic framework:
 \begin{enumerate}
 \item Predict new data with $\prob{d_i}{f(d_{i-1}, \ldots), h}$.
-Keep this evidence as a testing score for comparison against other models.
-(Practically, keep its equivalent $\log$ mean over independent samples.)
+Keep (the $\log$ of) this evidence as a testing score for comparison against
+other models.
 These $d_i$ may be only reductions of the data, such as class labels
@@
 Other inputs that are not predicted, such as associated images that the machine
-uses to predict the labels, are included in the context $h$.
+uses to predict the labels, are in the implicit context.
 \item Use those data to update your state
@@
 practice.
-Learning only part of the data is not wrong, it is just less precise than a
-divinely perfect learning algorithm, which we do not currently possess.
-\item Go to 1.
+Learning only part of the data is not wrong, but less precise.
+\item Go\,to 1.
 \end{enumerate}
-These sequential data $d_i$ might be labelled the `training', `validation',
-and `testing' sets, but other splittings are possible.
+These portions of data $d_i$ might be labelled the `training', `validation',
+and `testing' sets.
 This perspective highlights the importance of independent training and testing
-sets --- since $\prob{d}{d, \ldots} = 1$, if any memory of $d_i$ is kept in
-the $f(\ldots, d_i, \ldots)$ state, then it is a memory and not a prediction.
-Independent testing data permit the evidence interpretation of predictive
-probabilities.
+data --- information should not be reused since
+$\prob{d_2}{d_2, \ldots} = 1$, any memory of $d_2$ that is kept
+in the $f(\ldots, d_2, \ldots)$ state spoils the likelihood if $d_2$ is reused
+for testing after training~\cite{sivia2006data}.
+Independent testing data permit the clean interpretations of likelihood ratios.
 
@@
 Now that we have thoroughly covered probability theory, we are ready to return
-to the jelly bean game that is illustrated in Figure~\ref{fig:searches_beans}
-and asks you to guess how many beans are in the jar.
+to the jelly bean game of Figure~\ref{fig:searches_beans}.
 Given all you know about illustrations of jars and beans, you could make a
-vague guess without looking.
-That guess should only be refined after inspecting the image and updating your
-model to account for it.
+vague guess without looking, but and that guess would be refined after
+inspecting the image to update your implicit internal model.
 Although I do not propose that human brains \emph{do} use probability
-algebra~\cite{jaynes1988brain}, this inference process could, in principle,
-be modelled numerically by distributing probability over the positive integers
-and updating those probabilities in light of a likelihood function for the
-image data.
+algebra~\cite{jaynes1988brain}, this inference process could be modelled
+numerically by distributing probability over the bean numbers
+and updating those probabilities (from prior to posterior) in light of a
+likelihood function for the image.
+
 Any guess, however, does not depend only on that probability distribution,
 but on the risks and rewards associated with guessing.
-
-As written in the caption to Figure~\ref{fig:searches_beans}, there is a
-notional reward to the first exactly answer.
-You are therefore incentivized to answer a) quickly, before other good guesses,
+The game offers a notional reward to the first exactly answer, so
+you are therefore incentivized to answer a) quickly, before other good guesses,
 and b) exactly, with no reward for being close to the true answer.
@@
 But others following you, seeing your answer, should prefer to guess
-differently if they want to win (and they think any other guess is plausibly
-correct).
+differently if they want to win.
 
@@
 bad its results would be.
-Mathematical decision theory agrees with this intuition---
-a famous theorem by Wald~\cite{
+Mathematical decision theory agrees with this intuition ---
+a theorem by Wald~\cite{
 wald1947bayes,
@@
 \label{eqn:searches_bayes_decision_rule}
-R(t) = \int\! c(t, x) \,\mathrm{d}p(x)
+R(t) = \int\! c(t, x) \,\mathrm{d}\pi(x)
 ,
@@
 where we mean `expected' in the mathematical mean sense of the `mean' average.
-This cost function $c(t, x)$ evaluates the badness of outcomes for a given
+The cost function $c(t, x)$ evaluates the badness of outcomes for a given
 decision $t$ and states of the world $x$ to which context must assign prior
-probabilities $p(x)$.
+probabilities $\pi(x)$.
 For a given cost function and prior, one should choose the decision $t$ that
@@
 Although alternative ideas (such as minimax, minimizing a maximum
-loss~\cite{savage1951review}) can work well in certain contexts, this theorem
-states that restricting oneself to minimizing an expected cost function cannot
-deny one access to optimal decisions.
+loss~\cite{savage1951review}) can work well in certain contexts,
+Wald's theorem states that restricting oneself to minimizing an expected cost
+function cannot deny one access to optimal decisions.
 Alternative formulations can at best be equivalent.
@@
 in the risk;
-decision-making machines and organisms are capable of learning to make good
+decision-making machines (such as organisms) are do learn to make good
 decisions without constructing either cost or prior explicitly.
-Towards our goal of usefully reporting data, such decision theory has been
-applied to argue for ways to quote properties of posterior distributions.
-As examples, minimizing the expected absolute error $|x - \check{x}|$ leads
+Towards the goal of usefully reporting data, such decision theory has been
+applied to argue for ways to summarize posterior distributions.
+For example, minimizing the expected absolute error
+$c(t, x) = |t - x|$ leads
 to reporting the median, and minimizing the expected square error
-$(x - \check{x})^2$ leads to reporting the mean.
+$c(t, x) = (t - x)^2$ leads to reporting the mean.
 Had the rules of our bean counting game rewarded you based on such distances,
-these results could have been good guesses. But those weren't the rules.
+these results could have been good guesses.
+But those weren't the rules.
 
@@
 Although Bayes' rule is not seriously disputed, the system in which Wald was
-working defined probability as a frequency in random sampling, which cannot
-reasonably apply to subjective, non-random distributions of probability mass
+working defined probability via frequencies in random sampling, which cannot
+reasonably apply to the subjective, non-random distributions of probability
 that many prior distributions are.
 This frequency-restricted, `frequentist', philosophy dominated 20th century
-statistical research, and influences practices that \atlas\ continues to
-employ.
-We therefore review some features of frequentist statistics in the following
-section.
+statistical research, and influences practices still followed in \atlas.
+We therefore review some features of frequentist statistics and its history
+in the next section.
+% Would like something along the lines of:
+% `No formal derivations except for historical precedent,
+% so review with that history'
+% But feels too rude.
 
@@
 $\log\cls$ is non-positive with the $-n$ test statistic, and indeed non-negative
-for the opposite ordering $t(n) = n$.
+for the reverse ordering $t(n) = n$.
 The $\log$ likelihood ratio on the actual data does not feature this bias:

### 2022-09-19 main/chapter_searches.tex
add walk note; minor tweaks

@@
 which can assign their \pvalues\ through well-behaved analytic
-expressions~\cite{Cowan:2010js}.
+expressions~\cite{
+Cowan:2010js,
+wilks1938large,
+wald1943tests
+}.%
+\footnote{%
+Remarkably, the asymptotic limits employed by~\cite{Cowan:2010js} to
+approximate \pvalues\ derive from a theorem of the same Abraham~Wald who later
+proved the redundancy of \pvalue\ decision rules~\cite{wald1947bayes}.%
+}
 Without such formulae, the remaining practical approach to estimating
@@
 extracting interpretable results.
-A general \heplikelihood\ model is a function $H$ that takes some parameters $x$
-and outputs two things:
+A general \heplikelihood\ model is a function $H(x)$ of some parameters $x$
+that outputs two things:
 expected yields $y$ to predict Poisson data,
-and an objective value $E$ that should be minimized in a `fit' to choose the
-best prediction.
+and an objective value $E$ to be minimized (in a `fit') to choose a best
+prediction.
 That is,

### 2022-09-19 main/chapter_searches.tex
more review edits; add more ban citations

@@
 statistical research, and influences practices still followed in \atlas.
-We therefore review some features of frequentist statistics and its history
-in the next section.
+We shall therefore review some features of frequentist statistics and its
+history.
 % Would like something along the lines of:
@@
 
-We do think of many processes as random when considering scientific data, so
-this definition works widely for likelihood functions.
+We do think of many natural processes as random, so this definition works
+widely for likelihood functions on data.
 Non-data hypotheses, however, are rarely considered as actually random events,
-so the frequency definition cannot not apply.
-Prior predictions often apply probabilities to express subjective uncertainties
-that are manifestly non-random.
+so the frequency definition cannot apply ---
+prior predictions often express subjective uncertainties that are manifestly
+non-random.
 Under this frequency definition, those predictive probabilities are nonsense
@@
 $n$ (number of heads) to $N$ (number of tosses).
-Now suppose you are given a red envelope containing a coin;
+But suppose you are given a red envelope containing a coin;
 which way up is the coin in the envelope?
@@
 You do not know.
-In Neyman's words,
+In the words of Jerzy~Neyman,
 ``I have put the word ``~random~'' in inverted commas because it is very
@@
 
-Tom~Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
+Rosencrantz and Guildenstern, in their play by Tom~Stoppard~\cite{
 stoppard1967rosencrantz
 }
-opens with a clearer example ---
-the titular characters toss a coin ninety-two times, all landing heads.
-They then play a game with a third player, landing eight more heads.
+toss a coin ninety-two times, all landing heads.
+They then play a game with a third player and land eight more heads.
 Will the next toss be heads or tails?
-The answer is not random --- it is burned onto paper in print, so no frequency
-can be assigned.
-Were subjective probabilities permitted, you could predict the result, and that
-prediction would be subjectively informed by it being the one-hundred-and-first
-toss, our cultural entanglement with round numbers in base-10, and your reading
-of these words.
+The answer is not random --- it is burned onto paper in print.
+If you were to put probabilities on the outcomes, they would be subjectively
+informed by it being the one-hundred-and-first toss, our cultural enjoyment
+of round numbers in base-10, and your reading of these words.
 Frequentist theory, however, forbids such subjective probabilities of
-non-random unknowns, so is left to search for alternative modes of inference.
+non-random unknowns, and is left searching for alternative forms of inference.
 
@@
 Boole~\cite{boole1854investigation}, in reactions that may well have been
-propelled by reactions to flawed posterior-only methods.
+catalysed by the presence of flawed posterior-only methods.
 Modern frequentist theory has a rigorous set-theoretic formulation in the
@@
 
-Modern frequentist methods are largely attributable to work by
-Ronald~Fisher (for likelihood, $p$-values, and estimation criteria~\cite{
+Modern frequentist methods are largely attributable to work
+in the 1920s and '30s by
+Ronald~Fisher (for likelihood, \pvalues, and estimation criteria~\cite{
 fisher1912fitting,
@@
 neymanpearson1928max
-})
-in the 1920s and '30s.
+}).
 These Fisher and Neyman-Pearson schools do not form a unified framework, but
@@
 
-Some literature names this Fisher/Neyman-Pearson assemblage
+Some authors name this Fisher/Neyman-Pearson assemblage
 ``classical'' statistics~\cite{
@@
 Feldman:1997qc
-}, but that is simply inaccurate for 20th century creations that innovated on
-18th century ideas~\cite{bayes1763lii, laplace1774stigler}.
+},
+but that word is simply inaccurate for 20th century innovations on 18th century
+ideas~\cite{
+bayes1763lii,
+laplace1774stigler
+}.
 Jaynes instead names them ``orthodox''~\cite{
@@
 (heretical?) ideas have taken hold in research at the LHC ---
-the orthodoxy has moved on.
+orthodoxy has moved on.
 We therefore stick with `frequentist'.
@@
 $1/2$, $1/5$, and $1/15$ of its maximum, respectively.
-The extents of these lines could be tabulated to usefully communicate aspects
-of its profile.
+These lines could be tabulated to usefully summarize the profile of the
+likelihood function.
 }
@@
 we employ for the probability of observed data given varied
-contexts%
+contexts,%
 \footnote{%
@@
 likelihood is also a probability, just not a probability \emph{distribution}.
-},
+}
 and deserves credit for popularizing its use in interpreting and communicating
@@
 When recommending error bar recipes, \atlas' statistics committee appears
-to have not considered this option~\cite{cowan2011errorbars}, although
-our statistics experts are surely familiar with Fisher's work.
+to have not considered this option~\cite{cowan2011errorbars}.
+% , although
+% our statistics experts are surely familiar with Fisher's work.
 }
@@
 parameters being considered, and Fisher showed that maximum likelihood
-estimates satisfy some desirable criteria~\cite{fisher1922estimators}.
+estimates satisfy some desirable asymptotic
+criteria~\cite{fisher1922estimators}.
 These criteria do not include \emph{unbiasedness}, which states that the mean
@@
 Fisher rejected the bias criterion for this
-reason~\cite{jaynes2003probability}, and other definitions absorb its
+reason~\cite{jaynes2003probability}.
+Other definitions absorb its
 arbitrariness into an explicit loss function~\cite{lehmann2005testing}.
-Requiring unbiasedness can actively harm our results by preventing the use of
+Requiring unbiasedness can actively harm results by preventing the use of
 useful estimates,%
@@
 Problem~1.14 of~\cite{lehmann2005testing} has a delightful example of
-estimating an exponentiated a Poisson mean $e^\mu$ from $n$ data
+estimating an exponentiated a Poisson mean $\exp(\mu)$ from $n$ data
 where the unique unbiased estimate is $(-1)^{n + 1}$.
-Section~17.3 of~\cite{jaynes2003probability} finds similar nonsense for
+Section~17.3 of~\cite{jaynes2003probability} shows similar nonsense for
 unbiased estimates of $\mu^2$ and $\exp(-\mu)$.%
@@
 However, the hard rejection of priors on hypotheses also prevents evidence
-calculations, without which likelihoods alone become confounded in many
-parameters, perhaps making them nuisances.
+calculations, without which likelihoods alone can become confounded in many
+parameters
+(perhaps making those parameters seem feel more like `nuisances'
+than useful levers).
 To circumvent this, frequentist methods employ two key concepts:
 test statistics, which can reduce the data to remove unwanted parameters,
-and $p$-values, which divert the discourse
-from what the alternative hypotheses \emph{are}
-to what alternative data \emph{might have been}.
+and \pvalues, which divert the discourse
+from what the alternative \emph{hypotheses are}
+to what alternative \emph{data might have been}.
 
@@
 \label{eqn:searches_t_statistic}
-t(d) = \frac{\bar{\mu}(d) - \mu}{\bar{\sigma}(d)}\sqrt{n}
+t(d) = \frac{\bar{\mu}(d) - \mu}{\bar{\sigma}(d)/\sqrt{n}}
 ,
@@
 which shifts the sample mean $\bar{\mu}(d)$ by a known true mean $\mu$, and
-scales that difference down by the same standard deviation
-$\bar{\sigma}(d)$~\cite{student1908, fisher1925t}.
-By dividing out a (scale equivariant) estimate of the standard deviation, $t$
-is independent of common scale factors on the data, and that scale invariance
-has helped to drive the study and common use of this
-$t$-statistic~\cite{lehmann2011fisher}.
+cancels any common scale factors with a scale-equivariant estimate of the
+standard deviation $\bar{\sigma}(d)$~\cite{
+student1908,
+fisher1925t,
+lehmann2011fisher
+}.
 
@@
 to compare how well different hypotheses predict it.
-For unclear reasons, Fisher advocates a different use ---
-rather than considering likelihoods and likelihood ratios, he recommends
+Fisher, however, advocates a different use ---
+rather than considering likelihoods here, he recommends
 that researchers consider the ``probable error''~\cite{fisher1921probable},
-which is a cumulative distribution function in the negated test statistic.
-In modern jargon, a `probable error' at the observed test statistic
-$\tilde{t}$ and an hypothesis $h$ is a \textbf{\pvalue}
+which is a cumulative distribution function in the (negated) test statistic.
+In modern jargon, a `probable error' with observed test statistic
+$\tilde{t}$ and hypothesis $h$ is a \textbf{\pvalue}
 \begin{align}
@@
 the \pvalue\ $P(h)$ is the prior probability of data
-\emph{greater or equal to} the observed data when ordered by the value of the
-test statistic.
+\emph{greater or equal to} the observed data, if data are ordered by the value
+of the test statistic.
 Equation~\ref{eqn:searches_pvalue_likelihood} points out that the \pvalue\ is
@@
 $r=0.3$ and
-$r=0.95$, finding $p$-values
+$r=0.95$, finding \pvalues\
 $P(r=0.3)=0.142$ and
@@
 (``when measured on the $r$ scale'').
-By writing `likely', Fisher is carefully not referencing the probability
-of parameters, but their likelihoods in our definition.
+By writing `likely', Fisher is carefully referencing their likelihoods,
+not their probability.
 But this factor of $1000$ is a \pvalue\ ratio, not a likelihood ratio,
 so how can this make sense?
-Plausibly, it results from a tacit data reduction; comparing
+It could plausibly result from a tacit data reduction --- comparing
 Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
-shows that the \pvalue\ is a likelihood only if the result is reduced from
-$A: t =    \tilde{t}$ to
-$B: t \geq \tilde{t}$.
-And since this reduction is technically true --- $A$ implies $B$ --- the
-likelihood ratio on $B$ is logically valid and not \emph{necessarily}
+shows that the \pvalue\ is a likelihood if the data statement is reduced from
+$d_1: t =    \tilde{t}$ to
+$d_2: t \geq \tilde{t}$.
+And since this reduction is technically true --- $d_1$ implies $d_2$ --- the
+likelihood ratio on $d_2$ is logically valid and not \emph{necessarily}
 misleading if accurately reported.
 
-However, $B: t \geq \tilde{t}$ is a logical statement (true for all test
+However, $d_2: t \geq \tilde{t}$ is a logical statement (true for all test
 statistics greater than or equal to the observed value) and not a randomly
@@
 benefit --- it reduces the information content of the reported data without
-reducing the associated costs of reporting or interpreting those data.
-Both $A$ and $B$ are printed in the same number of characters, and $p$-values
-can be painfully expensive to evaluate.
-And it actively harms any result with small a test statistic by
-diluting the likelihood ratio, since if
+reducing the associated costs of communicating or interpreting those data.
+Both $d_1$ and $d_2$ are printed in the same number of characters, and
+\pvalues\ can be painfully expensive to evaluate.
+Furthermore, any result with small a test statistic is actively censored,
+since if
 \begin{equation}
@@
 \end{equation}
-then large tail probabilities squash the $\log$ likelihood ratio toward zero.
-
-Fisher's \pvalue\ ratio of $1000$ is, therefore, a $\log$ likelihood
-ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderately
-strong indication.
+then large tail probabilities can only shrink the $\log$ likelihood ratio
+towards zero.
+
+Fisher's \pvalue\ ratio of $1000$ can therefore be a $\log$ likelihood
+ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderate
+indication.
 A likelihood ratio on the test statistic itself, and not the inclusive
@@
 of both signal and background distributions, so with an $m_{4\ell}$ test
-statistic $P(\textrm{Higgs boson}) \approx P(ZZ^*) \approx 0.5$, so neither
-is ``suspect''.
+statistic $P(\textrm{Higgs boson}) \approx P(ZZ^*) \approx 0.5$ and neither
+is `suspect'.
 Local ratios between the two distributions clearly favour signal for this
-event, however; the $p$-values' failure is in their dependence on irrelevant
-tail areas, which are a feature of the histogram and not the data.
+event, however:
+the \pvalues' failure is in their dependence on irrelevant tail areas,
+which are a feature of the histogram and not the data.
 This $ZZ^*$ density peaks for $ZZ$ above a
 $2 m(Z) \approx 180\,\eV[G]$ kinematic edge,
-and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$;
-expanding the histogram to show either of these tails would change nothing
-local to our $m_{4\ell} = 125\,\eV[G]$ event, but would change our $p$-values.
+and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$,
+so expanding the histogram to show either of these tails would change nothing
+local to our $m_{4\ell} = 125\,\eV[G]$ event, but would pathologically change
+its \pvalues.
 
@@
 are optimal in a specific sense within their framework of decisions
-based on $p$-values.
+based on \pvalues.
 Precisely, at a fixed probability $\alpha$ of rejecting a true hypothesis,
@@
 Bernstein2016princess,
-wagenmakers2007practical
-}
-in fields where they are not banned~\cite{Trafimow2015ban}.
+wagenmakers2007practical,
+hauer2004harm,
+wasserstein2016asa
+}
+in fields where they are not yet banned~\cite{
+Trafimow2015ban,
+woolston2015phychology,
+chumra2019time,
+shrout1997should,
+hunter1997needed
+}.
 For example, the false presentation as a ``probability of compatibility'' with
@@
 Each $\cls$ is a \pvalue\ ratio.
-By reducing the observed data to an inequality, the $p$-value selectively
+By reducing the observed data to an inequality, the \pvalue\ selectively
 censors some of its information.
@@
 Each $\cls$ is \pvalue\ ratio.
-By reducing the observed data to an inequality, the $p$-value selectively
+By reducing the observed data to an inequality, the \pvalue\ selectively
 censors some of its information.

### 2022-09-19 main/chapter_searches.tex
rearrange cls n figure

@@
 pdg2022ynf
-}, but such a $p_b$ might not be a \pvalue\ when it excludes equality with the
+}
+where $p_b$ might not be a \pvalue\ if it excludes equality with the
 observed test statistic~\cite{cern2011procedure}.
-}%
+}
+
 As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
@@
 values of $s$ for which $\cls = 1 - \gamma$.
+
+Early variants of $\cls$ derived it from an identity with posterior tail
+probabilities in Poisson and Gaussian special cases~\cite{
+Helene1983upper,
+pdg1988,
+read2000modified,
+pdg2022ynf
+}
+with a contrived frequentist interpretation~\cite{zech1988cls}.
+For use in general cases, it is more recently described as a
+``modified frequentist'',
+``approximate confidence in the signal-only hypothesis''~\cite{
+read2000modified,
+Read2002cls
+}.
+High-profile usage in Higgs boson searches at the
+Large Electron–Positron Collider~\cite{
+read1997optimal,
+bock1998lower,
+etde1998prospects,
+junk1999confidence,
+lep2000searches,
+lep2003search
+}
+appears to have propelled $\cls$ to prominence;
+these early applications used data deficits to exclude their
+background-plus signal models, sometimes with the test statistic $t(n) = -n$,
+and disliked how deficits led to small \pvalues\ for the background-only model,
+with which they might have confusingly set negative limits on the Higgs boson
+signal cross-section.
+Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left)
+for small data counts;
+in contrast, $\cls$ can only favour the background-only model, due to its
+negative values that are shown on the right of the same figure.
 
@@
 \end{figure}
-
-Early variants of $\cls$ derived it from an identity with posterior tail
-probabilities in Poisson or Gaussian special cases~\cite{
-Helene1983upper,
-pdg1988,
-read2000modified,
-pdg2022ynf
-}
-with a contrived frequentist interpretation~\cite{zech1988cls}.
-For use in general cases, it is more recently described as a
-``modified frequentist'',
-``approximate confidence in the signal-only hypothesis''~\cite{
-read2000modified,
-Read2002cls
-}.
-High-profile usage in Higgs boson searches at the
-Large Electron–Positron Collider~\cite{
-read1997optimal,
-bock1998lower,
-etde1998prospects,
-junk1999confidence,
-lep2000searches,
-lep2003search
-}
-appears to have propelled $\cls$ to prominence;
-these early applications used data deficits to exclude their
-background-plus signal models, sometimes with the test statistic $t(n) = -n$,
-and disliked how deficits led to small \pvalues\ for the background-only model,
-with which they might have confusingly set negative limits on the Higgs boson
-signal cross-section.
-Such \pvalues\ are illustrated in Figure~\ref{fig:searches_sb_n} (left)
-for small data counts;
-in contrast, $\cls$ can only favour the background-only model, due to its
-negative values that are shown on the right of the same figure.
 

### 2022-09-19 main/chapter_2ljets.tex
remove significance section; this is already far too long; some tweaks
clean up author lists

--- 2022-09-15 main/chapter_2ljets.tex
@@
 backgrounds~\cite{atlas2022searches}.
-The lower panel shows $S_\mathrm{ATLAS}$ from
-Equation~\ref{eqn:significance_atlas}.
+The lower panel shows the \atlas-recommended significance
+measure~\cite{atlas_significance}.
 Control, validation and signal regions are shown from left to right, with the
@@
 This requirement is typically justified by the limitations of approximate
-statistical methods that work
-``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
+statistical methods which work
+``from as few as $\mathcal{O}(10)$ data events''~\cite{baak2015histfitter}.
 But the Poisson likelihood works for infinitesimally small
@@
 through the special parameter name
-\texttt{SigXSec}~\cite{histfitter2014, histfittergithub}.
+\texttt{SigXSec}~\cite{baak2015histfitter, histfittergithub}.
 

### 2022-09-19 main/chapter_searches.tex
clean up author lists

@@
 suspect after an experiment.
-For example, in Fisher's suggestion of $P(h) \in 0.1--0.9$ being not suspect,
-then $\alpha = 0.2$.%
+For example, in Fisher's suggestion of $P(h) \in 0.1\textrm{--}0.9$ being not
+suspect, then $\alpha = 0.2$.%
 \footnote{%
@@
 }
-otherwise meets the edges at $0.1$ and $0.9$ in its cumulative distribution).
 And that chance is independent of the hypothesis, its alternatives, and even
@@
 anything at all and still incur suspicion for unrelated hypotheses with the
-same prior probability.
-Your wall clock reads not less than $48$ minutes past the hour?
-The Standard Model is suspect.
+same prior probability ---
+your wall clock reads greater than or equal to $48$ minutes past the hour?
+If so, then the Standard Model is suspect.
 Figure~\ref{fig:searches_significant} illustrates this peculiarity and the
-associated difficulty in communicating the results.
+associated difficulties with communicating \pvalue\ results.
 
@@
 
-Neyman and Pearson formalized this \pvalue\ based suspicion into decision
-rules and thought about what it is that makes one test statistic better than
+Neyman and Pearson formalized this `suspicion' into decision rules,
+and considered what it is that makes one test statistic better than
 another~\cite{neymanpearson1928max, neymanpearson1933lemma}.
-Rather than a \pvalue\ arousing suspicion of an hypothesis, here it is used
+Rather than a \pvalue\ arousing suspicion of an hypothesis, they use it
 to decide on actions of `rejecting' or `accepting' an hypothesis.
-From the Bayesian arguments that we are presenting in
-Section~\ref{sec:searches_data_analysis},
-Neyman and Pearson argued for the interest in likelihood ratios, and further
-considered their use as test statistics~\cite{neymanpearson1928max};
+From the same Bayesian arguments of Section~\ref{sec:searches_data_analysis},
+Neyman and Pearson argued for the interest in likelihood ratios as promising
+test statistics~\cite{neymanpearson1928max};
 they also considered \emph{maximum} likelihood ratios
@@
 \end{equation}
-for a specified hypothesis $h_1$ against the largest likelihood from a set
-of alternatives.
+to compare a specified hypothesis $h_1$ against a set of alternatives.
 
@@
 usefully communicate relevant information from the data.
-And ass test statistics, they also win generality by avoid context-specific
+And as test statistics, they also win generality by avoid context-specific
 motivations, such as scale invariance for the $t$-statistic of
-Equation~\ref{eqn:searches_t_statistic} --- any situation with predictive
-models can construct likelihoods and maximum likelihood ratios.
+Equation~\ref{eqn:searches_t_statistic} --- any predictive models can have
+likelihoods and maximum likelihood ratios.
+
 ``Without claiming that this method is necessarily the ``best'' to adopt,''
@@
 neymanpearson1933lemma}.
-To this quote, Lehmann appends ``(sic!)''~\cite{lehmann2011fisher}, since it
-appears to conflate the likelihood for an hypothesis with its probability.
-But Neyman and Pearson's association of likelihood ratios with
+Lehmann appends ``(sic!)'' to this quote~\cite{lehmann2011fisher};
+it appears to mistakenly conflate the likelihood for an hypothesis with its
+probability, but Neyman and Pearson's association of likelihood ratios with
 ``more probable'' hypotheses may not be an error.
-The authors understood the algebra of how large likelihood ratios make
-hypotheses \emph{more probable} by changing their posterior probabilities,
-independent of whatever (non-zero) prior probabilities they are assigned.
-Indeed, they later state that
+The authors understood the probability algebra of how large likelihood ratios
+make hypotheses \emph{more probable} by scaling whatever (non-zero) prior
+probabilities may be assigned.
+Indeed they later state that
 \begin{displayquote}
@@
 \end{displayquote}
-Elsewhere also, they consider an example with two hypotheses $\mathrm{A}_1$
+Elsewhere, they also consider an example with two hypotheses $\mathrm{A}_1$
 and $\mathrm{A}_2$, and find a $\log$ likelihood ratio of
@@
 of event variables of the kind we use in \atlas\ searches.
-This can be seen from the signal and background distributions of
+This is seen from the signal and background distributions of
 Figures~\ref{fig:searches_sig_bkg_prior_likelihood}
-and~\ref{fig:searches_atlas_higgs_4l};
+and~\ref{fig:searches_atlas_higgs_4l}:
 the event variable in Figure~\ref{fig:searches_sig_bkg_prior_likelihood} is
-like a likelihood ratio in that any right-handed inequality selects a good
-signal purity relative to its background yield, but the signal bump in the
+like a likelihood ratio in that any right-handed selection contains a good
+signal purity relative to its background, but the signal bump in the
 middle of the background distribution in
@@
 window around the Higgs boson mass.
-Re-ordering $m_{4\ell}$ by a likelihood ratio could turn this selection window
+Re-ordering $m_{4\ell}$ by a likelihood ratio could turn the window
 into an equivalent and simpler one-sided selection, and the Neyman-Pearson
@@
 `confidence intervals', which are ranges of parameters that are
-accepted by the \pvalue\ decision rules~\cite{
+accepted by a \pvalue\ decision rule~\cite{
 clopper1934confidence,
@@
 probability mass, but this \pvalue\ form can always satisfy that through the
-free choice of test statistic and extends naturally beyond one dimension.%
+free choice of test statistic, and extends naturally beyond one dimension.%
 }
@@
 `probably' contains the unknown parameter, but that gets it backwards.
-For a known true parameter $x$, the confidence intervals of random repeated
-experiments should contain $x$ with probability $\gamma$;
+For a known true parameter $x$, the various confidence intervals
+$x_1\textrm{--}x_2$ from random repeated experiments should contain
+$x$ with probability $\gamma$;
 the interval makes a subtle statement about unknown alternative data for
@@
 an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
-likelihood function
+likelihood function:
 \begin{equation}
@@
 It can only detract.
-The recommendation by Fisher to directly communicate the likelihood function
+Fisher's recommendation to communicate the likelihood function itself
 remains simpler and clearer.
@@
 Data analysis in modern High Energy Physics does not exactly follow frequentist
-ideology, but is a hands-on, practical, and evolving mixture that has adapted
-to avoid some of the more glaring flaws.
+ideology, but is a hands-on, practical, and evolving ecosystem that has adapted
+to avoid some of the more glaring flaws when attacking real problems.
 We have seen in Section~\ref{sec:searches_data_analysis} how likelihood ratios
@@
 This section reviews how these objects are applied generally in conventional
-analysis at the LHC and specifically towards the \atlas\ search that this
+analysis at the LHC, and specifically towards the \atlas\ search that this
 thesis presents.
@@
 a model~\cite{HIGG-2018-04} appears to be spreading as a good
-meme~\cite{dawkins1989selfish} in High Energy Physics ~\cite{
+meme~\cite{dawkins1989selfish} in High Energy Physics~\cite{
 roth2007fit,
@@
 replacements, which $\log$ likelihood ratios can provide.
-And standard practice at the LHC contains the beginnings of such
+Standard practice at the LHC contains the beginnings of such
 replacements under the name $\cls$~\cite{
@@
 which is used for setting upper limits on disfavoured signal yields $s$
-above background yields $b$;
+above background yields $b$:
 \begin{equation}
@@
 \footnote{%
-Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
-not confidence levels ---
-a confidence level $\gamma$ is a decision boundary used, for example,
-to assign limits by solving for $x$ such that
-$\gamma = 1 - P(x)$~\cite{pdg1996}.
+Despite their $\mathrm{CL}$ labels, $\cls$, $\clsb$, and $\clb$ are not
+Confidence Levels ---
+a confidence level $\gamma$ is a threshold used to assign limits by solving
+for $x$ such that $\gamma = 1 - P(x)$~\cite{pdg1996}.
 In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
 at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
-by~\cite{pdg1984}) do make the distinction~\cite{
+by~\cite{pdg1984}) do distinction the two~\cite{
 eadie1971statistical,
 frodesen1979probability
-}.%
+}.%s
 }%
@@
 same data, so $\clsb$ and $\clb$ should use the same test statistic.
-Identical test statistics are used for all results of this work, but differing
-test statistics have been allowed elsewhere~\cite{bock1998lower}.
+Identical test statistics are used for all results in this thesis, but
+differing test statistics have been allowed elsewhere~\cite{bock1998lower}.
 }%
@@
 }.
-High-profile usage in Higgs boson searches at the
+High-profile usage in combined Higgs boson searches at the
 Large Electron–Positron Collider~\cite{
@@
 (linearly) related to the likelihood ratio.
-Each $\cls$ is a \pvalue\ ratio.
+Each $\cls$ is a \pvalue\ ratio with the stated test statistic.
 By reducing the observed data to an inequality, the \pvalue\ selectively
-censors some of its information.
+censors its information.
 }
@@
 
-Despite the statement in Section~\ref{sec:searches_fisher} that \pvalue\
+Despite my statement in Section~\ref{sec:searches_fisher} that \pvalue\
 ratios are not ``not \emph{necessarily} misleading'',
@@
 no data can give a $\cls$ that favours the signal model;
-$\log\cls$ is non-positive with the $-n$ test statistic, and indeed non-negative
+$\log\cls$ is non-positive with the $-n$ test statistic, and non-negative
 for the reverse ordering $t(n) = n$.
@@
 excesses.
-Since for Poisson models the $\log$ likelihood ratio is linear in $n$,
+The $\log$ likelihood ratio is linear in $n$ for these Poisson models,
 \begin{equation}
@@
 \end{equation}
-both $n$ and $-n$ can be optimal in the Neyman-Pearson sense.
-
-By using \pvalue\ ratios with Neyman-Pearson optimal test statistic, we have
+so both $n$ and $-n$ can be optimal in the Neyman-Pearson sense.
+
+By using $\cls$ with a Neyman-Pearson optimal test statistic, we have therefore
 constructed an analysis that can only ever favour one hypothesis:
@@
 be weaker support for the signal than $n=9$.
-Each $\cls$ is \pvalue\ ratio.
+Each $\cls$ is a \pvalue\ ratio with the stated test statistic.
 By reducing the observed data to an inequality, the \pvalue\ selectively
-censors some of its information.
+censors its information.
 }
@@
 predetermined threshold, say $5$ events, then models can predict whether
-$n \leq 5$, and the truth or falsehood in data can favour either side.%
-\footnote{%
-This context of a fixed inequality is one condition where the
-\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
-}
+$n \leq 5$, and the honest truth or falsehood in data can favour either side.
 Ultimately, the reader just distrust the reporter ---
@@
 identical, since one might infer from the former that in fact $n = 5$.
-The message itself is data --- what it says may not be the whole truth.
-
-With more than one Poisson bin to model, the $n$ test statistic is no longer
+The message itself is data, and what it says may not be the whole truth.
+
+With multiple Poisson bins, the $n$ test statistic is no longer
 sufficient.
@@
 such as $\Lambda$ of Equation~\ref{eqn:searches_max_like_ratio}, often useful.
-In the single Poisson case, $\Lambda$ shuffles the ordering of
+With a single bin, $\Lambda$ shuffles the ordering of
 data and leads to some odd features in the results, which are displayed in
 Figure~\ref{fig:searches_sb_t}.
-In particular, the choice $t(n) = -\Lambda(n)$ pushes data that are far from
-the background expectation into its tails, so can disfavour the background
+In particular, the choice $t(n) = -\Lambda(n)$ pushes any data that are far
+from the background expectation into its tails, so can disfavour the background
 model for small $n$.
 Probability masses from small data also interrupt its behaviour for excesses,
-leading to jagged changes, notably that $\log\cls$ reduces as the data
-increase from $9$ to $11$ --- counter to intuition and logic, increasing excess
-data can reduce the reported favour for the signal model.
-
-Instead of $\Lambda$ itself, various modified maximum likelihood ratio test
+and lead to jagged changes, notably that $\log\cls$ reduces as the data
+increase from $9$ to $11$.
+Counter to intuition and logic, this prescription can report decreased favour
+for the signal model when increasing the excess in data.
+
+Rather than $\Lambda$ itself, various modified maximum likelihood ratio test
 statistics enjoy standard usage at the LHC~\cite{cern2011procedure};
 these are variably defined as piecewise functions of the fitted signal
-parameter $s$, and can include likelihoods from the signal-plus-background
+parameter, and can include likelihoods from the signal-plus-background
 model with constrained maximizations which aim to reduce the impact of
@@
 
-These modified maximum likelihood ratios benefit from known approximations to
+Modified maximum likelihood ratios also benefit from known approximations to
 their sampling distributions in asymptotic (Gaussian) limits,
@@
 \pvalues\ is to generate Monte Carlo simulations of alternative data.
-Such simulations can be computationally expensive, ill-defined for
+However, such simulations can be computationally expensive, and ill-defined for
 ``likelihood'' models that blur the distinction between actual data and
@@
 \subsection{Like-lihood modelling}
-To represent the Standard Model and alternatives, LHC data analysis uses
+To represent the Standard Model and alternatives, LHC data analyses use
 parametric models to describe the models' Poisson expectations of
@@
 extracting interpretable results.
-A general \heplikelihood\ model is a function $H(x)$ of some parameters $x$
-that outputs two things:
-expected yields $y$ to predict Poisson data,
+A general \heplikelihood\ model is a function $H(x)$ that takes input
+parameters $x$ and outputs two things:
+expected yields $y(x)$ to predict Poisson data,
 and an objective value $E$ to be minimized (in a `fit') to choose a best
@@
 The minimization objective $E$ is the summed negative $\log$ likelihood
-(or regularizer) for all actual and auxiliary data;
-we choose the letter $E$ for \textbf{energy}, to highlight the analogy with
+(or regularizer) for all actual and auxiliary data.
+We choose the letter $E$ for \textbf{energy}, to highlight the analogy with
 statistical physics and the Boltzmann factor
@@
 \end{equation}
-with which a system collapses to its minimum energy at low temperature
-$T = 1/\beta$~\cite{
+with which a thermodynamic system collapses to its minimum energy at cold
+temperatures $T = 1/\beta$~\cite{
 skilling2017david,
@@
 measure~\cite{cranmer2012histfactory, skilling2017david},
-but in practice such distributions often coincide with neither intuition nor
-the best fit value.
+but in my experience with \heplikelihood\ models such distributions often
+coincide with neither intuition nor the best fit value.
 
@@
 and \pyhf~\cite{heinrich2021pyhf}.
-\histfactory\ models are specialized for working with discrete numbers of
-histogram bins.
+\histfactory\ models are specialized for working with discrete histogram bins.
 They constrain parameters with either Gaussian or Poisson%
@@
 \item minimize $E(x)$ to find the best-fit parameters $\check{x}$, then
-\item sample alternative Poisson (actual) data with rates $y(\check(x))$ and
+\item sample alternative Poisson (actual) data with rates $y(\check{x})$ and
 alternative auxiliary (fictitious) data from their assigned distributions,
@@
 \end{enumerate}
-The \pvalue\ is then estimated as the fraction of toys with test statistics
+Each \pvalue\ is estimated as the fraction of sampled toys with test statistics
 greater than or equal to the observed test statistic.
@@
 distribution depends on the observed data $d$ ---
-this problematic because it means that the approximated \pvalue\ is not
+this is problematic because it means that the approximated \pvalue\ is not
 $\prob{t \geq t(d)}{h}$, but depends on the data like
@@
 Tuning and analysis are performed in a conventional and sensible procedure,
-but their pieces have somewhat esoteric names that we define here.
-
-Our searches see data only through discrete histogram bins,
-or `regions', which are defined by logical selections in event variables.
+but their pieces have esoteric names that we define here.
+
+Our searches see data only through discrete histogram bins named `regions',
+which are defined by logical selections in event variables.
 To communicate their intended uses, we further categorize these bins into
@@
 disjoint (orthogonal) smaller bins.
-Adapted from \histfitter~\cite{histfitter2015}.
+Adapted from \histfitter~\cite{baak2015histfitter}.
 }
@@
 \paragraph{Control regions (CRs).}
-Control regions select large numbers of data to
-constrain global normalization factors that scale the major
-Standard Model backgrounds;
+Control regions select large numbers of data to constrain global normalization
+factors that scale Standard Model backgrounds;
 these normalization factors control the total cross-sections of
@@
 Before applying their normalization factors to regions that are sensitive
-to signal hypotheses, data in validation regions first
-test the quality of extrapolation.
+to signal hypotheses, data in validation regions are first used to test the
+quality of extrapolation.
 Each model is adjusted until validation region data are well predicted ---
@@
 (CR) and validation (VR) data.
-Testing is done by predicting of data in the signal regions,
-in which signal yields should significantly add on top of the Standard Model
-backgrounds.
+Final testing is done by predicting of data in the signal regions, in which
+signal models could have significant effects.
 Some signal regions are labelled as \textbf{discovery regions} (\textbf{DRs})
@@
 Standard Model, that its tuned extrapolation to validation regions works,
-and that other parameters are behaving according to intuition.
+and that other parameters are behaving as intended.
 
 After fitting the background-only model to observed signal region data,
-it can be useful to check how its parameters have been `pulled' away from their
-central values.
+it can be useful to check that its parameters have been `pulled' away from
+their central values into regimes where their approximated extrapolations may
+be invalid
 Post-fit expectations in signal regions are also available with error bars;
@@
 useful to judge the background-only model.
-Instead, they are an updated prediction for the expected yield in an
+Instead, they are an updated prediction of the expected yields from an
 hypothetical future repeat of the experiment.
@@
 parameters and present contiguous `exclusion contours' within which all models
-are marked as excluded at a given level.
+are marked as excluded.
 
@@
 Despite the likely superiority of $t(n) = n$ test statistics for these
-single-bin signal-plus-background experiments, standard maximum likelihood
-ratio test statistics are used, and may explain quirks in the discovery
-region results that we shall see in the results of
-Chapter~\ref{chapter:2ljets}.
+single-bin data, the standard use of maximum likelihood ratio test statistics
+and may explain quirks in the discovery region results that we shall see in the
+results of Chapter~\ref{chapter:2ljets}.
 

### 2022-09-19 main/chapter_searches.tex
overleaf tweaks; remove cls footnotes

@@
 data, and to report that message to the wider community.
-Many different messages could be extracted from given data,
-and those messages will have various values to their recipients.
+Various messages could be extracted from given data,
+and those messages will have differing values to their recipients.
 Data analysis is not an automatic process, but a design with many judgements
@@
 unbiased priors by embracing priors as active and explicit predictions.
-This likelihood perspective the modern, practical approach to Bayesian data
+This likelihood perspective is the modern, practical approach to Bayesian data
 analysis~\cite{
@@
 High Energy Physics, notably the Poisson and Gaussian distributions.
-We sketch their derivations here, and refer to textbooks for
-details~\cite{jaynes2003probability}.
-Both which use the exponential identity that
+This section sketches their derivations, both of which use the identity that
 \begin{equation}
@@
 $p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
-Using the exponential identity of Equation~\ref{eqn:searches_exponential},
+Using the identity of Equation~\ref{eqn:searches_exponential},
 this limit tends towards the Poisson distribution~\cite{jaynes2003probability}:
@@
 to the jelly bean game of Figure~\ref{fig:searches_beans}.
-Given all you know about illustrations of jars and beans, you could make a
-vague guess without looking, but and that guess would be refined after
-inspecting the image to update your implicit internal model.
+Given all you know about jars and beans, you could make a vague guess without
+looking at the figure, and you could refine that guess by inspecting the image.
 Although I do not propose that human brains \emph{do} use probability
@@
 in the risk;
-decision-making machines (such as organisms) are do learn to make good
+decision-making machines (such as organisms) do learn to make good
 decisions without constructing either cost or prior explicitly.
@@
 another~\cite{neymanpearson1928max, neymanpearson1933lemma}.
-Rather than a \pvalue\ arousing suspicion of an hypothesis, they use it
+Rather than a \pvalue\ arousing informal suspicion of an hypothesis, they use it
 to decide on actions of `rejecting' or `accepting' an hypothesis.
@@
 probabilities may be assigned.
-Indeed they later state that
+Indeed, they later state that
 \begin{displayquote}
@@
 results that can be expected to be within a specified range.''
-In fact it is the percentage of ranges
+In fact, it is the percentage of ranges
 (confidence intervals, which are test results)
@@
 gerasimov2021new
-},%
-\footnote{%
-A draft of~\cite{HIGG-2018-51} stated that a \pvalue\ determined
-``the probability that the background-only hypothesis is compatible with the
-observed data''.
-That text was removed in response to my internal comments, but
-``probability of compatibility'' remains in its figure captions.%
-}%
-\footnote{%
-Reference~\cite{HIGG-2018-57} accurately describes a \pvalue\ as a tail
-probability, but also (repeatedly) states that a
-``probability of compatibility\ldots\ corresponds to a \pvalue''.%
-}%
+},
 and is plausibly explained by mutations from vaguely accurate statements like
@@
 predetermined threshold, say $5$ events, then models can predict whether
-$n \leq 5$, and the honest truth or falsehood in data can favour either side.
+$n \leq 5$, and the honestly answered truth or falsehood in data can favour
+either side.
 Ultimately, the reader just distrust the reporter ---
@@
 \begin{equation}
-p(x) \propto e^{-\beta E(x)}
+f(x) \propto e^{-\beta E(x)}
 .

### 2022-09-19 main/draft.tex
clean up author lists
overleaf tweaks; remove cls footnotes

@@
 
-% similarity and difference?? TODO
+%
+\footnote{%
+This context of a fixed inequality is one condition where the
+\pvalue\ ratio is not misleading when interpreted as a likelihood ratio.%
+}
+
+%
+\footnote{%
+A draft of~\cite{HIGG-2018-51} stated that a \pvalue\ determined
+``the probability that the background-only hypothesis is compatible with the
+observed data''.
+That text was removed in response to my internal comments, but
+``probability of compatibility'' remains in its figure captions.%
+}%
+\footnote{%
+Reference~\cite{HIGG-2018-57} accurately describes a \pvalue\ as a tail
+probability, but also (repeatedly) states that a
+``probability of compatibility\ldots\ corresponds to a \pvalue''.%
+}%

### 2022-09-19 main/chapter_context.tex
merge theory and experiment chapters

--- 2022-09-15 main/chapter_experiment.tex
@@
 \begin{singlespacing}
-\chapter{The ATLAS experiment}
-\label{chapter:experiment}
+\chapter{Context: theory and experiment}
+\label{chapter:context}
+%
+\begin{epigraphs}
+\qitem{%
+The idea that theorems follow from the postulates does not correspond to simple
+observation.
+If the Pythagorean theorem were found to not follow from the postulates, we
+would again search for a way to alter the postulates until it was true.
+Euclid's postulates came from the Pythagorean theorem, not the other way.}%
+{Richard~Hamming,
+\textit{The Unreasonable Effectiveness of Mathematics},
+1980~\cite{hamming1980unreasonable}}
+\qitem{%
+One geometry can not be more true than another; it can only be more convenient.
+Geometry is not true, it is advantageous.}%
+{Robert~M.~Pirsig,
+\textit{Zen and the Art of Motorcycle Maintenance},
+1974~\cite{pirsig1999zen}}
+\end{epigraphs}
+\end{singlespacing}
+
+\begin{singlespacing}
+\section{The Standard Model and beyond}
+%
+\begin{epigraphs}
+\qitem{%
+One geometry can not be more true than another; it can only be more convenient.
+Geometry is not true, it is advantageous.}%
+{Robert~M.~Pirsig,
+\textit{Zen and the Art of Motorcycle Maintenance},
+1974~\cite{pirsig1999zen}}
+\end{epigraphs}
+\end{singlespacing}
+
+\TODO{Standard Model past present future}
+
+\subsection{Particles}
+
+\subsection{The Standard Model}
+
+\subsection{Supersymmetric models}
+
+\subsubsection{Minimal supersymmetric models}
+Lightest Supersymmetric Particle (LSP)
+
+\subsubsection{Guage Mediated Supersymmetry Breaking}
+Gauge Mediated Supersymmetry Breaking (GMSB)
+
+\begin{singlespacing}
+\section{The \atlas\ Experiment}
 %
@@
 \end{singlespacing}
-
-\TODO{merge into theory chapter?}
 
@@
 
-\section{Inner tracker}
+\subsection{Inner tracker}
 \label{sec:atlas_tracker}
@@
 
-\section{Calorimeters}
+\subsection{Calorimeters}
 \label{sec:atlas_calo}
@@
 
-\section{Muon system}
+\subsection{Muon system}
 \label{sec:atlas_muon}
@@
 
-\section{Trigger}
+\subsection{Trigger}
 \label{sec:atlas_trigger}
@@
 
-\section{Software}
+\subsection{Software}
 \label{sec:atlas_software}
@@
 
-\section{Data}
+\subsection{Data}
 \label{sec:atlas_data}
 % discuss collisions, pileup, met, metsig
+
+% LHC runs

### 2022-09-19 main/chapter_searches.tex
merge theory and experiment chapters

@@
 
-\subsection{Probability, abstract}
+\subsection{Probability: abstract}
 A probability $\prob{b}{a}$ is a number, which quantifies the proportion
@@
 
-\subsection{Probability, applied}
+\subsection{Probability: applied}
 \label{sec:searches_probability_applied}

### 2022-09-19 main/chapter_2ljets.tex
fix experiment chapter ref

@@
 But there are exceptions, and exceptions to the exceptions.
-One exception, as discussed in Chapter~\ref{chapter:experiment}, is if the muon
+One exception, as discussed in Chapter~\ref{chapter:context}, is if the muon
 is hard enough that its curvature through \atlas' magnetic fields is

### 2022-09-20 main/chapter_context.tex
add experimentation citations; begin detector chapter

@@
 
-from the paper~\cite{atlas2022searches}
+Towards explaining the main result of this thesis,
+a search for effects from supersymmetric models in
+\atlas~\cite{atlas2022searches},
+we first some relevant facts about the \atlas\ detector itself.
 
-\TODO{main detector}
+\atlas\ is a Titan.
+On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
+experiment, and alongside its majors neighbours
+\cms~\cite{cms2008experiment},
+\lhcb~\cite{lhcb2008experiment},
+and ALICE~\cite{alice2008experiment}
+it is a multi-kilotonne beast,
+and a wonder of scientific engineering~\cite{
+atlas2008experiment,
+atlas1999design1,
+atlas1999design2
+}.
+
+% design goals
+
+Our search paper~\cite{atlas2022searches} contains concise and standard
+descriptions of the \atlas\ detector and its relevant components.
+Through its many inexact reproductions in past \atlas\ publications, the text
+of these descriptions has become so highly refined that my attempts to improve
+upon it would be in vain.
+Instead, this section quotes our published text with explanatory commentary.
+
 \begin{displayquote}
@@
 
+The `A' in \atlas\ does not stand for Argon, but perhaps it should.
+
 \subsection{Muon system}
@@
 % how do we predict data?
-\TODO{software}
+\TODO{software and data?}
 \begin{displayquote}

### 2022-09-20 main/chapter_context.tex
add sec references
add atlas cutaway figure

@@
 
-\subsection{Inner tracker}
-\label{sec:atlas_tracker}
-\TODO{inner tracker}
+Figure~\ref{fig:atlas_cutaway}.
+
+Inner detector~\ref{sec:atlas_inner}
+
+Calorimeter systems~\ref{sec:atlas_calo}
+
+Muon chambers~\ref{sec:atlas_muon}
+
+Trigger~\ref{sec:atlas_trigger}
+
+Software and data~\ref{sec:atlas_software_data}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
+\caption[
+A slice of \atlas' design
+]{%
+A slice of \atlas' design, reproduced from~\cite{atlas1999design1},
+including a simulated event that interacts with several sensitive subsystems.
+}
+\label{fig:atlas_cutaway}
+\end{figure}
+
+\subsection{Inner detector}
+\label{sec:atlas_inner}
+\TODO{inner detector}
 \begin{displayquote}
@@
 \label{sec:atlas_trigger}
-% L1 L2 HLT
+\TODO{L1 HLT (L2, event thing)}
 \TODO{trigger}
@@
 
-\subsection{Software}
-\label{sec:atlas_software}
-% how do we predict data?
-\TODO{software and data?}
+\subsection{Software and data}
+\label{sec:atlas_software_data}
+\TODO{how do we predict data?}
+\TODO{software and data}
 \begin{displayquote}
@@
 
-\subsection{Data}
-\label{sec:atlas_data}
-% discuss collisions, pileup, met, metsig
+\TODO{discuss collisions, pileup, met}
 
-% LHC runs
+\TODO{LHC runs}

### 2022-09-20 main/chapter_context.tex
write some on design

@@
 if observing outer space gives us a view of the past, observing inner space
-would surely give us a glimpse into the future --- would be interesting if NASA
+would surely give us a glimpse into the future - would be interesting if NASA
 made a telescope for that%
@@
 \atlas~\cite{atlas2022searches},
-we first some relevant facts about the \atlas\ detector itself.
+we first some relevant facts about the \atlas\ detector ---
+its design and history.
 
-\atlas\ is a Titan.
+The \atlas\ detector did not assemble itself into a preordained form.
+It results from a large collaborative effort towards a design that would
+provide maximally useful and novel data
+without exceeding real-world budgets in construction or operation.
+Original designs hoped to construct an \atlas\ detector that could
+observe the Higgs boson (which was predicted but yet unseen),
+precisely measure known massive objects ($W^\pm$, $Z$, $t$, $b$),
+and possibly discover new particles such as those hypothesized by
+supersymmetric models~\cite{atlas1999design1}.
+We reviews those three goals here.
+
+\paragraph{Higgs boson discovery.}
+Prior to \atlas' construction, the Higgs boson mass (and existence) was
+poorly known, but loosely constrained by results like
+$m_H \gtrsim 90\,\eV[G]$ from direct searches at the
+Large Electron–Positron Collider (LEP),
+$m_H = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
+and $m_H \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
+atlas1999design2,
+ghinculov1998perturb,
+lep1999ewk
+}.
+
+\paragraph{Precise measurements of known objects.}
+
+\paragraph{Supersymmetry discovery.}
+This is not a failure. It is a success.
+
+\paragraph{Design.}
+Atlas is a Titan.
 On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
@@
 \caption[
-A slice of \atlas' design
+A central slice of \atlas' design
 ]{%
-A slice of \atlas' design, reproduced from~\cite{atlas1999design1},
+A central slice of \atlas' design, reproduced from~\cite{atlas1999design1},
 including a simulated event that interacts with several sensitive subsystems.

### 2022-09-20 main/chapter_searches.tex
write some on design

@@
 }.
-High-profile usage in combined Higgs boson searches at the
-Large Electron–Positron Collider~\cite{
+High-profile usage in combined Higgs boson searches at LEP~\cite{
 read1997optimal,

### 2022-09-20 main/chapter_context.tex
write some about higgs

@@
 }.
+
+The Higgs boson couples to an extraordinary variety of fundamental particles,
+and \atlas' designers achieved discovery sensitivity across the entire
+$m_H \in 100\textrm{--}1000\,\eV[G]$
+window by targetting decays including
+$H\to b\bar b$,
+$H\to \gamma\gamma$,
+$H\to ZZ^* \to 4\ell$, and
+$H\to ZZ \to \ell\ell\nu\nu$.
+This spectrum demands precise measurements of leptons, photons, and
+$b$-jets, as well as sensitivity to the missing transverse momentum of
+invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
+masses.
+
+% Famously, \atlas\ and \cms\ have since found a Higgs boson candidate at
+% $m_H = 125\,\eV[G]$~\cite{
+% }
+% with all of its predicted Standard Model properties~\cite{
+% }.
 

### 2022-09-20 main/chapter_context.tex
add summaryt of lhc higgs results with citations
some todos and tweaks

--- 2022-09-20 main/chapter_context.tex
@@
 observe the Higgs boson (which was predicted but yet unseen),
-precisely measure known massive objects ($W^\pm$, $Z$, $t$, $b$),
+precisely measure known massive objects ($W^\pm$, $Z$, $t$),
 and possibly discover new particles such as those hypothesized by
 supersymmetric models~\cite{atlas1999design1}.
-We reviews those three goals here.
+We now review those three goals.
 
@@
 poorly known, but loosely constrained by results like
-$m_H \gtrsim 90\,\eV[G]$ from direct searches at the
+$m_H \gtrsim 90\,\eV[G]$ from searches at the
 Large Electron–Positron Collider (LEP),
@@
 
+The Higgs boson couples to an extraordinary variety of fundamental particles,
+and \atlas' designers achieved discovery sensitivity across the entire
+$100$ to $1000\,\eV[G]$
+window by targetting decays including
+$H\to b\bar b$,
+$H\to \gamma\gamma$,
+$H\to ZZ^* \to 4\ell$, and
+$H\to ZZ \to \ell\ell\nu\nu$.
+This spectrum demands precise measurements of leptons, photons, and
+$b$-jets, as well as sensitivity to the missing transverse momentum of
+invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
+masses.
+
+Both \atlas\ and \cms\ have now found a Higgs boson candidate at
+$m_H = 125\,\eV[G]$~\cite{
+atlas2012higgs,
+atlas2012combined,
+cms2012higgs
+}
+and confirmed all of its tested Standard Model properties~\cite{
+combined2016higgs,
+atlas2022ten,
+cms2022ten
+}.
+
 \paragraph{Precise measurements of known objects.}
+\TODO{The top quark's almost-exclusive decay to a $b$ ... $b$ sensitivity}
+% top physics: tttt, ttH
+% rare ZZy ?
+% W mass measurement
 
@@
 This is not a failure. It is a success.
+Our goal is not to discover something that is not there, but to learn more
+about nature.
+The nonappearance of supersymmetric particles at the LHC is itself a discovery.
 
@@
 
+\TODO{upgraded with IBL}
+
 \subsection{Calorimeters}

### 2022-09-20 main/chapter_context.tex
condense a couple of sentences

@@
 The \atlas\ detector did not assemble itself into a preordained form.
-It results from a large collaborative effort towards a design that would
-provide maximally useful and novel data
-without exceeding real-world budgets in construction or operation.
+It was designed with intent to provide maximally useful and novel data
+within real-world cost budgets.
 Original designs hoped to construct an \atlas\ detector that could
@@
 supersymmetric models~\cite{atlas1999design1}.
-We now review those three goals.
+We now review these three goals with the subsequent news from \atlas'
+operations.
 
@@
 }.
-
 The Higgs boson couples to an extraordinary variety of fundamental particles,

### 2022-09-20 main/chapter_context.tex
add some crap sm notes

@@
 
-The \atlas\ detector did not assemble itself into a preordained form.
-It was designed with intent to provide maximally useful and novel data
-within real-world cost budgets.
+The \atlas\ detector was not delivered in a preordained form.
+It was designed with intent to provide maximally useful and novel data within
+real-world cost budgets.
 Original designs hoped to construct an \atlas\ detector that could
 observe the Higgs boson (which was predicted but yet unseen),
-precisely measure known massive objects ($W^\pm$, $Z$, $t$),
+precisely measure known massive objects ($W$, $Z$, and $t$),
 and possibly discover new particles such as those hypothesized by
@@
 }.
-The Higgs boson couples to an extraordinary variety of fundamental particles,
-and \atlas' designers achieved discovery sensitivity across the entire
-$100$ to $1000\,\eV[G]$
-window by targetting decays including
+The Higgs boson couples to an extraordinary variety of fundamental particles.
+For its discovery alone, \atlas' designers achieved sensitivity across the
+entire $100$ to $1000\,\eV[G]$ window by targetting decays including
 $H\to b\bar b$,
@@
 $H\to ZZ^* \to 4\ell$, and
-$H\to ZZ \to \ell\ell\nu\nu$.
-This spectrum demands precise measurements of leptons, photons, and
+$H\to ZZ \to \ell\ell\nu\nu$;
+these channels demand precise measurements of leptons, photons, and
 $b$-jets, as well as sensitivity to the missing transverse momentum of
@@
 
-\paragraph{Precise measurements of known objects.}
-\TODO{The top quark's almost-exclusive decay to a $b$ ... $b$ sensitivity}
+\paragraph{Measurement of known massive objects.}
+Other than the Higgs boson, the heavy Standard Model resonances are the
+$W$ and $Z$ bosons and the top quark, all of which are produced at
+high rates by LHC proton-proton collisions.
+
+The goal of measuring these heavy objects puts more emphasis on determining
+the total energy of hadronic jets than their precise hadronic compositions.
+
+\TODO{Although \atlas\ has an active B-physics programme, the stronger particle
+identification of the \lhcb\ detector makes it more competitive for hadronic
+physics.?}
+\TODO{Heavy ions and ALICE?}
+
+\TODO{standard model~\cite{atlas2021summarysm}}
+
+\TODO{top mass~\cite{atlas2022symmarytop, atlas2019topmass}}
+
+\TODO{Top factory. The top quark's almost-exclusive decay as $t\rightarrow Wb$ ... $b$ sensitivity}
 % top physics: tttt, ttH
@@
 % W mass measurement
+
+\TODO{$W$ mass~\cite{atlas2018wmass}}
+\TODO{`diboson' $WZ$ $W\gamma$}
+
+\TODO{B-physics (supplements dedicated results from \lhcb)}
+
+$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
 

### 2022-09-20 main/chapter_context.tex
scribble some more notes

@@
 \paragraph{Measurement of known massive objects.}
+\TODO{
+Argument:
+want to measure W, t,
+so measure jets and precise objects
+so weaker B and heavy ion programme.
+...
+Refer to competitive measurements of W and t masses
+and xsec of various rare processes (diboson, triboson)
+}
+
 Other than the Higgs boson, the heavy Standard Model resonances are the
@@
 The goal of measuring these heavy objects puts more emphasis on determining
-the total energy of hadronic jets than their precise hadronic compositions.
+the total energy of the hadronic jets produced by their hadronic decays
+than precise hadrons that those jets comprise.
 
@@
 
-\paragraph{Design.}
+\TODO{
+Decay to SM particles,
+emphasis on missing transverse momentum.
+...
+Many SUSY searches, many other exotics searches, no convincing evidence for
+new effects.
+}
+
+\subsection{Design}
 Atlas is a Titan.

### 2022-09-20 main/chapter_2ljets.tex
add ssfm cite; use auto bib for foundations and algorithms
more awful attempts at writing sm measurements; fix - in pages

@@
 $p = |\pm \vec p|$, the two particles have different energies since they
-have different masses: $E_Z^2 = m_Z^2 + p^2$, $E_g^2 = 0 + p^2$.
+have different masses: $E_Z^2 = m(Z)^2 + p^2$, $E_g^2 = 0 + p^2$.
 % hobbyhorse
@@
 $m(\neutralino_2) = m(\neutralino_1)$ and
-$m(\neutralino_2) = m(\neutralino_1) - m_Z$ diagonal lines.
+$m(\neutralino_2) = m(\neutralino_1) - m(Z)$ diagonal lines.
 }
@@
 Naturally, \atlas\ has an unblinding procedure that requires committee
-review and approval in accordance with internal documentation.
+review and approval in accordance with internal
+documentation~\cite{usa1944simple}.
 In the \atlas\ SUSY world, searches are nominally performed blind to signal

### 2022-09-20 main/chapter_context.tex
more awful attempts at writing sm measurements; fix - in pages

--- 2022-09-20 main/chapter_context.tex
@@
 real-world cost budgets.
-Original designs hoped to construct an \atlas\ detector that could
-observe the Higgs boson (which was predicted but yet unseen),
-precisely measure known massive objects ($W$, $Z$, and $t$),
-and possibly discover new particles such as those hypothesized by
-supersymmetric models~\cite{atlas1999design1}.
+Within general purpose utilities, \atlas' designs aimed to construct a
+detector that could
+\begin{itemize}
+\item find the Higgs boson (which was predicted but yet unseen),
+\item precisely measure known massive objects ($W$, $Z$, and $t$),
+\item discover new (supersymmetric) particles~\cite{atlas1999design1}.
+\end{itemize}
 We now review these three goals with the subsequent news from \atlas'
@@
 
-\paragraph{Higgs boson discovery.}
+\paragraph{Find the Higgs boson.}
 Prior to \atlas' construction, the Higgs boson mass (and existence) was
 poorly known, but loosely constrained by results like
-$m_H \gtrsim 90\,\eV[G]$ from searches at the
+$m(H) \gtrsim 90\,\eV[G]$ from searches at the
 Large Electron–Positron Collider (LEP),
-$m_H = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
-and $m_H \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
+$m(H) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
+and $m(H) \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
 atlas1999design2,
@@
 }.
+
 The Higgs boson couples to an extraordinary variety of fundamental particles.
@@
 Both \atlas\ and \cms\ have now found a Higgs boson candidate at
-$m_H = 125\,\eV[G]$~\cite{
+$m(H) = 125\,\eV[G]$~\cite{
 atlas2012higgs,
@@
 cms2022ten
-}.
+}
+such as its production mechanisms, decay rates,
+spin, and parity~\cite{HIGG-2013-01, HIGG-2013-17, HIGG-2014-06}.
 
-\paragraph{Measurement of known massive objects.}
+\paragraph{Precisely measure known massive objects.}
 Other than the Higgs boson, the heavy Standard Model resonances are the
-$W$ and $Z$ bosons and the top quark, all of which are produced at
-high rates by LHC proton-proton collisions.
+$W$ and $Z$ bosons and the top quark.
+All of these are produced at high rates in LHC proton-proton collisions,
+and present opportunities for \atlas\ to measure their properties
+(such as masses and production cross-sections) with unprecedented precision.
 
-The goal of measuring these heavy objects puts more emphasis on determining
-the total energy of hadronic jets than their precise hadronic compositions.
-
-\TODO{Although \atlas\ has an active B-physics programme, the stronger particle
-identification of the \lhcb\ detector makes it more competitive for hadronic
-physics.?}
-\TODO{Heavy ions and ALICE?}
+Sensitivity to hadronic momenta is crucial,
+both directly to measure heavy hadronic decays
+(like $t \rightarrow bq\bar{q}$),
+and indirectly for the missing transverse momenta from semi-invisible
+leptonic decays (like $W \to \ell\nu$).
+Since individual hadron identities are less relevant to these measurements,%
+\footnote{%
+Rare decays like
+$W^\pm\rightarrow \pi^\pm \gamma$ or $W^\pm\rightarrow \pi^\pm \pi^+ \pi^-$
+could give clean signals, but are not competitive due to their
+tiny production rates and difficult identification~\cite{
+cdf1996search,
+mangano2014wpiy,
+cms2021wpiy
+}.%
+}
+it makes sense for \atlas' design to forfeit hadron identification systems
+in favour of precise calorimetry with total angular coverage.
 
@@
 
-$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
+\TODO{w mass conflict~\cite{cdf2022high}}
 
-\paragraph{Supersymmetry discovery.}
+\paragraph{Discover new (supersymmetric) particles.}
 This is not a failure. It is a success.
@@
 
-\paragraph{Design.}
+\TODO{
+Decay to SM particles,
+emphasis on missing transverse momentum.
+...
+Many SUSY searches, many other exotics searches, no convincing evidence for
+new effects.
+}
+
+$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
+
+\subsection{Design}
 Atlas is a Titan.

### 2022-09-20 main/draft.tex
more awful attempts at writing sm measurements; fix - in pages

@@
 }%
+
+% awful awful tangled up day
+
+\TODO{B-hadrons and LHCb?}
+% more cost-effective to improve calorimetry at the expense of PID
+
+% measurements are supported by complete and precise calorimetry
+% PID is expensive and less useful in comparison
+
+To precise measure of these massive objects
+justifies a focus on complete and precise calorimetry
+at the expense of exact particle identification.
+
+\TODO{
+Argument:
+want to measure W, t,
+so measure jets and precise objects
+so weaker B and heavy ion programme.
+...
+Refer to competitive measurements of W and t masses
+and xsec of various rare processes (diboson, triboson)
+}
+
+Towards maximizing the quality of these precision measurements,
+
+sensitivity to
+exact hadron identities is comparatively less cost-effective.%
+
+To maximize the quality of measurements like $m(t)$ and $m(W)$, it therefore
+makes sense for \atlas' design to prefer complete and accurate calorimetry over
+hadron identification.
+
+Such hadronic momentum measurements contribute , both of which are
+leveraged for \atlas' measurements of $m(t)$ and $m(W)$~\cite{
+atlas2022symmarytop,
+atlas2019topmass
+}.
+
+Although \atlas\ has an active programmes in other Standard Model research such
+as B-physics and heavy-ion collisions, these are rarely competitive with the specialized experiments
+\lhcb\ and ALICE respectively.

### 2022-09-21 main/chapter_context.tex
write some awkward text for heavy sm, start adding citation

@@
 
-\TODO{standard model~\cite{atlas2021summarysm}}
-
-\TODO{top mass~\cite{atlas2022symmarytop, atlas2019topmass}}
-
-\TODO{Top factory. The top quark's almost-exclusive decay as $t\rightarrow Wb$ ... $b$ sensitivity}
-% top physics: tttt, ttH
-% rare ZZy ?
-% W mass measurement
-
-\TODO{$W$ mass~\cite{atlas2018wmass}}
-\TODO{`diboson' $WZ$ $W\gamma$}
-
-\TODO{B-physics (supplements dedicated results from \lhcb)}
-
-\TODO{w mass conflict~\cite{cdf2022high}}
+\atlas\ has now produced a plethora of competitively precise measurements
+of the heavy Standard Model~\cite{atlas2021summarysm},
+including $m(W)$~\cite{atlas2018wmass} and
+$m(t)$~\cite{atlas2022symmarytop, atlas2019topmass},
+but also details of their behaviours~\TODO{cite}
+and production cross-sections at various centre-of-mass energies.
+Rare processes have also been studied in new levels of detail,
+including
+electroweak diboson ($WW/WZ/ZZ$ etc.)~\cite{
+STDM-2015-21,
+STDM-2015-23,
+STDM-2017-09
+}
+and triboson~\cite{
+STDM-2016-06,
+STDM-2017-22,
+STDM-2019-09
+},
+and rare top processes
+($\ttbar Z$, $\ttbar\ttbar$ etc.)~\TODO{cite},
+These continue to not validate the Standard Model, and are aided by a total
+lack of new phenomena which could have overpowered these rare signals.
 

### 2022-09-21 main/chapter_searches.tex
write some awkward text for heavy sm, start adding citation

@@
 $2 m(Z) \approx 180\,\eV[G]$ kinematic edge,
-and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$,
+and $Z \to 4\ell$ around $m_{4\ell} \approx m(Z)$~\cite{STDM-2017-09},
 so expanding the histogram to show either of these tails would change nothing

### 2022-09-21 main/chapter_context.tex
add more citations for heavy standard model paragraph, begin notes for susy

--- 2022-09-20 main/chapter_context.tex
@@
 such as its production mechanisms, decay rates,
-spin, and parity~\cite{HIGG-2013-01, HIGG-2013-17, HIGG-2014-06}.
+spin, and parity~\cite{
+HIGG-2013-01,
+HIGG-2013-17,
+HIGG-2014-06
+}.
 
@@
 
-\TODO{standard model~\cite{atlas2021summarysm}}
-
-\TODO{top mass~\cite{atlas2022symmarytop, atlas2019topmass}}
-
-\TODO{Top factory. The top quark's almost-exclusive decay as $t\rightarrow Wb$ ... $b$ sensitivity}
-% top physics: tttt, ttH
-% rare ZZy ?
-% W mass measurement
-
-\TODO{$W$ mass~\cite{atlas2018wmass}}
-\TODO{`diboson' $WZ$ $W\gamma$}
-
-\TODO{B-physics (supplements dedicated results from \lhcb)}
-
-\TODO{w mass conflict~\cite{cdf2022high}}
+\atlas\ has now produced a plethora of competitively precise measurements
+of the heavy Standard Model~\cite{atlas2021summarysm},
+including $m(W)$~\cite{atlas2018wmass} and
+$m(t)$~\cite{atlas2022symmarytop, atlas2019topmass, TOPQ-2015-03},
+but also details of differential cross-sections~\cite{
+STDM-2016-11,
+STDM-2016-14,
+TOPQ-2018-15,
+TOPQ-2016-10
+}.
+Rare processes have also been studied in new levels of detail,
+including
+electroweak diboson ($WW/WZ/ZZ$ etc.)~\cite{
+STDM-2015-21,
+STDM-2015-23,
+STDM-2017-09
+}
+and triboson~\cite{
+STDM-2016-06,
+STDM-2017-22,
+STDM-2019-09
+},
+and rare top processes
+($\ttbar Z$, $\ttbar W$ etc.)~\cite{
+TOPQ-2013-05,
+TOPQ-2018-01,
+TOPQ-2020-03
+},
+These continue to not validate the Standard Model, and are aided by the lack of
+new phenomena, which could have overpowered such rare signals.
 
 \paragraph{Discover new (supersymmetric) particles.}
+\TODO{
+has discovery potential up to multi-TeV mass scales,
+expect decays to knwon particles, so complements Higgs,
+discovery and Standard Model measurements,
+indeed search of this analysis benefits both from ZZ llvv and Hbb,
+$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
+}
+
 This is not a failure. It is a success.
@@
 }
-
-$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
 

### 2022-09-21 main/chapter_context.tex
add mt bhadron footnote

@@
 both directly to measure heavy hadronic decays
-(like $t \rightarrow bq\bar{q}$),
+(like $t \to bq\bar{q}$),
 and indirectly for the missing transverse momenta from semi-invisible
@@
 Rare decays like
-$W^\pm\rightarrow \pi^\pm \gamma$ or $W^\pm\rightarrow \pi^\pm \pi^+ \pi^-$
+$W^\pm \to \pi^\pm \gamma$ or $W^\pm \to \pi^\pm \pi^+ \pi^-$
 could give clean signals, but are not competitive due to their
@@
 cms2021wpiy
+}.%
+Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but these
+do not require hadron reconstruction~\cite{
+CDF:2009mbf,
+CMS:2016ixg,
+ATLAS:2022jbw
 }.%

### 2022-09-21 main/chapter_context.tex
add discover susy text and citations

@@
 cms2021wpiy
-}.%
-Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but these
-do not require hadron reconstruction~\cite{
+}.
+Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but do not
+require hadron identification~\cite{
 CDF:2009mbf,
@@
 \paragraph{Discover new (supersymmetric) particles.}
-\TODO{
-has discovery potential up to multi-TeV mass scales,
-expect decays to knwon particles, so complements Higgs,
-discovery and Standard Model measurements,
-indeed search of this analysis benefits both from ZZ llvv and Hbb,
-$\twoljets$ piggyback on the $ZZ \to \ell\ell\nu\nu$ channel
+High-energy proton-proton collisions at the LHC therefore offer sensitivity
+to new particles up to multi-TeV mass scales.
+Heavy new particles are anticipated to either decay to known Standard Model
+objects, or to be invisible in \atlas.
+Designs that can discover such new particles are therefore neatly aligned with
+the other goals to refining the Standard Model
+and to discovering the Higgs boson at any mass.
+Visible decay products should already be caught by \atlas' abilities
+to measure the Higgs boson and other Standard Model particles,
+and the prospect of new invisible particles reinforces the need for precise
+measurements of the missing transverse momentum.
+
+Indeed, the main search presented in this thesis~\cite{atlas2022searches}
+joins nicely with the pre-existing needs for sensitivity to signatures
+including
+$H \to b\bar b$
+and
+$ZZ \to \ell\ell\nu\nu$.
+
+Dozens of searches for new phenomena have now been conducted in \atlas\ data,
+with~\cite{
+ATL-PHYS-PUB-2022-013
 }
-
+and without~\cite{
+ATL-PHYS-PUB-2022-007,
+ATL-PHYS-PUB-2022-012,
+ATL-PHYS-PUB-2022-036,
+ATL-PHYS-PUB-2022-034
+} supersymmetry,
+and show no convincing evidence for effects beyond the Standard Model.
 This is not a failure. It is a success.
-Our goal is not to discover something that is not there, but to learn more
-about nature.
-The nonappearance of supersymmetric particles at the LHC is itself a discovery.
-
-\TODO{
-Decay to SM particles,
-emphasis on missing transverse momentum.
-...
-Many SUSY searches, many other exotics searches, no convincing evidence for
-new effects.
-}
+Our goal in science is not to discover something that is not there,
+but to learn more about nature.
+Discovering no tangible new particles in LHC conditions is a discovery itself.
 
@@
 }.
-
+s
 % design goals

### 2022-09-21 main/chapter_context.tex
write detector intro; remove csquotes for plain quote; add footnote footnote

@@
 but to learn more about nature.
-Discovering no tangible new particles in LHC conditions is a discovery itself.
+The discovery no tangible new particles in LHC conditions is itself a
+discovery.
 
@@
 On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
-experiment, and alongside its majors neighbours
+experiment, and alongside its big neighbours
 \cms~\cite{cms2008experiment},
@@
 }.
-s
-% design goals
 
@@
 
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 $\Delta R \equiv \sqrt{(\Delta\eta)^{2} + (\Delta\phi)^{2}}$.%
-''~\cite{atlas2022searches}
+''\footnotemark~\cite{atlas2022searches}%
 }
@@
 ''~\cite{atlas2022searches}
-\end{displayquote}
+\end{quote}
+To illustrate this description, a vertical slice of \atlas\ is displayed in
+Figure~\ref{fig:atlas_cutaway}, and each subsystem is described in more detail
+in a subsection below:
+the inner detector in Section~\ref{sec:atlas_inner},
+the electromagnetic and hadronic calorimeter systems in
+Section~\ref{sec:atlas_calo},
+and
+the chambers of the muon spectrometer in Section~\ref{sec:atlas_muon}.
+The final connection to the data analysed in this thesis is made through the
+trigger system which is described in Section~\ref{sec:atlas_trigger},
+and details of the data themselves are given in
+Section~\ref{sec:atlas_data}
 
-Figure~\ref{fig:atlas_cutaway}.
-
-Inner detector~\ref{sec:atlas_inner}
-
-Calorimeter systems~\ref{sec:atlas_calo}
-
-Muon chambers~\ref{sec:atlas_muon}
-
-Trigger~\ref{sec:atlas_trigger}
-
-Software and data~\ref{sec:atlas_software_data}
+\footnotetext{%
+This footnote, which was perfected in~\cite{lester2013heffalon},
+``should be used in every \atlas\ publication''~\cite{atlas_footnote}.
+True to form, internal \atlas\ guidelines ask that
+``If you think you can improve on it, please consult with the
+Publications Committee''~\cite{atlas_coordinate}.
+We somehow added the sentence defining angular distance without such a
+consultation.%
+}
 
@@
 \TODO{inner detector}
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 ''~\cite{atlas2022searches}
-\end{displayquote}
+\end{quote}
 
@@
 \TODO{calo}
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 \TODO{muon system}
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 \TODO{trigger}
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 ''~\cite{atlas2022searches}
-\end{displayquote}
+\end{quote}
 
-\subsection{Software and data}
-\label{sec:atlas_software_data}
+\subsection{Data}
+\label{sec:atlas_data}
 \TODO{how do we predict data?}
 \TODO{software and data}
-\begin{displayquote}
+\begin{quote}
 ``%

### 2022-09-21 main/chapter_searches.tex
write detector intro; remove csquotes for plain quote; add footnote footnote

@@
 In his own words,%
-\begin{displayquote}
+\begin{quote}
 \small
@@
 ''~\cite{fisher1921probable},
-\end{displayquote}
+\end{quote}
 and furthermore we agree that in comparison to other prescriptions
@@
 Section~\ref{sec:searches_np}),
-\begin{displayquote}
+\begin{quote}
 \small
@@
 ''~\cite{fisher1956statistical}.
-\end{displayquote}
+\end{quote}
 For example, Figure~\ref{fig:searches_fisher_likelihood} displays a figure by
@@
 Indeed, they later state that
-\begin{displayquote}
+\begin{quote}
 ``%
@@
 ''~\cite{neymanpearson1933lemma}
-\end{displayquote}
+\end{quote}
 Elsewhere, they also consider an example with two hypotheses $\mathrm{A}_1$
@@
 Pratt succinctly explains the confidence interval by example:
-\begin{displayquote}
+\begin{quote}
 \small
@@
 ''~\cite{pratt1961testing}.
-\end{displayquote}
+\end{quote}
 Neyman and Pearson never claimed otherwise; they did precise and rigorous
@@
 Fisher also has strong words here:
-\begin{displayquote}
+\begin{quote}
 \small
@@
 ''~\cite{fisher1956statistical}
-\end{displayquote}
+\end{quote}
 Amen.

### 2022-09-21 main/chapter_context.tex
write some on inner tracker

@@
 Through its many inexact reproductions in past \atlas\ publications, the text
-of these descriptions has become so highly refined that my attempts to improve
-upon it would be in vain.
+of these descriptions has become so highly refined that attempts to improve
+it myself would probably be in vain.
 Instead, this section quotes our published text with explanatory commentary.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
+\caption[
+A central slice of \atlas' design
+]{%
+A central slice of \atlas' design, reproduced from~\cite{atlas1999design1},
+including a simulated event that interacts with several sensitive subsystems.
+}
+\label{fig:atlas_cutaway}
+\end{figure}
 
@@
 in a subsection below:
-the inner detector in Section~\ref{sec:atlas_inner},
+the inner tracking detector in Section~\ref{sec:atlas_inner},
 the electromagnetic and hadronic calorimeter systems in
@@
 Publications Committee''~\cite{atlas_coordinate}.
-We somehow added the sentence defining angular distance without such a
-consultation.%
+We added the sentence defining angular distance without such a consultation.%
 }
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
-\caption[
-A central slice of \atlas' design
-]{%
-A central slice of \atlas' design, reproduced from~\cite{atlas1999design1},
-including a simulated event that interacts with several sensitive subsystems.
-}
-\label{fig:atlas_cutaway}
-\end{figure}
 
@@
 \label{sec:atlas_inner}
-\TODO{inner detector}
+The inner tracking detectors begin \atlas' measurements of new collisions
+by precisely tracing the paths taken by energetic charged particles
+as they curve through the solenoidal magnetic field.
+Track curvature provides momentum information,
+tracks' associations into collision vertices helps to separate
+pile-up,
+and their associations into displaced vertices help to identify
+particle decays.
+
+Our description continues:
 \begin{quote}
@@
 \end{quote}
+The insertable B-layer is a silicon pixel detector that begins just
+$3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
+present and future high LHC luminosities,
+particularly for sensitivity to `Goldilocks Zone' particles
+(B-hadrons, occasionally tau leptons, and maybe new species) whose expected
+lifetimes are just right to decay inside the tracking system.
+Following radially outwards, the silicon pixel and silicon microstrip%
+\footnote{%
+Formerly known as the SemiConductor Tracker~\cite{atlas1999design1},
+the silicon microstrip system retains the
+SCT~\cite{atlas2008experiment}.%
+}
+detectors (which use silicon technology)
+and the transition radiation tracking detectors (which uses straw tubes).
 
-\TODO{upgraded with IBL}
+As for the whole \atlas\ detector, complete $\phi$ coverage and a large $\eta$
+range minimizes leakage.
+Towards \atlas' design goals, this inner tracking detector crucially provides
+measurements of electrons, B-hadrons, and the missing transverse momentum.
 
@@
 \label{sec:atlas_calo}
+The `A' in \atlas\ does not stand for Argon, but perhaps it should.
 \TODO{calo}
@@
 \end{quote}
-
-The `A' in \atlas\ does not stand for Argon, but perhaps it should.
 

### 2022-09-22 main/chapter_context.tex
write on calo

@@
 \label{sec:atlas_calo}
-The `A' in \atlas\ does not stand for Argon, but perhaps it should.
-\TODO{calo}
+Inelastic collisions cause energetic particles to scatter, distributing their
+energy into macroscopic particle showers.
+Calorimetry uses this principle to measure energies
+by interleaving dense absorbing materials (which initiate showers)
+with sensitive materials that convert showers into electrical signals.
+In \atlas,
 \begin{quote}
@@
 \end{quote}
+The EM calorimeters provide located electron/photon energy measurements,
+and the hadronic calorimeters subsequently aim to stop and measure hadronic
+activity for later analysis in jet clustering.
+Maximal angular coverage by the calorimeters further facilitates missing
+transverse momentum estimates.
+
+Neither `A' in \atlas\ stands for Argon, but perhaps one should.
+Liquid-argon is the chosen active material in the EM and forward calorimeters
+for its robust, linear response in converting showers to electrical signals
+for readout by electrodes.
+Liquid argon is poured as a liquid to fill the gaps between the various
+metallic absorbers, which are folded into zigzag ``accordion'' patterns that
+optimize uniformity around the detector.
+The steel/scintillator hadronic tile calorimeter is a
+``simple and cost effective''~\cite{atlas1996tile} design that uses
+photomultiplier tubes to convert light from its polystyrene scintillators into
+electrical signals.
+Particles lose some energy before reaching the calorimeters; a thin
+liquid-argon pre-sampler is installed before the central EM calorimeter
+to regain such loss.
+
+These calorimeter systems achieve attenuations of at least ten interaction
+lengths throughout their $|\eta| < 4.9$ range.
+Maximizing attenuation minimizes noise from hadronic `punch-through' into the
+muon chambers.
 

### 2022-09-22 main/chapter_context.tex
tweak footnote footnote

@@
 \footnotetext{%
-This footnote, which was perfected in~\cite{lester2013heffalon},
-``should be used in every \atlas\ publication''~\cite{atlas_footnote}.
+This footnote and appears in every \atlas\ publication, and was perfected
+in~\cite{lester2013heffalon}.
 True to form, internal \atlas\ guidelines ask that
 ``If you think you can improve on it, please consult with the
-Publications Committee''~\cite{atlas_coordinate}.
+Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
 We added the sentence defining angular distance without such a consultation.%

### 2022-09-22 main/chapter_context.tex
add magnet figure

@@
 \footnotetext{%
-This footnote and appears in every \atlas\ publication, and was perfected
+This footnote, which appears in every \atlas\ publication, was perfected
 in~\cite{lester2013heffalon}.
@@
 and their associations into displaced vertices help to identify
-particle decays.
-
-Our description continues:
+decays.
 \begin{quote}
@@
 \end{quote}
-The insertable B-layer is a silicon pixel detector that begins just
+This insertable B-layer is a silicon pixel detector that begins just
 $3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
 present and future high LHC luminosities,
-particularly for sensitivity to `Goldilocks Zone' particles
-(B-hadrons, occasionally tau leptons, and maybe new species) whose expected
-lifetimes are just right to decay inside the tracking system.
+particularly for sensitivity to particles whose lifetimes are just right to
+decay inside the tracking system;
+These `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
+maybe new species.
+
 Following radially outwards, the silicon pixel and silicon microstrip%
@@
 Liquid argon is poured as a liquid to fill the gaps between the various
-metallic absorbers, which are folded into zigzag ``accordion'' patterns that
+metallic absorbers, which are folded into zigzag ``accordion'' shapes that
 optimize uniformity around the detector.
@@
 \label{sec:atlas_muon}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
+\caption[
+Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+solenoid
+]{%
+Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+solenoid.
+Reproduced from~\cite{atlas2008experiment}.
+}
+\label{fig:atlas_magnets}
+\end{figure}
+
 \TODO{muon system}
@@
 \end{quote}
+\TODO{four kinds of chambers}
+
+Figure~\ref{fig:atlas_magnets}
 

### 2022-09-22 main/chapter_context.tex
add muon text; begin trigger

@@
 detectors (which use silicon technology)
-and the transition radiation tracking detectors (which uses straw tubes).
+and the transition radiation tracking detectors (which uses `straw' tubes
+gas tubes).
 
@@
 \label{sec:atlas_muon}
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
-\caption[
-Layout of \atlas' toroidal magnets around the tile calorimeter and inner
-solenoid
-]{%
-Layout of \atlas' toroidal magnets around the tile calorimeter and inner
-solenoid.
-Reproduced from~\cite{atlas2008experiment}.
-}
-\label{fig:atlas_magnets}
-\end{figure}
-
-\TODO{muon system}
+Muons penetrate matter well because of their large mass and non-interaction
+with the strong nuclear force.
+\atlas\ doesn't try to stop energetic muons;
+above around $3\,\eV[G]$, they pass right through the calorimeters, and are
+instead measured by their curved tracks in a toroidal magnetic field that
+envelopes dedicated detectors on \atlas' periphery.
 \begin{quote}
@@
 \end{quote}
-\TODO{four kinds of chambers}
+Muon subsystems come in four varieties, all of which operate by detecting
+charge scattered by the muons as they fly through gaseous media.
+These systems track the muons' paths as the magnets bend them in the
+longitudinal plane; these toroidal magnets generate fields outside the tile
+calorimeter around the central barrel and in the end-caps as illustrated in
+Figure~\ref{fig:atlas_magnets}.
 
-Figure~\ref{fig:atlas_magnets}
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
+\caption[
+Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+solenoid
+]{%
+Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+solenoid, including the eight barrel and end-cap coils.
+Figure reproduced from~\cite{atlas2008experiment}.
+The \atlas\ muon spectrometer measures muons' tracks as they bend through
+this peripheral magnetic field.
+}
+\label{fig:atlas_magnets}
+\end{figure}
+
+Measurements of the tracks' curvature yield estimates of the muons' momenta;
+although extremely energetic muons will have straight tracks,
+\atlas' design and calibration achieves a high precision that increases to
+only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse momenta~\cite{
+atlas2008experiment,
+ATL-PHYS-PUB-2015-037
+}.
 
@@
 \TODO{L1 HLT (L2, event thing)}
-\TODO{trigger}
+\TODO{trigger~\cite{atlas2016trigger}}
+% 25ns bunch spacing
+\TODO{
+With $25\,\mathrm{ns}$ bunch spacing and \atlas' $25\,\textrm{m}$ diameter,
+light-speed particles ejected from one crossing are still zipping through
+\atlas\ when the next crossing happens.
+}
+% fast decisions to not overflow buffer queue
 \begin{quote}

### 2022-09-22 main/chapter_context.tex
diameter to radius; add atlas proposal

@@
 On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
-experiment, and alongside its big neighbours
+experiment, and alongside its biggest neighbours
 \cms~\cite{cms2008experiment},
@@
 and a wonder of scientific engineering~\cite{
+atlas1994proposal,
 atlas2008experiment,
@@
 only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse momenta~\cite{
+atlas1994proposal,
 atlas2008experiment,
@@
 \TODO{
-With $25\,\mathrm{ns}$ bunch spacing and \atlas' $25\,\textrm{m}$ diameter,
+With $25\,\mathrm{ns}$ bunch spacing and \atlas' $12.5\,\textrm{m}$ radius,
 light-speed particles ejected from one crossing are still zipping through
-\atlas\ when the next crossing happens.
+\atlas\ when the next collision arrives.
 }

### 2022-09-22 main/chapter_2ljets.tex
deduplicate atlas experiment citation; write trigger

@@
 electrons, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
-decaying tau leptons~\cite{atlas_PERF_2007_01}.
+decaying tau leptons~\cite{atlas2008experiment}.
 

### 2022-09-22 main/chapter_context.tex
deduplicate atlas experiment citation; write trigger

@@
 ``%
-The ATLAS detector~\cite{PERF-2007-01} is a multipurpose particle detector with
-a forward--backward symmetric cylindrical geometry and a near $4\pi$ coverage
-in solid angle.%
+The ATLAS detector~\cite{atlas2008experiment} is a multipurpose particle
+detector with a forward--backward symmetric cylindrical geometry and a near
+$4\pi$ coverage in solid angle.%
 \footnote{%
@@
 although extremely energetic muons will have straight tracks,
-\atlas' design and calibration achieves a high precision that increases to
-only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse momenta~\cite{
+\atlas' construction and calibration achieves a high precision, which
+increases to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
+momenta~\cite{
 atlas1994proposal,
@@
 \label{sec:atlas_trigger}
-\TODO{L1 HLT (L2, event thing)}
-\TODO{trigger~\cite{atlas2016trigger}}
-% 25ns bunch spacing
-\TODO{
-With $25\,\mathrm{ns}$ bunch spacing and \atlas' $12.5\,\textrm{m}$ radius,
+The LHC's $25\,\mathrm{ns}$ bunch spacing is so fast that
 light-speed particles ejected from one crossing are still zipping through
-\atlas\ when the next collision arrives.
-}
-% fast decisions to not overflow buffer queue
+\atlas' $12.5\,\textrm{m}$ radius when the next collision arrives.
+This extreme engineering is necessary to achieve the luminosities we
+need to see rare effects, but to avoid becoming overwhelmed
+by data it also demands a strict data reduction mechanism.
+This mechanism is the trigger:
 \begin{quote}
@@
 \end{quote}
+The first-level (L1) trigger uses its fast preliminary information to accept
+interesting event and mark their regions of interest.
+Complete data in these regions of interest are passed to the second-level
+(high-level trigger, HLT) to inform its decision of whether to accept each
+event.
+Events accepted by the HLT are recorded for future analysis~\cite{
+atlas2016trigger,
+atlas2008experiment
+}.
 

### 2022-09-22 main/chapter_context.tex
begin some notes on data subsection

@@
 but to learn more about nature.
-The discovery no tangible new particles in LHC conditions is itself a
-discovery.
+Finding no tangible new particles in LHC conditions is itself a discovery.
 
@@
 Atlas is a Titan.
-On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
-experiment, and alongside its biggest neighbours
+\atlas\ is the largest detector experiment at the LHC,
+and alongside its biggest neighbours
 \cms~\cite{cms2008experiment},
@@
 \label{sec:atlas_data}
+
+% process into event variables
+% clean for conditions?
+% describe LHC Run 1 and Run 2
+
 \TODO{how do we predict data?}
@@
 
-\TODO{discuss collisions, pileup, met}
+\TODO{
+data quality; ignore defects, maximize the uniformity of detector
+operations~\cite{
+DAPR-2018-01,
+Golling:2011zy
+}
+}
 
-\TODO{LHC runs}
+\TODO{performance and calibration?}
+
+\TODO{LHC runs; run 1, run 2}

### 2022-09-22 main/chapter_context.tex
write some on data

@@
 Towards explaining the main result of this thesis,
-a search for effects from supersymmetric models in
-\atlas~\cite{atlas2022searches},
-we first some relevant facts about the \atlas\ detector ---
-its design and history.
+a search for effects from supersymmetric models
+in data from the \atlas\ detector
+on the Large Hadron Collider (LHC)~\cite{
+atlas2022searches,
+atlas2008experiment,
+lhc2008machine
+},
+we first some relevant facts about the design and motivations
+behind \atlas.
 
@@
 \atlas' $12.5\,\textrm{m}$ radius when the next collision arrives.
-This extreme engineering is necessary to achieve the luminosities we
+This extreme engineering is necessary to achieve the integrated luminosities we
 need to see rare effects, but to avoid becoming overwhelmed
@@
 \label{sec:atlas_data}
-
-% process into event variables
-% clean for conditions?
-% describe LHC Run 1 and Run 2
-
-\TODO{how do we predict data?}
-\TODO{software and data}
+Raw digital data are hard to interpret.
+Before use in publishable analysis projects, software systems first mutate
+those data into a workable form, of clean events with calibrated and
+interpretable event variables.
 \begin{quote}
@@
 \end{quote}
+Prediction is important in data analysis.
+This software suite helps to generates predictive simulations, and provides
+realism by passing both the real and simulated data through the same processing
+pipelines~\cite{SOFT-2010-01, geant2003}.
+Software models of the \atlas\ detector itself are refined with the results from
+many calibration studies and in-situ measurements~\cite{atlas2008experiment},
 
-\TODO{
-data quality; ignore defects, maximize the uniformity of detector
-operations~\cite{
+The search presented in this thesis uses data from proton-proton collisions at
+$\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
+LHC, in which \atlas\ collected data in the years
+$2015\textrm{--}2018$~\cite{atlas2022searches}.
+This second run has various refinements over the Run~1, including beam energies
+increasing from $\sqrt{s} = 7\textrm{ or }8\,\eV[T]$, and bunch spacings
+halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.
+
+Data collected under anomalous detector operations are hard to model.
+Lists are maintained indicating which datasets arise from `good', normal
+detector operations, and these lists are commonly employed to select only clean
+subsets of good data for analysis projects.
+Testament to the quality of \atlas' operations, this cleaning removes less
+than $5\%$ of the Run~2 data~\cite{
 DAPR-2018-01,
 Golling:2011zy
-}
-}
+}.
 
-\TODO{performance and calibration?}
+\TODO{closing comments?}
 
-\TODO{LHC runs; run 1, run 2}
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-09-22 main/chapter_searches.tex
fix exp +- error oops!
lester comments (THANK YOU); start cutting

@@
 Statistical data analysis suffers from dogma --- prescriptions and recipes that
-one memorizes and follows while knowing neither their reasons nor their
+one memorizes and follows without necessarily knowing their reasons or their
 limitations:
@@
 \end{equation}
-in which $m(x)$ is a measure (a non-negative quantity assigned to its
-argument, such as mass of tea in a cup), and the comma `,' means `and'
-(logical and or set intersection or any other equivalent operation).
+in which $m(x)$ is a measure, and the comma `,' means `and'
+(logical and or set intersection or anything equivalent).
+
 For example, $\prob{b}{a}$ could quantify the proportion of tea within
@@
 submerged in the cup.
-
-Note that, in this definition, all probabilities are conditional.
+All probabilities are conditional in this definition;
 Alone, no $\probc{b}$ exists unless some conditioning information is implicit
@@
 },
-although random processes are another useful applications.
-
-Two algebraic rules derive from Equation~\ref{eqn:searches_prob_ratio}:
-a \textbf{product rule}
+although random processes are a useful application.
+
+Standard probability algebra derives from
+Equation~\ref{eqn:searches_prob_ratio}, and comprises the \textbf{product rule}
 \begin{align}
@@
 \intertext{%
-describing nested proportions
+for nested proportions
 (a spoon in a teacup in a sink),
@@
 \end{align}
-describing combination in a union;
-the subtraction avoids double-counting of any overlap and is usually avoided by
-splitting any problem into disjoint, orthogonal states.
-For disjoint cases with labels $b_i$, normalization requires that
-\begin{equation}
-\label{eqn:searches_normalization}
-\sum_i \prob{b_i}{a} = 1.
-\end{equation}
-These basic rules are well known, but we state them here to support the
-following statements on how probabilities can applied for practical results.%
-\footnote{%
-The formulation with sums and products is not strictly unique, but all
-alternatives are equivalent up to an invertible
-transformation~\cite{axioms1010038}.
-This choice is often best because it uses ordinary addition and
-multiplication, but $\log$ probabilities can avoid overflows outside the ranges
-representable by floating-point numbers in numerical analysis,
-thus changing multiplication to addition and addition to the
-`$\log$-add-$\exp$' operation.%
-}
+for combinations in a union, with sums over exhaustive sets outcomes
+necessarily normalized to unity.
 
@@
 The $\mathrm{Poisson}(0\mid x)$ likelihood and exponential $e^{-x}\mathrm{d}x$
-prior have the same form (left), but the prior density changes to preserve its unit
-area relative to $u = \log x$.
-On changing to $\log$ coordinates (right), the likelihood slides horizontally,
-but the prior density scales to preserve area.
-Both maximum likelihood points are equivalent, but the maximum prior density
-moves.
-\\[.5em]
-Particle lifetimes follow exponential distributions and their maximum
-densities are at zero time, but it is incorrect to describe their most probable
-lifetimes as zero ---
-the `most probable' $\log$ lifetime is not $\log 0$, but turns out to coincide
-with the mean lifetime.
-\\[.5em]
-For zero Poisson data, the maximum likelihood rate is zero, but zero might not
-be the most probable rate (if, for example, an experiment has known backgrounds).
+prior have the same form (left).
+While the prior density changes to preserve its unit area relative to
+$u = \log x$,
+the likelihood slides horizontally (right).
+Both likelihood maxima are equivalent, but the maximum prior densities change.
 }
@@
 To connect event variables to bin indices, we begin with some notation:
-we have hypotheses $h$ (signal or background),
+we have an hypothesis label $h$ (signal or background),
 an event variable $x$ that we histogram ($x = \met$),
@@
 calculating another likelihood over hypotheses $h$ alone in the bin,
-independent of, by averaging the likelihood over the prior.
+which is independent of $x$, by averaging the likelihood over the prior.
 This abstracted, $x$-independent likelihood is labelled the \textbf{evidence}:
@@
 the likelihood must contain all relevant information from the data.
+The same is true for the evidence for its reuse as a likelihood elsewhere.
 To summarize data for scientific communications, we therefore find a
@@
 
-The infamy of Bayes' rule stems from a fixation on the posterior distributions
-from unjustified prior assignments.
+Many objections to Bayes' rule stem from a fixation on posterior
+distributions from unjustified prior assignments.
 Classical works by
@@
 \lim_{N \to \infty}\!
-\left(1 - \frac{x}{N}\right)^N
+\left(1 + \frac{x}{N}\right)^N
 .
@@
 \label{eqn:searches_binomial}
-\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
+\mathrm{Binomial}(n\mid N, p) = \binom{N}{n} p^n (1 - p)^{N - n}
 ,
@@
 We therefore stick with `frequentist'.
+Sections~\ref{sec:searches_fisher} and~\ref{sec:searches_np} now introduce
+these frequentist methods to explain the historical context for our modern
+methods, which are introduced in Section~\ref{sec:searches_practice}.
 
@@
 \label{eqn:searches_pvalue_likelihood}
-;
 \end{align}
@@
 Equation~\ref{eqn:searches_pvalue_likelihood} points out that the \pvalue\ is
-just is the likelihood plus some fluff from a tail of larger test statistics.
+just is the likelihood plus some fluff from a tail of larger test statistics
+(if the test statistic distribution is discrete).
 
@@
 usefully communicate relevant information from the data.
-And as test statistics, they also win generality by avoid context-specific
+And as test statistics, they also win generality by avoiding context-specific
 motivations, such as scale invariance for the $t$-statistic of
@@
 }
-in fields where they are not yet banned~\cite{
+in fields where they are not banned~\cite{
 Trafimow2015ban,
@@
 Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
+Data on top are ordered with $n$ from left to right, and data on the bottom are
+reordered to have increasing values of the test statistic $-\Lambda(n)$.
 }
@@
 either side.
-Ultimately, the reader just distrust the reporter ---
-the statements `\atlas\ reports $n \leq 5$' and `logically, $n \leq 5$' are not
-identical, since one might infer from the former that in fact $n = 5$.
+Ultimately, the reader may distrust the reporter ---
+the statements
+`logically, $n \leq 5$' and
+`\atlas\ reports $n \leq 5$' and  are not
+identical, since one might infer from the latter that in fact the original
+data were $n = 5$ before their reduction into a test statistic tail.
 The message itself is data, and what it says may not be the whole truth.
@@
 
-\subsection{Like-lihood modelling}
+\subsection{Likelihood modelling}
 To represent the Standard Model and alternatives, LHC data analyses use

### 2022-09-22 main/draft.tex
write some awkward text for heavy sm, start adding citation
write on calo
lester comments (THANK YOU); start cutting

@@
 \lhcb\ and ALICE respectively.
+
+(contributing to the recent storm around a conflicting
+result~\cite{cdf2022high})
+
+, notably with a special use of
+tungsten absorbers in the forward hadronic calorimeters.
+
+% prob cutting
+
+Two algebraic rules derive from Equation~\ref{eqn:searches_prob_ratio}:
+a \textbf{product rule}
+\begin{align}
+\prob{c, b}{a} &= \prob{c}{b, a}\times \prob{b}{a},
+\label{eqn:searches_product_rule}
+\intertext{%
+describing nested proportions
+(a spoon in a teacup in a sink),
+and a \textbf{sum rule}%
+}
+\prob{c\;\textrm{or}\;b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
+\label{eqn:searches_sum_rule}
+\end{align}
+describing combination in a union;
+the subtraction avoids double-counting of any overlap and is usually avoided by
+splitting any problem into disjoint, orthogonal states.
+For disjoint cases with labels $b_i$, normalization requires that
+\begin{equation}
+\label{eqn:searches_normalization}
+\sum_i \prob{b_i}{a} = 1.
+\end{equation}
+These basic rules are well known, but we state them here to support the
+following statements on how probabilities can applied for practical results.%
+\footnote{%
+The formulation with sums and products is not strictly unique, but all
+alternatives are equivalent up to an invertible
+transformation~\cite{axioms1010038}.
+This choice is often best because it uses ordinary addition and
+multiplication, but $\log$ probabilities can avoid overflows outside the ranges
+representable by floating-point numbers in numerical analysis,
+thus changing multiplication to addition and addition to the
+`$\log$-add-$\exp$' operation.%
+}
+
+\\[.5em]
+Particle lifetimes follow exponential distributions, and their maximum
+densities are at zero time, but it is incorrect to describe their most probable
+lifetimes as zero.
+the `most probable' $\log$ lifetime is not $\log 0$, but turns out to coincide
+with the mean lifetime.
+\\[.5em]
+For zero Poisson data, the maximum likelihood rate is zero, but zero might not
+be the most probable rate (if, for example, an experiment has known backgrounds).

### 2022-09-22 main/chapter_searches.tex
more cuts and tweaks to searches section

@@
 
-To clarify our later description of the \atlas\ SUSY $\twoljets$ search in
+To support our later description of the \atlas\ SUSY $\twoljets$ search in
 Chapter~\ref{chapter:2ljets}, this chapter introduces the basic theory and
@@
 Probability assignments are generally constrained only by normalization, but
-might be informed by calibration data or loosely motivated opinions,
+might be informed by calibration data, loosely motivated opinions,
 or by established theories such as quantum mechanics.
-But there are examples of unambiguously derived that are commonly used in
-High Energy Physics, notably the Poisson and Gaussian distributions.
-This section sketches their derivations, both of which use the identity that
+But there are examples of unambiguously derived distributions that are commonly
+used in High Energy Physics, notably the Poisson and Gaussian distributions,
+both of which can derive from the identity that
 \begin{equation}
@@
 \end{equation}
-
-The Large Hadron Collider collides proton bunches at
-$40\,\mathrm{GHz}$,
-but \atlas' trigger system accepts only around
-$1\,\mathrm{kHz}$~\cite{atlas2020trigger}:
-our scenario has a fixed large number $N$ of trials
-(bunch collisions)
+As discussed in Section~\ref{sec:atlas_trigger}, the trigger rejects most
+LHC collision events;
+our data have fixed large number $N$ of trials (bunch crossings)
 and a small probability of acceptance (triggering).
-The summed total number $n$ of accepted events is therefore described by the
-binomial distribution for $n$:
+The number of accepted events  $n$ is therefore described by the
+binomial distribution:
 \begin{equation}
 \label{eqn:searches_binomial}
-\mathrm{Binomial}(n\mid N, p) = \binom{N}{n} p^n (1 - p)^{N - n}
+\mathrm{Binomial}(n\mid N, p) = \binom{N}{n}\, p^n (1 - p)^{N - n}
 ,
@@
 accept/reject decisions, summed over all possible permutations (sum rule) of
-that sequence for $n$ accepted.
+the accepted events.
+
 With the tiny trigger rate, we are interested in the limit of small
-$p$ and large $N$, which result in a mean acceptance number of $\lambda = Np$.
+$p$ and large $N$, which result in a mean acceptance of $\lambda = Np$.
 Using the identity of Equation~\ref{eqn:searches_exponential},
@@
 \end{equation}
-which is used ubiquitously to interpret histograms.
+which we use ubiquitously to interpret histograms.
 
@@
 Gaussian likelihoods justify many least-squares methods.
-
 Here, $\kappa$ and $\sigma$ both relate to the second derivative of the target
@@
 
-This is not the famous Central Limit Theorem, which is related to the
-distribution of sums of independent quantities with finite mean and variance.
-But can be used to derive it: the distribution of a sum is a convolution, which
-becomes a product in Fourier-transformed space, and that product converges
-to a Gaussian function by this same derivation if the mean and variance exist.
+This is not the Central Limit Theorem, but can be used to derive it: the
+distribution of a sum is a convolution, which becomes a product in
+Fourier-transformed space, and that product converges to a Gaussian function by
+if the mean and variance exist.
 Another beautiful alternative proof of the Central Limit Theorem is given by
@@
 constructed an analysis that can only ever favour one hypothesis:
-we have built a propaganda machine that only ever speaks truth but censors any
-information that supports alternative ideas.
+we have built a propaganda machine, which only ever speaks truth,
+but which censors any information that supports alternative ideas.
 Reducing data to inequalities $t \geq \tilde{t}$ implements the censorship,
@@
 Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
-Data on top are ordered with $n$ from left to right, and data on the bottom are
-reordered to have increasing values of the test statistic $-\Lambda(n)$.
+Data on top are ordered with $n$ increasing from left to right,
+and on the bottom those same data are reordered to have increasing values of
+the test statistic $-\Lambda(n)$.
 }

### 2022-09-23 main/chapter_searches.tex
minor proofread

@@
 Standard probability algebra derives from
-Equation~\ref{eqn:searches_prob_ratio}, and comprises the \textbf{product rule}
+Equation~\ref{eqn:searches_prob_ratio}; it comprises a \textbf{product rule}
 \begin{align}
@@
 frodesen1979probability
-}.%s
+}.%
 }%
@@
 pdg2022ynf
-}
-where $p_b$ might not be a \pvalue\ if it excludes equality with the
+},
+in which $p_b$ might not be a \pvalue\ if it excludes equality with the
 observed test statistic~\cite{cern2011procedure}.
@@
 either side.
-Ultimately, the reader may distrust the reporter ---
-the statements
+Ultimately, a reader may choose to distrust the reporter ---
+the messages
 `logically, $n \leq 5$' and
@@
 identical, since one might infer from the latter that in fact the original
-data were $n = 5$ before their reduction into a test statistic tail.
+data were $n = 5$, before their reduction into a test statistic tail.
 The message itself is data, and what it says may not be the whole truth.

### 2022-09-23 main/chapter_searches.tex
overleaf autocheck

@@
 \end{align}
-for combinations in a union, with sums over exhaustive sets outcomes
+for combinations in a union, with sums over exhaustive sets of outcomes
 necessarily normalized to unity.
@@
 \end{equation}
-where we use the double bar, $\mid\mid$, to separate to any common contextual
+where we use the double bar, $\mid\mid$, to separate any common contextual
 information.
@@
 But we should proudly proclaim that our results are `biased' under this flawed
-statistical definition if that makes them better by any reasonable measure.
+definition if that makes them better by any reasonable measure.
 
@@
 Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
-Data on top are ordered with $n$ increasing from left to right,
+Data on the top are ordered with $n$ increasing from left to right,
 and on the bottom those same data are reordered to have increasing values of

### 2022-09-23 main/chapter_context.tex
add some cringy closing comment
detector autocheck overleaf

@@
 Designs that can discover such new particles are therefore neatly aligned with
-the other goals to refining the Standard Model
+our other goals of refining the current Standard Model
 and to discovering the Higgs boson at any mass.
@@
 Formerly known as the SemiConductor Tracker~\cite{atlas1999design1},
-the silicon microstrip system retains the
+the silicon microstrip system retains the acronym
 SCT~\cite{atlas2008experiment}.%
@@
 range ($|\eta| < 1.7$).
-The endcap and forward regions are instrumented with LAr calorimeters for both
+The end-cap and forward regions are instrumented with LAr calorimeters for both
 the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
@@
 The steel/scintillator hadronic tile calorimeter is a
-``simple and cost effective''~\cite{atlas1996tile} design that uses
+``simple and cost effective''~\cite{atlas1996tile} design, which uses
 photomultiplier tubes to convert light from its polystyrene scintillators into
@@
 Prediction is important in data analysis.
-This software suite helps to generates predictive simulations, and provides
+This software suite helps to generate predictive simulations, and provides
 realism by passing both the real and simulated data through the same processing
@@
 
-\TODO{closing comments?}
+% closing comment?
+Now that we have predictive models, in the Standard Model and supersymmetric
+alternatives,
+and interpretable data, from \atlas' measurements of LHC collisions,
+we are ready to begin to squeeze some useful research from those data.
 

### 2022-09-23 main/chapter_context.tex
reduce footnote footnote futrther; other tweaks from notes

@@
 \footnotetext{%
-This footnote, which appears in every \atlas\ publication, was perfected
-in~\cite{lester2013heffalon}.
-True to form, internal \atlas\ guidelines ask that
-``If you think you can improve on it, please consult with the
-Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
-We added the sentence defining angular distance without such a consultation.%
+This footnote appears in every \atlas\ publication and was perfected
+in~\cite{lester2013heffalon}.%
 }
@@
 \label{sec:atlas_data}
-Raw digital data are hard to interpret.
+Raw binary data are hard to interpret.
 Before use in publishable analysis projects, software systems first mutate

### 2022-09-26 main/chapter_context.tex
add some text to theory chapter

@@
 1980~\cite{hamming1980unreasonable}}
-\qitem{%
-One geometry can not be more true than another; it can only be more convenient.
-Geometry is not true, it is advantageous.}%
-{Robert~M.~Pirsig,
-\textit{Zen and the Art of Motorcycle Maintenance},
-1974~\cite{pirsig1999zen}}
+% \qitem{%
+% One geometry can not be more true than another; it can only be more convenient.
+% Geometry is not true, it is advantageous.}%
+% {Robert~M.~Pirsig,
+% \textit{Zen and the Art of Motorcycle Maintenance},
+% 1974~\cite{pirsig1999zen}}
 \end{epigraphs}
@@
 
+Theory is needed to understand experimental data,
+and data are needed to improve theoretical models.
+For the experimental work of this thesis to be better understood, this chapter
+introduces its theoretical and experimental contexts;
+Section~\ref{sec:context_sm_and_beyond} presents the current culmination of
+theoretical particle physics in the Standard Model, its limitations, and its
+possible future directions including supersymmetric ideas, and
+Section~\ref{sec:context_atlas} then describes the \atlas\ detector, with whose
+data we hope refine the Standard Model into its improved future forms.
+
 \begin{singlespacing}
 \section{The Standard Model and beyond}
+\label{sec:context_sm_and_beyond}
 %
@@
 \qitem{%
-One geometry can not be more true than another; it can only be more convenient.
-Geometry is not true, it is advantageous.}%
-{Robert~M.~Pirsig,
-\textit{Zen and the Art of Motorcycle Maintenance},
-1974~\cite{pirsig1999zen}}
+In fact, it is a defensible conjecture that no
+theory can be complete and inviolable that emerges from finitely equipped minds
+and survives finitely scoped experiment, which are the twin rickety foundations
+for all theories.%
+}%
+{James~D.~Wells,
+\textit{Discovery Beyond the Standard Model of~Elementary~Particle~Physics},
+2019~\cite{wells2020discovery}}
 \end{epigraphs}
@@
 
-\TODO{Standard Model past present future}
+The Standard Model is science's best current understanding of Nature's
+fundamental objects in their observable behaviours as particles.
+It has solid mathematical foundations, and accurately describes the
+physics of particle interactions across a very broad domain.
+But that domain is bounded, so the Standard Model is not an ultimate theory of
+everything and work continues to expand that boundary on both theoretical and
+experimental fronts.
+We describe the current Standard Model, its limitations and possible future
+directions in Section~\ref{sec:context_sm}.
 
-\subsection{Particles}
+Supersymmetry an idea that has been applied to build plausible alternatives
+to the Standard Model, which may explain future experimental discoveries.
+Since the main work of this thesis is a search for effects of supersymmetric
+phenomena, supersymmetry and the specific supersymmetric models that we study
+are presented in Section~\ref{sec:context_susy}.
 
 \subsection{The Standard Model}
+\label{sec:context_sm}
+\TODO{particle content figure/tables}
+
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$
+
+\TODO{Higgs mechanism}
+
+\TODO{fermion structure}
+\TODO{three generations}
+\TODO{rotation between weak and mass eigenstates: CKM matrix}
+
+\TODO{neutrinos}
+\TODO{rotation between weak and mass eigenstates: PMNS matrix}
+
+\subsubsection{Future Standard Models}
+
+\TODO{Limitations: uncertainties and unknowns}
+\TODO{uncertainties: lagrangian form for neutrino masses, mass values (hierarchy, mixing angles); composition of dark matter}
+\TODO{unknowns: lack of convincing solutions; unification of gravity with quantum mechanics, origin of structure and parameters of standard model}
+
+\paragraph{Neutrino masses.}
+\TODO{Dirac neutrinos, possibly supplemented with a Majorana term}
+\TODO{Weinberg operator for Majorana neutrinos}
+
+\paragraph{Gravitation.}
+
+\paragraph{Dark matter.}
+\TODO{
+primordial black holes,
+WIMPs,
+axions,
+dust?
+}
+\TODO{combinations!}
+
+\paragraph{Cosmology.}
+\TODO{dark matter, cosmological constant, inflation, baryogenesis (matter-antimatter asymmetry)}
+\TODO{note \lcdm}
+
+\paragraph{Origins.}
+
+\paragraph{Aesthetics.}
+\TODO{naturalness and fine-tuning}
+\TODO{strong CP problem}
+\TODO{hierarchy problem}
 
 \subsection{Supersymmetric models}
+\label{sec:context_susy}
 
@@
 \section{The \atlas\ Experiment}
+\label{sec:context_atlas}
 %
@@
 },
-These continue to not validate the Standard Model, and are aided by the lack of
-new phenomena, which could have overpowered such rare signals.
+These continue to not invalidate the Standard Model, and are aided in their
+sensitivity by the lack of new phenomena,
+which could have overpowered these rare signals.
 

### 2022-09-26 main/chapter_searches.tex
reduce footnote footnote futrther; other tweaks from notes
add some text to theory chapter

@@
 
-Data only enter the inputs to Bayes' rule through the likelihood function, so
-the likelihood must contain all relevant information from the data.
+Data only enter the \emph{inputs} to Bayes' rule through the likelihood
+function, so the likelihood must contain all relevant information from the
+data.
 The same is true for the evidence for its reuse as a likelihood elsewhere.
@@
 {John~W.~Pratt,
-\textit{Review: Testing Statistical Hypotheses},
+\textit{Review: Testing Statistical Hypotheses
+by E.~L.~Lehmann},
 1961~\cite{pratt1961testing}}
@@
 at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
-by~\cite{pdg1984}) do distinction the two~\cite{
+by~\cite{pdg1984}) do distinguish between the two~\cite{
 eadie1971statistical,
@@
 these early applications used data deficits to exclude their
-background-plus signal models, sometimes with the test statistic $t(n) = -n$,
+background-plus-signal models, sometimes with the test statistic $t(n) = -n$,
 and disliked how deficits led to small \pvalues\ for the background-only model,
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
+\includegraphics[width=0.6\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
 \caption{%
@@
 Despite the likely superiority of $t(n) = n$ test statistics for these
-single-bin data, the standard use of maximum likelihood ratio test statistics
+single-bin data, maximum likelihood ratio test statistics are used as standard,
 and may explain quirks in the discovery region results that we shall see in the

### 2022-09-26 main/draft.tex
add some text to theory chapter

@@
 be the most probable rate (if, for example, an experiment has known backgrounds).
+
+% excess philosophy
+Theory digests data to extract useful information.
+Those data include not only observations from scientific experiments,
+but the humanity's accumulated knowledge that one might glean from books,
+intuitions that one learns by exploring the world,
+and even fragments inherited in biology.
+The nutrients that theoretical work extracts are deeper concepts,
+such as postulates or mathematical models,
+which might reconstruct implications of the data and make useful or novel
+predictions.
+
+The current Standard Model of particle physics is a magnificent results from
+theoretical and experimental works which, over the last century or
+so, collaborated to understand and observe the actions of Nature's most
+fundamental known particles.
+It now comprises a solid mathematical theory that accurately describes the
+physics of particle interactions across a broad domain.
+But it is not an ultimate theory of everything; that domain is bounded, and
+work continues to expand those boundaries.
+
+To support the later work that this thesis presents on experimental data from
+the \atlas\ detector, this chapter introduces contextual information.
+ presents the Standard Model of particle
+physics, both in its current state and possible future directions.
+Section~\ref{sec:context_atlas} then describes the \atlas\ detector.
+\TODO{more}

### 2022-09-26 main/chapter_context.tex
add some citations and begin sm

@@
 
-Theory is needed to understand experimental data,
+\noindent Theory is needed to understand experimental data,
 and data are needed to improve theoretical models.
@@
 
-The Standard Model is science's best current understanding of Nature's
-fundamental objects in their observable behaviours as particles.
-It has solid mathematical foundations, and accurately describes the
-physics of particle interactions across a very broad domain.
+\noindent Our best current understanding of Nature's fundamental objects in
+their observable behaviours as particles is encoded in the Standard Model,
+which uses solid mathematical foundations to accurately describe the physics
+of particle interactions across a broad domain.
 But that domain is bounded, so the Standard Model is not an ultimate theory of
-everything and work continues to expand that boundary on both theoretical and
+everything, and work continues to expand that boundary on both theoretical and
 experimental fronts.
-We describe the current Standard Model, its limitations and possible future
-directions in Section~\ref{sec:context_sm}.
+The current Standard Model is described in Section~\ref{sec:context_sm},
+along with its limitations and possible future directions.
 
 Supersymmetry an idea that has been applied to build plausible alternatives
-to the Standard Model, which may explain future experimental discoveries.
+to the Standard Model which may explain future experimental discoveries.
 Since the main work of this thesis is a search for effects of supersymmetric
@@
 \label{sec:context_sm}
-\TODO{particle content figure/tables}
+% \TODO{particle content figure/tables}
 
-$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$
+The Standard Model is formulated as a quantized gauge field theory with the
+symmetry that its Lagrangian is invariant under local operations of the group
+\begin{equation}
+SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y
+\end{equation}
+on its fields.
+Here, the subscript $C$
+
+~\cite{
+cottingham2007greenwood,
+wells2020discovery,
+rubakov1999classical,
+thomson2013modern
+}
 

### 2022-09-26 main/chapter_2ljets.tex
some more text on SM, minor tweaks

@@
 
-The main messages that this thesis reports are my contributions to the recently
+\noindent The main messages that this thesis reports are my contributions to the recently
 published \atlas\ searches ``for new phenomena in events with two leptons,
@@
 
-From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
+\noindent From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
 have extracted a sharper image of the high energy processes that simultaneously

### 2022-09-26 main/chapter_context.tex
some more text on SM, minor tweaks

@@
 % \TODO{particle content figure/tables}
+The current Standard Model is characterized by three generations of fermions
+(spin-$1/2$) that interact through forces mediated by vector bosons
+(spin-$1$) and a scalar Higgs boson (spin-$0$).
+Force particles are categorized into the strong interaction
+via with massless gluons $g$,
+the electromagnetic interaction via the massless photon $\gamma$,
+and
+the weak interaction via the very massive $W^+$, $W^-$, and $Z$ bosons.
+Electroweak unification combines the weak and electromagnetic interactions
+with the Higgs mechanism, which simultaneously provides the weak boson masses
+and adds massive Higgs boson, whose Yukawa interactions with the fermions also
+implement their masses~\cite{thomson2013modern}
 
-The Standard Model is formulated as a quantized gauge field theory with the
-symmetry that its Lagrangian is invariant under local operations of the group
-\begin{equation}
-SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y
-\end{equation}
-on its fields.
-Here, the subscript $C$
+Mathematically, the Standard Model is described by a quantized gauge field
+theory with the symmetry group
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
+meaning that its Lagrangian is invariant under local operations of group
+elements on its fields in certain representations.
 
-~\cite{
-cottingham2007greenwood,
-wells2020discovery,
+Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
+indicates that they operate on colour charge triplets.
+
+Unified electroweak interactions come from $SU\!(2)_L$ and $U\!(1)_Y$, whose
+four bosons $W^1$, $W^2$, $W^3$, and $B$ respectively would be massless
+without the Higgs mechanism, which inserts four scalar fields and
+`spontaneously breaks' the gauge symmetry with a potential that moves the
+ground state of the model to a gauge-asymmetric configuration.
+From this new ground state, linear combinations of the electroweak fields
+acquire masses while ``eating up'' three of the four scalar fields to yield
+the three weak bosons and the photon:
+$(W^1, W^2)$ mix into $(W^+, W^-)$, and $(W^3, B)$ mix into $(Z, \gamma)$.
+The fourth and final scalar field is the Higgs boson $h$~\cite{
 rubakov1999classical,
-thomson2013modern
-}
+cottingham2007greenwood
+}.
 
-\TODO{Higgs mechanism}
+The fermions of the Standard Model are listed in Table~\TODO{}.
+Older Standard Models included only two fermion generations and no
+neutrino masses~\cite{
+wells2020discovery
+},\TODO{cite more}
+but a modern model must include both three generations and non-zero neutrino
+masses to be compatible with experiment:
+three fermion generations are directly established
+\TODO{kaon mixing CPV, direct observation of tau/top/bottom, Z width},
+as are neutrino masses from their flavour oscillations demonstrating that
+they experience proper time~\TODO{cite}.
 
@@
 \TODO{neutrinos}
-\TODO{rotation between weak and mass eigenstates: PMNS matrix}
+\TODO{rotation between weak and mass eigenstates: PMNS matrix!!}
 
-\subsubsection{Future Standard Models}
+\subsubsection{Opportunities for future Standard Models}
 
@@
 
-Towards explaining the main result of this thesis,
+\noindent Towards explaining the main result of this thesis,
 a search for effects from supersymmetric models

### 2022-09-26 main/chapter_searches.tex
some more text on SM, minor tweaks

@@
 
-Our task in scientific data analysis is to distil useful messages from the
+\noindent Our task in scientific data analysis is to distil useful messages from the
 data, and to report that message to the wider community.
@@
 
-Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
+\noindent Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
 proportion of stuff within a larger whole, which when formalized as a ratio
@@
 
-Data analysis in modern High Energy Physics does not exactly follow frequentist
+\noindent Data analysis in modern High Energy Physics does not exactly follow frequentist
 ideology, but is a hands-on, practical, and evolving ecosystem that has adapted

### 2022-09-26 main/chapter_context.tex
chris detector chapter responses

@@
 real-world cost budgets.
-Within general purpose utilities, \atlas' designs aimed to construct a
+Within general purpose utilities, \atlas's designs aimed to construct a
 detector that could
@@
 \end{itemize}
-We now review these three goals with the subsequent news from \atlas'
+We now review these three goals with the subsequent news from \atlas's
 operations.
@@
 \paragraph{Find the Higgs boson.}
-Prior to \atlas' construction, the Higgs boson mass (and existence) was
+Prior to \atlas's construction, the Higgs boson mass (and existence) was
 poorly known, but loosely constrained by results like
@@
 The Higgs boson couples to an extraordinary variety of fundamental particles.
-For its discovery alone, \atlas' designers achieved sensitivity across the
+For its discovery alone, \atlas's designers achieved sensitivity across the
 entire $100$ to $1000\,\eV[G]$ window by targetting decays including
@@
 }
-it makes sense for \atlas' design to forfeit hadron identification systems
+it makes sense for \atlas's design to forfeit hadron identification systems
 in favour of precise calorimetry with total angular coverage.
@@
 and to discovering the Higgs boson at any mass.
-Visible decay products should already be caught by \atlas' abilities
+Visible decay products should already be caught by \atlas's abilities
 to measure the Higgs boson and other Standard Model particles,
@@
 \caption[
-A central slice of \atlas' design
+A central slice of \atlas's design
 ]{%
-A central slice of \atlas' design, reproduced from~\cite{atlas1999design1},
+A central slice of \atlas's design, reproduced from~\cite{atlas1999design1},
 including a simulated event that interacts with several sensitive subsystems.
@@
 This footnote appears in every \atlas\ publication and was perfected
-in~\cite{lester2013heffalon}.%
+in~\cite{lester2013heffalon}.
+True to form, internal \atlas\ guidelines ask that
+``If you think you can improve on it, please consult with the
+Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
+We added the sentence defining angular distance without such a consultation.%
 }
@@
 \label{sec:atlas_inner}
-The inner tracking detectors begin \atlas' measurements of new collisions
+The inner tracking detectors begin \atlas's measurements of new collisions
 by precisely tracing the paths taken by energetic charged particles
@@
 decay inside the tracking system;
-These `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
+these `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
 maybe new species.
@@
 range minimizes leakage.
-Towards \atlas' design goals, this inner tracking detector crucially provides
+Towards \atlas's design goals, this inner tracking detector crucially provides
 measurements of electrons, B-hadrons, and the missing transverse momentum.
@@
 instead measured by their curved tracks in a toroidal magnetic field that
-envelopes dedicated detectors on \atlas' periphery.
+envelopes dedicated detectors on \atlas's periphery.
 \begin{quote}
@@
 \caption[
-Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+Layout of \atlas's toroidal magnets around the tile calorimeter and inner
 solenoid
 ]{%
-Layout of \atlas' toroidal magnets around the tile calorimeter and inner
+Layout of \atlas's toroidal magnets around the tile calorimeter and inner
 solenoid, including the eight barrel and end-cap coils.
@@
 although extremely energetic muons will have straight tracks,
-\atlas' construction and calibration achieves a high precision, which
+\atlas's construction and calibration achieves a high precision, which
 increases to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
@@
 light-speed particles ejected from one crossing are still zipping through
-\atlas' $12.5\,\textrm{m}$ radius when the next collision arrives.
+\atlas's $12.5\,\textrm{m}$ radius when the next collision arrives.
 This extreme engineering is necessary to achieve the integrated luminosities we
@@
 \label{sec:atlas_data}
-Raw binary data are hard to interpret.
+Raw binary data are hard for us humans to interpret.
 Before use in publishable analysis projects, software systems first mutate
@@
 subsets of good data for analysis projects.
-Testament to the quality of \atlas' operations, this cleaning removes less
+Testament to the quality of \atlas's operations, this cleaning removes less
 than $5\%$ of the Run~2 data~\cite{
@@
 alternatives,
-and interpretable data, from \atlas' measurements of LHC collisions,
+and interpretable data, from \atlas's measurements of LHC collisions,
 we are ready to begin to squeeze some useful research from those data.

### 2022-09-26 main/chapter_context.tex
Holly awesome comment changes

@@
 },
-we first some relevant facts about the design and motivations
+we first cover some relevant facts about the design and motivations
 behind \atlas.
@@
 these channels demand precise measurements of leptons, photons, and
-$b$-jets, as well as sensitivity to the missing transverse momentum of
-invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
-masses.
+$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
+transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
+at the very highest masses.
 
@@
 Track curvature provides momentum information,
-tracks' associations into collision vertices helps to separate
-pile-up,
+tracks' associations into collision vertices helps to separate the multiple
+scatterings per crossing (pile-up),
 and their associations into displaced vertices help to identify
@@
 Towards \atlas's design goals, this inner tracking detector crucially provides
-measurements of electrons, B-hadrons, and the missing transverse momentum.
+measurements of electrons, B-hadrons, and contributes to the reconstruction of
+missing transverse momentum.
 
@@
 \label{sec:atlas_trigger}
-The LHC's $25\,\mathrm{ns}$ bunch spacing is so fast that
-light-speed particles ejected from one crossing are still zipping through
-\atlas's $12.5\,\textrm{m}$ radius when the next collision arrives.
+The LHC's $25\,\mathrm{ns}$ time difference between proton bunches
+(bunch spacing) is so fast that light-speed particles ejected from one crossing
+are still zipping through \atlas's $12.5\,\textrm{m}$ radius when the next
+collision arrives.
 This extreme engineering is necessary to achieve the integrated luminosities we

### 2022-09-26 main/chapter_searches.tex
Holly awesome comment changes

@@
 approaches that we describe in Section~\ref{sec:searches_practice}.
-Searches for Supersymmetric signals in \atlas\ follow a standard strategy for
+Searches for supersymmetric signals in \atlas\ follow a standard strategy for
 their data analysis, whose procedures, nomenclature, and parametric modelling
@@
 
-We begin with what probability is: at its most abstract, a probability a
+We begin with what probability is: at its most abstract, a probability is a
 proportion of stuff within a larger whole.
@@
 across states $b$.
-They and need not be random distributions of fluctuating $b$s~\cite{
+They need not be random distributions of fluctuating $b$s~\cite{
 jaynes2003probability,
@@
 Priors on hypotheses express uncertain predictions, and
-priors on data are useful in experimental design
+priors on data are useful in experimental design.
 
@@
 \quad \quad \mid\mid h
-.
+,
 % \textbf{likelihood} \times \textbf{prior}
@@
 \end{align}
+where we use the double bar, $\mid\mid$, to separate common contextual
+information.
 This is (the infamous) Bayes' rule, derived as a consequence of the
@@
 \quad \quad \mid\mid \mathrm{context}
-,
+.
 \end{equation}
-where we use the double bar, $\mid\mid$, to separate any common contextual
-information.
 When the evidence ratio is independent of that context, as it usually is,
@@
 Finally, their corresponding posteriors are the normalized shapes those samples
-take within the observed histogram bin, with which we do not compare the
-models, but which could in principle be used for secondary investigations.
+take within the observed histogram bin
+(as if it were more finely re-binned after analysis),
+with which we do not compare the models,
+but which could in principle be used for secondary investigations.
 
@@
 where $n$ is the number observed to satisfy conditions $a$ and $b$ when
-$N$ have satisfied $b$.
+$N$ have satisfied $a$.
 Interpreted with the Binomial model of
@@
 \\
-&= L(h) + \prob{t > \tilde{t}}{h}
+&= L(h) + \prob{t > \tilde{t}}{h}.
 \label{eqn:searches_pvalue_likelihood}
 \end{align}
-the \pvalue\ $P(h)$ is the prior probability of data
+The \pvalue\ $P(h)$ is the prior probability of data
 \emph{greater or equal to} the observed data, if data are ordered by the value
@@
 \pvalues\ can be painfully expensive to evaluate.
-Furthermore, any result with small a test statistic is actively censored,
+Furthermore, any result with a small test statistic is actively censored,
 since if
@@
 Searches for new effects from supersymmetric models in \atlas\ face the problem
-that their signals can appear in obscure, extreme, or otherwise less-studied
+that their signals often appear in obscure, extreme, or otherwise less-studied
 corners of the data.
@@
 After fitting the background-only model to observed signal region data,
-it can be useful to check that its parameters have been `pulled' away from
+it can be useful to check whether its parameters have been `pulled' away from
 their central values into regimes where their approximated extrapolations may
-be invalid
+be invalid.
 Post-fit expectations in signal regions are also available with error bars;

### 2022-09-26 main/chapter_context.tex
add a bunch of citations for standard model

@@
 theory with the symmetry group
-$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
+$SU\!(3)_C \!\times\! SU\!(2)_L \!\times\! U\!(1)_Y$,
 meaning that its Lagrangian is invariant under local operations of group
@@
 The fermions of the Standard Model are listed in Table~\TODO{}.
+% Glashow-Salam-Weinberg
 Older Standard Models included only two fermion generations and no
@@
 three fermion generations are directly established
-\TODO{kaon mixing CPV, direct observation of tau/top/bottom, Z width},
+\TODO{kaon mixing CPV~\cite{
+cronin1964evidence,
+kabayasji1973cpv
+}, direct observation of tau/top/bottom~\cite{
+perl1977evidence,
+herb1977observation,
+abachi1995observation
+}, Z width~\cite{
+lep2006precision
+}
+},
 as are neutrino masses from their flavour oscillations demonstrating that
-they experience proper time~\TODO{cite}.
+they experience proper time~\cite{
+kamiokande1998measurement,
+superk1998evidence,
+superk1999measurement,
+lsnd1998evidence
+}.
 
@@
 \TODO{three generations}
-\TODO{rotation between weak and mass eigenstates: CKM matrix}
+\TODO{rotation between weak and mass eigenstates: CKM matrix~\cite{
+cabibbo1963unitary,
+kabayasji1973cpv
+}}
 
@@
 
-\paragraph{Gravitation.}
+\TODO{mass models: dirac, Majorana, seesaw~\cite{
+wells2020discovery,
+king2003neutrino,
+thomson2013modern
+}}
 
-\paragraph{Dark matter.}
+\paragraph{Gravitation and cosmology.}
+
+\TODO{dark matter: rotation curves~\cite{
+begeman1991rotation,
+garrett2010dark
+}}
+
 \TODO{
-primordial black holes,
-WIMPs,
-axions,
-dust?
+primordial black holes~\cite{bernard2019primordial},
+WIMPs~\cite{jungman1996wimp},
+axions~\cite{kim2008axions},
+dust?~\cite{garrett2010dark}
 }
@@
 
-\paragraph{Cosmology.}
-\TODO{dark matter, cosmological constant, inflation, baryogenesis (matter-antimatter asymmetry)}
+\TODO{
+dark matter,
+cosmological constant~\cite{riess1998observational},
+inflation~\cite{bezrukov2008higgs},
+baryogenesis (matter-antimatter asymmetry)
+}
 \TODO{note \lcdm}
@@
 \paragraph{Aesthetics.}
-\TODO{naturalness and fine-tuning}
+
+\TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
 \TODO{strong CP problem}
@@
 \label{sec:context_susy}
+
+\TODO{higgs mass $\lesssim 130\,\eV[G]$~\cite{
+quiros1997constraints,
+wells2020discovery
+}}
+
+\TODO{primer~\cite{
+martin2016primer
+}}
 

### 2022-09-27 main/chapter_context.tex
add glashow-salam-weinberg citations
add bjorken november citation

@@
 theory with the symmetry group
-$SU\!(3)_C \!\times\! SU\!(2)_L \!\times\! U\!(1)_Y$,
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
 meaning that its Lagrangian is invariant under local operations of group
@@
 
+\TODO{
+Glashow-Salam-Weinberg~\cite{
+glashow1959renorm,
+weinberg1967model,
+salam1959weak
+}
+}
+
+\TODO{charm november revolution~\cite{bjorken1985november}}
+
 The fermions of the Standard Model are listed in Table~\TODO{}.
-% Glashow-Salam-Weinberg
 Older Standard Models included only two fermion generations and no

### 2022-09-27 main/chapter_context.tex
searches tweaks; some breaking hacks to SM

@@
 % \TODO{particle content figure/tables}
-The current Standard Model is characterized by three generations of fermions
-(spin-$1/2$) that interact through forces mediated by vector bosons
-(spin-$1$) and a scalar Higgs boson (spin-$0$).
-Force particles are categorized into the strong interaction
-via with massless gluons $g$,
-the electromagnetic interaction via the massless photon $\gamma$,
-and
-the weak interaction via the very massive $W^+$, $W^-$, and $Z$ bosons.
-Electroweak unification combines the weak and electromagnetic interactions
+The current Standard Model has three generations of fermions
+that interact through forces mediated by vector bosons
+and the scalar Higgs boson.
+The strong force is mediated by
+Massless gluons $g$ mediate the strong force,
+the massless photon $\gamma$ mediates the electromagnetic force,
+and very massive $W^+$, $W^-$, and $Z$ bosons weadiate the weak force.
+\TODO{}
+
+the weak and electromagnetic interactions
 with the Higgs mechanism, which simultaneously provides the weak boson masses
@@
 }.
+
+\TODO{(Higgs) mechansim~\cite{
+higgs1964broken,
+englert1964broken
+}}
+
+\TODO{ABEGHHK'tH~\cite{close2011infinity}}
 
@@
 \footnotetext{%
-This footnote appears in every \atlas\ publication and was perfected
+This footnote appears in every \atlas\ publication, but was perfected
 in~\cite{lester2013heffalon}.
@@
 Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
-We added the sentence defining angular distance without such a consultation.%
 }
@@
 Muon subsystems come in four varieties, all of which operate by detecting
-charge scattered by the muons as they fly through gaseous media.
-These systems track the muons' paths as the magnets bend them in the
+charges scattered by the muons as they fly through gaseous media.
+Together, these systems track the muons' paths as the magnets bend them in the
 longitudinal plane; these toroidal magnets generate fields outside the tile
@@
 Layout of \atlas's toroidal magnets around the tile calorimeter and inner
-solenoid, including the eight barrel and end-cap coils.
-Figure reproduced from~\cite{atlas2008experiment}.
+solenoid, including the eight barrel and end-cap coils,
+reproduced from~\cite{atlas2008experiment}.
 The \atlas\ muon spectrometer measures muons' tracks as they bend through

### 2022-09-27 main/chapter_searches.tex
searches tweaks; some breaking hacks to SM

@@
 and on the bottom those same data are reordered to have increasing values of
-the test statistic $-\Lambda(n)$.
+test statistic $t(n) = -\Lambda(n)$.
 }
@@
 with which a thermodynamic system collapses to its minimum energy at cold
-temperatures $T = 1/\beta$~\cite{
+temperatures $T \propto 1/\beta$~\cite{
 skilling2017david,
@@
 \footnote{%
-The Poisson likelihoods typically use a continuous interpolation with
-$n!\rightarrow \Gamma(n + 1)$.%
+Poisson likelihoods in these software packages typically use a continuous
+interpolation with $n! \to \Gamma(n + 1)$.%
 }
@@
 \item minimize $E(x)$ to find the best-fit parameters $\check{x}$, then
-\item sample alternative Poisson (actual) data with rates $y(\check{x})$ and
-alternative auxiliary (fictitious) data from their assigned distributions,
-then
+\item at fixed $\check{x}$, sample alternative Poisson (actual) data with rates
+$y(\check{x})$ and alternative auxiliary (fictitious) data from their assigned
+distributions, then
 \item evaluate the test statistic for each simulated `toy'
@@
 Sampling alternative data from fixed $y(\check{x})$ is also suspicious because
-we are in fact uncertain of the expected yields
-(demonstrated by us drawing error bars on them).
+we are in fact uncertain of the expected yields,
+as demonstrated by the error bars we draw on them.
 

### 2022-09-27 main/chapter_context.tex
some improvements to SM; some misc tweaks

@@
 1980~\cite{hamming1980unreasonable}}
-% \qitem{%
-% One geometry can not be more true than another; it can only be more convenient.
-% Geometry is not true, it is advantageous.}%
-% {Robert~M.~Pirsig,
-% \textit{Zen and the Art of Motorcycle Maintenance},
-% 1974~\cite{pirsig1999zen}}
+\qitem{%
+if observing outer space gives us a view of the past, observing inner space
+would surely give us a glimpse into the future - would be interesting if NASA
+made a telescope for that%
+}%
+{Ken~M,
+\textit{Comment: Yahoo! News},
+2012~\cite{kenm2012inner}}
 \end{epigraphs}
@@
 and data are needed to improve theoretical models.
-For the experimental work of this thesis to be better understood, this chapter
-introduces its theoretical and experimental contexts;
+To clarify the experimental work of this thesis, this chapter introduces its
+theoretical and experimental contexts;
 Section~\ref{sec:context_sm_and_beyond} presents the current culmination of
@@
 
-\begin{singlespacing}
 \section{The Standard Model and beyond}
 \label{sec:context_sm_and_beyond}
-%
-\begin{epigraphs}
-\qitem{%
-In fact, it is a defensible conjecture that no
-theory can be complete and inviolable that emerges from finitely equipped minds
-and survives finitely scoped experiment, which are the twin rickety foundations
-for all theories.%
-}%
-{James~D.~Wells,
-\textit{Discovery Beyond the Standard Model of~Elementary~Particle~Physics},
-2019~\cite{wells2020discovery}}
-\end{epigraphs}
-\end{singlespacing}
-
 \noindent Our best current understanding of Nature's fundamental objects in
@@
 \label{sec:context_sm}
-% \TODO{particle content figure/tables}
-The current Standard Model has three generations of fermions
-that interact through forces mediated by vector bosons
+\TODO{particle content figure/tables}
+
+The current Standard Model features three generations of fermions,
+which interact through forces mediated by vector bosons
 and the scalar Higgs boson.
@@
 and very massive $W^+$, $W^-$, and $Z$ bosons weadiate the weak force.
-\TODO{}
 
-the weak and electromagnetic interactions
-with the Higgs mechanism, which simultaneously provides the weak boson masses
-and adds massive Higgs boson, whose Yukawa interactions with the fermions also
-implement their masses~\cite{thomson2013modern}
-
-Mathematically, the Standard Model is described by a quantized gauge field
-theory with the symmetry group
+Gauge field theory describes this standard model with the symmetry group
 $SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
-meaning that its Lagrangian is invariant under local operations of group
-elements on its fields in certain representations.
-
+meaning that its Langrangian is invariant under local operations of the group
+on its fields in certain representations.
 Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
 indicates that they operate on colour charge triplets.
+Weak and electromagnetic forces are unified by electroweak interactions of
+$SU\!(2)_L$ and $U\!(1)_Y$, whose
+four bosons $W^1$, $W^2$, $W^3$, and $B$ respectively would be massless
+if the symmetry were unbroken,
+but it is `spontaneously' broken with the
+Higgs mechanism~\cite{
+higgs1964broken,
+englert1964broken
+}%
+\footnote{%
+Historical accuracy prefers `Brout–Englert–Higgs mechanism', but to
+properly credit associated work, Peter~Higgs has reportedly suggested
+``ABEGHHKtH mechanism''~\cite{close2011infinity}.%
+}
 
-Unified electroweak interactions come from $SU\!(2)_L$ and $U\!(1)_Y$, whose
-four bosons $W^1$, $W^2$, $W^3$, and $B$ respectively would be massless
+\TODO{L and Y, charge Q = T3 - Y or whatever it is}
+
 without the Higgs mechanism, which inserts four scalar fields and
@@
 
-\TODO{(Higgs) mechansim~\cite{
-higgs1964broken,
-englert1964broken
-}}
+the weak and electromagnetic interactions
+with the Higgs mechanism, which simultaneously provides the weak boson masses
+and adds massive Higgs boson, whose Yukawa interactions with the fermions also
+implement their masses~\cite{thomson2013modern}
 
@@
 
-\begin{singlespacing}
 \section{The \atlas\ Experiment}
 \label{sec:context_atlas}
-%
-\begin{epigraphs}
-\qitem{%
-if observing outer space gives us a view of the past, observing inner space
-would surely give us a glimpse into the future - would be interesting if NASA
-made a telescope for that%
-}%
-{Ken~M,
-\textit{Comment: Yahoo! News},
-2012~\cite{kenm2012inner}}
-\end{epigraphs}
-\end{singlespacing}
-
-\noindent Towards explaining the main result of this thesis,
+Towards explaining the main result of this thesis,
 a search for effects from supersymmetric models

### 2022-09-27 main/chapter_searches.tex
some improvements to SM; some misc tweaks

@@
 In the words of Jerzy~Neyman,
-``I have put the word ``~random~'' in inverted commas because it is very
+``I have put the word ``\,random\,'' in inverted commas because it is very
 difficult to define what is meant by it in practice''~\cite{
@@
 Figure~7 of~\cite{lester2022hunting}.
-When recommending error bar recipes, \atlas' statistics committee appears
+When recommending error bar recipes, \atlas's statistics committee appears
 to have not considered this option~\cite{cowan2011errorbars}.

### 2022-09-27 main/draft.tex
some more text on SM, minor tweaks
some improvements to SM; some misc tweaks

@@
 % significance measures
-ATLAS' recommended significance measure~\cite{atlas_significance} is
+ATLAS's recommended significance measure~\cite{atlas_significance} is
 \begin{align}
@@
 \TODO{more}
+
+These are classified into the strong sector
+% quarks
+\begin{itemize}
+\item up-type: up, charm, top ($u$, $c$, $t$), and
+\item down-type: down, strange, bottom ($d$, $s$, $b$),
+\end{itemize}
+% leptons
+
+% ex quotes
+
+\begin{singlespacing}
+\section{The Standard Model and beyond}
+\label{sec:context_sm_and_beyond}
+%
+\begin{epigraphs}
+\qitem{%
+In fact, it is a defensible conjecture that no
+theory can be complete and inviolable that emerges from finitely equipped minds
+and survives finitely scoped experiment, which are the twin rickety foundations
+for all theories.%
+}%
+{James~D.~Wells,
+\textit{Discovery Beyond the Standard Model of~Elementary~Particle~Physics},
+2019~\cite{wells2020discovery}}
+\end{epigraphs}
+\end{singlespacing}
+
+\begin{singlespacing}
+\section{The \atlas\ Experiment}
+\label{sec:context_atlas}
+%
+\begin{epigraphs}
+\qitem{%
+if observing outer space gives us a view of the past, observing inner space
+would surely give us a glimpse into the future - would be interesting if NASA
+made a telescope for that%
+}%
+{Ken~M,
+\textit{Comment: Yahoo! News},
+2012~\cite{kenm2012inner}}
+\end{epigraphs}
+\end{singlespacing}
+
+% \qitem{%
+% One geometry can not be more true than another; it can only be more convenient.
+% Geometry is not true, it is advantageous.}%
+% {Robert~M.~Pirsig,
+% \textit{Zen and the Art of Motorcycle Maintenance},
+% 1974~\cite{pirsig1999zen}}

### 2022-09-27 main/chapter_context.tex
clean up some bib names

@@
 \label{sec:context_sm}
-\TODO{particle content figure/tables}
-
 The current Standard Model features three generations of fermions,
@@
 and very massive $W^+$, $W^-$, and $Z$ bosons weadiate the weak force.
+\TODO{particle content figure/tables}
 

### 2022-09-27 main/chapter_context.tex
improve ewk section

@@
 
-Gauge field theory describes this standard model with the symmetry group
-$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
-meaning that its Langrangian is invariant under local operations of the group
+As a gauge field theory, the Standard Model has the symmetry group
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
+meaning that its langrangian is invariant under local operations of that group
 on its fields in certain representations.
+
 Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
 indicates that they operate on colour charge triplets.
-Weak and electromagnetic forces are unified by electroweak interactions of
-$SU\!(2)_L$ and $U\!(1)_Y$, whose
-four bosons $W^1$, $W^2$, $W^3$, and $B$ respectively would be massless
-if the symmetry were unbroken,
-but it is `spontaneously' broken with the
-Higgs mechanism~\cite{
+
+Weak and electromagnetic forces are unified in electroweak interactions with
+four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
+$SU\!(2)_L$ and $U\!(1)_Y$ symmetries respectively.
+These would be massless if the symmetry were perfectly respected, but it is
+`spontaneously' broken with the Higgs mechanism~\cite{
 higgs1964broken,
@@
 \footnote{%
-Historical accuracy prefers `Brout–Englert–Higgs mechanism', but to
-properly credit associated work, Peter~Higgs has reportedly suggested
-``ABEGHHKtH mechanism''~\cite{close2011infinity}.%
+Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
+Peter~Higgs has reportedly suggested
+``ABEGHHKtH mechanism''~\cite{close2011infinity}
+to properly credit other associated work.%
 }
-
-\TODO{L and Y, charge Q = T3 - Y or whatever it is}
-
-without the Higgs mechanism, which inserts four scalar fields and
-`spontaneously breaks' the gauge symmetry with a potential that moves the
-ground state of the model to a gauge-asymmetric configuration.
-From this new ground state, linear combinations of the electroweak fields
-acquire masses while ``eating up'' three of the four scalar fields to yield
-the three weak bosons and the photon:
-$(W^1, W^2)$ mix into $(W^+, W^-)$, and $(W^3, B)$ mix into $(Z, \gamma)$.
-The fourth and final scalar field is the Higgs boson $h$~\cite{
+by adding four scalar fields with a potential that moves the ground state of
+to a gauge-\emph{asymmetric} configuration.
+Around this new ground state, linear combinations of the electroweak fields
+acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
+four scalar fields to yield the three weak bosons and the photon:
+$W^1$ and $W^2$ mix into $W^+, W^-$, while $W^3$ and $B$ mix into $Z$ and
+$\gamma$.
+The fourth scalar is the Higgs boson $h$~\cite{
+glashow1959renorm,
+weinberg1967model,
+salam1959weak,
 rubakov1999classical,
@@
 }.
+Simultaneously, Yukawa interactions between the Higgs field and the fermions
+also implement fermion masses~\cite{thomson2013modern}.
+The subscript $_L$ on $SU\!(2)_L$ indicates that the weak forces interact only
+with left-chiral fermions, which are arranged into doublets as shown in
+Table~\TODO{ref table}.
 
-the weak and electromagnetic interactions
-with the Higgs mechanism, which simultaneously provides the weak boson masses
-and adds massive Higgs boson, whose Yukawa interactions with the fermions also
-implement their masses~\cite{thomson2013modern}
-
-\TODO{ABEGHHK'tH~\cite{close2011infinity}}
-
-\TODO{
-Glashow-Salam-Weinberg~\cite{
-glashow1959renorm,
-weinberg1967model,
-salam1959weak
-}
-}
-
-\TODO{charm november revolution~\cite{bjorken1985november}}
-
-The fermions of the Standard Model are listed in Table~\TODO{}.
 Older Standard Models included only two fermion generations and no
 neutrino masses~\cite{
-wells2020discovery
-},\TODO{cite more}
+wells2020discovery,
+bjorken1985november
+},
 but a modern model must include both three generations and non-zero neutrino

### 2022-09-27 main/chapter_context.tex
add an awkward standard model particles table

@@
 \TODO{rotation between weak and mass eigenstates: PMNS matrix!!}
+
+% using https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\label{tab:context_particles_sm}
+\centering
+\begin{tabular}{rrc}
+\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
+                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
+                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
+                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
+\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
+                          & electroweak & \makebox[7em]{\hfill $W^\pm$\!\! \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
+\end{tabular}
+\caption[Particles of the Standard Model]{%
+Particles of the Standard Model.
+All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
+spin-$0$ Higgs boson $h$.%
+}
+\end{table}
+
+Table~\ref{tab:context_particles_sm} \TODO{}
 

### 2022-09-27 main/chapter_context.tex
add sparticles table

@@
 \multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
-                          & electroweak & \makebox[7em]{\hfill $W^\pm$\!\! \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
+                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
 \end{tabular}
@@
 All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
-spin-$0$ Higgs boson $h$.%
+spin-$0$ Higgs boson $h$.
+Fermions come in left- and right-chiralities, although right handed neutrinos
+are not experimentally tested.%
 }
@@
 }}
+
+% using https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\label{tab:context_particles_mssm}
+\centering
+\begin{tabular}{rrc}
+\multirow{5}{*}{bosons}   & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
+                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
+                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
+                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[1ex]
+                          & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[3ex]
+\multirow{3}{*}{fermions} & strong                      & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
+                          & \multirow{2}{*}{electroweak}  & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
+                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
+\end{tabular}
+\caption[Particles of the Minimal Supersymmetric Standard Model]{%
+Particles of the Minimal Supersymmetric Standard Model
+\TODO{relate to Standard Model}
+\TODO{adds goldstino and gravitino}
+\TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
+}
+\end{table}
+
+Table~\ref{tab:context_particles_mssm} \TODO{}
 

### 2022-09-27 main/chapter_context.tex
improve right chiral neutrino comment
minor theory additions

@@
 The subscript $_L$ on $SU\!(2)_L$ indicates that the weak forces interact only
-with left-chiral fermions, which are arranged into doublets as shown in
-Table~\TODO{ref table}.
+interact with fermions that belong to left-chiral doublets: $(\nu_e, e)_L$,
+$(u, d)_L$ and their repeats in the three generations.
 
@@
 spin-$0$ Higgs boson $h$.
-Fermions come in left- and right-chiralities, although right handed neutrinos
-are not experimentally tested.%
+Fermions come in left- and right-chiralities, although the existence of
+right-chiral neutrinos is not strictly confirmed.%
 }
@@
 \TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
+\TODO{In this table, all bosons are spin-$0$ scalars and all fermions are spin-$1/2$.}
+\TODO{The bracketed $h$ corresponds to the Standard Model Higgs boson.}
 }

### 2022-09-27 main/chapter_2ljets.tex
add pmns, write a little on SM

@@
 missing transverse momentum, and the signal expectations would be added on top
-of the backgrounds if Nature realized these supersymmetric models.
+of the backgrounds if nature realized these supersymmetric models.
 

### 2022-09-27 main/chapter_context.tex
add pmns, write a little on SM

@@
 \label{sec:context_sm_and_beyond}
-\noindent Our best current understanding of Nature's fundamental objects in
+\noindent Our best current understanding of nature's fundamental objects in
 their observable behaviours as particles is encoded in the Standard Model,
@@
 \label{sec:context_sm}
-The current Standard Model features three generations of fermions,
+The current Standard Model features fermions
+(of four types in three generations),
 which interact through forces mediated by vector bosons
@@
 the massless photon $\gamma$ mediates the electromagnetic force,
-and very massive $W^+$, $W^-$, and $Z$ bosons weadiate the weak force.
-\TODO{particle content figure/tables}
-
-As a gauge field theory, the Standard Model has the symmetry group
-$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
-meaning that its langrangian is invariant under local operations of that group
-on its fields in certain representations.
-
-Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
-indicates that they operate on colour charge triplets.
-
-Weak and electromagnetic forces are unified in electroweak interactions with
-four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
-$SU\!(2)_L$ and $U\!(1)_Y$ symmetries respectively.
-These would be massless if the symmetry were perfectly respected, but it is
-`spontaneously' broken with the Higgs mechanism~\cite{
-higgs1964broken,
-englert1964broken
-}%
-\footnote{%
-Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
-Peter~Higgs has reportedly suggested
-``ABEGHHKtH mechanism''~\cite{close2011infinity}
-to properly credit other associated work.%
-}
-by adding four scalar fields with a potential that moves the ground state of
-to a gauge-\emph{asymmetric} configuration.
-Around this new ground state, linear combinations of the electroweak fields
-acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
-four scalar fields to yield the three weak bosons and the photon:
-$W^1$ and $W^2$ mix into $W^+, W^-$, while $W^3$ and $B$ mix into $Z$ and
-$\gamma$.
-The fourth scalar is the Higgs boson $h$~\cite{
-glashow1959renorm,
-weinberg1967model,
-salam1959weak,
-rubakov1999classical,
-cottingham2007greenwood
-}.
-Simultaneously, Yukawa interactions between the Higgs field and the fermions
-also implement fermion masses~\cite{thomson2013modern}.
-The subscript $_L$ on $SU\!(2)_L$ indicates that the weak forces interact only
-interact with fermions that belong to left-chiral doublets: $(\nu_e, e)_L$,
-$(u, d)_L$ and their repeats in the three generations.
-
-Older Standard Models included only two fermion generations and no
-neutrino masses~\cite{
-wells2020discovery,
-bjorken1985november
-},
-but a modern model must include both three generations and non-zero neutrino
-masses to be compatible with experiment:
-three fermion generations are directly established
-\TODO{kaon mixing CPV~\cite{
-cronin1964evidence,
-kabayasji1973cpv
-}, direct observation of tau/top/bottom~\cite{
-perl1977evidence,
-herb1977observation,
-abachi1995observation
-}, Z width~\cite{
-lep2006precision
-}
-},
-as are neutrino masses from their flavour oscillations demonstrating that
-they experience proper time~\cite{
-kamiokande1998measurement,
-superk1998evidence,
-superk1999measurement,
-lsnd1998evidence
-}.
-
-\TODO{fermion structure}
-\TODO{three generations}
-\TODO{rotation between weak and mass eigenstates: CKM matrix~\cite{
-cabibbo1963unitary,
-kabayasji1973cpv
-}}
-
-\TODO{neutrinos}
-\TODO{rotation between weak and mass eigenstates: PMNS matrix!!}
+and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
+These fundamental particles of the Standard Model are summarized in
+Table~\ref{tab:context_particles_sm}.
 
@@
 \begin{table}[tp]
-\label{tab:context_particles_sm}
 \centering
@@
 }
+\label{tab:context_particles_sm}
 \end{table}
 
-Table~\ref{tab:context_particles_sm} \TODO{}
+As a gauge field theory, the Standard Model has the symmetry group
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
+meaning that its lagrangian is invariant under local operations of that group
+on its fields in certain representations.
 
-\subsubsection{Opportunities for future Standard Models}
+Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
+indicates that they operate on colour-charge triplets of quarks.
+
+Weak and electromagnetic forces are unified in electroweak interactions with
+four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
+$SU\!(2)_L$ and $U\!(1)_Y$ symmetries respectively.
+These would be massless if the symmetry were perfectly respected, but it is
+`spontaneously' broken with the Higgs mechanism~\cite{
+higgs1964broken,
+englert1964broken
+}%
+\footnote{%
+Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
+Peter~Higgs has reportedly suggested
+``ABEGHHKtH mechanism''~\cite{close2011infinity}
+to fully credit other associated work.%
+}
+by adding four scalar fields with a potential that moves the ground state of
+to a gauge-\emph{asymmetric} configuration.
+Around this new ground state, linear combinations of the electroweak fields
+acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
+four scalar fields to yield the three weak bosons and the photon:
+$W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
+$\gamma$.
+The fourth scalar is the Higgs boson $h$~\cite{
+glashow1959renorm,
+weinberg1967model,
+salam1959weak,
+rubakov1999classical,
+cottingham2007greenwood
+}.
+
+Fermion masses arise through their Yukawa interactions with the Higgs field.
+The subscript $_L$ on $SU\!(2)_L$ indicates that the weak interaction only
+couples to fermions in left-chiral doublets: $(\nu_e, e)_L$,
+$(u, d)_L$ and their repeats across the three generations.
+In each generational trio (such as $u$, $c$, $t$), all three fermions share the
+same quantum numbers, but have different masses.
+A second generational difference occurs in quarks and neutrinos through the
+difference between their weak flavour eigenstates and the mass eigenstates of
+their free propagation;
+flavour and mass eigenstates are linearly related through unitary matrices:
+the CKM matrix for the quarks~\cite{
+cabibbo1963unitary,
+kabayasji1973cpv
+}
+and the PMNS matrix for the neutrinos~\cite{
+maki1962remarks,
+thomson2013modern
+}.
+
+Early versions of the Standard Model had only two fermion generations and
+no neutrino masses~\cite{
+wells2020discovery,
+bjorken1985november
+},
+but the third generation and non-zero neutrino masses are now observed;
+for example, the third generation is shown by $\mathrm{CP}$ asymmetry in
+kaons~\cite{
+cronin1964evidence,
+kabayasji1973cpv
+},
+precise measurements of the $Z$ boson~\cite{
+lep2006precision
+},
+and direct observations~\cite{
+perl1977evidence,
+herb1977observation,
+abachi1995observation
+};
+neutrino masses are also necessary to explain their observed flavour
+oscillations~\cite{
+kamiokande1998measurement,
+superk1998evidence,
+superk1999measurement,
+lsnd1998evidence
+}.
+
+Various theories can describe neutrino masses in the Standard Model, and it is
+currently uncertain which form best describes nature;
+this is both a limitation of the current Standard Model and an
+opportunity for future refinement.
+With future results, the Standard Model will continue to improve in its
+correspondence with nature, and we discuss some such opportunities
+for improvement in the next section.
+
+\subsection{Opportunities for future Standard Models}
+Problems with the Standard Model mean it needs replacing with supersymmetry
+--- is a false statement.
+Such an argument could be tempting to motivate this thesis, but the
+Standard Model is actually very good within its domain.
+Newton's laws are also very good within theirs \TODO{}
 
@@
 \begin{table}[tp]
-\label{tab:context_particles_mssm}
 \centering
@@
 }
+\label{tab:context_particles_mssm}
 \end{table}

### 2022-09-27 main/draft.tex
add pmns, write a little on SM

@@
 theoretical and experimental works which, over the last century or
-so, collaborated to understand and observe the actions of Nature's most
+so, collaborated to understand and observe the actions of nature's most
 fundamental known particles.

### 2022-09-27 main/chapter_context.tex
move table; write some philosophy and neutrinos

@@
 
-% using https://www.tablesgenerator.com/latex_tables
-\begin{table}[tp]
-\centering
-\begin{tabular}{rrc}
-\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
-                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
-                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
-                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
-\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
-                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
-\end{tabular}
-\caption[Particles of the Standard Model]{%
-Particles of the Standard Model.
-All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
-spin-$0$ Higgs boson $h$.
-Fermions come in left- and right-chiralities, although the existence of
-right-chiral neutrinos is not strictly confirmed.%
-}
-\label{tab:context_particles_sm}
-\end{table}
-
 As a gauge field theory, the Standard Model has the symmetry group
@@
 
+% using https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\centering
+\begin{tabular}{rrc}
+\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
+                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
+                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
+                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
+\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
+                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
+\end{tabular}
+\caption[Particles of the Standard Model]{%
+Particles of the Standard Model.
+All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
+spin-$0$ Higgs boson $h$.
+Fermions come in left- and right-chiralities, although the existence of
+right-chiral neutrinos is not strictly confirmed.%
+}
+\label{tab:context_particles_sm}
+\end{table}
+
 Early versions of the Standard Model had only two fermion generations and
@@
 Problems with the Standard Model mean it needs replacing with supersymmetry
---- is a false statement.
+--- is currently a false statement.
 Such an argument could be tempting to motivate this thesis, but the
-Standard Model is actually very good within its domain.
-Newton's laws are also very good within theirs \TODO{}
+Standard Model is actually good in its domain.
+This discussion is therefore oriented towards how the Standard Model could do
+better where it already works, and how that remit might expand to increase its
+generality.
 
-\TODO{Limitations: uncertainties and unknowns}
-\TODO{uncertainties: lagrangian form for neutrino masses, mass values (hierarchy, mixing angles); composition of dark matter}
-\TODO{unknowns: lack of convincing solutions; unification of gravity with quantum mechanics, origin of structure and parameters of standard model}
+When `nobody knows' the answer to a riddle, the status lies between two
+extremes:
+many answers appear satisfactory and none is the clear best,
+or no satisfactory answer is yet known.
+To distinguish these two cases, we choose for them the respective labels
+\emph{uncertain} and \emph{unknown}:
+the model of neutrino masses is \emph{uncertain},
+but the reconciliation quantum field theory with general relativity is
+\emph{unknown}.
+
+A problem with the Standard Model is that algorithms to compute its
+general consequences
+(where perturbative and discrete lattice methods are impractical)
+are unknown.
+But this is not an opportunity because it is inherent to the framework
+(of quantum field theory) that all serious alternatives also use.
+
+Before introducing supersymmetry, this section briefly review some promising
+directions of change for futrue models.
 
 \paragraph{Neutrino masses.}
-\TODO{Dirac neutrinos, possibly supplemented with a Majorana term}
-\TODO{Weinberg operator for Majorana neutrinos}
-
-\TODO{mass models: dirac, Majorana, seesaw~\cite{
-wells2020discovery,
-king2003neutrino,
-thomson2013modern
-}}
+Neutrinos could be Dirac fermions with Yukawa masses and boring right-chiral
+fields.
+Or they could also have Majorana mass terms, for which the seesaw mechanism
+can simultaneously generate the observed light neutrinos and heavy `sterile'
+neutrinos.
+Or their masses could come from other operators without right-handed
+fields~\cite{
+thomson2013modern,
+chala2021neutrino,
+wells2020discovery
+}.
+This riddle is largely unknown, and new results from theory or experiment
+that pinpoint a correct model could improve the Standard Model in the neutrino
+domain.
 
 \paragraph{Gravitation and cosmology.}
+
+\TODO{gravity}
 
@@
 \TODO{
+sterile neutrinos~\cite{boyarsky2019sterile},
 primordial black holes~\cite{bernard2019primordial},

### 2022-09-27 main/chapter_context.tex
try to condense intro text

@@
 Standard Model is actually good in its domain.
-This discussion is therefore oriented towards how the Standard Model could do
-better where it already works, and how that remit might expand to increase its
-generality.
+This discussion therefore looks towards where an altered Standard Model could
+go to improve what it already does, or to expand its generality.
 
@@
 or no satisfactory answer is yet known.
-To distinguish these two cases, we choose for them the respective labels
+We distinguish these two cases with the respective labels
 \emph{uncertain} and \emph{unknown}:
-the model of neutrino masses is \emph{uncertain},
+the model of neutrino masses is mostly \emph{uncertain},
 but the reconciliation quantum field theory with general relativity is
-\emph{unknown}.
+mostly \emph{unknown}.
 
-A problem with the Standard Model is that algorithms to compute its
-general consequences
-(where perturbative and discrete lattice methods are impractical)
-are unknown.
+One problem with the Standard Model is that algorithms to compute its
+\emph{general} consequences are unknown
+where perturbative and discrete lattice methods are impractical.
 But this is not an opportunity because it is inherent to the framework
-(of quantum field theory) that all serious alternatives also use.
-
-Before introducing supersymmetry, this section briefly review some promising
-directions of change for futrue models.
+of quantum field theory, which all competitive alternatives also use.
+Before introducing supersymmetry, this section reviews some promising
+directions of change for future models.
 

### 2022-09-27 main/chapter_context.tex
write some on dark matter

@@
 Problems with the Standard Model mean it needs replacing with supersymmetry
---- is currently a false statement.
-Such an argument could be tempting to motivate this thesis, but the
-Standard Model is actually good in its domain.
+--- is a false statement.
+Such an argument could motivate this thesis, but the Standard Model is actually
+good in its domain.
 This discussion therefore looks towards where an altered Standard Model could
-go to improve what it already does, or to expand its generality.
+go to improve its current abilities or to expand its generality.
 
@@
 \emph{uncertain} and \emph{unknown}:
-the model of neutrino masses is mostly \emph{uncertain},
-but the reconciliation quantum field theory with general relativity is
+the model of neutrino masses is mostly \emph{uncertain}, but
+the reconciliation quantum field theory with general relativity is
 mostly \emph{unknown}.
@@
 }.
-This riddle is largely unknown, and new results from theory or experiment
-that pinpoint a correct model could improve the Standard Model in the neutrino
+The answer is uncertain, and new results from theory or experiment that
+pinpoint a correct model could improve the Standard Model in the neutrino
 domain.
@@
 \paragraph{Gravitation and cosmology.}
+Gravity does not feature in the standard model, nor does it measurably impact
+particle physics experiments.
+Although gravitation is in another domain, the particle content of the
+Standard Model does impact cosmology.
 
-\TODO{gravity}
-
-\TODO{dark matter: rotation curves~\cite{
+The universe contains dark matter which interacts with gravity but not much
+with other forces; it is observed through galactic rotation curves,
+gravitational lensing, and cosmic microwave background data,
+but its composition is uncertain~\cite{
 begeman1991rotation,
-garrett2010dark
-}}
+garrett2010dark,
+planck2020results
+}.
+Dark matter candidates include new particles such as
+sterile neutrinos~\cite{boyarsky2019sterile},
+WIMPs (Weakly Interacting Massive Particles)~\cite{jungman1996wimp},
+or
+axions~\cite{kim2008axions},
+but also macroscopic objects such as
+primordial black holes~\cite{bernard2019primordial}.
+Dark matter may well comprise various of different components, and it would
+be a success for future particle physics theories to pinpoint necessary dark
+matter particles.
 
 \TODO{
-sterile neutrinos~\cite{boyarsky2019sterile},
-primordial black holes~\cite{bernard2019primordial},
-WIMPs~\cite{jungman1996wimp},
-axions~\cite{kim2008axions},
-dust?~\cite{garrett2010dark}
-}
-\TODO{combinations!}
-
-\TODO{
-dark matter,
+Three more mysteries:
 cosmological constant~\cite{riess1998observational},
@@
 }
-\TODO{note \lcdm}
 

### 2022-09-27 main/chapter_context.tex
fiddle with cosmology a little; add sakharov

@@
 but the third generation and non-zero neutrino masses are now observed;
-for example, the third generation is shown by $\mathrm{CP}$ asymmetry in
+for example, the third generation is shown by charge-parity $CP$ asymmetry in
 kaons~\cite{
@@
 
-When `nobody knows' the answer to a riddle, the status lies between two
-extremes:
-many answers appear satisfactory and none is the clear best,
+When `nobody knows' the answer to a riddle, the status lies somewhere between
+two extremes:
+either many answers appear satisfactory and none is the clear best,
 or no satisfactory answer is yet known.
@@
 
-\TODO{
-Three more mysteries:
-cosmological constant~\cite{riess1998observational},
+Other mysteries of cosmology include:
+the positive cosmological constant,
+the inflation of the early universe,
+and baryogenesis~\cite{
+wells2020discovery,
+riess1998observational
+}.
+Each lies somewhere between uncertain and unknown, but may not be entirely
+out of reach;
+the Higgs field could plausibly have driven
 inflation~\cite{bezrukov2008higgs},
-baryogenesis (matter-antimatter asymmetry)
-}
+and the Standard Model does slightly violate $C$ and $CP$, both of which
+are required for baryogenesis~\cite{sakharov1991re}.
 
 \paragraph{Origins.}
+\TODO{parameters and structure}
 

### 2022-09-28 main/chapter_context.tex
write some on origins

@@
 This discussion therefore looks towards where an altered Standard Model could
-go to improve its current abilities or to expand its generality.
+improve its current abilities or to expand its generality to describe more of
+the universe.
 
@@
 two extremes:
-either many answers appear satisfactory and none is the clear best,
-or no satisfactory answer is yet known.
-We distinguish these two cases with the respective labels
-\emph{uncertain} and \emph{unknown}:
+either many answers are satisfactory with no clear best,
+or no satisfactory answers are yet known.
+We distinguish these two cases with the labels
+\emph{uncertain} and \emph{unknown} respectively:
 the model of neutrino masses is mostly \emph{uncertain}, but
@@
 
-One problem with the Standard Model is that algorithms to compute its
-\emph{general} consequences are unknown
-where perturbative and discrete lattice methods are impractical.
-But this is not an opportunity because it is inherent to the framework
-of quantum field theory, which all competitive alternatives also use.
-Before introducing supersymmetry, this section reviews some promising
-directions of change for future models.
+One real problem with the Standard Model is that algorithms to compute its
+general consequences are unknown, since practical perturbative and discrete
+lattice methods are limited.
+Such imprecise modelling weakens our inferences from experimental results,
+anomalous or not,
+but does not present much opportunity because it is common to the quantum field
+theory that is used by all known viable models.
+This section briefly reviews some other directions of positive change for
+future models.
 
@@
 fields.
-Or they could also have Majorana mass terms, for which the seesaw mechanism
-can simultaneously generate the observed light neutrinos and heavy `sterile'
+They could also have Majorana mass terms, for which the seesaw mechanism can
+simultaneously generate the observed light neutrinos with very heavy `sterile'
 neutrinos.
-Or their masses could come from other operators without right-handed
-fields~\cite{
+Or their masses could come from other operators without needing the
+right-chiral fields~\cite{
 thomson2013modern,
@@
 }.
-The answer is uncertain, and new results from theory or experiment that
+This answer is uncertain, and new results from theory or experiment that
 pinpoint a correct model could improve the Standard Model in the neutrino
@@
 sterile neutrinos~\cite{boyarsky2019sterile},
-WIMPs (Weakly Interacting Massive Particles)~\cite{jungman1996wimp},
+Weakly Interacting Massive Particles (WIMPs)~\cite{jungman1996wimp},
 or
-axions~\cite{kim2008axions},
+axion-like particles~\cite{kim2008axions},
 but also macroscopic objects such as
@@
 }.
-Each lies somewhere between uncertain and unknown, but may not be entirely
-out of reach;
-the Higgs field could plausibly have driven
+Answers to these mysteries all lie somewhere between uncertain and unknown,
+but may not be entirely out of reach;
+for example, the Higgs field could plausibly have driven
 inflation~\cite{bezrukov2008higgs},
@@
 \paragraph{Origins.}
-\TODO{parameters and structure}
+Why does the Standard Model have its structure?
+Why is it configured with the observed values of our parameters?
+Why do we even live in $3+1$ dimensional spacetime with quantum mechanics?
+Answers to these questions are largely unknown, but would be
+attractive features for a fundamental theory.
+
+Certain Standard Model structures are quite general and parametrized such
+that parts can be tuned to describe nature.
+For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$ whose
+non-zero values would induce $\mathrm{CP}$ violation.
+Observations find it very close to zero, suggesting that nature does not use
+that freedom, so one promising feature of axion-like models is their
+implication that $\theta_\mathrm{CP} = 0$~\cite{
+kim2008axions,
+thomson2013modern,
+martin2017particle
+}.
+
+Justified constraints on the model could have led to better predictions of
+past data and might extrapolate to new phenomena through their deeper
+understanding of nature.
+Attempts to relate empirical parameters~\cite{
+sato1979ratio,
+beg1982gauge,
+koide1983fermion,
+denterria2012gaussian
+}
+or to derive mathematical structures~\cite{
+goyal2010origin,
+skilling2021abc,
+axioms1010038
+}
+are curious, but will remain speculative until they provide
+real and novel benefits.
 
 \paragraph{Aesthetics.}
-
 \TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
@@
 \TODO{hierarchy problem}
+
+Naturalness is opinion and only nature decides what is really natural.
 
@@
 Particles of the Minimal Supersymmetric Standard Model
+\TODO{check and cite}
 \TODO{relate to Standard Model}

### 2022-09-28 main/chapter_context.tex
fix higgs H mixup; write on naturalness

@@
 Other mysteries of cosmology include:
-the positive cosmological constant,
+the small positive cosmological constant,
 the inflation of the early universe,
@@
 that parts can be tuned to describe nature.
-For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$ whose
-non-zero values would induce $\mathrm{CP}$ violation.
-Observations find it very close to zero, suggesting that nature does not use
-that freedom, so one promising feature of axion-like models is their
-implication that $\theta_\mathrm{CP} = 0$~\cite{
+For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$
+whose non-zero values would induce $\mathrm{CP}$ violation in the strong
+sector.
+Observations show that it cannot be far from zero, and one promising feature of
+axion-like models is their implication that $\theta_\mathrm{CP} = 0$~\cite{
 kim2008axions,
@@
 \paragraph{Aesthetics.}
-\TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
-\TODO{strong CP problem}
-\TODO{hierarchy problem}
+In the search for improved models, heuristic criteria can help to identify
+appealing candidates.
+Two such criteria are `fine-tuning' and `naturalness', both of which are
+aesthetic judgements about dimensionless model parameters.
+Naturalness is opinion; only nature decides what is really natural,
+but these criteria can still guide the discourse of our scientific searches
+towards useful directions~\cite{wells2020discovery}.
 
-Naturalness is opinion and only nature decides what is really natural.
+Fine-tuning criteria dislike tight constraints, and naturalness contains
+various ideas of which one is that nice numbers should have moderate orders of
+magnitude~\cite{giudice2008naturally}.
+For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned within its
+experimental constraint of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
+large $11$ in its exponent defies naturalness.
+
+Another example is the hierarchy problem of the Higgs boson mass $m(h)$,
+which notes that the energy scale of new heavy particles linearly modifies the
+$m(h)$ through direct or indirect couplings.
+With any particles near the Planck scale $\sim 10^{18}\,\eV[G]$, implementing
+the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires finely-tuned
+cancellations between positive and negative additions.
+New fermions subtract from $m(h)$ and new scalars add by similar amounts,
+so new models can implement this fine-tuning by inserting only matched pairs of
+fermions and scalars --- this is the strategy of supersymmetric models.
 
@@
 
-% using https://www.tablesgenerator.com/latex_tables
+% made using:
+% https://www.tablesgenerator.com/latex_tables
 \begin{table}[tp]
@@
 entire $100$ to $1000\,\eV[G]$ window by targetting decays including
-$H\to b\bar b$,
-$H\to \gamma\gamma$,
-$H\to ZZ^* \to 4\ell$, and
-$H\to ZZ \to \ell\ell\nu\nu$;
+$h \to b\bar b$,
+$h \to \gamma\gamma$,
+$h \to ZZ^* \to 4\ell$, and
+$h \to ZZ \to \ell\ell\nu\nu$;
 these channels demand precise measurements of leptons, photons, and
@@
 Both \atlas\ and \cms\ have now found a Higgs boson candidate at
-$m(H) = 125\,\eV[G]$~\cite{
+$m(h) = 125\,\eV[G]$~\cite{
 atlas2012higgs,
@@
 including
-$H \to b\bar b$
+$h \to b\bar b$
 and

### 2022-09-28 main/chapter_context.tex
begin sketching susy section

@@
 magnitude~\cite{giudice2008naturally}.
-For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned within its
-experimental constraint of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
-large $11$ in its exponent defies naturalness.
+For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned by an
+observation of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
+large $11$ in its exponent also defies naturalness.
 
 Another example is the hierarchy problem of the Higgs boson mass $m(h)$,
-which notes that the energy scale of new heavy particles linearly modifies the
-$m(h)$ through direct or indirect couplings.
+which notes that the energy scale of new heavy particles add quadratically to
+$m(h)^2$ through direct or indirect couplings.
 With any particles near the Planck scale $\sim 10^{18}\,\eV[G]$, implementing
@@
 \label{sec:context_susy}
+
+% supersymmetry
+
+% supersymmetric models
+
+% fermion boson multiplets
+
+% table of particle content
+
+% explain mixing
+
+% reproduce arguments for Wino-Bino C1 N2
+
+Lightest Supersymmetric Particle (LSP)
+
+Gauge Mediated Supersymmetry Breaking (GMSB)
+
+% reproduce arguments for goldsino and gravitino
+% spin-$3/2$ parts couples only gravitationally, see another spin-$1/2$ fermion
 
@@
 Table~\ref{tab:context_particles_mssm} \TODO{}
-
-\subsubsection{Minimal supersymmetric models}
-Lightest Supersymmetric Particle (LSP)
-
-\subsubsection{Guage Mediated Supersymmetry Breaking}
-Gauge Mediated Supersymmetry Breaking (GMSB)
 

### 2022-09-28 main/chapter_context.tex
add some susy text

@@
 
-Another example is the hierarchy problem of the Higgs boson mass $m(h)$,
-which notes that the energy scale of new heavy particles add quadratically to
-$m(h)^2$ through direct or indirect couplings.
-With any particles near the Planck scale $\sim 10^{18}\,\eV[G]$, implementing
-the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires finely-tuned
-cancellations between positive and negative additions.
+Another example is the `hierarchy' problem of the Higgs boson mass $m(h)$,
+which notes that new particles at a heavy mass scale $\Lambda$ induce changes
+of $\Delta m(h)^2 = \mathcal{O}(\Lambda^2)$,
+through direct or indirect couplings.
+With any particles near the Planck scale $\Lambda \sim 10^{18}\,\eV[G]$,
+implementing the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires
+finely-tuned cancellations between positive and negative contributions.
 New fermions subtract from $m(h)$ and new scalars add by similar amounts,
-so new models can implement this fine-tuning by inserting only matched pairs of
-fermions and scalars --- this is the strategy of supersymmetric models.
+so new models can implement this fine-tuning by inserting only matched sets of
+fermions and bosons.
+This solution is the strategy of supersymmetric models.
 
@@
 \label{sec:context_susy}
+Supersymmetry is the concept of symmetry under the operation of exchanging
+bosonic and fermionic states, and supersymmetric models use supersymmetry in
+their construction.
+Supersymmetry itself is not a model, but if our searches did hypothetically
+find evidence for any supersymmetric model, then that would be evidence for
+supersymmetry in nature.
 
-% supersymmetry
+Nature is not totally supersymmetric.
+If it were, then there would be bosonic `superpartners' for each of the known
+fermions (and vice versa) with the same masses and charges, and those
+`sparticles' would have been seen long ago.
+However, like the broken symmetry of the Standard Model gauge group,
+supersymmetry could still participate in a broken condition.
 
-% supersymmetric models
-
+Any new model must include the observational facts of the Standard Model,
+and with supersymmetry the minimum viable product is the
+Minimal Supersymmetric Standard Model (MSSM).
 % fermion boson multiplets

### 2022-09-28 main/chapter_context.tex
remove some EMs; minor susy tweaks

@@
 `sparticles' would have been seen long ago.
-However, like the broken symmetry of the Standard Model gauge group,
-supersymmetry could still participate in a broken condition.
+But supersymmetry could still participate in a broken condition where the
+sparticles are hidden at larger masses.
 
@@
 \end{quote}
-The EM calorimeters provide located electron/photon energy measurements,
+The electromagnetic calorimeters provide located electron/photon energy
+measurements,
 and the hadronic calorimeters subsequently aim to stop and measure hadronic
@@
 Neither `A' in \atlas\ stands for Argon, but perhaps one should.
-Liquid-argon is the chosen active material in the EM and forward calorimeters
-for its robust, linear response in converting showers to electrical signals
-for readout by electrodes.
+Liquid-argon is the chosen active material in the electromagnetic and forward
+calorimeters for its robust, linear response in converting showers to
+electrical signals for readout by electrodes.
 Liquid argon is poured as a liquid to fill the gaps between the various

### 2022-09-28 main/chapter_context.tex
less janky muon sentence

@@
 These calorimeter systems achieve attenuations of at least ten interaction
-lengths throughout their $|\eta| < 4.9$ range.
-Maximizing attenuation minimizes noise from hadronic `punch-through' into the
-muon chambers.
+lengths throughout their $|\eta| < 4.9$ range;
+by maximizing this attenuation, the design minimizes noise from hadronic
+`punch-through' of non-muons into the muon system.
 

### 2022-09-28 main/chapter_context.tex
add persint manual

@@
 \end{singlespacing}
+
+\TODO{split chapters again}
 
@@
 ]{%
-A central slice of \atlas's design, reproduced from~\cite{atlas1999design1},
+A central slice of \atlas's design, reproduced from~\cite{atlas1999design1,
+persint2014manual},
 including a simulated event that interacts with several sensitive subsystems.

### 2022-09-28 main/chapter_2ljets.tex
move higgs to the top, write some more susy junk

@@
 One may, of course, define a directed `transverse energy', but it will
-not be conserved (in the standard model) unless it is equivalent to a conserved
+not be conserved (in the Standard Model) unless it is equivalent to a conserved
 momentum.
@@
 Figure~\ref{fig:2ljets_presel_met}.
-What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
+What remains of the Standard Model in the high-$\metsig$ tail of $\twoljets$
 selections is almost pure \diboson\ --- $ZZ$ with a hard $Z\to \nu\nu$,
@@
 Lacking direct observation of the sleptons, or any other supersymmetric
-resonances, at the LHC, $\mttwo$ remains useful for separating standard model
+resonances, at the LHC, $\mttwo$ remains useful for separating Standard Model
 processes since, with ideal reconstruction,
@@
 Note that the region $\mjetone > 110$ is also checked in this validation region
-and seen to contain no data --- a good sign for the standard model.
+and seen to contain no data --- a good sign for the Standard Model.
 Since background estimates are near zero up there, non-zero data would have
@@
 triggers~\cite{atlas_susy_compressed_2l_2018_run2},
-we still have some sensitivity since standard model backgrounds drop off
+we still have some sensitivity since Standard Model backgrounds drop off
 below $\mll \approx m(Z)$.

### 2022-09-28 main/chapter_context.tex
move higgs to the top, write some more susy junk

@@
 }
-by adding four scalar fields with a potential that moves the ground state of
-to a gauge-\emph{asymmetric} configuration.
+by adding four scalar fields (as a complex doublet) with a potential that moves
+the ground state of to a gauge-\emph{asymmetric} configuration.
 Around this new ground state, linear combinations of the electroweak fields
@@
 \paragraph{Gravitation and cosmology.}
-Gravity does not feature in the standard model, nor does it measurably impact
+Gravity does not feature in the Standard Model, nor does it measurably impact
 particle physics experiments.
@@
 and with supersymmetry the minimum viable product is the
-Minimal Supersymmetric Standard Model (MSSM).
+Minimal Supersymmetric Standard Model (MSSM), which adds the school of
+new particles whose mass eigenstates are listed in
+Table~\ref{tab:context_particles_mssm},
+in which all sparticles are marked with tildes $\tilde{~}$.
+
 % fermion boson multiplets
+The MSSM matches each standard fermion to a pair of scalar
+`sfermions' (left and right)
+and each vector gauge field to a `gaugino' fermion
+($\tilde g$, $\tilde W^1$, $\tilde W^2$, $\tilde W^3$, and $\tilde B$).
+The Higgs field does not translate so directly, but is replaced with two sets
+of complex scalar fields
+($H^+_u$, $H^0_u$ and $H^0_d$, $H^\pm_d$) with
+gaugino superpartners
+($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
+These mix into uncertain mass eigenstates.
+Gaugino mixtures are the charged charginos $C^\pm_i$ and
+neutral neutralinos $N_i$.
+From the four complex scalars, three degrees of freedom are eaten by the
+weak bosons to leave five massive scalars $H^0$, $A^0$, $H^\pm$ and the
+reborn Standard Model Higgs boson $h$.
 
-% table of particle content
+None of these new particles is yet observed, but \TODO{}
 
-% explain mixing
+\TODO{R-parity}
 
-% reproduce arguments for Wino-Bino C1 N2
+\begin{equation}
+P_R = (-1)^{3(B - L) + 2s}
+\end{equation}
+\TODO{for baryon number $B$, lepton number $L$ and spin $s$.}
 
@@
 
-\TODO{higgs mass $\lesssim 130\,\eV[G]$~\cite{
+\TODO{consistent with a $m(h) \lesssim 130\,\eV[G]$~\cite{
 quiros1997constraints,
@@
 \begin{tabular}{rrc}
-\multirow{5}{*}{bosons}   & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
+\multirow{5}{*}{bosons}   & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[1ex]
+                          & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                           &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                           & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
-                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[1ex]
-                          & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[3ex]
-\multirow{3}{*}{fermions} & strong                        & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
-                          & \multirow{2}{*}{electroweak}  & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
+                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[3ex]
+\multirow{3}{*}{fermions} & gluinos                       & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
+                          & \multirow{2}{*}{gauginos}     & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                           &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
@@
 \caption[Particles of the Minimal Supersymmetric Standard Model]{%
-Particles of the Minimal Supersymmetric Standard Model
-\TODO{check and cite}
-\TODO{relate to Standard Model}
-\TODO{adds goldstino and gravitino}
-\TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
-\TODO{In this table, all bosons are spin-$0$ scalars and all fermions are spin-$1/2$.}
-\TODO{The bracketed $h$ corresponds to the Standard Model Higgs boson.}
+Particles of the Minimal Supersymmetric Standard Model in their mass
+eigenstates~\cite{martin2016primer}, in addition to the Standard Model
+particles of Table~\ref{tab:context_particles_sm}.
+The Standard Model Higgs boson $h$ (bracketed) is repeated here with its new
+pals.
+Sparticles are marked with tildes $\tilde{~}$.
+All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
+The charginos $\tilde C^\pm_i$ and neutralinos $\tilde N^i$ are alternatively
+labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches},
+and comprise mixtures of the charged and neutral superpartners to gauge fields
+respectively.
+Each squark and slepton has two versions from mixtures of the left and right
+fields.
+Locally broken supersymmetry also adds a spin-$3/2$ gravitino $\tilde G$.
 }
@@
 \end{table}
-
-Table~\ref{tab:context_particles_mssm} \TODO{}
 

### 2022-09-28 main/chapter_context.tex
more theory tweaks

@@
 lattice methods are limited.
-Such imprecise modelling weakens our inferences from experimental results,
-anomalous or not,
-but does not present much opportunity because it is common to the quantum field
+Imprecise modelling weakens experimental results (anomalous or not),
+but does not present an opportunity because it is common to the quantum field
 theory that is used by all known viable models.
@@
 }
-are curious, but will remain speculative until they provide
-real and novel benefits.
+are curious, but will remain speculative until they provide novel benefits.
 
@@
 their construction.
-Supersymmetry itself is not a model, but if our searches did hypothetically
-find evidence for any supersymmetric model, then that would be evidence for
-supersymmetry in nature.
-
 Nature is not totally supersymmetric.
@@
 ($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
-These mix into uncertain mass eigenstates.
-Gaugino mixtures are the charged charginos $C^\pm_i$ and
-neutral neutralinos $N_i$.
+These mix into uncertain mass eigenstates:
+and the gaugino mixtures are the charged charginos $C^\pm_i$ and
+neutral neutralinos $N_i$
+(also labelled $\chargino_i$ and
+$\neutralino_i$~\cite{atlas2022searches}),
+each squark and slepton has two versions from mixtures of the left and right
+fields.
 From the four complex scalars, three degrees of freedom are eaten by the
-weak bosons to leave five massive scalars $H^0$, $A^0$, $H^\pm$ and the
-reborn Standard Model Higgs boson $h$.
+weak bosons to leave the reborn Standard Model Higgs boson $h$
+and four new scalars.
 
-None of these new particles is yet observed, but \TODO{}
+None of these new particles is yet observed. \TODO{}
 
@@
 All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
-The charginos $\tilde C^\pm_i$ and neutralinos $\tilde N^i$ are alternatively
-labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches},
-and comprise mixtures of the charged and neutral superpartners to gauge fields
-respectively.
-Each squark and slepton has two versions from mixtures of the left and right
-fields.
 Locally broken supersymmetry also adds a spin-$3/2$ gravitino $\tilde G$.

### 2022-09-28 main/chapter_context.tex
write more on susy: rparity

@@
 \subsection{Opportunities for future Standard Models}
-Problems with the Standard Model mean it needs replacing with supersymmetry
---- is a false statement.
-Such an argument could motivate this thesis, but the Standard Model is actually
-good in its domain.
+The Standard Model works very well within its domain.
 This discussion therefore looks towards where an altered Standard Model could
@@
 `sparticles' would have been seen long ago.
-But supersymmetry could still participate in a broken condition where the
+But supersymmetry might still participate in a broken condition if the
 sparticles are hidden at larger masses.
@@
 Any new model must include the observational facts of the Standard Model,
-and with supersymmetry the minimum viable product is the
-Minimal Supersymmetric Standard Model (MSSM), which adds the school of
-new particles whose mass eigenstates are listed in
-Table~\ref{tab:context_particles_mssm},
-in which all sparticles are marked with tildes $\tilde{~}$.
+and the Minimal Supersymmetric Standard Model (MSSM) satisfies this while
+adding a school of new particles,
+which is listed in Table~\ref{tab:context_particles_mssm}.
 
@@
 ($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
-These mix into uncertain mass eigenstates:
-and the gaugino mixtures are the charged charginos $C^\pm_i$ and
+(All sparticles are marked with tildes $\tilde{~}$.)
+These fields mix into mass eigenstates which current data do not entirely
+determine:
+the gaugino mixtures are the charged charginos $C^\pm_i$ and
 neutral neutralinos $N_i$
-(also labelled $\chargino_i$ and
+(also labelled $\chargino_i$,
 $\neutralino_i$~\cite{atlas2022searches}),
-each squark and slepton has two versions from mixtures of the left and right
-fields.
+and each squark and slepton has two states that mix the left and right fields.
 From the four complex scalars, three degrees of freedom are eaten by the
 weak bosons to leave the reborn Standard Model Higgs boson $h$
-and four new scalars.
-
-None of these new particles is yet observed. \TODO{}
-
-\TODO{R-parity}
-
-\begin{equation}
-P_R = (-1)^{3(B - L) + 2s}
-\end{equation}
-\TODO{for baryon number $B$, lepton number $L$ and spin $s$.}
-
-Lightest Supersymmetric Particle (LSP)
-
-Gauge Mediated Supersymmetry Breaking (GMSB)
-
-% reproduce arguments for goldsino and gravitino
-% spin-$3/2$ parts couples only gravitationally, see another spin-$1/2$ fermion
-
-\TODO{consistent with a $m(h) \lesssim 130\,\eV[G]$~\cite{
-quiros1997constraints,
-wells2020discovery
-}}
-
-\TODO{primer~\cite{
+and four new scalars~\cite{
 martin2016primer
-}}
+}.
 
@@
 \end{tabular}
-\caption[Particles of the Minimal Supersymmetric Standard Model]{%
-Particles of the Minimal Supersymmetric Standard Model in their mass
-eigenstates~\cite{martin2016primer}, in addition to the Standard Model
-particles of Table~\ref{tab:context_particles_sm}.
-The Standard Model Higgs boson $h$ (bracketed) is repeated here with its new
-pals.
+\caption[%
+Particles and sparticles of the Minimal Supersymmetric Standard Model%
+]{%
+Particles and sparticles of the Minimal Supersymmetric Standard Model
+in their mass eigenstates~\cite{martin2016primer}, in addition to the
+Standard Model particles of Table~\ref{tab:context_particles_sm}.
+The standard Higgs boson ($h$) is repeated here with its new pals.
 Sparticles are marked with tildes $\tilde{~}$.
 All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
-Locally broken supersymmetry also adds a spin-$3/2$ gravitino $\tilde G$.
 }
@@
 \end{table}
+
+Various mechanisms could spontaneously break supersymmetry.
+Gravitational ideas exist, but one model considered in this thesis uses a
+Gauge-Mediated Supersymmetry Breaking (GMSB) mechanism,
+which adds yet more fields:
+a `secluded' sector which breaks the symmetry, and
+a `messenger' sector which
+communicates the effect back to the observable sector.
+Such models also add a spin-$3/2$ `gravitino' ($\gravitino$),
+which should be light and whose helicity $\pm1/2$ components would interact
+in collider physics~\cite{
+giudice1999gmsb,
+dine1982pheno,
+martin2016primer
+}.
+
+All models considered in this thesis conserve $R$-parity
+\begin{equation}
+P_R = (-1)^{3(B - L) + 2s}
+\end{equation}
+(with baryon number $B$, lepton number $L$ and spin $s$)
+in their interactions~\cite{
+farrar1978rparity,
+martin2016primer
+}.
+This conservation helps to prevent proton decays,
+but also makes the Lightest Supersymmetric Particle (LSP) a plausible
+WIMP dark matter candidate since it cannot decay to purely Standard Model
+particles.
+
+% current observations reaching up to $\eV[T]$ scale
+% hierarchy problem
+
+Broken supersymmetry does not perfectly resolve the hierarchy problem, but
+could still reduce its fine-tuning if the superpartner masses are below the
+$\lesssim 1\,\eV[T]$ scale~\cite{martin2016primer}.
+
+the MSSM has the success that it is consistent with the observed
+$m(h) \lesssim 135\,\eV[G]$~\cite{
+quiros1997constraints,
+wells2020discovery
+}
+
+None of these hypothetical particles is yet observed, but it remains possible
+that they could be invoked to describe future experimental data.
 

### 2022-09-28 main/chapter_context.tex
wrap susy section

@@
 
-Gluons correspond to the eight generators of $SU\!(3)_C$, whose subscript $_C$
-indicates that they operate on colour-charge triplets of quarks.
-
-Weak and electromagnetic forces are unified in electroweak interactions with
-four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
-$SU\!(2)_L$ and $U\!(1)_Y$ symmetries respectively.
+In this model, gluons correspond to the eight generators of $SU\!(3)_C$, whose
+subscript $_C$ indicates that they operate on colour-charge triplets of quarks.
+Weak and electromagnetic forces are also unified into electroweak interactions
+with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
+$SU\!(2)_L$ and $U\!(1)$ symmetries respectively.
 These would be massless if the symmetry were perfectly respected, but it is
@@
 }.
-Answers to these mysteries all lie somewhere between uncertain and unknown,
+Solutions to these mysteries all lie somewhere between uncertain and unknown,
 but may not be entirely out of reach;
@@
 
-Certain Standard Model structures are quite general and parametrized such
-that parts can be tuned to describe nature.
+Certain Standard Model structures are quite general, and are parametrized such
+that parts can be smoothly disabled if they do not describe nature.
 For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$
@@
 sector.
-Observations show that it cannot be far from zero, and one promising feature of
-axion-like models is their implication that $\theta_\mathrm{CP} = 0$~\cite{
+Observations find that $\theta_\mathrm{CP}$ cannot be far from zero, and one
+attractive feature of axion-like models is their requirement that
+$\theta_\mathrm{CP} = 0$~\cite{
 kim2008axions,
@@
 neutral neutralinos $N_i$
-(also labelled $\chargino_i$,
-$\neutralino_i$~\cite{atlas2022searches}),
+(also labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches}),
 and each squark and slepton has two states that mix the left and right fields.
@@
 \caption[%
-Particles and sparticles of the Minimal Supersymmetric Standard Model%
+Particles of the Minimal Supersymmetric Standard Model%
 ]{%
-Particles and sparticles of the Minimal Supersymmetric Standard Model
+Particles of the Minimal Supersymmetric Standard Model
 in their mass eigenstates~\cite{martin2016primer}, in addition to the
@@
 
-% current observations reaching up to $\eV[T]$ scale
-% hierarchy problem
-
-Broken supersymmetry does not perfectly resolve the hierarchy problem, but
-could still reduce its fine-tuning if the superpartner masses are below the
-$\lesssim 1\,\eV[T]$ scale~\cite{martin2016primer}.
-
-the MSSM has the success that it is consistent with the observed
-$m(h) \lesssim 135\,\eV[G]$~\cite{
+Supersymmetric modelling successfully predicted the lightest Higgs boson
+mass to be below $\lesssim 130\,\eV[G]$ long before its observation at
+$125\,\eV[G]$~\cite{
+espinosa1992higgs,
+espinosa1993upper,
 quiros1997constraints,
-wells2020discovery
-}
-
-None of these hypothetical particles is yet observed, but it remains possible
-that they could be invoked to describe future experimental data.
+wells2020discovery,
+atlas2012higgs,
+atlas2012combined,
+cms2012higgs
+}.
+This prediction, however, assumed a mass scale below $1\,\eV[T]$,
+and observations at the LHC are now limiting sparticle masses
+into super-$\eV[T]$ territory (in specific models).
+Such limits are also troublesome for the hierarchy problem, because large
+superpartner masses introduce fine-tuning problems of their
+own~\cite{martin2016primer}.
+Despite its current lack of observational evidence, however,
+supersymmetry remains a useful tool for building theoretically plausible
+models, and might help us to understand future discoveries.
 

### 2022-09-28 main/chapter_experiment.tex
type some on bkg-only and excl fits
split context back out into theory and exp

--- 2022-05-13 main/chapter_experiment.tex
@@
 \begin{singlespacing}
-\chapter{The ATLAS experiment}
+\chapter{The \atlas\ Experiment}
 \label{chapter:experiment}
@@
 if observing outer space gives us a view of the past, observing inner space
-would surely give us a glimpse into the future --- would be interesting if NASA
+would surely give us a glimpse into the future - would be interesting if NASA
 made a telescope for that%
 }%
-{Ken M}
+{Ken~M,
+\textit{Comment: Yahoo! News},
+2012~\cite{kenm2012inner}}
 \end{epigraphs}
@@
 
-Hello
+Towards explaining the main result of this thesis,
+a search for effects from supersymmetric models
+in data from the \atlas\ detector
+on the Large Hadron Collider (LHC)~\cite{
+atlas2022searches,
+atlas2008experiment,
+lhc2008machine
+},
+we first cover some relevant facts about the design and motivations
+behind \atlas.
+
+The \atlas\ detector was not delivered in a preordained form.
+It was designed with intent to provide maximally useful and novel data within
+real-world cost budgets.
+Within general purpose utilities, \atlas's designs aimed to construct a
+detector that could
+\begin{itemize}
+\item find the Higgs boson (which was predicted but yet unseen),
+\item precisely measure known massive objects ($W$, $Z$, and $t$),
+\item discover new (supersymmetric) particles~\cite{atlas1999design1}.
+\end{itemize}
+We now review these three goals with the subsequent news from \atlas's
+operations.
+
+\paragraph{Find the Higgs boson.}
+Prior to \atlas's construction, the Higgs boson mass (and existence) was
+poorly known, but loosely constrained by results like
+$m(H) \gtrsim 90\,\eV[G]$ from searches at the
+Large Electron–Positron Collider (LEP),
+$m(H) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
+and $m(H) \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
+atlas1999design2,
+ghinculov1998perturb,
+lep1999ewk
+}.
+
+The Higgs boson couples to an extraordinary variety of fundamental particles.
+For its discovery alone, \atlas's designers achieved sensitivity across the
+entire $100$ to $1000\,\eV[G]$ window by targetting decays including
+$h \to b\bar b$,
+$h \to \gamma\gamma$,
+$h \to ZZ^* \to 4\ell$, and
+$h \to ZZ \to \ell\ell\nu\nu$;
+these channels demand precise measurements of leptons, photons, and
+$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
+transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
+at the very highest masses.
+
+Both \atlas\ and \cms\ have now found a Higgs boson candidate at
+$m(h) = 125\,\eV[G]$~\cite{
+atlas2012higgs,
+atlas2012combined,
+cms2012higgs
+}
+and confirmed all of its tested Standard Model properties~\cite{
+combined2016higgs,
+atlas2022ten,
+cms2022ten
+}
+such as its production mechanisms, decay rates,
+spin, and parity~\cite{
+HIGG-2013-01,
+HIGG-2013-17,
+HIGG-2014-06
+}.
+
+\paragraph{Precisely measure known massive objects.}
+Other than the Higgs boson, the heavy Standard Model resonances are the
+$W$ and $Z$ bosons and the top quark.
+All of these are produced at high rates in LHC proton-proton collisions,
+and present opportunities for \atlas\ to measure their properties
+(such as masses and production cross-sections) with unprecedented precision.
+
+Sensitivity to hadronic momenta is crucial,
+both directly to measure heavy hadronic decays
+(like $t \to bq\bar{q}$),
+and indirectly for the missing transverse momenta from semi-invisible
+leptonic decays (like $W \to \ell\nu$).
+Since individual hadron identities are less relevant to these measurements,%
+\footnote{%
+Rare decays like
+$W^\pm \to \pi^\pm \gamma$ or $W^\pm \to \pi^\pm \pi^+ \pi^-$
+could give clean signals, but are not competitive due to their
+tiny production rates and difficult identification~\cite{
+cdf1996search,
+mangano2014wpiy,
+cms2021wpiy
+}.
+Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but do not
+require hadron identification~\cite{
+CDF:2009mbf,
+CMS:2016ixg,
+ATLAS:2022jbw
+}.%
+}
+it makes sense for \atlas's design to forfeit hadron identification systems
+in favour of precise calorimetry with total angular coverage.
+
+\atlas\ has now produced a plethora of competitively precise measurements
+of the heavy Standard Model~\cite{atlas2021summarysm},
+including $m(W)$~\cite{atlas2018wmass} and
+$m(t)$~\cite{atlas2022symmarytop, atlas2019topmass, TOPQ-2015-03},
+but also details of differential cross-sections~\cite{
+STDM-2016-11,
+STDM-2016-14,
+TOPQ-2018-15,
+TOPQ-2016-10
+}.
+Rare processes have also been studied in new levels of detail,
+including
+electroweak diboson ($WW/WZ/ZZ$ etc.)~\cite{
+STDM-2015-21,
+STDM-2015-23,
+STDM-2017-09
+}
+and triboson~\cite{
+STDM-2016-06,
+STDM-2017-22,
+STDM-2019-09
+},
+and rare top processes
+($\ttbar Z$, $\ttbar W$ etc.)~\cite{
+TOPQ-2013-05,
+TOPQ-2018-01,
+TOPQ-2020-03
+},
+These continue to not invalidate the Standard Model, and are aided in their
+sensitivity by the lack of new phenomena,
+which could have overpowered these rare signals.
+
+\paragraph{Discover new (supersymmetric) particles.}
+High-energy proton-proton collisions at the LHC therefore offer sensitivity
+to new particles up to multi-TeV mass scales.
+Heavy new particles are anticipated to either decay to known Standard Model
+objects, or to be invisible in \atlas.
+Designs that can discover such new particles are therefore neatly aligned with
+our other goals of refining the current Standard Model
+and to discovering the Higgs boson at any mass.
+Visible decay products should already be caught by \atlas's abilities
+to measure the Higgs boson and other Standard Model particles,
+and the prospect of new invisible particles reinforces the need for precise
+measurements of the missing transverse momentum.
+
+Indeed, the main search presented in this thesis~\cite{atlas2022searches}
+joins nicely with the pre-existing needs for sensitivity to signatures
+including
+$h \to b\bar b$
+and
+$ZZ \to \ell\ell\nu\nu$.
+
+Dozens of searches for new phenomena have now been conducted in \atlas\ data,
+with~\cite{
+ATL-PHYS-PUB-2022-013
+}
+and without~\cite{
+ATL-PHYS-PUB-2022-007,
+ATL-PHYS-PUB-2022-012,
+ATL-PHYS-PUB-2022-036,
+ATL-PHYS-PUB-2022-034
+} supersymmetry,
+and show no convincing evidence for effects beyond the Standard Model.
+This is not a failure. It is a success.
+Our goal in science is not to discover something that is not there,
+but to learn more about nature.
+Finding no tangible new particles in LHC conditions is itself a discovery.
+
+\section{Design}
+Atlas is a Titan.
+\atlas\ is the largest detector experiment at the LHC,
+and alongside its biggest neighbours
+\cms~\cite{cms2008experiment},
+\lhcb~\cite{lhcb2008experiment},
+and ALICE~\cite{alice2008experiment}
+it is a multi-kilotonne beast,
+and a wonder of scientific engineering~\cite{
+atlas1994proposal,
+atlas2008experiment,
+atlas1999design1,
+atlas1999design2
+}.
+
+Our search paper~\cite{atlas2022searches} contains concise and standard
+descriptions of the \atlas\ detector and its relevant components.
+Through its many inexact reproductions in past \atlas\ publications, the text
+of these descriptions has become so highly refined that attempts to improve
+it myself would probably be in vain.
+Instead, this section quotes our published text with explanatory commentary.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
+\caption[
+A central slice of \atlas's design
+]{%
+A central slice of \atlas's design, reproduced from~\cite{atlas1999design1,
+persint2014manual},
+including a simulated event that interacts with several sensitive subsystems.
+}
+\label{fig:atlas_cutaway}
+\end{figure}
+
+\begin{quote}
+``%
+The ATLAS detector~\cite{atlas2008experiment} is a multipurpose particle
+detector with a forward--backward symmetric cylindrical geometry and a near
+$4\pi$ coverage in solid angle.%
+\footnote{%
+``ATLAS uses a right-handed coordinate system with its origin at the nominal
+interaction point (IP) in the centre of the detector and the $z$-axis along the
+beam pipe.
+The $x$-axis points from the IP to the centre of the LHC ring, and the $y$-axis
+points upwards.
+Cylindrical coordinates $(r,\phi)$ are used in the transverse plane, $\phi$
+being the azimuthal angle around the $z$-axis.
+The pseudorapidity is defined in terms of the polar angle $\theta$ as
+$\eta = -\ln \tan(\theta/2)$.
+Angular distance is measured in units of
+$\Delta R \equiv \sqrt{(\Delta\eta)^{2} + (\Delta\phi)^{2}}$.%
+''\footnotemark~\cite{atlas2022searches}%
+}
+It consists of an inner tracking detector surrounded by a thin superconducting
+solenoid providing a $2\,\mathrm{T}$ axial magnetic field, electromagnetic and
+hadronic calorimeters, and a muon spectrometer.%
+''~\cite{atlas2022searches}
+\end{quote}
+To illustrate this description, a vertical slice of \atlas\ is displayed in
+Figure~\ref{fig:atlas_cutaway}, and each subsystem is described in more detail
+in a subsection below:
+the inner tracking detector in Section~\ref{sec:atlas_inner},
+the electromagnetic and hadronic calorimeter systems in
+Section~\ref{sec:atlas_calo},
+and
+the chambers of the muon spectrometer in Section~\ref{sec:atlas_muon}.
+The final connection to the data analysed in this thesis is made through the
+trigger system which is described in Section~\ref{sec:atlas_trigger},
+and details of the data themselves are given in
+Section~\ref{sec:atlas_data}
+
+\footnotetext{%
+This footnote appears in every \atlas\ publication, but was perfected
+in~\cite{lester2013heffalon}.
+True to form, internal \atlas\ guidelines ask that
+``If you think you can improve on it, please consult with the
+Publications Committee''~\cite{atlas_coordinate, atlas_footnote}.
+}
+
+\section{Inner detector}
+\label{sec:atlas_inner}
+The inner tracking detectors begin \atlas's measurements of new collisions
+by precisely tracing the paths taken by energetic charged particles
+as they curve through the solenoidal magnetic field.
+Track curvature provides momentum information,
+tracks' associations into collision vertices helps to separate the multiple
+scatterings per crossing (pile-up),
+and their associations into displaced vertices help to identify
+decays.
+\begin{quote}
+``%
+The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
+It consists of silicon pixel, silicon microstrip, and transition radiation
+tracking detectors.
+An additional layer of silicon pixels, the insertable
+B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
+''~\cite{atlas2022searches}
+\end{quote}
+This insertable B-layer is a silicon pixel detector that begins just
+$3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
+present and future high LHC luminosities,
+particularly for sensitivity to particles whose lifetimes are just right to
+decay inside the tracking system;
+these `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
+maybe new species.
+
+Following radially outwards, the silicon pixel and silicon microstrip%
+\footnote{%
+Formerly known as the SemiConductor Tracker~\cite{atlas1999design1},
+the silicon microstrip system retains the acronym
+SCT~\cite{atlas2008experiment}.%
+}
+detectors (which use silicon technology)
+and the transition radiation tracking detectors (which uses `straw' tubes
+gas tubes).
+
+As for the whole \atlas\ detector, complete $\phi$ coverage and a large $\eta$
+range minimizes leakage.
+Towards \atlas's design goals, this inner tracking detector crucially provides
+measurements of electrons, B-hadrons, and contributes to the reconstruction of
+missing transverse momentum.
+
+\section{Calorimeters}
+\label{sec:atlas_calo}
+Inelastic collisions cause energetic particles to scatter, distributing their
+energy into macroscopic particle showers.
+Calorimetry uses this principle to measure energies
+by interleaving dense absorbing materials (which initiate showers)
+with sensitive materials that convert showers into electrical signals.
+In \atlas,
+\begin{quote}
+``%
+Lead/liquid-argon (LAr) sampling calorimeters provide electromagnetic (EM)
+energy measurements with high granularity.
+A steel/scintillator-tile hadron calorimeter covers the central pseudorapidity
+range ($|\eta| < 1.7$).
+The end-cap and forward regions are instrumented with LAr calorimeters for both
+the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
+''~\cite{atlas2022searches}
+\end{quote}
+The electromagnetic calorimeters provide located electron/photon energy
+measurements,
+and the hadronic calorimeters subsequently aim to stop and measure hadronic
+activity for later analysis in jet clustering.
+Maximal angular coverage by the calorimeters further facilitates missing
+transverse momentum estimates.
+
+Neither `A' in \atlas\ stands for Argon, but perhaps one should.
+Liquid-argon is the chosen active material in the electromagnetic and forward
+calorimeters for its robust, linear response in converting showers to
+electrical signals for readout by electrodes.
+Liquid argon is poured as a liquid to fill the gaps between the various
+metallic absorbers, which are folded into zigzag ``accordion'' shapes that
+optimize uniformity around the detector.
+The steel/scintillator hadronic tile calorimeter is a
+``simple and cost effective''~\cite{atlas1996tile} design, which uses
+photomultiplier tubes to convert light from its polystyrene scintillators into
+electrical signals.
+Particles lose some energy before reaching the calorimeters; a thin
+liquid-argon pre-sampler is installed before the central EM calorimeter
+to regain such loss.
+
+These calorimeter systems achieve attenuations of at least ten interaction
+lengths throughout their $|\eta| < 4.9$ range;
+by maximizing this attenuation, the design minimizes noise from hadronic
+`punch-through' of non-muons into the muon system.
+
+\section{Muon system}
+\label{sec:atlas_muon}
+Muons penetrate matter well because of their large mass and non-interaction
+with the strong nuclear force.
+\atlas\ doesn't try to stop energetic muons;
+above around $3\,\eV[G]$, they pass right through the calorimeters, and are
+instead measured by their curved tracks in a toroidal magnetic field that
+envelopes dedicated detectors on \atlas's periphery.
+\begin{quote}
+``%
+The muon spectrometer surrounds the calorimeters and is based on three large
+superconducting air-core toroidal magnets with eight coils each.
+The field integral of the toroids ranges between $2.0$ and
+$6.0\,\mathrm{T\kern-0.15ex m}$ across most of the detector.
+The muon spectrometer includes a system of precision chambers for tracking and
+fast detectors for triggering.%
+''~\cite{atlas2022searches}
+\end{quote}
+Muon subsystems come in four varieties, all of which operate by detecting
+charges scattered by the muons as they fly through gaseous media.
+Together, these systems track the muons' paths as the magnets bend them in the
+longitudinal plane; these toroidal magnets generate fields outside the tile
+calorimeter around the central barrel and in the end-caps as illustrated in
+Figure~\ref{fig:atlas_magnets}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
+\caption[
+Layout of \atlas's toroidal magnets around the tile calorimeter and inner
+solenoid
+]{%
+Layout of \atlas's toroidal magnets around the tile calorimeter and inner
+solenoid, including the eight barrel and end-cap coils,
+reproduced from~\cite{atlas2008experiment}.
+The \atlas\ muon spectrometer measures muons' tracks as they bend through
+this peripheral magnetic field.
+}
+\label{fig:atlas_magnets}
+\end{figure}
+
+Measurements of the tracks' curvature yield estimates of the muons' momenta;
+although extremely energetic muons will have straight tracks,
+\atlas's construction and calibration achieves a high precision, which
+increases to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
+momenta~\cite{
+atlas1994proposal,
+atlas2008experiment,
+ATL-PHYS-PUB-2015-037
+}.
+
+\section{Trigger}
+\label{sec:atlas_trigger}
+The LHC's $25\,\mathrm{ns}$ time difference between proton bunches
+(bunch spacing) is so fast that light-speed particles ejected from one crossing
+are still zipping through \atlas's $12.5\,\textrm{m}$ radius when the next
+collision arrives.
+This extreme engineering is necessary to achieve the integrated luminosities we
+need to see rare effects, but to avoid becoming overwhelmed
+by data it also demands a strict data reduction mechanism.
+This mechanism is the trigger:
+\begin{quote}
+``%
+A two-level trigger system is used to select events.
+The first-level trigger is implemented in hardware and uses a subset of the
+detector information to accept events at a rate below $100\,\mathrm{kHz}$.
+This is followed by a software-based trigger that reduces the accepted event
+rate to $1\,\mathrm{kHz}$ on average depending on the data-taking conditions.%
+''~\cite{atlas2022searches}
+\end{quote}
+The first-level (L1) trigger uses its fast preliminary information to accept
+interesting event and mark their regions of interest.
+Complete data in these regions of interest are passed to the second-level
+(high-level trigger, HLT) to inform its decision of whether to accept each
+event.
+Events accepted by the HLT are recorded for future analysis~\cite{
+atlas2016trigger,
+atlas2008experiment
+}.
+
+\section{Data}
+\label{sec:atlas_data}
+Raw binary data are hard for us humans to interpret.
+Before use in publishable analysis projects, software systems first mutate
+those data into a workable form, of clean events with calibrated and
+interpretable event variables.
+\begin{quote}
+``%
+An extensive software suite~\cite{ATL-SOFT-PUB-2021-001} is used in the
+reconstruction and analysis of real and simulated data, in detector operations,
+and in the trigger and data acquisition systems of the experiment.%
+''~\cite{atlas2022searches}
+\end{quote}
+Prediction is important in data analysis.
+This software suite helps to generate predictive simulations, and provides
+realism by passing both the real and simulated data through the same processing
+pipelines~\cite{SOFT-2010-01, geant2003}.
+Software models of the \atlas\ detector itself are refined with the results from
+many calibration studies and in-situ measurements~\cite{atlas2008experiment},
+
+The search presented in this thesis uses data from proton-proton collisions at
+$\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
+LHC, in which \atlas\ collected data in the years
+$2015\textrm{--}2018$~\cite{atlas2022searches}.
+This second run has various refinements over the Run~1, including beam energies
+increasing from $\sqrt{s} = 7\textrm{ or }8\,\eV[T]$, and bunch spacings
+halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.
+
+Data collected under anomalous detector operations are hard to model.
+Lists are maintained indicating which datasets arise from `good', normal
+detector operations, and these lists are commonly employed to select only clean
+subsets of good data for analysis projects.
+Testament to the quality of \atlas's operations, this cleaning removes less
+than $5\%$ of the Run~2 data~\cite{
+DAPR-2018-01,
+Golling:2011zy
+}.
+
+% closing comment?
+Now that we have predictive models, in the Standard Model and supersymmetric
+alternatives,
+and interpretable data, from \atlas's measurements of LHC collisions,
+we are ready to begin to squeeze some useful research from those data.
+
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-09-28 main/chapter_theory.tex
first proofread of summaires section; add linking paragraph
split context back out into theory and exp

--- 2022-05-16 main/chapter_theory.tex
@@
 \qitem{%
-The idea that theorems follow from the postulates does not correspond to simple
-observation.
-If the Pythagorean theorem were found to not follow from the postulates, we
-would again search for a way to alter the postulates until it was true.
-Euclid's postulates came from the Pythagorean theorem, not the other way.}%
-{Richard Hamming~\cite{hamming1980unreasonable}}
-\qitem{%
-One geometry can not be more true than another; it can only be more convenient.
-Geometry is not true, it is advantageous.}%
-{Robert M. Pirsig~\cite{pirsig1999zen}}
+if observing outer space gives us a view of the past, observing inner space
+would surely give us a glimpse into the future - would be interesting if NASA
+made a telescope for that%
+}%
+{Ken~M,
+\textit{Comment: Yahoo! News},
+2012~\cite{kenm2012inner}}
 \end{epigraphs}
@@
 
-% Standard Model past present future
+\noindent Our best current understanding of nature's fundamental objects in
+their observable behaviours as particles is encoded in the Standard Model,
+which uses solid mathematical foundations to accurately describe the physics
+of particle interactions across a broad domain.
+But that domain is bounded, so the Standard Model is not an ultimate theory of
+everything, and work continues to expand that boundary on both theoretical and
+experimental fronts.
+The current Standard Model is described in Section~\ref{sec:context_sm},
+along with its limitations and possible future directions.
 
-Hello
-
-\section{Particles}
+Supersymmetry an idea that has been applied to build plausible alternatives
+to the Standard Model which may explain future experimental discoveries.
+Since the main work of this thesis is a search for effects of supersymmetric
+phenomena, supersymmetry and the specific supersymmetric models that we study
+are presented in Section~\ref{sec:context_susy}.
 
 \section{The Standard Model}
+\label{sec:context_sm}
+The current Standard Model features fermions
+(of four types in three generations),
+which interact through forces mediated by vector bosons
+and the scalar Higgs boson.
+The strong force is mediated by
+Massless gluons $g$ mediate the strong force,
+the massless photon $\gamma$ mediates the electromagnetic force,
+and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
+These fundamental particles of the Standard Model are summarized in
+Table~\ref{tab:context_particles_sm}.
+
+As a gauge field theory, the Standard Model has the symmetry group
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
+meaning that its lagrangian is invariant under local operations of that group
+on its fields in certain representations.
+
+Gluons in this model correspond to the eight generators of $SU\!(3)_C$, whose
+subscript $_C$ indicates that they operate on colour-charge triplets of quarks.
+
+Electromagnetic and weak forces are unified into the electroweak
+interactions with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
+$SU\!(2)_L$ and $U\!(1)$ symmetries respectively.
+These would be massless if the symmetry were respected, but it is
+`spontaneously' broken with the Higgs mechanism~\cite{
+higgs1964broken,
+englert1964broken
+}%
+\footnote{%
+Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
+Peter~Higgs has reportedly suggested
+``ABEGHHKtH mechanism''~\cite{close2011infinity}
+to fully credit other associated work.%
+}
+which adds four scalar fields (as a complex doublet) with a potential that
+moves the ground state to a gauge-\emph{asymmetric} configuration.
+Around this new ground state, linear combinations of the electroweak fields
+acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
+four scalar fields to yield the three weak bosons and the photon:
+$W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
+$\gamma$.
+The fourth scalar is the Higgs boson $h$~\cite{
+glashow1959renorm,
+weinberg1967model,
+salam1959weak,
+rubakov1999classical,
+cottingham2007greenwood
+}.
+
+Fermion masses arise through their Yukawa interactions with the Higgs field.
+The subscript $_L$ on $SU\!(2)_L$ indicates that the weak interaction only
+couples to fermions in left-chiral doublets: $(\nu_e, e)_L$,
+$(u, d)_L$ and their repeats across the three generations.
+In each generational trio (such as $u$, $c$, $t$), all three fermions share the
+same quantum numbers, but have different masses.
+A second generational difference occurs in quarks and neutrinos through the
+difference between their weak flavour eigenstates and the mass eigenstates of
+their free propagation;
+flavour and mass eigenstates are linearly related through unitary matrices:
+the CKM matrix for the quarks~\cite{
+cabibbo1963unitary,
+kabayasji1973cpv
+}
+and the PMNS matrix for the neutrinos~\cite{
+maki1962remarks,
+thomson2013modern
+}.
+
+% using https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\centering
+\begin{tabular}{rrc}
+\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
+                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
+                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
+                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
+\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
+                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
+\end{tabular}
+\caption[Particles of the Standard Model]{%
+Particles of the Standard Model.
+All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
+spin-$0$ Higgs boson $h$.
+Fermions come in left- and right-chiralities, although the existence of
+right-chiral neutrinos is not strictly confirmed.%
+}
+\label{tab:context_particles_sm}
+\end{table}
+
+Early versions of the Standard Model had only two fermion generations and
+no neutrino masses~\cite{
+wells2020discovery,
+bjorken1985november
+},
+but the third generation and non-zero neutrino masses are now observed;
+for example, the third generation is shown by charge-parity $CP$ asymmetry in
+kaons~\cite{
+cronin1964evidence,
+kabayasji1973cpv
+},
+precise measurements of the $Z$ boson~\cite{
+lep2006precision
+},
+and direct observations~\cite{
+perl1977evidence,
+herb1977observation,
+abachi1995observation
+};
+neutrino masses are also necessary to explain their observed flavour
+oscillations~\cite{
+kamiokande1998measurement,
+superk1998evidence,
+superk1999measurement,
+lsnd1998evidence
+}.
+
+Various theories can describe neutrino masses in the Standard Model, and it is
+currently uncertain which form best describes nature;
+this is both a limitation of the current Standard Model and an
+opportunity for future refinement.
+With future results, the Standard Model will continue to improve in its
+correspondence with nature, and we discuss some such opportunities
+for improvement in the next section.
+
+\section{Opportunities for future Standard Models}
+The Standard Model works very well within its domain.
+This discussion therefore looks towards where an altered Standard Model could
+improve its current abilities or to expand its generality to describe more of
+the universe.
+
+When `nobody knows' the answer to a riddle, the status lies somewhere between
+two extremes:
+either many answers are satisfactory with no clear best,
+or no satisfactory answers are yet known.
+We distinguish these two cases with the labels
+\emph{uncertain} and \emph{unknown} respectively:
+the model of neutrino masses is mostly \emph{uncertain}, but
+the reconciliation quantum field theory with general relativity is
+mostly \emph{unknown}.
+
+One real problem with the Standard Model is that algorithms to compute its
+general consequences are unknown, since practical perturbative and discrete
+lattice methods are limited.
+Imprecise modelling weakens experimental results (anomalous or not),
+but does not present an opportunity because it is common to the quantum field
+theory that is used by all known viable models.
+This section briefly reviews some other directions of positive change for
+future models.
+
+\paragraph{Neutrino masses.}
+Neutrinos could be Dirac fermions with Yukawa masses and boring right-chiral
+fields.
+They could also have Majorana mass terms, for which the seesaw mechanism can
+simultaneously generate the observed light neutrinos with very heavy `sterile'
+neutrinos.
+Or their masses could come from other operators without needing the
+right-chiral fields~\cite{
+thomson2013modern,
+chala2021neutrino,
+wells2020discovery
+}.
+This answer is uncertain, and new results from theory or experiment that
+pinpoint a correct model could improve the Standard Model in the neutrino
+domain.
+
+\paragraph{Gravitation and cosmology.}
+Gravity does not feature in the Standard Model, nor does it measurably impact
+particle physics experiments.
+Although gravitation is in another domain, the particle content of the
+Standard Model does impact cosmology.
+
+The universe contains dark matter which interacts with gravity but not much
+with other forces; it is observed through galactic rotation curves,
+gravitational lensing, and cosmic microwave background data,
+but its composition is uncertain~\cite{
+begeman1991rotation,
+garrett2010dark,
+planck2020results
+}.
+Dark matter candidates include new particles such as
+sterile neutrinos~\cite{boyarsky2019sterile},
+Weakly Interacting Massive Particles (WIMPs)~\cite{jungman1996wimp},
+or
+axion-like particles~\cite{kim2008axions},
+but also macroscopic objects such as
+primordial black holes~\cite{bernard2019primordial}.
+Dark matter may well comprise various of different components, and it would
+be a success for future particle physics theories to pinpoint necessary dark
+matter particles.
+
+Other mysteries of cosmology include:
+the small positive cosmological constant,
+the inflation of the early universe,
+and baryogenesis~\cite{
+wells2020discovery,
+riess1998observational
+}.
+Solutions to these mysteries all lie somewhere between uncertain and unknown,
+but may not be entirely out of reach;
+for example, the Higgs field could plausibly have driven
+inflation~\cite{bezrukov2008higgs},
+and the Standard Model does slightly violate $C$ and $CP$, both of which
+are required for baryogenesis~\cite{sakharov1991re}.
+
+\paragraph{Origins.}
+Why does the Standard Model have its structure?
+Why is it configured with the observed values of our parameters?
+Why do we even live in $3+1$ dimensional spacetime with quantum mechanics?
+Answers to these questions are largely unknown, but would be
+attractive features for a fundamental theory.
+
+Certain Standard Model structures are quite general, and are parametrized such
+that parts can be smoothly disabled if they do not describe nature.
+For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$
+whose non-zero values would induce $\mathrm{CP}$ violation in the strong
+sector.
+Observations find that $\theta_\mathrm{CP}$ cannot be far from zero, and one
+attractive feature of axion-like models is their requirement that
+$\theta_\mathrm{CP} = 0$~\cite{
+kim2008axions,
+thomson2013modern,
+martin2017particle
+}.
+
+Justified constraints on the model could have led to better predictions of
+past data and might extrapolate to new phenomena through their deeper
+understanding of nature.
+Attempts to relate empirical parameters~\cite{
+sato1979ratio,
+beg1982gauge,
+koide1983fermion,
+denterria2012gaussian
+}
+or to derive mathematical structures~\cite{
+goyal2010origin,
+skilling2021abc,
+axioms1010038
+}
+are curious, but will remain speculative until they provide novel benefits.
+
+\paragraph{Aesthetics.}
+In the search for improved models, heuristic criteria can help to identify
+appealing candidates.
+Two such criteria are `fine-tuning' and `naturalness', both of which are
+aesthetic judgements about dimensionless model parameters.
+Naturalness is opinion; only nature decides what is really natural,
+but these criteria can still guide the discourse of our scientific searches
+towards useful directions~\cite{wells2020discovery}.
+
+Fine-tuning criteria dislike tight constraints, and naturalness contains
+various ideas of which one is that nice numbers should have moderate orders of
+magnitude~\cite{giudice2008naturally}.
+For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned by an
+observation of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
+large $11$ in its exponent also defies naturalness.
+
+Another example is the `hierarchy' problem of the Higgs boson mass $m(h)$,
+which notes that new particles at a heavy mass scale $\Lambda$ induce changes
+of $\Delta m(h)^2 = \mathcal{O}(\Lambda^2)$,
+through direct or indirect couplings.
+With any particles near the Planck scale $\Lambda \sim 10^{18}\,\eV[G]$,
+implementing the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires
+finely-tuned cancellations between positive and negative contributions.
+New fermions subtract from $m(h)$ and new scalars add by similar amounts,
+so new models can implement this fine-tuning by inserting only matched sets of
+fermions and bosons.
+This solution is the strategy of supersymmetric models.
 
 \section{Supersymmetric models}
+\label{sec:context_susy}
+Supersymmetry is the concept of symmetry under the operation of exchanging
+bosonic and fermionic states, and supersymmetric models use supersymmetry in
+their construction.
+Nature is not totally supersymmetric.
+If it were, then there would be bosonic `superpartners' for each of the known
+fermions (and vice versa) with the same masses and charges, and those
+`sparticles' would have been seen long ago.
+But supersymmetry might still participate in a broken condition if the
+sparticles are hidden at larger masses.
 
-\subsection{Minimal supersymmetric models}
+Any new model must include the observational facts of the Standard Model,
+and the Minimal Supersymmetric Standard Model (MSSM) satisfies this while
+adding a school of new particles,
+which is listed in Table~\ref{tab:context_particles_mssm}.
 
-\subsection{Guage Mediated Supersymmetry Breaking}
+% fermion boson multiplets
+The MSSM matches each standard fermion to a pair of scalar
+`sfermions' (left and right)
+and each vector gauge field to a `gaugino' fermion
+($\tilde g$, $\tilde W^1$, $\tilde W^2$, $\tilde W^3$, and $\tilde B$).
+The Higgs field does not translate so directly, but is replaced with two sets
+of complex scalar fields
+($H^+_u$, $H^0_u$ and $H^0_d$, $H^\pm_d$) with
+gaugino superpartners
+($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
+(All sparticles are marked with tildes $\tilde{~}$.)
+These fields mix into mass eigenstates which current data do not entirely
+determine:
+the gaugino mixtures are the charged charginos $C^\pm_i$ and
+neutral neutralinos $N_i$
+(also labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches}),
+and each squark and slepton has two states that mix the left and right fields.
+From the four complex scalars, three degrees of freedom are eaten by the
+weak bosons to leave the reborn Standard Model Higgs boson $h$
+and four new scalars~\cite{
+martin2016primer
+}.
+
+% made using:
+% https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\centering
+\begin{tabular}{rrc}
+\multirow{5}{*}{bosons}   & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[1ex]
+                          & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
+                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
+                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
+                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[3ex]
+\multirow{3}{*}{fermions} & gluinos                       & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
+                          & \multirow{2}{*}{gauginos}     & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
+                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
+\end{tabular}
+\caption[%
+Particles of the Minimal Supersymmetric Standard Model%
+]{%
+Particles of the Minimal Supersymmetric Standard Model
+in their mass eigenstates~\cite{martin2016primer}, in addition to the
+Standard Model particles of Table~\ref{tab:context_particles_sm}.
+The standard Higgs boson ($h$) is repeated here with its new pals.
+Sparticles are marked with tildes $\tilde{~}$.
+All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
+}
+\label{tab:context_particles_mssm}
+\end{table}
+
+Various mechanisms could spontaneously break supersymmetry.
+Gravitational ideas exist, but one model considered in this thesis uses a
+Gauge-Mediated Supersymmetry Breaking (GMSB) mechanism,
+which adds yet more elusive fields:
+a `secluded' sector which breaks the symmetry, and
+a `messenger' sector which
+communicates that breaking effect back to the observable sector.
+Such models also add a spin-$3/2$ `gravitino' ($\gravitino$),
+which should be light and whose helicity $\pm1/2$ components would interact
+in collider physics~\cite{
+giudice1999gmsb,
+dine1982pheno,
+martin2016primer
+}.
+
+All supersymmetric models considered in this thesis conserve $R$-parity
+\begin{equation}
+P_R = (-1)^{3(B - L) + 2s}
+\end{equation}
+(with baryon number $B$, lepton number $L$ and spin $s$)
+in their interactions~\cite{
+farrar1978rparity,
+martin2016primer
+}.
+Conservation of $R$-parity helps by preventing proton decays,
+but also makes the Lightest Supersymmetric Particle (LSP) a plausible
+WIMP dark matter candidate since it cannot decay to purely Standard Model
+particles.
+Providing this LSP dark matter candidate through a production mechanism
+that could be observable at the LHC is a strong motivating factor
+both for the supersymmetric models
+and for the experimental searches of this thesis.
+
+Supersymmetric modelling successfully predicted the lightest Higgs boson
+mass to be below $\lesssim 130\,\eV[G]$ long before its observation at
+$125\,\eV[G]$~\cite{
+espinosa1992higgs,
+espinosa1993upper,
+quiros1997constraints,
+wells2020discovery,
+atlas2012higgs,
+atlas2012combined,
+cms2012higgs
+}.
+This prediction, however, assumed a mass scale below $1\,\eV[T]$,
+and observations at the LHC are now limiting sparticle masses
+into super-$\eV[T]$ territory (in specific models).
+Such limits are also troublesome for the hierarchy problem, because large
+superpartner masses introduce fine-tuning problems of their
+own~\cite{martin2016primer}.
+Despite its current lack of observational evidence, however,
+supersymmetry remains a tool for building theoretically plausible
+models, and might help us to understand future discoveries.
+
+% this comment activates enlarged spacing for the final paragraph lol

### 2022-09-28 main/draft.tex
move table; write some philosophy and neutrinos
more theory tweaks
split context back out into theory and exp

@@
 % 1974~\cite{pirsig1999zen}}
+
+So are Newton's laws in their domain; both models are useful in different
+applications.
+
+Supersymmetry itself is not a model, but if our searches did hypothetically
+find evidence for any supersymmetric model, then that would be evidence for
+supersymmetry in nature.
+
+% for preamble
+
+\noindent Theory is needed to understand experimental data,
+and data are needed to improve theoretical models.
+To clarify the experimental work of this thesis, this chapter introduces its
+theoretical and experimental contexts;
+Section~\ref{sec:context_sm_and_beyond} presents the current culmination of
+theoretical particle physics in the Standard Model, its limitations, and its
+possible future directions including supersymmetric ideas, and
+Section~\ref{sec:context_atlas} then describes the \atlas\ detector, with whose
+data we hope refine the Standard Model into its improved future forms.

### 2022-09-28 main/chapter_2ljets.tex
cut out preface junk; fix more atlas's; minor tweaks

@@
 \end{singlespacing}
-
-\noindent The main messages that this thesis reports are my contributions to the recently
+\noindent
+The main messages that this thesis reports are my contributions to the recently
 published \atlas\ searches ``for new phenomena in events with two leptons,
@@
 
-Each trigger in \atlas' arsenal activates on a selection comprising some
+Each trigger in \atlas's arsenal activates on a selection comprising some
 combination of sufficiently hard objects.
@@
 Unlike lepton flavour tagging, which can very cleanly separate electrons from
-muons, jet tagging is a rather more noisy affair due to \atlas' limited
+muons, jet tagging is a rather more noisy affair due to \atlas's limited
 ability to reconstruct the individual hadrons within jets.
@@
 One exception, as discussed in Chapter~\ref{chapter:context}, is if the muon
-is hard enough that its curvature through \atlas' magnetic fields is
+is hard enough that its curvature through \atlas's magnetic fields is
 imprecisely reconstructed.
@@
 Higgs boson mass at around $125\,\eV[G]$, and is chosen not only to
-accommodate \atlas' noisy reconstruction of $b$-jets, but also to
+accommodate \atlas's noisy reconstruction of $b$-jets, but also to
 capture $Z\to bb$ decays, which become significant for signal models
@@
 it would be wasteful and unproductive to consider them seriously.
-I am grateful that \atlas' statistical conventions allow us to ignore these
+I am grateful that \atlas's statistical conventions allow us to ignore these
 problems.
@@
 large-radius jets with $R=0.8$, since these require less extreme boosts to
-capture their targets, and the focus from \atlas' performance operations is
+capture their targets, and the focus from \atlas's performance operations is
 therefore reasonable to understand these common use-cases first.
@@
 \end{singlespacing}
-
-\noindent From \atlas' large and previous Run~2 dataset, our $\twoljets$ searches
+\noindent
+From \atlas's large and precious Run~2 dataset, our $\twoljets$ searches
 have extracted a sharper image of the high energy processes that simultaneously
@@
 datasets and found similar conclusions.
-By providing refined and updated interpretations, we have advanced \atlas' SUSY
-research programme in this two-lepton plus jets plus $\met$ channel.
+By providing refined and updated interpretations, we have advanced \atlas's
+SUSY research programme in this two-lepton plus jets plus $\met$ channel.
 My focus has been on the $\twoljets$-electroweak search within this project,
@@
 explicitly declared in the main publication~\cite{atlas2022searches},
-and might otherwise have remained hidden in \atlas' internal documentation.
+and might otherwise have remained hidden in \atlas's internal documentation.
 
 Perhaps we \emph{should not} be restrained by crude statistical
-methods, but the fact is that we \emph{are} limited to immutable standard
-prescriptions.
+methods, but the fact is that we \emph{are} limited to immutable prescriptions.
 And by ignoring that truth, my designs cost this search project complications

### 2022-09-28 main/chapter_experiment.tex
cut out preface junk; fix more atlas's; minor tweaks

@@
 \end{singlespacing}
-
+\noindent
 Towards explaining the main result of this thesis,

### 2022-09-28 main/chapter_searches.tex
remove an ugly comma
cut out preface junk; fix more atlas's; minor tweaks

@@
 One way is to make it so simple that there are \emph{obviously} no deficiencies
+\\
 and the other way is to make it so complicated that there are no \emph{obvious}
@@
 \end{singlespacing}
-
-\noindent Our task in scientific data analysis is to distil useful messages from the
+\noindent
+Our aim in scientific data analysis is to distil useful messages from the
 data, and to report that message to the wider community.
@@
 \);
-since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$,
+since $\sigma_{ab}^2 = \widetilde{\sigma}_{ab}^2 + \sigma_a^2 \sigma_b^2$
 the propagation formula assumes at least one small relative error.
@@
 statistical research, and influences practices still followed in \atlas.
-We shall therefore review some features of frequentist statistics and its
-history.
+We therefore review some features of frequentist statistics and its history.
 % Would like something along the lines of:
@@
 \end{singlespacing}
-
-\noindent Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
-proportion of stuff within a larger whole, which when formalized as a ratio
-of measures and has broad applications.
-Frequentist theory confines itself to the special case where those measures
-are numbers of randomly occurring events, such that a frequentist
-probability is a ratio of frequencies:
+\noindent
+Probability as defined in Equation~\ref{eqn:searches_prob_ratio}
+has broad applications, but frequentist theory confines itself to the special
+case where those measures are numbers of randomly occurring events, such that
+a probability becomes a ratio of frequencies:
 \begin{equation}
@@
 \end{singlespacing}
-
-\noindent Data analysis in modern High Energy Physics does not exactly follow frequentist
+\noindent
+Data analysis in modern High Energy Physics does not exactly follow frequentist
 ideology, but is a hands-on, practical, and evolving ecosystem that has adapted

### 2022-09-28 main/chapter_theory.tex
add mssm mass sentence
cut out preface junk; fix more atlas's; minor tweaks

@@
 \qitem{%
-if observing outer space gives us a view of the past, observing inner space
-would surely give us a glimpse into the future - would be interesting if NASA
-made a telescope for that%
-}%
-{Ken~M,
-\textit{Comment: Yahoo! News},
-2012~\cite{kenm2012inner}}
+The idea that theorems follow from the postulates does not correspond to simple
+observation.
+If the Pythagorean theorem were found to not follow from the postulates, we
+would again search for a way to alter the postulates until it was true.
+Euclid's postulates came from the Pythagorean theorem, not the other way.}%
+{Richard~Hamming,
+\textit{The Unreasonable Effectiveness of Mathematics},
+1980~\cite{hamming1980unreasonable}}
 \end{epigraphs}
 \end{singlespacing}
-
-\noindent Our best current understanding of nature's fundamental objects in
-their observable behaviours as particles is encoded in the Standard Model,
-which uses solid mathematical foundations to accurately describe the physics
-of particle interactions across a broad domain.
-But that domain is bounded, so the Standard Model is not an ultimate theory of
-everything, and work continues to expand that boundary on both theoretical and
-experimental fronts.
-The current Standard Model is described in Section~\ref{sec:context_sm},
+\noindent
+The Standard Model encodes out best understanding of nature's fundamental
+objects in their behaviours as particles.
+From solid mathematical foundations, it accurately describe the physics
+of particle interactions across a broad domain,
+but that domain is bounded, so the Standard Model is not an ultimate theory of
+everything.
+Work continues to expand that boundary on both theoretical and experimental
+fronts.
+The current Standard Model is described in Section~\ref{sec:theory_sm},
 along with its limitations and possible future directions.
@@
 phenomena, supersymmetry and the specific supersymmetric models that we study
-are presented in Section~\ref{sec:context_susy}.
+are presented in Section~\ref{sec:theory_susy}.
 
 \section{The Standard Model}
-\label{sec:context_sm}
+\label{sec:theory_sm}
 The current Standard Model features fermions
@@
 These fundamental particles of the Standard Model are summarized in
-Table~\ref{tab:context_particles_sm}.
+Table~\ref{tab:theory_particles_sm}.
 
@@
 }
-\label{tab:context_particles_sm}
+\label{tab:theory_particles_sm}
 \end{table}
@@
 
-\paragraph{Neutrino masses.}
+\subsection{Neutrino masses}
 Neutrinos could be Dirac fermions with Yukawa masses and boring right-chiral
@@
 neutrinos.
-Or their masses could come from other operators without needing the
+Or their masses could come from other operators necessarily introducing the
 right-chiral fields~\cite{
@@
 }.
-This answer is uncertain, and new results from theory or experiment that
+The answer is uncertain, and new results from theory or experiment that
 pinpoint a correct model could improve the Standard Model in the neutrino
@@
 
-\paragraph{Gravitation and cosmology.}
+\subsection{Gravitation and cosmology}
 Gravity does not feature in the Standard Model, nor does it measurably impact
@@
 
-\paragraph{Origins.}
+\subsection{Origins}
 Why does the Standard Model have its structure?
@@
 
-\paragraph{Aesthetics.}
+\subsection{Aesthetics}
 In the search for improved models, heuristic criteria can help to identify
@@
 \section{Supersymmetric models}
-\label{sec:context_susy}
+\label{sec:theory_susy}
 Supersymmetry is the concept of symmetry under the operation of exchanging
@@
 adding a school of new particles,
-which is listed in Table~\ref{tab:context_particles_mssm}.
+which is listed in Table~\ref{tab:theory_particles_mssm}.
 
-% fermion boson multiplets
 The MSSM matches each standard fermion to a pair of scalar
@@
 
+Current observations do limit the masses and mixings of this new
+school of MSSM particles, but their values remain very uncertain and in need
+of new experimental constraints.
+
 % made using:
@@
 in their mass eigenstates~\cite{martin2016primer}, in addition to the
-Standard Model particles of Table~\ref{tab:context_particles_sm}.
+Standard Model particles of Table~\ref{tab:theory_particles_sm}.
 The standard Higgs boson ($h$) is repeated here with its new pals.
@@
 }
-\label{tab:context_particles_mssm}
+\label{tab:theory_particles_mssm}
 \end{table}
@@
 Supersymmetric modelling successfully predicted the lightest Higgs boson
-mass to be below $\lesssim 130\,\eV[G]$ long before its observation at
+mass to be $\lesssim 130\,\eV[G]$ long before its observation at
 $125\,\eV[G]$~\cite{

### 2022-09-28 main/draft.tex
cut out preface junk; fix more atlas's; minor tweaks

@@
 data we hope refine the Standard Model into its improved future forms.
+
+% far too much intro waffle
+
+\TODO{This does not belong in an introduction and will be moved.}
+
+\TODO{CUT THIS SHORT}
+
+\subsection{Testing symmetries in data}
+Data-driven tests of physical symmetries are a speciality of the Cambridge
+\atlas\ group.
+Each test considers a theoretically motivated symmetry that states that
+distributions of data should be identical when transformed under its operation.
+This means that it can be possible to disprove a symmetry by searching only
+for differences between data and transformed versions of those data, without
+necessarily requiring other theoretical inputs.
+
+% emus
+Charge symmetry of electron-muon, `emu', pairs is a key example among these
+tests.
+Theoretically, it proposes that, in the absence of biases, data
+from proton-proton collisions at the LHC would include equal distributions
+of $e^+\mu^-$ and $e^-\mu^+$ events~\cite{Lester:2016qdv}.
+Real data can have emu biases --- one clear (but minor) example is that our
+detectors are made of matter containing negatively charged electrons, not
+positrons, so high energy electrons positrons have subtly different rates of
+energy loss in that detector matter.
+However, \cite{Lester:2016qdv} argues that any biases at the LHC should
+conspire only to favour the rate of $e^-\mu^+$ over the rate of $e^+\mu^-$
+events.
+Therefore defining an asymmetry
+\begin{equation}
+A^\mathrm{emu} = \frac{
+\sigma(e^+\mu^- + X)
+}{
+\sigma(e^-\mu^+ + X)
+}
+,
+\end{equation}
+one can predict $A^\mathrm{emu} \leq 1$ from the Standard Model at the LHC.
+Negative values could be explained by experimental biases, but
+$A^\mathrm{emu} > 1$ measured in data might indicate new physics, since it and
+can be caused by models of new physics that might invoke R-parity violating
+supersymmetry~\cite{Lester:2016qdv} or leptoquarks~\cite{EXOT-2018-29}.
+After its extensive development~\cite{Brunt:2674708,Pacey:2747774},
+this test has been performed in \atlas\ data, and found no ``significant
+model-independent evidence''~\cite{EXOT-2018-29} for $A^\mathrm{emu} > 1$.
+
+% parity violation
+Parity symmetry, which changes the signs of all spatial coordinates
+$\vec x\rightarrow -\vec x$, is a second key example.
+Nature does violate parity symmetry, as first shown by the Wu
+experiment~\cite{PhysRev.105.1413}, in the weak sector, but that is not
+directly observable at the LHC due to its non-polarized beams and
+polarization-insensitive detectors~\cite{Lester:2019bso}.
+Unknown physical effects could, however, generate forms of parity asymmetry
+that could be seen at the LHC.
+Although the charge symmetry of the emu test has known sources of bias,
+no such biases are so clearly expected for parity.
+
+General parity symmetry can therefore be tested at the LHC by considering a
+parity-odd event variable and comparing its distribution to its parity-flipped
+version, which simply swaps its sign.
+% alpha from lester schott
+One such test has been performed with LHC data~\cite{Lester:2019bso}.
+This test used \cms\ open data and the parity-odd event variable
+\begin{equation}
+\label{eqn:cms_alpha}
+\alpha
+= \arcsin\!
+\left(
+\frac{
+\vec p^{\,j_1} \times \vec p^{\,j_2}
+}{
+|\vec p^{\,j_1} \times \vec p^{\,j_2}|
+}
+\cdot
+\frac{
+\vec p^{\,j_3}
+}{
+|\vec p^{\,j_3}|
+}
+\right)
+,
+\end{equation}
+in which $p^{\,j_i}$ is the momentum of the $i$th hardest jet in a multi-jet
+event,
+and did not find evidence for parity violation.
+But that does not prove that no parity-violating effects are occuring in these
+data, only that it did not appear in the distribution of $\alpha$.
+
+Can a single parity-odd event variable be sensitive to all norms of parity
+violation?
+The answer, it turns out, is `no' if that variable is also required to be
+continuous~\cite{lesterChiralMeasurements2021}.
+In this framework, therefore, a joint analysis of multiple variables is
+required for any sufficiently complicated data.
+Furthermore, it is not generally easy to construct such universally-sensitive
+sets of event variables~\cite{
+Gripaios:2020hya,
+lester2021lorentz
+}.
+
+Rather than beginning from universally sensitive event variables, an
+alternative strategy is to build candidate variables to the data
+themselves~\cite{lester2021stressed}.
+If done wrong, such an idea could simply reinforce variations due to simple
+statistical noise.
+It can, however, be formulated as a rigorous test if one employs standard
+Machine Learning practice and splits the data into independent training and
+testing sets.
+Variables learned (by function approximation algorithms such as
+Boosted Decision Trees~\cite{xgboost} or
+Neural Networks~\cite{MurphyKevinP.2012Mlap})
+from the testing set can correspond to hypotheses about possible asymmetries
+in the data distribution, and if those hypotheses make accurate predictions in
+independent testing data, then they are evidence against the symmetry.
+
+My contribution to this field is to adapt this idea of learning parity
+variables,
+\emph{which was initiated by Christopher~G.~Lester}~\cite{lester2021stressed},
+into a probabilistic description, in which the algorithms predict
+`which is real?' between real data an their parity-transformed
+copies~\cite{tombs2021which}.
+This approach derives an objective function
+(which is a special case of the binary
+cross-entropy~\cite{MurphyKevinP.2012Mlap}),
+from which learning algorithms
+can approximate the odds ratio between the data distribution and its
+parity-transformed version.
+Importantly, it also allows one to model inefficiencies or holes in the
+distribution, which exist in practice from kinematic selections and imperfect
+hardware, and incorporate those into the training and testing without biasing
+the result.
+We have also demonstrated this method on simulated multi-jet data in an
+\atlas-like detector~\cite{lester2022hunting}.
+By simulating parity-violating events from a Lorentz-violating extension
+to the Standard Model, we show that this method build variables that are much
+more sensitive than the $\alpha$ of Equation~\ref{eqn:cms_alpha} while
+retaining its other attractive symmetries by encoding them into our function
+architectures~\cite{lester2021stressed, tombs2021which}.
+
+This work demonstrates that the `which is real?' method is ready to test for
+parity violation in real LHC data.
+Much like the emu work, however, its realization would require great diligence
+to handle important considerations of systematic uncertainties and statistical
+interpretations before it could be published through \atlas.
+
+The `which is real?' method is general, and can test not only parity, but
+symmetry under any transformation of the data, irrespective of whether it is
+discrete or continuous.
+In the context of modern machine learning, it is an example of
+`self-supervised' learning, in which the training objective derives from the
+data and not from additional labels~\cite{
+Noroozi2016jigsaw,
+multitaskself2017,
+devlin2019bert
+}, and quite similar to the `CLIP' objective, which also attempts to identify
+a real examples from collections of alternatives~\cite{pmlr-v139-radford21a}.
+What is special in our work is our application for scientific self-supervised
+testing.

### 2022-09-28 main/frontmatter.tex
cut out preface junk; fix more atlas's; minor tweaks

@@
-\titlepage[of Homerton College]{%
+\titlepage[Homerton College]{%
 A dissertation submitted to the University of Cambridge\\
@@
 \begin{abstract}
-\atlas\ is a detector experiment at \cern. $14~\eV[T]$.
+\TODO{searches for new effects in atlas are a core reason for its operations}
+
+\TODO{this thesis presents such a search~\cite{atlas2022searches}}
+
+\TODO{after briefly introducing its theoretical and experimental context}
+
+\TODO{For given modelling of theory and experiment,
+data analysis procedures dominate, and they take a very particular form
+which we attempt to understand in its historical context
+}
+
+\TODO{
+The result is compatible with many other observations,
+that the Standard Model works very well at the LHC.
+}
+
+\TODO{supersymmetry is not falsifiable?}
+
 \end{abstract}
@@
 \begin{declaration}
-This dissertation is the result of my own work, except where explicit
+\noindent This dissertation is the result of my own work, except where explicit
 reference is made to the work of others, and has not been submitted
-for another qualification to this or any other university. This
-dissertation does not exceed the word limit for the respective Degree
-Committee.
+for another qualification to this or any other university.
+This dissertation does not exceed the word limit for the respective
+Degree Committee.
 \vspace*{1cm}
 \begin{flushright}
-Rupert Tombs
+Rupert~Tombs
 \end{flushright}
@@
 \begin{acknowledgements}
-Of the many people who deserve thanks, some are particularly prominent,
-such as my supervisor\dots
-\\
+<3
+% Holly
+% Dan
+% Chris
+
+% Friends who have come and gone from Homerton over the years
+
+% Sarah Williams
+% Ben Hooberman
+% Jon Long
+
+% Family, alive and dead.
+
 This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
@@
 
-This thesis describes my research on various hoozywotsists.%
-\footnote{test.}
+Data Intensive Science in High-Energy Physics
 
-\noindent
-For this example, I'll just mention Chapter~\ref{chapter:searches}.
-and Chapter~\ref{chapter:searches}..
+\TODO{Context of Standard Model theory, directions in which future developments
+could improve it, and supersymmetric models}
+Chapter~\ref{chapter:theory}
+
+\TODO{the design of the \atlas\ detector}
+Chapter~\ref{chapter:experiment}
+
+\TODO{theory and practice of data science, as it could be for clarity
+and how it is in our \atlas\ SUSY conventions}
+Chapter~\ref{chapter:searches}
+
+\TODO{the main event: the search with many of its gory details in addition
+to those presented publically in the paper~\cite{atlas2022searches}}
+Chapter~\ref{chapter:2ljets}
+
+\paragraph{Other work.}
+\TODO{Use some of the data analysis ideas of Chapter~\ref{chapter:searches}
+to build predictive models
+and test them to test for symmetries}
+\TODO{citations}
+
 \end{preface}

### 2022-09-28 main/chapter_searches.tex
fix quote linebreak; define SUSY in preface

@@
 One way is to make it so simple that there are \emph{obviously} no deficiencies
-\\
-and the other way is to make it so complicated that there are no \emph{obvious}
+and the\\
+other way is to make it so complicated that there are no \emph{obvious}
 deficiencies.%

### 2022-09-28 main/frontmatter.tex
fix quote linebreak; define SUSY in preface

@@
 could improve it, and supersymmetric models}
+supersymmery (SUSY)
 Chapter~\ref{chapter:theory}

### 2022-09-28 main/chapter_2ljets.tex
some overleaf tweaks

@@
 But there are exceptions, and exceptions to the exceptions.
-One exception, as discussed in Chapter~\ref{chapter:context}, is if the muon
+One exception, as discussed in Chapter~\ref{chapter:experiment}, is if the muon
 is hard enough that its curvature through \atlas's magnetic fields is

### 2022-09-28 main/chapter_experiment.tex
some overleaf tweaks

@@
 High-energy proton-proton collisions at the LHC therefore offer sensitivity
-to new particles up to multi-TeV mass scales.
+to new particles up to multi-$\eV[T]$ mass scales.
 Heavy new particles are anticipated to either decay to known Standard Model

### 2022-09-28 main/chapter_theory.tex
some overleaf tweaks

@@
 \noindent
-The Standard Model encodes out best understanding of nature's fundamental
+The Standard Model encodes our best understanding of nature's fundamental
 objects in their behaviours as particles.
-From solid mathematical foundations, it accurately describe the physics
+From solid mathematical foundations, it accurately describes the physics
 of particle interactions across a broad domain,
@@
 
-Supersymmetry an idea that has been applied to build plausible alternatives
+Supersymmetry is an idea that has been applied to build plausible alternatives
 to the Standard Model which may explain future experimental discoveries.
@@
 All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
-spin-$0$ Higgs boson $h$.
+spin-$0$ scalar Higgs boson $h$.
 Fermions come in left- and right-chiralities, although the existence of
@@
 the model of neutrino masses is mostly \emph{uncertain}, but
-the reconciliation quantum field theory with general relativity is
+the reconciliation of quantum field theory with general relativity is
 mostly \emph{unknown}.
@@
 neutrinos.
-Or their masses could come from other operators necessarily introducing the
+Or their masses could come from other operators without  introducing the
 right-chiral fields~\cite{
@@
 Fine-tuning criteria dislike tight constraints, and naturalness contains
-various ideas of which one is that nice numbers should have moderate orders of
+various ideas, of which one is that nice numbers should have moderate orders of
 magnitude~\cite{giudice2008naturally}.

### 2022-09-29 main/chapter_theory.tex
write abstract; quote tweaks; move higgs citation

@@
 Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
-Peter~Higgs has reportedly suggested
-``ABEGHHKtH mechanism''~\cite{close2011infinity}
-to fully credit other associated work.%
+Peter~Higgs has reportedly suggested ``ABEGHHKtH mechanism'' to fully credit
+associated work~\cite{close2011infinity}.%
 }

### 2022-09-29 main/frontmatter.tex
change from tabs to spaces; add frontmatter to contents with roman numbers
write abstract; quote tweaks; move higgs citation

@@
 \begin{abstract}
-\TODO{searches for new effects in atlas are a core reason for its operations}
+To search for new phenomena is an essential endeavour of experimental physics,
+and is enhanced by plausible theoretical predictions.
+This thesis presents such a search \atlas\ data for signs of high-energy
+protons scattering to massive resonances, whose dramatic decays are
+predicted by supersymmetric models~\cite{atlas2022searches}.
 
-\TODO{this thesis presents such a search~\cite{atlas2022searches}}
+Nothing new appears, and supports the Standard Model in its continuing
+successful extrapolation to the describe the most extreme behaviours colliding
+protons.
+Supersymmetry as a concept as not falsified by these results,
+nor by the many similar observations.
+But it is not confirmed.
 
-\TODO{after briefly introducing its theoretical and experimental context}
-
-\TODO{For given modelling of theory and experiment,
-data analysis procedures dominate, and they take a very particular form
-which we attempt to understand in its historical context
-}
-
-\TODO{
-The result is compatible with many other observations,
-that the Standard Model works very well at the LHC.
-}
-
-\TODO{supersymmetry is not falsifiable?}
-
+Data are best interpreted if their theoretical and experimental background is
+understood, so this thesis begins by introducing the theoretical and
+experimental background to its search,
+and the subtleties of its data-analysis procedures in their historical context.
 \end{abstract}
@@
 \begin{declaration}
-\noindent This dissertation is the result of my own work, except where explicit
+\noindent
+This dissertation is the result of my own work, except where explicit
 reference is made to the work of others, and has not been submitted
@@
 {Donald~E.~Knuth,
-\textit{The Art of Computer Programming: Volume~1},
+\textit{The Art of Computer Programming, Volume~1},
 1975~\cite{knuth1975art}}
@@
 
+\afterpage{%
+\phantomsection
+\addcontentsline{toc}{chapter}{Contents}%
+}
 \tableofcontents

### 2022-09-29 main/frontmatter.tex
minor text tweaks; some abstract work

@@
 nor by the many similar observations.
-But it is not confirmed.
+But it is also not confirmed.
 
-Data are best interpreted if their theoretical and experimental background is
-understood, so this thesis begins by introducing the theoretical and
-experimental background to its search,
+Data are best interpreted with an understanding of their background, so we
+begin by introducing the theoretical and experimental background to our search,
 and the subtleties of its data-analysis procedures in their historical context.

### 2022-09-29 main/frontmatter.tex
abstract tweaks

@@
 
-Nothing new appears, and supports the Standard Model in its continuing
-successful extrapolation to the describe the most extreme behaviours colliding
-protons.
-Supersymmetry as a concept as not falsified by these results,
-nor by the many similar observations.
+Nothing new appears in the results.
+This observation supports the Standard Model in its continuing successful
+extrapolation to the describe the most extreme behaviours of proton collisions.
+Supersymmetry as a concept as not falsified by this lack of visible new
+phenomena, nor by the many similar observations.
 But it is also not confirmed.

### 2022-09-29 main/chapter_theory.tex
add some on simplified models

@@
 }
-which adds four scalar fields (as a complex doublet) with a potential that
-moves the ground state to a gauge-\emph{asymmetric} configuration.
-Around this new ground state, linear combinations of the electroweak fields
+which a complex doublet (equivalently four scalar fields) with an associated
+potential.
+Minima of the potential lie in a gauge-symmetric set, and the choice made by
+a system as it settles into one particular ground state spontaneously breaks
+the symmetry.
+Around the chosen ground state, linear combinations of the electroweak fields
 acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
@@
 understanding of nature.
-Attempts to relate empirical parameters~\cite{
+Attempts to relate the empirical parameters~\cite{
 sato1979ratio,
@@
 }
-or to derive mathematical structures~\cite{
+or to derive the mathematical structures~\cite{
 goyal2010origin,
@@
 }
-are curious, but will remain speculative until they provide novel benefits.
+are curious, but will remain only curiosities unless they provide novel
+benefits.
 
@@
 
-Current observations do limit the masses and mixings of this new
-school of MSSM particles, but their values remain very uncertain and in need
-of new experimental constraints.
-
 % made using:
@@
 \end{table}
+
+Current observations do limit the masses and mixings of this new
+school of MSSM particles, but their values remain very uncertain and in need
+of new experimental constraints.
+Diverse observable effects can be induced by parameter choices,
+so we use simplified models to coarsely characterize interesting selections
+of plausible effects.
+Such simplified models feature only a few free variables, and decouple other
+effects by, for example, sending masses to extremely large values or assuming
+no mixing~\cite{
+atlas2022searches,
+alwall2009simplified,
+alves20212simplified
+}.
 
@@
 supersymmetry remains a tool for building theoretically plausible
-models, and might help us to understand future discoveries.
+models, and might help us to understand future discoveries,
+which could perhaps be made in \atlas\ data.
 

### 2022-09-29 main/chapter_theory.tex
susy tweaks

@@
 This prediction, however, assumed a mass scale below $1\,\eV[T]$,
-and observations at the LHC are now limiting sparticle masses
-into super-$\eV[T]$ territory (in specific models).
+and observations at the LHC are now limiting sparticle masses,
+in specific simplified models, into super-$\eV[T]$ territory.
 Such limits are also troublesome for the hierarchy problem, because large
@@
 supersymmetry remains a tool for building theoretically plausible
-models, and might help us to understand future discoveries,
-which could perhaps be made in \atlas\ data.
+models, and might help us to understand future discoveries
+which could be made with data from the \atlas\ detector.
 

### 2022-09-29 main/chapter_theory.tex
autoformat bib; some text tweaks

@@
 }.
-The answer is uncertain, and new results from theory or experiment that
-pinpoint a correct model could improve the Standard Model in the neutrino
-domain.
+
+The answer is uncertain, and new results from theory or experiment could
+improve the Standard Model in the neutrino sector
+by pinpointing a uniquely correct model.
 
@@
 and the Standard Model does slightly violate $C$ and $CP$, both of which
-are required for baryogenesis~\cite{sakharov1991re}.
+must be broken for baryogenesis~\cite{sakharov1991re}.
 
@@
 }
-are curious, but will remain only curiosities unless they provide novel
+are curious, but will remain as curiosities only until they provide novel
 benefits.
@@
 magnitude~\cite{giudice2008naturally}.
-For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned by an
-observation of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
-large $11$ in its exponent also defies naturalness.
+For example, $\theta_\mathrm{CP}$ is fine-tuned by the observational constraint
+$\theta_\mathrm{CP} \lesssim 10^{-11}$, and the large $11$ in its exponent
+defies naturalness.
 
@@
 This prediction, however, assumed a mass scale below $1\,\eV[T]$,
-and observations at the LHC are now limiting sparticle masses,
-in specific simplified models, into super-$\eV[T]$ territory.
-Such limits are also troublesome for the hierarchy problem, because large
-superpartner masses introduce fine-tuning problems of their
+and observations at the LHC are now limiting sparticle masses into
+super-$\eV[T]$ territory (in specific simplified models).
+Such observational limits are also troublesome for the hierarchy problem,
+because large superpartner masses introduce fine-tuning problems of their
 own~\cite{martin2016primer}.

### 2022-09-29 main/frontmatter.tex
autoformat bib; some text tweaks

--- 2022-09-29 main/frontmatter.tex
@@
 \begin{abstract}
-To search for new phenomena is an essential endeavour of experimental physics,
-and is enhanced by plausible theoretical predictions.
-This thesis presents such a search \atlas\ data for signs of high-energy
-protons scattering to massive resonances, whose dramatic decays are
-predicted by supersymmetric models~\cite{atlas2022searches}.
+Searching for new phenomena is an essential endeavour of experimental physics,
+and is helped by guidance from plausible theoretical predictions.
+This thesis presents such a search for signals of high-energy protons
+scattering to massive resonances, whose dramatic decays are predicted by
+supersymmetric models~\cite{atlas2022searches}.
 
-Nothing new appears, and supports the Standard Model in its continuing
-successful extrapolation to the describe the most extreme behaviours colliding
-protons.
-Supersymmetry as a concept as not falsified by these results,
+Nothing new appears.
+Our search finds support the Standard Model in its continuing successful
+extrapolation to describe the most extreme behaviours of proton collisions.
+Supersymmetry as a concept as not falsified by this lack of new phenomena,
 nor by the many similar observations.
@@
 
-Data are best interpreted with an understanding of their background, so we
-begin by introducing the theoretical and experimental background to our search,
-and the subtleties of its data-analysis procedures in their historical context.
+Data are best interpreted with an understanding of their origins, so we begin
+by introducing the theoretical and experimental background to our search, and
+the subtleties of our data-analysis procedures in their historical context.
 \end{abstract}

### 2022-09-29 main/frontmatter.tex
begin acknowledgements and other forematter tweaks

@@
 
+\begin{singlespacing}
 \begin{declaration}
 \noindent
+\TODO{
 This dissertation is the result of my own work, except where explicit
@@
 Degree Committee.
+}
 \vspace*{1cm}
@@
 \end{declaration}
+\end{singlespacing}
 
 \begin{acknowledgements}
-<3
-% Holly
-% Dan
-% Chris
+Warning: many thank-yous incoming.
+Thank you, thank you, thank you, thank you
+to all the friends and family who have come and gone during this project;
+I feel much love for you all <3.
 
-% Friends who have come and gone from Homerton over the years
+From the era of optimal office 958:
+thank you to Daniel~Noel for the years of comradery, surviving this with me
+--- I hope you enjoyed crushing me over the board;
+thank you to Holly~Pacey and not only for antics with the non-sofa ---
+you are an inspirational human and academic and I owe you endless thanks for
+putting up with the ravings of my most miserable months on this
+project\ldots\ oops!
 
+\TODO{supervision}
 % Sarah Williams
-% Ben Hooberman
-% Jon Long
+% for the opportunity to teach at Queens' and valuable experience from
+% Christopher Gorham lester
 
-% Family, alive and dead.
+\TODO{analysis team despite our time differences}
 
-This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
+\TODO{Cantab}
+% Tom Gillam
+% Faridka Mustafazade
+% Fred Desobry
+
+\TODO{Cambridge: Tina, Richard, working group formerly known as SUSY}
+
+For reviewing drafts of this thesis:
+thank you to
+Christopher Gorham Lester,
+Holly Pacey, and
+Daniel No\"el.
+Your comments all had stunningly sharp precision and positive impacts on this
+thesis and on my morale.
+
+The Science and Technology Facilities Research Council supported this project.
+This thesis template uses parts of the \textsc{hepthesis}
+class~\cite{hepthesis}.
 \end{acknowledgements}

### 2022-09-29 main/chapter_2ljets.tex
minor text tweaks; some abstract work
first reasonable draft of acknowledgements

@@
 \emph{This jet mass modelling was discussed with, and approved by, the \atlas\
-Jet/Etmiss conveners. Special thanks to Chris~Delitzsch for assistance
-preparing and sharing the official jet mass scale variations.}
+Jet/Etmiss conveners. Special thanks to Chris~Delitzsch for preparing and
+sharing the official jet mass scale variations.}
 
@@
 and $\pm13\%$ for \topother.
-\emph{Thanks to Jon Long for researching and assigning these uncertainties.}
+\emph{Thanks to Jon~Long for researching and assigning these uncertainties.}
 

### 2022-09-29 main/frontmatter.tex
first reasonable draft of acknowledgements

--- 2022-09-29 main/frontmatter.tex
@@
 
-Nothing new appears.
-Our search finds support the Standard Model in its continuing successful
-extrapolation to describe the most extreme behaviours of proton collisions.
+Nothing new appears in the results, so they support the Standard Model in its
+continuing successful extrapolation to describe the most extreme behaviours of
+proton collisions.
 Supersymmetry as a concept as not falsified by this lack of new phenomena,
@@
 \end{declaration}
+\end{singlespacing}
 
 \begin{acknowledgements}
-<3
-% Holly
-% Dan
-% Chris
+Warning: many thank-yous incoming.
+Friends and family made these years worthwhile;
+deepest love and thanks to all of you who have come and gone.
 
-% Friends who have come and gone from Homerton over the years
+\paragraph{958 2018/19.}
+Thank you Daniel~Noel for the years of comradery and surviving this with me
+--- I hope you enjoyed crushing me OTB.
+Thank you Holly~Pacey, not only for non-sofa antics
+--- you are an inspirational human and academic and you deserve endless thanks
+for putting up with the ravings of my most miserable months on this
+project\ldots\ oops!
 
-% Sarah Williams
-% Ben Hooberman
-% Jon Long
+\paragraph{Supervision.}
+Thank you Christopher~Gorham~Lester for guidance and encouragement,
+and for frequent reminders of the fun and intrigue outside \atlas.
+Thank you Sarah~Williams for your advice and support inside \atlas, and
+for the invaluable experience of teaching at Queens' College.
 
-% Family, alive and dead.
+\paragraph{\atlas.}
+Thank you to the Cambridge \atlas\ group and especially Tina~Potter for
+sharing help and wisdom.
+Thanks to all collaborators mentioned in Chapter~\ref{chapter:2ljets},
+to Ben~Hooberman and Sarah~Williams for leadership, and especially to
+Jon~Long for our teamwork in the push for publication,
 
-This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
+\paragraph{Misc.}
+Thank you Christopher~Gorham~Lester, Holly~Pacey, and Daniel~No\"el for your
+stunningly precise and valuable comments on drafts of this thesis.
+Thank you to Homerton College and Melanie~Keene.
+This document uses parts of `hepthesis'~\cite{hepthesis}.
 \end{acknowledgements}

### 2022-09-29 main/frontmatter.tex
set legalese nonsense

@@
 \titlepage[Homerton College]{%
-A dissertation submitted to the University of Cambridge\\
-for the degree of Doctor of Philosophy%
+September 2022\\
+This thesis is submitted for the degree of Doctor of Philosophy%
 }
@@
 
-Nothing new appears in the results, so they support the Standard Model in its
-continuing successful extrapolation to describe the most extreme behaviours of
-proton collisions.
+No new phenomena appear in the results, so they support the Standard Model in
+its continuing successful extrapolation to describe the most extreme behaviours
+of proton collisions.
 Supersymmetry as a concept as not falsified by this lack of new phenomena,
@@
 \noindent
-\TODO{
-This dissertation is the result of my own work, except where explicit
-reference is made to the work of others, and has not been submitted
-for another qualification to this or any other university.
-This dissertation does not exceed the word limit for the respective
+This thesis is the result of my own work and includes nothing which is the
+outcome of work done in collaboration except as
+%declared in the preface and
+specified in the text.
+%
+It is not substantially the same as any work that has already been submitted
+before for any degree or other qualification.
+%except as declared in the preface and specified in the text.
+%
+It includes neither more nor fewer than
+two squared times three cubed
+%one hundred and eight
+jelly beans.
+%
+It does not exceed the prescribed word limit for the
+Physics and Chemistry
 Degree Committee.
-}
 \vspace*{1cm}
@@
 Thank you to Homerton College and Melanie~Keene.
-This document uses parts of `hepthesis'~\cite{hepthesis}.
+This document uses parts of the `hepthesis' class~\cite{hepthesis}.
 \end{acknowledgements}

### 2022-09-29 main/chapter_theory.tex
write a proper preface

@@
 but the third generation and non-zero neutrino masses are now observed;
-for example, the third generation is shown by charge-parity $CP$ asymmetry in
-kaons~\cite{
+for example, the third generation is shown by charge-parity ($\mathrm{CP}$
+asymmetry in kaons~\cite{
 cronin1964evidence,
@@
 inflation~\cite{bezrukov2008higgs},
-and the Standard Model does slightly violate $C$ and $CP$, both of which
-must be broken for baryogenesis~\cite{sakharov1991re}.
+and the Standard Model does slightly violate $\mathrm{C}$ and $\mathrm{CP}$,
+both of which must be broken for baryogenesis~\cite{sakharov1991re}.
 

### 2022-09-29 main/frontmatter.tex
write a proper preface

@@
 Data are best interpreted with an understanding of their origins, so we begin
-by introducing the theoretical and experimental background to our search, and
-the subtleties of our data-analysis procedures in their historical context.
+by describing the theoretical and experimental background to our search.
+Our results depend strongly on the interpretation of data by conventional
+analysis procedures, whose properties and explanations are not obvious.
+To explain these procedures, we therefore also describe their historical
+context and the basic theory of data analysis.
 \end{abstract}
@@
 It does not exceed the prescribed word limit for the
+%
 Physics and Chemistry
+%
 Degree Committee.
@@
 \paragraph{\atlas.}
-Thank you to the Cambridge \atlas\ group and especially Tina~Potter for
+Thank you to the whole Cambridge \atlas\ group, and especially Tina~Potter for
 sharing help and wisdom.
@@
 Thank you to Homerton College and Melanie~Keene.
-This document uses parts of the `hepthesis' class~\cite{hepthesis}.
+This work was supported by the Science and Technology Facilities Council.
+The template for this document derives from `hepthesis'~\cite{hepthesis}.
 \end{acknowledgements}
@@
 \end{singlespacing}
+\noindent
+``Two leptons, jets, and missing transverse momentum'' of our title
+refers an experimental signature of new phenomena which was targetted by our
+recent search in \atlas\ data~\cite{atlas2022searches}.
+The subtitle ``yet~more non-evidence for supersymmetry'' refers to how its
+negative result is consistent with all other research in the area.
+The wording deliberately avoids suggesting evidence against supersymmetry
+(SUSY);
+we see evidence against specific supersymmetric models, but the abstract
+concept of supersymmetry remains at large.
 
-Data Intensive Science in High-Energy Physics
+Chapter~\ref{chapter:2ljets} dissects this search project and my contributions
+to it.
+Earlier chapters introduce relevant content:
+Chapter~\ref{chapter:searches} details data analysis,
+its theory and conventional practice,
+Chapter~\ref{chapter:experiment} describes the design of the \atlas\ detector,
+and
+Chapter~\ref{chapter:theory} introduces the theory of Standard Model, its
+promising future directions, and supersymmetry.
+Please consider reading them in forwards order.
 
-\TODO{Context of Standard Model theory, directions in which future developments
-could improve it, and supersymmetric models}
-supersymmery (SUSY)
-Chapter~\ref{chapter:theory}
-
-\TODO{the design of the \atlas\ detector}
-Chapter~\ref{chapter:experiment}
-
-\TODO{theory and practice of data science, as it could be for clarity
-and how it is in our \atlas\ SUSY conventions}
-Chapter~\ref{chapter:searches}
-
-\TODO{the main event: the search with many of its gory details in addition
-to those presented publically in the paper~\cite{atlas2022searches}}
-Chapter~\ref{chapter:2ljets}
-
-\paragraph{Other work.}
-\TODO{Use some of the data analysis ideas of Chapter~\ref{chapter:searches}
-to build predictive models
-and test them to test for symmetries}
-\TODO{citations}
-
+Another product of this PhD (not further explored in this thesis),
+was the development of a general method for testing symmetries in data~\cite{
+lester2021stressed,
+tombs2021which
+}
+and its application to parity violation in hadron colliders~\cite{
+lester2022hunting,
+Lester:2019bso,
+lesterChiralMeasurements2021,
+Gripaios:2020hya,
+lester2021lorentz
+}.
+The result is a practical solution which employs the principles of predicting
+data and reporting likelihoods as advocated by Chapter~\ref{chapter:searches}.
+It uses `self-supervised' machine learning to train predictive models on a
+task that can only be solved if the proposed symmetry is violated~\cite{
+tombs2021which,
+pmlr-v139-radford21a,
+Noroozi2016jigsaw,
+multitaskself2017,
+devlin2019bert
+}.
 \end{preface}

### 2022-09-29 main/chapter_experiment.tex
begin acknowledgements and other forematter tweaks
some searches section cutting :-(

@@
 real-world cost budgets.
-Within general purpose utilities, \atlas's designs aimed to construct a
-detector that could
+Within ambitions for general-purpose utility, \atlas's designs aimed to
+construct a detector that could
 \begin{itemize}
@@
 \end{itemize}
-We now review these three goals with the subsequent news from \atlas's
-operations.
+We review these three goals with news \atlas's subsequent operations.
 

### 2022-09-29 main/chapter_searches.tex
write abstract; quote tweaks; move higgs citation
some searches section cutting :-(

@@
 sivia2006data,
-skilling2010foundations
+skilling2010foundations,
+skilling2008rant
 }.
@@
 
-Phrased succinctly by Skilling~\cite{skilling2008rant},
-``we SHOULD use Bayes by assigning subjective priors and inspecting evidence
-values.''
-This framing of probabilistic data analysis agrees exactly with the standard
-practice of histogramming simulations in High Energy Physics:
-signal and background models define priors, and their (normalized) bin yields
-are the evidence values that we use to compare the models.
-Finally, their corresponding posteriors are the normalized shapes those samples
-take within the observed histogram bin
-(as if it were more finely re-binned after analysis),
-with which we do not compare the models,
-but which could in principle be used for secondary investigations.
-
 \subsection{Probability assignments}
@@
 should be initially assigned.
-Probability assignments are generally constrained only by normalization, but
-might be informed by calibration data, loosely motivated opinions,
-or by established theories such as quantum mechanics.
-But there are examples of unambiguously derived distributions that are commonly
-used in High Energy Physics, notably the Poisson and Gaussian distributions,
+Such assignments are must be normalized, but might be loosely informed informed
+by calibration data, established theories like quantum mechanics,
+or vaguely motivated opinions,
+But there are examples of unambiguously derived distributions with common use
+in High-energy physics: the Poisson and Gaussian distributions,
 both of which can derive from the identity that
@@
 As discussed in Section~\ref{sec:atlas_trigger}, the trigger rejects most
-LHC collision events;
-our data have fixed large number $N$ of trials (bunch crossings)
-and a small probability of acceptance (triggering).
+LHC collision events so our data have fixed large number $N$ of trials
+(bunch crossings) and a small probability of acceptance (triggering).
 The number of accepted events  $n$ is therefore described by the
@@
 
-With the tiny trigger rate, we are interested in the limit of small
+With the tiny trigger rate, we work near the limit of small
 $p$ and large $N$, which result in a mean acceptance of $\lambda = Np$.
@@
 \end{equation}
-which we use ubiquitously to interpret histograms.
+which we use everywhere to interpret histograms.
 
 The Gaussian distribution has many justifications:
-as the unique distribution that is spherically symmetric and factors linearly
-into independent dimensions~\cite{
+as the unique distribution that is spherically symmetric and factors into
+linearly independent dimensions~\cite{
 jaynes2003probability,
@@
 },
-as a practical approximation that permits analysis in linear algebra,
+as an approximation that permits analysis in linear algebra,
 and as the maximum entropy~\cite{PhysRev.106.620} distribution for a
@@
 
-As a limiting approximation, the Gaussian form arises from products of many
-likelihood functions from the combination of independent measurements.
-At its maximum $\check{x}$, the first derivative of a smooth function $f(x)$
-vanishes and its second derivative is negative;
-to leading order, the Taylor expansion of a product of $f$s looks like
+The Gaussian form arises from products of many functions, which often arise
+from the products of likelihood functions on independent data.
+Consider a smooth (likelihood) function $f(x)$.
+At its maximum $\check{x}$, its first derivative vanishes and its second
+derivative is negative, so to leading order a product of $f$s looks like
 \begin{equation}
@@
 \end{equation}
-which, again by Equation~\ref{eqn:searches_exponential}, goes into the Gaussian
-form
+which by Equation~\ref{eqn:searches_exponential} goes into the Gaussian form
 \begin{equation}
@@
 $\mathrm{d}x$ terms as reminders that it is a density function in $x$.
-On a $\log$ scale, the Gaussian form is quadratic in $\mu$, so
-Gaussian likelihoods justify many least-squares methods.
-Here, $\kappa$ and $\sigma$ both relate to the second derivative of the target
-function; this relationship extends to multiple dimensions in the Hessian
-matrix and its inverse, and permits the useful approximation of estimate error
-bars from second derivatives.
+Since the Gaussian form is an exponentiated quadratic, Gaussian likelihoods
+justify many least-squares methods.
 
 This is not the Central Limit Theorem, but can be used to derive it: the
-distribution of a sum is a convolution, which becomes a product in
-Fourier-transformed space, and that product converges to a Gaussian function by
+distribution of a sum is a convolution which becomes a product in
+Fourier space, and that product converges to a Gaussian function by
 if the mean and variance exist.
-Another beautiful alternative proof of the Central Limit Theorem is given by
+Another more beautiful proof of the Central Limit Theorem is given by
 Jaynes~\cite{jaynes2003probability}
@@
 In the current year, it would be remiss to not mention machine learning
-algorithms.
-Among many applications, learning algorithms have practical successes in
-probability assignments, such as to predict labels in classification,
-or to predict next words in text generation~\cite{
-MurphyKevinP.2012Mlap,
-radford2019language
-}.
-Machine learning tunes flexible functions by consuming data and testing
-predictions in this probabilistic framework:
+algorithms, which tune flexible functions by consuming data and testing
+their results.
+Their probabilistic scheme works as follows:
 \begin{enumerate}
@@
 or index permutations~\cite{Noroozi2016jigsaw, multitaskself2017}.
-Other inputs that are not predicted, such as associated images that the machine
-uses to predict the labels, are in the implicit context.
+Other inputs (such as images to which the labels belong) that are not predicted
+are in the implicit context.
 \item Use those data to update your state
 $f(d_{i-1}, \ldots) \to f(d_i, d_{i-1}, \ldots)$.
-Unlike ideal posterior updating, this encodes only a function of the data, not
+Unlike ideal posterior updating, this encodes only a function of the data and
 its whole --- one usually cannot afford to consume all aspects of data, so
@@
 practice.
-Learning only part of the data is not wrong, but less precise.
+Learning only part of the data is not wrong, just less precise.
 \item Go\,to 1.
@@
 These portions of data $d_i$ might be labelled the `training', `validation',
-and `testing' sets.
-This perspective highlights the importance of independent training and testing
-data --- information should not be reused since
-$\prob{d_2}{d_2, \ldots} = 1$, any memory of $d_2$ that is kept
-in the $f(\ldots, d_2, \ldots)$ state spoils the likelihood if $d_2$ is reused
-for testing after training~\cite{sivia2006data}.
-Independent testing data permit the clean interpretations of likelihood ratios.
+and `testing' sets, and they must be independent for rigorous testing
+because $\prob{d_1}{d_1, \ldots} = 1$:
+if you study (training) from a known exam script (testing), your answers are
+memory, not prediction.
+Independent testing data permits the cleanest interpretations of
+likelihoods~\cite{sivia2006data}.
 
@@
 
-This theory does not, however, define how to construct that cost function.
+This theory does not define how to construct that cost function.
 Like the prior, it must come from context, and indeed the two are intertwined
 in the risk;
-decision-making machines (such as organisms) do learn to make good
+decision-making machines (or organisms) do learn to make good
 decisions without constructing either cost or prior explicitly.
-Towards the goal of usefully reporting data, such decision theory has been
+Towards the goal of usefully reporting data, decision theory has been
 applied to argue for ways to summarize posterior distributions.
@@
 But those weren't the rules.
-
-When arguing that we should report likelihoods and evidence values for our
-data, we have not explicitly constructed cost functions, nor have we made
-quantitative predictions of future events.
-But we can qualitatively consider them in this decision theory framework ---
-we imagine that future scientists will want to reuse our data, and would be
-pleased (attain a small cost) if they were able to reinterpret the results for
-their own models.
-We can therefore reason that likelihoods would usually have more utility than
-alternatives such as point posterior averages.
 
@@
 This frequency-restricted, `frequentist', philosophy dominated 20th century
-statistical research, and influences practices still followed in \atlas.
-We therefore review some features of frequentist statistics and its history.
-% Would like something along the lines of:
-% `No formal derivations except for historical precedent,
-% so review with that history'
-% But feels too rude.
+statistical research, and influences practices that are still followed in
+\atlas.
+To understand these procedures, we therefore review some features of
+frequentist statistics and its history.
 
@@
 {John~W.~Pratt,
-\textit{Review: Testing Statistical Hypotheses
-by E.~L.~Lehmann},
+\textit{Review: Testing Statistical Hypotheses},
 1961~\cite{pratt1961testing}}

### 2022-09-29 main/draft.tex
some searches section cutting :-(

@@
 testing.
+
+Phrased succinctly by Skilling~\cite{skilling2008rant},
+``we SHOULD use Bayes by assigning subjective priors and inspecting evidence
+values.''
+This f
+
+% Here, $\kappa$ and $\sigma$ both relate to the second derivative of the target
+% function; this relationship extends to multiple dimensions in the Hessian
+% matrix and its inverse, and permits the useful approximation of estimate error
+% bars from second derivatives.
+
+% Among many applications, learning algorithms have practical successes in
+% probability assignments, such as to predict labels in classification,
+% or to predict next words in text generation~\cite{
+% MurphyKevinP.2012Mlap,
+% radford2019language
+% }.
+which
+
+When arguing that we should report likelihoods and evidence values for our
+data, we have not explicitly constructed cost functions, nor have we made
+quantitative predictions of future events.
+But we can qualitatively consider them in this decision theory framework ---
+we imagine that future scientists will want to reuse our data, and would be
+pleased (attain a small cost) if they were able to reinterpret the results for
+their own models.
+We can therefore reason that likelihoods would usually have more utility than
+alternatives such as point posterior averages.
+
+% Would like something along the lines of:
+% `No formal derivations except for historical precedent,
+% so review with that history'
+% But feels too rude.

### 2022-09-29 main/frontmatter.tex
some searches section cutting :-(

@@
 \noindent
-``Two leptons, jets, and missing transverse momentum'' of our title
-refers an experimental signature of new phenomena which was targetted by our
-recent search in \atlas\ data~\cite{atlas2022searches}.
+``Two leptons, jets, and missing transverse momentum'' from the title of this
+thesis refers an experimental signature of new phenomena which was targetted by
+our recent search in \atlas\ data~\cite{atlas2022searches}.
 The subtitle ``yet~more non-evidence for supersymmetry'' refers to how its
 negative result is consistent with all other research in the area.
-The wording deliberately avoids suggesting evidence against supersymmetry
-(SUSY);
-we see evidence against specific supersymmetric models, but the abstract
-concept of supersymmetry remains at large.
+The wording avoids suggesting evidence against supersymmetry (SUSY);
+we see evidence against specific simplified supersymmetric models, but the
+abstract concept of supersymmetry remains at large.
 
@@
 
-Another product of this PhD (not further explored in this thesis),
+Another product of this PhD,
 was the development of a general method for testing symmetries in data~\cite{
@@
 }.
+This project is not further explored in this thesis.
 \end{preface}

### 2022-09-30 main/chapter_searches.tex
overleaf tweask

@@
 When the evidence ratio is independent of that context, as it usually is,
-it separates from the prior (odds) ratio,
+it separates from the prior odds,
 and the $\log$ likelihood (or evidence) ratios are constant additive terms,
-which update any $\log$ prior odds ratio by the same addition.
+which update any $\log$ prior odds by the same addition.
 
@@
 should be initially assigned.
-Such assignments are must be normalized, but might be loosely informed informed
+Such assignments are must be normalized, but might be loosely informed
 by calibration data, established theories like quantum mechanics,
@@
 As discussed in Section~\ref{sec:atlas_trigger}, the trigger rejects most
-LHC collision events so our data have fixed large number $N$ of trials
+LHC collision events, so our data have fixed large number $N$ of trials
 (bunch crossings) and a small probability of acceptance (triggering).
@@
 a model~\cite{HIGG-2018-04} appears to be spreading as a good
-meme~\cite{dawkins1989selfish} in High Energy Physics~\cite{
+meme~\cite{dawkins1989selfish} in High-energy physics~\cite{
 roth2007fit,
@@
 By construction, they are uniformly distributed between $0$ and $1$, so the
-probability that $P(h) < \varepsilon$ is equal to the probability that
+probability that $P(h) < \varepsilon$ equals the probability that
 $0.4 \leq P(h) < 0.4 + \varepsilon$ or that it lands in any
-$\varepsilon$-sized set
-(if the data are indeed sampled under $h$).
+$\varepsilon$-sized set,
+if the data are sampled under $h$.
 What matters to our interpretation is whether an alternative hypothesis

### 2022-09-30 main/frontmatter.tex
overleaf tweask

@@
 This thesis is the result of my own work and includes nothing which is the
-outcome of work done in collaboration except as
+outcome of work done in collaboration, except as
 %declared in the preface and
@@
 \paragraph{958 2018/19.}
-Thank you Daniel~Noel for the years of comradery and surviving this with me
+Thank you Daniel~Noel for the years of camaraderie, surviving this with me
 --- I hope you enjoyed crushing me OTB.
 Thank you Holly~Pacey, not only for non-sofa antics
---- you are an inspirational human and academic and you deserve endless thanks
-for putting up with the ravings of my most miserable months on this
-project\ldots\ oops!
+--- you are an inspiration and you deserve endless thanks for putting up
+with the ravings of my most miserable months on this project\ldots\ oops!
 
@@
 ``Two leptons, jets, and missing transverse momentum'' from the title of this
-thesis refers an experimental signature of new phenomena which was targetted by
-our recent search in \atlas\ data~\cite{atlas2022searches}.
+thesis refers to an experimental signature of new phenomena which was targetted
+by our recent search in \atlas\ data~\cite{atlas2022searches}.
 The subtitle ``yet~more non-evidence for supersymmetry'' refers to how its
@@
 and
-Chapter~\ref{chapter:theory} introduces the theory of Standard Model, its
+Chapter~\ref{chapter:theory} introduces theory for the Standard Model, its
 promising future directions, and supersymmetry.

### 2022-09-30 main/chapter_theory.tex
overleaf tweaks: frontmatter and theory

@@
 Since the main work of this thesis is a search for effects of supersymmetric
-phenomena, supersymmetry and the specific supersymmetric models that we study
-are presented in Section~\ref{sec:theory_susy}.
+phenomena, Section~\ref{sec:theory_susy} discusses supersymmetry and
+supersymmetric models.
 
@@
 and the scalar Higgs boson.
-The strong force is mediated by
 Massless gluons $g$ mediate the strong force,
@@
 
-Gluons in this model correspond to the eight generators of $SU\!(3)_C$, whose
+Gluons correspond to the eight generators of $SU\!(3)_C$, whose
 subscript $_C$ indicates that they operate on colour-charge triplets of quarks.
@@
 These would be massless if the symmetry were respected, but it is
-`spontaneously' broken with the Higgs mechanism~\cite{
+`spontaneously' broken with the Higgs mechanism
+\footnote{%
+Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
+Peter~Higgs has reportedly suggested ``ABEGHHKtH mechanism'' to fully credit
+contributing work~\cite{close2011infinity}.%
+}~\cite{
 higgs1964broken,
@@
 }%
-\footnote{%
-Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
-Peter~Higgs has reportedly suggested ``ABEGHHKtH mechanism'' to fully credit
-associated work~\cite{close2011infinity}.%
-}
-which a complex doublet (equivalently four scalar fields) with an associated
-potential.
+which adds a complex doublet (equivalently four scalar fields) with an
+associated potential to the model.
 Minima of the potential lie in a gauge-symmetric set, and the choice made by
-a system as it settles into one particular ground state spontaneously breaks
-the symmetry.
-Around the chosen ground state, linear combinations of the electroweak fields
+settling into one particular ground state spontaneously breaks the symmetry.
+Around that chosen ground state, linear combinations of the electroweak fields
 acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
-four scalar fields to yield the three weak bosons and the photon:
+four scalar fields to yield the three massive weak bosons and the photon:
 $W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
@@
 Gravity does not feature in the Standard Model, nor does it measurably impact
-particle physics experiments.
+the particles in our experimental data.
 Although gravitation is in another domain, the particle content of the
@@
 sterile neutrinos~\cite{boyarsky2019sterile},
-Weakly Interacting Massive Particles (WIMPs)~\cite{jungman1996wimp},
+weakly interacting massive particles (WIMPs)~\cite{jungman1996wimp},
 or
@@
 primordial black holes~\cite{bernard2019primordial}.
-Dark matter may well comprise various of different components, and it would
+Dark matter may well comprise a mixture of different components, and it would
 be a success for future particle physics theories to pinpoint necessary dark
@@
 Solutions to these mysteries all lie somewhere between uncertain and unknown,
-but may not be entirely out of reach;
-for example, the Higgs field could plausibly have driven
+but may not be entirely out of reach for future Standard Models.
+For example, the Higgs field could plausibly have driven
 inflation~\cite{bezrukov2008higgs},
@@
 Justified constraints on the model could have led to better predictions of
-past data and might extrapolate to new phenomena through their deeper
+past data, and might extrapolate to new phenomena with their deeper
 understanding of nature.
@@
 }
-or to derive the mathematical structures~\cite{
+or to derive the mathematical structures of physics~\cite{
 goyal2010origin,
@@
 Naturalness is opinion; only nature decides what is really natural,
-but these criteria can still guide the discourse of our scientific searches
-towards useful directions~\cite{wells2020discovery}.
+but these criteria can still guide the scientific discourse towards useful
+outcomes~\cite{wells2020discovery}.
 
@@
 which notes that new particles at a heavy mass scale $\Lambda$ induce changes
-of $\Delta m(h)^2 = \mathcal{O}(\Lambda^2)$,
+of $\Delta m(h)^2 = \mathcal{O}(\Lambda^2)$
 through direct or indirect couplings.
-With any particles near the Planck scale $\Lambda \sim 10^{18}\,\eV[G]$,
+With any new particles near the Planck scale $\Lambda \sim 10^{18}\,\eV[G]$,
 implementing the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires
@@
 so new models can implement this fine-tuning by inserting only matched sets of
-fermions and bosons.
-This solution is the strategy of supersymmetric models.
+fermions and bosons --- this is the strategy of supersymmetric models.
 
@@
 and the Minimal Supersymmetric Standard Model (MSSM) satisfies this while
-adding a school of new particles,
-which is listed in Table~\ref{tab:theory_particles_mssm}.
+adding a school of new particles
+who are registered in Table~\ref{tab:theory_particles_mssm}.
 
@@
 ($\tilde g$, $\tilde W^1$, $\tilde W^2$, $\tilde W^3$, and $\tilde B$).
-The Higgs field does not translate so directly, but is replaced with two sets
-of complex scalar fields
+The Higgs field does not translate so directly, but is replaced with four
+complex scalar fields
 ($H^+_u$, $H^0_u$ and $H^0_d$, $H^\pm_d$) with
@@
 ($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
-(All sparticles are marked with tildes $\tilde{~}$.)
-These fields mix into mass eigenstates which current data do not entirely
-determine:
+These fields mix into mass eigenstates whose compositions are uncertain:
 the gaugino mixtures are the charged charginos $C^\pm_i$ and
@@
 and each squark and slepton has two states that mix the left and right fields.
-From the four complex scalars, three degrees of freedom are eaten by the
-weak bosons to leave the reborn Standard Model Higgs boson $h$
-and four new scalars~\cite{
-martin2016primer
-}.
+The four complex Higgs fields have eight degrees of freedom, of which three are
+again eaten by the weak bosons to leave the reborn Standard Model Higgs boson $h$
+and four new scalars~\cite{martin2016primer}.
 
@@
 Diverse observable effects can be induced by parameter choices,
-so we use simplified models to coarsely characterize interesting selections
-of plausible effects.
+so we use simplified models to coarsely characterize interesting and plausible
+examples.
 Such simplified models feature only a few free variables, and decouple other
@@
 Various mechanisms could spontaneously break supersymmetry.
-Gravitational ideas exist, but one model considered in this thesis uses a
-Gauge-Mediated Supersymmetry Breaking (GMSB) mechanism,
+Gravitational mechanism exist, but one model considered in this thesis uses
+Gauge-Mediated Supersymmetry Breaking (GMSB),
 which adds yet more elusive fields:
@@
 }.
-Conservation of $R$-parity helps by preventing proton decays,
+Conservation of $R$-parity usefully preventing proton decays,
 but also makes the Lightest Supersymmetric Particle (LSP) a plausible
@@
 Providing this LSP dark matter candidate through a production mechanism
-that could be observable at the LHC is a strong motivating factor
-both for the supersymmetric models
-and for the experimental searches of this thesis.
+that is accessible at the LHC is a strong motivating factor both for
+supersymmetric models and for the experimental search of this thesis.
 
@@
 own~\cite{martin2016primer}.
-Despite its current lack of observational evidence, however,
+Despite its current lack of observational evidence,
 supersymmetry remains a tool for building theoretically plausible

### 2022-09-30 main/frontmatter.tex
overleaf tweaks: frontmatter and theory

@@
 This thesis presents such a search for signals of high-energy protons
-scattering to massive resonances, whose dramatic decays are predicted by
+scattering to massive resonances with dramatic decays, which are predicted by
 supersymmetric models~\cite{atlas2022searches}.
@@
 by describing the theoretical and experimental background to our search.
-Our results depend strongly on the interpretation of data by conventional
+Our results also depend strongly on the interpretation of data by conventional
 analysis procedures, whose properties and explanations are not obvious.
-To explain these procedures, we therefore also describe their historical
-context and the basic theory of data analysis.
+To explain these procedures, we also discuss their historical context and the
+basic theory of data analysis.
 \end{abstract}
@@
 Warning: many thank-yous incoming.
-Friends and family made these years worthwhile;
-deepest love and thanks to all of you who have come and gone.
+Friends and family made these years worthwhile:
+mu deepest love and thanks to all of you who have come and gone.
 
 \paragraph{958 2018/19.}
-Thank you Daniel~Noel for the years of camaraderie, surviving this with me
---- I hope you enjoyed crushing me OTB.
-Thank you Holly~Pacey, not only for non-sofa antics
---- you are an inspiration and you deserve endless thanks for putting up
-with the ravings of my most miserable months on this project\ldots\ oops!
+Thank you Daniel~Noel for the years of camaraderie, surviving this with me.
+I hope you enjoyed crushing me OTB.
+Thank you Holly~Pacey, not only for non-sofa antics; you are an inspiration
+and you deserve endless thanks for putting up with the ravings of my most
+miserable months on this project\ldots\ oops!
 
@@
 This work was supported by the Science and Technology Facilities Council.
-The template for this document derives from `hepthesis'~\cite{hepthesis}.
+This thesis template derives from `hepthesis'~\cite{hepthesis}.
 \end{acknowledgements}
@@
 \noindent
-``Two leptons, jets, and missing transverse momentum'' from the title of this
-thesis refers to an experimental signature of new phenomena which was targetted
+``Two leptons, jets, and missing transverse momentum'' from our title thesis
+refers to an experimental signature of new phenomena which was targetted
 by our recent search in \atlas\ data~\cite{atlas2022searches}.
 The subtitle ``yet~more non-evidence for supersymmetry'' refers to how its
-negative result is consistent with all other research in the area.
-The wording avoids suggesting evidence against supersymmetry (SUSY);
+negative result is consistent with all other research in the area,
+and avoids suggesting evidence against supersymmetry (SUSY);
 we see evidence against specific simplified supersymmetric models, but the
@@
 
-Another product of this PhD,
-was the development of a general method for testing symmetries in data~\cite{
+Another product of this PhD was the development, with collaborators, of a
+general method for testing symmetries in data~\cite{
 lester2021stressed,
@@
 The result is a practical solution which employs the principles of predicting
-data and reporting likelihoods as advocated by Chapter~\ref{chapter:searches}.
-It uses `self-supervised' machine learning to train predictive models on a
-task that can only be solved if the proposed symmetry is violated~\cite{
+data and reporting likelihoods that Chapter~\ref{chapter:searches} advocates.
+It uses `self-supervised' machine learning to train predictive models to
+perform tasks that can only be solved if the tested symmetry is
+violated~\cite{
 tombs2021which,
@@
 }.
-This project is not further explored in this thesis.
+This thesis does not explore that project further.
 \end{preface}

### 2022-09-30 main/chapter_experiment.tex
overleaf tweaks to experiment

@@
 It was designed with intent to provide maximally useful and novel data within
-real-world cost budgets.
-Within ambitions for general-purpose utility, \atlas's designs aimed to
+real-world budgets.
+Alongside ambitions for general-purpose utility, \atlas's designs aimed to
 construct a detector that could
@@
 \item find the Higgs boson (which was predicted but yet unseen),
-\item precisely measure known massive objects ($W$, $Z$, and $t$),
+\item precisely measure known massive objects ($W$, $Z$, and $t$), and possibly
 \item discover new (supersymmetric) particles~\cite{atlas1999design1}.
@@
 \paragraph{Find the Higgs boson.}
-Prior to \atlas's construction, the Higgs boson mass (and existence) was
-poorly known, but loosely constrained by results like
-$m(H) \gtrsim 90\,\eV[G]$ from searches at the
+Before \atlas's construction, the Higgs boson mass (and existence) was
+uncertain, but loosely constrained by results like
+$m(h) \gtrsim 90\,\eV[G]$ from searches at the
 Large Electron–Positron Collider (LEP),
-$m(H) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
-and $m(H) \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
+$m(h) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
+and $m(h) \lesssim 1\,\eV[T]$ from unitarity arguments~\cite{
 atlas1999design2,
@@
 For its discovery alone, \atlas's designers achieved sensitivity across the
-entire $100$ to $1000\,\eV[G]$ window by targetting decays including
+entire $100\,\eV[G]$ to $\,\eV[T]$ window by targetting decays including
 $h \to b\bar b$,
@@
 $h \to ZZ^* \to 4\ell$, and
-$h \to ZZ \to \ell\ell\nu\nu$;
-these channels demand precise measurements of leptons, photons, and
-$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
-transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
-at the very highest masses.
+$h \to ZZ \to \ell\ell\nu\nu$.
+These channels demand precise measurements of leptons, photons, and
+$b$-jets (jets seeded by $b$ quarks), as well as sensitivity to the missing
+transverse momentum of invisible neutrinos.
 
-Both \atlas\ and \cms\ have now found a Higgs boson candidate at
+\atlas\ and \cms\ have now found a Higgs boson candidate with
 $m(h) = 125\,\eV[G]$~\cite{
@@
 could give clean signals, but are not competitive due to their
-tiny production rates and difficult identification~\cite{
+tiny cross-sections~\cite{
 cdf1996search,
@@
 }.
-Leptonic $B$ hadron decays are useful in $m(t)$ measurements, but do not
+Leptonic $B$ hadron decays are useful for $m(t)$ measurements, but do not
 require hadron identification~\cite{
@@
 },
-These continue to not invalidate the Standard Model, and are aided in their
-sensitivity by the lack of new phenomena,
-which could have overpowered these rare signals.
+These measurements continue to refine the Standard Model, and their precision
+is aided by no new phenomena getting in the way.
 
 \paragraph{Discover new (supersymmetric) particles.}
-High-energy proton-proton collisions at the LHC therefore offer sensitivity
+High-energy proton-proton collisions at the LHC offer sensitivity
 to new particles up to multi-$\eV[T]$ mass scales.
@@
 our other goals of refining the current Standard Model
-and to discovering the Higgs boson at any mass.
+and to discovering the Higgs boson.
 Visible decay products should already be caught by \atlas's abilities
@@
 
-Indeed, the main search presented in this thesis~\cite{atlas2022searches}
-joins nicely with the pre-existing needs for sensitivity to signatures
+Indeed, the main search of this thesis~\cite{atlas2022searches}
+meets nicely with the pre-existing need for sensitivity to signatures
 including
@@
 $ZZ \to \ell\ell\nu\nu$.
-
 Dozens of searches for new phenomena have now been conducted in \atlas\ data,
-with~\cite{
-ATL-PHYS-PUB-2022-013
-}
-and without~\cite{
+with and without supersymmetry~\cite{
+ATL-PHYS-PUB-2022-013,
 ATL-PHYS-PUB-2022-007,
@@
 ATL-PHYS-PUB-2022-034
-} supersymmetry,
-and show no convincing evidence for effects beyond the Standard Model.
+},
+and show no convincing sign of effects beyond the Standard Model.
 This is not a failure. It is a success.
-Our goal in science is not to discover something that is not there,
+Our scientific goal is not to discover something that is not there,
 but to learn more about nature.
-Finding no tangible new particles in LHC conditions is itself a discovery.
+Discovering no tangible new particles in LHC conditions is itself a discovery.
 
@@
 
-Our search paper~\cite{atlas2022searches} contains concise and standard
-descriptions of the \atlas\ detector and its relevant components.
-Through its many inexact reproductions in past \atlas\ publications, the text
-of these descriptions has become so highly refined that attempts to improve
-it myself would probably be in vain.
-Instead, this section quotes our published text with explanatory commentary.
-
 \begin{figure}[tp]
@@
 
+Our search paper~\cite{atlas2022searches} contains concise and standard
+descriptions of the \atlas\ detector and its relevant components.
+Through its evolution in many past \atlas\ publications, these descriptions
+have become so highly refined that attempts to improve them myself would
+probably be in vain.
+Instead, this section quotes our published text with explanatory commentary.
 \begin{quote}
@@
 \end{quote}
-To illustrate this description, a vertical slice of \atlas\ is displayed in
-Figure~\ref{fig:atlas_cutaway}, and each subsystem is described in more detail
-in a subsection below:
+To illustrate this summary, Figure~\ref{fig:atlas_cutaway} shows a vertical slice
+of \atlas.
+Each subsystem is described in more detail in a subsection below:
 the inner tracking detector in Section~\ref{sec:atlas_inner},
-the electromagnetic and hadronic calorimeter systems in
+the electromagnetic and hadronic calorimeters in
 Section~\ref{sec:atlas_calo},
 and
-the chambers of the muon spectrometer in Section~\ref{sec:atlas_muon}.
-The final connection to the data analysed in this thesis is made through the
+the muon spectrometer in Section~\ref{sec:atlas_muon}.
+The final connection to data analysed in this thesis is made through the
 trigger system which is described in Section~\ref{sec:atlas_trigger},
 and details of the data themselves are given in
-Section~\ref{sec:atlas_data}
+Section~\ref{sec:atlas_data}.
 
@@
 in~\cite{lester2013heffalon}.
-True to form, internal \atlas\ guidelines ask that
+True to form, internal \atlas\ guidelines request that
 ``If you think you can improve on it, please consult with the
@@
 
-Following radially outwards, the silicon pixel and silicon microstrip%
+Following radially outwards are the silicon pixel and silicon microstrip%
 \footnote{%
@@
 detectors (which use silicon technology)
-and the transition radiation tracking detectors (which uses `straw' tubes
-gas tubes).
+and the transition radiation tracking detector
+(which uses `straw' gas tubes).
 
 As for the whole \atlas\ detector, complete $\phi$ coverage and a large $\eta$
-range minimizes leakage.
-Towards \atlas's design goals, this inner tracking detector crucially provides
-measurements of electrons, B-hadrons, and contributes to the reconstruction of
-missing transverse momentum.
+range minimizes leakage from the inner detector.
+Towards design goals, this tracking detector provides crucial
+measurements of electrons and B-hadrons,
+and contributes to the reconstruction of missing transverse momentum.
 
@@
 Maximal angular coverage by the calorimeters further facilitates missing
-transverse momentum estimates.
+transverse momentum measurements.
 
-Neither `A' in \atlas\ stands for Argon, but perhaps one should.
+Neither `A' in \atlas\ stands for Argon, but perhaps one could.
 Liquid-argon is the chosen active material in the electromagnetic and forward
@@
 liquid-argon pre-sampler is installed before the central EM calorimeter
-to regain such loss.
+to regain these losses.
 
-These calorimeter systems achieve attenuations of at least ten interaction
+The calorimeter systems achieve attenuations of at least ten interaction
 lengths throughout their $|\eta| < 4.9$ range;
@@
 instead measured by their curved tracks in a toroidal magnetic field that
-envelopes dedicated detectors on \atlas's periphery.
+permeates dedicated muon detectors on \atlas's periphery.
 \begin{quote}
@@
 although extremely energetic muons will have straight tracks,
-\atlas's construction and calibration achieves a high precision, which
-increases to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
+\atlas's construction and calibration achieves a high precision which
+loosens to only $10\%$ relative error on tracks with $1\,\eV[T]$ transverse
 momenta~\cite{
@@
 
-The search presented in this thesis uses data from proton-proton collisions at
+The search of this thesis uses data from proton-proton collisions at
 $\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
@@
 % closing comment?
-Now that we have predictive models, in the Standard Model and supersymmetric
-alternatives,
-and interpretable data, from \atlas's measurements of LHC collisions,
+Now that we have predictive models in the Standard Model and supersymmetric
+alternatives
+and interpretable data from \atlas's measurements of LHC collisions,
 we are ready to begin to squeeze some useful research from those data.

### 2022-09-30 main/chapter_theory.tex
less neutrino mass: use an atlas measurement

@@
 \emph{uncertain} and \emph{unknown} respectively:
-the model of neutrino masses is mostly \emph{uncertain}, but
+the top quark mass is slightly \emph{uncertain}, but
 the reconciliation of quantum field theory with general relativity is

### 2022-09-30 main/chapter_searches.tex
overleaf readthrough

@@
 data, and to report that message to the wider community.
-Various messages could be extracted from given data,
-and those messages will have differing values to their recipients.
+Various messages could be extracted from any given data,
+and those messages will have varied values to their recipients.
 Data analysis is not an automatic process, but a design with many judgements
@@
 
-To support our later description of the \atlas\ SUSY $\twoljets$ search in
-Chapter~\ref{chapter:2ljets}, this chapter introduces the basic theory and
-practice of data analysis and its conventional form in the \atlas\ SUSY
+To support our later description of the \atlas\ SUSY `$\twoljets$' search in
+Chapter~\ref{chapter:2ljets}, this chapter introduces the basic theory of
+data analysis and its conventional form in the \atlas\ SUSY
 community, parts of which can only be explained in their historical context.
@@
 How many jelly beans are in the jar of Figure~\ref{fig:searches_beans}?
-Your answer depends not only on what you observe from the figure, but also on
+Your guess depends not only on what you observe from the figure, but also on
 what else you know about cartoon jars of beans, and on what reward is offered.
@@
 and on which numbers of beans you subjectively infer to be probable.
-We shall return to this example after developing its analysis in probability
+We return to this example after developing its analysis in probability
 and decision theory.
@@
 Probabilistic data analysis follows the ordinary structure of science:
-write down predictive models, test those predictions against data,
-then update your models (or invent new ones) to refine their predictions
-to accommodate what has been learned.
-Likelihood functions are particularly important to this process because they
-describe our predictions.
-But as we shall see, not everything is a likelihood.
+write down predictive models, test their predictions against data,
+then update your models (or invent new ones) to refine their predictions.
+Likelihood functions are particularly important to this process because
+they quantify predictions.
+But not everything is a likelihood.
 
@@
 proportion of stuff within a larger whole.
-Probabilities are useful because proportions have applications
-(particularly to logical reasoning with uncertainty), and because
-they can be manipulated and interrelated with an intuitive linear algebra.
-The remainder of this section uses a standard example from High Energy Physics
---- a histogram ---
-to introduce this probabilistic structure and other parts of its technical
-language.
+Probabilities are useful because proportions have important applications,
+and because they can be manipulated and interrelated with an intuitive linear
+algebra.
+The remainder of this section uses the omnipresent tool of high-energy
+physics --- the histogram --- to introduce the structure of probability and
+its technical language.
 
 \subsection{Probability: abstract}
-A probability $\prob{b}{a}$ is a number, which quantifies the proportion
-of $b$ within $a$ and can be expressed as a ratio~\cite{axioms1010038}
+A probability $\prob{b}{a}$ is a number which quantifies the proportion
+of $b$ within $a$, and can be expressed as a ratio
 \begin{equation}
@@
 in which $m(x)$ is a measure, and the comma `,' means `and'
-(logical and or set intersection or anything equivalent).
+(logical `and', or set intersection or any other equivalent
+thing)~\cite{axioms1010038}.
 
@@
 and $\prob{b}{a} = 0$ when $a$ contradicts $b$.
-Indeed, Cox's theorem shows that probability algebra is uniquely consistent
+Indeed Cox's theorem shows that probability algebra is uniquely consistent
 with Boolean logic~\cite{
@@
 
-Standard probability algebra derives from
-Equation~\ref{eqn:searches_prob_ratio}; it comprises a \textbf{product rule}
+Probability algebra derives from
+Equation~\ref{eqn:searches_prob_ratio}; it has a \textbf{product rule}
 \begin{align}
@@
 convenient to represent them as density functions.
-Densities are stated relative to coordinates (or occasionally measures),
+Densities are stated relative to coordinates (or measures),
 so changing coordinates changes their values as their unit probability mass
 is sloshed around.
-Likelihoods are not densities in the conditioning variable that they explore,
+Likelihoods are not densities,
 so changing coordinates instead slides them horizontally like beads on an
@@
 
-To illustrate this, consider a histogram that bins event yields in
+To illustrate this, we consider a histogram that bins event yields in
 $\met$, with distributions from background and (supersymmetric) signal samples
@@
 }
-
 To preserve the area of each bin, which shows its probability mass,
-bin heights therefore change to adapt to their widths that change due to the
-differing values of $\met$ and $\sqrt{\met}$.
-This coordinate-dependence of the prior density means that its point of
+bin heights change to adapt to their width in each coordinate.
+Coordinate-dependence of the prior density means that its point of
 maximum density is not always meaningful.
@@
 
-Standard methods approximate such histograms from Monte Carlo simulations,
-which do not directly indicate bin indices but state the values for event
-variables.
-To connect event variables to bin indices, we begin with some notation:
+Standard methods approximate such histograms from Monte Carlo simulations.
+Simulations do not directly indicate bin indices, but state the values for
+event variables;
+to connect event variables to bin indices, we begin with some notation:
 we have an hypothesis label $h$ (signal or background),
@@
 filter that only rejects events that are contradicted by the data.
-Other likelihoods that are not binary act similarly, but as inefficient filters
-that reject events with some probability.
+Other likelihoods that are not binary act as inefficient filters
+which reject events with some probability.
 
@@
 The name `evidence' relates to its application for model
-comparison~\cite{mackay2003information}.
-
-Suppose that we have this one event with $\met > 100\,\eV[G]$ ---
-did it come from the background or signal?%
+comparison~\cite{
+mackay2003information,
+skilling2004nested,
+skilling2006nested
+}.
+
+Suppose that we have one event with $\met > 100\,\eV[G]$.
+Did it come from the background or signal?%
 \footnote{%
@@
 support the same maximum likelihood of
-$L(\varepsilon + 100\,\eV[G]) = 1$.
-These evidence values here differ only by their prior masses in that bin.
+$L(101\,\eV[G]) = 1$.
+These evidence values differ only by their prior masses in the bin.
 
@@
 function, so the likelihood must contain all relevant information from the
-data.
-The same is true for the evidence for its reuse as a likelihood elsewhere.
+data; the same is true for the evidence for its reuse elsewhere as an
+abstracted likelihood.
 To summarize data for scientific communications, we therefore find a
@@
 \begin{itemize}
-\item the data themselves, with instructions for their interpretation,
+\item the data themselves with instructions for their interpretation,
 \item the likelihood function, or functions for varied interpretations, and
@@
 \end{itemize}
-Each is a compression of the data, which is easier to interpret in more refined
-contexts.
-Note that summaries of these objects also be useful, such as maxima
-of the likelihood function,
+Each step compresses the data, into a smaller form that is easier to interpret
+in more refined contexts.
+Summaries of these objects are useful, such as maxima of the likelihood function,
 or regions in which it exceeds meaningful thresholds.
@@
 
-Many objections to Bayes' rule stem from a fixation on posterior
+Common objections to Bayes' rule stem from a fixation on posterior
 distributions from unjustified prior assignments.
@@
 Prior assignments are arbitrary.
-You are welcome to propose any prediction in any circumstances, and the
-scientific response is to test how well it predicts our observations of nature.
 Our emphasis on the likelihood functions and evidence values that quantify
@@
 is removed;
-Figure~\ref{fig:searches_baton_roue_bayes} illustrates this analogy.
-A unicycle is more of a circus trick than a practical transport.
+Figure~\ref{fig:searches_baton_roue_bayes} illustrates this analogy;
+a unicycle is more of a circus trick than a practical transport.
 
@@
 But there are examples of unambiguously derived distributions with common use
-in High-energy physics: the Poisson and Gaussian distributions,
+in high-energy physics: the Poisson and Gaussian distributions,
 both of which can derive from the identity that
@@
 At its maximum $\check{x}$, its first derivative vanishes and its second
-derivative is negative, so to leading order a product of $f$s looks like
+derivative is negative, so at leading order a product of $f$s near its maximum
+looks like
 \begin{equation}
@@
 
-This is not the Central Limit Theorem, but can be used to derive it: the
+This is not the famous central limit theorem, but can be used to derive it: the
 distribution of a sum is a convolution which becomes a product in
@@
 if the mean and variance exist.
-Another more beautiful proof of the Central Limit Theorem is given by
+A more beautiful proof of the central limit theorem is given by
 Jaynes~\cite{jaynes2003probability}
-(Section ``7.16 The central limit theorem''), but would be spoiled by a
+(Section ``7.16 The central limit theorem''), but would be spoiled by
 reproduction here.
@@
 \subsection{Decision theory}
-Now that we have thoroughly covered probability theory, we are ready to return
+Now that we have thoroughly covered probability theory, we  return
 to the jelly bean game of Figure~\ref{fig:searches_beans}.
@@
 Wald's theorem states that restricting oneself to minimizing an expected cost
-function cannot deny one access to optimal decisions.
+function cannot deny access to optimal decisions.
 Alternative formulations can at best be equivalent.
 
-This theory does not define how to construct that cost function.
-Like the prior, it must come from context, and indeed the two are intertwined
-in the risk;
+This theory does not define how to construct a cost function.
+Like probability assignments, it must come from context, and indeed the two
+are intertwined in the risk;
 decision-making machines (or organisms) do learn to make good
@@
 and must be rejected.
-The frequency definition does not change probability algebra, it only restricts
-it to a permitted scope.
+The frequency definition does not change probability algebra,
+but it does restrict its scope.
 
@@
 toss a coin ninety-two times, all landing heads.
-They then play a game with a third player and land eight more heads.
-Will the next toss be heads or tails?
+They then play a game and land eight more heads.
+Will the next toss land heads or tails?
 The answer is not random --- it is burned onto paper in print.
@@
 of round numbers in base-10, and your reading of these words.
-Frequentist theory, however, forbids such subjective probabilities of
+Frequentist theory forbids such subjective probabilities of
 non-random unknowns, and is left searching for alternative forms of inference.
@@
 We therefore stick with `frequentist'.
-Sections~\ref{sec:searches_fisher} and~\ref{sec:searches_np} now introduce
-these frequentist methods to explain the historical context for our modern
+Sections~\ref{sec:searches_fisher} and~\ref{sec:searches_np} introduce
+these frequentist methods to explain the historical context of our modern
 methods, which are introduced in Section~\ref{sec:searches_practice}.
@@
 or by obfuscating those estimates with bias corrections.
-Remaining uses of unbiasedness in High Energy Physics~\cite{
+
+Lingering uses of unbiasedness in high-energy physics~\cite{
 pdg2022ynf,
@@
 are largely explained by the linguistic trickery of its name --- our
-results should not be biased by forces outside the data and relevant theory,
-and we like to report that to be true.
+results should not be biased, and we like to report that to be true.
 But we should proudly proclaim that our results are `biased' under this flawed
@@
 Fisher's \pvalue\ ratio of $1000$ can therefore be a $\log$ likelihood
-ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a moderate
+ratio of $+6.9$ in favour of $r=0.3$ over $r=0.95$, which is a notable
 indication.
-A likelihood ratio on the test statistic itself, and not the inclusive
-tail statement, would however encode more information and could provide a
-more precise answer.
+A likelihood ratio on the test statistic or data itself would however
+encode more information and could provide a more precise answer.
 
@@
 \end{quote}
-Elsewhere, they also consider an example with two hypotheses $\mathrm{A}_1$
+Elsewhere also, they consider an example with two hypotheses $\mathrm{A}_1$
 and $\mathrm{A}_2$, and find a $\log$ likelihood ratio of
@@
 % Hypothesis $\mathrm{A}_1$ to that of $\mathrm{A}_2$ is $1.24 \times 10^{-16}$.
-$-37$, concluding that
+$-37$ and conclude that
 ``%
@@
 \noindent
-Data analysis in modern High Energy Physics does not exactly follow frequentist
+Data analysis in modern high-energy physics does not exactly follow frequentist
 ideology, but is a hands-on, practical, and evolving ecosystem that has adapted
@@
 a model~\cite{HIGG-2018-04} appears to be spreading as a good
-meme~\cite{dawkins1989selfish} in High-energy physics~\cite{
+meme~\cite{dawkins1989selfish} in high-energy physics~\cite{
 roth2007fit,
@@
 Likelihood ratios on the actual data escape immediately by bypassing the
-censor, but even inequalities need not be biased --- if taken against a
-predetermined threshold, say $5$ events, then models can predict whether
-$n \leq 5$, and the honestly answered truth or falsehood in data can favour
-either side.
-Ultimately, a reader may choose to distrust the reporter ---
-the messages
-`logically, $n \leq 5$' and
-`\atlas\ reports $n \leq 5$' and  are not
-identical, since one might infer from the latter that in fact the original
-data were $n = 5$, before their reduction into a test statistic tail.
-The message itself is data, and what it says may not be the whole truth.
+censor, but even inequalities need not be biased if models can ask
+questions of the data and be given honest answers.
+For example, if asking whether the data exceed a predetermined threshold
+of  $5$ events, then the truth or falsehood of `$n \leq 5$'
+can favour either side.
 
@@
 ``likelihood'' models that blur the distinction between actual data and
-auxiliary constraints on the models' parameters.
+auxiliary constraints.
 
@@
 statistical physics and the Boltzmann factor
-\begin{equation}
-f(x) \propto e^{-\beta E(x)}
-.
-\end{equation}
+$f(x) \propto e^{-E(x)/T}$,
 with which a thermodynamic system collapses to its minimum energy at cold
-temperatures $T \propto 1/\beta$~\cite{
+temperatures~\cite{
 skilling2017david,
@@
 At unit temperature, this can induce a probability distribution
-for the parameters relative to an appropriate (prior)
-measure~\cite{cranmer2012histfactory, skilling2017david},
+relative to an appropriate measure~\cite{
+cranmer2012histfactory,
+skilling2017david
+},
 but in my experience with \heplikelihood\ models such distributions often

### 2022-09-30 main/chapter_2ljets.tex
overleaf skim; holy shit this is long and boring

@@
 \noindent
-The main messages that this thesis reports are my contributions to the recently
+The main results of this thesis are my contributions to the recently
 published \atlas\ searches ``for new phenomena in events with two leptons,
@@
 particularly to its electroweak part.
-For brevity, I will call this publication $\twoljets$ and its electroweak
+For brevity, we will call this publication $\twoljets$ and its electroweak
 part $\twoljets$-electroweak.
@@
 This unconventional ordering makes sense for two reasons.
-Firstly, it avoids iterating over those elements twice, once before unblinding
+First, it avoids iterating over those elements twice, once before unblinding
 and again after;
@@
 of the chapter.
-Secondly, and sadly, many design decisions can only be explained in light of
+Second, many design decisions can only be explained in light of
 the unblinded results, since major aspects of the search were reworked in the
@@
 When first meeting an \atlas\ search paper,
-there are certain plots and tables that I, personally, prefer to find before
-delving into its written details.
+there are certain plots and tables that I personally prefer to find before
+delving into its details.
 Specifically, these include the summary plot with all signal regions,
@@
 perpendicular components~\cite{atlas_met_significance}.
-This is the square of a `Mahalanobis' (standard-deviations)
-distance~\cite{mahalanobis1936generalised} between the origin and $\ptmiss$.
+This is a squared `Mahalanobis' (standard-deviations)
+distance~\cite{mahalanobis1936generalised} between $\ptmiss$ and the origin.
 
@@
 mass $\mttwo$ is a partner to the `transverse mass' $m_T$
-(to be defined shortly)
+(to be defined shortly),
 which is useful in searches for supersymmetric particles.
@@
 always chosen as the two hardest jets. Why?
-
 One reason is that this choice matches the convention of SR2-int and SR2-high
@@
 All $\twoljets$-electroweak regions are defined within a loose `Pre-selection'
-requirement of the basic $Z\to \ell^+\ell^-$ plus jets plus $\met$
+requirement of the basic $Z\to \ell\ell$ plus jets plus $\met$
 conditions that we target.
@@
 \emph{%
-Thanks are due to Knut~Vadla for important contributions in the initial
-exploration of signal region designs,
-to Ben~Hooberman for suggesting the investigation of massive jets,
-and to the whole $\twoljets$ team for collaboration in their finalization.
+Thanks to Knut~Vadla for important contributions in the initial signal region
+designs, to Ben~Hooberman for suggesting the investigation of massive jets,
+and to the whole $\twoljets$ team for collaboration in their finalization.5%
 }
@@
 Diboson backgrounds and C1N2 signals have more in common than just $\met$;
-both also produce resonant $Z\to \ell^+\ell^-$ pairs
+both also produce resonant $Z\to \ell\ell$ pairs
 and do not enhance the rate of heavy-flavour jets, as displayed in
@@
 
-The high regions, defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
-SR-High-1J, and \srllbb.
-
 \begin{table}[tp]
@@
 
-Here, the standard wide $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement
+The high regions, which defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
+SR-High-1J, and \srllbb.
+Their wide standard $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement
 captures both $W\to jj$ and $Z\to jj$ for C1N2 and GMSB models
@@
 
-Boosted $Z\to \ell^+\ell^-$ decays will tend to have small $\rll$.
+Boosted $Z\to \ell\ell$ decays will tend to have small $\rll$.
 Similarly to $\rjj$ in the High regions, this can usefully select signals,
-but here also works to reject other $\ell^+\ell^-$ production mechanisms which
+but here also works to reject other $\ell\ell$ production mechanisms which
 produce leptons in less correlated directions, particularly dileptonic
@@
 Signals trend towards small $\rll$, plausibly due to their more boosted
-$Z\to \ell^+\ell^-$ decays; small $\rll$ also suppresses $\ttbar$.
+$Z\to \ell\ell$ decays; small $\rll$ also suppresses $\ttbar$.
 The mass of signals seen at small $\mttwoll$ in (b) motivates the design of
@@
 supersymmetric resonance decays to $Z+\mathrm{invisible}$ and sends both
-the $Z\to \ell^+\ell^-$ and invisible particle, which becomes
+the $Z\to \ell\ell$ and invisible particle, which becomes
 $\ptmiss$, in the same direction.
@@
 Hadronic resonances that decay to lepton pairs,
-such as the various charmonium ($J/\psi$) bottomonium ($\Upsilon$) states,
+such as the various charmonium ($J/\psi$) and bottomonium ($\Upsilon$) states,
 become important below around $10\,\eV[G]$, and we did not have simulations
@@
 The requirement $\mll > 12\,\eV[G]$ is therefore imposed to robustly remove
-this un-modelled regime with a $2\,\eV[G]$ margin to avoid edge effects.
+this un-modelled regime with that $2\,\eV[G]$ margin to reduce edge effects.
 
@@
 
-Personally, I see little benefit from having tens of discovery regions.
+Personally, I see little benefit from having tens of similar discovery regions.
 Yes, given a pet model, one can choose the region with its best expected
@@
 
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.38\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRHigh.png}
+\caption{DR-High}
+\end{subfigure}
+\quad\quad
+\begin{subfigure}{0.38\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRllbb.png}
+\caption{DR-$\llbb$}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.38\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRInt.png}
+\caption{DR-Int}
+\end{subfigure}
+\quad\quad
+\begin{subfigure}{0.38\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRLow.png}
+\caption{DR-Low}
+\end{subfigure}
+\\[0.5em]
+\begin{subfigure}{0.38\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DROffShell.png}
+\caption{DR-OffShell}
+\end{subfigure}
+\caption[
+Pre-fit yields in discovery regions
+]{%
+Pre-fit yields in discovery regions.
+Errors are statistical.
+}
+\label{fig:2ljets_disco_prefit}
+\end{figure}
+
 As with other designs, simplicity has value.
@@
 
-\begin{figure}[tp]
-\centering
-\begin{subfigure}{0.38\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRHigh.png}
-\caption{DR-High}
-\end{subfigure}
-\quad\quad
-\begin{subfigure}{0.38\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRllbb.png}
-\caption{DR-$\llbb$}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.38\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRInt.png}
-\caption{DR-Int}
-\end{subfigure}
-\quad\quad
-\begin{subfigure}{0.38\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DRLow.png}
-\caption{DR-Low}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.38\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/2ljets_disco_plot_DROffShell.png}
-\caption{DR-OffShell}
-\end{subfigure}
-\caption[
-Pre-fit yields in discovery regions
-]{%
-Pre-fit yields in discovery regions.
-Errors are statistical.
-}
-\label{fig:2ljets_disco_prefit}
-\end{figure}
-
-\FloatBarrier
 \section{Modelling}
@@
 workspaces~\cite{cranmer2012histfactory, heinrich2021pyhf}.
-\emph{Thanks to Jon~Long for collaboration in preparing this \hepdata\ archive,
+\emph{Thanks to Jonathan~Long for collaboration in preparing this \hepdata\ archive,
 and to Alessia~Murrone for invaluable feedback on, and validation of, its
@@
 \emph{
-Thanks to Eirik~Gramstad for developing the Matrix Method for our
+Thanks to Eirik~Gramstad for providing the Matrix Method for our
 $\twoljets$ searches, documenting it, providing its samples, and showing us how
@@
 difference (between central and high) for the combined high variation, and
-sets the low ($-1$-sigma) variation to zero.
+sets the low variation to zero.
 
@@
 Although $p$-values \emph{do} depend on what might have happened with
-alternative data
-(those with test statistics greater than or equal to that observed),
-it would be wasteful and unproductive to consider them seriously.
+alternative data, it is unproductive to consider them seriously.
 I am grateful that \atlas's statistical conventions allow us to ignore these
@@
 Amusingly, tuning to data specific to each experiment means that each
-experiment can simulate different fundamental physics~\cite{
+experiment simulates different fundamental physics~\cite{
 herwig2000parameter,
@@
 \end{itemize}
-An `if A then B' statement does not define what to do if A is false, so
-this recipe does not define what sign to give $\sigma$ if the deviation is not
-statistically significant.
+An `if $\mathrm{A}$ then $\mathrm{B}$' statement does not define what to do if
+$\mathrm{A}$ is false, so this recipe does not define what sign to give
+$\sigma$ if the deviation is not statistically significant.
 Since all variations have direction
@@
 with calibrated noise to approximate reconstruction-level results.
-\emph{Thanks to Jon~Long for preparing our initial
+\emph{Thanks to Jonathan~Long for preparing our initial
 \simpleanalysis\ implementations and sharing software to evaluate these
@@
 and $\pm13\%$ for \topother.
-\emph{Thanks to Jon~Long for researching and assigning these uncertainties.}
+\emph{Thanks to Jonathan~Long for researching and assigning these uncertainties.}
 
@@
 alternative simulations~\cite{Anders:2125718, atlas_twiki_susytheoretical}.
-\emph{Thanks to Jon~Long and Yumeng~Cao for collaboration in preparing these
+\emph{Thanks to Jonathan~Long and Yumeng~Cao for collaboration in preparing these
 variations}.
@@
 whose excellent scripts were adapted to simulate these signal
-samples, and thanks to Jon~Long for contributing to this study.
+samples, and thanks to Jonathan~Long for contributing to this study.
 }
@@
 Unfortunately, these requirements limit only \emph{when} data-driven estimation
-procedures are defined, and do not limit the \emph{what} those procedures
-\emph{do}.
+procedures are \emph{defined},
+and do not limit the \emph{what} those procedures \emph{do}.
 The rules explicitly allow for methods like the Matrix Method
@@
 
-To some extent, our published exclusion contours are an artistic choice.
+Our published exclusion contours are to some extent an artistic choice.
 What we know are the values of statistics in an irregular grid of
@@
 Not only do these simulations take computational time, they suffer
-torrential memory leaks that halt progress beyond small sample sizes.
+torrential memory leaks that halted progress beyond small sample sizes.
 To work around this, we evaluate and combine many small samples (with different
@@
 background model.
-Thankfully, this problem is somewhat remedied by the $\cls$-prescribed ratio
+Thankfully, this problem is remedied by the $\cls$-prescribed ratio
 against $\clsb$, since both $p$-values suffer from similar effects.
@@
 process of playing with them they begin to harden and set until they're
-immutable like rock, and at that point you can't change the world ---
+immutable like rock, and at that point you can't change the world ---\\
 not without breaking it into pieces and starting  fresh with new ideas.%

### 2022-09-30 main/draft.tex
resolve last todos; fix ref [1] noise

@@
 \madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
-\TODO{Should these posters be included as figures? Maybe appendix}
+\NO_DOTO{Should these posters be included as figures? Maybe appendix}
 However, real events have substantial additional noise from QCD
@@
 
-\TODO{\ldots}
+\NO_DOTO{\ldots}
 
@@
 % }
-\TODO{examples of use and misuse in hep}
+\NO_DOTO{examples of use and misuse in hep}
 
-\TODO{p-value footnote:
+\NO_DOTO{p-value footnote:
 The chance of a small \pvalue\ is small.
@@
 
-\TODO{log maximum likelihood ratio}
+\NO_DOTO{log maximum likelihood ratio}
 $\log$ maximum likelihood ratio $\kappa$
@@
 
-\TODO{tail area}
+\NO_DOTO{tail area}
 
@@
 \end{align}
-\TODO{significance measures}
+\NO_DOTO{significance measures}
 ~\cite{
@@
 
-\TODO{B-hadrons and LHCb?}
+\NO_DOTO{B-hadrons and LHCb?}
 % more cost-effective to improve calorimetry at the expense of PID
@@
 
-\TODO{
+\NO_DOTO{
 Argument:
@@
 Section~\ref{sec:context_atlas} then describes the \atlas\ detector.
-\TODO{more}
+\NO_DOTO{more}
 
@@
 
-\TODO{This does not belong in an introduction and will be moved.}
+\NO_DOTO{This does not belong in an introduction and will be moved.}
 
-\TODO{CUT THIS SHORT}
+\NO_DOTO{CUT THIS SHORT}
 

### 2023-02-11 main/frontmatter.tex
abstract tweaks

@@
 scattering to massive resonances with dramatic decays, which are predicted by
-supersymmetric models~\cite{atlas2022searches}.
+supersymmetric models.%
+% hack: hidden reference to constrain the reference numbering
+\sbox0{\cite{atlas2022searches}}
 
@@
 of proton collisions.
-Supersymmetry as a concept as not falsified by this lack of new phenomena,
+Supersymmetry as a concept is not falsified by this lack of new phenomena,
 nor by the many similar observations.
@@
 Friends and family made these years worthwhile:
-mu deepest love and thanks to all of you who have come and gone.
+my deepest love and thanks to all of you who have come and gone.
 

### 2023-02-11 main/chapter_theory.tex
chapter 1 text tweaks

@@
 Supersymmetry is an idea that has been applied to build plausible alternatives
-to the Standard Model which may explain future experimental discoveries.
+to the Standard Model that may explain future experimental discoveries.
 Since the main work of this thesis is a search for effects of supersymmetric
@@
 $SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
-meaning that its lagrangian is invariant under local operations of that group
+meaning that its Lagrangian is invariant under local operations of that group
 on its fields in certain representations.
@@
 interactions with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
-$SU\!(2)_L$ and $U\!(1)$ symmetries respectively.
+$SU\!(2)_L$ and $U\!(1)$ symmetries, respectively.
 These would be massless if the symmetry were respected, but it is
@@
 englert1964broken
-}%
+},
 which adds a complex doublet (equivalently four scalar fields) with an
@@
 We distinguish these two cases with the labels
-\emph{uncertain} and \emph{unknown} respectively:
-the top quark mass is slightly \emph{uncertain}, but
-the reconciliation of quantum field theory with general relativity is
+\emph{uncertain} and \emph{unknown}, respectively:
+the top quark mass is slightly \emph{uncertain},%
+\footnote{%
+The latest PDG meta-analysis estimates
+$m(t) = 172.69 \pm 0.30\,\eV[G]$~\cite{pdg2022ynf}.
+}
+but the reconciliation of quantum field theory with general relativity is
 mostly \emph{unknown}.
@@
 Various mechanisms could spontaneously break supersymmetry.
-Gravitational mechanism exist, but one model considered in this thesis uses
+Gravitational mechanisms exist, but one model considered in this thesis uses
 Gauge-Mediated Supersymmetry Breaking (GMSB),
@@
 }.
-Conservation of $R$-parity usefully preventing proton decays,
+Conservation of $R$-parity usefully prevents proton decays,
 but also makes the Lightest Supersymmetric Particle (LSP) a plausible
@@
 models, and might help us to understand future discoveries
-which could be made with data from the \atlas\ detector.
+that could be made with data from the \atlas\ detector.
 

### 2023-02-11 main/chapter_experiment.tex
add roi; more tweaks; fix make diff

@@
 \atlas\ has now produced a plethora of competitively precise measurements
-of the heavy Standard Model~\cite{atlas2021summarysm},
+of the heaviest particles in the Standard Model~\cite{atlas2021summarysm},
 including $m(W)$~\cite{atlas2018wmass} and
@@
 TOPQ-2020-03
-},
+}.
 These measurements continue to refine the Standard Model, and their precision
@@
 The final connection to data analysed in this thesis is made through the
-trigger system which is described in Section~\ref{sec:atlas_trigger},
+trigger system, which is described in Section~\ref{sec:atlas_trigger},
 and details of the data themselves are given in
@@
 The first-level (L1) trigger uses its fast preliminary information to accept
-interesting event and mark their regions of interest.
-Complete data in these regions of interest are passed to the second-level
+interesting events, and mark regions of interest.
+A Region-of-Interest (RoI) is an angular region of a detector subsystem.
+Each RoI is assigned in $\eta\textrm{--}\phi$ coordinates to contain detector materials
+that should contain more data on the effects seen in L1.
+RoI sizes vary from
+$\Delta\eta \times \Delta\phi \approx 0.1 \times 0.1$ for electron/photon
+reconstruction, up to
+$\Delta\eta \times \Delta\phi = 0.8 \times 0.8$ for
+hadronic jets~\cite{atlas2016trigger}.
+Complete data in the regions of interest are passed to the second-level
 (high-level trigger, HLT) to inform its decision of whether to accept each
 event.
+To reduce processing time, HLT algorithms can choose to inspect data only in,
+or near to, each regions of interest.
 Events accepted by the HLT are recorded for future analysis~\cite{

### 2023-02-11 main/chapter_searches.tex
add roi; more tweaks; fix make diff

@@
 \subsection{Probability: abstract}
-A probability $\prob{b}{a}$ is a number which quantifies the proportion
+A probability $\prob{b}{a}$ is a number that quantifies the proportion
 of $b$ within $a$, and can be expressed as a ratio
@@
 Other likelihoods that are not binary act as inefficient filters
-which reject events with some probability.
+that reject events with some probability.
 
@@
 Posterior probabilities are model-dependent objects that can usefully
-digesting data to make inferences, but are not pure statements of the data ---
+digest data to make inferences, but are not pure statements of the data ---
 reporting likelihood functions and evidence values allows readers to assign

### 2023-02-11 main/chapter_theory.tex
add roi; more tweaks; fix make diff

@@
 Supersymmetry is an idea that has been applied to build plausible alternatives
-to the Standard Model that may explain future experimental discoveries.
+to the Standard Model. and that may explain future experimental discoveries.
 Since the main work of this thesis is a search for effects of supersymmetric

### 2023-02-11 main/chapter_searches.tex
tweak searches

@@
 ,
-% \textbf{likelihood} \times \textbf{prior}
-% &= \textbf{posterior} \times \textbf{evidence} .\nonumber
 \end{align}
@@
 Equation~\ref{eqn:searches_pvalue_likelihood} points out that the \pvalue\ is
-just is the likelihood plus some fluff from a tail of larger test statistics
+just the likelihood plus some fluff from a tail of larger test statistics
 (if the test statistic distribution is discrete).
@@
 }
-A parameter corresponds to an hypothesis that assigns a prior distribution to
+A parameter corresponds to a hypothesis that assigns a prior distribution to
 plausible data, and those data for which an obeisant researcher accepts the
@@
 and is plausibly explained by mutations from vaguely accurate statements like
-``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
+``compatible with the SM prediction with a p-value of $0.10\%$''~\cite{
 lhcb2021test
@@
 \subsection{\texorpdfstring{$\cls$}{CLs}}
+\label{sec:searches_cls}
 Such misrepresentations of our results reinforces the need for clearer
@@
 How can one escape this trap?
-Likelihood ratios on the actual data escape immediately by bypassing the
+Likelihood ratios on the actual data escape immediately by by-passing the
 censor, but even inequalities need not be biased if models can ask
@@
 In such general cases, maximum likelihood ratio test statistics,
-such as $\Lambda$ of Equation~\ref{eqn:searches_max_like_ratio}, often useful.
+such as $\Lambda$ of Equation~\ref{eqn:searches_max_like_ratio}, are often
+useful.
 With a single bin, $\Lambda$ shuffles the ordering of
@@
 parameter, and can include likelihoods from the signal-plus-background
-model with constrained maximizations which aim to reduce the impact of
+model with constrained maximizations, which aim to reduce the impact of
 unphysical results such as negative Poisson expectations~\cite{
@@
 This transformation of priors to auxiliary likelihoods is generally possible
---- in a prior density $f(x)\,\mathrm{d}x$ the function $f(x)$ is a likelihood
+--- in a prior density $f(x)\,\mathrm{d}x$, the function $f(x)$ is a likelihood
 (strictly, a Radon-Nikodym derivative~\cite{billingsley2008probability})
@@
 \end{equation}
-The minimization objective $E$ is the summed negative $\log$ likelihood
-(or regularizer) for all actual and auxiliary data.
+The minimization objective $E$ is the summed negative $\log$ \heplikelihood\
+for all actual and auxiliary data.
 We choose the letter $E$ for \textbf{energy}, to highlight the analogy with
 statistical physics and the Boltzmann factor
-$f(x) \propto e^{-E(x)/T}$,
+$f(x) \propto e^{-E(x)/(k_B T)}$,
 with which a thermodynamic system collapses to its minimum energy at cold
@@
 but in my experience with \heplikelihood\ models such distributions often
-coincide with neither intuition nor the best fit value.
+coincide with neither intuition nor the best fit value.%
+\footnote{%
+One can reproduce this experience by sampling a public \pyhf\ \heplikelihood\
+model
+(for example, try SR-HM(disc.) of $1\ell bb$~\cite{SUSY-2019-08,hepdata.90607})
+with standard Markov-Chain Monte Carlo algorithms.
+I recommend first estimating the covariance by expanding about the maximum
+\heplikelihood, and linearly transforming your coordinates to whiten that
+covariance to the identity matrix.
+Then, algorithms provided by
+TensorFlow Probability~\cite{tensorflow2015-whitepaper}
+(the Metropolis-Adjusted Langevin Algorithm~\cite{mala1998, XIFARA201414}
+is a good start)
+can sample quite efficiently with tensor evaluation through
+\pyhf~\cite{heinrich2021pyhf}.
+Histograms of samples for the signal strength in the discovery fits
+can differ substantially from profile-\heplikelihood\ curves.
+}
 
@@
 $\prob{t \geq t(d)}{f(d, h)}$.
-And since the \pvalue\ is a censored likelihood, this is an error of early
-unblinding, or, in machine-learning speak, training on the test set.
+And since the \pvalue\ is a censored likelihood value%
+\footnote{%
+As argued in Section~\ref{sec:searches_cls}.%
+}, this is an error of early
+unblinding, or, in machine-learning-speak, training on the test set.
 It can make little difference when the yields are insensitive to the observed
@@
 factors.
-Likelihoods for validation region data are not included in the \heplikelihood\
-models, but their data are important and indirectly used to update the
-model in this manner.
+No Poisson likelihoods for validation region data are included in the
+\heplikelihood\ models, but their data are important and indirectly used to
+update the model in this manner.
 

### 2023-02-25 main/chapter_2ljets.tex
2ljets up to p64

@@
 Example effects of these signal models are displayed in signal regions of the
-$\twoljets$-electroweak search are displayed in
+$\twoljets$-electroweak search in
 Figure~\ref{fig:2ljets_signal_examples}.
@@
 cross-sections are varied up and down by factors intended to describe the size
-theoretical uncertainties.
+of the theoretical uncertainties.
 \\[0.5em]
@@
 All contours are interpolated from a sparse grid.
+The mass splitting is notated by
+$\Delta m = m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
 }
@@
 from asymptotic approximations to the distribution of the test statistics.
-Grey areas are observed limits from the two-lepton results from previous
+No previous limits have been set on this model by two-lepton
 \atlas\ searches~\cite{atlas_23l_SUSY_2016_24, atlas_2l_SUSY_2013_11}.

### 2023-02-25 main/chapter_2ljets.tex
2ljets p65

@@
 $S_{\mathrm{obs}}^{95}$ in the figure.
+\footnotetext{%
+The visible cross-section is defined such that the expected event count is
+given by the product of $A\epsilon\sigma\times L$, where $L$ is the integrated
+luminosity (typically in $\mathrm{fb}^{-1}$ units).
+This particular presentation is uses an `acceptance' $A$, an
+`efficiency' $\epsilon$, and an observable cross-section $\sigma_{\mathrm{obs}}$
+(presented here at at $95\%$ limit).
+The acceptance and efficiency are defined using truth-level simulations, in
+which the $A$ is the fraction of truth-level events that pass selection
+requirements, and $\epsilon$ is the ratio from the truth-level yield to the
+reconstruction-level yield.
+Because detector effects can increase yields, this $\epsilon$ can exceed
+one~\cite{susy_results_presentation_papers}.
+}
 
@@
 % https://stackoverflow.com/a/2132998
-\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
+\begin{tabular*}{\textwidth}{l|r@{$~\pm~$}lcccccc}
 & \multicolumn{2}{c}{Fitted}
@@
 & $p_{s\!=\!0}$
-\\[1.5ex]
+\\
+\hline
 DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
@@
 the observed $95\%$ $\cls$ upper limit on the visible cross-section
-$A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$,
+(labelled $A\epsilon\sigma_{\mathrm{obs}}^{95}\,\mathrm{fb}$\footnotemark),
 its corresponding signal expectation $S_\mathrm{obs}^{95}$,
@@
 $\pm1$-sigma variations,
-$\clb$ evaluated with the signal expectation set to the observed upper limit,
+$\clb$ (as defined in Section~\ref{sec:searches_cls}) evaluated with the signal
+expectation set to the observed upper limit,
 and the discovery $p$-value capped at $0.5$.

### 2023-02-25 main/chapter_2ljets.tex
up to p102

@@
 \item Production of main data inputs for all three $\twoljets$ searches
-and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09}.
+and the Recursive Jigsaw Reconstruction (RJR) $3\ell$
+search~\cite{atlas_rjr_3l_SUSY_2019_09}.
 Production of all experimental systematic variations for the background
@@
 this search project exists to supplement those results with information from
-the full LHC Run~2 data-set that was collected between in the years $2015$
+the full LHC Run~2 data-set that was collected in the years $2015$
 through to $2018$ and contains
@@
 Leveraging the increased data quantity is one key contribution of $\twoljets$.
-How well those data are used depends on how well the search designed, so
+How well those data are used depends on how well the search is designed, so
 $\twoljets$ further benefits from lessons learned from the experiences of
@@
 \emph{This figure was made by the RJR team.}~\cite{atlas2022searches}.
-Both diagrams represent processes which are targetted in the electroweak
+Both diagrams represent processes that are targetted in the electroweak
 analysis.
@@
 Summary plots for RJR searches:
-(a) and (b)~two- and three- lepton selections respectively,
+(a) and (b)~two- and three- lepton selections, respectively,
 \emph{reproduced from}~\cite{atlas_rjr_23l_SUSY_2017_03},
@@
 and opposite charge, like those from $Z$ boson decays, but they have different
-mass dilepton mass characteristics.
+dilepton mass characteristics.
 
@@
 \centering
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.62\textwidth}
 \centering
 \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_gluino_slepton.pdf}
-\caption{gluino-slepton}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
+\end{subfigure}
+\\
+\begin{subfigure}{0.62\textwidth}
 \centering
 \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_gluino_z.pdf}
-\caption{gluino-$Z^{(*)}$}
-\end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.48\textwidth}
+\end{subfigure}
+\\
+\begin{subfigure}{0.62\textwidth}
 \centering
 \includegraphics[width=\textwidth]{figures/2ljets_strong_contours_squark_z.pdf}
-\caption{squark-$Z^{(*)}$}
 \end{subfigure}
@@
 \\[0.5em]
+(Top) gluino-slepton,
+(Middle) gluino-$Z^{(*)}$,
+(Bottom) squark-$Z^{(*)}$.
+\\
 Each model point uses fit results from the signal region with the
@@
 relying less on loosely justified extrapolations,
-and orthogonal regions have independent likelihood functions that can be
-cleanly combined, by multiplication, to leverage the joint information in their
-data.
+and orthogonal regions have independent Poisson likelihood functions that can
+be cleanly combined, by multiplication, to leverage the joint information in
+their data.
 
@@
 search~\cite{atlas_23l_SUSY_2016_24} chose one signal region for each signal
-point from a measure expected sensitivity measure and drew contours around
+point from an expected sensitivity measure, and drew contours around
 results from the chosen fit results.
@@
 various variables.
-This leads to specifications which are challenging to present or understand,
+This leads to specifications that are challenging to present or understand,
 and which require region names with numerous suffixes to specify where they
@@
 $\met > 100\,\eV[G]$~\cite{atlas_twiki_lowest_unprescaled}.
-Leptons, however, are striking signals which are produced as smaller background
-rates, as they suppressed by electroweak couplings.
+Leptons, however, are striking signals that are produced at smaller
+background rates, as they are suppressed by electroweak couplings.
 \atlas\ can therefore trigger on leptons with much softer requirements;
-lepton triggers allow us to use softer jets and $\met$, that are relied upon
+lepton triggers allow us to use softer jets and $\met$, which are relied upon
 throughout the $\twoljets$ analyses.
 Although $\twoljets$-electroweak requires $\met > 100\,\eV[G]$ in all of its
-regions, the same is not true of $\twoljets$-RJR not various validation
+regions, the same is not true of $\twoljets$-RJR, nor various validation
 studies.
@@
 As mentioned at the beginning of this Section~\ref{sec:2ljets_events},
-`leptons' means `electron-or-muons' in the context of $\twoljets$,.
+`leptons' means `electron-or-muons' in the context of $\twoljets$.
 This is because taus leptons decay before reconstruction so are used only
@@
 
-Isolation of leptons from nearby hadronic activity is another feature which
+Isolation of leptons from nearby hadronic activity is another feature that
 can be used to reduce misidentification rates.
@@
 to have exactly two baseline leptons and exactly two signal leptons, meaning
-that the event has two reconstructed leptons which pass both sets of criteria.
+that the event has two reconstructed leptons that pass both sets of criteria.
 This vetoes events with soft or low quality third leptons, ensuring that those
@@
 Larger radius jets can also use taggers for heavy decays, such as those from
-$W$, Higgs or top resonance~\cite{JETM-2018-06}, but large radius were
+$W$, Higgs or top resonance~\cite{JETM-2018-06}, but large radius jets were
 considered too much additional work to be used in $\twoljets$,
@@
 or a calorimeter energy cluster could be attributed to both an electron and a
-photon. (Although we do not directly use photons in $\twoljets$, we do not veto
+photon. Although we do not directly use photons in $\twoljets$, we do not veto
 events containing them --- photons are accounted for in the momentum balance
-of $\ptmiss$.)
+of $\ptmiss$.
 
@@
 Every event has $\ptmiss$.
-Unlike leptons, jets, and other objects which we must select for, the $\ptmiss$
+Unlike leptons, jets, and other objects that we must select for, the $\ptmiss$
 of an event is a natural result of the conservation of momentum, which \atlas'
@@
 Our supersymmetric signals, however, do.
-This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
+This makes the missing transverse momentum $\ptmiss$ and its magnitude $\met$
 important features for separating signal and background events in all parts
@@
 The experimental resolution of $\ptmiss$, however, is imprecise, and
-importantly that resolution varies significantly from one event to another.
+importantly varies significantly from one event to another.
 
@@
 \centering
-\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_logy.png}
-\hfill
-\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_sig_logy.png}
+\includegraphics[width=0.62\textwidth]{figures/2ljets_presel_met_logy.png}
+\\
+\includegraphics[width=0.62\textwidth]{figures/2ljets_presel_met_sig_logy.png}
 \caption[
@@
 events at smaller $\met$.
-\\[0.5em]
+\\
 This selection requires two same-flavour opposite-sign signal leptons with
@@
 $\metsig > 6$.
-\\[0.5em]
+\\
 ``Significance'' displayed in the lower subplot uses the Binomial significance
@@
 the alternative jet assignments had some potential to define secondary signal
-regions to catch those signal events which fall outside the primary
+regions to catch those signal events that fall outside the primary
 $\mjj$ window.
@@
 \vspace{0.5em}
-\item $\dphillmet$: azimuthal angle between the lepton pair and $\ptmiss$
+\item $\dphillmet$: azimuthal angle between $\vec p_{\ell\ell}$ and $\ptmiss$,
+where $\vec p_{\ell\ell}$ is the vector sum of the momenta of the leptons in
+the lepton pair.
 \item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
@@
 larger $\metsig$.
-Initial binning in $\metsig$ therefore offers improves sensitivity to
+Initial binning in $\metsig$ therefore offers improved sensitivity to
 various parameter points.
@@
 designs, to Ben~Hooberman for suggesting the investigation of massive jets,
-and to the whole $\twoljets$ team for collaboration in their finalization.5%
+and to the whole $\twoljets$ team for collaboration in their finalization.%
 }
@@
 additional sensitivity in High regions by loose requirements
-of $\mll \in 71\textrm{--}111\,\eV[G]$ and $\nbtag \leq 1$.
+of $\mll \in 71\textrm{--}111\,\eV[G]$%
+\footnote{%
+We use interval notation in which
+en-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$, and
+concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
+with boundaries at $a$, $b$, and $c$.
+This notation is consistent with the $\twoljets$
+paper~\cite{atlas2022searches}.
+}
+and $\nbtag \leq 1$.
 This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
@@
 $\metsig$ bins are suffixed -a and -b for $\metsig \in 18\textrm{--}21$
-and $\metsig > 21$ respectively.
+and $\metsig > 21$, respectively.
 }
@@
 \centering
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.62\textwidth}
 \centering
@@
 \end{subfigure}
-\hfill
-\begin{subfigure}{0.48\textwidth}
+\\
+\begin{subfigure}{0.62\textwidth}
 \centering
@@
 
-The high regions, which defined in Table~\ref{tab:2ljets_high}, comprise SR-High,
-SR-High-1J, and \srllbb.
+The high regions, which are defined in Table~\ref{tab:2ljets_high}, comprise
+SR-High, SR-High-1J, and \srllbb.
 Their wide standard $\mjj \in 60\textrm{--}110\,\eV[G]$ requirement
-captures both $W\to jj$ and $Z\to jj$ for C1N2 and GMSB models
-respectively and is applied similarly to all other on-shell
+captures both $W\to jj$ and $Z\to jj$ for C1N2 and GMSB models,
+respectively, and is applied similarly to all other on-shell
 $\twoljets$-electroweak signal regions.
@@
 \centering
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \hfill
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering

### 2023-02-25 main/chapter_2ljets.tex
enlarge figures; progress to p110

@@
 \centering
-\includegraphics[width=0.48\textwidth]{figures/2ljets_vr_llbb_met_sig.pdf}
+\includegraphics[width=0.62\textwidth]{figures/2ljets_vr_llbb_met_sig.pdf}
 \caption[
@@
 Some figures label \crvz\ as `\crz Z'; the region was renamed late in this
-analysis project when it was show to contain substantial fractions of
+analysis project when it was shown to contain substantial fractions of
 $W\kern-0.4exZ$ as well as $ZZ$, as illustrated in
@@
 background becomes large since its fake $\met$, which arises from
-from mismeasurements or rare interactions, is no longer eliminated.
+mismeasurements or rare interactions, is no longer eliminated.
 Backgrounds from $\ttbar$ similarly increase, and the selections inherited from
@@
 Similarly to $\rjj$ in the High regions, this can usefully select signals,
-but here also works to reject other $\ell\ell$ production mechanisms which
+but here also works to reject other $\ell\ell$ production mechanisms that
 produce leptons in less correlated directions, particularly dileptonic
 $\ttbar$.
-As elsewhere, sensitivity is increased by binning SR-Low with
+As elsewhere, sensitivity is increased by binning SR-Low
 to form SR-Low-a with $\metsig \in 6\textrm{--}9$ and
@@
 exactly two jets.
-Abandoned region designs which used more jets are discussed below in
+Abandoned region designs, which used more jets, are discussed below in
 Section~\ref{sec:2ljets_low_isr}.
@@
 \centering
-\begin{subfigure}{0.38\textwidth}
+\begin{subfigure}{0.46\textwidth}
 \centering
@@
 \end{subfigure}
-\quad\quad
-\begin{subfigure}{0.38\textwidth}
+\hfill
+\begin{subfigure}{0.46\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.38\textwidth}
+\begin{subfigure}{0.46\textwidth}
 \centering
@@
 \centering
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \includegraphics[width=\textwidth]{figures/2ljets_C1N2_WZ_200p0_100p0_2L2J_presel_jetM[0]_50.png}
@@
 \\[1.5ex]
-\begin{subfigure}{0.48\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering

### 2023-02-25 main/chapter_2ljets.tex
up to p170

@@
 
-The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low;
+The gap in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low;
 their union therefore has a continuous $\mjj$ distribution that is displayed
@@
 overestimated.
-This plot contradicts that: most of the $\zjets$ distribution is fine, and
-the strange data deficit happens where \diboson\ becomes large near $m(W)$.
+This plot contradicts that: most of the $\zjets$ distribution is fine;
+the strange data deficit is local to $\mjj \approx m(W)$, which is where the
+\diboson\ sample contributes a majority of the simulated background.
 
@@
 These are CR-Low-ISR-Z and CR-Low-ISR-DF, which select $\zjets$ and $\ttbar$
-backgrounds and find smaller deficits of $20\%$ and $10\%$ respectively,
+backgrounds and find smaller deficits of $20\%$ and $10\%$, respectively,
 as shown in Figure~\ref{fig:2ljets_low_isr_old_more}.
@@
 
-A complicated likelihood model was constructed which normalized \diboson,
+A complicated likelihood model was constructed, which normalized \diboson,
 $\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,
@@
 Section~\ref{sec:2ljets_low} for Low regions,
-and which seeks hard jets which boost the supersymmetric particles and their
+and which seeks hard jets that boost the supersymmetric particles and their
 decay products.
@@
 \centering
-\begin{subfigure}{0.62\textwidth}
+\begin{subfigure}{0.7\textwidth}
 \centering
@@
 \\
-\begin{subfigure}{0.62\textwidth}
+\begin{subfigure}{0.7\textwidth}
 \centering
@@
 Various approaches have been taken to discovery regions in \atlas\ searches
---- although it is normal to have one per signal regions~\cite{SUSY-2018-02},
+--- although it is normal to have one per signal region~\cite{SUSY-2018-02},
 some searches study dozens of discovery regions comprising signal region
@@
 ~\ref{fig:2ljets_disco_trials_llbb}~and~\ref{fig:2ljets_disco_trials_int}
-for High, $\llbb$ and Int categories respectively.
+for High, $\llbb$ and Int categories, respectively.
 Similar figures were produced for the other categories, but they are used older
@@
 Although prompt strong effects may only rarely fake leptons, they have very
-large cross-sections and are not exoected to be modelled well.
+large cross-sections and are not expected to be modelled well.
 This motivates the use of data-driven FNP estimates to replace the need for
@@
 \times \prob{t}{r,2}$
-in which the numerical index points to the first and second leptons
+in which the numerical index points to the first and second leptons,
 respectively.
@@
 For comparison, this table also includes the simulated background without
-FNP leptons under the heading `MC Prompt' and the and only
-the simulated FNP contributions under the heading `MC FNP'.
+FNP leptons under the heading `MC Prompt' and the only
+simulated FNP contributions under the heading `MC FNP'.
 Negative or void estimates are an easy consequence of the use of signed
@@
 difference (between central and high) for the combined high variation, and
-sets the low variation to zero.
+set the low variation to zero.
 
@@
 An extraordinary feature of this analysis is the use of $\mjetone$ for
-SR-High-1J and VR-High-1J which, as described in Section~\ref{sec:2ljets_high},
+SR-High-1J and VR-High-1J, which, as described in Section~\ref{sec:2ljets_high},
 require a boosted, small-radius jet with a large mass
@@
 \item The $\zjets$ sample is normalized to \crz\ with an unconstrained
-normalization factor. This means that variations to the $\zjets$ sample in \crz
-are absorbed by scaling all of its modifications in other regions.
+normalization factor. This means that variations to the $\zjets$ sample in
+\crz\ are absorbed by scaling all of its modifications in other regions.
 \item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
@@
 deficit from the paper.
-Despite bring real and significant, this deficit was erased from public view
+Despite being real and significant, this deficit was erased from public view
 and, to my knowledge, is only reported through inner details of
@@
 To mitigate this, we instead take the variations from looser regions
-VR-sys-main and VR-sys-OffShell, that target on- and off-shell regions
-respectively and hope to improve the statistical precision.
+VR-sys-main and VR-sys-OffShell, that target on- and off-shell regions,
+respectively, and hope to improve the statistical precision.
 Both of these loose VR-sys-* regions require the pre-selection from
@@
 In the looser regions VR-sys-main and VR-sys-OffShell,
-which receive $1784$ and $7015$ truth events respectively, each variation
+which receive $1784$ and $7015$ truth events, respectively, each variation
 reduces to around $10\%$ and remains dominated by the statistical uncertainty.
@@
 From these benchmark signal models, we evaluate a relative uncertainty due
-due to each component, and extrapolate those variations from these benchmark
+to each component, and extrapolate those variations from these benchmark
 examples to others in the parameter plane.

### 2023-02-25 main/chapter_2ljets.tex
finish text tweaks

@@
 
-Although one might expect our selected \diboson\ sample to comprise mostly
-$ZZ$, with one $Z\to \ell\ell$ and the other invisibly to
-$Z\to \nu\nu$ to generate our required $Z$-like lepton pair with
-$\met$.
-Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the three-lepton
-$W\kern-0.4exZ$-like process has similar equal yields.
+One might expect our selected \diboson\ sample to comprise mostly
+$ZZ$, with one $Z\to \ell\ell$ to yield our required lepton pair, and the other
+decaying invisibly as $Z\to \nu\nu$ to yield our required $\met$.
+However, Figure~\ref{fig:2ljets_splits_diboson} shows that in fact the
+three-lepton $W\kern-0.4exZ$-like process has similarly large yields.
 Those $W\kern-0.4exZ$ events must therefore have a missing third lepton from
@@
 But they unblind the data, notably in high-$\met$ where we expect signals.
-When commenting on this, the resonse was that it does not count as unblinding
+When commenting on this, the response was that it does not count as unblinding
 because we are inspecting a loose selection.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/2ljets_fit_corrmatrix.pdf}
+\includegraphics[height=\textwidth,angle=90,origin=c]{figures/2ljets_fit_corrmatrix.pdf}
 \caption[
@@
 
-Parameters named ``\texttt{$\alpha$\_*}'' are centred on zero with unit initial width,
-and others are initially relative to $1$ with variable error bars.
+Parameters named ``\texttt{$\alpha$\_*}'' are centred on zero with unit initial
+width, and others are initially relative to $1$ with variable error bars.
 Within the \atlas\ SUSY group, plots like this are considered good if all
@@
 \centering
-\includegraphics[height=0.81\textheight]{figures/2ljets_fit_parameters.pdf}
+\adjincludegraphics[width=0.495\textwidth,trim={0 {0.49\height} 0 0},clip]{figures/2ljets_fit_parameters.pdf}
+\hfill
+\adjincludegraphics[width=0.495\textwidth,trim={0 0 0 {0.49\height}},clip]{figures/2ljets_fit_parameters.pdf}
 \caption[
@@
 since it exceeds $1.5$, it otherwise disappeared off the edge of this plot.
+\\[0.5em]
+This figure is split to increase its scale; the top of its right half continues
+from the bottom of its left half.
 }
@@
 What should perhaps be the most meaningful result of each discovery region is
-is observed upper limit $S_{\mathrm{obs}}^{95}$.
+observed upper limit $S_{\mathrm{obs}}^{95}$.
 The analysis of each region sets this upper limit by scanning different values
@@
 $\mu_\textrm{discovery}=3$, $\mu_\textrm{discovery}=10$,
-and $\mu_\textrm{discovery}=14$ respectively.
+and $\mu_\textrm{discovery}=14$, respectively.
 \\[0.5em]

### 2023-03-05 main/chapter_theory.tex
begin sm particle properties table

@@
 right-chiral neutrinos is not strictly confirmed.%
+See Table~\ref{tab:theory_particles_sm_properties} for details of their
+properties.
 }
@@
 for improvement in the next section.
+
+% using https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\centering
+\begin{tabular}{ccccccc}
+                            & Mass$~\eV[G]$                       & $Q_{EM}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\
+\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+$e$                         & $5.11\times10^{-4}$                 & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
+$\nu_e$                     & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\
+$\mu$                       & $105.7$                             & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
+$\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\
+$\tau$                      & $1.777$                             & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
+$\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\
+\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+$u$                         & $2.16^{+0.49}_{-0.26}\times10^{-3}$ & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
+$d$                         & $4.67^{+0.48}_{-0.17}\times10^{-3}$ & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
+$c$                         & $1.27 \pm 0.02$                     & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
+$s$                         & $93.4^{+8.6}_{-3.4}\times10^{-3}$   & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
+$t$                         & $172.7\pm0.3$                       & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
+$b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
+\multicolumn{1}{l}{Bosons}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+$g$                         & $0$                                 & $0$                  &                      &                      &                      &                      \\
+$W^+$                       & $80.4$                              & $+1$                 &                      &                      &                      &                      \\
+$W^-$                       & $80.4$                              & $-1$                 &                      &                      &                      &                      \\
+$Z$                         & $91.2$                              & $0$                  &                      &                      &                      &                      \\
+$\gamma$                    & $0$                                 & $0$                  &                      &                      &                      &                      \\
+$h$                         & $125.3 \pm 0.2$                     & $0$                  &                      &                      &                      &
+\end{tabular}
+\caption[Properties of Standard Model particles]{%
+Properties of Standard Model particles.
+% masses
+Masses are given as determined in a recent review by the
+Particle Data Group~\cite{pdg2022ynf}, which one should refer to for more
+precision and more uncertainty estimates.
+Uncertainty estimates are included only where they are relatively large.
+% quark mass schemes
+Quark masses are reported in the $\overline{\mathrm{MS}}$ scheme~\cite{
+pdg2022ynf,
+PhysRevD.18.3998
+}.
+% TODO reference for properties: textbooks?
+}
+\label{tab:theory_particles_sm_properties}
+\end{table}
 

### 2023-03-05 main/chapter_theory.tex
finish particle properties table

@@
 Table~\ref{tab:theory_particles_sm}.
+The masses and charges of these particles are further detailed in
+Table~\ref{tab:theory_particles_sm_properties}.
 
@@
 \begin{tabular}{ccccccc}
-                            & Mass$~\eV[G]$                       & $Q_{EM}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\
-\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+Fermions                    & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\[0.2em]
+\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]
 $e$                         & $5.11\times10^{-4}$                 & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
@@
 $\tau$                      & $1.777$                             & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
-$\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\
-\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+$\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\[1em]
+\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]
 $u$                         & $2.16^{+0.49}_{-0.26}\times10^{-3}$ & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
@@
 $s$                         & $93.4^{+8.6}_{-3.4}\times10^{-3}$   & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
-$t$                         & $172.7\pm0.3$                       & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
-$b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
-\multicolumn{1}{l}{Bosons}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-$g$                         & $0$                                 & $0$                  &                      &                      &                      &                      \\
-$W^+$                       & $80.4$                              & $+1$                 &                      &                      &                      &                      \\
-$W^-$                       & $80.4$                              & $-1$                 &                      &                      &                      &                      \\
-$Z$                         & $91.2$                              & $0$                  &                      &                      &                      &                      \\
-$\gamma$                    & $0$                                 & $0$                  &                      &                      &                      &                      \\
-$h$                         & $125.3 \pm 0.2$                     & $0$                  &                      &                      &                      &
+$t$                         & $172.69\pm0.30$                     & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
+$b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\[1em]
+\multicolumn{1}{l}{Bosons}  & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y$                  & $T^3$                &                      &                      \\[0.2em]
+$g$                         & $0$                                 & $0$                  & $0$                  & $0$                  &                      &                      \\
+$W^+$                       & $80.4$                              & $+1$                 & $0$                  & $+1$                 &                      &                      \\
+$W^-$                       & $80.4$                              & $-1$                 & $0$                  & $-1$                 &                      &                      \\
+$Z$                         & $91.2$                              & $0$                  & $0$                  & $0$                  &                      &                      \\
+$\gamma$                    & $0$                                 & $0$                  & $0$                  & $0$                  &                      &                      \\
+$h$                         & $125.3 \pm 0.2$                     & $0$                  & $+\frac{1}{2}$       & $-\frac{1}{2}$       &                      &
 \end{tabular}
 \caption[Properties of Standard Model particles]{%
-Properties of Standard Model particles.
+Properties of Standard Model particles~\cite{thomson2013modern,pdg2022ynf}.
 % masses
@@
 }.
-% TODO reference for properties: textbooks?
+% electric charge
+The electric charge is given by $Q_{\mathrm{EM}} = Y + T^3$; one can check
+that the left (L) and right (R) -handed fermions have equal electric charges
+corresponding to their left and right weak hypercharges ($Y_L$ and $Y_R$)
+and third weak isospin components ($T^3_L$ and $T^3_R$).
 }

### 2023-03-05 main/chapter_experiment.tex
add luminosity/hl-lhc note

@@
 $3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
-present and future high LHC luminosities,
+high luminosity of the current LHC,%
+\footnote{%
+Instantaneous luminosity in the Run~2 LHC peaked at
+$19\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2018~\cite{ATLAS:2022hro},
+which is much higher than previous conditions such as the Run~1 peak
+instantaneous luminosity of
+$3.6\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2012~\cite{ATLAS:2013tpw}.
+The future High Luminosity LHC (HL-LHC), however, might achieve
+$750\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
+For the HL-LHC, the current inner tracking and insertable B-layer will be
+replaced by a new Inner Trac\emph{k}er (ITk) which is designed to handle such
+extreme luminosity conditions~\cite{CERN-LHCC-2017-021}.
+}
 particularly for sensitivity to particles whose lifetimes are just right to

### 2023-03-05 main/chapter_theory.tex
tidy up table
add susy particles mass limits

@@
 Fermions come in left- and right-chiralities, although the existence of
-right-chiral neutrinos is not strictly confirmed.%
+right-chiral neutrinos is not strictly confirmed.
 See Table~\ref{tab:theory_particles_sm_properties} for details of their
-properties.
+properties.%
 }
@@
 \begin{tabular}{ccccccc}
-Fermions                    & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\[0.2em]
-\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]
+fermions                    & mass$\,\eV[G]$                      & $Q_{\mathrm{EM}}$    & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\[0.2em]
+leptons                     &                                     &                      &                      &                      &                      &                      \\[0.2em]
 $e$                         & $5.11\times10^{-4}$                 & $-1$                 & $-\frac{1}{2}$       & $-\frac{1}{2}$       & $-1$                 & $0$                  \\
@@
 $\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\[1em]
-\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]
+quarks                      &                                     &                      &                      &                      &                      &                      \\[0.2em]
 $u$                         & $2.16^{+0.49}_{-0.26}\times10^{-3}$ & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
@@
 $b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\[1em]
-\multicolumn{1}{l}{Bosons}  & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y$                  & $T^3$                &                      &                      \\[0.2em]
+bosons                      & mass$\,\eV[G]$                      & $Q_{\mathrm{EM}}$    & $Y$                  & $T^3$                &                      &                      \\[0.2em]
 $g$                         & $0$                                 & $0$                  & $0$                  & $0$                  &                      &                      \\
@@
 \footnote{%
-The latest PDG meta-analysis estimates
-$m(t) = 172.69 \pm 0.30\,\eV[G]$~\cite{pdg2022ynf}.
+The latest meta-analysis by the Particle Data Group estimates
+$m(t) = 172.69 \pm 0.30\,\eV[G]$~\cite{pdg2022ynf}, as listed in
+Table~\ref{tab:theory_particles_sm_properties}.
 }
@@
 }.
+Experimental data only provide lower bounds on the masses of these new
+particles, and those bounds depend strongly on the simplifying assumptions
+made in each case. Some moderately-general bounds have been assigned however,
+and examples of these bounds are listed in
+Table~\ref{tab:theory_particles_mssm_mass_limits}.
+Longer listings of limits in many special cases can be found in the
+Particle Data Group review~\cite{pdg2022ynf}.
+
+% made using:
+% https://www.tablesgenerator.com/latex_tables
+\begin{table}[tp]
+\centering
+\begin{tabular}{cc}
+fermions                                        & mass limit$\,\eV[G]$ \\[0.2em]
+$\neutralino_1$ (stable)                        & $> 46$         \\
+$\neutralino_1$ (unstable)                      & $> 380$        \\
+$\neutralino_2$,$\neutralino_3$,$\neutralino_4$ & $> 750$        \\
+$\chargino_1$,$\chargino_2$                     & $> 94$         \\[1em]
+bosons                                          & mass limit$\,\eV[G]$ \\[0.2em]
+$\tilde \nu$                                    & $> 94$         \\
+$\tilde e$                                      & $> 107$        \\
+$\tilde \mu$                                    & $> 94$         \\
+$\tilde \tau$                                   & $> 81.9$       \\
+$\tilde q$                                      & $> 1220$       \\
+$\tilde b$                                      & $> 1270$       \\
+$\tilde t$                                      & $> 1310$       \\
+$\tilde g$                                      & $> 2300$
+\end{tabular}
+\caption[%
+Mass limits for particles of the Minimal Supersymmetric Standard Model%
+]{%
+Mass limits for particles of the Minimal Supersymmetric Standard Model as
+reported under by the Particle Data Group under loose
+assumptions, quoted at ``$\mathrm{CL}=95.0\%$''~\cite{pdg2022ynf}.
+The charges of these particles equal the charges of their Standard Model
+partners, which are listed in
+Table~\ref{tab:theory_particles_sm_properties}.\\[0.5em]
+All of these limits assume R-parity conservation and various other,
+instance-specific constraints.
+For example, the limit of
+$m(\neutralino_2,\neutralino_3,\neutralino_4) > 750\,\eV[G]$
+assumes that $m(\neutralino_1) < 100\,\eV[G]$.
+Many stronger constraints are reported within the constraints of simplified
+models.
+Limits on R-parity violating models differ, and can be either stronger or
+weaker.
+In the stable $\neutralino_1$ case, the $\neutralino_1$ is the lightest
+supersymmetric particle and a dark matter candidate;
+in the unstable $\neutralino_1$ case, it rapidly decays.
+Stronger limits can be set for particles that live long enough to decay within
+the sensitive materials of collider detectors.\\[0.5em]
+All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
+}
+\label{tab:theory_particles_mssm_mass_limits}
+\end{table}
 

### 2023-03-05 main/chapter_2ljets.tex
squeeze a bit more figure size
leptons postit

@@
 \centering
-\begin{subfigure}{0.6\textwidth}
+\begin{subfigure}{0.62\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.52\textwidth}
+\begin{subfigure}{0.525\textwidth}
 \centering
@@
 % \hfill
-\begin{subfigure}{0.46\textwidth}
+\begin{subfigure}{0.465\textwidth}
 \centering
@@
 Other, more specialized, criteria exist in \atlas, but they are not relevant
-$\twoljets$.
+to $\twoljets$.
 
@@
 improved total sensitivity.
-Baseline requirements are softer and lower quality than signal requirements.
+Baseline requirements are softer and lower quality than signal requirements;
+the signal leptons are a subset of the baseline leptons.
+Therefore if a lepton is a signal lepton, then it is also a baseline lepton,
+but for clarity we often prefer to state that both requirements are met.
 
@@
 that the event has two reconstructed leptons that pass both sets of criteria.
-This vetoes events with soft or low quality third leptons, ensuring that those
-three-lepton events can be used independently in other studies.
+Requiring exactly two baseline leptons ensures that we veto events with soft
+or low quality third leptons, thus ensuring that our data do not overlap with
+other \atlas\ analysis that use three-lepton events.
+Enforcing no overlap in this way makes the analysis more independent, and
+therefore more cleanly combined in future meta-analyses.
 

### 2023-03-05 main/chapter_2ljets.tex
amenable
variables postit

@@
 Enforcing no overlap in this way makes the analysis more independent, and
-therefore more cleanly combined in future meta-analyses.
+therefore amenable to cleaner combinations in future meta-analyses.
 
@@
 thesis.
+This list is roughly sorted to have the most important variables at the top.
 \begin{itemize}
@@
 
+Broadly speaking, some of these variables select necessary conditions for
+signal-like events, and others help to improve signal purity within those
+selection requirements.
+Specifically, our signals produce $\met$, lepton pairs with
+$\mll \approx m(Z)$, and jets leading to positive $\njet$.
+Requiring appropriate conditions on these variables rejects large
+Standard Model background yields without harming signal efficiency.
+Other variables help to reduce background contamination with some trade-off
+against signal efficiency; among these, $\metsig$ is the most generally strong
+separating variable, and other variables are chosen in contexts where
+they aid the local designs that we detail in Section~\ref{sec:2ljets_design}.
+
 \FloatBarrier
@@
 This notation is consistent with the $\twoljets$
-paper~\cite{atlas2022searches}.
+paper~\cite{atlas2022searches}.%
 }

### 2023-03-05 main/chapter_2ljets.tex
postit p128

@@
 for High, $\llbb$ and Int categories, respectively.
-Similar figures were produced for the other categories, but they are used older
-region definitions, are identical in style, and are not particularly
-illuminating.
+In these figures, the upper panels histogram simulated yields from signal and
+background contributions, and the power panels show various sensitivity
+heuristics (significance measures) for which larger values indicate larger
+expected sensitivity to the signal in our future statistical analyses.
+We want to choose designs that have large expected sensitivity to a broad
+variety of interesting signal models.
+Similar figures were produced for the other categories, but they are not
+included because they used outdated region definitions, are identical in style,
+and are not particularly illuminating.
 
@@
 The region SR-High-8, labelled `\texttt{High8\_1+\_2}' here, is chosen for its
-simplicity and good sensitivity to both benchmark signal models.
-Upper plots show background and signal yields.
-Lower plots show various measures of projected
-sensitivity~\cite{cousins2008evaluation}.
+simplicity and good sensitivity to both benchmark signal models.\\[0.5em]
+Upper panels show expected background and signal yields from simulations, with
+all standard model backgrounds summed into ``Total Bkg'', and a $20\%$
+relative systematic uncertainty added in quadrature to the statistical
+uncertainty from the finite Monte Carlo simulations.
 Yields were derived in an early iteration of the analysis and do not exactly
-match those in the final result.
+match those in the final result.\\[0.5em]
+Lower panels show various measures of projected
+sensitivity~\cite{cousins2008evaluation}, in which larger values are intended
+to indicate a larger expected experimental sensitivity to the proposed signal
+being added to the background.
+All significance measures roughly agree with one another.
 }
@@
 The region \srllbb, labelled `SRllbb' here, is chosen for
-sensitivity, simplicity, and to avoid overlapping with DR-High.
-Upper plots show background and signal yields.
-Lower plots show various measures of projected
-sensitivity~\cite{cousins2008evaluation}.
+sensitivity, simplicity, and to avoid overlapping with DR-High.\\[0.5em]
+Upper panels show expected background and signal yields from simulations, with
+all standard model backgrounds summed into ``Total Bkg'', and a $20\%$
+relative systematic uncertainty added in quadrature to the statistical
+uncertainty from the finite Monte Carlo simulations.
 Yields were derived in an early iteration of the analysis and do not exactly
-match those in the final result.
+match those in the final result.\\[0.5em]
+Lower panels show various measures of projected
+sensitivity~\cite{cousins2008evaluation}, in which larger values are intended
+to indicate a larger expected experimental sensitivity to the proposed signal
+being added to the background.
+All significance measures roughly agree with one another.
 }
@@
 The region SR-Int, labelled `Int\_1$+$\_2' here, is chosen for its simplicity
-and broad sensitivity.
-Upper plots show background and signal yields.
-Lower plots show various measures of projected
-sensitivity~\cite{cousins2008evaluation}.
+and broad sensitivity.\\[0.5em]
+Upper panels show expected background and signal yields from simulations, with
+all standard model backgrounds summed into ``Total Bkg'', and a $20\%$
+relative systematic uncertainty added in quadrature to the statistical
+uncertainty from the finite Monte Carlo simulations.
 Yields were derived in an early iteration of the analysis and do not exactly
-match those in the final result.
+match those in the final result.\\[0.5em]
+Lower panels show various measures of projected
+sensitivity~\cite{cousins2008evaluation}, in which larger values are intended
+to indicate a larger expected experimental sensitivity to the proposed signal
+being added to the background.
+All significance measures roughly agree with one another.
 }

### 2023-03-05 main/chapter_2ljets.tex
postit 131: split figure, expand caption

@@
 \centering
-\begin{subfigure}{0.46\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \hfill
-\begin{subfigure}{0.46\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \end{subfigure}
-\\[0.5em]
-\begin{subfigure}{0.46\textwidth}
+\caption[
+Pre-fit yields in discovery regions in DR-High and DR-$\llbb$
+]{%
+Pre-fit yields in discovery regions in DR-High and DR-$\llbb$.
+Errors are statistical.
+These are single-bin histograms.
+Background yields are stacked and filled by colour.
+Signal yields appear as lines and would add to the backgrounds.
+Unblinded data are appear in black in regions that observe more than zero
+events.%
+}
+\label{fig:2ljets_disco_prefit_higher}
+\end{figure}
+
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \\[0.5em]
-\begin{subfigure}{0.46\textwidth}
+\begin{subfigure}{0.495\textwidth}
 \centering
@@
 \caption[
-Pre-fit yields in discovery regions
-]{%
-Pre-fit yields in discovery regions.
+Pre-fit yields in discovery regions in DR-Int, DR-Low, and DR-OffShell
+]{%
+Pre-fit yields in discovery regions in DR-Int, DR-Low, and DR-OffShell.
 Errors are statistical.
-}
-\label{fig:2ljets_disco_prefit}
+These are single-bin histograms.
+Background yields are stacked and filled by colour.
+Signal yields appear as lines and would add to the backgrounds.
+Unblinded data are appear in black in regions that observe more than zero
+events.%
+}
+\label{fig:2ljets_disco_prefit_lower}
 \end{figure}
@@
 Histograms of our five discovery regions are displayed in
-Figure~\ref{fig:2ljets_disco_prefit}
+Figures~\ref{fig:2ljets_disco_prefit_higher}
+and~\ref{fig:2ljets_disco_prefit_lower}
 with pre-fit backgrounds and statistical uncertainties only.
-The unblinded data in this figure repeat the SR-High and SR-Low deficits
+The unblinded data in these figures repeat the SR-High and SR-Low deficits
 mentioned previously.

### 2023-03-05 main/chapter_2ljets.tex
postit p162

@@
 where rate top processes contribute around $5\%$ of the total background,
-the \topother\ variation are
+the \topother\ variations are
 $\pm8\%$ (hard scatter) and
 $\pm13\%$ (parton shower).
+These variations are assigned from differences between yields from varied
+simulated samples.
 Most other regions have larger fractional errors due to our limited
@@
 \paragraph{Cross-sections.}
-As described in the paper~\cite{atlas2022searches},
+As described in the paper,
 background samples that are not normalized to data are assigned cross-section
 uncertainties as global scale-factors.
-These are $\pm10\%$ for Higgs processes,
+These uncertainties are assigned from theoretical uncertainties in
+the samples' computational simulation, and are $\pm10\%$ for Higgs processes,
 $\pm32\%$ for triboson,
 $\pm6\%$ for single-top,
-and $\pm13\%$ for \topother.
+and $\pm13\%$ for \topother~\cite{atlas2022searches}.
 \emph{Thanks to Jonathan~Long for researching and assigning these uncertainties.}

### 2023-03-05 main/chapter_2ljets.tex
postit p166

@@
 
-My extrapolation scheme is fairly arbitrary, and is performed as follows.
+Most extrapolation schemes are fairly arbitrary, and our choice here is taken
+for some heuristically attractive properties to be explained here.
 Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
@@
 }
+,
 \end{equation}
 in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
-physical understanding into the extrapolation,
-and the sum is over all benchmark points.
-
-Parent mass $m(\chargino_1, \neutralino_2) = m_a$
-and mass splitting $m_a - m_b$ are the most
-characteristic features in C1N2.
-From some experimentation, I found that a Manhattan distance in those features
-extrapolated nicely.
-Being a linear sum, the exponentiated Manhattan distance beneficially
-separating into a product of terms, so contours of constant $r_j(\vec x)$ are
+physical understanding into the extrapolation, and the sum is over all
+benchmark points.
+One benefit of this extrapolation scheme is that its results $r_j(\vec x)$
+are bounded by range of the observed points $r_{ij}$;
+alternatives, such as Radial Basis
+Functions~\cite{2020SciPy-NMeth,fasshauer2007meshfree}, can extrapolate without
+bound to large or small (negative) outputs, to extreme outputs, which do not
+seem physically justified in this case.
+
+This formulation of extrapolation as a weighted mixture has some freedom in
+which to assign the distance function $d(\vec x, \vec y)$, and within that
+freedom we can make choices that encode some physical understanding.
+In the case of our C1N2 model, two characteristic features are the
+parent mass $m(\chargino_1, \neutralino_2) = m_a$
+and the mass splitting $m_a - m_b$, and we want an extrapolation scheme which
+aligns changes in these variables with changes in the extrapolated value.
+From some experimentation, I found that a Manhattan distance in these features
+implements such attractive extrapolation:
+being a linear sum, the exponentiated Manhattan distance beneficially
+separates into a product of terms, so contours of constant $r_j(\vec x)$ are
 segmented into straight lines that relate to the chosen coordinates.
 Alternatives, such as the Euclidean distance, produced curved contours that did
-not align as well with my intuitions.
-Since the mass distance is dimensionful, it needs to be scaled to a reference
+not align as well with the $m_a$ and $m_a - m_b$ directions.
+Since the mass distance is dimensionful, it must be scaled to a reference
 value, which I chose as $100\,\eV[G]$; therefore

### 2023-03-12 main/chapter_2ljets.tex
begin scribbling some notes on future expectations
begin future passage

@@
 Here, `ISR' stands for Initial-State Radiation, following the naming
-of previous \atlas\ searches including partial Run-2 $2\ell$/$3\ell$
+of previous \atlas\ searches including partial Run~2 $2\ell$/$3\ell$
 search which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
@@
 the impressive off-shell exclusion contours are reproduced
-from the \atlas\ Run-2 $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09} in
+from the \atlas\ Run~2 $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09} in
 Figure~\ref{fig:ljets_offshell_3l_exclusion}.
@@
 \caption[
-Contours showing off-shell C1N2 exclusion from the \atlas\ Run-2 $3\ell$ search
+Contours showing off-shell C1N2 exclusion from the \atlas\ Run~2 $3\ell$ search
 ]{%
 Contours showing off-shell C1N2 exclusion in the three lepton channel,
-reproduced from the \atlas\ Run-2 $3\ell$
+reproduced from the \atlas\ Run~2 $3\ell$
 search~\cite{atlas_rjr_3l_SUSY_2019_09, hepdata.95751}.
@@
 
+% TODO future expectations
+% """
+% • Sec 4.8 (Discussion): Expand on future expectations e.g. luminosities,
+% heavier daughter particles with current data, HL-LHC and affect of pile-up,
+% FCC etc.
+% """
+
+No end is in sight for SUSY searches, but future LHC collision data will
+continue to push the high-mass frontier of sensitivity for the supersymmetric
+models considered in this thesis.
+% - second wave not happening
+\atlas\ has published `second-wave' Run~2 papers for some similar searches,
+which re-analyse the Run~2 data with refined techniques or more focussed
+objectives~\cite{
+ATLAS:2022hbt,
+ATLAS:2022nrb
+},
+but the recent date of $\twoljets$'s publication suggests that \atlas\ will
+use Run~3 data in its next paper on the topic.
+% - Run 3 expected luminosity
+% - HL-LHC
+TODO~\cite{ZurbanoFernandez:2020cco}
+% - If a future collider is constructed, then...
+
 Perhaps we \emph{should not} be restrained by crude statistical

### 2023-03-12 main/chapter_2ljets.tex
add some future lhc content

@@
 but the recent date of $\twoljets$'s publication suggests that \atlas\ will
-use Run~3 data in its next paper on the topic.
+use Run~3 data in its next paper on our topic.
 % - Run 3 expected luminosity
+In Run~3, the LHC is only operating at a marginally increased centre-of-mass energy
+of $\sqrt{s} = 13.6\,\eV[T]$~\cite{run3sqrts};
+its main benefit will therefore be the increased data quantity,
+for which a a recent projection indicates a target integrated luminosity of
+$250\,\mathrm{fb}^{-1}$ for $pp$ collisions by the year
+2025~\cite{run3luminosity},
+on top of the $139\,\mathrm{fb}^{-1}$ of Run~2.
+Given similar collision conditions, this will allow sensitivity to rare signals
+with proportionately small cross-sections.
+
 % - HL-LHC
+TODO
+- around $300\,\mathrm{fb}^{-1}$ to $400\,\mathrm{fb}^{-1}$ per year
+- pile-up of around $200$ interactions per bunch crossing
+- relative to around $50$ interactions per bunch crossing in Run~2~\cite{ATLAS:2022hro}
 TODO~\cite{ZurbanoFernandez:2020cco}

### 2023-03-12 main/chapter_2ljets.tex
add fcc etc writing

--- 2023-03-12 main/chapter_2ljets.tex
@@
 % • Sec 4.8 (Discussion): Expand on future expectations e.g. luminosities,
-% heavier daughter particles with current data, HL-LHC and affect of pile-up,
+% heavier daughter particles with current data,
+%
+% HL-LHC and affect of pile-up,
 % FCC etc.
@@
 which re-analyse the Run~2 data with refined techniques or more focussed
-objectives~\cite{
-ATLAS:2022hbt,
-ATLAS:2022nrb
-},
+objectives~\cite{ATLAS:2022hbt, ATLAS:2022nrb},
 but the recent date of $\twoljets$'s publication suggests that \atlas\ will
-use Run~3 data in its next paper on the topic.
+use Run~3 data in its next paper on our topic.
 % - Run 3 expected luminosity
-% - HL-LHC
-TODO~\cite{ZurbanoFernandez:2020cco}
-% - If a future collider is constructed, then...
+In Run~3, the LHC is only operating at a marginally increased centre-of-mass energy
+of $\sqrt{s} = 13.6\,\eV[T]$~\cite{run3sqrts};
+its main benefit will therefore be the increased data quantity,
+for which a a recent projection indicates a target integrated luminosity of
+$250\,\mathrm{fb}^{-1}$ for proton-proton collisions by the year
+2025~\cite{run3luminosity},
+on top of the $139\,\mathrm{fb}^{-1}$ of Run~2.
+Given similar collision conditions, this will allow sensitivity to rare signals
+with proportionately small cross-sections.
+
+% HL-LHC
+Following Run~3, the LHC will be upgraded to the
+High-Luminosity Large Hadron Collider (HL-LHC), whose improved operations
+should yield around $300\,\mathrm{fb}^{-1}$ to $400\,\mathrm{fb}^{-1}$ of
+$14\,\eV[T]$ $pp$ collisions per year.
+This accelerated data rates comes at the cost of quadrupled instantaneous
+luminosity (as described in Section~\ref{sec:atlas_inner}), which will push
+the pile-up background from around $50$ interactions per bunch crossing in
+Run~2 to around $200$ interactions per bunch crossing~\cite{
+ATLAS:2022hro,
+ZurbanoFernandez:2020cco
+}.
+The HL-LHC could yield $4000\,\mathrm{fb}^{-1}$ of proton-proton collisions,
+but for corresponding sensitivity, upgrades to \atlas's hardware and software
+must succeed in managing the increased pile-up background, both by
+identifying pile-up events and by modelling them in simulation.
+
+Collision energies ultimately bound the masses of heavy particles that can
+be produced in collisions, so stronger accelerators will be necessary
+to expand the search in that direction.
+Future prospects for the coming decades at CERN include the
+High-Energy LHC (HE-LHC), which could achieve $\sqrt{s}=27\,\eV[T]$ with new
+hardware in the current LHC tunnel~\cite{FCC:2018byv},
+and a Future Circular Collider (FCC), whose new $100\,\mathrm{km}$
+`quasi-circular' tunnel would permit proton collisions at
+$\sqrt{s}=100\,\eV[T]$~\cite{FCC:2018vvp}.
+Just as the LHC follows the LEP electron-positron collider, this FCC hadron
+collider (FCC-hh) might be proceeded by an electron-positron collider (FCC-ee)
+with centre-of-mass energies high enough (up to $\sqrt{s}=350\,\eV[T]$) to
+pair-produce all Standard Model particles~\cite{FCC:2018evy}.
+Such scrutiny of the Standard Model is also a search for new physical effects,
+and including some that could be explained with supersymmetric
+models~\cite{FCC:2018byv, Gounaris1999}.
 

### 2023-03-12 main/chapter_2ljets.tex
weaken diboson statement

@@
 \label{sec:2ljets_diboson_alt}
-No complete alternative samples were available for comparison with our central
-\diboson\ samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
-electroweak vertices.
-This limitation is simply a result of how \atlas\ has decided to allocate
-its limited resources on, and quite reasonably so, since full simulation is
-computationally expensive and \diboson\ processes are relatively rare.
-We did, however, have access to the cheaper alternative set of small,
+When executing this analysis project, no complete alternative \diboson\ samples
+were available in the specific post-processed format required for our analysis'
+software framework.
+Our central \diboson\ samples are simulated with \sherpa~2.2.2 at NLO with up
+to six electroweak vertices, and although \diboson\ sample comparisons are
+conditionally recommended by internal
+documentation~\cite{atlas_twiki_susytheoretical}, we failed to acquire
+alternative samples for this comparison.
+This limitation is in part a result of how \atlas\ and our local SUSY group has
+decided to allocate its limited resources, and quite reasonably so, since
+full simulation and data derivation is expensive in computation and data
+storage regards, and \diboson\ processes are relatively rare.
+Instead, we followed advice to use the cheaper alternative set of small,
 truth-level samples, that were simulated with the same \sherpa\ setup but with

### 2023-03-12 main/chapter_2ljets.tex
explain signal figures

@@
 aligns changes in these variables with changes in the extrapolated value.
-From some experimentation, I found that a Manhattan distance in these features
+From some experimentation, I found that a Manhattan distance in these variables
 implements such attractive extrapolation:
@@
 To account for its drastically different kinematics, the C1N2 off-shell space
-is treated separately.
+is treated separately;
+the boundaries of this off-shell space are marked with black lines.
+\\[0.4em]
+In these figures, the black points mark the locations where uncertainties have
+been evaluated, and the coloured regions indicate the extrapolated
+uncertainty assignments.
+Colour indicates the assigned uncertainty relative to the central
+signal yield, and its scale varies from $0$ to $100\%$ relative uncertainty
+in all figures.
+The coloured lines indicate contours of constant uncertainty assignment.
 }

### 2023-03-12 main/chapter_2ljets.tex
clarify no_sr summary plot differences

@@
 Figure~\ref{fig:2ljets_summary}.
+Unlike Figure~\ref{fig:2ljets_summary} the this figure does not adjust its
+background estimates to match the observed data in signal regions; notably,
+this figure has much larger background predictions in SR-Low-a and SR-Low-b,
+with correspondingly larger negative significances in those regions.
+The two figures also have aesthetic differences in the colouring and grouping
+of background samples.
+\\[0.4em]
 Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
@@
 Because the post-fit yields in SR-Low achieve a small negative significance,
-all mentions of its deficit were removed from the paper.
+all mentions of its data deficit were removed from the paper.
 }

### 2023-03-12 main/chapter_2ljets.tex
write overlap removal procedure

@@
 and $\pt$.
+
+This sequential overlap procedure is described as follows:
+\begin{enumerate}
+\item Remove any electron candidate that shares an inner-detector track with a
+baseline muon.
+This reduces backgrounds from events with
+$\mu\to\mu\gamma$, $\gamma\to e^+ e^-$ radiations in the detector.
+\item Remove any baseline jet within $\Delta R \leq 0.2$ of a baseline electron.
+\item Remove any electron within
+$\Delta R \leq \min(0.04 + (10\,\eV[G]) / p_T^e, 0.4)$
+of a baseline jet.
+This suppresses backgrounds from heavy-flavour hadron decays.
+\item Remove any baseline jet that is within $\Delta R \leq 0.2$ of a baseline
+muon, or that shares a track with a muon.
+\item Remove any muon within
+$\Delta R \leq \min(0.04 + (10\,\eV[G]) / p_T^\mu, 0.4)$
+of a baseline jet.
+This is matches the electron removal instruction above.
+\end{enumerate}
+The $\ptmiss$ calculation also ignores any electron that shares a calorimeter
+cluster with a photon~\cite{atlas2022searches}.
 

### 2023-03-12 main/chapter_2ljets.tex
explain why low-isr is included

@@
 Here, `ISR' stands for Initial-State Radiation, following the naming
-of previous \atlas\ searches including partial Run~2 $2\ell$/$3\ell$
-search which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
+of previous \atlas\ searches including the partial Run~2 $2\ell$/$3\ell$
+search, which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
 $\twoljets$-electroweak search.
+Although these Low-ISR regions were removed from the analysis and we never
+inspected the data in SR-Low-ISR (never unblinded), I include them here because
+they reveal some interesting context to the surviving analysis design.
+Specifically, prototype validation regions for Low-ISR designs revealed some
+bad modelling by our Standard Model background samples in phase space that is
+close to SR-Low, which is a region where we observe significant data deficits.
+Secondly, these regions were an impactful part of the analysis design process
+and led to the final design of SR-Low;
+describing my experiences with them therefore reveals more about the true
+design process.
 
@@
 
-A complicated likelihood model was constructed, which normalized \diboson,
+A complicated \heplikelihood\ model was constructed, which normalized \diboson,
 $\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,

### 2023-03-12 main/chapter_2ljets.tex
explain why offshell happened

@@
 }.
+
+Our $\twoljets$ data selections are sub-optimal for targetting compressed
+signal regimes because they require two hard leptons and only use two-lepton
+triggers, while compressed signals tend to produce soft decay products and
+can benefit from other $\met$ triggers for ISR scenarios.
+Therefore, the goal of this off-shell study cannot be to contribute
+competitive sensitivity.
+It is also not an independent verification of the dedicated compressed
+searches because our selections overlap.
+Our meagre remaining goals are therefore to discover what sensitivity is
+available in the $\twoljets$ data, and to provide a continuous extension of our
+exclusion contours across the $\Delta m = m(Z)$ threshold
+(as shown in the lower left of Figure~\ref{fig:2ljets_contours_c1n2}).
 

### 2023-03-12 main/chapter_2ljets.tex
clean up delta m and figure ref

@@
 available in the $\twoljets$ data, and to provide a continuous extension of our
-exclusion contours across the $\Delta m = m(Z)$ threshold
-(as shown in the lower left of Figure~\ref{fig:2ljets_contours_c1n2}).
+exclusion contours across the
+$m(\chargino_1, \neutralino_2) - m(\neutralino_1) = m(Z)$ threshold.
+The resulting extension appears in the lower left of
+Figure~\ref{fig:2ljets_contours_c1n2}.
 

### 2023-03-12 main/chapter_2ljets.tex
describe region subsets

@@
 
+The signal region subsets whose exclusions are presented in
+Figure~\ref{fig:2ljets_fit_multi_contours} are described as follows:
+\begin{itemize}
+\item `High' contains only the bins of SR-High and SR-High-1J.
+\item `HighNo4' contains only the bins of SR-High only.
+It excludes SR-High-1J such that the difference between `High' and `HighNo4'
+evaluates the impact of including the mono-jet region SR-High-1J.
+\item `$\llbb$' contains only SR-$\llbb$.
+It is not independently sensitive to the C1N2 models, but demonstrates the
+unique reach gained by SR-$\llbb$ for GMSB models.
+\item `Int' contains only the bins of SR-Int.
+\item `Low` contains only the bins of SR-Low.
+\item `OffShell` contains only the bins of SR-OffShell.
+\end{itemize}
+
 \begin{figure}[tp]
@@
 appropriate parameter spaces.
-Although figures were produced before unblinding, with older samples and region
-designs, their later updates do not change the general message.
-Values of $\cls$ from expected data in the combined fit are overlaid.
+Although these figures were produced before unblinding, with older samples and
+region designs, their later updates do not change the general message.
+Values of $\cls$ from expected data in the combined fit are overlaid in tiny
+grey text.
 }

### 2023-03-12 main/chapter_2ljets.tex
explain why no light gmsb

@@
 contamination does not entirely invalidate the result.
-We do not consider off-shell GMSB models, so SR-OffShell regions do not exclude
-any points alone here.
+We do not consider off-shell GMSB models because they are strongly excluded by
+existing data%
+\footnote{%
+Each leg of our GMSB decay process has $\neutralino_1\to h/Z\gravitino$ with a
+light $\gravitino$; $m(\gravitino) = 1$.
+An off-shell decay therefore requires
+$m(\neutralino_1) < m(Z) + 1\,\eV[G] \approx 92.2\,\eV[G]$.
+This simplified GMSB model assumes $m(\neutralino_1) \approx m(\chargino_1)$,
+and such a light $\chargino_1$ is excluded by LEP~2
+searches~\cite{lepsusycombined}.
+The CMS Collaboration has also published stronger constraints using proton
+collision data from Run~1 of the LHC\cite{
+cms_susy_2018_partial_run2_combination,
+PhysRevD.90.092007
+}.%
+},
+so SR-OffShell regions do not exclude any points alone here.
 
@@
 
-% TODO future expectations
-% """
-% • Sec 4.8 (Discussion): Expand on future expectations e.g. luminosities,
-% heavier daughter particles with current data,
-%
-% HL-LHC and affect of pile-up,
-% FCC etc.
-% """
-
 No end is in sight for SUSY searches, but future LHC collision data will

### 2023-03-12 main/chapter_2ljets.tex
explain (un)blinding

@@
 
+Prediction is essential to the act of data analysis, because if a model
+incorrectly predicts observations, then it follows logically that that model is
+wrong.
+Blinding hides the data until predictions are made, and is a powerful way to
+ensure predictiveness.
+A model that must see an observation before pretending to predict it is not
+predicting the observation --- such a model is not blind.
+Similarly, a model that cannot see an observation is blind and can at best
+attempt predict it.
+In our context, a blind analysis must fully determine all predictions in signal
+regions before knowing what data lie in those signal regions.
+We can remain blind while using other data, such as those in validation
+regions, if those regions are unaffected by the signals, but to prematurely
+peek into the signal regions would break the blinding and threaten to break our
+models' predictiveness.
+
+Unblinding is the act of looking at previously blinded data, and should be
+undertaken only after our predictions are finalized.
+
 Blindness to data is useful, but not necessary ---

### 2023-03-12 main/chapter_2ljets.tex
begin variables defn; use 0.4em spacing

@@
 Control, validation and signal regions are shown from left to right, with the
-regions within each category ordered approximately by their typical $\met$.
+regions within each category ordered approximately by their typical
+missing transverse momentum ($\met$).
 Validation region likelihoods do not directly constrain the fit.
@@
 }~\cite{atlas2022searches, atlas_susy_feynman}.
-\\[0.5em]
+\\[0.4em]
 (a) C1N2, where the initial $\chargino_1\textrm{--}\neutralino_2$ pair
@@
 $m(\chargino_1, \neutralino_2)$ and $m(\neutralino_1)$.
-\\[0.5em]
+\\[0.4em]
 (b) GMSB, where the initial $\neutralino_1\textrm{--}\neutralino_1$ pair
@@
 \end{subfigure}
-\\[0.5em]
+\\[0.4em]
 \begin{subfigure}{0.495\textwidth}
@@
 Unblinded data are shown --- regions in the top two plots observe zero data
-and the error bars for zero events are hidden for aesthetic reasons.%
+and the error bars for zero events are hidden for aesthetic reasons.
+\\[0.4em]
+The event variable $\metsig$ is the missing transverse momentum significance,
+defined in Section~\ref{sec:2ljets_metsig}, and $\mbb$ is the invariant mass of
+the vector sum of the hardest two $b$-tagged jets.
 }
@@
 search~\cite{atlas2022searches}.
-\\[0.5em]
+\\[0.4em]
 Space below the solid red line is labelled as excluded, and its dotted
@@
 of the theoretical uncertainties.
-\\[0.5em]
+\\[0.4em]
 The yellow band shows the $\pm1$-sigma region of exclusion contours
@@
 $\Delta m = m(\chargino_1, \neutralino_2) - m(\neutralino_1)$.
+\\[0.4em]
 }
@@
 Limits are intended to reflect constrains on additive signal contributions.
-\\[0.5em]
+\\[0.4em]
 Left to right:
@@
 integrated luminosity $139.0\,\mathrm{fb}^{-1}$.
-\\[0.5em]
+\\[0.4em]
 Upper limits use the one-sided profile likelihood test
@@
 \end{subfigure}
-\\[0.5em]
+\\[0.4em]
 \begin{subfigure}{0.9\textwidth}
@@
 \end{subfigure}
-\\[0.5em]
+\\[0.4em]
 \begin{subfigure}{0.48\textwidth}
@@
 }~\cite{atlas2022searches, atlas_susy_feynman}.
-\\[0.5em]
+\\[0.4em]
 All three produce same-flavour, opposite-sign lepton pairs, along with jets
@@
 }~\cite{atlas2022searches}.
-\\[0.5em]
+\\[0.4em]
 (Top) gluino-slepton,
@@
 \end{subfigure}
-\\[0.5em]
+\\[0.4em]
 \begin{subfigure}{0.525\textwidth}
@@
 selections~\cite{atlas_23l_SUSY_2016_24, hepdata.81996}.
-\\[0.5em]
+\\[0.4em]
 (a) Histogrammed distribution of $\met$ in and below SR2-int and SR2-high with
@@
 SR2-int requires $\met > 150\,\eV[G]$ and SR2-high requires $\met > 250\,\eV[G]$.
-\\[0.5em]
+\\[0.4em]
 (b) Contours for the combined $\twoljets$ and $3\ell$ channels.
@@
 data~\cite{atlas_2l_SUSY_2013_11}.
-\\[0.5em]
+\\[0.4em]
 (c) Signal regions selected for each signal point in the grid.
@@
 \item $\ptjone$: transverse momentum of the hardest jet%
-\vspace{0.5em}
+\vspace{0.4em}
 \item $\mll$: invariant mass of the hardest two leptons
@@
 described in Section~\ref{sec:2ljets_mt2}
-\vspace{0.5em}
+\vspace{0.4em}
 \item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
 \item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
-\vspace{0.5em}
+\vspace{0.4em}
 \item $\njet$: number of signal jets
@@
 described in Section~\ref{sec:2ljets_btagging}
-\vspace{0.5em}
+\vspace{0.4em}
 \item $\dphillmet$: azimuthal angle between $\vec p_{\ell\ell}$ and $\ptmiss$,
@@
 $n_\mathrm{leptons}^\mathrm{signal} = n_\mathrm{leptons}^\mathrm{baseline} = 2$,
-\\[0.5em]
+\\[0.4em]
 same flavour, opposite charge leptons,
-\\[0.5em]
+\\[0.4em]
 $\njet \geq 1$,
-\\[0.5em]
+\\[0.4em]
 $p_\mathrm{T}^{\,\ell_2} > 25$,
-\\[0.5em]
+\\[0.4em]
 \hphantom{~and}$\metsig > 6$, and
-\\[0.5em]
+\\[0.4em]
 $\met > 100$.
@@
 & $0\textrm{--}0.8\textrm{--}1.6$
-\\[0.5em]
+\\[0.4em]
 \: VR-High
@@
 & $< 1.6$
-\\[0.5em]
+\\[0.4em]
 \: VR-High-R
@@
 &
-\\[0.5em]
+\\[0.4em]
 \: VR-High-1J
@@
 &
-\\[0.5em]
+\\[0.4em]
 \: VR-$\llbb$
@@
 Differences between regions are \uwave{underlined}.
-\\[0.5em]
+\\[0.4em]
 The $\rll$ bins of SR-High are labelled SR-High-8 and SR-High-16, and their
@@
 & $> 60$
-\\[0.5em]
+\\[0.4em]
 \: VR-Int
@@
 & $\uwave{\hphantom{< 60}}$
-\\[0.5em]
+\\[0.4em]
 \crtt
@@
 &
-\\[0.5em]
+\\[0.4em]
 \: VR-Low
@@
 & $< 0.6$
-\\[0.5em]
+\\[0.4em]
 \: VR-Low-2
@@
 (c, d) SR-Low-2 removes $\dphillmet$.
-\\[0.5em]
+\\[0.4em]
 Signals trend towards small $\rll$, plausibly due to their more boosted
@@
 & $>2$
-\\[0.5em]
+\\[0.4em]
 \: VR-OffShell
@@
 The region SR-High-8, labelled `\texttt{High8\_1+\_2}' here, is chosen for its
-simplicity and good sensitivity to both benchmark signal models.\\[0.5em]
+simplicity and good sensitivity to both benchmark signal models.\\[0.4em]
 Upper panels show expected background and signal yields from simulations, with
@@
 Yields were derived in an early iteration of the analysis and do not exactly
-match those in the final result.\\[0.5em]
+match those in the final result.\\[0.4em]
 Lower panels show various measures of projected
@@
 The region \srllbb, labelled `SRllbb' here, is chosen for
-sensitivity, simplicity, and to avoid overlapping with DR-High.\\[0.5em]
+sensitivity, simplicity, and to avoid overlapping with DR-High.\\[0.4em]
 Upper panels show expected background and signal yields from simulations, with
@@
 The region SR-Int, labelled `Int\_1$+$\_2' here, is chosen for its simplicity
-and broad sensitivity.\\[0.5em]
+and broad sensitivity.\\[0.4em]
 Upper panels show expected background and signal yields from simulations, with
@@
 The Monte Carlo estimate is clearly more accurate.
-\\[0.5em]
+\\[0.4em]
 % thesis_project/2ljets/vrss_plots/PAM_FNP_50pc.pdf
@@
 & MC Prompt (stat) & MC FNP (stat) & MM FNP (stat, sys)
-\\[0.5em]
+\\[0.4em]
 \crz\ & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
@@
 & \quad\underline{Max}\hphantom{foo}
-\\[0.5em]
+\\[0.4em]
 \crz\ & $143.6 \pm 4.6$
@@
 and the signal model is simulated with fast simulation.
-\\[0.5em]
+\\[0.4em]
 Yes, this really is a screenshot of a ROOT TBrowser~\cite{ROOT}, with no
@@
 simulation.
-\\[0.5em]
+\\[0.4em]
 This selection requires one signal lepton (electron or muon) with
@@
 & $\sigma_2$
-\\[0.5em]
+\\[0.4em]
 \crdy
@@
 & $1.6$
-\\[0.5em]
+\\[0.4em]
 VR-Low
@@
 & $3.2$
-\\[0.5em]
+\\[0.4em]
 \underline{SR-Low-a}
@@
 since it exceeds $1.5$, it otherwise disappeared off the edge of this plot.
-\\[0.5em]
+\\[0.4em]
 This figure is split to increase its scale; the top of its right half continues
@@
 crosses $0.05$.
-\\[0.5em]
+\\[0.4em]
 Coloured backgrounds illustrate what would have been observed with data at
@@
 Figure~\ref{fig:2ljets_test_stat_distributions}.
-\\[0.5em]
+\\[0.4em]
 For visual effect, the test statistic for $\mu_\textrm{discovery}=16$
@@
 This aesthetic change does not affect the results.
-\\[0.5em]
+\\[0.4em]
 \emph{These figures were produced with \histfitter\ tools through
@@
 and $\mu_\textrm{discovery}=14$, respectively.
-\\[0.5em]
+\\[0.4em]
 Axis limits are set from the minimum and maximum test statistics recorded.
@@
 observation would indicate their likelihood ratio.
-\\[0.5em]
+\\[0.4em]
 Results are:
@@
 and (c) $\clb = 0.68$ and $\clsb = 0.037$.
-\\[0.5em]
+\\[0.4em]
 \emph{These figures were adapted from plots produced with \histfitter\ tools

### 2023-03-12 main/chapter_theory.tex
begin variables defn; use 0.4em spacing

@@
 partners, which are listed in
-Table~\ref{tab:theory_particles_sm_properties}.\\[0.5em]
+Table~\ref{tab:theory_particles_sm_properties}.
+\\[0.4em]
 All of these limits assume R-parity conservation and various other,
@@
 Stronger limits can be set for particles that live long enough to decay within
-the sensitive materials of collider detectors.\\[0.5em]
+the sensitive materials of collider detectors.
+\\[0.4em]
 All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.

### 2023-03-12 main/chapter_2ljets.tex
define variables in figure captions

@@
 \\[0.4em]
+This simplified model assumes that
+$m(\chargino_1) = m(\neutralino_2) = m(\chargino_1, \neutralino_2)$.
 }
@@
 selections~\cite{atlas_23l_SUSY_2016_24, hepdata.81996}.
+The event variable $\met$ is the missing transverse momentum.
+The label $m(\chargino_1)/m(\neutralino_2)$ indicates the mass
+$m(\chargino_1, \neutralino_2)$ under the simplifying assumption that
+$m(\chargino_1, \neutralino_2) = m(\chargino_1) = m(\neutralino_2)$.
 \\[0.4em]
@@
 \centering
-\includegraphics[width=0.62\textwidth]{figures/2ljets_presel_met_logy.png}
+\includegraphics[width=0.6\textwidth]{figures/2ljets_presel_met_logy.png}
 \\
-\includegraphics[width=0.62\textwidth]{figures/2ljets_presel_met_sig_logy.png}
+\includegraphics[width=0.6\textwidth]{figures/2ljets_presel_met_sig_logy.png}
 \caption[
@@
 Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
-models.
+models; $\met$ is the missing transverse momentum, and $\metsig$ is its
+significance as defined in Section~\ref{sec:2ljets_metsig}.
 At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
@@
 $\met > 100\,\eV[G]$, and
-$\metsig > 6$.
+$\metsig > 6$;
+these variables are defined in Section~\ref{sec:2ljets_cheatsheet}.
 \\
-``Significance'' displayed in the lower subplot uses the Binomial significance
-$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
-uncertainty and yields taken to the right of a given value.
-Larger values indicate greater expected sensitivity to the signal model.
+``Significance'' in the lower plot is
+$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ uncertainty and
+yields taken to the right of a given value.
 }

### 2023-03-12 main/chapter_searches.tex
explain ntuple

@@
 it simply uses the likelihood to filter simulated samples
-(like `skimming an ntuple'),
+(like `skimming an ntuple'%
+\footnote{%
+This is a reference to colloquial particle physics jargon;
+an `ntuple' is a ROOT TTree, which is a generic data structure peculiar to the
+ROOT software framework~\cite{ROOT}.
+A TTree resembles a columnar collection of named arrays and jagged arrays,
+and is entirely unlike the `tree' structures of mathematics, computer
+science, or dendrology.
+}),
 then re-normalizes them to a probability distribution within the constraint.

### 2023-03-12 main/chapter_2ljets.tex
fix lines
add forward blinding reference; tidy jagged arrays

@@
 project --- blind to data ---
-followed later by the results --- unblinded.
+followed later by the results --- unblinded.%
+\footnote{%
+Blinding, unblinding, and their implementation in this analysis are discussed
+more thoroughly in Section~\ref{sec:2ljets_unblinding}.%
+}
 Instead, this chapter first reports the unblinded results and follows those
@@
 imprecisely reconstructed.
-A further exception to that rule can occur, however, if the jet originates from a
-$b$ quark, such that the weak decays of its hadrons tend to lose momentum in
+A further exception to that rule can occur, however, if the jet originates from
+a $b$ quark, such that the weak decays of its hadrons tend to lose momentum in
 neutrinos or soft muons that are not reconstructed.
@@
 and $W^\pm Z$ where the charged lepton in $W^\pm\to\ell^\pm\nu$
-is not reconstructed; these subcomponents of the \diboson\ process are discussed
-later in Section~\ref{sec:2ljets_validation} and displayed in
+is not reconstructed; these subcomponents of the \diboson\ process are
+discussed later in Section~\ref{sec:2ljets_validation} and displayed in
 Figure~\ref{fig:2ljets_splits_diboson}.

### 2023-03-12 main/chapter_searches.tex
add forward blinding reference; tidy jagged arrays

@@
 ROOT software framework~\cite{ROOT}.
-A TTree resembles a columnar collection of named arrays and jagged arrays,
+A TTree resembles a columnar collection of named (jagged) arrays,
 and is entirely unlike the `tree' structures of mathematics, computer
-science, or dendrology.
+science, and dendrology.
 }),

### 2023-03-13 main/chapter_theory.tex
comma neutrinos

@@
 bosons                                          & mass limit$\,\eV[G]$ \\[0.2em]
-$\tilde \nu$                                    & $> 94$         \\
+$\tilde \nu_e$,$\tilde \nu_\mu$,$\tilde \nu_\tau$ & $> 94$         \\
 $\tilde e$                                      & $> 107$        \\

### 2023-03-13 main/chapter_theory.tex
expand light quarks, add spaces after commas

--- 2023-03-12 main/chapter_theory.tex
@@
 $\neutralino_1$ (unstable)                      & $> 380$        \\
-$\neutralino_2$,$\neutralino_3$,$\neutralino_4$ & $> 750$        \\
-$\chargino_1$,$\chargino_2$                     & $> 94$         \\[1em]
+$\neutralino_2$, $\neutralino_3$, $\neutralino_4$ & $> 750$        \\
+$\chargino_1$, $\chargino_2$                     & $> 94$         \\[1em]
 bosons                                          & mass limit$\,\eV[G]$ \\[0.2em]
-$\tilde \nu$                                    & $> 94$         \\
+$\tilde \nu_e$, $\tilde \nu_\mu$, $\tilde \nu_\tau$ & $> 94$       \\
 $\tilde e$                                      & $> 107$        \\
@@
 $\tilde \tau$                                   & $> 81.9$       \\
-$\tilde q$                                      & $> 1220$       \\
+$\tilde u$, $\tilde d$, $\tilde c$, $\tilde s$     & $> 1220$       \\
 $\tilde b$                                      & $> 1270$       \\

### 2023-03-13 main/chapter_2ljets.tex
matrix method probabilities and uncertainties

@@
 
+The `real efficiency' transition probabilities, $\prob{t}{r,i}$, are estimated
+from simulation by matching truth-level and reconstructed leptons; they
+are assigned as functions of $|\eta|$ and $\pt$, and increase from
+$\prob{t}{r,i} \approx 70\%$ at $\pt = 25\,\eV[G]$ to
+$\prob{t}{r,i} \approx 95\%$ for $\pt > 100\,\eV[G]$ with small uncertainties.
+The `fake rate' transition probabilities, $\prob{t}{f,i}$, are estimated
+from the difference between simulation and data in selections that enhance the
+fake lepton contribution; they are also assigned as functions of
+$|\eta|$ and $\pt$, and are mostly small at around
+$\prob{t}{f,i} \approx 10\%$.
+Larger fake rates are seen for hard electrons with
+$\pt > 100\,\eV[G]$, for which $\prob{t}{f,i} \approx 40\%$ due to
+contributions from heavy-flavour decays.
+Subtracting simulated backgrounds introduces an uncertainty of
+around $\pm10\%$ on all fake rate estimates.
+
 To make FNP estimates in our regions, the Matrix Method is
@@
 are varied according to various uncertainties in the Matrix Method procedure,
-and statistical errors are assigned in the standard way from the samples
+and statistical errors are assigned in the standard way from the samples'
 weights.
+These systematic and statistical uncertainties are propagated separately,
+and do impact our final results through their impact on the region-specific
+fake estimates, which are described in
+Section~\ref{sec:2ljets_matrix_method_ewk}.
 

### 2023-03-13 main/chapter_2ljets.tex
write about simulation

@@
 
+Throughout $\twoljets$-electroweak, simulated Monte Carlo samples are the
+primary estimates of background and signal processes.
+We first simulate samples at `truth-level' by random sampling from
+distributions that are implemented by various software packages to approximate
+theoretical predictions from quantum field theory.
+Simulating events from high-order perturbative quantum field theory is
+difficult and computationally expensive, so these implementations use
+lower-order approximations, which we normalize to cross-section predictions
+that are evaluated at higher precision.
+Parton shower programs are used to model the soft QCD evolution of
+strongly-interacting particles from the hard scatter down to the low energy
+scales of hadronization; because of colour confinement, only these
+(colour-neutral) hadrons can survive to interact with the detector.
+The various Monte Carlo simulation programs used in this search are summarized
+in Table~\ref{tab:2ljets_paper_mc}, which is adapted from the $\twoljets$
+paper.
+Truth-level results are finally passed through a full simulation of the
+\atlas\ detector to simulate `reconstruction-level'
+data~\cite{SOFT-2010-01, geant2003}.
+
 To interpret the results seriously, we need to understand our uncertainties ---
@@
 drafts.}
+
+\begin{table}[tp]
+\begin{center}
+\scriptsize
+\begin{tabular}{l l l l l l }
+% \hline
+% Physical process &  Generator  & Parton & Cross-section & Tune & PDF set\\
+%                 &             & shower & order in pQCD &      & \\
+% %\hline\hline
+% \noalign{\smallskip}\hline\noalign{\smallskip}
+% \ttbar~\cite{ATL-PHYS-PUB-2016-004}         & \POWHEGBOX[v2] \cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}    & \Pythia[8.230] \cite{Sjostrand:2014zea} & NNLO+NNLL \cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}                   & A14 \cite{ATL-PHYS-PUB-2014-021}   & \textsc{NNPDF2.3lo} \cite{Ball:2012cx}\\
+% $Wt$~\cite{ATL-PHYS-PUB-2016-004}  & \POWHEGBOX[v2]    & \Pythia[8.230] & NNLO+NNLL \cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}        &A14    & \textsc{NNPDF3.0nlo} \cite{Ball:2014uwa}\\
+% $VV$~\cite{ATL-PHYS-PUB-2016-002} & \Sherpa[2.2.2] \cite{Bothmann:2019yzt} & \Sherpa[2.2.2] & NLO \cite{diboson1,diboson2} & \textsc{Sherpa}\ default & \textsc{NNPDF3.0nnlo} \\
+% $Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}& \Sherpa[2.2.1]           & \Sherpa[2.2.1]  &NNLO \cite{DYNNLO1,DYcrosssection}       & \textsc{Sherpa}\ default     &\textsc{NNPDF3.0nnlo}\\
+% Higgs ($t\bar{t}H$) & \POWHEGBOX[v2]    & \Pythia[8.230] & NLO \cite{deFlorian:2016spz} & A14    & \textsc{NPDF2.3lo}\\
+% Higgs (VBF, $Vh$) & \POWHEGBOX[v2]     & \Pythia[8.212] & NNLO \cite{deFlorian:2016spz} & AZNLO \cite{STDM-2012-23} &  \textsc{CTEQ6L1} \cite{Pumplin:2002vw} \\
+% Higgs (ggF) & \POWHEGBOX[v2]     & \Pythia[8.212] & NNNLO \cite{deFlorian:2016spz} & AZNLO & \textsc{CTEQ6L1} \\
+% $ttt$, $tttt$, $t\bar{t}$\,+\,$WW$ & \textsc{MG5\_aMC@NLO} \cite{Alwall:2014hca} & \Pythia[8.186] \cite{Skands:2014pea} & NLO \cite{Frederix:2017wme}, \cite{Alwall:2014hca} & A14 & \textsc{NNPDF2.3lo}\\
+% $t\bar{t}$\,+\,$V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}& \textsc{MG5\_aMC@NLO}        & \Pythia[8.210] & NLO \cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca} & \textsc{A14} & \textsc{NNPDF2.3lo}\\
+% $VVV$~\cite{ATL-PHYS-PUB-2016-002} & \Sherpa[2.2.2] & \Sherpa[2.2.2] &  NLO \cite{Bothmann:2019yzt} & \textsc{Sherpa}\ default     & \textsc{NNPDF3.0nnlo}\\
+% \noalign{\smallskip}\hline\noalign{\smallskip}
+\end{tabular}
+\caption{%
+Caption TODO
+% TODO rewrite
+% Simulated background event samples used in this analysis with the corresponding matrix element and parton shower generators,
+% cross-section order in perturbative QCD used to normalize the event yield, underlying-event tune and PDF set.
+% $V$ represents either a $W$ or $Z$ boson.%
+}
+\label{tab:2ljets_paper_mc}
+\end{center}
+\end{table}
 

### 2023-03-13 main/chapter_2ljets.tex
begin table copy

--- 2023-03-13 main/chapter_2ljets.tex
@@
 drafts.}
+
+\begin{sidewaystable}[tp]
+\begin{center}
+\scriptsize
+\begin{tabular}{llllll}
+Physical process &  Generator  & Parton & Cross-section & Tune & PDF set\\
+                &             & shower & order in pQCD &      &
+\\[0.2em]
+\hline
+\\[0.2em]
+$\ttbar$~\cite{ATL-PHYS-PUB-2016-004}         & \powhegbox\,v2~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}    & \pythia\,8.230~\cite{Sjostrand:2014zea} & NNLO+NNLL~\cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}                   & A14~\cite{ATL-PHYS-PUB-2014-021}   & \textsc{NNPDF2.3lo}~\cite{Ball:2012cx}\\
+$Wt$~\cite{ATL-PHYS-PUB-2016-004}  & \powhegbox\,v2    & \pythia\,8.230 & NNLO+NNLL~\cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}        &A14    & \textsc{NNPDF3.0nlo}~\cite{Ball:2014uwa}\\
+$VV$~\cite{ATL-PHYS-PUB-2016-002} & \sherpa\,2.2.2~\cite{Bothmann:2019yzt} & \sherpa\,2.2.2 & NLO~\cite{diboson1,diboson2} & \textsc{Sherpa}\ default & \textsc{NNPDF3.0nnlo} \\
+$Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}& \sherpa\,2.2.1           & \sherpa\,2.2.1  &NNLO~\cite{DYNNLO1,DYcrosssection}       & \textsc{Sherpa}\ default     &\textsc{NNPDF3.0nnlo}\\
+Higgs ($\ttbar h$) & \powhegbox\,v2    & \pythia\,8.230 & NLO~\cite{deFlorian:2016spz} & A14    & \textsc{NPDF2.3lo}\\
+Higgs (VBF, $Vh$) & \powhegbox\,v2     & \pythia\,8.212 & NNLO~\cite{deFlorian:2016spz} & AZNLO~\cite{STDM-2012-23} &  \textsc{CTEQ6L1}~\cite{Pumplin:2002vw} \\
+Higgs (ggF) & \powhegbox\,v2     & \pythia\,8.212 & NNNLO~\cite{deFlorian:2016spz} & AZNLO & \textsc{CTEQ6L1} \\
+$ttt$, $tttt$, $\ttbar$\,+\,$WW$ & \amcatnlo~\cite{Alwall:2014hca} & \pythia\,8.186~\cite{Skands:2014pea} & NLO~\cite{Frederix:2017wme,Alwall:2014hca} & A14 & \textsc{NNPDF2.3lo}\\
+$\ttbar$\,+\,$V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}& \amcatnlo        & \pythia\,8.210 & NLO~\cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca} & \textsc{A14} & \textsc{NNPDF2.3lo}\\
+$VVV$~\cite{ATL-PHYS-PUB-2016-002} & \sherpa\,2.2.2 & \sherpa\,2.2.2 &  NLO~\cite{Bothmann:2019yzt} & \textsc{Sherpa}\ default     & \textsc{NNPDF3.0nnlo}\\
+\end{tabular}
+\caption{%
+Caption TODO
+% TODO define VBF
+% TODO rewrite
+% Simulated background event samples used in this analysis with the corresponding matrix element and parton shower generators,
+% cross-section order in perturbative QCD used to normalize the event yield, underlying-event tune and PDF set.
+% $V$ represents either a $W$ or $Z$ boson.%
+}
+\label{tab:2ljets_paper_mc}
+\end{center}
+\end{sidewaystable}
 
@@
 in the union of \crz\ and SR-Low, which form a continuous region in $\mjj$.
-Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
+Our central $\zjets$ samples use \sherpa\,2.2.1~\cite{
 Bothmann:2019yzt,
@@
 software framework.
-Our central \diboson\ samples are simulated with \sherpa~2.2.2 at NLO with up
+Our central \diboson\ samples are simulated with \sherpa\,2.2.2 at NLO with up
 to six electroweak vertices, and although \diboson\ sample comparisons are

### 2023-03-13 main/chapter_2ljets.tex
add mc citations

@@
 \\[0.2em]
-$\ttbar$~\cite{ATL-PHYS-PUB-2016-004}         & \powhegbox\,v2~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}    & \pythia\,8.230~\cite{Sjostrand:2014zea} & NNLO+NNLL~\cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}                   & A14~\cite{ATL-PHYS-PUB-2014-021}   & \textsc{NNPDF2.3lo}~\cite{Ball:2012cx}\\
-$Wt$~\cite{ATL-PHYS-PUB-2016-004}  & \powhegbox\,v2    & \pythia\,8.230 & NNLO+NNLL~\cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}        &A14    & \textsc{NNPDF3.0nlo}~\cite{Ball:2014uwa}\\
-$VV$~\cite{ATL-PHYS-PUB-2016-002} & \sherpa\,2.2.2~\cite{Bothmann:2019yzt} & \sherpa\,2.2.2 & NLO~\cite{diboson1,diboson2} & \textsc{Sherpa}\ default & \textsc{NNPDF3.0nnlo} \\
-$Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}& \sherpa\,2.2.1           & \sherpa\,2.2.1  &NNLO~\cite{DYNNLO1,DYcrosssection}       & \textsc{Sherpa}\ default     &\textsc{NNPDF3.0nnlo}\\
-Higgs ($\ttbar h$) & \powhegbox\,v2    & \pythia\,8.230 & NLO~\cite{deFlorian:2016spz} & A14    & \textsc{NPDF2.3lo}\\
-Higgs (VBF, $Vh$) & \powhegbox\,v2     & \pythia\,8.212 & NNLO~\cite{deFlorian:2016spz} & AZNLO~\cite{STDM-2012-23} &  \textsc{CTEQ6L1}~\cite{Pumplin:2002vw} \\
-Higgs (ggF) & \powhegbox\,v2     & \pythia\,8.212 & NNNLO~\cite{deFlorian:2016spz} & AZNLO & \textsc{CTEQ6L1} \\
-$ttt$, $tttt$, $\ttbar$\,+\,$WW$ & \amcatnlo~\cite{Alwall:2014hca} & \pythia\,8.186~\cite{Skands:2014pea} & NLO~\cite{Frederix:2017wme,Alwall:2014hca} & A14 & \textsc{NNPDF2.3lo}\\
-$\ttbar$\,+\,$V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}& \amcatnlo        & \pythia\,8.210 & NLO~\cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca} & \textsc{A14} & \textsc{NNPDF2.3lo}\\
-$VVV$~\cite{ATL-PHYS-PUB-2016-002} & \sherpa\,2.2.2 & \sherpa\,2.2.2 &  NLO~\cite{Bothmann:2019yzt} & \textsc{Sherpa}\ default     & \textsc{NNPDF3.0nnlo}\\
+$\ttbar$~\cite{ATL-PHYS-PUB-2016-004}
+    & \powhegbox\,v2~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
+    & \pythia\,8.230~\cite{Sjostrand:2014zea}
+    & NNLO+NNLL~\cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}
+    & A14~\cite{ATL-PHYS-PUB-2014-021}
+    & \textsc{NNPDF2.3lo}~\cite{Ball:2012cx}
+\\
+$Wt$~\cite{ATL-PHYS-PUB-2016-004}
+    & \powhegbox\,v2
+    & \pythia\,8.230
+    & NNLO+NNLL~\cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}
+    & A14
+    & \textsc{NNPDF3.0nlo}~\cite{Ball:2014uwa}
+\\
+$VV$~\cite{ATL-PHYS-PUB-2016-002}
+    & \sherpa\,2.2.2~\cite{Bothmann:2019yzt}
+    & \sherpa\,2.2.2 & NLO~\cite{diboson1,diboson2}
+    & \textsc{Sherpa}~default
+    & \textsc{NNPDF3.0nnlo}
+\\
+$Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}
+    & \sherpa\,2.2.1
+    & \sherpa\,2.2.1
+    & NNLO~\cite{DYNNLO1,DYcrosssection}
+    & \textsc{Sherpa}~default
+    &\textsc{NNPDF3.0nnlo}
+\\
+Higgs ($\ttbar h$)
+    & \powhegbox\,v2
+    & \pythia\,8.230
+    & NLO~\cite{deFlorian:2016spz}
+    & A14
+    & \textsc{NPDF2.3lo}
+\\
+Higgs (VBF, $Vh$)
+    & \powhegbox\,v2
+    & \pythia\,8.212
+    & NNLO~\cite{deFlorian:2016spz}
+    & AZNLO~\cite{STDM-2012-23}
+    & \textsc{CTEQ6L1}~\cite{Pumplin:2002vw}
+\\
+Higgs (ggF)
+    & \powhegbox\,v2
+    & \pythia\,8.212
+    & NNNLO~\cite{deFlorian:2016spz}
+    & AZNLO
+    & \textsc{CTEQ6L1}
+\\
+$ttt$, $tttt$, $\ttbar$\,+\,$WW$
+    & \amcatnlo~\cite{Alwall:2014hca}
+    & \pythia\,8.186~\cite{Skands:2014pea}
+    & NLO~\cite{Frederix:2017wme,Alwall:2014hca}
+    & A14
+    & \textsc{NNPDF2.3lo}
+\\
+$\ttbar$\,+\,$V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}
+    & \amcatnlo
+    & \pythia\,8.210
+    & NLO~\cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca}
+    & \textsc{A14}
+    & \textsc{NNPDF2.3lo}
+\\
+$VVV$~\cite{ATL-PHYS-PUB-2016-002}
+    & \sherpa\,2.2.2
+    & \sherpa\,2.2.2
+    &  NLO~\cite{Bothmann:2019yzt}
+    & \textsc{Sherpa}~default
+    & \textsc{NNPDF3.0nnlo}
+\\
 \end{tabular}
@@
 Caption TODO
+% TODO import citations
+% TODO rewrite
 % TODO define VBF
-% TODO rewrite
 % Simulated background event samples used in this analysis with the corresponding matrix element and parton shower generators,
@@
 % $V$ represents either a $W$ or $Z$ boson.%
+Adapted from~\cite{atlas2022searches}.
 }

### 2023-03-13 main/chapter_2ljets.tex
write generators caption; improve format

@@
 We first simulate samples at `truth-level' by random sampling from
-distributions that are implemented by various software packages to approximate
-theoretical predictions from quantum field theory.
+distributions with parton distribution functions and scattering matrix elements
+that are implemented by various software packages to approximate theoretical
+predictions from quantum field theory.
 Simulating events from high-order perturbative quantum field theory is
@@
 \begin{tabular}{llllll}
-Physical process &  Generator  & Parton & Cross-section & Tune & PDF set\\
-                &             & shower & order in pQCD &      &
+Physical process
+    & Generator
+    & Parton Shower
+    & Cross-section order in QCD
+    & Tune
+    & PDF set
 \\[0.2em]
@@
 \\
-$ttt$, $tttt$, $\ttbar$\,+\,$WW$
+$ttt$, $tttt$, $\ttbar WW$
     & \amcatnlo~\cite{Alwall:2014hca}
@@
 \\
-$\ttbar$\,+\,$V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}
+$\ttbar V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}
     & \amcatnlo
@@
 \caption{%
-Caption TODO
-% TODO import citations
-% TODO rewrite
-% TODO define VBF
-% Simulated background event samples used in this analysis with the corresponding matrix element and parton shower generators,
-% cross-section order in perturbative QCD used to normalize the event yield, underlying-event tune and PDF set.
-% $V$ represents either a $W$ or $Z$ boson.%
-Adapted from~\cite{atlas2022searches}.
+Summary of the simulation tools and configurations used for central background
+samples in $\twoljets$.
+Each row specifies a combination of matrix element generator, parton shower,
+and Parton Distribution Function (PDF) components, whose combination we use to
+simulate the named physical process.
+The tune specifies many parameters of the simulation tools to achieve a good
+description of some historical calibration data.
+Total sample cross-sections are normalized to a reference value evaluated
+at the specified order in perturbative QCD (pQCD).
+Symbols:
+$V$ is $W$ or $Z$,
+VBF is Vector Boson Fusion,
+ggF is gluon-gluon Fusion.
+This table is adapted from~\cite{atlas2022searches}.
 }

### 2023-03-13 main/chapter_2ljets.tex
add uncertainties summary plot and commentary

@@
 in Table~\ref{tab:2ljets_paper_mc}, which is adapted from the $\twoljets$
-paper.
+paper~\cite{atlas2022searches}.
 Truth-level results are finally passed through a full simulation of the
@@
 \end{tabular}
-\caption{%
+\caption[
+Summary of the simulation tools and configurations
+]{%
 Summary of the simulation tools and configurations used for central background
@@
 ggF is gluon-gluon Fusion.
-This table is adapted from~\cite{atlas2022searches}.
+This table is adapted from the $\twoljets$ paper~\cite{atlas2022searches}.
 }
@@
 
+These parameters control the yield assigned in each region by our
+\heplikelihood\ model as described in
+Section~\ref{sec:searches_heplikelihood}.
+Uncertainties in the parameters therefore propagate to uncertainties in the
+background expectations.
+To illustrate which categories of parameters contribute the most uncertainty
+to which regions,
+Figure~\ref{fig:2ljets_fit_uncertainty} presents a summary of the relative
+uncertainties attributable to sources of uncertainty in each of region of the
+$\twoljets$-electroweak search.
+This figure is reproduced from the $\twoljets$ paper~\cite{atlas2022searches}.
+No single source of uncertainty dominates in all regions.
+Theoretical uncertainties are large in many regions, most dramatically
+in SR-High-1J and SR-$\llbb$. These two regions have small yields and large
+variations due to alternative sample comparisons and uncertainties that are
+not constrained in control regions;
+see Section~\ref{sec:2ljets_jet_mass} for details of the extra uncertainties
+in SR-High-1J.
+Large relative contributions from FNP backgrounds drive the uncertainties in
+SR-High-16, both directly in the data-driven uncertainty category and
+indirectly through their poor statistical precision in the statistical
+uncertainty category.
+Experimental and normalization uncertainties make middling contributions
+in all regions.
+
+\begin{figure}[tp]
+\centering
+\adjincludegraphics[width=0.98\textwidth]{figures/2ljets_syst_summary_plot_ewk}
+\caption[
+Summary of the contributions to uncertainty on total backgrounds
+]{%
+Summary of the contributions to uncertainty on total backgrounds, represented
+as relative errors for all signal regions of the $\twoljets$-electroweak
+search.
+Uncertainty attributions are assigned by a standard method that explores the
+variations in the \heplikelihood\ model with respect to individual parameters.
+Experimental uncertainties (``Exp. Unc.'') include detector effects in object
+identification and reconstruction, pile-up, and triggering.
+Statistical uncertainties (``Bkg. Sample Stats'') arise from the
+finite number of weighted Monte Carlo events in our background samples
+and from statistical uncertainties on the Matrix Method FNP estimate.
+Data-driven background uncertainties (``DD Bkg. Unc.'') include only effects
+from the FNP background estimate in this search.
+Theoretical uncertainties (``Theory Unc.'') include only theoretical
+uncertainties on the Standard Model backgrounds, and not on any signals.
+Background normalization uncertainties (``Bkg. Norm.'') derive from the
+free normalization factors that should be constrained by control regions.
+The total uncertainty (``Total Unc.'') is a sum in quadrature of all
+contributions.
+\\[0.4em]
+This figure is reproduced from the $\twoljets$ paper~\cite{atlas2022searches}.
+}
+\label{fig:2ljets_fit_uncertainty}
+\end{figure}
+
 \paragraph{Contours.}

### 2023-03-13 main/chapter_searches.tex
add uncertainties summary plot and commentary

@@
 \subsection{Likelihood modelling}
+\label{sec:searches_heplikelihood}
 To represent the Standard Model and alternatives, LHC data analyses use

### 2023-03-13 main/chapter_2ljets.tex
add int trigger table

@@
 used in $\twoljets$.
+A more detailed listing of our trigger choices is presented in
+Table~\ref{tab:2ljets_trigger_int}.
 
@@
 $\twoljets$-electroweak regions are required to have $\pt > 25\,\eV[G]$.
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{ccll}
+Year & $\ell\ell$ & Trigger & Comment
+\\[0.2em]
+\hline
+2015 & $ee$     & {\footnotesize HLT\_2e12\_lhloose\_L12EM10VH} & \\
+2015 & $\mu\mu$ & {\footnotesize HLT\_2mu10} & \\
+2015 & $\mu\mu$ & {\footnotesize HLT\_mu18\_mu8noL1} & \\
+2015 & $e\mu$   & {\footnotesize HLT\_e17\_lhloose\_mu14} & \\
+2015 & $e\mu$   & {\footnotesize HLT\_e7\_lhmedium\_mu24} & \\[0.8em]
+%
+2016 & $ee$     & {\footnotesize HLT\_2e15\_lhvloose\_nod0\_L12EM13VH} & except run 300540 \\
+2016 & $ee$     & {\footnotesize HLT\_2e17\_lhvloose\_nod0} & run 300540, periods D4-- \\
+2016 & $\mu\mu$ & {\footnotesize HLT\_2mu10} & period A \\
+2016 & $\mu\mu$ & {\footnotesize HLT\_2mu14} & from period B \\
+2016 & $\mu\mu$ & {\footnotesize HLT\_mu20\_mu8noL1} & periods A-D3 \\
+2016 & $\mu\mu$ & {\footnotesize HLT\_mu22\_mu8noL1} & periods from D4 \\
+2016 & $e\mu$   & {\footnotesize HLT\_e17\_lhloose\_nod0\_mu14} & \\
+2016 & $e\mu$   & {\footnotesize HLT\_e24\_lhmedium\_nod0\_L1EM20VHI\_mu8noL1}\hspace{-3em} & \\
+2016 & $e\mu$   & {\footnotesize HLT\_e7\_lhmedium\_nod0\_mu24} & \\[0.8em]
+%
+2017 & $ee$     & {\footnotesize HLT\_2e17\_lhvloose\_nod0\_L12EM15VHI} & except periods B5--B8 \\
+2017 & $ee$     & {\footnotesize HLT\_2e24\_lhvloose\_nod0} & \\
+2017 & $\mu\mu$ & {\footnotesize HLT\_2mu14} & \\
+2017 & $\mu\mu$ & {\footnotesize HLT\_mu22\_mu8noL1} & \\
+2017 & $e\mu$   & {\footnotesize HLT\_e17\_lhloose\_nod0\_mu14} & \\
+2017 & $e\mu$   & {\footnotesize HLT\_e26\_lhmedium\_nod0\_mu8noL1} & \\
+2017 & $e\mu$   & {\footnotesize HLT\_e7\_lhmedium\_nod0\_mu24} & \\[0.8em]
+%
+2018 & $ee$     & {\footnotesize HLT\_2e17\_lhvloose\_nod0\_L12EM15VHI} & \\
+2018 & $ee$     & {\footnotesize HLT\_2e24\_lhvloose\_nod0} & \\
+2018 & $\mu\mu$ & {\footnotesize HLT\_2mu14} & \\
+2018 & $\mu\mu$ & {\footnotesize HLT\_mu22\_mu8noL1} & \\
+2018 & $e\mu$   & {\footnotesize HLT\_e17\_lhloose\_nod0\_mu14} & \\
+2018 & $e\mu$   & {\footnotesize HLT\_e26\_lhmedium\_nod0\_mu8noL1} & \\
+2018 & $e\mu$   & {\footnotesize HLT\_e7\_lhmedium\_nod0\_mu24} & \\
+\end{tabular}
+\caption[
+Detailed trigger specifications
+]{
+Detailed trigger specifications for the $\twoljets$ analysis in all four
+years of Run~2.
+These are the dilepton triggers with the weakest available requirements.
+Some information can be scraped from the name of each trigger.
+Run numbers and periods mentioned in the ``Comment'' column indicate specific
+parts of each LHC run, and are chosen on the advice of internal \atlas\
+documentation and relate to local operating
+conditions~\cite{atlas_twiki_lowest_unprescaled}.
+This table is adapted from supporting material for the $\twoljets$ search
+paper~\cite{atlas2022searches}.
+}
+\label{tab:2ljets_trigger_int}
+\end{table}
 

### 2023-03-13 main/chapter_2ljets.tex
add aux cutflows appendix

@@
 Detailed trigger specifications
-]{
+]{%
 Detailed trigger specifications for the $\twoljets$ analysis in all four
@@
 models
-]{
+]{%
 Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal

### 2023-03-23 main/chapter_2ljets.tex
add cutflow and rejection factors to cheat-sheet

@@
 
-Signal regions of the $\twoljets$-electroweak analysis are combined to form
-five discovery regions.
-Results of their model-independent interpretations are presented in
-Table~\ref{tab:2ljets_discovery}.
-The same data deficits drive stronger bounds on possible signals that manifest
-as smaller limits on the signal expectations, which are labelled
-$S_{\mathrm{obs}}^{95}$ in the figure.
-\footnotetext{%
-The visible cross-section is defined such that the expected event count is
-given by the product of $A\epsilon\sigma\times L$, where $L$ is the integrated
-luminosity (typically in $\mathrm{fb}^{-1}$ units).
-This particular presentation is uses an `acceptance' $A$, an
-`efficiency' $\epsilon$, and an observable cross-section $\sigma_{\mathrm{obs}}$
-(presented here at at $95\%$ limit).
-The acceptance and efficiency are defined using truth-level simulations, in
-which the $A$ is the fraction of truth-level events that pass selection
-requirements, and $\epsilon$ is the ratio from the truth-level yield to the
-reconstruction-level yield.
-Because detector effects can increase yields, this $\epsilon$ can exceed
-one~\cite{susy_results_presentation_papers}.
-}
-
 % upper limits
@@
 \end{table}
+
+Signal regions of the $\twoljets$-electroweak analysis are combined to form
+five discovery regions.
+Results of their model-independent interpretations are presented in
+Table~\ref{tab:2ljets_discovery}.
+The same data deficits drive stronger bounds on possible signals that manifest
+as smaller limits on the signal expectations, which are labelled
+$S_{\mathrm{obs}}^{95}$ in the figure.
+\footnotetext{%
+The visible cross-section is defined such that the expected event count is
+given by the product of $A\epsilon\sigma\times L$, where $L$ is the integrated
+luminosity (typically in $\mathrm{fb}^{-1}$ units).
+This particular presentation is uses an `acceptance' $A$, an
+`efficiency' $\epsilon$, and an observable cross-section $\sigma_{\mathrm{obs}}$
+(presented here at at $95\%$ limit).
+The acceptance and efficiency are defined using truth-level simulations, in
+which the $A$ is the fraction of truth-level events that pass selection
+requirements, and $\epsilon$ is the ratio from the truth-level yield to the
+reconstruction-level yield.
+Because detector effects can increase yields, this $\epsilon$ can exceed
+one~\cite{susy_results_presentation_papers}.
+}
 
@@
 they aid the local designs that we detail in Section~\ref{sec:2ljets_design}.
+
+The power of each of these variables to select signals and exclude backgrounds
+is hugely variable across our many signal models and regions of phase space.
+For illustration, Table~\ref{tab:2ljets_ewk_acceff} reproduces some signal
+efficiency numbers for the sum of yields in all $\twoljets$-electroweak signal
+regions.
+It finds a that only $0.0003\%$ of signal events are captured under the light
+C1N2 model with
+$m(\chargino_1, \neutralino_2) = 100\,\eV[G]$ and
+$m(\neutralino_1) = 60\,\eV[G]$.
+The reduced cross-sections for models with large masses are offset by their
+more spectacular appearances, with a much larger capture rate of
+for more extreme models.
+For example, the corresponding capture rate it $0.64\%$ for the C1N2 model with
+$m(\chargino_1, \neutralino_2) = 650\,\eV[G]$ and
+$m(\neutralino_1) = 250\,\eV[G]$.
+Detailed cut-flow tables,
+Tables~\ref{aux:ewk:cutflowSRllbb} to~\ref{aux:ewk:cutflowSROffShell_2} of
+Appendix~\ref{app:2ljets_cut_flows},
+give more granular information into the effects of many intermediate
+selection requirements across all signal regions for several representative
+signal models.
+Notably, $\mttwoll$ rejects substantial amounts of signal, although this
+rejection is worthwhile for its larger rate of background rejection.
+
+On background rejection performance,
+Tables~\ref{tab:cutflows_CR-VZ} to~\ref{tab:cutflows_CR-DY} of
+Appendix~\ref{app:2ljets_cut_flows} present cut-flow tables in the four control
+regions of $\twoljets$-electroweak with background sample rates included.
+These tables show that $\mttwoll$ selections are powerful for removing $\ttbar$
+backgrounds, as expected, and that $\metsig$ selections beyond the
+pre-selection requirement of $\metsig > 6$ drastically reduce most other
+backgrounds.
+Summed across all four control regions with data given in these tables,
+the control region selections reduce background rates with a survival fraction
+of $1.5\%$ relative to the pre-selection requirements of
+Table~\ref{tab:2ljets_presel}.
+
+\begin{table}[tp]
+\centering
+\begin{tabular}{ccl}
+\multicolumn{3}{c}{C1N2}   \\
+$m(\chargino_1, \neutralino_2)$ ($\eV[G]$)
+& $m(\neutralino_1)$ ($\eV[G]$)
+& $A\times\epsilon$ (\%)
+\\[1.5ex]
+\hline
+% 0.0000028408852207408747
+100 & 60 & 0.0003       \\
+% 0.000007322904893685767
+150 & 70 & 0.0007       \\
+% 0.00041538437098791273
+250 & 100 & 0.042       \\
+% 0.002093724286276909
+400 & 200 & 0.21     \\
+% 0.0063561141231079175
+650 & 250 & 0.64     \\
+\end{tabular}
+\\[2em]
+\begin{tabular}{ccl}
+\multicolumn{3}{c}{GMSB}  \\
+$m(\neutralino_1)$  ($\eV[G]$)
+& $B(\neutralino_1 \to h \gravitino)$
+& $A\times\epsilon$ (\%)  \\[1.5ex]
+\hline
+% 0.00023984225213678648
+200 & 50 & 0.02  \\
+% 0.002095514599015893
+500 & 80 & 0.21   \\
+% 0.0161139463793303
+700 & 0  & 1.6   \\
+\end{tabular}
+\caption[
+]{%
+Acceptance times efficiency (fraction of total production yields) in the union
+of all $\twoljets$-electroweak signal regions, displayed for example C1N2 and
+GMSB signal models.
+Acceptance $A$ and efficiency $\epsilon$ are defined via truth-level
+simulations as specified in the footnote from Table~\ref{tab:2ljets_discovery}.
+This table is adapted from the $\twoljets$ paper~\cite{atlas2022searches}.%
+}
+\label{tab:2ljets_ewk_acceff}
+\end{table}
 

### 2023-03-27 main/chapter_2ljets.tex
add background feynman diagrams

@@
 deficits in data that we have shown in Figure~\ref{fig:2ljets_summary}.
+
+Any sensitivity to these signals requires them to be visible above the
+background noise.
+For our purposes, that background noise arises from known processes of the
+Standard Model.
+After selecting parts of data that are sensitive to our C1N2 and GMSB signals,
+it turns out that the most important backgrounds are describable as
+$ZZ$-like \diboson,
+$\zyjets$,
+or $\ttbar$;
+these processes are illustrated in Figure~\ref{fig:2ljets_feynman}.
+Each of these Standard Model processes produces visible lepton pairs, jets, and
+missing transverse momentum to mimic our signal models' effects.
+Although $\zyjets$ events lack prompt invisible particles for missing
+transverse momentum, this is offset by its large production cross-section
+with missing transverse momentum sourced from indirect effects such as
+mismeasured jets.
+
+% background diagrams
+\begin{figure}[tp]
+\centering
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_feynman_diboson_zz.png}
+\caption{$ZZ$-like \diboson}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.46\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_feynman_zyjets_slash.png}
+\caption{$\zyjets$}
+\end{subfigure}
+\\
+\begin{subfigure}{0.4\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_feynman_ttbar.png}
+\caption{$\ttbar$}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.46\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/2ljets_feynman_top_decay.png}
+\caption{Top quark decay}
+\end{subfigure}
+\caption[
+Feynman diagrams for the Standard Model background processes
+]{%
+Feynman diagrams for the Standard Model background processes of important
+background contributions in the $\twoljets$-electroweak search.
+In \diboson\ events we expect one $Z$ boson-like decay to charged leptons, and
+the other to neutrinos; additional jets can arise from initial state radiation
+or additional strong radiations from the hard scatter.
+Missing transverse momentum is absent from prompt emissions in $\zyjets$
+events, as illustrated in (b), but arises from rare mismeasurements or
+non-prompt decays.
+Top quarks rapidly decay weakly, mainly to $b$-flavoured hadrons due to the
+small weak mixing with the third generation;
+panel (d) illustrates this with a leptonic decay that we expect to select.
+}
+\label{fig:2ljets_feynman}
+\end{figure}
 

### 2023-03-27 main/chapter_2ljets.tex
cr/vr/sr usage in fits and decision metrics

@@
 vaguely-specified desires, and are informed with incomplete information.
+When designing the regions described in this section, I conducted a manual
+search and optimization with human feedback from the $\twoljets$ team.
+Expected significance statistics were used to assess the sensitivity of a
+region design to particular representative signals, but final decisions were
+made to maximise the extend of projected exclusion contours as evaluated by our
+statistical software.
 

### 2023-03-27 main/chapter_searches.tex
cr/vr/sr usage in fits and decision metrics

@@
 which are defined by logical selections in event variables.
+Each search has its own desiderata, so many different criteria can be used
+to choose these logical selections; broadly, however, selections are made to
+separate signals from backgrounds, such that data appearing in the regions are
+informative as to whether signals are present and what the values of unknown
+parameters might best explain the data.
 To communicate their intended uses, we further categorize these bins into
@@
 of their results in Chapter~\ref{chapter:2ljets}.
+Control regions are included in all fits with an intent to constrain the
+normalization of Standard Model background samples.
+Validation regions are not directly included as likelihood constraints in any
+fits, but are used for visual inspection only, such that the analysts can
+iteratively adapt their modelling to look good in validation regions without
+directly biasing final results.
+Signal regions are included only once the backgrounds are constrained and
+validated and only in fits where one is testing for the presence of signals
+as detailed below.
 
@@
 defined models to the \heplikelihood\ model.
-These fits can use multiple signal regions simultaneously, and couple the
-effect of parametrized modifications to the signals and backgrounds between
-the bins of those regions.
+These fits can use multiple signal regions and control regions simultaneously,
+and couple the effect of parametrized modifications to the signals and
+backgrounds between the bins of those regions.
 

### 2023-03-27 main/chapter_searches.tex
add heplikelihood formula

@@
 }.
-At unit temperature, this can induce a probability distribution
+At temperature $1/k_B$, this can induce a probability distribution
 relative to an appropriate measure~\cite{
@@
 }
+
+The formula for a standard \roofit\ \heplikelihood\ function can be
+summarised as
+\begin{equation}
+\label{eqn:searches_heplikelihood_e}
+e^{E(x)} =
+\prod_{b \in \mathrm{bins}}\mathrm{Poisson}(n_b \mid y_b(x))
+\times \prod_{i \in \mathrm{parameters}} f_i(x_i),
+\end{equation}
+in which $\mathrm{Poisson}(n\mid \mu)$ is the Poisson likelihood function
+of Equation~\ref{eqn:searches_poisson}, $y_b(x)$ is the yield in bin $b$
+for parameters $x$, and $f_i$ is the constraint function for parameter $i$.
+Constraint functions $f_i$ are typically of
+Gaussian (Equation~\ref{eqn:searches_gaussian}) or
+Poissonian (Equation~\ref{eqn:searches_poisson} or a continuous interpolation)
+forms, and often interpreted as likelihoods on fictional data.
+The mappings from parameters to yields $y_b(x)$ are parameterised to
+non-linearly interpolate between given `low', `central', and `high' variations
+of the yields corresponding to certain parameters; many such
+interpolation/extrapolation functions are available, but the most common
+interpolates with a sixth-order polynomial that is smoothly attached to
+exponential tails for extrapolation.
+To avoid the numerical difficulties associated with taking a product over many
+terms, one typically handles $E(x)$, not $e^{E(x)}$.
+The bins may be further grouped into channels that share
+terms~\cite{verkerke2003roofit}.
 

### 2023-03-27 main/chapter_experiment.tex
add fcc etc writing
add breakdown of lumi/conditions by year

@@
 $3.6\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2012~\cite{ATLAS:2013tpw}.
-The future High Luminosity LHC (HL-LHC), however, might achieve
-$750\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
-For the HL-LHC, the current inner tracking and insertable B-layer will be
-replaced by a new Inner Trac\emph{k}er (ITk) which is designed to handle such
+The future High-Luminosity LHC (HL-LHC), however, might achieve
+$75\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
+For the HL-LHC, the current inner tracking and insertable B-layer will
+be replaced by a new Inner Trac\emph{k}er (ITk) which is designed to handle such
 extreme luminosity conditions~\cite{CERN-LHCC-2017-021}.
@@
 halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.
+A breakdown of the LHC proton-proton data conditions and quantity is given
+in Table~\ref{tab:experiment_run2_state}.
+Broadly, instantaneous luminosity was increased over time, leading to increases
+in both data quantity and pile-up interactions alongside those data.
+
+\begin{sidewaystable}[tp]
+\centering
+\begin{tabular}{lcccc}
+Observable                                                        & 2015         & 2016         & 2017                & 2018         \\
+\hline
+Maximum number of colliding bunch-pairs                           & $2232$       & $2208$       & $2544$ ($1909$)     & $2544$       \\
+Bunch spacing ($\mathrm{ns}$)                                     & $25$         & $25$         & $25$ (---)          & $25$         \\
+Typical number of protons per bunch ($\times 10^{11}$)            & $1.1$        & $1.1$        & $1.1$ ($1.2$)       & $1.1$        \\
+Peak luminosity ($\times 10^{33}\mathrm{cm}^{-2}\mathrm{s}^{-1}$) & $5$          & $13$         & $16$                & $19$         \\
+Peak number of inelastic interactions per crossing                & $\approx 16$ & $\approx 41$ & $\approx 45$ ($60$) & $\approx 55$ \\
+Luminosity-weighted average inelastic interactions per crossing   & $13$         & $25$         & $38$                & $36$         \\
+Integrated luminosity delivered ($\mathrm{fb}^{-1}$)              & $4$          & $39$         & $50.6$              & $63.8$
+\end{tabular}
+\caption[
+LHC data rates and conditions by year in Run~2
+]{%
+LHC data rates and conditions by year in Run~2.
+This table is adapted from~\cite{ATLAS:2022hro}.
+Bracketed numbers under 2017 refer to an alternative LHC operating condition
+which alternated in a pattern of eight bunches with the usual
+$25\,\mathrm{ns}$ separation followed by a gap of four empty bunch slots.
+Alongside a primary interaction, the inelastic interactions in a crossing are
+pile-up events and add noise to our measurements.
+}
+\label{tab:experiment_run2_state}
+\end{sidewaystable}
 
@@
 }.
+After selecting for data quality requirements, we are left with a total of
+$139.0\,\mathrm{fb}^{-1}$ \atlas\ proton-proton collision data to analyse.
+This integrated luminosity comprises
+% lumi_15_16 = 36.2 # [1/fb] 15+16
+% lumi_17 = 44.3 # [1/fb] 17
+% lumi_18 = 58.5 # [1/fb] 18
+$36.2\,\mathrm{fb}^{-1}$ from the combination of 2015 and 2016,
+$44.3\,\mathrm{fb}^{-1}$ from 2017, and
+$58.5\,\mathrm{fb}^{-1}$ from 2018.
 
-% closing comment?
+% closing comment
 Now that we have predictive models in the Standard Model and supersymmetric

### 2023-03-27 main/chapter_experiment.tex
add variable, 8b4e

@@
 Maximum number of colliding bunch-pairs                           & $2232$       & $2208$       & $2544$ ($1909$)     & $2544$       \\
-Bunch spacing ($\mathrm{ns}$)                                     & $25$         & $25$         & $25$ (---)          & $25$         \\
+Bunch spacing ($\mathrm{ns}$)                                     & $25$         & $25$         & $25$ (variable)          & $25$         \\
 Typical number of protons per bunch ($\times 10^{11}$)            & $1.1$        & $1.1$        & $1.1$ ($1.2$)       & $1.1$        \\
@@
 This table is adapted from~\cite{ATLAS:2022hro}.
-Bracketed numbers under 2017 refer to an alternative LHC operating condition
-which alternated in a pattern of eight bunches with the usual
+Bracketed numbers under 2017 refer to an alternative `8b4e' LHC operating
+condition which alternated in a pattern of eight bunches with the usual
 $25\,\mathrm{ns}$ separation followed by a gap of four empty bunch slots.

### 2023-03-27 main/chapter_experiment.tex
fix table formatting
begin trigger summary table
add trigger summary table

@@
 
+\TODO{paragraph for Table~\ref{tab:experiment_trigger_breakdown}.}
+
+\begin{table}[tp]
+\centering
+\footnotesize
+\begin{adjustbox}{width=\textwidth}
+\begin{tabular}{llcc}
+Trigger                        & Summary                                                          & L1 Peak Rate ($\mathrm{kHz}$) & HLT Peak Rate ($\mathrm{Hz}$) \\
+\hline
+\multirow{5}{*}{Single lepton} & Isolated $\mu$, $\pt > 27\,\eV[G]$                               & $16$                          & $218$                         \\
+                               & Isolated tight $e$, $\pt > 27\,\eV[G]$                           & $31$                          & $195$                         \\
+                               & $\mu$, $\pt > 52\,\eV[G]$                                        & $16$                          & $70$                          \\
+                               & $e$, $\pt > 61\,\eV[G]$                                          & $28$                          & $20$                          \\
+                               & $\tau$, $\pt > 170\,\eV[G]$                                      & $1.4$                         & $42$                          \\
+\hline
+\multirow{9}{*}{Dilepton}      & \textbf{$\mu\mu$, both $\pt > 15\,\eV[G]$}                       & \textbf{$2.2$}                & \textbf{$30$}                 \\
+                               & \textbf{$\mu\mu$, $\pt > 23, 9\,\eV[G]$}                         & \textbf{$16$}                 & \textbf{$47$}                 \\
+                               & \textbf{loose $ee$, both $\pt > 18\,\eV[G]$}                     & \textbf{$2.0$}                & \textbf{$13$}                 \\
+                               & \textbf{$e\mu$, $\pt > 18, 15\,\eV[G]$}                          & \textbf{$16$}                 & \textbf{$6$}                  \\
+                               & \textbf{loose $e$, $\mu$, $\pt > 18, 15\,\eV[G]$}                & \textbf{$2.6$}                & \textbf{$5$}                  \\
+                               & \textbf{$e\mu$, $\pt > 29, 9\,\eV[G]$}                           & \textbf{$21$}                 & \textbf{$4$}                  \\
+                               & $\tau\tau$, $\pt > 40, 30\,\eV[G]$                               & $5.7$                         & $93$                          \\
+                               & $\tau$, isolated $\mu$, $\pt > 30, 15\,\eV[G]$                   & $2.4$                         & $17$                          \\
+                               & $\tau$, isolated $e$, $\pt > 30, 18\,\eV[G]$                     & $4.6$                         & $19$                          \\
+\hline
+\multirow{5}{*}{Trilepton}     & loose $eee$, $\pt > 25,13,13\,\eV[G]$                            & $1.6$                         & $0.1$                         \\
+                               & $\mu\mu\mu$, $\pt > 7\,\eV[G]$                                   & $0.2$                         & $7$                           \\
+                               & $\mu\mu\mu$, $\pt > 21,5,5\,\eV[G]$                              & $16$                          & $9$                           \\
+                               & $\mu\mu$, loose $e$, $\pt > 11,11,13\,\eV[G]$                    & $2.2$                         & $0.5$                         \\
+                               & loose $ee$, $\mu$, $\pt > 13,13,11\,\eV[G]$                      & $2.3$                         & $0.1$                         \\
+\hline
+Single photon                  & loose $\gamma$, $\pt > 145\,\eV[G]$                              & $24$                          & $47$                          \\
+\hline
+\multirow{3}{*}{Diphoton}      & loose $\gamma\gamma$, both $\pt > 55\,\eV[G]$                    & $3.0$                         & $7$                           \\
+                               & $\gamma\gamma$, $\pt > 40,30\,\eV[G]$                            & $3.0$                         & $21$                          \\
+                               & isolated tight $\gamma\gamma$, both $\pt > 25\,\eV[G]$           & $2.0$                         & $15$                          \\
+\hline
+\multirow{3}{*}{Single jet}    & jet ($R = 0.4$), $\pt > 435\,\eV[G]$                             & $3.7$                         & $35$                          \\
+                               & jet ($R = 1.0$), $\pt > 480\,\eV[G]$                             & $2.6$                         & $42$                          \\
+                               & jet ($R = 1.0$), $\pt > 450\,\eV[G]$, $m_j > 45\,\eV[G]$         & $2.6$                         & $36$                          \\
+\hline
+\multirow{5}{*}{$b$-jet}       & $b$-tag @ $60\%$, $\pt > 285\,\eV[G]$                            & $3.6$                         & $15$                          \\
+                               & two $b$-tag @ $60\%$, $\pt > 185,70\,\eV[G]$                     & $3.6$                         & $11$                          \\
+                               & $b$-tag @ $60\%$, three jets, all $\pt > 85\,\eV[G]$             & $1.5$                         & $14$                          \\
+                               & two $b$-tag @ $60\%$, one jet, $\pt > 65, 65,160\,\eV[G]$        & $1.3$                         & $17$                          \\
+                               & two $b$-tag @ $60\%$, two jets, all $\pt > 65\,\eV[G]$           & $3.2$                         & $15$                          \\
+\hline
+\multirow{4}{*}{Multi-jet}     & four jets, all $\pt > 125\,\eV[G]$                               & $0.5$                         & $16$                          \\
+                               & five jets, all $\pt > 95\,\eV[G]$                                & $4.8$                         & $10$                          \\
+                               & six jets, all $\pt > 80\,\eV[G]$                                 & $4.8$                         & $4$                           \\
+                               & six jets, all $|\eta| < 2$, all $\pt > 60\,\eV[G]$               & $4.8$                         & $15$                          \\
+\hline
+$\met$                         & $\met > 200\,\eV[G]$                                             & $5.1$                         & $94$                          \\
+\hline
+\multirow{4}{*}{$B$-physics}   & $\mu\mu$, $\pt > 11, 6\,\eV[G]$, $0.1 < m_{\mu\mu} < 14\,\eV[G]$ & $2.9$                         & $55$                          \\
+                               & $\mu\mu$, $\pt > 6, 6\,\eV[G]$, $2.5 < m_{\mu\mu} < 4.0\,\eV[G]$ & $1.4$                         & $55$                          \\
+                               & $\mu\mu$, $\pt > 6, 6\,\eV[G]$, $4.7 < m_{\mu\mu} < 5.9\,\eV[G]$ & $1.4$                         & $6$                           \\
+                               & $\mu\mu$, $\pt > 6, 6\,\eV[G]$, $7 < m_{\mu\mu} < 12\,\eV[G]$    & $1.2$                         & $12$
+\end{tabular}
+\end{adjustbox}
+\caption[
+Summary of \atlas\ trigger rates in 2018 data
+]{%
+Summary of \atlas\ trigger rates in 2018 data at the peak instantaneous
+luminosity of $2\times10^{34}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
+The search presented in this thesis uses only dilepton $e$ and $\mu$ triggers,
+which are highlighted in bold.
+Table adapted from~\cite{ATL-DAQ-PUB-2019-001}.
+Rates overlap between triggers.
+}
+\label{tab:experiment_trigger_breakdown}
+\end{table}
+
 \section{Data}

### 2023-03-27 main/chapter_experiment.tex
add trigger yap paragraph

@@
 
-\TODO{paragraph for Table~\ref{tab:experiment_trigger_breakdown}.}
+\atlas\ implements a menu of many different triggers for specific categories of
+interesting events in a representative time period.
+This trigger menu includes triggers for combinations of up to three charged
+leptons, one or two photons, up to six jets, up to two $b$-tagged jets, missing
+transverse momentum, or hadronic states for $B$ physics.
+Of these, Table~\ref{tab:experiment_trigger_breakdown} summarises the most
+important trigger categories and their trigger rates under peak Run~2
+operations.
+In total, the peak HLT trigger rate was $1750\,\mathrm{Hz}$ for the main
+physics categories, and $200\,\mathrm{Hz}$ for the $B$-physics trigger category
+(which also triggers on other, non-$B$, light
+states)~\cite{ATL-DAQ-PUB-2019-001}.
+The $\twoljets$ search presented in this thesis uses only dilepton triggers
+for electrons and muons.
+As shown in Table~\ref{tab:experiment_trigger_breakdown}, these have a total
+HLT trigger rate less than $105\,\mathrm{Hz}$%
+\footnote{%
+The listed rates overlap, so the sum of their peak rates is only an upper
+bound on their union.
+},
+meaning that our chosen triggers cover only a few percent of \atlas's data.
 

### 2023-03-29 main/chapter_experiment.tex
add muon resolution

@@
 the muon spectrometer in Section~\ref{sec:atlas_muon}.
+These detectors' resolutions to relevant physical objects is described in
+Section~\label{sec:atlas_object_resolutions}.
 The final connection to data analysed in this thesis is made through the
@@
 }.
+
+\section{Object resolutions}
+\label{sec:atlas_object_resolutions}
+Ultimately, the majority usage of most of \atlas's data happens through their
+reconstructions into interpretable objects, such as muons, electrons, jets,
+photons, or taus.
+From one perspective, these reconstructions are themselves exact data, as they
+are nothing less than functions of observed data.
+However, reconstructed objects do not exactly coincide with some proposed
+underlying truth --- reconstructions more closely resemble that truth perturbed
+by some noise.
+\atlas's experimental resolution determines the scale of that noise.
+
+Since the $\twoljets$ analysis is most concerned with muons, electrons and
+jets.
+\TODO{and missing transverse momentum},
+this section reviews \atlas's experimental resolution for these objects.
+
+\paragraph{Muons.}
+Since hard muons are not stopped inside \atlas, their reconstruction derives
+from the curved tracks they leave in the inner detector and muon system as they
+traverse \atlas's strong magnetic fields.
+High tracking fidelity allows us to reliably reconstruct muons across most of
+their expected momentum spectrum, with $\sigma(\pt) / \pt \lesssim 10\%$ up to
+$\pt \approx 1\,\eV[T]$.
+But resolution must degrade as these tracks' curvatures reduce at very high
+energies, and Figure~\ref{fig:atlas_cutaway} illustrates this degradation.
+The relative momentum resolution is approximately Gaussian in
+$\sigma(\pt) / \pt$ for muons across this momentum range
+(and therefore non-Gaussian in $\pt$ itself)~\cite{atlas1999design1}.
+
+At a reference transverse momentum of $\pt = 100\,\eV[G]$, \atlas's typical
+relative muon momentum resolution is at the $\sigma(\pt) / \pt \approx 2\%$
+level.
+However, non-uniformities in the detector geometry cause this to spike up to
+$\approx 10\%$ in local areas of the $\eta\textrm{--}\phi$ angular space;
+these degraded local areas include regions containing cryostat hardware or
+the detector's supporting feet~\cite{atlas1999design1, CERN-LHCC-97-022}.
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/atlas_muon_resolution.pdf}
+\caption[
+Relative muon momentum resolution in \atlas\ as a function of $\pt$
+]{%
+Relative muon momentum resolution in \atlas\ as a function of $\pt$, for
+muons selected within a pseudorapidity window of $|\eta| < 1.5$ and averaged
+over that window.
+The dots mark resolution values extracted from simulations, and the line
+represents a semi-empirical description of the relative
+resolution~\cite{CERN-LHCC-97-022}.
+This figure is reproduced from~\cite{atlas1999design1}.
+}
+\label{fig:atlas_cutaway}
+\end{figure}
+
+\paragraph{Electrons.}
+
+\paragraph{Jets.}
+
+\paragraph{Missing transverse momentum.}
+\TODO{}
 

### 2023-03-29 main/chapter_experiment.tex
add electron resolution

@@
 The relative momentum resolution is approximately Gaussian in
-$\sigma(\pt) / \pt$ for muons across this momentum range
+$1 / \pt$ for muons across this momentum range
 (and therefore non-Gaussian in $\pt$ itself)~\cite{atlas1999design1}.
@@
 \paragraph{Electrons.}
+Electrons are measured by both tracking in the inner detector and calorimetry
+in the electromagnetic calorimeter.
+These two measurements give two complementary inputs that aid in electron
+reconstruction.
+As for muons, the resolution from tracking scales with $1/\pt$, but the
+calorimetry error is dominated by Poisson statistics with a number of charged
+particles that scales in proportion to the electron energy $E$.
+Combining these two factors with a constant baseline noise, one can therefore
+model the relative electron resolution as
+\begin{equation}
+\label{eqn:experiment_electron_resolution}
+\frac{\sigma(E)}{E}
+\approx
+\frac{a_1}{\sqrt{E}}
+\oplus
+\frac{a_2}{\pt}
+\oplus
+a_3
+\end{equation}
+in which $\oplus$ indicates a sum in quadrature.
+In \atlas, these $a_i$ parameters are roughly
+$a_1 \approx 0.1\,\sqrt{\eV[G]}$,
+$a_2 \approx 0.25\,\eV[G]$, and
+$a_3 \approx 0.007$~\cite{nackenhorst2010top, Lester:705139, ERDMANN201418}.
+Although resolutions vary across the solid angle, this formula roughly
+reproduces the reported resolutions of
+$\frac{\sigma(E)}{E} \approx 4\%$ at $E = \pt = 10\,\eV[G]$,
+$\frac{\sigma(E)}{E} \approx 1\%$ at $E = \pt = 100\,\eV[G]$, and
+$\frac{\sigma(E)}{E} \approx 0.1\%$ at
+$E = \pt = 1\,\eV[T]$ for electrons~\cite{atlas2019electron}.
+
+Energy losses from high-energy electrons are dominated by Bremsstrahlung
+radiations, in which the electron radiates a photon and recoils from its
+momentum.
+Electron reconstruction algorithms efficiently recover this lost Bremsstrahlung
+momentum by associating nearby photon calorimeter clusters back to their source
+electron many cases~\cite{atlas1994proposal, atlas1999design1}.
 
 \paragraph{Jets.}
+\begin{equation}
+\label{eqn:experiment_jet_resolution}
+\frac{\sigma(E)}{E}
+\approx
+\frac{b_1}{\sqrt{E}}
+\oplus
+b_2
+\end{equation}
+with
+$b_1 \approx 0.5\,\sqrt{\eV[G]}$ and
+$b_2 \approx 0.03$.
 
@@
 \TODO{}
+% Reference from Holly
+%  ATLAS collaboration. ET
+% miss performance in the ATLAS detector using 2015–2016 LHC
+% pp collisions. ATLAS-CONF-2018-023, 2016.
 

### 2023-03-29 main/chapter_experiment.tex
add jet resolution

@@
 
-At a reference transverse momentum of $\pt = 100\,\eV[G]$, \atlas's typical
-relative muon momentum resolution is at the $\sigma(\pt) / \pt \approx 2\%$
-level.
-However, non-uniformities in the detector geometry cause this to spike up to
-$\approx 10\%$ in local areas of the $\eta\textrm{--}\phi$ angular space;
-these degraded local areas include regions containing cryostat hardware or
-the detector's supporting feet~\cite{atlas1999design1, CERN-LHCC-97-022}.
-
 \begin{figure}[tp]
@@
 
+At a reference transverse momentum of $\pt = 100\,\eV[G]$, \atlas's typical
+relative muon momentum resolution is at the $\sigma(\pt) / \pt \approx 2\%$
+level.
+However, non-uniformities in the detector geometry cause this to spike up to
+$\approx 10\%$ in local areas of the $\eta\textrm{--}\phi$ angular space;
+these degraded local areas include regions containing cryostat hardware or
+the detector's supporting feet~\cite{atlas1999design1, CERN-LHCC-97-022}.
+
 \paragraph{Electrons.}
@@
 calorimetry error is dominated by Poisson statistics with a number of charged
-particles that scales in proportion to the electron energy $E$.
+particles that scales in proportion to the electron's energy.
 Combining these two factors with a constant baseline noise, one can therefore
@@
 \label{eqn:experiment_electron_resolution}
-\frac{\sigma(E)}{E}
+\frac{\sigma(\pt)}{\pt}
 \approx
-\frac{a_1}{\sqrt{E}}
+\frac{a_1}{\sqrt{\pt}}
 \oplus
@@
 a_3
+,
 \end{equation}
 in which $\oplus$ indicates a sum in quadrature.
-In \atlas, these $a_i$ parameters are roughly
+To describe \atlas, these $a_i$ parameters are roughly
 $a_1 \approx 0.1\,\sqrt{\eV[G]}$,
@@
 Although resolutions vary across the solid angle, this formula roughly
-reproduces the reported resolutions of
-$\frac{\sigma(E)}{E} \approx 4\%$ at $E = \pt = 10\,\eV[G]$,
-$\frac{\sigma(E)}{E} \approx 1\%$ at $E = \pt = 100\,\eV[G]$, and
-$\frac{\sigma(E)}{E} \approx 0.1\%$ at
-$E = \pt = 1\,\eV[T]$ for electrons~\cite{atlas2019electron}.
+reproduces the reported electron resolutions of
+$\frac{\sigma(\pt)}{\pt} \approx 4\%$ at $\pt = 10\,\eV[G]$,
+$\frac{\sigma(\pt)}{\pt} \approx 1\%$ at $\pt = 100\,\eV[G]$, and
+$\frac{\sigma(\pt)}{\pt} \approx 0.1\%$ at
+$\pt = 1\,\eV[T]$~\cite{atlas2019electron}.
 
@@
 \paragraph{Jets.}
+Although charged hadrons are tracked in the inner detector, the total energy of
+jets also depends vitally on neutral hadrons, which are only observed through
+their hadronic interactions in the calorimeter system.
+With less influence from tracking, calorimetry noise is the dominant
+contribution to the experimental resolution on hadronic jets.
+However, other noise contributions arising from electrical signals and pile-up
+scale in proportion in $1/\pt$.
+one can therefore again approximate \atlas's relative jet energy resolution as
 \begin{equation}
 \label{eqn:experiment_jet_resolution}
-\frac{\sigma(E)}{E}
+\frac{\sigma(\pt)}{\pt}
 \approx
-\frac{b_1}{\sqrt{E}}
+\frac{b_1}{\sqrt{\pt}}
 \oplus
-b_2
+\frac{b_2}{\pt}
+\oplus
+b_3
+,
 \end{equation}
+the same functional form as for electrons in
+Equation~\ref{eqn:experiment_electron_resolution},
 with
-$b_1 \approx 0.5\,\sqrt{\eV[G]}$ and
-$b_2 \approx 0.03$.
+$b_1 \approx 1.0\,\sqrt{\eV[G]}$
+$b_2 \approx 0.5\,\eV[G]$, and
+$b_3 \approx 0.03$~\cite{atlas2021jetresolution}.
+This formula approximately describes the measured jet resolutions of
+$\frac{\sigma(\pt)}{\pt} \approx 20\%$ at $\pt = 30\,\eV[G]$,
+$\frac{\sigma(\pt)}{\pt} \approx 8\%$ at $\pt = 200\,\eV[G]$, and
+$\frac{\sigma(\pt)}{\pt} \approx 4\%$ at
+$\pt = 2\,\eV[T]$~\cite{atlas2021jetresolution}.
+
+Note that hadronic weak decays can produce real, not-prompt missing momentum
+in jets.
+This effect can skew jet reconstruction distributions towards low energies,
+and is particularly prominent for heavy-flavour jets~\cite{nackenhorst2010top}.
 

### 2023-03-29 main/chapter_experiment.tex
add met resolution

@@
 Since the $\twoljets$ analysis is most concerned with muons, electrons and
-jets.
-\TODO{and missing transverse momentum},
-this section reviews \atlas's experimental resolution for these objects.
+jets and missing transverse momentum,
+this section reviews \atlas's experimental resolutions for these objects.
 
@@
 $b_2 \approx 0.5\,\eV[G]$, and
-$b_3 \approx 0.03$~\cite{atlas2021jetresolution}.
-This formula approximately describes the measured jet resolutions of
+$b_3 \approx 0.03$.
+With these parameters, formula approximately describes the measured relative
+jet resolutions of
 $\frac{\sigma(\pt)}{\pt} \approx 20\%$ at $\pt = 30\,\eV[G]$,
@@
 \paragraph{Missing transverse momentum.}
-\TODO{}
-% Reference from Holly
-%  ATLAS collaboration. ET
-% miss performance in the ATLAS detector using 2015–2016 LHC
-% pp collisions. ATLAS-CONF-2018-023, 2016.
+Measurement resolutions of all other objects propagate into the missing
+transverse momentum; this is necessary as it is not directly observed itself,
+but derived from other objects' transverse momenta.
+But collisions also emit significant amounts of momentum in softer particles,
+which can leave tracks and calorimeter hits without being clustered into hard
+physical objects such as jets or electrons.
+The $\met$ resolution therefore depends not only on the resolutions of hard
+objects, but also on the measurement of this soft contribution.
+Various different $\met$ definitions, known as `working points', use different
+quality criteria to define the $\met$ composition, and each working point
+balances resolution against accuracy for different context-specific
+objectives~\cite{ATLAS-CONF-2018-023, Pacey:2747774}.
+
+Pile-up is a major contributor to the $\met$ resolution, because secondary
+interactions produce both tracks and calorimeter clusters that might can be
+indistinguishable from products of the hard-scatter primary vertex
+(the main interaction of interest in a bunch crossing).
+All misattributed momentum from these additional vertices causes $\met$ noise.
+The \atlas\ $\met$ resolution is reported in two orthogonal transverse
+components: $\sigma(\metx)$ and $\sigma(\mety)$.%
+\footnote{%
+The directional decomposition of $\ptmiss$ uncertainty is important for the
+$\met$ significance variable $\metsig$, which we shall discuss in
+Section~\ref{sec:2ljets_metsig}.
+}%
+\footnote{%
+Despite their $E$ symbols, these variables $\metx$ and $\mety$ are momenta
+because they have spatial directions.%
+}
+In reference samples of simulated $Z \to \mu\mu$ events, both transverse
+$\met$ resolution components increase with the number of primary vertices in a
+bunch crossing ($n_{PV}$), from
+$\sigma(\metx) = \sigma(\mety) \approx 10\,\eV[G]$ at $n_{PV} = 1$, to
+$\sigma(\metx) = \sigma(\mety) \approx 20\,\eV[G]$ at $n_{PV} = 30$.
+Note that these resolutions vary by around $\pm 20\%$ with the $\met$
+working point choice~\cite{ATLAS-CONF-2018-023}.
 

### 2023-04-01 main/chapter_experiment.tex
begin lhc section; add complex figure

@@
 
+Without high-energy hadron collisions in its centre, \atlas\ would be neutered.
+For all of our data, we are therefore indebted to the LHC for so reliably
+providing these exciting particle interactions.
+We therefore begin this chapter with a description of the LHC itself in
+Section~\ref{sec:atlas_lhc} before detailing the design of \atlas\ in
+Section~\ref{sec:atlas_design}.
+
+\section{The Large Hadron Collider}
+\label{sec:atlas_lhc}
+
+% the LHC
+% where the protons come from
+Figure~\ref{fig:atlas_cern_complex}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.98\textwidth]{figures/atlas_lhc_complex.png}
+\caption[
+The CERN accelerator and detector complex, with the LHC as its largest piece
+]{%
+The CERN accelerator and detector complex, with the LHC as its largest piece
+shown in dark blue-grey.
+The complex uses a sequence of increasingly large and powerful accelerators to
+boost ionised particles to their final collision energies in the LHC and to
+yield beams for other CERN experiments.
+This poster is adapted from~\cite{Haffner:1621894}.
+}
+\label{fig:atlas_cern_complex}
+\end{figure}
+
+% the accelerator complex
+% RF cavities
+% magnets (dipole, quadrupole etc)
+% size of beam
+% luminosity and its uncertainty
+% pile-up
+
 \section{Design}
+\label{sec:atlas_design}
 Atlas is a Titan.

### 2023-04-01 main/chapter_experiment.tex
add pileup plot

@@
 Without high-energy hadron collisions in its centre, \atlas\ would be neutered.
-For all of our data, we are therefore indebted to the LHC for so reliably
-providing these exciting particle interactions.
+We are therefore indebted to the LHC for so reliably providing the exciting
+hadron collisions that are necessary for all of our proton collision data.
 We therefore begin this chapter with a description of the LHC itself in
-Section~\ref{sec:atlas_lhc} before detailing the design of \atlas\ in
+Section~\ref{sec:atlas_lhc}, before detailing the design of \atlas\ in
 Section~\ref{sec:atlas_design}.
@@
 % where the protons come from
-Figure~\ref{fig:atlas_cern_complex}
+Figure~\ref{fig:experiment_cern_complex}
 
@@
 \centering
-\includegraphics[width=0.98\textwidth]{figures/atlas_lhc_complex.png}
+\includegraphics[width=0.99\textwidth]{figures/atlas_lhc_complex.png}
 \caption[
@@
 }
-\label{fig:atlas_cern_complex}
+\label{fig:experiment_cern_complex}
 \end{figure}
@@
 % pile-up
+\TODO{Distribution in the mean number of primary vertices per crossing, where the
+variability in that mean arises from differing beam conditions over run-time.}
+Figure~\ref{fig:experiment_run_2_mu}
+
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.62\textwidth]{figures/atlas_pileup_mu_2015_2018.pdf}
+\caption[
+Mean number interactions per bunch crossing in Run~2 of the LHC
+]{%
+Mean number interactions per bunch crossing in Run~2 of the LHC;
+rates presented are weighted by luminosity and sub-divided by year of
+operation.
+This figure includes all data that \atlas\ recorded with stable LHC proton
+beams.
+The reported integrated luminosities therefore exceed those we present later
+in Table~\ref{tab:experiment_run2_state}, which are further filtered to require
+other data-goodness criteria.
+This figure is reproduced from~\cite{atlas_public_luminosity}.
+}
+\label{fig:experiment_run_2_mu}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/atlas_cutaway_volume_1.pdf}
+\includegraphics[width=0.62\textwidth]{figures/atlas_cutaway_volume_1.pdf}
 \caption[
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/atlas_magnets.pdf}
+\includegraphics[width=0.62\textwidth]{figures/atlas_magnets.pdf}
 \caption[

### 2023-04-08 main/chapter_experiment.tex
type some lhc notes from reading
begin lhc accelerator complex words

@@
 
-\section{The Large Hadron Collider}
+\section{The Large Hadron Collider at CERN}
 \label{sec:atlas_lhc}
+% The LHC
+% largest piece
+The Large Hadron Collider (LHC) is the largest and most powerful piece of
+CERN's extensive particle accelerator complex.
+It lives in a $26.7\,\mathrm{km}$ (circumference) tunnel, which is buried
+around $100\,\mathrm{m}$ under France and Switzerland at CERN, Geneva.
+It and comprises two rings that contain and accelerate counter-rotating bunches
+of particles.
+These accelerated bunches are brought into near-head-on collisions at
+the four large detector experiments:
+\atlas~\cite{atlas2008experiment},
+\lhcb~\cite{lhcb2008experiment},
+\cms~\cite{cms2008experiment}, and
+\alice~\cite{alice2008experiment}.
+Although proton-proton collisions are the LHC's primary subject, it is also
+operated for heavy ion collisions including lead ($^{208}\mathrm{Pb}^{82+}$)
+and oxygen ($^{16}\mathrm{O}^{4+}$)~\cite{
+Brewer:2021kiv,
+lhc2008machine,
+lhcDesignreport2004
+}.
+Heavy ion collisions are primarily studied by the \alice\ experiments, but are
+also analysed by the general-purpose detectors \atlas\ and \cms.
+Since proton collision data are the topic of this thesis, this section focuses
+on proton operations.
 
-% the LHC
-% where the protons come from
-Figure~\ref{fig:experiment_cern_complex}
+% does not stand alone
+The LHC does not stand alone.
+It implements the final boosts and collisions of particles, but requires
+earlier stages of CERN's systems to inject the protons at suitably high
+energies for it to continue their acceleration.
+In brief, the accelerator chain flows as follows:
+% Accelerator complex
+\begin{enumerate}
+% TODO where the protons come from
+\item The protons come from
+\end{enumerate}
+This sequence is a subset of CERN's full accelerator complex,
+which is illustrated in
+Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894}.
 
@@
 % the accelerator complex
-% RF cavities
-% magnets (dipole, quadrupole etc)
-% size of beam
-% luminosity and its uncertainty
-% pile-up
+% - sequence of contributing accelerators and their beam energies
+% beampipe restricted by size of lep tunnel
+% - required particle count makes antiprotons hard
+% - proton-proton same charge requires separate beampipes
+% - size means most cost efficient to put both beams in one cryostat
+% - beampipes therefore mechanically and magnetically coupled
+% - superconducting magnets at around 2 to 4 Kelvin
+% - cooled with superfluid helium
+% - low heat capacity, danger of heating from beam radiation
+% Radiofrequency cavities
+% - radio waves at 400 MHz such that slow protons accelerated, fast decelerated
+% - in LHC, accelerates from 450 GeV to 6.5 TeV (7 target)
+% magnets
+% - dipole for steering
+% - quadrupole for focussing (pairs)
+% - higher orders why? septupole at beam entry
+% TODO size of beam
+% TODO luminosity and its uncertainty
+% TODO pile-up
 \TODO{Distribution in the mean number of primary vertices per crossing, where the
@@
 \centering
-\includegraphics[width=0.62\textwidth]{figures/atlas_pileup_mu_2015_2018.pdf}
+\includegraphics[width=0.7\textwidth]{figures/atlas_pileup_mu_2015_2018.pdf}
 \caption[

### 2023-04-08 main/chapter_experiment.tex
some on pre-accelerators; split run3 upgrade

@@
 \begin{enumerate}
-% TODO where the protons come from
-\item The protons come from
+% where the protons come from
+\item The protons come from hydrogen gas, which is initially ionised into a
+plasma of protons and $e^-$ ions in the strong electric fields of a
+duoplasmotron ion source, in which the protons are initially accelerated by
+electric fields towards a cathode~\cite{lhc50years2013}.
+% linac2
+\item After ionisation, a linear accelerator named Linac2 provides the first
+major proton acceleration in the chain, reaching $50\,\eV[M]$
+energies~\cite{lhc50years2013}.
+% PSB
+\item Secondly, the Proton Synchrotron Booster (PSB) accelerates these protons
+to $1.4\,\eV[G]$ energies.
+The PSB comprises four rings with $160\,\mathrm{m}$
+circumference~\cite{linac4protonsource,lhc50years2013}.
+% PS
+\item Thirdly, protons from the PSB are accelerated by the
+Proton Synchrotron (PS) to $25\,\eV[G]$ energies~\cite{lhc2000psInjector}.
+TODO
+% SPS
+\item SPS
+% LHC
+\item LHC
 \end{enumerate}
@@
 Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894}.
+Each of these accelerators was upgraded from its previous operations to achieve
+the performance demanded by designs for the LHC's high energy and luminosity.
+
+Note that this sequence describes the LHC accelerator complex as it was for
+operations that generated the data used in this thesis, in `Run~2' of the LHC.
+It has since been upgraded.
+For example, since the year 2020 (for Run~3 onwards), Linac2 is replaced by a
+new linear accelerator named Linac4,
+which accelerates $\mathrm{H}^-$ ions to $160\,\eV[M]$;
+these $\mathrm{H}^-$ ions are later stripped of their electrons,
+to bare protons by passage through a carbon foil.
+Linac4 will supply higher luminosities for future operations of the accelerator
+complex~\cite{linac4techreport,linac4protonsource}.
+The PSB is since also upgraded to reach $2\,\eV[G]$ energies.
 

### 2023-04-09 main/chapter_experiment.tex
add rf cavities

@@
 % where the protons come from
-\item The protons come from hydrogen gas, which is initially ionised into a
+\item[0.] The protons come from hydrogen gas, which is initially ionised into a
 plasma of protons and $e^-$ ions in the strong electric fields of a
-duoplasmotron ion source, in which the protons are initially accelerated by
+`duoplasmotron' ion source, in which the protons are initially accelerated by
 electric fields towards a cathode~\cite{lhc50years2013}.
@@
 \item Thirdly, protons from the PSB are accelerated by the
+$628\,\mathrm{m}$-circumference
 Proton Synchrotron (PS) to $25\,\eV[G]$ energies~\cite{lhc2000psInjector}.
-TODO
 % SPS
-\item SPS
+\item Fourthly, protons from the PS are further boosted by the
+$6.9\,\mathrm{km}$-circumference
+Super Proton Synchrotron (SPS) to $450\,\eV[G]$
+energies~\cite{synchroton:445034}.
 % LHC
-\item LHC
+\item Finally, protons from the SPS are injected the LHC, which finishes
+their acceleration from $250\,\eV[G]$ to $6.5\,\eV[T]$, achieving the
+world-leading $\sqrt{s} = 13\,\eV[T]$ proton-proton collisions that are studied
+in this thesis.
 \end{enumerate}
 This sequence is a subset of CERN's full accelerator complex,
-which is illustrated in
-Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894}.
-Each of these accelerators was upgraded from its previous operations to achieve
-the performance demanded by designs for the LHC's high energy and luminosity.
+illustrated in Figure~\ref{fig:experiment_cern_complex}~\cite{Haffner:1621894},
+which also supports various other engineering and research experiments.
+Each of these accelerators (Linac2, PSB, PS, and SPS) was upgraded from its
+previous operations to achieve the performance demanded by designs for the
+LHC's high energy and luminosity.
 
 Note that this sequence describes the LHC accelerator complex as it was for
-operations that generated the data used in this thesis, in `Run~2' of the LHC.
+proton-proton operations that generated the data used in this thesis,
+in `Run~2' of the LHC.
 It has since been upgraded.
@@
 complex~\cite{linac4techreport,linac4protonsource}.
-The PSB is since also upgraded to reach $2\,\eV[G]$ energies.
 
@@
 
-% the accelerator complex
-% - sequence of contributing accelerators and their beam energies
-% beampipe restricted by size of lep tunnel
-% - required particle count makes antiprotons hard
-% - proton-proton same charge requires separate beampipes
-% - size means most cost efficient to put both beams in one cryostat
-% - beampipes therefore mechanically and magnetically coupled
-% - superconducting magnets at around 2 to 4 Kelvin
-% - cooled with superfluid helium
-% - low heat capacity, danger of heating from beam radiation
-% Radiofrequency cavities
-% - radio waves at 400 MHz such that slow protons accelerated, fast decelerated
-% - in LHC, accelerates from 450 GeV to 6.5 TeV (7 target)
-% magnets
+The LHC's physical design is strongly influenced by the layout of its tunnel,
+which was previously occupied by the Large Electron-Positron Collider (LEP).
+Colliders of oppositely-charged particle pairs, such as electrons and positrons
+or protons and antiprotons, can accelerate both partners in the same beam-pipe.
+The same is not true of like-charged pairs, such as protons against protons.
+This means that the LHC must squeeze two beam-pipes into the narrow tunnel
+where LEP's one beam-pipe previously sat.
+The LHC's solution is to put both beams next to each other, in
+the same cryostatic container.
+This ``two-in-one'' solution fits both beams into the physical space, but adds
+engineering challenges because the two beam channels are mechanically and
+magnetically coupled due to their
+proximity~\cite{lhcDesignreport2004,lhc2008machine}.
+
+Radiofrequency (RF) cavities are the technology used to accelerate protons in
+the LHC.
+They operate by setting up electron beams to oscillate at a frequency that
+resonates with a specially designed cavity, this producing strong oscillating
+electromagnetic fields.
+These electromagnetic fields oscillate at $400\,\mathrm{MHz}$ in the LHC, and
+must be carefully tuned jointly with the incident proton beams to align in
+their timing.
+When perfectly timed, the RF cavity will accelerate protons that are less
+energetic that the target beam energy, and decelerate any that are more
+energetic.
+This achieves a dynamically stable acceleration.
+The LHC uses $16$ superconducting RF cavities contained in four separate
+cryostatic modules~\cite{lhcDesignreport2004}.
+
+% TODO magnets
 % - dipole for steering

### 2023-04-09 main/chapter_experiment.tex
write about magnets

@@
 This achieves a dynamically stable acceleration.
-The LHC uses $16$ superconducting RF cavities contained in four separate
-cryostatic modules~\cite{lhcDesignreport2004}.
+The LHC uses sixteen superconducting RF cavities, which are contained in four
+separate cryostatic modules~\cite{lhcDesignreport2004}.
 
-% TODO magnets
+% Magnets
+Magnets are used to steer and focus the LHC's proton beams.
 % - dipole for steering
+Dipole magnets conduct steering by creating constant magnetic fields, with
+north and south magnetic poles on diametrically-opposite sides of the beam.
+These dipole fields steer the beam by applying constant Lorentz forces
+perpendicular to the beam's motion.
 % - quadrupole for focussing (pairs)
-% - higher orders why? septupole at beam entry
+Quadrupole magnets focus the beam to counteract its natural dispersion, with
+alternating magnetic poles placed after each $90^\circ$ rotation around the
+beam.
+A quadrupole magnet is focussing in one axis, and defocussing in a
+perpendicular axis.
+However, alternating pairs of quadrupole magnets can achieve
+a total focussing effect; their effects are not linear.
+
+% higher orders
+Sextupole, octupole, and decapole magnets are used to implement detailed
+corrections with each of the dipoles magnets.
+Chromatic aberrations in optics are errors that arise from the colour-dependent
+behaviour of light.
+The analogous momentum-dependent behaviour in particle beams is known as
+`chromaticity'.
+Sextupole magnets are used in the LHC to correct for chromaticity at each
+quadrupole~\cite{Guiducci:1992kb, lhc2008machine}.
+
 % TODO size of beam

### 2023-04-09 main/chapter_experiment.tex
add beam size

@@
 
-\section{The Large Hadron Collider at CERN}
+\section{The Large Hadron Collider}
 \label{sec:atlas_lhc}
@@
 `chromaticity'.
-Sextupole magnets are used in the LHC to correct for chromaticity at each
+Sextupole magnets are also used in the LHC to correct for chromaticity at each
 quadrupole~\cite{Guiducci:1992kb, lhc2008machine}.
 
-% TODO size of beam
+% size of beam
+By its peak luminosity design specifications, each bunch of an LHC beam
+should contain $1.15\times10^{11}$ protons.
+The ring can contain up to $2808$ bunches.
+After acceleration to collision energy, each bunch has a root-mean-square
+(RMS, standard-deviation) length of $7.55\,\mathrm{cm}$.
+The beam has a transverse RMS width of $16.7\,\mathrm{\mu m}$ at
+interaction points 1 and 5 (\atlas\ and \cms),
+and of $70.9\,\mathrm{\mu m}$ at
+interaction points 2 and 8 (\alice\ and \lhcb)%
+~\cite{lhcDesignreport2004}.
+
 % TODO luminosity and its uncertainty

### 2023-04-09 main/chapter_experiment.tex
add lumi measurement and uncertainty

@@
 
-% TODO luminosity and its uncertainty
+% luminosity and its uncertainty
+The luminosity delivered to each detector varies with its local beam controls.
+For example, beams at the \lhcb\ and \alice\ interaction points are
+deliberately offset to reduce data rates and `pile-up' backgrounds from
+multiple interactions~\cite{Follin:2014nva}
+The peak instantaneous luminosity delivered by the LHC to \atlas\ in Run~2 was
+$1.9\times10^{34}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2018;
+see Table~\ref{tab:experiment_run2_state} for more details here.
+% Van Der Meer scans
+Luminosity is measured in \atlas\ during dedicated low-luminosity runs of the
+LHC, using the ``van der Meer'' method, which measures interaction rates while
+scanning the beams across each other in transverse
+space~\cite{vanderMeer:296752,ATLAS:2022hro}.
+These \atlas\ luminosity measurements primarily rely on data from the
+LUCID2 Cherenkov detector, which comprises two parts that are offset from the
+interaction point by about $\pm 17\,\mathrm{m}$ along the beam-pipe~\cite{
+SBARRA2019152,
+Avoni_2018
+}.
+The relative uncertainty on \atlas's Run~2 integrated luminosity is $\pm1.7\%$.
+
 % TODO pile-up

### 2023-04-09 main/chapter_experiment.tex
add pile-up

@@
 
-% TODO pile-up
-\TODO{Distribution in the mean number of primary vertices per crossing, where the
-variability in that mean arises from differing beam conditions over run-time.}
-Figure~\ref{fig:experiment_run_2_mu}
+% pile-up
+From a theoretical standpoint, we would like to study individual proton
+scattering events. But from a practical standpoint, to get enough interesting
+data with limited resources we must increase the luminosity.
+We therefore expect many proton scatterings per bunch crossing, and hope that
+one of them is interesting enough to record.
+On reconstruction, each proton-proton scattering in a bunch crossing produces
+a vertex.%
+\footnote{%
+Sometimes each vertex is described as a primary vertex, to distinguish it from
+secondary vertices from decaying products. Here, we reserve `primary' for the
+main hard scatter of interest in an event.%
+}
+The number of vertices in a bunch crossing is the pile-up number.
+The primary vertex is the main interaction of interest.
+It is typically the hardest hard scatter, and is identified in \atlas\ as the
+vertex with the largest sum of squared transverse momenta of its
+tracks~\cite{atlas2022searches}.
 
@@
 \end{figure}
+
+% pile-up
+Pile-up interactions cause tracks and other detector hits, which are noise
+atop our measurements of the primary vertex.
+If not removed, these pile-up signals can skew measurements and create an
+illusion missing transverse momentum.
+Although precise measurements can help to remove this noise, it reduces the
+information content of our data and cannot be cleaned entirely.
+In Run~2 data, the number of pile-up vertices per bunch crossing varies around
+several tens of vertices, as displayed in
+Figure~\ref{fig:experiment_run_2_mu}~\cite{atlas_public_luminosity}.
+Note that the variability in the mean pile-up count arises from differing beam
+conditions over run-time; for example, the instantaneous luminosity can only
+decrease during each LHC beam fill, and the expected pile-up rate decreases
+with it.
+
+Having reviewed the LHC on which our detectors are constructed, we now move on
+to the design of the \atlas\ detector itself.
 

### 2023-04-09 main/chapter_theory.tex
begin lagrangian form

@@
 As a gauge field theory, the Standard Model has the symmetry group
-$SU\!(3)_C \times SU\!(2)_L \times U\!(1)$,
+$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
 meaning that its Lagrangian is invariant under local operations of that group
 on its fields in certain representations.
+The subscript $_C$ refers to colour, $_L$ refers to left-chiral fields, and
+$_Y$ refers to the weak hypercharge.
 
@@
 interactions with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
-$SU\!(2)_L$ and $U\!(1)$ symmetries, respectively.
+$SU\!(2)_L$ and $U\!(1)_Y$ symmetries, respectively.
 These would be massless if the symmetry were respected, but it is
@@
 \end{table}
+
+\subsection{Lagrangian formulation}
+\label{sec:theory_sm_lagrangian}
+This section briefly reviews the Lagrangian of the Standard Model.
+It is convenient to express such complicated mathematical expressions as sums
+(and (possibly nested) sums)
+of separate, labelled parts.
+We employ this strategy.
+
+The Standard Model Lagrangian can be decomposed into this sum of
+(partial Lagrangian) terms:
+\begin{equation}
+\label{eqn:theory_sm_lagrangian_top_sum}
+\mathcal{L}_\mathrm{SM} =
+\mathcal{L}_\mathrm{YM}
++ \mathcal{L}_\mathrm{WD}
++ \mathcal{L}_\mathrm{Yu}
++ \mathcal{L}_\mathrm{H}
+,
+\end{equation}
+which we define below;
+$\mathcal{L}_\mathrm{YM}$ corresponds to the gauge bosons,
+$\mathcal{L}_\mathrm{WD}$ corresponds to the fermions and their gauge
+interactions,
+$\mathcal{L}_\mathrm{Yu}$ corresponds to the Yukawa interactions between the
+fermions and the Higgs,
+and
+$\mathcal{L}_\mathrm{H}$ corresponds to the Higgs boson and its potential.
+
+\paragraph{Yang-Mills.}
+The Yang-Mills summand $\mathcal{L}_\mathrm{YM}$ implements the gauge groups
+$SU\!(3)_C$ (colour), $SU\!(2)_L$ (weak isospin), and $U\!(1)_Y$ (hypercharge),
+each with a nested term respectively:
+\begin{equation}
+\label{eqn:theory_sm_lagrangian_top_ym}
+\mathcal{L}_\mathrm{YM} =
+\mathcal{L}_\mathrm{QCD}
++ \mathcal{L}_\mathrm{I_W}
++ \mathcal{L}_\mathrm{Y}
+.
+\end{equation}
+All three Yang-Mills terms take similar forms:
+\begin{align}
+\label{eqn:theory_sm_lagrangian_top_ym_terms}
+\mathcal{L}_\mathrm{QCD} &=
+-\frac{1}{4g^2_3}
+\sum_{A=1}^8 G^A_{\mu\nu} G^{\mu\nu A}
+,\\
+\mathcal{L}_\mathrm{I_W} &=
+-\frac{1}{4g^2_2}
+\sum_{a=1}^3 A^a_{\mu\nu} F^{\mu\nu a}
+,\textrm{ and}\\
+\mathcal{L}_\mathrm{Y} &=
+-\frac{1}{4g^2_1} B{\mu\nu} B^{\mu\nu}
+.
+\end{align}
+Here, $g_3$, $g_2$, and $g_3$ are dimensionless constants.
+They relate to colour, weak isospin, and hypercharge, respectively.
+Furthermore, for colour,
+\begin{equation}
+G^A_{\mu\nu} =
+\partial_\mu A^A_\nu - \partial_\nu A^A_\mu - f^{ABC} A^B_\mu A^C_\nu
+,
+\end{equation}
+with $A, B, C \in \{1, 2, ..., 8\}$ indexing the eight gluon fields $A^B_\mu$ and
+the $SU\!(3)_C$ structure function $f^{ABC}$.
+For weak isospin,
+\begin{equation}
+F^a_{\mu\nu} =
+\partial_\mu W^a_\nu - \partial_\nu W^a_\mu - \epsilon^{abc} W^b_\mu W^c_\nu
+,
+\end{equation}
+with $a, b, c \in \{1, 2, 3\}$ indexing the vector bosons $W^a_\mu$ and
+the $SU\!(2)_L$ structure function $\epsilon^{abc}$.
+For hypercharge with the vector boson $B$,
+\begin{equation}
+B_{\mu\nu} =
+\partial_\mu B_\nu - \partial_\nu B_\mu
+.
+\end{equation}
+
+\paragraph{Weyl-Dirac.}
+The Weyl-Dirac summand $\mathcal{L}_\mathrm{WD}$ implements the fermions and
+their gauge interactions, and can be written:
+\begin{equation}
+\begin{split}
+\mathcal{L}_\mathrm{WD} =
+\sum_i^3
+\Bigl(
+& L^\dagger_i \sigma^\mu \mathcal{D}_\mu L_i
++ \bar e^\dagger_i \sigma^\mu \mathcal{D}_\mu \bar e_i \\
+& + \mathbf{Q}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{Q}_i
++ \mathbf{\bar u}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar u}_i
++ \mathbf{\bar d}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar d}_i
+\Bigr)
+,
+\end{split}
+\end{equation}
+where the $\mathcal{D}_\mu$ expressions corresponding to the fermion fields:
+$L_i$ for lepton doublets,
+$\bar e_i$ for lepton singlets,
+$\mathbf{Q}_i$ for quark doublets,
+and $\mathbf{\bar u}_i$ and $\mathbf{\bar d}_i$ for up- and down-type quark
+singlets, respectively.
+The three fermion generations are indexed by $i$.
+The quark fields come in colour triplets.
+
+\paragraph{Yukawa.} \TODO{}
+
+\paragraph{Higgs.} \TODO{}
+
+\TODO{add citations}
+%http://www.physics.ufl.edu/~ramond/JourneysChapter2_CUP.pdf
 

### 2023-04-09 main/chapter_theory.tex
add higgs and citations

@@
 + \mathcal{L}_\mathrm{WD}
-+ \mathcal{L}_\mathrm{Yu}
-+ \mathcal{L}_\mathrm{H}
++ \mathcal{L}_\mathrm{Yukawa}
++ \mathcal{L}_\mathrm{Higgs}
 ,
@@
 interactions,
-$\mathcal{L}_\mathrm{Yu}$ corresponds to the Yukawa interactions between the
+$\mathcal{L}_\mathrm{Yukawa}$ corresponds to the Yukawa interactions between the
 fermions and the Higgs,
 and
-$\mathcal{L}_\mathrm{H}$ corresponds to the Higgs boson and its potential.
+$\mathcal{L}_\mathrm{Higgs}$ corresponds to the Higgs boson and its
+potential~\cite{ramond1999journeys,pdg2022ynf}.
 
@@
 \mathcal{L}_\mathrm{Y} &=
--\frac{1}{4g^2_1} B{\mu\nu} B^{\mu\nu}
+-\frac{1}{4g^2_1} B_{\mu\nu} B^{\mu\nu}
 .
@@
 The three fermion generations are indexed by $i$.
-The quark fields come in colour triplets.
+The quark fields are in colour triplets.
 
-\paragraph{Yukawa.} \TODO{}
+\paragraph{Yukawa.}
+The Yukawa term implements masses for the charged leptons by coupling
+them to the scalar Higgs field $H$:
+\begin{equation}
+\mathcal{L}_\mathrm{Yukawa} =
+i\hat L_i \bar e_j H^* Y^{[e]}_{ij}
++ i\mathbf{\widehat Q}_i \mathbf{\bar d}_j H^* Y^{[d]}_{ij}
++ i\mathbf{\widehat Q}_i \mathbf{\bar u}_j \tau_2 H Y^{[u]}_{ij}
++ [\textrm{c.c.}]
+,
+\end{equation}
+in which $\tau_2$ contains weak isospin indices,
+and $Y^{[e]}_{ij}$, $Y^{[d]}_{ij}$, $Y^{[u]}_{ij}$ are Yukawa coupling matrices
+with shape $3\times3$ for the three fermion generations.
 
-\paragraph{Higgs.} \TODO{}
-
-\TODO{add citations}
-%http://www.physics.ufl.edu/~ramond/JourneysChapter2_CUP.pdf
+\paragraph{Higgs.}
+Finally, the Higgs term,
+\begin{equation}
+\mathcal{L}_\mathrm{Higgs} =
+(\mathcal{D}_\mu H)^\dagger (\mathcal{D}^\mu H)
+- \left[ -\mu^2 H^\dagger H + \lambda (H^\dagger H)^2 \right]
+,
+\end{equation}
+contains the dynamics and potential for the scalar Higgs doublet.
+The potential terms are isolated into square brackets to highlight the negative
+sign on the mass term $\mu^2$, which with the quartic term contributes to
+spontaneous symmetry breaking~\cite{rubakov1999classical}.
 

### 2023-04-10 main/chapter_theory.tex
clarify notation

@@
 
-The Standard Model Lagrangian can be decomposed into this sum of
+The Standard Model Lagrangian (density) can be decomposed into this sum of
 (partial Lagrangian) terms:
@@
 potential~\cite{ramond1999journeys,pdg2022ynf}.
+Note that this is only a condensed form of the full Standard Model Lagrangian,
+which neglects much of its complexity.
 
@@
 singlets, respectively.
+The notation $\bar x$ indicates the Dirac adjoint of $x$.
 The three fermion generations are indexed by $i$.
@@
 with shape $3\times3$ for the three fermion generations.
+The notation $H^*$ indicates the complex conjugate of the complex scalar
+Higgs field, and `c.c.' implies the complex conjugate of the expression to
+its left.
 
@@
 contains the dynamics and potential for the scalar Higgs doublet.
+The symbol $\mathcal{D}_\mu$ is a gauge-covariant derivative operator.
+The notation $x^\dagger$ indicates the hermitian conjugate of $x$.
 The potential terms are isolated into square brackets to highlight the negative

### 2023-04-10 main/chapter_theory.tex
write some lagrangian for mssm

@@
 
+\subsection{Lagrangian formulation}
+\label{sec:theory_susy_lagrangian}
+In Section~\ref{sec:theory_sm_lagrangian}, we presented a highly compressed
+summary of the Standard Model Lagrangian.
+Despite its minimality, the MSSM is a far more complicated theory than the
+Standard Model, and the expression of its Lagrangian formulation is
+far larger.
+This section therefore presents and even more compressed summary of the
+Lagrangian formulation of the MSSM.
+
+Although any viable theory must include the Standard Model in an appropriate
+appropriation, the Standard Model's Lagrangian need not be explicitly included.
+Indeed, it is not convenient to express the MSSM's Lagrangian as an add-on to
+the Standard Model. It stands alone.
+
+A convenient first decomposition of the MSSM Lagrangian's parts splits it into
+four parts:
+\begin{equation}
+\mathcal{L}_\mathrm{MSSM} =
+\mathcal{L}_\mathrm{SUSY}
++ \mathcal{L}_\mathrm{soft}
++ \mathcal{L}_\mathrm{gf}
++ \mathcal{L}_\mathrm{ghost}
+,
+\end{equation}
+in which
+$\mathcal{L}_\mathrm{SUSY}$ is the supersymmetric part,
+$\mathcal{L}_\mathrm{soft}$ is the a `softly' breaking part
+(which is necessary because supersymmetry is not an unbroken symmetry of
+nature),
+and
+$\mathcal{L}_\mathrm{gf}$ and $\mathcal{L}_\mathrm{ghost}$
+are two additions that are necessary for quantization of the
+theory~\cite{kuroda1999complete}.
+
+\paragraph{Supersymmetry.}
+The supersymmetric summand $\mathcal{L}_\mathrm{SUSY}$ contains numerous terms
+that implement the main particle content of the MSSM.
+These terms are not trivially expressed as Lagrangian densities in
+four-dimensional spacetime, but are more fundamentally also integrated over
+`superspace' coordinates~\cite{martin2016primer, kuroda1999complete}.
+It contains kinetic terms for the gauge bosons, gauginos, fermions, sfermions,
+Higgs, and higgsinos fields.
+It also implements interactions between these fields,
+including interactions between the fermions and gauge bosons,
+Yukawa interactions between the fermions and Higgs fields and higgsinos,
+and interactions between the gauge bosons and Higgs fields.
+The gauge-Higgs interactions produce masses for the gauge bosons.
+Other terms directly implement masses for mixtures of the higgsino fields.
+
+\paragraph{Soft breaking.}
+With no known mechanism for supersymmetry breaking,
+the term $\mathcal{L}_\mathrm{soft}$ is expressed as a maximally general form
+with $30$ free parameters, which have only weak constraints to ensure
+symmetries in the theory.
+It includes possible mass terms for the constituent fields, as well as scalar
+interaction vertices.
+
+\paragraph{Others.}
+The gauge-fixing term $\mathcal{L}_\mathrm{gf}$ ensures that the gauge boson
+propagators take the Standard Model form.
+The ghost term $\mathcal{L}_\mathrm{ghost}$ introduces special scalar fields
+that interact with bosons in the theory, but which have no effect at tree-level
+but are necessary for self-consistency of the theory in loop diagrams.
+
+\subsection{Specific simplified models}
+\TODO{C1N2 and GMSB for $\twoljets$}
+
+\TODO{Cross-sections}
+
 % this comment activates enlarged spacing for the final paragraph lol

### 2023-04-10 main/chapter_theory.tex
begin waffle on c1n2; clarify covariant derivatives

@@
 \end{equation}
-where the $\mathcal{D}_\mu$ expressions corresponding to the fermion fields:
+where each of the the $\mathcal{D}_\mu$ operators expresses a unique covariant
+derivative that corresponds to the operand field.
+These fermion fields are labelled
 $L_i$ for lepton doublets,
@@
 contains the dynamics and potential for the scalar Higgs doublet.
-The symbol $\mathcal{D}_\mu$ is a gauge-covariant derivative operator.
+The symbol $\mathcal{D}_\mu$ is a gauge-covariant derivative operator
+specific to the Higgs field.
 The notation $x^\dagger$ indicates the hermitian conjugate of $x$.
@@
 + \mathcal{L}_\mathrm{soft}
-+ \mathcal{L}_\mathrm{gf}
++ \mathcal{L}_\mathrm{GF}
 + \mathcal{L}_\mathrm{ghost}
@@
 and
-$\mathcal{L}_\mathrm{gf}$ and $\mathcal{L}_\mathrm{ghost}$
+$\mathcal{L}_\mathrm{GF}$ and $\mathcal{L}_\mathrm{ghost}$
 are two additions that are necessary for quantization of the
@@
 \paragraph{Others.}
-The gauge-fixing term $\mathcal{L}_\mathrm{gf}$ ensures that the gauge boson
+The gauge-fixing term $\mathcal{L}_\mathrm{GF}$ ensures that the gauge boson
 propagators take the Standard Model form.
@@
 that interact with bosons in the theory, but which have no effect at tree-level
-but are necessary for self-consistency of the theory in loop diagrams.
+but are necessary for self-consistency of the theory in loop
+diagrams~\cite{martin2016primer, kuroda1999complete}.
 
 \subsection{Specific simplified models}
-\TODO{C1N2 and GMSB for $\twoljets$}
+Two specific simplified supersymmetric models are studied in this thesis.
+These models were selected for the \atlas\ `$\twoljets$'
+search~\cite{atlas2022searches},
+presented in Chapter~\ref{chapter:2ljets},
+because they are well-studied but not-yet excluded models that can affect the
+data in its targetted signature of data with two charged leptons,
+hadronic jets, and missing transverse momentum.
+Within the context of $\twoljets$, we name these two simplified models
+`C1N2' and `GMSB'.
+\TODO{add citations}
+
+\paragraph{C1N2.}
+`C1N2' is a restricted subset of the MSSM in which the only active SUSY
+particles
+are the lightest chargino ($\chargino_1$)
+and the lightest two neutralinos ($\neutralino_1$, $\neutralino_2$).
+The name `C1N2` is an abbreviation to
+`Chargino 1' concatenated with `Neutralino 2'.
+
+\paragraph{GMSB.}
+`GMSB' is \TODO{}.
 

### 2023-04-10 main/chapter_theory.tex
c1n2 waffle more

@@
 \subsection{Specific simplified models}
+\TODO{add citations}
 Two specific simplified supersymmetric models are studied in this thesis.
@@
 `C1N2' and `GMSB'.
-\TODO{add citations}
+Many \atlas\ searches distribute their efforts across many simplified models.
+This search patches just one part in the larger \atlas\ SUSY programme.
 
@@
 `Chargino 1' concatenated with `Neutralino 2'.
+This model is active in the electroweak sector of the MSSM, which has received
+increased interest in recent years because the LHC is more sensitive to models
+in the strong sector.
+Existing LHC have have already ruled out many strongly-interacting models
+up to multi-$\eV[T]$ mass scales, but some viable phase space remains in the
+electroweak sector.
+
+Chargino-neutralino pairs could be produced in proton collisions via off-shell
+$W$ boson propagators.
+The targeted experimental signature is achieved by decays of these gauginos
+to the $\neutralino_1$ LSP via emission of Standard Model particles.
 

### 2023-04-10 main/chapter_theory.tex
begin GMSB waffle

@@
 Two specific simplified supersymmetric models are studied in this thesis.
-These models were selected for the \atlas\ `$\twoljets$'
+These models were selected for the \atlas\ $\twoljets$
 search~\cite{atlas2022searches},
@@
 `C1N2' and `GMSB'.
-Many \atlas\ searches distribute their efforts across many simplified models.
-This search patches just one part in the larger \atlas\ SUSY programme.
+
+Both of these simplified models are micro-examples of complete new theories
+that offer solutions to the hierarchy
+problem~\cite{Sakai:1981gr,SUSY-2018-02}.
+Both models also conserve R-parity, so they satisfy constraints from proton
+decay and introduce a stable, neutral dark matter candidate in their respective
+LSPs~\cite{Ellis:1983ew}.
+Many \atlas\ searches distribute their efforts across many simplified models;
+this search embodies just one tile in the mosaic of \atlas's SUSY programme.
 
@@
 in the strong sector.
-Existing LHC have have already ruled out many strongly-interacting models
-up to multi-$\eV[T]$ mass scales, but some viable phase space remains in the
-electroweak sector.
+Existing LHC results have have already ruled out many strongly-interacting
+models up to multi-$\eV[T]$ mass scales, but some interesting phase space
+remains in the electroweak sector.
+We choose to study C1N2 to attend to this remaining phase space.
 
-Chargino-neutralino pairs could be produced in proton collisions via off-shell
-$W$ boson propagators.
+Chargino-neutralino pairs are produced in proton collisions via off-shell
+$W$ boson propagators in this model.
 The targeted experimental signature is achieved by decays of these gauginos
@@
 \paragraph{GMSB.}
-`GMSB' is \TODO{}.
+As described earlier in Section~\ref{sec:theory_susy}, GMSB models offer a
+general mechanism to break supersymmetry by communicating some symmetry
+breaking down from high mass scales~\cite{Meade:2008wd,SUSY-2018-02}.
+The specific GMSB model used in the \atlas\ $\twoljets$ search
+(and referred to as simply GMSB in its context) is a simplified model that
+captures broad patterns of plausible behaviours in an experimentally pragmatic
+format.
+Under this GMSB model, visible signatures are generated by decays higgsino-like
+neutralinos to Standard Model particles and a gravitino LSP, specifically
+$\neutralino_1 \to Z/h + \gravitino$.
+% TODO diverse decay channels
+% TODO diverse initial states
 
+\subsubsection{Cross-sections}
 \TODO{Cross-sections}

### 2023-04-10 main/chapter_theory.tex
begin cross-section

@@
 \subsection{Specific simplified models}
-\TODO{add citations}
 Two specific simplified supersymmetric models are studied in this thesis.
@@
 neutralinos to Standard Model particles and a gravitino LSP, specifically
-$\neutralino_1 \to Z/h + \gravitino$.
-% TODO diverse decay channels
-% TODO diverse initial states
+$\neutralino_1 \to Z/h + \gravitino$ from each side of a
+$\neutralino_1 \neutralino_1$ pair production event.
+
+We are experimentally interested in this decay mode because the $Z$ and Higgs
+bosons can decay to diverse Standard Model decay products.
+Again, the \atlas\ $\twoljets$ search chooses to target this model because
+a significant decay mode includes $Z \to \ell^+\ell^-$, jets, and invisible
+particles giving missing transverse momentum.
+This simplified GMSB model also chooses to make the SUSY mass spectrum
+`compressed', meaning that there are small mass differences between
+$\neutralino_1$ and the slightly-heavier $\neutralino_2$ and $\chargino_2$.
+This choice creates diverse initial states, since the hard scatter events can
+emit various different pairs of heavier gaugino species and they will all decay
+via unobservably soft emissions to produce the same
+$\neutralino_1 \neutralino_1$ target
+pair~\cite{atlas2022searches,hepdata.116034}.
 
 \subsubsection{Cross-sections}
-\TODO{Cross-sections}
+A substantial production cross-section is a necessary condition for a process
+to be observable with finite data.
+The Run~2 dataset studied by our $\twoljets$ search comprises
+$L = 139\,\mathrm{fb}^{-1}$
+of proton-proton collisions, meaning that a process with production
+cross-section $\sigma = 1\,\mathrm{fb}$
+should occur in the data with a Poisson mean of $\mu = L \sigma = 139$ times.
+To achieve mean $\mu \geq 1$ event in the dataset, a process needs a cross-section
+at least $1 / L \approx 7\,\mathrm{pb}$.
+
+Cross-sections for the $\twoljets$ C1N2 and GMSB models are determined by two
+factors: production cross-section and the decay branching fraction.
+Each process pair-produces some heavy parent particles, so their production
+cross-sections scale down with the parent particle mass
+as both parton distribution hard-scatter matrix element
+are pushed to harder scales.
+For mass scale $q$, these three factors combine to an approximate scaling
+of $\sigma \sim q^{-6}$ in the several-hundred-$\,\eV[G]$ regime.
+Branching fractions in decay further reduce the observable cross-sections in
+our selected decay channels; however, these are constant factors that do not
+scale with parent mass.
+
+\TODO{figure}
+
+\TODO{example numbers}
 

### 2023-04-10 main/chapter_theory.tex
add pdg figure

@@
 particles giving missing transverse momentum.
-This simplified GMSB model also chooses to make the SUSY mass spectrum
-`compressed', meaning that there are small mass differences between
-$\neutralino_1$ and the slightly-heavier $\neutralino_2$ and $\chargino_2$.
+This simplified GMSB model also chooses a
+`compressed' mass spectrum, meaning small differences between the masses of
+$\neutralino_1$ and the slightly-heavier $\chargino_1$ and $\neutralino_2$.
 This choice creates diverse initial states, since the hard scatter events can
@@
 cross-section $\sigma = 1\,\mathrm{fb}$
-should occur in the data with a Poisson mean of $\mu = L \sigma = 139$ times.
-To achieve mean $\mu \geq 1$ event in the dataset, a process needs a cross-section
-at least $1 / L \approx 7\,\mathrm{pb}$.
+yields a Poisson mean of $\mu = L \sigma = 139$ events.
+To achieve a mean of one event in the dataset,
+below which it becomes exceedingly hard to observe,
+a process needs a cross-section at least $1 / L \approx 7\,\mathrm{pb}$.
 
@@
 are pushed to harder scales.
-For mass scale $q$, these three factors combine to an approximate scaling
-of $\sigma \sim q^{-6}$ in the several-hundred-$\,\eV[G]$ regime.
+For mass scale $q$, these three factors combine to an rapidly decaying scaling
+of cross-section with parent mass, which goes very roughly like
+$\sigma \sim q^{-6}$ in the several-hundred-$\,\eV[G]$ regime.
+Example production cross-sections for similar models are displayed in
+Figure~\ref{fig:theory_susy_xs}.
 Branching fractions in decay further reduce the observable cross-sections in
@@
 
-\TODO{figure}
+\begin{figure}[tp]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/theory_pdg_susy_xsec.pdf}
+\caption[
+Cross-section scaling of some supersymmetric pair production modes
+]{%
+Cross-section scaling of some supersymmetric pair production modes
+in models similar to the $\twoljets$ C1N2 and GMSB simplified models.
+Note the similar scaling parent particle mass  for all gaugino and slepton
+production modes shown;
+this effect is due to fundamental properties of parton distribution functions
+and scattering matrix-elements, and is not unique to supersymmetry.
+These cross-sections do not include branching fraction factors, which are
+needed to determine the C1N2 and GMSB cross-sections.
+This figure is reproduced from~\cite{pdg2022ynf}.
+\\[0.4em]
+The label `wino` indicates that the gaugino mixtures mostly comprise the
+superpartners of the $W$ fields, and `higgsino` indicates that they instead
+comprise higgsino fields.
+\\[0.4em]
+Cross-sections below around $7\,\mathrm{pb}$ are expected to produce no
+events in our $139\,\mathrm{fb}^{-1}$ of data.
+}
+\label{fig:theory_susy_xs}
+\end{figure}
 

### 2023-04-10 main/chapter_theory.tex
paste some cross-section numbers

@@
 yields a Poisson mean of $\mu = L \sigma = 139$ events.
-To achieve a mean of one event in the dataset,
-below which it becomes exceedingly hard to observe,
-a process needs a cross-section at least $1 / L \approx 7\,\mathrm{pb}$.
-
-Cross-sections for the $\twoljets$ C1N2 and GMSB models are determined by two
-factors: production cross-section and the decay branching fraction.
-Each process pair-produces some heavy parent particles, so their production
-cross-sections scale down with the parent particle mass
-as both parton distribution hard-scatter matrix element
-are pushed to harder scales.
-For mass scale $q$, these three factors combine to an rapidly decaying scaling
-of cross-section with parent mass, which goes very roughly like
-$\sigma \sim q^{-6}$ in the several-hundred-$\,\eV[G]$ regime.
-Example production cross-sections for similar models are displayed in
-Figure~\ref{fig:theory_susy_xs}.
-Branching fractions in decay further reduce the observable cross-sections in
-our selected decay channels; however, these are constant factors that do not
-scale with parent mass.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/theory_pdg_susy_xsec.pdf}
+\includegraphics[width=0.98\textwidth]{figures/theory_pdg_susy_xsec.pdf}
 \caption[
@@
 These cross-sections do not include branching fraction factors, which are
-needed to determine the C1N2 and GMSB cross-sections.
+needed to determine both C1N2 and GMSB model cross-sections.
+The C1N2 model is wino-like, and the GMSB model is higgsino-like.
 This figure is reproduced from~\cite{pdg2022ynf}.
@@
 
-\TODO{example numbers}
+Cross-sections for the $\twoljets$ C1N2 and GMSB models are determined by two
+factors: production cross-section and the decay branching fraction.
+Each process pair-produces some heavy parent particles, so their production
+cross-sections scale down with the parent particle mass
+as both parton distribution hard-scatter matrix element
+are pushed to harder scales.
+For mass scale $q$, these three factors combine to an rapidly decaying scaling
+of cross-section with parent mass, which goes very roughly like
+$\sigma \sim q^{-6}$ in the several-hundred-$\,\eV[G]$ regime.
+Production cross-sections for these models are displayed in
+Figure~\ref{fig:theory_susy_xs};
+in this figure,
+cross-sections for the C1N2 model are correspond to those labelled
+`$\chargino_1 \neutralino_2$ (wino)', and
+cross-sections for the GMSB model are a little smaller, and correspond to those
+labelled `$\tilde \chi \tilde \chi$ (higgsino)'.
+
+Typical scenarios studied in $\twoljets$ parent masses ranging from
+around $200\,\eV[G]$ to $800\,\eV[G]$ under both of its C1N2 and GMSB models.
+The production cross-section for signals from its C1N2 model is
+$1800\,\mathrm{fb}$ at $m(\mathrm{parent}) = 200\,\eV[G]$,
+and $4.8\,\mathrm{fb}$ at $m(\mathrm{parent}) = 800\,\eV[G]$.
+The production cross-section for signals from its GMSB model is
+$1300\,\mathrm{fb}$ at $m(\mathrm{parent}) = 200\,\eV[G]$,
+and $3.5\,\mathrm{fb}$ at $m(\mathrm{parent}) = 800\,\eV[G]$~\cite{
+Fuks:2012qx,Fuks:2013vua}.
+These values are rounded to two significant figures.
+Cross-section uncertainties are shaded in  Figure~\ref{fig:theory_susy_xs},
+but appear very small on its log scale;
+the relative uncertainties increase with parent mass from about
+$\pm 6\%$ at $m(\mathrm{parent}) = 200\,\eV[G]$.
+to about
+$\pm 12\%$ at $m(\mathrm{parent}) = 800\,\eV[G]$
+in both models.
+
+Although these cross-sections imply that several hundred signal events are
+expected even at $m(\mathrm{parent}) = 800\,\eV[G]$, the raw number of signal
+events does not mean that the signal is visible.
+Visible signals must necessarily also be recorded as data,
+distinguishable from Standard Model backgrounds.
+Indeed, much of the work on a search like $\twoljets$ pertains to this
+data selection and the separation of signals from backgrounds.
+Furthermore, these are raw production cross-sections;
+the yield of signal events that also decay into selected products is much
+smaller ---
+for starters, a $Z$ boson decays to electrons or muons with only around
+$10\%$ branching fraction~\cite{pdg2022ynf}.
+Such diversity in decay modes motivates the wide variety of targets in
+\atlas's many SUSY searches.
 

### 2023-04-10 main/chapter_theory.tex
begin majorana neutrino bit

@@
 by pinpointing a uniquely correct model.
+
+\subsubsection{Future Majorana-Dirac possibilities}
+% TODO deleteme
+% • Opportunities for future SMs – include description of experiemental
+% expectations for Majorana vs Dirac neutrinos (double-beta decay, accelerator
+% neutrino experiments (DUNE, HyperK), reactors etc.
+Future experimental results might resolve the nature of neutrinos.
+% 0vBB
+If neutrinos have a Majorana nature, then nuclear isotopes that naturally
+undergo double-beta decay (emitting two electrons and two neutrinos)
+cold rarely undergo neutrinoless double-beta decay ($0\nu\beta\beta$).
+In $0\nu\beta\beta$ events, the two neutrinos annihilate to leave the electrons
+emitted back-to-back without missing momentum.
+Current results have not observed $0\nu\beta\beta$ decays, but have set limits
+on $0\nu\beta\beta$ lifetimes in excess of $10^{21}\,\mathrm{years}$
+for many nuclear isotopes~\cite{Bilenky:2014uka}.
+Future experiments, such as SuperNEMO~\cite{Barabash:2011row},
+will study $0\nu\beta\beta$ with increased precision.
+
+% accelerator experiments
+Large upcoming neutrino experiments,
+DUNE~\cite{DUNE:2016hlj} and
+Hyper-Kamiokande~\cite{Hyper-Kamiokande:2018ofw},
+will use accelerator-produced neutrinos to make precise measurements of the
+neutrinos' Dirac CP violation,
+their mass ordering, and their flavour oscillations;
+although these experiments do not directly probe the neutrinos' possible
+Majorana components, they will be indirectly sensitive through possible new
+effects~\cite{Carrasco-Martinez:2020mlg}.
+
+% reactors
+Modern experiments that detect neutrinos from nuclear reactors,
+such as JUNO % ~\cite{TODO},
+are not also
+\TODO{reactors}
+% etc??
+\TODO{etc??}
+
+Other approaches to testing neutrinos' Majorana nature are not competitive with
+$0\nu\beta\beta$. These proposed alternatives include \TODO{}
+% https://link.springer.com/article/10.1007/JHEP01(2019)122
+% https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.112.131802
+% https://www.sciencedirect.com/science/article/pii/S0370269318310001?via%3Dihub
+% https://journals.aps.org/prd/abstract/10.1103/PhysRevD.96.075006
 

### 2023-04-10 main/chapter_theory.tex
do reactors, begin etc

@@
 \subsubsection{Future Majorana-Dirac possibilities}
-% TODO deleteme
-% • Opportunities for future SMs – include description of experiemental
-% expectations for Majorana vs Dirac neutrinos (double-beta decay, accelerator
-% neutrino experiments (DUNE, HyperK), reactors etc.
 Future experimental results might resolve the nature of neutrinos.
@@
 Modern experiments that detect neutrinos from nuclear reactors,
-such as JUNO % ~\cite{TODO},
-are not also
-\TODO{reactors}
-% etc??
-\TODO{etc??}
+such as JUNO~\cite{JUNO:2015sjr},
+are also not directly sensitive to the question of the neutrinos' Majorana
+nature.
+Along with accelerator experiments, their precise measurements of oscillation
+behaviours will, however, help the $0\nu\beta\beta$ to more precisely
+interpret their data~\cite{Cao:2017drk}.
 
-Other approaches to testing neutrinos' Majorana nature are not competitive with
-$0\nu\beta\beta$. These proposed alternatives include \TODO{}
+% etc
+Other approaches to testing neutrinos' Majorana nature include studies of \TODO{}
 % https://link.springer.com/article/10.1007/JHEP01(2019)122
@@
 % https://journals.aps.org/prd/abstract/10.1103/PhysRevD.96.075006
+However, these methods are not competitive with the next generation of
+$0\nu\beta\beta$ experiments and currently show no Majorana
+signal~\cite{Carrasco-Martinez:2020mlg}.
 

### 2023-04-10 main/chapter_theory.tex
finish up neutrino majorana etc

--- 2023-04-10 main/chapter_theory.tex
@@
 by pinpointing a uniquely correct model.
+
+\subsubsection{Future Majorana-Dirac possibilities}
+Future experimental results might resolve the nature of neutrinos.
+% 0vBB
+If neutrinos have a Majorana nature, then nuclear isotopes that naturally
+undergo double-beta decay (emitting two electrons and two neutrinos)
+could rarely perform neutrinoless double-beta decay ($0\nu\beta\beta$).
+In $0\nu\beta\beta$ events, the two neutrinos annihilate to leave the electrons
+emitted back-to-back without missing momentum.
+Current results have not observed $0\nu\beta\beta$ decays, but have set limits
+on $0\nu\beta\beta$ lifetimes in excess of $10^{21}\,\mathrm{years}$
+for many nuclear isotopes~\cite{Bilenky:2014uka}.
+Future experiments, such as SuperNEMO~\cite{Barabash:2011row},
+will study $0\nu\beta\beta$ with increased precision.
+
+% accelerator experiments
+Large upcoming neutrino experiments,
+DUNE~\cite{DUNE:2016hlj} and
+Hyper-Kamiokande~\cite{Hyper-Kamiokande:2018ofw},
+will use accelerator-produced neutrinos to make precise measurements of the
+neutrinos' Dirac CP violation,
+their mass ordering, and their flavour oscillations.
+Although these experiments do not directly probe the neutrinos' possible
+Majorana components, they will be indirectly sensitive via the possibility
+of exotic new physical effects~\cite{Carrasco-Martinez:2020mlg}.
+
+% reactors
+Modern experiments that detect neutrinos from nuclear reactors,
+such as JUNO~\cite{JUNO:2015sjr},
+are also not directly sensitive to the question of the neutrinos' Majorana
+nature.
+Along with accelerator experiments, their precise measurements of oscillation
+behaviours will, however, help the $0\nu\beta\beta$ studies to more precisely
+interpret their data~\cite{Cao:2017drk}.
+
+% etc
+Other approaches to testing neutrinos' Majorana nature include studies of
+proton-proton collisions~\cite{CMS:2018jxx} or
+rare $B$-hadron decays~\cite{LHCb:2014osd,Shuve:2016muy}.
+However, these methods are not competitive with the next generation of
+$0\nu\beta\beta$ experiments and currently show no Majorana
+signal~\cite{Carrasco-Martinez:2020mlg}.
 

### 2023-04-29 main/chapter_2ljets.tex
refer to new atlas fnp paper
incorporate Val's round 2 comments

@@
 \caption[
-Feynman diagrams for the Standard Model background processes
-]{%
-Feynman diagrams for the Standard Model background processes of important
-background contributions in the $\twoljets$-electroweak search.
+Example Feynman diagrams for the Standard Model background processes
+]{%
+Example Feynman diagrams for the Standard Model background processes of
+important background contributions in the $\twoljets$-electroweak search.
 In \diboson\ events we expect one $Z$ boson-like decay to charged leptons, and
@@
 as smaller limits on the signal expectations, which are labelled
-$S_{\mathrm{obs}}^{95}$ in the figure.
+$S_{\mathrm{obs}}^{95}$ in this table.
 \footnotetext{%
@@
 the signal leptons are a subset of the baseline leptons.
-Therefore if a lepton is a signal lepton, then it is also a baseline lepton,
+Therefore, if a lepton is a signal lepton, then it is also a baseline lepton,
 but for clarity we often prefer to state that both requirements are met.
@@
 for more extreme models.
-For example, the corresponding capture rate it $0.64\%$ for the C1N2 model with
+For example, the corresponding capture rate is $0.64\%$ for the C1N2 model with
 $m(\chargino_1, \neutralino_2) = 650\,\eV[G]$ and
@@
 Appendix~\ref{app:2ljets_cut_flows} present cut-flow tables in the four control
-regions of $\twoljets$-electroweak with background sample rates included.
+regions of $\twoljets$-electroweak, with background sample rates included.
 These tables show that $\mttwoll$ selections are powerful for removing $\ttbar$
@@
 Although these Low-ISR regions were removed from the analysis and we never
-inspected the data in SR-Low-ISR (never unblinded), I include them here because
-they reveal some interesting context to the surviving analysis design.
+inspected the data in SR-Low-ISR (this signal region was never unblinded),
+I include them here because they reveal some interesting context to the
+surviving analysis design.
 Specifically, prototype validation regions for Low-ISR designs revealed some
@@
 Since it includes actual data, this method unblinds all regions,
-including signal regions, early.
+including signal regions, early~\cite{ATLAS:2022swp}.
 And because it knows and uses the data we are supposed to be predicting, it is
@@
 parent mass $m(\chargino_1, \neutralino_2) = m_a$
-and the mass splitting $m_a - m_b$, and we want an extrapolation scheme which
+and the mass splitting $m_a - m_b$, and we want an extrapolation scheme that
 aligns changes in these variables with changes in the extrapolated value.
@@
 Figure~\ref{fig:2ljets_summary}.
-Unlike Figure~\ref{fig:2ljets_summary} the this figure does not adjust its
+Unlike Figure~\ref{fig:2ljets_summary}, this figure does not adjust its
 background estimates to match the observed data in signal regions; notably,
@@
 Just as the LHC follows the LEP electron-positron collider, this FCC hadron
-collider (FCC-hh) might be proceeded by an electron-positron collider (FCC-ee)
-with centre-of-mass energies high enough (up to $\sqrt{s}=350\,\eV[T]$) to
+collider (FCC-hh) might be preceeded by an electron-positron collider (FCC-ee)
+with centre-of-mass energies high enough (up to $\sqrt{s}=350\,\eV[G]$) to
 pair-produce all Standard Model particles~\cite{FCC:2018evy}.

### 2023-04-29 main/chapter_experiment.tex
incorporate Val's round 2 comments

@@
 around $100\,\mathrm{m}$ under France and Switzerland at CERN, Geneva.
-It and comprises two rings that contain and accelerate counter-rotating bunches
+It comprises two rings that contain and accelerate counter-rotating bunches
 of particles.
@@
 Heavy ion collisions are primarily studied by the \alice\ experiments, but are
-also analysed by the general-purpose detectors \atlas\ and \cms.
+also analysed by the general-purpose detectors \atlas\ and \cms,
+as well as \lhcb.
 Since proton collision data are the topic of this thesis, this section focuses
@@
 % LHC
-\item Finally, protons from the SPS are injected the LHC, which finishes
+\item Finally, protons from the SPS are injected into the LHC, which finishes
 their acceleration from $250\,\eV[G]$ to $6.5\,\eV[T]$, achieving the
@@
 They operate by setting up electron beams to oscillate at a frequency that
-resonates with a specially designed cavity, this producing strong oscillating
+resonates with a specially designed cavity, thus producing strong oscillating
 electromagnetic fields.
@@
 a total focussing effect; their effects are not linear.
+% "Give number of magnets" - internal examiner
+The LHC includes
+$1232$ dipole magnets
+and $4800$ higher-pole `corrector' magnets~\cite{lhc2008machine}.
 
@@
 % size of beam
-By its peak luminosity design specifications, each bunch of an LHC beam
+At its peak luminosity by design specifications, each bunch of an LHC beam
 should contain $1.15\times10^{11}$ protons.
@@
 }.
-The relative uncertainty on \atlas's Run~2 integrated luminosity is $\pm1.7\%$.
+The relative uncertainty on \atlas's Run~2 integrated luminosity is
+$\pm1.7\%$~\cite{atlas2022searches}.
 
@@
 If not removed, these pile-up signals can skew measurements and create an
-illusion missing transverse momentum.
+illusion of missing transverse momentum.
 Although precise measurements can help to remove this noise, it reduces the
@@
 These detectors' resolutions to relevant physical objects is described in
-Section~\label{sec:atlas_object_resolutions}.
+Section~\ref{sec:atlas_object_resolutions}.
 The final connection to data analysed in this thesis is made through the
@@
 Electron reconstruction algorithms efficiently recover this lost Bremsstrahlung
-momentum by associating nearby photon calorimeter clusters back to their source
-electron many cases~\cite{atlas1994proposal, atlas1999design1}.
+momentum by associating nearby photon-induced calorimeter clusters back to
+their source electron many cases~\cite{atlas1994proposal, atlas1999design1}.
 
@@
 $b_3 \approx 0.03$.
-With these parameters, formula approximately describes the measured relative
-jet resolutions of
+With these parameters, this formula approximately describes the measured
+relative jet resolutions of
 $\frac{\sigma(\pt)}{\pt} \approx 20\%$ at $\pt = 30\,\eV[G]$,
@@
 The search presented in this thesis uses only dilepton $e$ and $\mu$ triggers,
-which are highlighted in bold.
-Table adapted from~\cite{ATL-DAQ-PUB-2019-001}.
+which are highlighted in \textbf{bold}.
 Rates overlap between triggers.
+This table is adapted from~\cite{ATL-DAQ-PUB-2019-001}.
 }
@@
 Bracketed numbers under 2017 refer to an alternative `8b4e' LHC operating
-condition which alternated in a pattern of eight bunches with the usual
+condition, which alternated in a pattern of eight bunches with the usual
 $25\,\mathrm{ns}$ separation followed by a gap of four empty bunch slots.

### 2023-04-29 main/chapter_searches.tex
incorporate Val's round 2 comments

@@
 model
-(for example, try SR-HM(disc.) of $1\ell bb$~\cite{SUSY-2019-08,hepdata.90607})
+(for example, I recommend trying SR-HM(disc.) of
+$1\ell bb$~\cite{SUSY-2019-08,hepdata.90607})
 with standard Markov-Chain Monte Carlo algorithms.

### 2023-04-29 main/chapter_theory.tex
incorporate Val's round 2 comments

@@
 Supersymmetry is an idea that has been applied to build plausible alternatives
-to the Standard Model. and that may explain future experimental discoveries.
+to the Standard Model, and which may explain future experimental discoveries.
 Since the main work of this thesis is a search for effects of supersymmetric
@@
 
-A convenient first decomposition of the MSSM Lagrangian's parts splits it into
-four parts:
+A convenient first decomposition of the MSSM Lagrangian splits it into four
+parts:
 \begin{equation}
@@
 \\[0.4em]
-The label `wino` indicates that the gaugino mixtures mostly comprise the
-superpartners of the $W$ fields, and `higgsino` indicates that they instead
+The label `wino' indicates that the gaugino mixtures mostly comprise the
+superpartners of the $W$ fields, and `higgsino' indicates that they instead
 comprise higgsino fields.
@@
 cross-sections scale down with the parent particle mass
-as both parton distribution hard-scatter matrix element
+as both parton distribution functions hard-scatter matrix elements
 are pushed to harder scales.
-For mass scale $q$, these three factors combine to an rapidly decaying scaling
+For mass scale $q$, these three factors combine to a rapidly decaying scaling
 of cross-section with parent mass, which goes very roughly like
@@
 in this figure,
-cross-sections for the C1N2 model are correspond to those labelled
+cross-sections for the C1N2 model correspond to those labelled
 `$\chargino_1 \neutralino_2$ (wino)', and
@@
 
-Typical scenarios studied in $\twoljets$ parent masses ranging from
-around $200\,\eV[G]$ to $800\,\eV[G]$ under both of its C1N2 and GMSB models.
+Typical scenarios studied in $\twoljets$ have parent particle masses ranging
+from around $200\,\eV[G]$ to $800\,\eV[G]$, under both of its C1N2 and GMSB
+models.
 The production cross-section for signals from its C1N2 model is
@@
 Cross-section uncertainties are shaded in  Figure~\ref{fig:theory_susy_xs},
-but appear very small on its log scale;
+but appear very small on its logarithmic scale;
 the relative uncertainties increase with parent mass from about

### 2023-04-29 main/chapter_experiment.tex
incorporate Karolos's suggestions; format bib

@@
 }.
-The relative uncertainty on \atlas's Run~2 integrated luminosity is
-$\pm1.7\%$~\cite{atlas2022searches}.
+The relative uncertainty on \atlas's Run~2 integrated luminosity,
+in the form that we use in the $\twoljets$ search,
+is $\pm1.7\%$~\cite{atlas2022searches,ATLAS-CONF-2019-021}.
+More recent measurements have reduced this uncertainty to
+$\pm0.83\%$~\cite{ATLAS:2022hro}.
 

### 2023-05-05 main/frontmatter.tex
swap declaration-abstract order; add title and author

@@
 }
-
-\begin{abstract}
-Searching for new phenomena is an essential endeavour of experimental physics,
-and is helped by guidance from plausible theoretical predictions.
-This thesis presents such a search for signals of high-energy protons
-scattering to massive resonances with dramatic decays, which are predicted by
-supersymmetric models.%
-% hack: hidden reference to constrain the reference numbering
-\sbox0{\cite{atlas2022searches}}
-
-No new phenomena appear in the results, so they support the Standard Model in
-its continuing successful extrapolation to describe the most extreme behaviours
-of proton collisions.
-Supersymmetry as a concept is not falsified by this lack of new phenomena,
-nor by the many similar observations.
-But it is also not confirmed.
-
-Data are best interpreted with an understanding of their origins, so we begin
-by describing the theoretical and experimental background to our search.
-Our results also depend strongly on the interpretation of data by conventional
-analysis procedures, whose properties and explanations are not obvious.
-To explain these procedures, we also discuss their historical context and the
-basic theory of data analysis.
-\end{abstract}
 
@@
 \end{singlespacing}
+
+\begin{abstract}
+\textbf{Title:} Two leptons, jets, and missing transverse momentum in ATLAS:
+yet more non-evidence for supersymmetry
+
+\noindent
+\textbf{Author:} Rupert~Tombs
+\\
+
+Searching for new phenomena is an essential endeavour of experimental physics,
+and is helped by guidance from plausible theoretical predictions.
+This thesis presents such a search for signals of high-energy protons
+scattering to massive resonances with dramatic decays, which are predicted by
+supersymmetric models.%
+% hack: hidden reference to constrain the reference numbering
+\sbox0{\cite{atlas2022searches}}
+
+No new phenomena appear in the results, so they support the Standard Model in
+its continuing successful extrapolation to describe the most extreme behaviours
+of proton collisions.
+Supersymmetry as a concept is not falsified by this lack of new phenomena,
+nor by the many similar observations.
+But it is also not confirmed.
+
+Data are best interpreted with an understanding of their origins, so we begin
+by describing the theoretical and experimental background to our search.
+Our results also depend strongly on the interpretation of data by conventional
+analysis procedures, whose properties and explanations are not obvious.
+To explain these procedures, we also discuss their historical context and the
+basic theory of data analysis.
+\end{abstract}
 
